### 2023-06-06 phd-cs-us.tex
init

\documentclass[10pt]{article}
\usepackage[margin=1.1in]{geometry}
%\usepackage[small,compact]{titlesec} %very powerful
\usepackage{tcolorbox}
\usepackage{enumitem}
\usepackage[T1]{fontenc}
\usepackage{cite}
%\usepackage{caption}
%\captionsetup{font=small}
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{hyperref}
\hypersetup{
    colorlinks,
    citecolor=black,
    filecolor=black,
    linkcolor=blue,
    urlcolor=blue,
}
\usepackage{booktabs}

\newcommand{\tool}{\textsf{NeuralSAT}}
\def\Section{\S}

\newcommand{\mycomment}[3][\color{black}]{{#1{{#2}: {#3}}}}
\newcommand{\tvn}[1]{\mycomment{TVN}{#1}}{}
\newcommand{\red}[1]{{\color{red}{#1}}}
\renewcommand{\figurename}{Fig.}
\renewcommand{\tablename}{Tab.}
\title{Demystifying the Computer Science Ph.D. Admission in the US}
\date{}
\author{ThanhVu (Vu) Nguyen with inputs from faculty at GMU and other universities}
%\setcounter{secnumdepth}{3}
\begin{document}
\maketitle

\begin{abstract}
Having been involved in PhD admissions for many years and after
numerious interaction international (especially Vietnamese) students, I've
realized that international students lack a clear understanding of
the Computer Science PhD admission process at US universities. This confusion not only
discourages students from applying but also creates the perception that
getting admitted is difficult compared to CS PhD programs in other countries.

So I want to share my opinions and advice for those who are interested in applying for a \textbf{PhD in Computer Science in the US}.
While I wrote this document for CS applicants, it should be applicable to various STEM fields.
Moreover, while many examples given are for GMU, which is ranked 33rd on CSRankings (June 2016), the writing should be generalize to most other R1 universities  (though \emph{very} top schools migh be very extreme, e.g., see the admission process at \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{CMU}.

I wish you the best of luck. And if you follow these advice,
you will at least have a good chance at GMU (see
\href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-GMU}{why
you want to study at GMU}). Happy hunting!

If you have suggestions or comments, feel free to create a \href{https://github.com/dynaroars/dynaroars.github.io/issues}{GitHub issue} for discussion.
\end{abstract}

\section{Should You Apply?}

First, I want to emphasize that PhD students in Computer
Science \emph{do not} need to worry about funding, especially at R1
universities in the US. If you are admitted, you will almost certainly
receive \emph{full funding} to support your study, including tuition,
health insurance, and 9-month stipend. Moreover, depending on the university,
you may even receive additional benefits like summer salary. Note that
funding is typically more available for PhD students than 
Masters. % [[#funding][Read more about funding here]]

Second, I believe that applying to a good US university \emph{should not} be any
harder than at good schools in other countries. If you think you have a
chance at schools in other countries, e.g., South Korea, Singapore, Germany, UK, Japan and Australia, then you surely have a chance in
the US as well.

\begin{tcolorbox}[left=1pt,right=1pt,top=1pt,bottom=1pt]
Vu: One of the reasons I create this post is that several of my colleagues are interested in 
recruiting Vietnamese students and were surprised when seeing very few Vietnamese applications (compared to other countries). In general the number of
PhD applications from Vietnam to US universities is very low and  more would be very welcomed. 
\end{tcolorbox}

%% Vu: *what's a PhD?*  This [[https://matt.might.net/articles/phd-school-in-pictures/][series of pictures]] from [[https://matt.might.net][Matt Might]] illustrates what a PhD means.
%% \end{tcolorbox}

\section{How is your application evaluated?}

After you submit your PhD application (usually in December), it will be first screened
for general requirements, e.g., did you submit your transcripts and standard scores? did your writters submit their letters?
Then your application will be reviewed by a
\textbf{PhD admission committee} consisting of faculty members in CS. Each application is reviewed by about \emph{three} faculty members, who will evaluate your profile and try to reach a consensus about your case.

Many PhD committee members are \emph{new assistant professors}. This helps junior faculty, who typically don't have students yet, the opportunities to recruit students. The chair of the committee will be a senior professor, but they will not review individual applicaitions and instead assign them to committee members. The chair will look at various factors such as keywords/interests, research areas, or mentioning faculty names, and assign the application to appropriate faculty.

\begin{tcolorbox}[left=1pt,right=1pt,top=1pt,bottom=1pt]
Vu: We usually decide that a candidate is either (i) admit with funding (TA or RA) or (ii) rejected. In other words, in most cases, we either
admit you with full funding, or we don't. In some rare cases, we may admit
you with no funding because you have funding on your own (e.g., you are
supported by your government or have external grants). Also, we justify
our decision with a short summary about your application, where we list
strengths (e.g., well-known undergrad school) and weaknesses (e.g., weak
LORs).
\end{tcolorbox}

\begin{tcolorbox}[left=1pt,right=1pt,top=1pt,bottom=1pt]
    Hakan: At GMU, for full consideration, students should make sure to submit \textbf{ALL} required documents by the application deadline (December 1st), and should never assume that some required documents (such as official TOEFL scores or official diplomas/transcripts) will be waived by the admissions office. If something is listed and not marked as "optional", it is mandatory and they should plan for submitting all those.  
\end{tcolorbox}
\paragraph{Why we do not waive application fee?}  This is typically a requirement of the university. Individual departments/programs do not have the flexibility to waive the application fee, even if they want to.

In my opinion, requiring applicants to pay the fee helps ensure their seriousness, as it filters out non-serious candidates. If the application process were free for everyone, we would receive an overwhelming number of applications to review.

\section{Application}\label{sec:application}

The primary focus of the admissions committee is to \textbf{evaluate your background and interest in research} since a PhD in Computer Science
is a research degree. To assess your research capability, we consider
the following key indicators.

\subsection{Research Ability}

The most effective evidence of research ability is having \textbf{published papers in reputable international conferences or journals}.
Having published good papers is a sign that the applicant was involved in research.

It's important to aim for the top conferences in your field, which you can
find on places such as CSRankings~\cite{csrankings}, designed specifically to help CS PhD
applicants. Local conferences and non-English journals or conferences do
not carry as much weight since their quality is often unknown to US
faculty.

However, I understand that many international students do not have the opportunities to publish in top places, so general IEEE/ACM/USENIX confs/journals would suffice.  But be sure to talk about it in your \hyperref[sec:research-statement]{statement}.

\begin{tcolorbox}[left=1pt,right=1pt,top=1pt,bottom=1pt]
Vu: Vietnamese students often mention Scopus Q1, which consists of diversed journals from IEEE, Elsevier, and several other publishers unfamiliar to me.  Honestly, I don't know/recognize the majority of journals listed in Scopus Q1. So this might be something to be mindful of, as CS faculty in the US might not be too familiar with Scopus or journals listed in there, so devote sometime in your statement to discuss their significance.
\end{tcolorbox}

\begin{tcolorbox}[left=1pt,right=1pt,top=1pt,bottom=1pt]
Craig: GMU and many other universities allow you to upload your published papers. In many cases, even if the papers were not published at top places, we can still determine their quality by simply skimming over the paper.  
\end{tcolorbox}

Additionally, \textbf{work experiences at renowned research laboratories}, such as Microsoft Research, can significantly strengthen your
application.  Unfortunately, many good international research places, e.g., VinAI in Vietnam, remain relatively unknown to most universities in the US. So you should explicit say something about them in your statement.

Finally, \textbf{participating internationally recognized competitions} can also demonstrate your research potential.
For example, participating in Math Olympiads if you want to do theory or  winning ACM programming contests if you want to ``build'' stuff, e.g., software analysis.

\subsection{Letters of Recommendation (LORs)}

Most CS PhD applications will require at least \textbf{two LORs}. Having a letter from an internationally recognized researcher can greatly strengthen your application. However, obtaining such letters
can be challenging for international students, who might not have much interactions with such experts. So it is acceptable to have a letter from professors that know you well enough to talk about \emph{your specific research experience and capabilities}.

Many students have letters written by the applicants themselves and signed by their professors. These have little
values (we can easily recognize them) and will consider them weakness.
Similarly, many professors write very generic letters for students (a common example is that the students didn't do any
research or make any impression for the professor to write about). These
letters are also not useful and considered weak.

Many students get letters from someone from company where they did internship or are
working at. It is OK as long as it is a research-based personalized
letter (once again, we are talking about PhD applications, not MS).

\begin{tcolorbox}[left=1pt,right=1pt,top=1pt,bottom=1pt]
Vu: For me personally, it's better to have a good personalized
letter about your own research ability from someone who is less
well-known than a generic/weak letter from a well-known person.
\end{tcolorbox}

\subsection{Research Statement}\label{sec:research-statement}

While you might not be able to get good LORs or change
\hyperref[sec:your-school]{where your go to school}, you have control of your
statement! So write it well because we do take it seriously.
A well-written LOR also shows that you can communicate, which is very important in research, and that you can effectively teach and communicate with students, which is important for TA funding (see \S\ref{sec:funding}).
%is important because if you need (GTA) funding, it will provide evidence
%that you can teach and communicate with students.

There are many guides on writing research statement, e.g.,~\cite{blattman2022writing},
so I will not talk too much about it. In short, discuss about your research vision and convince us that you can achieve it through your experience, e.g., published papers, or if you work on some projects by yourself, talk about it.

Finally, this is something very easy to do, but is missed by many
applicants: \textbf{customize the statement} for the school you're applying to,
e.g., why do you apply there ? name several professors who you're
interested in working with (in some cases your application will be sent to them for evaluation).
Be careful not to send wrong statement to wrong school or mixing
facts (e.g., talking about school X but mentioned about working with
profs. at school Y; and definitely do not talk about George Washington when applying to George Mason). I have seen such statements more time that I
should.

\begin{tcolorbox}[left=1pt,right=1pt,top=1pt,bottom=1pt]
Vu: I always read the research statement first and then LORs. If I am
persuaded by then, I would skim over other factors and advocate for
admission (unless I see red flags in other parts). If I am not
convinced, then I will likely recommend rejection (unless I see
something standout in other parts).
\end{tcolorbox}

\subsection{Your School}\label{sec:your-school}

Graduating from top universities \emph{that we recognize} helps (but still not as much as having research experience).
However, committee members might not know much about schools in your country and just treat your school as
\emph{``unknown foreign''}, which can be a minus point (if your school is well-known, then it is \emph{``top foreign''}, which is definitely a plus).

So what can you do about this? several things such as ask your CS dept to put itself on CSRankings (it's the easiest way to get CS people to know about the school), explain about your school in your statement (and ask your LOR writer to do that), and of course, if you're Vietnamese, consider a CS PhD program that has \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{Vietnamese professors}.

\begin{tcolorbox}[left=1pt,right=1pt,top=1pt,bottom=1pt]
Vu: Sometime PhD admission committee will share a document such as \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Foreign-Top-Schools}{this one}, which lists the top schools in several countries. I have looked at Vietnamese applications (whether they are assigned to me or not) and provide inputs to the reviews of those applications, e.g., X is the top tech school in Vietnam and so it should be /top/ instead of /unknown/ foreign, which makes a huge difference.
\end{tcolorbox}

\subsection{Grades/GREs}\label{sec:grades}
Having good grades is important, but, in general, unless your school is well-known, having top grades/ranks
usually will not help. This is simply because we cannot evaluate them.

This can be an issue for students in many international top universities 
where the competition is so high that very good students students can still have low ranking (and be overlooked by Admission committee).
So what to do with this? well, same as \hyperref[sec:your-school]{before}, e.g., put a note about this in your statement.

Note that while having good grades at unknown school might not help,
having very bad grades will be \red{red flag} (unless your LORs or
statements give proper explanation). This is especially true if you
have bad grades in relevant courses.

\paragraph{GRE} Most CS programs in the US \emph{no longer require GREs}, so you don't need to
take them. However, they might be useful for international students from programs we are not familiar with.

You also do need to
take standarized English test. Just do well enough to pass minimum
requirement set by the university.

\begin{tcolorbox}[left=1pt,right=1pt,top=1pt,bottom=1pt]
Vu: The minimum for GMU CS (being above this might not mean much, but below is a \red{red flag}).
\begin{itemize}
\item GPA: $\ge 3.0$ in your undergrad
  \begin{itemize}
  \item but also consider the rank/prestige of your schools
  \end{itemize}
\item GRE: not required, most CS graduate programs no longer require GRE
%    \item but if you want to use it, then we expect a total (V+Q) of $\ge 311$ (with a $\ge 157$ Q) and A $\ge 3.0-3.5$.
\item English requirement tests (one of the below)
  \begin{itemize}    
  \item TOEF: 88 pts in total AND $\ge 20$ points in each subsection OR
  \item IELTS: $\ge 6.5$ OR
  \item DuoLingo Graduate English: $\ge 120$ OR 
  \item Pearson Test of Academic English: $\ge 67$
  \end{itemize}  
\end{itemize}
\end{tcolorbox}

\subsection{CV/Resume}
This should be a summary of the accomplishments of the applicant.  I would quickly scan this document to see if something stand out (e.g., Publications, Programming Competition Awards, Teaching Experiences).

\subsection{Interview}

Sometime a faculty wants to interview an applicant to make a decision. Typically this means they lean toward admitting you (if we don't like your application, we will not bother doing the interview).

An interview lasts about 20--30 mins, and one implicit thing you will be evaluated on is whether you can communicate effectively (i.e., speak/understand English).  You will also get chance to ask questions about the university so think of something to ask (just the same as you interview at a company).

\begin{tcolorbox}[left=1pt,right=1pt,top=1pt,bottom=1pt]
At GMU, we are encouraged to interview candidates. In most cases we interview very strong candidates, i.e., the interview is actually to recruit them.  In some rare cases a faculty interviews a candidate that they see potentials and want argue for admission, i.e., without the interview, that candidate is definitely rejected. In any case, getting interview means you have a very good chance of getting admitted.
\end{tcolorbox}

\section{Getting Admitted and Choosing the Right School}

Around March you should hear back from most PhD programs that you applied (if not, send email and ask). You will have to make your decision by around April 15.
If you have offers, congratulations!  Now you're at a different game because the schools that admit you will now try to get you to accept them!

Most schools will have an \textbf{Open House}, which is a great resource to learn about the dept, faculty, area, etc. During the Open House, you get a chance to talk directly to multiple faculty and current students.  Take notes of faculty who make you excited, count those that are taking in new students (if they meet you, likely they are considering new students!).  Talk to students about their advisors, the dept, the area, funding situation etc.  Ask about anything you want to determine that they deserve \emph{you}.

In short, if you can come to the Open House, do come.  But if you're international student outside of the US, then likely you cannot come.  So see if you can attend it virtually and ask to meet with individual faculty if you can.

\begin{tcolorbox}[left=1pt,right=1pt,top=1pt,bottom=1pt]
GMU has Virtual Open House, e.g., \url{https://cs-gmu.github.io/cs-phd-voh-s23/}, which I've co-organized in the last two years. We invite all admitted PhD students to the VOH to learn about the CS program, the department, GMU, and the DC area in general. Students also get opportunies to chat with professors and current students.  Everything is through Zoom.
\end{tcolorbox}

\section{Funding}\label{sec:funding}

As mentioned, if you're admitted to a \emph{good} CS PhD program, you should not have to worry about funding!  
In the US, the common types of funding for PhD (in any fields) are \emph{graduate teaching assistant} (GTA or TA), \emph{graduate research assistant} (GRA or RA), and \emph{Fellowship}.
RA is the type of funding paid by a prof. for you to do their research. TA is paid by the department for you to help with teaching. Finally, fellowship is an independent funding that can come the school, a company, or organization. Tab.~\ref{tab:funding} summarizes the differences.

\begin{table}
  \centering
  \caption{Different types of fundings}\label{tab:funding}
  \begin{tabular}{c|c|c|c}
    \toprule
    &\textbf{TA}&\textbf{RA}&\textbf{Fellowship}\\
    \midrule
    \textbf{From} & School & Profs. & School/External\\
    \textbf{For}                  & Teaching Assistant       & Research                        & Research                              \\
    \textbf{Tuition/Ins./Stipend} & Yes                      & Yes                             & Yes                                   \\
    \textbf{Cover Summer?}              & No                       & Maybe                           & Yes                                   \\
    \midrule
    \textbf{Pros}                 & Research Freedom         & Get to do research              & Research Freedom                      \\
    \textbf{Cons}                 & TA, Uncertain            & Research restriction, Uncertain & Competitive, limited             \\
    \bottomrule
  \end{tabular}
\end{table}

\subsection{Graduating Assistantship (TA/RA)}
The most common type of funding is \textbf{graduate assistanship}, which is either TA or RA. Both TA and RA come with tuition waiving (you don't have to pay tuition), health insurance (this takes care of your insurance, which is a must have in the US), and most importantly, your stipend (i.e., your salary). Some universities also pay insurance for spouse/chilren (or give very good discount)

A couple of notes about stipend. First, the amount of stipend depends on the university, which in turns depend on various factors such as location (e.g., a stipend in Washington DC is likely higher than in Lincoln, Nebraska due to higher living cost). Second, a school year is (typically) 9-month in the US, so stipend is for 9 months (so divide by 9 for each month). Third, like for most source of income in the US, you will have to pay tax on your stipend. Finally, private universities might pay more for stipend.

\begin{tcolorbox}[left=1pt,right=1pt,top=1pt,bottom=1pt]
At GMU, TA and RA have similiar benefits in tuition waiving and insurance.  For stipend, depending on the college and department, a 9-month graduate assistant stipend is set.  TA and RA will usually be that ammount (TA will definitely be that, RA might flunctuate a bit depending on the stage of the student and the prof.). 
\end{tcolorbox}

\subsubsection{Teaching Assistant (TA)}

TA is common in the beginning when you haven't found your advisor who would pay you RA. As a TA, you help professors with their classes (e.g., grading or teaching labs/recitation). Your TAship is paid through the department, i.e., they hire you to help teach.  During a semester, a TA mgith work with several courses and professors (not necessary their advisor).  TA funding typically is not available during the summer, which has no school.

\paragraph{How to get TA?}  Unless you have other fundings such as RA or Fellowships, TA is typically a default thing. When you apply to be a full-time student, you also sate that you need financial assistant. It is common that the PhD committee will either admit you and give you GTA, or reject you; i.e., we do not admit a student without supporting them.

\begin{tcolorbox}[left=1pt,right=1pt,top=1pt,bottom=1pt]
At GMU CS, students admitted with TA have  4 years of GTA guaranteed and also receive  stipend for the \textbf{first} summer.
\end{tcolorbox}

Even if you have other fundings and do not need TA, you still should do TA at least once.  This allows you to see what teaching is like, which is especially helpful for research career where you often have to give talks and tell people about your work. Note that GMU sometimes has classes that a more seninor student can teach.  In that case, you will be paid as a lecturer, which is higher than GTA.  This is a good opportunity for students to get teaching experience and also get paid more.
\subsection{Research Assistant (RA)}
RA is provided through a professor through their own funding so you can work on their project.  
You do not need to teach as an RA, so you can focus on your research. Depending on the professor, RA may be available during the summer.

\textbf{How to get RA?} If a professor recruits you, they will likely give you RA right away (e.g., when you apply).  A common scenario is that you get admitted with TA, and then after a year or two find an advisor to support you with RA.

\begin{tcolorbox}[left=1pt,right=1pt,top=1pt,bottom=1pt]
Vu: If you're lucky and got recruited by a prof. who would give you RA right away, it's very likely you will get admitted.  For example, if a prof., even if not in PhD admissin committee, wants to work with and funds you, the PhD admission committee will respect that decision and admit your application (unless your application has many red flags).
\end{tcolorbox}

\subsection{Fellowship/Scholarship}

Fellowship is another type of funding in which the student applies for (e.g., from school, industries, government). Fellowships are typically competitive and generous, and gives pretty much all benefits tuition/insurance that a graduate assistantship has.  Moreover, it often gives higher stipend (including summer) and opens doors for job opportunities (e.g., internship).  For example, a student with a Microsoft fellowship will likely get an internship at Microsoft.

In general, fellowship is prestigious, and you will stand out if you get one.  Every PhD student has pubs, but only superstars have NSF grad or Microsoft fellowship. In fact, these are so prestigious that even if you didn't get it but make it to the final round, school will still mention you on their website and you still should put it on your CV.

\textbf{How to get Fellowship?} You apply for them.  The US government has many fellowships but these likely require US citizenship or residency.  However, tech companies including Google, Microsoft, Facebook, IBM all have fellowships that international students can apply for. Prestigious fellowships typically require a clear and good research plan, so it is usueful to have research experience and even publication for a better chance (remember, you're competiting with the top Ph.D. students at top universities worldwide).

\begin{tcolorbox}[left=1pt,right=1pt,top=1pt,bottom=1pt]
At GMU, Ph.D. applicants are automatically eligible for a Presidential Fellowship.  It is at least as good as GTA but the most important thing is that as a fellowship it is trully free money (i.e., you are not depending on any prof. or TA duties).  PhD admission committee members nominate applicants for this fellowship and the committee will vote and give the fellowship to the top 2.
\end{tcolorbox}

\section{Miscs and FAQs}

\begin{itemize}
    \item   Is an MS degree required for admission to PhD?
    \begin{description}[before=\small]
    \item No. In fact, student with BS can get MS degree "along the way" to PhD.  However, MS can help if it gives research experience or is from a more well-known school than your undergrad institution.
    \end{description}

    \item Can I apply to PhD to CS if my undergrad was in Math, Biology, Finance, etc?
    \begin{description}[before=\small]
        \item Yes, as long as you can demonstrate you are ready for CS PhD research through research experiences, LoRs, statements, etc as mentioned.
    \end{description}

    \item What can you do to increase your admission chance?
    \begin{description}[before=\small]
        \item Show something that makes you \textbf{stand out}, e.g., are you a female or a minority in CS (research for this on Google)? Do you participate in outreach activities that help increase diversity and inclusion in CS?  All of these are important in CS in the US.

        Also, even if you do not have formal research experience, you can talk about your personal project. If it is used by many people, have lots of stars in Github, etc, it would certainly worth talking about.  If you write technical, research-like blogs, that would also help.
    \end{description}

    \item How do I address a professor Firstname Lastname?
    \begin{description}[before=\small]
        \item If you don't know that professor (e.g., first email contact), then use \textbf{Prof. Lastname} or \textbf{Dr. Lastname}. I've seen many international students write \textbf{Prof.} or \textbf{Dr.} \textbf{Firstname Lastname}.  Writing like that makes it like you copy and paste the names, so no need to do so,  just Prof. or Dr. Lastname.

        Also do not use Mr. or Mrs., or just write Firstname. May be it is OK with others but I find it a bit disrespectful. As you know that prof. better and depends on their preference, you can call them by their Firstname.

\begin{tcolorbox}[left=1pt,right=1pt,top=1pt,bottom=1pt]
    Vu: I've been called Dr. Vu and I find it a bit amusing but am totally fine with it.
    \end{tcolorbox}
    \end{description}
\end{itemize}

% **

\bibliographystyle{abbrv}
\bibliography{phd-cs-us.bib}
\newpage

\end{document}

# - Asking professors if they can get in or if they take students.\\
#   > Many international students will write professors asking if they can
#   get in, or if they are taking students. Unlikely anyone will reply to
#   such email because the short answer is we simply cannot evaluate your
#   profile ourselves, it needs to go to the whole committee as explained.
#   Even if a professor is seeking for students (almost everyone is
#   looking for good students), they simply cannot take one in without
#   going through the application process.

# - See this link for a large list of [[https://github.com/dynaroars/dynaroars.github.io/wiki/Funding][funding sources for PhD students]]

# ** Self-funded
# - Some PhD students are funded through their company
#   - The students work full time for the company and do their PhD part-time
#   - Typically will take longer (because you have to work full time) and require some serious committement from the students

# - Some PhD students are funded through country, e.g., GMU has several students from Saudi Arabia that are funded through their government.

# ** Others  
# - There are also various kind of one-time thing and small awards from industry or organization, e.g., a $2000 award for the summer.
Unlike the B.A., where you or your parents pay many tens of thousands of dollars,
or the M.S., where you typically work as a teaching assistant and possibly continue
to pay many tens of thousands of dollars, the Ph.D. is a time where funding is not a
concern to you. At most schools, you will not pay tuition during the time that you
are getting a Ph.D.. Typically, you will also receive a living stipend – on the order
of $2000 per month, from which you will pay your living expenses. Ideally, your
only responsibility will be research. This is called doing an RAship (Research
Assistantship).
The Ph.D. is a tremendous opportunity. You get to pick an advisor in any re-
search area you like and then you get to do research in that area, receive mentoring,
think deeply on problems, publish papers, become famous, while paying zero tu-
ition for 6 years and receiving a salary. Your advisor is paying for this opportunity
by writing grant proposals to companies and to the government to ask for fund-
ing. A single graduate student can cost an advisor upwards of 80K per year (given
the cost of tuition, the stipend, the overhead tax charged by the school, cost of
equipment and physical space, etc.).
Important note 1: At most schools, you can only do an RAship if you have
an advisor who has funding for you. Since some advisors don’t apply for grants
or are in areas which aren’t well-funded, you may have to work as a teaching
assistant every semester to get your stipend. This is called a TAship (Teaching
Assistantship). When I was a graduate student, I had a few friends who were
forced to TA 13 semesters, to fund their way through school! Alternatively, you
will have to restrict your choice of advisors to those who have funding. At CMU,
every Ph.D. student is guaranteed a stipend plus tuition regardless of which advisor
she chooses to work with.
Important note 2: There are many companies and government organizations
which offer Graduate Fellowships for Ph.D. students. If you are lucky enough
to get one of these, they will cover most of your way through graduate school,
and you will never have to worry about whether your advisor has funding or not.
Details about graduate fellowships will be discussed in Section 4.

** Current 2023 Rankings of CS PhD programs in the U.S.
  Below is the 2023 Rankings from [[https://www.csrankings.org][CSRankings.org]] as of <2023-04-09 Sun>.

1. Carnegie Mellon University 
2. Univ. of Illinois at Urbana-Champaign 
3. Univ. of California - San Diego  
4. Massachusetts Institute of Technology 
5. Stanford University 
5. University of Michigan  
7. University of Washington 
8. Cornell University  
8. Univ. of California - Berkeley  
10. Georgia Institute of Technology 
11. Northeastern University  
11. University of Maryland - College Park 
13. University of Wisconsin - Madison  
14. Purdue University  closed chart
15. University of Texas at Austin  
16. Columbia University  
16. University of Pennsylvania 
18. New York University  
18. Princeton University 
20. University of Massachusetts Amherst  
21. Univ. of California - Los Angeles
22. University of Chicago 
23. Stony Brook University
23. Univ. of California - Santa Barbara 
25. Rutgers University  
26. Univ. of California - Irvine 
27. University of Southern California 
28. Duke University 
29. Univ. of California - Riverside
30. Pennsylvania State University 
31. Northwestern University 
32. Ohio State University 
33. George Mason University 
33. Harvard University  
35. Brown University  
35. University of Utah 
37. Texas A&M University
37. Univ. of California - Santa Cruz  
37. Yale University  
40. Boston University 
40. University at Buffalo
40. University of Colorado Boulder 
43. North Carolina State University
44. Rice University 
44. University of Illinois at Chicago  
46. University of North Carolina 
46. University of Virginia 
48. Arizona State University 
48. University of Minnesota 
48. Virginia Tech  
51. Oregon State University  
51. Univ. of California - Davis

* Links
:PROPERTIES:
:CUSTOM_ID: links
:END:
- https://idleprocess.wordpress.com/2009/12/07/why-go-to-graduate-school-and-how-to-get-into-the-program-of-your-dreams/
- https://mycsphd.org/_pages/application-parts.html
- https://emeryberger.com/admission-notes/
- https://www.elsevier.com/connect/9-things-you-should-consider-before-embarking-on-a-phd
- https://grad.uchicago.edu/wp-content/uploads/2019/08/Applying-to-PhD-Programs-Guide-2019.pdf
- https://www.pathwaystoscience.org/pdf/CIC_GradSchoolGuide.pdf
- https://www.science.org/content/article/applying-phd-these-10-tips-can-help-you-succeed
- https://chrisblattman.com/blog/2022/03/25/faqs-on-phd-applications/
- https://www.cs.cmu.edu/~harchol/gradschooltalk.pdf

# * Types of Funding for CS PhDs

# * How does GRA work?
# While you might get a 9-month $28K GRA stipend,  your advisor typically has to budget at least twice that much.  Because they have to pay your tuitition, insurance, and overhead that the university takes.  The chart below shows that each CS GRA costs about $70K at GMU.

# - Example budget of a PhD student in CS at GMU.  Using stipend and tuition rate in 2023. Conference Registration and Travel are estimates.
# - It is also **very likely** that GMU faculty won't have to pay student tuition starting Fall 2025, when this happens then we can take out  Tuition and total cost is 46.5K.

# | Budget                                 | Cost    |                                                                                           |
# |----------------------------------------+---------+-------------------------------------------------------------------------------------------|
# | GRA (9-month)                          | 28K     |                                                                                           |
# | GRA (summer)                           | 9K      | 3-month, 20hrs/week                                                                       |
# | Fringe Benefits (7.3%)                 | 0       | for faculty mostly (non-wage expenses such as social security, medicare)                  |
# | **Total Salary**                       | 37K     |                                                                                           |
# | Health Insurance                       | 3K      | Full year                                                                                 |
# | Tuition (In-State) for Fall and Spring | 15K     | ($680/ Credit + $150/Student Fee/ Credit)* 9 credits = $7470 ($6120 + $1350) per semester |
# | **Total**                              | 18K     | Full year tuition + insurance                                                             |
# | Total Materials & Supplies             | 0       |                                                                                           |
# | Conference Registration                | 500     |                                                                                           |
# | International Travel                   | 1.8K    |                                                                                           |
# | Domestic Travel                        | 700     |                                                                                           |
# | **Total Travel**                       | 3K      |                                                                                           |
# | Total **Direct** Cost                  | 58K     | Salary 37K + Travel 3K  + Health 3K + Tuition 18K                                         |
# | F & A (MTDC)                           | 21K     | Direct Cost - GRA Salary                                                                  |
# | Total **Indirect** Cost                | 12K     | **58.9%** of MTDC                                                                         |
# | **Total** (Direct + Indirect)          | **70K** |                                                                                           |

# - F & A = Facilities & Administrative Cost Base 
# - MTDC = Modified Total Direct Cost

# * Miscs

# - Asking professors if they can get in or if they take students.\\
#   > Many international students will write professors asking if they can
#   get in, or if they are taking students. Unlikely anyone will reply to
#   such email because the short answer is we simply cannot evaluate your
#   profile ourselves, it needs to go to the whole committee as explained.
#   Even if a professor is seeking for students (almost everyone is
#   looking for good students), they simply cannot take one in without
#   going through the application process.

# - How long for a CS PhD in US ? > 5-6 years on avg.

# - Does having an MS help? > It can help if the MS gives research
#   experience or is from a more well-known school than your
#   undergraduate. But not having an MS does not hurt either. Nowadays it
#   is normal for students with bachelor degrees apply directly to
#   PhD in CS.

### 2023-06-06 phd-cs-us.tex
Updates from Overleaf

@@
 
-\newcommand{\tool}{\textsf{NeuralSAT}}
 \def\Section{\S}
@@
 \date{}
-\author{ThanhVu (Vu) Nguyen with inputs from faculty at GMU and other universities}
+\author{ThanhVu (Vu) Nguyen, George Mason University}
 %\setcounter{secnumdepth}{3}
@@
 Having been involved in PhD admissions for many years and after
-numerious interaction international (especially Vietnamese) students, I've
-realized that international students lack a clear understanding of
+numerous interaction with  students (especially Vietnamese), I've
+realized that \emph{international} students lack a clear understanding of
 the Computer Science PhD admission process at US universities. This confusion not only
@@
 While I wrote this document for CS applicants, it should be applicable to various STEM fields.
-Moreover, while many examples given are for GMU, which is ranked 33rd on CSRankings (June 2016), the writing should be generalize to most other R1 universities  (though \emph{very} top schools migh be very extreme, e.g., see the admission process at \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{CMU}.
+Moreover, while many examples given are for GMU, which is 33rd on CSRankings (June 2016), the writing should be generalize to most other R1\footnote{An \href{https://en.wikipedia.org/wiki/List_of_research_universities_in_the_United_States}{R1 institution} in the US refers to a research-intensive university with a high level of research activity across various disciplines.} universities  (though \emph{very} top schools might be very extreme, e.g., see the admission process at \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{CMU}).
 
@@
 
-If you have suggestions or comments, feel free to create a \href{https://github.com/dynaroars/dynaroars.github.io/issues}{GitHub issue} for discussion.
+If you have suggestions or comments, feel free to create a \href{https://github.com/nguyenthanhvuh/phd-cs-us/issues}{GitHub issue} for discussion.
 \end{abstract}
@@
 
-\section{How is your application evaluated?}
+\section{How is Your Application Evaluated?}
 
 After you submit your PhD application (usually in December), it will be first screened
-for general requirements, e.g., did you submit your transcripts and standard scores? did your writters submit their letters?
+for general requirements, e.g., did you submit your transcripts and standard scores? did your reference writers submit their letters?
 Then your application will be reviewed by a
 \textbf{PhD admission committee} consisting of faculty members in CS.
-Each application is reviewed by about \emph{three} faculty members, who will evaluate your profile and try to reach a consensus about your case.
+Each application is assigned to about \emph{three} faculty members, who will evaluate your profile and try to reach a consensus about your case.
+Note that while in most cases the assigned reviewers will be the main ones deciding your application, every faculty will have access to your application and can give inputs on your profile.
 
-Many PhD committee members are \emph{new assistant professors}.
-This helps junior faculty, who typically don't have students yet, the opportunities to recruit students.
-The chair of the committee will be a senior professor, but they will not review individual applicaitions and instead assign them to committee members.
-The chair will look at various factors such as keywords/interests, research areas, or mentioning faculty names, and assign the application to appropriate faculty.
+In many cases, the admission committee will involve junior faculty (e.g., assistant professors) in the department.
+While this requires  a significant time investment, it provides junior faculty the opportunities to recruit students.
+The chair of the committee will be a senior professor, but they will not review individual applications and instead assign them to committee members.
+The chair will look at various factors such as research interests or mentioning faculty names, and assign the applications to appropriate faculty.
 
@@
 In my opinion, requiring applicants to pay the fee helps ensure their seriousness, as it filters out non-serious candidates.
-If the application process were free for everyone, we would receive an overwhelming number of applications to review.
+Also, if the application process were free for everyone, we would receive an overwhelming number of applications to review.
 
@@
 is a research degree. To assess your research capability, we consider
-the following key indicators.
+the following key indicators, listed in order of importance.
 
@@
 \begin{tcolorbox}[left=1pt,right=1pt,top=1pt,bottom=1pt]
-Vu: Vietnamese students often mention Scopus Q1, which consists of diversed journals from IEEE, Elsevier, and several other publishers unfamiliar to me.
+Vu: Vietnamese students often mention Scopus Q1, which consists of diverse journals from IEEE, Elsevier, and several other publishers unfamiliar to me.
 Honestly, I don't know/recognize the majority of journals listed in Scopus Q1.
-So this might be something to be mindful of, as CS faculty in the US might not be too familiar with Scopus or journals listed in there, so devote sometime in your statement to discuss their significance.
+So this might be something to be mindful of, as \textbf{CS} faculty might not be too familiar with Scopus or journals listed in there, so devote sometime in your statement to discuss the significance of your papers.
 \end{tcolorbox}
@@
 \begin{tcolorbox}[left=1pt,right=1pt,top=1pt,bottom=1pt]
-Craig: GMU and many other universities allow you to upload your published papers.
+Craig: GMU and many other universities allow you to upload your published papers and other writing samples.
 In many cases, even if the papers were not published at top places, we can still determine their quality by simply skimming over the paper.
@@
 applicants: \textbf{customize the statement} for the school you're applying to,
-e.g., why do you apply there ? name several professors who you're
-interested in working with (in some cases your application will be sent to them for evaluation).
+e.g., why do you apply here? talk about a couple of professors who you're interested in working with (in many cases your application will be forwarded to them for evaluation).
 Be careful not to send wrong statement to wrong school or mixing
@@
 
-Graduating from top universities \emph{that we recognize} helps (but still not as much as having research experience).
-However, committee members might not know much about schools in your country and just treat your school as
+Graduating from top universities \emph{that we recognize} helps.
+However, if committee members do not know much about schools in your country, they will likely treat your school as
 \emph{``unknown foreign''}, which can be a minus point (if your school is well-known, then it is \emph{``top foreign''}, which is definitely a plus).
@@
 So what can you do about this?
-several things such as ask your CS dept to put itself on CSRankings (it's the easiest way to get CS people to know about the school), explain about your school in your statement (and ask your LOR writer to do that), and of course, if you're Vietnamese, consider a CS PhD program that has \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{Vietnamese professors}.
+several things such as asking your CS dept to put itself on CSRankings (it's the easiest way to get CS people to know about the school), explaining about your school in your statement (and asking your LOR writer to do that too), and of course, if you're Vietnamese, considering a CS PhD program that has \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{Vietnamese professors}.
 
@@
 Vu: Sometime PhD admission committee will share a document such as \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Foreign-Top-Schools}{this one}, which lists the top schools in several countries.
-I have looked at Vietnamese applications (whether they are assigned to me or not) and provide inputs to the reviews of those applications, e.g., X is the top tech school in Vietnam and so it should be /top/ instead of /unknown/ foreign, which makes a huge difference.
+I have looked at Vietnamese applications (whether they are assigned to me or not) and provide inputs to the reviewers of those applications, e.g., X is the top tech school in Vietnam and so it should be \emph{top} instead of \emph{unknown} foreign, which makes a huge difference.
 \end{tcolorbox}
@@
 
-This can be an issue for students in many international top universities 
-where the competition is so high that very good students students can still have low ranking (and be overlooked by Admission committee).
+This can be an issue for students in many top international universities where the competition is so high that very good students students can still have low ranking (and be overlooked by Admission committee).
 So what to do with this? well, same as \hyperref[sec:your-school]{before}, e.g., put a note about this in your statement.
@@
 You also do need to
-take standarized English test. Just do well enough to pass minimum
+take standardized English test. Just do well enough to pass minimum
 requirement set by the university.
@@
 \begin{itemize}
-\item GPA: $\ge 3.0$ in your undergrad
-  \begin{itemize}
-  \item but also consider the rank/prestige of your schools
-  \end{itemize}
-\item GRE: not required, most CS graduate programs no longer require GRE
+\item GPA: $\ge 3.0$ in your undergrad (but as mentioned, we also consider the rank/prestige of your school)
+\item GRE: not required (most CS graduate programs no longer require GRE, though it can help boost your profile)
 %    \item but if you want to use it, then we expect a total (V+Q) of $\ge 311$ (with a $\ge 157$ Q) and A $\ge 3.0-3.5$.
@@
 
-An interview lasts about 20--30 mins, and one implicit thing you will be evaluated on is whether you can communicate effectively (i.e., speak/understand English).
+An interview lasts about 15--30 mins, and one implicit thing you will be evaluated on is whether you can communicate effectively (i.e., speak/understand English).
 You will also get chance to ask questions about the university so think of something to ask (just the same as you interview at a company).
@@
 At GMU, we are encouraged to interview candidates.
-In most cases we interview very strong candidates, i.e., the interview is actually to recruit them.
+For very strong candidates, the interview is actually to recruit them.
 In some rare cases a faculty interviews a candidate that they see potentials and want argue for admission, i.e., without the interview, that candidate is definitely rejected.
-In any case, getting interview means you have a very good chance of getting admitted.
+In any case, getting interview means you have a very good chance of being admitted.
 \end{tcolorbox}
@@
 
-Most schools will have an \textbf{Open House}, which is a great resource to learn about the dept, faculty, area, etc.
-During the Open House, you get a chance to talk directly to multiple faculty and current students.
+Most schools will have an \textbf{Open House}, which is a great resource to learn about the school, department, faculty, research, living, etc.
+During the Open House, you get a chance to talk to individual faculty and current students.
 Take notes of faculty who make you excited, count those that are taking in new students (if they meet you, likely they are considering new students!).
@@
 \begin{tcolorbox}[left=1pt,right=1pt,top=1pt,bottom=1pt]
-GMU has Virtual Open House, e.g., \url{https://cs-gmu.github.io/cs-phd-voh-s23/}, which I've co-organized in the last two years.
-We invite all admitted PhD students to the VOH to learn about the CS program, the department, GMU, and the DC area in general.
-Students also get opportunies to chat with professors and current students.
-Everything is through Zoom.
+Vu: GMU has Virtual Open House, e.g., \url{https://cs-gmu.github.io/cs-phd-voh-s23/}, which I've co-organized in the last two years.
+We invite all admitted PhD students to the VOH through Zoom to learn about the CS program, the department, GMU, and the DC area in general.
+Students also get opportunities to chat with professors and current students.
 \end{tcolorbox}
@@
 As mentioned, if you're admitted to a \emph{good} CS PhD program, you should not have to worry about funding!  
-In the US, the common types of funding for PhD (in any fields) are \emph{graduate teaching assistant} (GTA or TA), \emph{graduate research assistant} (GRA or RA), and \emph{Fellowship}.
+In the US, the common types of funding for PhD are \emph{graduate teaching assistant} (GTA or TA), \emph{graduate research assistant} (GRA or RA), and \emph{Fellowship}.
 RA is the type of funding paid by a prof.
@@
   \centering
-  \caption{Different types of fundings}\label{tab:funding}
+  \caption{Different types of PhD funding}\label{tab:funding}
   \begin{tabular}{c|c|c|c}
@@
 Both TA and RA come with tuition waiving (you don't have to pay tuition), health insurance (this takes care of your insurance, which is a must have in the US), and most importantly, your stipend (i.e., your salary).
-Some universities also pay insurance for spouse/chilren (or give very good discount)
+Some universities also pay insurance for spouse/children (or give very good discount).
 
-A couple of notes about stipend.
+Several about stipend.
 First, the amount of stipend depends on the university, which in turns depend on various factors such as location (e.g., a stipend in Washington DC is likely higher than in Lincoln, Nebraska due to higher living cost).
@@
 \begin{tcolorbox}[left=1pt,right=1pt,top=1pt,bottom=1pt]
-At GMU, TA and RA have similiar benefits in tuition waiving and insurance.
+At GMU, TA and RA have similar benefits in tuition waiving and insurance.
 For stipend, depending on the college and department, a 9-month graduate assistant stipend is set.
-TA and RA will usually be that ammount (TA will definitely be that, RA might flunctuate a bit depending on the stage of the student and the prof.).
+TA and RA will usually be that amount (TA will definitely be that, RA might fluctuate a bit depending on the stage of the student and the prof.).
 \end{tcolorbox}
@@
 Your TAship is paid through the department, i.e., they hire you to help teach.
-During a semester, a TA mgith work with several courses and professors (not necessary their advisor).
+During a semester, a TA might work with several courses and professors (not necessary their advisor).
 TA funding typically is not available during the summer, which has no school.
 
-\paragraph{How to get TA?}  Unless you have other fundings such as RA or Fellowships, TA is typically a default thing.
+\paragraph{How to get TA?}  Unless you have other funding such as RA or Fellowships, TA is typically a default thing.
 When you apply to be a full-time student, you also sate that you need financial assistant.
@@
 
-Even if you have other fundings and do not need TA, you still should do TA at least once.
+Even if you have other funding and do not need TA, you still should do TA at least once.
 This allows you to see what teaching is like, which is especially helpful for research career where you often have to give talks and tell people about your work.
-Note that GMU sometimes has classes that a more seninor student can teach.
+Note that GMU sometimes has classes that a more senior student can teach.
 In that case, you will be paid as a lecturer, which is higher than GTA.
@@
 
-\textbf{How to get RA?} If a professor recruits you, they will likely give you RA right away (e.g., when you apply).
-A common scenario is that you get admitted with TA, and then after a year or two find an advisor to support you with RA.
+\textbf{How to get RA?} When a professor recruits you, they will likely give you RA right away (e.g., when you apply).
+A common scenario is that you first get admitted with TA, and then after a year or two find an advisor to support you with RA.
 
@@
 who would give you RA right away, it's very likely you will get admitted.
-For example, if a prof., even if not in PhD admissin committee, wants to work with and funds you, the PhD admission committee will respect that decision and admit your application (unless your application has many red flags).
+For example, if a prof., even if not in PhD admission committee, wants to work with and funds you, the PhD admission committee will respect that decision and admit your application (unless your application has many red flags).
 \end{tcolorbox}
@@
 Fellowship is another type of funding in which the student applies for (e.g., from school, industries, government).
-Fellowships are typically competitive and generous, and gives pretty much all benefits tuition/insurance that a graduate assistantship has.
+Fellowships are typically competitive and generous, and gives pretty much all benefits tuition/insurance that a TA/RA has.
 Moreover, it often gives higher stipend (including summer) and opens doors for job opportunities (e.g., internship).
@@
 \textbf{How to get Fellowship?} You apply for them.
-The US government has many fellowships but these likely require US citizenship or residency.
-However, tech companies including Google, Microsoft, Facebook, IBM all have fellowships that international students can apply for.
-Prestigious fellowships typically require a clear and good research plan, so it is usueful to have research experience and even publication for a better chance (remember, you're competiting with the top Ph.D.
-students at top universities worldwide).
+The US government has many fellowships but these would likely require US citizenship or residency.
+However, tech companies including Google, Microsoft, Facebook, IBM have fellowships that international students can apply for.
+
+Prestigious fellowships typically require a clear and good research plan, so it is a good idea to wait until at least your second year to have research experience and even publication before applying.
+Remember, you're competiting with the top Ph.D.
+students at top universities worldwide.
 
@@
 applicants are automatically eligible for a Presidential Fellowship.
-It is at least as good as GTA but the most important thing is that as a fellowship it is trully free money (i.e., you are not depending on any prof.
+It is at least as good as GTA but the most important thing is that as a fellowship it is truly free money (i.e., you are not depending on any prof.
 or TA duties).
@@
 
-\begin{itemize}
+\begin{enumerate}
     \item   Is an MS degree required for admission to PhD?
-    \begin{description}[before=\small]
+    \begin{description}
     \item No.
@@
     \item Can I apply to PhD to CS if my undergrad was in Math, Biology, Finance, etc?
-    \begin{description}[before=\small]
+    \begin{description}
         \item Yes, as long as you can demonstrate you are ready for CS PhD research through research experiences, LoRs, statements, etc as mentioned.
@@
     \item What can you do to increase your admission chance?
-    \begin{description}[before=\small]
+    \begin{description}
         \item Show something that makes you \textbf{stand out}, e.g., are you a female or a minority in CS (research for this on Google)?
@@
     \item How do I address a professor Firstname Lastname?
-    \begin{description}[before=\small]
+    \begin{description}
         \item If you don't know that professor (e.g., first email contact), then use \textbf{Prof.
@@
     \end{description}
-\end{itemize}
+\end{enumerate}
 
-% **
+\section{Acknowledgement}
+
+The following people contributed to this document: Craig Yu (GMU), Hakan Aydin (GMU).
 

### 2023-06-06 phd-cs-us.tex
updates

@@
 
+\subsection{Current 2023 Rankings of CS PhD programs in the U.S}
+  Below is the 2023 Rankings from \href{https://www.csrankings.org}{CSRankings.org}, a ranking system that is based on top CS conferences, as of April 2023.
+
+\begin{itemize}
+\item Carnegie Mellon University 
+\item Univ. of Illinois at Urbana-Champaign 
+\item Univ. of California - San Diego  
+\item Massachusetts Institute of Technology 
+\item Stanford University         
+\item University of Michigan  
+\item University of Washington 
+\item Cornell University  
+\item Univ. of California - Berkeley  
+\item Georgia Institute of Technology 
+\item Northeastern University  
+\item University of Maryland - College Park 
+\item University of Wisconsin - Madison  
+\item Purdue University  closed chart
+\item University of Texas at Austin  
+\item Columbia University  
+\item University of Pennsylvania 
+\item New York University  
+\item Princeton University 
+\item University of Massachusetts Amherst  
+\item Univ. of California - Los Angeles
+\item University of Chicago 
+\item Stony Brook University
+\item Univ. of California - Santa Barbara 
+\item Rutgers University  
+\item Univ. of California - Irvine 
+\item University of Southern California 
+\item Duke University 
+\item Univ. of California - Riverside
+\item Pennsylvania State University 
+\item Northwestern University 
+\item Ohio State University 
+\item George Mason University 
+\item Harvard University  
+\item Brown University  
+\item University of Utah 
+\item Texas A\&M University
+\item Univ. of California - Santa Cruz  
+\item Yale University  
+\item Boston University 
+\item University at Buffalo
+\item University of Colorado Boulder 
+\item North Carolina State University
+\item Rice University 
+\item University of Illinois at Chicago  
+\item University of North Carolina 
+\item University of Virginia 
+\item Arizona State University 
+\item University of Minnesota 
+\item Virginia Tech  
+\item  Oregon State University  
+\item Univ. of California - Davis
+\end{itemize}
+
 \section{Acknowledgement}
-
-The following people contributed to this document: Craig Yu (GMU), Hakan Aydin (GMU).
-
+Many faculty and students have contributed to this document.
+Craig Yu (GMU) and Hakan Aydin (GMU) contributed to the early version of this document. Other GMU faculty also have provided valuable feedbacks.  Thanks to all of them.
 \bibliographystyle{abbrv}
@@
 Details about graduate fellowships will be discussed in Section 4.
-
-** Current 2023 Rankings of CS PhD programs in the U.S.
-  Below is the 2023 Rankings from [[https://www.csrankings.org][CSRankings.org]] as of <2023-04-09 Sun>.
-
-1. Carnegie Mellon University 
-2. Univ. of Illinois at Urbana-Champaign 
-3. Univ. of California - San Diego  
-4. Massachusetts Institute of Technology 
-5. Stanford University 
-5. University of Michigan  
-7. University of Washington 
-8. Cornell University  
-8. Univ. of California - Berkeley  
-10. Georgia Institute of Technology 
-11. Northeastern University  
-11. University of Maryland - College Park 
-13. University of Wisconsin - Madison  
-14. Purdue University  closed chart
-15. University of Texas at Austin  
-16. Columbia University  
-16. University of Pennsylvania 
-18. New York University  
-18. Princeton University 
-20. University of Massachusetts Amherst  
-21. Univ. of California - Los Angeles
-22. University of Chicago 
-23. Stony Brook University
-23. Univ. of California - Santa Barbara 
-25. Rutgers University  
-26. Univ. of California - Irvine 
-27. University of Southern California 
-28. Duke University 
-29. Univ. of California - Riverside
-30. Pennsylvania State University 
-31. Northwestern University 
-32. Ohio State University 
-33. George Mason University 
-33. Harvard University  
-35. Brown University  
-35. University of Utah 
-37. Texas A&M University
-37. Univ. of California - Santa Cruz  
-37. Yale University  
-40. Boston University 
-40. University at Buffalo
-40. University of Colorado Boulder 
-43. North Carolina State University
-44. Rice University 
-44. University of Illinois at Chicago  
-46. University of North Carolina 
-46. University of Virginia 
-48. Arizona State University 
-48. University of Minnesota 
-48. Virginia Tech  
-51. Oregon State University  
-51. Univ. of California - Davis
 

### 2023-06-06 phd-cs-us.tex
updates

@@
-\documentclass[10pt]{article}
+\documentclass[11pt]{article}
 \usepackage[margin=1.1in]{geometry}
@@
 \title{Demystifying the Computer Science Ph.D. Admission in the US}
-\date{}
+\date{\today}
 \author{ThanhVu (Vu) Nguyen, George Mason University}
@@
 Having been involved in PhD admissions for many years and after
-numerous interaction with  students (especially Vietnamese), I've
-realized that \emph{international} students lack a clear understanding of
+numerous interaction with  students, I've
+realized that \emph{international} students, especially Vietnamese, lack a clear understanding of
 the Computer Science PhD admission process at US universities. This confusion not only
@@
 So I want to share my opinions and advice for those who are interested in applying for a \textbf{PhD in Computer Science in the US}.
-While I wrote this document for CS applicants, it should be applicable to various STEM fields.
-Moreover, while many examples given are for GMU, which is 33rd on CSRankings (June 2016), the writing should be generalize to most other R1\footnote{An \href{https://en.wikipedia.org/wiki/List_of_research_universities_in_the_United_States}{R1 institution} in the US refers to a research-intensive university with a high level of research activity across various disciplines.} universities  (though \emph{very} top schools might be very extreme, e.g., see the admission process at \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{CMU}).
+While I wrote this document for Vietnamese students interested in CS, it should be applicable to students in various countries interested in STEM.
+Moreover, while many examples given are for GMU, which is 33rd on CSRankings (see \S\ref{sec:ranking}), the writing should be generalize to most other R1\footnote{An \href{https://en.wikipedia.org/wiki/List_of_research_universities_in_the_United_States}{R1 institution} in the US refers to a research-intensive university with a high level of research activity across various disciplines.} universities  (though \emph{very} top schools might be very extreme, e.g., see the admission process at \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{CMU}).
 
@@
 funding is typically more available for PhD students than 
-Masters. % [[#funding][Read more about funding here]]
+Masters and especially undergraduate studies, which typically have no funding at all. \S\ref{sec:funding} provides more details on funding.
 
 Second, I believe that applying to a good US university \emph{should not} be any
-harder than at good schools in other countries. If you think you have a
-chance at schools in other countries, e.g., South Korea, Singapore, Germany, UK, Japan and Australia, then you surely have a chance in
-the US as well.
+harder than at schools in other countries. If you think you have a
+chance at schools in other countries, e.g., South Korea, Singapore, Germany, UK, Japan and Australia, then you will surely have a chance in the US as well.
 
@@
 Vu: One of the reasons I create this post is that several of my colleagues are interested in 
-recruiting Vietnamese students and were surprised when seeing very few Vietnamese applications (compared to other countries). In general the number of
+recruiting Vietnamese students and were surprised when seeing very few applications in Vietnam (compared to other countries). In general the number of
 PhD applications from Vietnam to US universities is very low and  more would be very welcomed. 
@@
 \begin{tcolorbox}[left=1pt,right=1pt,top=1pt,bottom=1pt]
-Vu: For me personally, it's better to have a good personalized
+Vu: It's better to have a good personalized
 letter about your own research ability from someone who is less
 well-known than a generic/weak letter from a well-known person.
+
+Vietnamese students sometimes go through great length just to get letters from well-known senior professors in their school, but the letters are very generic and weak and therefore carry little value, in fact, red flags.
+Moreover, a top professor in Vietnamese might not be well-known to US faculty (see more details in \S\ref{sec:your-school}).
+So save the trouble and just get letters from \emph{any} professors who knows you well and can write a good letter about your research ability.
 \end{tcolorbox}
@@
 This can be an issue for students in many top international universities where the competition is so high that very good students students can still have low ranking (and be overlooked by Admission committee).
-So what to do with this? well, same as \hyperref[sec:your-school]{before}, e.g., put a note about this in your statement.
+So what to do with this? well, same as \hyperref[sec:your-school]{before}, e.g., put a note about this in your statement and ask your LoR writers to talk about it.
 
@@
 statements give proper explanation). This is especially true if you
-have bad grades in relevant courses.
+have bad grades in relevant, e.g., CS and Math, courses.
 
@@
 
-You also do need to
-take standardized English test. Just do well enough to pass minimum
-requirement set by the university.
+\paragraph{English Test} Unless you're from certain countries (e.g., English speaking ones), you will need to
+take standardized English test. Just do well enough to pass minimum requirement set by the university, which nowadays has many options for you to choose from.
 
 \begin{tcolorbox}[left=1pt,right=1pt,top=1pt,bottom=1pt]
-Vu: The minimum for GMU CS (being above this might not mean much, but below is a \red{red flag}).
+Vu: The minimum for GMU (being above this might not mean much, but below is a \red{red flag}).
 \begin{itemize}
 \item GPA: $\ge 3.0$ in your undergrad (but as mentioned, we also consider the rank/prestige of your school)
-\item GRE: not required (most CS graduate programs no longer require GRE, though it can help boost your profile)
+\item GRE: not required, though it can help boost your profile
 %    \item but if you want to use it, then we expect a total (V+Q) of $\ge 311$ (with a $\ge 157$ Q) and A $\ge 3.0-3.5$.
@@
 This should be a summary of the accomplishments of the applicant.
-I would quickly scan this document to see if something stand out (e.g., Publications, Programming Competition Awards, Teaching Experiences).
+It should allow the reviewers to quickly scan to identify standout achievements (e.g., Publications, Programming Competition Awards, Teaching Experience).
 
@@
 \begin{tcolorbox}[left=1pt,right=1pt,top=1pt,bottom=1pt]
-At GMU, we are encouraged to interview candidates.
+Vu: At GMU, we are encouraged to interview candidates.
 For very strong candidates, the interview is actually to recruit them.
@@
 \begin{tcolorbox}[left=1pt,right=1pt,top=1pt,bottom=1pt]
-At GMU, TA and RA have similar benefits in tuition waiving and insurance.
+Vu: TA and RA at GMU have similar benefits in tuition waiving and insurance.
 For stipend, depending on the college and department, a 9-month graduate assistant stipend is set.
@@
 
-\subsection{Current 2023 Rankings of CS PhD programs in the U.S}
+\subsection{Current 2023 Rankings of CS PhD programs in the U.S}\label{sec:ranking}
   Below is the 2023 Rankings from \href{https://www.csrankings.org}{CSRankings.org}, a ranking system that is based on top CS conferences, as of April 2023.

### 2023-06-09 phd-cs-us.tex
updates csrankings table

@@
 \begin{abstract}
-Having been involved in PhD admissions for many years and after
+Having been involved in PhD admissions for many years and having
 numerous interaction with  students, I've
-realized that \emph{international} students, especially Vietnamese, lack a clear understanding of
+realized that \emph{international} students, especially \textbf{Vietnamese}, lack a clear understanding of
 the Computer Science PhD admission process at US universities. This confusion not only
@@
 First, I want to emphasize that PhD students in Computer
-Science \emph{do not} need to worry about funding, especially at R1
+Science \emph{do not} need to worry about funding, especially at good R1
 universities in the US. If you are admitted, you will almost certainly
 receive \emph{full funding} to support your study, including tuition,
-health insurance, and 9-month stipend. Moreover, depending on the university,
-you may even receive additional benefits like summer salary. Note that
+health insurance, and stipend. Moreover, depending on the university,
+you may even receive additional benefits such as summer salary and laptops. Note that
 funding is typically more available for PhD students than 
-Masters and especially undergraduate studies, which typically have no funding at all. \S\ref{sec:funding} provides more details on funding.
+Masters and especially undergraduate studies, which typically have no funding. \S\ref{sec:funding} provides more details on funding.
 
@@
 Vu: One of the reasons I create this post is that several of my colleagues are interested in 
-recruiting Vietnamese students and were surprised when seeing very few applications in Vietnam (compared to other countries). In general the number of
+recruiting Vietnamese students and were surprised when seeing very few applications in Vietnam (e.g., our CS PhD program has on average 3 applications from Vietnam each year). In general the number of
 PhD applications from Vietnam to US universities is very low and  more would be very welcomed. 
@@
 
-In many cases, the admission committee will involve junior faculty (e.g., assistant professors) in the department.
-While this requires  a significant time investment, it provides junior faculty the opportunities to recruit students.
-The chair of the committee will be a senior professor, but they will not review individual applications and instead assign them to committee members.
-The chair will look at various factors such as research interests or mentioning faculty names, and assign the applications to appropriate faculty.
+In many cases, the admission committee can involve current students and assistant professors in the department.
+This provides junior faculty the opportunities to recruit students.
+The chair of the committee will be a senior professor, but they likely will not review individual applications and instead assign them to committee members.
+The chair will look at various factors such as research interests or mentioning faculty names to assign the applications to appropriate faculty.
 
@@
 \subsection{Current 2023 Rankings of CS PhD programs in the U.S}\label{sec:ranking}
-  Below is the 2023 Rankings from \href{https://www.csrankings.org}{CSRankings.org}, a ranking system that is based on top CS conferences, as of April 2023.
+  Tab.~\ref{tab:ranking} lists the top 50 CS programs in the US from \href{https://www.csrankings.org}{CSRankings.org}, a ranking system that is based on CS conferences.
 
-\begin{itemize}
-\item Carnegie Mellon University 
-\item Univ. of Illinois at Urbana-Champaign 
-\item Univ. of California - San Diego  
-\item Massachusetts Institute of Technology 
-\item Stanford University         
-\item University of Michigan  
-\item University of Washington 
-\item Cornell University  
-\item Univ. of California - Berkeley  
-\item Georgia Institute of Technology 
-\item Northeastern University  
-\item University of Maryland - College Park 
-\item University of Wisconsin - Madison  
-\item Purdue University  closed chart
-\item University of Texas at Austin  
-\item Columbia University  
-\item University of Pennsylvania 
-\item New York University  
-\item Princeton University 
-\item University of Massachusetts Amherst  
-\item Univ. of California - Los Angeles
-\item University of Chicago 
-\item Stony Brook University
-\item Univ. of California - Santa Barbara 
-\item Rutgers University  
-\item Univ. of California - Irvine 
-\item University of Southern California 
-\item Duke University 
-\item Univ. of California - Riverside
-\item Pennsylvania State University 
-\item Northwestern University 
-\item Ohio State University 
-\item George Mason University 
-\item Harvard University  
-\item Brown University  
-\item University of Utah 
-\item Texas A\&M University
-\item Univ. of California - Santa Cruz  
-\item Yale University  
-\item Boston University 
-\item University at Buffalo
-\item University of Colorado Boulder 
-\item North Carolina State University
-\item Rice University 
-\item University of Illinois at Chicago  
-\item University of North Carolina 
-\item University of Virginia 
-\item Arizona State University 
-\item University of Minnesota 
-\item Virginia Tech  
-\item  Oregon State University  
-\item Univ. of California - Davis
-\end{itemize}
+  \begin{table}
+    \centering
+    \small
+    \caption{Top 50 CS PhD programs in the U.S. (CSRankings, April 2023)}\label{tab:ranking}
+  \begin{tabular}{rl|rl}
+    \toprule
+    1 & Carnegie Mellon University & 26 & Univ. of California - Irvine \\
+    2 & Univ. of Illinois at Urbana-Champaign  & 27 & University of Southern California \\
+    3 & Univ. of California - San Diego & 28 &  Duke University \\
+    4 & Massachusetts Institute of Technology & 29 & Univ. of California - Riverside\\
+    5 & Stanford University         & 30 & Pennsylvania State University \\
+    6 & University of Michigan - Ann Arbor & 31 & Northwestern University \\
+    7 & University of Washington   & 32& Ohio State University \\  
+    8 & Cornell University       &33 & George Mason University \\
+    9 & Univ. of California - Berkeley  &34& Harvard University  \\
+    10 & Georgia Institute of Technology &35& Brown University  \\
+    11 & Northeastern University  &36& University of Utah \\
+    12 & University of Maryland - College Park &37& Texas A\&M University\\
+    13 & University of Wisconsin - Madison  &38& Univ. of California - Santa Cruz  \\
+    14 &   Purdue University  &39 & Yale University  \\
+    15 & University of Texas at Austin  &40& Boston University \\
+    16 &   Columbia University  &41 & University at Buffalo\\
+    17 & University of Pennsylvania &42& University of Colorado-Boulder \\
+    18 & New York University  & 43& North Carolina State University\\
+    19 & Princeton University & 44& Rice University \\
+    20 & University of Massachusetts Amherst &45& University of Illinois at Chicago  \\
+    21 & Univ. of California - Los Angeles &46& University of North Carolina \\
+    22 & University of Chicago &47& University of Virginia \\
+    23 & Stony Brook University &48& Arizona State University \\
+    24 & Univ. of California - Santa Barbara &49& University of Minnesota \\
+    25 & Rutgers University &50& Virginia Tech  \\
+    \bottomrule
+    \end{tabular}
+\end{table}
 

### 2023-06-09 phd-cs-us.tex
updates csrankings table

@@
 
-Sometime a faculty wants to interview an applicant to make a decision. Typically this means they lean toward admitting you (if we don't like your application, we will not bother doing the interview).
+Sometime a faculty wants to interview an applicant to make a decision. This means they are leaning toward admitting you (if we don't like your application, we will not bother doing the interview).
 
@@
 For very strong candidates, the interview is actually to recruit them.
-In some rare cases a faculty interviews a candidate that they see potentials and want argue for admission, i.e., without the interview, that candidate is definitely rejected.
+In some cases a faculty interviews a candidate that they see potentials and wants argue for admission, i.e., without the interview, that candidate is definitely rejected.
 In any case, getting interview means you have a very good chance of being admitted.
@@
   \centering
+  \footnotesize
   \caption{Different types of PhD funding}\label{tab:funding}

### 2023-06-10 phd-cs-us.tex
Updates from Overleaf

@@
 \usepackage{booktabs}
+
+\newtcolorbox{cbox}[1][]
+{
+left=1pt,right=1pt,top=1pt,bottom=1pt,
+colback=gray!10,
+boxrule=0.5pt,
+title=#1 % I would like to make this (one of these in general) assignment optional depending on #1, #2...
+}
+
+ \newenvironment{commentbox}[1][]{
+ \small
+    \begin{cbox}[#1]
+ }{
+   \end{cbox}
+}
 
@@
 
-\begin{tcolorbox}[left=1pt,right=1pt,top=1pt,bottom=1pt]
+\begin{commentbox}[]
 Vu: One of the reasons I create this post is that several of my colleagues are interested in 
@@
 PhD applications from Vietnam to US universities is very low and  more would be very welcomed. 
-\end{tcolorbox}
+\end{commentbox}
 
@@
 
-\begin{tcolorbox}[left=1pt,right=1pt,top=1pt,bottom=1pt]
-Vu: We usually decide that a candidate is either (i) admit with funding (TA or RA) or (ii) rejected. In other words, in most cases, we either
+\begin{commentbox}[]
+Vu: We usually decide that a PhD candidate is either (i) admit with funding (TA or RA) or (ii) rejected. In other words, in most cases, we either
 admit you with full funding, or we don't. In some rare cases, we may admit
 you with no funding because you have funding on your own (e.g., you are
-supported by your government or have external grants). Also, we justify
+supported by your government or have external grants). We justify
 our decision with a short summary about your application, where we list
@@
 LORs).
-\end{tcolorbox}
+\end{commentbox}
 
-\begin{tcolorbox}[left=1pt,right=1pt,top=1pt,bottom=1pt]
-    Hakan: At GMU, for full consideration, students should make sure to submit \textbf{ALL} required documents by the application deadline (December 1st), and should never assume that some required documents (such as official TOEFL scores or official diplomas/transcripts) will be waived by the admissions office.
+\begin{commentbox}[]
+    Hakan: At GMU, for full consideration, students should make sure to submit \textbf{ALL} required documents by the application deadline, and should never assume that some required documents (such as official TOEFL scores or official diplomas/transcripts) will be waived by the admissions office.
 If something is listed and not marked as "optional", it is mandatory and they should plan for submitting all those.
-\end{tcolorbox}
+\end{commentbox}
 \paragraph{Why we do not waive application fee?}  This is typically a requirement of the university.
@@
 
-\begin{tcolorbox}[left=1pt,right=1pt,top=1pt,bottom=1pt]
-Vu: Vietnamese students often mention Scopus Q1, which consists of diverse journals from IEEE, Elsevier, and several other publishers unfamiliar to me.
-Honestly, I don't know/recognize the majority of journals listed in Scopus Q1.
+\begin{commentbox}[]
+Vu: Vietnamese students often mention Scopus Q1, which consists of diverse journals from IEEE, Elsevier, and many other publishers unfamiliar to me.
+Honestly, I don't know/recognize many of journals listed in Scopus Q1 and.
 So this might be something to be mindful of, as \textbf{CS} faculty might not be too familiar with Scopus or journals listed in there, so devote sometime in your statement to discuss the significance of your papers.
-\end{tcolorbox}
+\end{commentbox}
 
-\begin{tcolorbox}[left=1pt,right=1pt,top=1pt,bottom=1pt]
+\begin{commentbox}[]
 Craig: GMU and many other universities allow you to upload your published papers and other writing samples.
 In many cases, even if the papers were not published at top places, we can still determine their quality by simply skimming over the paper.
-\end{tcolorbox}
+\end{commentbox}
 
@@
 
-\subsection{Research Statement}\label{sec:research-statement}
+\subsection{Research or Personal Statement}\label{sec:research-statement}
 
-While you might not be able to get good LORs or change
-\hyperref[sec:your-school]{where your go to school}, you have control of your
+While you might not have control over LORs \hyperref[sec:your-school]{where your go to school}, you do over your
 statement! So write it well because we do take it seriously.
@@
 
-Finally, this is something very easy to do, but is missed by many
+Finally, this is something easy to do, but is missed by many
 applicants: \textbf{customize the statement} for the school you're applying to,
 e.g., why do you apply here? talk about a couple of professors who you're interested in working with (in many cases your application will be forwarded to them for evaluation).
+This shows that you're serious and have done homework on where you're applying to.
 Be careful not to send wrong statement to wrong school or mixing
@@
 \begin{tcolorbox}[left=1pt,right=1pt,top=1pt,bottom=1pt]
+\small
 Vu: I always read the research statement first and then LORs. If I am

### 2023-06-10 phd-cs-us.tex
updates

@@
 
-\newtcolorbox{cbox}[1][]
+\newtcolorbox{cbox}[]
 {
@@
 colback=gray!10,
-boxrule=0.5pt,
-title=#1 % I would like to make this (one of these in general) assignment optional depending on #1, #2...
+boxrule=0.5pt
 }
 
- \newenvironment{commentbox}[1][]{
+ \newenvironment{commentbox}{
  \small
-    \begin{cbox}[#1]
+    \begin{cbox}
  }{
@@
 While I wrote this document for Vietnamese students interested in CS, it should be applicable to students in various countries interested in STEM.
-Moreover, while many examples given are for GMU, which is 33rd on CSRankings (see \S\ref{sec:ranking}), the writing should be generalize to most other R1\footnote{An \href{https://en.wikipedia.org/wiki/List_of_research_universities_in_the_United_States}{R1 institution} in the US refers to a research-intensive university with a high level of research activity across various disciplines.} universities  (though \emph{very} top schools might be very extreme, e.g., see the admission process at \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{CMU}).
+Moreover, while many examples given are for GMU, which is 33rd on CSRankings (see \S\ref{sec:ranking}), the writing should be generalize to most other R1\footnote{An \href{https://en.wikipedia.org/wiki/List_of_research_universities_in_the_United_States}{R1 institution} in the US is a research-intensive university with a high level of research activity across various disciplines.} universities  (though \emph{very} top schools might be very extreme, e.g., see the admission process at \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{CMU}).
 
@@
 
-\begin{commentbox}[]
+\begin{commentbox}
 Vu: One of the reasons I create this post is that several of my colleagues are interested in 
@@
 %% Vu: *what's a PhD?*  This [[https://matt.might.net/articles/phd-school-in-pictures/][series of pictures]] from [[https://matt.might.net][Matt Might]] illustrates what a PhD means.
-%% \end{tcolorbox}
+%% \end{commentbox}
 
@@
 
-\begin{commentbox}[]
+\begin{commentbox}
 Vu: We usually decide that a PhD candidate is either (i) admit with funding (TA or RA) or (ii) rejected. In other words, in most cases, we either
@@
 
-\begin{commentbox}[]
+\begin{commentbox}
     Hakan: At GMU, for full consideration, students should make sure to submit \textbf{ALL} required documents by the application deadline, and should never assume that some required documents (such as official TOEFL scores or official diplomas/transcripts) will be waived by the admissions office.
@@
 
-\begin{commentbox}[]
+\begin{commentbox}
 Vu: Vietnamese students often mention Scopus Q1, which consists of diverse journals from IEEE, Elsevier, and many other publishers unfamiliar to me.
@@
 
-\begin{commentbox}[]
+\begin{commentbox}
 Craig: GMU and many other universities allow you to upload your published papers and other writing samples.
@@
 
-\begin{tcolorbox}[left=1pt,right=1pt,top=1pt,bottom=1pt]
+\begin{commentbox}
 Vu: It's better to have a good personalized
@@
 So save the trouble and just get letters from \emph{any} professors who knows you well and can write a good letter about your research ability.
-\end{tcolorbox}
+\end{commentbox}
 
@@
 
-\begin{tcolorbox}[left=1pt,right=1pt,top=1pt,bottom=1pt]
+\begin{commentbox}
 \small
@@
 something standout in other parts).
-\end{tcolorbox}
+\end{commentbox}
 
@@
 
-\begin{tcolorbox}[left=1pt,right=1pt,top=1pt,bottom=1pt]
+\begin{commentbox}
 Vu: Sometime PhD admission committee will share a document such as \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Foreign-Top-Schools}{this one}, which lists the top schools in several countries.
 I have looked at Vietnamese applications (whether they are assigned to me or not) and provide inputs to the reviewers of those applications, e.g., X is the top tech school in Vietnam and so it should be \emph{top} instead of \emph{unknown} foreign, which makes a huge difference.
-\end{tcolorbox}
+\end{commentbox}
 
@@
 
-\begin{tcolorbox}[left=1pt,right=1pt,top=1pt,bottom=1pt]
+\begin{commentbox}
 Vu: The minimum for GMU (being above this might not mean much, but below is a \red{red flag}).
@@
 \end{itemize}
-\end{tcolorbox}
+\end{commentbox}
 
@@
 
-\begin{tcolorbox}[left=1pt,right=1pt,top=1pt,bottom=1pt]
+\begin{commentbox}
 Vu: At GMU, we are encouraged to interview candidates.
@@
 In any case, getting interview means you have a very good chance of being admitted.
-\end{tcolorbox}
+\end{commentbox}
 
@@
 
-\begin{tcolorbox}[left=1pt,right=1pt,top=1pt,bottom=1pt]
+\begin{commentbox}
 Vu: GMU has Virtual Open House, e.g., \url{https://cs-gmu.github.io/cs-phd-voh-s23/}, which I've co-organized in the last two years.
@@
 Students also get opportunities to chat with professors and current students.
-\end{tcolorbox}
+\end{commentbox}
 
@@
 
-\begin{tcolorbox}[left=1pt,right=1pt,top=1pt,bottom=1pt]
+\begin{commentbox}
 Vu: TA and RA at GMU have similar benefits in tuition waiving and insurance.
@@
 TA and RA will usually be that amount (TA will definitely be that, RA might fluctuate a bit depending on the stage of the student and the prof.).
-\end{tcolorbox}
+\end{commentbox}
 
@@
 
-\begin{tcolorbox}[left=1pt,right=1pt,top=1pt,bottom=1pt]
+\begin{commentbox}
 At GMU CS, students admitted with TA have  4 years of GTA guaranteed and also receive  stipend for the \textbf{first} summer.
-\end{tcolorbox}
+\end{commentbox}
 
@@
 
-\begin{tcolorbox}[left=1pt,right=1pt,top=1pt,bottom=1pt]
+\begin{commentbox}
 Vu: If you're lucky and got recruited by a prof.
@@
 For example, if a prof., even if not in PhD admission committee, wants to work with and funds you, the PhD admission committee will respect that decision and admit your application (unless your application has many red flags).
-\end{tcolorbox}
+\end{commentbox}
 
@@
 
-\begin{tcolorbox}[left=1pt,right=1pt,top=1pt,bottom=1pt]
+\begin{commentbox}
 At GMU, Ph.D.
@@
 PhD admission committee members nominate applicants for this fellowship and the committee will vote and give the fellowship to the top 2.
-\end{tcolorbox}
+\end{commentbox}
 
@@
 \begin{enumerate}   
-    \item   Is an MS degree required for admission to PhD?
-    \begin{description}
-    \item No.
-In fact, student with BS can get MS degree "along the way" to PhD.
-However, MS can help if it gives research experience or is from a more well-known school than your undergrad institution.
-    \end{description}
-
-    \item Can I apply to PhD to CS if my undergrad was in Math, Biology, Finance, etc?
-    \begin{description}
-        \item Yes, as long as you can demonstrate you are ready for CS PhD research through research experiences, LoRs, statements, etc as mentioned.
-    \end{description}
-
     \item What can you do to increase your admission chance?
@@
 If you write technical, research-like blogs, that would also help.
+    \end{description}
+
+    \item Can I apply to PhD to CS if my undergrad was not in CS or any related areas?
+    \begin{description}
+        \item Yes, as long as you can demonstrate you are ready for CS PhD research through research experiences, LoRs, statements, etc as mentioned.
+You might be even able leverage this to make your profile stand out.
+    \end{description}
+
+    \item   Is an MS degree required for admission to PhD?
+    \begin{description}
+    \item No.
+In fact, student with BS can get MS degree "along the way" to PhD.
+However, MS can help if it gives research experience or is from a more well-known school than your undergrad institution.
     \end{description}
@@
 
-\begin{tcolorbox}[left=1pt,right=1pt,top=1pt,bottom=1pt]
+\begin{commentbox}
     Vu: I've been called Dr. Vu and I find it a bit amusing but am totally fine with it.
-    \end{tcolorbox}
+    \end{commentbox}
     \end{description}

### 2023-06-10 phd-cs-us.tex
Updates from Overleaf

@@
 
-\newtcolorbox{cbox}[]
+\newtcolorbox{cbox}
 {
@@
 \renewcommand{\tablename}{Tab.}
-\title{Demystifying the Computer Science Ph.D. Admission in the US}
+\title{Demystifying the Computer Science Ph.D. Admission in the US \\{\Large A guide for Vietnamese and International Students}}
 \date{\today}
-\author{ThanhVu (Vu) Nguyen, George Mason University}
+\author{\small ThanhVu (Vu) Nguyen, George Mason University}
 %\setcounter{secnumdepth}{3}
@@
 While I wrote this document for Vietnamese students interested in CS, it should be applicable to students in various countries interested in STEM.
-Moreover, while many examples given are for GMU, which is 33rd on CSRankings (see \S\ref{sec:ranking}), the writing should be generalize to most other R1\footnote{An \href{https://en.wikipedia.org/wiki/List_of_research_universities_in_the_United_States}{R1 institution} in the US is a research-intensive university with a high level of research activity across various disciplines.} universities  (though \emph{very} top schools might be very extreme, e.g., see the admission process at \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{CMU}).
+Moreover, while many examples given are for GMU, which is 32rd in the US according CSRankings (see \S\ref{sec:ranking}), the writing should be generalize to most other R1\footnote{An \href{https://en.wikipedia.org/wiki/List_of_research_universities_in_the_United_States}{R1 institution} in the US is a research-intensive university with a high level of research activity across various disciplines.} universities  (though \emph{very} top schools might be very extreme, e.g., see the admission process at \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{CMU}).
 
@@
 admit you with full funding, or we don't. In some rare cases, we may admit
-you with no funding because you have funding on your own (e.g., you are
-supported by your government or have external grants). We justify
-our decision with a short summary about your application, where we list
-strengths (e.g., well-known undergrad school) and weaknesses (e.g., weak
+you without funding because you have funding on your own (e.g., you are
+supported by your government or have external grants). We also justify
+our decision with a summary about your application, where we list
+strengths (e.g., well-known school) and weaknesses (e.g., weak
 LORs).
@@
 
-\paragraph{English Test} Unless you're from certain countries (e.g., English speaking ones), you will need to
+\paragraph{English Test} Unless your degrees are from certain countries such as \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-GMU#standard-tests-waiver-eligible-countries}{these}, you will need to
 take standardized English test. Just do well enough to pass minimum requirement set by the university, which nowadays has many options for you to choose from.
@@
 \begin{itemize}
-\item GPA: $\ge 3.0$ in your undergrad (but as mentioned, we also consider the rank/prestige of your school)
+\item GPA: $\ge 3.0$ in your undergrad (but we also consider the rank/prestige of your school)
 \item GRE: not required, though it can help boost your profile
@@
 
-Fellowship is another type of funding in which the student applies for (e.g., from school, industries, government).
+Fellowship is another type of funding that the student needs to apply for (e.g., from school, industries, government).
 Fellowships are typically competitive and generous, and gives pretty much all benefits tuition/insurance that a TA/RA has.
@@
 \textbf{How to get Fellowship?} You apply for them.
-The US government has many fellowships but these would likely require US citizenship or residency.
+The US government has many fellowships that would likely require US citizenship or residency.
 However, tech companies including Google, Microsoft, Facebook, IBM have fellowships that international students can apply for.
@@
 \subsection{Current 2023 Rankings of CS PhD programs in the U.S}\label{sec:ranking}
-  Tab.~\ref{tab:ranking} lists the top 50 CS programs in the US from \href{https://www.csrankings.org}{CSRankings.org}, a ranking system that is based on CS conferences.
+  Tab.~\ref{tab:ranking} lists the top 50 CS programs in the US from \href{https://www.csrankings.org}{CSRankings.org}, a ranking system  based on top CS conferences.
 
-  \begin{table}
+  \begin{table}[h!]
     \centering
     \small
-    \caption{Top 50 CS PhD programs in the U.S. (CSRankings, April 2023)}\label{tab:ranking}
+    \caption{Top 50 CS PhD programs in the U.S. (CSRankings, June 2023)}\label{tab:ranking}
   \begin{tabular}{rl|rl}
     \toprule
-    1 & Carnegie Mellon University & 26 & Univ. of California - Irvine \\
-    2 & Univ. of Illinois at Urbana-Champaign  & 27 & University of Southern California \\
-    3 & Univ. of California - San Diego & 28 &  Duke University \\
-    4 & Massachusetts Institute of Technology & 29 & Univ. of California - Riverside\\
-    5 & Stanford University         & 30 & Pennsylvania State University \\
-    6 & University of Michigan - Ann Arbor & 31 & Northwestern University \\
-    7 & University of Washington   & 32& Ohio State University \\  
-    8 & Cornell University       &33 & George Mason University \\
-    9 & Univ. of California - Berkeley  &34& Harvard University  \\
-    10 & Georgia Institute of Technology &35& Brown University  \\
-    11 & Northeastern University  &36& University of Utah \\
-    12 & University of Maryland - College Park &37& Texas A\&M University\\
-    13 & University of Wisconsin - Madison  &38& Univ. of California - Santa Cruz  \\
-    14 &   Purdue University  &39 & Yale University  \\
-    15 & University of Texas at Austin  &40& Boston University \\
-    16 &   Columbia University  &41 & University at Buffalo\\
-    17 & University of Pennsylvania &42& University of Colorado-Boulder \\
-    18 & New York University  & 43& North Carolina State University\\
-    19 & Princeton University & 44& Rice University \\
-    20 & University of Massachusetts Amherst &45& University of Illinois at Chicago  \\
-    21 & Univ. of California - Los Angeles &46& University of North Carolina \\
-    22 & University of Chicago &47& University of Virginia \\
-    23 & Stony Brook University &48& Arizona State University \\
-    24 & Univ. of California - Santa Barbara &49& University of Minnesota \\
-    25 & Rutgers University &50& Virginia Tech  \\
+    1 & Carnegie Mellon & 26 & Univ. of California - Irvine \\
+    2 & Univ. of Illinois at Urbana-Champaign  & 27 &  Duke University \\
+    3 & Univ. of California-San Diego & 28 & Rutgers University \\
+    4 & MIT & 29 & Univ. of California - Riverside\\
+    5 & Georgia Institute of Technology         & 30 & Northwestern University\\
+    6 & Stanford University& 31 & Pennsylvania State University  \\
+    7 & University of Michigan - Ann Arbor   & 32& George Mason University\\  
+    8 & University of Washington      &33 &  Harvard University \\
+    9 &  Univ. of California - Berkeley  &34&  Univ. of California - Santa Cruz \\
+    10 & Cornell University  & 35 &  Yale University \\
+    11 & University of Maryland - College Park &  36& Brown University \\ 
+    12 & Northeastern University 37&  Ohio State University\\
+    13 & University of Wisconsin - Madison  &38& Texas A\&M University \\
+    14 & Columbia University   &39 & Boston University  \\
+    15 &   Purdue University  &40& North Carolina State University\\\  
+    16 & University of Texas at Austin   &41 & University of Utah \\
+    17 & University of Pennsylvania &42 & University at Buffalo\\
+    18 & Princeton University  & 43& Rice University\\
+    19 & University of Massachusetts-Amherst & 44&  University of Colorado-Boulder \\
+    20 &  New York University  &45& University of Illinois at Chicago  \\
+    21 & Univ. of California - Los Angeles &46& Virginia Tech  \\
+    22 & University of Southern California &47&  Arizona State University \\
+    23 & Stony Brook University &48&University of Minnesota \\
+    24 & University of Chicago &49& University of Virginia \\
+    25 & Univ. of California - Santa Barbara &50& University of North Carolina \\
     \bottomrule

### 2023-06-11 phd-cs-us.tex
updates

--- 2023-06-10 phd-cs-us.tex
@@
 \usepackage{booktabs}
-
-\newtcolorbox{cbox}[1][]
-{
-left=1pt,right=1pt,top=1pt,bottom=1pt,
-colback=gray!10,
-boxrule=0.5pt,
-title=#1 % I would like to make this (one of these in general) assignment optional depending on #1, #2...
-}
-
- \newenvironment{commentbox}[1][]{
- \small
-    \begin{cbox}[#1]
- }{
-   \end{cbox}
-}
 
@@
 
-\begin{commentbox}[]
+\begin{tcolorbox}[left=1pt,right=1pt,top=1pt,bottom=1pt]
 Vu: One of the reasons I create this post is that several of my colleagues are interested in 
@@
 PhD applications from Vietnam to US universities is very low and  more would be very welcomed. 
-\end{commentbox}
+\end{tcolorbox}
 
@@
 
-\begin{commentbox}[]
-Vu: We usually decide that a PhD candidate is either (i) admit with funding (TA or RA) or (ii) rejected. In other words, in most cases, we either
+\begin{tcolorbox}[left=1pt,right=1pt,top=1pt,bottom=1pt]
+Vu: We usually decide that a candidate is either (i) admit with funding (TA or RA) or (ii) rejected. In other words, in most cases, we either
 admit you with full funding, or we don't. In some rare cases, we may admit
 you with no funding because you have funding on your own (e.g., you are
-supported by your government or have external grants). We justify
+supported by your government or have external grants). Also, we justify
 our decision with a short summary about your application, where we list
@@
 LORs).
-\end{commentbox}
+\end{tcolorbox}
 
-\begin{commentbox}[]
-    Hakan: At GMU, for full consideration, students should make sure to submit \textbf{ALL} required documents by the application deadline, and should never assume that some required documents (such as official TOEFL scores or official diplomas/transcripts) will be waived by the admissions office.
+\begin{tcolorbox}[left=1pt,right=1pt,top=1pt,bottom=1pt]
+    Hakan: At GMU, for full consideration, students should make sure to submit \textbf{ALL} required documents by the application deadline (December 1st), and should never assume that some required documents (such as official TOEFL scores or official diplomas/transcripts) will be waived by the admissions office.
 If something is listed and not marked as "optional", it is mandatory and they should plan for submitting all those.
-\end{commentbox}
+\end{tcolorbox}
 \paragraph{Why we do not waive application fee?}  This is typically a requirement of the university.
@@
 
-\begin{commentbox}[]
-Vu: Vietnamese students often mention Scopus Q1, which consists of diverse journals from IEEE, Elsevier, and many other publishers unfamiliar to me.
-Honestly, I don't know/recognize many of journals listed in Scopus Q1 and.
+\begin{tcolorbox}[left=1pt,right=1pt,top=1pt,bottom=1pt]
+Vu: Vietnamese students often mention Scopus Q1, which consists of diverse journals from IEEE, Elsevier, and several other publishers unfamiliar to me.
+Honestly, I don't know/recognize the majority of journals listed in Scopus Q1.
 So this might be something to be mindful of, as \textbf{CS} faculty might not be too familiar with Scopus or journals listed in there, so devote sometime in your statement to discuss the significance of your papers.
-\end{commentbox}
+\end{tcolorbox}
 
-\begin{commentbox}[]
+\begin{tcolorbox}[left=1pt,right=1pt,top=1pt,bottom=1pt]
 Craig: GMU and many other universities allow you to upload your published papers and other writing samples.
 In many cases, even if the papers were not published at top places, we can still determine their quality by simply skimming over the paper.
-\end{commentbox}
+\end{tcolorbox}
 
@@
 
-\subsection{Research or Personal Statement}\label{sec:research-statement}
+\subsection{Research Statement}\label{sec:research-statement}
 
-While you might not have control over LORs \hyperref[sec:your-school]{where your go to school}, you do over your
+While you might not be able to get good LORs or change
+\hyperref[sec:your-school]{where your go to school}, you have control of your
 statement! So write it well because we do take it seriously.
@@
 
-Finally, this is something easy to do, but is missed by many
+Finally, this is something very easy to do, but is missed by many
 applicants: \textbf{customize the statement} for the school you're applying to,
 e.g., why do you apply here? talk about a couple of professors who you're interested in working with (in many cases your application will be forwarded to them for evaluation).
-This shows that you're serious and have done homework on where you're applying to.
 Be careful not to send wrong statement to wrong school or mixing
@@
 \begin{tcolorbox}[left=1pt,right=1pt,top=1pt,bottom=1pt]
-\small
 Vu: I always read the research statement first and then LORs. If I am
@@
 
-  \begin{table}
+  \begin{table}[ht!]
     \centering

### 2023-06-11 phd-cs-us.tex

--- 2023-06-10 phd-cs-us.tex
@@
 \newcommand{\tvn}[1]{\mycomment{TVN}{#1}}{}
+\newcommand{\tl}[1]{\mycomment{ThanhLe}{#1}}{}
 \newcommand{\red}[1]{{\color{red}{#1}}}
@@
 Similarly, many professors write very generic letters for students (a common example is that the students didn't do any
-research or make any impression for the professor to write about). These
-letters are also not useful and considered weak.
+research or make any impression for the professor to write about). These letters are also not useful and considered weak.
 
-Many students get letters from someone from company where they did internship or are
+Many students get letters from supervisors from company where they did internship or are
 working at. It is OK as long as it is a research-based personalized
@@
 \begin{commentbox}
-Vu: It's better to have a good personalized
+Vu: In Vietnam some letter writers ask the applicants to write their own letters so they can just sign.
+As mentioned, this will hurt the applicants (admission committee are actually quite good at determining this)!
+I was told that some people did this to ensure the applicants will only go to places where they want to the students to go to (e.g., stay in Vietnam or go to specific places where they have networks).
+\end{commentbox}
+
+\begin{commentbox}
+Vu: Vietnamese students sometimes go through great length just to get letters from well-known senior professors in their school, but the letters are generic and carry little value, in fact, \red{red flags}.
+Moreover, a top professor in Vietnamese might not be well-known to US faculty (see more details in \S\ref{sec:your-school}).
+So save the trouble and just get letters from \emph{any} professors/supervisors who knows you well and can write a good letter about your research ability.
+
+In short, it's better to have a good personalized
 letter about your own research ability from someone who is less
 well-known than a generic/weak letter from a well-known person.
-
-Vietnamese students sometimes go through great length just to get letters from well-known senior professors in their school, but the letters are very generic and weak and therefore carry little value, in fact, red flags.
-Moreover, a top professor in Vietnamese might not be well-known to US faculty (see more details in \S\ref{sec:your-school}).
-So save the trouble and just get letters from \emph{any} professors who knows you well and can write a good letter about your research ability.
 \end{commentbox}
@@
 Prestigious fellowships typically require a clear and good research plan, so it is a good idea to wait until at least your second year to have research experience and even publication before applying.
-Remember, you're competiting with the top Ph.D.
+Remember, you're competing with the top Ph.D.
 students at top universities worldwide.
@@
 
-        Also, even if you do not have formal research experience, you can talk about your personal project.
-If it is used by many people, have lots of stars in Github, etc, it would certainly worth talking about.
-If you write technical, research-like blogs, that would also help.
+    \item Even if you do not have formal research experience, you can talk about your personal projects.
+For example, if you have an open-source project on Github that is used by many people, has lots of stars in Github, do talk about it.
+If you write technical, research-like blogs, talk about them too.
+    \end{description}
+
+   \item How to rank or select a CS program?
+
+\begin{description}
+\item International students not familiar with US universities often put them into \emph{two} bins:  (i) very top school that they dream about such as Stanford, MIT, Princeton, Harvard and (ii) everything else.
+Sometimes they apply reputations of schools for non-CS programs, e.g., medical or physics, to rank CS programs.
+In some cases they rank universities based on popular states they know in the US, e.g., California, New York, and Texas.  Let's just say there are so many thing wrong with these methods.
+
+\item You can learn about CS programs and research expertise of faculty using resources such as \href{https://csrankings.org}{CSRankings.org}, which is designed specifically to help prospective PhD students in Computer Science!
+You will be very surprised to learn that a school X that you don't know much about has very strong research in your interested field Y.
+This is also a good way to learn about individual faculty (who works on what) and the well-known CS conferences\footnote{In CS (and only in CS), conferences, not journals, are oten the main venue to publish research finding.}.
+\begin{commentbox}
+    Dat: many Vietnamese students, even those from very top engineering schools, do not know about CSRankings.  Only few applicants who worked at top research places such as VinAI might know about it.
+\end{commentbox}
+
+\item But more generally, rankings is superficial and you will want to do more research to be informed and make better decision.
+For example, if you get admissions to several places, you should contact profs.
+that you're interested in at those place and talk to them.
+They would be more willing to chat to you now that you have been admitted.
+Ask them questions about their work, how they manage students, their expectations.
+You can even ask to contact their students.
 \end{description}
@@
 \section{Acknowledgement}
-Many faculty and students have contributed to this document.
-Craig Yu (GMU) and Hakan Aydin (GMU) contributed to the early version of this document. Other GMU faculty also have provided valuable feedbacks.  Thanks to all of them.
+
+Many people have contributed to this document.
+Craig Yu (GMU) and Hakan Aydin (GMU) provided valuable input in the early version.
+Other GMU faculty members also have provided feedback and contributions.
+Many Vietnamese students have provided feedback and contributed questions, which provide contents for the writing.
 \bibliographystyle{abbrv}

### 2023-06-11 phd-cs-us.tex
updates

@@
 \date{\today}
-\author{\small ThanhVu (Vu) Nguyen, George Mason University}
+\author{\small \href{https://nguyenthanhvuh.github.io}{ThanhVu (Vu) Nguyen}, \href{https://cs.gmu.edu}{George Mason University}}
 %\setcounter{secnumdepth}{3}
@@
 
-If you have suggestions or comments, feel free to create a \href{https://github.com/nguyenthanhvuh/phd-cs-us/issues}{GitHub issue} for discussion.
+If you have questions or comments, feel free to create a \href{https://github.com/nguyenthanhvuh/phd-cs-us/issues}{GitHub issue} for discussion.
 \end{abstract}
@@
 Vu: One of the reasons I create this post is that several of my colleagues are interested in 
-recruiting Vietnamese students and were surprised when seeing very few applications in Vietnam (e.g., our CS PhD program has on average 3 applications from Vietnam each year). In general the number of
-PhD applications from Vietnam to US universities is very low and  more would be very welcomed. 
+recruiting Vietnamese students and were surprised when seeing very few applications in Vietnam (e.g., our CS PhD program has on average 5-6 applications from Vietnam each year).
+In general the number of
+PhD applications from Vietnam to US universities is few and  more would be very welcomed. 
 \end{commentbox}
@@
 \begin{commentbox}
-Vu: Vietnamese students often mention Scopus Q1, which consists of diverse journals from IEEE, Elsevier, and many other publishers unfamiliar to me.
+Vu: Many international students mention Scopus Q1, which consists of diverse journals from IEEE, Elsevier, and many other publishers unfamiliar to me.
 Honestly, I don't know/recognize many of journals listed in Scopus Q1 and.
@@
 \begin{commentbox}
-Vu: In Vietnam some letter writers ask the applicants to write their own letters so they can just sign.
+Vu: Some letter writers ask the applicants to write their own letters so they can just sign.
 As mentioned, this will hurt the applicants (admission committee are actually quite good at determining this)!
-I was told that some people did this to ensure the applicants will only go to places where they want to the students to go to (e.g., stay in Vietnam or go to specific places where they have networks).
+I was told that some people did this to ensure the applicants don't go abraod or only go to places where they want to the students to go to.
 \end{commentbox}
@@
 \begin{commentbox}
-Vu: Vietnamese students sometimes go through great length just to get letters from well-known senior professors in their school, but the letters are generic and carry little value, in fact, \red{red flags}.
-Moreover, a top professor in Vietnamese might not be well-known to US faculty (see more details in \S\ref{sec:your-school}).
+Vu: Sometimes students would go through great length just to get letters from well-known senior professors in their school, but the letters are generic and carry little value, in fact, \red{red flags}.
+Moreover, a top professor in Vietnam might not be well-known to US faculty (see more details in \S\ref{sec:your-school}).
 So save the trouble and just get letters from \emph{any} professors/supervisors who knows you well and can write a good letter about your research ability.
@@
 Other GMU faculty members also have provided feedback and contributions.
-Many Vietnamese students have provided feedback and contributed questions, which provide contents for the writing.
+Many students have provided feedback and contributed questions, which provide contents for the writing.
 \bibliographystyle{abbrv}

### 2023-06-11 phd-cs-us.tex
updates

@@
 So what can you do about this?
-several things such as asking your CS dept to put itself on CSRankings (it's the easiest way to get CS people to know about the school), explaining about your school in your statement (and asking your LOR writer to do that too), and of course, if you're Vietnamese, considering a CS PhD program that has \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{Vietnamese professors}.
+several things such as asking your CS dept to put itself on CSRankings (it's the easiest way to get CS people to know about the school), explaining about your school in your statement (and asking your LOR writer to do that too), and of course, if you're interested in working with Vietnamese professors, consider a PhD program in the US that  \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{have them}.
 
@@
 Vu: Sometime PhD admission committee will share a document such as \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Foreign-Top-Schools}{this one}, which lists the top schools in several countries.
-I have looked at Vietnamese applications (whether they are assigned to me or not) and provide inputs to the reviewers of those applications, e.g., X is the top tech school in Vietnam and so it should be \emph{top} instead of \emph{unknown} foreign, which makes a huge difference.
+I have looked at Vietnamese applications (whether they are assigned to me or not) and provide input to the reviewers of those applications, e.g., X is the top tech school in Vietnam and so it should be \emph{top} instead of \emph{unknown} foreign, which makes a huge difference.
 \end{commentbox}
@@
     11 & University of Maryland - College Park &  36& Brown University \\ 
-    12 & Northeastern University 37&  Ohio State University\\
+    12 & Northeastern University &37&  Ohio State University\\
     13 & University of Wisconsin - Madison  &38& Texas A\&M University \\

### 2023-06-11 phd-cs-us.tex
updates

@@
 TA is paid by the department for you to help with teaching.
-Finally, fellowship is an independent funding that can come the school, a company, or organization.
+Finally, fellowship is an independent funding that can come from a school, a company, or an organization.
 Tab.~\ref{tab:funding} summarizes the differences.
@@
 \paragraph{How to get TA?}  Unless you have other funding such as RA or Fellowships, TA is typically a default thing.
-When you apply to be a full-time student, you also sate that you need financial assistant.
+When you apply to be a full-time student,  state that you need financial assistant.
 It is common that the PhD committee will either admit you and give you GTA, or reject you; i.e., we do not admit a student without supporting them.
@@
     \item What can you do to increase your admission chance?
-    \begin{description}
+      \begin{itemize}[leftmargin=*]
         \item Show something that makes you \textbf{stand out}, e.g., are you a female or a minority in CS (research for this on Google)?
@@
 If you write technical, research-like blogs, talk about them too.
-    \end{description}
+    \end{itemize}
 
@@
 
-\begin{description}
+\begin{itemize}[leftmargin=*]
 \item International students not familiar with US universities often put them into \emph{two} bins:  (i) very top school that they dream about such as Stanford, MIT, Princeton, Harvard and (ii) everything else.
@@
 You can even ask to contact their students.
-\end{description}
+\end{itemize}
 
     \item Can I apply to PhD to CS if my undergrad was not in CS or any related areas?
-    \begin{description}
+    \begin{itemize}[leftmargin=*]
         \item Yes, as long as you can demonstrate you are ready for CS PhD research through research experiences, LoRs, statements, etc as mentioned.
 You might be even able leverage this to make your profile stand out.
-    \end{description}
+    \end{itemize}
 
     \item   Is an MS degree required for admission to PhD?
-    \begin{description}
+    \begin{itemize}[leftmargin=*]
     \item No.
@@
 However, MS can help if it gives research experience or is from a more well-known school than your undergrad institution.
-    \end{description}
+    \end{itemize}
 
     \item How do I address a professor Firstname Lastname?
-    \begin{description}
+    \begin{itemize}[leftmargin=*]
         \item If you don't know that professor (e.g., first email contact), then use \textbf{Prof.
@@
     \end{commentbox}
-    \end{description}
+    \end{itemize}
 \end{enumerate}

### 2023-06-11 phd-cs-us.tex
updates

@@
 \renewcommand{\tablename}{Tab.}
-\title{Demystifying the Computer Science Ph.D. Admission in the US \\{\Large A guide for Vietnamese and International Students}}
+\title{Demystifying the Computer Science Ph.D. Admission in \\US Universities\\{\Large A guide for Vietnamese and International Students}}
 \date{\today}
@@
 So I want to share my opinions and advice for those who are interested in applying for a \textbf{PhD in Computer Science in the US}.
-While I wrote this document for Vietnamese students interested in CS, it should be applicable to students in various countries interested in STEM.
-Moreover, while many examples given are for GMU, which is 32rd in the US according CSRankings (see \S\ref{sec:ranking}), the writing should be generalize to most other R1\footnote{An \href{https://en.wikipedia.org/wiki/List_of_research_universities_in_the_United_States}{R1 institution} in the US is a research-intensive university with a high level of research activity across various disciplines.} universities  (though \emph{very} top schools might be very extreme, e.g., see the admission process at \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{CMU}).
+While I wrote this document for Vietnamese students interested in CS, it should be applicable to students in various countries.
+Moreover, while many examples given are for GMU, which is 32nd in the US according CSRankings (see \S\ref{sec:ranking}), the writing should be generalize to most other R1\footnote{An \href{https://en.wikipedia.org/wiki/List_of_research_universities_in_the_United_States}{R1 institution} in the US is a research-intensive university with a high level of research activity across various disciplines.} universities  (though \emph{very} top schools is much more selective, e.g., see the admission process at \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{CMU}).
 
@@
 
-        Also do not use Mr.
-or Mrs., or just write Firstname.
-May be it is OK with others but I find it a bit disrespectful.
+ \item If you don't know the prof., do not use Mr.
+or Mrs., or Firstname.
+To me it seems a bit disrespectful.
 As you know that prof.

### 2023-06-11 phd-cs-us.tex
updates
Updates from Overleaf

@@
 \usepackage{graphicx}
-\usepackage{wrapfig}
 \usepackage{amsmath}
@@
 \usepackage{hyperref}
+\usepackage{wrapfig}
 \usepackage{xcolor}
@@
 While I wrote this document for Vietnamese students interested in CS, it should be applicable to students in various countries.
-Moreover, while many examples given are for GMU, which is 32nd in the US according CSRankings (see \S\ref{sec:ranking}), the writing should be generalize to most other R1\footnote{An \href{https://en.wikipedia.org/wiki/List_of_research_universities_in_the_United_States}{R1 institution} in the US is a research-intensive university with a high level of research activity across various disciplines.} universities  (though \emph{very} top schools is much more selective, e.g., see the admission process at \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{CMU}).
+Moreover, while many examples given are for GMU, which is 32nd in the US according CSRankings (see \S\ref{sec:ranking}), the writing should be generalize to most other R1\footnote{An \href{https://en.wikipedia.org/wiki/List_of_research_universities_in_the_United_States}{R1 institution} in the US is a research-intensive university with a high level of research activity across various disciplines.} universities  (though \emph{very} top schools can be very selective, e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU).
 
@@
 
+This document is on \href{https://github.com/nguyenthanhvuh/phd-cs-us}{Github}.
 If you have questions or comments, feel free to create a \href{https://github.com/nguyenthanhvuh/phd-cs-us/issues}{GitHub issue} for discussion.
@@
 
-However, I understand that many international students do not have the opportunities to publish in top places, so general IEEE/ACM/USENIX confs/journals would suffice.
-But be sure to talk about it in your \hyperref[sec:research-statement]{statement}.
+However, I understand that many international students do not have the opportunities to publish in top places, so general confs/journals would suffice.
+But be sure to upload your papers with your application and talk about it in your \hyperref[sec:research-statement]{statement}.
 
@@
 
-Most CS PhD applications will require at least \textbf{two LORs}. Having a letter from an internationally recognized researcher can greatly strengthen your application. However, obtaining such letters
+Most CS PhD programs will require at least \textbf{two LORs}. Having a letter from an internationally recognized researcher can greatly strengthen your application. However, obtaining such letters
 can be challenging for international students, who might not have much interactions with such experts.
@@
 \begin{commentbox}
-Vu: Some letter writers ask the applicants to write their own letters so they can just sign.
-As mentioned, this will hurt the applicants (admission committee are actually quite good at determining this)!
-I was told that some people did this to ensure the applicants don't go abraod or only go to places where they want to the students to go to.
+Vu: Some letter writers ask the applicants to write their own letters for them to sign.
+As mentioned, this will hurt the applicants (admission committee members are actually quite good at determining this)!
+I was also told that some people did this so that the applicants do not go abroad or only go to places where they want the students to go to.
 \end{commentbox}
@@
 \begin{commentbox}
-Vu: Sometimes students would go through great length just to get letters from well-known senior professors in their school, but the letters are generic and carry little value, in fact, \red{red flags}.
+Vu: Sometimes students would go through great length just to get letters from well-known professors in their school, but the letters are generic and carry little value, in fact, \red{red flags}.
 Moreover, a top professor in Vietnam might not be well-known to US faculty (see more details in \S\ref{sec:your-school}).
 So save the trouble and just get letters from \emph{any} professors/supervisors who knows you well and can write a good letter about your research ability.
-
-In short, it's better to have a good personalized
+It's better to have a good personalized
 letter about your own research ability from someone who is less
@@
 
+\begin{wrapfigure}{r}{0.5\textwidth}
+  \begin{center}
+    %\includegraphics[width=0.48\textwidth]{birds}
+  \end{center}
+  \caption{Birds}
+\end{wrapfigure}
 While you might not have control over LORs \hyperref[sec:your-school]{where your go to school}, you do over your
@@
 
-There are many guides on writing research statement, e.g.,~\cite{blattman2022writing},
-so I will not talk too much about it.
+There are many guides on writing statement, e.g.,~\cite{blattman2022writing}, and\href{https://cs-sop.org/}{example statements},  so I will not talk too much about it.
 In short, discuss about your research vision and convince us that you can achieve it through your experience, e.g., published papers, or if you work on some projects by yourself, talk about it.
@@
 
+\begin{commentbox}
+Vu: having health insurance is a must and in many cases required in US universities.
+So do not think that you're young and healthy and therefore can get around with not having insurance.
+At GMU, and most CS PhD programs at good schools, your GTA or GRA \emph{will always} come with full insurance.
+In fact, at GMU, your spouse/children will get significant discount rate for health insurance.
+So you will never have to worry much about health issues for you or your family here.
+\end{commentbox}
+
 \subsubsection{Teaching Assistant (TA)}
@@
 As a TA, you help professors with their classes (e.g., grading or teaching labs/recitation).
-Your TAship is paid through the department, i.e., they hire you to help teach.
+Your TAship is paid through the school/department, i.e., they hire you to help teach.
 During a semester, a TA might work with several courses and professors (not necessary their advisor).

### 2023-06-12 phd-cs-us.tex
Updates from Overleaf

@@
 \renewcommand{\tablename}{Tab.}
-\title{Demystifying the Computer Science Ph.D. Admission in \\US Universities\\{\Large A guide for Vietnamese and International Students}}
+\title{Demystifying the Computer Science \\Ph.D. Admission in US Universities\\{\large A guide for Vietnamese and International Students}}
 \date{\today}
@@
 
-This document is on \href{https://github.com/nguyenthanhvuh/phd-cs-us}{Github}.
+This document is available on \href{https://github.com/nguyenthanhvuh/phd-cs-us}{Github}.
 If you have questions or comments, feel free to create a \href{https://github.com/nguyenthanhvuh/phd-cs-us/issues}{GitHub issue} for discussion.
@@
 
-\begin{wrapfigure}{r}{0.5\textwidth}
-  \begin{center}
-    %\includegraphics[width=0.48\textwidth]{birds}
-  \end{center}
-  \caption{Birds}
-\end{wrapfigure}
 While you might not have control over LORs \hyperref[sec:your-school]{where your go to school}, you do over your
@@
 
-There are many guides on writing statement, e.g.,~\cite{blattman2022writing}, and\href{https://cs-sop.org/}{example statements},  so I will not talk too much about it.
+There are various guides on writing statement, e.g.,~\cite{blattman2022writing}, and \href{https://cs-sop.org/}{many example statements},  so I will not talk too much about it.
 In short, discuss about your research vision and convince us that you can achieve it through your experience, e.g., published papers, or if you work on some projects by yourself, talk about it.
 
+\begin{wrapfigure}{r}{0.47\textwidth}
+\vspace{-0.4in}
+  \begin{center}
+    \includegraphics[width=0.47\textwidth]{c2.png}
+  \end{center}
+\vspace{-0.35in}  
+  %\caption{Birds}
+\end{wrapfigure}
 Finally, this is something easy to do, but is missed by many
@@
 e.g., why do you apply here? talk about a couple of professors who you're interested in working with (in many cases your application will be forwarded to them for evaluation).
-This shows that you're serious and have done homework on where you're applying to.
+This shows that you're serious and have done homework on places you're applying to.
 Be careful not to send wrong statement to wrong school or mixing
@@
 
+   \begin{center}
+    \includegraphics[scale=0.5]{c1.png}
+   \end{center}
+
 \begin{itemize}[leftmargin=*]
 \item  International students not familiar with US universities often put them into \emph{two} bins:  (i) very top school that they dream about such as Stanford, MIT, Princeton, Harvard and (ii) everything else.
-Sometimes they apply reputations of schools for non-CS programs, e.g., medical or physics, to rank CS programs.
+Sometimes they rank CS programs using the reputations non-CS programs, e.g., medical or physics.
 In some cases they rank universities based on popular states they know in the US, e.g., California, New York, and Texas.  Let's just say there are so many thing wrong with these methods.
@@
 
-\item But more generally, rankings is superficial and you will want to do more research to be informed and make better decision.
+\item More generally, rankings can be superficial and you will want to do more research to be informed and make better decision.
 For example, if you get admissions to several places, you should contact profs.
@@
 
-    \item Can I apply to PhD to CS if my undergrad was not in CS or any related areas?
+    \item Can I apply to PhD in CS if my undergrad was not in CS or any related areas?
     \begin{itemize}[leftmargin=*]

### 2023-06-12 phd-cs-us.tex
updates

@@
 \title{Demystifying the Computer Science \\Ph.D. Admission in US Universities\\{\large A guide for Vietnamese and International Students}}
-\date{\today}
-\author{\small \href{https://nguyenthanhvuh.github.io}{ThanhVu (Vu) Nguyen}, \href{https://cs.gmu.edu}{George Mason University}}
+%\date{\today}
+\author{\small \href{https://nguyenthanhvuh.github.io}{ThanhVu (Vu) Nguyen}\\
+\href{https://cs.gmu.edu}{George Mason University}}
 %\setcounter{secnumdepth}{3}
@@
 
-\begin{enumerate}   
-    \item What can you do to increase your admission chance?
-      \begin{itemize}[leftmargin=*]
-        \item Show something that makes you \textbf{stand out}, e.g., are you a female or a minority in CS (research for this on Google)?
+\subsection{What can you do to increase your admission chance?}
+ Show something that makes you \textbf{stand out}, e.g., are you a female or a minority in CS (research for this on Google)?
 Do you participate in outreach activities that help increase diversity and inclusion in CS?
@@
 
-    \item Even if you do not have formal research experience, you can talk about your personal projects.
+Even if you do not have research experience, you can talk about your personal projects.
 For example, if you have an open-source project on Github that is used by many people, has lots of stars in Github, do talk about it.
 If you write technical, research-like blogs, talk about them too.
-    \end{itemize}
 
-   \item How to rank or select a CS program?
-
+\subsection{How to rank or select a CS program?}
    \begin{center}
@@
 
-\begin{itemize}[leftmargin=*]
-\item  International students not familiar with US universities often put them into \emph{two} bins:  (i) very top school that they dream about such as Stanford, MIT, Princeton, Harvard and (ii) everything else.
+International students not familiar with US universities often put them into \emph{two} bins:  (i) very top school that they dream about such as Stanford, MIT, Princeton, Harvard and (ii) everything else.
 Sometimes they rank CS programs using the reputations non-CS programs, e.g., medical or physics.
-In some cases they rank universities based on popular states they know in the US, e.g., California, New York, and Texas.  Let's just say there are so many thing wrong with these methods.
+In some cases they rank universities based on popular states they know in the US, e.g., California and New York.  Let's just say there are so many thing wrong with these methods.
 
-\item You can learn about CS programs and research expertise of faculty using resources such as \href{https://csrankings.org}{CSRankings.org}, which is designed specifically to help prospective PhD students in Computer Science!
+You can learn about CS programs and research expertise of faculty using resources such as \href{https://csrankings.org}{CSRankings.org}, which is designed specifically to help prospective PhD students in Computer Science!
 You will be very surprised to learn that a school X that you don't know much about has very strong research in your interested field Y.
@@
 \begin{commentbox}
-    Dat: many Vietnamese students, even those from very top engineering schools, do not know about CSRankings.  Only few applicants who worked at top research places such as VinAI might know about it.
+    Dat: many Vietnamese students, including those from very top engineering schools, do not know about CSRankings.
+May be applicants who worked at top research places such as VinAI might know about it.
 \end{commentbox}
 
-\item More generally, rankings can be superficial and you will want to do more research to be informed and make better decision.
+More generally, rankings can be superficial and you will want to do more research to be informed and make better decision.
 For example, if you get admissions to several places, you should contact profs.
@@
 You can even ask to contact their students.
-\end{itemize}
 
-    \item Can I apply to PhD in CS if my undergrad was not in CS or any related areas?
-    \begin{itemize}[leftmargin=*]
-        \item Yes, as long as you can demonstrate you are ready for CS PhD research through research experiences, LoRs, statements, etc as mentioned.
+\subsection{Can I apply to PhD in CS if my undergrad was not in CS or any related areas?}
+
+Yes, as long as you can demonstrate you are ready for CS PhD research through research experiences, LoRs, statements, etc as mentioned.
 You might be even able leverage this to make your profile stand out.
-    \end{itemize}
 
-    \item   Is an MS degree required for admission to PhD?
-    \begin{itemize}[leftmargin=*]
-    \item No.
-In fact, student with BS can get MS degree "along the way" to PhD.
-However, MS can help if it gives research experience or is from a more well-known school than your undergrad institution.
-    \end{itemize}
+\subsection{Is an MS degree required for admission to PhD?}
+No. In fact, student with BS can get MS degree "along the way" to PhD.  However, MS can help if it gives research experience or is from a more well-known school than your undergrad institution.
 
-    \item How do I address a professor Firstname Lastname?
-    \begin{itemize}[leftmargin=*]
-        \item If you don't know that professor (e.g., first email contact), then use \textbf{Prof.
+\subsection{How do I address a professor Firstname Lastname?}
+If you don't know that professor (e.g., first email contact), then use \textbf{Prof.
 Lastname} or \textbf{Dr.
@@
 
- \item If you don't know the prof., do not use Mr.
+If you don't know the prof., do not use Mr.
 or Mrs., or Firstname.
@@
 \end{commentbox}
-    \end{itemize}
-\end{enumerate}
 

### 2023-06-12 phd-cs-us.tex
Updates from Overleaf

@@
 But if you're international student outside of the US, then likely you cannot come.
-So see if you can attend it virtually and ask to meet with individual faculty if you can.
+So see if you can attend it virtually and meet with individual faculty.
 
@@
 You will be very surprised to learn that a school X that you don't know much about has very strong research in your interested field Y.
-This is also a good way to learn about individual faculty (who works on what) and the well-known CS conferences\footnote{In CS (and only in CS), conferences, not journals, are oten the main venue to publish research finding.}.
+This is also a good way to learn about individual faculty (who works on what) and the well-known CS conferences\footnote{In CS (and only in CS), conferences, not journals, are often the main venue to publish research finding.}.
+\S\ref{sec:ranking} gives the top 50 CS programs in the US according to CSRankings.
 \begin{commentbox}
-    Dat: many Vietnamese students, including those from very top engineering schools, do not know about CSRankings.
-May be applicants who worked at top research places such as VinAI might know about it.
+    Dat: many Vietnamese students, including those from top engineering schools, do not know about CSRankings.  May be applicants who worked at top research places such as VinAI would know about it.
 \end{commentbox}
@@
 
-\subsection{Current 2023 Rankings of CS PhD programs in the U.S}\label{sec:ranking}
+\section{Rankings of CS PhD programs in the U.S}\label{sec:ranking}
   Tab.~\ref{tab:ranking} lists the top 50 CS programs in the US from \href{https://www.csrankings.org}{CSRankings.org}, a ranking system  based on top CS conferences.
@@
     \small
-    \caption{Top 50 CS PhD programs in the U.S. (CSRankings, June 2023)}\label{tab:ranking}
+    \caption{Top 50 CS PhD programs in the U.S.
+(CSRankings, June 2023).
+$^*$ indicates the university has a \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{Vietnamese prof.} that can advise CS Ph.D.
+students.}\label{tab:ranking}
   \begin{tabular}{rl|rl}
@@
     1 & Carnegie Mellon & 26 & Univ. of California - Irvine \\
-    2 & Univ. of Illinois at Urbana-Champaign  & 27 &  Duke University \\
-    3 & Univ. of California-San Diego & 28 & Rutgers University \\
+    2 & Univ. of Illinois at Urbana-Champaign$^*$  & 27 &  Duke University \\
+    3 & Univ. of California-San Diego & 28 & Rutgers University$^*$ \\
     4 & MIT & 29 & Univ. of California - Riverside\\
@@
     6 & Stanford University& 31 & Pennsylvania State University  \\
-    7 & University of Michigan - Ann Arbor   & 32& George Mason University\\  
+    7 & University of Michigan - Ann Arbor   & 32& George Mason University$^*$\\  
     8 & University of Washington      &33 &  Harvard University \\
@@
     11 & University of Maryland - College Park &  36& Brown University \\ 
-    12 & Northeastern University &37&  Ohio State University\\
-    13 & University of Wisconsin - Madison  &38& Texas A\&M University \\
+    12 & Northeastern University$^*$ &37&  Ohio State University\\
+    13 & University of Wisconsin - Madison$^*$  &38& Texas A\&M University$^*$ \\
     14 & Columbia University   &39 & Boston University  \\
@@
     16 & University of Texas at Austin   &41 & University of Utah \\
-    17 & University of Pennsylvania &42 & University at Buffalo\\
+    17 & University of Pennsylvania$^*$ &42 & University at Buffalo\\
     18 & Princeton University  & 43& Rice University\\
-    19 & University of Massachusetts-Amherst & 44&  University of Colorado-Boulder \\
+    19 & University of Massachusetts-Amherst$^*$ & 44&  University of Colorado-Boulder \\
     20 &  New York University  &45& University of Illinois at Chicago  \\
-    21 & Univ. of California - Los Angeles &46& Virginia Tech  \\
-    22 & University of Southern California &47&  Arizona State University \\
-    23 & Stony Brook University &48&University of Minnesota \\
+    21 & Univ. of California - Los Angeles &46& Virginia Tech$^*$  \\
+    22 & University of Southern California &47&  Arizona State University$^*$ \\
+    23 & Stony Brook University$^*$ &48&University of Minnesota \\
     24 & University of Chicago &49& University of Virginia \\
-    25 & Univ. of California - Santa Barbara &50& University of North Carolina \\
+    25 & Univ. of California - Santa Barbara &50& University of North Carolina$^*$ \\
     \bottomrule

### 2023-06-12 phd-cs-us.tex
updates

@@
 \usepackage{cite}
-%\usepackage{caption}
-%\captionsetup{font=small}
+\usepackage{caption}
+\captionsetup{font=small}
 \usepackage{graphicx}
@@
 (CSRankings, June 2023).
-$^*$ indicates the university has a \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{Vietnamese prof.} that can advise CS Ph.D.
+\red{$^*$} indicates the university has a \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{Vietnamese prof.} that can advise CS Ph.D.
 students.}\label{tab:ranking}
@@
     1 & Carnegie Mellon & 26 & Univ. of California - Irvine \\
-    2 & Univ. of Illinois at Urbana-Champaign$^*$  & 27 &  Duke University \\
-    3 & Univ. of California-San Diego & 28 & Rutgers University$^*$ \\
+    2 & Univ. of Illinois at Urbana-Champaign\red{$^*$}  & 27 &  Duke University \\
+    3 & Univ. of California-San Diego & 28 & Rutgers University\red{$^*$} \\
     4 & MIT & 29 & Univ. of California - Riverside\\
@@
     6 & Stanford University& 31 & Pennsylvania State University  \\
-    7 & University of Michigan - Ann Arbor   & 32& George Mason University$^*$\\  
+    7 & University of Michigan - Ann Arbor\red{$^*$}   & 32& George Mason University\red{$^*$}\\  
     8 & University of Washington      &33 &  Harvard University \\
@@
     11 & University of Maryland - College Park &  36& Brown University \\ 
-    12 & Northeastern University$^*$ &37&  Ohio State University\\
-    13 & University of Wisconsin - Madison$^*$  &38& Texas A\&M University$^*$ \\
+    12 & Northeastern University\red{$^*$} &37&  Ohio State University\\
+    13 & University of Wisconsin - Madison\red{$^*$}  &38& Texas A\&M University\red{$^*$} \\
     14 & Columbia University   &39 & Boston University  \\
@@
     16 & University of Texas at Austin   &41 & University of Utah \\
-    17 & University of Pennsylvania$^*$ &42 & University at Buffalo\\
+    17 & University of Pennsylvania\red{$^*$} &42 & University at Buffalo\red{$^*$}\\
     18 & Princeton University  & 43& Rice University\\
-    19 & University of Massachusetts-Amherst$^*$ & 44&  University of Colorado-Boulder \\
+    19 & University of Massachusetts-Amherst\red{$^*$} & 44&  University of Colorado-Boulder \\
     20 &  New York University  &45& University of Illinois at Chicago  \\
-    21 & Univ. of California - Los Angeles &46& Virginia Tech$^*$  \\
-    22 & University of Southern California &47&  Arizona State University$^*$ \\
-    23 & Stony Brook University$^*$ &48&University of Minnesota \\
+    21 & Univ. of California - Los Angeles &46& Virginia Tech\red{$^*$}  \\
+    22 & University of Southern California &47&  Arizona State University\red{$^*$} \\
+    23 & Stony Brook University\red{$^*$} &48&University of Minnesota \\
     24 & University of Chicago &49& University of Virginia \\
-    25 & Univ. of California - Santa Barbara &50& University of North Carolina$^*$ \\
+    25 & Univ. of California - Santa Barbara &50& University of North Carolina\red{$^*$} \\
     \bottomrule

### 2023-06-12 phd-cs-us.tex
updates

@@
 \renewcommand{\tablename}{Tab.}
-\title{Demystifying the Computer Science \\Ph.D. Admission in US Universities\\{\large A guide for Vietnamese and International Students}}
+\title{Demystifying the Computer Science \\PhD Admission in US Universities\\{\large A guide for Vietnamese and International Students}}
 %\date{\today}
@@
 Prestigious fellowships typically require a clear and good research plan, so it is a good idea to wait until at least your second year to have research experience and even publication before applying.
-Remember, you're competing with the top Ph.D.
-students at top universities worldwide.
+Remember, you're competing with the top PhD students at top universities worldwide.
 
 \begin{commentbox}
-At GMU, Ph.D.
-applicants are automatically eligible for a Presidential Fellowship.
+At GMU, PhD applicants are automatically eligible for a Presidential Fellowship.
 It is at least as good as GTA but the most important thing is that as a fellowship it is truly free money (i.e., you are not depending on any prof.
@@
 (CSRankings, June 2023).
-\red{$^*$} indicates the university has a \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{Vietnamese prof.} that can advise CS Ph.D.
-students.}\label{tab:ranking}
+\red{$^*$} indicates the university has a \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{Vietnamese prof.} that can advise CS PhD students.}\label{tab:ranking}
   \begin{tabular}{rl|rl}
@@
 or the M.S., where you typically work as a teaching assistant and possibly continue
-to pay many tens of thousands of dollars, the Ph.D. is a time where funding is not a
+to pay many tens of thousands of dollars, the PhD is a time where funding is not a
 concern to you. At most schools, you will not pay tuition during the time that you
-are getting a Ph.D.. Typically, you will also receive a living stipend – on the order
+are getting a PhD. Typically, you will also receive a living stipend – on the order
 of $2000 per month, from which you will pay your living expenses. Ideally, your
@@
 Assistantship).
-The Ph.D. is a tremendous opportunity. You get to pick an advisor in any re-
+The PhD is a tremendous opportunity. You get to pick an advisor in any re-
 search area you like and then you get to do research in that area, receive mentoring,
@@
 will have to restrict your choice of advisors to those who have funding. At CMU,
-every Ph.D. student is guaranteed a stipend plus tuition regardless of which advisor
+every PhD student is guaranteed a stipend plus tuition regardless of which advisor
 she chooses to work with.
 Important note 2: There are many companies and government organizations
-which offer Graduate Fellowships for Ph.D. students. If you are lucky enough
+which offer Graduate Fellowships for PhD students. If you are lucky enough
 to get one of these, they will cover most of your way through graduate school,

### 2023-06-12 phd-cs-us.tex
updates

@@
    \begin{center}
-    \includegraphics[scale=0.5]{c1.png}
+    \includegraphics[scale=0.4]{c1.png}
    \end{center}
@@
 Yes, as long as you can demonstrate you are ready for CS PhD research through research experiences, LoRs, statements, etc as mentioned.
-You might be even able leverage this to make your profile stand out.
+You might be even able to leverage this to make your profile stand out.
 
@@
 
-\subsection{How do I address a professor Firstname Lastname?}
+\subsection{How do I address a professor?}
 If you don't know that professor (e.g., first email contact), then use \textbf{Prof.
@@
 (CSRankings, June 2023).
-\red{$^*$} indicates the university has a \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{Vietnamese prof.} that can advise CS PhD students.}\label{tab:ranking}
+\red{$^*$} indicates that the university has \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{Vietnamese prof.} that can advise CS PhD students.}\label{tab:ranking}
   \begin{tabular}{rl|rl}

### 2023-06-12 phd-cs-us.tex
updates

@@
 
-\subsection{Letters of Recommendation (LORs)}
+\subsection{Letter of Recommendation (LOR)}\label{sec:lor}
 
@@
 
+\begin{wrapfigure}{r}{0.45\textwidth}
+    \vspace{-0.3in}
+      \begin{center}
+        \includegraphics[width=0.45\textwidth]{c6.png}
+      \end{center}
+    \vspace{-0.4in}
+    \end{wrapfigure}
 Many students have letters written by the applicants themselves and signed by their professors. These have little
@@
 \subsection{How do I address a professor?}
+
+\begin{wrapfigure}{l}{0.47\textwidth}
+    \vspace{-0.3in}
+      \begin{center}
+        \includegraphics[width=0.47\textwidth]{c5.png}
+      \end{center}
+    \vspace{-0.6in}
+    \end{wrapfigure}
 If you don't know that professor (e.g., first email contact), then use \textbf{Prof.

### 2023-06-12 phd-cs-us.tex
updates

@@
 \subsection{What can you do to increase your admission chance?}
+
  Show something that makes you \textbf{stand out}, e.g., are you a female or a minority in CS (research for this on Google)?

### 2023-06-12 phd-cs-us.tex
updates

@@
 
+\subsection{How long does it take to complete the PhD program?}
+
+\begin{wrapfigure}{r}{0.30\textwidth}
+    \vspace{-0.2in}
+      \begin{center}
+        \includegraphics[width=0.30\textwidth]{c4a.png}
+      \end{center}
+    \vspace{-0.2in}
+    \end{wrapfigure}
+    Typically, 5-7 years for Ph.D.
+in CS at a US university.
+The first 1.5--2 years you spend on coursework and research.
+The next 2--3 years you focus on your research, forming dissertation topic, and getting results published.
+The last 1-2 years you write and defend your dissertation and look for job.
+The figure on the right is from PhDComics and shows the "ambition" level of a PhD student over their study.
+
 \subsection{How do I address a professor?}
 
-\begin{wrapfigure}{l}{0.47\textwidth}
-    \vspace{-0.3in}
+\begin{wrapfigure}{l}{0.5\textwidth}
+    \vspace{-0.4in}
       \begin{center}
-        \includegraphics[width=0.47\textwidth]{c5.png}
+        \includegraphics[width=0.5\textwidth]{c5.png}
       \end{center}
-    \vspace{-0.6in}
+    \vspace{-0.9in}
     \end{wrapfigure}

### 2023-06-12 phd-cs-us.tex
updates

@@
 \begin{abstract}
-Having been involved in PhD admissions for many years and having
-numerous interaction with  students, I've
-realized that \emph{international} students, especially \textbf{Vietnamese}, lack a clear understanding of
+Having been involved in PhD admissions for many years and frequent interactions with international students, I've
+realized that these students, especially \emph{Vietnamese}, lack a clear understanding of
 the Computer Science PhD admission process at US universities. This confusion not only
@@
 So I want to share my opinions and advice for those who are interested in applying for a \textbf{PhD in Computer Science in the US}.
-While I wrote this document for Vietnamese students interested in CS, it should be applicable to students in various countries.
-Moreover, while many examples given are for GMU, which is 32nd in the US according CSRankings (see \S\ref{sec:ranking}), the writing should be generalize to most other R1\footnote{An \href{https://en.wikipedia.org/wiki/List_of_research_universities_in_the_United_States}{R1 institution} in the US is a research-intensive university with a high level of research activity across various disciplines.} universities  (though \emph{very} top schools can be very selective, e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU).
+While I wrote this document for Vietnamese students interested in CS, it should be applicable to students in other countries.
+Moreover, while many examples given are for GMU, which is 32nd in the US according CSRankings (see \S\ref{sec:ranking}), the writing should be generalize to most other R1\footnote{An \href{https://en.wikipedia.org/wiki/List_of_research_universities_in_the_United_States}{R1 institution} in the US is a research-intensive university with a high level of research activity across various disciplines.} universities  (though \emph{very} top schools could be very selective, e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU).
 
@@
 health insurance, and stipend. Moreover, depending on the university,
-you may even receive additional benefits such as summer salary and laptops. Note that
-funding is typically more available for PhD students than 
-Masters and especially undergraduate studies, which typically have no funding. \S\ref{sec:funding} provides more details on funding.
+you may even receive additional benefits such as summer salary and laptops. \S\ref{sec:funding} provides more details on funding.
 
@@
 Vu: One of the reasons I create this post is that several of my colleagues are interested in 
-recruiting Vietnamese students and were surprised when seeing very few applications in Vietnam (e.g., our CS PhD program has on average 5-6 applications from Vietnam each year).
+recruiting Vietnamese students and were surprised when seeing very few applications in Vietnam (e.g., our CS PhD program has on average 5--6 applications from Vietnam each year).
 In general the number of
@@
 Each application is assigned to about \emph{three} faculty members, who will evaluate your profile and try to reach a consensus about your case.
-Note that while in most cases the assigned reviewers will be the main ones deciding your application, every faculty will have access to your application and can give inputs on your profile.
+Note that while the assigned reviewers are likely the main ones deciding your application, \emph{every} faculty in the department will have access to your application and can give inputs on your profile.
 
-In many cases, the admission committee can involve current students and assistant professors in the department.
+In many cases, the admission committee involves assistant professors in the department.
 This provides junior faculty the opportunities to recruit students.
@@
 Finally, fellowship is an independent funding that can come from a school, a company, or an organization.
-Tab.~\ref{tab:funding} summarizes the differences.
+Note that funding is typically more available for PhD students than
+Masters and especially undergraduate studies, which typically have no funding.  Tab.~\ref{tab:funding} summarizes the differences.
 
@@
 
-\subsection{Can I apply to PhD in CS if my undergrad was not in CS or any related areas?}
+\subsection{Tenured or tenure-track faculty?} 
+The short answer is that tenured-track faculty (e.g., assistant professors) are more likely to be young and very active in research (they have to, in order to get tenure).
+Thus, they will likely have more time to work with you and push you to do research and publish.
+The downside is that they may not have as much experience in managing students and may not have as much funding as tenured faculty.
+Tenured faculty (e.g., associate and full profs.) are more likely to be older, more well-known, and have more experience in managing students.
+However, they might not push you as hard (they don't have to, they already got tenured) and expect you to figure things out yourself (so you need to be very independent).
+Some tenured faculty are also no longer active in research, e.g., they are more involved in administrative duties.
+You can learn about faculty and their level of research activity through the faculty's website and CSRankings.
+
+\subsection{Can I apply to CS PhD if my undergrad was not in CS or related areas?}
 
@@
 
-\begin{wrapfigure}{r}{0.30\textwidth}
-    \vspace{-0.2in}
+\begin{wrapfigure}{r}{0.28\textwidth}
+    \vspace{-0.4in}
       \begin{center}
-        \includegraphics[width=0.30\textwidth]{c4a.png}
+        \includegraphics[width=0.28\textwidth]{c4a.png}
       \end{center}
@@
     \end{wrapfigure}
-    Typically, 5-7 years for Ph.D.
-in CS at a US university.
+    Typically, 5-7 years for PhD in CS at a US university.
 The first 1.5--2 years you spend on coursework and research.
@@
 The last 1-2 years you write and defend your dissertation and look for job.
-The figure on the right is from PhDComics and shows the "ambition" level of a PhD student over their study.
+The PhDComics figure on the right shows the "ambition" level of a PhD student over their years of study.
 

### 2023-06-12 phd-cs-us.tex
Updates from Overleaf

@@
 Having been involved in PhD admissions for many years and frequent interactions with international students, I've
-realized that these students, especially \emph{Vietnamese}, lack a clear understanding of
+realized that most of these students, especially \emph{Vietnamese}, lack a clear understanding of
 the Computer Science PhD admission process at US universities. This confusion not only
@@
 While I wrote this document for Vietnamese students interested in CS, it should be applicable to students in other countries.
-Moreover, while many examples given are for GMU, which is 32nd in the US according CSRankings (see \S\ref{sec:ranking}), the writing should be generalize to most other R1\footnote{An \href{https://en.wikipedia.org/wiki/List_of_research_universities_in_the_United_States}{R1 institution} in the US is a research-intensive university with a high level of research activity across various disciplines.} universities  (though \emph{very} top schools could be very selective, e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU).
+Moreover, while many examples given are for George Mason University, the writing should be generalize to most other R1\footnote{An \href{https://en.wikipedia.org/wiki/List_of_research_universities_in_the_United_States}{R1 institution} in the US is a research-intensive university with a high level of research activity across various disciplines.} universities  (though \emph{very} top schools could be very selective, e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU).
 
@@
 \begin{commentbox}
-Vu: One of the reasons I create this post is that several of my colleagues are interested in 
-recruiting Vietnamese students and were surprised when seeing very few applications in Vietnam (e.g., our CS PhD program has on average 5--6 applications from Vietnam each year).
+Vu: One of the reasons I create this post is that my colleagues at GMU are interested in 
+recruiting Vietnamese students and were surprised when seeing very few applications in Vietnam (e.g., each year our CS program receives more than 350 PhD applications, most of which are international but only 3--4 are from Vietnam).
 In general the number of
@@
 Then your application will be reviewed by a
-\textbf{PhD admission committee} consisting of faculty members in CS.
+\textbf{PhD admission committee} that consists of faculty members in CS.
 Each application is assigned to about \emph{three} faculty members, who will evaluate your profile and try to reach a consensus about your case.
@@
 Vu: We usually decide that a PhD candidate is either (i) admit with funding (TA or RA) or (ii) rejected. In other words, in most cases, we either
-admit you with full funding, or we don't. In some rare cases, we may admit
-you without funding because you have funding on your own (e.g., you are
-supported by your government or have external grants). We also justify
+admit you with full funding, or we don't. In some rare cases, we admit
+without funding because you have funding on your own (e.g.,
+supported by your government or having external grants). We also justify
 our decision with a summary about your application, where we list
@@
 \begin{commentbox}
-    Dat: many Vietnamese students, including those from top engineering schools, do not know about CSRankings.  May be applicants who worked at top research places such as VinAI would know about it.
+    Dat: MOst Vietnamese students, including those from top schools, do not know about CSRankings.  May be applicants who worked at top research places such as VinAI would know about it.
 \end{commentbox}
@@
 
-\subsection{Tenured or tenure-track faculty?} 
+\subsection{Tenured or tenure-track faculty? Who do you choose?} 
 The short answer is that tenured-track faculty (e.g., assistant professors) are more likely to be young and very active in research (they have to, in order to get tenure).
@@
 Some tenured faculty are also no longer active in research, e.g., they are more involved in administrative duties.
+
 You can learn about faculty and their level of research activity through the faculty's website and CSRankings.
+Ultimately, choose one that fits you the most by communicating with them, meeting them, asking them questions, even talking to their students.
 
@@
     \end{wrapfigure}
-    Typically, 5-7 years for PhD in CS at a US university.
-The first 1.5--2 years you spend on coursework and research.
-The next 2--3 years you focus on your research, forming dissertation topic, and getting results published.
+    Typically, 5-7 years for PhD in CS at US universities.
+The first 1.5--2 years you spend on coursework and learning research.
+The next 2--3 years you focus on your research, formi dissertation topic, and get results published.
 The last 1-2 years you write and defend your dissertation and look for job.
-The PhDComics figure on the right shows the "ambition" level of a PhD student over their years of study.
+The PhDComics figure on the right shows the "ambition" level of a PhD student over their years of study (they miss the 6-7th Year where the ambition is now "Just get me the ...
+out of here").
 
@@
 
-  \begin{table}[h!]
+  \begin{table}
     \centering

### 2023-06-12 phd-cs-us.tex
updates

@@
 \usepackage{enumitem}
+\usepackage{epigraph}
 \usepackage[T1]{fontenc}
@@
 \section{Should You Apply?}
+\epigraph{Don't make fun of graduate students. They just made a terrible life choice.}{The Simpsons}
 
@@
 
-\section{Rankings of CS PhD programs in the U.S}\label{sec:ranking}
+\section{How much do \emph{you} cost?}
+PhD students often wonder how their salary is so low compared to their advisor's ludicrous grants or that why their offer letters sometime mentioned their benefits worth way more than what they actually receive (i.e., stipend).
+This section aims to shed some light to these questions by breaking down how much a GRA costs to their advisor.
+
+Tab.~\ref{tab:cost} shows the budget breakdown for a GRA per year.
+These numbers are based on my experience at public universities in the US.
+Private universities may have different numbers.
+For simplicity, I will assume the department has a 9-month stipend of \$27000 (GMU actually pays more) and therefore a 3-month summer of \$9000.
+I will also use GMU tuition rate of about \$15000/year for full-time study and a 58.9\% \emph{indirect cost}, which is what GMU charges for administrative costs (yes, after all, universities are businesses!).
+Finally, I assume the student will make 2 conference trips per year, one domestic and one international (so conf.
+registration, airline tickets, taxi, meals, etc are all included).
+At the end, the total budget comes out to be \$69K.
+\emph{The summary is that while you're paid X, your advisor probably pays 2X for you}.
+
+\begin{table}
+    \centering
+    \small
+    \caption{GRA cost breakdown. F \& A is Facilities \& Administrative Cost Base and 
+    MTDC is Modified Total Direct Cost.}\label{tab:cost}
+    \begin{tabular}{rcl}
+        \textbf{Budget} & \textbf{Cost} & \textbf{Notes} \\
+        \midrule
+        GRA (9-month) & 27K & \\
+        GRA (summer)  &9K	& 3-month, 20hrs/week\\
+        \textbf{Total Salary} &\red{36K}	&\\
+        \midrule
+        Health Insurance	&3K	& full year\\
+        Tuition (In-State) &	15K	& (\$680/ Credit + \$150/Student Fee/ Credit)* 9 credits = \\
+        &&\$7470 (\$6120 + \$1350) per semester\\
+        \textbf{Total Tuition \& Insurance}	&\red{18K}	&Full year tuition + insurance\\
+        \midrule
+        Conference Registration	& 500 & \\
+        International Travel &	1800& \\
+        Domestic Travel	& 700	& \\
+        \textbf{Total Travel}&	\red{3K}	\\
+        \midrule
+     Total Direct Cost	& \red{57K}	&Salary + Travel + Health + Tuition \\
+     F \& A (MTDC)	& 21K	& Direct Cost - GRA Salary\\
+    Total Indirect Cost	& \red{12K}	&58.9\% of MTDC\\
+    \textbf{Total (Direct + Indirect)} &	\red{69K}	& Budget for a GRA\\
+        \bottomrule
+    \end{tabular}
+\end{table}
+
+\section{Ranking of CS PhD programs in the U.S}\label{sec:ranking}
   Tab.~\ref{tab:ranking} lists the top 50 CS programs in the US from \href{https://www.csrankings.org}{CSRankings.org}, a ranking system  based on top CS conferences.

### 2023-06-12 phd-cs-us.tex
updates

@@
 However, I understand that many international students do not have the opportunities to publish in top places, so general confs/journals would suffice.
-But be sure to upload your papers with your application and talk about it in your \hyperref[sec:research-statement]{statement}.
+But be sure to upload your papers with your application and talk about them in your \hyperref[sec:research-statement]{statement}.
 
@@
 
-There are various guides on writing statement, e.g.,~\cite{blattman2022writing}, and \href{https://cs-sop.org/}{many example statements},  so I will not talk too much about it.
+There are various guides on writing statement, e.g.,~\cite{blattman2022writing}, and \href{https://cs-sop.org/}{many example statements} are available.
+So I will not talk too much about statements.
 In short, discuss about your research vision and convince us that you can achieve it through your experience, e.g., published papers, or if you work on some projects by yourself, talk about it.
@@
 RA is provided through a professor through their own funding so you can work on their project.  
-You do not need to teach as an RA, so you can focus on your research. Depending on the professor, RA may be available during the summer.
+You do not need to teach as an RA, so you can focus on your research. Depending on the professor, RA may be available during the summer. \S\ref{sec:ra-cost} gives more details on RA budget.
 
@@
 \begin{commentbox}
-    Dat: MOst Vietnamese students, including those from top schools, do not know about CSRankings.  May be applicants who worked at top research places such as VinAI would know about it.
+    Dat: Most Vietnamese students, including those from top schools, do not know about CSRankings.  May be applicants who worked at top research places such as VinAI would know about it.
 \end{commentbox}
@@
 
-\section{How much do \emph{you} cost?}
-PhD students often wonder how their salary is so low compared to their advisor's ludicrous grants or that why their offer letters sometime mentioned their benefits worth way more than what they actually receive (i.e., stipend).
-This section aims to shed some light to these questions by breaking down how much a GRA costs to their advisor.
+\section{How much do \emph{you} cost?}\label{sec:ra-cost}
+PhD students often wonder how their salary is so low compared to ludicrous grants their advisors get or that why their offer letters sometime mentioned that their benefits worth way more than what they actually receive (i.e., stipend).
+This section aims to shed some light to these questions.
 
-Tab.~\ref{tab:cost} shows the budget breakdown for a GRA per year.
+Tab.~\ref{tab:cost} shows the budget breakdown for a GRA per year (this level of details is what faculty actually uses when applying for funding).
 These numbers are based on my experience at public universities in the US.
@@
     \caption{GRA cost breakdown. F \& A is Facilities \& Administrative Cost Base and 
-    MTDC is Modified Total Direct Cost.}\label{tab:cost}
+    MTDC is Modified Total Direct Cost. These are things that the university can charge overhead to.}\label{tab:cost}
     \begin{tabular}{rcl}
-        \textbf{Budget} & \textbf{Cost} & \textbf{Notes} \\
+        \textbf{Budget} & \textbf{Cost \$} & \textbf{Notes} \\
         \midrule

### 2023-06-12 phd-cs-us.tex
updates

@@
 So I want to share my opinions and advice for those who are interested in applying for a \textbf{PhD in Computer Science in the US}.
-While I wrote this document for Vietnamese students interested in CS, it should be applicable to students in other countries.
-Moreover, while many examples given are for George Mason University, the writing should be generalize to most other R1\footnote{An \href{https://en.wikipedia.org/wiki/List_of_research_universities_in_the_United_States}{R1 institution} in the US is a research-intensive university with a high level of research activity across various disciplines.} universities  (though \emph{very} top schools could be very selective, e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU).
+While this document is written for Vietnamese students interested in CS, it should applies to students in other countries as well.
+Moreover, while many examples given are for George Mason University, the writing should generalize to other R1\footnote{An \href{https://en.wikipedia.org/wiki/List_of_research_universities_in_the_United_States}{R1 institution} in the US is a research-intensive university with a high level of research activity across various disciplines.} universities  (though \emph{very} top schools could be very selective, e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU).
 
@@
 
-Around March you should hear back from most PhD programs that you applied (if not, send email and ask). You will have to make your decision by around April 15.
-If you have offers, congratulations!  Now you're at a different game because the schools that admit you will now try to get you to accept them!
+Around March you should hear back from most PhD programs that you applied to (if not, send email and ask).
+If you have offers, congratulations!
+Now you're at a different game because the schools that admit you will now try to get you to accept them!
+You will likely have to make your decision by around April 15.
 
@@
 \begin{commentbox}
-Vu: GMU has Virtual Open House, e.g., \url{https://cs-gmu.github.io/cs-phd-voh-s23/}, which I've co-organized in the last two years.
+Vu: GMU has \emph{Virtual} Open House (VOH), e.g., \url{https://cs-gmu.github.io/cs-phd-voh-s23/}, which I've co-organized in the last two years.
 We invite all admitted PhD students to the VOH through Zoom to learn about the CS program, the department, GMU, and the DC area in general.
@@
 You can learn about CS programs and research expertise of faculty using resources such as \href{https://csrankings.org}{CSRankings.org}, which is designed specifically to help prospective PhD students in Computer Science!
-You will be very surprised to learn that a school X that you don't know much about has very strong research in your interested field Y.
-This is also a good way to learn about individual faculty (who works on what) and the well-known CS conferences\footnote{In CS (and only in CS), conferences, not journals, are often the main venue to publish research finding.}.
+You will be very surprised to learn that a school X that you didn't know much about has very strong research in your interested field Y.
+This is also a good way to learn about individual faculty (who works on what) and well-known CS conferences\footnote{In CS (and only in CS), conferences, not journals, are often the main venue to publish research finding.}.
 \S\ref{sec:ranking} gives the top 50 CS programs in the US according to CSRankings.

### 2023-06-14 phd-cs-us.tex
Updates from Overleaf

@@
 
-\newtcolorbox{cbox}
+\newtcolorbox{cbox}[1][]
 {
@@
 
- \newenvironment{commentbox}{
+\newenvironment{commentbox}[1][]{
 \small
     \begin{cbox}
+    \textbf{#1}: 
  }{
@@
 
-\newcommand{\mycomment}[3][\color{black}]{{#1{{#2}: {#3}}}}
+\newcommand{\mycomment}[3][\color{blue}]{{#1{{#2}: {#3}}}}
 \newcommand{\tvn}[1]{\mycomment{TVN}{#1}}{}
+\newcommand{\didi}[1]{\mycomment{Didier}{#1}}{}
 \newcommand{\tl}[1]{\mycomment{ThanhLe}{#1}}{}
@@
 \begin{abstract}
-Having been involved in PhD admissions for many years and frequent interactions with international students, I've
-realized that most of these students, especially \emph{Vietnamese}, lack a clear understanding of
+Having been involved in PhD admissions for many years, I've
+realized that many international students, especially \emph{Vietnamese}, lack a clear understanding of
 the Computer Science PhD admission process at US universities. This confusion not only
@@
 
-So I want to share my opinions and advice for those who are interested in applying for a \textbf{PhD in Computer Science in the US}.
-While this document is written for Vietnamese students interested in CS, it should applies to students in other countries as well.
+So I want to reveal some details about the PhD Admission process and share my opinions and advice for those who are interested in applying for a \textbf{PhD in Computer Science in the US}.
+While this document is written for Vietnamese students interested in Computer Science, it should apply to students in many other countries and other majors.
 Moreover, while many examples given are for George Mason University, the writing should generalize to other R1\footnote{An \href{https://en.wikipedia.org/wiki/List_of_research_universities_in_the_United_States}{R1 institution} in the US is a research-intensive university with a high level of research activity across various disciplines.} universities  (though \emph{very} top schools could be very selective, e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU).
+
+In addition, this document can help \emph{US faculty} to understand more about international students, e.g., questions, concerns, and various cultural differences.
+By understanding, appreciating, and even leveraging these information, CS programs in the US can attract larger and stronger application pools from Vietnam and other countries.
 
@@
 \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-GMU}{why
-you want to study at GMU}). Happy hunting!
+you want to study at GMU}). Happy school hunting!
 
@@
 health insurance, and stipend. Moreover, depending on the university,
-you may even receive additional benefits such as summer salary and laptops. \S\ref{sec:funding} provides more details on funding.
+you may even receive additional benefits such as summer salary, laptops, (conference/workshop) traveling. \S\ref{sec:funding} provides more details on funding.
 
@@
 
-\begin{commentbox}
-Vu: One of the reasons I create this post is that my colleagues at GMU are interested in 
-recruiting Vietnamese students and were surprised when seeing very few applications in Vietnam (e.g., each year our CS program receives more than 350 PhD applications, most of which are international but only 3--4 are from Vietnam).
+\begin{commentbox}[Vu]
+One of the reasons I create this document is that my (non-Vietnamese) colleagues at GMU are interested in 
+recruiting Vietnamese students and are surprised when seeing very few applications in Vietnam (e.g., each year our CS program receives more than 350 PhD applications, most of which are international but only 3--4 are from Vietnam).
 In general the number of
@@
 
-\begin{commentbox}
-Vu: We usually decide that a PhD candidate is either (i) admit with funding (TA or RA) or (ii) rejected. In other words, in most cases, we either
+\begin{commentbox}[Vu] 
+We usually decide that a PhD candidate is either (i) admit with funding (TA or RA) or (ii) rejected. In other words, in most cases, we either
 admit you with full funding, or we don't. In some rare cases, we admit
@@
 
-\begin{commentbox}
-    Hakan: At GMU, for full consideration, students should make sure to submit \textbf{ALL} required documents by the application deadline, and should never assume that some required documents (such as official TOEFL scores or official diplomas/transcripts) will be waived by the admissions office.
+%\didi{Is there more information on typical strengths and common weaknesses of applications.
+This is especially useful to sophomore and junior students as they still have time to work on those strengths.}
+% tvn: the main thing is research experiences
+
+\begin{commentbox}[Hakan]
+At GMU, for full consideration, students should make sure to submit \textbf{ALL} required documents by the application deadline, and should never assume that some required documents (such as official TOEFL scores or official diplomas/transcripts) will be waived by the admissions office.
 If something is listed and not marked as "optional", it is mandatory and they should plan for submitting all those.
@@
 
-The most effective evidence of research ability is having \textbf{published papers in reputable international conferences or journals}.
-Having published good papers is a sign that the applicant was involved in research.
+The most effective evidence of research ability is having \textbf{published papers in reputable international journals or conferences}.
+Having published good papers is a sign that the applicant has successfully involved in research.
 
-It's important to aim for the top conferences in your field, which you can
-find on places such as CSRankings~\cite{csrankings}, designed specifically to help CS PhD
-applicants. Local conferences and non-English journals or conferences do
-not carry as much weight since their quality is often unknown to US
-faculty.
+It's important to aim for the top venues in your field, which you can
+find on places such as CSRankings~\cite{csrankings}, which ranks CS programs based on how their faculty publish at top \emph{conferences}. Local conferences and non-English journals or conferences do
+not carry as much weight since their quality is often unknown to US faculty.
 
@@
 
-\begin{commentbox}
-Vu: Many international students mention Scopus Q1, which consists of diverse journals from IEEE, Elsevier, and many other publishers unfamiliar to me.
-Honestly, I don't know/recognize many of journals listed in Scopus Q1 and.
-So this might be something to be mindful of, as \textbf{CS} faculty might not be too familiar with Scopus or journals listed in there, so devote sometime in your statement to discuss the significance of your papers.
+\begin{commentbox}[Vu]
+Many international students mention Scopus Q1, which consists of various journals from IEEE, Elsevier, and many other publishers.
+I don't know/recognize many of journals listed in Scopus Q1 and.
+This might be something to be mindful of, as \textbf{CS} faculty might not be too familiar with Scopus or journals listed in there, so devote sometime in your statement to discuss the significance of your papers.
 \end{commentbox}
 
-\begin{commentbox}
-Craig: GMU and many other universities allow you to upload your published papers and other writing samples.
+\begin{commentbox}[Craig]
+GMU and many other universities allow you to upload your published papers and other writing samples.
 In many cases, even if the papers were not published at top places, we can still determine their quality by simply skimming over the paper.
@@
 application.
-Unfortunately, many good international research places, e.g., VinAI in Vietnam, remain relatively unknown to most universities in the US.
+Unfortunately, many good research places in your countries, e.g., VinAI in Vietnam, remain relatively unknown to most universities in the US.
 So you should explicit say something about them in your statement.
@@
 For example, participating in Math Olympiads if you want to do theory or  winning ACM programming contests if you want to ``build'' stuff, e.g., software analysis.
+
+\begin{commentbox}[Thanh]
+Due to academic culture, professors in Vietnam usually aim for (international) journals instead of conferences.
+Could you give some tips on how to know whether a journal is good (CSrankings, unfortunately, only consider conferences)?.
+\tcblower
+Vu: One way is looking at what well-known researchers publish at.
+For example, if you are interested in a field X, you can use CSRankings to look at active faculty in X, and then look at their websites to see what journals they publish at.
+\end{commentbox}
 
@@
 
-\begin{commentbox}
-Vu: Some letter writers ask the applicants to write their own letters for them to sign.
+\begin{commentbox}[Vu]
+Some letter writers ask the applicants to write their own letters for them to sign.
 As mentioned, this will hurt the applicants (admission committee members are actually quite good at determining this)!
 I was also told that some people did this so that the applicants do not go abroad or only go to places where they want the students to go to.
-\end{commentbox}
-
-\begin{commentbox}
-Vu: Sometimes students would go through great length just to get letters from well-known professors in their school, but the letters are generic and carry little value, in fact, \red{red flags}.
+\tcblower
+Sometimes students would go through great length just to get letters from well-known professors in their school, but the letters are generic and carry little value, in fact, \red{red flags}.
 Moreover, a top professor in Vietnam might not be well-known to US faculty (see more details in \S\ref{sec:your-school}).
@@
 well-known than a generic/weak letter from a well-known person.
+\end{commentbox}
+
+\begin{commentbox}[Didier]
+\emph{Should letter writers have PhDs?}  In Rwanda, a lot of students interact more with teaching faculty who might not have PhD.
+\tcblower
+\textbf{Vu}: In general, at least one writer should have PhD to properly evaluate your research ability.
+However, I think this is an interesting and useful detail that US faculty might not be aware of and students should mention about this in their statements.
 \end{commentbox}
@@
 In short, discuss about your research vision and convince us that you can achieve it through your experience, e.g., published papers, or if you work on some projects by yourself, talk about it.
+Also, use the statement to talk about stuff that admission committee members might not know about, e.g., your Github project with 1K+ stars or  your regular contributions to well-known open-source projects.
 
-\begin{wrapfigure}{r}{0.47\textwidth}
+\begin{wrapfigure}{r}{0.50\textwidth}
 \vspace{-0.4in}
   \begin{center}
-    \includegraphics[width=0.47\textwidth]{c2.png}
+    \includegraphics[width=0.50\textwidth]{c2.png}
   \end{center}
-\vspace{-0.35in}  
-  %\caption{Birds}
+\vspace{-0.4in}  
 \end{wrapfigure}
@@
 This shows that you're serious and have done homework on places you're applying to.
-Be careful not to send wrong statement to wrong school or mixing
+Be careful not to send statements to wrong schools or mixing
 facts (e.g., talking about school X but mentioned about working with
@@
 
-\begin{commentbox}
-\small
-Vu: I always read the research statement first and then LORs. If I am
+\begin{commentbox}[Vu]
+I always read the research statement first and then LORs. If I am
 persuaded by then, I would skim over other factors and advocate for
@@
 
-\begin{commentbox}
-Vu: Sometime PhD admission committee will share a document such as \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Foreign-Top-Schools}{this one}, which lists the top schools in several countries.
-I have looked at Vietnamese applications (whether they are assigned to me or not) and provide input to the reviewers of those applications, e.g., X is the top tech school in Vietnam and so it should be \emph{top} instead of \emph{unknown} foreign, which makes a huge difference.
+\begin{commentbox}[Vu]
+Sometime PhD admission committee will share a document such as \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Foreign-Top-Schools}{this one}, which lists the top schools in several countries.
+Personally I have looked at Vietnamese applications (whether they are assigned to me or not) and provide input to their reviewers, e.g., X is the top tech school in Vietnam and so it should be \emph{top} instead of \emph{unknown} foreign, which makes a huge difference.
 \end{commentbox}
@@
 \subsection{Grades/GREs}\label{sec:grades}
-Having good grades is important, but, in general, unless your school is well-known, having top grades/ranks
+
+Having good grades is important, but unless your school is well-known, having top grades/ranks
 usually will not help. This is simply because we cannot evaluate them.
@@
 
+\begin{commentbox}[Thanh]
+Vietnamese universities typically offer several specialized programs, such as the talented engineer program at HUST, that have highly competitive entrance exams and a limited number of available slots (e.g., 30 per year).
+However, these programs may present several challenges.
+Professors within these programs often set higher requirements for students, including more demanding tests and assignments, resulting in lower GPAs and overall rankings.
+For example, an 3.5 GPA students from such talented programs are easily much better than a 4.0 GPA students not in those programs.
+Similarly, variations in GPA standards exist among different universities, with technical universities generally having lower GPAs than economical universities.
+All of these make gaining admission in the US difficult as US faculty are not familiar with these issues.
+I wonder if you have and advice and tips for the situations?
+\tcblower
+\textbf{Vu} Many Vietnamese students and faculty lament how this competitive/low-grading from top Vietnamese universities hurts its own students in applying abroad.
+One way to mitigate this is making these issues known to admission committee.
+Schools with Vietnamese profs are probably aware of them, but in general your letter writers/you can explicitly mention these in their letters/ your statement.
+\end{commentbox}
+
 \paragraph{GRE} Most CS programs in the US \emph{no longer require GREs}, so you don't need to
@@
 
-\begin{commentbox}
-Vu: The minimum for GMU (being above this might not mean much, but below is a \red{red flag}).
+\begin{commentbox}[Vu]
+The minimum for GMU (being above this might not mean much, but below is a \red{red flag}).
 \begin{itemize}
@@
 
-\begin{commentbox}
-Vu: At GMU, we are encouraged to interview candidates.
+\begin{commentbox}[Vu]
+At GMU, we are encouraged to interview candidates.
 For very strong candidates, the interview is actually to recruit them.
@@
 
-\begin{commentbox}
-Vu: GMU has \emph{Virtual} Open House (VOH), e.g., \url{https://cs-gmu.github.io/cs-phd-voh-s23/}, which I've co-organized in the last two years.
+\begin{commentbox}[Vu]
+GMU has \emph{Virtual} Open House (VOH), e.g., \url{https://cs-gmu.github.io/cs-phd-voh-s23/}, which I've co-organized in the last two years.
 We invite all admitted PhD students to the VOH through Zoom to learn about the CS program, the department, GMU, and the DC area in general.
@@
 
-\begin{commentbox}
-Vu: TA and RA at GMU have similar benefits in tuition waiving and insurance.
+\begin{commentbox}[Vu]
+TA and RA at GMU have similar benefits in tuition waiving and insurance.
 For stipend, depending on the college and department, a 9-month graduate assistant stipend is set.
 TA and RA will usually be that amount (TA will definitely be that, RA might fluctuate a bit depending on the stage of the student and the prof.).
-\end{commentbox}
-
-\begin{commentbox}
-Vu: having health insurance is a must and in many cases required in US universities.
-So do not think that you're young and healthy and therefore can get around with not having insurance.
-At GMU, and most CS PhD programs at good schools, your GTA or GRA \emph{will always} come with full insurance.
-In fact, at GMU, your spouse/children will get significant discount rate for health insurance.
+\tcblower
+Having health insurance is required at many US universities.
+So do not think that you're young and healthy and ignore insurance.
+At GMU, and at most good CS PhD programs, your GTA or GRA \emph{will always} come with full insurance.
+In fact, at GMU your spouse/children will get significant discount rate for health insurance.
 So you will never have to worry much about health issues for you or your family here.
@@
 
-\begin{commentbox}
+\begin{commentbox}[Vu]
 At GMU CS, students admitted with TA have  4 years of GTA guaranteed and also receive  stipend for the \textbf{first} summer.
@@
 
-\begin{commentbox}
-Vu: If you're lucky and got recruited by a prof.
+\begin{commentbox}[Vu]
+If you got recruited by a prof.
 who would give you RA right away, it's very likely you will get admitted.
@@
 
-\begin{commentbox}
+\begin{commentbox}[Vu]
 At GMU, PhD applicants are automatically eligible for a Presidential Fellowship.
@@
 \S\ref{sec:ranking} gives the top 50 CS programs in the US according to CSRankings.
-\begin{commentbox}
-    Dat: Most Vietnamese students, including those from top schools, do not know about CSRankings.  May be applicants who worked at top research places such as VinAI would know about it.
+\begin{commentbox}[Dat] Most Vietnamese students, including those from top schools, do not know about CSRankings.
+May be applicants who worked at top research places such as VinAI would know about it.
 \end{commentbox}
@@
 \subsection{Tenured or tenure-track faculty? Who do you choose?}
+
 The short answer is that tenured-track faculty (e.g., assistant professors) are more likely to be young and very active in research (they have to, in order to get tenure).
@@
 Ultimately, choose one that fits you the most by communicating with them, meeting them, asking them questions, even talking to their students.
+
+% \subsubsection{Strong Faculty vs High-ranked School}
+% \begin{commentbox}[Thanh]
+% When considering PhD programs, we often wonder if we should prioritize a high-ranking university or a professor with a strong reputation?
+Of course, both are great, but it is hard to achieve both.
+I believe receiving some guidance from you  would be incredibly valuable.
+% \end{commentbox}
+
+% This is a good question and there doesn't seem to be any best answer or specific algorithm that you can use.
+But I'll throw out something for you to think about.
+To me, there are many good schools in the US and so you have some leeway in choosing.
+For example, I don't believe there's much difference among schools in top 5 (e.g., CMU, UIUC, UCSD, MIT, GTech are in the same equivalent class), top 10 (e.g., Michigan,  Stanford, UWash, UC-Berkeley, Cornell is another equivalent class), top 20, top 30, top 40, and so on (though I would say schools above 100 are pretty much in the same equivalent class).
+It's hard to say a top X is much much better than a top X+20 (e.g.,  Purdue at 15 is stronger than Yale at 35, but not \emph{that much} stronger as you would think).
+So keep that in mind, while ranking does matter (e.g., more active faculty, stronger students), the lower ranked universities are still pretty good, might have sub fields that are extremely strong (NCState is 40th, but in Software Engineering it is easily top 5), and many unique things.
+This is in contrast with universities in Vietnam where there are large gaps among the top 5, 10, and so on.
 
@@
 The last 1-2 years you write and defend your dissertation and look for job.
-The PhDComics figure on the right shows the "ambition" level of a PhD student over their years of study (they miss the 6-7th Year where the ambition is now "Just get me the ...
-out of here").
+The PhDComics figure on the right shows the "ambition" level of a PhD student over their years of study (they miss the 6-7th Year where the ambition is "Please just get me graduate").
 
@@
 \section{How much do \emph{you} cost?}\label{sec:ra-cost}
-PhD students often wonder how their salary is so low compared to ludicrous grants their advisors get or that why their offer letters sometime mentioned that their benefits worth way more than what they actually receive (i.e., stipend).
+PhD students often wonder why their salary is so low compared to ludicrous grants their advisors get or why their offer letters sometime mentioned that their benefits worth way more than what they actually receive (i.e., stipend).
 This section aims to shed some light to these questions.
@@
 For simplicity, I will assume the department has a 9-month stipend of \$27000 (GMU actually pays more) and therefore a 3-month summer of \$9000.
-I will also use GMU tuition rate of about \$15000/year for full-time study and a 58.9\% \emph{indirect cost}, which is what GMU charges for administrative costs (yes, after all, universities are businesses!).
-Finally, I assume the student will make 2 conference trips per year, one domestic and one international (so conf.
+I will also use GMU tuition rate of about \$15,000/year for full-time study (which is quite cheap compared to private universities, e.g., MIT charges around \$50K) and a 58.9\% (overhead) rate on \emph{indirect cost}, which is what GMU charges for administrative costs (yes, after all, universities are businesses!).
+Finally, I assume the student makes conference trips per year, one domestic and one international (so conf.
 registration, airline tickets, taxi, meals, etc are all included).
@@
 Many students have provided feedback and contributed questions, which provide contents for the writing.
+
 \bibliographystyle{abbrv}
@@
 \newpage
+
+\section{Virtual Research Opportunities Beyond Physical Boundaries}
+\didi{In my experience, prospective PhD students (from Rwanda) are often unaware of the information provided in this document throughout their undergraduate.
+For this reason they often find themselves unprepared for PhD admission process by the time they finish undergraduate and have to spend a year or more to improve their application profiles.
+I think it would be beneficial to expand the document to also target students who are not yet ready for the application process to help them improve on all criteria considered during admission.
+I drafted the short essay on how to gain research experience remotely in order to improve their research profile and potentially get LOR from experts (hopefully faculty)}\\
+In the realm of computer science, research experience plays a crucial role in securing admissions to top-tier Ph.D.
+programs.
+However, students from underrepresented small colleges, both in the United States and internationally, often encounter limited research opportunities within their institutions.
+Thankfully, there exists an alternative pathway for these aspiring researchers to gain valuable research experience, even without being physically present at a university.
+
+\textbf{Virtual Research Programs}:
+Several universities, organizations, and research institutions offer virtual internships and research programs aimed at providing hands-on research experience.
+These programs often involve working remotely under the guidance of experienced mentors and collaborating with a team of fellow researchers.
+For instance, \href{https://docs.google.com/forms/d/1btIwt4HwjyKMOUk-EMy3rbkfWzFxv2lNrMm_zkd0pA4/viewform?edit_requested=true}{UIUC+ Summer Undergraduate Research in Software Engineering}  offers an unpaid remote internship for software engineering students all over the world to collaborate with mentors from University of Illinois at Urbana Champaign.
+Virtual internships offer opportunities to contribute to ongoing research projects, conduct experiments, analyze data, and write research reports.
+Students can search for such programs through university websites, research institutes, or as of recently chatGPT (or Bard).
+
+\textbf{Online Research Communities and Open Source Contributions}:
+Online research communities and platforms offer a wealth of opportunities for aspiring researchers.
+Platforms like GitHub, GitLab, and Bitbucket host repositories spanning diverse research domains.
+Students can explore repositories related to their areas of interest, collaborate with other contributors, and actively engage in open source research projects.
+By contributing code, fixing bugs, implementing new features, or providing documentation, students can gain practical research experience and interact with experienced developers and researchers in the field.
+
+\textbf{Online Conferences and Workshops}:
+Attending online conferences and workshops is another way to gain exposure to cutting-edge research and establish connections with experts in the field.
+Many conferences now provide virtual participation options, enabling students to access research talks, poster sessions, and panel discussions and sometimes access designated chat rooms or networking events where participants can engage with researchers, ask questions, and seek potential research collaborations.
+It is beneficial to also create profiles academic collaboration platforms such as ResearchGate.
+By creating profiles, joining relevant research groups, and participating in discussions, students can connect with established researchers, contribute to ongoing projects, and potentially collaborate on publications or research proposals as they provide access
+
+In conclusion, for students lacking research opportunities at their small colleges, the virtual realm offers a wealth of possibilities to gain valuable research experience.
+Engaging with open source projects, joining online research communities, utilizing academic collaboration platforms, participating in virtual internships, and attending online conferences are all avenues to contribute to research, connect with experts, and enhance one's research profile.
 

### 2023-06-14 phd-cs-us.tex
updates

--- 2023-06-12 phd-cs-us.tex
@@
 
-Additionally, \textbf{work experiences at renowned research laboratories}, such as Microsoft Research, can significantly strengthen your
+Additionally, \textbf{work experiences at renown research laboratories}, such as Microsoft Research, can significantly strengthen your
 application.
@@
 
-\begin{wrapfigure}{r}{0.45\textwidth}
-    \vspace{-0.3in}
+\begin{wrapfigure}{r}{0.47\textwidth}
+    \vspace{-0.35in}
       \begin{center}
-        \includegraphics[width=0.45\textwidth]{c6.png}
+        \includegraphics[width=0.47\textwidth]{c6.png}
       \end{center}
@@
 \subsection{Tenured or tenure-track faculty? Who do you choose?}
-The short answer is that tenured-track faculty (e.g., assistant professors) are more likely to be young and very active in research (they have to, in order to get tenure).
+
+\begin{wrapfigure}{r}{0.45\textwidth}
+    \vspace{-0.35in}
+      \begin{center}
+        \includegraphics[width=0.45\textwidth]{c8.png}
+      \end{center}
+    \vspace{-0.3in}
+    \end{wrapfigure}The short answer is that tenured-track/untenured faculty (e.g., assistant professors) are more likely to be young and very active in research (they have to, in order to get tenure).
 Thus, they will likely have more time to work with you and push you to do research and publish.
 The downside is that they may not have as much experience in managing students and may not have as much funding as tenured faculty.
-Tenured faculty (e.g., associate and full profs.) are more likely to be older, more well-known, and have more experience in managing students.
-However, they might not push you as hard (they don't have to, they already got tenured) and expect you to figure things out yourself (so you need to be very independent).
+Tenured faculty (e.g., associate and full profs.) are more likely to be older, well-known, and have more experience in managing students.
+However, they might not push you as hard (they don't have to, they already got tenure) and expect you to figure things out yourself (so you need to be very independent).
 Some tenured faculty are also no longer active in research, e.g., they are more involved in administrative duties.
@@
 #   - The students work full time for the company and do their PhD part-time
-#   - Typically will take longer (because you have to work full time) and require some serious committement from the students
+#   - Typically will take longer (because you have to work full time) and require some serious commitment from the students
 

### 2023-06-14 phd-cs-us.tex

--- 2023-06-14 phd-cs-us.tex
@@
 \subsection{Tenured or tenure-track faculty? Who do you choose?} 
-
 The short answer is that tenured-track faculty (e.g., assistant professors) are more likely to be young and very active in research (they have to, in order to get tenure).

### 2023-06-14 phd-cs-us.tex
updates

@@
 All of these make gaining admission in the US difficult as US faculty are not familiar with these issues.
-I wonder if you have and advice and tips for the situations?
 \tcblower
@@
 In the US, the common types of funding for PhD are \emph{graduate teaching assistant} (GTA or TA), \emph{graduate research assistant} (GRA or RA), and \emph{Fellowship}.
-RA is the type of funding paid by a prof.
+RA is paid by a prof.
 for you to do their research.
 TA is paid by the department for you to help with teaching.
-Finally, fellowship is an independent funding that can come from a school, a company, or an organization.
+Finally, fellowship is an independent funding that come from a school, a company, or an organization.
+Tab.~\ref{tab:funding} summarizes the differences.
+
 Note that funding is typically more available for PhD students than 
-Masters and especially undergraduate studies, which typically have no funding.  Tab.~\ref{tab:funding} summarizes the differences.
+Masters and especially undergraduate studies, which typically have no funding (and they also have to pay international tuition, which is very expensive!).
 

### 2023-06-14 phd-cs-us.tex
updates

@@
 \begin{commentbox}[Thanh]
-Vietnamese universities typically offer several specialized programs, such as the talented engineer program at HUST, that have highly competitive entrance exams and a limited number of available slots (e.g., 30 per year).
+Vietnamese universities typically offer specialized programs, such as the talented engineer program at HUST, that have higly competitive entrance exams and a limited number of available slots (e.g., 30 per year).
 However, these programs may present several challenges.
 Professors within these programs often set higher requirements for students, including more demanding tests and assignments, resulting in lower GPAs and overall rankings.
-For example, an 3.5 GPA students from such talented programs are easily much better than a 4.0 GPA students not in those programs.
+For example, an 3.5 GPA students from such talented programs are typically much better than a 4.0 GPA students not in those programs.
 Similarly, variations in GPA standards exist among different universities, with technical universities generally having lower GPAs than economical universities.

### 2023-06-14 phd-cs-us.tex
Updates from Overleaf

@@
 \tcblower
-\textbf{Vu} Many Vietnamese students and faculty lament how this competitive/low-grading from top Vietnamese universities hurts its own students in applying abroad.
+\textbf{Vu}: Many Vietnamese students and faculty lament how this competitive/low-grading from top Vietnamese universities hurts its own students in applying abroad.
 One way to mitigate this is making these issues known to admission committee.
@@
 
-More generally, rankings can be superficial and you will want to do more research to be informed and make better decision.
+More generally, rankings can be superficial and you need to do more research to be informed and make better decision.
 For example, if you get admissions to several places, you should contact profs.
@@
 
-\section{How much do \emph{you} cost?}\label{sec:ra-cost}
+\subsection{How much do \emph{you} cost?}\label{sec:ra-cost}
 PhD students often wonder why their salary is so low compared to ludicrous grants their advisors get or why their offer letters sometime mentioned that their benefits worth way more than what they actually receive (i.e., stipend).
@@
 \end{table}
+
+\subsection{Having fun during a PhD?}
+PhD students \emph{and faculty} probably find it amusing about the notion that students, especially international ones, can genuinely enjoy their PhD studies.
+In fact, after reading posts after posts on VietPhD.org on how PhD students are commonly mistreated, stressed, it seems being miserable is a norm during a PhD study.
+
+There are many advice on surviving PhD that you can follow. But here I just list a few that works for me and what I advice my students to do.\tvn{TODO}
 
@@
 Other GMU faculty members also have provided feedback and contributions.
-Many students have provided feedback and contributed questions, which provide contents for the writing.
+Many students including Didier (GMU), Thanh (Melbourne), Dat (Melbourne) have contributed valuable questions and feedback.
+\textbf{Thank you!}
 

### 2023-06-15 phd-cs-us.tex
Updates from Overleaf

@@
 \subsection{Tenured or tenure-track faculty? Who do you choose?} 
-The short answer is that tenured-track faculty (e.g., assistant professors) are more likely to be young and very active in research (they have to, in order to get tenure).
+The short answer is that tenured-track faculty (e.g., assistant professors) are more likely to be young and active in research (they have to, in order to get tenure).
 Thus, they will likely have more time to work with you and push you to do research and publish.
-The downside is that they may not have as much experience in managing students and may not have as much funding as tenured faculty.
+However, they may not have as much experience in managing students and may not have as much funding as tenured faculty.
 Tenured faculty (e.g., associate and full profs.) are more likely to be older, more well-known, and have more experience in managing students.
@@
     \end{wrapfigure}
-If you don't know that professor (e.g., first email contact), then use \textbf{Prof.
+If you don't know the professor (e.g., first email contact), then use \textbf{Prof.
 Lastname} or \textbf{Dr.
@@
 
-\begin{commentbox}
-    Vu: I've been called Dr. Vu and I find it a bit amusing but am totally fine with it.
+\begin{commentbox}[Vu]
+    I've been called Dr. Vu and I find it a bit amusing but am totally fine with it.
 \end{commentbox}
@@
 
-\subsection{Having fun during a PhD?}
-PhD students \emph{and faculty} probably find it amusing about the notion that students, especially international ones, can genuinely enjoy their PhD studies.
+%\subsection{Having fun during a PhD?}
+%PhD students \emph{and faculty} probably find it amusing about the notion that students, especially international ones, can genuinely enjoy their PhD studies.
 In fact, after reading posts after posts on VietPhD.org on how PhD students are commonly mistreated, stressed, it seems being miserable is a norm during a PhD study.
 
-There are many advice on surviving PhD that you can follow. But here I just list a few that works for me and what I advice my students to do.\tvn{TODO}
+%There are many advice on surviving PhD that you can follow. But here I just list a few that works for me and what I advice my students to do.\tvn{TODO}
 

### 2023-06-15 phd-cs-us.tex
Updates from Overleaf

@@
 \documentclass[11pt]{article}
-\usepackage[margin=1.1in]{geometry}
+\usepackage[T1]{fontenc}
+\usepackage[margin=1.0in]{geometry}
+\usepackage{soul}
 %\usepackage[small,compact]{titlesec} %very powerful
@@
 \usepackage{epigraph}
-\usepackage[T1]{fontenc}
+
 \usepackage{cite}
@@
 \renewcommand{\tablename}{Tab.}
-\title{Demystifying the Computer Science \\PhD Admission in US Universities\\{\large A guide for Vietnamese and International Students}}
+\title{\vspace{-1in} Demystifying the Computer Science \\PhD Admission in US Universities\\{\large A guide for Vietnamese and International Students}}
 %\date{\today}
-\author{\small \href{https://nguyenthanhvuh.github.io}{ThanhVu (Vu) Nguyen}\\
+\author{\small \href{https://nguyenthanhvuh.github.io}{ThanhVu (Vu) Nguyen} (\href{mailto:tvn@gmu.edu}{tvn@gmu.edu})\\
  \href{https://cs.gmu.edu}{George Mason University}}
-%\setcounter{secnumdepth}{3}
+
 \begin{document}
@@
 Having been involved in PhD admissions for many years, I've
-realized that many international students, especially \emph{Vietnamese}, lack a clear understanding of
+realized that many \textbf{international students}, especially those from smaller countries such as \emph{Vietnamese}, lack a clear understanding of
 the Computer Science PhD admission process at US universities. This confusion not only
@@
 
-So I want to reveal some details about the PhD Admission process and share my opinions and advice for those who are interested in applying for a \textbf{PhD in Computer Science in the US}.
+So I want to give some details about the PhD Admission process and share my opinions and advice for those who are interested in applying for a \textbf{PhD in Computer Science in the US}.
 While this document is written for Vietnamese students interested in Computer Science, it should apply to students in many other countries and other majors.
@@
 
-In addition, this document can help \emph{US faculty} to understand more about international students, e.g., questions, concerns, and various cultural differences.
+In addition, this document can help \emph{US faculty and admission committee members} understand more about international students, e.g., questions, concerns, and various cultural differences.
 By understanding, appreciating, and even leveraging these information, CS programs in the US can attract larger and stronger application pools from Vietnam and other countries.
@@
 \section{Should You Apply?}
-\epigraph{Don't make fun of graduate students. They just made a terrible life choice.}{The Simpsons}
+\epigraph{\vspace{-0.2in} Don't make fun of graduate students. They just made a terrible life choice.}{The Simpsons}
 
@@
 Second, I believe that applying to a good US university \emph{should not} be any
-harder than at schools in other countries. If you think you have a
-chance at schools in other countries, e.g., South Korea, Singapore, Germany, UK, Japan and Australia, then you will surely have a chance in the US as well.
+harder than at schools in other countries.
+If you think you have a chance in other countries, e.g., South Korea, Singapore, Germany, UK, Japan and Australia, then you will surely have a chance in the US as well.
 
@@
 \begin{commentbox}[Thanh]
-Vietnamese universities typically offer specialized programs, such as the talented engineer program at HUST, that have higly competitive entrance exams and a limited number of available slots (e.g., 30 per year).
-However, these programs may present several challenges.
-Professors within these programs often set higher requirements for students, including more demanding tests and assignments, resulting in lower GPAs and overall rankings.
+Vietnamese universities typically offer specialized programs, such as the talented engineer program at HUST, that have highly competitive entrance exams and a limited number of available slots (e.g., 30 per year).
+However, these programs often set higher requirements for students, including more demanding tests and assignments, resulting in lower GPAs and overall rankings.
 For example, an 3.5 GPA students from such talented programs are typically much better than a 4.0 GPA students not in those programs.
 Similarly, variations in GPA standards exist among different universities, with technical universities generally having lower GPAs than economical universities.
-All of these make gaining admission in the US difficult as US faculty are not familiar with these issues.
+These make gaining admission in the US difficult as US faculty are not familiar with these issues.
 \tcblower
@@
 
-Several about stipend.
+Several words about stipend.
 First, the amount of stipend depends on the university, which in turns depend on various factors such as location (e.g., a stipend in Washington DC is likely higher than in Lincoln, Nebraska due to higher living cost).
@@
 TA and RA at GMU have similar benefits in tuition waiving and insurance.
-For stipend, depending on the college and department, a 9-month graduate assistant stipend is set.
-TA and RA will usually be that amount (TA will definitely be that, RA might fluctuate a bit depending on the stage of the student and the prof.).
+For stipend, the college and department will set a 9-month graduate assistant stipend.
+TA and RA will usually be that amount (TA will likely be that amount but RA might be higher depending on the stage of the student (1st year vs ABD\footnote{All but dissertation: really close to graduate.}) and the prof.).
 \tcblower
@@
 TA is common in the beginning when you haven't found your advisor who would pay you RA.
-As a TA, you help professors with their classes (e.g., grading or teaching labs/recitation).
-Your TAship is paid through the school/department, i.e., they hire you to help teach.
+As a TA, you spend up to 20 hrs/week and help professors with their classes (e.g., grading or teaching labs/recitation).
+Your TA is paid through the school/department, i.e., they hire you to help teach.
 During a semester, a TA might work with several courses and professors (not necessary their advisor).
@@
 Even if you have other funding and do not need TA, you still should do TA at least once.
-This allows you to see what teaching is like, which is especially helpful for research career where you often have to give talks and tell people about your work.
+This allows you to see what teaching is like, which is especially helpful for research career where you often give talks and tell people about your work.
 Note that GMU sometimes has classes that a more senior student can teach.
@@
 \subsection{Tenured or tenure-track faculty? Who do you choose?}
+
+\begin{wrapfigure}{r}{0.48\textwidth}
+    \vspace{-0.3in}
+      \begin{center}
+        \includegraphics[width=0.48\textwidth]{c8.png}
+      \end{center}
+    \vspace{-0.2in}
+    \end{wrapfigure}
 The short answer is that tenured-track faculty (e.g., assistant professors) are more likely to be young and active in research (they have to, in order to get tenure).

### 2023-06-15 phd-cs-us.tex
Updates from Overleaf
updates

@@
 %\date{\today}
-\author{\small \href{https://nguyenthanhvuh.github.io}{ThanhVu (Vu) Nguyen} (\href{mailto:tvn@gmu.edu}{tvn@gmu.edu})\\
- \href{https://cs.gmu.edu}{George Mason University}}
+\author{\small \href{https://nguyenthanhvuh.github.io}{ThanhVu (Vu) Nguyen} (\href{mailto:tvn@gmu.edu}{tvn@gmu.edu})}
 
@@
 \begin{commentbox}[Vu] 
-We usually decide that a PhD candidate is either (i) admit with funding (TA or RA) or (ii) rejected. In other words, in most cases, we either
+We usually decide that a full-time PhD candidate is either (i) admit with funding (TA or RA) or (ii) rejected. In other words, in most cases, we either
 admit you with full funding, or we don't. In some rare cases, we admit

### 2023-06-15 phd-cs-us.tex
updates

@@
 For simplicity, I will assume the department has a 9-month stipend of \$27000 (GMU actually pays more) and therefore a 3-month summer of \$9000.
-I will also use GMU tuition rate of about \$15,000/year for full-time study (which is quite cheap compared to private universities, e.g., MIT charges around \$50K) and a 58.9\% (overhead) rate on \emph{indirect cost}, which is what GMU charges for administrative costs (yes, after all, universities are businesses!).
-Finally, I assume the student makes conference trips per year, one domestic and one international (so conf.
+I will also use GMU tuition rate of about \$15,000/year for full-time study (which is quite cheap compared to private universities, e.g., MIT charges around \$50K) and a 58.9\% rate on \emph{indirect cost}, which is what GMU charges for overhead or administrative costs (yes, after all, universities are businesses!).
+Finally, I assume the student makes two conference trips per year, one domestic and one international (so conf.
 registration, airline tickets, taxi, meals, etc are all included).
+I do not include budgets such as computer equipments.
 At the end, the total budget comes out to be \$69K.

### 2023-06-16 phd-cs-us.tex
updates

@@
 Having been involved in PhD admissions for many years, I've
-realized that many \textbf{international students}, especially those from smaller countries such as \emph{Vietnamese}, lack a clear understanding of
+realized that many \textbf{international students}, especially those from smaller countries such as \emph{Vietnam}, lack a clear understanding of
 the Computer Science PhD admission process at US universities. This confusion not only
@@
 
-So I want to give some details about the PhD Admission process and share my opinions and advice for those who are interested in applying for a \textbf{PhD in Computer Science in the US}.
-While this document is written for Vietnamese students interested in Computer Science, it should apply to students in many other countries and other majors.
-Moreover, while many examples given are for George Mason University, the writing should generalize to other R1\footnote{An \href{https://en.wikipedia.org/wiki/List_of_research_universities_in_the_United_States}{R1 institution} in the US is a research-intensive university with a high level of research activity across various disciplines.} universities  (though \emph{very} top schools could be very selective, e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU).
+So I want to share some details about the PhD admission process and advice for those who are interested in applying for a \textbf{PhD in Computer Science in the US}.
+While this document is intended for Vietnamese students interested in Computer Science, it may be applicable to students in from various countries and disciplines.
+Moreover, although many examples are for George Mason University, the writing should be generalizable to other R1\footnote{An \href{https://en.wikipedia.org/wiki/List_of_research_universities_in_the_United_States}{R1 institution} in the US is a research-intensive university with a high level of research activity across various disciplines.} universities  (though \emph{very} top schools could be very selective, e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU).
 
-In addition, this document can help \emph{US faculty and admission committee members} understand more about international students, e.g., questions, concerns, and various cultural differences.
-By understanding, appreciating, and even leveraging these information, CS programs in the US can attract larger and stronger application pools from Vietnam and other countries.
+In addition, this document can help \emph{US faculty and admission committee} understand more about international students and their cultural differences.
+By understanding, appreciating, and even leveraging these differences, CS programs in the US can attract larger and more competitive application pools from Vietnam and other countries.
 
-I wish you the best of luck. And if you follow these advice,
+I wish you the best of luck. And if you follow this guideline,
 you will at least have a good chance at GMU (see
@@
 for general requirements, e.g., did you submit your transcripts and standard scores? did your reference writers submit their letters?
-Then your application will be reviewed by a
-\textbf{PhD admission committee} that consists of faculty members in CS.
-Each application is assigned to about \emph{three} faculty members, who will evaluate your profile and try to reach a consensus about your case.
-Note that while the assigned reviewers are likely the main ones deciding your application, \emph{every} faculty in the department will have access to your application and can give inputs on your profile.
+
+Application will be reviewed by a
+\textbf{PhD admission committee} that consists of faculty members in CS (in some cases the committee can involve affiliated faculty from different disciplines).
+These faculty have a wide-range of expertise and background to ensure diverse perspectives in the evaluation process.
+The size of the committee depends on the department size, e.g., around 15 at GMU.
+Each application is assigned to about \emph{three} faculty members, who will evaluate your profile and reach a consensus.
+Note that while the assigned reviewers are likely the main ones deciding your application, \emph{every} faculty in the department will have access to your application and can provide inputs and opinions on your profile.
 
@@
 The primary focus of the admissions committee is to \textbf{evaluate your background and interest in research} since a PhD in Computer Science
-is a research degree. To assess your research capability, we consider
+is a research degree and determine if you would \textbf{fit} the program. To evaluate your profile, we consider
 the following key indicators, listed in order of importance.
@@
 
-\begin{wrapfigure}{r}{0.50\textwidth}
-\vspace{-0.4in}
+\begin{wrapfigure}{r}{0.52\textwidth}
+\vspace{-0.35in}
   \begin{center}
-    \includegraphics[width=0.50\textwidth]{c2.png}
+    \includegraphics[width=0.52\textwidth]{c2.png}
   \end{center}
@@
 applicants: \textbf{customize the statement} for the school you're applying to,
-e.g., why do you apply here? talk about a couple of professors who you're interested in working with (in many cases your application will be forwarded to them for evaluation).
+e.g., why do you apply here? mention the names of professors who you're interested in (in many cases your application will be forwarded to them for evaluation).
 This shows that you're serious and have done homework on places you're applying to.
-Be careful not to send statements to wrong schools or mixing
-facts (e.g., talking about school X but mentioned about working with
-profs. at school Y; and definitely do not talk about George Washington when applying to George Mason). I have seen such statements more time that I
-should.
+Admission committee will look for this part at the end of the statement, so do not skip it.
 
@@
 something standout in other parts).
+\tcblower
+Be careful not to send statements to wrong schools or mixing
+facts (e.g., talking about school X but mentioned about working with
+profs. at school Y; and definitely do not talk about George Washington when applying to George Mason). I have seen such statements more time that I
+should.
 \end{commentbox}
@@
 An interview lasts about 15--30 mins, and one implicit thing you will be evaluated on is whether you can communicate effectively (i.e., speak/understand English).
-You will also get chance to ask questions about the university so think of something to ask (just the same as you interview at a company).
+You will also get chance to ask questions about the university so think of something to ask (just the same as interviewing at a company).
 
@@
 For very strong candidates, the interview is actually to recruit them.
-In some cases a faculty interviews a candidate that they see potentials and wants argue for admission, i.e., without the interview, that candidate is definitely rejected.
+In some cases a faculty interviews a candidate that they see potentials and wants argue for admission, i.e., without the interview, that application is likely rejected.
 In any case, getting interview means you have a very good chance of being admitted.
@@
 \S\ref{sec:ranking} gives the top 50 CS programs in the US according to CSRankings.
+
 \begin{commentbox}[Dat] Most Vietnamese students, including those from top schools, do not know about CSRankings.
@@
 \subsection{Tenured or tenure-track faculty? Who do you choose?}
-
+\begin{commentbox}[Thanh]
+In my opinion, having a well-suited advisor is crucial for a successful PhD and research career.
+One effective approach to finding a suitable professor is by working with a professor during your undergraduate studies.
+An exemplary instance is VinAI's residency program, where residents collaborate with professors from the US for two years before applying to PhD programs.
+Many VinAI residents have achieved remarkable results and gained admission to prestigious US universities.
+Unfortunately, VinAI's resident program is limited to AI research.
+In other fields, e.g.
+SE,
+Vietnamese students face challenges in reaching US professors. Do you have any tips for Vietnamese students who want to connect with US professors and work as research assistants?
+\end{commentbox}
 \begin{wrapfigure}{r}{0.48\textwidth}
@@
 However, they may not have as much experience in managing students and may not have as much funding as tenured faculty.
+
 Tenured faculty (e.g., associate and full profs.) are more likely to be older, more well-known, and have more experience in managing students.
@@
 \begin{wrapfigure}{r}{0.28\textwidth}
-    \vspace{-0.4in}
+    \vspace{-0.2in}
       \begin{center}
@@
     Typically, 5-7 years for PhD in CS at US universities.
-The first 1.5--2 years you spend on coursework and learning research.
-The next 2--3 years you focus on your research, formi dissertation topic, and get results published.
-The last 1-2 years you write and defend your dissertation and look for job.
-The PhDComics figure on the right shows the "ambition" level of a PhD student over their years of study (they miss the 6-7th Year where the ambition is "Please just get me graduate").
+The first 2 years you spend on coursework, finding an adviser, learning research.
+The next 2--3 years you focus on your research, form dissertation topic, and get results published.
+The last 1-2 years you continue to publish, write and defend your dissertation, and look for job.
+The PhDComics figure on the right shows the "ambition" level of a PhD student over their years of study (they miss the 6-7th Year where the ambition is "Just get me graduate").
 
@@
 registration, airline tickets, taxi, meals, etc are all included).
-I do not include budgets such as computer equipments.
 At the end, the total budget comes out to be \$69K.

### 2023-06-16 phd-cs-us.tex
Updates from Overleaf

@@
 \begin{commentbox}[Vu] 
-We usually decide that a full-time PhD candidate is either (i) admit with funding (TA or RA) or (ii) rejected. In other words, in most cases, we either
+We usually decide that a PhD candidate is either (i) admit with funding (TA or RA) or (ii) rejected. In other words, in most cases, we either
 admit you with full funding, or we don't. In some rare cases, we admit
@@
 For simplicity, I will assume the department has a 9-month stipend of \$27000 (GMU actually pays more) and therefore a 3-month summer of \$9000.
-I will also use GMU tuition rate of about \$15,000/year for full-time study (which is quite cheap compared to private universities, e.g., MIT charges around \$50K) and a 58.9\% rate on \emph{indirect cost}, which is what GMU charges for overhead or administrative costs (yes, after all, universities are businesses!).
-Finally, I assume the student makes two conference trips per year, one domestic and one international (so conf.
+I will also use GMU tuition rate of about \$15,000/year for full-time study (which is quite cheap compared to private universities, e.g., MIT charges around \$50K) and a 58.9\% (overhead) rate on \emph{indirect cost}, which is what GMU charges for administrative costs (yes, after all, universities are businesses!).
+Finally, I assume the student makes conference trips per year, one domestic and one international (so conf.
 registration, airline tickets, taxi, meals, etc are all included).

### 2023-06-17 phd-cs-us.tex
Updates from Overleaf

--- 2023-06-16 phd-cs-us.tex
@@
 
-The primary focus of the admissions committee is to \textbf{evaluate your background and interest in research} since a PhD in Computer Science
-is a research degree and determine if you would \textbf{fit} the program. To evaluate your profile, we consider
+The primary focus of the admissions committee is to \textbf{evaluate your background and interest in research} since a PhD in Computer Science and determine if you would \textbf{fit into} the program.
+To evaluate your profile, we consider
 the following key indicators, listed in order of importance.
@@
 
+\subsection{Should you contact a US professor? How to do so "correctly"?}
+
+\tvn{TODO: How to write an email that will get a reply?}
+
 \subsection{Can I apply to CS PhD if my undergrad was not in CS or related areas?}

### 2023-06-19 phd-cs-us.tex

@@
 In addition, this document can help \emph{US faculty and admission committee} understand more about international students and their cultural differences.
-By understanding, appreciating, and even leveraging these differences, CS programs in the US can attract larger and more competitive application pools from Vietnam and other countries.
+By understanding, appreciating, and even leveraging these differences, CS programs in the US can attract larger and more competitive application pools from international students.
 
@@
 \paragraph{Why we do not waive application fee?}  This is typically a requirement of the university.
-Individual departments/programs do not have the flexibility to waive the application fee, even if they want to.
+Individual departments and programs do not have the flexibility to waive the application fee, even if they want to.
 
@@
 While you might not have control over LORs \hyperref[sec:your-school]{where your go to school}, you do over your
-statement! So write it well because we do take it seriously.
+statement! So write it well because we take it seriously.
+It is something that can make your application unique and stand out, which can make you "fit"  the CS program you're applying to, even if you don't have very strong research experience.
 A well-written LOR also shows that you can communicate, which is very important in research, and that you can effectively teach and communicate with students, which is important for TA funding (see \S\ref{sec:funding}).
@@
 applicants: \textbf{customize the statement} for the school you're applying to,
-e.g., why do you apply here? mention the names of professors who you're interested in (in many cases your application will be forwarded to them for evaluation).
+e.g., why do you apply here? provide names of professors who you're interested in (in many cases your application will be forwarded to them for evaluation).
 This shows that you're serious and have done homework on places you're applying to.
@@
 \tcblower
-Be careful not to send statements to wrong schools or mixing
+Do careful research on professors, don't mention \emph{emeritus} or  adjunct faculty. 
+Also, be careful not to send statements to wrong schools or mixing
 facts (e.g., talking about school X but mentioned about working with
@@
 \tcblower
-\textbf{Vu}: Many Vietnamese students and faculty lament how this competitive/low-grading from top Vietnamese universities hurts its own students in applying abroad.
-One way to mitigate this is making these issues known to admission committee.
+\textbf{Vu}: Vietnamese students and even faculty often lament how this competitive/low-grading system hurts Vietnamese students when applying abroad.
+One way to mitigate this is making these issues known to admission committee in your statement.
 Schools with Vietnamese profs are probably aware of them, but in general your letter writers/you can explicitly mention these in their letters/ your statement.
@@
 For very strong candidates, the interview is actually to recruit them.
-In some cases a faculty interviews a candidate that they see potentials and wants argue for admission, i.e., without the interview, that application is likely rejected.
+In some cases a faculty interviews a candidate that they see potentials and wants to argue for admission, i.e., without the interview, that application is likely rejected.
 In any case, getting interview means you have a very good chance of being admitted.

### 2023-06-19 phd-cs-us.tex
updates

@@
 \begin{commentbox}[Vu]
-One of the reasons I create this document is that my (non-Vietnamese) colleagues at GMU are interested in 
+One of the reasons I create this document is that my colleagues at GMU are interested in 
 recruiting Vietnamese students and are surprised when seeing very few applications in Vietnam (e.g., each year our CS program receives more than 350 PhD applications, most of which are international but only 3--4 are from Vietnam).
@@
 At GMU, for full consideration, students should make sure to submit \textbf{ALL} required documents by the application deadline, and should never assume that some required documents (such as official TOEFL scores or official diplomas/transcripts) will be waived by the admissions office.
-If something is listed and not marked as "optional", it is mandatory and they should plan for submitting all those.
+If something is listed and not marked as ``optional'', it is mandatory and they should plan for submitting all those.
 \end{commentbox}
@@
 statement! So write it well because we take it seriously.
-It is something that can make your application unique and stand out, which can make you "fit"  the CS program you're applying to, even if you don't have very strong research experience.
+In your statement, you have the opportunity to make  make your application stand out and unique, which can make you \emph{fit} the CS program you're applying to, even if you don't have very strong research experience.
 A well-written LOR also shows that you can communicate, which is very important in research, and that you can effectively teach and communicate with students, which is important for TA funding (see \S\ref{sec:funding}).
@@
 TA and RA at GMU have similar benefits in tuition waiving and insurance.
-For stipend, the college and department will set a 9-month graduate assistant stipend.
-TA and RA will usually be that amount (TA will likely be that amount but RA might be higher depending on the stage of the student (1st year vs ABD\footnote{All but dissertation: really close to graduate.}) and the prof.).
+The college and department will set the rate for 9-month graduate assistant stipend.
+TA will likely be that amount but RA might be higher depending on the stage of the student (1st year vs ABD\footnote{All but dissertation: really close to graduate.}) and the prof.
 \tcblower
@@
 \subsection{Tenured or tenure-track faculty? Who do you choose?}
-\begin{commentbox}[Thanh]
-In my opinion, having a well-suited advisor is crucial for a successful PhD and research career.
-One effective approach to finding a suitable professor is by working with a professor during your undergraduate studies.
-An exemplary instance is VinAI's residency program, where residents collaborate with professors from the US for two years before applying to PhD programs.
-Many VinAI residents have achieved remarkable results and gained admission to prestigious US universities.
-Unfortunately, VinAI's resident program is limited to AI research.
-In other fields, e.g.
-SE,
-Vietnamese students face challenges in reaching US professors. Do you have any tips for Vietnamese students who want to connect with US professors and work as research assistants?
-\end{commentbox}
+
 \begin{wrapfigure}{r}{0.48\textwidth}
@@
 
+\begin{commentbox}[Thanh]
+    In my opinion, having a well-suited advisor is crucial for a successful PhD and research career.
+One effective approach to finding a suitable professor is by working with a professor during your undergraduate studies.
+An exemplary instance is VinAI's residency program, where residents collaborate with professors from the US for two years before applying to PhD programs.
+Many VinAI residents have achieved remarkable results and gained admission to prestigious US universities.
+Unfortunately, VinAI's resident program is limited to AI research.
+In other fields, e.g.
+SE,
+    Vietnamese students face challenges in reaching US professors. Do you have any tips for Vietnamese students who want to connect with US professors and work as research assistants?
+    \end{commentbox}
+
 % \subsubsection{Strong Faculty vs High-ranked School}
@@
 
-\subsection{Should you contact a US professor? How to do so "correctly"?}
+\subsection{Should you contact a US professor? How to do so ``correctly''?}
 
-\tvn{TODO: How to write an email that will get a reply?}
+If you want to contact a prof.
+to \emph{ask about your admission chance}, then \textbf{don't}.
+We don't know, we can't answer because as explained throughout this document, we don't make individual decisions and might not even be assigned to evaluate your application.
+It is the same as sending a paper draft to a journal editor and ask them if your paper will be accepted.
+
+If you want to contact a prof.
+to ask about \emph{research opportunities}, then \emph{may be}.
+First, you should read the prof's website, see if they are saying something about contacting them.
+Many profs.
+indicate what you should do to contact them (e.g., specific email subjects).
+In general cases, you need to \emph{customize your email}.
+For example, read their papers, and see if you are interested in their work.
+If you are, then you can send them an email talking how/why their work would match yours (give details and evidence, don't just say I am a software developer and therefore match with your interests in software design) and ask about research opportunities.
 
@@
 \subsection{Is an MS degree required for admission to PhD?}
-No. In fact, student with BS can get MS degree "along the way" to PhD.  However, MS can help if it gives research experience or is from a more well-known school than your undergrad institution.
+No. In fact, student with BS can get MS degree ``along the way'' to PhD.  However, MS can help if it gives research experience or is from a more well-known school than your undergrad institution.
+
+If you have an MS then some course work \emph{might be} transferred for course credits, which save a bit of time. But overall don't count on this, especially if your MS is not from the US.
+
+% \begin{commentbox}[Thanh]
+% \end{commentbox}
 
@@
     \end{wrapfigure}
-    Typically, 5-7 years for PhD in CS at US universities.
+    Typically, 5--7 years for PhD in CS in the US.
+This can be longer than CS PhD at universities in other countries, which might require MS first (recall that US PhD programs do not require MS and you can get MS along the way to PhD).
+Within these 5--7 years, CS PhD students in the US also often take a ``leave of absence'' for 1--2 semesters to do internship at companies and research labs.
+
     The first 2 years you spend on coursework, finding an adviser, learning research.
@@
 The last 1-2 years you continue to publish, write and defend your dissertation, and look for job.
-The PhDComics figure on the right shows the "ambition" level of a PhD student over their years of study (they miss the 6-7th Year where the ambition is "Just get me graduate").
+    In many cases you might take a summer or two off to do internship to get additional research opportunities.
+    The PhDComics figure on the right shows the ``ambition'' level of a PhD student over their years of study (they miss the 6-7th Year where the ambition is \emph{``Just let me graduate''}).
 

### 2023-06-20 phd-cs-us.tex
Updates from Overleaf

@@
 
-Graduating from top universities \emph{that we recognize} helps.
+Graduating from top universities \emph{that we recognize} helps. For example, if your school is well-known, then it is \emph{``top foreign''}, which is definitely a plus.
 However, if committee members do not know much about schools in your country, they will likely treat your school as
-\emph{``unknown foreign''}, which can be a minus point (if your school is well-known, then it is \emph{``top foreign''}, which is definitely a plus).
+\emph{``unknown foreign''}, which can be a minus point because we are uncertain about the quality of your school.
 
 So what can you do about this?
-several things such as asking your CS dept to put itself on CSRankings (it's the easiest way to get CS people to know about the school), explaining about your school in your statement (and asking your LOR writer to do that too), and of course, if you're interested in working with Vietnamese professors, consider a PhD program in the US that  \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{have them}.
+several things such as asking your CS dept to put itself on CSRankings (it's the easiest way to get CS people to know about the school) and explaining about your school in your statement (and asking your LOR writer to do that too).
+Of course, if you're interested in working with Vietnamese professors, consider a PhD program in the US that  \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{have them}.
 
@@
 Sometime PhD admission committee will share a document such as \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Foreign-Top-Schools}{this one}, which lists the top schools in several countries.
-Personally I have looked at Vietnamese applications (whether they are assigned to me or not) and provide input to their reviewers, e.g., X is the top tech school in Vietnam and so it should be \emph{top} instead of \emph{unknown} foreign, which makes a huge difference.
+In some cases we ask people (faculty, students) if we think they know about the place.
+For example, when I was a postdoc at UMD, people in admission committee ask me to evaluate applicants from Vietnam.
+During my time at UNL and now here at GMU, I have looked at Vietnamese applications (whether they are assigned to me or not) and provide input to their reviewers, e.g., X is the top tech school in Vietnam and so it should be \emph{top} instead of \emph{unknown} foreign, which makes a huge difference.
 \end{commentbox}
@@
 
+This is one of the most common questions that international students often have.
+Faculty often received many mails from international students about admission, TA, RA and various related topics.
+Most of the time, we ignore these emails; but sometimes, we do answer them.
+So under what cases we would answer and would not?
+
 If you want to contact a prof.
-to \emph{ask about your admission chance}, then \textbf{don't}.
-We don't know, we can't answer because as explained throughout this document, we don't make individual decisions and might not even be assigned to evaluate your application.
-It is the same as sending a paper draft to a journal editor and ask them if your paper will be accepted.
+to \emph{ask about your admission chance}, \textbf{don't}.
+We don't know and can't answer because as explained throughout this document, we don't make individual decisions and might not even be assigned to evaluate your application.
+It is the same as sending a paper draft to a journal editor and ask them if your paper has a chance.
+
+So what to do if you want someone to look at your profile and give input?
+You could ask your professors, people at your school, or contact those who have previously applied and get feedback.
+For these kind of feedback,  don't ask strangers like random profs., instead ask someone you have some personal connection to.
 

### 2023-06-24 phd-cs-us.tex
Updates from Overleaf

@@
 
-This is one of the most common questions that international students often have.
+This is a common question that international students have.
 Faculty often received many mails from international students about admission, TA, RA and various related topics.
 Most of the time, we ignore these emails; but sometimes, we do answer them.
-So under what cases we would answer and would not?
+So when we usually respond and when we do not?
 
@@
 If you want to contact a prof.
-to ask about \emph{research opportunities}, then \emph{may be}.
-First, you should read the prof's website, see if they are saying something about contacting them.
-Many profs.
-indicate what you should do to contact them (e.g., specific email subjects).
-In general cases, you need to \emph{customize your email}.
+to ask about \emph{research opportunities}, or \emph{GTA/GRA} support, then \emph{may be}.
+The best way to catch the prof.'s attention you is to \emph{customize your email} to that specific faculty.
 For example, read their papers, and see if you are interested in their work.
 If you are, then you can send them an email talking how/why their work would match yours (give details and evidence, don't just say I am a software developer and therefore match with your interests in software design) and ask about research opportunities.
+Also, don't feel bad if they do not reply.
+
+You should also read the prof's website, see if they are saying something about contacting them. Many profs. indicate what you should do to contact them (e.g., specific email subjects).
 

### 2023-07-05 phd-cs-us.tex
updates

--- 2023-06-20 phd-cs-us.tex
@@
 \usepackage{epigraph}
-
 \usepackage{cite}
@@
 So I want to share some details about the PhD admission process and advice for those who are interested in applying for a \textbf{PhD in Computer Science in the US}.
-While this document is intended for Vietnamese students interested in Computer Science, it may be applicable to students in from various countries and disciplines.
-Moreover, although many examples are for George Mason University, the writing should be generalizable to other R1\footnote{An \href{https://en.wikipedia.org/wiki/List_of_research_universities_in_the_United_States}{R1 institution} in the US is a research-intensive university with a high level of research activity across various disciplines.} universities  (though \emph{very} top schools could be very selective, e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU).
+While this document is primarily intended for Vietnamese interested in computer science, it may be relevant to students from various countries and disciplines.
+Moreover, although many examples are specifics for George Mason University, the information should be generalizable to other R1\footnote{An \href{https://en.wikipedia.org/wiki/List_of_research_universities_in_the_United_States}{R1 institution} in the US is a research-intensive university with a high level of research activity across various disciplines.} institutions in the US  (though \emph{very} top schools could be very selective, e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU).
 
-In addition, this document can help \emph{US faculty and admission committee} understand more about international students and their cultural differences.
-By understanding, appreciating, and even leveraging these differences, CS programs in the US can attract larger and more competitive application pools from international students.
+In addition, this document can help \emph{US faculty and admission committee} gain a better understanding of international students and their cultural differences.
+By appreciating and leveraging these differences, CS programs in the US can attract larger and more competitive application pools from international students.
 
@@
 receive \emph{full funding} to support your study, including tuition,
-health insurance, and stipend. Moreover, depending on the university,
-you may even receive additional benefits such as summer salary, laptops, (conference/workshop) traveling. \S\ref{sec:funding} provides more details on funding.
+health insurance, and stipend (monthly salary). Moreover, depending on the university,
+you may even receive additional benefits such as summer pay, laptops, (conference/workshop) traveling. \S\ref{sec:funding} provides more details on funding.
 
@@
 
-Application will be reviewed by a
+Then applications will be reviewed by a
 \textbf{PhD admission committee} that consists of faculty members in CS (in some cases the committee can involve affiliated faculty from different disciplines).
@@
 However, I understand that many international students do not have the opportunities to publish in top places, so general confs/journals would suffice.
-But be sure to upload your papers with your application and talk about them in your \hyperref[sec:research-statement]{statement}.
+But be sure to upload your papers with your application and talk about them in your statement (see \S\ref{sec:research-statement}).
 
@@
 
-Most CS PhD programs will require at least \textbf{two LORs}. Having a letter from an internationally recognized researcher can greatly strengthen your application. However, obtaining such letters
+CS PhD programs will require at least \textbf{two LORs}. Having a letter from an internationally recognized researcher can greatly strengthen your application. However, obtaining such letters
 can be challenging for international students, who might not have much interactions with such experts.
@@
     \end{wrapfigure}
-Many students have letters written by the applicants themselves and signed by their professors. These have little
-values (we can easily recognize them) and will consider them weakness.
-Similarly, many professors write very generic letters for students (a common example is that the students didn't do any
-research or make any impression for the professor to write about). These letters are also not useful and considered weak.
+Many students have letters written by the themselves and signed by their professors. These have little
+values. We can easily recognize them and will consider them weakness.
+Similarly, many professors write generic letters for students, e.g., the students didn't do any
+research or make any impression for the professor to write about. These letters are also not useful and considered weak.
 
@@
 International students not familiar with US universities often put them into \emph{two} bins:  (i) very top school that they dream about such as Stanford, MIT, Princeton, Harvard and (ii) everything else.
-Sometimes they rank CS programs using the reputations non-CS programs, e.g., medical or physics.
+Sometimes they rank CS programs using the reputations non-CS programs such as medical, math, or physics.
 In some cases they rank universities based on popular states they know in the US, e.g., California and New York.  Let's just say there are so many thing wrong with these methods.
@@
 You can learn about CS programs and research expertise of faculty using resources such as \href{https://csrankings.org}{CSRankings.org}, which is designed specifically to help prospective PhD students in Computer Science!
-You will be very surprised to learn that a school X that you didn't know much about has very strong research in your interested field Y.
+You will be very surprised to learn that a school X that you didn't know much about has very strong research in your interested field Y (and vice versa, a school you thought highly off has no one works in the research field you're interested in).
 This is also a good way to learn about individual faculty (who works on what) and well-known CS conferences\footnote{In CS (and only in CS), conferences, not journals, are often the main venue to publish research finding.}.
@@
 However, they might not push you as hard (they don't have to, they already got tenured) and expect you to figure things out yourself (so you need to be very independent).
-Some tenured faculty are also no longer active in research, e.g., they are more involved in administrative duties.
+Some tenured faculty are also no longer active in research and more involved with administrative duties or with their startup companies.
 
@@
 Unfortunately, VinAI's resident program is limited to AI research.
+
     In other fields, e.g.
-SE,
-    Vietnamese students face challenges in reaching US professors. Do you have any tips for Vietnamese students who want to connect with US professors and work as research assistants?
+Software Engineering, Vietnamese students face challenges in reaching US professors.
+Do you have any tips for Vietnamese students who want to connect with US professors and work as research assistants?
     \end{commentbox}
@@
 
-This is one of the most common questions that international students often have.
-Faculty often received many mails from international students about admission, TA, RA and various related topics.
-Most of the time, we ignore these emails; but sometimes, we do answer them.
-So under what cases we would answer and would not?
+Faculty received many "cold" e-mails from international students seeking for admission, TA, and RA.
+Most of the time, we ignore these emails; but in rare occasion we do answer them.
+So how to write an email that can get our attention?
 
-If you want to contact a prof.
-to \emph{ask about your admission chance}, \textbf{don't}.
+First, if you want to contact a prof.
+to \emph{ask about your admission chance}, please \textbf{don't}.
 We don't know and can't answer because as explained throughout this document, we don't make individual decisions and might not even be assigned to evaluate your application.
@@
 So what to do if you want someone to look at your profile and give input?
-You could ask your professors, people at your school, or contact those who have previously applied and get feedback.
-For these kind of feedback,  don't ask strangers like random profs., instead ask someone you have some personal connection to.
+You could ask your professors, collaborators, or those who have previously applied.
+For these kind of feedback,  don't ask strangers like random profs., instead ask someone you have personal connection with.
 
-If you want to contact a prof.
-to ask about \emph{research opportunities}, then \emph{may be}.
-First, you should read the prof's website, see if they are saying something about contacting them.
-Many profs.
-indicate what you should do to contact them (e.g., specific email subjects).
-In general cases, you need to \emph{customize your email}.
-For example, read their papers, and see if you are interested in their work.
-If you are, then you can send them an email talking how/why their work would match yours (give details and evidence, don't just say I am a software developer and therefore match with your interests in software design) and ask about research opportunities.
+If you want to contact a prof. to ask about \emph{research opportunities}, or \emph{GTA/GRA} support, then \emph{yes}, I believe you should, but you really have to write your mail in a certain way. 
+First, read the prof's website, see if they are saying something about contacting them. Many profs. indicate what you should do to contact them (e.g., specific email subjects). 
+In general, the best way to catch the prof.'s attention is to \emph{customize your email} to that specific faculty.
+For example, read their papers, know what they work on, and see if you are interested in their research.
+If you are, then you can send them an email talking how/why their work would match yours.
+In contrast, if you write a generic email that can be sent to multiple professors (e.g., if you just change some names and keywords in the email), you will not get any response.
+
+Below is a good example that I would definitely reply to.
+
+\begin{commentbox}[Good example]
+
+Dear Prof. Nguyen,
+
+I read your TSE paper on numerical invariant generation, and I am interested in this line of dynamic invariant research.
+I have worked (optional: with prof.
+Y at Z) on static analysis on C programs and I think it would be a good fit to tackle the spurious issues mentioned in your paper.
+I have a small paper at conference/workshop C and a project on symbolic execution at Github G.
+
+...
+\tcblower
+This is a good example because it is clearly written just for me.  It shows that the student knows about my work on invariant generation and has related  background (paper C and project G).
+\end{commentbox}
+
+Finally, profs. are very busy so don't take it personally if you don't get anything from them (though I would be very surprised if such thoughtful emails get no replies!).
 

### 2023-07-05 phd-cs-us.tex
updates

@@
 
-If you want to contact a prof. to ask about \emph{research opportunities}, or \emph{GTA/GRA} support, then \emph{yes}, I believe you should, but you really have to write your mail in a certain way. 
-First, read the prof's website, see if they are saying something about contacting them. Many profs. indicate what you should do to contact them (e.g., specific email subjects). 
-In general, the best way to catch the prof.'s attention is to \emph{customize your email} to that specific faculty.
+If you want to contact a prof. to ask about \emph{research opportunities}, or \emph{GTA/GRA} support, then \emph{yes}, I believe you should---it is \emph{worth it}. 
+First, read the prof's website, see if they say something about contacting them. Many profs. indicate how prospective students should contact them (e.g., specific email subjects). 
+In general, the best way to catch the prof.'s attention is to \emph{customize your email} to that faculty.
 For example, read their papers, know what they work on, and see if you are interested in their research.
-If you are, then you can send them an email talking how/why their work would match yours.
-In contrast, if you write a generic email that can be sent to multiple professors (e.g., if you just change some names and keywords in the email), you will not get any response.
+Then send them an email talking how/why their work would match yours.
+In contrast, if you write a generic email that can be sent to multiple professors (e.g., if you just change some names and keywords in the email), you will not likely get a response.
 
@@
 
-I read your TSE paper on numerical invariant generation, and I am interested in this line of dynamic invariant research.
+I read your TSE'21 paper on numerical invariant generation, and I am interested in this line of dynamic invariant research.
 I have worked (optional: with prof.
-Y at Z) on static analysis on C programs and I think it would be a good fit to tackle the spurious issues mentioned in your paper.
+Y at Z) on static program analysis and I think it could be used to tackle the spurious issues mentioned in your paper.
 I have a small paper at conference/workshop C and a project on symbolic execution at Github G.

### 2023-07-08 phd-cs-us.tex
updates

@@
 \usepackage{wrapfig}
+\setlength\intextsep{0pt} % remove extra space above and below in-line float
 \usepackage{xcolor}
@@
 
-First, I want to emphasize that PhD students in Computer
-Science \emph{do not} need to worry about funding, especially at good R1
+First, I want to emphasize that PhD students in Computer Science \emph{do not} need to worry about funding, especially at good R1
 universities in the US. If you are admitted, you will almost certainly
@@
 \begin{commentbox}[Vu]
-One of the reasons I create this document is that my colleagues at GMU are interested in 
-recruiting Vietnamese students and are surprised when seeing very few applications in Vietnam (e.g., each year our CS program receives more than 350 PhD applications, most of which are international but only 3--4 are from Vietnam).
+One of the reasons I create this document is that my colleagues at GMU are interested in recruiting Vietnamese students and are surprised when seeing very few applications in Vietnam (e.g., each year our CS program receives more than 350 PhD applications, most of which are international but only 3--4 are from Vietnam).
 In general the number of
@@
 These faculty have a wide-range of expertise and background to ensure diverse perspectives in the evaluation process.
-The size of the committee depends on the department size, e.g., around 15 at GMU.
-Each application is assigned to about \emph{three} faculty members, who will evaluate your profile and reach a consensus.
+The size and the review load of the committee depends on the department size.
+At GMU, the PhD committee typically has 15--20 faculty, and each committee member is assigned from 25--30 applications.
+Note that at GMU (and probably many other schools) we have a separate committee for the MS program.
+
+Each application is then assigned to about \emph{three} faculty members, who will evaluate your profile and reach a consensus.
 Note that while the assigned reviewers are likely the main ones deciding your application, \emph{every} faculty in the department will have access to your application and can provide inputs and opinions on your profile.
@@
 
-It's important to aim for the top venues in your field, which you can
-find on places such as CSRankings~\cite{csrankings}, which ranks CS programs based on how their faculty publish at top \emph{conferences}. Local conferences and non-English journals or conferences do
+It's important to aim for the top venues in your field. You can
+find top CS conferences at places such as CSRankings~\cite{csrankings}, which ranks CS programs based on how their faculty publish at top \emph{conferences}.
+Local conferences and non-English journals or conferences do
 not carry as much weight since their quality is often unknown to US faculty.
@@
 Many international students mention Scopus Q1, which consists of various journals from IEEE, Elsevier, and many other publishers.
-I don't know/recognize many of journals listed in Scopus Q1 and.
-This might be something to be mindful of, as \textbf{CS} faculty might not be too familiar with Scopus or journals listed in there, so devote sometime in your statement to discuss the significance of your papers.
+I don't know/recognize many of journals listed in Scopus Q1.
+This might be something to be mindful of, as \textbf{CS} faculty might not be too familiar with Scopus or journals listed there, so devote sometime in your statement to discuss the significance of your papers.
 \end{commentbox}
@@
 
-\begin{wrapfigure}{r}{0.47\textwidth}
-    \vspace{-0.35in}
+\begin{wrapfigure}{r}{0.47\linewidth}
       \begin{center}
-        \includegraphics[width=0.47\textwidth]{c6.png}
+        \includegraphics[width=1\linewidth]{c6.png}
       \end{center}
-    \vspace{-0.4in}
 \end{wrapfigure}
-Many students have letters written by the themselves and signed by their professors. These have little
-values. We can easily recognize them and will consider them weakness.
+Many students have letters written by themselves and signed by their professors. These have little
+values. We can recognize them and will consider them weakness.
 Similarly, many professors write generic letters for students, e.g., the students didn't do any
@@
 
+\begin{wrapfigure}{r}{0.52\linewidth}
+    \begin{center}
+      \includegraphics[width=1\linewidth]{c2.png}
+    \end{center}
+  \end{wrapfigure}
 While you might not have control over LORs \hyperref[sec:your-school]{where your go to school}, you do over your
@@
 
-\begin{wrapfigure}{r}{0.52\textwidth}
-\vspace{-0.35in}
-  \begin{center}
-    \includegraphics[width=0.52\textwidth]{c2.png}
-  \end{center}
-\vspace{-0.4in}  
-\end{wrapfigure}
 Finally, this is something easy to do, but is missed by many
@@
 So what can you do about this?
-several things such as asking your CS dept to put itself on CSRankings (it's the easiest way to get CS people to know about the school) and explaining about your school in your statement (and asking your LOR writer to do that too).
+several things including asking your CS dept to put itself on CSRankings (it's the easiest way to get CS people to know about the school) and explaining about your school in your statement (ask your LOR writer to do that too).
 Of course, if you're interested in working with Vietnamese professors, consider a PhD program in the US that  \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{have them}.
@@
 \begin{commentbox}[Vu]
-Sometime PhD admission committee will share a document such as \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Foreign-Top-Schools}{this one}, which lists the top schools in several countries.
-In some cases we ask people (faculty, students) if we think they know about the place.
+Sometime PhD admission committee in the US will share a document such as \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Foreign-Top-Schools}{this one}, which lists the top schools in several countries.
+In some cases we ask other faculty and students if we think they know about the place.
 For example, when I was a postdoc at UMD, people in admission committee ask me to evaluate applicants from Vietnam.
@@
 take them. However, they might be useful for international students from programs we are not familiar with.
+
+\begin{commentbox}[Vu]
+Personally I do not pay attention to GRE scores as I do not believe they are good indicator for research. However, if you have good GRE scores then you should list them in your application.
+\end{commentbox}
 
@@
 \begin{wrapfigure}{r}{0.48\textwidth}
-    \vspace{-0.3in}
       \begin{center}
@@
       \end{center}
-    \vspace{-0.2in}
     \end{wrapfigure}
@@
 \begin{wrapfigure}{r}{0.28\textwidth}
-    \vspace{-0.2in}
       \begin{center}
@@
 \begin{wrapfigure}{l}{0.5\textwidth}
-    \vspace{-0.4in}
       \begin{center}
@@
       \end{center}
-    \vspace{-0.9in}
     \end{wrapfigure}

### 2023-07-08 phd-cs-us.tex
updates

@@
 
-Several words about stipend.
+Some words stipend.
 First, the amount of stipend depends on the university, which in turns depend on various factors such as location (e.g., a stipend in Washington DC is likely higher than in Lincoln, Nebraska due to higher living cost).
 Second, a school year is (typically) 9-month in the US, so stipend is for 9 months (so divide by 9 for each month).
-Third, like for most source of income in the US, you will have to pay tax on your stipend.
-Finally, private universities might pay more for stipend.
+Third, like most source of income in the US, you will have to pay tax on your stipend.
+Fourth, CS department typically has higher stipend comparing to other study fields.
+Finally, private universities might pay more for stipend (but they could have higher "activity" or some other hidden fee, or you will have to pay some amount for each class credits).
+
+Students often complain their stipend being too low, but it is actually not bad and you can live comfortably with it.
+In many cases, it is enough to support your spouse and kids (many CS PhD students have their spouse and kids with them).
+So don't worry too much about stipend.
+If you're admitted to a good CS PhD program, you will be fine.
+
+For a full breakdown of how much a graduate student costs, see \S\ref{sec:ra-cost}.
 
@@
 You can learn about CS programs and research expertise of faculty using resources such as \href{https://csrankings.org}{CSRankings.org}, which is designed specifically to help prospective PhD students in Computer Science!
-You will be very surprised to learn that a school X that you didn't know much about has very strong research in your interested field Y (and vice versa, a school you thought highly off has no one works in the research field you're interested in).
+You will be very surprised to learn that a school that you didn't know much about has very strong research in your interested field (and vice versa, a school you thought highly off has no one works in the research field you're interested in).
 This is also a good way to learn about individual faculty (who works on what) and well-known CS conferences\footnote{In CS (and only in CS), conferences, not journals, are often the main venue to publish research finding.}.
@@
 
-\section{Acknowledgement}
+\section{History and Acknowledgement}
 
-Many people have contributed to this document.
+\paragraph{History} This document was conceived during a lunch with Craig Yu at GMU.
+We talked on about why we were not able to attract good Vietnamese international students, despite having a much stronger CS program than many schools that these students want to go to (part of the reason is described in \S\ref{sec:ranking}) and wished there were a way for international students to know about the US PhD programs (as well as for US faculty to understand more about international students and therefore have better chance of recruiting and working with them).
+I was also  a member of the large VietPhD group on Facebook and saw many questions from students about PhD programs.
+However, most active participant are students (and some few faculty) in non-CS fields or not in US.
+Like ChatGPT, their answers to CS PhD program in the US could be informative and helpful, but unfortunately not always accurate and sometimes leading to more confusion.
+So I thought it would be useful to have a document that is specific to CS PhD programs in the US from an insider prospective.
+
+I started writing this document in May 2023 and have been updating it since then (mostly around deadline time when I tend to procastinate :p).
+I have shared it with many students and faculty members and received many valuable feedback.
+I also hope that it will be updated and improved by many people in the future.
+I have put the source code of this document on \href{https://github.com/nguyenthanhvuh/phd-cs-us}{GitHub} so that anyone can contribute to it.
+I also hope that this document will be translated to Vietnamese and other languages so that it can reach more students.
+
+\paragraph{Acknowledgement} Many people have contributed to this document.
 Craig Yu (GMU) and Hakan Aydin (GMU) provided valuable input in the early version.

### 2023-08-13 phd-cs-us.tex
apply comments from Hung Le.

@@
 \renewcommand{\tablename}{Tab.}
-\title{\vspace{-1in} Demystifying the Computer Science \\PhD Admission in US Universities\\{\large A guide for Vietnamese and International Students}}
+\title{\vspace{-1in} Demystifying the Computer Science \\PhD Admission in US Universities\\{\large A guide for International Students}}
 %\date{\today}
@@
 Having been involved in PhD admissions for many years, I've
-realized that many \textbf{international students}, especially those from smaller countries such as \emph{Vietnam}, lack a clear understanding of
+realized that many \textbf{international students}, especially those from smaller countries such as Vietnam, lack a clear understanding of
 the Computer Science PhD admission process at US universities. This confusion not only
@@
 So I want to share some details about the PhD admission process and advice for those who are interested in applying for a \textbf{PhD in Computer Science in the US}.
-While this document is primarily intended for Vietnamese interested in computer science, it may be relevant to students from various countries and disciplines.
+While this document is primarily intended for students in Vietnam interested in CS, it may be relevant to students from various countries and disciplines.
 Moreover, although many examples are specifics for George Mason University, the information should be generalizable to other R1\footnote{An \href{https://en.wikipedia.org/wiki/List_of_research_universities_in_the_United_States}{R1 institution} in the US is a research-intensive university with a high level of research activity across various disciplines.} institutions in the US  (though \emph{very} top schools could be very selective, e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU).
@@
 At GMU, the PhD committee typically has 15--20 faculty, and each committee member is assigned from 25--30 applications.
-Note that at GMU (and probably many other schools) we have a separate committee for the MS program.
+Note GMU and UMass (and probably many other schools) have a separate committee for the MS program.
 
@@
 Also, if the application process were free for everyone, we would receive an overwhelming number of applications to review.
+
+\begin{commentbox}[Hung]
+    The application fee varies between schools, and is about \$70--\$100.
+I found this number pretty ridiculous.
+In most cases, there is no proper justification for that amount.
+Yes, the application fee could act as a filter, but a nominal amount would do the job.
+While a blanket waiver might not be feasible, universities should extend the eligibility of fee waiver applications to more countries.
+\end{commentbox}
 
@@
 
+\begin{commentbox}[Hung]
+The reputation of VinAI has been increasing steadily over the past few years; many of my colleagues heard about VinAI.
+\end{commentbox}
+
 Finally, \textbf{participating internationally recognized competitions} can also demonstrate your research potential.
@@
 \begin{wrapfigure}{r}{0.47\linewidth}
+\vspace{-0.2in}
       \begin{center}
@@
       \end{center}
+      \vspace{-0.1in}
 \end{wrapfigure}
@@
 
+\begin{commentbox}[Hung]
+A sad reality is that most professors in Vietnam \textbf{DO NOT} know how to write a good letter, or are lazy in writing letters hence delegate the writing to the students.
+Unfortunately, there is no easy solution to this problem.
+\end{commentbox}
+
 \begin{commentbox}[Didier]
@@
 \begin{wrapfigure}{r}{0.52\linewidth}
+\vspace{-0.2in}
     \begin{center}
@@
     \end{center}
+    \vspace{-0.2in}
   \end{wrapfigure}
@@
 So I will not talk too much about statements.
-In short, discuss about your research vision and convince us that you can achieve it through your experience, e.g., published papers, or if you work on some projects by yourself, talk about it.
+In short, discuss about your research goal and vision, and convince us that you can achieve it through your experience, e.g., published papers, or if you work on some projects by yourself, talk about it.
 Also, use the statement to talk about stuff that admission committee members might not know about, e.g., your Github project with 1K+ stars or  your regular contributions to well-known open-source projects.
+
+\begin{commentbox}[Hung]
+I think asking for a “research vision” from a Ph.D.
+applicant is too much.
+Even people graduated with a Ph.D.
+have a great difficulty in explicating their researh vision.
+In my opinion, you should focus on showing (not telling) your research passion (if you do have).
+\end{commentbox}
 
@@
 
+\begin{commentbox}[Hung]
+I personally don’t care about which schools students attended. There are many other ways to evaluate an applicant.
+\end{commentbox}
+
 \subsection{Grades/GREs}\label{sec:grades}
@@
 
-This can be an issue for students in many top international universities where the competition is so high that very good students students can still have low ranking (and be overlooked by Admission committee).
+This can be an issue for students in many top international universities where the competition is so high that very good students students can still have low ranking from these schools (and be overlooked by Admission committee).
 So what to do with this? well, same as \hyperref[sec:your-school]{before}, e.g., put a note about this in your statement and ask your LoR writers to talk about it.
@@
 \subsection{Interview}
+Sometimes a faculty wants to interview an applicant to make a decision.
+Typically, this means that they are considering admitting you.
+If they are not interested in your application, they will not proceed with an interview.
 
-Sometime a faculty wants to interview an applicant to make a decision. This means they are leaning toward admitting you (if we don't like your application, we will not bother doing the interview).
-
-An interview lasts about 15--30 mins, and one implicit thing you will be evaluated on is whether you can communicate effectively (i.e., speak/understand English).
-You will also get chance to ask questions about the university so think of something to ask (just the same as interviewing at a company).
+An interview often lasts about 15-30 minutes, and one important aspect of evaluation is your ability to effectively communicate, including speaking and understanding English.
+In addition, during the interview, you will have the opportunity to ask questions about the university.
+It's essential to prepare some thoughtful questions, similar to interviewing at a company.
 
@@
 For very strong candidates, the interview is actually to recruit them.
-In some cases a faculty interviews a candidate that they see potentials and wants to argue for admission, i.e., without the interview, that application is likely rejected.
-In any case, getting interview means you have a very good chance of being admitted.
+In some cases a faculty interviews a candidate that they see potentials and want to advocate for their admission.
+Without the interview, such applications may be more likely to be rejected.
+Overall, being selected for interview means a very good chance of being admitted.
 \end{commentbox}
@@
 
-Around March you should hear back from most PhD programs that you applied to (if not, send email and ask).
-If you have offers, congratulations!
-Now you're at a different game because the schools that admit you will now try to get you to accept them!
+By around March you should hear back from most PhD programs you applied to . If you haven't heard back, reach our via email and ask about the status of your application.
+If you receive offers, congratulations!
+Now you're at a different game because the schools that have admitted you will try to get you to accept them!
 You will likely have to make your decision by around April 15.
@@
 
- Show something that makes you \textbf{stand out}, e.g., are you a female or a minority in CS (research for this on Google)?
+ Show something that makes you \textbf{stand out}, e.g., do you have a degree or background in music and want to integrate that to CS?
+are you a female or a minority in CS (research for this on Google)?
 Do you participate in outreach activities that help increase diversity and inclusion in CS?
-All of these are important in CS in the US.
+All of these are unique and would get noticed from reviewers.
 
@@
 \begin{wrapfigure}{r}{0.48\textwidth}
+\vspace{-0.2in}
       \begin{center}
@@
       \end{center}
+      \vspace{-0.2in}
     \end{wrapfigure}
@@
 Tenured faculty (e.g., associate and full profs.) are more likely to be older, more well-known, and have more experience in managing students.
-However, they might not push you as hard (they don't have to, they already got tenured) and expect you to figure things out yourself (so you need to be very independent).
+However, they might not push you as hard (they don't have to, they already got tenured).
+They might also expect you to figure things out yourself (so you need to be very independent).
 Some tenured faculty are also no longer active in research and more involved with administrative duties or with their startup companies.
@@
 
-\subsection{Should you contact a US professor? How to do so ``correctly''?}
+\subsection{Should you contact a US professor? What to do to get a desired reply?}
 
 Faculty received many "cold" e-mails from international students seeking for admission, TA, and RA.
-Most of the time, we ignore these emails; but in rare occasion we do answer them.
+Most of the time, we ignore these emails, but in some rare occasions we do answer them.
 So how to write an email that can get our attention?
@@
 
-I read your TSE'21 paper on numerical invariant generation, and I am interested in this line of dynamic invariant research.
+I am writing to inquire about potential research opportunities as a GRA in your group at GMU. Currently I am an undergraduate student in Computer Science at UNIV and plan to graduate in May 2023.
+
+I have read your TSE'21 paper on numerical invariant generation, and I am interested in this line of dynamic invariant research.
 I have worked (optional: with prof.
@@
 
-\subsection{How long does it take to complete the PhD program?}
+\subsection{How long does it take to complete the CS PhD program?}
 
 \begin{wrapfigure}{r}{0.28\textwidth}
+\vspace{-0.2in}
 \begin{center}
@@
 \end{center}
+\vspace{-0.1in}
 \end{wrapfigure}
@@
 
+\subsection{PhD in other related field such as CE, IST, Cybersecurity, and Stats}
+
+In many cases you do not need to do a PhD in CS to work in your area of interested.
+For example, in addition to having a traditional CS department, GMU has IST and Cybersecurity departments with faculty who have PhD in CS and work on CS topics (e.g., AI, Security, Robotics).
+In short, it is totally possible that you still get to do CS research and publish in CS-related venues even if you're not in a traditional CS program.
+It is very common to see faculty with CS PhD in a non-CS department as well as faculty with non-CS PhD in CS department.
+
+It is worth it to note that if your intention is a PhD in CS, then you likely need to be in the CS dept \emph{and} advised by a CS faculty.
+For example, if your advisor is not in CS, then you likely cannot not get a CS degree.
+In fact, if a faculty is not in CS, they likely cannot be the main advisor of a CS PhD student---they may \emph{co-advise} or be in the PhD dissertation committee, but your main advisor will need to be a tenured or tenure-track faculty in CS.
+If in doubt, you should check with the CS department for their requirements.
+
+Note that for the purpose of helping students interested in a PhD in CS, CSRankings only includes tenured or tenured track faculty who can advise CS PhD students.
+I also have compiled a \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{list} of Vietnamese faculty who can advise PhD students.
+
 \subsection{How do I address a professor?}
 
-\begin{wrapfigure}{l}{0.5\textwidth}
+\begin{wrapfigure}{l}{0.5\linewidth}
+\vspace{-0.3in}
       \begin{center}
-        \includegraphics[width=0.5\textwidth]{c5.png}
+        \includegraphics[width=1.0\linewidth]{c5.png}
       \end{center}
+      \vspace{-0.3in}
     \end{wrapfigure}
@@
 
-I started writing this document in May 2023 and have been updating it since then (mostly around deadline time when I tend to procastinate :p).
-I have shared it with many students and faculty members and received many valuable feedback.
-I also hope that it will be updated and improved by many people in the future.
+I started writing this document in May 2023 and have been updating it since then (mostly around deadline time when I tend to procrastinate!).
 I have put the source code of this document on \href{https://github.com/nguyenthanhvuh/phd-cs-us}{GitHub} so that anyone can contribute to it.
-I also hope that this document will be translated to Vietnamese and other languages so that it can reach more students.
 
 \paragraph{Acknowledgement} Many people have contributed to this document.
-Craig Yu (GMU) and Hakan Aydin (GMU) provided valuable input in the early version.
+Craig Yu (GMU), Hakan Aydin (GMU), Hung Le (UMass) provided valuable input in the early version.
 Other GMU faculty members also have provided feedback and contributions.
-Many students including Didier (GMU), Thanh (Melbourne), Dat (Melbourne) have contributed valuable questions and feedback.
+Many students including Didier (GMU), Thanh (Melbourne), and Dat (Melbourne) have contributed valuable questions and feedback.
 \textbf{Thank you!}

### 2023-08-13 phd-cs-us.tex
Updates from Overleaf

@@
 statements give proper explanation). This is especially true if you
-have bad grades in relevant, e.g., CS and Math, courses.
+have bad grades in relevant courses (e.g., Math and CS).
 
@@
 One way to mitigate this is making these issues known to admission committee in your statement.
-Schools with Vietnamese profs are probably aware of them, but in general your letter writers/you can explicitly mention these in their letters/ your statement.
-\end{commentbox}
-
-\paragraph{GRE} Most CS programs in the US \emph{no longer require GREs}, so you don't need to
-take them. However, they might be useful for international students from programs we are not familiar with.
+Universities with Vietnamese profs are probably aware of them, but in general your letter writers \emph{and} you can explicitly mention these in their letters and your statement.
+\end{commentbox}
+
+\paragraph{GRE} Most CS programs in the US \emph{no longer require GREs}, so you don't need to take them.
+However, they might be useful for international students from programs we are not familiar with.
 
@@
 
-By around March you should hear back from most PhD programs you applied to . If you haven't heard back, reach our via email and ask about the status of your application.
+By around March you should hear back from most PhD programs you applied to . If you haven't heard back, reach out via email and ask about the status of your application.
 If you receive offers, congratulations!
 Now you're at a different game because the schools that have admitted you will try to get you to accept them!
-You will likely have to make your decision by around April 15.
+You will have to make your decision by around April 15.
 
@@
 TA is paid by the department for you to help with teaching.
-Finally, fellowship is an independent funding that come from a school, a company, or an organization.
+Finally, fellowship is an independent funding that can come from a school, a company, or an organization.
 Tab.~\ref{tab:funding} summarizes the differences.
@@
 Note that funding is typically more available for PhD students than 
-Masters and especially undergraduate studies, which typically have no funding (and they also have to pay international tuition, which is very expensive!).
+Masters. Undergraduate students typically have no funding and also have to pay international tuition, which is very expensive!.
 
@@
 
-Some words stipend.
+Several things about stipends.
 First, the amount of stipend depends on the university, which in turns depend on various factors such as location (e.g., a stipend in Washington DC is likely higher than in Lincoln, Nebraska due to higher living cost).
@@
 Fourth, CS department typically has higher stipend comparing to other study fields.
-Finally, private universities might pay more for stipend (but they could have higher "activity" or some other hidden fee, or you will have to pay some amount for each class credits).
+Finally, private universities might pay more for stipend (but they could have higher "activity" or some other hidden fee, or you will have to pay some fee for each class credits).
 
@@
 If you're admitted to a good CS PhD program, you will be fine.
+A good school would know that it has to be competitive, otherwise it will loose good students to other schools.
+For example, at GMU, every year we discuss about improving the benefits, and especially stipend, for our graduate students.
 
@@
 The college and department will set the rate for 9-month graduate assistant stipend.
-TA will likely be that amount but RA might be higher depending on the stage of the student (1st year vs ABD\footnote{All but dissertation: really close to graduate.}) and the prof.
+TA, which is paid by the department, will likely be that amount but RA might be higher depending on the stage of the student (1st year vs ABD\footnote{All but dissertation: really close to graduate.}) and the prof.
 \tcblower
@@
 During a semester, a TA might work with several courses and professors (not necessary their advisor).
-TA funding typically is not available during the summer, which has no school.
+TA funding typically is not available during the summer, which has much fewer courses.
 
@@
 who would give you RA right away, it's very likely you will get admitted.
-For example, if a prof., even if not in PhD admission committee, wants to work with and funds you, the PhD admission committee will respect that decision and admit your application (unless your application has many red flags).
+For example, if a prof., even if not in PhD admission committee, wants to work with and funds you, the PhD admission committee will respect that decision and admit you (unless your application has many red flags).
 \end{commentbox}
@@
 
-\subsection{What can you do to increase your admission chance?}
-
- Show something that makes you \textbf{stand out}, e.g., do you have a degree or background in music and want to integrate that to CS?
+\subsection{What can you do to increase your admission chance?}\label{sec:improve-your-chance}
+
+ Show something that makes you \textbf{stand out}, e.g., do you have a degree or background in music and want to integrate that with CS?
 are you a female or a minority in CS (research for this on Google)?
@@
 You can learn about CS programs and research expertise of faculty using resources such as \href{https://csrankings.org}{CSRankings.org}, which is designed specifically to help prospective PhD students in Computer Science!
-You will be very surprised to learn that a school that you didn't know much about has very strong research in your interested field (and vice versa, a school you thought highly off has no one works in the research field you're interested in).
+You will be very surprised to learn that a school that you didn't know much about can have very strong research in your interested topic (and vice versa, a school you thought highly about has no one works in the research field you're interested in).
 This is also a good way to learn about individual faculty (who works on what) and well-known CS conferences\footnote{In CS (and only in CS), conferences, not journals, are often the main venue to publish research finding.}.
@@
 
-\begin{commentbox}[Dat] Most Vietnamese students, including those from top schools, do not know about CSRankings.
+\begin{commentbox}[Dat] Most Vietnamese students, including those from top schools, \textbf{do not know} about CSRankings.
 May be applicants who worked at top research places such as VinAI would know about it.
@@
 
-More generally, rankings can be superficial and you need to do more research to be informed and make better decision.
+However, in general, rankings can be superficial and you need to do more research to be informed and make better decision.
 For example, if you get admissions to several places, you should contact profs.
@@
 
-If you want to contact a prof. to ask about \emph{research opportunities}, or \emph{GTA/GRA} support, then \emph{yes}, I believe you should---it is \emph{worth it}. 
+If you want to contact a prof.
+to ask about \emph{research opportunities}, or \emph{GTA/GRA} support, then \emph{yes}, I believe you should---it is \emph{worth it}.
+However, you need to do it in a right way.
+
 First, read the prof's website, see if they say something about contacting them. Many profs. indicate how prospective students should contact them (e.g., specific email subjects). 
@@
 Then send them an email talking how/why their work would match yours.
-In contrast, if you write a generic email that can be sent to multiple professors (e.g., if you just change some names and keywords in the email), you will not likely get a response.
+In contrast, if you write a generic email that can be sent to multiple professors (e.g., if you just change some names and keywords in the email), you will not get a response.
 
@@
 Yes, as long as you can demonstrate you are ready for CS PhD research through research experiences, LoRs, statements, etc as mentioned.
-You might be even able to leverage this to make your profile stand out.
-
-\subsection{Is an MS degree required for admission to PhD?}
+You might be even able to leverage this to make your profile stand out as mentioned in \S\ref{sec:improve-your-chance}.
+
+\subsection{Is an MS degree required for admission to PhD in CS?}
 No. In fact, student with BS can get MS degree ``along the way'' to PhD.  However, MS can help if it gives research experience or is from a more well-known school than your undergrad institution.
@@
 If in doubt, you should check with the CS department for their requirements.
-
-Note that for the purpose of helping students interested in a PhD in CS, CSRankings only includes tenured or tenured track faculty who can advise CS PhD students.
+For this specific reason,  CSRankings only includes tenured or tenured track faculty who can advise CS PhD students.
 I also have compiled a \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{list} of Vietnamese faculty who can advise PhD students.
@@
       \end{center}
-      \vspace{-0.3in}
+      \vspace{-0.5in}
     \end{wrapfigure}
@@
 I've seen many international students write \textbf{Prof.} or \textbf{Dr.} \textbf{Firstname Lastname}.
-Writing like that makes it like you copy and paste the names, so no need to do so,  just Prof.
+Writing like that makes it like you copy and paste names, so no need to do so,  just Prof.
 or Dr.
@@
 As you know that prof.
-better and depends on their preference, you can call them by their Firstname.
+better and depends on their preference, you may call them by their Firstname.
+
+Note that at some universities, especially those that do not offer PhD studies, will use the formal \textbf{Dr.} title, i.e., Dr. Lastname.  So just observe and follow how it is done at your place.
 

### 2023-08-15 phd-cs-us.tex
Updates from Overleaf

@@
 
+\renewcommand{\figurename}{Fig.}
+\renewcommand{\tablename}{Tab.}
 \def\Section{\S}
-
 \newcommand{\mycomment}[3][\color{blue}]{{#1{{#2}: {#3}}}}
@@
 \newcommand{\red}[1]{{\color{red}{#1}}}
-\renewcommand{\figurename}{Fig.}
-\renewcommand{\tablename}{Tab.}
+
 \title{\vspace{-1in} Demystifying the Computer Science \\PhD Admission in US Universities\\{\large A guide for International Students}}
@@
 So I want to share some details about the PhD admission process and advice for those who are interested in applying for a \textbf{PhD in Computer Science in the US}.
-While this document is primarily intended for students in Vietnam interested in CS, it may be relevant to students from various countries and disciplines.
-Moreover, although many examples are specifics for George Mason University, the information should be generalizable to other R1\footnote{An \href{https://en.wikipedia.org/wiki/List_of_research_universities_in_the_United_States}{R1 institution} in the US is a research-intensive university with a high level of research activity across various disciplines.} institutions in the US  (though \emph{very} top schools could be very selective, e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU).
-
-In addition, this document can help \emph{US faculty and admission committee} gain a better understanding of international students and their cultural differences.
+While this document is primarily intended for students interested in CS, it may be relevant to students from various disciplines.
+Furthermore, although many examples are specifics for George Mason University, the information should be generalizable to other R1\footnote{An \href{https://en.wikipedia.org/wiki/List_of_research_universities_in_the_United_States}{R1 institution} in the US is a research-intensive university with a high level of research activity across various disciplines.} institutions in the US  (though \emph{very} top schools could be very selective, e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU).
+
+In addition, this document can help \textbf{US faculty and admission committee} gain a better understanding of international students and their cultural differences.
 By appreciating and leveraging these differences, CS programs in the US can attract larger and more competitive application pools from international students.
@@
 
-\begin{commentbox}[Hung]
-    The application fee varies between schools, and is about \$70--\$100.
+%\begin{commentbox}[Hung]
+%    The application fee varies between schools, and is about \$70--\$100.
 I found this number pretty ridiculous.
@@
 While a blanket waiver might not be feasible, universities should extend the eligibility of fee waiver applications to more countries.
-\end{commentbox}
+%\end{commentbox}
 
@@
 As mentioned, this will hurt the applicants (admission committee members are actually quite good at determining this)!
-I was also told that some people did this so that the applicants do not go abroad or only go to places where they want the students to go to.
+Multiple students mentioned that  some professor did this so that the applicants do not go abroad or only go to places where they want the students to go to.
 \tcblower
@@
 \begin{commentbox}[Vu]
-I always read the research statement first and then LORs. If I am
+I always read the research statement (and the LORs) first. If I am
 persuaded by then, I would skim over other factors and advocate for
@@
 several things including asking your CS dept to put itself on CSRankings (it's the easiest way to get CS people to know about the school) and explaining about your school in your statement (ask your LOR writer to do that too).
-Of course, if you're interested in working with Vietnamese professors, consider a PhD program in the US that  \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{have them}.
+Of course, if you're interested in working with Vietnamese professors, consider  \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{CS PhD programs in the US that have them}.
+
+\begin{commentbox}[Hung]
+I personally don’t care about which schools students attended. There are many other ways to evaluate an applicant.
+\end{commentbox}
 
@@
 During my time at UNL and now here at GMU, I have looked at Vietnamese applications (whether they are assigned to me or not) and provide input to their reviewers, e.g., X is the top tech school in Vietnam and so it should be \emph{top} instead of \emph{unknown} foreign, which makes a huge difference.
-\end{commentbox}
-
-\begin{commentbox}[Hung]
-I personally don’t care about which schools students attended. There are many other ways to evaluate an applicant.
 \end{commentbox}
@@
 
-\paragraph{GRE} Most CS programs in the US \emph{no longer require GREs}, so you don't need to take them.
+\paragraph{GRE} Most CS programs in the US \textbf{no longer require GREs}, so you don't need to take them.
 However, they might be useful for international students from programs we are not familiar with.
@@
 
-Most schools will have an \textbf{Open House}, which is a great resource to learn about the school, department, faculty, research, living, etc.
+Most schools will have \textbf{Open House} events, which are a great resource to learn about the school, department, faculty, research, living, etc.
 During the Open House, you get a chance to talk to individual faculty and current students.
@@
 
-In short, if you can come to the Open House, do come.
-But if you're international student outside of the US, then likely you cannot come.
-So see if you can attend it virtually and meet with individual faculty.
+In short, if you can, do come to Open House.  Even if you can't come in person, you still can attend virtually and meet with individual faculty.
 
@@
     \textbf{Pros}                 & Research Freedom         & Get to do research              & Research Freedom                      \\
-    \textbf{Cons}                 & TA, Uncertain            & Research restriction, Uncertain & Competitive, limited             \\
+    \textbf{Cons}                 & Teaching Duties, Uncertain            & Research Restriction, Uncertain & Competitive, Limited             \\
     \bottomrule
@@
 
-Fellowship is another type of funding that the student needs to apply for (e.g., from school, industries, government).
+Fellowship is another type of funding that students can apply for (e.g., from school, industries, government).
 Fellowships are typically competitive and generous, and gives pretty much all benefits tuition/insurance that a TA/RA has.
-Moreover, it often gives higher stipend (including summer) and opens doors for job opportunities (e.g., internship).
+Moreover, they often give higher stipend (including summer) and open doors for job opportunities (e.g., internship).
 For example, a student with a Microsoft fellowship will likely get an internship at Microsoft.
@@
 
-\textbf{How to get Fellowship?} You apply for them.
-The US government has many fellowships that would likely require US citizenship or residency.
-However, tech companies including Google, Microsoft, Facebook, IBM have fellowships that international students can apply for.
+\textbf{How to get Fellowship?} You need to apply for them.
+The US government has many fellowships, though they would likely require US citizenship or residency.
+However, tech companies including Google, Microsoft, Facebook have fellowships that international students can apply for.
 
@@
 
- Show something that makes you \textbf{stand out}, e.g., do you have a degree or background in music and want to integrate that with CS?
-are you a female or a minority in CS (research for this on Google)?
+ Show something that makes you \textbf{stand out}, e.g., do you have a degree or background in \emph{biology} or \emph{music} and want to integrate them with CS?
+are you a female or a minority in CS (research for "URM minority in CS in the US" on Google)?
 Do you participate in outreach activities that help increase diversity and inclusion in CS?
@@
 
-Even if you do not have research experience, you can talk about your personal projects.
+Even if you do not have research experience, you can talk about your personal projects, as long as they can help show you can do research.
 For example, if you have an open-source project on Github that is used by many people, has lots of stars in Github, do talk about it.
@@
 
-\subsection{How to rank or select a CS program?}
+\begin{commentbox}[Vu]
+In his \href{https://matt.might.net/articles/how-to-apply-and-get-in-to-graduate-school-in-science-mathematics-engineering-or-computer-science/}{post}, Matt Might was initially unsure about an application.
+However, upon learning that the applicant had led a 100km hike in the Himalayas, he decided to accept the applicant.
+
+\tcblower
+This shows the student has the persistence and determination required for research. I would also advocate for accepting that student!
+\end{commentbox}
+
+%It is worth noting that in many cases,  an application could be rejected by most reviewers, however have a "champion", who doesn't even have to be in the Admission comittee, that strongly advocate for your profile.
+
+\subsection{How to rank or select a CS graduate program?}
    \begin{center}
@@
 
-International students not familiar with US universities often put them into \emph{two} bins:  (i) very top school that they dream about such as Stanford, MIT, Princeton, Harvard and (ii) everything else.
+International students not familiar with US universities often put them into \emph{two} bins:  (i) very top schools that they dream about such as Stanford, MIT, Princeton, Harvard and (ii) everything else.
 Sometimes they rank CS programs using the reputations non-CS programs such as medical, math, or physics.
@@
 You will be very surprised to learn that a school that you didn't know much about can have very strong research in your interested topic (and vice versa, a school you thought highly about has no one works in the research field you're interested in).
-This is also a good way to learn about individual faculty (who works on what) and well-known CS conferences\footnote{In CS (and only in CS), conferences, not journals, are often the main venue to publish research finding.}.
+This is also a good way to learn about individual faculty (who works on what) and well-known CS conferences\footnote{In CS (and probablyo nly in CS), conferences, not journals, are often the main venue to publish research finding.}.
 \S\ref{sec:ranking} gives the top 50 CS programs in the US according to CSRankings.
@@
     \end{wrapfigure}
-The short answer is that tenured-track faculty (e.g., assistant professors) are more likely to be young and active in research (they have to, in order to get tenure).
+The short answer is that tenured-track faculty such as assistant professors are more likely to be young and active in research (they have to, in order to get tenure).
 Thus, they will likely have more time to work with you and push you to do research and publish.
-However, they may not have as much experience in managing students and may not have as much funding as tenured faculty.
-
-Tenured faculty (e.g., associate and full profs.) are more likely to be older, more well-known, and have more experience in managing students.
+However, they may not have as much experience in managing students and may not have as much funding (yet).
+
+Tenured faculty, e.g., associate and full profs., are more likel older, more well-known, and have more experience in managing students.
 However, they might not push you as hard (they don't have to, they already got tenured).
@@
 
-First, read the prof's website, see if they say something about contacting them. Many profs. indicate how prospective students should contact them (e.g., specific email subjects). 
+First, read the prof's website, see if they say something about contacting them. Many profs. indicate how prospective students should contact them (e.g., using specific email subjects). 
 In general, the best way to catch the prof.'s attention is to \emph{customize your email} to that faculty.
@@
 
-Yes, as long as you can demonstrate you are ready for CS PhD research through research experiences, LoRs, statements, etc as mentioned.
+Absolutely, as long as you can demonstrate you are ready for CS PhD research through research experiences, LoRs, statements, etc as mentioned.
 You might be even able to leverage this to make your profile stand out as mentioned in \S\ref{sec:improve-your-chance}.
@@
 In many cases you do not need to do a PhD in CS to work in your area of interested.
-For example, in addition to having a traditional CS department, GMU has IST and Cybersecurity departments with faculty who have PhD in CS and work on CS topics (e.g., AI, Security, Robotics).
+For example, in addition to having a traditional CS department, GMU has IST and Cybersecurity departments, which have faculty  with PhD in CS and work on CS topics (e.g., AI, Security, Robotics).
 In short, it is totally possible that you still get to do CS research and publish in CS-related venues even if you're not in a traditional CS program.
@@
 
-It is worth it to note that if your intention is a PhD in CS, then you likely need to be in the CS dept \emph{and} advised by a CS faculty.
-For example, if your advisor is not in CS, then you likely cannot not get a CS degree.
-In fact, if a faculty is not in CS, they likely cannot be the main advisor of a CS PhD student---they may \emph{co-advise} or be in the PhD dissertation committee, but your main advisor will need to be a tenured or tenure-track faculty in CS.
+However, if your intention is a PhD in CS, then you likely need to be in the CS dept \emph{and} advised by a CS faculty.
+In fact, if a faculty is not in CS, it is unlikely that they can be the \emph{main} advisor of a CS PhD student---they may \emph{co-advise} or be in the PhD dissertation committee, but your main advisor will need to be a tenured or tenure-track faculty in CS.
 If in doubt, you should check with the CS department for their requirements.
-For this specific reason,  CSRankings only includes tenured or tenured track faculty who can advise CS PhD students.
+For this specific reason,  CSRankings includes only tenured or tenured track faculty who can advise CS PhD students.
 I also have compiled a \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{list} of Vietnamese faculty who can advise PhD students.
@@
 \paragraph{Acknowledgement} Many people have contributed to this document.
-Craig Yu (GMU), Hakan Aydin (GMU), Hung Le (UMass) provided valuable input in the early version.
+Profs.
+Craig Yu (GMU), Hakan Aydin (GMU), and Hung Le (UMass) provided valuable input in the early version.
 Other GMU faculty members also have provided feedback and contributions.
-Many students including Didier (GMU), Thanh (Melbourne), and Dat (Melbourne) have contributed valuable questions and feedback.
+Many students including Didier (GMU), Thanh (Melbourne), and Dat (Melbourne) also have contributed valuable questions and feedback.
 \textbf{Thank you!}

### 2023-08-22 phd-cs-us.tex
Updates from Overleaf

@@
 
-\title{\vspace{-1in} Demystifying the Computer Science \\PhD Admission in US Universities\\{\large A guide for International Students}}
+\title{\vspace{-1in} Demystifying the Computer Science \\PhD Admission in US Universities\\{\large A Guide for International Students}}
 %\date{\today}
@@
 \end{commentbox}
+
+For Vietnamese students, it's worth highlighting the \href{https://vef2.org/}{VEF 2.0 program}, which has played a crucial role in aiding many good students in gaining acceptance to top-tier PhD programs in the US.
+VEF 2.0 follows an intriguing model where  US faculty members from leading institutions are invited to conduct rigorous interviews with VEF students and subsequently provide reference letters on their behalf.
+Despite the limited interaction between the interviewer and the interviewee (primarily confined to the interview itself), the impact of these reference letters is remarkably positive.
+This fact is underscored by the noteworthy success of numerous VEF 2.0 students in securing admission to top CS PhD programs across the US.
 
@@
 \tcblower
-This shows the student has the persistence and determination required for research. I would also advocate for accepting that student!
+I would also advocate for that student as this shows they have the persistence and determination required for research.
 \end{commentbox}
@@
    \begin{center}
-    \includegraphics[scale=0.4]{c1.png}
+    \includegraphics[scale=0.5]{c1.png}
    \end{center}
@@
 You will be very surprised to learn that a school that you didn't know much about can have very strong research in your interested topic (and vice versa, a school you thought highly about has no one works in the research field you're interested in).
-This is also a good way to learn about individual faculty (who works on what) and well-known CS conferences\footnote{In CS (and probablyo nly in CS), conferences, not journals, are often the main venue to publish research finding.}.
+This is also a good way to learn about individual faculty (who works on what) and well-known CS conferences\footnote{In CS (and probably only in CS), conferences, not journals, are often the main venue to publish research finding.}.
 \S\ref{sec:ranking} gives the top 50 CS programs in the US according to CSRankings.
@@
 
+\begin{commentbox}[Hung]
+I always encourage the students I admitted to talk with my students and the students of other faculty in other schools who admitted them.
+You will unlikely hear straight-out complaints from current students in a professor’s group.
+But sometimes what is important are things that they (current students) don’t tell you.
+Pay attention to their "level of excitement" being in the group.
+\end{commentbox}
+
 \subsection{Tenured or tenure-track faculty? Who do you choose?}
@@
       \begin{center}
-        \includegraphics[width=0.48\textwidth]{c8.png}
+        \includegraphics[width=1\linewidth]{c8.png}
       \end{center}
@@
 
-Tenured faculty, e.g., associate and full profs., are more likel older, more well-known, and have more experience in managing students.
+Tenured faculty, e.g., associate and full profs., are likely older, more well-known, and have more experience in managing students.
 However, they might not push you as hard (they don't have to, they already got tenured).
@@
 
-Absolutely, as long as you can demonstrate you are ready for CS PhD research through research experiences, LoRs, statements, etc as mentioned.
+Yes, as long as you can demonstrate you are ready for CS PhD research through research experiences, LoRs, statements, etc as mentioned.
 You might be even able to leverage this to make your profile stand out as mentioned in \S\ref{sec:improve-your-chance}.
@@
 
-\begin{wrapfigure}{r}{0.28\textwidth}
+\begin{wrapfigure}{r}{0.35\textwidth}
 \vspace{-0.2in}
 \begin{center}
-\includegraphics[width=0.28\textwidth]{c4a.png}
+\includegraphics[width=1\linewidth]{c4a.png}
 \end{center}
-\vspace{-0.1in}
+\vspace{-0.3in}
 \end{wrapfigure}
@@
 
-    The first 2 years you spend on coursework, finding an adviser, learning research.
+    The first 2 years you spend on coursework (somewhat equivalent to an MS study), finding an adviser, learning research.
 The next 2--3 years you focus on your research, form dissertation topic, and get results published.
@@
 
-\subsection{PhD in other related field such as CE, IST, Cybersecurity, and Stats}
+\subsection{PhD in other related field such as CE, IST, Cybersecurity}
 
@@
 For example, in addition to having a traditional CS department, GMU has IST and Cybersecurity departments, which have faculty  with PhD in CS and work on CS topics (e.g., AI, Security, Robotics).
-In short, it is totally possible that you still get to do CS research and publish in CS-related venues even if you're not in a traditional CS program.
+So it is totally possible that you still get to do CS research and publish in CS-focused venues even if you're not in a traditional CS program.
 It is very common to see faculty with CS PhD in a non-CS department as well as faculty with non-CS PhD in CS department.
@@
       \begin{center}
-        \includegraphics[width=1.0\linewidth]{c5.png}
+        \includegraphics[width=1\linewidth]{c5.png}
       \end{center}

### 2023-09-04 phd-cs-us.tex
Updates from Overleaf

@@
 
-\begin{wrapfigure}{r}{0.52\linewidth}
-\vspace{-0.2in}
     \begin{center}
-      \includegraphics[width=1\linewidth]{c2.png}
+      \includegraphics[scale=0.5]{c2.png}
     \end{center}
-    \vspace{-0.2in}
-  \end{wrapfigure}
+
 While you might not have control over LORs \hyperref[sec:your-school]{where your go to school}, you do over your
@@
 
-\begin{wrapfigure}{r}{0.48\textwidth}
-\vspace{-0.2in}
       \begin{center}
-        \includegraphics[width=1\linewidth]{c8.png}
+        \includegraphics[scale=0.3]{c8.png}
       \end{center}
-      \vspace{-0.2in}
-    \end{wrapfigure}
+
 The short answer is that tenured-track faculty such as assistant professors are more likely to be young and active in research (they have to, in order to get tenure).
@@
 
-\begin{wrapfigure}{r}{0.35\textwidth}
-\vspace{-0.2in}
 \begin{center}
-\includegraphics[width=1\linewidth]{c4a.png}
+\includegraphics[scale=0.3]{c4a.png}
 \end{center}
-\vspace{-0.3in}
-\end{wrapfigure}
-    Typically, 5--7 years for PhD in CS in the US.
-This can be longer than CS PhD at universities in other countries, which might require MS first (recall that US PhD programs do not require MS and you can get MS along the way to PhD).
-Within these 5--7 years, CS PhD students in the US also often take a ``leave of absence'' for 1--2 semesters to do internship at companies and research labs.
+
+    Typically, it takes 5--7 years for CS PhD in the US.
+This can be longer than CS PhDs at universities in other countries, which might require MS first (recall that US PhD programs do not require MS and you can get MS along the way to PhD).
+Within these 5--7 years, CS PhD students often take a ``leave of absence'' for 1--2 semesters to do internship at companies and research labs.
 
@@
 
-\begin{wrapfigure}{l}{0.5\linewidth}
-\vspace{-0.3in}
       \begin{center}
-        \includegraphics[width=1\linewidth]{c5.png}
+        \includegraphics[scale=0.5]{c5.png}
       \end{center}
-      \vspace{-0.5in}
-    \end{wrapfigure}
+
 If you don't know the professor (e.g., first email contact), then use \textbf{Prof.
@@
 Lastname.
-
-If you don't know the prof., do not use Mr.
+Moreover, if you don't know the prof., do not use Mr.
 or Mrs., or Firstname.

### 2023-09-04 phd-cs-us.tex
Updates from Overleaf

@@
 
+\epigraph{"How is education supposed to make me feel smarter?
+Besides, every time I learn something new, it pushes some of the old stuff out of my brain.
+Remember when I took that home wine making course, and I forgot how to drive?"}{The Simpsons}
+
 After you submit your PhD application (usually in December), it will be first screened
@@
 
+\epigraph{Son, if you really want something in this life, you have to work for it. Now quiet! They’re about to announce the lottery numbers.}{The Simpsons}
+
 The primary focus of the admissions committee is to \textbf{evaluate your background and interest in research} since a PhD in Computer Science and determine if you would \textbf{fit into} the program.
@@
 It's important to aim for the top venues in your field. You can
-find top CS conferences at places such as CSRankings~\cite{csrankings}, which ranks CS programs based on how their faculty publish at top \emph{conferences}.
+find top CS conferences at places such as \href{https://csrankings.org}{CSRankings}, which ranks CS programs based on how their faculty publish at top \emph{conferences}.
 Local conferences and non-English journals or conferences do
@@
 
-For Vietnamese students, it's worth highlighting the \href{https://vef2.org/}{VEF 2.0 program}, which has played a crucial role in aiding many good students in gaining acceptance to top-tier PhD programs in the US.
-VEF 2.0 follows an intriguing model where  US faculty members from leading institutions are invited to conduct rigorous interviews with VEF students and subsequently provide reference letters on their behalf.
-Despite the limited interaction between the interviewer and the interviewee (primarily confined to the interview itself), the impact of these reference letters is remarkably positive.
-This fact is underscored by the noteworthy success of numerous VEF 2.0 students in securing admission to top CS PhD programs across the US.
+For Vietnamese students, it's worth mentioning the \href{https://vef2.org/}{VEF 2.0 program}, which has helped many good students in gaining acceptance to top-tier PhD programs in the US.
+VEF 2.0 follows an interesting model where US faculty members from leading institutions are invited to conduct rigorous interviews with VEF students and subsequently provide reference letters on their behalf.
+Despite the limited interaction between the interviewer and the interviewee (primarily confined to the interview itself), the impact of these reference letters can be very positive as they have helped numerous VEF 2.0 students getting into top CS PhD programs in the US.
 
@@
 \begin{commentbox}[Vu]
-I always read the research statement (and the LORs) first. If I am
+I often read the research statement (and the LORs) first. If I am
 persuaded by then, I would skim over other factors and advocate for
@@
 
+\epigraph{"Oh...
+and how is education supposed to make me feel smarter?
+Besides, every time I learn something new, it pushes some old stuff out of my brain.
+Remember when I took that home wine-making course and I forgot how to drive?"}{The Simpsons}
+
 By around March you should hear back from most PhD programs you applied to . If you haven't heard back, reach out via email and ask about the status of your application.
@@
 \section{Funding}\label{sec:funding}
+
+\epigraph{Bart, with \$10,000, we’d be millionaires! We could buy all kinds of useful things like…love!’}{The Simpson}
 
@@
 \section{Miscs and FAQs}
+\epigraph{"I want to share something with you – the three little sentences that will get you through life; number 1: Cover for me, number 2: Oh, good idea, Boss, and number 3: It was like that when I got here."}{The Simpsons}
 
@@
 International students not familiar with US universities often put them into \emph{two} bins:  (i) very top schools that they dream about such as Stanford, MIT, Princeton, Harvard and (ii) everything else.
-Sometimes they rank CS programs using the reputations non-CS programs such as medical, math, or physics.
-In some cases they rank universities based on popular states they know in the US, e.g., California and New York.  Let's just say there are so many thing wrong with these methods.
+In many cases, they use resources such as US News, but their ranking is \emph{highly} questionable~\footnote{\url{https://cra.org/cra-statement-us-news-world-report-rankings-computer-science-universities/}}.
+Sometimes these students rank CS programs using the reputations non-CS programs such as medical, math, or physics.
+In some cases they rank universities based on popular states they know in the US, e.g., California and New York.  Clearly, there are so many thing wrong with these methods.
 
@@
       \begin{center}
-        \includegraphics[scale=0.3]{c8.png}
+        \includegraphics[scale=0.4]{c8.png}
       \end{center}
@@
 
-Below is a good example that I would definitely reply to.
+Below is a good example that I would reply to.
 
@@
 \begin{commentbox}[Vu]
-I've been called Dr. Vu and I find it a bit amusing but am totally fine with it.
+Dr. Vu is a bit amusing but I am totally fine with it.
 \end{commentbox}
@@
 
-\section{Virtual Research Opportunities Beyond Physical Boundaries}
+\section{Virtual Research Opportunities Beyond Physical Boundaries (written by Didier)}
 \didi{In my experience, prospective PhD students (from Rwanda) are often unaware of the information provided in this document throughout their undergraduate.

### 2023-09-17 phd-cs-us.tex
Updates from Overleaf

@@
 At GMU, the PhD committee typically has 15--20 faculty, and each committee member is assigned from 25--30 applications.
-Note GMU and UMass (and probably many other schools) have a separate committee for the MS program.
+Note GMUU, UMass, and probably many other schools, have a separate committee for the MS program.
 
@@
 It's important to aim for the top venues in your field. You can
-find top CS conferences at places such as \href{https://csrankings.org}{CSRankings}, which ranks CS programs based on how their faculty publish at top \emph{conferences}.
+find top CS conferences at places such as \href{https://csrankings.org}{CSRankings}, which ranks CS programs based on how their faculty publish at top \emph{conferences} (see more in \S\ref{sec:ranking}).
 Local conferences and non-English journals or conferences do
@@
 
-For Vietnamese students, it's worth mentioning the \href{https://vef2.org/}{VEF 2.0 program}, which has helped many good students in gaining acceptance to top-tier PhD programs in the US.
-VEF 2.0 follows an interesting model where US faculty members from leading institutions are invited to conduct rigorous interviews with VEF students and subsequently provide reference letters on their behalf.
-Despite the limited interaction between the interviewer and the interviewee (primarily confined to the interview itself), the impact of these reference letters can be very positive as they have helped numerous VEF 2.0 students getting into top CS PhD programs in the US.
+For Vietnamese students, it's worth mentioning the \href{https://vef2.org/}{VEF2.0 program}, which has helped many good students in gaining acceptance to top PhD programs in the U.
+VEF2.0 follows an interesting model where US faculty members from leading institutions are invited to conduct rigorous interviews with VEF students and subsequently provide reference letters on their behalf.
+Despite the limited interaction between the interviewer and the interviewee (primarily confined to the interview itself), the impact of these reference letters can be positive as they have helped many students getting into top PhD programs in the US.
 
@@
 In many cases, they use resources such as US News, but their ranking is \emph{highly} questionable~\footnote{\url{https://cra.org/cra-statement-us-news-world-report-rankings-computer-science-universities/}}.
-Sometimes these students rank CS programs using the reputations non-CS programs such as medical, math, or physics.
+Sometimes these students evaluate CS programs using the reputations non-CS programs such as medical, math, or physics.
 In some cases they rank universities based on popular states they know in the US, e.g., California and New York.  Clearly, there are so many thing wrong with these methods.
@@
 \bibliography{phd-cs-us.bib}
-\newpage
-
-\section{Virtual Research Opportunities Beyond Physical Boundaries (written by Didier)}
-\didi{In my experience, prospective PhD students (from Rwanda) are often unaware of the information provided in this document throughout their undergraduate.
+
+% \newpage
+% \section{Virtual Research Opportunities Beyond Physical Boundaries (written by Didier)}
+% \didi{In my experience, prospective PhD students (from Rwanda) are often unaware of the information provided in this document throughout their undergraduate.
 For this reason they often find themselves unprepared for PhD admission process by the time they finish undergraduate and have to spend a year or more to improve their application profiles.
@@
 I drafted the short essay on how to gain research experience remotely in order to improve their research profile and potentially get LOR from experts (hopefully faculty)}\\
-In the realm of computer science, research experience plays a crucial role in securing admissions to top-tier Ph.D.
+% In the realm of computer science, research experience plays a crucial role in securing admissions to top-tier Ph.D.
 programs.
@@
 
-\textbf{Virtual Research Programs}:
-Several universities, organizations, and research institutions offer virtual internships and research programs aimed at providing hands-on research experience.
+% \textbf{Virtual Research Programs}:
+% Several universities, organizations, and research institutions offer virtual internships and research programs aimed at providing hands-on research experience.
 These programs often involve working remotely under the guidance of experienced mentors and collaborating with a team of fellow researchers.
 For instance, \href{https://docs.google.com/forms/d/1btIwt4HwjyKMOUk-EMy3rbkfWzFxv2lNrMm_zkd0pA4/viewform?edit_requested=true}{UIUC+ Summer Undergraduate Research in Software Engineering}  offers an unpaid remote internship for software engineering students all over the world to collaborate with mentors from University of Illinois at Urbana Champaign.
-Virtual internships offer opportunities to contribute to ongoing research projects, conduct experiments, analyze data, and write research reports.
+% Virtual internships offer opportunities to contribute to ongoing research projects, conduct experiments, analyze data, and write research reports.
 Students can search for such programs through university websites, research institutes, or as of recently chatGPT (or Bard).
 
-\textbf{Online Research Communities and Open Source Contributions}:
-Online research communities and platforms offer a wealth of opportunities for aspiring researchers.
+% \textbf{Online Research Communities and Open Source Contributions}:
+% Online research communities and platforms offer a wealth of opportunities for aspiring researchers.
 Platforms like GitHub, GitLab, and Bitbucket host repositories spanning diverse research domains.
@@
 
-\textbf{Online Conferences and Workshops}:
-Attending online conferences and workshops is another way to gain exposure to cutting-edge research and establish connections with experts in the field.
+% \textbf{Online Conferences and Workshops}:
+% Attending online conferences and workshops is another way to gain exposure to cutting-edge research and establish connections with experts in the field.
 Many conferences now provide virtual participation options, enabling students to access research talks, poster sessions, and panel discussions and sometimes access designated chat rooms or networking events where participants can engage with researchers, ask questions, and seek potential research collaborations.
@@
 
-In conclusion, for students lacking research opportunities at their small colleges, the virtual realm offers a wealth of possibilities to gain valuable research experience.
+% In conclusion, for students lacking research opportunities at their small colleges, the virtual realm offers a wealth of possibilities to gain valuable research experience.
 Engaging with open source projects, joining online research communities, utilizing academic collaboration platforms, participating in virtual internships, and attending online conferences are all avenues to contribute to research, connect with experts, and enhance one's research profile.

### 2023-09-17 demystify.tex
Updates from Overleaf
update

--- 2023-09-17 phd-cs-us.tex
@@
 While this document is primarily intended for students interested in CS, it may be relevant to students from various disciplines.
-Furthermore, although many examples are specifics for George Mason University, the information should be generalizable to other R1\footnote{An \href{https://en.wikipedia.org/wiki/List_of_research_universities_in_the_United_States}{R1 institution} in the US is a research-intensive university with a high level of research activity across various disciplines.} institutions in the US  (though \emph{very} top schools could be very selective, e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU).
+Furthermore, although many examples are specifics for George Mason University (GMU), the information should be generalizable to other R1\footnote{An \href{https://en.wikipedia.org/wiki/List_of_research_universities_in_the_United_States}{R1 institution} in the US is a research-intensive university with a high level of research activity across various disciplines.} institutions in the US  (though \emph{very} top schools could be very selective, e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU).
 
@@
 %There are many advice on surviving PhD that you can follow. But here I just list a few that works for me and what I advice my students to do.\tvn{TODO}
+
+\section{Cultural Differences and Tips}
+
+This section lists various things related to studying the US that I notice international students often get surprised or confused.
+
+\begin{enumerate}
+    \item \textbf{Cheating}: Cheating, e.g., exams, coding, is a big no-no in the US.
+If you're caught cheating, you will face very heavy consequence and likely be expelled from the university (e.g., after the second time).
+This is very different from many international countries where cheating is common and often tolerated.
+Faculty is extremely good at detecting cheating (how would we not?
+we have deal with these situations so many times over so many years), and \emph{will} report cheating cases.
+In short, whatever you do, don't cheat---not worth it.
+    \item \textbf{Illegal Software}: Using illegal/cracked software is very common in many countries.
+But do not use/install them on university computer (e.g., computers given by your department or advisor).
+It is unlikely that the university will track you down, but it is the \textbf{software company} that will.
+They have very sophisticated tools to detect illegal software and will sue your university/department (imagine your department or advisor being sued for half a million dollars, and it is you who caused it).
+If you need software, ask your advisor or the department.
+\end{enumerate}
 
@@
 \bibliographystyle{abbrv}
-\bibliography{phd-cs-us.bib}
+\bibliography{demystify.bib}
 

### 2023-09-18 demystify.tex
update

@@
     \item \textbf{Cheating}: Cheating, e.g., exams, coding, is a big no-no in the US.
-If you're caught cheating, you will face very heavy consequence and likely be expelled from the university (e.g., after the second time).
+If you're caught cheating, you will face very heavy consequence and likely be expelled from the university (e.g., after the second time at GMU).
 This is very different from many international countries where cheating is common and often tolerated.
-Faculty is extremely good at detecting cheating (how would we not?
-we have deal with these situations so many times over so many years), and \emph{will} report cheating cases.
+Faculty is extremely good at detecting cheating (how can we not?
+we deal with these situations so many times over so many years), and \emph{will} report cheating cases.
 In short, whatever you do, don't cheat---not worth it.
+
     \item \textbf{Illegal Software}: Using illegal/cracked software is very common in many countries.
@@
 It is unlikely that the university will track you down, but it is the \textbf{software company} that will.
-They have very sophisticated tools to detect illegal software and will sue your university/department (imagine your department or advisor being sued for half a million dollars, and it is you who caused it).
+They have very sophisticated tools to detect illegal software and will sue your university/department.
+Imagine your department or advisor being sued for half a million dollars, and it is you who caused it.
 If you need software, ask your advisor or the department.

### 2023-09-18 demystify.tex
update

@@
 This is very different from many international countries where cheating is common and often tolerated.
-Faculty is extremely good at detecting cheating (how can we not?
-we deal with these situations so many times over so many years), and \emph{will} report cheating cases.
+Faculty is extremely good at detecting cheating (we have been dealing with these situations so many times over so many years), and \emph{will} report cheating cases.
 In short, whatever you do, don't cheat---not worth it.

### 2023-09-19 demystify.tex
Updates from Overleaf

@@
 International students not familiar with US universities often put them into \emph{two} bins:  (i) very top schools that they dream about such as Stanford, MIT, Princeton, Harvard and (ii) everything else.
-In many cases, they use resources such as US News, but their ranking is \emph{highly} questionable~\footnote{\url{https://cra.org/cra-statement-us-news-world-report-rankings-computer-science-universities/}}.
+In many cases, they use resources such as rankings from US News, which are \emph{highly} questionable\footnote{\url{https://cra.org/cra-statement-us-news-world-report-rankings-computer-science-universities/}}.
 Sometimes these students evaluate CS programs using the reputations non-CS programs such as medical, math, or physics.
-In some cases they rank universities based on popular states they know in the US, e.g., California and New York.  Clearly, there are so many thing wrong with these methods.
+They even rank universities based on popular states they know in the US, e.g., California and New York.  Clearly, there are so many thing wrong with these methods.
 
 You can learn about CS programs and research expertise of faculty using resources such as \href{https://csrankings.org}{CSRankings.org}, which is designed specifically to help prospective PhD students in Computer Science!
-You will be very surprised to learn that a school that you didn't know much about can have very strong research in your interested topic (and vice versa, a school you thought highly about has no one works in the research field you're interested in).
+You will be very surprised to learn that a school that you didn't know much about can have very strong research in your interested topic (and vice versa, a school you thought highly about might have no faculty working in the research field you're interested in).
 This is also a good way to learn about individual faculty (who works on what) and well-known CS conferences\footnote{In CS (and probably only in CS), conferences, not journals, are often the main venue to publish research finding.}.
@@
 Tenured faculty, e.g., associate and full profs., are likely older, more well-known, and have more experience in managing students.
-However, they might not push you as hard (they don't have to, they already got tenured).
-They might also expect you to figure things out yourself (so you need to be very independent).
-Some tenured faculty are also no longer active in research and more involved with administrative duties or with their startup companies.
-
-You can learn about faculty and their level of research activity through the faculty's website and CSRankings.
-Ultimately, choose one that fits you the most by communicating with them, meeting them, asking them questions, even talking to their students.
+However, they might not push you as hard (they don't have to, they already got tenure).
+They might also expect you to figure things out yourself, i.e., so you need to be very independent.
+Some tenured faculty are also no longer active in research and more involved with administrative duties or with their startup companies (this means they will likely not take new students).
+
+You can learn about faculty and their level of research activity through the faculty's website or even from CSRankings.
+Ultimately, choose one that fits you the most by communicating with them, meeting them, asking them questions, and even talking to their students.
 
@@
 to ask about \emph{research opportunities}, or \emph{GTA/GRA} support, then \emph{yes}, I believe you should---it is \emph{worth it}.
-However, you need to do it in a right way.
+However, you need to do it the \emph{right way}.
 
@@
 
-    The first 2 years you spend on coursework (somewhat equivalent to an MS study), finding an adviser, learning research.
+    The first 2 years you typically take coursework (somewhat equivalent to an MS study), find an adviser, learn how to do research.
 The next 2--3 years you focus on your research, form dissertation topic, and get results published.
@@
     In many cases you might take a summer or two off to do internship to get additional research opportunities.
-    The PhDComics figure on the right shows the ``ambition'' level of a PhD student over their years of study (they miss the 6-7th Year where the ambition is \emph{``Just let me graduate''}).
+    The PhDComics figure on top shows the ``ambition'' level of a PhD student over their years of study (they miss the 6-7th year where the ambition is \emph{``Just let me graduate''}).
 
@@
 \subsection{How much do \emph{you} cost?}\label{sec:ra-cost}
-PhD students often wonder why their salary is so low compared to ludicrous grants their advisors get or why their offer letters sometime mentioned that their benefits worth way more than what they actually receive (i.e., stipend).
+PhD students often ask why their salary is so low compared to ludicrous grants their advisors get.
+They also wonder why their offer letters sometime mentioned that their benefits worth way more than what they actually receive (i.e., stipend).
 This section aims to shed some light to these questions.
@@
 Other GMU faculty members also have provided feedback and contributions.
-Many students including Didier (GMU), Thanh (Melbourne), and Dat (Melbourne) also have contributed valuable questions and feedback.
+Many students including Didier (GMU), Thanh (Melbourne), and Dat (Melbourne) have contributed valuable questions and feedback.
 \textbf{Thank you!}

### 2023-09-19 demystify.tex
Updates from Overleaf

@@
 discourages students from applying but also creates the perception that
-getting admitted is difficult compared to CS PhD programs in other countries.
+getting admitted to a CS PhD program in the US is difficult compared to other countries.
 
@@
 \end{document}
-
-# - Asking professors if they can get in or if they take students.\\
-#   > Many international students will write professors asking if they can
-#   get in, or if they are taking students. Unlikely anyone will reply to
-#   such email because the short answer is we simply cannot evaluate your
-#   profile ourselves, it needs to go to the whole committee as explained.
-#   Even if a professor is seeking for students (almost everyone is
-#   looking for good students), they simply cannot take one in without
-#   going through the application process.
-
-# - See this link for a large list of [[https://github.com/dynaroars/dynaroars.github.io/wiki/Funding][funding sources for PhD students]]
-
-# ** Self-funded
-# - Some PhD students are funded through their company
-#   - The students work full time for the company and do their PhD part-time
-#   - Typically will take longer (because you have to work full time) and require some serious commitment from the students
-
-# - Some PhD students are funded through country, e.g., GMU has several students from Saudi Arabia that are funded through their government.
-
-# ** Others  
-# - There are also various kind of one-time thing and small awards from industry or organization, e.g., a $2000 award for the summer.
-Unlike the B.A., where you or your parents pay many tens of thousands of dollars,
-or the M.S., where you typically work as a teaching assistant and possibly continue
-to pay many tens of thousands of dollars, the PhD is a time where funding is not a
-concern to you. At most schools, you will not pay tuition during the time that you
-are getting a PhD. Typically, you will also receive a living stipend – on the order
-of $2000 per month, from which you will pay your living expenses. Ideally, your
-only responsibility will be research. This is called doing an RAship (Research
-Assistantship).
-The PhD is a tremendous opportunity. You get to pick an advisor in any re-
-search area you like and then you get to do research in that area, receive mentoring,
-think deeply on problems, publish papers, become famous, while paying zero tu-
-ition for 6 years and receiving a salary. Your advisor is paying for this opportunity
-by writing grant proposals to companies and to the government to ask for fund-
-ing. A single graduate student can cost an advisor upwards of 80K per year (given
-the cost of tuition, the stipend, the overhead tax charged by the school, cost of
-equipment and physical space, etc.).
-Important note 1: At most schools, you can only do an RAship if you have
-an advisor who has funding for you. Since some advisors don’t apply for grants
-or are in areas which aren’t well-funded, you may have to work as a teaching
-assistant every semester to get your stipend. This is called a TAship (Teaching
-Assistantship). When I was a graduate student, I had a few friends who were
-forced to TA 13 semesters, to fund their way through school! Alternatively, you
-will have to restrict your choice of advisors to those who have funding. At CMU,
-every PhD student is guaranteed a stipend plus tuition regardless of which advisor
-she chooses to work with.
-Important note 2: There are many companies and government organizations
-which offer Graduate Fellowships for PhD students. If you are lucky enough
-to get one of these, they will cover most of your way through graduate school,
-and you will never have to worry about whether your advisor has funding or not.
-Details about graduate fellowships will be discussed in Section 4.
-
-* Links
-:PROPERTIES:
-:CUSTOM_ID: links
-:END:
-- https://idleprocess.wordpress.com/2009/12/07/why-go-to-graduate-school-and-how-to-get-into-the-program-of-your-dreams/
-- https://mycsphd.org/_pages/application-parts.html
-- https://emeryberger.com/admission-notes/
-- https://www.elsevier.com/connect/9-things-you-should-consider-before-embarking-on-a-phd
-- https://grad.uchicago.edu/wp-content/uploads/2019/08/Applying-to-PhD-Programs-Guide-2019.pdf
-- https://www.pathwaystoscience.org/pdf/CIC_GradSchoolGuide.pdf
-- https://www.science.org/content/article/applying-phd-these-10-tips-can-help-you-succeed
-- https://chrisblattman.com/blog/2022/03/25/faqs-on-phd-applications/
-- https://www.cs.cmu.edu/~harchol/gradschooltalk.pdf
-
-# * Types of Funding for CS PhDs
-
-# * How does GRA work?
-# While you might get a 9-month $28K GRA stipend,  your advisor typically has to budget at least twice that much.
-Because they have to pay your tuitition, insurance, and overhead that the university takes.
-The chart below shows that each CS GRA costs about $70K at GMU.
-
-# - Example budget of a PhD student in CS at GMU.  Using stipend and tuition rate in 2023. Conference Registration and Travel are estimates.
-# - It is also **very likely** that GMU faculty won't have to pay student tuition starting Fall 2025, when this happens then we can take out  Tuition and total cost is 46.5K.
-
-# | Budget                                 | Cost    |                                                                                           |
-# |----------------------------------------+---------+-------------------------------------------------------------------------------------------|
-# | GRA (9-month)                          | 28K     |                                                                                           |
-# | GRA (summer)                           | 9K      | 3-month, 20hrs/week                                                                       |
-# | Fringe Benefits (7.3%)                 | 0       | for faculty mostly (non-wage expenses such as social security, medicare)                  |
-# | **Total Salary**                       | 37K     |                                                                                           |
-# | Health Insurance                       | 3K      | Full year                                                                                 |
-# | Tuition (In-State) for Fall and Spring | 15K     | ($680/ Credit + $150/Student Fee/ Credit)* 9 credits = $7470 ($6120 + $1350) per semester |
-# | **Total**                              | 18K     | Full year tuition + insurance                                                             |
-# | Total Materials & Supplies             | 0       |                                                                                           |
-# | Conference Registration                | 500     |                                                                                           |
-# | International Travel                   | 1.8K    |                                                                                           |
-# | Domestic Travel                        | 700     |                                                                                           |
-# | **Total Travel**                       | 3K      |                                                                                           |
-# | Total **Direct** Cost                  | 58K     | Salary 37K + Travel 3K  + Health 3K + Tuition 18K                                         |
-# | F & A (MTDC)                           | 21K     | Direct Cost - GRA Salary                                                                  |
-# | Total **Indirect** Cost                | 12K     | **58.9%** of MTDC                                                                         |
-# | **Total** (Direct + Indirect)          | **70K** |                                                                                           |
-
-# - F & A = Facilities & Administrative Cost Base 
-# - MTDC = Modified Total Direct Cost
-
-# * Miscs
-
-# - Asking professors if they can get in or if they take students.\\
-#   > Many international students will write professors asking if they can
-#   get in, or if they are taking students. Unlikely anyone will reply to
-#   such email because the short answer is we simply cannot evaluate your
-#   profile ourselves, it needs to go to the whole committee as explained.
-#   Even if a professor is seeking for students (almost everyone is
-#   looking for good students), they simply cannot take one in without
-#   going through the application process.
-
-# - How long for a CS PhD in US ? > 5-6 years on avg.
-
-# - Does having an MS help? > It can help if the MS gives research
-#   experience or is from a more well-known school than your
-#   undergraduate. But not having an MS does not hurt either. Nowadays it
-#   is normal for students with bachelor degrees apply directly to
-#   PhD in CS.

### 2023-09-19 demystify.tex
update spellings

@@
 Having been involved in PhD admissions for many years, I've
-realized that many \textbf{international students}, especially those from smaller countries such as Vietnam, lack a clear understanding of
+realized that many \textbf{international students}, especially those from smaller countries like Vietnam, lack a clear understanding of
 the Computer Science PhD admission process at US universities. This confusion not only
@@
 In addition, this document can help \textbf{US faculty and admission committee} gain a better understanding of international students and their cultural differences.
-By appreciating and leveraging these differences, CS programs in the US can attract larger and more competitive application pools from international students.
+By recognizing and leveraging these differences, CS programs in the US can attract larger and more competitive application pools from international students.
 
-I wish you the best of luck. And if you follow this guideline,
-you will at least have a good chance at GMU (see
+I wish you the best of luck. And if you follow this guidance, you will at least have a good chance at GMU (see
 \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-GMU}{why
@@
 harder than at schools in other countries.
-If you think you have a chance in other countries, e.g., South Korea, Singapore, Germany, UK, Japan and Australia, then you will surely have a chance in the US as well.
+If you believe you have a chance in other countries, e.g., South Korea, Singapore, Germany, UK, Japan and Australia, then you will surely have a chance in the US as well.
 
@@
 In general the number of
-PhD applications from Vietnam to US universities is few and  more would be very welcomed. 
+PhD applications from Vietnam to US universities is low and more would be very welcomed. 
 \end{commentbox}
@@
 
-Then applications will be reviewed by a
-\textbf{PhD admission committee} that consists of faculty members in CS (in some cases the committee can involve affiliated faculty from different disciplines).
+Then, applications are reviewed by a \textbf{PhD admission committee} that consists of faculty members in CS (in some cases the committee can involve affiliated faculty from different disciplines).
 These faculty have a wide-range of expertise and background to ensure diverse perspectives in the evaluation process.
@@
 We usually decide that a full-time PhD candidate is either (i) admit with funding (TA or RA) or (ii) rejected. In other words, in most cases, we either
-admit you with full funding, or we don't. In some rare cases, we admit
+admit you with full funding, or we reject your application. In some rare cases, we admit
 without funding because you have funding on your own (e.g.,
 supported by your government or having external grants). We also justify
-our decision with a summary about your application, where we list
-strengths (e.g., well-known school) and weaknesses (e.g., weak
+our decision with a summary of your application, where we list
+strengths (e.g., a well-known school) and weaknesses (e.g., weak
 LORs). 
@@
 
-Additionally, \textbf{work experiences at renown research laboratories}, such as Microsoft Research, can significantly strengthen your
+Additionally, \textbf{work experiences at renowned research laboratories}, such as Microsoft Research, can significantly strengthen your
 application.
@@
 So I will not talk too much about statements.
-In short, discuss about your research goal and vision, and convince us that you can achieve it through your experience, e.g., published papers, or if you work on some projects by yourself, talk about it.
+In short, talk about your research goal and vision, and convince us that you can achieve it through your experience, e.g., published papers, or if you work on some projects by yourself, talk about it.
 Also, use the statement to talk about stuff that admission committee members might not know about, e.g., your Github project with 1K+ stars or  your regular contributions to well-known open-source projects.
@@
 several things including asking your CS dept to put itself on CSRankings (it's the easiest way to get CS people to know about the school) and explaining about your school in your statement (ask your LOR writer to do that too).
-Of course, if you're interested in working with Vietnamese professors, consider  \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{CS PhD programs in the US that have them}.
+Of course, if you're interested in working with Vietnamese, consider  \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{CS PhD programs in the US that have Vietnamese professors}.
 
@@
 
-Having good grades is important, but unless your school is well-known, having top grades/ranks
+Having good grades is important, but unless your school is well-known, having top grades or rankings
 usually will not help. This is simply because we cannot evaluate them.
 
-This can be an issue for students in many top international universities where the competition is so high that very good students students can still have low ranking from these schools (and be overlooked by Admission committee).
+This can be an issue for students in many top international universities where the competition is so high that very good students students can still have low rankings from these schools (and be overlooked by Admission committee).
 So what to do with this? well, same as \hyperref[sec:your-school]{before}, e.g., put a note about this in your statement and ask your LoR writers to talk about it.
@@
 having very bad grades will be \red{red flag} (unless your LORs or
-statements give proper explanation). This is especially true if you
+statement give a proper explanation). This is especially true if you
 have bad grades in relevant courses (e.g., Math and CS).
@@
 \begin{commentbox}[Vu]
-Personally I do not pay attention to GRE scores as I do not believe they are good indicator for research. However, if you have good GRE scores then you should list them in your application.
+Personally I do not pay attention to GRE scores as I do not believe they are good indicator for research. However, if you have good GRE scores then you should include them in your application.
 \end{commentbox}
 
-\paragraph{English Test} Unless your degrees are from certain countries such as \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-GMU#standard-tests-waiver-eligible-countries}{these}, you will need to
-take standardized English test. Just do well enough to pass minimum requirement set by the university, which nowadays has many options for you to choose from.
+\paragraph{English Test} Unless your degrees are from specific countries such as \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-GMU#standard-tests-waiver-eligible-countries}{these}, you will need to
+take standardized English test. Just do well enough to pass minimum requirement set by the university, which nowadays has many options for you to select from.
 
@@
 \subsection{CV/Resume}
-This should be a summary of the accomplishments of the applicant.
+The CV/Resume should provide a summary of the applicant's achievements.
 It should allow the reviewers to quickly scan to identify standout achievements (e.g., Publications, Programming Competition Awards, Teaching Experience).
@@
 
-An interview often lasts about 15-30 minutes, and one important aspect of evaluation is your ability to effectively communicate, including speaking and understanding English.
+Typically, an interview takes about 15-30 minutes, and one important aspect of evaluation is your ability to effectively communicate, including speaking and understanding English.
 In addition, during the interview, you will have the opportunity to ask questions about the university.
-It's essential to prepare some thoughtful questions, similar to interviewing at a company.
+It's essential to prepare some thoughtful questions, similar to a job interview.
 
@@
 
-In short, if you can, do come to Open House.  Even if you can't come in person, you still can attend virtually and meet with individual faculty.
+In summary, if possible, attend the Open House.  Even if you can't come in person, you still can attend virtually and meet with individual faculty.
 

### 2023-09-19 demystify.tex
various updates

@@
 Fourth, CS department typically has higher stipend comparing to other study fields.
-Finally, private universities might pay more for stipend (but they could have higher "activity" or some other hidden fee, or you will have to pay some fee for each class credits).
+Finally, private universities might pay more for stipend (but they could have higher "activity" or some other hidden fee, or you may be required to pay fees for each credit hour).
 
@@
 They might also expect you to figure things out yourself, i.e., so you need to be very independent.
-Some tenured faculty are also no longer active in research and more involved with administrative duties or with their startup companies (this means they will likely not take new students).
+Some tenured faculty are also no longer active in research and more involved with administrative responsibilities or with their startup companies (this means they will likely not take new students).
 
 You can learn about faculty and their level of research activity through the faculty's website or even from CSRankings.
-Ultimately, choose one that fits you the most by communicating with them, meeting them, asking them questions, and even talking to their students.
+Ultimately, choose one that fits you the most by communicating with them, meeting them, asking them questions, and even talking to their current students.
 
@@
 
-\begin{enumerate}
-    \item \textbf{Cheating}: Cheating, e.g., exams, coding, is a big no-no in the US.
+\paragraph{Diversity} US universities prioritize diversity and inclusion.
+It's crucial for students to respect and appreciate varied opinions, backgrounds, and cultures.
+Unlike some countries where certain voices are marginalized, in the US, all perspectives are valued equally (especially at universities).
+Racism or discrimination can have serious consequences, including academic and disciplinary actions.
+
+\paragraph{Academic Integrity} Plagiarism and cheating, e.g., exams, coding, is a big no-no in the US.
 If you're caught cheating, you will face very heavy consequence and likely be expelled from the university (e.g., after the second time at GMU).
@@
 
-    \item \textbf{Illegal Software}: Using illegal/cracked software is very common in many countries.
-But do not use/install them on university computer (e.g., computers given by your department or advisor).
+Here is the typically steps: (i) a faculty suspecting a cheating case will report it to the Office of Academic Integrity (OAI) at the university;  (ii)  OAI will completely take over and investigate the case; and finally (iii) OAI make the final decision.
+It is important to note that after receiving the report from your prof., OAI completely takes over (so begging your professor will not help because they simply are no longer involved).
+
+\paragraph{Illegal Software} Using illegal/cracked software is very common in many countries.
+But do not install or use them on university computer (e.g., computers given by your department or advisor).
 It is unlikely that the university will track you down, but it is the \textbf{software company} that will.
@@
 If you need software, ask your advisor or the department.
-\end{enumerate}
 

### 2023-09-19 demystify.tex
various updates

@@
     \begin{cbox}
-    \textbf{#1}: 
+    \textbf{#1} 
  }{
@@
 \section{Should You Apply?}
-\epigraph{\vspace{-0.2in} Don't make fun of graduate students. They just made a terrible life choice.}{The Simpsons}
+\epigraph{\vspace{-0.2in} Don't make fun of graduate students. They just made a terrible life choice.}{\textsc{The Simpsons}}
 
@@
 One of the reasons I create this document is that my colleagues at GMU are interested in recruiting Vietnamese students and are surprised when seeing very few applications in Vietnam (e.g., each year our CS program receives more than 350 PhD applications, most of which are international but only 3--4 are from Vietnam).
-In general the number of
-PhD applications from Vietnam to US universities is low and more would be very welcomed. 
+In general the number of CS PhD applications from Vietnam to US universities is low and more would be very welcomed.
 \end{commentbox}
@@
 
-\epigraph{"How is education supposed to make me feel smarter?
+\epigraph{How is education supposed to make me feel smarter?
 Besides, every time I learn something new, it pushes some of the old stuff out of my brain.
-Remember when I took that home wine making course, and I forgot how to drive?"}{The Simpsons}
+Remember when I took that home wine making course, and I forgot how to drive?}{\textsc{The Simpsons}}
 
@@
 
-\epigraph{Son, if you really want something in this life, you have to work for it. Now quiet! They’re about to announce the lottery numbers.}{The Simpsons}
+\epigraph{Son, if you really want something in this life, you have to work for it. Now quiet! They’re about to announce the lottery numbers.}{\textsc{The Simpsons}}
 
@@
 Besides, every time I learn something new, it pushes some old stuff out of my brain.
-Remember when I took that home wine-making course and I forgot how to drive?"}{The Simpsons}
+Remember when I took that home wine-making course and I forgot how to drive?"}{\textsc{The Simpsons}}
 
@@
 
-\epigraph{Bart, with \$10,000, we’d be millionaires! We could buy all kinds of useful things like…love!’}{The Simpson}
+\epigraph{Bart, with \$10,000, we’d be millionaires! We could buy all kinds of useful things like…love!’}{\textsc{The Simpsons}}
 
@@
 \section{Miscs and FAQs}
-\epigraph{"I want to share something with you – the three little sentences that will get you through life; number 1: Cover for me, number 2: Oh, good idea, Boss, and number 3: It was like that when I got here."}{The Simpsons}
+\epigraph{"I want to share something with you – the three little sentences that will get you through life; number 1: Cover for me, number 2: Oh, good idea, Boss, and number 3: It was like that when I got here."}{\textsc{The Simpsons}}
 
@@
 
-\begin{commentbox}[Vu]
 In his \href{https://matt.might.net/articles/how-to-apply-and-get-in-to-graduate-school-in-science-mathematics-engineering-or-computer-science/}{post}, Matt Might was initially unsure about an application.
 However, upon learning that the applicant had led a 100km hike in the Himalayas, he decided to accept the applicant.
-
-\tcblower
-I would also advocate for that student as this shows they have the persistence and determination required for research.
-\end{commentbox}
+This is a good example of being "stand out", and I would also advocate for that student as this shows they have the persistence and determination required for research.
 
@@
 
-\begin{commentbox}[Good example]
-
-Dear Prof. Nguyen,
+\begin{commentbox}[Dear Prof. Nguyen,]
 
@@
 ...
-\tcblower
+
 This is a good example because it is clearly written just for me.  It shows that the student knows about my work on invariant generation and has related  background (paper C and project G). 
@@
 
-\subsection{PhD in other related field such as CE, IST, Cybersecurity}
+\subsection{PhD in other related fields: CE, IST, Cybersecurity}
 
-In many cases you do not need to do a PhD in CS to work in your area of interested.
-For example, in addition to having a traditional CS department, GMU has IST and Cybersecurity departments, which have faculty  with PhD in CS and work on CS topics (e.g., AI, Security, Robotics).
+In many cases you do not have to do a PhD in CS to work in your area of interested.
+For example, in addition to a traditional CS department, GMU has IST and Cybersecurity departments, which have faculty  with PhD in CS and work on CS topics (e.g., AI, Security, Robotics).
 So it is totally possible that you still get to do CS research and publish in CS-focused venues even if you're not in a traditional CS program.
-It is very common to see faculty with CS PhD in a non-CS department as well as faculty with non-CS PhD in CS department.
+It is  common to see faculty with CS PhD in a non-CS department as well as faculty with non-CS PhD in CS department.
 
@@
 As you know that prof.
-better and depends on their preference, you may call them by their Firstname.
+better and depends on their preference, you may call them by their firstname.
+Many students call me \emph{Dr.
+Vu}, which I find a bit amusing but am totally fine with it (and Vu is way easier to pronounce than Nguyen).
 
-Note that at some universities, especially those that do not offer PhD studies, will use the formal \textbf{Dr.} title, i.e., Dr. Lastname.  So just observe and follow how it is done at your place.
-
-\begin{commentbox}[Vu]
-Dr. Vu is a bit amusing but I am totally fine with it.
-\end{commentbox}
+Note that at some universities, especially those that do not offer PhD studies, will use the formal \textbf{Dr.} title, i.e., Dr.
+Lastname.
+So just observe and follow how it is done at your place.
+Also note that some faculty might not have a Ph.D., so they might not be Dr.
+In that case, you can use Prof.
+Lastname.
 
@@
 \paragraph{Diversity} US universities prioritize diversity and inclusion.
-It's crucial for students to respect and appreciate varied opinions, backgrounds, and cultures.
+Students need to respect and appreciate varied opinions, backgrounds, and cultures.
 Unlike some countries where certain voices are marginalized, in the US, all perspectives are valued equally (especially at universities).
-Racism or discrimination can have serious consequences, including academic and disciplinary actions.
+Racism or discrimination will have serious consequences, including academic and disciplinary actions.
 
-\paragraph{Academic Integrity} Plagiarism and cheating, e.g., exams, coding, is a big no-no in the US.
+\paragraph{Academic Integrity} Plagiarism and cheating (e.g., exams, coding) is a big no-no in the US.
 If you're caught cheating, you will face very heavy consequence and likely be expelled from the university (e.g., after the second time at GMU).
@@
 
-Here is the typically steps: (i) a faculty suspecting a cheating case will report it to the Office of Academic Integrity (OAI) at the university;  (ii)  OAI will completely take over and investigate the case; and finally (iii) OAI make the final decision.
-It is important to note that after receiving the report from your prof., OAI completely takes over (so begging your professor will not help because they simply are no longer involved).
+Here is the typically steps: (i) a faculty suspecting a cheating case will report it to the Office of Academic Integrity (OAI) at the university;  (ii)  OAI will take over and investigate the case; and finally (iii) OAI will make the final decision.
+It is important to note that after receiving the report from your prof., OAI \emph{completely} takes over (so begging your professor will not help because they simply are no longer involved and can't do much here).
 
@@
 But do not install or use them on university computer (e.g., computers given by your department or advisor).
-It is unlikely that the university will track you down, but it is the \textbf{software company} that will.
+It is unlikely that the university will track you down, but it is the \emph{software company} that will.
 They have very sophisticated tools to detect illegal software and will sue your university/department.

### 2023-09-19 demystify.tex
prof references

@@
 better and depends on their preference, you may call them by their firstname.
-Many students call me \emph{Dr.
-Vu}, which I find a bit amusing but am totally fine with it (and Vu is way easier to pronounce than Nguyen).
+For example, my students and colleague call me Vu.
+Some students call me \emph{Dr.
+Vu}, which I find a bit amusing but am totally fine with it.
 
-Note that at some universities, especially those that do not offer PhD studies, will use the formal \textbf{Dr.} title, i.e., Dr.
+Note that at some universities, especially those that do not offer PhD degrees, will use the formal \textbf{Dr.} title, i.e., Dr.
 Lastname.
 So just observe and follow how it is done at your place.
-Also note that some faculty might not have a Ph.D., so they might not be Dr.
-In that case, you can use Prof.
-Lastname.
+Also note that some faculty might not have a Ph.D.
+and so you can use Prof.
+Lastname instead.
+
+A related topic is to consider is how to refer to a professor you know.
+Here, because you are already familiar with these individuals, you can informally use their firstname (or Dr./Prof., if you want to be formal) as mentioned above.
+You can also include their institution if it makes it more precise.
+For example, I did my postdoc with Jeff Foster at Univ.
+of Maryland.
+Do not include ranking (e.g., Assistant, Associate, Scientist, ...) when referring to someone.
+I see many students, especially Vietnamese, include a lengthy title of people they know, e.g., my advisors are Asst.
+Prof.
+X and Assoc.
+Prof.
+Y., and I also collaborate with ACM Distinguished Scientist Z.
+This is \emph{not necessary} and makes it look like you're trying to show off your connections.
+In short, these nuances represent some cultural differences that you may encounter and will gradually adapt to (more on cultural differences in \S\ref{sec:cultural}).
 
@@
 
-\section{Cultural Differences and Tips}
+\section{Cultural Differences and Tips}\label{sec:cultural}
 

### 2023-09-23 demystify.tex
Updates from Overleaf

@@
 This shows that you're serious and have done homework on places you're applying to.
-Admission committee will look for this part at the end of the statement, so do not skip it.
+Admission committee will look for this part (typically written at the end of your statement), so do not skip it.
 
@@
 
-\begin{commentbox}[Vu]
 Sometime PhD admission committee in the US will share a document such as \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Foreign-Top-Schools}{this one}, which lists the top schools in several countries.
@@
 During my time at UNL and now here at GMU, I have looked at Vietnamese applications (whether they are assigned to me or not) and provide input to their reviewers, e.g., X is the top tech school in Vietnam and so it should be \emph{top} instead of \emph{unknown} foreign, which makes a huge difference.
-\end{commentbox}
 
@@
 
-\begin{commentbox}[Vu]
-The minimum for GMU (being above this might not mean much, but below is a \red{red flag}).
+The following lists the minimum for GMU (being above this might not mean much, but below is a \red{red flag}).
 \begin{itemize}
@@
 \end{itemize}
-\end{commentbox}
 
@@
 If you're admitted to a good CS PhD program, you will be fine.
-A good school would know that it has to be competitive, otherwise it will loose good students to other schools.
+A good school knows that it has to be competitive, otherwise it will lose good students to other schools.
 For example, at GMU, every year we discuss about improving the benefits, and especially stipend, for our graduate students.
@@
 The college and department will set the rate for 9-month graduate assistant stipend.
-TA, which is paid by the department, will likely be that amount but RA might be higher depending on the stage of the student (1st year vs ABD\footnote{All but dissertation: really close to graduate.}) and the prof.
+TA, which is paid by the department, will likely be that amount set by the college, but RA might be higher depending on the stage of the student (1st year vs ABD\footnote{All but dissertation: really close to graduate.}) and the prof.
 \tcblower
@@
 
-\begin{commentbox}[Vu]
-If you got recruited by a prof.
+It is worth noting that if you got recruited by a prof.
 who would give you RA right away, it's very likely you will get admitted.
 For example, if a prof., even if not in PhD admission committee, wants to work with and funds you, the PhD admission committee will respect that decision and admit you (unless your application has many red flags).
-\end{commentbox}
 
@@
 
-Faculty received many "cold" e-mails from international students seeking for admission, TA, and RA.
+Faculty often receive  "cold" e-mails from international students seeking for admission, TA, and RA.
 Most of the time, we ignore these emails, but in some rare occasions we do answer them.
@@
 Students need to respect and appreciate varied opinions, backgrounds, and cultures.
-Unlike some countries where certain voices are marginalized, in the US, all perspectives are valued equally (especially at universities).
+Unlike some countries where certain voices are ignored, in the US, all perspectives are valued equally (especially at universities).
 Racism or discrimination will have serious consequences, including academic and disciplinary actions.
@@
 
-\paragraph{Illegal Software} Using illegal/cracked software is very common in many countries.
+\paragraph{Illegal Software} Using illegal software or applying "cracks" to software is very common in many countries.
 But do not install or use them on university computer (e.g., computers given by your department or advisor).
@@
 They have very sophisticated tools to detect illegal software and will sue your university/department.
-Imagine your department or advisor being sued for half a million dollars, and it is you who caused it.
+Imagine your department or advisor being sued for million dollars, and it is you who caused it.
 If you need software, ask your advisor or the department.

### 2023-09-23 demystify.tex

@@
 At GMU, the PhD committee typically has 15--20 faculty, and each committee member is assigned from 25--30 applications.
-Note GMUU, UMass, and probably many other schools, have a separate committee for the MS program.
+Note that GMU, UMass, and probably many other schools, have separate committees for MS programs.
 
@@
 
-In many cases, the admission committee involves assistant professors in the department.
+In many cases, the PhD admission committee involves assistant professors in the department.
 This provides junior faculty the opportunities to recruit students.
@@
 
-\begin{commentbox}[Vu] 
-We usually decide that a full-time PhD candidate is either (i) admit with funding (TA or RA) or (ii) rejected. In other words, in most cases, we either
+At GMU, we usually decide that a full-time PhD candidate is either (i) admit with funding (TA or RA) or (ii) rejected. In other words, in most cases, we either
 admit you with full funding, or we reject your application. In some rare cases, we admit
@@
 LORs).
-\end{commentbox}
 
@@
 Several things about stipends.
-First, the amount of stipend depends on the university, which in turns depend on various factors such as location (e.g., a stipend in Washington DC is likely higher than in Lincoln, Nebraska due to higher living cost).
-Second, a school year is (typically) 9-month in the US, so stipend is for 9 months (so divide by 9 for each month).
+First, the amount of stipend depends on the university, which in turn depends on various factors such as location (e.g., a stipend in Washington DC is likely higher than in Lincoln, Nebraska due to higher living cost).
+Second, a school year is typically 9-month in the US, so stipend is for 9 months (so divide by 9 for each month).
 Third, like most source of income in the US, you will have to pay tax on your stipend.
@@
 Having health insurance is required at many US universities.
-So do not think that you're young and healthy and ignore insurance.
+Do not assume that you're young and healthy and ignore insurance.
 At GMU, and at most good CS PhD programs, your GTA or GRA \emph{will always} come with full insurance.
@@
 
-A related topic is to consider is how to refer to a professor you know.
-Here, because you are already familiar with these individuals, you can informally use their firstname (or Dr./Prof., if you want to be formal) as mentioned above.
+\paragraph{Referring to professors you know} Here, because you are already familiar with these individuals, you can just informally use their firstname (or Dr./Prof., if you want to be formal) as mentioned above.
 You can also include their institution if it makes it more precise.
@@
 Do not include ranking (e.g., Assistant, Associate, Scientist, ...) when referring to someone.
-I see many students, especially Vietnamese, include a lengthy title of people they know, e.g., my advisors are Asst.
+I see many international students include a lengthy title of people they know, e.g., my advisor is Asst.
 Prof.
-X and Assoc.
-Prof.
-Y., and I also collaborate with ACM Distinguished Scientist Z.
+X, and I also collaborate with ACM Distinguished Scientist Y.
 This is \emph{not necessary} and makes it look like you're trying to show off your connections.
-In short, these nuances represent some cultural differences that you may encounter and will gradually adapt to (more on cultural differences in \S\ref{sec:cultural}).
+In short, these nuances represent some cultural differences that you may encounter and will gradually adapt to.
+More on cultural differences in \S\ref{sec:cultural}.
 
@@
 This section aims to shed some light to these questions.
-
-Tab.~\ref{tab:cost} shows the budget breakdown for a GRA per year (this level of details is what faculty actually uses when applying for funding).
-These numbers are based on my experience at public universities in the US.
-Private universities may have different numbers.
-For simplicity, I will assume the department has a 9-month stipend of \$27000 (GMU actually pays more) and therefore a 3-month summer of \$9000.
-I will also use GMU tuition rate of about \$15,000/year for full-time study (which is quite cheap compared to private universities, e.g., MIT charges around \$50K) and a 58.9\% rate on \emph{indirect cost}, which is what GMU charges for overhead or administrative costs (yes, after all, universities are businesses!).
-Finally, I assume the student makes two conference trips per year, one domestic and one international (so conf.
-registration, airline tickets, taxi, meals, etc are all included).
-At the end, the total budget comes out to be \$69K.
-\emph{The summary is that while you're paid X, your advisor probably pays 2X for you}.
 
@@
 \end{table}
+
+Tab.~\ref{tab:cost} shows the budget breakdown for a GRA per year (this level of details is what faculty actually uses when applying for funding).
+These numbers are based on my experience at public universities in the US.
+Private universities may have different numbers.
+For simplicity, I will assume the department has a 9-month stipend of \$27000 (GMU actually pays more) and therefore a 3-month summer of \$9000.
+I will also use GMU tuition rate of about \$15,000/year for full-time study (which is quite cheap compared to private universities, e.g., MIT charges around \$50K) and a 58.9\% rate on \emph{indirect cost}, which is what GMU charges for overhead or administrative costs (yes, after all, universities are businesses!).
+Finally, I assume the student makes two conference trips per year, one domestic and one international (so conf.
+registration, airline tickets, taxi, meals, etc are all included).
+At the end, the total budget comes out to be \$69K.
+\emph{The summary is that while you're paid X, your advisor probably pays 2X for you}.
 

### 2023-09-23 demystify.tex
Updates from Overleaf

@@
 This shows that you're serious and have done homework on places you're applying to.
-Admission committee will look for this part (typically written at the end of your statement), so do not skip it.
+Admission committee will look for this part at the end of the statement, so do not skip it.
 
@@
 
+\begin{commentbox}[Vu]
 Sometime PhD admission committee in the US will share a document such as \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Foreign-Top-Schools}{this one}, which lists the top schools in several countries.
@@
 During my time at UNL and now here at GMU, I have looked at Vietnamese applications (whether they are assigned to me or not) and provide input to their reviewers, e.g., X is the top tech school in Vietnam and so it should be \emph{top} instead of \emph{unknown} foreign, which makes a huge difference.
+\end{commentbox}
 
@@
 
-The following lists the minimum for GMU (being above this might not mean much, but below is a \red{red flag}).
+\begin{commentbox}[Vu]
+The minimum for GMU (being above this might not mean much, but below is a \red{red flag}).
 \begin{itemize}
@@
 \end{itemize}
+\end{commentbox}
 
@@
 If you're admitted to a good CS PhD program, you will be fine.
-A good school knows that it has to be competitive, otherwise it will lose good students to other schools.
+A good school would know that it has to be competitive, otherwise it will loose good students to other schools.
 For example, at GMU, every year we discuss about improving the benefits, and especially stipend, for our graduate students.
@@
 The college and department will set the rate for 9-month graduate assistant stipend.
-TA, which is paid by the department, will likely be that amount set by the college, but RA might be higher depending on the stage of the student (1st year vs ABD\footnote{All but dissertation: really close to graduate.}) and the prof.
+TA, which is paid by the department, will likely be that amount but RA might be higher depending on the stage of the student (1st year vs ABD\footnote{All but dissertation: really close to graduate.}) and the prof.
 \tcblower
@@
 
-It is worth noting that if you got recruited by a prof.
+\begin{commentbox}[Vu]
+If you got recruited by a prof.
 who would give you RA right away, it's very likely you will get admitted.
 For example, if a prof., even if not in PhD admission committee, wants to work with and funds you, the PhD admission committee will respect that decision and admit you (unless your application has many red flags).
+\end{commentbox}
 
@@
 
-Faculty often receive  "cold" e-mails from international students seeking for admission, TA, and RA.
+Faculty received many "cold" e-mails from international students seeking for admission, TA, and RA.
 Most of the time, we ignore these emails, but in some rare occasions we do answer them.
@@
 Students need to respect and appreciate varied opinions, backgrounds, and cultures.
-Unlike some countries where certain voices are ignored, in the US, all perspectives are valued equally (especially at universities).
+Unlike some countries where certain voices are marginalized, in the US, all perspectives are valued equally (especially at universities).
 Racism or discrimination will have serious consequences, including academic and disciplinary actions.
@@
 
-\paragraph{Illegal Software} Using illegal software or applying "cracks" to software is very common in many countries.
+\paragraph{Illegal Software} Using illegal/cracked software is very common in many countries.
 But do not install or use them on university computer (e.g., computers given by your department or advisor).
@@
 They have very sophisticated tools to detect illegal software and will sue your university/department.
-Imagine your department or advisor being sued for million dollars, and it is you who caused it.
+Imagine your department or advisor being sued for half a million dollars, and it is you who caused it.
 If you need software, ask your advisor or the department.

### 2023-09-25 demystify.tex
formatting, minor

@@
 
-\begin{commentbox}[Vu]
+\begin{commentbox}[Vu:]
 One of the reasons I create this document is that my colleagues at GMU are interested in recruiting Vietnamese students and are surprised when seeing very few applications in Vietnam (e.g., each year our CS program receives more than 350 PhD applications, most of which are international but only 3--4 are from Vietnam).
-In general the number of CS PhD applications from Vietnam to US universities is low and more would be very welcomed.
+In general the number of CS PhD applications from Vietnam to US universities is low and more would be welcomed.
 \end{commentbox}
@@
 
-\begin{commentbox}[Hakan]
+\begin{commentbox}[Hakan:]
 At GMU, for full consideration, students should make sure to submit \textbf{ALL} required documents by the application deadline, and should never assume that some required documents (such as official TOEFL scores or official diplomas/transcripts) will be waived by the admissions office.
@@
 
-\begin{commentbox}[Vu]
+\begin{commentbox}[Vu:]
 Many international students mention Scopus Q1, which consists of various journals from IEEE, Elsevier, and many other publishers.
@@
 
-\begin{commentbox}[Craig]
+\begin{commentbox}[Craig:]
 GMU and many other universities allow you to upload your published papers and other writing samples.
@@
 
-\begin{commentbox}[Hung]
+\begin{commentbox}[Hung:]
 The reputation of VinAI has been increasing steadily over the past few years; many of my colleagues heard about VinAI.
@@
 
-\begin{commentbox}[Thanh]
+\begin{commentbox}[Thanh:]
 Due to academic culture, professors in Vietnam usually aim for (international) journals instead of conferences.
@@
 
-\begin{commentbox}[Vu]
+\begin{commentbox}[Vu:]
 Some letter writers ask the applicants to write their own letters for them to sign.
@@
 
-\begin{commentbox}[Hung]
+\begin{commentbox}[Hung:]
 A sad reality is that most professors in Vietnam \textbf{DO NOT} know how to write a good letter, or are lazy in writing letters hence delegate the writing to the students.
@@
 
-\begin{commentbox}[Hung]
+\begin{commentbox}[Hung:]
 I think asking for a “research vision” from a Ph.D.
@@
 
-\begin{commentbox}[Vu]
+\begin{commentbox}[Vu:]
 I often read the research statement (and the LORs) first. If I am
@@
 
-\begin{commentbox}[Hung]
+\begin{commentbox}[Hung:]
 I personally don’t care about which schools students attended. There are many other ways to evaluate an applicant.
@@
 
-\begin{commentbox}[Vu]
+\begin{commentbox}[Vu:]
 Sometime PhD admission committee in the US will share a document such as \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Foreign-Top-Schools}{this one}, which lists the top schools in several countries.
@@
 
-\begin{commentbox}[Thanh]
+\begin{commentbox}[Thanh:]
 Vietnamese universities typically offer specialized programs, such as the talented engineer program at HUST, that have highly competitive entrance exams and a limited number of available slots (e.g., 30 per year).
@@
 
-\paragraph{GRE} Most CS programs in the US \textbf{no longer require GREs}, so you don't need to take them.
+\paragraph{GRE} Most good CS PhD programs in the US \textbf{no longer require GREs}, so you don't need to take them.
 However, they might be useful for international students from programs we are not familiar with.
 
-\begin{commentbox}[Vu]
+\begin{commentbox}[Vu:]
 Personally I do not pay attention to GRE scores as I do not believe they are good indicator for research. However, if you have good GRE scores then you should include them in your application.
@@
 
-\begin{commentbox}[Vu]
+\begin{commentbox}[Vu:]
 The minimum for GMU (being above this might not mean much, but below is a \red{red flag}).
@@
 
-\begin{commentbox}[Vu]
+\begin{commentbox}[Vu:]
 At GMU, we are encouraged to interview candidates.
@@
 
-\begin{commentbox}[Vu]
+\begin{commentbox}[Vu:]
 GMU has \emph{Virtual} Open House (VOH), e.g., \url{https://cs-gmu.github.io/cs-phd-voh-s23/}, which I've co-organized in the last two years.
@@
 
-\begin{commentbox}[Vu]
+\begin{commentbox}[Vu:]
 TA and RA at GMU have similar benefits in tuition waiving and insurance.
@@
 
-\paragraph{How to get TA?}  Unless you have other funding such as RA or Fellowships, TA is typically a default thing.
+\paragraph{How to get TA?}  Unless you have other funding such as RA or Fellowships, TA is typically a default thing for good CS PhD programs.
 When you apply to be a full-time student,  state that you need financial assistant.
@@
 
-\begin{commentbox}[Vu]
+\begin{commentbox}[Vu:]
 At GMU CS, students admitted with TA have  4 years of GTA guaranteed and also receive  stipend for the \textbf{first} summer.
@@
 This allows you to see what teaching is like, which is especially helpful for research career where you often give talks and tell people about your work.
-Note that GMU sometimes has classes that a more senior student can teach.
-In that case, you will be paid as a lecturer, which is higher than GTA.
-This is a good opportunity for students to get teaching experience and also get paid more.
+GMU sometimes has classes that a more senior student can teach.
+In that case, you will be paid as GTA or even sometimes as a lecturer.
+This is a good opportunity for students to get teaching experience and also get paid more (as a lecturer).
+
 \subsection{Research Assistant (RA)}
@@
 
-\begin{commentbox}[Vu]
+\begin{commentbox}[Vu:]
 If you got recruited by a prof.
@@
 
-\begin{commentbox}[Vu]
+\begin{commentbox}[Vu:]
 At GMU, PhD applicants are automatically eligible for a Presidential Fellowship.
@@
 
-International students not familiar with US universities often put them into \emph{two} bins:  (i) very top schools that they dream about such as Stanford, MIT, Princeton, Harvard and (ii) everything else.
+International students not familiar with US universities often put them into \emph{two} bins:  (i) very top schools that they dream about, and (ii) everything else.
 In many cases, they use resources such as rankings from US News, which are \emph{highly} questionable\footnote{\url{https://cra.org/cra-statement-us-news-world-report-rankings-computer-science-universities/}}.
@@
 
-\begin{commentbox}[Hung]
+\begin{commentbox}[Hung:]
 I always encourage the students I admitted to talk with my students and the students of other faculty in other schools who admitted them.
@@
 
-The short answer is that tenured-track faculty such as assistant professors are more likely to be young and active in research (they have to, in order to get tenure).
+%The tenure system in universities in the US (and other countries) can be confusing.
+In most cases, you probably don't need to know much about it.
+But it is helpful to know a few things so you know how to find PhD advisor.
+First, faculty of a CS (and other) department can be divided into 2 types:  (i) tenured and tenure-track and (ii) non-tenured or teaching faculty.
+Tenured/Tenure -track faculty focus on research and related tasks such as mentoring and getting grants to support students, etc.
+Their teaching load is very light, typically they just teach 1 or at most 2 classes a semester.
+Non-tenured/teaching faculty focuses on teaching, and often teach 3-4 classes per semester.
+They can mentor Ph.D.
+students but they typically cannot serve as a formal advisor of Ph.D.
+students (i.e., they cannot graduate PhD students).
+Note that there is also affiliated faculty who is tenured/tenure-track profs.
+from another department.
+They can serve in Ph.D.
+dissertation committee but they likely cannot serve as the main advisor of CS students.
+
+The short answer is that tenure-track faculty such as assistant professors are more likely to be young and active in research (they have to, in order to get tenure).
 Thus, they will likely have more time to work with you and push you to do research and publish.
@@
 
-\begin{commentbox}[Thanh]
+\begin{commentbox}[Thanh:]
     In my opinion, having a well-suited advisor is crucial for a successful PhD and research career.
@@
 % \subsubsection{Strong Faculty vs High-ranked School}
-% \begin{commentbox}[Thanh]
+% \begin{commentbox}[Thanh:]
 % When considering PhD programs, we often wonder if we should prioritize a high-ranking university or a professor with a strong reputation?
@@
 
-% \begin{commentbox}[Thanh]
+% \begin{commentbox}[Thanh:]
 % \end{commentbox}
@@
 Prof.
-X, and I also collaborate with ACM Distinguished Scientist Y.
+X, and I also collaborate with NASA Distinguished Scientist Y.
 This is \emph{not necessary} and makes it look like you're trying to show off your connections.

### 2023-10-05 demystify.tex
Updates from Overleaf

--- 2023-09-23 demystify.tex
@@
 
-For Vietnamese students, it's worth mentioning the \href{https://vef2.org/}{VEF2.0 program}, which has helped many good students in gaining acceptance to top PhD programs in the U.
+For Vietnamese students, it's worth mentioning the \href{https://vef2.org/}{VEF2.0 program}, which has helped many good students in gaining acceptance to top PhD programs in the US.
 VEF2.0 follows an interesting model where US faculty members from leading institutions are invited to conduct rigorous interviews with VEF students and subsequently provide reference letters on their behalf.
@@
 Note that funding is typically more available for PhD students than 
-Masters. Undergraduate students typically have no funding and also have to pay international tuition, which is very expensive!.
+Masters. Undergraduates typically have no funding and also have to pay international tuition, which is very expensive!.
 
@@
 
-\section{Miscs and FAQs}
-\epigraph{"I want to share something with you – the three little sentences that will get you through life; number 1: Cover for me, number 2: Oh, good idea, Boss, and number 3: It was like that when I got here."}{\textsc{The Simpsons}}
-
-\subsection{What can you do to increase your admission chance?}\label{sec:improve-your-chance}
-
- Show something that makes you \textbf{stand out}, e.g., do you have a degree or background in \emph{biology} or \emph{music} and want to integrate them with CS?
-are you a female or a minority in CS (research for "URM minority in CS in the US" on Google)?
-Do you participate in outreach activities that help increase diversity and inclusion in CS?
-All of these are unique and would get noticed from reviewers.
-
-Even if you do not have research experience, you can talk about your personal projects, as long as they can help show you can do research.
-For example, if you have an open-source project on Github that is used by many people, has lots of stars in Github, do talk about it.
-If you write technical, research-like blogs, talk about them too.
-
-In his \href{https://matt.might.net/articles/how-to-apply-and-get-in-to-graduate-school-in-science-mathematics-engineering-or-computer-science/}{post}, Matt Might was initially unsure about an application.
-However, upon learning that the applicant had led a 100km hike in the Himalayas, he decided to accept the applicant.
-This is a good example of being "stand out", and I would also advocate for that student as this shows they have the persistence and determination required for research.
-
-%It is worth noting that in many cases,  an application could be rejected by most reviewers, however have a "champion", who doesn't even have to be in the Admission comittee, that strongly advocate for your profile.
-
-\subsection{How to rank or select a CS graduate program?}
+\section{Choosing Schools and Professors}
+\subsection{Selecting and Ranking Schools for CS}
    \begin{center}
@@
 Pay attention to their "level of excitement" being in the group.
+\end{commentbox}
+
+\begin{commentbox}[Xiaokuan]
+Chinese students often only look at USNews rankings when selecting their Ph.D. universities (I did that, too, when I was applying for Ph.D. positions).
+Now that I am a professor, I find it to be the least promising way.
+%
+The reason is that USNews does not provide a good metric for evaluating the quality of the Ph.D. program.
+%
+If you want to do great research, CSRankings is the best way to find good and active professors (which did not exist when I was applying),
+since it solely focuses on publications at top-tier CS conferences.
+%
+Also, 
+I think Ph.D. is not only about research; 
+you need to also consider your daily life there, since you will (probably) stay for at least five years.
+%
+You might regret it if you did not consider this seriously before applying.
+%
+Whether your research interest matches that of your advisor is also important;
+if there is a mismatch,
+either you or your advisor has to make compromises,
+which often leads to disagreements or conflicts.
+
 \end{commentbox}
@@
 
+\section{Miscs and FAQs}
+\epigraph{"I want to share something with you – the three little sentences that will get you through life; number 1: Cover for me, number 2: Oh, good idea, Boss, and number 3: It was like that when I got here."}{\textsc{The Simpsons}}
+
+\subsection{What can you do to increase your admission chance?}\label{sec:improve-your-chance}
+
+ Show something that makes you \textbf{stand out}, e.g., do you have a degree or background in \emph{biology} or \emph{music} and want to integrate them with CS?
+are you a female or a minority in CS (research for "URM minority in CS in the US" on Google)?
+Do you participate in outreach activities that help increase diversity and inclusion in CS?
+All of these are unique and would get noticed from reviewers.
+
+Even if you do not have research experience, you can talk about your personal projects, as long as they can help show you can do research.
+For example, if you have an open-source project on Github that is used by many people, has lots of stars in Github, do talk about it.
+If you write technical, research-like blogs, talk about them too.
+
+In his \href{https://matt.might.net/articles/how-to-apply-and-get-in-to-graduate-school-in-science-mathematics-engineering-or-computer-science/}{post}, Matt Might was initially unsure about an application.
+However, upon learning that the applicant had led a 100km hike in the Himalayas, he decided to accept the applicant.
+This is a good example of being "stand out", and I would also advocate for that student as this shows they have the persistence and determination required for research.
+
+%It is worth noting that in many cases,  an application could be rejected by most reviewers, however have a "champion", who doesn't even have to be in the Admission comittee, that strongly advocate for your profile.
+
 \subsection{Can I apply to CS PhD if my undergrad was not in CS or related areas?}
@@
 Moreover, if you don't know the prof., do not use Mr.
-or Mrs., or Firstname.
-To me it seems a bit disrespectful.
-As you know that prof.
+or Mrs., or Firstname---to me it seems a bit disrespectful.
+As you get to know that prof.
 better and depends on their preference, you may call them by their firstname.
@@
 \paragraph{Acknowledgement} Many people have contributed to this document.
-Profs.
-Craig Yu (GMU), Hakan Aydin (GMU), and Hung Le (UMass) provided valuable input in the early version.
+Profs. Craig Yu (GMU), Hakan Aydin (GMU), 
+Xiaokuan Zhang (GMU), and Hung Le (UMass) provided valuable input in the early version.
 Other GMU faculty members also have provided feedback and contributions.

### 2023-10-06 demystify.tex
update

--- 2023-09-25 demystify.tex
@@
 
-\subsection{Should you contact a US professor? What to do to get a desired reply?}
+\subsection{Should you contact a professor? What to do to get a desired reply?}
 
@@
 If you want to contact a prof.
-to ask about \emph{research opportunities}, or \emph{GTA/GRA} support, then \emph{yes}, I believe you should---it is \emph{worth it}.
+to ask about \emph{research opportunities}, or \emph{funding} support, then \emph{yes}, I believe you should---it is \emph{worth it}.
 However, you need to do it the \emph{right way}.
 
-First, read the prof's website, see if they say something about contacting them. Many profs. indicate how prospective students should contact them (e.g., using specific email subjects). 
-In general, the best way to catch the prof.'s attention is to \emph{customize your email} to that faculty.
+First, read the prof's website, see if they say something about contacting them. Many profs. explicitly indicate how prospective students should contact them (e.g., using specific email subjects). 
+In general, the best way to catch the prof.'s attention is to \emph{customize your email} for them.
 For example, read their papers, know what they work on, and see if you are interested in their research.

### 2023-10-06 demystify.tex
Updates from Overleaf

--- 2023-10-05 demystify.tex
@@
 
-\section{Choosing Schools and Professors}
+\section{Choosing Schools and Professors}\label{sec:schoolsandprofs}
+
+Choosing a school and an advisor is clearly one of the most important things in your mind when you apply to and especially when you get admitted to a PhD program in CS.
+This is further complicated due to cultural differences and unfamiliarity of  international students to the US system and faculty.
+I hope this section will clear some confusions and help you make more informed decision.
+
+\subsection{Universities}
+
+There are many types of universities in the US (as well as other countries). But here I mainly focus on the object of obtaining a Ph.D degree in CS.
+
+\subsection{Schools that offer Ph.D. in CS}  Most university will have a "Computer Science" or CS-like programs. 
+However, while many universities offer Ph.D.
+degree in their CS programs, many do not.
+These universities might offer just Bachelor degrees (e.g., BS) and no graduate studies (i.e., no MS or Ph.D degrees), or they must just offer MS programs (but no PhD).
+
+Thus, if your goal is Ph.D.
+in CS, you will have to filter out schools that do not offer such degree.
+This might sound obvious but many students (even those in the US) get confused due to the large number of universities.
+For example, Penn State in University Park has Ph.D.
+in CS,  but Penn State in Harrisburg only has BS and MS in CS and Penn State in York only has BS in CS.
+On the other hand, multiplication locations of the University of Texas, e.g., Austin, Dallas, Arlington, all have PhDs in CS.
+Also, some schools might offer PhDs in related fields, but not CS.
+
+\subsubsection{R1, R2, ...}
+
 \subsection{Selecting and Ranking Schools for CS}
@@
 
-\subsection{Tenured or tenure-track faculty? Who do you choose?}
+\subsection{Choosing a Professor}
+
+\subsubsection{Tenured, Tenure Track, Teaching,  Assistant Profs., Associate Profs}
+
+\subsubsection{Tenured or tenure-track faculty? Who do you choose?}
 

### 2023-10-06 demystify.tex
updates

@@
 
-\section{Getting Admitted and Choosing the Right School}
+\section{Getting Admitted}
 
@@
 
-Choosing a school and an advisor is clearly one of the most important things in your mind when you apply to and especially when you get admitted to a PhD program in CS.
+Choosing a school and an advisor is clearly one of the most important things in your mind when you apply to and especially when you get admitted multiple programs.
 This is further complicated due to cultural differences and unfamiliarity of  international students to the US system and faculty.
-I hope this section will clear some confusions and help you make more informed decision.
+This section aims to mitigate some confusions and help you make more informed decision.
 
@@
 
-There are many types of universities in the US (as well as other countries). But here I mainly focus on the object of obtaining a Ph.D degree in CS.
-
-\subsection{Schools that offer Ph.D. in CS}  Most university will have a "Computer Science" or CS-like programs. 
+There are many types of universities in the US (as well as other countries). But here I focus the discussion in the context of obtaining a Ph.D degree in CS.
+
+\subsection{Schools that offer Ph.D. in CS}  Most university will have CS programs. 
 However, while many universities offer Ph.D.

### 2023-10-15 demystify.tex
Updates from Overleaf

--- 2023-10-06 demystify.tex
@@
 \newcommand{\red}[1]{{\color{red}{#1}}}
+\newcommand{\xz}[1]{\mycomment{Xiaokuan}{[#1]}}{}
 
@@
 %
-Whether your research interest matches that of your advisor is also important;
-if there is a mismatch,
-either you or your advisor has to make compromises,
-which often leads to disagreements or conflicts.
-
 \end{commentbox}
@@
 
+\begin{commentbox}[Xiaokuan]
+Whether the student's research interest matches that of the advisor is very important;
+if there is a mismatch,
+either the student or the advisor has to make compromises,
+which often leads to disagreements or conflicts. 
+IMO, the advisor should be the one who {\it guides}  students to do research while allowing students to pursue their own interest,
+instead of {\it dictating} their research.
+
+\end{commentbox}
 % \subsubsection{Strong Faculty vs High-ranked School}
@@
 However, you need to do it the \emph{right way}.
+
+\xz{Applying for Ph.D. and contacting a potential Ph.D. advisor is a classic {\bf `why me, why you'} problem, 
+similar to looking for a job in a company.
+On a high level,
+you need to show that you have done your {\it homework}
+regarding the professor and the university,
+and clearly explain:
+1) why do you think you are a good fit in professor A's group?
+2) why do you want to be advised by professor A, not B? 
+3) why do you want to apply for university X, not Y?
+If you don't want to spend time to do your homework,
+the chance of getting a reply is close to zero.
+}
 

### 2023-10-15 demystify.tex

--- 2023-10-06 demystify.tex
@@
 If you want to contact a prof.
-to ask about \emph{research opportunities}, or \emph{funding} support, then \emph{yes}, I believe you should---it is \emph{worth it}.
+to ask about \emph{research opportunities}, or \emph{GTA/GRA} support, then \emph{yes}, I believe you should---it is \emph{worth it}.
 However, you need to do it the \emph{right way}.
+
+\xz{Applying for Ph.D. and contacting a potential Ph.D. advisor is a classic {\bf `why me, why you'} problem, 
+similar to looking for a job in a company.
+On a high level,
+you need to show that you have done your {\it homework}
+regarding the professor and the university,
+and clearly explain:
+1) why do you think you are a good fit in professor A's group?
+2) why do you want to be advised by professor A, not B? 
+3) why do you want to apply for university X, not Y?
+If you don't want to spend time to do your homework,
+the chance of getting a reply is close to zero.
+}
 

### 2023-10-19 demystify.tex
updates

@@
 
+\tableofcontents
+
 \section{Should You Apply?}
@@
 
-\section{How is Your Application Evaluated?}
+\section{How is Your Application Evaluated?}\label{sec:evalapps}
 
@@
 
-\begin{commentbox}[Hung:]
-I personally don’t care about which schools students attended. There are many other ways to evaluate an applicant.
-\end{commentbox}
+% \begin{commentbox}[Hung:]
+% I personally don’t care about which schools students attended. There are many other ways to evaluate an applicant.
+% \end{commentbox}
 
@@
 
-\section{Getting Admitted}
+\section{Getting Admitted}\label{sec:accepted}
 
@@
 
-\subsection{Research Assistant (RA)}
+\subsubsection{Research Assistant (RA)}
 RA is provided through a professor through their own funding so you can work on their project.  
@@
 A common scenario is that you first get admitted with TA, and then after a year or two find an advisor to support you with RA.
-
-\begin{commentbox}[Vu:]
-If you got recruited by a prof.
-who would give you RA right away, it's very likely you will get admitted.
+See how to contact a prof.
+in \S\ref{sec:contact} for research opportunities.
+
+\begin{commentbox}[Vu:]
+If you got recruited by a prof., who would give you RA right away, it's very likely you will get admitted.
 For example, if a prof., even if not in PhD admission committee, wants to work with and funds you, the PhD admission committee will respect that decision and admit you (unless your application has many red flags).
@@
 
-There are many types of universities in the US (as well as other countries). But here I focus the discussion in the context of obtaining a Ph.D degree in CS.
-
-\subsection{Schools that offer Ph.D. in CS}  Most university will have CS programs. 
+This section aims to explain different types of universities in the US in the context of obtaining a CS Ph.D degree.
+
+\subsection{Schools that offer Ph.D. in CS}
+
+Most universities the US will have CS programs. 
 However, while many universities offer Ph.D.
 degree in their CS programs, many do not.
-These universities might offer just Bachelor degrees (e.g., BS) and no graduate studies (i.e., no MS or Ph.D degrees), or they must just offer MS programs (but no PhD).
+These universities might offer just Bachelor degrees (e.g., BS) and no graduate studies (i.e., no MS or Ph.D degrees), or they just offer MS programs (but no PhD).
 
 Thus, if your goal is Ph.D.
-in CS, you will have to filter out schools that do not offer such degree.
+in CS, you will have to filter out schools that do not offer such a degree.
 This might sound obvious but many students (even those in the US) get confused due to the large number of universities.
 For example, Penn State in University Park has Ph.D.
-in CS,  but Penn State in Harrisburg only has BS and MS in CS and Penn State in York only has BS in CS.
+in CS,  but Penn State in Harrisburg only has BS and MS in CS, and Penn State in York only has BS in CS.
 On the other hand, multiplication locations of the University of Texas, e.g., Austin, Dallas, Arlington, all have PhDs in CS.
-Also, some schools might offer PhDs in related fields, but not CS.
-
-\subsubsection{R1, R2, ...}
+
+While this can be confusing at first, a little research, e.g., searching for PhD in CS from the school website, will help you find out.
+All schools listed in \S\ref{sec:ranking} have PhD studies in CS, so you can start there.
+
+%\subsubsection{R1, R2, ...}
 
@@
 However, in general, rankings can be superficial and you need to do more research to be informed and make better decision.
-For example, if you get admissions to several places, you should contact profs.
+For example, if you get admissions to several places, you should consider attending Open Houses and contact profs.
 that you're interested in at those place and talk to them.
-They would be more willing to chat to you now that you have been admitted.
+They would be more willing to chat with you now that you have been admitted.
 Ask them questions about their work, how they manage students, their expectations.
 You can even ask to contact their students.
+See more in \S\ref{sec:accepted} on what to do after getting admitted.
 
@@
 
-\subsubsection{Tenured, Tenure Track, Teaching,  Assistant Profs., Associate Profs}
+%\subsubsection{Tenured, Tenure Track, Teaching,  Assistant Profs., Associate Profs}
 
@@
 Do you have any tips for Vietnamese students who want to connect with US professors and work as research assistants?
-    \end{commentbox}
-
-\begin{commentbox}[Xiaokuan]
+\textbf{Vu}: see sec ~\ref{sec:contact} to contact a professor for research opportunities.
+    \end{commentbox}
+
+\begin{commentbox}[Xiaokuan:]
 Whether the student's research interest matches that of the advisor is very important;
@@
 
-\subsection{Should you contact a professor? What to do to get a desired reply?}
+\subsection{Should you contact a professor? What to do to get a desired reply?}\label{sec:contact}
 
@@
 to \emph{ask about your admission chance}, please \textbf{don't}.
-We don't know and can't answer because as explained throughout this document, we don't make individual decisions and might not even be assigned to evaluate your application.
+We don't know and can't answer because as explained in \S\ref{sec:evalapps}, we don't make individual decisions and might not even be assigned to evaluate your application.
 It is the same as sending a paper draft to a journal editor and ask them if your paper has a chance.
@@
 
-\xz{Applying for Ph.D. and contacting a potential Ph.D. advisor is a classic {\bf `why me, why you'} problem, 
+First, read the prof's website, see if they say something about contacting them. Many profs. explicitly indicate how prospective students should contact them (e.g., using specific email subjects). 
+In general, the best way to catch the prof.'s attention is to \emph{customize your email} for them.
+For example, read their papers, know what they work on, and see if you are interested in their research.
+Then send them an email talking how/why their work would match yours.
+In contrast, if you write a generic email that can be sent to multiple professors (e.g., if you just change some names and keywords in the email), you will not get a response.
+
+Below is a good example that I would reply to.
+
+\begin{commentbox}[Dear Prof. Nguyen,]
+
+I am writing to inquire about potential research opportunities as a GRA in your group at GMU. Currently I am an undergraduate student in Computer Science at UNIV and plan to graduate in May 2023.
+
+I have read your TSE'21 paper on numerical invariant generation, and I am interested in this line of dynamic invariant research.
+I have worked (optional: with prof.
+Y at Z) on static program analysis and I think it could be used to tackle the spurious issues mentioned in your paper.
+I have a small paper at conference/workshop C and a project on symbolic execution at Github G.
+
+...
+
+This is a good example because it is clearly written just for me.  It shows that the student knows about my work on invariant generation and has related  background (paper C and project G). 
+\end{commentbox}
+
+Finally, profs. are very busy so don't take it personally if you don't get anything from them (though I would be very surprised if such thoughtful emails get no replies!).
+
+\begin{commentbox}[Xiaokuan]
+    Applying for Ph.D. and contacting a potential Ph.D. advisor is a classic {\bf `why me, why you'} problem, 
 similar to looking for a job in a company.
@@
 the chance of getting a reply is close to zero.
-}
-
-First, read the prof's website, see if they say something about contacting them. Many profs. explicitly indicate how prospective students should contact them (e.g., using specific email subjects). 
-In general, the best way to catch the prof.'s attention is to \emph{customize your email} for them.
-For example, read their papers, know what they work on, and see if you are interested in their research.
-Then send them an email talking how/why their work would match yours.
-In contrast, if you write a generic email that can be sent to multiple professors (e.g., if you just change some names and keywords in the email), you will not get a response.
-
-Below is a good example that I would reply to.
-
-\begin{commentbox}[Dear Prof. Nguyen,]
-
-I am writing to inquire about potential research opportunities as a GRA in your group at GMU. Currently I am an undergraduate student in Computer Science at UNIV and plan to graduate in May 2023.
-
-I have read your TSE'21 paper on numerical invariant generation, and I am interested in this line of dynamic invariant research.
-I have worked (optional: with prof.
-Y at Z) on static program analysis and I think it could be used to tackle the spurious issues mentioned in your paper.
-I have a small paper at conference/workshop C and a project on symbolic execution at Github G.
-
-...
-
-This is a good example because it is clearly written just for me.  It shows that the student knows about my work on invariant generation and has related  background (paper C and project G). 
-\end{commentbox}
-
-Finally, profs. are very busy so don't take it personally if you don't get anything from them (though I would be very surprised if such thoughtful emails get no replies!).
+\end{commentbox}
 
@@
 
- Show something that makes you \textbf{stand out}, e.g., do you have a degree or background in \emph{biology} or \emph{music} and want to integrate them with CS?
+ Show something that makes you \textbf{stand out}, e.g., do you have a degree or background in \emph{dance} or \emph{music} and want to integrate them with CS?
 are you a female or a minority in CS (research for "URM minority in CS in the US" on Google)?
@@
     Typically, it takes 5--7 years for CS PhD in the US.
-This can be longer than CS PhDs at universities in other countries, which might require MS first (recall that US PhD programs do not require MS and you can get MS along the way to PhD).
+This can be longer than CS PhDs in other countries, which might require MS first (recall that US PhD programs do not require MS and you can get MS along the way to PhD).
 Within these 5--7 years, CS PhD students often take a ``leave of absence'' for 1--2 semesters to do internship at companies and research labs.
@@
 PhD students often ask why their salary is so low compared to ludicrous grants their advisors get.
-They also wonder why their offer letters sometime mentioned that their benefits worth way more than what they actually receive (i.e., stipend).
+They also wonder why their offer letters sometime mentioned that their benefits appear better more than what they actually receive (i.e., stipend).
 This section aims to shed some light to these questions.
@@
 
-\section{Cultural Differences and Tips}\label{sec:cultural}
-
-This section lists various things related to studying the US that I notice international students often get surprised or confused.
+\section{Cultural Differences and Other Issues}\label{sec:cultural}
+
+This section lists various issues that international students should pay attention to.
 
@@
 
-\paragraph{Academic Integrity} Plagiarism and cheating (e.g., exams, coding) is a big no-no in the US.
+\paragraph{Academic Integrity} Plagiarism and cheating (e.g., exams, coding) is a BIG no-no in the US.
 If you're caught cheating, you will face very heavy consequence and likely be expelled from the university (e.g., after the second time at GMU).
-This is very different from many international countries where cheating is common and often tolerated.
+This is quite different from many international countries where cheating is common and often tolerated.
 Faculty is extremely good at detecting cheating (we have been dealing with these situations so many times over so many years), and \emph{will} report cheating cases.
@@
 Here is the typically steps: (i) a faculty suspecting a cheating case will report it to the Office of Academic Integrity (OAI) at the university;  (ii)  OAI will take over and investigate the case; and finally (iii) OAI will make the final decision.
-It is important to note that after receiving the report from your prof., OAI \emph{completely} takes over (so begging your professor will not help because they simply are no longer involved and can't do much here).
-
-\paragraph{Illegal Software} Using illegal/cracked software is very common in many countries.
-But do not install or use them on university computer (e.g., computers given by your department or advisor).
+It is important to note that after receiving the report from your prof., OAI \emph{completely} takes over and makes decision.
+This means begging your professor will not help because they simply are no longer involved in the case.
+
+\paragraph{Illegal Software} Using illegal/cracked software is very common in many countries (and even in the US).
+However, \emph{do not} install or use them on university computers (e.g., those given by your department or advisor).
 It is unlikely that the university will track you down, but it is the \emph{software company} that will.
 They have very sophisticated tools to detect illegal software and will sue your university/department.
-Imagine your department or advisor being sued for half a million dollars, and it is you who caused it.
+Imagine your department or advisor being sued for half a million dollars, and it is \emph{you} who caused it.
 If you need software, ask your advisor or the department.
@@
 \paragraph{History} This document was conceived during a lunch with Craig Yu at GMU.
-We talked on about why we were not able to attract good Vietnamese international students, despite having a much stronger CS program than many schools that these students want to go to (part of the reason is described in \S\ref{sec:ranking}) and wished there were a way for international students to know about the US PhD programs (as well as for US faculty to understand more about international students and therefore have better chance of recruiting and working with them).
+We talked on about why GMU were not able to attract good Vietnamese and other international students, despite having a much stronger CS program than many schools that these students want to go to (part of the reason is described in \S\ref{sec:ranking}) and wished there were a way for international students to know about the US PhD programs (as well as for US faculty to understand more about international students and therefore have better chance of recruiting and working with them).
 I was also  a member of the large VietPhD group on Facebook and saw many questions from students about PhD programs.
-However, most active participant are students (and some few faculty) in non-CS fields or not in US.
+However, most active participant are students in non-CS fields or not in US.
 Like ChatGPT, their answers to CS PhD program in the US could be informative and helpful, but unfortunately not always accurate and sometimes leading to more confusion.

### 2023-10-19 demystify.tex
Updates from Overleaf

--- 2023-10-15 demystify.tex
@@
 
-For Vietnamese students, it's worth mentioning the \href{https://vef2.org/}{VEF2.0 program}, which has helped many good students in gaining acceptance to top PhD programs in the US.
+For Vietnamese students, it's worth mentioning about the \href{https://vef2.org/}{VEF2.0 program}, which has helped many good students in gaining acceptance to top PhD programs in the US.
 VEF2.0 follows an interesting model where US faculty members from leading institutions are invited to conduct rigorous interviews with VEF students and subsequently provide reference letters on their behalf.
@@
 
-Finally, this is something easy to do, but is missed by many
+Finally, this is something easy to do, but is neglected by many
 applicants: \textbf{customize the statement} for the school you're applying to,
-e.g., why do you apply here? provide names of professors who you're interested in (in many cases your application will be forwarded to them for evaluation).
+e.g., why do you apply \emph{here}? provide names of professors who you're interested in (in many cases your application will be forwarded to them for evaluation).
 This shows that you're serious and have done homework on places you're applying to.
@@
 
-So what can you do about this?
-several things including asking your CS dept to put itself on CSRankings (it's the easiest way to get CS people to know about the school) and explaining about your school in your statement (ask your LOR writer to do that too).
-Of course, if you're interested in working with Vietnamese, consider  \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{CS PhD programs in the US that have Vietnamese professors}.
-
 \begin{commentbox}[Hung:]
@@
 \end{commentbox}
+
+So what can you do about this?
+several things including asking your CS dept to put itself on CSRankings (it's the easiest way to get CS people to know about the school and its CS faculty)  and explaining about your school in your statement (ask your LOR writer to do that too).
+Of course, if you're interested in working with Vietnamese, consider  \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{CS PhD programs in the US that have Vietnamese professors}.
 

### 2023-10-19 demystify.tex

--- 2023-10-19 demystify.tex
@@
 So what can you do about this?
-several things including asking your CS dept to put itself on CSRankings (it's the easiest way to get CS people to know about the school) and explaining about your school in your statement (ask your LOR writer to do that too).
+several things including asking your CS dept to put itself on CSRankings (it's the easiest way to get CS people to know about the school and its CS faculty)  and explaining about your school in your statement (ask your LOR writer to do that too).
 Of course, if you're interested in working with Vietnamese, consider  \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{CS PhD programs in the US that have Vietnamese professors}.
-
-% \begin{commentbox}[Hung:]
-% I personally don’t care about which schools students attended. There are many other ways to evaluate an applicant.
-% \end{commentbox}
 

### 2023-10-24 demystify.tex
updates
Updates from Overleaf

@@
 
-If you don't know the professor (e.g., first email contact), then use \textbf{Prof.
-Lastname} or \textbf{Dr.
-Lastname}.
-I've seen many international students write \textbf{Prof.} or \textbf{Dr.} \textbf{Firstname Lastname}.
-Writing like that makes it like you copy and paste names, so no need to do so,  just Prof.
+If you're reaching out to a professor for the first time,  address them as either Prof.
+Lastname or Dr.
+Lastname.
+I've noticed that many international students tend to use the titles Prof.
+or Dr.
+FirstName LastName, but this can come across as if you're simply copying and pasting names.
+It's not necessary, so stick with Prof.
 or Dr.
 Lastname.
-Moreover, if you don't know the prof., do not use Mr.
-or Mrs., or Firstname---to me it seems a bit disrespectful.
-As you get to know that prof.
-better and depends on their preference, you may call them by their firstname.
-For example, my students and colleague call me Vu.
-Some students call me \emph{Dr.
-Vu}, which I find a bit amusing but am totally fine with it.
-
-Note that at some universities, especially those that do not offer PhD degrees, will use the formal \textbf{Dr.} title, i.e., Dr.
-Lastname.
-So just observe and follow how it is done at your place.
-Also note that some faculty might not have a Ph.D.
-and so you can use Prof.
-Lastname instead.
+
+Furthermore, it's important to avoid using Mr.
+or Mrs., or the professor's first name if you're not acquainted with them yet.
+This approach may be seen as somewhat disrespectful, especially in academic fields where most professors hold advanced degrees.
+As you become more familiar with a professor and depending on their preferences, you may transition to addressing them by their first name.
+For instance, my students and colleagues call me Vu.
+Some students also refer to me as Dr.
+Vu, which I find somewhat amusing but am perfectly fine with.
+
+Note that in some universities, e.g., those that do not offer PhD degrees, the formal title Dr.
+Lastname is commonly used.
+It's essential to observe and follow the conventions at your particular institution.
+Additionally, be aware that not all faculty members may hold a Ph.D., in which case using Prof.
+Lastname is a suitable alternative.
 

### 2023-10-24 demystify.tex
updates

@@
 Do you have any tips for Vietnamese students who want to connect with US professors and work as research assistants?
-\textbf{Vu}: see sec ~\ref{sec:contact} to contact a professor for research opportunities.
+\textbf{Vu}: \S\ref{sec:contact} shows how to contact a professor for research opportunities.
     \end{commentbox}
@@
       \begin{center}
-        \includegraphics[scale=0.5]{c5.png}
+        \includegraphics[scale=0.7]{c5.png}
       \end{center}
@@
 Lastname.
-I've noticed that many international students tend to use the titles Prof.
+Many international students use Prof.
 or Dr.
@@
 
-Furthermore, it's important to avoid using Mr.
+Furthermore, avoid using Mr.
 or Mrs., or the professor's first name if you're not acquainted with them yet.
-This approach may be seen as somewhat disrespectful, especially in academic fields where most professors hold advanced degrees.
 As you become more familiar with a professor and depending on their preferences, you may transition to addressing them by their first name.
-For instance, my students and colleagues call me Vu.
-Some students also refer to me as Dr.
-Vu, which I find somewhat amusing but am perfectly fine with.
-
-Note that in some universities, e.g., those that do not offer PhD degrees, the formal title Dr.
-Lastname is commonly used.
-It's essential to observe and follow the conventions at your particular institution.
+For example, my students and colleague call me Vu. Some students call me \emph{Dr. Vu}, which I find a bit amusing but am totally fine with it.
+
+Note that in some universities the formal title Dr.
+Lastname is preferred.
+You just need to observe and follow the conventions at your particular institution.
 Additionally, be aware that not all faculty members may hold a Ph.D., in which case using Prof.
@@
 
-\paragraph{Referring to professors you know} Here, because you are already familiar with these individuals, you can just informally use their firstname (or Dr./Prof., if you want to be formal) as mentioned above.
+\paragraph{Referring to professors you know} Here, because you are already familiar with these individuals, you can just informally use their firstname as mentioned above (or Dr./Prof., if you want to be formal).
 You can also include their institution if it makes it more precise.
@@
 Prof.
-X, and I also collaborate with NASA Distinguished Scientist Y.
+X, and I also collaborate with Distinguished Scientist Y.
 This is \emph{not necessary} and makes it look like you're trying to show off your connections.
@@
 PhD students often ask why their salary is so low compared to ludicrous grants their advisors get.
-They also wonder why their offer letters sometime mentioned that their benefits appear better more than what they actually receive (i.e., stipend).
+They also wonder why their offer letters sometime mentioned that their benefits appear better than what they actually receive (i.e., stipend).
 This section aims to shed some light to these questions.

### 2023-10-24 demystify.tex
updates

@@
 Moreover, they often give higher stipend (including summer) and open doors for job opportunities (e.g., internship).
-For example, a student with a Microsoft fellowship will likely get an internship at Microsoft.
 
@@
 
-%\subsubsection{Tenured, Tenure Track, Teaching,  Assistant Profs., Associate Profs}
-
-\subsubsection{Tenured or tenure-track faculty? Who do you choose?}
+\subsubsection{Tenured, Tenure Track, Teaching,  Assistant Profs., Associate Profs}
+There are several types of faculty in a university that has a CS PhD program.
+For example, you might hear about tenured,  tenure track, teaching faculty.
+You might also see assistant, associate, full, adjunct, emeritus, teaching, research professors.
+Here is a primer on these titles.
+Note that these are general and might not apply to all universities.
+
+\paragraph{Tenure-line and teaching faculty} Tenure line faculty, consisting of tenured and tenured-track profs., focus on research, which includes publishing papers, obtaining grants, mentoring Ph.D students.
+They often have very low teaching load (e.g., 1 per semester).
+\emph{Tenured} faculty are professors who have been promoted to have a permanent position (informally, very hard to fire them).
+\emph{Tenure-track} faculty are (more likely young) professors who are on the track to get tenure.
+Both tenure and tenured track faculty can serve as a formal advisor of Ph.D.
+students .
+This is very important because you need to have a tenure line faculty as your main PhD advisor.
+\S\ref{sec:tenure-vs-tenure-track} talks more about choosing between tenure and tenured-track professor as your advisor.
+
+In contrast, \emph{teaching} (or instructional faculty or professor of practice) mainly focus on teaching.
+They typically teach 3-4 classes per semester and \emph{do not} have research responsibilities (e.g., they do not have to worry much about publishing papers, obtaining grants).
+They can mentor Ph.D.
+students but they typically cannot serve as a formal advisor of Ph.D.
+students (i.e., they \emph{cannot graduate} PhD students).
+
+\paragraph{Rankings} Faculty have a ranking, e.g., assistant, associate, full, regardless if they are tenure line or teaching.
+Briefly, assistant means new faculty, associate means they have been promoted, and full means they are senior.
+A tenure track faculty starts with being an assistant professor and then gets tenure and promoted to associate (typically after 6 years) and then full (time varies, some become full within 3-4 years, some 10+ years, some remain associate).
+
+Adjunct means they are not full-time faculty, e.g., they might be working in industry and teach a class or two.
+Emeritus means they are retired but still have some affiliation with the university.
+Research faculty (or scientist) are typically non-tenure line faculty who focus on research.
+Due to their roles, adjunct, emeritus, and research faculty typically do not serve as a formal advisor of Ph.D.
+students.
+
+\subsubsection{Tenured or tenure-track faculty? Who do you choose?}\label{sec:tenure-vs-tenure-track}
 

### 2023-10-24 demystify.tex
updates

@@
     So I want to share some details about the PhD admission process and advice for those who are interested in applying for a \textbf{PhD in Computer Science in the US}.
-    While this document is primarily intended for students interested in CS, it may be relevant to students from various disciplines.
-    Furthermore, although many examples are specifics for George Mason University (GMU), the information should be generalizable to other R1\footnote{An \href{https://en.wikipedia.org/wiki/List_of_research_universities_in_the_United_States}{R1 institution} in the US is a research-intensive university with a high level of research activity across various disciplines.} institutions in the US  (though \emph{very} top schools could be very selective, e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU).
+    While this document is primarily intended for students interested in CS, it may be relevant to students from various disciplines and should be 
+    Furthermore, although many examples are specifics for schools that I and other contributors of this document know about, the information should be generalizable to other R1\footnote{An \href{https://en.wikipedia.org/wiki/List_of_research_universities_in_the_United_States}{R1 institution} in the US is a research-intensive university with a high level of research activity across various disciplines.} institutions in the US  (though \emph{very} top schools could be very selective, e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU).
 
@@
     \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-GMU}{why
-    you want to study at GMU}). Happy school hunting!
+    you want to study at here}). Happy school hunting!
 
@@
 
-This can be an issue for students in many top international universities where the competition is so high that very good students students can still have low rankings from these schools (and be overlooked by Admission committee).
+This can be an issue for students in many top international universities where the competition is so high that very good students students can still have low rankings from these schools (and be overlooked by the admission committee).
 So what to do with this? well, same as \hyperref[sec:your-school]{before}, e.g., put a note about this in your statement and ask your LoR writers to talk about it.
@@
 However, they might be useful for international students from programs we are not familiar with.
-
-\begin{commentbox}[Vu:]
-Personally I do not pay attention to GRE scores as I do not believe they are good indicator for research. However, if you have good GRE scores then you should include them in your application.
-\end{commentbox}
+If you have good GRE scores then you should include them in your application.
 
@@
 
-Choosing a school and an advisor is clearly one of the most important things in your mind when you apply to and especially when you get admitted multiple programs.
-This is further complicated due to cultural differences and unfamiliarity of  international students to the US system and faculty.
-This section aims to mitigate some confusions and help you make more informed decision.
+Choosing a school and an advisor is clearly one of the most important things in your mind when you apply to and especially when you get admitted.
+This is further complicated due to cultural differences and unfamiliarity of  international students to the education system in the .
+This section aims to mitigate some confusions and help you make informed decision.
 
@@
 However, while many universities offer Ph.D.
-degree in their CS programs, many do not.
+in their CS programs, many do not.
 These universities might offer just Bachelor degrees (e.g., BS) and no graduate studies (i.e., no MS or Ph.D degrees), or they just offer MS programs (but no PhD).
-
-Thus, if your goal is Ph.D.
-in CS, you will have to filter out schools that do not offer such a degree.
-This might sound obvious but many students (even those in the US) get confused due to the large number of universities.
 For example, Penn State in University Park has Ph.D.
@@
 
-While this can be confusing at first, a little research, e.g., searching for PhD in CS from the school website, will help you find out.
+Thus, if your goal is Ph.D.
+in CS, you have to aim only for schools offering such a degree.
+%This might sound obvious but many students (even those in the US) get confused due to the large number of universities.
+While this can be confusing due to the large number of universities in the US, a little research, e.g., searching for PhD in CS from the school website, will help you find out.
 All schools listed in \S\ref{sec:ranking} have PhD studies in CS, so you can start there.
@@
 International students not familiar with US universities often put them into \emph{two} bins:  (i) very top schools that they dream about, and (ii) everything else.
-In many cases, they use resources such as rankings from US News, which are \emph{highly} questionable\footnote{\url{https://cra.org/cra-statement-us-news-world-report-rankings-computer-science-universities/}}.
+In many cases, they use resources such as rankings from US News, which are not very transparent and \emph{highly} questionable\footnote{\url{https://cra.org/cra-statement-us-news-world-report-rankings-computer-science-universities/}}.
 Sometimes these students evaluate CS programs using the reputations non-CS programs such as medical, math, or physics.
@@
 
-\begin{commentbox}[Dat] Most Vietnamese students, including those from top schools, \textbf{do not know} about CSRankings.
+\begin{commentbox}[Dat:] Most Vietnamese students, including those from top schools, \textbf{do not know} about CSRankings.
 May be applicants who worked at top research places such as VinAI would know about it.
@@
 
-\begin{commentbox}[Xiaokuan]
+\begin{commentbox}[Xiaokuan:]
 Chinese students often only look at USNews rankings when selecting their Ph.D. universities (I did that, too, when I was applying for Ph.D. positions).
@@
 
+\begin{center}
+    \includegraphics[scale=0.2]{alpinist-climbing-peak-mountain-comic-hand-drawn-vector-illustration.jpg}
+  \end{center}
+
  Show something that makes you \textbf{stand out}, e.g., do you have a degree or background in \emph{dance} or \emph{music} and want to integrate them with CS?
@@
 
-\paragraph{Referring to professors you know} Here, because you are already familiar with these individuals, you can just informally use their firstname as mentioned above (or Dr./Prof., if you want to be formal).
+\paragraph{Referring to professors you know} Because you are already familiar with these individuals, you can just informally use their firstname as mentioned above (or Dr./Prof., if you want to be formal).
 You can also include their institution if it makes it more precise.
@@
 Do not include ranking (e.g., Assistant, Associate, Scientist, ...) when referring to someone.
-I see many international students include a lengthy title of people they know, e.g., my advisor is Asst.
+I see many international students include a lengthy title of people they know, e.g., I am advised by Asst.
 Prof.

### 2023-10-26 demystify.tex
updates

@@
 
-While you might not have control over LORs \hyperref[sec:your-school]{where your go to school}, you do over your
-statement! So write it well because we take it seriously.
+While you might not have control over LORs or \hyperref[sec:your-school]{where your go to school}, you do over your
+statement! So write it well because it could make a big difference.
 In your statement, you have the opportunity to make  make your application stand out and unique, which can make you \emph{fit} the CS program you're applying to, even if you don't have very strong research experience.
@@
 Also, use the statement to talk about stuff that admission committee members might not know about, e.g., your Github project with 1K+ stars or  your regular contributions to well-known open-source projects.
+Finally, talk about what makes you different than the other 1000 applicants that we are reviewing (e.g., see \S\ref{sec:improve-your-chance} for increasing your admission chance).
 

### 2023-10-26 demystify.tex
Updates from Overleaf

--- 2023-10-24 demystify.tex
@@
 
-This section lists various issues that international students should pay attention to.
+This section lists various cultural issues that international students should pay attention to.
 
@@
 They have very sophisticated tools to detect illegal software and will sue your university/department.
-Imagine your department or advisor being sued for half a million dollars, and it is \emph{you} who caused it.
+Imagine your department or advisor being sued for a large sum of money, and it is \emph{you} who caused it.
 If you need software, ask your advisor or the department.

### 2023-10-26 demystify.tex
Updates from Overleaf

@@
 
-After you submit your PhD application (usually in December), it will be first screened
+After you submit your PhD application (usually December 15), it will be first screened
 for general requirements, e.g., did you submit your transcripts and standard scores? did your reference writers submit their letters?
@@
 There are several types of faculty in a university that has a CS PhD program.
-For example, you might hear about tenured,  tenure track, teaching faculty.
+You might hear about tenured,  tenure track, teaching faculty.
 You might also see assistant, associate, full, adjunct, emeritus, teaching, research professors.
 Here is a primer on these titles.
-Note that these are general and might not apply to all universities.
 
@@
 
+% \section{Additional Resources}
+% \begin{itemize}
+% \item \bibentry{nlpphd}
+% \end{itemize}
+
 \section{History and Acknowledgement}
@@
 \paragraph{History} This document was conceived during a lunch with Craig Yu at GMU.
-We talked on about why GMU were not able to attract good Vietnamese and other international students, despite having a much stronger CS program than many schools that these students want to go to (part of the reason is described in \S\ref{sec:ranking}) and wished there were a way for international students to know about the US PhD programs (as well as for US faculty to understand more about international students and therefore have better chance of recruiting and working with them).
+We talked on about why GMU were not able to attract good Vietnamese and other international students, despite having a much stronger CS program than many schools that these students want to go to (part of the reason is described in \S\ref{sec:ranking}).
+We wished there were a way for international students to know about the US PhD programs (as well as for US faculty to understand more about international students and therefore have better chance of recruiting and working with them).
 I was also  a member of the large VietPhD group on Facebook and saw many questions from students about PhD programs.

### 2023-11-01 demystify.tex
updates
Updates from Overleaf

@@
 Choosing a school and an advisor is clearly one of the most important things in your mind when you apply to and especially when you get admitted.
-This is further complicated due to cultural differences and unfamiliarity of  international students to the education system in the .
+This is further complicated due to cultural differences and unfamiliarity of  international students to the US higher education system.
 This section aims to mitigate some confusions and help you make informed decision.
@@
 
-This section aims to explain different types of universities in the US in the context of obtaining a CS Ph.D degree.
-
-\subsection{Schools that offer Ph.D. in CS}
+We will first discuss about universities in the US that offer PhD in CS.
+
+\subsubsection{Schools that offer PhD in CS}
 
 Most universities the US will have CS programs. 
-However, while many universities offer Ph.D.
-in their CS programs, many do not.
+However, while many universities offer PhD in their CS programs, many do not.
 These universities might offer just Bachelor degrees (e.g., BS) and no graduate studies (i.e., no MS or Ph.D degrees), or they just offer MS programs (but no PhD).
@@
 
-\subsection{Selecting and Ranking Schools for CS}
+\subsubsection{Selecting and Ranking Schools for CS}
    \begin{center}
@@
 
+Now we discuss about finding advisors.
+
 \subsubsection{Tenured, Tenure Track, Teaching,  Assistant Profs., Associate Profs}
@@
 You might hear about tenured,  tenure track, teaching faculty.
-You might also see assistant, associate, full, adjunct, emeritus, teaching, research professors.
-Here is a primer on these titles.
+You might also things such as assistant, associate, full, adjunct, emeritus, teaching, research professors.
+Here is a primer on these terminologies.
 
@@
 \paragraph{Rankings} Faculty have a ranking, e.g., assistant, associate, full, regardless if they are tenure line or teaching.
-Briefly, assistant means new faculty, associate means they have been promoted, and full means they are senior.
+Assistant means new faculty, associate means they have been promoted, and full means they are senior.
 A tenure track faculty starts with being an assistant professor and then gets tenure and promoted to associate (typically after 6 years) and then full (time varies, some become full within 3-4 years, some 10+ years, some remain associate).
@@
       \end{center}
-
-%The tenure system in universities in the US (and other countries) can be confusing.
-In most cases, you probably don't need to know much about it.
-But it is helpful to know a few things so you know how to find PhD advisor.
-First, faculty of a CS (and other) department can be divided into 2 types:  (i) tenured and tenure-track and (ii) non-tenured or teaching faculty.
-Tenured/Tenure -track faculty focus on research and related tasks such as mentoring and getting grants to support students, etc.
-Their teaching load is very light, typically they just teach 1 or at most 2 classes a semester.
-Non-tenured/teaching faculty focuses on teaching, and often teach 3-4 classes per semester.
-They can mentor Ph.D.
-students but they typically cannot serve as a formal advisor of Ph.D.
-students (i.e., they cannot graduate PhD students).
-Note that there is also affiliated faculty who is tenured/tenure-track profs.
-from another department.
-They can serve in Ph.D.
-dissertation committee but they likely cannot serve as the main advisor of CS students.
 
@@
 
-\subsection{Should you contact a professor? What to do to get a desired reply?}\label{sec:contact}
+\subsubsection{Should you contact a professor? What to do to get a desired reply?}\label{sec:contact}
 
@@
 
+\subsubsection{PhD in other related fields: CE, IST, Cybersecurity}
+
+In many cases you do not have to do a PhD in CS to work in your area of interested.
+For example, in addition to a traditional CS department, GMU has IST and Cybersecurity departments, which have faculty  with PhD in CS and work on CS topics (e.g., AI, Security, Robotics).
+So it is totally possible that you still get to do CS research and publish in CS-focused venues even if you're not in a traditional CS program.
+It is  common to see faculty with CS PhD in a non-CS department as well as faculty with non-CS PhD in CS department.
+
+However, if your intention is a PhD in CS, then you likely need to be in the CS dept \emph{and} advised by a CS faculty.
+In fact, if a faculty is not in CS, it is unlikely that they can be the \emph{main} advisor of a CS PhD student---they may \emph{co-advise} or be in the PhD dissertation committee, but your main advisor will need to be a tenured or tenure-track faculty in CS.
+If in doubt, you should check with the CS department for their requirements.
+For this specific reason,  CSRankings includes only tenured or tenured track faculty who can advise CS PhD students.
+I also have compiled a \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{list} of Vietnamese faculty who can advise PhD students.
+
 \section{Miscs and FAQs}
@@
     The PhDComics figure on top shows the ``ambition'' level of a PhD student over their years of study (they miss the 6-7th year where the ambition is \emph{``Just let me graduate''}).
-
-\subsection{PhD in other related fields: CE, IST, Cybersecurity}
-
-In many cases you do not have to do a PhD in CS to work in your area of interested.
-For example, in addition to a traditional CS department, GMU has IST and Cybersecurity departments, which have faculty  with PhD in CS and work on CS topics (e.g., AI, Security, Robotics).
-So it is totally possible that you still get to do CS research and publish in CS-focused venues even if you're not in a traditional CS program.
-It is  common to see faculty with CS PhD in a non-CS department as well as faculty with non-CS PhD in CS department.
-
-However, if your intention is a PhD in CS, then you likely need to be in the CS dept \emph{and} advised by a CS faculty.
-In fact, if a faculty is not in CS, it is unlikely that they can be the \emph{main} advisor of a CS PhD student---they may \emph{co-advise} or be in the PhD dissertation committee, but your main advisor will need to be a tenured or tenure-track faculty in CS.
-If in doubt, you should check with the CS department for their requirements.
-For this specific reason,  CSRankings includes only tenured or tenured track faculty who can advise CS PhD students.
-I also have compiled a \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{list} of Vietnamese faculty who can advise PhD students.
 
@@
 or Mrs., or the professor's first name if you're not acquainted with them yet.
-As you become more familiar with a professor and depending on their preferences, you may transition to addressing them by their first name.
-For example, my students and colleague call me Vu. Some students call me \emph{Dr. Vu}, which I find a bit amusing but am totally fine with it.
+As you become more familiar with your prof and depending on their preferences, you may transition to addressing them by their first name.
+For example, I prefer that my students and colleague call me Vu. Some students call me \emph{Dr. Vu}, which I find a bit amusing but am totally fine with it.
 
@@
 You can also include their institution if it makes it more precise.
-For example, I did my postdoc with Jeff Foster at Univ.
-of Maryland.
+For example, I can say:  \emph{"I did my postdoc with Jeff Foster at Univ.
+of Maryland"}.
+
 Do not include ranking (e.g., Assistant, Associate, Scientist, ...) when referring to someone.
-I see many international students include a lengthy title of people they know, e.g., I am advised by Asst.
+I see many international students include a lengthy title of people they know, e.g., \emph{I am advised by Asst.
 Prof.
-X, and I also collaborate with Distinguished Scientist Y.
+X, and I also collaborate with Distinguished Scientist Y}.
 This is \emph{not necessary} and makes it look like you're trying to show off your connections.
-In short, these nuances represent some cultural differences that you may encounter and will gradually adapt to.
+These nuances represent some cultural differences that you may encounter and will gradually adapt to.
 More on cultural differences in \S\ref{sec:cultural}.

### 2023-11-01 demystify.tex
updates

--- 2023-10-26 demystify.tex
@@
 \begin{document}
+\setcounter{section}{-1};
+
 \maketitle
@@
 \tableofcontents
+
+\section{Summary}\label{sec:summary}
+Here we summarize the main points of the document. This allows you to quickly skim through the document and decide which parts you want to read in detail.
+\begin{enumerate}
+\item Should you apply?
+\begin{itemize}
+    \item Yes, definitely.  CS PhD study in the US is fully funded and admission into good universities is not any harder than non-US schools.
+\end{itemize}
+\item How is your application evaluated?
+\begin{itemize}
+    \item Applications are evaluated by a committee and each application is reviewed by typically 3 faculty.
+    \item Individually faculty cannot directly admit a student (so do not email and ask if you have a chance), but they can \emph{advocate} for a student (so do email and introduce yourself if you share the same research interest).
+    \item The committee will look at various factors, but the most important ones are research ability, e.g., publications, and recommendation letters.
+    \item LORs are important, but only if they are personalized and talk about your research ability.
+    \item GRE \emph{is not} required by most schools.
+    \item Grades are important, but only if your school is well-known. Otherwise, they are not useful.
+\end{itemize}
+\end{enumerate}
+
+\begin{itemize}
+    \item 
+\end{itemize}
 
@@
 
-\begin{wrapfigure}{r}{0.47\linewidth}
-\vspace{-0.2in}
 \begin{center}
-        \includegraphics[width=1\linewidth]{c6.png}
+\includegraphics[width=0.6\linewidth]{c6.png}
 \end{center}
-      \vspace{-0.1in}
-\end{wrapfigure}
+
 Many students have letters written by themselves and signed by their professors. These have little
@@
 
-Many students get letters from supervisors from company where they did internship or are
-working at. It is OK as long as it is a research-based personalized
-letter (once again, we are talking about PhD applications, not MS).
+Many students get letters from supervisors from companies where they did internships or are working at.
+It is OK as long as it is a research-based personalized letter (once again, we are talking about PhD applications, not MS).
 

### 2023-11-01 demystify.tex
updates

@@
 \begin{enumerate}
-\item Should you apply?
+\item Should you apply to a PhD in CS in the US?
 \begin{itemize}
-    \item Yes, definitely.  CS PhD study in the US is fully funded and admission into good universities is not any harder than non-US schools.
+    \item \emph{Yes, definitely}.  CS PhD study in the US is fully funded and admission into good universities is not any harder than non-US schools (\S\ref{sec:should}).
 \end{itemize}
@@
 \begin{itemize}
-    \item Applications are evaluated by a committee and each application is reviewed by typically 3 faculty.
-    \item Individually faculty cannot directly admit a student (so do not email and ask if you have a chance), but they can \emph{advocate} for a student (so do email and introduce yourself if you share the same research interest).
+    \item Applications are evaluated by the \emph{PhD Admission} committee and each application is reviewed by typically 3 faculty (\S\ref{sec:evalapps}).
+    \item Individually faculty \emph{cannot directly admit} a student---so do not email and ask if you have a chance.
+However, faculty can \emph{advocate} for a student and therefore increase their admission chance---so contact and introduce yourself (\S\ref{sec:contact}).
     \item The committee will look at various factors, but the most important ones are research ability, e.g., publications, and recommendation letters.
-    \item LORs are important, but only if they are personalized and talk about your research ability.
-    \item GRE \emph{is not} required by most schools.
-    \item Grades are important, but only if your school is well-known. Otherwise, they are not useful.
+    \item LORs are important, but only if they are personalized and talk about your research ability (\S\ref{sec:lor}).
+    \item GRE \emph{is not} required by most schools (\S\ref{sec:grades}).
+    \item Grades are important, but only if your school is well-known (\S\ref{sec:grades}).
+    \item Personal statement is very important. Write it in such a way that makes you \emph{stand out} (\S\ref{sec:research-statement} and \S\ref{sec:improve-your-chance})
 \end{itemize}
+\item What to do after getting admitted?
+\begin{itemize}
+    \item Attend \emph{Open House} to learn more about the place and \emph{interview} profs---they would be much more willing to talk to you now (\S\ref{sec:accepted}).
+\end{itemize}
+\item Funding
 \end{enumerate}
 
-\begin{itemize}
-    \item 
-\end{itemize}
-
-\section{Should You Apply?}
+\section{Should You Apply to do a PhD in CS in the US?}\label{sec:should}
+
 \epigraph{\vspace{-0.2in} Don't make fun of graduate students. They just made a terrible life choice.}{\textsc{The Simpsons}}
@@
 
-\subsection{Is an MS degree required for admission to PhD in CS?}
+\subsection{Is an MS degree required for admission to PhD in CS?}\label{sec:msrequirement}
 No. In fact, student with BS can get MS degree ``along the way'' to PhD.  However, MS can help if it gives research experience or is from a more well-known school than your undergrad institution.
@@
 
-\subsection{How long does it take to complete the CS PhD program?}
+\subsection{How long does it take to complete the CS PhD program?}\label{sec:time}
 
@@
     In many cases you might take a summer or two off to do internship to get additional research opportunities.
-    The PhDComics figure on top shows the ``ambition'' level of a PhD student over their years of study (they miss the 6-7th year where the ambition is \emph{``Just let me graduate''}).
+    The PhDComics figure on top shows the ``ambition'' level of a PhD student over their years of study (they miss the 6--7th year where the ambition is \emph{``Just let me graduate''}).
+
+\subsection{Difference between PhD in the US and other countries}
+Here I try to list the main differences between CS PhD in the US and other countries.  This is based on my experience in the US and what I heard from others who did PhD in other countries.
+
+\emph{MS Degree requirement}:  as mentioned in \S\ref{sec:msrequirement} and \ref{sec:time}, US PhD programs do not require MS degree.
+In contrast, many other countries do require MS degree before joining a PhD program.
+This means that US PhD programs are longer (5--7 years, 2 of which are coursework) than other countries (3--4 years, no coursework).
+
+\emph{Project proposal}: in many countries, you have to choose a project and advisor \emph{during} the application process (e.g., you write a proposal for potential advisor).
+In the US, you start your PhD without a project and advisor and find them later.
+This is also why US PhD programs are longer.
+
+\emph{Course work}: in the US you will spend the first couple of years taking classes and explore potential advisor and research topics.
+In other countries, you often do not do course work and start your research right away, e.g., you immediately work on the research project you propose with the advisor you chose.
+In the US you also have to pass a series of "exams", e.g., qualifying exam, comprehensive exam, thesis proposal defense.
+In other countries, you do not have to do these exams or only do few of them.
+For example, ABD (all but dissertation) is used in the US to refer to a PhD candidate who have finished all course work and exams and only need to write and defend their dissertation.
+
+\emph{Funding}:  In Europe, stipend (salary) comes from the university or from Gov research organizations and often have a fixed duration (e.g., 3 years).
+In the US, stipend (e.g., RA) comes directly from your advisor (and no fixed duration).
+There are fewer TA opportunities in the European universities compared to the US.
+
+\emph{Faculty Position after PhD}: In other countries, PhD graduates typically apply for research positions at research labs, i.e., postdocs, and then consider faculty position.
+In the US, PhD graduates interested in academia can apply directly for faculty position.
+This is why US PhD programs are longer and more focused on research.
+
+\emph{Work-life balance}: PhD students are often overworked in the US (seriously!) compared to other countries, e.g., in Europe.
+This is partly due to the longer PhD program and that US PhD students are often paid through TA, which requires them to work for the dept.
+In contrast, PhD students in other countries are often paid through fellowships, which do not require them to do TA work.
 
@@
 This section aims to shed some light to these questions.
+
+      \begin{center}
+        \includegraphics[scale=0.5]{c6.png}
+      \end{center}
 
@@
 %There are many advice on surviving PhD that you can follow. But here I just list a few that works for me and what I advice my students to do.\tvn{TODO}
+
+\subsection{Will I be miserable during my a PhD?}
+Many posts from VietPhD, a Facebook group for Vietnamese PhD students, are about how PhD students are mistreated, stressed, and miserable.
+Sure there are issues with PhD and academic in general, but then there are also issues any jobs, family, and life.
+Certainly there are horrible professors, bad relationships with co-authors and officemates, feeling left out or discriminated (e.g., because you're an international student), and it is good to be aware of those.
+However, in reality there are many good mentors, supportive labmates and department, happy students, etc.
+Don't let social media make you feel pessimistic and destroy your quest to advance knowledge.
 

### 2023-11-02 demystify.tex
add stuff for nonCS/nonSTEM

@@
 \begin{document}
+
+\maketitle
 \setcounter{section}{-1};
-
-\maketitle
 
@@
 
-Yes, as long as you can demonstrate you are ready for CS PhD research through research experiences, LoRs, statements, etc as mentioned.
+Yes, \emph{as long as} you can demonstrate you are ready for CS PhD research through research experiences, LoRs, statements, etc as mentioned.
 You might be even able to leverage this to make your profile stand out as mentioned in \S\ref{sec:improve-your-chance}.
+
+One main concern for non-CS or non-STEM students is if you have the sufficient technical background, typically obtained through core CS courses.
+So you will want to show that you have such knowledge through your coursework, projects, or research.
+For example, if you have taken a course on Algorithms, even online ones like Coursera, you can talk about it in your statement.
+If you have done a project or have a professional certification (e.g., A+) through work that requires knowledge of OS, you can talk about it in your statement.
+If you have done research that requires knowledge of Discrete Maths, you can talk about it in your statement.
+You can also ask your LoR writers to talk about your technical background.
+In summary, in your application, convince us that you have the technical background to do CS PhD research. 
+Note that many online courses teaching about AI does not really help much as these are not core CS knowledge.
+
+In short, you \emph{do not need} to formally taking CS courses, you just need to show that you have these essential knowledge through ways as mentioned.
+Many universities are well-aware that incoming graduate students might not have all the technical background, so they often have a "bridge" courses to help students catch up.
+For example, GMU has 4 bridge courses (Data Structure and Algorithm, Computer Systems, Discrete Maths, and Programming Foundation) that incoming students can take to catch up on their CS knowledge.
+
+\begin{commentbox}[Vu:]
+    I would strongly advocate to admit a student with a non-STEM background (e.g., business, English major) but can show that they have a strong drive for CS by studying core CS knowledge through various channels (e.g., self-study through online courses, projects, etc).
+I have seen many students with non-CS background who are very successful in CS PhD.
+In fact, I have seen many students with CS background who are not successful in CS PhD.
+So it is not about your background, it is about your drive and passion for CS research.
+\end{commentbox}
 
@@
 \subsection{Difference between PhD in the US and other countries}
-Here I try to list the main differences between CS PhD in the US and other countries.  This is based on my experience in the US and what I heard from others who did PhD in other countries.
+Here we summarize the main differences between CS PhD in the US and other countries.  This is based on my experience and what I heard from others who did PhD in other countries.
 
@@
 
-\emph{Project proposal}: in many countries, you have to choose a project and advisor \emph{during} the application process (e.g., you write a proposal for potential advisor).
-In the US, you start your PhD without a project and advisor and find them later.
+\emph{Project proposal}: in many countries, you have to choose a project and advisor \emph{during} the application process (e.g., you write a proposal to a potential advisor).
+In the US, you often start your PhD without a project and advisor and find them later.
 This is also why US PhD programs are longer.
@@
 
-\emph{Funding}:  In Europe, stipend (salary) comes from the university or from Gov research organizations and often have a fixed duration (e.g., 3 years).
-In the US, stipend (e.g., RA) comes directly from your advisor (and no fixed duration).
-There are fewer TA opportunities in the European universities compared to the US.
-
-\emph{Faculty Position after PhD}: In other countries, PhD graduates typically apply for research positions at research labs, i.e., postdocs, and then consider faculty position.
+\emph{Funding}:  In Europe, stipend (salary) comes from the university or from gov't research organizations and often have a fixed duration (e.g., 3 years).
+In the US, stipend (e.g., RA) comes directly from your advisor (no fixed duration).
+There are also fewer TA opportunities in the European universities compared to the US.
+
+\emph{Faculty position after PhD}: In other countries, PhD graduates typically apply for research positions at research labs, i.e., postdocs, and then consider faculty position.
 In the US, PhD graduates interested in academia can apply directly for faculty position.
@@
       \begin{center}
-        \includegraphics[scale=0.5]{c6.png}
+        \includegraphics[scale=0.3]{c6.png}
       \end{center}
@@
 
-Tab.~\ref{tab:cost} shows the budget breakdown for a GRA per year (this level of details is what faculty actually uses when applying for funding).
+Tab.~\ref{tab:cost} shows the budget breakdown for a GRA per year (this level of details is what faculty actually uses when applying for grants).
 These numbers are based on my experience at public universities in the US.
@@
 I will also use GMU tuition rate of about \$15,000/year for full-time study (which is quite cheap compared to private universities, e.g., MIT charges around \$50K) and a 58.9\% rate on \emph{indirect cost}, which is what GMU charges for overhead or administrative costs (yes, after all, universities are businesses!).
-Finally, I assume the student makes two conference trips per year, one domestic and one international (so conf.
+Finally, I assume the student take two conference trips per year, one domestic and one international (conf.
 registration, airline tickets, taxi, meals, etc are all included).
-At the end, the total budget comes out to be \$69K.
-\emph{The summary is that while you're paid X, your advisor probably pays 2X for you}.
+
+At the end, the total budget comes out to be \$69K/year to support a PhD student.
+\emph{The summary is that over your 5-6 year of your PhD, you cost about \$350K--400K, and while you're paid X, your advisor probably pays 2X for you}.
 
@@
 
-\subsection{Will I be miserable during my a PhD?}
+\subsection{Will I be miserable during my a PhD?}\label{sec:miserable}
 Many posts from VietPhD, a Facebook group for Vietnamese PhD students, are about how PhD students are mistreated, stressed, and miserable.
-Sure there are issues with PhD and academic in general, but then there are also issues any jobs, family, and life.
-Certainly there are horrible professors, bad relationships with co-authors and officemates, feeling left out or discriminated (e.g., because you're an international student), and it is good to be aware of those.
+Indeed there are issues with PhD and academic in general, but then there are also issues with jobs, family, and life.
+Certainly there are bad relationships with professors, co-authors, and officemates, and toxic working environment, feeling left out or discriminated (e.g., because you're an international student), and it is good to be aware of those.
 However, in reality there are many good mentors, supportive labmates and department, happy students, etc.
-Don't let social media make you feel pessimistic and destroy your quest to advance knowledge.
+So don't let social media make you feel pessimistic and destroy your quest to advance knowledge.
 
@@
 
-I started writing this document in May 2023 and have been updating it since then (mostly around deadline time when I tend to procrastinate!).
+I started writing this document in May 2023 and have been updating it since then (mostly around deadline time when I tend to procrastinate, i.e., \emph{productive procrastination}!).
 I have put the source code of this document on \href{https://github.com/nguyenthanhvuh/phd-cs-us}{GitHub} so that anyone can contribute to it.

### 2023-11-02 demystify.tex
updates

@@
 %\date{\today}
-\author{\small \href{https://nguyenthanhvuh.github.io}{ThanhVu (Vu) Nguyen} (\href{mailto:tvn@gmu.edu}{tvn@gmu.edu})}
+\author{\href{https://nguyenthanhvuh.github.io}{ThanhVu (Vu) Nguyen} (\href{mailto:tvn@gmu.edu}{tvn@gmu.edu})\\{\small George Mason University, Dept. of Computer Science}}
 
@@
 
-    This document is available on \href{https://github.com/nguyenthanhvuh/phd-cs-us}{Github}.
+    The latest version of document is at \url{https://nguyenthanhvuh.github.io/phd-cs-us/demystify.pdf} and its \LaTeX{} source is on \href{https://github.com/nguyenthanhvuh/phd-cs-us}{Github}.
 If you have questions or comments, feel free to create a \href{https://github.com/nguyenthanhvuh/phd-cs-us/issues}{GitHub issue} for discussion.
@@
 harder than at schools in other countries.
-If you believe you have a chance in other countries, e.g., South Korea, Singapore, Germany, UK, Japan and Australia, then you will surely have a chance in the US as well.
+If you believe you have a chance in other countries, e.g., South Korea, Singapore, Germany, UK, Japan and Australia, then you will surely have a chance in the US as well\footnote{\S\ref{sec:non-us-differences} compares CS PhD study in the US to other countries.}.
 
@@
 
-\paragraph{Rankings} Faculty have a ranking, e.g., assistant, associate, full, regardless if they are tenure line or teaching.
+\paragraph{Rankings} Faculty have ranking, e.g., assistant, associate, full, regardless if they are tenure line or teaching.
 Assistant means new faculty, associate means they have been promoted, and full means they are senior.
@@
 
-\subsection{Difference between PhD in the US and other countries}
+\subsection{Difference between PhD in the US and other countries}\label{sec:non-us-differences}
 Here we summarize the main differences between CS PhD in the US and other countries.  This is based on my experience and what I heard from others who did PhD in other countries.

### 2023-11-02 demystify.tex
updates

@@
 However, faculty can \emph{advocate} for a student and therefore increase their admission chance---so contact and introduce yourself (\S\ref{sec:contact}).
+\end{itemize}
+\item Application Materials
+\begin{itemize}
     \item The committee will look at various factors, but the most important ones are research ability, e.g., publications, and recommendation letters.
@@
     \item Personal statement is very important. Write it in such a way that makes you \emph{stand out} (\S\ref{sec:research-statement} and \S\ref{sec:improve-your-chance})
+    \item Getting an interview is typically a \emph{good sign} (no one interviews weak candidates).
 \end{itemize}
@@
 \item Funding
+\begin{itemize}
+    \item TA and RA are two main sources of funding.
+TA (teaching assistantship) is provided by the department to help profs.
+with classes (e.g., grading).
+RA (research assistantship) is provided by profs.
+to help them with their research (\S\ref{sec:funding}).
+\end{itemize}
+\item Choosing School and Professors
+\begin{itemize}
+    \item Many schools do not offer PhD studies in CS and many CS professors \emph{cannot} be advisor, i.e., cannot graduate PhD students  (\S\ref{sec:schoolsandprofs}).
+    \item Contacting a prof. for research opportunities is recommended, but do it \emph{properly} (\S\ref{sec:contact}).
+\end{itemize}
+\item Miscs and FAQS
+\begin{itemize}
+    \item Increasing your admission chance by being unique and standing out (\S\ref{sec:improve-your-chance}).
+    \item You can successfully apply to CS PhD even if you have non-STEM background (\S\ref{sec:non-stem}).
+    \item Compared to other countries, CS PhD in the US does not require an MS degree but has longer PhD study time (\S\ref{sec:non-us-differences}).
+    \item Your PhD costs about \$400K in total, but you do not pay for it (\S\ref{sec:ra-cost}).
+    \item Despite many miserable stories on social media, many PhD students have good mentors, supportive labmates, enjoytable working environment ... and are happy (\S\ref{sec:happy}).
+\end{itemize}
 \end{enumerate}
@@
 Now you're at a different game because the schools that have admitted you will try to get you to accept them!
+Several important factors to consider are the reputation of schools and professors (\S\ref{sec:schoolsandprofs}), and of course, funding (\S\ref{sec:funding}).
 You will have to make your decision by around April 15.
 
-Most schools will have \textbf{Open House} events, which are a great resource to learn about the school, department, faculty, research, living, etc.
+\paragraph{Open House} Most schools will have \textbf{Open House} events, which are a great resource to learn about the school, department, faculty, research, living, etc.
 During the Open House, you get a chance to talk to individual faculty and current students.
@@
 Ask about anything you want to determine that they deserve \emph{you}.
-
 In summary, if possible, attend the Open House.  Even if you can't come in person, you still can attend virtually and meet with individual faculty.
@@
 
-\subsection{Can I apply to CS PhD if my undergrad was not in CS or related areas?}
+\subsection{Can I apply to CS PhD if my undergrad was not in CS or related areas?}\label{sec:non-stem}
 
@@
 
-\subsection{Will I be miserable during my a PhD?}\label{sec:miserable}
+\subsection{Will I be miserable during my a PhD?}\label{sec:happy}
 Many posts from VietPhD, a Facebook group for Vietnamese PhD students, are about how PhD students are mistreated, stressed, and miserable.

### 2023-11-02 demystify.tex
updates

@@
 
-By around March you should hear back from most PhD programs you applied to . If you haven't heard back, reach out via email and ask about the status of your application.
+By around March you should hear back from most PhD programs you applied to. If you haven't heard back, reach out via email and ask about the status of your application.
 If you receive offers, congratulations!
 Now you're at a different game because the schools that have admitted you will try to get you to accept them!
-Several important factors to consider are the reputation of schools and professors (\S\ref{sec:schoolsandprofs}), and of course, funding (\S\ref{sec:funding}).
-You will have to make your decision by around April 15.
+Important factors to consider are the reputation of schools and professors (\S\ref{sec:schoolsandprofs}), and of course, funding availability (\S\ref{sec:funding}).
+You will have to make your decision by around \emph{April 15}.
 
 \paragraph{Open House} Most schools will have \textbf{Open House} events, which are a great resource to learn about the school, department, faculty, research, living, etc.
+Even if you can't come in person, you still can attend virtually and meet with individual faculty.
 During the Open House, you get a chance to talk to individual faculty and current students.
@@
 Ask about anything you want to determine that they deserve \emph{you}.
-In summary, if possible, attend the Open House.  Even if you can't come in person, you still can attend virtually and meet with individual faculty.
-
-\begin{commentbox}[Vu:]
-GMU has \emph{Virtual} Open House (VOH), e.g., \url{https://cs-gmu.github.io/cs-phd-voh-s23/}, which I've co-organized in the last two years.
+
+\begin{commentbox}[Vu:]
+GMU has \emph{Virtual} Open House (VOH), e.g., \url{https://cs-gmu.github.io/cs-phd-voh-s23/}, which I've co-organized in recent years.
 We invite all admitted PhD students to the VOH through Zoom to learn about the CS program, the department, GMU, and the DC area in general.
@@
 \emph{Project proposal}: in many countries, you have to choose a project and advisor \emph{during} the application process (e.g., you write a proposal to a potential advisor).
-In the US, you often start your PhD without a project and advisor and find them later.
+In the US, you often start your PhD without an advisor or project and find them later.
 This is also why US PhD programs are longer.
@@
 \emph{Course work}: in the US you will spend the first couple of years taking classes and explore potential advisor and research topics.
-In other countries, you often do not do course work and start your research right away, e.g., you immediately work on the research project you propose with the advisor you chose.
-In the US you also have to pass a series of "exams", e.g., qualifying exam, comprehensive exam, thesis proposal defense.
+In other countries, you often do not do course work and start your research right away, e.g., you immediately work on the research project you proposed with the advisor you chose.
+Also, in the US you also have to pass a series of "exams", e.g., qualifying exam, comprehensive exam, thesis proposal defense\footnote{The word "ABD" (all but dissertation) is used in the US to refer to a PhD candidate who have finished all course work and exams and only need to write and defend their dissertation.}.
 In other countries, you do not have to do these exams or only do few of them.
-For example, ABD (all but dissertation) is used in the US to refer to a PhD candidate who have finished all course work and exams and only need to write and defend their dissertation.
 

### 2023-11-03 demystify.tex
updates

@@
 \maketitle
-\setcounter{section}{-1};
+\setcounter{section}{-1}
 
@@
 \begin{itemize}
-    \item The committee will look at various factors, but the most important ones are research ability, e.g., publications, and recommendation letters.
+    \item The committee will look at various factors, but the most important ones are research ability, e.g., publications, personal statement, and recommendation letters.
     \item LORs are important, but only if they are personalized and talk about your research ability (\S\ref{sec:lor}).
@@
     \item Your PhD costs about \$400K in total, but you do not pay for it (\S\ref{sec:ra-cost}).
-    \item Despite many miserable stories on social media, many PhD students have good mentors, supportive labmates, enjoytable working environment ... and are happy (\S\ref{sec:happy}).
+    \item Despite some miserable stories on social media, many PhD students have good mentors, supportive labmates, healthy working environment ... and are happy (\S\ref{sec:happy}).
 \end{itemize}
@@
 can be challenging for international students, who might not have much interactions with such experts.
-So it is acceptable to have a letter from professors that know you well enough to talk about \emph{your specific research experience and capabilities}.
+So it is fine to have a letter from people (e.g., profs, researchers, mentors) that know you well enough to talk about \emph{your specific research experience and capabilities}.
+This person can be a postdoc or even graduate student who has mentored you.
 
@@
 values. We can recognize them and will consider them weakness.
-Similarly, many professors write generic letters for students, e.g., the students didn't do any
-research or make any impression for the professor to write about. These letters are also not useful and considered weak.
+Also, many professors write generic letters for students, e.g., the students didn't do any
+research or make any impression for the professor to write about.
+These letters are also not useful and considered weak.
+It might be a good idea to directly ask if the prof.
+is willing to write a \emph{strong} letter for you.
+If not, then you should ask someone else.
+For example, if a student I don't know well ask me to write letter for them, I will explicitly tell them I don't know them that well to write about them in details, and such a short, generic, and weak letter will hurt their chance.
 
@@
 \tcblower
-\textbf{Vu}: In general, at least one writer should have PhD to properly evaluate your research ability.
-However, I think this is an interesting and useful detail that US faculty might not be aware of and students should mention about this in their statements.
+\textbf{Vu}: This is an interesting and useful detail that US faculty might not be aware of and students should mention about this in their statements.
+In general, I think it should be fine as long as that person can properly evaluate your research ability.
 \end{commentbox}

### 2023-11-03 demystify.tex
updates

@@
 \begin{itemize}
-    \item Many schools do not offer PhD studies in CS and many CS professors \emph{cannot} be advisor, i.e., cannot graduate PhD students  (\S\ref{sec:schoolsandprofs}).
+    \item Many schools do not offer PhD studies in CS and many CS professors \emph{cannot} be adviser, i.e., cannot graduate PhD students  (\S\ref{sec:schoolsandprofs}).
     \item Contacting a prof. for research opportunities is recommended, but do it \emph{properly} (\S\ref{sec:contact}).
@@
 Take notes of faculty who make you excited, count those that are taking in new students (if they meet you, likely they are considering new students!).
-Talk to students about their advisors, the dept, the area, funding situation etc.
+Talk to students about their advisers, the dept, the area, funding situation etc.
 Ask about anything you want to determine that they deserve \emph{you}.
@@
 
-\subsubsection{Teaching Assistant (TA)}
-
-TA is common in the beginning when you haven't found your advisor who would pay you RA.
+\subsubsection{Teaching Assistant (TA)}\label{sec:ta}
+
+TA is common in the beginning when you haven't found your adviser who would pay you RA.
 As a TA, you spend up to 20 hrs/week and help professors with their classes (e.g., grading or teaching labs/recitation).
 Your TA is paid through the school/department, i.e., they hire you to help teach.
-During a semester, a TA might work with several courses and professors (not necessary their advisor).
+During a semester, a TA might work with several courses and professors (not necessary their adviser).
 TA funding typically is not available during the summer, which has much fewer courses.
@@
 \textbf{How to get RA?} When a professor recruits you, they will likely give you RA right away (e.g., when you apply).
-A common scenario is that you first get admitted with TA, and then after a year or two find an advisor to support you with RA.
+A common scenario is that you first get admitted with TA, and then after a year or two find an adviser to support you with RA.
 See how to contact a prof.
@@
 
-Choosing a school and an advisor is clearly one of the most important things in your mind when you apply to and especially when you get admitted.
+Choosing a school and an adviser is clearly one of the most important things in your mind when you apply to and especially when you get admitted.
 This is further complicated due to cultural differences and unfamiliarity of  international students to the US higher education system.
@@
 
-Now we discuss about finding advisors.
+Now we discuss about finding advisers.
+
+\subsubsection{Finding the right adviser}
+Obviously, there is no one-size-fit-all answer to this question.
+The best adviser is the one that you can work well with, i.e., \emph{fits} you.
+This is difficult and takes time.
+Fortunately, unlike many countries that require finding an adviser and research topic before starting the PhD (\S\ref{sec:non-us-differences}), CS PhD programs in the US will typically give you a couple of years to search for advisers and research topics.
+This is especially true if you're admitted with TA (\S\ref{sec:ta}), which gives you time to explore and find the right adviser.
+
+\begin{commentbox}[Xiaokuan:]
+    Whether the student's research interest matches that of the adviser is very important;
+    if there is a mismatch,
+    either the student or the adviser has to make compromises,
+    which often leads to disagreements or conflicts. 
+    IMO, the adviser should be the one who {\it guides}  students to do research while allowing students to pursue their own interest,
+    instead of {\it dictating} their research.
+    \end{commentbox}
+
+Here are some general advice to find professor.
+First, search for profs.
+that share similar research interests.
+You can learn about faculty and their level of research activity through the faculty's website or even from CSRankings.
+For example, in CSRankings, if you want to work with PL, you can search for those publish in PL conferences.
+If you want to work with SE \emph{and} AI, you can search for faculty who work in both SE and AI.
+You can also go to their website, look at their papers and projects.
+In any case, you should contact the professor and talk to them.
+For example, you can (cold) email them and ask about their research and if they are taking new students and schedule appointments to chat with them (office hrs are often the best time).
+See \S\ref{sec:contact} for more details on contacting professors.
+
+Another very effective way is taking graduate level courses in the topics you are interested in (remember: you have those 2 years to explore around).
+Professors teaching these \emph{special topics courses} and \emph{research seminars} typically are recruiting students---giving you even a higher chance.
+Do well in the class, answer questions, talk to the profs after classes, etc---being stand out.
+Many professors, including myself, prefer taking in new students this way.
+It gives both the professor and student more time, e.g., a whole semester, to work and evaluate the relationship before making any commitment (sounds a bit like a marriage!).
+You can also ask if you can do an independent study or research with a prof.
+This can be informal (no credit) and takes place during the summer or winter break.
+For example, I work with multiple students, many of whom are undergrads, this way.
+Many will drop because they find they don't like my research, but some will stay and become my PhD students.
+
+Ultimately, choose one that fits you the most by communicating with them, taking their courses, meeting them, asking them questions, and even talking to their current students.
+Indeed all this takes a lot of time and effort, but remember, you will be working with this person for 4-5 years, so it is important to try to find the right one.
+
+\begin{commentbox}[Thanh:]
+    In my opinion, having a well-suited adviser is crucial for a successful PhD and research career.
+One effective approach to finding a suitable professor is by working with a professor during your undergraduate studies.
+An exemplary instance is VinAI's residency program, where residents collaborate with professors from the US for two years before applying to PhD programs.
+Many VinAI residents have achieved remarkable results and gained admission to prestigious US universities.
+Unfortunately, VinAI's resident program is limited to AI research.
+
+    In other fields, e.g.
+Software Engineering, Vietnamese students face challenges in reaching US professors.
+Do you have any tips for Vietnamese students who want to connect with US professors and work as research assistants?
+    \tcblower
+    \textbf{Vu}: \S\ref{sec:contact} shows how to contact a professor for research opportunities.
+Many probably will say no as they do not have the bandwidth to take on random strangers, but some may say yes if they see you have a matching profile.
+    \end{commentbox}
 
 \subsubsection{Tenured, Tenure Track, Teaching,  Assistant Profs., Associate Profs}
-There are several types of faculty in a university that has a CS PhD program.
-You might hear about tenured,  tenure track, teaching faculty.
+Not every faculty in the CS program can serve as your formal adviser.
+Let's try to understand the different types of faculty and their roles.
+For example, you might hear about tenured,  tenure track, teaching faculty.
 You might also things such as assistant, associate, full, adjunct, emeritus, teaching, research professors.
@@
 \emph{Tenure-track} faculty are (more likely young) professors who are on the track to get tenure.
-Both tenure and tenured track faculty can serve as a formal advisor of Ph.D.
-students .
-This is very important because you need to have a tenure line faculty as your main PhD advisor.
-\S\ref{sec:tenure-vs-tenure-track} talks more about choosing between tenure and tenured-track professor as your advisor.
+\textbf{Both tenure and tenured track faculty can serve as a formal adviser of Ph.D.
+students}.
+This is very important because you need to have a tenure line faculty as your main PhD adviser.
+\S\ref{sec:tenure-vs-tenure-track} talks more about choosing between tenure and tenured-track professor as your adviser.
 
@@
 They can mentor Ph.D.
-students but they typically cannot serve as a formal advisor of Ph.D.
+students but they typically cannot serve as a formal adviser of Ph.D.
 students (i.e., they \emph{cannot graduate} PhD students).
 
-\paragraph{Rankings} Faculty have ranking, e.g., assistant, associate, full, regardless if they are tenure line or teaching.
+\paragraph{Titles} Faculty have titles, e.g., assistant, associate, full, regardless if they are tenure line or teaching.
 Assistant means new faculty, associate means they have been promoted, and full means they are senior.
@@
 
-Adjunct means they are not full-time faculty, e.g., they might be working in industry and teach a class or two.
+Adjunct faculty not full-time faculty, e.g., they might be working in industry and teach a class or two for fun.
 Emeritus means they are retired but still have some affiliation with the university.
 Research faculty (or scientist) are typically non-tenure line faculty who focus on research.
-Due to their roles, adjunct, emeritus, and research faculty typically do not serve as a formal advisor of Ph.D.
+Due to their roles, adjunct, emeritus, and research faculty typically do not advise Ph.D.
 students.
@@
 
-You can learn about faculty and their level of research activity through the faculty's website or even from CSRankings.
-Ultimately, choose one that fits you the most by communicating with them, meeting them, asking them questions, and even talking to their current students.
-
-\begin{commentbox}[Thanh:]
-    In my opinion, having a well-suited advisor is crucial for a successful PhD and research career.
-One effective approach to finding a suitable professor is by working with a professor during your undergraduate studies.
-An exemplary instance is VinAI's residency program, where residents collaborate with professors from the US for two years before applying to PhD programs.
-Many VinAI residents have achieved remarkable results and gained admission to prestigious US universities.
-Unfortunately, VinAI's resident program is limited to AI research.
-
-    In other fields, e.g.
-Software Engineering, Vietnamese students face challenges in reaching US professors.
-Do you have any tips for Vietnamese students who want to connect with US professors and work as research assistants?
-\textbf{Vu}: \S\ref{sec:contact} shows how to contact a professor for research opportunities.
-    \end{commentbox}
-
-\begin{commentbox}[Xiaokuan:]
-Whether the student's research interest matches that of the advisor is very important;
-if there is a mismatch,
-either the student or the advisor has to make compromises,
-which often leads to disagreements or conflicts. 
-IMO, the advisor should be the one who {\it guides}  students to do research while allowing students to pursue their own interest,
-instead of {\it dictating} their research.
-
-\end{commentbox}
 % \subsubsection{Strong Faculty vs High-ranked School}
@@
 \begin{commentbox}[Xiaokuan]
-    Applying for Ph.D. and contacting a potential Ph.D. advisor is a classic {\bf `why me, why you'} problem, 
+    Applying for Ph.D. and contacting a potential Ph.D. adviser is a classic {\bf `why me, why you'} problem, 
 similar to looking for a job in a company.
@@
 However, if your intention is a PhD in CS, then you likely need to be in the CS dept \emph{and} advised by a CS faculty.
-In fact, if a faculty is not in CS, it is unlikely that they can be the \emph{main} advisor of a CS PhD student---they may \emph{co-advise} or be in the PhD dissertation committee, but your main advisor will need to be a tenured or tenure-track faculty in CS.
+In fact, if a faculty is not in CS, it is unlikely that they can be the \emph{main} adviser of a CS PhD student---they may \emph{co-advise} or be in the PhD dissertation committee, but your main adviser will need to be a tenured or tenure-track faculty in CS.
 If in doubt, you should check with the CS department for their requirements.
@@
 
-\emph{Project proposal}: in many countries, you have to choose a project and advisor \emph{during} the application process (e.g., you write a proposal to a potential advisor).
-In the US, you often start your PhD without an advisor or project and find them later.
+\emph{Project proposal}: in many countries, you have to choose a project and adviser \emph{during} the application process (e.g., you write a proposal to a potential adviser).
+In the US, you often start your PhD without an adviser or project and find them later.
 This is also why US PhD programs are longer.
 
-\emph{Course work}: in the US you will spend the first couple of years taking classes and explore potential advisor and research topics.
-In other countries, you often do not do course work and start your research right away, e.g., you immediately work on the research project you proposed with the advisor you chose.
+\emph{Course work}: in the US you will spend the first couple of years taking classes and explore potential adviser and research topics.
+In other countries, you often do not do course work and start your research right away, e.g., you immediately work on the research project you proposed with the adviser you chose.
 Also, in the US you also have to pass a series of "exams", e.g., qualifying exam, comprehensive exam, thesis proposal defense\footnote{The word "ABD" (all but dissertation) is used in the US to refer to a PhD candidate who have finished all course work and exams and only need to write and defend their dissertation.}.
@@
 \emph{Funding}:  In Europe, stipend (salary) comes from the university or from gov't research organizations and often have a fixed duration (e.g., 3 years).
-In the US, stipend (e.g., RA) comes directly from your advisor (no fixed duration).
+In the US, stipend (e.g., RA) comes directly from your adviser (no fixed duration).
 There are also fewer TA opportunities in the European universities compared to the US.
@@
 \subsection{How much do \emph{you} cost?}\label{sec:ra-cost}
-PhD students often ask why their salary is so low compared to ludicrous grants their advisors get.
+PhD students often ask why their salary is so low compared to ludicrous grants their advisers get.
 They also wonder why their offer letters sometime mentioned that their benefits appear better than what they actually receive (i.e., stipend).
@@
 At the end, the total budget comes out to be \$69K/year to support a PhD student.
-\emph{The summary is that over your 5-6 year of your PhD, you cost about \$350K--400K, and while you're paid X, your advisor probably pays 2X for you}.
+\emph{The summary is that over your 5-6 year of your PhD, you cost about \$350K--400K, and while you're paid X, your adviser probably pays 2X for you}.
 
@@
 \paragraph{Illegal Software} Using illegal/cracked software is very common in many countries (and even in the US).
-However, \emph{do not} install or use them on university computers (e.g., those given by your department or advisor).
+However, \emph{do not} install or use them on university computers (e.g., those given by your department or adviser).
 It is unlikely that the university will track you down, but it is the \emph{software company} that will.
 They have very sophisticated tools to detect illegal software and will sue your university/department.
-Imagine your department or advisor being sued for a large sum of money, and it is \emph{you} who caused it.
-If you need software, ask your advisor or the department.
+Imagine your department or adviser being sued for a large sum of money, and it is \emph{you} who caused it.
+If you need software, ask your adviser or the department.
 

### 2023-11-04 demystify.tex
updates

@@
 
-\subsection{How do I address a professor?}
+\subsection{How do I address a professor?}\label{sec:address}
 
@@
 If you need software, ask your adviser or the department.
+
+\paragraph{Gifts and professionalism} In many countries, it's customary to give professors gifts, often during holidays.
+These gifts can be costly and expected by professors.
+In the US, it's not a common practice, and it's perfectly acceptable not to give gifts.
+However, if you'd like to offer small souvenir-like tokens, it's a thoughtful gesture that's appreciated.
+Professors often display such gifts proudly, and they can come from students, colleagues, or when they travel.
+In summary, small gifts are fine, but avoid anything that might make your professors uncomfortable.
+
+There appears to be a misconception that in the US things are strictly business-oriented, e.g., profs are your bosses who pay students for their work.
+And if you do something nice, then your action can be interpreted as you want something in return.
+However, the reality is quite the opposite.
+While in the US people can be a bit more straightforward and direct (e.g., your prof.
+will let you know quickly if they don't like something), they are also friendly and informal.
+For example, you can call your prof by their first name (\S\ref{sec:address}), you can disagree with them (and actually gain lots of respect doing so), you can ask them for help (even on personal issues---they might not be able to help, but they can listen and might have good advice), you can come to their house for parties, and of course you can even give them small thoughtful gifts that they put on their desks.
+Note that many students and professors have long lifetime relationships beyond the student's academic journey.
+For example, I still keep in touch with many of my mentors (even my 9th grade teacher in Hawaii), and friends through cards, emails, and phone calls during holidays or simply whenever I see something remind me of them.
+These practices are genuine and serve as a meaningful way to stay connected with others.
+
+% There is a common misconception that the United States maintains a strictly business-oriented and purely professional atmosphere, where professors are viewed as bosses who pay students for their work.
+However, the reality is quite the opposite.
+While it's true that Americans tend to be more straightforward and direct in their interactions, they also value friendliness and informality.
+
+% For instance, in the United States, it's perfectly acceptable to address your professor by their first name (as discussed in Section \ref{sec:address}).
+Students can also express their disagreements with professors, and in doing so, they can actually earn a great deal of respect.
+Moreover, you can approach your professors for assistance, even on personal matters.
+While they might not always have the answers, they are often willing to listen and offer valuable advice.
+
+% A unique aspect of the American academic culture is the opportunity for social interaction with professors.
+It's not uncommon for students to be invited to their professors' homes for parties and gatherings, such as Thanksgiving dinners, which can foster a sense of community and camaraderie.
+
+% In addition, small, thoughtful gifts are often exchanged between students and professors, and it's heartwarming to see these tokens displayed on professors' desks, symbolizing the strong connections that can develop between them.
+
+% What's particularly noteworthy is the enduring relationships that many students and professors maintain well beyond the student's academic journey.
+These connections can span a lifetime, with individuals staying in touch through cards, emails, and phone calls during holidays or simply when something reminds them of one another.
+This practice serves as a genuine and meaningful way to stay connected with mentors and friends, reflecting the deep-rooted value of maintaining relationships in American society.
 

### 2023-11-05 demystify.tex
updates

--- 2023-11-03 demystify.tex
@@
 If you need software, ask your adviser or the department.
+
+\paragraph{Gifts and Professionalism} In many countries, it's customary to give professors gifts, often during holidays.
+These gifts can be costly and profs.
+sometimes expected them.
+In the US, it's not a common practice, and it's perfectly OK to not to give gifts.
+\emph{However}, if you'd like to offer small souvenir-like tokens, it's a thoughtful gesture that's appreciated.
+Some professors proudly display their gifts, which can come from students and colleagues (e.g., when they travel to their home countries or conferences).
+In summary, small gifts are fine, but avoid anything that might make your professors uncomfortable.
+
+There's a misconception that in the US it's all business, with professors as bosses who pay students for their work, and if you do something nice implies you expect something in return.
+However, the reality is quite the opposite.
+While people can be straightforward and direct, they are also friendly and informal.
+You can call your professor by their first name (\S\ref{sec:address}), respectfully disagree and argue with that (and can actually gain respect doing so), seek their help (even on personal matters), come to their houses for parties or gathering, and give them small thoughtful gifts that they put on their desks.
+Many students and professors maintain lifelong relationships, staying in touch through cards, emails, and calls, even after their academic journey ends.
+These connections are genuine and meaningful.
 

### 2023-11-06 demystify.tex
Updates from Overleaf

--- 2023-11-03 demystify.tex
@@
 Also, use the statement to talk about stuff that admission committee members might not know about, e.g., your Github project with 1K+ stars or  your regular contributions to well-known open-source projects.
-Finally, talk about what makes you different than the other 1000 applicants that we are reviewing (e.g., see \S\ref{sec:improve-your-chance} for increasing your admission chance).
+Finally, talk about what makes you different than the other applicants (e.g., see \S\ref{sec:improve-your-chance} for increasing your admission chance).
 
@@
 
-TA is common in the beginning when you haven't found your adviser who would pay you RA.
-As a TA, you spend up to 20 hrs/week and help professors with their classes (e.g., grading or teaching labs/recitation).
+TA is common in the beginning when you haven't found an adviser who would pay you RA.
+As a TA, you spend up to 20 hrs/week and help with classes (e.g., grading or teaching labs/recitation).
 Your TA is paid through the school/department, i.e., they hire you to help teach.
 During a semester, a TA might work with several courses and professors (not necessary their adviser).
-TA funding typically is not available during the summer, which has much fewer courses.
-
-\paragraph{How to get TA?}  Unless you have other funding such as RA or Fellowships, TA is typically a default thing for good CS PhD programs.
+TA funding is not typically available during the summer, which has few courses.
+
+\paragraph{How to get TA?}  Unless you have other funding such as RA or Fellowships, TA is typically default for good CS PhD programs.
 When you apply to be a full-time student,  state that you need financial assistant.
@@
 \begin{commentbox}[Vu:]
-At GMU CS, students admitted with TA have  4 years of GTA guaranteed and also receive  stipend for the \textbf{first} summer.
+At GMU CS, students admitted with TA have  4 years of GTA guaranteed, and in some cases also receive  stipend for the first summer.
 \end{commentbox}
@@
 Many posts from VietPhD, a Facebook group for Vietnamese PhD students, are about how PhD students are mistreated, stressed, and miserable.
-Indeed there are issues with PhD and academic in general, but then there are also issues with jobs, family, and life.
-Certainly there are bad relationships with professors, co-authors, and officemates, and toxic working environment, feeling left out or discriminated (e.g., because you're an international student), and it is good to be aware of those.
-However, in reality there are many good mentors, supportive labmates and department, happy students, etc.
-So don't let social media make you feel pessimistic and destroy your quest to advance knowledge.
+Certainly there are bad relationships with professors, co-authors, and lab mates, and toxic working environment, feeling left out or discriminated (e.g., because you're an international student), and it is good to be aware of those.
+However, in reality there are many good mentors, supportive lab mates and department, happy students, etc.
+So don't let social media make you feel pessimistic and deter your quest to advance knowledge.
 
@@
 
-This section lists various cultural issues that international students should pay attention to.
+This section lists various cultural issues that international students might want to pay attention to.
 
@@
 If you need software, ask your adviser or the department.
+
+\paragraph{Gifts} In many countries, it is customary to give professors gifts, often during holidays.
+These gifts can be costly and profs.
+sometimes expected them.
+In the US, it is uncommon and perfectly OK to not to give gifts.
+\emph{However}, if you'd like to offer small souvenir-like tokens, it's a thoughtful gesture that's appreciated.
+Some professors proudly display their gifts, which can come from students and colleagues (e.g., when they travel to their home countries or conferences).
+In summary, small gifts are fine, but avoid anything that might make your professors uncomfortable.
+
+There's a misconception that in the US it is all business, with professors as bosses who pay students for their work, and doing nice thing implies you expect something in return.
+However, the reality is quite the opposite.
+While people can be straightforward and direct, they are also friendly and informal.
+You can call your professor by their first name (\S\ref{sec:address}), disagree and argue with that (and can actually gain respect doing so), seek their help (even on personal matters), come to their houses for parties or gathering, and give them small thoughtful gifts that they put on their desks.
+Many students and professors also maintain lifelong relationships, staying in touch through cards, emails, and calls, even after their academic journey ends.
 

### 2023-11-06 demystify.tex
updates

--- 2023-11-05 demystify.tex
@@
 %\usepackage[small,compact]{titlesec} %very powerful
-\usepackage{tcolorbox}
+\usepackage[most]{tcolorbox}
 \usepackage{enumitem}
@@
 
-\newtcolorbox{cbox}[1][]
-{
-left=1pt,right=1pt,top=1pt,bottom=1pt,
-colback=gray!10,
-boxrule=0.5pt
+% \newtcolorbox{cbox}[1][]
+% {
+% left=1pt,right=1pt,top=1pt,bottom=1pt,
+% colback=gray!10,
+% boxrule=0.5pt
+% }
+
+\newtcolorbox{mybox}{
+enhanced,
+boxrule=0pt,frame hidden,
+borderline west={2pt}{0pt}{green!75!black},
+colback=green!10!white,
+sharp corners
 }
@@
 \small
-    \begin{cbox}
-    \textbf{#1} 
+    \begin{mybox}
+    {\small \textbf{#1}}
  }{
-   \end{cbox}
+   \end{mybox}
 }
@@
 \section{Summary}\label{sec:summary}
+
 Here we summarize the main points of the document. This allows you to quickly skim through the document and decide which parts you want to read in detail.
@@
 In short, talk about your research goal and vision, and convince us that you can achieve it through your experience, e.g., published papers, or if you work on some projects by yourself, talk about it.
-Also, use the statement to talk about stuff that admission committee members might not know about, e.g., your Github project with 1K+ stars or  your regular contributions to well-known open-source projects.
-Finally, talk about what makes you different than the other 1000 applicants that we are reviewing (e.g., see \S\ref{sec:improve-your-chance} for increasing your admission chance).
+Also, use the statement to talk about things that admission committee members might not know about, e.g., your Github project with 1K+ stars or your regular contributions to well-known open-source projects.
+Finally, talk about what makes you different than the other 1000 applicants that we are considering (e.g., see \S\ref{sec:improve-your-chance} for increasing your admission chance).
 
@@
 \tcblower
-\textbf{Vu}: Vietnamese students and even faculty often lament how this competitive/low-grading system hurts Vietnamese students when applying abroad.
+\textbf{Vu}: Vietnamese students and even faculty often lament how this grading system hurts Vietnamese students applying abroad.
 One way to mitigate this is making these issues known to admission committee in your statement.
@@
 
-Another very effective way is taking graduate level courses in the topics you are interested in (remember: you have those 2 years to explore around).
+Another very effective way is taking graduate level courses in the topics you are interested in (remember: you have those 2 years to explore).
 Professors teaching these \emph{special topics courses} and \emph{research seminars} typically are recruiting students---giving you even a higher chance.
@@
 
-There's a misconception that in the US it's all business, with professors as bosses who pay students for their work, and if you do something nice implies you expect something in return.
+There's a misconception that in the US it's all business, with professors as bosses who pay students for their work, and you doing something nice implying you expect something in return.
 However, the reality is quite the opposite.

### 2023-11-06 demystify.tex

@@
 
-\paragraph{Gifts and Professionalism} In many countries, it's customary to give professors gifts, often during holidays.
+\paragraph{Gifts} In many countries, it is customary to give professors gifts, often during holidays.
 These gifts can be costly and profs.
 sometimes expected them.
-In the US, it's not a common practice, and it's perfectly OK to not to give gifts.
+In the US, it is uncommon and perfectly OK to not to give gifts.
 \emph{However}, if you'd like to offer small souvenir-like tokens, it's a thoughtful gesture that's appreciated.

### 2023-11-06 demystify.tex
updates

@@
 
-\subsection{Research or Personal Statement}\label{sec:research-statement}
+\subsection{Statement of Purpose (SOP)}\label{sec:research-statement}
 
@@
 While you might not have control over LORs or \hyperref[sec:your-school]{where your go to school}, you do over your
-statement! So write it well because it could make a big difference.
+statement of purpose (SOP) or personal statement! So write it well because it could make a big difference.
 In your statement, you have the opportunity to make  make your application stand out and unique, which can make you \emph{fit} the CS program you're applying to, even if you don't have very strong research experience.
@@
 Even people graduated with a Ph.D.
-have a great difficulty in explicating their researh vision.
+have a great difficulty in explicating their research vision.
 In my opinion, you should focus on showing (not telling) your research passion (if you do have).
@@
 As a TA, you spend up to 20 hrs/week and help with classes (e.g., grading or teaching labs/recitation).
-Your TA is paid through the school/department, i.e., they hire you to help teach.
+Your TA is paid through the school or department, i.e., they hire you to help teach.
 During a semester, a TA might work with several courses and professors (not necessary their adviser).
@@
 
-Choosing a school and an adviser is clearly one of the most important things in your mind when you apply to and especially when you get admitted.
+Choosing a school and an adviser is clearly one of the most important things in your mind when you apply and especially when you get admitted.
 This is further complicated due to cultural differences and unfamiliarity of  international students to the US higher education system.
@@
 
-We will first discuss about universities in the US that offer PhD in CS.
+We will first discuss about universities in the US that offer PhD in CS. Then we will talk about how to rank and select them.
 
@@
     \textbf{Vu}: \S\ref{sec:contact} shows how to contact a professor for research opportunities.
-Many probably will say no as they do not have the bandwidth to take on random strangers, but some may say yes if they see you have a matching profile.
+Many probably will say no as they do not have the bandwidth to take on random students, but some may say yes if they see potential fits.
 \end{commentbox}
@@
 students (i.e., they \emph{cannot graduate} PhD students).
+
+It is worth noting that professors in a non-CS department also unlikely can serve as formal PhD advisers for CS students (see \S\ref{sec:related-fields}).
 
@@
 
-\subsubsection{PhD in other related fields: CE, IST, Cybersecurity}
+\subsubsection{PhD in other related fields: CE, IST, Cybersecurity}\label{sec:related-fields}
 
@@
 
+\paragraph{About Me} I am an assistant professor at George Mason University (GMU).
+Prior to GMU, I was at the University of Nebraska-Lincoln (UNL).
+I have been involved in the PhD admission process at GMU and UNL for many years and currently I am the program director of the Masters program in Software Engineering at GMU (thus also have some experience with the MS admission process--which is quite different than PhD).
+I got my undergraduate and MS degrees from Penn State.
+I did my PhD at the University of New Mexico-Albuquerque, and then a postdoc at the University of Maryland-College Park.
+I had also worked at the Naval Research Lab and Lockheed Martin.
+
+While I grew up in the US, I am familiar with international students, e.g., many of my students are international students.
+I have many international collaborators and also have been a mentor for many students from Vietnam.
+I also have many colleague and friends who were international students and have heard their stories.
+I hope to capture their challenges and experiences in this document so that it can be useful for prospective students.
+
 \paragraph{Acknowledgement} Many people have contributed to this document.

### 2023-11-06 demystify.tex
updates

@@
 For example, if you have taken a course on Algorithms, even online ones like Coursera, you can talk about it in your statement.
-If you have done a project or have a professional certification (e.g., A+) through work that requires knowledge of OS, you can talk about it in your statement.
+If you have done a project that requires knowledge of OS or have a professional certification (e.g., A+) through work, you can talk about it in your statement.
 If you have done research that requires knowledge of Discrete Maths, you can talk about it in your statement.
@@
 Many universities are well-aware that incoming graduate students might not have all the technical background, so they often have a "bridge" courses to help students catch up.
-For example, GMU has 4 bridge courses (Data Structure and Algorithm, Computer Systems, Discrete Maths, and Programming Foundation) that incoming students can take to catch up on their CS knowledge.
+For example, GMU has four bridge courses (Data Structures and Algorithms, Computer Systems, Discrete Math, and Programming Foundation) that incoming students can take to catch up on their CS knowledge.
 
 \begin{commentbox}[Vu:]
-    I would strongly advocate to admit a student with a non-STEM background (e.g., business, English major) but can show that they have a strong drive for CS by studying core CS knowledge through various channels (e.g., self-study through online courses, projects, etc).
+    I would strongly advocate for a non-STEM student who shows that they have a strong drive for CS by studying core CS knowledge through various channels (e.g., self-study through online courses, projects, etc).
 I have seen many students with non-CS background who are very successful in CS PhD.
-In fact, I have seen many students with CS background who are not successful in CS PhD.
+I also have seen many students with CS background who are not successful in CS PhD.
 So it is not about your background, it is about your drive and passion for CS research.
@@
 
-\paragraph{About Me} I am an assistant professor at George Mason University (GMU).
+\paragraph{About Me} I am an assistant professor in the CS dept at George Mason University (GMU).
 Prior to GMU, I was at the University of Nebraska-Lincoln (UNL).
-I have been involved in the PhD admission process at GMU and UNL for many years and currently I am the program director of the Masters program in Software Engineering at GMU (thus also have some experience with the MS admission process--which is quite different than PhD).
+I have served in the PhD admission process at GMU and UNL for many years.
+Currently I am also the program director of the Masters program in Software Engineering at GMU (thus also have some experience with the MS admission process--which is quite different than PhD).
 I got my undergraduate and MS degrees from Penn State.
-I did my PhD at the University of New Mexico-Albuquerque, and then a postdoc at the University of Maryland-College Park.
-I had also worked at the Naval Research Lab and Lockheed Martin.
+I completed my PhD at the University of New Mexico-Albuquerque and a postdoc at the University of Maryland-College Park.
 
 While I grew up in the US, I am familiar with international students, e.g., many of my students are international students.
-I have many international collaborators and also have been a mentor for many students from Vietnam.
+I have many collaborators from various countries and have been a mentor for many students from Vietnam.
 I also have many colleague and friends who were international students and have heard their stories.
-I hope to capture their challenges and experiences in this document so that it can be useful for prospective students.
+I hope to capture their challenges and experiences in this document so that it can be useful for prospective international students.
 

### 2023-11-06 demystify.tex
updates

@@
 
-\paragraph{Diversity} US universities prioritize diversity and inclusion.
+%\paragraph{Diversity} US universities prioritize diversity and inclusion.
 Students need to respect and appreciate varied opinions, backgrounds, and cultures.

### 2023-11-06 demystify.tex
updates

@@
 
-\emph{Funding}:  In Europe, stipend (salary) comes from the university or from gov't research organizations and often have a fixed duration (e.g., 3 years).
+\emph{Funding}:  In many countries stipend comes from the university or from gov't research organizations and often have a fixed duration (e.g., 3 years).
 In the US, stipend (e.g., RA) comes directly from your adviser (no fixed duration).
@@
 
-\emph{Faculty position after PhD}: In other countries, PhD graduates typically apply for research positions at research labs, i.e., postdocs, and then consider faculty position.
-In the US, PhD graduates interested in academia can apply directly for faculty position.
+\emph{Faculty position after PhD}: In other countries, PhD graduates who are interested academia typically apply for research positions at research labs, i.e., postdocs, and then consider faculty position.
+In the US, PhD graduates often apply directly for faculty position (postdoc for US graduates is no longer a popular option as it was before).
 This is why US PhD programs are longer and more focused on research.
 
-\emph{Work-life balance}: PhD students are often overworked in the US (seriously!) compared to other countries, e.g., in Europe.
-This is partly due to the longer PhD program and that US PhD students are often paid through TA, which requires them to work for the dept.
+\emph{Work-life balance}: I believe that PhD students are often overworked in the US compared to other countries, e.g., in Europe.
+This is partly due to the longer PhD program and that US PhD students are often paid through TA, which requires them to do TA in addition to their own research.
 In contrast, PhD students in other countries are often paid through fellowships, which do not require them to do TA work.
@@
 
-Note that in some universities the formal title Dr.
-Lastname is preferred.
+Note that in some universities the formal title \emph{Dr.} Lastname is preferred.
 You just need to observe and follow the conventions at your particular institution.
@@
 
-\paragraph{Referring to professors you know} Because you are already familiar with these individuals, you can just informally use their firstname as mentioned above (or Dr./Prof., if you want to be formal).
+\paragraph{Referring to professors you know} Because you are already familiar with these individuals, you can just informally use their names if they are OK with it as mentioned above (or Dr./Prof., if you want to be formal).
 You can also include their institution if it makes it more precise.
@@
 Do not include ranking (e.g., Assistant, Associate, Scientist, ...) when referring to someone.
-I see many international students include a lengthy title of people they know, e.g., \emph{I am advised by Asst.
+I see many students include a lengthy title of people they know, e.g., \emph{I am advised by Asst.
 Prof.
@@
 Many posts from VietPhD, a Facebook group for Vietnamese PhD students, are about how PhD students are mistreated, stressed, and miserable.
-Certainly there are bad relationships with professors, co-authors, and lab mates, and toxic working environment, feeling left out or discriminated (e.g., because you're an international student), and it is good to be aware of those.
+Certainly issues including bad relationships with professors, conflicts with co-authors and lab mates, and toxic working environment, feeling left out or discriminated (e.g., because you're an international student) \emph{do} exist, and it is good to be aware of those.
 However, in reality there are many good mentors, supportive lab mates and department, happy students, etc.
@@
 
-Here is the typically steps: (i) a faculty suspecting a cheating case will report it to the Office of Academic Integrity (OAI) at the university;  (ii)  OAI will take over and investigate the case; and finally (iii) OAI will make the final decision.
+Here is the typically steps: (i) a faculty suspecting a cheating case will report it to the Office of Academic Integrity (OAI) at the university---the report often has supporting evidence and suggested penalty (e.g., receive a failing grade);  (ii)  OAI will take over and investigate the case; and (iii) OAI will make the final decision.
 It is important to note that after receiving the report from your prof., OAI \emph{completely} takes over and makes decision.
@@
 While people can be straightforward and direct, they are also friendly and informal.
-You can call your professor by their first name (\S\ref{sec:address}), respectfully disagree and argue with that (and can actually gain respect doing so), seek their help (even on personal matters), come to their houses for parties or gathering, and give them small thoughtful gifts that they put on their desks.
-Many students and professors maintain lifelong relationships, staying in touch through cards, emails, and calls, even after their academic journey ends.
-These connections are genuine and meaningful.
+You can call your professor by their first name (\S\ref{sec:address}), disagree with them and argue (and can actually gain respect doing so), seek their help (even on personal matters), come to their houses for parties or gathering, and give them small thoughtful gifts that they put on their desks.
+Many students and professors maintain lifelong relationships, staying in touch through cards, emails, and calls, even after their academic journey ends.% These connections are genuine and meaningful.
 
@@
 Prior to GMU, I was at the University of Nebraska-Lincoln (UNL).
-I have served in the PhD admission process at GMU and UNL for many years.
-Currently I am also the program director of the Masters program in Software Engineering at GMU (thus also have some experience with the MS admission process--which is quite different than PhD).
+I have been in the PhD admission process at GMU and UNL for many years.
+Currently I serve as the program director of the Masters program in Software Engineering at GMU (thus also have some experience with the MS admission process--which is quite different than PhD).
 I got my undergraduate and MS degrees from Penn State.
@@
 
-While I grew up in the US, I am familiar with international students, e.g., many of my students are international students.
-I have many collaborators from various countries and have been a mentor for many students from Vietnam.
+While I am not an international student, many of my students and collaborators are international.
+I also mentor many students from Vietnam.
 I also have many colleague and friends who were international students and have heard their stories.

### 2023-11-06 demystify.tex
updates
using memoir class

@@
-\documentclass[11pt]{article}
+\documentclass[oneside,11pt]{memoir}
+\setsecnumdepth{subsection}
+
 \usepackage[T1]{fontenc}
-\usepackage[margin=1.0in]{geometry}
+\usepackage[margin=1.2in]{geometry}
 \usepackage{soul}
@@
 \usepackage{graphicx}
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{amsthm}
 \usepackage{hyperref}
@@
 }
-
 \renewcommand{\figurename}{Fig.}
@@
 
-\title{\vspace{-1in} Demystifying the Computer Science \\PhD Admission in US Universities\\{\large A Guide for International Students}}
-%\date{\today}
+\title{Demystifying the Computer Science \\PhD Admission in US Universities\\{\large A Guide for International Students}}
+
 \author{\href{https://nguyenthanhvuh.github.io}{ThanhVu (Vu) Nguyen} (\href{mailto:tvn@gmu.edu}{tvn@gmu.edu})\\{\small George Mason University, Dept. of Computer Science}}
@@
 \begin{document}
+\frontmatter
 
 \maketitle
-\setcounter{section}{-1}
-
-\begin{abstract}
+\chapter{Preface}
 Having been involved in PhD admissions for many years, I've
@@
 
-    The latest version of document is at \url{https://nguyenthanhvuh.github.io/phd-cs-us/demystify.pdf} and its \LaTeX{} source is on \href{https://github.com/nguyenthanhvuh/phd-cs-us}{Github}.
+The latest version of this document is at \url{nguyenthanhvuh.github.io/phd-cs-us/demystify.pdf} and its \LaTeX{} source is on \href{https://github.com/nguyenthanhvuh/phd-cs-us}{Github}.
 If you have questions or comments, feel free to create a \href{https://github.com/nguyenthanhvuh/phd-cs-us/issues}{GitHub issue} for discussion.
-\end{abstract}
-
-\tableofcontents
-
-\section{Summary}\label{sec:summary}
-
+
+\newpage
+\tableofcontents*
+
+\chapter{Summary}\label{sec:summary}
 Here we summarize the main points of the document. This allows you to quickly skim through the document and decide which parts you want to read in detail.
@@
 
-\section{Should You Apply to do a PhD in CS in the US?}\label{sec:should}
+\mainmatter
+\chapter{Should You Apply to a PhD program in CS in the US?}\label{sec:should}
 
@@
 
-\section{How is Your Application Evaluated?}\label{sec:evalapps}
+\chapter{How is Your Application Evaluated?}\label{sec:evalapps}
 
@@
 
-\section{Application}\label{sec:application}
+\chapter{Application}\label{sec:application}
 
@@
 
-\subsection{Research Ability}
+\section{Research Ability}
 
@@
 
-\subsection{Letter of Recommendation (LOR)}\label{sec:lor}
+\section{Letter of Recommendation (LOR)}\label{sec:lor}
 
@@
 
-\subsection{Statement of Purpose (SOP)}\label{sec:research-statement}
+\section{Statement of Purpose (SOP)}\label{sec:research-statement}
 
@@
 
-\subsection{Your School}\label{sec:your-school}
+\section{Your School}\label{sec:your-school}
 
@@
 
-\subsection{Grades/GREs}\label{sec:grades}
+\section{Grades/GREs}\label{sec:grades}
 
@@
 
-\subsection{CV/Resume}
+\section{CV/Resume}
 The CV/Resume should provide a summary of the applicant's achievements.
@@
 
-\subsection{Interview}
+\section{Interview}
 Sometimes a faculty wants to interview an applicant to make a decision.
@@
 
-\section{Getting Admitted}\label{sec:accepted}
+\chapter{Getting Admitted}\label{sec:accepted}
 
@@
 
-\section{Funding}\label{sec:funding}
+\chapter{Funding}\label{sec:funding}
 
@@
 
-\subsection{Graduating Assistantship (TA/RA)}
+\section{Graduating Assistantship (TA/RA)}
 The most common type of funding is \textbf{graduate assistanship}, which is either TA or RA.
@@
 
-\subsubsection{Teaching Assistant (TA)}\label{sec:ta}
+\subsection{Teaching Assistant (TA)}\label{sec:ta}
 
@@
 
-\subsubsection{Research Assistant (RA)}
+\subsection{Research Assistant (RA)}
 RA is provided through a professor through their own funding so you can work on their project.  
@@
 
-\subsection{Fellowship/Scholarship}
+\section{Fellowship/Scholarship}
 
@@
 
-\section{Choosing Schools and Professors}\label{sec:schoolsandprofs}
+\chapter{Choosing Schools and Professors}\label{sec:schoolsandprofs}
 
@@
 
-\subsection{Universities}
+\section{Choosing a University}
 
@@
 
-\subsubsection{Schools that offer PhD in CS}
+\subsection{Schools that offer PhD in CS}
 
@@
 
-%\subsubsection{R1, R2, ...}
-
-\subsubsection{Selecting and Ranking Schools for CS}
+%\subsection{R1, R2, ...}
+
+\subsection{Selecting and Ranking Schools for CS}
    \begin{center}
@@
 
-\subsection{Choosing a Professor}
+\section{Choosing a Professor}
 
@@
 
-\subsubsection{Finding the right adviser}
+\subsection{Finding the right adviser}
 Obviously, there is no one-size-fit-all answer to this question.
@@
 Fortunately, unlike many countries that require finding an adviser and research topic before starting the PhD (\S\ref{sec:non-us-differences}), CS PhD programs in the US will typically give you a couple of years to search for advisers and research topics.
-This is especially true if you're admitted with TA (\S\ref{sec:ta}), which gives you time to explore and find the right adviser.
+This is especially true if you're admitted with TA (\S\ref{sec:ta}), which gives you time to explore and find adviser.
 
@@
     which often leads to disagreements or conflicts. 
-    IMO, the adviser should be the one who {\it guides}  students to do research while allowing students to pursue their own interest,
-    instead of {\it dictating} their research.
+    IMO, the adviser should be the one who \emph{guides}  students to do research while allowing students to pursue their own interest,
+    instead of \emph{dictating} their research.
     \end{commentbox}
@@
 
-\subsubsection{Tenured, Tenure Track, Teaching,  Assistant Profs., Associate Profs}
+\subsection{Tenured, Tenure Track, Teaching,  Assistant Profs., Associate Profs}
 Not every faculty in the CS program can serve as your formal adviser.
@@
 
-\subsubsection{Tenured or tenure-track faculty? Who do you choose?}\label{sec:tenure-vs-tenure-track}
+\subsection{Tenured or tenure-track faculty? Who do you choose?}\label{sec:tenure-vs-tenure-track}
 
@@
 
-% \subsubsection{Strong Faculty vs High-ranked School}
+% \subsection{Strong Faculty vs High-ranked School}
 % \begin{commentbox}[Thanh:]
@@
 
-\subsubsection{Should you contact a professor? What to do to get a desired reply?}\label{sec:contact}
+\subsection{Should you contact a professor? What to do to get a desired reply?}\label{sec:contact}
 
@@
 \begin{commentbox}[Xiaokuan]
-    Applying for Ph.D. and contacting a potential Ph.D. adviser is a classic {\bf `why me, why you'} problem, 
+    Applying for Ph.D. and contacting a potential Ph.D. adviser is a classic \textbf{`why me, why you'} problem, 
 similar to looking for a job in a company.
 On a high level,
-you need to show that you have done your {\it homework}
+you need to show that you have done your \emph{homework}
 regarding the professor and the university,
@@
 
-\subsubsection{PhD in other related fields: CE, IST, Cybersecurity}\label{sec:related-fields}
+\subsection{PhD in other related fields: CE, IST, Cybersecurity}\label{sec:related-fields}
 
@@
 
-\section{Miscs and FAQs}
+\chapter{Miscs and FAQs}
 \epigraph{"I want to share something with you – the three little sentences that will get you through life; number 1: Cover for me, number 2: Oh, good idea, Boss, and number 3: It was like that when I got here."}{\textsc{The Simpsons}}
 
-\subsection{What can you do to increase your admission chance?}\label{sec:improve-your-chance}
+\section{What can you do to increase your admission chance?}\label{sec:improve-your-chance}
 
@@
 
-\subsection{Can I apply to CS PhD if my undergrad was not in CS or related areas?}\label{sec:non-stem}
+\section{Can I apply to CS PhD if my undergrad was not in CS or related areas?}\label{sec:non-stem}
 
@@
 
-\subsection{Is an MS degree required for admission to PhD in CS?}\label{sec:msrequirement}
+\section{Is an MS degree required for admission to PhD in CS?}\label{sec:msrequirement}
 No. In fact, student with BS can get MS degree ``along the way'' to PhD.  However, MS can help if it gives research experience or is from a more well-known school than your undergrad institution.
@@
 
-\subsection{How long does it take to complete the CS PhD program?}\label{sec:time}
+\section{How long does it take to complete the CS PhD program?}\label{sec:time}
 
@@
 
-\subsection{Difference between PhD in the US and other countries}\label{sec:non-us-differences}
+\section{Difference between PhD in the US and other countries}\label{sec:non-us-differences}
 Here we summarize the main differences between CS PhD in the US and other countries.  This is based on my experience and what I heard from others who did PhD in other countries.
@@
 
-\subsection{How do I address a professor?}\label{sec:address}
+\section{How do I address a professor?}\label{sec:address}
 
@@
 
-\subsection{How much do \emph{you} cost?}\label{sec:ra-cost}
+\section{How much do \emph{you} cost?}\label{sec:ra-cost}
 PhD students often ask why their salary is so low compared to ludicrous grants their advisers get.
@@
 
-%\subsection{Having fun during a PhD?}
+%\section{Having fun during a PhD?}
 %PhD students \emph{and faculty} probably find it amusing about the notion that students, especially international ones, can genuinely enjoy their PhD studies.
@@
 
-\subsection{Will I be miserable during my a PhD?}\label{sec:happy}
+\section{Will I be miserable during my a PhD?}\label{sec:happy}
 Many posts from VietPhD, a Facebook group for Vietnamese PhD students, are about how PhD students are mistreated, stressed, and miserable.
@@
 
-\section{Cultural Differences and Other Issues}\label{sec:cultural}
+\chapter{Cultural Differences and Other Issues}\label{sec:cultural}
 
@@
 
-\paragraph{Academic Integrity} Plagiarism and cheating (e.g., exams, coding) is a BIG no-no in the US.
+\section{Academic Integrity}
+
+Plagiarism and cheating (e.g., exams, coding) is a BIG no-no in the US.
 If you're caught cheating, you will face very heavy consequence and likely be expelled from the university (e.g., after the second time at GMU).
@@
 
-\paragraph{Illegal Software} Using illegal/cracked software is very common in many countries (and even in the US).
+\section{Illegal Software} Using illegal/cracked software is very common in many countries (and even in the US).
 However, \emph{do not} install or use them on university computers (e.g., those given by your department or adviser).
@@
 
-\paragraph{Gifts} In many countries, it is customary to give professors gifts, often during holidays.
+\section{Gifts} In many countries, it is customary to give professors gifts, often during holidays.
 These gifts can be costly and profs.
@@
 
-\section{Ranking of CS PhD programs in the U.S}\label{sec:ranking}
+\appendix
+
+\chapter{Ranking of CS PhD programs}\label{sec:ranking}
+
 Tab.~\ref{tab:ranking} lists the top 50 CS programs in the US from \href{https://www.csrankings.org}{CSRankings.org}, a ranking system  based on top CS conferences.
-
     \begin{table}
@@
 
-\section{History and Acknowledgement}
+\chapter{History and Acknowledgement}
 

### 2023-11-06 demystify.tex
Updates from Overleaf

@@
 
-After you submit your PhD application (usually December 15), it will be first screened
+After you submit your PhD application (usually November--December), it will be first screened
 for general requirements, e.g., did you submit your transcripts and standard scores? did your reference writers submit their letters?
@@
 The size and the review load of the committee depends on the department size.
-At GMU, the PhD committee typically has 15--20 faculty, and each committee member is assigned from 25--30 applications.
+At GMU, the PhD committee typically has 15--20 faculty, and each committee member is assigned to review 25--30 applications.
 Note that GMU, UMass, and probably many other schools, have separate committees for MS programs.
 
-Each application is then assigned to about \emph{three} faculty members, who will evaluate your profile and reach a consensus.
+Each application is assigned to about \emph{three} faculty members, who will evaluate your profile and reach a consensus.
 Note that while the assigned reviewers are likely the main ones deciding your application, \emph{every} faculty in the department will have access to your application and can provide inputs and opinions on your profile.
 
-In many cases, the PhD admission committee involves assistant professors in the department.
+The PhD admission committee typically involves assistant professors in the department (see \S\ref{sec:faculty-types} for various faculty titles).
 This provides junior faculty the opportunities to recruit students.
@@
 
-At GMU, we usually decide that a full-time PhD candidate is either (i) admit with funding (TA or RA) or (ii) rejected. In other words, in most cases, we either
+At GMU, we usually decide that a full-time PhD candidate is either (i) admit with funding (TA or RA, see \S\ref{sec:funding}) or (ii) rejected. In other words, in most cases, we either
 admit you with full funding, or we reject your application. In some rare cases, we admit
@@
 
-However, I understand that many international students do not have the opportunities to publish in top places, so general confs/journals would suffice.
+However, it is understandable that many students do not have the opportunities to publish in top places.
+Thus, general confs/journals, even those under submissions or rejected, would suffice.
 But be sure to upload your papers with your application and talk about them in your statement (see \S\ref{sec:research-statement}).
@@
 \begin{commentbox}[Vu:]
-GMU has \emph{Virtual} Open House (VOH), e.g., \url{https://cs-gmu.github.io/cs-phd-voh-s23/}, which I've co-organized in recent years.
+GMU has \emph{Virtual} Open House (VOH), e.g., \url{https://cs-gmu.github.io/cs-phd-voh-s23/}.
 We invite all admitted PhD students to the VOH through Zoom to learn about the CS program, the department, GMU, and the DC area in general.
@@
 Tab.~\ref{tab:funding} summarizes the differences.
-
 Note that funding is typically more available for PhD students than 
-Masters. Undergraduates typically have no funding and also have to pay international tuition, which is very expensive!.
+Masters. %Undergraduates typically have no funding and also have to pay international tuition, which is very expensive!.
 
@@
 \begin{commentbox}[Vu:]
-At GMU, PhD applicants are automatically eligible for a Presidential Fellowship.
+PhD applicants at GMU are automatically eligible for a \emph{Presidential Fellowship}.
 It is at least as good as GTA but the most important thing is that as a fellowship it is truly free money (i.e., you are not depending on any prof.
-or TA duties).
+or TA).
 PhD admission committee members nominate applicants for this fellowship and the committee will vote and give the fellowship to the top 2.
@@
 
-\subsection{Tenured, Tenure Track, Teaching,  Assistant Profs., Associate Profs}
+\subsection{Tenured, Tenure Track, Teaching,  Assistant Profs., Associate Profs}\label{sec:faculty-types}
+
 Not every faculty in the CS program can serve as your formal adviser.

### 2023-11-07 demystify.tex
updates

@@
 \usepackage[T1]{fontenc}
-\usepackage[margin=1.2in]{geometry}
+\usepackage[margin=1.5in]{geometry}
 \usepackage{soul}
@@
 
-\title{Demystifying the Computer Science \\PhD Admission in US Universities\\{\large A Guide for International Students}}
-
-\author{\href{https://nguyenthanhvuh.github.io}{ThanhVu (Vu) Nguyen} (\href{mailto:tvn@gmu.edu}{tvn@gmu.edu})\\{\small George Mason University, Dept. of Computer Science}}
+\title{Demystifying the Computer Science PhD Admission\\in US Universities\\{\large A Guide for International Students}}
+
+\author{\href{https://nguyenthanhvuh.github.io}{ThanhVu (Vu) Nguyen}\\{\small George Mason University, Dept. of Computer Science}}
+
+\chapterstyle{pedersen}
+
+\makeatletter
+\def\maketitle{%
+  \null
+  \thispagestyle{empty}%
+  \vfill
+  \begin{center}\leavevmode
+    \normalfont
+    {\LARGE\raggedright \@title\par}%
+    \vfill%
+    %\hrulefill\par
+    {\Large \@author\par}%
+    \vfill%
+    {\large\raggedleft \@date\par}%
+  \end{center}%
+  \vfill
+  \null
+  \cleardoublepage
+  }
+\makeatother
+%\author{ThanhVu H. Nguyen}
+%\title{TITLE OF BOOK}
+%\date{}
 
 \begin{document}
+\maketitle
 \frontmatter
 
-\maketitle
 \chapter{Preface}
@@
 
-\section{Difference between PhD in the US and other countries}\label{sec:non-us-differences}
-Here we summarize the main differences between CS PhD in the US and other countries.  This is based on my experience and what I heard from others who did PhD in other countries.
+\section{Differences between PhD in the US and other countries}\label{sec:non-us-differences}
+This summarizes the main differences between CS PhD in the US and other countries. % This is based on my experience and what I heard from others who did PhD in other countries.
 
@@
 
-\section{How do I address a professor?}\label{sec:address}
+\section{How do I call or address a professor?}\label{sec:address}
 
 \begin{center}
-        \includegraphics[scale=0.7]{c5.png}
+\includegraphics[scale=0.6]{c5.png}
 \end{center}

### 2023-11-08 demystify.tex
updates

@@
 So I want to share some details about the PhD admission process and advice for those who are interested in applying for a \textbf{PhD in Computer Science in the US}.
-While this document is primarily intended for students interested in CS, it may be relevant to students from various disciplines and should be 
-Furthermore, although many examples are specifics for schools that I and other contributors of this document know about, the information should be generalizable to other R1\footnote{An \href{https://en.wikipedia.org/wiki/List_of_research_universities_in_the_United_States}{R1 institution} in the US is a research-intensive university with a high level of research activity across various disciplines.} institutions in the US  (though \emph{very} top schools could be very selective, e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU).
+While this document is primarily intended for students interested in CS, it may be relevant to students from various disciplines.
+Furthermore, although many examples are specifics for schools that I and other contributors of this document know about, the information should be generalizable to other R1\footnote{An \href{https://en.wikipedia.org/wiki/List_of_research_universities_in_the_United_States}{R1 institution} in the US is a research-intensive university with a high level of research activity across various disciplines.
+Currently, 146 (out of 4000) universities are classified as R1.} institutions in the US  (though \emph{very} top schools could be very selective, e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU).
 
@@
 \begin{enumerate}
-\item Should you apply to a PhD in CS in the US?
+\item Should you apply to a CS PhD in the US?
 \begin{itemize}
@@
 \mainmatter
-\chapter{Should You Apply to a PhD program in CS in the US?}\label{sec:should}
+\chapter{Should You Apply to a CS PhD Program in the US?}\label{sec:should}
 

### 2023-11-09 demystify.tex
updates

@@
 
-\title{Demystifying the Computer Science PhD Admission\\in US Universities\\{\large A Guide for International Students}}
+\title{Demystifying the Computer Science PhD Admission in the US\\{\large A Guide for International Students}}
 
@@
     \normalfont
-    {\LARGE\raggedright \@title\par}%
+    {\LARGE\raggedright \textbf{\@title}\par}%
     \vfill%
@@
     \item The committee will look at various factors, but the most important ones are research ability, e.g., publications, personal statement, and recommendation letters.
-    \item LORs are important, but only if they are personalized and talk about your research ability (\S\ref{sec:lor}).
-    \item GRE \emph{is not} required by most schools (\S\ref{sec:grades}).
-    \item Grades are important, but only if your school is well-known (\S\ref{sec:grades}).
+    \item LORs are very important, but only if they are personalized and talk about your research ability (\S\ref{sec:lor}).
     \item Personal statement is very important. Write it in such a way that makes you \emph{stand out} (\S\ref{sec:research-statement} and \S\ref{sec:improve-your-chance})
+    \item GRE \emph{is not} required (\S\ref{sec:grades}). Spend your time on something else.
+    \item Grades are important, but depend on the reputation of your school (\S\ref{sec:grades}).
     \item Getting an interview is typically a \emph{good sign} (no one interviews weak candidates).
@@
 harder than at schools in other countries.
-If you believe you have a chance in other countries, e.g., South Korea, Singapore, Germany, UK, Japan and Australia, then you will surely have a chance in the US as well\footnote{\S\ref{sec:non-us-differences} compares CS PhD study in the US to other countries.}.
+If you believe you have a chance in other countries, e.g., South Korea, Singapore, Germany, UK, Japan and Australia, then you will surely have a chance in the US as well.
+\S\ref{sec:non-us-differences} compares CS PhD study in the US to other countries.
 
 \begin{commentbox}[Vu:]
-One of the reasons I create this document is that my colleagues at GMU are interested in recruiting Vietnamese students and are surprised when seeing very few applications in Vietnam (e.g., each year our CS program receives more than 350 PhD applications, most of which are international but only 3--4 are from Vietnam).
-In general the number of CS PhD applications from Vietnam to US universities is low and more would be welcomed.
+One of the reasons I create this document is that my colleagues at GMU are interested in recruiting Vietnamese students and are surprised when seeing very few applications in Vietnam (see \S\ref{sec:ack}).
+Each year our CS program receives more than 350 PhD applications, most of which are international but only 3--4 are from Vietnam.
+In general the number of CS PhD applications from Vietnam to US universities is very low and more would be welcome.
 \end{commentbox}
@@
 
-\chapter{Application}\label{sec:application}
+\chapter{Your Application}\label{sec:application}
 
@@
 
-\section{Research Ability}
+\section{Research Ability}\label{sec:research-ability}
 
@@
     \textbf{Vu}: \S\ref{sec:contact} shows how to contact a professor for research opportunities.
-Many probably will say no as they do not have the bandwidth to take on random students, but some may say yes if they see potential fits.
+Many probably will say no as they do not have the bandwidth to take on random students, but some may say yes if they see potential fit.
 \end{commentbox}
@@
 
-Plagiarism and cheating (e.g., exams, coding) is a BIG no-no in the US.
+Plagiarism and cheating (e.g., exams, assignments) is a BIG no-no in the US.
 If you're caught cheating, you will face very heavy consequence and likely be expelled from the university (e.g., after the second time at GMU).
@@
 
-Here is the typically steps: (i) a faculty suspecting a cheating case will report it to the Office of Academic Integrity (OAI) at the university---the report often has supporting evidence and suggested penalty (e.g., receive a failing grade);  (ii)  OAI will take over and investigate the case; and (iii) OAI will make the final decision.
+Here is the typically steps: (i) a faculty suspecting a cheating case will report it to the Office of Academic Integrity (OAI) at the university---the report often has supporting evidence and suggested penalty (e.g., a failing grade);  (ii)  OAI will take over and investigate the case; and (iii) OAI will make the final decision.
 It is important to note that after receiving the report from your prof., OAI \emph{completely} takes over and makes decision.
@@
 
-\chapter{History and Acknowledgement}
+\chapter{History and Acknowledgement}\label{sec:ack}
 

### 2023-11-09 demystify.tex
updates

@@
   \footnotesize
-  \caption{Different types of PhD funding}\label{tab:funding}
+  \caption{Different types of PhD funding. All Covered: includes tuition, insurance, and stipend.}\label{tab:funding}
   \begin{tabular}{c|c|c|c}
@@
     \textbf{From} & School & Profs. & School/External\\
-    \textbf{For}                  & Teaching Assistant       & Research                        & Research                              \\
-    \textbf{Tuition/Ins./Stipend} & Yes                      & Yes                             & Yes                                   \\
-    \textbf{Cover Summer?}              & No                       & Maybe                           & Yes                                   \\
+    \textbf{For}                  & Teaching Assist.       & Research                        & Research                              \\
+    \textbf{Cover All?} & Yes                      & Yes                             & Yes                                   \\
+    \textbf{Summer?}              & No                       & Maybe                           & Yes                                   \\
     \midrule
     \textbf{Pros}                 & Research Freedom         & Get to do research              & Research Freedom                      \\
-    \textbf{Cons}                 & Teaching Duties, Uncertain            & Research Restriction, Uncertain & Competitive, Limited             \\
+    \textbf{Cons}                 & Teaching Duties           & Research Restriction & Competitive, Limited             \\
     \bottomrule

### 2023-11-09 demystify.tex
updates

@@
 Having been involved in PhD admissions for many years, I've
-realized that many \textbf{international students}, especially those from smaller countries like Vietnam, lack a clear understanding of
+realized that many \textbf{international students} (and even domestic ones) lack a clear understanding of
 the Computer Science PhD admission process at US universities. This confusion not only
@@
 
+%though \emph{very} top schools could be very selective, e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU
 So I want to share some details about the PhD admission process and advice for those who are interested in applying for a \textbf{PhD in Computer Science in the US}.
@@
 Furthermore, although many examples are specifics for schools that I and other contributors of this document know about, the information should be generalizable to other R1\footnote{An \href{https://en.wikipedia.org/wiki/List_of_research_universities_in_the_United_States}{R1 institution} in the US is a research-intensive university with a high level of research activity across various disciplines.
-Currently, 146 (out of 4000) universities are classified as R1.} institutions in the US  (though \emph{very} top schools could be very selective, e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU).
+Currently, 146 (out of 4000) universities are classified as R1.} institutions in the US (and also likely universities in other countries).
 
@@
 \paragraph{English Test} Unless your degrees are from specific countries such as \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-GMU#standard-tests-waiver-eligible-countries}{these}, you will need to
-take standardized English test. Just do well enough to pass minimum requirement set by the university, which nowadays has many options for you to select from.
+take standardized English test. Just do well enough to pass minimum requirement set by the university, which has many options for you to select from.
 

### 2023-11-19 demystify.tex
updates

@@
 \usepackage[T1]{fontenc}
-\usepackage[margin=1.5in]{geometry}
+%\usepackage[margin=1.5in]{geometry}
 \usepackage{soul}
@@
 }
+
+\newtcolorbox{mydomesticbox}{
+enhanced,
+boxrule=0pt,frame hidden,
+borderline west={2pt}{0pt}{red!75!black},
+colback=blue!10!white,
+sharp corners
+}
+
+\newenvironment{domesticbox}[1][]{
+\small
+    \begin{mydomesticbox}
+    {\small \textbf{#1}}
+ }{
+   \end{mydomesticbox}
+}
+
 \renewcommand{\figurename}{Fig.}
@@
 
+\begin{domesticbox}[Domestic students:] If you're a domestic student, you actually have several advantages in your application.
+First, we already know about your school (see \S\ref{sec:your-school}), and you are also more familiar with the US education system and culture.
+You also have more opportunities for funding, e.g., through government scholarships for US citizens and residents.
+Finally, compared to international students, PhD applications in CS from domestic students are also few and that can help your case.
+\end{domesticbox}
+
 %% Vu: *what's a PhD?*  This [[https://matt.might.net/articles/phd-school-in-pictures/][series of pictures]] from [[https://matt.might.net][Matt Might]] illustrates what a PhD means.
@@
 \begin{center}
-\includegraphics[width=0.6\linewidth]{c6.png}
+\includegraphics[width=0.6\linewidth]{files/c6.png}
 \end{center}
@@
     \begin{center}
-      \includegraphics[scale=0.5]{c2.png}
+      \includegraphics[scale=0.5]{files/c2.png}
     \end{center}
@@
 \end{commentbox}
+
+\paragraph{If you do not get admitted} If you do not get admitted to any schools (or don't want to go to the ones that you got it), try again next time.
+Graduate admission can involve randomness and noise.
+Don't bother asking for feedback, you will not likely get any.
+In the meantime, you can work on improving your profile, e.g., publish more papers, get more research experiences, etc.
+
+You can also consider applying to MS programs, which are typically easier to get in (but you need to pay).
+If you get in an MS at a school of your choice, you can contact professors to work with them.
+If you do well, you can ask the professor to support you to convert to PhD.
+This is a common path for many students.
 
@@
    \begin{center}
-    \includegraphics[scale=0.5]{c1.png}
+    \includegraphics[scale=0.5]{files/c1.png}
    \end{center}
@@
 In any case, you should contact the professor and talk to them.
-For example, you can (cold) email them and ask about their research and if they are taking new students and schedule appointments to chat with them (office hrs are often the best time).
-See \S\ref{sec:contact} for more details on contacting professors.
-
-Another very effective way is taking graduate level courses in the topics you are interested in (remember: you have those 2 years to explore).
+For example, you can (cold) email them and ask about their research and if they are taking new students (\S\ref{sec:contact}) and schedule appointments to chat with them (office hrs are often the best time).
+
+Another effective way is taking graduate level courses in the topics you are interested in (remember: you have those 2 years to explore).
 Professors teaching these \emph{special topics courses} and \emph{research seminars} typically are recruiting students---giving you even a higher chance.
@@
       \begin{center}
-        \includegraphics[scale=0.4]{c8.png}
+        \includegraphics[scale=0.4]{files/c8.png}
       \end{center}
@@
 
+\begin{center}
+\includegraphics[scale=0.5]{files/emails.png}
+\end{center}
+
 Faculty received many "cold" e-mails from international students seeking for admission, TA, and RA.
 Most of the time, we ignore these emails, but in some rare occasions we do answer them.
-So how to write an email that can get our attention?
+So why do we ignore some but reply to others and how to write an email that get our attention?
 
@@
 \begin{center}
-    \includegraphics[scale=0.2]{alpinist-climbing-peak-mountain-comic-hand-drawn-vector-illustration.jpg}
+    \includegraphics[scale=0.2]{files/alpinist-climbing-peak-mountain-comic-hand-drawn-vector-illustration.jpg}
   \end{center}
@@
 \begin{center}
-\includegraphics[scale=0.3]{c4a.png}
+\includegraphics[scale=0.3]{files/c4a.png}
 \end{center}
@@
 \begin{center}
-\includegraphics[scale=0.6]{c5.png}
+\includegraphics[scale=0.6]{files/c5.png}
 \end{center}
@@
       \begin{center}
-        \includegraphics[scale=0.3]{c6.png}
+        \includegraphics[scale=0.3]{files/c6.png}
       \end{center}
@@
 
+\chapter{References}
+
 \chapter{History and Acknowledgement}\label{sec:ack}

### 2023-11-24 demystify.tex
updates

@@
 \usepackage[T1]{fontenc}
-%\usepackage[margin=1.5in]{geometry}
+\usepackage[margin=1.5in]{geometry}
 \usepackage{soul}
-%\usepackage[small,compact]{titlesec} %very powerful
+% \usepackage[small,compact]{titlesec} %very powerful
 \usepackage[most]{tcolorbox}
@@
     \vfill%
-    %\hrulefill\par
+    % \hrulefill\par
     {\Large \@author\par}%
@@
 \makeatother
-%\author{ThanhVu H. Nguyen}
-%\title{TITLE OF BOOK}
-%\date{}
+% \author{ThanhVu H. Nguyen}
+% \title{TITLE OF BOOK}
+% \date{}
 
@@
 Having been involved in PhD admissions for many years, I've
-realized that many \textbf{international students} (and even domestic ones) lack a clear understanding of
+realized that many \textbf{international students}, especially those in  smaller countries or less well-known universities, lack a clear understanding of
 the Computer Science PhD admission process at US universities. This confusion not only
@@
 
-%though \emph{very} top schools could be very selective, e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU
+% though \emph{very} top schools could be very selective, e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU
 So I want to share some details about the PhD admission process and advice for those who are interested in applying for a \textbf{PhD in Computer Science in the US}.
@@
 \chapter{Summary}\label{sec:summary}
-Here we summarize the main points of the document. This allows you to quickly skim through the document and decide which parts you want to read in detail.
+
+Here, we outline the main points of the document. This gives you an overview to decide which specific topics you want to explore more thoroughly.
+
 \begin{enumerate}
@@
 
-%\didi{Is there more information on typical strengths and common weaknesses of applications.
+% \didi{Is there more information on typical strengths and common weaknesses of applications.
 This is especially useful to sophomore and junior students as they still have time to work on those strengths.}
@@
 
-%\begin{commentbox}[Hung]
+% \begin{commentbox}[Hung]
 %   The application fee varies between schools, and is about \$70--\$100.
@@
 While a blanket waiver might not be feasible, universities should extend the eligibility of fee waiver applications to more countries.
-%\end{commentbox}
+% \end{commentbox}
 
@@
 
-It's important to aim for the top venues in your field. You can
-find top CS conferences at places such as \href{https://csrankings.org}{CSRankings}, which ranks CS programs based on how their faculty publish at top \emph{conferences} (see more in \S\ref{sec:ranking}).
-Local conferences and non-English journals or conferences do
+However, it is understandable that many students do not have the opportunities to publish in top places.
+Thus, general writings, even those under submissions or rejected, would still be good (and a lot better than nothing).
+But be sure to upload your papers with your application and talk about them in your statement (see \S\ref{sec:research-statement}).
+Note that local conferences and non-English journals or conferences do
 not carry as much weight since their quality is often unknown to US faculty.
-
-However, it is understandable that many students do not have the opportunities to publish in top places.
-Thus, general confs/journals, even those under submissions or rejected, would suffice.
-But be sure to upload your papers with your application and talk about them in your statement (see \S\ref{sec:research-statement}).
+However, if you have published in such places, you should still mention them in your statement and explain why they are good.
+
+% In CS, conferences are important and top conferences are prestigious.
+You can find top CS conferences at places such as \href{https://csrankings.org}{CSRankings}, which ranks CS programs based on how their faculty publish at top \emph{conferences} (see more in \S\ref{sec:ranking}).
 
@@
 
-CS PhD programs will require at least \textbf{two LORs}. Having a letter from an internationally recognized researcher can greatly strengthen your application. However, obtaining such letters
+Most CS PhD programs will require at least \textbf{two LORs}. Having a letter from an internationally recognized researcher can greatly strengthen your application. However, obtaining such letters
 can be challenging for international students, who might not have much interactions with such experts.
@@
 
-For Vietnamese students, it's worth mentioning about the \href{https://vef2.org/}{VEF2.0 program}, which has helped many good students in gaining acceptance to top PhD programs in the US.
-VEF2.0 follows an interesting model where US faculty members from leading institutions are invited to conduct rigorous interviews with VEF students and subsequently provide reference letters on their behalf.
-Despite the limited interaction between the interviewer and the interviewee (primarily confined to the interview itself), the impact of these reference letters can be positive as they have helped many students getting into top PhD programs in the US.
+\paragraph{Always Waive your Right to Review the LORs}  Choosing not to look at a reference letter is pretty standard in job and school applications.
+When you waive your right to see the letter, it adds a layer of trust, showing you're confident in your choice of referees and that you're not trying to twist their words.
+It's also about keeping things open and honest between you and your letter writters, and encourages them to be real about your strengths and qualifications.
+Plus, it keeps things private.
+
+Reviewers might raise concerns (a red flag) about  a letter that is not waived, e.g., they will think that if you do not trust your letter writers, then you should find someone else to write for you.
+In short, it's a standard practice and a way of keeping things straightforward and respectful in the whole recommendation game.
 
@@
 
+\paragraph{VEF 2.0} For Vietnamese students, it's worth mentioning about the \href{https://vef2.org/}{VEF2.0 program}, which has helped many good students in gaining acceptance to top PhD programs in the US.
+VEF2.0 follows an interesting model where US faculty members from leading institutions are invited to conduct rigorous interviews with VEF students and subsequently provide reference letters on their behalf.
+Despite the limited interaction between the interviewer and the interviewee (primarily confined to the interview itself), the impact of these reference letters can be positive as they have helped many students getting into top PhD programs in the US.
+
 \section{Statement of Purpose (SOP)}\label{sec:research-statement}
@@
 \begin{center}
-      \includegraphics[scale=0.5]{files/c2.png}
+  \includegraphics[scale=0.4]{files/c2.png}
 \end{center}
@@
 statement of purpose (SOP) or personal statement! So write it well because it could make a big difference.
-In your statement, you have the opportunity to make  make your application stand out and unique, which can make you \emph{fit} the CS program you're applying to, even if you don't have very strong research experience.
+In your statement, you have the opportunity to make your application stand out and unique, which can make you \emph{fit} the CS program you're applying to, even if you don't have very strong research experience.
 A well-written LOR also shows that you can communicate, which is very important in research, and that you can effectively teach and communicate with students, which is important for TA funding (see \S\ref{sec:funding}).
-%is important because if you need (GTA) funding, it will provide evidence
-%that you can teach and communicate with students.
+% is important because if you need (GTA) funding, it will provide evidence
+% that you can teach and communicate with students.
 
@@
 
+% \subsection{Outline of a LOR}\label{sec:lor-outline}
+
 \section{Your School}\label{sec:your-school}
@@
 Having good grades is important, but unless your school is well-known, having top grades or rankings
-usually will not help. This is simply because we cannot evaluate them.
+usually will not help because we cannot evaluate them.
 
@@
 
-\paragraph{If you do not get admitted} If you do not get admitted to any schools (or don't want to go to the ones that you got it), try again next time.
+\paragraph{If you do not get admitted} If you do not get admitted to any schools (or don't want to go to the ones that admit you), try again next time.
 Graduate admission can involve randomness and noise.
 Don't bother asking for feedback, you will not likely get any.
-In the meantime, you can work on improving your profile, e.g., publish more papers, get more research experiences, etc.
+In the meantime, you can work on improving your profile, e.g., get more research experiences, publish more papers, improve your connections for better LoR writers, etc.
 
 You can also consider applying to MS programs, which are typically easier to get in (but you need to pay).
-If you get in an MS at a school of your choice, you can contact professors to work with them.
+If you get in an MS program at a school of your choice, you can contact professors to work with them.
 If you do well, you can ask the professor to support you to convert to PhD.
@@
 
-%\subsection{R1, R2, ...}
-
-\subsection{Selecting and Ranking Schools for CS}
+% \subsection{R1, R2, ...}
+
+\subsection{Selecting and Ranking Schools for CS}\label{sec:selecting-ranking-schools}
 \begin{center}
@@
 \section{Choosing a Professor}
-
-Now we discuss about finding advisers.
-
-\subsection{Finding the right adviser}
 Obviously, there is no one-size-fit-all answer to this question.
@@
 
-\begin{commentbox}[Xiaokuan:]
-    Whether the student's research interest matches that of the adviser is very important;
-    if there is a mismatch,
-    either the student or the adviser has to make compromises,
-    which often leads to disagreements or conflicts. 
-    IMO, the adviser should be the one who \emph{guides}  students to do research while allowing students to pursue their own interest,
-    instead of \emph{dictating} their research.
-    \end{commentbox}
+\subsection{Finding the right adviser}
 
@@
 For example, you can (cold) email them and ask about their research and if they are taking new students (\S\ref{sec:contact}) and schedule appointments to chat with them (office hrs are often the best time).
+
+\begin{commentbox}[Xiaokuan:]
+  Whether the student's research interest matches that of the adviser is very important;
+  if there is a mismatch,
+  either the student or the adviser has to make compromises,
+  which often leads to disagreements or conflicts.
+  IMO, the adviser should be the one who \emph{guides}  students to do research while allowing students to pursue their own interest,
+  instead of \emph{dictating} their research.
+\end{commentbox}
 
@@
 
-%It is worth noting that in many cases,  an application could be rejected by most reviewers, however have a "champion", who doesn't even have to be in the Admission comittee, that strongly advocate for your profile.
+% It is worth noting that in many cases,  an application could be rejected by most reviewers, however have a "champion", who doesn't even have to be in the Admission comittee, that strongly advocate for your profile.
 
@@
 \begin{center}
-\includegraphics[scale=0.6]{files/c5.png}
+  \includegraphics[scale=0.5]{files/c5.png}
 \end{center}
 
-If you're reaching out to a professor for the first time,  address them as either Prof.
+If you're reaching out to a professor for the first time,  address them as Prof.
 Lastname or Dr.
@@
 
-Note that in some universities the formal title \emph{Dr.} Lastname is preferred.
+Note that in some universities the formal title Dr.
+Lastname is preferred over Prof.
 You just need to observe and follow the conventions at your particular institution.
@@
 
-%\section{Having fun during a PhD?}
-%PhD students \emph{and faculty} probably find it amusing about the notion that students, especially international ones, can genuinely enjoy their PhD studies.
+% \section{Having fun during a PhD?}
+% PhD students \emph{and faculty} probably find it amusing about the notion that students, especially international ones, can genuinely enjoy their PhD studies.
 In fact, after reading posts after posts on VietPhD.org on how PhD students are commonly mistreated, stressed, it seems being miserable is a norm during a PhD study.
 
-%There are many advice on surviving PhD that you can follow. But here I just list a few that works for me and what I advice my students to do.\tvn{TODO}
+% There are many advice on surviving PhD that you can follow. But here I just list a few that works for me and what I advice my students to do.\tvn{TODO}
 
 \section{Will I be miserable during my a PhD?}\label{sec:happy}
-Many posts from VietPhD, a Facebook group for Vietnamese PhD students, are about how PhD students are mistreated, stressed, and miserable.
+There are many stories on how students are mistreated, stressed, and miserable.
 Certainly issues including bad relationships with professors, conflicts with co-authors and lab mates, and toxic working environment, feeling left out or discriminated (e.g., because you're an international student) \emph{do} exist, and it is good to be aware of those.
@@
 
-%\paragraph{Diversity} US universities prioritize diversity and inclusion.
+% \paragraph{Diversity} US universities prioritize diversity and inclusion.
 Students need to respect and appreciate varied opinions, backgrounds, and cultures.
@@
 
-Tab.~\ref{tab:ranking} lists the top 50 CS programs in the US from \href{https://www.csrankings.org}{CSRankings.org}, a ranking system  based on top CS conferences.
 \begin{table}
@@
 
+Tab.~\ref{tab:ranking} lists the top 50 CS programs in the US from \href{https://www.csrankings.org}{CSRankings.org}, a ranking system  based on top CS conferences.
+
 % \section{Additional Resources}
@@
 \paragraph{History} This document was conceived during a lunch with Craig Yu at GMU.
-We talked on about why GMU were not able to attract good Vietnamese and other international students, despite having a much stronger CS program than many schools that these students want to go to (part of the reason is described in \S\ref{sec:ranking}).
+We talked on about why GMU were not able to attract good Vietnamese and other international students, despite having a much stronger CS program than many schools that these students want to go to (part of the reason is described in \S\ref{sec:selecting-ranking-schools}).
 We wished there were a way for international students to know about the US PhD programs (as well as for US faculty to understand more about international students and therefore have better chance of recruiting and working with them).
@@
 However, most active participant are students in non-CS fields or not in US.
-Like ChatGPT, their answers to CS PhD program in the US could be informative and helpful, but unfortunately not always accurate and sometimes leading to more confusion.
+Like ChatGPT, their answers to questions on CS PhD program in the US could be informative and helpful, but unfortunately not always accurate and sometimes leading to more confusion.
 So I thought it would be useful to have a document that is specific to CS PhD programs in the US from an insider prospective.
@@
 I have been in the PhD admission process at GMU and UNL for many years.
-Currently I serve as the program director of the Masters program in Software Engineering at GMU (thus also have some experience with the MS admission process--which is quite different than PhD).
-I got my undergraduate and MS degrees from Penn State.
+Currently I serve as the program director of the MS program in Software Engineering at GMU (thus also have some experience with the MS admission process--which is quite different than PhD).
+%I got my undergraduate and MS degrees from Penn State.
 I completed my PhD at the University of New Mexico-Albuquerque and a postdoc at the University of Maryland-College Park.

### 2023-11-24 demystify.tex
Updates from Overleaf

--- 2023-11-09 demystify.tex
@@
 \end{commentbox}
+\begin{commentbox}[Deepak:] 
+If an applicant is anxious about their school not being known outside their country, they can provide information about their school and department, with independent sources where such informationa can be verified.
+\end{commentbox}
 
@@
 See \S\ref{sec:contact} for more details on contacting professors.
+\begin{commentbox}
+    [Deepak:]
+    In my view, cold emails are not welcome by most faculty members and should be avoided.
+However, if one is already admitted to a program in some department, by all means, send an email to the faculty you may be interested in working with, but do mention right at the beginning that you are already admitted to the program as well as several other universities.
+State specific areas (preferably specific topics-ML, robotics instead of AI).
+\end{commentbox}
 
@@
 For example, I prefer that my students and colleague call me Vu. Some students call me \emph{Dr. Vu}, which I find a bit amusing but am totally fine with it.
+
+\begin{commentbox}[DK:]
+I was amused to read this as if I recall correctly, you never called me by my first name when you were at UNM. You always called me Prof. And, many times, I would jokingly call you back as Prof. Vu.
+\end{commentbox}
 
@@
 Many students and professors maintain lifelong relationships, staying in touch through cards, emails, and calls, even after their academic journey ends.% These connections are genuine and meaningful.
+
+\begin{commentbox}
+    [DK:] Here are a few items you may consider addressing: putting international students in touch with other students from their countries, assuring them that they would be picked up from airports upon their arrival and that their initial stay will be taken care of.
+Most universities have other resources for these, but it is worth mentioning that they would be taken care of upon arrival and can get help during the transition phase.
+Learning to cook was a big deal when I arrived over 50 years ago-August 1973.
+But things have changed as one can find many ethnic food places, a big contrast from 1973, when there were two Indian restaurants in Greater Boston area.
+\end{commentbox}
 

### 2023-11-24 demystify.tex

--- 2023-11-24 demystify.tex
@@
 
+\begin{commentbox}
+    [Deepak:]
+    In my view, cold emails are not welcome by most faculty members and should be avoided.
+However, if one is already admitted to a program in some department, by all means, send an email to the faculty you may be interested in working with, but do mention right at the beginning that you are already admitted to the program as well as several other universities.
+State specific areas (preferably specific topics-ML, robotics instead of AI).
+\end{commentbox}
+
 Another effective way is taking graduate level courses in the topics you are interested in (remember: you have those 2 years to explore).

### 2023-11-24 demystify.tex
updates

@@
 \begin{commentbox}[Deepak:] 
-If an applicant is anxious about their school not being known outside their country, they can provide information about their school and department, with independent sources where such informationa can be verified.
+If an applicant is anxious about their school not being known outside their country, they can provide information about their school and department, with independent sources where such information can be verified.
 \end{commentbox}
@@
 This can be an issue for students in many top international universities where the competition is so high that very good students students can still have low rankings from these schools (and be overlooked by the admission committee).
-So what to do with this? well, same as \hyperref[sec:your-school]{before}, e.g., put a note about this in your statement and ask your LoR writers to talk about it.
+So what to do with this? as mentioned in \S\ref{sec:your-school} you can put a note about this in your statement and ask your LoR writers to talk about it.
 
@@
 Another effective way is taking graduate level courses in the topics you are interested in (remember: you have those 2 years to explore).
-Professors teaching these \emph{special topics courses} and \emph{research seminars} typically are recruiting students---giving you even a higher chance.
-Do well in the class, answer questions, talk to the profs after classes, etc---being stand out.
+Professors teaching these \emph{special topics courses} and \emph{research seminars} might be recruiting students---giving you even a higher chance.
+Do well in the class, answer questions, talk to the prof after classes, etc---being stand out.
 Many professors, including myself, prefer taking in new students this way.
@@
 This can be informal (no credit) and takes place during the summer or winter break.
-For example, I work with multiple students, many of whom are undergrads, this way.
+For example, I do this with multiple students, many of whom are undergrads.
 Many will drop because they find they don't like my research, but some will stay and become my PhD students.
@@
   \textbf{Vu}: \S\ref{sec:contact} shows how to contact a professor for research opportunities.
-Many probably will say no as they do not have the bandwidth to take on random students, but some may say yes if they see potential fit.
+Many probably will say no (or do not reply) as they do not have the bandwidth to take on random students, but some may say yes if they see potential fit.
 \end{commentbox}
@@
   I was amused to read this as if I recall correctly, you never called me by my first name when you were at UNM. You always called me Prof. And, many times, I would jokingly call you back as Prof. Vu.
-\end{commentbox}
+  \tcblower
+  \textbf{Vu}: Yes, for some reason I enjoy addressing you as ``Prof.'' (without appending a last or first name).
+I call Steph and everyone else by their their first names.
+The use of Prof.
+Vu may have foreshadowed my future in academia.
+\end{commentbox}
+
+Certainly, there's a peculiar satisfaction in addressing you as "Prof." without appending a last or first name.
+While I refer to Steph and everyone else by their first names, perhaps addressing Prof.
+Vu in this manner hinted at my future in academia.
 
@@
 While people can be straightforward and direct, they are also friendly and informal.
-You can call your professor by their first name (\S\ref{sec:address}), disagree with them and argue (and can actually gain respect doing so), seek their help (even on personal matters), come to their houses for parties or gathering, and give them small thoughtful gifts that they put on their desks.
+You can call your professor by their first name (\S\ref{sec:address}), disagree with them and argue (and gain respect doing so), seek their help (even on personal matters), come to their houses for parties or gathering, and give them small thoughtful gifts that they put on their desks.
 Many students and professors maintain lifelong relationships, staying in touch through cards, emails, and calls, even after their academic journey ends.% These connections are genuine and meaningful.

### 2023-11-24 demystify.tex
updates

@@
 While I am not an international student, many of my students and collaborators are international.
-I also mentor many students from Vietnam.
-I also have many colleague and friends who were international students and have heard their stories.
+I also mentor multiple students from Vietnam.
+I also have various colleague and friends who were international students and have heard their stories.
 I hope to capture their challenges and experiences in this document so that it can be useful for prospective international students.
@@
 Profs. Craig Yu (GMU), Hakan Aydin (GMU), 
-Xiaokuan Zhang (GMU), and Hung Le (UMass) provided valuable input in the early version.
+Xiaokuan Zhang (GMU), Hung Le (UMass), and Deepak Kapur (UNM) provided valuable input in the early version.
 Other GMU faculty members also have provided feedback and contributions.

### 2023-11-24 demystify.tex
updates

@@
 
-\begin{commentbox}
-    [Deepak:]
-    In my view, cold emails are not welcome by most faculty members and should be avoided.
-However, if one is already admitted to a program in some department, by all means, send an email to the faculty you may be interested in working with, but do mention right at the beginning that you are already admitted to the program as well as several other universities.
-State specific areas (preferably specific topics-ML, robotics instead of AI).
-\end{commentbox}
-
 Another effective way is taking graduate level courses in the topics you are interested in (remember: you have those 2 years to explore).
@@
 \begin{center}
-  \includegraphics[scale=0.5]{files/emails.png}
+  \includegraphics[scale=0.4]{files/emails.png}
 \end{center}
@@
 to ask about \emph{research opportunities}, or \emph{GTA/GRA} support, then \emph{yes}, I believe you should---it is \emph{worth it}.
-However, you need to do it the \emph{right way}.
-
-First, read the prof's website, see if they say something about contacting them. Many profs. explicitly indicate how prospective students should contact them (e.g., using specific email subjects). 
+However, you need to put effort into it and really do it the \emph{right way}.
+
+First, read the prof's website, see if they say something about contacting them.
+Many profs.
+explicitly indicate how prospective students should (or should not) contact them (e.g., using specific email subjects).
 In general, the best way to catch the prof.'s attention is to \emph{customize your email} for them.
@@
 Then send them an email talking how/why their work would match yours.
-In contrast, if you write a generic email that can be sent to multiple professors (e.g., if you just change some names and keywords in the email), you will not get a response.
+In contrast, if you write a generic email that can be sent to multiple professors (e.g., if you just change some names and keywords in the email or copy and paste paper titles), you will not get a response.
 
@@
   the chance of getting a reply is close to zero.
+\end{commentbox}
+
+\begin{commentbox}
+    [Deepak:]
+    In my view, cold emails are not welcome by most faculty members and should be avoided.
+However, if one is already admitted to a program in some department, by all means, send an email to the faculty you may be interested in working with, but do mention right at the beginning that you are already admitted to the program as well as several other universities.
+State specific areas (preferably specific topics-ML, robotics instead of AI).
 \end{commentbox}
@@
 
-While I am not an international student, many of my students and collaborators are international.
-I also mentor multiple students from Vietnam.
-I also have various colleague and friends who were international students and have heard their stories.
-I hope to capture their challenges and experiences in this document so that it can be useful for prospective international students.
+While I am not an international student, many of my students and collaborators are.
+I also mentor multiple students from Vietnan, and have various colleague and friends who were international students.
+I have heard their stories, and hope to capture their challenges and experiences in this document so that it can be useful for prospective international students.
 

### 2023-11-24 demystify.tex
updaes

@@
 
-Second, I believe that applying to a good US university \emph{should not} be any
-harder than at schools in other countries.
-If you believe you have a chance in other countries, e.g., South Korea, Singapore, Germany, UK, Japan and Australia, then you will surely have a chance in the US as well.
+Second, applying to a good US university \emph{should not} be any harder than at schools in other countries.
+In fact, it might even be more flexible since CS PhD in the US \emph{do not} require having an MS (or having to know a topic/proposal/adviser in advance).
+If you believe you have a chance in other countries, e.g., South Korea, Singapore, Germany, UK, Japan, and Australia, then you will surely have a chance in the US as well.
 \S\ref{sec:non-us-differences} compares CS PhD study in the US to other countries.

### 2023-11-24 demystify.tex
updates

@@
 
-The primary focus of the admissions committee is to \textbf{evaluate your background and interest in research} since a PhD in Computer Science and determine if you would \textbf{fit into} the program.
+The primary focus of the admissions committee is to \textbf{evaluate your background and interest in research} (a PhD in Computer Science is a research degree!), and determine if you would \textbf{fit into} the program.
 To evaluate your profile, we consider
@@
 If not, then you should ask someone else.
-For example, if a student I don't know well ask me to write letter for them, I will explicitly tell them I don't know them that well to write about them in details, and such a short, generic, and weak letter will hurt their chance.
+For example, if a student I don't know well ask me to write letter for them, I will explicitly tell them I don't know them that well to write much about them, and such a short, generic, and weak letter will not help their profile.
 
@@
 
-Reviewers might raise concerns (a red flag) about  a letter that is not waived, e.g., they will think that if you do not trust your letter writers, then you should find someone else to write for you.
+Reviewers might raise concerns (a red flag) about  a letter that is not waived, e.g., if you do not trust your letter writers, then you should find someone else to write for you.
 In short, it's a standard practice and a way of keeping things straightforward and respectful in the whole recommendation game.

### 2023-11-25 demystify.tex
update

@@
 
-\begin{domesticbox}[Domestic students:] If you're a domestic student, you actually have several advantages in your application.
+\begin{domesticbox}[Domestic students:] If you're a domestic student, you have several advantages in your application.
 First, we already know about your school (see \S\ref{sec:your-school}), and you are also more familiar with the US education system and culture.
 You also have more opportunities for funding, e.g., through government scholarships for US citizens and residents.
-Finally, compared to international students, PhD applications in CS from domestic students are also few and that can help your case.
+Finally, compared to international students, PhD applications in CS from domestic students are few and that can help your case.
 \end{domesticbox}
@@
 
-\section{Graduating Assistantship (TA/RA)}
-The most common type of funding is \textbf{graduate assistanship}, which is either TA or RA.
+\section{Graduate Assistantship (TA/RA)}
+The most common type of funding is \textbf{graduate assistantship}, which is either TA or RA.
 Both TA and RA come with tuition waiving (you don't have to pay tuition), health insurance (this takes care of your insurance, which is a must have in the US), and most importantly, your stipend (i.e., your salary).
@@
 Students often complain their stipend being too low, but it is actually not bad and you can live comfortably with it.
-In many cases, it is enough to support your spouse and kids (many CS PhD students have their spouse and kids with them).
+In many cases, it is also enough to support your spouse and kids (many CS PhD students have their family with them).
 So don't worry too much about stipend.
 If you're admitted to a good CS PhD program, you will be fine.
-A good school would know that it has to be competitive, otherwise it will loose good students to other schools.
+A good school would know that it has to be competitive to attract students.
 For example, at GMU, every year we discuss about improving the benefits, and especially stipend, for our graduate students.
@@
 
-Choosing a school and an adviser is clearly one of the most important things in your mind when you apply and especially when you get admitted.
+Choosing a school and an adviser is clearly among the most important things in your mind when you apply and especially when you get admitted.
 This is further complicated due to cultural differences and unfamiliarity of  international students to the US higher education system.
@@
 
-Ultimately, choose one that fits you the most by communicating with them, taking their courses, meeting them, asking them questions, and even talking to their current students.
+Ultimately, choose a prof.
+that fits you the most by communicating with them, taking their courses, meeting them, asking them questions, and even talking to their current students.
 Indeed all this takes a lot of time and effort, but remember, you will be working with this person for 4-5 years, so it is important to try to find the right one.
@@
 
-\subsection{Tenured, Tenure Track, Teaching,  Assistant Profs., Associate Profs}\label{sec:faculty-types}
+\subsection{Tenured, Tenure Track, Teaching,  Assistant Profs., Associate Profs, and O
+ther Fancy Academic Titles}\label{sec:faculty-types}
 
@@
 
-% \begin{commentbox}[Thanh:]
-% \end{commentbox}
-
 \section{How long does it take to complete the CS PhD program?}\label{sec:time}
@@
 
-\section{Differences between PhD in the US and other countries}\label{sec:non-us-differences}
+\section{Differences between PhD in the US and Other Countries}\label{sec:non-us-differences}
+
 This summarizes the main differences between CS PhD in the US and other countries. % This is based on my experience and what I heard from others who did PhD in other countries.
@@
 
-\emph{Work-life balance}: I believe that PhD students are often overworked in the US compared to other countries, e.g., in Europe.
+\emph{Work-life balance}: I believe that PhD students in the US are overworked compared to other countries, e.g., in Europe.
 This is partly due to the longer PhD program and that US PhD students are often paid through TA, which requires them to do TA in addition to their own research.
-In contrast, PhD students in other countries are often paid through fellowships, which do not require them to do TA work.
+In contrast, PhD students in other countries are often paid through fellowships, which do not require them to do TA.
 
@@
 If you're reaching out to a professor for the first time,  address them as Prof.
-Lastname or Dr.
+or Dr.
 Lastname.

### 2023-11-25 demystify.tex
update

@@
 
+\section{References}
+\begin{itemize}
+    \item \href{https://parentheticallyspeaking.org/articles/us-cs-phd-faq/}{Getting a Computer Science PhD in the USA} by Shriram Krishnamurthi
+\end{itemize}
+
 \chapter{How is Your Application Evaluated?}\label{sec:evalapps}
@@
 
+\subsection{References}
+\begin{itemize}
+\item \href{http://www.pl-enthusiast.net/2022/10/03/how-to-write-a-grad-school-personal-statement/}{How to Write a Grad School Personal Statement} by Mike Hicks, 10/03/22
+\end{itemize}
 % \subsection{Outline of a LOR}\label{sec:lor-outline}
@@
 
-\subsection{Tenured, Tenure Track, Teaching,  Assistant Profs., Associate Profs, and O
-ther Fancy Academic Titles}\label{sec:faculty-types}
+\subsection{Tenured, Tenure Track, Teaching,  Assistant Profs., Associate Profs, and Other Academic Titles}\label{sec:faculty-types}
 
@@
 \end{commentbox}
+
+\subsubsection{References}
+\begin{itemize}
+    \item \href{https://yonatanbisk.com/emailing_professors.html}{Emailing Professors} by Yonatan Bisk
+\end{itemize}
 

### 2023-11-26 demystify.tex
update

@@
 \begin{itemize}
-\item \href{http://www.pl-enthusiast.net/2022/10/03/how-to-write-a-grad-school-personal-statement/}{How to Write a Grad School Personal Statement} by Mike Hicks, 10/03/22
+\item \href{http://www.pl-enthusiast.net/2022/10/03/how-to-write-a-grad-school-personal-statement/}{How to Write a Grad School Personal Statement} by Mike Hicks
 \end{itemize}
@@
 that fits you the most by communicating with them, taking their courses, meeting them, asking them questions, and even talking to their current students.
-Indeed all this takes a lot of time and effort, but remember, you will be working with this person for 4-5 years, so it is important to try to find the right one.
+Doing these takes a lot of time and effort, but remember, you will be working with this person for 5+ years, so it is important to try to find the right one.
 
@@
 
+\subsubsection{References}
+\begin{itemize}
+    \item \href{https://www.cs.columbia.edu/wp-content/uploads/2019/03/Get-Advisor.pdf}{The Definitive "what do I ask/look for" in a PhD Advisor Guide}
+\end{itemize}
 \subsection{Tenured, Tenure Track, Teaching,  Assistant Profs., Associate Profs, and Other Academic Titles}\label{sec:faculty-types}
@@
 \begin{itemize}
-    \item \href{https://yonatanbisk.com/emailing_professors.html}{Emailing Professors} by Yonatan Bisk
+    \item \href{https://yonatanbisk.com/emailing_professors.html}{A Note about Emailing Professors} by Yonatan Bisk
 \end{itemize}

### 2023-11-27 demystify.tex
Updates from Overleaf

@@
 
+\section{Preparing Materials}
+\subsection{Sharing spreadsheets}
+Many students use a spreadsheet to help them and their letter writers to keep track of their applications.
+You can create a personal and a shared spreadsheet but a simplest one is just a shared one (e.g., using Google sheet).
+Here are some information to put on the sheet.
+\begin{itemize}
+\item \textbf{Your info}: Full name, email, phone, link to website/CV).  This is helpful for the writers just in case they want to quickly get some information about you.
+\item \textbf{Applications details}: University, Dept, Application System URL, Submission Deadlines, Application Status (e.g., submitted,  rejected, wait-listed, accepted)
+\item \textbf{LoRs}: Writer 1, Writer 2, Writer 3.
+If this is a shared document, you might want to omit names and just use writer X.
+Under each is their status, e.g., sent/not yet/reminder needed, etc.
+\end{itemize}
+
 \chapter{Getting Admitted}\label{sec:accepted}

### 2023-11-27 demystify.tex
update

@@
 \setsecnumdepth{subsection}
+\setcounter{tocdepth}{3}
 
@@
 
-\section{References}
+\section*{References}
 \begin{itemize}
@@
 
-\subsection{References}
+\subsection*{References}
 \begin{itemize}
@@
 
-\section{Choosing a Professor}
+\section{Choosing an Advisor}
 Obviously, there is no one-size-fit-all answer to this question.
@@
 Ultimately, choose a prof.
-that fits you the most by communicating with them, taking their courses, meeting them, asking them questions, and even talking to their current students.
-Doing these takes a lot of time and effort, but remember, you will be working with this person for 5+ years, so it is important to try to find the right one.
+that fits you the most by communicating with them, taking their courses, meeting and asking them questions, and even talking to their current students.
+It will take time and effort, but remember, you will be working with this person for 5+ years, so it is important to try to find the right one.
 
@@
 
-\subsubsection{References}
+\subsubsection*{References}
 \begin{itemize}
@@
 
-\chapter{References}
+%\chapter{References}
 

### 2023-11-29 demystify.tex
updates

@@
 In any case, you should contact the professor and talk to them.
-For example, you can (cold) email them and ask about their research and if they are taking new students (\S\ref{sec:contact}) and schedule appointments to chat with them (office hrs are often the best time).
+For example, you can (cold) email them and ask about their research and if they are taking new students (\S\ref{sec:contact}), and schedule appointments to chat with them (office hrs are often the best time).
 
@@
 \end{itemize}
-\subsection{Tenured, Tenure Track, Teaching,  Assistant Profs., Associate Profs, and Other Academic Titles}\label{sec:faculty-types}
-
-Not every faculty in the CS program can serve as your formal adviser.
+\subsection{Types of Faculty: Who can serve as a PhD adviser?}\label{sec:faculty-types}
+
+Not every faculty can serve as your formal adviser.
 Let's try to understand the different types of faculty and their roles.
@@
 In contrast, \emph{teaching} (or instructional faculty or professor of practice) mainly focus on teaching.
-They typically teach 3-4 classes per semester and \emph{do not} have research responsibilities (e.g., they do not have to worry much about publishing papers, obtaining grants).
+They typically teach 3--4 classes per semester (which is quite a lot) and \emph{do not} have research responsibilities (e.g., they do not have to worry much about publishing papers, obtaining grants).
 They can mentor Ph.D.

### 2023-12-05 demystify.tex
updates

@@
 To evaluate your profile, we consider
-the following key indicators, listed in order of importance.
+the following key indicators.
 
@@
 The most effective evidence of research ability is having \textbf{published papers in reputable international journals or conferences}.
-Having published good papers is a sign that the applicant has successfully involved in research.
+Having papers published is a sign that the applicant has successfully involved in research.
+
+\begin{commentbox}[Vu:]
+    Many international students mention Scopus Q1, which consists of various journals from IEEE, Elsevier, and many other publishers.
+I don't know/recognize many of journals listed in Scopus Q1.
+This might be something to be mindful of, as \textbf{CS} faculty might not be too familiar with Scopus or journals listed there, so devote some part in your statement to discuss the significance of your papers.
+\end{commentbox}
 
 However, it is understandable that many students do not have the opportunities to publish in top places.
-Thus, general writings, even those under submissions or rejected, would still be good (and a lot better than nothing).
+Thus, general writings, even those under submissions or rejected, would still be good (and much better than nothing).
 But be sure to upload your papers with your application and talk about them in your statement (see \S\ref{sec:research-statement}).
@@
 You can find top CS conferences at places such as \href{https://csrankings.org}{CSRankings}, which ranks CS programs based on how their faculty publish at top \emph{conferences} (see more in \S\ref{sec:ranking}).
-
-\begin{commentbox}[Vu:]
-  Many international students mention Scopus Q1, which consists of various journals from IEEE, Elsevier, and many other publishers.
-I don't know/recognize many of journals listed in Scopus Q1.
-This might be something to be mindful of, as \textbf{CS} faculty might not be too familiar with Scopus or journals listed there, so devote sometime in your statement to discuss the significance of your papers.
-\end{commentbox}
 
@@
 
-Most CS PhD programs will require at least \textbf{two LORs}. Having a letter from an internationally recognized researcher can greatly strengthen your application. However, obtaining such letters
-can be challenging for international students, who might not have much interactions with such experts.
-So it is fine to have a letter from people (e.g., profs, researchers, mentors) that know you well enough to talk about \emph{your specific research experience and capabilities}.
-This person can be a postdoc or even graduate student who has mentored you.
-
 \begin{center}
@@
 
-Many students have letters written by themselves and signed by their professors. These have little
-values. We can recognize them and will consider them weakness.
+Most CS PhD programs will require at least \textbf{two LORs}. Having a letter from an internationally recognized researcher can greatly strengthen your application. However, obtaining such letters
+can be challenging for international students, who might not have much interactions with such experts.
+So it is fine to have a letter from people (e.g., profs, researchers, mentors) that know you well enough to talk about \emph{your research experience and capabilities}.
+This person can be a postdoc who has mentored you.
+
+\begin{commentbox}[Vu:]
+    If you are from a good school from your country or your grading system is not standard.
+You can ask your reference writers to explain about that in their letters.
+For example, "Bac Khoa" are the top universities in Vietnam for CS but few people outside Vietnam know about them.
+So if you are from there, you should ask your reference writers to mention that in their letters.
+\end{commentbox}
+
+Many students have letters written by themselves and signed by their professors. Such letters have little
+values and would be considered weak (you cannot even find someone cares enough about you to write a candid personal reference letter?).
+Instead of the reference writer writing about you, in this case it is you who write about yourself (and they just sign the letter).
+
 Also, many professors write generic letters for students, e.g., the students didn't do any
@@
 
-\paragraph{Always Waive your Right to Review the LORs}  Choosing not to look at a reference letter is pretty standard in job and school applications.
+\paragraph{Always Waive your Right to Review the LORs}  Choosing not to look at a reference letter is pretty standard in school and job applications.
 When you waive your right to see the letter, it adds a layer of trust, showing you're confident in your choice of referees and that you're not trying to twist their words.
-It's also about keeping things open and honest between you and your letter writters, and encourages them to be real about your strengths and qualifications.
+It's also about keeping things open and honest between you and your letter writers, and encourages them to be real about your strengths and qualifications.
 Plus, it keeps things private.
@@
 
+\paragraph{Reminding ref.
+writers to submit their letters} After you submit your application, you can tell your writers that you have submitted your application and remind them to submit their letters if they haven't done so.
+But don't send too many reminders as that can be annoying to the writers.
+
+Note that most places only have deadlines for you, the applicant, but are very flexible with the letter writers (in many cases do not really give them any deadline or a very general one).
+Also, many places do not begin the admission review process right after the deadline and instead won't start until the next semester (mid January).
+
 \paragraph{VEF 2.0} For Vietnamese students, it's worth mentioning about the \href{https://vef2.org/}{VEF2.0 program}, which has helped many good students in gaining acceptance to top PhD programs in the US.
 VEF2.0 follows an interesting model where US faculty members from leading institutions are invited to conduct rigorous interviews with VEF students and subsequently provide reference letters on their behalf.
-Despite the limited interaction between the interviewer and the interviewee (primarily confined to the interview itself), the impact of these reference letters can be positive as they have helped many students getting into top PhD programs in the US.
+Despite the limited interaction between the interviewer and the interviewee (primarily confined to the interview itself), these reference letters are generally helpful as they have helped many students getting into top PhD programs in the US.
 

### 2023-12-06 demystify.tex
updates

@@
 
-Most CS PhD programs will require at least \textbf{two LORs}. Having a letter from an internationally recognized researcher can greatly strengthen your application. However, obtaining such letters
+Most PhD programs will require at least \textbf{two LORs}. Having a letter from an internationally recognized researcher can greatly strengthen your application. However, obtaining such letters
 can be challenging for international students, who might not have much interactions with such experts.
-So it is fine to have a letter from people (e.g., profs, researchers, mentors) that know you well enough to talk about \emph{your research experience and capabilities}.
-This person can be a postdoc who has mentored you.
+So it is fine to have a letter from people (e.g., profs, researchers, postdocs who mentored you) that know you well enough to talk about \emph{your research experience and capabilities}.
+Many students get letters from supervisors from companies where they did internships or are working at.
+It is OK as long as it is a research-based personalized letter (once again, we are talking about PhD applications, not MS).
 
 \begin{commentbox}[Vu:]
-    If you are from a good school from your country or your grading system is not standard.
-You can ask your reference writers to explain about that in their letters.
-For example, "Bac Khoa" are the top universities in Vietnam for CS but few people outside Vietnam know about them.
-So if you are from there, you should ask your reference writers to mention that in their letters.
-\end{commentbox}
-
-Many students have letters written by themselves and signed by their professors. Such letters have little
-values and would be considered weak (you cannot even find someone cares enough about you to write a candid personal reference letter?).
+    If your grading system is not US standard or you are from a good school unknown outside of your country, you can ask your reference writers to explain about that in their letters.
+For example, "Bach Khoa" are the top universities in Vietnam for STEM studies but few people outside Vietnam know about them.
+So if you are from there, you should ask your reference writers to mention that.
+\end{commentbox}
+
+\paragraph{Self-written Letters} Many international students write letters themselves (typically due to the requests of their profs.) and have them signed by their profs.
+Such letters have little values and would be considered weak by reviewers, who can question why you cannot even find someone cares enough about you to write a candid personal reference letter.
 Instead of the reference writer writing about you, in this case it is you who write about yourself (and they just sign the letter).
 
-Also, many professors write generic letters for students, e.g., the students didn't do any
-research or make any impression for the professor to write about.
-These letters are also not useful and considered weak.
+\paragraph{Generic Letters} When the writers do not know much about the applicants (e.g., just taking some course with them or not making any impression to write about), they might write a generic letter, which is not useful and also considered weak.
 It might be a good idea to directly ask if the prof.
@@
 If not, then you should ask someone else.
-For example, if a student I don't know well ask me to write letter for them, I will explicitly tell them I don't know them that well to write much about them, and such a short, generic, and weak letter will not help their profile.
-
-Many students get letters from supervisors from companies where they did internships or are working at.
-It is OK as long as it is a research-based personalized letter (once again, we are talking about PhD applications, not MS).
+For example, if a student I don't know well ask me to write letter for them, I will explicitly tell them I don't know them that well to write much about them, and such a short, generic, and weak letter will not help their case.
 
@@
 
-\paragraph{Always Waive your Right to Review the LORs}  Choosing not to look at a reference letter is pretty standard in school and job applications.
+\paragraph{Waiving your Right to Review LORs}  Choosing not to look at a reference letter is pretty standard in school and job applications.
 When you waive your right to see the letter, it adds a layer of trust, showing you're confident in your choice of referees and that you're not trying to twist their words.
@@
 
-\paragraph{Reminding ref.
-writers to submit their letters} After you submit your application, you can tell your writers that you have submitted your application and remind them to submit their letters if they haven't done so.
+\paragraph{Reminding Writers to Submit Letters} After you submit your application, you can tell your writers that and remind them to submit their letters if they haven't done so.
 But don't send too many reminders as that can be annoying to the writers.
 
-Note that most places only have deadlines for you, the applicant, but are very flexible with the letter writers (in many cases do not really give them any deadline or a very general one).
-Also, many places do not begin the admission review process right after the deadline and instead won't start until the next semester (mid January).
+Note that most places only have deadlines for the applicant, but are very flexible with the letter writers (in many cases do not even give them any deadline).
+Also, many places do not begin the admission review process right after the deadline and work on application reviews in the next semester (mid January).
 
@@
 Typically, it takes 5--7 years for CS PhD in the US.
-This can be longer than CS PhDs in other countries, which might require MS first (recall that US PhD programs do not require MS and you can get MS along the way to PhD).
+This can be longer than CS PhDs in other countries, which might require MS first (recall that CS PhD programs in the US do not require MS and you can get MS along the way to PhD).
 Within these 5--7 years, CS PhD students often take a ``leave of absence'' for 1--2 semesters to do internship at companies and research labs.
@@
 
-\emph{MS Degree requirement}:  as mentioned in \S\ref{sec:msrequirement} and \ref{sec:time}, US PhD programs do not require MS degree.
+\emph{MS Degree requirement}:  as mentioned in \S\ref{sec:msrequirement} and \ref{sec:time}, CS PhD programs in the US do not require MS degree.
 In contrast, many other countries do require MS degree before joining a PhD program.
@@
 \emph{Course work}: in the US you will spend the first couple of years taking classes and explore potential adviser and research topics.
-In other countries, you often do not do course work and start your research right away, e.g., you immediately work on the research project you proposed with the adviser you chose.
+In other countries, you (who already have an MS) start your research right away, e.g., you immediately work on the research project you proposed with the adviser you chose.
 Also, in the US you also have to pass a series of "exams", e.g., qualifying exam, comprehensive exam, thesis proposal defense\footnote{The word "ABD" (all but dissertation) is used in the US to refer to a PhD candidate who have finished all course work and exams and only need to write and defend their dissertation.}.
-In other countries, you do not have to do these exams or only do few of them.
-
-\emph{Funding}:  In many countries stipend comes from the university or from gov't research organizations and often have a fixed duration (e.g., 3 years).
+In other countries, you do not have to do these exams or only do a few of them.
+
+\emph{Funding}:  In many countries stipend comes from the university or from the gov't.
+These funding often have a fixed duration (e.g., 3 years).
 In the US, stipend (e.g., RA) comes directly from your adviser (no fixed duration).
@@
 
-\emph{Work-life balance}: I believe that PhD students in the US are overworked compared to other countries, e.g., in Europe.
+\emph{Work-life balance}: PhD students in the US are often said to be overworked compared to other countries, e.g., in Europe.
 This is partly due to the longer PhD program and that US PhD students are often paid through TA, which requires them to do TA in addition to their own research.
@@
   \tcblower
-  \textbf{Vu}: Yes, for some reason I enjoy addressing you as ``Prof.'' (without appending a last or first name).
-I call Steph and everyone else by their their first names.
-The use of Prof.
-Vu may have foreshadowed my future in academia.
-\end{commentbox}
-
-Certainly, there's a peculiar satisfaction in addressing you as "Prof." without appending a last or first name.
-While I refer to Steph and everyone else by their first names, perhaps addressing Prof.
-Vu in this manner hinted at my future in academia.
+  \textbf{Vu}: Yes, for some reason I enjoy addressing you as ``Prof.'' (without appending a last or first name).  The use of Prof. Vu may have foreshadowed my future in academia.
+\end{commentbox}
 

### 2023-12-06 demystify.tex
updates

@@
 Do not include ranking (e.g., Assistant, Associate, Scientist, ...) when referring to someone.
-I see many students include a lengthy title of people they know, e.g., \emph{I am advised by Asst.
+I see many international students include a lengthy title of people they know, e.g., \emph{I am advised by Asst.
 Prof.
@@
 PhD students often ask why their salary is so low compared to ludicrous grants their advisers get.
-They also wonder why their offer letters sometime mentioned that their benefits appear better than what they actually receive (i.e., stipend).
+They also wonder why their offer letters sometimes show that their benefits higher than what they actually receive (i.e., stipend).
 This section aims to shed some light to these questions.
@@
 Private universities may have different numbers.
-For simplicity, I will assume the department has a 9-month stipend of \$27000 (GMU actually pays more) and therefore a 3-month summer of \$9000.
+For simplicity, in this table I will assume the department has a 9-month stipend of \$27000 (GMU actually pays more) and therefore a 3-month summer of \$9000.
 I will also use GMU tuition rate of about \$15,000/year for full-time study (which is quite cheap compared to private universities, e.g., MIT charges around \$50K) and a 58.9\% rate on \emph{indirect cost}, which is what GMU charges for overhead or administrative costs (yes, after all, universities are businesses!).

### 2023-12-06 demystify.tex
updates

@@
 
-\begin{commentbox}
-    [DK:] Here are a few items you may consider addressing: putting international students in touch with other students from their countries, assuring them that they would be picked up from airports upon their arrival and that their initial stay will be taken care of.
+% \begin{commentbox}
+%     [DK:] Here are a few items you may consider addressing: putting international students in touch with other students from their countries, assuring them that they would be picked up from airports upon their arrival and that their initial stay will be taken care of.
 Most universities have other resources for these, but it is worth mentioning that they would be taken care of upon arrival and can get help during the transition phase.
@@
 But things have changed as one can find many ethnic food places, a big contrast from 1973, when there were two Indian restaurants in Greater Boston area.
-\end{commentbox}
+% \end{commentbox}
 

### 2023-12-10 demystify.tex
Updates from Overleaf

@@
 
-Additionally, \textbf{work experiences at renowned research laboratories}, such as Microsoft Research, can significantly strengthen your
+Additionally, \textbf{work experiences at renowned research laboratories}, such as Microsoft Research, can strengthen your
 application.
@@
 Finally, \textbf{participating internationally recognized competitions} can also demonstrate your research potential.
-For example, participating in Math Olympiads if you want to do theory or  winning ACM programming contests if you want to ``build'' stuff, e.g., software analysis.
+For example, participating in Math Olympiads if you want to do theory or  winning ACM programming contests if you want to ``build'' systems, e.g., software analysis.
 
@@
 
-Having good grades is important, but unless your school is well-known, having top grades or rankings
+Having good grades is good, but unless your school is well-known, having top grades or rankings
 usually will not help because we cannot evaluate them.
 
-This can be an issue for students in many top international universities where the competition is so high that very good students students can still have low rankings from these schools (and be overlooked by the admission committee).
+This can be an issue for students in many top international universities where the competition is so high that very good students can still have low rankings from these schools (and be overlooked by the admission committee).
 So what to do with this? as mentioned in \S\ref{sec:your-school} you can put a note about this in your statement and ask your LoR writers to talk about it.
@@
 Note that while having good grades at unknown school might not help,
-having very bad grades will be \red{red flag} (unless your LORs or
+having bad grades will be \red{red flag} (unless your LORs or
 statement give a proper explanation). This is especially true if you
@@
 The CV/Resume should provide a summary of the applicant's achievements.
-It should allow the reviewers to quickly scan to identify standout achievements (e.g., Publications, Programming Competition Awards, Teaching Experience).
+It should allow the reviewers to quickly scan to identify major achievements (e.g., Publications, Programming Competition Awards, Teaching Experience).
 
@@
 Now you're at a different game because the schools that have admitted you will try to get you to accept them!
-Important factors to consider are the reputation of schools and professors (\S\ref{sec:schoolsandprofs}), and of course, funding availability (\S\ref{sec:funding}).
+Important factors to consider are the reputation of schools and professors (\S\ref{sec:schoolsandprofs}), and funding availability (\S\ref{sec:funding}).
 You will have to make your decision by around \emph{April 15}.
@@
 \paragraph{Open House} Most schools will have \textbf{Open House} events, which are a great resource to learn about the school, department, faculty, research, living, etc.
-Even if you can't come in person, you still can attend virtually and meet with individual faculty.
+Even if you can't come in person, you should attend virtually and meet with individual faculty.
 During the Open House, you get a chance to talk to individual faculty and current students.
@@
 
-\paragraph{If you do not get admitted} If you do not get admitted to any schools (or don't want to go to the ones that admit you), try again next time.
+\paragraph{If you do not get admitted} If you do not get admitted to any schools or don't want to go to the ones that admit you, try again next time.
 Graduate admission can involve randomness and noise.

### 2023-12-20 demystify.tex
Updates from Overleaf

@@
 admit you with full funding, or we reject your application. In some rare cases, we admit
-without funding because you have funding on your own (e.g.,
-supported by your government or having external grants). We also justify
+without funding because you have funding on your own, e.g.,
+supported by your government or having external fellowships. We also justify
 our decision with a summary of your application, where we list
@@
 
-\paragraph{Waiving your Right to Review LORs}  Choosing not to look at a reference letter is pretty standard in school and job applications.
+\paragraph{Waiving your Right}  Choosing not to look at a reference letter is pretty standard in school and job applications.
 When you waive your right to see the letter, it adds a layer of trust, showing you're confident in your choice of referees and that you're not trying to twist their words.
@@
 
-\paragraph{Reminding Writers to Submit Letters} After you submit your application, you can tell your writers that and remind them to submit their letters if they haven't done so.
+\paragraph{Reminding Writers} After you submit your application, you can tell your writers that and remind them to submit their letters if they haven't done so.
 But don't send too many reminders as that can be annoying to the writers.
@@
 
-\section{Preparing Materials}
-\subsection{Sharing spreadsheets}
-Many students use a spreadsheet to help them and their letter writers to keep track of their applications.
-You can create a personal and a shared spreadsheet but a simplest one is just a shared one (e.g., using Google sheet).
+\section{Preparing and Tracking Applications}
+
+\paragraph{Shared Spreadsheet} Many students use a spreadsheet to help them and their letter writers to keep track of their applications.
+They often use a shared spreadsheet (e.g., using Google Docs or Sheets).
 Here are some information to put on the sheet.
@@
 Under each is their status, e.g., sent/not yet/reminder needed, etc.
+
+You can then share this sheet with your reference writers and remind them to submit LoRs on your behalf (see \S\ref{sec:lor}).
 \end{itemize}
+
+%\paragraph{Communicating with LoR writers and People Who Support You}
+
+%stressfull time, rely on support from others
+%ask ppl to review your SOPs
+%provide information for LORs, ..
+%
 

### 2023-12-23 demystify.tex
Updates from Overleaf

@@
         \begin{itemize}
-          \item Applications are evaluated by the \emph{PhD Admission} committee and each application is reviewed by typically 3 faculty (\S\ref{sec:evalapps}).
+          \item Applications are evaluated by the \emph{PhD Admission} committee and each application is reviewed by typically three faculty (\S\ref{sec:evalapps}).
           \item Individually faculty \emph{cannot directly admit} a student---so do not email and ask if you have a chance.
@@
           \item Compared to other countries, CS PhD in the US does not require an MS degree but has longer PhD study time (\S\ref{sec:non-us-differences}).
-          \item Your PhD costs about \$400K in total, but you do not pay for it (\S\ref{sec:ra-cost}).
-          \item Despite some miserable stories on social media, many PhD students have good mentors, supportive labmates, healthy working environment ... and are happy (\S\ref{sec:happy}).
+          \item Your PhD costs about \$400K in total, but you \emph{do not} pay for it (\S\ref{sec:ra-cost}).
+          \item Despite some miserable stories on social media, many PhD students have good mentors, supportive lab mates, healthy working environment ... and are happy (\S\ref{sec:happy}).
         \end{itemize}
@@
 Second, applying to a good US university \emph{should not} be any harder than at schools in other countries.
-In fact, it might even be more flexible since CS PhD in the US \emph{do not} require having an MS (or having to know a topic/proposal/adviser in advance).
+In fact, it might even be more flexible since CS PhD in the US \emph{do not} require having an MS or knowing a topic/proposal/adviser in advance (by the way, GRE is also not required!).
 If you believe you have a chance in other countries, e.g., South Korea, Singapore, Germany, UK, Japan, and Australia, then you will surely have a chance in the US as well.
@@
 
-In many cases you do not have to do a PhD in CS to work in your area of interested.
+In many cases you do not need to do a PhD in CS to work in your area of interested.
 For example, in addition to a traditional CS department, GMU has IST and Cybersecurity departments, which have faculty  with PhD in CS and work on CS topics (e.g., AI, Security, Robotics).
@@
 
-However, if your intention is a PhD in CS, then you likely need to be in the CS dept \emph{and} advised by a CS faculty.
-In fact, if a faculty is not in CS, it is unlikely that they can be the \emph{main} adviser of a CS PhD student---they may \emph{co-advise} or be in the PhD dissertation committee, but your main adviser will need to be a tenured or tenure-track faculty in CS.
+However, if your goal is a PhD in CS, then you likely need to be in the CS dept \emph{and} advised by a CS faculty.
+In fact, if a faculty is not in CS, it is unlikely that they can be the \emph{main} adviser of a CS PhD student.
+They may \emph{co-advise} or be in the PhD dissertation committee, but your main adviser will need to be a tenured or tenure-track faculty in CS.
 If in doubt, you should check with the CS department for their requirements.
@@
 In the US, you often start your PhD without an adviser or project and find them later.
+This is because as someone who just finished your undergraduate study, it is unlikely that you know enough about research to do proposal or picking an adviser.
 This is also why US PhD programs are longer.
@@
 
-\emph{Faculty position after PhD}: In other countries, PhD graduates who are interested academia typically apply for research positions at research labs, i.e., postdocs, and then consider faculty position.
+\emph{Faculty position after PhD}: In other countries, PhD graduates interested academia typically apply for additional research appointments, i.e., postdocs, and then consider faculty position.
 In the US, PhD graduates often apply directly for faculty position (postdoc for US graduates is no longer a popular option as it was before).
-This is why US PhD programs are longer and more focused on research.
+%This is also why US PhD programs are longer (remember, 5--7 years).
 
@@
 
-This section lists various cultural issues that international students might want to pay attention to.
+This section lists some cultural issues that international students might want to pay attention to. These issues are not specific to just the PhD admission process but in general.
 
@@
 
+\section{Relationships with your Profs.}
 There's a misconception that in the US it's all business, with professors as bosses who pay students for their work, and you doing something nice implying you expect something in return.

### 2023-12-23 demystify.tex
updates

--- 2023-12-10 demystify.tex
@@
 
-Additionally, \textbf{work experiences at renowned research laboratories}, such as Microsoft Research, can strengthen your
+Additionally, \textbf{work experiences at renowned research laboratories}, such as Microsoft Research, can significantly strengthen your
 application.
@@
 Finally, \textbf{participating internationally recognized competitions} can also demonstrate your research potential.
-For example, participating in Math Olympiads if you want to do theory or  winning ACM programming contests if you want to ``build'' systems, e.g., software analysis.
+For example, participating in Math Olympiads if you want to do theory or  winning ACM programming contests if you want to ``build'' stuff, e.g., software analysis.
 
@@
 
-\section{Grades/GREs}\label{sec:grades}
-
-Having good grades is good, but unless your school is well-known, having top grades or rankings
+\section{Grades and GREs}\label{sec:grades}
+
+Having good grades is important, but unless your school is well-known, having top grades or rankings
 usually will not help because we cannot evaluate them.
 
-This can be an issue for students in many top international universities where the competition is so high that very good students can still have low rankings from these schools (and be overlooked by the admission committee).
+This can be an issue for students in many top international universities where the competition is so high that very good students students can still have low rankings from these schools (and be overlooked by the admission committee).
 So what to do with this? as mentioned in \S\ref{sec:your-school} you can put a note about this in your statement and ask your LoR writers to talk about it.
@@
 Note that while having good grades at unknown school might not help,
-having bad grades will be \red{red flag} (unless your LORs or
+having very bad grades will be \red{red flag} (unless your LORs or
 statement give a proper explanation). This is especially true if you
@@
 The CV/Resume should provide a summary of the applicant's achievements.
-It should allow the reviewers to quickly scan to identify major achievements (e.g., Publications, Programming Competition Awards, Teaching Experience).
+It should allow the reviewers to quickly scan to identify standout achievements (e.g., Publications, Programming Competition Awards, Teaching Experience).
 
@@
 Now you're at a different game because the schools that have admitted you will try to get you to accept them!
-Important factors to consider are the reputation of schools and professors (\S\ref{sec:schoolsandprofs}), and funding availability (\S\ref{sec:funding}).
+Important factors to consider are the reputation of schools and professors (\S\ref{sec:schoolsandprofs}), and of course, funding availability (\S\ref{sec:funding}).
 You will have to make your decision by around \emph{April 15}.
@@
 \paragraph{Open House} Most schools will have \textbf{Open House} events, which are a great resource to learn about the school, department, faculty, research, living, etc.
-Even if you can't come in person, you should attend virtually and meet with individual faculty.
+Even if you can't come in person, you still can attend virtually and meet with individual faculty.
 During the Open House, you get a chance to talk to individual faculty and current students.
@@
 
-\paragraph{If you do not get admitted} If you do not get admitted to any schools or don't want to go to the ones that admit you, try again next time.
+\paragraph{If you do not get admitted} If you do not get admitted to any schools (or don't want to go to the ones that admit you), try again next time.
 Graduate admission can involve randomness and noise.

### 2023-12-23 demystify.tex
updates

--- 2023-12-23 demystify.tex
@@
 \setcounter{tocdepth}{3}
-
 \usepackage[T1]{fontenc}
@@
 \usepackage{booktabs}
-
-% \newtcolorbox{cbox}[1][]
-% {
-% left=1pt,right=1pt,top=1pt,bottom=1pt,
-% colback=gray!10,
-% boxrule=0.5pt
-% }
 
@@
 
-\section{Grades/GREs}\label{sec:grades}
+\section{Grades and GREs}\label{sec:grades}
 
@@
 
-Most universities the US will have CS programs. 
+Most universities the US have CS programs. 
 However, while many universities offer PhD in their CS programs, many do not.
@@
 in CS,  but Penn State in Harrisburg only has BS and MS in CS, and Penn State in York only has BS in CS.
-On the other hand, multiplication locations of the University of Texas, e.g., Austin, Dallas, Arlington, all have PhDs in CS.
+On the other hand, multiple locations of the University of Texas, e.g., Austin, Dallas, Arlington, have PhDs in CS.
 

### 2023-12-31 demystify.tex
Updates from Overleaf

@@
 
-\subsection{Should you contact a professor? What to do to get a desired reply?}\label{sec:contact}
+\subsection{Should you contact a professor? How to get a desired reply?}\label{sec:contact}
 
@@
 If in doubt, you should check with the CS department for their requirements.
-For this specific reason,  CSRankings includes only tenured or tenured track faculty who can advise CS PhD students.
+For this specific reason,  CSRankings includes only tenure line faculty who can advise CS PhD students.
 I also have compiled a \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{list} of Vietnamese faculty who can advise PhD students.

### 2024-01-02 demystify.tex
minor fix changing LOR to SOP

@@
 In your statement, you have the opportunity to make your application stand out and unique, which can make you \emph{fit} the CS program you're applying to, even if you don't have very strong research experience.
-A well-written LOR also shows that you can communicate, which is very important in research, and that you can effectively teach and communicate with students, which is important for TA funding (see \S\ref{sec:funding}).
+A well-written SOP also shows that you can communicate, which is very important in research, and that you can effectively teach and communicate with students, which is important for TA funding (see \S\ref{sec:funding}).
 % is important because if you need (GTA) funding, it will provide evidence

### 2024-01-02 demystify.tex
updates

--- 2023-12-31 demystify.tex
@@
     3 & Univ. of California-San Diego & 28 & Rutgers University\red{$^*$} \\
-    4 & MIT & 29 & Univ. of California - Riverside\\
-    5 & Georgia Institute of Technology         & 30 & Northwestern University\\
-    6 & Stanford University& 31 & Pennsylvania State University  \\
+    4 & Georgia Institute of Technology & 29 & Univ. of California - Riverside\\
+    4 & MIT         & 30 & Northwestern University\\
+    6 & Univ. of California - Berkeley& 31 & Pennsylvania State University  \\
     7 & University of Michigan - Ann Arbor\red{$^*$}   & 32& George Mason University\red{$^*$}\\  
     8 & University of Washington      &33 &  Harvard University \\
-    9 &  Univ. of California - Berkeley  &34&  Univ. of California - Santa Cruz \\
+    9 &  Stanford University  &34&  Univ. of California - Santa Cruz \\
     10 & Cornell University  & 35 &  Yale University \\
@@
     13 & University of Wisconsin - Madison\red{$^*$}  &38& Texas A\&M University\red{$^*$} \\
-    14 & Columbia University   &39 & Boston University  \\
-    15 &   Purdue University  &40& North Carolina State University\\\  
-    16 & University of Texas at Austin   &41 & University of Utah \\
-    17 & University of Pennsylvania\red{$^*$} &42 & University at Buffalo\red{$^*$}\\
+    13 & Purdue University  &39 & Boston University  \\
+    15 & University of Texas at Austin &40& North Carolina State University\\\  
+    16 & University of Pennsylvania &41 & University of Utah \\
+    17 & Columbia University \red{$^*$} &42 & University at Buffalo\red{$^*$}\\
     18 & Princeton University  & 43& Rice University\\
-    19 & University of Massachusetts-Amherst\red{$^*$} & 44&  University of Colorado-Boulder \\
-    20 &  New York University  &45& University of Illinois at Chicago  \\
+    19 & New York University  & 44&  University of Colorado-Boulder \\
+    20 &  University of Massachusetts-Amherst\red{$^*$} &45& University of Illinois at Chicago  \\
     21 & Univ. of California - Los Angeles &46& Virginia Tech\red{$^*$}  \\

### 2024-01-02 demystify.tex
updates

@@
 Furthermore, although many examples are specifics for schools that I and other contributors of this document know about, the information should be generalizable to other R1\footnote{An \href{https://en.wikipedia.org/wiki/List_of_research_universities_in_the_United_States}{R1 institution} in the US is a research-intensive university with a high level of research activity across various disciplines.
-Currently, 146 (out of 4000) universities are classified as R1.} institutions in the US (and also likely universities in other countries).
+Currently, 146 (out of 4000) universities are classified as R1.} institutions in the US (and universities in other countries).
 
@@
 
-The latest version of this document is at \url{nguyenthanhvuh.github.io/phd-cs-us/demystify.pdf} and its \LaTeX{} source is on \href{https://github.com/nguyenthanhvuh/phd-cs-us}{Github}.
+This document is available at \url{nguyenthanhvuh.github.io/phd-cs-us/demystify.pdf} and its \LaTeX{} source is on \href{https://github.com/nguyenthanhvuh/phd-cs-us}{Github}.
 If you have questions or comments, feel free to create a \href{https://github.com/nguyenthanhvuh/phd-cs-us/issues}{GitHub issue} for discussion.
@@
 
-Here, we outline the main points of the document. This gives you an overview to decide which specific topics you want to explore more thoroughly.
+The main points of this guideline are summarized here. This gives you an overview to decide which specific topics you want to explore more thoroughly.
 
@@
 Second, applying to a good US university \emph{should not} be any harder than at schools in other countries.
-In fact, it might even be more flexible since CS PhD in the US \emph{do not} require having an MS or knowing a topic/proposal/adviser in advance (by the way, GRE is also not required!).
+In fact, it might even be more flexible since CS PhD in the US \emph{do not} require having an MS or knowing a topic/proposal/adviser in advance (by the way, GRE is not required!).
 If you believe you have a chance in other countries, e.g., South Korea, Singapore, Germany, UK, Japan, and Australia, then you will surely have a chance in the US as well.
@@
 You also have more opportunities for funding, e.g., through government scholarships for US citizens and residents.
-Finally, compared to international students, PhD applications in CS from domestic students are few and that can help your case.
+Finally, compared to international students, PhD applications in CS from domestic students are few and that can really help your case.
 \end{domesticbox}
@@
 
-\section*{References}
+\section*{Additional Resources}
 \begin{itemize}
@@
 
-After you submit your PhD application (usually November--December), it will be first screened
-for general requirements, e.g., did you submit your transcripts and standard scores? did your reference writers submit their letters?
-
-Then, applications are reviewed by a \textbf{PhD admission committee} that consists of faculty members in CS (in some cases the committee can involve affiliated faculty from different disciplines).
+After you submit your PhD application (usually November--December), it will first be screened
+for general requirements, e.g., did you submit your transcripts and standard scores?
+did your reference writers submit their letters?
+Usually, this screen process is done through a central university system (i.e., not by the CS department).
+If you pass this screen, your application will be forwarded to the CS department for further evaluation.
+If you don't, then the system will inform you what is missing and what you need to do.
+
+\begin{commentbox}[Hakan:]
+    At GMU, for full consideration, students should make sure to submit \textbf{ALL} required documents by the application deadline, and should never assume that some required documents (such as official TOEFL scores or official diplomas/transcripts) will be waived by the admissions office.
+If something is listed and not marked as ``optional'', it is mandatory and they should plan for submitting all those.
+  \end{commentbox}
+
+Then, applications are reviewed by a \textbf{PhD admission committee} (adcom) that consists of faculty members in CS (in some cases the committee can involve affiliated faculty from different disciplines).
 These faculty have a wide-range of expertise and background to ensure diverse perspectives in the evaluation process.
-The size and the review load of the committee depends on the department size.
-At GMU, the PhD committee typically has 15--20 faculty, and each committee member is assigned to review 25--30 applications.
-Note that GMU, UMass, and probably many other schools, have separate committees for MS programs.
+The size and the review load of the adcom depends on the department size.
+At GMU, the PhD adcom typically has 15--20 faculty, and each committee member is assigned to review 25--30 applications.
+Note that GMU, UMass, and probably many other schools, have separate adcom for MS programs.
 
@@
 
-The PhD admission committee typically involves assistant professors in the department (see \S\ref{sec:faculty-types} for various faculty titles).
+The PhD adcom typically involves assistant professors in the department (see \S\ref{sec:faculty-types} for various faculty titles).
 This provides junior faculty the opportunities to recruit students.
@@
 
-\begin{commentbox}[Hakan:]
-  At GMU, for full consideration, students should make sure to submit \textbf{ALL} required documents by the application deadline, and should never assume that some required documents (such as official TOEFL scores or official diplomas/transcripts) will be waived by the admissions office.
-If something is listed and not marked as ``optional'', it is mandatory and they should plan for submitting all those.
-\end{commentbox}
 \paragraph{Why we do not waive application fee?}  This is typically a requirement of the university.

### 2024-01-02 demystify.tex
updates

@@
 
-\paragraph{Self-written Letters} Many international students write letters themselves (typically due to the requests of their profs.) and have them signed by their profs.
-Such letters have little values and would be considered weak by reviewers, who can question why you cannot even find someone cares enough about you to write a candid personal reference letter.
-Instead of the reference writer writing about you, in this case it is you who write about yourself (and they just sign the letter).
+\paragraph{Self-written Letters} Many international students write letters themselves, typically due to the requests of their profs., and have them signed by their profs.
+Such letters have little values and are considered weak by reviewers (why can you not even find someone who cares or knows enough about you to write a candid personal reference letter?).
+Instead of the reference writer talking about you, in this case it is you who write about yourself (and they just sign the letter).
 
@@
 
+\begin{commentbox}[Hung:]
+  A sad reality is that most professors in Vietnam \textbf{DO NOT} know how to write a good letter, or are lazy in writing letters hence delegate the writing to the students.
+Unfortunately, there is no easy solution to this problem.
+\end{commentbox}
+
 \begin{commentbox}[Vu:]
-  Some letter writers ask the applicants to write their own letters for them to sign.
-As mentioned, this will hurt the applicants (admission committee members are actually quite good at determining this)!
-Multiple students mentioned that  some professor did this so that the applicants do not go abroad or only go to places where they want the students to go to.
+
+    Another sad reality is that some professors are unwilling to write letters or write weak ones because they do not want (good) students to go abroad or only go to places where they want the students to go to.
+If you are in this situation, you should find someone else to write for you.
     \tcblower
-  Sometimes students would go through great length just to get letters from well-known professors in their school, but the letters are generic and carry little value, in fact, \red{red flags}.
+
+    Sometimes students would go through great length just to get letters from well-known professors in their school (e.g., department head).
+But as mentioned, if these professors do not know you, the letters are generic and carry little value, in fact, \red{red flags}.
 Moreover, a top professor in Vietnam might not be well-known to US faculty (see more details in \S\ref{sec:your-school}).
-So save the trouble and just get letters from \emph{any} professors/supervisors who knows you well and can write a good letter about your research ability.
+So save the trouble and just get letters from \emph{any} professors/supervisors who knows you well and can write a good letter about \emph{your} research ability.
 It's better to have a good personalized
@@
     well-known than a generic/weak letter from a well-known person.
-\end{commentbox}
-
-\begin{commentbox}[Hung:]
-  A sad reality is that most professors in Vietnam \textbf{DO NOT} know how to write a good letter, or are lazy in writing letters hence delegate the writing to the students.
-Unfortunately, there is no easy solution to this problem.
+
   \end{commentbox}

### 2024-01-04 demystify.tex
updates

@@
         \begin{itemize}
-          \item Many schools do not offer PhD studies in CS and many CS professors \emph{cannot} be adviser, i.e., cannot graduate PhD students  (\S\ref{sec:schoolsandprofs}).
+          \item Many schools do not offer PhD studies in CS and many CS professors \emph{cannot} advise or graduate PhD students  (\S\ref{sec:schoolsandprofs}).
           \item Contacting a prof. for research opportunities is recommended, but do it \emph{properly} (\S\ref{sec:contact}).
@@
 
-\chapter{Cultural Differences and Other Issues}\label{sec:cultural}
-
-This section lists some cultural issues that international students might want to pay attention to. These issues are not specific to just the PhD admission process but in general.
-
-% \paragraph{Diversity} US universities prioritize diversity and inclusion.
-Students need to respect and appreciate varied opinions, backgrounds, and cultures.
-Unlike some countries where certain voices are marginalized, in the US, all perspectives are valued equally (especially at universities).
-Racism or discrimination will have serious consequences, including academic and disciplinary actions.
-
-\section{Academic Integrity}
-
-Plagiarism and cheating (e.g., exams, assignments) is a BIG no-no in the US.
-If you're caught cheating, you will face very heavy consequence and likely be expelled from the university (e.g., after the second time at GMU).
-This is quite different from many international countries where cheating is common and often tolerated.
-Faculty is extremely good at detecting cheating (we have been dealing with these situations so many times over so many years), and \emph{will} report cheating cases.
-In short, whatever you do, don't cheat---not worth it.
-
-Here is the typically steps: (i) a faculty suspecting a cheating case will report it to the Office of Academic Integrity (OAI) at the university---the report often has supporting evidence and suggested penalty (e.g., a failing grade);  (ii)  OAI will take over and investigate the case; and (iii) OAI will make the final decision.
-It is important to note that after receiving the report from your prof., OAI \emph{completely} takes over and makes decision.
-This means begging your professor will not help because they simply are no longer involved in the case.
-
-\section{Illegal Software} Using illegal/cracked software is very common in many countries (and even in the US).
-However, \emph{do not} install or use them on university computers (e.g., those given by your department or adviser).
-It is unlikely that the university will track you down, but it is the \emph{software company} that will.
-They have very sophisticated tools to detect illegal software and will sue your university/department.
-Imagine your department or adviser being sued for a large sum of money, and it is \emph{you} who caused it.
-If you need software, ask your adviser or the department.
-
-\section{Gifts} In many countries, it is customary to give professors gifts, often during holidays.
-These gifts can be costly and profs.
-sometimes expected them.
-In the US, it is uncommon and perfectly OK to not to give gifts.
-\emph{However}, if you'd like to offer small souvenir-like tokens, it's a thoughtful gesture that's appreciated.
-Some professors proudly display their gifts, which can come from students and colleagues (e.g., when they travel to their home countries or conferences).
-In summary, small gifts are fine, but avoid anything that might make your professors uncomfortable.
-
-\section{Relationships with your Profs.}
-There's a misconception that in the US it's all business, with professors as bosses who pay students for their work, and you doing something nice implying you expect something in return.
-However, the reality is quite the opposite.
-While people can be straightforward and direct, they are also friendly and informal.
-You can call your professor by their first name (\S\ref{sec:address}), disagree with them and argue (and gain respect doing so), seek their help (even on personal matters), come to their houses for parties or gathering, and give them small thoughtful gifts that they put on their desks.
-Many students and professors maintain lifelong relationships, staying in touch through cards, emails, and calls, even after their academic journey ends.% These connections are genuine and meaningful.
-
-% \begin{commentbox}
-%     [DK:] Here are a few items you may consider addressing: putting international students in touch with other students from their countries, assuring them that they would be picked up from airports upon their arrival and that their initial stay will be taken care of.
-Most universities have other resources for these, but it is worth mentioning that they would be taken care of upon arrival and can get help during the transition phase.
-Learning to cook was a big deal when I arrived over 50 years ago-August 1973.
-But things have changed as one can find many ethnic food places, a big contrast from 1973, when there were two Indian restaurants in Greater Boston area.
-% \end{commentbox}
-
 \appendix
 
-\chapter{Ranking of CS PhD programs}\label{sec:ranking}
+\chapter{Rankings of CS PhD programs}\label{sec:ranking}
 
@@
   \caption{Top 50 CS PhD programs in the U.S.
-(CSRankings, June 2023).
+(\href{https://csrankings.org}{CSRankings}, Jan.
+2024).
 \red{$^*$} indicates that the university has \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{Vietnamese prof.} that can advise CS PhD students.}\label{tab:ranking}
@@
     \toprule
-    1 & Carnegie Mellon & 26 & Univ. of California - Irvine \\
-    2 & Univ. of Illinois at Urbana-Champaign\red{$^*$}  & 27 &  Duke University \\
+    1 & Carnegie Mellon & 26 & Duke University \\
+    2 & Univ. of Illinois at Urbana-Champaign\red{$^*$}  & 27 & Univ. of California - Santa Barbara \\  
     3 & Univ. of California-San Diego & 28 & Rutgers University\red{$^*$} \\
     4 & Georgia Institute of Technology & 29 & Univ. of California - Riverside\\
-    4 & MIT         & 30 & Northwestern University\\
-    6 & Univ. of California - Berkeley& 31 & Pennsylvania State University  \\
+    5 & MIT                            & 30 & Pennsylvania State University  \\
+    6 & Univ. of California - Berkeley& 31 & Northwestern University\\
     7 & University of Michigan - Ann Arbor\red{$^*$}   & 32& George Mason University\red{$^*$}\\  
-    8 & University of Washington      &33 &  Harvard University \\
-    9 &  Stanford University  &34&  Univ. of California - Santa Cruz \\
-    10 & Cornell University  & 35 &  Yale University \\
-    11 & University of Maryland - College Park &  36& Brown University \\ 
-    12 & Northeastern University\red{$^*$} &37&  Ohio State University\\
-    13 & University of Wisconsin - Madison\red{$^*$}  &38& Texas A\&M University\red{$^*$} \\
-    13 & Purdue University  &39 & Boston University  \\
-    15 & University of Texas at Austin &40& North Carolina State University\\\  
-    16 & University of Pennsylvania &41 & University of Utah \\
-    17 & Columbia University \red{$^*$} &42 & University at Buffalo\red{$^*$}\\
+    8 & University of Washington      &33 &  Texas A\&M University\red{$^*$} \\
+    9 &  Stanford University  &34&  University of Utah \\
+    10 & Cornell University  & 35 &  North Carolina State University\\\  
+    11 & University of Maryland - College Park &  36& Ohio State \\ 
+    12 & Northeastern University\red{$^*$} &37& University of Virginia  \\ 
+    13 & Purdue University &38& Yale University \\ 
+    14 & University of Wisconsin - Madison\red{$^*$} &39 & Univ. of California - Santa Cruz \\
+    15 & University of Texas at Austin &40& Brown University \\ 
+    16 & University of Pennsylvania &41 & Harvard University \\
+    17 & Columbia University \red{$^*$} &42 & Boston University  \\
     18 & Princeton University  & 43& Rice University\\
-    19 & New York University  & 44&  University of Colorado-Boulder \\
-    20 &  University of Massachusetts-Amherst\red{$^*$} &45& University of Illinois at Chicago  \\
-    21 & Univ. of California - Los Angeles &46& Virginia Tech\red{$^*$}  \\
-    22 & University of Southern California &47&  Arizona State University\red{$^*$} \\
-    23 & Stony Brook University\red{$^*$} &48&University of Minnesota \\
-    24 & University of Chicago &49& University of Virginia \\
-    25 & Univ. of California - Santa Barbara &50& University of North Carolina\red{$^*$} \\
+    19 & New York University  & 44&  University at Buffalo\red{$^*$}\\
+    20 &  University of Massachusetts-Amherst\red{$^*$} &45& University of Colorado-Boulder \\
+    21 & Univ. of California - Los Angeles &46& University of Illinois at Chicago  \\
+    22 & University of Southern California &47& University of Minnesota \\ 
+    23 & University of Chicago &48& University of North Carolina\red{$^*$} \\
+    24 & Stony Brook University\red{$^*$} &49& Arizona State University\red{$^*$} \\
+    25 &  Univ. of California - Irvine&50& Univ. Of California - Davis \\
     \bottomrule
@@
 \end{table}
-
-Tab.~\ref{tab:ranking} lists the top 50 CS programs in the US from \href{https://www.csrankings.org}{CSRankings.org}, a ranking system  based on top CS conferences.
+%Virginia Tech\red{$^*$}  \\
+Tab.~\ref{tab:ranking} lists the top 50 CS programs in the US from \href{https://www.csrankings.org}{CSRankings.org}, a ranking system  based on faculty publications at top CS conferences.
 
@@
 
-\chapter{History and Acknowledgement}\label{sec:ack}
-
-\paragraph{History} This document was conceived during a lunch with Craig Yu at GMU.
-We talked on about why GMU were not able to attract good Vietnamese and other international students, despite having a much stronger CS program than many schools that these students want to go to (part of the reason is described in \S\ref{sec:selecting-ranking-schools}).
-We wished there were a way for international students to know about the US PhD programs (as well as for US faculty to understand more about international students and therefore have better chance of recruiting and working with them).
-I was also  a member of the large VietPhD group on Facebook and saw many questions from students about PhD programs.
-However, most active participant are students in non-CS fields or not in US.
-Like ChatGPT, their answers to questions on CS PhD program in the US could be informative and helpful, but unfortunately not always accurate and sometimes leading to more confusion.
-So I thought it would be useful to have a document that is specific to CS PhD programs in the US from an insider prospective.
-
-I started writing this document in May 2023 and have been updating it since then (mostly around deadline time when I tend to procrastinate, i.e., \emph{productive procrastination}!).
-I have put the source code of this document on \href{https://github.com/nguyenthanhvuh/phd-cs-us}{GitHub} so that anyone can contribute to it.
-
-\paragraph{About Me} I am an assistant professor in the CS dept at George Mason University (GMU).
-Prior to GMU, I was at the University of Nebraska-Lincoln (UNL).
-I have been in the PhD admission process at GMU and UNL for many years.
-Currently I serve as the program director of the MS program in Software Engineering at GMU (thus also have some experience with the MS admission process--which is quite different than PhD).
-%I got my undergraduate and MS degrees from Penn State.
-I completed my PhD at the University of New Mexico-Albuquerque and a postdoc at the University of Maryland-College Park.
-
-While I am not an international student, many of my students and collaborators are.
-I also mentor multiple students from Vietnan, and have various colleague and friends who were international students.
-I have heard their stories, and hope to capture their challenges and experiences in this document so that it can be useful for prospective international students.
-
-\paragraph{Acknowledgement} Many people have contributed to this document.
-Profs. Craig Yu (GMU), Hakan Aydin (GMU), 
-Xiaokuan Zhang (GMU), Hung Le (UMass), and Deepak Kapur (UNM) provided valuable input in the early version.
-Other GMU faculty members also have provided feedback and contributions.
-Many students including Didier (GMU), Thanh (Melbourne), and Dat (Melbourne) have contributed valuable questions and feedback.
-\textbf{Thank you!}
-
-\bibliographystyle{abbrv}
-\bibliography{demystify.bib}
-
-          %           \newpage
+% \chapter{Undergraduate Research Opportunities}
+
+% Prospective PhD students are often unaware of the importance of research experience in their application.
+This provides both concrete research experiences and potential LOR writers that can talk about your research ability.
+% %This is especially true for students from small colleges where research opportunities are limited.  
+% This section aims to provide some guidance on how to gain research experience as an undergraduate student.
+%This section is written together Didier, a PhD student at GMU who did his undergraduate study in Rwanda.
+
+% \paragraph{Locally} There are many ways to find research opportunities right in your own institution.
+If you did well or like a class, you can check with the professor of that class for research opportunities.
+Sometimes professors would mention that they look for undergraduates on their websites, so do check those out!
+Even if they say they are looking for graduate students, you still should contact them and ask.
+You can also ask your academic adviser or other faculty members for suggestions.
+Finally, you can also ask your peers who are already doing research.
+You can use the method described in \S\ref{sec:contact} to contact professors
+
+% In short, the most important thing is that you need to \emph{ask}!
+Don't be shy, don't be afraid of rejection, and don't be afraid of being ignored or that you're not good enough.
+Just ask!
+The worst thing that can happen is that you get a ``no'' or no response.
+But you will never know if you don't ask.
+And if you get a ``no'', you can always ask again later.
+There are students who got rejected the first time but got accepted the second time.
+So don't give up!
+
+% \paragraph{Online Research Communities and Open Source Contributions}:
+%                     Places such as Github offer many opportunities for research, even though they might not appear so (e.g., they do not explicitly mention research).
+Students can explore repositories related to their areas of interest, collaborate with other contributors, and actively engage in open source research projects.
+By contributing code, fixing bugs, implementing new features, or providing documentation, students can gain practical research experience and interact with experienced developers and researchers in the field.
+
+% \paragraph{Paid or Unpaid}  travel opportunities, publishing opportunities, etc
+
 %                    \section{Virtual Research Opportunities Beyond Physical Boundaries (written by Didier)}
@@
 
-          %           \textbf{Online Research Communities and Open Source Contributions}:
-          %           Online research communities and platforms offer a wealth of opportunities for aspiring researchers.
-Platforms like GitHub, GitLab, and Bitbucket host repositories spanning diverse research domains.
-Students can explore repositories related to their areas of interest, collaborate with other contributors, and actively engage in open source research projects.
-By contributing code, fixing bugs, implementing new features, or providing documentation, students can gain practical research experience and interact with experienced developers and researchers in the field.
-
 %                     \textbf{Online Conferences and Workshops}:
@@
 
+\chapter{Cultural Differences and Other Issues}\label{sec:cultural}
+
+This section lists some cultural issues that international students might want to pay attention to. These issues are not specific to just the PhD admission process but in general.
+
+% \paragraph{Diversity} US universities prioritize diversity and inclusion.
+Students need to respect and appreciate varied opinions, backgrounds, and cultures.
+Unlike some countries where certain voices are marginalized, in the US, all perspectives are valued equally (especially at universities).
+Racism or discrimination will have serious consequences, including academic and disciplinary actions.
+
+\section{Accepting, Postponing, and Rejecting Offers}
+
+Students, especially international ones, often get confused on what to do after they get an offer from a professor, e.g., how to postpone or accept/reject offers, and most importantly, how to do so without offending anyone.
+In short, if you are not serious about an offer for any reason (e.g., still waiting for other offers), do not accept it.
+If you are not sure, ask for more time to decide.
+But do not accept an offer and then later reject it.
+This is considered rude and unprofessional.
+If you do so, you will likely burn bridges with that professor.
+
+\paragraph{Accepting an offer} If you decide to accept an offer, you can say something like: \emph{``Thank you for the offer.
+I am very excited about it and would like to accept it.
+I look forward to working with you.
+Can you please send me more details about the offer and what to do next?''} .
+Here you are not only accepting the offer but also asking for more details.
+This is perfectly fine and professors will understand and appreciate your enthusiasm.
+They will likely send you more details about the offer and what to do next.
+If you decide to accept an offer, you should do so quickly.
+
+ If you have better offers, you might mention them to the professor and ask if they could match them.
+It is unlikely they can do so as RA might be fixed by the department and not negotiable.
+They might recommend you for departmental fellowships or some additional perks.
+
+\paragraph{Postponing an offer} If you need more time to decide, you can ask for more time.
+You can say something like: \emph{``Thank you for the offer.
+I am very excited about it.
+However, I am still waiting for other offers and need more time to decide.
+Would it be possible to postpone the decision for a few weeks?''}.
+This is perfectly fine and professors will understand and might even appreciate your honesty.
+They will likely give you a few weeks to decide.
+If you need more time, you can ask for more time.
+But do not ask for too much time, e.g., more than a month.
+You also should not postpone the offers multiple times, which will likely annoy people.
+
+\paragraph{Rejecting an offer} If you decide to reject an offer, you can say something like: \emph{``Thank you for the offer.
+I am very excited about it.
+However, I have decided to accept another offer.
+I appreciate your time and consideration.
+I hope we can work together in the future.''}  Professors will understand and will likely wish you good luck.
+If you decide to reject an offer, you should do so quickly.
+
+This is very similar to a professional job offer.
+You should be serious about it and communicate promptly.
+And most importantly, do not accept an offer and then later reject it.
+If you do this with a company, you might face legal consequences.
+If you do this with a professor, they will likely not do anything about it (i.e., not worth their time) but they will not want to work with you in the future.
+
+%1.
+I'm fine if I extend an offer and you request a delay, even if it's because you're exploring other programs.
+No issue here, it's similar to job offerings.
+Of course, I may not be able to wait indefinitely and may extend the offer to another student, but I'll let you know that.
+%2.
+I'm also okay if you find a program or professor that's a better fit.
+Similar to the first point, no hard feelings at all.
+However, I'd appreciate a prompt notification of your decision so I can extend an offer to another student.
+%3.
+I'm NOT okay if you accept my offer but then, upon finding something better, tell me of your decision to withdraw Just like in job offers, once you are committed to an offer, there should be a compelling reason to withdraw (and claiming "I found somewhere better" is the weakest excuse).
+%While I won't take any actions against you (a company might), I naturally will remember this and may be cautious in future interactions.
+I'll also reconsider making similar offers to students in similar situations.
+
+\section{Academic Integrity (Cheating and Plagiarism)}
+
+Plagiarism and cheating (e.g., exams, assignments) is a BIG no-no in the US.
+If you're caught cheating, you will face very heavy consequence and likely be expelled from the university (e.g., after the second time at GMU).
+This is quite different from many international countries where cheating is common and often tolerated.
+Faculty is extremely good at detecting cheating (we have been dealing with these situations so many times over so many years), and \emph{will} report cheating cases.
+In short, whatever you do, don't cheat---not worth it.
+
+Here is the typically steps: (i) a faculty suspecting a cheating case will report it to the Office of Academic Integrity (OAI) at the university---the report often has supporting evidence and suggested penalty (e.g., a failing grade);  (ii)  OAI will take over and investigate the case; and (iii) OAI will make the final decision.
+It is important to note that after receiving the report from your prof., OAI \emph{completely} takes over and makes decision.
+This means begging your professor will not help because they simply are no longer involved in the case.
+
+\section{Illegal Software} Using illegal/cracked software is very common in many countries (and even in the US).
+However, \emph{do not} install or use them on university computers (e.g., those given by your department or adviser).
+It is unlikely that the university will track you down, but it is the \emph{software company} that will.
+They have very sophisticated tools to detect illegal software and will sue your university/department.
+Imagine your department or adviser being sued for a large sum of money, and it is \emph{you} who caused it.
+If you need software, ask your adviser or the department.
+
+\section{Gifts} In many countries, it is customary to give professors gifts, often during holidays.
+These gifts can be costly and profs.
+sometimes expected them.
+In the US, it is uncommon and perfectly OK to not to give gifts.
+\emph{However}, if you'd like to offer small souvenir-like tokens, it's a thoughtful gesture that's appreciated.
+Some professors proudly display their gifts, which can come from students and colleagues (e.g., when they travel to their home countries or conferences).
+In summary, small gifts are fine, but avoid anything that might make your professors uncomfortable.
+
+\section{Relationships with your Profs.}
+There's a misconception that in the US it's all business, with professors as bosses who pay students for their work, and you doing something nice implying you expect something in return.
+However, the reality is quite the opposite.
+While people can be straightforward and direct, they are also friendly and informal.
+You can call your professor by their first name (\S\ref{sec:address}), disagree with them and argue (and gain respect doing so), seek their help (even on personal matters), come to their houses for parties or gathering, and give them small thoughtful gifts that they put on their desks.
+Many students and professors maintain lifelong relationships, staying in touch through cards, emails, and calls, even after their academic journey ends.% These connections are genuine and meaningful.
+
+% \begin{commentbox}
+%     [DK:] Here are a few items you may consider addressing: putting international students in touch with other students from their countries, assuring them that they would be picked up from airports upon their arrival and that their initial stay will be taken care of.
+Most universities have other resources for these, but it is worth mentioning that they would be taken care of upon arrival and can get help during the transition phase.
+Learning to cook was a big deal when I arrived over 50 years ago-August 1973.
+But things have changed as one can find many ethnic food places, a big contrast from 1973, when there were two Indian restaurants in Greater Boston area.
+% \end{commentbox}
+
+\chapter{History and Acknowledgement}\label{sec:ack}
+
+\paragraph{History} This document was conceived during a lunch with Craig Yu at GMU.
+We talked on about why GMU were not able to attract good Vietnamese and other international students, despite having a much stronger CS program than many schools that these students want to go to (part of the reason is described in \S\ref{sec:selecting-ranking-schools}).
+We wished there were a way for international students to know about the US PhD programs (as well as for US faculty to understand more about international students and therefore have better chance of recruiting and working with them).
+I was also  a member of the large VietPhD group on Facebook and saw many questions from students about PhD programs.
+However, most active participant are students in non-CS fields or not in US.
+Like ChatGPT, their answers to questions on CS PhD program in the US could be informative and helpful, but unfortunately not always accurate and sometimes leading to more confusion.
+So I thought it would be useful to have a document that is specific to CS PhD programs in the US from an insider prospective.
+
+I started writing this document in May 2023 and have been updating it since then (mostly around deadline time when I tend to procrastinate, i.e., \emph{productive procrastination}!).
+I have put the source code of this document on \href{https://github.com/nguyenthanhvuh/phd-cs-us}{GitHub} so that anyone can contribute to it.
+
+\paragraph{About Me} I am an assistant professor in the CS dept at George Mason University (GMU).
+Prior to GMU, I was at the University of Nebraska-Lincoln (UNL).
+I have been in the PhD admission process at GMU and UNL for many years.
+Currently I serve as the program director of the MS program in Software Engineering at GMU (thus also have some experience with the MS admission process--which is quite different than PhD).
+%I got my undergraduate and MS degrees from Penn State.
+I completed my PhD at the University of New Mexico-Albuquerque and a postdoc at the University of Maryland-College Park.
+
+While I am not an international student, many of my students and collaborators are.
+I also mentor multiple students from Vietnan, and have various colleague and friends who were international students.
+I have heard their stories, and hope to capture their challenges and experiences in this document so that it can be useful for prospective international students.
+
+\paragraph{Acknowledgement} Many people have contributed to this document.
+Profs. Craig Yu (GMU), Hakan Aydin (GMU), 
+Xiaokuan Zhang (GMU), Hung Le (UMass), and Deepak Kapur (UNM) provided valuable input in the early version.
+Other GMU faculty members also have provided feedback and contributions.
+Many students including Didier (GMU), Thanh (Melbourne), and Dat (Melbourne) have contributed valuable questions and feedback.
+\textbf{Thank you!}
+
+\bibliographystyle{abbrv}
+\bibliography{demystify.bib}
+
 \end{document}

### 2024-01-04 demystify.tex
updates

@@
 
-\subsection*{References}
+\subsection*{Additional Resources}
 \begin{itemize}
@@
 
-\subsubsection*{References}
+\subsubsection*{Additional Resources}
 \begin{itemize}
@@
 
-\subsubsection*{References}
+\subsubsection*{Additional Resources}
 \begin{itemize}
     \item \href{https://yonatanbisk.com/emailing_professors.html}{A Note about Emailing Professors} by Yonatan Bisk
+    \item \S\ref{sec:accept-postpone-decline} How to accept/postpone/decline PhD offers (and do it gracefully)
 \end{itemize}
@@
 I also have compiled a \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{list} of Vietnamese faculty who can advise PhD students.
+\S\ref{sec:faculty-types} talks more about who can serve as your PhD adviser.
 
@@
 
-%\chapter{References}
+%\chapter{Additional Resources}
 
@@
 
-\section{Accepting, Postponing, and Rejecting Offers}
+\section{Accepting, Postponing, and Decline Offers}\label{sec:accept-postpone-decline}
 
@@
 
-\paragraph{Rejecting an offer} If you decide to reject an offer, you can say something like: \emph{``Thank you for the offer.
+\paragraph{Declining an offer} If you decide to decline or reject an offer, you can say something like: \emph{``Thank you for the offer.
 I am very excited about it.

### 2024-01-04 demystify.tex
updates

@@
 
-This document is available at \url{nguyenthanhvuh.github.io/phd-cs-us/demystify.pdf} and its \LaTeX{} source is on \href{https://github.com/nguyenthanhvuh/phd-cs-us}{Github}.
+This document is available at
+
+\begin{center}
+\url{nguyenthanhvuh.github.io/phd-cs-us/demystify.pdf}, 
+\end{center}
+
+\noindent and its \LaTeX{} source is also on \href{https://github.com/nguyenthanhvuh/phd-cs-us}{Github}.
 If you have questions or comments, feel free to create a \href{https://github.com/nguyenthanhvuh/phd-cs-us/issues}{GitHub issue} for discussion.
@@
 There are many stories on how students are mistreated, stressed, and miserable.
-Certainly issues including bad relationships with professors, conflicts with co-authors and lab mates, and toxic working environment, feeling left out or discriminated (e.g., because you're an international student) \emph{do} exist, and it is good to be aware of those.
-However, in reality there are many good mentors, supportive lab mates and department, happy students, etc.
+Issues including bad relationships with professors, conflicts with co-authors and lab mates, feeling discriminated (e.g., because you're an international student), etc \emph{do} exist, and it is good to be aware of those.
+However, in reality there are many good mentors, supportive lab mates and department, and so on.
 So don't let social media make you feel pessimistic and deter your quest to advance knowledge.
@@
 
-This section lists some cultural issues that international students might want to pay attention to. These issues are not specific to just the PhD admission process but in general.
+This section lists some cultural issues that international students might want to pay attention to. These issues might not specific to just the PhD admission process but in general.
 
@@
 Students, especially international ones, often get confused on what to do after they get an offer from a professor, e.g., how to postpone or accept/reject offers, and most importantly, how to do so without offending anyone.
-In short, if you are not serious about an offer for any reason (e.g., still waiting for other offers), do not accept it.
+In short, if you are not serious about an offer for any reason (e.g., still waiting for other offers), do not accept, i.e., commit to it.
 If you are not sure, ask for more time to decide.
 But do not accept an offer and then later reject it.
-This is considered rude and unprofessional.
-If you do so, you will likely burn bridges with that professor.
+This is considered \emph{rude and unprofessional}, and will likely burn bridges with that professor.
 
@@
 You should be serious about it and communicate promptly.
-And most importantly, do not accept an offer and then later reject it.
+Importantly, do not accept an offer and then later reject it.
 If you do this with a company, you might face legal consequences.
@@
 \section{Illegal Software} Using illegal/cracked software is very common in many countries (and even in the US).
-However, \emph{do not} install or use them on university computers (e.g., those given by your department or adviser).
+However, \emph{do not} install or use them on university computers, even those given to you from your adviser.
 It is unlikely that the university will track you down, but it is the \emph{software company} that will.
@@
 Imagine your department or adviser being sued for a large sum of money, and it is \emph{you} who caused it.
-If you need software, ask your adviser or the department.
-
-\section{Gifts} In many countries, it is customary to give professors gifts, often during holidays.
+If you need to purchase software,  ask your adviser or the department.
+
+\section{Gifts}\label{sec:gifts} In many countries, it is customary to give professors gifts, often during holidays.
 These gifts can be costly and profs.
@@
 While people can be straightforward and direct, they are also friendly and informal.
-You can call your professor by their first name (\S\ref{sec:address}), disagree with them and argue (and gain respect doing so), seek their help (even on personal matters), come to their houses for parties or gathering, and give them small thoughtful gifts that they put on their desks.
-Many students and professors maintain lifelong relationships, staying in touch through cards, emails, and calls, even after their academic journey ends.% These connections are genuine and meaningful.
+You can call your professor by their first name (\S\ref{sec:address}), disagree with them and argue (and gain respect doing so), seek their help (even on personal matters), come to their houses for parties or gathering, and give them small thoughtful gifts that they put on their desks (\S\ref{sec:gifts}).
+Many students and professors maintain lifelong relationships, staying in touch through cards, emails, and calls, even after their academic journey ends.
+%These connections are genuine and meaningful.
 

### 2024-01-04 demystify.tex
updates

@@
 We talked on about why GMU were not able to attract good Vietnamese and other international students, despite having a much stronger CS program than many schools that these students want to go to (part of the reason is described in \S\ref{sec:selecting-ranking-schools}).
-We wished there were a way for international students to know about the US PhD programs (as well as for US faculty to understand more about international students and therefore have better chance of recruiting and working with them).
+We wished there were a way for international students to know about the US PhD programs (also for US faculty to understand more about international students and therefore have better chance of recruiting and working with them).
 I was also a member of the large VietPhD group on Facebook and saw many questions from students about PhD programs.
-However, most active participant are students in non-CS fields or not in US.
-Like ChatGPT, their answers to questions on CS PhD program in the US could be informative and helpful, but unfortunately not always accurate and sometimes leading to more confusion.
+However, most active participant are students in non-CS fields or not in US, and their answers are unfortunately not always accurate and sometimes leading to more confusion.
 So I thought it would be useful to have a document that is specific to CS PhD programs in the US from an insider prospective.
@@
 Currently I serve as the program director of the MS program in Software Engineering at GMU (thus also have some experience with the MS admission process--which is quite different than PhD).
-%I got my undergraduate and MS degrees from Penn State.
-I completed my PhD at the University of New Mexico-Albuquerque and a postdoc at the University of Maryland-College Park.
+My personal and lab website is at \href{https://dynaroars.github.io}{dynaroars.github.io}.
 

### 2024-01-05 demystify.tex
updates

@@
 
-\paragraph{Shared Spreadsheet} Many students use a spreadsheet to help them and their letter writers to keep track of their applications.
-They often use a shared spreadsheet (e.g., using Google Docs or Sheets).
+\paragraph{Shared Spreadsheet} Many students use a (shared) spreadsheet, e.g., Google Sheets/Docs, to help them and their letter writers to keep track of their applications.
 Here are some information to put on the sheet.
 \begin{itemize}
-\item \textbf{Your info}: Full name, email, phone, link to website/CV).  This is helpful for the writers just in case they want to quickly get some information about you.
+  \item \textbf{Your info}: Full name, email, phone, link to website/CV.  This is helpful for the writers just in case they want to quickly get some information about you.
+
   \item \textbf{Applications details}: University, Dept, Application System URL, Submission Deadlines, Application Status (e.g., submitted,  rejected, wait-listed, accepted)
+
   \item \textbf{LoRs}: Writer 1, Writer 2, Writer 3.

### 2024-01-09 demystify.tex
updates

@@
 \documentclass[oneside,11pt]{memoir}
-\setsecnumdepth{subsection}
-\setcounter{tocdepth}{3}
+\chapterstyle{pedersen}
+
+\usepackage[utf8]{inputenc}
 \usepackage[T1]{fontenc}
@@
 \usepackage[most]{tcolorbox}
+\setsecnumdepth{subsection}
+\setcounter{tocdepth}{3}
 \usepackage{enumitem}
@@
 \setlength\intextsep{0pt} % remove extra space above and below in-line float
-\usepackage{xcolor}
+\usepackage[svgnames]{xcolor}
 \usepackage{hyperref}
@@
 
-\title{Demystifying the Computer Science PhD Admission in the US\\{\large A Guide for International Students}}
+\title{Demystifying the Computer Science \\
+  PhD Admission in the US\\{\large A Guideline for International Students}}
 
 \author{\href{https://nguyenthanhvuh.github.io}{ThanhVu (Vu) Nguyen}\\{\small George Mason University, Dept. of Computer Science}}
-
-\chapterstyle{pedersen}
 
@@
 \makeatother
-% \author{ThanhVu H. Nguyen}
-% \title{TITLE OF BOOK}
-% \date{}
 
@@
 \begin{center}
-\url{nguyenthanhvuh.github.io/phd-cs-us/demystify.pdf}, 
+  \href{https://nguyenthanhvuh.github.io/phd-cs-us/demystify.pdf}{nguyenthanhvuh.github.io/phd-cs-us/demystify.pdf},
 \end{center}
@@
 \begin{enumerate}
-  \item Should you apply to a CS PhD in the US?
+  \item Should you apply?
         \begin{itemize}
@@
         \begin{itemize}
+                \item  Celebrating! Now it is your turn to evaluate the school!
           \item Attend \emph{Open House} to learn more about the place and \emph{interview} profs---they would be much more willing to talk to you now (\S\ref{sec:accepted}).
@@
         \begin{itemize}
-          \item TA and RA are two main sources of funding.
+          \item TA and RA are two main funding sources.
 TA (teaching assistantship) is provided by the department to help profs.
@@
 
-%\paragraph{Communicating with LoR writers and People Who Support You}
-
-%stressfull time, rely on support from others
-%ask ppl to review your SOPs
-%provide information for LORs, ..
+% \paragraph{Communicating with LoR writers and People Who Support You}
+
+% stressfull time, rely on support from others
+% ask ppl to review your SOPs
+% provide information for LORs, ..
 %
@@
 \end{table}
-%Virginia Tech\red{$^*$}  \\
+% Virginia Tech\red{$^*$}  \\
 Tab.~\ref{tab:ranking} lists the top 50 CS programs in the US from \href{https://www.csrankings.org}{CSRankings.org}, a ranking system  based on faculty publications at top CS conferences.
@@
 
-%\chapter{Additional Resources}
+% \chapter{Additional Resources}
 
@@
 This provides both concrete research experiences and potential LOR writers that can talk about your research ability.
-% %This is especially true for students from small colleges where research opportunities are limited.  
+% % This is especially true for students from small colleges where research opportunities are limited.
 % This section aims to provide some guidance on how to gain research experience as an undergraduate student.
@@
 
-%1.
+% 1.
 I'm fine if I extend an offer and you request a delay, even if it's because you're exploring other programs.
@@
 Of course, I may not be able to wait indefinitely and may extend the offer to another student, but I'll let you know that.
-%2.
+% 2.
 I'm also okay if you find a program or professor that's a better fit.
@@
 However, I'd appreciate a prompt notification of your decision so I can extend an offer to another student.
-%3.
+% 3.
 I'm NOT okay if you accept my offer but then, upon finding something better, tell me of your decision to withdraw Just like in job offers, once you are committed to an offer, there should be a compelling reason to withdraw (and claiming "I found somewhere better" is the weakest excuse).
-%While I won't take any actions against you (a company might), I naturally will remember this and may be cautious in future interactions.
+% While I won't take any actions against you (a company might), I naturally will remember this and may be cautious in future interactions.
 I'll also reconsider making similar offers to students in similar situations.

### 2024-01-09 demystify.tex
updates

@@
   \item \href{https://parentheticallyspeaking.org/articles/us-cs-phd-faq/}{Getting a Computer Science PhD in the USA} by Shriram Krishnamurthi
+  \item  \href{https://pg.ucsd.edu/PhD-application-tips.htm}{Blog post on PhD Admission} by Philip Guo
+  \item   \href{https://talkingtorobots.com/yonatanbisk.html}{PhD Admission post} by Yonatan Bisk (customized for CMU admission)
 \end{itemize}
@@
 So I will not talk too much about statements.
-In short, talk about your research goal and vision, and convince us that you can achieve it through your experience, e.g., published papers, or if you work on some projects by yourself, talk about it.
-Also, use the statement to talk about things that admission committee members might not know about, e.g., your Github project with 1K+ stars or your regular contributions to well-known open-source projects.
+In short, focus on your research goal and vision, and convince us that you can achieve it through your experience, e.g., published papers, or if you work on some projects by yourself, talk about it.
+Also, talk about things that adcom members might not know about, e.g., your Github project with 1K+ stars or your regular contributions to well-known open-source projects.
 Finally, talk about what makes you different than the other 1000 applicants that we are considering (e.g., see \S\ref{sec:improve-your-chance} for increasing your admission chance).
@@
 This shows that you're serious and have done homework on places you're applying to.
-Admission committee will look for this part at the end of the statement, so do not skip it.
+Adcom will look for this part at the end of the statement, so do not skip it.
 
@@
   \item \href{http://www.pl-enthusiast.net/2022/10/03/how-to-write-a-grad-school-personal-statement/}{How to Write a Grad School Personal Statement} by Mike Hicks
+  \item     \href{https://cs-sop.notion.site/cs-sop/CS-PhD-Statements-of-Purpose-df39955313834889b7ac5411c37b958d?p=f5d5980a71524ebaa4e6ae57266b847c&pm=s}{CS PhD SOP database} by cs-sop.org
 \end{itemize}
@@
 \begin{commentbox}[Vu:]
-  Sometime PhD admission committee in the US will share a document such as \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Foreign-Top-Schools}{this one}, which lists the top schools in several countries.
+  Sometime PhD adcom in the US will share a document such as \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Foreign-Top-Schools}{this one}, which lists the top schools in several countries.
 In some cases we ask other faculty and students if we think they know about the place.
-For example, when I was a postdoc at UMD, people in admission committee ask me to evaluate applicants from Vietnam.
+For example, when I was a postdoc at UMD, members in their CS PhD adcom ask me to evaluate applicants from Vietnam.
 During my time at UNL and now here at GMU, I have looked at Vietnamese applications (whether they are assigned to me or not) and provide input to their reviewers, e.g., X is the top tech school in Vietnam and so it should be \emph{top} instead of \emph{unknown} foreign, which makes a huge difference.
@@
 
-This can be an issue for students in many top international universities where the competition is so high that very good students can still have low rankings from these schools (and be overlooked by the admission committee).
+This can be an issue for students in many top international universities where the competition is so high that very good students can still have low rankings from these schools (and be overlooked by the adcom).
 So what to do with this? as mentioned in \S\ref{sec:your-school} you can put a note about this in your statement and ask your LoR writers to talk about it.
@@
   \textbf{Vu}: Vietnamese students and even faculty often lament how this grading system hurts Vietnamese students applying abroad.
-One way to mitigate this is making these issues known to admission committee in your statement.
+One way to mitigate this is making these issues known in your SOP.
 Universities with Vietnamese profs are probably aware of them, but in general your letter writers \emph{and} you can explicitly mention these in their letters and your statement.
@@
   If you got recruited by a prof., who would give you RA right away, it's very likely you will get admitted.
-For example, if a prof., even if not in PhD admission committee, wants to work with and funds you, the PhD admission committee will respect that decision and admit you (unless your application has many red flags).
+For example, if a prof., even if not in PhD adcom, wants to work with and funds you, the PhD adcom will respect that decision and admit you (unless your application has many red flags).
 \end{commentbox}
@@
 or TA).
-PhD admission committee members nominate applicants for this fellowship and the committee will vote and give the fellowship to the top 2.
+PhD adcom members nominate applicants for this fellowship and the committee will vote and give the fellowship to the top 2.
 \end{commentbox}

### 2024-01-20 demystify.tex
Updates from Overleaf

@@
 \setlength\intextsep{0pt} % remove extra space above and below in-line float
-\usepackage[svgnames]{xcolor}
 \usepackage{hyperref}
@@
 \usepackage{booktabs}
+\usepackage{xcolor}
 
@@
 
-\section{Choosing an Advisor}
+\section{Choosing an Adviser}
 Obviously, there is no one-size-fit-all answer to this question.
@@
 
-While I am not an international student, many of my students and collaborators are.
-I also mentor multiple students from Vietnan, and have various colleague and friends who were international students.
-I have heard their stories, and hope to capture their challenges and experiences in this document so that it can be useful for prospective international students.
+Though I'm not an international student, many of my students and collaborators are.
+I also mentor multiple students from Vietnam, and have close colleagues and friends who were once international students.
+I hope to capture the diverse challenges and experiences they've faced in this document, so that it  can be a valuable resource for prospective international students.
+Finally, my upbringing in the US provides a perspective aligned with American culture, allowing me to shed light on various issues, particularly those related to cultural differences (\S\ref{sec:cultural}).
 

### 2024-01-23 demystify.tex
Updates from Overleaf

@@
           \item Compared to other countries, CS PhD in the US does not require an MS degree but has longer PhD study time (\S\ref{sec:non-us-differences}).
-          \item Your PhD costs about \$400K in total, but you \emph{do not} pay for it (\S\ref{sec:ra-cost}).
+          \item Your 5--6 year PhD program costs about \$400K in total, but you \emph{do not} pay for it (\S\ref{sec:ra-cost}).
           \item Despite some miserable stories on social media, many PhD students have good mentors, supportive lab mates, healthy working environment ... and are happy (\S\ref{sec:happy}).
@@
 Second, applying to a good US university \emph{should not} be any harder than at schools in other countries.
-In fact, it might even be more flexible since CS PhD in the US \emph{do not} require having an MS or knowing a topic/proposal/adviser in advance (by the way, GRE is not required!).
+In fact, it might even be more flexible since CS PhD in the US \emph{do not} require having an MS or a topic/proposal/adviser in advance.
 If you believe you have a chance in other countries, e.g., South Korea, Singapore, Germany, UK, Japan, and Australia, then you will surely have a chance in the US as well.
@@
 
-After you submit your PhD application (usually November--December), it will first be screened
+After you submit your PhD application (usually around December), it will first be screened
 for general requirements, e.g., did you submit your transcripts and standard scores?
@@
 
-Then, applications are reviewed by a \textbf{PhD admission committee} (adcom) that consists of faculty members in CS (in some cases the committee can involve affiliated faculty from different disciplines).
+Then, applications are reviewed by a \textbf{PhD admission committee} (\emph{adcom}) that consists of faculty members in CS (in some cases the committee can involve affiliated faculty from different disciplines).
 These faculty have a wide-range of expertise and background to ensure diverse perspectives in the evaluation process.
@@
 without funding because you have funding on your own, e.g.,
-supported by your government or having external fellowships. We also justify
+supported by your government or having external fellowships. We justify
 our decision with a summary of your application, where we list

### 2024-01-23 demystify.tex
updates

@@
 \section{Interview}
-Sometimes a faculty wants to interview an applicant to make a decision.
-Typically, this means that they are considering admitting you.
-If they are not interested in your application, they will not proceed with an interview.
+
+Many PhD programs have the tradition of interviewing applicants.
+The most common case is that a professor is interested in working with a student and so want to chat with them to make decision (e.g., to offer them RA).
+In some case, the interview is done by several professors, e.g., to see if a student fits to their group or to recruit a very strong student to their program.
 
 Typically, an interview takes about 15-30 minutes, and one important aspect of evaluation is your ability to effectively communicate, including speaking and understanding English.
-In addition, during the interview, you will have the opportunity to ask questions about the university.
+A professor might want to know more about your research background or even ask you to read a paper (so be sure to do it!).
+Finally, during the interview, you will have the opportunity to ask questions about the university.
 It's essential to prepare some thoughtful questions, similar to a job interview.
+
+In short, getting an interview is a good sign; it means that someone is considering you. If they are not interested in your application, they will not proceed with an interview.
+However, do not worry much if you do not get an interview! For example, at GMU, most of our admitted students with TA do not go through any interviews.
 

### 2024-01-23 demystify.tex
updates

@@
           \item Grades are important, but depend on the reputation of your school (\S\ref{sec:grades}).
-          \item Getting an interview is typically a \emph{good sign} (no one interviews weak candidates).
+          \item Getting an interview is typically a \emph{good sign} (no one wants to interview weak candidates) (\S\ref{sec:interview}).
         \end{itemize}
@@
 
-\section{Interview}
+\section{Interview}\label{sec:interview}
 
 Many PhD programs have the tradition of interviewing applicants.
-The most common case is that a professor is interested in working with a student and so want to chat with them to make decision (e.g., to offer them RA).
+The most common case is that a professor is interested in working with a student and want to chat with them to make decision (e.g., to offer them RA).
+The professor might include one of their current students as part of the interview.
 In some case, the interview is done by several professors, e.g., to see if a student fits to their group or to recruit a very strong student to their program.
 
-Typically, an interview takes about 15-30 minutes, and one important aspect of evaluation is your ability to effectively communicate, including speaking and understanding English.
+Typically, an interview takes about 15--30 minutes, and one important aspect of evaluation is your ability to effectively communicate, including speaking and understanding English.
 A professor might want to know more about your research background or even ask you to read a paper (so be sure to do it!).
@@
 It's essential to prepare some thoughtful questions, similar to a job interview.
-
-In short, getting an interview is a good sign; it means that someone is considering you. If they are not interested in your application, they will not proceed with an interview.
-However, do not worry much if you do not get an interview! For example, at GMU, most of our admitted students with TA do not go through any interviews.
 
@@
 Without the interview, such applications may be more likely to be rejected.
-Overall, being selected for interview means a very good chance of being admitted.
-\end{commentbox}
+\end{commentbox}
+
+In short, getting an interview is a good sign; it means that someone is considering you. If they are not interested in your application, they will not proceed with an interview.
+However, do not worry much if you do not get an interview! For example, at GMU, most of our admitted students with TA do not go through any interviews.
 

### 2024-01-25 demystify.tex

@@
 While you might not have control over LORs or \hyperref[sec:your-school]{where your go to school}, you do over your
-statement of purpose (SOP) or personal statement! So write it well because it could make a big difference.
+statement of purpose (SOP) or personal statement\footnote{Some schools separate these documents and ask you to write both: SOP, which focuses on research experiences, and Personal Statement, which is everything more personal, e.g., why PhD, challenges, diversity, etc}!
+So write it well because it could make a big difference.
 In your statement, you have the opportunity to make your application stand out and unique, which can make you \emph{fit} the CS program you're applying to, even if you don't have very strong research experience.

### 2024-01-25 demystify.tex
Updates from Overleaf

--- 2024-01-23 demystify.tex
@@
 \section{Interview}
-
-Many PhD programs have the tradition of interviewing applicants.
-The most common case is that a professor is interested in working with a student and so want to chat with them to make decision (e.g., to offer them RA).
-In some case, the interview is done by several professors, e.g., to see if a student fits to their group or to recruit a very strong student to their program.
+Sometimes a faculty wants to interview an applicant to make a decision.
+Typically, this means that they are considering admitting you.
+If they are not interested in your application, they will not proceed with an interview.
 
 Typically, an interview takes about 15-30 minutes, and one important aspect of evaluation is your ability to effectively communicate, including speaking and understanding English.
-A professor might want to know more about your research background or even ask you to read a paper (so be sure to do it!).
-Finally, during the interview, you will have the opportunity to ask questions about the university.
+In addition, during the interview, you will have the opportunity to ask questions about the university.
 It's essential to prepare some thoughtful questions, similar to a job interview.
-
-In short, getting an interview is a good sign; it means that someone is considering you. If they are not interested in your application, they will not proceed with an interview.
-However, do not worry much if you do not get an interview! For example, at GMU, most of our admitted students with TA do not go through any interviews.
 

### 2024-01-27 demystify.tex
Updates from Overleaf

--- 2024-01-25 demystify.tex
@@
 applicants: \textbf{customize the statement} for the school you're applying to,
-e.g., why do you apply \emph{here}? provide names of professors who you're interested in (in many cases your application will be forwarded to them for evaluation).
+e.g., why do you apply \emph{here}?
+provide names of professors who you're interested in (if they are not already in the adcom, your application might get forwarded to them for evaluation; and they will likely be those who interview you).
 This shows that you're serious and have done homework on places you're applying to.
@@
   \tcblower
-  Do careful research on professors, don't mention \emph{emeritus} or  adjunct faculty.
+  Do careful research on professors, don't mention \emph{emeritus} or  adjunct faculty (see more about various types of faculty in \S\ref{sec:faculty-types}).
   Also, be careful not to send statements to wrong schools or mixing
@@
 However, if committee members do not know much about schools in your country, they will likely treat your school as
-\emph{``unknown foreign''}, which can be a minus point because we are uncertain about the quality of your school.
+\emph{``unknown foreign''}, which can be a minus point because they are uncertain about the quality of your school.
 

### 2024-01-27 demystify.tex
updates

@@
 \def\Section{\S}
+\renewcommand{\figureautorefname}{Fig.}
+\renewcommand{\tableautorefname}{Tab.}
+\makeatletter
+\renewcommand{\chapterautorefname}{\S\@gobble}
+\renewcommand{\sectionautorefname}{\S\@gobble}
+\renewcommand{\subsectionautorefname}{\S\@gobble}
+\makeatother
+
 \newcommand{\mycomment}[3][\color{blue}]{{#1{{#2}: {#3}}}}
@@
 % though \emph{very} top schools could be very selective, e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU
-So I want to share some details about the PhD admission process and advice for those who are interested in applying for a \textbf{PhD in Computer Science in the US}.
-While this document is primarily intended for students interested in CS, it may be relevant to students from various disciplines. 
+So I want to share some details about the admission process and advice for those who are interested in applying for a \textbf{PhD in Computer Science in the US}.
+While this document is primarily intended for students interested in CS, it might be relevant to students from various disciplines.
 Furthermore, although many examples are specifics for schools that I and other contributors of this document know about, the information should be generalizable to other R1\footnote{An \href{https://en.wikipedia.org/wiki/List_of_research_universities_in_the_United_States}{R1 institution} in the US is a research-intensive university with a high level of research activity across various disciplines.
@@
 
-The main points of this guideline are summarized here. This gives you an overview to decide which specific topics you want to explore more thoroughly.
+Below we summarize the main points of this guideline. This gives you an overview to decide which specific topics you want to explore more thoroughly.
 
@@
         \begin{itemize}
-          \item \emph{Yes, definitely}.  CS PhD study in the US is fully funded and admission into good universities is not any harder than non-US schools (\S\ref{sec:should}).
+          \item \emph{Yes, definitely}.  CS PhD study in the US is fully funded and admission into good universities is not any harder than non-US schools (\autoref{sec:should}).
         \end{itemize}
@@
         \begin{itemize}
-          \item Applications are evaluated by the \emph{PhD Admission} committee and each application is reviewed by typically three faculty (\S\ref{sec:evalapps}).
+          \item Applications are evaluated by the \emph{PhD Admission} committee and each application is reviewed by typically three faculty (\autoref{sec:evalapps}).
           \item Individually faculty \emph{cannot directly admit} a student---so do not email and ask if you have a chance.
-However, faculty can \emph{advocate} for a student and therefore increase their admission chance---so contact and introduce yourself (\S\ref{sec:contact}).
+However, faculty can \emph{advocate} for a student and therefore increase their admission chance---so do contact and introduce yourself (\autoref{sec:contact}).
          \end{itemize}
@@
         \begin{itemize}
-          \item The committee will look at various factors, but the most important ones are research ability, e.g., publications, personal statement, and recommendation letters.
-          \item LORs are very important, but only if they are personalized and talk about your research ability (\S\ref{sec:lor}).
-          \item Personal statement is very important. Write it in such a way that makes you \emph{stand out} (\S\ref{sec:research-statement} and \S\ref{sec:improve-your-chance})
-          \item GRE \emph{is not} required (\S\ref{sec:grades}). Spend your time on something else.
-          \item Grades are important, but depend on the reputation of your school (\S\ref{sec:grades}).
-          \item Getting an interview is typically a \emph{good sign} (no one wants to interview weak candidates) (\S\ref{sec:interview}).
+          \item The committee will look at various factors, but the most important ones are research ability, e.g., publications, statement of purpose (SOP), and  letters of recommendation (LORs).
+          \item LORs are very important, but only if they are personalized and talk about your research ability (\autoref{sec:lor}).
+          \item SOP is very important. Write it in such a way that makes you \emph{stand out} (\autoref{sec:research-statement} and \autoref{sec:improve-your-chance})
+          \item GRE \emph{is not} required (\autoref{sec:grades}). Spend your time on something else!
+          \item Grades are important, but depend on the reputation of your school (\autoref{sec:grades}).
+          \item Getting an interview is typically a \emph{good sign} as no one wants to interview weak candidates (\autoref{sec:interview}).
         \end{itemize}
@@
                 \item  Celebrating! Now it is your turn to evaluate the school!
-          \item Attend \emph{Open House} to learn more about the place and \emph{interview} profs---they would be much more willing to talk to you now (\S\ref{sec:accepted}).
+          \item Attend \emph{Open House} to learn more about the place and \emph{interview} profs---they would be much more willing to talk to you now (\autoref{sec:accepted}).
         \end{itemize}
@@
 RA (research assistantship) is provided by profs.
-to help them with their research (\S\ref{sec:funding}).
+to help with their research (\autoref{sec:funding}).
         \end{itemize}
@@
         \begin{itemize}
-          \item Many schools do not offer PhD studies in CS and many CS professors \emph{cannot} advise or graduate PhD students  (\S\ref{sec:schoolsandprofs}).
-          \item Contacting a prof. for research opportunities is recommended, but do it \emph{properly} (\S\ref{sec:contact}).
+          \item Many schools do not offer PhD studies in CS and many CS professors \emph{cannot} formally advise or graduate PhD students  (\autoref{sec:schoolsandprofs}).
+          \item Contacting a prof. for research opportunities is recommended, but do it \emph{properly} (\autoref{sec:contact}).
         \end{itemize}
@@
         \begin{itemize}
-          \item Increasing your admission chance by being unique and standing out (\S\ref{sec:improve-your-chance}).
-          \item You can successfully apply to CS PhD even if you have non-STEM background (\S\ref{sec:non-stem}).
-          \item Compared to other countries, CS PhD in the US does not require an MS degree but has longer PhD study time (\S\ref{sec:non-us-differences}).
-          \item Your 5--6 year PhD program costs about \$400K in total, but you \emph{do not} pay for it (\S\ref{sec:ra-cost}).
-          \item Despite some miserable stories on social media, many PhD students have good mentors, supportive lab mates, healthy working environment ... and are happy (\S\ref{sec:happy}).
+          \item Increasing your admission chance by being unique and standing out (\autoref{sec:improve-your-chance}).
+          \item You can successfully apply to CS PhD even if you have non-STEM background (\autoref{sec:non-stem}).
+          \item Compared to other countries, CS PhD in the US does not require an MS degree but has longer PhD study time (\autoref{sec:non-us-differences}).
+          \item Your 5--6 year PhD program costs about \$400K in total, but you \emph{do not} pay for it (\autoref{sec:ra-cost}).
+          \item Despite some miserable stories on social media, many PhD students have good mentors, supportive lab mates, healthy working environment ... and are happy (\autoref{sec:happy}).
         \end{itemize}
@@
 health insurance, and stipend (monthly salary). Moreover, depending on the university,
-you may even receive additional benefits such as summer pay, laptops, (conference/workshop) traveling. \S\ref{sec:funding} provides more details on funding.
+you may even receive additional benefits such as summer pay, laptops, (conference/workshop) traveling. \autoref{sec:funding} provides more details on funding.
 
@@
 If you believe you have a chance in other countries, e.g., South Korea, Singapore, Germany, UK, Japan, and Australia, then you will surely have a chance in the US as well.
-\S\ref{sec:non-us-differences} compares CS PhD study in the US to other countries.
+\autoref{sec:non-us-differences} compares CS PhD study in the US to other countries.
 
 \begin{commentbox}[Vu:]
-  One of the reasons I create this document is that my colleagues at GMU are interested in recruiting Vietnamese students and are surprised when seeing very few applications in Vietnam (see \S\ref{sec:ack}).
+  One of the reasons I create this document is that my colleagues at GMU are interested in recruiting Vietnamese students and are surprised when seeing very few applications in Vietnam (see \autoref{sec:ack}).
 Each year our CS program receives more than 350 PhD applications, most of which are international but only 3--4 are from Vietnam.
@@
 \begin{domesticbox}[Domestic students:] If you're a domestic student, you have several advantages in your application.
-First, we already know about your school (see \S\ref{sec:your-school}), and you are also more familiar with the US education system and culture.
+First, we already know about your school (see \autoref{sec:your-school}), and you are also more familiar with the US education system and culture.
 You also have more opportunities for funding, e.g., through government scholarships for US citizens and residents.
@@
 
-The PhD adcom typically involves assistant professors in the department (see \S\ref{sec:faculty-types} for various faculty titles).
+The PhD adcom typically involves assistant professors in the department (see \autoref{sec:faculty-types} for various faculty titles).
 This provides junior faculty the opportunities to recruit students.
@@
 
-At GMU, we usually decide that a full-time PhD candidate is either (i) admit with funding (TA or RA, see \S\ref{sec:funding}) or (ii) rejected. In other words, in most cases, we either
+At GMU, we usually decide that a full-time PhD candidate is either (i) admit with funding (TA or RA, see \autoref{sec:funding}) or (ii) rejected. In other words, in most cases, we either
 admit you with full funding, or we reject your application. In some rare cases, we admit
@@
 Thus, general writings, even those under submissions or rejected, would still be good (and much better than nothing).
-But be sure to upload your papers with your application and talk about them in your statement (see \S\ref{sec:research-statement}).
+But be sure to upload your papers with your application and talk about them in your statement (see \autoref{sec:research-statement}).
 Note that local conferences and non-English journals or conferences do
@@
 % In CS, conferences are important and top conferences are prestigious.
-You can find top CS conferences at places such as \href{https://csrankings.org}{CSRankings}, which ranks CS programs based on how their faculty publish at top \emph{conferences} (see more in \S\ref{sec:ranking}).
+You can find top CS conferences at places such as \href{https://csrankings.org}{CSRankings}, which ranks CS programs based on how their faculty publish at top \emph{conferences} (see more in \autoref{sec:ranking}).
 
@@
 But as mentioned, if these professors do not know you, the letters are generic and carry little value, in fact, \red{red flags}.
-Moreover, a top professor in Vietnam might not be well-known to US faculty (see more details in \S\ref{sec:your-school}).
+Moreover, a top professor in Vietnam might not be well-known to US faculty (see more details in \autoref{sec:your-school}).
 So save the trouble and just get letters from \emph{any} professors/supervisors who knows you well and can write a good letter about \emph{your} research ability.
@@
 In your statement, you have the opportunity to make your application stand out and unique, which can make you \emph{fit} the CS program you're applying to, even if you don't have very strong research experience.
-A well-written SOP also shows that you can communicate, which is very important in research, and that you can effectively teach and communicate with students, which is important for TA funding (see \S\ref{sec:funding}).
+A well-written SOP also shows that you can communicate, which is very important in research, and that you can effectively teach and communicate with students, which is important for TA funding (see \autoref{sec:funding}).
 % is important because if you need (GTA) funding, it will provide evidence
@@
 Also, talk about things that adcom members might not know about, e.g., your Github project with 1K+ stars or your regular contributions to well-known open-source projects.
-Finally, talk about what makes you different than the other 1000 applicants that we are considering (e.g., see \S\ref{sec:improve-your-chance} for increasing your admission chance).
+Finally, talk about what makes you different than the other 1000 applicants that we are considering (e.g., see \autoref{sec:improve-your-chance} for increasing your admission chance).
 
@@
   \tcblower
-  Do careful research on professors, don't mention \emph{emeritus} or  adjunct faculty (see more about various types of faculty in \S\ref{sec:faculty-types}).
+  Do careful research on professors, don't mention \emph{emeritus} or  adjunct faculty (see more about various types of faculty in \autoref{sec:faculty-types}).
   Also, be careful not to send statements to wrong schools or mixing
@@
 Graduating from top universities \emph{that we recognize} helps. For example, if your school is well-known, then it is \emph{``top foreign''}, which is definitely a plus.
-However, if committee members do not know much about schools in your country, they will likely treat your school as
-\emph{``unknown foreign''}, which can be a minus point because they are uncertain about the quality of your school.
+However, if we do not know much about schools in your country, then we are uncertain about the quality of your school and  likely treat your school as \emph{``unknown foreign''}, which can be a minus point.
 
@@
 This can be an issue for students in many top international universities where the competition is so high that very good students can still have low rankings from these schools (and be overlooked by the adcom).
-So what to do with this? as mentioned in \S\ref{sec:your-school} you can put a note about this in your statement and ask your LoR writers to talk about it.
+So what to do with this? as mentioned in \autoref{sec:your-school} you can put a note about this in your statement and ask your LoR writers to talk about it.
 
@@
 
-In short, getting an interview is a good sign; it means that someone is considering you. If they are not interested in your application, they will not proceed with an interview.
+In short, getting an interview is a good sign; it means that someone is considering you. If we are not interested in your application, we will not proceed with an interview.
 However, do not worry much if you do not get an interview! For example, at GMU, most of our admitted students with TA do not go through any interviews.
@@
 
-        You can then share this sheet with your reference writers and remind them to submit LoRs on your behalf (see \S\ref{sec:lor}).
+        You can then share this sheet with your reference writers and remind them to submit LoRs on your behalf (see \autoref{sec:lor}).
 \end{itemize}
@@
 Now you're at a different game because the schools that have admitted you will try to get you to accept them!
-Important factors to consider are the reputation of schools and professors (\S\ref{sec:schoolsandprofs}), and funding availability (\S\ref{sec:funding}).
+Important factors to consider are the reputation of schools and professors (\autoref{sec:schoolsandprofs}), and funding availability (\autoref{sec:funding}).
 You will have to make your decision by around \emph{April 15}.
@@
 Finally, fellowship is an independent funding that can come from a school, a company, or an organization.
-Tab.~\ref{tab:funding} summarizes the differences.
+\autoref{tab:funding} summarizes the differences.
 Note that funding is typically more available for PhD students than 
@@
 
-For a full breakdown of how much a graduate student costs, see \S\ref{sec:ra-cost}.
+For a full breakdown of how much a graduate student costs, see \autoref{sec:ra-cost}.
 
@@
 RA is provided through a professor through their own funding so you can work on their project.  
-You do not need to teach as an RA, so you can focus on your research. Depending on the professor, RA may be available during the summer. \S\ref{sec:ra-cost} gives more details on RA budget.
+You do not need to teach as an RA, so you can focus on your research. Depending on the professor, RA may be available during the summer. \autoref{sec:ra-cost} gives more details on RA budget.
 
@@
 See how to contact a prof.
-in \S\ref{sec:contact} for research opportunities.
+in \autoref{sec:contact} for research opportunities.
 
@@
 While this can be confusing due to the large number of universities in the US, a little research, e.g., searching for PhD in CS from the school website, will help you find out.
-All schools listed in \S\ref{sec:ranking} have PhD studies in CS, so you can start there.
+All schools listed in \autoref{sec:ranking} have PhD studies in CS, so you can start there.
 
@@
 This is also a good way to learn about individual faculty (who works on what) and well-known CS conferences\footnote{In CS (and probably only in CS), conferences, not journals, are often the main venue to publish research finding.}.
-\S\ref{sec:ranking} gives the top 50 CS programs in the US according to CSRankings.
+\autoref{sec:ranking} gives the top 50 CS programs in the US according to CSRankings.
 
@@
 You can even ask to contact their students.
-See more in \S\ref{sec:accepted} on what to do after getting admitted.
+See more in \autoref{sec:accepted} on what to do after getting admitted.
 
@@
 This is difficult and takes time.
-Fortunately, unlike many countries that require finding an adviser and research topic before starting the PhD (\S\ref{sec:non-us-differences}), CS PhD programs in the US will typically give you a couple of years to search for advisers and research topics.
-This is especially true if you're admitted with TA (\S\ref{sec:ta}), which gives you time to explore and find adviser.
+Fortunately, unlike many countries that require finding an adviser and research topic before starting the PhD (\autoref{sec:non-us-differences}), CS PhD programs in the US will typically give you a couple of years to search for advisers and research topics.
+This is especially true if you're admitted with TA (\autoref{sec:ta}), which gives you time to explore and find adviser.
 
@@
 In any case, you should contact the professor and talk to them.
-For example, you can (cold) email them and ask about their research and if they are taking new students (\S\ref{sec:contact}), and schedule appointments to chat with them (office hrs are often the best time).
+For example, you can (cold) email them and ask about their research and if they are taking new students (\autoref{sec:contact}), and schedule appointments to chat with them (office hrs are often the best time).
 
@@
   \tcblower
-  \textbf{Vu}: \S\ref{sec:contact} shows how to contact a professor for research opportunities.
+  \textbf{Vu}: \autoref{sec:contact} shows how to contact a professor for research opportunities.
 Many probably will say no (or do not reply) as they do not have the bandwidth to take on random students, but some may say yes if they see potential fit.
@@
 Let's try to understand the different types of faculty and their roles.
-For example, you might hear about tenured,  tenure track, teaching faculty.
-You might also things such as assistant, associate, full, adjunct, emeritus, teaching, research professors.
+For example, you might hear about tenured,  tenure track, and teaching faculty.
+You might also hear about assistant, associate, full, adjunct, emeritus, teaching, research professors.
 Here is a primer on these terminologies.
@@
 This is very important because you need to have a tenure line faculty as your main PhD adviser.
-\S\ref{sec:tenure-vs-tenure-track} talks more about choosing between tenure and tenured-track professor as your adviser.
+\autoref{sec:tenure-vs-tenure-track} talks more about choosing between tenure and tenured-track professor as your adviser.
 
 In contrast, \emph{teaching} (or instructional faculty or professor of practice) mainly focus on teaching.
-They typically teach 3--4 classes per semester (which is quite a lot) and \emph{do not} have research responsibilities (e.g., they do not have to worry much about publishing papers, obtaining grants).
+They typically teach 3--4 classes per semester (which is quite a lot) and \emph{do not} have research responsibilities (e.g., they do not have to worry much about publishing papers or obtaining grants).
 They can mentor Ph.D.
@@
 
-It is worth noting that professors in a non-CS department also unlikely can serve as formal PhD advisers for CS students (see \S\ref{sec:related-fields}).
-
-\paragraph{Titles} Faculty have titles, e.g., assistant, associate, full, regardless if they are tenure line or teaching.
+It is worth noting that professors in a non-CS department also unlikely can serve as formal PhD advisers for CS students (see \autoref{sec:related-fields}).
+
+\paragraph{Titles} Faculty have rankings, e.g., assistant, associate, full, regardless if they are tenure line or teaching.
 Assistant means new faculty, associate means they have been promoted, and full means they are senior.
@@
 
-Adjunct faculty not full-time faculty, e.g., they might be working in industry and teach a class or two for fun.
+Adjunct faculty is not full-time, e.g., they might be working in industry and teach a class or two for fun.
 Emeritus means they are retired but still have some affiliation with the university.
-Research faculty (or scientist) are typically non-tenure line faculty who focus on research.
+Research faculty (or research scientist) are typically non-tenure line faculty who focus on research.
 Due to their roles, adjunct, emeritus, and research faculty typically do not advise Ph.D.
@@
 to \emph{ask about your admission chance}, please \textbf{don't}.
-We don't know and can't answer because as explained in \S\ref{sec:evalapps}, we don't make individual decisions and might not even be assigned to evaluate your application.
+We don't know and can't answer because as explained in \autoref{sec:evalapps}, we don't make individual decisions and might not even be assigned to evaluate your application.
 It is the same as sending a paper draft to a journal editor and ask them if your paper has a chance.
@@
   \item \href{https://yonatanbisk.com/emailing_professors.html}{A Note about Emailing Professors} by Yonatan Bisk
-  \item \S\ref{sec:accept-postpone-decline} How to accept/postpone/decline PhD offers (and do it gracefully)
+  \item \autoref{sec:accept-postpone-decline} How to accept/postpone/decline PhD offers (and do it gracefully)
 \end{itemize}
@@
 I also have compiled a \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{list} of Vietnamese faculty who can advise PhD students.
-\S\ref{sec:faculty-types} talks more about who can serve as your PhD adviser.
+\autoref{sec:faculty-types} talks more about who can serve as your PhD adviser.
 
@@
 Yes, \emph{as long as} you can demonstrate you are ready for CS PhD research through research experiences, LoRs, statements, etc as mentioned.
-You might be even able to leverage this to make your profile stand out as mentioned in \S\ref{sec:improve-your-chance}.
+You might be even able to leverage this to make your profile stand out as mentioned in \autoref{sec:improve-your-chance}.
 
@@
 
-\emph{MS Degree requirement}:  as mentioned in \S\ref{sec:msrequirement} and \ref{sec:time}, CS PhD programs in the US do not require MS degree.
+\emph{MS Degree requirement}:  as mentioned in \autoref{sec:msrequirement} and \ref{sec:time}, CS PhD programs in the US do not require MS degree.
 In contrast, many other countries do require MS degree before joining a PhD program.
@@
 These nuances represent some cultural differences that you may encounter and will gradually adapt to.
-More on cultural differences in \S\ref{sec:cultural}.
+More on cultural differences in \autoref{sec:cultural}.
 
@@
 
-Tab.~\ref{tab:cost} shows the budget breakdown for a GRA per year (this level of details is what faculty actually uses when applying for grants).
+\autoref{tab:cost} shows the budget breakdown for a GRA per year (this level of details is what faculty actually uses when applying for grants).
 These numbers are based on my experience at public universities in the US.
@@
     16 & University of Pennsylvania &41 & Harvard University \\
-    17 & Columbia University \red{$^*$} &42 & Boston University  \\
-    18 & Princeton University  & 43& Rice University\\
+    17 & Columbia University\red{$^*$} &42 & Boston University  \\
+    18 & Princeton University\red{$^*$}  & 43& Rice University\\
     19 & New York University  & 44&  University at Buffalo\red{$^*$}\\
@@
 % Virginia Tech\red{$^*$}  \\
-Tab.~\ref{tab:ranking} lists the top 50 CS programs in the US from \href{https://www.csrankings.org}{CSRankings.org}, a ranking system  based on faculty publications at top CS conferences.
+\autoref{tab:ranking} lists the top 50 CS programs in the US from \href{https://www.csrankings.org}{CSRankings.org}, a ranking system  based on faculty publications at top CS conferences.
 
@@
 Finally, you can also ask your peers who are already doing research.
-You can use the method described in \S\ref{sec:contact} to contact professors
+You can use the method described in \autoref{sec:contact} to contact professors
 
@@
 While people can be straightforward and direct, they are also friendly and informal.
-You can call your professor by their first name (\S\ref{sec:address}), disagree with them and argue (and gain respect doing so), seek their help (even on personal matters), come to their houses for parties or gathering, and give them small thoughtful gifts that they put on their desks (\S\ref{sec:gifts}).
+You can call your professor by their first name (\autoref{sec:address}), disagree with them and argue (and gain respect doing so), seek their help (even on personal matters), come to their houses for parties or gathering, and give them small thoughtful gifts that they put on their desks (\autoref{sec:gifts}).
 Many students and professors maintain lifelong relationships, staying in touch through cards, emails, and calls, even after their academic journey ends.
@@
 \paragraph{History} This document was conceived during a lunch with Craig Yu at GMU.
-We talked on about why GMU were not able to attract good Vietnamese and other international students, despite having a much stronger CS program than many schools that these students want to go to (part of the reason is described in \S\ref{sec:selecting-ranking-schools}).
+We talked on about why GMU were not able to attract good Vietnamese and other international students, despite having a much stronger CS program than many schools that these students want to go to (part of the reason is described in \autoref{sec:selecting-ranking-schools}).
 We wished there were a way for international students to know about the US PhD programs (also for US faculty to understand more about international students and therefore have better chance of recruiting and working with them).
@@
 I hope to capture the diverse challenges and experiences they've faced in this document, so that it  can be a valuable resource for prospective international students.
-Finally, my upbringing in the US provides a perspective aligned with American culture, allowing me to shed light on various issues, particularly those related to cultural differences (\S\ref{sec:cultural}).
+Finally, my upbringing in the US provides a perspective aligned with American culture, allowing me to shed light on various issues, particularly those related to cultural differences (\autoref{sec:cultural}).
 

### 2024-02-02 demystify.tex
updates

--- 2024-01-27 demystify.tex
@@
 While you might not have control over LORs or \hyperref[sec:your-school]{where your go to school}, you do over your
-statement of purpose (SOP) or personal statement\footnote{Some schools separate these documents and ask you to write both: SOP, which focuses on research experiences, and Personal Statement, which is everything more personal, e.g., why PhD, challenges, diversity, etc}!
-So write it well because it could make a big difference.
+statement of purpose (SOP) or personal statement! So write it well because it could make a big difference.
 In your statement, you have the opportunity to make your application stand out and unique, which can make you \emph{fit} the CS program you're applying to, even if you don't have very strong research experience.
@@
 applicants: \textbf{customize the statement} for the school you're applying to,
-e.g., why do you apply \emph{here}?
-provide names of professors who you're interested in (if they are not already in the adcom, your application might get forwarded to them for evaluation; and they will likely be those who interview you).
+e.g., why do you apply \emph{here}? provide names of professors who you're interested in (in many cases your application will be forwarded to them for evaluation).
 This shows that you're serious and have done homework on places you're applying to.
@@
   \tcblower
-  Do careful research on professors, don't mention \emph{emeritus} or  adjunct faculty (see more about various types of faculty in \S\ref{sec:faculty-types}).
+  Do careful research on professors, don't mention \emph{emeritus} or  adjunct faculty.
   Also, be careful not to send statements to wrong schools or mixing
@@
 However, if committee members do not know much about schools in your country, they will likely treat your school as
-\emph{``unknown foreign''}, which can be a minus point because they are uncertain about the quality of your school.
+\emph{``unknown foreign''}, which can be a minus point because we are uncertain about the quality of your school.
 
@@
 Many PhD programs have the tradition of interviewing applicants.
-The most common case is that a professor is interested in working with a student and want to chat with them to make decision (e.g., to offer them RA).
+The most common case is that a professor is interested in working with a student and wants to chat with them to make decision (e.g., to offer them RA).
 The professor might include one of their current students as part of the interview.
-In some case, the interview is done by several professors, e.g., to see if a student fits to their group or to recruit a very strong student to their program.
+In some case, the interview is done by several professors, e.g., to see if a student would fit in their group or to recruit a very strong student to their program.
 

### 2024-02-10 demystify.tex
Updates from Overleaf

@@
 \usepackage{xcolor}
+\usepackage[makeroom]{cancel}
 
@@
 \title{Demystifying the Computer Science \\
-  PhD Admission in the US\\{\large A Guideline for International Students}}
+  PhD Admission in the US\\{\large A Guideline for International \emph{and Domestic} Students}}
 
@@
 Having been involved in PhD admissions for many years, I've
-realized that many \textbf{international students}, especially those in  smaller countries or less well-known universities, lack a clear understanding of
+realized that many \textbf{international students} (and also domestic ones), especially those in  smaller countries or less well-known universities, lack a clear understanding of
 the Computer Science PhD admission process at US universities. This confusion not only

### 2024-02-10 demystify.tex
updates

@@
 
-Most PhD programs will require at least \textbf{two LORs}. Having a letter from an internationally recognized researcher can greatly strengthen your application. However, obtaining such letters
+Most PhD programs will require at least \textbf{two LORs}.
+Having a letter from an internationally recognized researcher will greatly strengthen your application.
+In fact, a strong letter from a well-known researcher can outweigh the lack of publications.
+
+However, obtaining such letters
 can be challenging for international students, who might not have much interactions with such experts.
@@
 Note that funding is typically more available for PhD students than 
-Masters. %Undergraduates typically have no funding and also have to pay international tuition, which is very expensive!.
+Masters.
 
@@
 Both TA and RA come with tuition waiving (you don't have to pay tuition), health insurance (this takes care of your insurance, which is a must have in the US), and most importantly, your stipend (i.e., your salary).
-Some universities also pay insurance for spouse/children (or give very good discount).
+Some universities also give significant discount or pay insurance for spouse/children.
 
 Several things about stipends.
-First, the amount of stipend depends on the university, which in turn depends on various factors such as location (e.g., a stipend in Washington DC is likely higher than in Lincoln, Nebraska due to higher living cost).
-Second, a school year is typically 9-month in the US, so stipend is for 9 months (so divide by 9 for each month).
-Third, like most source of income in the US, you will have to pay tax on your stipend.
-Fourth, CS department typically has higher stipend comparing to other study fields.
-Finally, private universities might pay more for stipend (but they could have higher "activity" or some other hidden fee, or you may be required to pay fees for each credit hour).
+First, the amount of stipend \emph{varies} and depends on various factors such as location (e.g., a stipend in Washington DC is likely higher than in Lincoln, Nebraska due to higher living cost).
+Second, an academic year (AY)  year is typically \emph{9-month} in the US, so stipend is for 9 months (so divide by 9 for each month).
+Third, you might get paid over the summer through funding from your professor or fellowship (typicaly no TA over the summer).
+Fourth, like most source of income in the US, you will have to \emph{pay tax} on your stipend.
+Finally, private universities might pay more for stipend, but they could have higher "activity" or some other hidden fee, or you may be required to pay fees for each credit hour.
 
@@
 
-Fellowship is another type of funding that students can apply for (e.g., from school, industries, government).
+Fellowship is another type of funding that students can get (e.g., from the university, industries, government).
 Fellowships are typically competitive and generous, and gives pretty much all benefits tuition/insurance that a TA/RA has.
@@
 
-In general, fellowship is prestigious, and you will stand out if you get one.
-Every PhD student has pubs, but only superstars have NSF grad or Microsoft fellowship.
-In fact, these are so prestigious that even if you didn't get it but make it to the final round, school will still mention you on their website and you still should put it on your CV.
-
-\textbf{How to get Fellowship?} You need to apply for them.
-The US government has many fellowships, though they would likely require US citizenship or residency.
-However, tech companies including Google, Microsoft, Facebook have fellowships that international students can apply for.
-
-Prestigious fellowships typically require a clear and good research plan, so it is a good idea to wait until at least your second year to have research experience and even publication before applying.
+\textbf{How to get Fellowship?} You need to apply for them.  Many schools provide such fellowships to attract admitted students.
+External fellowships including from the US government and tech companies (e.g., Google, Microsoft, Facebook).
+Fellowships from US government would likely require US citizenship or residency.
+% However, tech companies including Google, Microsoft, Facebook have fellowships that international students can apply for.
+
+Prestigious external fellowships typically require a clear and good research plan, so it is a good idea to wait until at least your second year to have research experience and even publication before applying.
 Remember, you're competing with the top PhD students at top universities worldwide.
@@
 or TA).
-PhD adcom members nominate applicants for this fellowship and the committee will vote and give the fellowship to the top 2.
-\end{commentbox}
+PhD adcom members nominate applicants for this fellowship and the whole committee will decide.
+\end{commentbox}
+
+In general, fellowship, especially competitive ones from the government or company, is prestigious, and you will stand out if you get one.
+Every PhD student has pubs, but only superstars have NSF grad or Microsoft fellowship.
+In fact, these are so prestigious that even if you didn't get it but make it to the final round or even ``honorable mentioning'', school will still mention you on their website and you still should put it on your CV.
 

### 2024-02-10 demystify.tex
updates

--- 2024-02-02 demystify.tex
@@
 
-The most effective evidence of research ability is having \textbf{published papers in reputable international journals or conferences}.
-Having papers published is a sign that the applicant has successfully involved in research.
+The most concrete evidence of research ability is having \textbf{published papers in reputable international journals or conferences}.
+Having published papers, especially at top venues, is a sign that the applicant has successfully involved in research.
 
@@
 not carry as much weight since their quality is often unknown to US faculty.
-However, if you have published in such places, you should still mention them in your statement and explain why they are good.
+However, if you have published in such places, you should still upload them, mention them in your statement, and explain why they are good.
 
@@
 
-Most PhD programs will require at least \textbf{two LORs}. Having a letter from an internationally recognized researcher can greatly strengthen your application. However, obtaining such letters
+Most PhD programs will require at least \textbf{two LORs}.
+Having a letter from an internationally recognized researcher can \textbf{greatly} strengthen your application.
+In fact, it is arguable at many top CS programs, strong LoRs from well-known researchers are even more important than published papers.
+
+However, obtaining such letters
 can be challenging for international students, who might not have much interactions with such experts.
@@
 
-\paragraph{Waiving your Right}  Choosing not to look at a reference letter is pretty standard in school and job applications.
+\paragraph{Waiving Your Right}  Choosing not to look at a reference letter is pretty standard in school and job applications.
 When you waive your right to see the letter, it adds a layer of trust, showing you're confident in your choice of referees and that you're not trying to twist their words.
@@
 
-\paragraph{VEF 2.0} For Vietnamese students, it's worth mentioning about the \href{https://vef2.org/}{VEF2.0 program}, which has helped many good students in gaining acceptance to top PhD programs in the US.
+\paragraph{VEF 2.0} For Vietnamese students, it's worth mentioning about the \href{https://vef2.org/}{VEF2.0 program}, which has helped many good students in gaining admission to top CS PhD programs.
 VEF2.0 follows an interesting model where US faculty members from leading institutions are invited to conduct rigorous interviews with VEF students and subsequently provide reference letters on their behalf.
-Despite the limited interaction between the interviewer and the interviewee (primarily confined to the interview itself), these reference letters are generally helpful as they have helped many students getting into top PhD programs in the US.
+Despite the limited interaction between the interviewer and the interviewee (primarily confined to the interview itself), these reference letters are generally effective as they have helped many students getting into top PhD programs in the US.
 
@@
 
-\section{Interview}\label{sec:interview}
-
-Many PhD programs have the tradition of interviewing applicants.
-The most common case is that a professor is interested in working with a student and wants to chat with them to make decision (e.g., to offer them RA).
+\section{Interview and the Waiting Game}\label{sec:interview}
+
+After you submitted your applications, the waiting game begins! For many students, this is an excruciating time. This section provides some information and tips to help you get through this time.
+
+% but you should try to relax and focus on your study and work.
+You can also use this time to improve your profile, e.g., by working on research projects, publishing papers, or improving your connections for better LoR writers.
+
+\paragraph{Interviews} Many PhD programs have the tradition of interviewing applicants (just like job applications).
+The most common case is that a professor is interested in working with a student and want to chat with them to make decision (e.g., to offer them RA).
 The professor might include one of their current students as part of the interview.
-In some case, the interview is done by several professors, e.g., to see if a student would fit in their group or to recruit a very strong student to their program.
+
+In some case, the interview is done by several professors, e.g., to see if a student fits to their group or to recruit a very strong student to their program.
+The interview can also be done by Adcom members to see if the student fit their program.
 
@@
 A professor might want to know more about your research background or even ask you to read a paper (so be sure to do it!).
-Finally, during the interview, you will have the opportunity to ask questions about the university.
+During the interview, you will have the opportunity to ask questions about the university.
 It's essential to prepare some thoughtful questions, similar to a job interview.
@@
 \begin{commentbox}[Vu:]
-  At GMU, we are encouraged to interview candidates.
+  At GMU, faculty are encouraged to interview candidates.
 For very strong candidates, the interview is actually to recruit them.
@@
 Without the interview, such applications may be more likely to be rejected.
-\end{commentbox}
-
+  \tcblower
   In short, getting an interview is a good sign; it means that someone is considering you. If we are not interested in your application, we will not proceed with an interview.
-However, do not worry much if you do not get an interview! For example, at GMU, most of our admitted students with TA do not go through any interviews.
+
+\end{commentbox}
+
+\paragraph{Not getting interviews} While it is generally good to get interview, do not feel so despair if you do not get any.
+Many programs do not have the tradition of interviewing applicants.
+For example, at GMU, most of our admitted students with TA do not go through any interviews.
+
+\paragraph{Timeline} The timeline for interviews and notification letters varies.
+Faculty setup interviews based on their (very busy) schedule.
+So do not be surprised if you get an interview invitation in the last minute (also don't get surprised if your interview is on a Uber ride!).
+And again, some schools do not do interviews at all.
+
+Some schools send out admission letters in batches, some do rolling admission, and some do not send anything out (e.g., you're on their wait list).
+However, in general, you should hear back from most schools by around March.
+
+In short, you should not worry too much about the timeline. Also do not send emails asking about interviews or status.
+
+\paragraph{Updating your profile} In general, you should not send emails to update your profile.
+However, if you have new publications or other achievements, you can send an email asking to update your profile.
 
@@
 They may \emph{co-advise} or be in the PhD dissertation committee, but your main adviser will need to be a tenured or tenure-track faculty in CS.
-If in doubt, you should check with the CS department for their requirements.
+If in doubt, check with the CS department for their requirements.
 For this specific reason,  CSRankings includes only tenure line faculty who can advise CS PhD students.
@@
 
-\autoref{tab:cost} shows the budget breakdown for a GRA per year (this level of details is what faculty actually uses when applying for grants).
+\autoref{tab:cost} shows the budget breakdown for a GRA per year.
 These numbers are based on my experience at public universities in the US.
@@
 
-This section lists some cultural issues that international students might want to pay attention to. These issues might not specific to just the PhD admission process but in general.
+This section lists some cultural issues that international students might want to pay attention to. These issues apply in general and not just PhD admission.
 
@@
 
-\paragraph{Accepting an offer} If you decide to accept an offer, you can say something like: \emph{``Thank you for the offer.
-I am very excited about it and would like to accept it.
-I look forward to working with you.
-Can you please send me more details about the offer and what to do next?''} .
+\paragraph{Accepting an offer} If you decide to accept an offer, you can say: \emph{``Thank you for the offer.
+I would like to accept it and look forward to working with you.
+Could you please send me more details about the offer and what to do next?''}.
 Here you are not only accepting the offer but also asking for more details.
-This is perfectly fine and professors will understand and appreciate your enthusiasm.
-They will likely send you more details about the offer and what to do next.
-If you decide to accept an offer, you should do so quickly.
+The prof.
+will likely send you more details about the offer and what to do next.
+If you decide to accept an offer, do so quickly.
 
@@
 
-\paragraph{Postponing an offer} If you need more time to decide, you can ask for more time.
-You can say something like: \emph{``Thank you for the offer.
+\paragraph{Postponing an offer} If you need more time to decide, you can ask for more time: \emph{``Thank you for the offer.
 I am very excited about it.
@@
 But do not ask for too much time, e.g., more than a month.
-You also should not postpone the offers multiple times, which will likely annoy people.
-
-\paragraph{Declining an offer} If you decide to decline or reject an offer, you can say something like: \emph{``Thank you for the offer.
-I am very excited about it.
+You also should not postpone the offers multiple times, which will annoy people.
+
+\paragraph{Declining an offer} If you decide to decline or reject an offer, you can say: \emph{``Thank you for the offer.
 However, I have decided to accept another offer.
 I appreciate your time and consideration.
-I hope we can work together in the future.''}  Professors will understand and will likely wish you good luck.
-If you decide to reject an offer, you should do so quickly.
+I hope we can work together in the future.''}  Professors will understand and wish you good luck.
+If you decide to reject an offer, do so quickly.
 

### 2024-02-10 demystify.tex

@@
 Most PhD programs will require at least \textbf{two LORs}.
-Having a letter from an internationally recognized researcher can \textbf{greatly} strengthen your application.
-In fact, it is arguable at many top CS programs, strong LoRs from well-known researchers are even more important than published papers.
+Having a letter from an internationally recognized researcher will greatly strengthen your application.
+In fact, a strong letter from a well-known researcher can outweigh the lack of publications.
 

### 2024-02-21 demystify.tex
updates

@@
 Most PhD programs will require at least \textbf{two LORs}.
-Having a letter from an internationally recognized researcher will greatly strengthen your application.
-In fact, a strong letter from a well-known researcher can outweigh the lack of publications.
+Having a strong letter from an internationally recognized researcher will greatly strengthen your application.
+In fact, a strong letter from a well-known researcher can outweight the lack of publications.
 
@@
 \paragraph{Self-written Letters} Many international students write letters themselves, typically due to the requests of their profs., and have them signed by their profs.
-Such letters have little values and are considered weak by reviewers (why can you not even find someone who cares or knows enough about you to write a candid personal reference letter?).
+Such letters have little values and are considered weak by reviewers---why can you not even find someone who cares or knows enough about you to write a candid personal reference letter?.
 Instead of the reference writer talking about you, in this case it is you who write about yourself (and they just sign the letter).
@@
 
-\paragraph{Interviews} Many PhD programs have the tradition of interviewing applicants (just like job applications).
-The most common case is that a professor is interested in working with a student and want to chat with them to make decision (e.g., to offer them RA).
+\paragraph{Interviews} Just like a job application, you can get interviews from some schools.
+The most common case is that a professor is interested in working with you and want to chat to make decision (e.g., to offer RA).
 The professor might include one of their current students as part of the interview.
@@
 In some case, the interview is done by several professors, e.g., to see if a student fits to their group or to recruit a very strong student to their program.
-The interview can also be done by Adcom members to see if the student fit their program.
+The interview can also be done by adcom members to see if the student fit their program.
 
@@
   In short, getting an interview is a good sign; it means that someone is considering you. If we are not interested in your application, we will not proceed with an interview.
-
 \end{commentbox}
@@
 
+\end{itemize}
+
 You can then share this sheet with your reference writers and remind them to submit LoRs on your behalf (see \autoref{sec:lor}).
-\end{itemize}
+Also, you should periodically update your writers with your status; a simple note such as \emph{``I am getting interviewed by prof.
+X at school Y.
+Do you have any advice?''} can get you a lot of help.
 
@@
 Now you're at a different game because the schools that have admitted you will try to get you to accept them!
-Important factors to consider are the reputation of schools and professors (\autoref{sec:schoolsandprofs}), and funding availability (\autoref{sec:funding}).
+Important factors to consider include the reputation of schools and professors (\autoref{sec:schoolsandprofs}), and funding availability (\autoref{sec:funding}).
 You will have to make your decision by around \emph{April 15}.
@@
 % There are many advice on surviving PhD that you can follow. But here I just list a few that works for me and what I advice my students to do.\tvn{TODO}
+
+\section{Buying Computer Equipments}\label{sec:buying-equipment} Many students, including myself, get very excited about getting to buy new computer equipments and electronics when they start their PhD.
+However, you should check with your adviser first.
+Many profs.
+have funding to buy you a computer and other equipments (e.g., keyboards, headphones, tablets, etc).
+Also, do not worry that this is a "waste" of your adviser's money.
+This is part of the funding that your adviser gets to support you and your research.
+So use it to get what you need to do your research (and it's also OK to use them for personal purposes!).
 

### 2024-02-21 faculty.tex
updates

--- 2024-02-10 demystify.tex
@@
 \mainmatter
-\chapter{Should You Apply to a CS PhD Program in the US?}\label{sec:should}
 
-\epigraph{\vspace{-0.2in} Don't make fun of graduate students. They just made a terrible life choice.}{\textsc{The Simpsons}}
+\chapter{Research}
+\chapter{Obtaining Funding}
+\section{NSF and other Government Funding}
 
-First, I want to emphasize that PhD students in Computer Science \emph{do not} need to worry about funding, especially at good R1
-universities in the US. If you are admitted, you will almost certainly
-receive \emph{full funding} to support your study, including tuition,
-health insurance, and stipend (monthly salary). Moreover, depending on the university,
-you may even receive additional benefits such as summer pay, laptops, (conference/workshop) traveling. \autoref{sec:funding} provides more details on funding.
+\paragraph{Other Government Funding}
 
-Second, applying to a good US university \emph{should not} be any harder than at schools in other countries.
-In fact, it might even be more flexible since CS PhD in the US \emph{do not} require having an MS or a topic/proposal/adviser in advance.
-If you believe you have a chance in other countries, e.g., South Korea, Singapore, Germany, UK, Japan, and Australia, then you will surely have a chance in the US as well.
-\autoref{sec:non-us-differences} compares CS PhD study in the US to other countries.
+\section{Internal Funding}
 
-\begin{commentbox}[Vu:]
-  One of the reasons I create this document is that my colleagues at GMU are interested in recruiting Vietnamese students and are surprised when seeing very few applications in Vietnam (see \autoref{sec:ack}).
-Each year our CS program receives more than 350 PhD applications, most of which are international but only 3--4 are from Vietnam.
-In general the number of CS PhD applications from Vietnam to US universities is very low and more would be welcome.
-\end{commentbox}
+\section{Industrial Gifts}
 
-\begin{domesticbox}[Domestic students:] If you're a domestic student, you have several advantages in your application.
-First, we already know about your school (see \autoref{sec:your-school}), and you are also more familiar with the US education system and culture.
-You also have more opportunities for funding, e.g., through government scholarships for US citizens and residents.
-Finally, compared to international students, PhD applications in CS from domestic students are few and that can really help your case.
-\end{domesticbox}
+\chapter{Teaching}
+\section{Teaching Load}
+\section{Teaching Evaluation}
+\section{Teaching Philosophy}
+\section{Teaching Awards}
 
-%% Vu: *what's a PhD?*  This [[https://matt.might.net/articles/phd-school-in-pictures/][series of pictures]] from [[https://matt.might.net][Matt Might]] illustrates what a PhD means.
-%% \end{commentbox}
+\chapter{Mentoring}
+\section{Mentoring Undergraduate Students}
+\section{Mentoring Graduate Students}
+\section{Mentoring Postdocs}
+\section{Mentoring Junior Faculty}
 
-\section*{Additional Resources}
-\begin{itemize}
-  \item \href{https://parentheticallyspeaking.org/articles/us-cs-phd-faq/}{Getting a Computer Science PhD in the USA} by Shriram Krishnamurthi
-  \item  \href{https://pg.ucsd.edu/PhD-application-tips.htm}{Blog post on PhD Admission} by Philip Guo
-  \item   \href{https://talkingtorobots.com/yonatanbisk.html}{PhD Admission post} by Yonatan Bisk (customized for CMU admission)
-\end{itemize}
+\chapter{Services}
 
-\chapter{How is Your Application Evaluated?}\label{sec:evalapps}
+\section{Service to your Professional Community}
+\section{Service to the Department}
+\section{Service to the University}
+\section{Service to the Community}
 
-\epigraph{How is education supposed to make me feel smarter?
-Besides, every time I learn something new, it pushes some of the old stuff out of my brain.
-Remember when I took that home wine making course, and I forgot how to drive?}{\textsc{The Simpsons}}
-
-After you submit your PhD application (usually around December), it will first be screened
-for general requirements, e.g., did you submit your transcripts and standard scores?
-did your reference writers submit their letters?
-Usually, this screen process is done through a central university system (i.e., not by the CS department).
-If you pass this screen, your application will be forwarded to the CS department for further evaluation.
-If you don't, then the system will inform you what is missing and what you need to do.
-
-\begin{commentbox}[Hakan:]
-  At GMU, for full consideration, students should make sure to submit \textbf{ALL} required documents by the application deadline, and should never assume that some required documents (such as official TOEFL scores or official diplomas/transcripts) will be waived by the admissions office.
-If something is listed and not marked as ``optional'', it is mandatory and they should plan for submitting all those.
-\end{commentbox}
-
-Then, applications are reviewed by a \textbf{PhD admission committee} (\emph{adcom}) that consists of faculty members in CS (in some cases the committee can involve affiliated faculty from different disciplines).
-These faculty have a wide-range of expertise and background to ensure diverse perspectives in the evaluation process.
-The size and the review load of the adcom depends on the department size.
-At GMU, the PhD adcom typically has 15--20 faculty, and each committee member is assigned to review 25--30 applications.
-Note that GMU, UMass, and probably many other schools, have separate adcom for MS programs.
-
-Each application is assigned to about \emph{three} faculty members, who will evaluate your profile and reach a consensus.
-Note that while the assigned reviewers are likely the main ones deciding your application, \emph{every} faculty in the department will have access to your application and can provide inputs and opinions on your profile.
-
-The PhD adcom typically involves assistant professors in the department (see \autoref{sec:faculty-types} for various faculty titles).
-This provides junior faculty the opportunities to recruit students.
-The chair of the committee will be a senior professor, but they likely will not review individual applications and instead assign them to committee members.
-The chair will look at various factors such as research interests or mentioning faculty names to assign the applications to appropriate faculty.
-
-At GMU, we usually decide that a full-time PhD candidate is either (i) admit with funding (TA or RA, see \autoref{sec:funding}) or (ii) rejected. In other words, in most cases, we either
-admit you with full funding, or we reject your application. In some rare cases, we admit
-without funding because you have funding on your own, e.g.,
-supported by your government or having external fellowships. We justify
-our decision with a summary of your application, where we list
-strengths (e.g., a well-known school) and weaknesses (e.g., weak
-LORs).
-
-% \didi{Is there more information on typical strengths and common weaknesses of applications.
-This is especially useful to sophomore and junior students as they still have time to work on those strengths.}
-% tvn: the main thing is research experiences
-
-\paragraph{Why we do not waive application fee?}  This is typically a requirement of the university.
-Individual departments and programs do not have the flexibility to waive the application fee, even if they want to.
-
-In my opinion, requiring applicants to pay the fee helps ensure their seriousness, as it filters out non-serious candidates.
-Also, if the application process were free for everyone, we would receive an overwhelming number of applications to review.
-
-% \begin{commentbox}[Hung]
-%   The application fee varies between schools, and is about \$70--\$100.
-I found this number pretty ridiculous.
-In most cases, there is no proper justification for that amount.
-Yes, the application fee could act as a filter, but a nominal amount would do the job.
-While a blanket waiver might not be feasible, universities should extend the eligibility of fee waiver applications to more countries.
-% \end{commentbox}
-
-\chapter{Your Application}\label{sec:application}
-
-\epigraph{Son, if you really want something in this life, you have to work for it. Now quiet! They’re about to announce the lottery numbers.}{\textsc{The Simpsons}}
-
-The primary focus of the admissions committee is to \textbf{evaluate your background and interest in research} (a PhD in Computer Science is a research degree!), and determine if you would \textbf{fit into} the program.
-To evaluate your profile, we consider
-the following key indicators.
-
-\section{Research Ability}\label{sec:research-ability}
-
-The most concrete evidence of research ability is having \textbf{published papers in reputable international journals or conferences}.
-Having published papers, especially at top venues, is a sign that the applicant has successfully involved in research.
-
-\begin{commentbox}[Vu:]
-  Many international students mention Scopus Q1, which consists of various journals from IEEE, Elsevier, and many other publishers.
-I don't know/recognize many of journals listed in Scopus Q1.
-This might be something to be mindful of, as \textbf{CS} faculty might not be too familiar with Scopus or journals listed there, so devote some part in your statement to discuss the significance of your papers.
-\end{commentbox}
-
-However, it is understandable that many students do not have the opportunities to publish in top places.
-Thus, general writings, even those under submissions or rejected, would still be good (and much better than nothing).
-But be sure to upload your papers with your application and talk about them in your statement (see \autoref{sec:research-statement}).
-Note that local conferences and non-English journals or conferences do
-not carry as much weight since their quality is often unknown to US faculty.
-However, if you have published in such places, you should still upload them, mention them in your statement, and explain why they are good.
-
-% In CS, conferences are important and top conferences are prestigious.
-You can find top CS conferences at places such as \href{https://csrankings.org}{CSRankings}, which ranks CS programs based on how their faculty publish at top \emph{conferences} (see more in \autoref{sec:ranking}).
-
-\begin{commentbox}[Craig:]
-  GMU and many other universities allow you to upload your published papers and other writing samples.
-In many cases, even if the papers were not published at top places, we can still determine their quality by simply skimming over the paper.
-\end{commentbox}
-
-Additionally, \textbf{work experiences at renowned research laboratories}, such as Microsoft Research, can strengthen your
-application.
-Unfortunately, many good research places in your countries, e.g., VinAI in Vietnam, remain relatively unknown to most universities in the US.
-So you should explicit say something about them in your statement.
-
-\begin{commentbox}[Hung:]
-  The reputation of VinAI has been increasing steadily over the past few years; many of my colleagues heard about VinAI.
-\end{commentbox}
-
-Finally, \textbf{participating internationally recognized competitions} can also demonstrate your research potential.
-For example, participating in Math Olympiads if you want to do theory or  winning ACM programming contests if you want to ``build'' systems, e.g., software analysis.
-
-\begin{commentbox}[Thanh:]
-  Due to academic culture, professors in Vietnam usually aim for (international) journals instead of conferences.
-Could you give some tips on how to know whether a journal is good (CSrankings, unfortunately, only consider conferences)?.
-  \tcblower
-  Vu: One way is looking at what well-known researchers publish at.
-For example, if you are interested in a field X, you can use CSRankings to look at active faculty in X, and then look at their websites to see what journals they publish at.
-\end{commentbox}
-
-\section{Letter of Recommendation (LOR)}\label{sec:lor}
-
-\begin{center}
-  \includegraphics[width=0.6\linewidth]{files/c6.png}
-\end{center}
-
-Most PhD programs will require at least \textbf{two LORs}.
-Having a letter from an internationally recognized researcher will greatly strengthen your application.
-In fact, a strong letter from a well-known researcher can outweigh the lack of publications.
-
-However, obtaining such letters
-can be challenging for international students, who might not have much interactions with such experts.
-So it is fine to have a letter from people (e.g., profs, researchers, postdocs who mentored you) that know you well enough to talk about \emph{your research experience and capabilities}.
-Many students get letters from supervisors from companies where they did internships or are working at.
-It is OK as long as it is a research-based personalized letter (once again, we are talking about PhD applications, not MS).
-
-\begin{commentbox}[Vu:]
-  If your grading system is not US standard or you are from a good school unknown outside of your country, you can ask your reference writers to explain about that in their letters.
-For example, "Bach Khoa" are the top universities in Vietnam for STEM studies but few people outside Vietnam know about them.
-So if you are from there, you should ask your reference writers to mention that.
-\end{commentbox}
-
-\paragraph{Self-written Letters} Many international students write letters themselves, typically due to the requests of their profs., and have them signed by their profs.
-Such letters have little values and are considered weak by reviewers (why can you not even find someone who cares or knows enough about you to write a candid personal reference letter?).
-Instead of the reference writer talking about you, in this case it is you who write about yourself (and they just sign the letter).
-
-\paragraph{Generic Letters} When the writers do not know much about the applicants (e.g., just taking some course with them or not making any impression to write about), they might write a generic letter, which is not useful and also considered weak.
-It might be a good idea to directly ask if the prof.
-is willing to write a \emph{strong} letter for you.
-If not, then you should ask someone else.
-For example, if a student I don't know well ask me to write letter for them, I will explicitly tell them I don't know them that well to write much about them, and such a short, generic, and weak letter will not help their case.
-
-\begin{commentbox}[Hung:]
-  A sad reality is that most professors in Vietnam \textbf{DO NOT} know how to write a good letter, or are lazy in writing letters hence delegate the writing to the students.
-Unfortunately, there is no easy solution to this problem.
-\end{commentbox}
-
-\begin{commentbox}[Vu:]
-
-  Another sad reality is that some professors are unwilling to write letters or write weak ones because they do not want (good) students to go abroad or only go to places where they want the students to go to.
-If you are in this situation, you should find someone else to write for you.
-  \tcblower
-
-  Sometimes students would go through great length just to get letters from well-known professors in their school (e.g., department head).
-But as mentioned, if these professors do not know you, the letters are generic and carry little value, in fact, \red{red flags}.
-Moreover, a top professor in Vietnam might not be well-known to US faculty (see more details in \autoref{sec:your-school}).
-So save the trouble and just get letters from \emph{any} professors/supervisors who knows you well and can write a good letter about \emph{your} research ability.
-It's better to have a good personalized
-  letter about your own research ability from someone who is less
-  well-known than a generic/weak letter from a well-known person.
-
-\end{commentbox}
-
-\paragraph{Waiving Your Right}  Choosing not to look at a reference letter is pretty standard in school and job applications.
-When you waive your right to see the letter, it adds a layer of trust, showing you're confident in your choice of referees and that you're not trying to twist their words.
-It's also about keeping things open and honest between you and your letter writers, and encourages them to be real about your strengths and qualifications.
-Plus, it keeps things private.
-
-Reviewers might raise concerns (a red flag) about  a letter that is not waived, e.g., if you do not trust your letter writers, then you should find someone else to write for you.
-In short, it's a standard practice and a way of keeping things straightforward and respectful in the whole recommendation game.
-
-\begin{commentbox}[Didier]
-  \emph{Should letter writers have PhDs?}  In Rwanda, a lot of students interact more with teaching faculty who might not have PhD.
-  \tcblower
-  \textbf{Vu}: This is an interesting and useful detail that US faculty might not be aware of and students should mention about this in their statements.
-In general, I think it should be fine as long as that person can properly evaluate your research ability.
-\end{commentbox}
-
-\paragraph{Reminding Writers} After you submit your application, you can tell your writers that and remind them to submit their letters if they haven't done so.
-But don't send too many reminders as that can be annoying to the writers.
-
-Note that most places only have deadlines for the applicant, but are very flexible with the letter writers (in many cases do not even give them any deadline).
-Also, many places do not begin the admission review process right after the deadline and work on application reviews in the next semester (mid January).
-
-\paragraph{VEF 2.0} For Vietnamese students, it's worth mentioning about the \href{https://vef2.org/}{VEF2.0 program}, which has helped many good students in gaining admission to top CS PhD programs.
-VEF2.0 follows an interesting model where US faculty members from leading institutions are invited to conduct rigorous interviews with VEF students and subsequently provide reference letters on their behalf.
-Despite the limited interaction between the interviewer and the interviewee (primarily confined to the interview itself), these reference letters are generally effective as they have helped many students getting into top PhD programs in the US.
-
-\section{Statement of Purpose (SOP)}\label{sec:research-statement}
-
-\begin{center}
-  \includegraphics[scale=0.4]{files/c2.png}
-\end{center}
-
-While you might not have control over LORs or \hyperref[sec:your-school]{where your go to school}, you do over your
-statement of purpose (SOP) or personal statement\footnote{Some schools separate these documents and ask you to write both: SOP, which focuses on research experiences, and Personal Statement, which is everything more personal, e.g., why PhD, challenges, diversity, etc}!
-So write it well because it could make a big difference.
-In your statement, you have the opportunity to make your application stand out and unique, which can make you \emph{fit} the CS program you're applying to, even if you don't have very strong research experience.
-A well-written SOP also shows that you can communicate, which is very important in research, and that you can effectively teach and communicate with students, which is important for TA funding (see \autoref{sec:funding}).
-% is important because if you need (GTA) funding, it will provide evidence
-% that you can teach and communicate with students.
-
-There are various guides on writing statement, e.g.,~\cite{blattman2022writing}, and \href{https://cs-sop.org/}{many example statements} are available.
-So I will not talk too much about statements.
-In short, focus on your research goal and vision, and convince us that you can achieve it through your experience, e.g., published papers, or if you work on some projects by yourself, talk about it.
-Also, talk about things that adcom members might not know about, e.g., your Github project with 1K+ stars or your regular contributions to well-known open-source projects.
-Finally, talk about what makes you different than the other 1000 applicants that we are considering (e.g., see \autoref{sec:improve-your-chance} for increasing your admission chance).
-
-\begin{commentbox}[Hung:]
-  I think asking for a “research vision” from a Ph.D.
-applicant is too much.
-Even people graduated with a Ph.D.
-have a great difficulty in explicating their research vision.
-In my opinion, you should focus on showing (not telling) your research passion (if you do have).
-\end{commentbox}
-
-Finally, this is something easy to do, but is neglected by many
-applicants: \textbf{customize the statement} for the school you're applying to,
-e.g., why do you apply \emph{here}?
-provide names of professors who you're interested in (if they are not already in the adcom, your application might get forwarded to them for evaluation; and they will likely be those who interview you).
-This shows that you're serious and have done homework on places you're applying to.
-Adcom will look for this part at the end of the statement, so do not skip it.
-
-\begin{commentbox}[Vu:]
-  I often read the research statement (and the LORs) first. If I am
-  persuaded by then, I would skim over other factors and advocate for
-  admission (unless I see red flags in other parts). If I am not
-  convinced, then I will likely recommend rejection (unless I see
-  something standout in other parts).
-  \tcblower
-  Do careful research on professors, don't mention \emph{emeritus} or  adjunct faculty (see more about various types of faculty in \autoref{sec:faculty-types}).
-  Also, be careful not to send statements to wrong schools or mixing
-  facts (e.g., talking about school X but mentioned about working with
-  profs. at school Y; and definitely do not talk about George Washington when applying to George Mason). I have seen such statements more time that I
-  should.
-\end{commentbox}
-
-\subsection*{Additional Resources}
-\begin{itemize}
-  \item \href{http://www.pl-enthusiast.net/2022/10/03/how-to-write-a-grad-school-personal-statement/}{How to Write a Grad School Personal Statement} by Mike Hicks
-  \item     \href{https://cs-sop.notion.site/cs-sop/CS-PhD-Statements-of-Purpose-df39955313834889b7ac5411c37b958d?p=f5d5980a71524ebaa4e6ae57266b847c&pm=s}{CS PhD SOP database} by cs-sop.org
-\end{itemize}
-% \subsection{Outline of a LOR}\label{sec:lor-outline}
-
-\section{Your School}\label{sec:your-school}
-
-Graduating from top universities \emph{that we recognize} helps. For example, if your school is well-known, then it is \emph{``top foreign''}, which is definitely a plus.
-However, if we do not know much about schools in your country, then we are uncertain about the quality of your school and  likely treat your school as \emph{``unknown foreign''}, which can be a minus point.
-
-So what can you do about this?
-several things including asking your CS dept to put itself on CSRankings (it's the easiest way to get CS people to know about the school and its CS faculty)  and explaining about your school in your statement (ask your LOR writer to do that too).
-Of course, if you're interested in working with Vietnamese, consider  \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{CS PhD programs in the US that have Vietnamese professors}.
-
-\begin{commentbox}[Vu:]
-  Sometime PhD adcom in the US will share a document such as \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Foreign-Top-Schools}{this one}, which lists the top schools in several countries.
-In some cases we ask other faculty and students if we think they know about the place.
-For example, when I was a postdoc at UMD, members in their CS PhD adcom ask me to evaluate applicants from Vietnam.
-During my time at UNL and now here at GMU, I have looked at Vietnamese applications (whether they are assigned to me or not) and provide input to their reviewers, e.g., X is the top tech school in Vietnam and so it should be \emph{top} instead of \emph{unknown} foreign, which makes a huge difference.
-\end{commentbox}
-\begin{commentbox}[Deepak:] 
-  If an applicant is anxious about their school not being known outside their country, they can provide information about their school and department, with independent sources where such information can be verified.
-\end{commentbox}
-
-\section{Grades and GREs}\label{sec:grades}
-
-Having good grades is good, but unless your school is well-known, having top grades or rankings
-usually will not help because we cannot evaluate them.
-
-This can be an issue for students in many top international universities where the competition is so high that very good students can still have low rankings from these schools (and be overlooked by the adcom).
-So what to do with this? as mentioned in \autoref{sec:your-school} you can put a note about this in your statement and ask your LoR writers to talk about it.
-
-Note that while having good grades at unknown school might not help,
-having bad grades will be \red{red flag} (unless your LORs or
-statement give a proper explanation). This is especially true if you
-have bad grades in relevant courses (e.g., Math and CS).
-
-\begin{commentbox}[Thanh:]
-  Vietnamese universities typically offer specialized programs, such as the talented engineer program at HUST, that have highly competitive entrance exams and a limited number of available slots (e.g., 30 per year).
-However, these programs often set higher requirements for students, including more demanding tests and assignments, resulting in lower GPAs and overall rankings.
-For example, an 3.5 GPA students from such talented programs are typically much better than a 4.0 GPA students not in those programs.
-Similarly, variations in GPA standards exist among different universities, with technical universities generally having lower GPAs than economical universities.
-These make gaining admission in the US difficult as US faculty are not familiar with these issues.
-  \tcblower
-  \textbf{Vu}: Vietnamese students and even faculty often lament how this grading system hurts Vietnamese students applying abroad.
-One way to mitigate this is making these issues known in your SOP.
-Universities with Vietnamese profs are probably aware of them, but in general your letter writers \emph{and} you can explicitly mention these in their letters and your statement.
-\end{commentbox}
-
-\paragraph{GRE} Most good CS PhD programs in the US \textbf{no longer require GREs}, so you don't need to take them.
-However, they might be useful for international students from programs we are not familiar with.
-If you have good GRE scores then you should include them in your application.
-
-\paragraph{English Test} Unless your degrees are from specific countries such as \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-GMU#standard-tests-waiver-eligible-countries}{these}, you will need to
-take standardized English test. Just do well enough to pass minimum requirement set by the university, which has many options for you to select from.
-
-\begin{commentbox}[Vu:]
-  The minimum for GMU (being above this might not mean much, but below is a \red{red flag}).
-  \begin{itemize}
-    \item GPA: $\ge 3.0$ in your undergrad (but we also consider the rank/prestige of your school)
-    \item GRE: not required, though it can help boost your profile
-    % \item but if you want to use it, then we expect a total (V+Q) of $\ge 311$ (with a $\ge 157$ Q) and A $\ge 3.0-3.5$.
-    \item English requirement tests (one of the below)
-          \begin{itemize}
-            \item TOEF: 88 pts in total AND $\ge 20$ points in each subsection OR
-            \item IELTS: $\ge 6.5$ OR
-            \item DuoLingo Graduate English: $\ge 120$ OR
-            \item Pearson Test of Academic English: $\ge 67$
-          \end{itemize}
-  \end{itemize}
-\end{commentbox}
-
-\section{CV/Resume}
-The CV/Resume should provide a summary of the applicant's achievements.
-It should allow the reviewers to quickly scan to identify major achievements (e.g., Publications, Programming Competition Awards, Teaching Experience).
-
-\section{Interview and the Waiting Game}\label{sec:interview}
-
-After you submitted your applications, the waiting game begins! For many students, this is an excruciating time. This section provides some information and tips to help you get through this time.
-
-% but you should try to relax and focus on your study and work.
-You can also use this time to improve your profile, e.g., by working on research projects, publishing papers, or improving your connections for better LoR writers.
-
-\paragraph{Interviews} Many PhD programs have the tradition of interviewing applicants (just like job applications).
-The most common case is that a professor is interested in working with a student and want to chat with them to make decision (e.g., to offer them RA).
-The professor might include one of their current students as part of the interview.
-
-In some case, the interview is done by several professors, e.g., to see if a student fits to their group or to recruit a very strong student to their program.
-The interview can also be done by Adcom members to see if the student fit their program.
-
-Typically, an interview takes about 15--30 minutes, and one important aspect of evaluation is your ability to effectively communicate, including speaking and understanding English.
-A professor might want to know more about your research background or even ask you to read a paper (so be sure to do it!).
-During the interview, you will have the opportunity to ask questions about the university.
-It's essential to prepare some thoughtful questions, similar to a job interview.
-
-\begin{commentbox}[Vu:]
-  At GMU, faculty are encouraged to interview candidates.
-For very strong candidates, the interview is actually to recruit them.
-In some cases a faculty interviews a candidate that they see potentials and want to advocate for their admission.
-Without the interview, such applications may be more likely to be rejected.
-  \tcblower
-  In short, getting an interview is a good sign; it means that someone is considering you. If we are not interested in your application, we will not proceed with an interview.
-
-\end{commentbox}
-
-\paragraph{Not getting interviews} While it is generally good to get interview, do not feel so despair if you do not get any.
-Many programs do not have the tradition of interviewing applicants.
-For example, at GMU, most of our admitted students with TA do not go through any interviews.
-
-\paragraph{Timeline} The timeline for interviews and notification letters varies.
-Faculty setup interviews based on their (very busy) schedule.
-So do not be surprised if you get an interview invitation in the last minute (also don't get surprised if your interview is on a Uber ride!).
-And again, some schools do not do interviews at all.
-
-Some schools send out admission letters in batches, some do rolling admission, and some do not send anything out (e.g., you're on their wait list).
-However, in general, you should hear back from most schools by around March.
-
-In short, you should not worry too much about the timeline. Also do not send emails asking about interviews or status.
-
-\paragraph{Updating your profile} In general, you should not send emails to update your profile.
-However, if you have new publications or other achievements, you can send an email asking to update your profile.
-
-\section{Preparing and Tracking Applications}
-
-\paragraph{Shared Spreadsheet} Many students use a (shared) spreadsheet, e.g., Google Sheets/Docs, to help them and their letter writers to keep track of their applications.
-Here are some information to put on the sheet.
-\begin{itemize}
-  \item \textbf{Your info}: Full name, email, phone, link to website/CV.  This is helpful for the writers just in case they want to quickly get some information about you.
-
-  \item \textbf{Applications details}: University, Dept, Application System URL, Submission Deadlines, Application Status (e.g., submitted,  rejected, wait-listed, accepted)
-
-  \item \textbf{LoRs}: Writer 1, Writer 2, Writer 3.
-If this is a shared document, you might want to omit names and just use writer X.
-Under each is their status, e.g., sent/not yet/reminder needed, etc.
-
-        You can then share this sheet with your reference writers and remind them to submit LoRs on your behalf (see \autoref{sec:lor}).
-\end{itemize}
-
-% \paragraph{Communicating with LoR writers and People Who Support You}
-
-% stressfull time, rely on support from others
-% ask ppl to review your SOPs
-% provide information for LORs, ..
-%
-
-\chapter{Getting Admitted}\label{sec:accepted}
-
-\epigraph{"Oh...
-and how is education supposed to make me feel smarter?
-Besides, every time I learn something new, it pushes some old stuff out of my brain.
-Remember when I took that home wine-making course and I forgot how to drive?"}{\textsc{The Simpsons}}
-
-By around March you should hear back from most PhD programs you applied to. If you haven't heard back, reach out via email and ask about the status of your application.
-If you receive offers, congratulations!
-Now you're at a different game because the schools that have admitted you will try to get you to accept them!
-Important factors to consider are the reputation of schools and professors (\autoref{sec:schoolsandprofs}), and funding availability (\autoref{sec:funding}).
-You will have to make your decision by around \emph{April 15}.
-
-\paragraph{Open House} Most schools will have \textbf{Open House} events, which are a great resource to learn about the school, department, faculty, research, living, etc.
-Even if you can't come in person, you should attend virtually and meet with individual faculty.
-During the Open House, you get a chance to talk to individual faculty and current students.
-Take notes of faculty who make you excited, count those that are taking in new students (if they meet you, likely they are considering new students!).
-Talk to students about their advisers, the dept, the area, funding situation etc.
-Ask about anything you want to determine that they deserve \emph{you}.
-
-\begin{commentbox}[Vu:]
-  GMU has \emph{Virtual} Open House (VOH), e.g., \url{https://cs-gmu.github.io/cs-phd-voh-s23/}.
-We invite all admitted PhD students to the VOH through Zoom to learn about the CS program, the department, GMU, and the DC area in general.
-Students also get opportunities to chat with professors and current students.
-\end{commentbox}
-
-\paragraph{If you do not get admitted} If you do not get admitted to any schools or don't want to go to the ones that admit you, try again next time.
-Graduate admission can involve randomness and noise.
-Don't bother asking for feedback, you will not likely get any.
-In the meantime, you can work on improving your profile, e.g., get more research experiences, publish more papers, improve your connections for better LoR writers, etc.
-
-You can also consider applying to MS programs, which are typically easier to get in (but you need to pay).
-If you get in an MS program at a school of your choice, you can contact professors to work with them.
-If you do well, you can ask the professor to support you to convert to PhD.
-This is a common path for many students.
-
-\chapter{Funding}\label{sec:funding}
-
-\epigraph{Bart, with \$10,000, we’d be millionaires! We could buy all kinds of useful things like…love!’}{\textsc{The Simpsons}}
-
-As mentioned, if you're admitted to a \emph{good} CS PhD program, you should not have to worry about funding!  
-In the US, the common types of funding for PhD are \emph{graduate teaching assistant} (GTA or TA), \emph{graduate research assistant} (GRA or RA), and \emph{Fellowship}.
-RA is paid by a prof.
-for you to do their research.
-TA is paid by the department for you to help with teaching.
-Finally, fellowship is an independent funding that can come from a school, a company, or an organization.
-\autoref{tab:funding} summarizes the differences.
-Note that funding is typically more available for PhD students than 
-Masters.
-
-\begin{table}
-  \centering
-  \footnotesize
-  \caption{Different types of PhD funding. All Covered: includes tuition, insurance, and stipend.}\label{tab:funding}
-  \begin{tabular}{c|c|c|c}
-    \toprule
-    &\textbf{TA}&\textbf{RA}&\textbf{Fellowship}\\
-    \midrule
-    \textbf{From} & School & Profs. & School/External\\
-    \textbf{For}                  & Teaching Assist.       & Research                        & Research                              \\
-    \textbf{Cover All?} & Yes                      & Yes                             & Yes                                   \\
-    \textbf{Summer?}              & No                       & Maybe                           & Yes                                   \\
-    \midrule
-    \textbf{Pros}                 & Research Freedom         & Get to do research              & Research Freedom                      \\
-    \textbf{Cons}                 & Teaching Duties           & Research Restriction & Competitive, Limited             \\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
-\section{Graduate Assistantship (TA/RA)}
-The most common type of funding is \textbf{graduate assistantship}, which is either TA or RA.
-Both TA and RA come with tuition waiving (you don't have to pay tuition), health insurance (this takes care of your insurance, which is a must have in the US), and most importantly, your stipend (i.e., your salary).
-Some universities also give significant discount or pay insurance for spouse/children.
-
-Several things about stipends.
-First, the amount of stipend \emph{varies} and depends on various factors such as location (e.g., a stipend in Washington DC is likely higher than in Lincoln, Nebraska due to higher living cost).
-Second, an academic year (AY)  year is typically \emph{9-month} in the US, so stipend is for 9 months (so divide by 9 for each month).
-Third, you might get paid over the summer through funding from your professor or fellowship (typicaly no TA over the summer).
-Fourth, like most source of income in the US, you will have to \emph{pay tax} on your stipend.
-Finally, private universities might pay more for stipend, but they could have higher "activity" or some other hidden fee, or you may be required to pay fees for each credit hour.
-
-Students often complain their stipend being too low, but it is actually not bad and you can live comfortably with it.
-In many cases, it is also enough to support your spouse and kids (many CS PhD students have their family with them).
-So don't worry too much about stipend.
-If you're admitted to a good CS PhD program, you will be fine.
-A good school would know that it has to be competitive to attract students.
-For example, at GMU, every year we discuss about improving the benefits, and especially stipend, for our graduate students.
-
-For a full breakdown of how much a graduate student costs, see \autoref{sec:ra-cost}.
-
-\begin{commentbox}[Vu:]
-  TA and RA at GMU have similar benefits in tuition waiving and insurance.
-The college and department will set the rate for 9-month graduate assistant stipend.
-TA, which is paid by the department, will likely be that amount but RA might be higher depending on the stage of the student (1st year vs ABD\footnote{All but dissertation: really close to graduate.}) and the prof.
-  \tcblower
-  Having health insurance is required at many US universities.
-Do not assume that you're young and healthy and ignore insurance.
-At GMU, and at most good CS PhD programs, your GTA or GRA \emph{will always} come with full insurance.
-In fact, at GMU your spouse/children will get significant discount rate for health insurance.
-So you will never have to worry much about health issues for you or your family here.
-\end{commentbox}
-
-\subsection{Teaching Assistant (TA)}\label{sec:ta}
-
-TA is common in the beginning when you haven't found an adviser who would pay you RA.
-As a TA, you spend up to 20 hrs/week and help with classes (e.g., grading or teaching labs/recitation).
-Your TA is paid through the school or department, i.e., they hire you to help teach.
-During a semester, a TA might work with several courses and professors (not necessary their adviser).
-TA funding is not typically available during the summer, which has few courses.
-
-\paragraph{How to get TA?}  Unless you have other funding such as RA or Fellowships, TA is typically default for good CS PhD programs.
-When you apply to be a full-time student,  state that you need financial assistant.
-It is common that the PhD committee will either admit you and give you GTA, or reject you; i.e., we do not admit a student without supporting them.
-
-\begin{commentbox}[Vu:]
-  At GMU CS, students admitted with TA have  4 years of GTA guaranteed, and in some cases also receive  stipend for the first summer.
-\end{commentbox}
-
-Even if you have other funding and do not need TA, you still should do TA at least once.
-This allows you to see what teaching is like, which is especially helpful for research career where you often give talks and tell people about your work.
-GMU sometimes has classes that a more senior student can teach.
-In that case, you will be paid as GTA or even sometimes as a lecturer.
-This is a good opportunity for students to get teaching experience and also get paid more (as a lecturer).
-
-\subsection{Research Assistant (RA)}
-RA is provided through a professor through their own funding so you can work on their project.  
-You do not need to teach as an RA, so you can focus on your research. Depending on the professor, RA may be available during the summer. \autoref{sec:ra-cost} gives more details on RA budget.
-
-\textbf{How to get RA?} When a professor recruits you, they will likely give you RA right away (e.g., when you apply).
-A common scenario is that you first get admitted with TA, and then after a year or two find an adviser to support you with RA.
-See how to contact a prof.
-in \autoref{sec:contact} for research opportunities.
-
-\begin{commentbox}[Vu:]
-  If you got recruited by a prof., who would give you RA right away, it's very likely you will get admitted.
-For example, if a prof., even if not in PhD adcom, wants to work with and funds you, the PhD adcom will respect that decision and admit you (unless your application has many red flags).
-\end{commentbox}
-
-\section{Fellowship/Scholarship}
-
-Fellowship is another type of funding that students can get (e.g., from the university, industries, government).
-Fellowships are typically competitive and generous, and gives pretty much all benefits tuition/insurance that a TA/RA has.
-Moreover, they often give higher stipend (including summer) and open doors for job opportunities (e.g., internship).
-
-\textbf{How to get Fellowship?} You need to apply for them.  Many schools provide such fellowships to attract admitted students.
-External fellowships including from the US government and tech companies (e.g., Google, Microsoft, Facebook).
-Fellowships from US government would likely require US citizenship or residency.
-% However, tech companies including Google, Microsoft, Facebook have fellowships that international students can apply for.
-
-Prestigious external fellowships typically require a clear and good research plan, so it is a good idea to wait until at least your second year to have research experience and even publication before applying.
-Remember, you're competing with the top PhD students at top universities worldwide.
-
-\begin{commentbox}[Vu:]
-  PhD applicants at GMU are automatically eligible for a \emph{Presidential Fellowship}.
-It is at least as good as GTA but the most important thing is that as a fellowship it is truly free money (i.e., you are not depending on any prof.
-or TA).
-PhD adcom members nominate applicants for this fellowship and the whole committee will decide.
-\end{commentbox}
-
-In general, fellowship, especially competitive ones from the government or company, is prestigious, and you will stand out if you get one.
-Every PhD student has pubs, but only superstars have NSF grad or Microsoft fellowship.
-In fact, these are so prestigious that even if you didn't get it but make it to the final round or even ``honorable mentioning'', school will still mention you on their website and you still should put it on your CV.
-
-\chapter{Choosing Schools and Professors}\label{sec:schoolsandprofs}
-
-Choosing a school and an adviser is clearly among the most important things in your mind when you apply and especially when you get admitted.
-This is further complicated due to cultural differences and unfamiliarity of  international students to the US higher education system.
-This section aims to mitigate some confusions and help you make informed decision.
-
-\section{Choosing a University}
-
-We will first discuss about universities in the US that offer PhD in CS. Then we will talk about how to rank and select them.
-
-\subsection{Schools that offer PhD in CS}
-
-Most universities the US have CS programs. 
-However, while many universities offer PhD in their CS programs, many do not.
-These universities might offer just Bachelor degrees (e.g., BS) and no graduate studies (i.e., no MS or Ph.D degrees), or they just offer MS programs (but no PhD).
-For example, Penn State in University Park has Ph.D.
-in CS,  but Penn State in Harrisburg only has BS and MS in CS, and Penn State in York only has BS in CS.
-On the other hand, multiple locations of the University of Texas, e.g., Austin, Dallas, Arlington, have PhDs in CS.
-
-Thus, if your goal is Ph.D.
-in CS, you have to aim only for schools offering such a degree.
-%This might sound obvious but many students (even those in the US) get confused due to the large number of universities.
-While this can be confusing due to the large number of universities in the US, a little research, e.g., searching for PhD in CS from the school website, will help you find out.
-All schools listed in \autoref{sec:ranking} have PhD studies in CS, so you can start there.
-
-% \subsection{R1, R2, ...}
-
-\subsection{Selecting and Ranking Schools for CS}\label{sec:selecting-ranking-schools}
-\begin{center}
-  \includegraphics[scale=0.5]{files/c1.png}
-\end{center}
-
-International students not familiar with US universities often put them into \emph{two} bins:  (i) very top schools that they dream about, and (ii) everything else.
-In many cases, they use resources such as rankings from US News, which are not very transparent and \emph{highly} questionable\footnote{\url{https://cra.org/cra-statement-us-news-world-report-rankings-computer-science-universities/}}.
-Sometimes these students evaluate CS programs using the reputations non-CS programs such as medical, math, or physics.
-They even rank universities based on popular states they know in the US, e.g., California and New York.  Clearly, there are so many thing wrong with these methods.
-
-You can learn about CS programs and research expertise of faculty using resources such as \href{https://csrankings.org}{CSRankings.org}, which is designed specifically to help prospective PhD students in Computer Science!
-You will be very surprised to learn that a school that you didn't know much about can have very strong research in your interested topic (and vice versa, a school you thought highly about might have no faculty working in the research field you're interested in).
-This is also a good way to learn about individual faculty (who works on what) and well-known CS conferences\footnote{In CS (and probably only in CS), conferences, not journals, are often the main venue to publish research finding.}.
-\autoref{sec:ranking} gives the top 50 CS programs in the US according to CSRankings.
-
-\begin{commentbox}[Dat:] Most Vietnamese students, including those from top schools, \textbf{do not know} about CSRankings.
-May be applicants who worked at top research places such as VinAI would know about it.
-\end{commentbox}
-
-However, in general, rankings can be superficial and you need to do more research to be informed and make better decision.
-For example, if you get admissions to several places, you should consider attending Open Houses and contact profs.
-that you're interested in at those place and talk to them.
-They would be more willing to chat with you now that you have been admitted.
-Ask them questions about their work, how they manage students, their expectations.
-You can even ask to contact their students.
-See more in \autoref{sec:accepted} on what to do after getting admitted.
-
-\begin{commentbox}[Hung:]
-  I always encourage the students I admitted to talk with my students and the students of other faculty in other schools who admitted them.
-You will unlikely hear straight-out complaints from current students in a professor’s group.
-But sometimes what is important are things that they (current students) don’t tell you.
-Pay attention to their "level of excitement" being in the group.
-\end{commentbox}
-
-\begin{commentbox}[Xiaokuan:]
-  Chinese students often only look at USNews rankings when selecting their Ph.D. universities (I did that, too, when I was applying for Ph.D. positions).
-  Now that I am a professor, I find it to be the least promising way.
-  %
-  The reason is that USNews does not provide a good metric for evaluating the quality of the Ph.D. program.
-  %
-  If you want to do great research, CSRankings is the best way to find good and active professors (which did not exist when I was applying),
-  since it solely focuses on publications at top-tier CS conferences.
-  %
-  Also,
-  I think Ph.D. is not only about research;
-  you need to also consider your daily life there, since you will (probably) stay for at least five years.
-  %
-  You might regret it if you did not consider this seriously before applying.
-  %
-\end{commentbox}
-
-\section{Choosing an Adviser}
-Obviously, there is no one-size-fit-all answer to this question.
-The best adviser is the one that you can work well with, i.e., \emph{fits} you.
-This is difficult and takes time.
-Fortunately, unlike many countries that require finding an adviser and research topic before starting the PhD (\autoref{sec:non-us-differences}), CS PhD programs in the US will typically give you a couple of years to search for advisers and research topics.
-This is especially true if you're admitted with TA (\autoref{sec:ta}), which gives you time to explore and find adviser.
-
-\subsection{Finding the right adviser}
-
-Here are some general advice to find professor.
-First, search for profs.
-that share similar research interests.
-You can learn about faculty and their level of research activity through the faculty's website or even from CSRankings.
-For example, in CSRankings, if you want to work with PL, you can search for those publish in PL conferences.
-If you want to work with SE \emph{and} AI, you can search for faculty who work in both SE and AI.
-You can also go to their website, look at their papers and projects.
-In any case, you should contact the professor and talk to them.
-For example, you can (cold) email them and ask about their research and if they are taking new students (\autoref{sec:contact}), and schedule appointments to chat with them (office hrs are often the best time).
-
-\begin{commentbox}[Xiaokuan:]
-  Whether the student's research interest matches that of the adviser is very important;
-  if there is a mismatch,
-  either the student or the adviser has to make compromises,
-  which often leads to disagreements or conflicts.
-  IMO, the adviser should be the one who \emph{guides}  students to do research while allowing students to pursue their own interest,
-  instead of \emph{dictating} their research.
-\end{commentbox}
-
-Another effective way is taking graduate level courses in the topics you are interested in (remember: you have those 2 years to explore).
-Professors teaching these \emph{special topics courses} and \emph{research seminars} might be recruiting students---giving you even a higher chance.
-Do well in the class, answer questions, talk to the prof after classes, etc---being stand out.
-Many professors, including myself, prefer taking in new students this way.
-It gives both the professor and student more time, e.g., a whole semester, to work and evaluate the relationship before making any commitment (sounds a bit like a marriage!).
-You can also ask if you can do an independent study or research with a prof.
-This can be informal (no credit) and takes place during the summer or winter break.
-For example, I do this with multiple students, many of whom are undergrads.
-Many will drop because they find they don't like my research, but some will stay and become my PhD students.
-
-Ultimately, choose a prof.
-that fits you the most by communicating with them, taking their courses, meeting and asking them questions, and even talking to their current students.
-It will take time and effort, but remember, you will be working with this person for 5+ years, so it is important to try to find the right one.
-
-\begin{commentbox}[Thanh:]
-  In my opinion, having a well-suited adviser is crucial for a successful PhD and research career.
-One effective approach to finding a suitable professor is by working with a professor during your undergraduate studies.
-An exemplary instance is VinAI's residency program, where residents collaborate with professors from the US for two years before applying to PhD programs.
-Many VinAI residents have achieved remarkable results and gained admission to prestigious US universities.
-Unfortunately, VinAI's resident program is limited to AI research.
-
-  In other fields, e.g.
-Software Engineering, Vietnamese students face challenges in reaching US professors.
-Do you have any tips for Vietnamese students who want to connect with US professors and work as research assistants?
-  \tcblower
-  \textbf{Vu}: \autoref{sec:contact} shows how to contact a professor for research opportunities.
-Many probably will say no (or do not reply) as they do not have the bandwidth to take on random students, but some may say yes if they see potential fit.
-\end{commentbox}
-
-\subsubsection*{Additional Resources}
-\begin{itemize}
-  \item \href{https://www.cs.columbia.edu/wp-content/uploads/2019/03/Get-Advisor.pdf}{The Definitive "what do I ask/look for" in a PhD Advisor Guide}
-\end{itemize}
-\subsection{Types of Faculty: Who can serve as a PhD adviser?}\label{sec:faculty-types}
-
-Not every faculty can serve as your formal adviser.
-Let's try to understand the different types of faculty and their roles.
-For example, you might hear about tenured,  tenure track, and teaching faculty.
-You might also hear about assistant, associate, full, adjunct, emeritus, teaching, research professors.
-Here is a primer on these terminologies.
-
-\paragraph{Tenure-line and teaching faculty} Tenure line faculty, consisting of tenured and tenured-track profs., focus on research, which includes publishing papers, obtaining grants, mentoring Ph.D students.
-They often have very low teaching load (e.g., 1 per semester).
-\emph{Tenured} faculty are professors who have been promoted to have a permanent position (informally, very hard to fire them).
-\emph{Tenure-track} faculty are (more likely young) professors who are on the track to get tenure.
-\textbf{Both tenure and tenured track faculty can serve as a formal adviser of Ph.D.
-students}.
-This is very important because you need to have a tenure line faculty as your main PhD adviser.
-\autoref{sec:tenure-vs-tenure-track} talks more about choosing between tenure and tenured-track professor as your adviser.
-
-In contrast, \emph{teaching} (or instructional faculty or professor of practice) mainly focus on teaching.
-They typically teach 3--4 classes per semester (which is quite a lot) and \emph{do not} have research responsibilities (e.g., they do not have to worry much about publishing papers or obtaining grants).
-They can mentor Ph.D.
-students but they typically cannot serve as a formal adviser of Ph.D.
-students (i.e., they \emph{cannot graduate} PhD students).
-
-It is worth noting that professors in a non-CS department also unlikely can serve as formal PhD advisers for CS students (see \autoref{sec:related-fields}).
-
-\paragraph{Titles} Faculty have rankings, e.g., assistant, associate, full, regardless if they are tenure line or teaching.
-Assistant means new faculty, associate means they have been promoted, and full means they are senior.
-A tenure track faculty starts with being an assistant professor and then gets tenure and promoted to associate (typically after 6 years) and then full (time varies, some become full within 3-4 years, some 10+ years, some remain associate).
-
-Adjunct faculty is not full-time, e.g., they might be working in industry and teach a class or two for fun.
-Emeritus means they are retired but still have some affiliation with the university.
-Research faculty (or research scientist) are typically non-tenure line faculty who focus on research.
-Due to their roles, adjunct, emeritus, and research faculty typically do not advise Ph.D.
-students.
-
-\subsection{Tenured or tenure-track faculty? Who do you choose?}\label{sec:tenure-vs-tenure-track}
-
-\begin{center}
-  \includegraphics[scale=0.4]{files/c8.png}
-\end{center}
-
-The short answer is that tenure-track faculty such as assistant professors are more likely to be young and active in research (they have to, in order to get tenure).
-Thus, they will likely have more time to work with you and push you to do research and publish.
-However, they may not have as much experience in managing students and may not have as much funding (yet).
-
-Tenured faculty, e.g., associate and full profs., are likely older, more well-known, and have more experience in managing students.
-However, they might not push you as hard (they don't have to, they already got tenure).
-They might also expect you to figure things out yourself, i.e., so you need to be very independent.
-Some tenured faculty are also no longer active in research and more involved with administrative responsibilities or with their startup companies (this means they will likely not take new students).
-
-% \subsection{Strong Faculty vs High-ranked School}
-% \begin{commentbox}[Thanh:]
-%   When considering PhD programs, we often wonder if we should prioritize a high-ranking university or a professor with a strong reputation?
-Of course, both are great, but it is hard to achieve both.
-I believe receiving some guidance from you  would be incredibly valuable.
-% \end{commentbox}
-
-% This is a good question and there doesn't seem to be any best answer or specific algorithm that you can use.
-But I'll throw out something for you to think about.
-To me, there are many good schools in the US and so you have some leeway in choosing.
-For example, I don't believe there's much difference among schools in top 5 (e.g., CMU, UIUC, UCSD, MIT, GTech are in the same equivalent class), top 10 (e.g., Michigan,  Stanford, UWash, UC-Berkeley, Cornell is another equivalent class), top 20, top 30, top 40, and so on (though I would say schools above 100 are pretty much in the same equivalent class).
-It's hard to say a top X is much much better than a top X+20 (e.g.,  Purdue at 15 is stronger than Yale at 35, but not \emph{that much} stronger as you would think).
-So keep that in mind, while ranking does matter (e.g., more active faculty, stronger students), the lower ranked universities are still pretty good, might have sub fields that are extremely strong (NCState is 40th, but in Software Engineering it is easily top 5), and many unique things.
-This is in contrast with universities in Vietnam where there are large gaps among the top 5, 10, and so on.
-
-\subsection{Should you contact a professor? How to get a desired reply?}\label{sec:contact}
-
-\begin{center}
-  \includegraphics[scale=0.4]{files/emails.png}
-\end{center}
-
-Faculty received many "cold" e-mails from international students seeking for admission, TA, and RA.
-Most of the time, we ignore these emails, but in some rare occasions we do answer them.
-So why do we ignore some but reply to others and how to write an email that get our attention?
-
-First, if you want to contact a prof.
-to \emph{ask about your admission chance}, please \textbf{don't}.
-We don't know and can't answer because as explained in \autoref{sec:evalapps}, we don't make individual decisions and might not even be assigned to evaluate your application.
-It is the same as sending a paper draft to a journal editor and ask them if your paper has a chance.
-
-So what to do if you want someone to look at your profile and give input?
-You could ask your professors, collaborators, or those who have previously applied.
-For these kind of feedback,  don't ask strangers like random profs., instead ask someone you have personal connection with.
-
-If you want to contact a prof.
-to ask about \emph{research opportunities}, or \emph{GTA/GRA} support, then \emph{yes}, I believe you should---it is \emph{worth it}.
-However, you need to put effort into it and really do it the \emph{right way}.
-
-First, read the prof's website, see if they say something about contacting them.
-Many profs.
-explicitly indicate how prospective students should (or should not) contact them (e.g., using specific email subjects).
-In general, the best way to catch the prof.'s attention is to \emph{customize your email} for them.
-For example, read their papers, know what they work on, and see if you are interested in their research.
-Then send them an email talking how/why their work would match yours.
-In contrast, if you write a generic email that can be sent to multiple professors (e.g., if you just change some names and keywords in the email or copy and paste paper titles), you will not get a response.
-
-Below is a good example that I would reply to.
-
-\begin{commentbox}[Dear Prof. Nguyen,]
-
-  I am writing to inquire about potential research opportunities as a GRA in your group at GMU. Currently I am an undergraduate student in Computer Science at UNIV and plan to graduate in May 2023.
-
-  I have read your TSE'21 paper on numerical invariant generation, and I am interested in this line of dynamic invariant research.
-I have worked (optional: with prof.
-Y at Z) on static program analysis and I think it could be used to tackle the spurious issues mentioned in your paper.
-I have a small paper at conference/workshop C and a project on symbolic execution at Github G.
-
-  ...
-
-  This is a good example because it is clearly written just for me.  It shows that the student knows about my work on invariant generation and has related  background (paper C and project G).
-\end{commentbox}
-
-Finally, profs. are very busy so don't take it personally if you don't get anything from them (though I would be very surprised if such thoughtful emails get no replies!).
-
-\begin{commentbox}[Xiaokuan]
-  Applying for Ph.D. and contacting a potential Ph.D. adviser is a classic \textbf{`why me, why you'} problem,
-  similar to looking for a job in a company.
-  On a high level,
-  you need to show that you have done your \emph{homework}
-  regarding the professor and the university,
-  and clearly explain:
-  1) why do you think you are a good fit in professor A's group?
-  2) why do you want to be advised by professor A, not B?
-  3) why do you want to apply for university X, not Y?
-  If you don't want to spend time to do your homework,
-  the chance of getting a reply is close to zero.
-\end{commentbox}
-
-\begin{commentbox}
-  [Deepak:]
-  In my view, cold emails are not welcome by most faculty members and should be avoided.
-However, if one is already admitted to a program in some department, by all means, send an email to the faculty you may be interested in working with, but do mention right at the beginning that you are already admitted to the program as well as several other universities.
-State specific areas (preferably specific topics-ML, robotics instead of AI).
-\end{commentbox}
-
-\subsubsection*{Additional Resources}
-\begin{itemize}
-  \item \href{https://yonatanbisk.com/emailing_professors.html}{A Note about Emailing Professors} by Yonatan Bisk
-  \item \autoref{sec:accept-postpone-decline} How to accept/postpone/decline PhD offers (and do it gracefully)
-\end{itemize}
-
-\subsection{PhD in other related fields: CE, IST, Cybersecurity}\label{sec:related-fields}
-
-In many cases you do not need to do a PhD in CS to work in your area of interested.
-For example, in addition to a traditional CS department, GMU has IST and Cybersecurity departments, which have faculty  with PhD in CS and work on CS topics (e.g., AI, Security, Robotics).
-So it is totally possible that you still get to do CS research and publish in CS-focused venues even if you're not in a traditional CS program.
-It is  common to see faculty with CS PhD in a non-CS department as well as faculty with non-CS PhD in CS department.
-
-However, if your goal is a PhD in CS, then you likely need to be in the CS dept \emph{and} advised by a CS faculty.
-In fact, if a faculty is not in CS, it is unlikely that they can be the \emph{main} adviser of a CS PhD student.
-They may \emph{co-advise} or be in the PhD dissertation committee, but your main adviser will need to be a tenured or tenure-track faculty in CS.
-If in doubt, check with the CS department for their requirements.
-For this specific reason,  CSRankings includes only tenure line faculty who can advise CS PhD students.
-I also have compiled a \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{list} of Vietnamese faculty who can advise PhD students.
-\autoref{sec:faculty-types} talks more about who can serve as your PhD adviser.
-
-\chapter{Miscs and FAQs}
-\epigraph{"I want to share something with you – the three little sentences that will get you through life; number 1: Cover for me, number 2: Oh, good idea, Boss, and number 3: It was like that when I got here."}{\textsc{The Simpsons}}
-
-\section{What can you do to increase your admission chance?}\label{sec:improve-your-chance}
-
-\begin{center}
-  \includegraphics[scale=0.2]{files/alpinist-climbing-peak-mountain-comic-hand-drawn-vector-illustration.jpg}
-\end{center}
-
-Show something that makes you \textbf{stand out}, e.g., do you have a degree or background in \emph{dance} or \emph{music} and want to integrate them with CS?
-are you a female or a minority in CS (research for "URM minority in CS in the US" on Google)?
-Do you participate in outreach activities that help increase diversity and inclusion in CS?
-All of these are unique and would get noticed from reviewers.
-
-Even if you do not have research experience, you can talk about your personal projects, as long as they can help show you can do research.
-For example, if you have an open-source project on Github that is used by many people, has lots of stars in Github, do talk about it.
-If you write technical, research-like blogs, talk about them too.
-
-In his \href{https://matt.might.net/articles/how-to-apply-and-get-in-to-graduate-school-in-science-mathematics-engineering-or-computer-science/}{post}, Matt Might was initially unsure about an application.
-However, upon learning that the applicant had led a 100km hike in the Himalayas, he decided to accept the applicant.
-This is a good example of being "stand out", and I would also advocate for that student as this shows they have the persistence and determination required for research.
-
-% It is worth noting that in many cases,  an application could be rejected by most reviewers, however have a "champion", who doesn't even have to be in the Admission comittee, that strongly advocate for your profile.
-
-\section{Can I apply to CS PhD if my undergrad was not in CS or related areas?}\label{sec:non-stem}
-
-Yes, \emph{as long as} you can demonstrate you are ready for CS PhD research through research experiences, LoRs, statements, etc as mentioned.
-You might be even able to leverage this to make your profile stand out as mentioned in \autoref{sec:improve-your-chance}.
-
-One main concern for non-CS or non-STEM students is if you have the sufficient technical background, typically obtained through core CS courses.
-So you will want to show that you have such knowledge through your coursework, projects, or research.
-For example, if you have taken a course on Algorithms, even online ones like Coursera, you can talk about it in your statement.
-If you have done a project that requires knowledge of OS or have a professional certification (e.g., A+) through work, you can talk about it in your statement.
-If you have done research that requires knowledge of Discrete Maths, you can talk about it in your statement.
-You can also ask your LoR writers to talk about your technical background.
-In summary, in your application, convince us that you have the technical background to do CS PhD research. 
-Note that many online courses teaching about AI does not really help much as these are not core CS knowledge.
-
-In short, you \emph{do not need} to formally taking CS courses, you just need to show that you have these essential knowledge through ways as mentioned.
-Many universities are well-aware that incoming graduate students might not have all the technical background, so they often have a "bridge" courses to help students catch up.
-For example, GMU has four bridge courses (Data Structures and Algorithms, Computer Systems, Discrete Math, and Programming Foundation) that incoming students can take to catch up on their CS knowledge.
-
-\begin{commentbox}[Vu:]
-  I would strongly advocate for a non-STEM student who shows that they have a strong drive for CS by studying core CS knowledge through various channels (e.g., self-study through online courses, projects, etc).
-I have seen many students with non-CS background who are very successful in CS PhD.
-I also have seen many students with CS background who are not successful in CS PhD.
-So it is not about your background, it is about your drive and passion for CS research.
-\end{commentbox}
-
-\section{Is an MS degree required for admission to PhD in CS?}\label{sec:msrequirement}
-No. In fact, student with BS can get MS degree ``along the way'' to PhD.  However, MS can help if it gives research experience or is from a more well-known school than your undergrad institution.
-
-If you have an MS then some course work \emph{might be} transferred for course credits, which save a bit of time. But overall don't count on this, especially if your MS is not from the US.
-
-\section{How long does it take to complete the CS PhD program?}\label{sec:time}
-
-\begin{center}
-  \includegraphics[scale=0.3]{files/c4a.png}
-\end{center}
-
-Typically, it takes 5--7 years for CS PhD in the US.
-This can be longer than CS PhDs in other countries, which might require MS first (recall that CS PhD programs in the US do not require MS and you can get MS along the way to PhD).
-Within these 5--7 years, CS PhD students often take a ``leave of absence'' for 1--2 semesters to do internship at companies and research labs.
-
-The first 2 years you typically take coursework (somewhat equivalent to an MS study), find an adviser, learn how to do research.
-The next 2--3 years you focus on your research, form dissertation topic, and get results published.
-The last 1-2 years you continue to publish, write and defend your dissertation, and look for job.
-In many cases you might take a summer or two off to do internship to get additional research opportunities.
-The PhDComics figure on top shows the ``ambition'' level of a PhD student over their years of study (they miss the 6--7th year where the ambition is \emph{``Just let me graduate''}).
-
-\section{Differences between PhD in the US and Other Countries}\label{sec:non-us-differences}
-
-This summarizes the main differences between CS PhD in the US and other countries. % This is based on my experience and what I heard from others who did PhD in other countries.
-
-\emph{MS Degree requirement}:  as mentioned in \autoref{sec:msrequirement} and \ref{sec:time}, CS PhD programs in the US do not require MS degree.
-In contrast, many other countries do require MS degree before joining a PhD program.
-This means that US PhD programs are longer (5--7 years, 2 of which are coursework) than other countries (3--4 years, no coursework).
-
-\emph{Project proposal}: in many countries, you have to choose a project and adviser \emph{during} the application process (e.g., you write a proposal to a potential adviser).
-In the US, you often start your PhD without an adviser or project and find them later.
-This is because as someone who just finished your undergraduate study, it is unlikely that you know enough about research to do proposal or picking an adviser.
-This is also why US PhD programs are longer.
-
-\emph{Course work}: in the US you will spend the first couple of years taking classes and explore potential adviser and research topics.
-In other countries, you (who already have an MS) start your research right away, e.g., you immediately work on the research project you proposed with the adviser you chose.
-Also, in the US you also have to pass a series of "exams", e.g., qualifying exam, comprehensive exam, thesis proposal defense\footnote{The word "ABD" (all but dissertation) is used in the US to refer to a PhD candidate who have finished all course work and exams and only need to write and defend their dissertation.}.
-In other countries, you do not have to do these exams or only do a few of them.
-
-\emph{Funding}:  In many countries stipend comes from the university or from the gov't.
-These funding often have a fixed duration (e.g., 3 years).
-In the US, stipend (e.g., RA) comes directly from your adviser (no fixed duration).
-There are also fewer TA opportunities in the European universities compared to the US.
-
-\emph{Faculty position after PhD}: In other countries, PhD graduates interested academia typically apply for additional research appointments, i.e., postdocs, and then consider faculty position.
-In the US, PhD graduates often apply directly for faculty position (postdoc for US graduates is no longer a popular option as it was before).
-%This is also why US PhD programs are longer (remember, 5--7 years).
-
-\emph{Work-life balance}: PhD students in the US are often said to be overworked compared to other countries, e.g., in Europe.
-This is partly due to the longer PhD program and that US PhD students are often paid through TA, which requires them to do TA in addition to their own research.
-In contrast, PhD students in other countries are often paid through fellowships, which do not require them to do TA.
-
-\section{How do I call or address a professor?}\label{sec:address}
-
-\begin{center}
-  \includegraphics[scale=0.5]{files/c5.png}
-\end{center}
-
-If you're reaching out to a professor for the first time,  address them as Prof.
-or Dr.
-Lastname.
-Many international students use Prof.
-or Dr.
-FirstName LastName, but this can come across as if you're simply copying and pasting names.
-It's not necessary, so stick with Prof.
-or Dr.
-Lastname.
-
-Furthermore, avoid using Mr.
-or Mrs., or the professor's first name if you're not acquainted with them yet.
-As you become more familiar with your prof and depending on their preferences, you may transition to addressing them by their first name.
-For example, I prefer that my students and colleague call me Vu. Some students call me \emph{Dr. Vu}, which I find a bit amusing but am totally fine with it.
-
-\begin{commentbox}[DK:]
-  I was amused to read this as if I recall correctly, you never called me by my first name when you were at UNM. You always called me Prof. And, many times, I would jokingly call you back as Prof. Vu.
-  \tcblower
-  \textbf{Vu}: Yes, for some reason I enjoy addressing you as ``Prof.'' (without appending a last or first name).  The use of Prof. Vu may have foreshadowed my future in academia.
-\end{commentbox}
-
-Note that in some universities the formal title Dr.
-Lastname is preferred over Prof.
-You just need to observe and follow the conventions at your particular institution.
-Additionally, be aware that not all faculty members may hold a Ph.D., in which case using Prof.
-Lastname is a suitable alternative.
-
-\paragraph{Referring to professors you know} Because you are already familiar with these individuals, you can just informally use their names if they are OK with it as mentioned above (or Dr./Prof., if you want to be formal).
-You can also include their institution if it makes it more precise.
-For example, I can say:  \emph{"I did my postdoc with Jeff Foster at Univ.
-of Maryland"}.
-
-Do not include ranking (e.g., Assistant, Associate, Scientist, ...) when referring to someone.
-I see many international students include a lengthy title of people they know, e.g., \emph{I am advised by Asst.
-Prof.
-X, and I also collaborate with Distinguished Scientist Y}.
-This is \emph{not necessary} and makes it look like you're trying to show off your connections.
-These nuances represent some cultural differences that you may encounter and will gradually adapt to.
-More on cultural differences in \autoref{sec:cultural}.
-
-\section{How much do \emph{you} cost?}\label{sec:ra-cost}
-PhD students often ask why their salary is so low compared to ludicrous grants their advisers get.
-They also wonder why their offer letters sometimes show that their benefits higher than what they actually receive (i.e., stipend).
-This section aims to shed some light to these questions.
-
-\begin{center}
-  \includegraphics[scale=0.3]{files/c6.png}
-\end{center}
-
-\begin{table}
-  \centering
-  \small
-  \caption{GRA cost breakdown. F \& A is Facilities \& Administrative Cost Base and
-    MTDC is Modified Total Direct Cost. These are things that the university can charge overhead to.}\label{tab:cost}
-  \begin{tabular}{rcl}
-    \textbf{Budget} & \textbf{Cost \$} & \textbf{Notes} \\
-    \midrule
-    GRA (9-month) & 27K & \\
-    GRA (summer)  &9K	& 3-month, 20hrs/week\\
-    \textbf{Total Salary} &\red{36K}	&\\
-    \midrule
-    Health Insurance	&3K	& full year\\
-    Tuition (In-State) &	15K	& (\$680/ Credit + \$150/Student Fee/ Credit)* 9 credits = \\
-                    &&\$7470 (\$6120 + \$1350) per semester\\
-    \textbf{Total Tuition \& Insurance}	&\red{18K}	&Full year tuition + insurance\\
-    \midrule
-    Conference Registration	& 500 & \\
-    International Travel &	1800& \\
-    Domestic Travel	& 700	& \\
-    \textbf{Total Travel}&	\red{3K}	\\
-    \midrule
-    Total Direct Cost	& \red{57K}	&Salary + Travel + Health + Tuition \\
-    F \& A (MTDC)	& 21K	& Direct Cost - GRA Salary\\
-    Total Indirect Cost	& \red{12K}	&58.9\% of MTDC\\
-    \textbf{Total (Direct + Indirect)} &	\red{69K}	& Budget for a GRA\\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
-\autoref{tab:cost} shows the budget breakdown for a GRA per year.
-These numbers are based on my experience at public universities in the US.
-Private universities may have different numbers.
-For simplicity, in this table I will assume the department has a 9-month stipend of \$27000 (GMU actually pays more) and therefore a 3-month summer of \$9000.
-I will also use GMU tuition rate of about \$15,000/year for full-time study (which is quite cheap compared to private universities, e.g., MIT charges around \$50K) and a 58.9\% rate on \emph{indirect cost}, which is what GMU charges for overhead or administrative costs (yes, after all, universities are businesses!).
-Finally, I assume the student take two conference trips per year, one domestic and one international (conf.
-registration, airline tickets, taxi, meals, etc are all included).
-
-At the end, the total budget comes out to be \$69K/year to support a PhD student.
-\emph{The summary is that over your 5-6 year of your PhD, you cost about \$350K--400K, and while you're paid X, your adviser probably pays 2X for you}.
-
-% \section{Having fun during a PhD?}
-% PhD students \emph{and faculty} probably find it amusing about the notion that students, especially international ones, can genuinely enjoy their PhD studies.
-In fact, after reading posts after posts on VietPhD.org on how PhD students are commonly mistreated, stressed, it seems being miserable is a norm during a PhD study.
-
-% There are many advice on surviving PhD that you can follow. But here I just list a few that works for me and what I advice my students to do.\tvn{TODO}
-
-\section{Will I be miserable during my a PhD?}\label{sec:happy}
-There are many stories on how students are mistreated, stressed, and miserable.
-Issues including bad relationships with professors, conflicts with co-authors and lab mates, feeling discriminated (e.g., because you're an international student), etc \emph{do} exist, and it is good to be aware of those.
-However, in reality there are many good mentors, supportive lab mates and department, and so on.
-So don't let social media make you feel pessimistic and deter your quest to advance knowledge.
-
-\appendix
-
-\chapter{Rankings of CS PhD programs}\label{sec:ranking}
-
-\begin{table}
-  \centering
-  \small
-  \caption{Top 50 CS PhD programs in the U.S.
-(\href{https://csrankings.org}{CSRankings}, Jan.
-2024).
-\red{$^*$} indicates that the university has \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{Vietnamese prof.} that can advise CS PhD students.}\label{tab:ranking}
-  \begin{tabular}{rl|rl}
-    \toprule
-    1 & Carnegie Mellon & 26 & Duke University \\
-    2 & Univ. of Illinois at Urbana-Champaign\red{$^*$}  & 27 & Univ. of California - Santa Barbara \\  
-    3 & Univ. of California-San Diego & 28 & Rutgers University\red{$^*$} \\
-    4 & Georgia Institute of Technology & 29 & Univ. of California - Riverside\\
-    5 & MIT                            & 30 & Pennsylvania State University  \\
-    6 & Univ. of California - Berkeley& 31 & Northwestern University\\
-    7 & University of Michigan - Ann Arbor\red{$^*$}   & 32& George Mason University\red{$^*$}\\  
-    8 & University of Washington      &33 &  Texas A\&M University\red{$^*$} \\
-    9 &  Stanford University  &34&  University of Utah \\
-    10 & Cornell University  & 35 &  North Carolina State University\\\  
-    11 & University of Maryland - College Park &  36& Ohio State \\ 
-    12 & Northeastern University\red{$^*$} &37& University of Virginia  \\ 
-    13 & Purdue University &38& Yale University \\ 
-    14 & University of Wisconsin - Madison\red{$^*$} &39 & Univ. of California - Santa Cruz \\
-    15 & University of Texas at Austin &40& Brown University \\ 
-    16 & University of Pennsylvania &41 & Harvard University \\
-    17 & Columbia University\red{$^*$} &42 & Boston University  \\
-    18 & Princeton University\red{$^*$}  & 43& Rice University\\
-    19 & New York University  & 44&  University at Buffalo\red{$^*$}\\
-    20 &  University of Massachusetts-Amherst\red{$^*$} &45& University of Colorado-Boulder \\
-    21 & Univ. of California - Los Angeles &46& University of Illinois at Chicago  \\
-    22 & University of Southern California &47& University of Minnesota \\ 
-    23 & University of Chicago &48& University of North Carolina\red{$^*$} \\
-    24 & Stony Brook University\red{$^*$} &49& Arizona State University\red{$^*$} \\
-    25 &  Univ. of California - Irvine&50& Univ. Of California - Davis \\
-    \bottomrule
-  \end{tabular}
-\end{table}
-% Virginia Tech\red{$^*$}  \\
-\autoref{tab:ranking} lists the top 50 CS programs in the US from \href{https://www.csrankings.org}{CSRankings.org}, a ranking system  based on faculty publications at top CS conferences.
-
-% \section{Additional Resources}
-% \begin{itemize}
-%   \item \bibentry{nlpphd}
-% \end{itemize}
-
-% \chapter{Additional Resources}
-
-% \chapter{Undergraduate Research Opportunities}
-
-% Prospective PhD students are often unaware of the importance of research experience in their application.
-This provides both concrete research experiences and potential LOR writers that can talk about your research ability.
-% % This is especially true for students from small colleges where research opportunities are limited.
-% This section aims to provide some guidance on how to gain research experience as an undergraduate student.
-%This section is written together Didier, a PhD student at GMU who did his undergraduate study in Rwanda.
-
-% \paragraph{Locally} There are many ways to find research opportunities right in your own institution.
-If you did well or like a class, you can check with the professor of that class for research opportunities.
-Sometimes professors would mention that they look for undergraduates on their websites, so do check those out!
-Even if they say they are looking for graduate students, you still should contact them and ask.
-You can also ask your academic adviser or other faculty members for suggestions.
-Finally, you can also ask your peers who are already doing research.
-You can use the method described in \autoref{sec:contact} to contact professors
-
-% In short, the most important thing is that you need to \emph{ask}!
-Don't be shy, don't be afraid of rejection, and don't be afraid of being ignored or that you're not good enough.
-Just ask!
-The worst thing that can happen is that you get a ``no'' or no response.
-But you will never know if you don't ask.
-And if you get a ``no'', you can always ask again later.
-There are students who got rejected the first time but got accepted the second time.
-So don't give up!
-
-% \paragraph{Online Research Communities and Open Source Contributions}:
-% Places such as Github offer many opportunities for research, even though they might not appear so (e.g., they do not explicitly mention research).
-Students can explore repositories related to their areas of interest, collaborate with other contributors, and actively engage in open source research projects.
-By contributing code, fixing bugs, implementing new features, or providing documentation, students can gain practical research experience and interact with experienced developers and researchers in the field.
-
-% \paragraph{Paid or Unpaid}  travel opportunities, publishing opportunities, etc
-
-% \section{Virtual Research Opportunities Beyond Physical Boundaries (written by Didier)}
-% \didi{In my experience, prospective PhD students (from Rwanda) are often unaware of the information provided in this document throughout their undergraduate.
-For this reason they often find themselves unprepared for PhD admission process by the time they finish undergraduate and have to spend a year or more to improve their application profiles.
-I think it would be beneficial to expand the document to also target students who are not yet ready for the application process to help them improve on all criteria considered during admission.
-I drafted the short essay on how to gain research experience remotely in order to improve their research profile and potentially get LOR from experts (hopefully faculty)}\\
-% In the realm of computer science, research experience plays a crucial role in securing admissions to top-tier Ph.D.
-programs.
-However, students from underrepresented small colleges, both in the United States and internationally, often encounter limited research opportunities within their institutions.
-Thankfully, there exists an alternative pathway for these aspiring researchers to gain valuable research experience, even without being physically present at a university.
-
-% \textbf{Virtual Research Programs}:
-% Several universities, organizations, and research institutions offer virtual internships and research programs aimed at providing hands-on research experience.
-These programs often involve working remotely under the guidance of experienced mentors and collaborating with a team of fellow researchers.
-For instance, \href{https://docs.google.com/forms/d/1btIwt4HwjyKMOUk-EMy3rbkfWzFxv2lNrMm_zkd0pA4/viewform?edit_requested=true}{UIUC+ Summer Undergraduate Research in Software Engineering}  offers an unpaid remote internship for software engineering students all over the world to collaborate with mentors from University of Illinois at Urbana Champaign.
-% Virtual internships offer opportunities to contribute to ongoing research projects, conduct experiments, analyze data, and write research reports.
-Students can search for such programs through university websites, research institutes, or as of recently chatGPT (or Bard).
-
-% \textbf{Online Conferences and Workshops}:
-% Attending online conferences and workshops is another way to gain exposure to cutting-edge research and establish connections with experts in the field.
-Many conferences now provide virtual participation options, enabling students to access research talks, poster sessions, and panel discussions and sometimes access designated chat rooms or networking events where participants can engage with researchers, ask questions, and seek potential research collaborations.
-It is beneficial to also create profiles academic collaboration platforms such as ResearchGate.
-By creating profiles, joining relevant research groups, and participating in discussions, students can connect with established researchers, contribute to ongoing projects, and potentially collaborate on publications or research proposals as they provide access
-
-% In conclusion, for students lacking research opportunities at their small colleges, the virtual realm offers a wealth of possibilities to gain valuable research experience.
-Engaging with open source projects, joining online research communities, utilizing academic collaboration platforms, participating in virtual internships, and attending online conferences are all avenues to contribute to research, connect with experts, and enhance one's research profile.
-
-\chapter{Cultural Differences and Other Issues}\label{sec:cultural}
-
-This section lists some cultural issues that international students might want to pay attention to. These issues apply in general and not just PhD admission.
-
-% \paragraph{Diversity} US universities prioritize diversity and inclusion.
-Students need to respect and appreciate varied opinions, backgrounds, and cultures.
-Unlike some countries where certain voices are marginalized, in the US, all perspectives are valued equally (especially at universities).
-Racism or discrimination will have serious consequences, including academic and disciplinary actions.
-
-\section{Accepting, Postponing, and Decline Offers}\label{sec:accept-postpone-decline}
-
-Students, especially international ones, often get confused on what to do after they get an offer from a professor, e.g., how to postpone or accept/reject offers, and most importantly, how to do so without offending anyone.
-In short, if you are not serious about an offer for any reason (e.g., still waiting for other offers), do not accept, i.e., commit to it.
-If you are not sure, ask for more time to decide.
-But do not accept an offer and then later reject it.
-This is considered \emph{rude and unprofessional}, and will likely burn bridges with that professor.
-
-\paragraph{Accepting an offer} If you decide to accept an offer, you can say: \emph{``Thank you for the offer.
-I would like to accept it and look forward to working with you.
-Could you please send me more details about the offer and what to do next?''}.
-Here you are not only accepting the offer but also asking for more details.
-The prof.
-will likely send you more details about the offer and what to do next.
-If you decide to accept an offer, do so quickly.
-
-If you have better offers, you might mention them to the professor and ask if they could match them.
-It is unlikely they can do so as RA might be fixed by the department and not negotiable.
-They might recommend you for departmental fellowships or some additional perks.
-
-\paragraph{Postponing an offer} If you need more time to decide, you can ask for more time: \emph{``Thank you for the offer.
-I am very excited about it.
-However, I am still waiting for other offers and need more time to decide.
-Would it be possible to postpone the decision for a few weeks?''}.
-This is perfectly fine and professors will understand and might even appreciate your honesty.
-They will likely give you a few weeks to decide.
-If you need more time, you can ask for more time.
-But do not ask for too much time, e.g., more than a month.
-You also should not postpone the offers multiple times, which will annoy people.
-
-\paragraph{Declining an offer} If you decide to decline or reject an offer, you can say: \emph{``Thank you for the offer.
-However, I have decided to accept another offer.
-I appreciate your time and consideration.
-I hope we can work together in the future.''}  Professors will understand and wish you good luck.
-If you decide to reject an offer, do so quickly.
-
-This is very similar to a professional job offer.
-You should be serious about it and communicate promptly.
-Importantly, do not accept an offer and then later reject it.
-If you do this with a company, you might face legal consequences.
-If you do this with a professor, they will likely not do anything about it (i.e., not worth their time) but they will not want to work with you in the future.
-
-% 1.
-I'm fine if I extend an offer and you request a delay, even if it's because you're exploring other programs.
-No issue here, it's similar to job offerings.
-Of course, I may not be able to wait indefinitely and may extend the offer to another student, but I'll let you know that.
-% 2.
-I'm also okay if you find a program or professor that's a better fit.
-Similar to the first point, no hard feelings at all.
-However, I'd appreciate a prompt notification of your decision so I can extend an offer to another student.
-% 3.
-I'm NOT okay if you accept my offer but then, upon finding something better, tell me of your decision to withdraw Just like in job offers, once you are committed to an offer, there should be a compelling reason to withdraw (and claiming "I found somewhere better" is the weakest excuse).
-% While I won't take any actions against you (a company might), I naturally will remember this and may be cautious in future interactions.
-I'll also reconsider making similar offers to students in similar situations.
-
-\section{Academic Integrity (Cheating and Plagiarism)}
-
-Plagiarism and cheating (e.g., exams, assignments) is a BIG no-no in the US.
-If you're caught cheating, you will face very heavy consequence and likely be expelled from the university (e.g., after the second time at GMU).
-This is quite different from many international countries where cheating is common and often tolerated.
-Faculty is extremely good at detecting cheating (we have been dealing with these situations so many times over so many years), and \emph{will} report cheating cases.
-In short, whatever you do, don't cheat---not worth it.
-
-Here is the typically steps: (i) a faculty suspecting a cheating case will report it to the Office of Academic Integrity (OAI) at the university---the report often has supporting evidence and suggested penalty (e.g., a failing grade);  (ii)  OAI will take over and investigate the case; and (iii) OAI will make the final decision.
-It is important to note that after receiving the report from your prof., OAI \emph{completely} takes over and makes decision.
-This means begging your professor will not help because they simply are no longer involved in the case.
-
-\section{Illegal Software} Using illegal/cracked software is very common in many countries (and even in the US).
-However, \emph{do not} install or use them on university computers, even those given to you from your adviser.
-It is unlikely that the university will track you down, but it is the \emph{software company} that will.
-They have very sophisticated tools to detect illegal software and will sue your university/department.
-Imagine your department or adviser being sued for a large sum of money, and it is \emph{you} who caused it.
-If you need to purchase software,  ask your adviser or the department.
-
-\section{Gifts}\label{sec:gifts} In many countries, it is customary to give professors gifts, often during holidays.
-These gifts can be costly and profs.
-sometimes expected them.
-In the US, it is uncommon and perfectly OK to not to give gifts.
-\emph{However}, if you'd like to offer small souvenir-like tokens, it's a thoughtful gesture that's appreciated.
-Some professors proudly display their gifts, which can come from students and colleagues (e.g., when they travel to their home countries or conferences).
-In summary, small gifts are fine, but avoid anything that might make your professors uncomfortable.
-
-\section{Relationships with your Profs.}
-There's a misconception that in the US it's all business, with professors as bosses who pay students for their work, and you doing something nice implying you expect something in return.
-However, the reality is quite the opposite.
-While people can be straightforward and direct, they are also friendly and informal.
-You can call your professor by their first name (\autoref{sec:address}), disagree with them and argue (and gain respect doing so), seek their help (even on personal matters), come to their houses for parties or gathering, and give them small thoughtful gifts that they put on their desks (\autoref{sec:gifts}).
-Many students and professors maintain lifelong relationships, staying in touch through cards, emails, and calls, even after their academic journey ends.
-%These connections are genuine and meaningful.
-
-% \begin{commentbox}
-%   [DK:] Here are a few items you may consider addressing: putting international students in touch with other students from their countries, assuring them that they would be picked up from airports upon their arrival and that their initial stay will be taken care of.
-Most universities have other resources for these, but it is worth mentioning that they would be taken care of upon arrival and can get help during the transition phase.
-Learning to cook was a big deal when I arrived over 50 years ago-August 1973.
-But things have changed as one can find many ethnic food places, a big contrast from 1973, when there were two Indian restaurants in Greater Boston area.
-% \end{commentbox}
+\section{Serving on NSF Panels}
+You should serve on NSF panels.  It is a great way to learn about the grant process: how to distinguish between good and bad proposals and how people review propsoals.
+It is also a good way learn what other people are doing, what is hot, and what is not.
 

### 2024-02-22 demystify.tex
updates

@@
 did your reference writers submit their letters?
-Usually, this screen process is done through a central university system (i.e., not by the CS department).
+Usually, this screen process is done through a central university system, i.e., not by the CS department.
 If you pass this screen, your application will be forwarded to the CS department for further evaluation.
-If you don't, then the system will inform you what is missing and what you need to do.
+If you don't,  the system will tell you what is missing and what you need to do.
+So pay attention to your email and check your application status regularly.
 
@@
 
-Then, applications are reviewed by a \textbf{PhD admission committee} (\emph{adcom}) that consists of faculty members in CS (in some cases the committee can involve affiliated faculty from different disciplines).
+Then, applications are reviewed by a \emph{PhD admission committee} (\textbf{adcom}) that consists of faculty members in CS (in some cases the committee can involve affiliated faculty from different disciplines).
 These faculty have a wide-range of expertise and background to ensure diverse perspectives in the evaluation process.
@@
 At GMU, the PhD adcom typically has 15--20 faculty, and each committee member is assigned to review 25--30 applications.
-Note that GMU, UMass, and probably many other schools, have separate adcom for MS programs.
-
-Each application is assigned to about \emph{three} faculty members, who will evaluate your profile and reach a consensus.
-Note that while the assigned reviewers are likely the main ones deciding your application, \emph{every} faculty in the department will have access to your application and can provide inputs and opinions on your profile.
+Note that many schools, including GMU, have separate adcoms for MS programs.
+
+Each application is assigned to 2--3 faculty members, who will evaluate your profile and reach a consensus.
+Note that while the assigned reviewers are likely the main ones deciding your application, other faculty in the department will likely have access to your application and can provide inputs and opinions on your profile.
 
@@
 This provides junior faculty the opportunities to recruit students.
-The chair of the committee will be a senior professor, but they likely will not review individual applications and instead assign them to committee members.
+Some members can be tenured faculty.
+The chair of the committee will likely a senior faculty, but they likely will not review individual applications and instead assign them to committee members.
 The chair will look at various factors such as research interests or mentioning faculty names to assign the applications to appropriate faculty.
@@
 our decision with a summary of your application, where we list
-strengths (e.g., a well-known school) and weaknesses (e.g., weak
-LORs).
+strengths, e.g., a well-known school, and weaknesses, e.g., weak
+LORs.
 
@@
 \epigraph{"I want to share something with you – the three little sentences that will get you through life; number 1: Cover for me, number 2: Oh, good idea, Boss, and number 3: It was like that when I got here."}{\textsc{The Simpsons}}
+
+\section{Why did I get rejected?}
+Many possible reasons!  Other than the obvious ones (e.g., weak research pontentials, low GPA, weak LoRs, unclear SOP), here are some that you might not expect:
+\begin{itemize}
+  \item  \emph{Your English scores (e.g., IELTS or TOEFL) are not high enough}.
+While professors do not care much about these, there is a certain minimum set by the college or the university that you need to pass to get admitted, especially if you want to get funding (e.g., you need to be able to communicate well with students as a TA).
+Note this is mostly for international students, as domestic one do not have to take these tests.
+
+  \item \emph{No faculty available to advise you}.
+You could have an excellent profile, but if there is no faculty who can advise you (e.g., they already have too many students), you will not be admitted.
+This is actually for your own good because you do not want to be in a program where you cannot find an adviser.
+\end{itemize}
+
+\section{Should I ask for feedback?}
+No, don't bother.
+You will likely not get any useful feedback.
+We cannot reveal your evaluation results (e.g., if we say because you have weak LORs, then that would reveal something about your writers) or give you advice on how to improve.
+So it is best to just move on.
+If you really want feedback, ask your professors, collaborators, ref writers, or those who have previously applied.
 

### 2024-02-25 ms-swe-handbook.tex
Updates from Overleaf

\documentclass[oneside,11pt]{memoir}
\chapterstyle{pedersen}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[margin=1.5in]{geometry}
\usepackage{soul}
% \usepackage[small,compact]{titlesec} %very powerful
\usepackage[most]{tcolorbox}
\setsecnumdepth{subsection}
\setcounter{tocdepth}{3}
\usepackage{enumitem}
\usepackage{epigraph}
\usepackage{cite}
\usepackage{caption}
\captionsetup{font=small}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{wrapfig}
\setlength\intextsep{0pt} % remove extra space above and below in-line float
\usepackage{hyperref}
\hypersetup{
  colorlinks,
  citecolor=black,
  filecolor=black,
  linkcolor=blue,
  urlcolor=blue,
}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage[makeroom]{cancel}

\newtcolorbox{mybox}{
  enhanced,
  boxrule=0pt,frame hidden,
  borderline west={2pt}{0pt}{green!75!black},
  colback=green!10!white,
  sharp corners
}

\newenvironment{commentbox}[1][]{
  \small
  \begin{mybox}
    {\small \textbf{#1}}
  }{
  \end{mybox}
}

\newtcolorbox{mydomesticbox}{
  enhanced,
  boxrule=0pt,frame hidden,
  borderline west={2pt}{0pt}{red!75!black},
  colback=blue!10!white,
  sharp corners
}

\newenvironment{domesticbox}[1][]{
  \small
  \begin{mydomesticbox}
    {\small \textbf{#1}}
  }{
  \end{mydomesticbox}
}

\renewcommand{\figurename}{Fig.}
\renewcommand{\tablename}{Tab.}
\def\Section{\S}
\renewcommand{\figureautorefname}{Fig.}
\renewcommand{\tableautorefname}{Tab.}
\makeatletter
\renewcommand{\chapterautorefname}{\S\@gobble}
\renewcommand{\sectionautorefname}{\S\@gobble}
\renewcommand{\subsectionautorefname}{\S\@gobble}
\makeatother

\newcommand{\mycomment}[3][\color{blue}]{{#1{{#2}: {#3}}}}
\newcommand{\tvn}[1]{\mycomment{TVN}{#1}}{}
\newcommand{\didi}[1]{\mycomment{Didier}{#1}}{}
\newcommand{\tl}[1]{\mycomment{ThanhLe}{#1}}{}
\newcommand{\red}[1]{{\color{red}{#1}}}
\newcommand{\xz}[1]{\mycomment{Xiaokuan}{[#1]}}{}

\title{Demystifying the Computer Science \\
  PhD Admission in the US\\{\large A Guideline for International \emph{and Domestic} Students}}

\author{\href{https://nguyenthanhvuh.github.io}{ThanhVu (Vu) Nguyen}\\{\small George Mason University, Dept. of Computer Science}}

\makeatletter
\def\maketitle{%
  \null
  \thispagestyle{empty}%
  \vfill
  \begin{center}\leavevmode
    \normalfont
    {\LARGE\raggedright \textbf{\@title}\par}%
    \vfill%
    % \hrulefill\par
    {\Large \@author\par}%
    \vfill%
    {\large\raggedleft \@date\par}%
  \end{center}%
  \vfill
  \null
  \cleardoublepage
}
\makeatother

\begin{document}
\maketitle
\frontmatter

\chapter{Preface}

\newpage
\tableofcontents*

\mainmatter

\chapter{Admission}

Admission is competitive among students who fulfill admission requirements for graduate study at the Computer Science Department of George Mason University.
Students seeking admission to the MS in Software Engineering program must satisfy the following requirements:
\begin{itemize}
    \item Hold a four-year (120-credit) baccalaureate degree from an accredited institution
    \item Have a cumulative grade point average of 3.0 for the last two years of undergraduate work (60 credits)
    \item Submit transcripts of all post-secondary education
        \begin{itemize}
            \item Unofficial copies of transcripts are acceptable for application review. However, upon admission, applicants will need to provide official transcripts before enrolling in the program.
        \end{itemize}
    \item A one-page goals statement, and a work résumé
        \begin{itemize}
            \item Your goals statement should be one to two pages, including a statement of career goals.
        \end{itemize}
    \item The MS in Software Engineering program \textbf{does not} require letters of recommendation.
    \item Complete the self-evaluation section of the online application
    \begin{itemize}
        \item This information is used by the admissions committee to assess an applicant’s academic preparation for the MS program. In some cases, students with deficiencies in preparation may be admitted provisionally. See “Foundation Requirements” below for more information.
        \end{itemize}
    \item GRE is \textbf{not required}
    \item TOEFL or IELTS is required of all applicants whose native language is not English, and who have not received a degree from a university in the United States, Canada, United Kingdom, Australia, or New Zealand.
    \begin{itemize}
        \item Required TOEFL iBT score: \textbf{80 points total overall with no section minimum}
        \item Required IELTS score: \textbf{6.5 overall}
    \end{itemize}
\end{itemize}

% Foundation Requirements

% To ensure that students have an adequate background in mathematical methods and computer science, the program requires the following four foundation courses, or their equivalents:

%     INFS 501 Discrete and Logical Structures for Information Systems
%     INFS 515 Computer Organization
%     INFS 519 Program Design and Data Structures
%     SWE 510 Object Oriented Programming in Java

% Prospective students are asked to indicate on their application if previously taken courses may satisfy these foundation requirements. Provisional admission is offered if a student has some deficiencies in preparation, but is otherwise a strong applicant. (Note: Due to federal requirements, students on F1/J1 visas are not eligible for provisional admission.) In such cases, students are advised of the necessary foundation courses to be satisfactorily completed with a grade of B or better before beginning the core curriculum. Foundation courses do not earn credit toward the MS degree.

% Students have one opportunity to test out of their required foundation courses before beginning their first semester. The exams are given before classes begin in January and August and may only be taken once. Students failing any one of the exams must take the equivalent course before enrolling in the core curriculum courses.

% More information regarding the foundation courses and test out exams can be found on this page.

\chapter{Foundation and Required Courses}

Students entering the MS-ISA, MS-ISYS, and MS-SWE programs must have course work or equivalent knowledge in the following five foundation areas: (1) introductory programming in any programming language; (2) knowledge of an object-oriented programming language such as Java and C++; (3) data structures and algorithms; (4) machine organization such as those given in computer system architecture or assembly language courses; (5) and topics in discrete mathematics, including sets, propositional and predicate logic,  relations, functions, trees, graphs, and inductive proofs. The level of knowledge required in these areas is equivalent to that taught in undergraduate courses, and may be achieved by taking the following George Mason University graduate courses:

\begin{itemize}
\item COMP 501 Computer Programming Foundations I or SWE 510 Object Oriented Programming in Java
\item COMP 502 Mathematical Foundations of Computing I or INFS 501 Discrete and Logical Structures for Information Systems
\item COMP 503 Computer Systems Foundations I or INFS 515 Computer Organization
\item COMP 511 Computer Programming Foundations II or INFS 519 Program Design and Data Structures
\end{itemize}

\textbf{Provisional admission} is offered if a student has some deficiencies in preparation, but is otherwise a strong applicant. (Note: \emph{Due to federal requirements, students on F1/J1 visas are not eligible for provisional admission.}) In such cases, students are advised of the necessary foundation courses to be satisfactorily completed with a grade of B or better before beginning the core curriculum. Foundation courses \emph{do not} earn credit toward the MS degree. 
%More information regarding the foundation courses and test out exams can be found on this page.

Students have one opportunity to \emph{test out} of their required foundation courses before beginning their first semester. 
The exams are given before classes begin in January and August and may only be taken once. Students failing any one of the exams must take the equivalent course before enrolling in the core curriculum courses.

Additionally, if a student feels they have taken an equivalent course that was overlooked, an \emph{appeal process} is available. More information on these options can be found on the \href{https://cs.gmu.edu/current-students/ms-students/foundation-courses/policies-and-procedures/}{policies and procedures page}.

\autoref{tab:foundation-courses} shows undergraduate courses offered at GMU serving as equivalents for the corresponding foundation courses.

\begin{table}[]
    \centering
    \footnotesize
    \begin{tabular}{c|c|c|c|c}
      Foundation Topics   & GMU Grad  &	GMU Undergrad CS &	GMU Undergrad IT &	NVCC\\
      \midrule
      OO programming & 	COMP 501 or SWE 510 & 	CS 211 & 	IT 206 or IT 209 &	CSC 202\\
      Discrete math              &	COMP 502 or INFS 501 & 	MATH 125 & 	MATH 125 (not MATH 112) & 	MATH 288\\
      Machine org. 	& COMP 503 or INFS 515 &	CS 367 	&IT 342 	& none\\
      Data structures &	COMP 511 or INFS 519 &	CS 310 	& IT 306 or IT 309 &	none\\
    \end{tabular}
    \caption{Foundation Courses and Equivalences}
    \label{tab:foundation-courses}
\end{table}

\chapter{MS SWE}

The MS in Software Engineering (MS-SWE) program prepares students to become leaders in engineering high quality, large scale, computing solutions to real life problems. The MS-SWE program prioritizes working professionals—as such, all SWE courses are taught \emph{late afternoon} or \emph{early evening}.

\section{Degree Requirements}

Students are required to complete 30 credits corresponding to 10 graduate courses. Students are encouraged to download the course planner spreadsheet, and update it as they proceed through the program.

The course work is divided into three categories: 12 credits of core courses, software engineering related courses of 9 credits, and 9 credits of elective courses.

\subsection{Core Courses (12 credits)}

\textbf{Four core courses} are required of all MS-SWE graduates:

\begin{enumerate}
    \item SWE 619 Object-Oriented Software Specification and Construction
    \item SWE 621 Software Modeling and Architectural Design
    \item SWE 632 User Interface Design and Development
    \item SWE 637 Software Testing
\end{enumerate}

\subsection{Software Engineering Related Courses (9 credits)}

Students must take \textbf{three courses} from the following list:

\begin{itemize} 
\item Software Engineering (SWE)
\begin{itemize}
    \item Any SWE courses at the \href{https://catalog.gmu.edu/courses/swe/}{600 level or above}
\end{itemize}
\item Computer Science (CS)

\begin{itemize}
    \item CS 540 Language Processors
    \item CS 550 Database Systems
    \item CS 555 Computer Communications and Networking
    \item CS 571 Operating Systems
    \item CS 584 Theory and Applications of Data Mining
    \item CS 675 Distributed Systems
    \item CS 678 Advanced Natural Language Processing
\end{itemize}

\item Information Security \& Assurance (ISA)

\begin{itemize}
    \item ISA 562 Information Security Theory and Practice
    \item ISA 650 Security Policy
    \item ISA 673 Operating Systems Security
\end{itemize}

\item Information Systems (INFS)
\begin{itemize}
    \item INFS 740 Database Programming for the World Wide Web
\end{itemize}

\item Operations Research (OR)
\begin{itemize}
    \item OR 542 Operations Research: Stochastic Models
\end{itemize}

\end{itemize}

\textbf{Note}: Credit will not be given for both INFS 614 and CS 550, or both SWE 622 and CS 675

\subsection{Elective Courses (9 credits)}

Students may select the remaining courses from the following list. Students may select courses not on this list with approval from the faculty advisor. Students, with the consent of a faculty sponsor and faculty advisor, may also complete a 6-credit thesis, which is primarily intended for students considering pursuing a PhD.

\begin{itemize}
    \item All Software Engineering (SWE) courses at the 600-level or above.
    \item All Information Security and Assurance (ISA) courses at the 500-level or above.
    \item All Information Systems (INFS) courses at the 600-level or above.
    \item Any of the following Electrical and Computer Engineering courses:
    \begin{itemize}
        \item ECE 542 Computer Network Architectures and Protocols
        \item ECE 612 Real-Time Embedded Systems
\end{itemize}        
    \item Any of the following Operations Research courses:
    \begin{itemize}
        \item OR 531 Analytics and Decision Analysis
        \item OR 541 Operations Research: Deterministic Models
        \item OR 542 Operations Research: Stochastic Models
\end{itemize}        
    \item Any of the following Statistics courses:
    \begin{itemize}
        \item STAT 544 Applied Probability
        \item STAT 554 Applied Statistics I
\end{itemize}        
    \item Any of the following Systems Engineering courses:
    \begin{itemize}
        \item SYST 560 Introduction to Air Traffic Control
        \item SYST 659 Topics in Systems Engineering
        \item SYST 680 Principles of Command, Control, Communications, Computing, and Intelligence (C4I)
\end{itemize}        
    \item Any of the following Psychology courses:
    \begin{itemize}
        \item PSYC 530 Cognitive Engineering: Cognitive Science Applied to Human Factors
        \item PSYC 734 Seminar in Human Factors and Applied Cognition
\end{itemize}        
\end{itemize}
Additional Information

For additional information on the degree requirements of the MS in Software Engineering
\begin{itemize}
    \item The MS-SWE section of the Mason Catalog is the official source for the degree requirements of the program.
    \item These slides from the orientation for new MS students provide an overview of the program, as well as additional useful information.
\end{itemize}

\section{Academic Advising}

A plan of study form for the MS Software Engineering degree should be completed by the student and approved by their academic advisor before the end of their second semester in the program. This serves as a planning guide for the student and should be kept up to date by regular consultation with their academic advisor. A final signed version of the plan must be included when the student submits a graduation application.

Plan of Study forms for all the MS degrees offered by the CS department are available at this \href{https://cs.gmu.edu/resources/student-forms/}{web page}.

For more information, please see the \href{https://cs.gmu.edu/current-students/ms-students/advising/}{academic advising pages} and \href{https://cs.gmu.edu/current-students/ms-students/faqs/}{the FAQ for Masters students}.

\chapter{Obtaining Funding}
\section{NSF and other Government Funding}

\paragraph{Other Government Funding}

\section{Internal Funding}

\section{Industrial Gifts}

\chapter{Teaching}
\section{Teaching Load}
\section{Teaching Evaluation}
\section{Teaching Philosophy}
\section{Teaching Awards}

\chapter{Mentoring}
\section{Mentoring Undergraduate Students}
\section{Mentoring Graduate Students}
\section{Mentoring Postdocs}
\section{Mentoring Junior Faculty}

\chapter{Services}

\section{Service to your Professional Community}
\section{Service to the Department}
\section{Service to the University}
\section{Service to the Community}

\chapter{History and Acknowledgement}\label{sec:ack}

\paragraph{Acknowledgement} Many people have contributed to this document.
\textbf{Thank you!}

\bibliographystyle{abbrv}
\bibliography{demystify.bib}

\end{document}

### 2024-02-26 demystify.tex
Updates from Overleaf

@@
 
-The most concrete evidence of research ability is having \textbf{published papers in reputable international journals or conferences}.
+The most concrete evidence of research ability is having \textbf{papers in reputable international journals or conferences}.
 Having published papers, especially at top venues, is a sign that the applicant has successfully involved in research.
+Students  accepted to top schools often have multiple first-authored papers at top venues.
 
@@
 
-However, it is understandable that many students do not have the opportunities to publish in top places.
+However, it is understandable that many students do not have the opportunities to publish papers.
 Thus, general writings, even those under submissions or rejected, would still be good (and much better than nothing).
@@
 Unfortunately, many good research places in your countries, e.g., VinAI in Vietnam, remain relatively unknown to most universities in the US.
-So you should explicit say something about them in your statement.
+So you or your LoR writers should explicit say something about them in your statement.
 
@@
 
-\section{Letter of Recommendation (LOR)}\label{sec:lor}
+\section{Letters of Recommendation (LOR)}\label{sec:lor}
 
@@
 It is OK as long as it is a research-based personalized letter (once again, we are talking about PhD applications, not MS).
+Again, the emphasis here is \emph{research}, i.e., the letters should describe your research experiences and potentials.
+Letters focusing your course work or non-research projects at a company won't carry much weight.
 
@@
 
-\paragraph{Reminding Writers} After you submit your application, you can tell your writers that and remind them to submit their letters if they haven't done so.
+\paragraph{Reminding Writers} After you submit your application, you should tell your writers about that and remind them to submit their letters if they haven't done so.
 But don't send too many reminders as that can be annoying to the writers.
@@
 So I will not talk too much about statements.
-In short, focus on your research goal and vision, and convince us that you can achieve it through your experience, e.g., published papers, or if you work on some projects by yourself, talk about it.
-Also, talk about things that adcom members might not know about, e.g., your Github project with 1K+ stars or your regular contributions to well-known open-source projects.
-Finally, talk about what makes you different than the other 1000 applicants that we are considering (e.g., see \autoref{sec:improve-your-chance} for increasing your admission chance).
-
-\begin{commentbox}[Hung:]
-  I think asking for a “research vision” from a Ph.D.
+In short, focus on your research motivation and goal, and convince us that you can achieve it through your experience, e.g., published papers.
+Also, talk about things that adcom members might not know about, e.g., your personal Github project with 1K+ stars or your regular contributions to well-known open-source projects.
+Finally, talk about what makes you different than the other 1000 applicants (e.g., see \autoref{sec:improve-your-chance} for increasing your admission chance).
+
+% \begin{commentbox}[Hung:]
+%   I think asking for a “research vision” from a Ph.D.
 applicant is too much.
@@
 In my opinion, you should focus on showing (not telling) your research passion (if you do have).
-\end{commentbox}
+% \end{commentbox}
 
@@
 e.g., why do you apply \emph{here}?
-provide names of professors who you're interested in (if they are not already in the adcom, your application might get forwarded to them for evaluation; and they will likely be those who interview you).
+provide names of professors who you're interested in (if they are not already in the adcom, your application might get forwarded to them for evaluation; and they might be interested in interviewing and recruiting you).
 This shows that you're serious and have done homework on places you're applying to.
@@
 \begin{commentbox}[Vu:]
-  I often read the research statement (and the LORs) first. If I am
+  I often read the research statement and LORs first. If I am
   persuaded by then, I would skim over other factors and advocate for
-  admission (unless I see red flags in other parts). If I am not
+  admission (unless I see red flags in other parts). However, if I am not
   convinced, then I will likely recommend rejection (unless I see
@@
 So what can you do about this?
-several things including asking your CS dept to put itself on CSRankings (it's the easiest way to get CS people to know about the school and its CS faculty)  and explaining about your school in your statement (ask your LOR writer to do that too).
+You can explain about your school in your statement (ask your LOR writer to do that too).
 Of course, if you're interested in working with Vietnamese, consider  \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{CS PhD programs in the US that have Vietnamese professors}.
+It might also be helpful to have your CS dept to put itself on CSRankings so that others know about the dept, its faculty, and research.
 
@@
 This can be an issue for students in many top international universities where the competition is so high that very good students can still have low rankings from these schools (and be overlooked by the adcom).
-So what to do with this? as mentioned in \autoref{sec:your-school} you can put a note about this in your statement and ask your LoR writers to talk about it.
+So what to do with this? As mentioned in \autoref{sec:your-school} you can put a note about this in your statement and ask your LoR writers to talk about it.
 
@@
 \paragraph{English Test} Unless your degrees are from specific countries such as \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-GMU#standard-tests-waiver-eligible-countries}{these}, you will need to
-take standardized English test. Just do well enough to pass minimum requirement set by the university, which has many options for you to select from.
+take standardized English test. Just do well enough to pass minimum requirement set by the university, which is required for assistantships, e.g., TA~\autoref{sec:funding}.
 

### 2024-02-26 faculty.tex
Updates from Overleaf

@@
 \mainmatter
+\chapter{Intro}
+
+\section{The 5 jobs of an assistant professor}
+At an R1 university, an assistant professor in CS (and likely in other fields) typically has \emph{five} jobs:
+
+\begin{enumerate}
+\item[Research] You need to do research and publish papers.
+To establish yourself as an independent research, you should publish papers with your students (and not solely with your PhD or postdoc adviser).
+
+\item[Funding] You need to write proposal and secure funding from sponsors including government agencies (e.g., NSF) or industry.
+
+\item[Mentoring] You need to recruit, mentor, and graduate PhD students.
+
+\item[Teaching] You need to teach!
+
+\item[Service]  You need to serve for your professional research communities (e.g., PC, review journals, NSF panels) and for your institution (e.g., PhD admission committees, dissertation committee).
+You also need to do various other things including writing LoRs for your students (and even your peers)
+
+\end{enumerate}
+
+So, there you go.
+A new assistant prof.
+opening their own lab starting their academic career typically has these 5 duties.
+In some sense having your own lab is like opening your own start up company.
+You have to work on raising funding, recruiting employees, do the actual work, etc.
+You don't have a boss, but you are responsible for everything.
+
+\section{Pros and Cons}
+
+Pros:
+\begin{itemize}
+\item Freedom.
+You don’t have a boss.
+Seriously, you don't.
+You can work on any research topics you want (preferably something that can get grants), choose your own students (not that easy to find good students) and collaborators, work on your own time, and have plenty of free time(technically I only have to be "at work" for 3 hrs a week for my class from 4:15PM--7:10PM).
+
+\item Students. It's a great award when seeing how your students evolve to be more mature and become an expert in their research.
+
+\item Flexibility.
+You can do pretty much anything.
+Travels, give talks, write paper/blogs/books (like this one).
+Create new courses and even online courses.
+Collaborate  with academia, government, research lab, industry.
+Start your own company when you retire!
+
+\item Time with family. This might seem like a weird pro, but as a professor, I have much more
+control over my time than many other jobs. It lets me schedule my work so that I can
+spend more with my family. For example, when my ﬁrst kid was born, my department
+gave me teaching time off so that I could spend time with him. I spent six months with
+my son – irreplaceable time that is harder to get in many other jobs (this is especially
+true in the USA where paternity and maternity leave is minimal in many jobs).
+\end{itemize}
+
+Cons:
+\begin{itemize}
+\item Responsibility: The role carries high responsibility, not only for one's own career but also for the careers of PhD students under their supervision.
+\item Stress: The initial years are particularly stressful as one learns various aspects of the job while under pressure.
+\item Poor financial compensation: Assistant professors typically earn less than they would in industry, leading to a significant opportunity cost over time.
+\item Funding: Securing research funds requires substantial time and effort, with no guarantee of success, given the competitive nature of grant applications.
+\item Culture of unpaid labor: Academia in the USA often relies on unpaid work, both from faculty and students, which can be a significant drawback compared to industry where compensation is typically fairer.
+\end{itemize}
 
@@
 \end{document}
+
+https://vijay03.github.io/asstprofbook/chapters/job.pdf

### 2024-02-26 demystify.tex
updates

@@
 
-There are various guides on writing statement, e.g.,~\cite{blattman2022writing}, and \href{https://cs-sop.org/}{many example statements} are available.
+There are various guides on writing SOP, e.g.,~\cite{blattman2022writing}, and \href{https://cs-sop.org/}{many example statements} are available.
 So I will not talk too much about statements.
@@
 This shows that you're serious and have done homework on places you're applying to.
-Adcom will look for this part at the end of the statement, so do not skip it.
+Adcom will look for this part, so do not skip it.
 
@@
 \paragraph{English Test} Unless your degrees are from specific countries such as \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-GMU#standard-tests-waiver-eligible-countries}{these}, you will need to
-take standardized English test. Just do well enough to pass minimum requirement set by the university, which is required for assistantships, e.g., TA~\autoref{sec:funding}.
+take standardized English test. Just do well enough to pass minimum requirement set by the university, which is required for TA (\autoref{sec:funding}).
 
@@
 
-After you submitted your applications, the waiting game begins! For many students, this is an excruciating time. This section provides some information and tips to help you get through this time.
-
-% but you should try to relax and focus on your study and work.
-You can also use this time to improve your profile, e.g., by working on research projects, publishing papers, or improving your connections for better LoR writers.
-
-\paragraph{Interviews} Just like a job application, you can get interviews from some schools.
-The most common case is that a professor is interested in working with you and want to chat to make decision (e.g., to offer RA).
-The professor might include one of their current students as part of the interview.
-
-In some case, the interview is done by several professors, e.g., to see if a student fits to their group or to recruit a very strong student to their program.
+After you submitted your applications, the waiting game begins! For many students, this is a very stressful time. This section provides some information and tips to help you get through this time.
+
+\paragraph{Interviews} Just like a job application, you might get interviews from schools.
+The most common case is that a professor is interested in working with you and wants to chat to make decision (e.g., to offer RA).
+In some cases, the interview is done by several professors, e.g., to see if a student fits to their group or to recruit a very strong student to their program.
 The interview can also be done by adcom members to see if the student fit their program.
+%The professor might include one of their current students as part of the interview.
 
@@
 
-\paragraph{Not getting interviews} While it is generally good to get interview, do not feel so despair if you do not get any.
+\paragraph{Not getting interviews} While it is generally good to get interview, not getting an interview does not mean you're out.
 Many programs do not have the tradition of interviewing applicants.
@@
 
+However, not getting an interview typically means you do not get directly RA support (\autoref{sec:funding}) by an individual faculty (if they want to pay you to do research for them, then likely they want to talk to you first).
+In this case, your application will be reviewed by the adcom and they will make a decision based on your application.
+
 \paragraph{Timeline} The timeline for interviews and notification letters varies.
 Faculty setup interviews based on their (very busy) schedule.
-So do not be surprised if you get an interview invitation in the last minute (also don't get surprised if your interview is on a Uber ride!).
-And again, some schools do not do interviews at all.
-
-Some schools send out admission letters in batches, some do rolling admission, and some do not send anything out (e.g., you're on their wait list).
+Do not be surprised if you get an interview invitation in the last minute.
+Again, some schools do not do interviews at all.
+
+Some schools send out admission letters in batches, some do rolling admission, and some do not send anything out (e.g., you're on their waitlist, or they might not send rejection letters).
 However, in general, you should hear back from most schools by around March.
@@
 \paragraph{Updating your profile} In general, you should not send emails to update your profile.
-However, if you have new publications or other achievements, you can send an email asking to update your profile.
-
-\section{Preparing and Tracking Applications}
-
-\paragraph{Shared Spreadsheet} Many students use a (shared) spreadsheet, e.g., Google Sheets/Docs, to help them and their letter writers to keep track of their applications.
+However, if you have new publications or other big achievements, you can send an email asking to update your profile.
+
+% \section{Preparing and Tracking Applications}
+
+% \paragraph{Shared Spreadsheet} Many students use a (shared) spreadsheet, e.g., Google Sheets/Docs, to help them and their letter writers to keep track of their applications.
 Here are some information to put on the sheet.
-\begin{itemize}
-  \item \textbf{Your info}: Full name, email, phone, link to website/CV.  This is helpful for the writers just in case they want to quickly get some information about you.
-
-  \item \textbf{Applications details}: University, Dept, Application System URL, Submission Deadlines, Application Status (e.g., submitted,  rejected, wait-listed, accepted)
-
-  \item \textbf{LoRs}: Writer 1, Writer 2, Writer 3.
+% \begin{itemize}
+%   \item \textbf{Your info}: Full name, email, phone, link to website/CV.  This is helpful for the writers just in case they want to quickly get some information about you.
+
+%   \item \textbf{Applications details}: University, Dept, Application System URL, Submission Deadlines, Application Status (e.g., submitted,  rejected, wait-listed, accepted)
+
+%   \item \textbf{LoRs}: Writer 1, Writer 2, Writer 3.
 If this is a shared document, you might want to omit names and just use writer X.
@@
 
-\end{itemize}
-
-You can then share this sheet with your reference writers and remind them to submit LoRs on your behalf (see \autoref{sec:lor}).
+% \end{itemize}
+
+% You can then share this sheet with your reference writers and remind them to submit LoRs on your behalf (see \autoref{sec:lor}).
 Also, you should periodically update your writers with your status; a simple note such as \emph{``I am getting interviewed by prof.
@@
 
+\paragraph{What else to do?} Make decision (\autoref{sec:accept-postpone-decline})!
+accept, reject, or defer the offers (if you have multiple options).
+Ask to meet with potential advisers (e.g., through Open House or separately) and even their students.
+Ask about computer equipment and software, office space, and other resources; in many case these will be provided for free by your adviser or department (\autoref{sec:buying-equipment}).
+
 \paragraph{If you do not get admitted} If you do not get admitted to any schools or don't want to go to the ones that admit you, try again next time.
@@
 
-You can also consider applying to MS programs, which are typically easier to get in (but you need to pay).
+You can also consider applying to MS programs, which are typically easier to get in; but you need to pay.
 If you get in an MS program at a school of your choice, you can contact professors to work with them.
 If you do well, you can ask the professor to support you to convert to PhD.
-This is a common path for many students.
+This is not uncommon for many students.
 
@@
 Second, an academic year (AY)  year is typically \emph{9-month} in the US, so stipend is for 9 months (so divide by 9 for each month).
-Third, you might get paid over the summer through funding from your professor or fellowship (typicaly no TA over the summer).
+Third, you might get paid over the summer through funding from your professor or fellowship (typically no TA over the summer).
 Fourth, like most source of income in the US, you will have to \emph{pay tax} on your stipend.
-Finally, private universities might pay more for stipend, but they could have higher "activity" or some other hidden fee, or you may be required to pay fees for each credit hour.
+Finally, private universities might pay more for stipend, but they could have higher "activity" or some other hidden fee (e.g., you may need to pay some fees for each credit hour).
 
 Students often complain their stipend being too low, but it is actually not bad and you can live comfortably with it.
-In many cases, it is also enough to support your spouse and kids (many CS PhD students have their family with them).
+In most cases, it is also enough to support your spouse and kids (many CS PhD students have their family with them).
 So don't worry too much about stipend.
@@
 
-\section{Buying Computer Equipments}\label{sec:buying-equipment} Many students, including myself, get very excited about getting to buy new computer equipments and electronics when they start their PhD.
+\section{Buying Computer Equipment}\label{sec:buying-equipment} Many students, including myself, get very excited about getting to buy new computer equipment and electronics when they start their PhD.
 However, you should check with your adviser first.

### 2024-02-27 faculty.tex
updates

@@
 
-\title{Demystifying the Computer Science \\
-  PhD Admission in the US\\{\large A Guideline for International \emph{and Domestic} Students}}
+\title{A Handbook for Tenured Track Faculty in Computer Science in the US}
 
@@
 \chapter{Summary}\label{sec:summary}
-
-Below we summarize the main points of this guideline. This gives you an overview to decide which specific topics you want to explore more thoroughly.
-
-\begin{enumerate}
-  \item Should you apply?
-        \begin{itemize}
-          \item \emph{Yes, definitely}.  CS PhD study in the US is fully funded and admission into good universities is not any harder than non-US schools (\autoref{sec:should}).
-        \end{itemize}
-  \item How is your application evaluated?
-        \begin{itemize}
-          \item Applications are evaluated by the \emph{PhD Admission} committee and each application is reviewed by typically three faculty (\autoref{sec:evalapps}).
-          \item Individually faculty \emph{cannot directly admit} a student---so do not email and ask if you have a chance.
-However, faculty can \emph{advocate} for a student and therefore increase their admission chance---so do contact and introduce yourself (\autoref{sec:contact}).
-         \end{itemize}
-  \item Application Materials
-        \begin{itemize}
-          \item The committee will look at various factors, but the most important ones are research ability, e.g., publications, statement of purpose (SOP), and  letters of recommendation (LORs).
-          \item LORs are very important, but only if they are personalized and talk about your research ability (\autoref{sec:lor}).
-          \item SOP is very important. Write it in such a way that makes you \emph{stand out} (\autoref{sec:research-statement} and \autoref{sec:improve-your-chance})
-          \item GRE \emph{is not} required (\autoref{sec:grades}). Spend your time on something else!
-          \item Grades are important, but depend on the reputation of your school (\autoref{sec:grades}).
-          \item Getting an interview is typically a \emph{good sign} as no one wants to interview weak candidates (\autoref{sec:interview}).
-        \end{itemize}
-  \item What to do after getting admitted?
-        \begin{itemize}
-                \item  Celebrating! Now it is your turn to evaluate the school!
-          \item Attend \emph{Open House} to learn more about the place and \emph{interview} profs---they would be much more willing to talk to you now (\autoref{sec:accepted}).
-        \end{itemize}
-  \item Funding
-        \begin{itemize}
-          \item TA and RA are two main funding sources.
-TA (teaching assistantship) is provided by the department to help profs.
-with classes (e.g., grading).
-RA (research assistantship) is provided by profs.
-to help with their research (\autoref{sec:funding}).
-        \end{itemize}
-  \item Choosing School and Professors
-        \begin{itemize}
-          \item Many schools do not offer PhD studies in CS and many CS professors \emph{cannot} formally advise or graduate PhD students  (\autoref{sec:schoolsandprofs}).
-          \item Contacting a prof. for research opportunities is recommended, but do it \emph{properly} (\autoref{sec:contact}).
-        \end{itemize}
-  \item Miscs and FAQS
-        \begin{itemize}
-          \item Increasing your admission chance by being unique and standing out (\autoref{sec:improve-your-chance}).
-          \item You can successfully apply to CS PhD even if you have non-STEM background (\autoref{sec:non-stem}).
-          \item Compared to other countries, CS PhD in the US does not require an MS degree but has longer PhD study time (\autoref{sec:non-us-differences}).
-          \item Your 5--6 year PhD program costs about \$400K in total, but you \emph{do not} pay for it (\autoref{sec:ra-cost}).
-          \item Despite some miserable stories on social media, many PhD students have good mentors, supportive lab mates, healthy working environment ... and are happy (\autoref{sec:happy}).
-        \end{itemize}
-\end{enumerate}
-
 \mainmatter
@@
 \section{The 5 jobs of an assistant professor}
-At an R1 university, an assistant professor in CS (and likely in other fields) typically has \emph{five} jobs:
+An assistant professor in CS (and likely in other fields) At an R1 university typically has \emph{five} components:
 
 \begin{enumerate}
-\item[Research] You need to do research and publish papers.
+\item[\textbf{Research}] You need to do research and publish papers.
 To establish yourself as an independent research, you should publish papers with your students (and not solely with your PhD or postdoc adviser).
 
-\item[Funding] You need to write proposal and secure funding from sponsors including government agencies (e.g., NSF) or industry.
+\item[\textbf{Funding}] You need to write (many) proposals and secure funding. Typical sponsors include government agencies (e.g., NSF) and industry.
 
-\item[Mentoring] You need to recruit, mentor, and graduate PhD students.
+\item[\textbf{Mentoring}] You need to recruit, mentor, and graduate PhD students.  You likely need to graduate at least a 1 PhD by the time you go up for tenure.
 
-\item[Teaching] You need to teach!
+\item[\textbf{Teaching}] You need to teach! Typically, you will teach about 1--2 courses per semester.
 
-\item[Service]  You need to serve for your professional research communities (e.g., PC, review journals, NSF panels) and for your institution (e.g., PhD admission committees, dissertation committee).
-You also need to do various other things including writing LoRs for your students (and even your peers)
+  \item[\textbf{Service}]  You need to serve for your research community (e.g., PC, review journals, NSF panels) for your institution (e.g., PhD admission committees, dissertation committee).
+You also need to do other things including writing LoRs for your students (and even your peers).
 
@@
 You don’t have a boss.
-Seriously, you don't.
-You can work on any research topics you want (preferably something that can get grants), choose your own students (not that easy to find good students) and collaborators, work on your own time, and have plenty of free time(technically I only have to be "at work" for 3 hrs a week for my class from 4:15PM--7:10PM).
+Seriously!
+You can work on any research topics you want, choose your own students and collaborators, work on your own time, and freedom to travel and work from anywhere.
 
@@
 
+Chapter 3: Funding
+Let’s talk about what, for many professors, is the least favorite part of the job: making sure you
+have enough money to keep the lights on in the lab. Essentially, a professor applies for grants
+(money from the government, industry, or other charitable foundations) by writing a proposal as
+the Principal Investigator (PI). The funding agency reviews the proposal. If they decide to fund it,
+they send the money to your university. The university then takes its cut of the funding
+(essentially to pay for common services like buildings, admins, etc) and gives you the rest. You
+then spend it on your summer salary, the salaries of your students, equipment, travel,
+registration costs and other expenses associated with your research.
+There are some nuances about the different types of funding, and how you can spend different
+kinds of funding. So let us get into the details. Unfortunately, these details tend to vary a lot by
+country, so everything I’m talking about in this chapter applies only to the US.
+3.1 An example budget
+Let me walk you through an example budget. Let us say you are writing a proposal for \$100,000
+USD (100K USD). The budget will talk about how you plan to spend this money. The following
+table (Table 1) shows the example budget.
+Budget
+Graduate Research Assistant - 1
+Total Salary 40,000
+Fringe Benefits 12,000
+Total Salary and Fringe Benefits 52,000
+Total Materials \& Supplies 0
+International travel to a conference 2500
+Domestic travel to a conference 1481
+Total Travel Costs 3,981
+Tuition for Fall or Spring 4,746
+Tuition for Summer 1,778
+GRA Tuition for 1 year 11,270
+Total Direct Costs 67,251
+Modified Total Direct Cost (MTDC) 55,981
+Indirect Costs 58.5% of MTDC 32,749
+Total Cost (Direct + indirect costs) 100,000
+37
+I will explain the various parts of this example budget. I will be talking about terms such as
+direct costs and indirect costs as defined by the University of Texas – I believe it is similar at
+most other US universities, but there might be differences. Please check with your university to
+make sure.
+Fringe. Let us say you want to support a student on this grant. In addition to paying for the
+student’s salary, you also have to pay for non-wage expenses: things like health insurance, social
+security, and medicare. This is usually expressed as a percentage of the student’s salary. At the
+University of Texas at Austin, this was 30% in 2022 (For a comparison point, it was 32% at
+Princeton). The stipend of a graduate research assistant (GRA) at UT is about \$30K USD for the
+fall and spring semesters ($40K USD including summer). So salary and fringe for a GRA for a
+year is \$52K.
+You also have to pay the tuition of the GRA, and at UT, this is about \$11.2K USD for a year. To
+keep things simple, I am using a single number for the tuition, but this depends on factors such
+as whether you are in-state or out-of-state, your progress in the PhD program (if you are
+all-but-dissertation, your tuition is lower), and whether you are a US citizen or not. It also
+depends on the fraction of time spent as GRA/TA: if this drops below 20 hours a week
+(considered full-time for students), tuition might have to be charged at the out-of-state rate. If
+the student is only doing 5 hours a week as GRA, it might also impact how much of their tuition
+can be covered by the grant; at 5 hours a week, the student might not be eligible for insurance. I
+highly recommend talking to your admin if you are accounting for less than 20 hours a week for
+your student on the project. Some schools also calculate tuition as a % of the stipend.
+So the total cost for supporting a GRA for one year is \$63.2K USD. While you are paying \$63.2K
+for each student for a year, the student's take-home pay is only the stipend part (minus taxes
+that are withheld): \$35.2K assuming a 12% tax rate.
+Note that fringe also applies to you (the PI) when you pay yourself out of your grants for your
+summer months.
+Direct costs. The direct cost part of your budget includes salaries, fringe benefits, travel, and
+equipment and supplies that contribute directly towards performing the proposed research.
+Direct costs are exclusive to each proposal: a piece of equipment can only be a direct cost on
+one proposal, and a GRA can be assigned only to one proposal at a given time. In our example,
+let us assume the only direct costs are supporting the student and paying for the student’s
+travel (including registration, transportation, lodging, and food) for one international and one
+domestic conference.
+Modified Total Direct Costs (MTDC). MTDC is a portion of the total direct costs used for
+calculating the indirect costs. It is the direct costs minus tuition, fellowships, equipment, and
+capital expenditures. Travel also counts as MTDC. In this budget, we are not spending anything
+38
+on materials and supplies in this budget, but if we did, it would be part of MTDC. Check with
+your university to determine exactly what is counted as MTDC – this can be surprisingly tricky.
+For example, you might expect computers to not count towards MTDC as it is “equipment”, but it
+may be counted as a “supply” if the university is unable to amortize its cost. In the example
+above, total direct costs is \$67.2K USD (63.2K for one GRA plus 4K for travel), but MTDC is only
+$56K (leaving out the \$11.2K for tuition).
+Indirect costs. The university takes a portion of all the money you receive, to pay for centralized
+functions such as building costs, utilities, admin, and libraries. This is termed as the “indirect”
+cost, usually expressed as a percentage of the MTDC. At most universities in the US, indirect
+costs are more than 50% of the MTDC. At UT, it is 58.5% (Mike Freedman reports it is 63% at
+Princeton). Thus, in our budget, we would have .585 * 56K = \$32.7K added as indirect costs. As
+a result, it would cost 67.2 + 32.7 = \$100K USD for supporting a single GRA and some travel on
+this grant. Note that even though the 58.5% number is scary, you are only paying 33% of your
+budget for indirect costs; in other words, you don’t lose 58.5% of the total budget to indirect
+costs.
+As you can see from the preceding table, the \$100K USD budget only covers one student for one
+year. At some universities, a GRA costs more, requiring more funding.
+This should give you a picture of how expensive it is to maintain large groups! If you had ten
+students, you would need to pull in a million dollars each year to fund your students.
+It is useful to think of your budget in terms of student-years. \$400K is simply two students for
+two years. A dissertation, roughly five student years, is \$500K. If you get the NSF CAREER grant,
+you would typically be able to support one dissertation. This helps determine the scope of your
+proposal: if you need five students to work on it for two years, a grant that provides \$250K will
+not be a good match.
+Summer months in your budget. PIs at most R1 universities in the US are expected to fund
+themselves via grants in the summer. When you are drawing up your budget, you are expected
+to estimate how much of your summer months you will dedicate to this project. This is taken as
+an indication of PI commitment at the funding agency – you don’t want to write a budget where
+your estimated time involvement is zero. So if you have one NSF grant where you estimate your
+time commitment to be one month, you can only put in one more month for another NSF grant
+proposal (since NSF limits total PI compensation to be two months per year across all NSF
+grants).
+3.2 Types of grants
+Broadly speaking, there are four different kinds of grants or funding, each with different
+restrictions on how you can use the money. The differences come down to what you can spend
+the money on, how long you can keep the money, and what the overhead is.
+39
+We can term the four kinds of funding as grants, contracts, university funding, and gift funding.
+Let me go over each type of funding.
+Grants. The National Science Foundation (NSF), the Defense Advanced Research Projects
+Agency (DARPA), the Army, the Navy, and the Air Force all fund research related to computer
+science. These grants are competitive – NSF has a 25% acceptance rate (though this varies by
+the directorate), and I imagine the other programs are similar.
+This money can typically only be used for funding students, yourself, travel, and equipment
+purchases that you had already put in the budget. For example, if you suddenly want to
+purchase equipment not in the original budget, it needs to be approved by the Office of
+Sponsored Research at your university (which makes sure you are spending the money in the
+right way). You cannot use government funding to pay for things like a meal with your students
+during the semester (meals during travel is allowed).
+Government funding has strict timelines by which the money has to be spent. Typically, the
+money is released on a yearly basis. You have to show that you have used the previous
+installment to get the next round of funding. This sometimes results in professors making last
+minute purchases to exhaust their grant!
+In some cases, you can ask your funding program manager for a no-cost extension to your
+grant. This typically means that you don’t need extra money, but you want extra time to use the
+money. This is usually considered on a case-by-case basis, and I would not recommend
+counting on getting a no-cost extension.
+Indirect costs are levied on government funding (more generally, on all funding that goes
+through the Office of Sponsored Research). The example budget I showed in the previous
+section assumes funding is via sponsored research.
+Grant proposals typically have reporting requirements (for example, you have to submit a report
+once a year for an NSF grant). However, grants typically do not expect any concrete deliverables
+to be provided at the end of the funding period. Essentially, the grant allows you to carry out the
+proposed research: the act of publishing itself is the desired outcome. As a result, grants have a
+strong alignment between the PIs and the funding agency: they both just want to get research
+done.
+Contracts. The second kind of funding is via contracts. Contract funding comes with an explicit
+contract you must fulfill to keep and obtain more funding. For example, there may be quarterly
+or monthly reports, or on-site visits by the sponsor. Contract funding typically also has
+deliverables that are expected at regular intervals during the funding period. One example of
+contract funding is DARPA funding, which comes with a number of deliverables and reporting
+requirements. Contract funding is less aligned with the PIs than grant funding: apart from doing
+the core research, the PI’s students will have to spend time on creating the deliverables, and the
+PI herself has to spend a significant amount of time on reports and meetings.
+40
+NSF vs DARPA. NSF is a good example of grant funding, while DARPA is a good example of
+contract funding. Both these agencies fund a lot of computer science research. DARPA requires
+deliverables and reports to be produced regularly during the grant duration. NSF only requires a
+report to be submitted for each year of the grant duration. You can apply for NSF funds as
+someone on a visa, while DARPA and other similar government funding programs might require
+you to be a US citizen or a permanent resident.
+University funding. Your university, or your college, or your department may choose to give you
+funding. Typically when you start, you get a “startup” package that allows you to kickstart your
+research group. University funding is “internal” funding in a way, and as such has fewer
+restrictions than government funding. But this really depends on the specific university. I know
+universities where startup funding cannot be used for covering summer salary of PIs. Apart
+from startup funding, your university might have internal competitions to award “seed” funding
+(typically under 50K) that you can then use to get preliminary results and later apply for
+traditional grants.
+University funding is also associated with timelines. For example, startup funding usually has to
+be spent before you get tenure. However, university funding deadlines might be a little more
+flexible than grant or contract funding deadlines.
+University funding typically does not have associated indirect costs (as it is already university
+money). You still have to pay fringe support for your students and your own summer salary.
+An interesting sub-category of university funding are Research Centers. A research center is
+funded by sponsors to carry out a certain mission over a span of around 5 years. The research
+center has a program director who is responsible for allocating the funding. PIs at the university
+apply for funding from the research center (a sort of internal grant), and are allocated funding.
+Obtaining funding in this manner is easier than competing for an external grant. Particularly for
+assistant professors who are just starting, research centers can provide seed funding before the
+first grants are obtained. Check out the ADA center at Michigan for a good example.
+Gift funding or unrestricted research funding. Companies or charitable foundations may
+choose to fund your research in the form of “gifts” or unrestricted research grants. This is the
+best kind of funding, for three reasons: indirect costs are typically not levied on these gifts, there
+is no associated timeline by when you should spend them, and you can use them for a broader
+range of things than university and grant/contract funding.
+Indirect costs are not levied because the gift-giver can specify that it cannot be used for indirect
+costs (the university has to honor this). Since it is an unrestricted gift, there are no limits on
+what you can spend them on, or by when you should spend them. This doesn’t mean you can
+spend the money on personal expenses. All expenses still have to be approved by the university,
+and should be used for research-related expenses. But things like taking your research group
+out for a meal can be expensed during gift funding (but not government funding).
+41
+Indirect costs not being levied can make a huge difference. For example, in our budget from
+earlier, if you were using gift funding to fund the student, you don’t have to pay indirect costs,
+and thus get over \$30K back. Considering a student costs \$63K, you get to sponsor the student
+for 1.5 years rather than 1 year!
+How should you use your funding? Let us say that you are lucky enough to have all four kinds of
+funding. And you have an expense that could be assigned to any of these funding (note that this
+is often not the case: you can’t use a grant on topic X to support a student working on Y if X and
+Y are not closely related). What should you use?
+The typical rule is “Use the most restricted funding that will expire the earliest”. Thus, you should
+always use the appropriate grant/contract funding first, then your university funding, and finally
+gift funding. Gift funding is the most flexible, so use it only when you have no other option.
+Always spend your gift funding on expenses that will have overhead (if funded from grants),
+such as students. Do not spend it on things like equipment (be careful to ask what counts as
+equipment at your university) which do not have overhead even when expensed to grants.
+Writing a grant with other PIs. While you could write smaller grants as the sole PI, typically for
+larger grants, you need to collaborate with other PIs and submit as a team. For example, you
+cannot submit NSF Large grants as a solo PI. One thing to remember though: all the funding is
+shared between the PIs. For example, imagine a grant for one million USD. It sounds amazing!
+However, if the grant was awarded to a team of two PIs, and the grant duration is five years,
+each PI can only support one student on it for the grant duration (100K * 2 * 5). Similarly, if a
+grant of 10 million USD was won by a team of five PIs, each PI can only support one student for
+four years (100K * 5 * 4). So in the end, the share of each PI is similar or lower than that of an
+NSF CAREER.
+This is not to say you should not collaborate with other PIs to write grants. Collaboration often
+brings new ideas and research directions that are unlikely or impossible when working alone.
+And bigger grants often provide prestige: for example, landing an NSF Large grant is significant
+and puts you on the map. But it is useful to calculate how much you expect to receive if a grant
+proposal is funded.
+3.3 The grant application process
+So how does one go about submitting a grant? There are several steps in the process:
+1. The funding agency first puts out a call for proposals. For example, check out the NSF
+call for CRII grants, a startup grant meant for new PIs who have not yet received a grant.
+Note that PIs at many R1 universities are not eligible for CRII. This was posted on May 18,
+2022. Each call has a solicitation page with details about the program and how the
+proposal should be structured – you should read this carefully.
+42
+2. In some cases, there are webinars where the program managers explain the grant
+opportunity. CRII had such a webinar on Jul 13 2022. I highly recommend attending
+these webinars if it is your first time submitting for a particular grant opportunity.
+3. In some cases, the funding agency might want you to submit a preliminary proposal.
+They will review the preliminary proposals and inform you if they would like you to go
+ahead with the full proposal. This is not the case for CRII.
+4. There will be a due date by which you need to submit the proposal. For the CRII, this is
+Sep 19 2022. The proposal will have a number of parts, and the solicitation will clearly lay
+out what the proposal should look like.
+5. You write the proposal, and sometime before you submit, you send it to the Office of
+Sponsored Research (OSP or OSR) at your university. They have to approve the proposal
+before you can submit, and they may request to see your proposal a week or 10 days
+before you submit. Please make sure you get a draft to the OSP on time!
+6. Finally, you submit the proposal, and you wait. To be exact, your admin/OSP submits for
+you, so you want to let them know when you are done. You and your admin/OSP will be
+submitting a lot of proposals together; make sure you get to know the folks doing this. I
+only have experience with NSF proposals, but these usually take 6-8 months to be
+reviewed. In some cases, it might take even longer.
+7. If you are lucky enough to get your proposal funded, you might get an email or a call from
+the NSF program manager. They might ask you to slightly modify your budget. For
+example, I was asked to reduce my NSF CAREER budget (and appropriately reduce the
+scope of my tasks as well). You will be asked to provide an abstract that will be displayed
+on the NSF website. For example, check out the abstract for my CAREER award.
+8. Finally, you will get an official email that your grant has been awarded, it will become
+public on the NSF website, and you will get the first installment delivered to your
+university. You can then begin spending it in accordance with what you proposed.
+9. In the unlucky case that your proposal is not funded, you will get back three or more
+reviews from NSF. These reviews tend to be short, and nowhere as detailed as paper
+reviews. As such, it can be hard to figure out what to do after getting a rejection. My
+advice is to go talk to senior faculty in your department if this happens, and they should
+be able to guide you in the right direction.
+3.3 Writing your first grant proposal
+Writing your first grant can be a stressful experience, since it is quite different from writing a
+research paper. There are a number of differences.
+First, the audience for a grant proposal is broader than that of a research paper; it is guaranteed
+that non-experts in your area will be reading and evaluating your proposal (for example, if I write
+43
+a grant proposal about storage research, I can expect it to be read by someone working in
+systems broadly, but not necessarily someone doing storage research). As a result, it has to be
+much broader and more accessible than a research paper. Proposal reviewers might also have a
+smaller amount of time to spend on each proposal – some funding agencies ask reviewers to
+read a dozen proposals in around a month when you serve on their panel. You can imagine how
+much time each proposal would get. Thus, your proposal must be easy to skim and read, must
+get to the major points quickly, and must be accessible to a broad audience.
+Second, in a research paper, you are reporting on something that is already done. The focus is
+on the technical details: how you were able to achieve the result, what are the drawbacks, etc.
+However, in a grant proposal, you are talking about things that you will do in the future. In a grant
+proposal, the most important thing is to be able to inspire the reviewer, and convince them that
+this is a thing worth doing (and funding). In this respect, it is much more similar to a founder
+pitching before an investor, than a professor talking to her peers about her research. While the
+details do matter to some extent, what is much more important is that you paint a big picture of
+how your proposed research will make the world a better place. This is quite different from
+writing a research paper, and takes some practice.
+For writing your first grant proposal, I’d recommend two things. First, talk to folks who have
+gotten the grant you are submitting for, and ask if you could see their grant proposal. Many folks
+are happy to share privately. Some folks, like Jeff Bigham and Gillian Hayes, are publicly sharing
+some of their proposals. Second, write up your draft proposal, and show it to senior folks in your
+department. They would be able to catch errors that could cause rejection.
+When submitting grant proposals, do not be disheartened if it gets rejected. As far as I can tell,
+most proposals, even from experienced faculty, get rejected a few times before getting
+accepted. Grant rejections are a bit more painful because the waiting time is high, you might
+only get to submit to a grant opportunity once per year. But with time and repeated submissions,
+many grant proposals do get funded.
+
+3.4 Serving on grant panels
+44
+3.5 Getting funding from industry
+Since I started at UT, I have been fortunate to receive funding from a number of companies:
+Meta, Google, VMware, and Toyota. How did I get this funding? Unlike with government
+agencies, there is no set process to follow in general (exception being Google with the Google
+Research award).
+Giving talks and networking. The process is more organic: I give talks at various companies
+about my research, and talk to their engineers about how my group’s research can impact them.
+This is useful to gain an understanding of how our work will play out in the real world, and to
+understand the problems they are facing.
+Sometimes after talks like these, the company reaches out to see if they can sponsor our
+research. At other times, the company is looking for academic groups to invest in, and my talk
+puts our group on their radar. Note that you might sometimes receive funding even without
+doing this – good work is always noticed and rewarded (especially if close to the company’s
+interests).
+In rare cases, if you know the folks at the company really well, you can also reach out and say I
+am trying to achieve research goal X, is your company looking to fund anybody in this space.
+But you only do this after building up strong ties with the company over a long period of time.
+You can’t start out by saying “can you give me some money”.
+Networking and building connections always pay out over the long term. So you shouldn’t have
+the mindset of “I gave a talk at company X, I should get some funding from them soon”. Rather,
+you should aim to get your group’s research out there and known by companies in that space.
+The funding will follow later, but time spent networking and giving talks is never wasted.
+3.6 Financial planning for your research group
+Finally, I want to talk a little bit about financial planning for your group. As a professor, you are
+now a manager, and you should ensure that the students working with you are adequately
+funded. If you are unable to fund your graduate students, your department might be able to fund
+them as TAs, though the dept would need some advance notice. For postdocs, this option is not
+there, so you need to make sure you have funding for your postdocs.
+Runway. In general, it is good practice to think six months to 1 year ahead, and make sure you
+have enough funding for existing students/postdocs and folks you are planning to hire. Initially,
+you will have your startup funding, so you start with a full tank of gas. For every
+student/postdoc you hire, you should check how much runway your group has: how long until
+your group runs out of funding. You should ensure you have a runway of 6 months to 1 year.
+45
+This means that if you have 1 year of runway, you need to write grant proposals now (since it
+usually takes about a year to be reviewed and funded). At the start of every semester, review
+your finances with your admin and make sure you have enough of a runway.
+Savings. An important part of financial planning is to always have a little bit extra in your coffers
+than your anticipated expenses. For example, imagine that you got a paper accepted, and there
+is no associated grant you can expense it to. If you don’t have some money left over, you would
+not be able to sponsor your student attending the conference and presenting the paper.
+Summers. Part of this financial planning is also deciding what to do in your summers. I strongly
+advocate ensuring that you can both pay yourself for the summer, and pay your students. This
+might require being conservative in taking on students until you are sure you can pay them.
+There is the temptation to cut back on your own summer salary so that you can pay your
+students; I would recommend avoiding this. While it can be a temporary fix, it is necessary for
+your long-term health, and that of your group, if you maintain your own summer payment.
+If you are able to work at a company for the summer, this can reduce the strain on your group’s
+finances while also helping your research become more grounded. I spent one summer at
+VMware after joining UT Austin, and that time spent was useful for my research. You would
+probably need to reach out much earlier to set this up – the group hosting you would need to
+make sure they have the budget to host you for the summer.
+Coordinating with admins. You want to make sure you and your admin are on the same page
+regarding what funding to use for each expense (sometimes it can be hard to move expenses
+between accounts). Most admins are experienced and get the picture quickly once you describe
+your overall strategy.
+Acknowledgements
+Thanks to Gabriel Parmer, Indranil Gupta, Jeff Bigham, Michael Ekstrand, Michael Freedman,
+Satish Puri, Siddharth Joshi, Solel Pirelli, and Stefan Savage for pointing out errors and providing
+suggestions for the chapter.
+Summary
+In this chapter, we talked about how grants work, what the application process looks like, and
+different sources of funding for research. We went over an example budget that shows all the
+different overheads involved in funding research.
+We have now discussed two of the four main big roles of a professor: managing students and
+managing funding. Next, we will discuss teaching
+
 \chapter{Teaching}
@@
 \chapter{Services}
+These are the services that you are expected to do as a faculty member.
+They are not as important as research, but they are still important.
+They are also a good way to get to know people in your department and in your field.
+
+Moreover, while some services are required (e.g., when your department chair assigns you to some committee), some are not and you can decline (e.g., an invitation to be a PC member).
+You should be strategic about which services you accept.
 
 \section{Service to your Professional Community}
+\section{Service to the Community}
 \section{Service to the Department}
 \section{Service to the University}
-\section{Service to the Community}
+\section{Serving on NSF Panels}
 
-\section{Serving on NSF Panels}
 You should serve on NSF panels.  It is a great way to learn about the grant process: how to distinguish between good and bad proposals and how people review propsoals.
-It is also a good way learn what other people are doing, what is hot, and what is not.
+It is also a good way learn what other people are doing, what is hot, and what is not. In short, it is a good way to learn how to write successful proposals.
+
+\paragraph{How to get on an NSF panel} You can cold email progarm managers in your area and ask to be on a panel.
+You can also ask your colleagues to make recommend for you.
+Same way you do for getting invited to be on a PC.
+
+\paragraph{How much work is it?} It is a lot of work.
+You will be asked to review a dozen proposals in a month or so.
+It will be similar to a program committee meeting, where all the proposals will be discussed.
+You will be expected to share your view of the paper, and argue for why it should be funded or not.
+
+% Once you become a PI, you will get invited to serve on grant review committees. You can directly
+% email program managers and say you are interested in reviewing, and they will see if they can
+% recruit you for the next reviewing cycle. I highly recommend doing this if you can: it is really
+% eye-opening to see how the sausage gets made! Serving on a grant review committee is
+% educational: you will immediately pick up what you need to do to make your proposals better.
+% If you do volunteer for grant proposal reviewing, expect this to take up a lot of time. You will be
+% asked to review a dozen proposals (or more) in a month or so. It will be similar to a program
+% committee meeting, where all the proposals will be discussed. You will be expected to share
+% your view of the paper, and argue for why it should be funded or not.
+
+One part of the professor job that doesn’t get a lot of attention is service. However,
+understanding and managing this component is important, both for your own tenure case and
+for the wider community. Let us start by talking about what service is, and why it is important.
+7.1 Defining and managing service
+So what do I mean when I say service? I would loosely define service as optional tasks that your
+community requests from you. You are free to decline these tasks without a direct impact on
+your career (there might be indirect impacts). These tasks are important to keep the community
+running.
+For example, your peer researchers may ask you to serve on the program committee for a
+conference, or review a paper for a journal. You will get a lot of these requests, and you are not
+expected to say yes to all of them. Your department will expect you to do some service each
+year, and there is an expectation that you get invited to serve on the PCs of the top conferences
+in your research area. But beyond that, you don’t get any benefits by doing a lot of service.
+Why should you do service? To put it in a nutshell, doing service is part of being a good citizen,
+both inside your department, and inside your wider academic community. This is because of the
+way academia operates, fully dependent on freely volunteered labor.
+For example, imagine that you want to run a conference. You need someone to organize the
+conference, and someone to review submissions and select papers for the program. One way
+to do this is to recruit reviewers, pay them for their time, and pay other folks to handle the
+logistics of running the conference. Speakers would get paid. All the money would come from a
+combination of sponsors and the fees that attendees provide. This is how industrial
+conferences operate.
+Academia uses a different model. People volunteer their time to act as reviewers or conference
+organizers. This allows the conference organizers to reduce the cost that needs to be paid by
+attendees (even so, the cost is usually several hundreds of US dollars). USENIX uses a slightly
+different model where the conference organizers are paid professionals, but the technical
+reviewing is still done by volunteers. This results in a slightly higher registration fee for
+attendees (compared to when everything is done by volunteers).
+Thus, almost everything in academia runs on volunteer labor. All the events you attend, the
+conferences and journals you publish in, the panels you enjoy, the awards that are given out, all
+depend on people volunteering their time. Since you get a net benefit from the time volunteered
+by others, it is only fair that you do your share of volunteering as well.
+72
+There are some other fringe benefits to volunteering. One is that you get your name out. When
+you are an assistant professor and you are struggling to establish yourself, it can be useful to
+have folks know you and recognize your name. The second benefit is that you get to meet folks
+in the course of doing service and form your network; this is particularly useful when you are
+just starting out.
+Why should you not do service? While it is important to do service, I’ve seen folks go overboard
+on this (and did too much myself in my first year). Doing service feels good: you are doing
+something for your community, and it feels like you are getting your name out. And it is hard to
+say no when folks you respect request you to join a committee.
+But you should always, always remember: you will not get tenure for service. No matter how
+good the service is. The first consideration at R1 universities is always research. If your research
+is not good, having done amazing service will not get you tenure.
+Apart from not helping you with tenure, taking on too many service commitments will make you
+bad at service. If you say yes to too many concurrent PCs, you will have a tough time doing a
+good job at reviewing papers for all of them. Nobody wins in this situation: the conferences and
+authors get crappy reviews, and you get a reputation for writing bad reviews.
+Managing service commitments. As an assistant professor, you should do enough service to be
+a good citizen, but not so much that it affects your other job responsibilities. You should
+conserve your limited energy for research and teaching. Be thoughtful and intentional about the
+service requests you accept – always think about how much time you are committing down the
+line (even if it is several months or a year away). The way to think about it is that you can do two
+to four big service requests each year. A big service request is something like an organizational
+role at a conference, or a program committee assignment where you will be reviewing 10+
+papers. You should really be doing only a few of these each year. Smaller service requests, such
+as being an external reviewer for a conference, you can do more of; but be careful that they don’t
+pile up.
+Saying no. How do you say no politely, especially when it is someone you respect and look up
+to? You should remember that folks who run conferences expect a lot of people to say no. It
+happens all the time, and if you decline, they will not take it personally. You don’t need to worry,
+“If I say no this year, maybe they will not invite me ever again”. I can relate because I had this
+exact worry at the beginning. I’m happy to report that I have declined to review at conferences,
+and still been invited the following year. As long as you do good research and write good
+reviews, you will be invited back. So do not let this consideration make you say yes when you
+don’t have the time to do a good job.
+I usually say something like: “Thank you for inviting me. Unfortunately, I am overloaded on
+service commitments this year, and I don’t think I will have enough time to do a good job.
+Therefore I will have to decline”. As you get a bit more senior, you could recommend a junior
+faculty member they could invite in your place.
+73
+7.2 Serving on program committees and editorial boards
+When you get started as an assistant professor, serving on program committees will form the
+bulk of your service. At first, it can be super flattering to be invited to a program committee: I
+definitely remember my first such invitation. You now have a voice at the table in determining
+the program! You can review papers alongside researchers you looked up to. As a result, you
+pretty much say yes to all the invitations you initially get.
+This is not a great strategy, though: you end up over-committing yourself in your initial years. I
+would recommend saying yes only in the following cases:
+1) The conference is your home venue, where you have been publishing in grad school, and
+where you hope to publish in the future.
+2) It is the top (or one of the top) conferences in your area.
+3) It is a community that you want to join, or an area that you want to publish in.
+4) It is a workshop specializing in your area.
+For example, I publish regularly at systems conferences such as ATC/FAST, so I tend to review
+papers for them. SOSP and OSDI are the top conferences in my area, so I almost always accept
+PC invitations for them. HotOS and HotStorage are the relevant workshops for my area, so I
+review papers for them as well.
+Don’t accept invitations to be on the PC for conferences that you have never published in, and
+are unlikely to do so in the future. Don’t accept invitations just because you know the PC chair if
+you are already committed to other PCs.
+Overall, a good number to aim for is about 1-2 big PCs each year (10+ papers per PC) and 1-2
+smaller ones (5+ papers per PC). Between research and teaching, it will be hard to do a good job
+with more PCs.
+University expectations regarding PCs. Generally, your department or university likes to see that
+you are being invited to serve on the PCs of the top conferences in your area. If you become an
+assistant professor and receive no PC invitations, that can be a bit of a red flag. However,
+beyond expecting to see some service (a few PCs each year), the department/university doesn’t
+really care how much service you do – they assume you will manage your service workload
+efficiently.
+How much time does reviewing a paper take? This really varies by area. I would say anywhere
+between one to four hours per paper is common. Do not spend more than half a day on a single
+paper. You should read actively, trying to get the main ideas and insights, which will enable you
+to write a good review. You should aim to read the paper only once (this gets easier with
+practice), taking notes as you go. Having questions you want to answer as you read will help
+make reviewing easier.
+74
+If you find yourself spending more than 1 day a week on average (or 8 hours per week) for
+reviewing papers, you have accepted too many PC invitations. It can be hard to properly track
+this though, since you have a lot of activity near the reviewing deadlines, and not much on other
+weeks.
+I’ve mostly talked about program committees since conferences are considered the primary
+publication venue in computer science, but much of the advice applies to journals as well. Only
+review for journals where you publish regularly, and track and manage how many submissions
+you accept to review.
+7.3 Serving on dissertation committees
+Another form of service is participating on dissertation committees, where you read the
+dissertation and attend the thesis proposal/defense. This can take up a lot more time, since you
+have to read an entire dissertation. It helps if you already know the students work, since that
+will form a significant portion of the dissertation. Still, serving on a dissertation committee can
+be an entire day’s worth of work.
+Requests for serving on dissertation committees can come from within the department or
+externally. The internal ones are harder to say no, since these are the colleagues you work with
+on a daily basis. Nevertheless, track how much time you are spending on this. Don’t say yes to
+more than one or two dissertation committees per semester/quarter.
+From the university viewpoint, I don’t think there are concrete expectations regarding serving on
+dissertation committees. Not being on dissertation committees early on in your career will not
+be seen as a red flag (to the same extent that not being on PCs would be).
+7.4 Serving as a session chair
+Another form of service is serving as the Session Chair at a conference. The session chair will
+introduce the session (typically 1 to 1.5 hours) at the front, introduce the speakers, and
+moderate the questions. The session chair also kicks off the QnA if there are no questions from
+the audience at first.
+Being a session chair is a fantastic service opportunity, and you should definitely sign up
+whenever you get the chance. It gets you face-time in front of the community, you get your name
+out, and the work required for being a good session chair is modest. Grab session chair
+opportunities with both hands when you are starting out!
+7.5 Organizational service roles
+As you become a bit more experienced, you will be offered organizational roles such as being
+the Publicity Chair, the Web Chair, the Poster Chair, or the Program Committee chair. Of these,
+being the chair of the program committee is most prestigious, since you are responsible for
+75
+delivering a high quality program in coordination with your chosen program committee. So if you
+are invited to be the PC chair of a conference or workshop you normally publish in, you should
+accept the invitation. Your university or department will care about whether you have been
+invited to be the PC chair (once you become a bit more senior).
+However, the other roles involve a lot of work and little recognition. Ideally, these would be done
+by professionals (as in the case of USENIX). However, for non-USENIX conferences, someone
+has to step up and do them. Be aware of the work-reward ratio if you choose to accept an
+assignment like this though.
+During my time as an assistant professor, I co-chaired HotStorage 2020 and SyStor 2021. I was
+also one of the General Chairs for SOSP 21. I enjoyed doing all three roles, but it was definitely a
+lot of work! I would not recommend doing any of these roles (especially being General Chair)
+early on as an assistant professor; it is a better fit when you are close to going up for tenure.
+7.6 Departmental and university service
+Apart from serving on program committees, editorial boards, and organizing conferences and
+workshops, you are also expected to do “local” service for your department and university. This
+can take multiple forms.
+At the department level, this means you will serve on one or more committees. Assistant
+professors are usually assigned to either the PhD admissions committee or the faculty
+recruiting committee. Both committees are extremely important and directly influence the future
+of the department. Other committees include Faculty Evaluation Committees, Budget
+Committees, and Promotion Committees that you can participate on once you become more
+senior. Expect to spend at least a couple of weeks (put together, it will be spread over the
+semester) on departmental committee work each year.
+At the university level, there will be committees deciding overall university policy. Typically, your
+department or college will need to be represented on these committees. As an assistant
+professor, it will be rare for you to be nominated by your department for these committees;
+typically, more senior professors participate in these committees.
+Summary
+The most important thing to remember about service is that it is like chores around the house;
+you need to regularly clean your house, water your plants, etc. You do this because you want a
+nice place to live in; similarly, service provides you a nice community to do research in. However,
+carefully manage the amount of service you do. You will be bombarded with requests, and it is
+all too easy to say yes to too many requests. Don’t accept so many service requests that it
+impacts your other responsibilities as a professor.
+76
 

### 2024-02-27 demystify.tex
updates

@@
 
-So what can you do about this?
+What can you do about this?
 You can explain about your school in your statement (ask your LOR writer to do that too).
 Of course, if you're interested in working with Vietnamese, consider  \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{CS PhD programs in the US that have Vietnamese professors}.
-It might also be helpful to have your CS dept to put itself on CSRankings so that others know about the dept, its faculty, and research.
+%It might also be helpful to have your CS dept to put itself on CSRankings so that others know about the dept, its faculty, and research.
 
@@
 
-You can learn about CS programs and research expertise of faculty using resources such as \href{https://csrankings.org}{CSRankings.org}, which is designed specifically to help prospective PhD students in Computer Science!
+You can learn about CS programs and research expertise of faculty using resources such as \href{https://csrankings.org}{CSRankings.org}.
 You will be very surprised to learn that a school that you didn't know much about can have very strong research in your interested topic (and vice versa, a school you thought highly about might have no faculty working in the research field you're interested in).
@@
 Y at Z) on static program analysis and I think it could be used to tackle the spurious issues mentioned in your paper.
-I have a small paper at conference/workshop C and a project on symbolic execution at Github G.
+I have a short paper at conference/workshop C and a project on symbolic execution (Github repo G).
 
@@
 In fact, if a faculty is not in CS, it is unlikely that they can be the \emph{main} adviser of a CS PhD student.
-They may \emph{co-advise} or be in the PhD dissertation committee, but your main adviser will need to be a tenured or tenure-track faculty in CS.
+They may \emph{co-advise} or be in the PhD dissertation committee, but your main adviser will likely be a tenured or tenure-track faculty in CS.
 If in doubt, check with the CS department for their requirements.
+
 For this specific reason,  CSRankings includes only tenure line faculty who can advise CS PhD students.
@@
 
-\section{Differences between PhD in the US and Other Countries}\label{sec:non-us-differences}
+\section{PhDs in the US vs. Other Countries}\label{sec:non-us-differences}
 
@@
 Students, especially international ones, often get confused on what to do after they get an offer from a professor, e.g., how to postpone or accept/reject offers, and most importantly, how to do so without offending anyone.
-In short, if you are not serious about an offer for any reason (e.g., still waiting for other offers), do not accept, i.e., commit to it.
-If you are not sure, ask for more time to decide.
-But do not accept an offer and then later reject it.
-This is considered \emph{rude and unprofessional}, and will likely burn bridges with that professor.
+In short, if you are not serious about an offer for any reason (e.g., still waiting for other offers), do not accept it.
+If you are not sure, ask for more time to decide.  But do not accept an offer and then later reject it.  This is considered \emph{rude and unprofessional}, and will likely burn bridges with that prof.
 
@@
 Could you please send me more details about the offer and what to do next?''}.
-Here you are not only accepting the offer but also asking for more details.
 The prof.
@@
 
-If you have better offers, you might mention them to the professor and ask if they could match them.
+% If you have better offers, you might mention them to the professor and ask if they could match them.
 It is unlikely they can do so as RA might be fixed by the department and not negotiable.
@@
 I started writing this document in May 2023 and have been updating it since then (mostly around deadline time when I tend to procrastinate, i.e., \emph{productive procrastination}!).
-I have put the source code of this document on \href{https://github.com/nguyenthanhvuh/phd-cs-us}{GitHub} so that anyone can contribute to it.
+I have put the source code of this document on \href{https://github.com/nguyenthanhvuh/phd-cs-us}{GitHub} so that others can contribute to it.
 

### 2024-02-27 demystify.tex
updates

@@
 Ultimately, choose a prof.
-that fits you the most by communicating with them, taking their courses, meeting and asking them questions, and even talking to their current students.
-It will take time and effort, but remember, you will be working with this person for 5+ years, so it is important to try to find the right one.
+that fits you the most by communicating with them, taking their courses, meeting and asking them questions, and talking to their current students.
+It will take time and effort, but since you will be working with this person for 5+ years, it is important to try to find the right one.
 
@@
 
-\paragraph{Tenure-line and teaching faculty} Tenure line faculty, consisting of tenured and tenured-track profs., focus on research, which includes publishing papers, obtaining grants, mentoring Ph.D students.
+\paragraph{Tenure-line and teaching faculty} Tenure line faculty, consisting of tenured and tenured-track profs., focus on research, which includes publishing papers, obtaining grants, and mentoring Ph.D students.
 They often have very low teaching load (e.g., 1 per semester).
@@
 In contrast, \emph{teaching} (or instructional faculty or professor of practice) mainly focus on teaching.
-They typically teach 3--4 classes per semester (which is quite a lot) and \emph{do not} have research responsibilities (e.g., they do not have to worry much about publishing papers or obtaining grants).
+They typically teach 3--4 classes per semester (which is quite a lot) and \emph{do not} have research responsibilities, e.g., they do not have to worry much about publishing papers or obtaining grants.
 They can mentor Ph.D.
 students but they typically cannot serve as a formal adviser of Ph.D.
-students (i.e., they \emph{cannot graduate} PhD students).
-
-It is worth noting that professors in a non-CS department also unlikely can serve as formal PhD advisers for CS students (see \autoref{sec:related-fields}).
-
-\paragraph{Titles} Faculty have rankings, e.g., assistant, associate, full, regardless if they are tenure line or teaching.
+students, i.e., they \emph{cannot graduate} PhD students.
+
+It is worth noting that professors in a non-CS department unlikely can serve as formal PhD advisers for CS students (see \autoref{sec:related-fields}).
+
+\paragraph{Titles} Faculty have titles such as assistant, associate, full, regardless if they are tenure line or teaching.
 Assistant means new faculty, associate means they have been promoted, and full means they are senior.
 A tenure track faculty starts with being an assistant professor and then gets tenure and promoted to associate (typically after 6 years) and then full (time varies, some become full within 3-4 years, some 10+ years, some remain associate).
+A non-tenure track faculty, e.g., teaching faculty or lecturer, also starts with assistant, then promoted to associated, and full.
 
@@
 Tenured faculty, e.g., associate and full profs., are likely older, more well-known, and have more experience in managing students.
-However, they might not push you as hard (they don't have to, they already got tenure).
-They might also expect you to figure things out yourself, i.e., so you need to be very independent.
+However, they might not push you as hard.
+They might also expect you to figure things out yourself, i.e., so you need to be independent.
 Some tenured faculty are also no longer active in research and more involved with administrative responsibilities or with their startup companies (this means they will likely not take new students).
@@
 
-Faculty received many "cold" e-mails from international students seeking for admission, TA, and RA.
+Faculty received many "cold" e-mails from prospective students seeking for admission, TA, and RA.
 Most of the time, we ignore these emails, but in some rare occasions we do answer them.
@@
 to ask about \emph{research opportunities}, or \emph{GTA/GRA} support, then \emph{yes}, I believe you should---it is \emph{worth it}.
-However, you need to put effort into it and really do it the \emph{right way}.
+However, you really need to put effort into it and do it the \emph{right way}.
 
@@
 
-\emph{MS Degree requirement}:  as mentioned in \autoref{sec:msrequirement} and \ref{sec:time}, CS PhD programs in the US do not require MS degree.
+\emph{MS Degree requirement}:  as mentioned in \autoref{sec:msrequirement} and \autoref{sec:time}, CS PhD programs in the US do not require MS degree.
 In contrast, many other countries do require MS degree before joining a PhD program.

### 2024-02-28 demystify.tex
updates

@@
 
-The short answer is that tenure-track faculty such as assistant professors are more likely to be young and active in research (they have to, in order to get tenure).
+The short answer is that tenure-track faculty such as assistant professors are likely young and active in research (they have to, in order to get tenure).
 Thus, they will likely have more time to work with you and push you to do research and publish.

### 2024-02-28 demystify.tex
updates

@@
 
-Most universities the US have CS programs. 
+Most universities the US have a CS program.
 However, while many universities offer PhD in their CS programs, many do not.
@@
   \item \emph{No faculty available to advise you}.
-You could have an excellent profile, but if there is no faculty who can advise you (e.g., they already have too many students), you will not be admitted.
+You could have an excellent profile, but if there is no faculty who is willing to advise you (e.g., they already have too many students), you will not be admitted.
 This is actually for your own good because you do not want to be in a program where you cannot find an adviser.
@@
 
-\section{Should I ask for feedback?}
+\paragraph{Should I ask for feedback?}
 No, don't bother.
@@
 One main concern for non-CS or non-STEM students is if you have the sufficient technical background, typically obtained through core CS courses.
-So you will want to show that you have such knowledge through your coursework, projects, or research.
+You will want to show that you have such knowledge through your coursework, projects, or research.
 For example, if you have taken a course on Algorithms, even online ones like Coursera, you can talk about it in your statement.
@@
 
-In short, you \emph{do not need} to formally taking CS courses, you just need to show that you have these essential knowledge through ways as mentioned.
+In short, you \emph{do not need} to formally taking CS courses, you just need to show that you have these essential knowledge, e.g., through the mentioned ways.
 Many universities are well-aware that incoming graduate students might not have all the technical background, so they often have a "bridge" courses to help students catch up.
@@
 
-\section{Is an MS degree required for admission to PhD in CS?}\label{sec:msrequirement}
+\section{Is an MS required for admission to PhD in CS?}\label{sec:msrequirement}
 No. In fact, student with BS can get MS degree ``along the way'' to PhD.  However, MS can help if it gives research experience or is from a more well-known school than your undergrad institution.

### 2024-02-29 demystify.tex
updates

@@
 \title{Demystifying the Computer Science \\
-  PhD Admission in the US\\{\large A Guideline for International \emph{and Domestic} Students}}
+  PhD Admission in the US\\{\large A Handbook for International \emph{and Domestic} Students}}
 
@@
 So I want to share some details about the admission process and advice for those who are interested in applying for a \textbf{PhD in Computer Science in the US}.
-While this document is primarily intended for students interested in CS, it might be relevant to students from various disciplines.
+While this handbook is primarily intended for students interested in CS, it might be relevant to students from various disciplines.
 Furthermore, although many examples are specifics for schools that I and other contributors of this document know about, the information should be generalizable to other R1\footnote{An \href{https://en.wikipedia.org/wiki/List_of_research_universities_in_the_United_States}{R1 institution} in the US is a research-intensive university with a high level of research activity across various disciplines.
@@
 
-In addition, this document can help \textbf{US faculty and admission committee} gain a better understanding of international students and their cultural differences.
+In addition, this handbook can help \textbf{US faculty and admission committee} gain a better understanding of international students and their cultural differences.
 By recognizing and leveraging these differences, CS programs in the US can attract larger and more competitive application pools from international students.
@@
 
-This document is available at
+This handbook is available at
 
@@
 
+\chapter{Domestic Students}\label{sec:domestic-students}
+
 \chapter{Cultural Differences and Other Issues}\label{sec:cultural}
@@
 
-\paragraph{History} This document was conceived during a lunch with Craig Yu at GMU.
+\paragraph{History} This handbook was conceived during a lunch with Craig Yu at GMU.
 We talked on about why GMU were not able to attract good Vietnamese and other international students, despite having a much stronger CS program than many schools that these students want to go to (part of the reason is described in \autoref{sec:selecting-ranking-schools}).
@@
 
-I started writing this document in May 2023 and have been updating it since then (mostly around deadline time when I tend to procrastinate, i.e., \emph{productive procrastination}!).
+I started writing this handbook in May 2023 and have been updating it since then (mostly around deadline time when I tend to procrastinate, i.e., \emph{productive procrastination}!).
 I have put the source code of this document on \href{https://github.com/nguyenthanhvuh/phd-cs-us}{GitHub} so that others can contribute to it.
@@
 I also mentor multiple students from Vietnam, and have close colleagues and friends who were once international students.
-I hope to capture the diverse challenges and experiences they've faced in this document, so that it  can be a valuable resource for prospective international students.
+I hope to capture the diverse challenges and experiences they've faced in this handbook, so that it  can be a valuable resource for prospective international students.
 Finally, my upbringing in the US provides a perspective aligned with American culture, allowing me to shed light on various issues, particularly those related to cultural differences (\autoref{sec:cultural}).
 
-\paragraph{Acknowledgement} Many people have contributed to this document.
+\paragraph{Acknowledgement} Many people have contributed to this work.
 Profs. Craig Yu (GMU), Hakan Aydin (GMU), 

### 2024-02-29 demystify.tex
updates

@@
 
-\begin{domesticbox}[Domestic students:] If you're a domestic student, you have several advantages in your application.
-First, we already know about your school (see \autoref{sec:your-school}), and you are also more familiar with the US education system and culture.
-You also have more opportunities for funding, e.g., through government scholarships for US citizens and residents.
-Finally, compared to international students, PhD applications in CS from domestic students are few and that can really help your case.
-\end{domesticbox}
-
 %% Vu: *what's a PhD?*  This [[https://matt.might.net/articles/phd-school-in-pictures/][series of pictures]] from [[https://matt.might.net][Matt Might]] illustrates what a PhD means.
@@
 
+Most of what is written in this document is applicable to both domestic\footnote{Domestic simply means you did your undergrad (or MS) at a US university.} and international students.
+However, there are some differences that domestic students should be aware of.
+This section lists some of these differences.
+
+\section{Application Materials}
+
+\paragraph{Fee Waiving} Some school might offer application fee waiver for domestic students.  You should check with the school you're applying to.
+
+\paragraph{School \autoref{sec:your-school}} we already know about your school, which is a plus (see \autoref{sec:your-school}).
+You are also more familiar with the US education system and culture (so hopefully you are aware of the cultural issues listed in \autoref{sec:cultural}).
+
+\paragraph{Standard Tests \autoref{sec:grades}} You do not need to take TOEFL or IELTS because you already did your undergrad (or MS).
+
+\paragraph{Transcripts} You do not need to get your transcripts evaluated by WES or other services (which can be a hassle and inaccurate).
+You can just send your official transcripts directly to the school you're applying to.
+
+\paragraph{Funding \S\ref{sec:funding}} You have more opportunities for funding, e.g., through government scholarships for US citizens and residents.
+You can also apply to specific programs \emph{before} you start your PhD, e.g., NSF Graduate Research Fellowship Program (GRFP) and Hertz Foundation Fellowship.
+These fellowships can be very competitive and will improve your admission chance if you get them.
+
+\paragraph{Research Experiences} You might have more research experiences than international students.
+For example, you likely have many opportunities to do research as an undergraduates, e.g., through REUs and internships.
+You should highlight these experiences in your application.
+
+\paragraph{Standing out} There are very few domestic applications compared to international ones (i.e., domestic students are minority in the CS PhD application pool).
+That makes you stand out and can really help your case.
+
+% \begin{domesticbox}[Domestic students:] If you're a domestic student, you have several advantages in your application.
+You also have more opportunities for funding, e.g., through government scholarships for US citizens and residents.
+Finally,
+% \end{domesticbox}
+
 \chapter{Cultural Differences and Other Issues}\label{sec:cultural}

### 2024-03-01 demystify.tex
updates

@@
 
-Additionally, \textbf{work experiences at renowned research laboratories}, such as Microsoft Research, can strengthen your
+Additionally, \textbf{work experiences at well-known research laboratories}, such as Microsoft Research, can strengthen your
 application.
@@
 As mentioned, if you're admitted to a \emph{good} CS PhD program, you should not have to worry about funding!  
-In the US, the common types of funding for PhD are \emph{graduate teaching assistant} (GTA or TA), \emph{graduate research assistant} (GRA or RA), and \emph{Fellowship}.
+In the US, the common types of funding for PhDs are \emph{graduate teaching assistant} (GTA or TA), \emph{graduate research assistant} (GRA or RA), and \emph{Fellowship}.
 RA is paid by a prof.
@@
 Several things about stipends.
-First, the amount of stipend \emph{varies} and depends on various factors such as location (e.g., a stipend in Washington DC is likely higher than in Lincoln, Nebraska due to higher living cost).
+First, the amount of stipend \emph{varies} and depends on factors such as location (e.g., a stipend in Washington DC is likely higher than in Lincoln, Nebraska due to higher living cost).
 Second, an academic year (AY)  year is typically \emph{9-month} in the US, so stipend is for 9 months (so divide by 9 for each month).
@@
 Fourth, like most source of income in the US, you will have to \emph{pay tax} on your stipend.
-Finally, private universities might pay more for stipend, but they could have higher "activity" or some other hidden fee (e.g., you may need to pay some fees for each credit hour).
-
-Students often complain their stipend being too low, but it is actually not bad and you can live comfortably with it.
+Finally, private universities might pay more for stipend, but they could have extra "activity" fee or some other hidden ones (e.g., you may need to pay some fees for each credit hour).
+
+\paragraph{Low Stipend?} Students often complain their stipend being too low, but it is actually not bad and you can live comfortably with it.
 In most cases, it is also enough to support your spouse and kids (many CS PhD students have their family with them).
@@
 \begin{commentbox}[Vu:]
-  At GMU CS, students admitted with TA have  4 years of GTA guaranteed, and in some cases also receive  stipend for the first summer.
+  At GMU CS, PhD students admitted TA has a 4 years of GTA guaranteed, and in also receive  stipend for the first summer.
 \end{commentbox}
@@
 
-\textbf{How to get RA?} When a professor recruits you, they will likely give you RA right away (e.g., when you apply).
+\textbf{How to get RA?} When a professor recruits you, they might offer you an RA right away (e.g., when you apply).
 A common scenario is that you first get admitted with TA, and then after a year or two find an adviser to support you with RA.
@@
 \begin{commentbox}[Vu:]
-  If you got recruited by a prof., who would give you RA right away, it's very likely you will get admitted.
-For example, if a prof., even if not in PhD adcom, wants to work with and funds you, the PhD adcom will respect that decision and admit you (unless your application has many red flags).
+  If you got recruited an offered an RA by a prof., it's very likely you will get admitted.
+For example, if a prof., even if not in PhD adcom, wants to fund you, the PhD adcom will respect that decision and admit you (unless your application has many red flags).
 \end{commentbox}
@@
 In general, fellowship, especially competitive ones from the government or company, is prestigious, and you will stand out if you get one.
-Every PhD student has pubs, but only superstars have NSF grad or Microsoft fellowship.
+Every PhD student has pubs, but only few  would have NSF Graduate or Microsoft fellowship.
 In fact, these are so prestigious that even if you didn't get it but make it to the final round or even ``honorable mentioning'', school will still mention you on their website and you still should put it on your CV.
@@
 \section{Why did I get rejected?}
-Many possible reasons!  Other than the obvious ones (e.g., weak research pontentials, low GPA, weak LoRs, unclear SOP), here are some that you might not expect:
+Many possible reasons!  Other than having weak profile (e.g., research potentials, GPA, LoRs, SOP), here are some that you might not expect:
 \begin{itemize}
   \item  \emph{Your English scores (e.g., IELTS or TOEFL) are not high enough}.
-While professors do not care much about these, there is a certain minimum set by the college or the university that you need to pass to get admitted, especially if you want to get funding (e.g., you need to be able to communicate well with students as a TA).
+While professors might not care much about these, there is a certain minimum set by the college or the university that you need to pass to get admitted, especially for TA funding (e.g., you need to be able to communicate well with students as a TA).
+Moreover, while professors are willing to fight to get you admitted, they might be reluctant to go against the standard requirement of the university or college on English scores.
 Note this is mostly for international students, as domestic one do not have to take these tests.
@@
 You will likely not get any useful feedback.
-We cannot reveal your evaluation results (e.g., if we say because you have weak LORs, then that would reveal something about your writers) or give you advice on how to improve.
+We are not willing to reveal your evaluation results or give you advice on how to improve.
 So it is best to just move on.
@@
 
-\section{What can you do to increase your admission chance?}\label{sec:improve-your-chance}
+\section{Increasing your admission chance}\label{sec:improve-your-chance}
 
@@
 
-Show something that makes you \textbf{stand out}, e.g., do you have a degree or background in \emph{dance} or \emph{music} and want to integrate them with CS?
+Given the high number of quality applicants and limited number of spots, other than having a good application profile, you really want to show something that makes you \textbf{stand out}, e.g., do you have a degree or background in \emph{dance} or \emph{music} and want to integrate them with CS?
 are you a female or a minority in CS (research for "URM minority in CS in the US" on Google)?

### 2024-03-02 demystify.tex
Updates from Overleaf

@@
 Obviously, there is no one-size-fit-all answer to this question.
-The best adviser is the one that you can work well with, i.e., \emph{fits} you.
+The best adviser is the one that you can work well with.
 This is difficult and takes time.
@@
 Here are some general advice to find professor.
-First, search for profs.
+Assuming you're not familiar with any particular one, then first, search for profs.
 that share similar research interests.
@@
 You can also go to their website, look at their papers and projects.
+
 In any case, you should contact the professor and talk to them.
@@
 
-Another effective way is taking graduate level courses in the topics you are interested in (remember: you have those 2 years to explore).
+Another effective way is taking graduate level courses in the topics you are interested in.
 Professors teaching these \emph{special topics courses} and \emph{research seminars} might be recruiting students---giving you even a higher chance.
@@
 
-Not every faculty can serve as your formal adviser.
+Not every faculty can serve as your official PhD adviser.
 Let's try to understand the different types of faculty and their roles.
@@
 \paragraph{Tenure-line and teaching faculty} Tenure line faculty, consisting of tenured and tenured-track profs., focus on research, which includes publishing papers, obtaining grants, and mentoring Ph.D students.
-They often have very low teaching load (e.g., 1 per semester).
+They often have low teaching load (e.g., 1 per semester).
 \emph{Tenured} faculty are professors who have been promoted to have a permanent position (informally, very hard to fire them).
 \emph{Tenure-track} faculty are (more likely young) professors who are on the track to get tenure.
-\textbf{Both tenure and tenured track faculty can serve as a formal adviser of Ph.D.
-students}.
-This is very important because you need to have a tenure line faculty as your main PhD adviser.
+At most universities, \textbf{only} tenure and tenured track faculty can serve as a formal adviser of Ph.D.
+students.
 \autoref{sec:tenure-vs-tenure-track} talks more about choosing between tenure and tenured-track professor as your adviser.
@@
 
-It is worth noting that professors in a non-CS department unlikely can serve as formal PhD advisers for CS students (see \autoref{sec:related-fields}).
+Also note that professors in a non-CS department unlikely can serve as formal PhD advisers for CS students (see \autoref{sec:related-fields}).
 
@@
 \begin{center}
-  \includegraphics[scale=0.4]{files/emails.png}
+  \includegraphics[scale=0.6]{files/emails.png}
 \end{center}
 
-Faculty received many "cold" e-mails from prospective students seeking for admission, TA, and RA.
+Faculty often receive many "cold" e-mails from prospective students.
 Most of the time, we ignore these emails, but in some rare occasions we do answer them.
-So why do we ignore some but reply to others and how to write an email that get our attention?
+So how to write an email that get our attention?
 
@@
 However, if your goal is a PhD in CS, then you likely need to be in the CS dept \emph{and} advised by a CS faculty.
-In fact, if a faculty is not in CS, it is unlikely that they can be the \emph{main} adviser of a CS PhD student.
-They may \emph{co-advise} or be in the PhD dissertation committee, but your main adviser will likely be a tenured or tenure-track faculty in CS.
+A non-CS may \emph{co-advise} or be in the PhD dissertation committee, but your main PhD adviser will likely be a tenured or tenure-track faculty in CS (\autoref{sec:faculty-types}).
 If in doubt, check with the CS department for their requirements.

### 2024-03-02 demystify.tex
Updates from Overleaf

@@
 
-\paragraph{Why we do not waive application fee?}  This is typically a requirement of the university.
+\paragraph{Why CS depts do not waive application fee?}  Note that some universities do waive application fee, however most do not.
+In any case, application fee is typically a requirement of the university.
 Individual departments and programs do not have the flexibility to waive the application fee, even if they want to.
@@
 In my opinion, requiring applicants to pay the fee helps ensure their seriousness, as it filters out non-serious candidates.
-Also, if the application process were free for everyone, we would receive an overwhelming number of applications to review.
+Also, if the application process were free for everyone, most good CS programs would receive an overwhelming number of applications to review.
 
@@
 In short, focus on your research motivation and goal, and convince us that you can achieve it through your experience, e.g., published papers.
-Also, talk about things that adcom members might not know about, e.g., your personal Github project with 1K+ stars or your regular contributions to well-known open-source projects.
-Finally, talk about what makes you different than the other 1000 applicants (e.g., see \autoref{sec:improve-your-chance} for increasing your admission chance).
+You should also back up your claims with concrete evidences.
+For example, if you say you have experience with teaching, then show what you did, e.g., undergrad TA or mentoring someone.
+If you say you work on project X, then show some results (paper submitted (or even rejected), achieved X\% performance improvement over ..).
+Do talk about things that adcom members might not know about and can help make you different than the other 1000 applicants, e.g.,  your personal Github project with 1K+ stars or your regular contributions to well-known open-source projects (see \autoref{sec:improve-your-chance} for increasing your admission chance).
 
@@
 
-Finally, this is something easy to do, but is neglected by many
+This is something easy to do, but is neglected by many
 applicants: \textbf{customize the statement} for the school you're applying to,
@@
 Adcom will look for this part, so do not skip it.
+
+Finally, you should get your SOP reviewed by LoR writers, professors (especially those who have served in adcom), or even graduate PhD students (because they have been through this process).
 
@@
 
-Graduating from top universities \emph{that we recognize} helps. For example, if your school is well-known, then it is \emph{``top foreign''}, which is definitely a plus.
+Graduating from top universities \emph{that we recognize} helps.
+For example, if you are an international student and your school is well-known, then it is \emph{``top foreign''}, which is definitely a plus.
 However, if we do not know much about schools in your country, then we are uncertain about the quality of your school and  likely treat your school as \emph{``unknown foreign''}, which can be a minus point.
@@
   \footnotesize
-  \caption{Different types of PhD funding. All Covered: includes tuition, insurance, and stipend.}\label{tab:funding}
+  \caption{Different types of PhD funding. All cover tuition, insurance, and stipend.}\label{tab:funding}
   \begin{tabular}{c|c|c|c}
@@
 
-\textbf{How to get Fellowship?} You need to apply for them.  Many schools provide such fellowships to attract admitted students.
-External fellowships including from the US government and tech companies (e.g., Google, Microsoft, Facebook).
+\textbf{How to get Fellowship?}   Many schools provide fellowships to attract admitted students.
+You likely will not need to do anything and the adcom will recommend such fellowships to admitted students.
+Some schools automatically offer a fellowship to all accepted students, while others only award it to a limited number of admitted students, such as the top percentile.
+
+You likely need to apply for external fellowships including those from the US government and tech companies (e.g., Google, Microsoft, Facebook).
 Fellowships from US government would likely require US citizenship or residency.
 % However, tech companies including Google, Microsoft, Facebook have fellowships that international students can apply for.
-
-Prestigious external fellowships typically require a clear and good research plan, so it is a good idea to wait until at least your second year to have research experience and even publication before applying.
+Prestigious external fellowships typically require a clear and good research plan, so it might be a good idea to wait until at least your second year to have research experience and even publication before applying.
 Remember, you're competing with the top PhD students at top universities worldwide.

### 2024-03-03 demystify.tex

@@
 Having a strong letter from an internationally recognized researcher will greatly strengthen your application.
-In fact, a strong letter from a well-known researcher can outweight the lack of publications.
+In fact, a strong letter from a well-known researcher can outweigh the lack of publications.
 
@@
 
+\paragraph{How to help your LOR writer} Your LoR writer can ask you to provide additional information to help them write your letter.
+For example, they might ask your GPA, research and work experiences, papers (if any), or anything you want them to mention.
+It might also be a good idea to provide them such information even when they do not ask, especially when you think there are additional information that should be mentioned.
+
 \paragraph{Self-written Letters} Many international students write letters themselves, typically due to the requests of their profs., and have them signed by their profs.
@@
 There are various guides on writing SOP, e.g.,~\cite{blattman2022writing}, and \href{https://cs-sop.org/}{many example statements} are available.
-So I will not talk too much about statements.
 In short, focus on your research motivation and goal, and convince us that you can achieve it through your experience, e.g., published papers.
@@
 If you say you work on project X, then show some results (paper submitted (or even rejected), achieved X\% performance improvement over ..).
-Do talk about things that adcom members might not know about and can help make you different than the other 1000 applicants, e.g.,  your personal Github project with 1K+ stars or your regular contributions to well-known open-source projects (see \autoref{sec:improve-your-chance} for increasing your admission chance).
-
-% \begin{commentbox}[Hung:]
-%   I think asking for a “research vision” from a Ph.D.
-applicant is too much.
-Even people graduated with a Ph.D.
-have a great difficulty in explicating their research vision.
-In my opinion, you should focus on showing (not telling) your research passion (if you do have).
-% \end{commentbox}
+
+You should also talk about things that adcom members might not know about and can help make you different than the other 1000 applicants, e.g.,  your personal Github project with 1K+ stars or your regular contributions to well-known open-source projects (see \autoref{sec:improve-your-chance} for increasing your admission chance).
 
@@
 What can you do about this?
-You can explain about your school in your statement (ask your LOR writer to do that too).
-Of course, if you're interested in working with Vietnamese, consider  \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{CS PhD programs in the US that have Vietnamese professors}.
+You can explain about your school in your SOP (and also ask your LOR writer to do that).
+Of course, if you're interested in working with Vietnamese, consider  \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{CS programs in the US that have Vietnamese professors}.
 %It might also be helpful to have your CS dept to put itself on CSRankings so that others know about the dept, its faculty, and research.
@@
 \paragraph{Timeline} The timeline for interviews and notification letters varies.
-Faculty setup interviews based on their (very busy) schedule.
+Faculty setup interviews based on their (busy) schedule.
 Do not be surprised if you get an interview invitation in the last minute.
-Again, some schools do not do interviews at all.
+Some schools do not do interviews at all.
 
@@
 Important factors to consider include the reputation of schools and professors (\autoref{sec:schoolsandprofs}), and funding availability (\autoref{sec:funding}).
-You will have to make your decision by around \emph{April 15}.
+You will have to make your decision by around \emph{April 15} (see \href{https://cgsnet.org/wp-content/uploads/2024/01/CGS_April15_Resolution_Jan312024.pdf}{why?}).
 
@@
   \centering
-  \footnotesize
+  \small
   \caption{Different types of PhD funding. All cover tuition, insurance, and stipend.}\label{tab:funding}

### 2024-03-03 demystify.tex
updates

--- 2024-03-02 demystify.tex
@@
 \title{Demystifying the Computer Science \\
-  PhD Admission in the US\\{\large A Handbook for International \emph{and Domestic} Students}}
+  PhD Admission in the US\\{\large A Guideline for International \emph{and Domestic} Students}}
 
@@
 So I want to share some details about the admission process and advice for those who are interested in applying for a \textbf{PhD in Computer Science in the US}.
-While this handbook is primarily intended for students interested in CS, it might be relevant to students from various disciplines.
+While this document is primarily intended for students interested in CS, it might be relevant to students from various disciplines.
 Furthermore, although many examples are specifics for schools that I and other contributors of this document know about, the information should be generalizable to other R1\footnote{An \href{https://en.wikipedia.org/wiki/List_of_research_universities_in_the_United_States}{R1 institution} in the US is a research-intensive university with a high level of research activity across various disciplines.
@@
 
-In addition, this handbook can help \textbf{US faculty and admission committee} gain a better understanding of international students and their cultural differences.
+In addition, this document can help \textbf{US faculty and admission committee} gain a better understanding of international students and their cultural differences.
 By recognizing and leveraging these differences, CS programs in the US can attract larger and more competitive application pools from international students.
@@
 
-This handbook is available at
+This document is available at
 
@@
 
+\begin{domesticbox}[Domestic students:] If you're a domestic student, you have several advantages in your application.
+First, we already know about your school (see \autoref{sec:your-school}), and you are also more familiar with the US education system and culture.
+You also have more opportunities for funding, e.g., through government scholarships for US citizens and residents.
+Finally, compared to international students, PhD applications in CS from domestic students are few and that can really help your case.
+\end{domesticbox}
+
 %% Vu: *what's a PhD?*  This [[https://matt.might.net/articles/phd-school-in-pictures/][series of pictures]] from [[https://matt.might.net][Matt Might]] illustrates what a PhD means.
@@
 
-Additionally, \textbf{work experiences at well-known research laboratories}, such as Microsoft Research, can strengthen your
+Additionally, \textbf{work experiences at renowned research laboratories}, such as Microsoft Research, can strengthen your
 application.
@@
 As mentioned, if you're admitted to a \emph{good} CS PhD program, you should not have to worry about funding!  
-In the US, the common types of funding for PhDs are \emph{graduate teaching assistant} (GTA or TA), \emph{graduate research assistant} (GRA or RA), and \emph{Fellowship}.
+In the US, the common types of funding for PhD are \emph{graduate teaching assistant} (GTA or TA), \emph{graduate research assistant} (GRA or RA), and \emph{Fellowship}.
 RA is paid by a prof.
@@
 Several things about stipends.
-First, the amount of stipend \emph{varies} and depends on factors such as location (e.g., a stipend in Washington DC is likely higher than in Lincoln, Nebraska due to higher living cost).
+First, the amount of stipend \emph{varies} and depends on various factors such as location (e.g., a stipend in Washington DC is likely higher than in Lincoln, Nebraska due to higher living cost).
 Second, an academic year (AY)  year is typically \emph{9-month} in the US, so stipend is for 9 months (so divide by 9 for each month).
@@
 Fourth, like most source of income in the US, you will have to \emph{pay tax} on your stipend.
-Finally, private universities might pay more for stipend, but they could have extra "activity" fee or some other hidden ones (e.g., you may need to pay some fees for each credit hour).
-
-\paragraph{Low Stipend?} Students often complain their stipend being too low, but it is actually not bad and you can live comfortably with it.
+Finally, private universities might pay more for stipend, but they could have higher "activity" or some other hidden fee (e.g., you may need to pay some fees for each credit hour).
+
+Students often complain their stipend being too low, but it is actually not bad and you can live comfortably with it.
 In most cases, it is also enough to support your spouse and kids (many CS PhD students have their family with them).
@@
 \begin{commentbox}[Vu:]
-  At GMU CS, PhD students admitted TA has a 4 years of GTA guaranteed, and in also receive  stipend for the first summer.
+  At GMU CS, students admitted with TA have  4 years of GTA guaranteed, and in some cases also receive  stipend for the first summer.
 \end{commentbox}
@@
 
-\textbf{How to get RA?} When a professor recruits you, they might offer you an RA right away (e.g., when you apply).
+\textbf{How to get RA?} When a professor recruits you, they will likely give you RA right away (e.g., when you apply).
 A common scenario is that you first get admitted with TA, and then after a year or two find an adviser to support you with RA.
@@
 \begin{commentbox}[Vu:]
-  If you got recruited an offered an RA by a prof., it's very likely you will get admitted.
-For example, if a prof., even if not in PhD adcom, wants to fund you, the PhD adcom will respect that decision and admit you (unless your application has many red flags).
+  If you got recruited by a prof., who would give you RA right away, it's very likely you will get admitted.
+For example, if a prof., even if not in PhD adcom, wants to work with and funds you, the PhD adcom will respect that decision and admit you (unless your application has many red flags).
 \end{commentbox}
@@
 In general, fellowship, especially competitive ones from the government or company, is prestigious, and you will stand out if you get one.
-Every PhD student has pubs, but only few  would have NSF Graduate or Microsoft fellowship.
+Every PhD student has pubs, but only superstars have NSF grad or Microsoft fellowship.
 In fact, these are so prestigious that even if you didn't get it but make it to the final round or even ``honorable mentioning'', school will still mention you on their website and you still should put it on your CV.
@@
 Obviously, there is no one-size-fit-all answer to this question.
-The best adviser is the one that you can work well with.
+The best adviser is the one that you can work well with, i.e., \emph{fits} you.
 This is difficult and takes time.
@@
 Here are some general advice to find professor.
-Assuming you're not familiar with any particular one, then first, search for profs.
+First, search for profs.
 that share similar research interests.
@@
 If you want to work with SE \emph{and} AI, you can search for faculty who work in both SE and AI.
-You can also go to their website, look at their papers and projects.
-
-In any case, you should contact the professor and talk to them.
-For example, you can (cold) email them and ask about their research and if they are taking new students (\autoref{sec:contact}), and schedule appointments to chat with them (office hrs are often the best time).
+Then you can contact the professor you're interested in (\autoref{sec:contact}).
+%For example, you can (cold) email them and ask about their research and if they are taking new students (\autoref{sec:contact}), and schedule appointments to chat with them (office hrs are often the best time).
 
@@
 
-Another effective way is taking graduate level courses in the topics you are interested in.
+Another effective way is taking graduate level courses in the topics you are interested in (remember: you have those 2 years to explore).
 Professors teaching these \emph{special topics courses} and \emph{research seminars} might be recruiting students---giving you even a higher chance.
@@
 
-Not every faculty can serve as your official PhD adviser.
+Not every faculty can serve as your formal adviser.
 Let's try to understand the different types of faculty and their roles.
@@
 \paragraph{Tenure-line and teaching faculty} Tenure line faculty, consisting of tenured and tenured-track profs., focus on research, which includes publishing papers, obtaining grants, and mentoring Ph.D students.
-They often have low teaching load (e.g., 1 per semester).
+They often have very low teaching load (e.g., 1 per semester).
 \emph{Tenured} faculty are professors who have been promoted to have a permanent position (informally, very hard to fire them).
 \emph{Tenure-track} faculty are (more likely young) professors who are on the track to get tenure.
-At most universities, \textbf{only} tenure and tenured track faculty can serve as a formal adviser of Ph.D.
-students.
+\textbf{Both tenure and tenured track faculty can serve as a formal adviser of Ph.D.
+students}.
+This is very important because you need to have a tenure line faculty as your main PhD adviser.
 \autoref{sec:tenure-vs-tenure-track} talks more about choosing between tenure and tenured-track professor as your adviser.
@@
 
-Also note that professors in a non-CS department unlikely can serve as formal PhD advisers for CS students (see \autoref{sec:related-fields}).
+It is worth noting that professors in a non-CS department unlikely can serve as formal PhD advisers for CS students (see \autoref{sec:related-fields}).
 
@@
 \begin{center}
-  \includegraphics[scale=0.6]{files/emails.png}
+  \includegraphics[scale=0.4]{files/emails.png}
 \end{center}
 
-Faculty often receive many "cold" e-mails from prospective students.
+Faculty received many "cold" e-mails from prospective students seeking for admission, TA, and RA.
 Most of the time, we ignore these emails, but in some rare occasions we do answer them.
-So how to write an email that get our attention?
+So why do we ignore some but reply to others and how to write an email that get our attention?
 
@@
 However, if your goal is a PhD in CS, then you likely need to be in the CS dept \emph{and} advised by a CS faculty.
-A non-CS may \emph{co-advise} or be in the PhD dissertation committee, but your main PhD adviser will likely be a tenured or tenure-track faculty in CS (\autoref{sec:faculty-types}).
+In fact, if a faculty is not in CS, it is unlikely that they can be the \emph{main} adviser of a CS PhD student.
+They may \emph{co-advise} or be in the PhD dissertation committee, but your main adviser will likely be a tenured or tenure-track faculty in CS.
 If in doubt, check with the CS department for their requirements.
@@
 \section{Why did I get rejected?}
-Many possible reasons!  Other than having weak profile (e.g., research potentials, GPA, LoRs, SOP), here are some that you might not expect:
+Many possible reasons!  Other than the obvious ones (e.g., weak research pontentials, low GPA, weak LoRs, unclear SOP), here are some that you might not expect:
 \begin{itemize}
   \item  \emph{Your English scores (e.g., IELTS or TOEFL) are not high enough}.
-While professors might not care much about these, there is a certain minimum set by the college or the university that you need to pass to get admitted, especially for TA funding (e.g., you need to be able to communicate well with students as a TA).
-Moreover, while professors are willing to fight to get you admitted, they might be reluctant to go against the standard requirement of the university or college on English scores.
+While professors do not care much about these, there is a certain minimum set by the college or the university that you need to pass to get admitted, especially if you want to get funding (e.g., you need to be able to communicate well with students as a TA).
 Note this is mostly for international students, as domestic one do not have to take these tests.
@@
 You will likely not get any useful feedback.
-We are not willing to reveal your evaluation results or give you advice on how to improve.
+We cannot reveal your evaluation results (e.g., if we say because you have weak LORs, then that would reveal something about your writers) or give you advice on how to improve.
 So it is best to just move on.
@@
 
-\section{Increasing your admission chance}\label{sec:improve-your-chance}
+\section{What can you do to increase your admission chance?}\label{sec:improve-your-chance}
 
@@
 
-Given the high number of quality applicants and limited number of spots, other than having a good application profile, you really want to show something that makes you \textbf{stand out}, e.g., do you have a degree or background in \emph{dance} or \emph{music} and want to integrate them with CS?
+Show something that makes you \textbf{stand out}, e.g., do you have a degree or background in \emph{dance} or \emph{music} and want to integrate them with CS?
 are you a female or a minority in CS (research for "URM minority in CS in the US" on Google)?
@@
 
-\chapter{Domestic Students}\label{sec:domestic-students}
-
-Most of what is written in this document is applicable to both domestic\footnote{Domestic simply means you did your undergrad (or MS) at a US university.} and international students.
-However, there are some differences that domestic students should be aware of.
-This section lists some of these differences.
-
-\section{Application Materials}
-
-\paragraph{Fee Waiving} Some school might offer application fee waiver for domestic students.  You should check with the school you're applying to.
-
-\paragraph{School \autoref{sec:your-school}} we already know about your school, which is a plus (see \autoref{sec:your-school}).
-You are also more familiar with the US education system and culture (so hopefully you are aware of the cultural issues listed in \autoref{sec:cultural}).
-
-\paragraph{Standard Tests \autoref{sec:grades}} You do not need to take TOEFL or IELTS because you already did your undergrad (or MS).
-
-\paragraph{Transcripts} You do not need to get your transcripts evaluated by WES or other services (which can be a hassle and inaccurate).
-You can just send your official transcripts directly to the school you're applying to.
-
-\paragraph{Funding \S\ref{sec:funding}} You have more opportunities for funding, e.g., through government scholarships for US citizens and residents.
-You can also apply to specific programs \emph{before} you start your PhD, e.g., NSF Graduate Research Fellowship Program (GRFP) and Hertz Foundation Fellowship.
-These fellowships can be very competitive and will improve your admission chance if you get them.
-
-\paragraph{Research Experiences} You might have more research experiences than international students.
-For example, you likely have many opportunities to do research as an undergraduates, e.g., through REUs and internships.
-You should highlight these experiences in your application.
-
-\paragraph{Standing out} There are very few domestic applications compared to international ones (i.e., domestic students are minority in the CS PhD application pool).
-That makes you stand out and can really help your case.
-
-% \begin{domesticbox}[Domestic students:] If you're a domestic student, you have several advantages in your application.
-You also have more opportunities for funding, e.g., through government scholarships for US citizens and residents.
-Finally,
-% \end{domesticbox}
-
 \chapter{Cultural Differences and Other Issues}\label{sec:cultural}
@@
 
-\paragraph{History} This handbook was conceived during a lunch with Craig Yu at GMU.
+\paragraph{History} This document was conceived during a lunch with Craig Yu at GMU.
 We talked on about why GMU were not able to attract good Vietnamese and other international students, despite having a much stronger CS program than many schools that these students want to go to (part of the reason is described in \autoref{sec:selecting-ranking-schools}).
@@
 
-I started writing this handbook in May 2023 and have been updating it since then (mostly around deadline time when I tend to procrastinate, i.e., \emph{productive procrastination}!).
+I started writing this document in May 2023 and have been updating it since then (mostly around deadline time when I tend to procrastinate, i.e., \emph{productive procrastination}!).
 I have put the source code of this document on \href{https://github.com/nguyenthanhvuh/phd-cs-us}{GitHub} so that others can contribute to it.
@@
 I also mentor multiple students from Vietnam, and have close colleagues and friends who were once international students.
-I hope to capture the diverse challenges and experiences they've faced in this handbook, so that it  can be a valuable resource for prospective international students.
+I hope to capture the diverse challenges and experiences they've faced in this document, so that it  can be a valuable resource for prospective international students.
 Finally, my upbringing in the US provides a perspective aligned with American culture, allowing me to shed light on various issues, particularly those related to cultural differences (\autoref{sec:cultural}).
 
-\paragraph{Acknowledgement} Many people have contributed to this work.
+\paragraph{Acknowledgement} Many people have contributed to this document.
 Profs. Craig Yu (GMU), Hakan Aydin (GMU), 

### 2024-03-03 demystify.tex

--- 2024-03-03 demystify.tex
@@
 For example, they might ask your GPA, research and work experiences, papers (if any), or anything you want them to mention.
-It might also be a good idea to provide them such information even when they do not ask, especially when you think there are additional information that should be mentioned.
+You can also provide them with a draft of your SOP so that they can see what you are saying about yourself and can complement that with their own perspective.
+It might also be a good idea to provide them such information even when they do not ask, especially if you have not interacted with them much or have not done much research with them.
 
@@
 If you want to work with SE \emph{and} AI, you can search for faculty who work in both SE and AI.
-You can also go to their website, look at their papers and projects.
-
-In any case, you should contact the professor and talk to them.
-For example, you can (cold) email them and ask about their research and if they are taking new students (\autoref{sec:contact}), and schedule appointments to chat with them (office hrs are often the best time).
+Then you do research about that professor and contact them (\autoref{sec:contact}).
 

### 2024-03-03 demystify.tex
updates

@@
 One way to mitigate this is making these issues known in your SOP.
-Universities with Vietnamese profs are probably aware of them, but in general your letter writers \emph{and} you can explicitly mention these in their letters and your statement.
+\href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{Universities with Vietnamese profs} are probably aware of them, but in general your letter writers \emph{and} you can explicitly mention these in their letters and your statement.
 \end{commentbox}
@@
 
-\paragraph{Open House} Most schools will have \textbf{Open House} events, which are a great resource to learn about the school, department, faculty, research, living, etc.
+\paragraph{Open House} Most schools will have \textbf{Open House} or visit events, which are a great resource to learn about the school, department, faculty, research, living, etc.
 Even if you can't come in person, you should attend virtually and meet with individual faculty.

### 2024-03-03 demystify.tex
Updates from Overleaf
updates

--- 2024-03-03 demystify.tex
@@
 
-\paragraph{Why CS depts do not waive application fee?}  Note that some universities do waive application fee, however most do not.
-In any case, application fee is typically a requirement of the university.
+\paragraph{Why we do not waive application fee?}  This is typically a requirement of the university.
 Individual departments and programs do not have the flexibility to waive the application fee, even if they want to.
@@
 In my opinion, requiring applicants to pay the fee helps ensure their seriousness, as it filters out non-serious candidates.
-Also, if the application process were free for everyone, most good CS programs would receive an overwhelming number of applications to review.
+Also, if the application process were free for everyone, we would receive an overwhelming number of applications to review.
 
@@
 Having a strong letter from an internationally recognized researcher will greatly strengthen your application.
-In fact, a strong letter from a well-known researcher can outweigh the lack of publications.
+In fact, a strong letter from a well-known researcher can outweight the lack of publications.
 
@@
 
-\paragraph{How to help your LOR writer} Your LoR writer can ask you to provide additional information to help them write your letter.
-For example, they might ask your GPA, research and work experiences, papers (if any), or anything you want them to mention.
-You can also provide them with a draft of your SOP so that they can see what you are saying about yourself and can complement that with their own perspective.
-It might also be a good idea to provide them such information even when they do not ask, especially if you have not interacted with them much or have not done much research with them.
-
 \paragraph{Self-written Letters} Many international students write letters themselves, typically due to the requests of their profs., and have them signed by their profs.
@@
 There are various guides on writing SOP, e.g.,~\cite{blattman2022writing}, and \href{https://cs-sop.org/}{many example statements} are available.
+So I will not talk too much about statements.
 In short, focus on your research motivation and goal, and convince us that you can achieve it through your experience, e.g., published papers.
-You should also back up your claims with concrete evidences.
-For example, if you say you have experience with teaching, then show what you did, e.g., undergrad TA or mentoring someone.
-If you say you work on project X, then show some results (paper submitted (or even rejected), achieved X\% performance improvement over ..).
-
-You should also talk about things that adcom members might not know about and can help make you different than the other 1000 applicants, e.g.,  your personal Github project with 1K+ stars or your regular contributions to well-known open-source projects (see \autoref{sec:improve-your-chance} for increasing your admission chance).
-
-This is something easy to do, but is neglected by many
+Also, talk about things that adcom members might not know about, e.g., your personal Github project with 1K+ stars or your regular contributions to well-known open-source projects.
+Finally, talk about what makes you different than the other 1000 applicants (e.g., see \autoref{sec:improve-your-chance} for increasing your admission chance).
+
+% \begin{commentbox}[Hung:]
+%   I think asking for a “research vision” from a Ph.D.
+applicant is too much.
+Even people graduated with a Ph.D.
+have a great difficulty in explicating their research vision.
+In my opinion, you should focus on showing (not telling) your research passion (if you do have).
+% \end{commentbox}
+
+Finally, this is something easy to do, but is neglected by many
 applicants: \textbf{customize the statement} for the school you're applying to,
@@
 Adcom will look for this part, so do not skip it.
-
-Finally, you should get your SOP reviewed by LoR writers, professors (especially those who have served in adcom), or even graduate PhD students (because they have been through this process).
 
@@
 
-Graduating from top universities \emph{that we recognize} helps.
-For example, if you are an international student and your school is well-known, then it is \emph{``top foreign''}, which is definitely a plus.
+Graduating from top universities \emph{that we recognize} helps. For example, if your school is well-known, then it is \emph{``top foreign''}, which is definitely a plus.
 However, if we do not know much about schools in your country, then we are uncertain about the quality of your school and  likely treat your school as \emph{``unknown foreign''}, which can be a minus point.
@@
 What can you do about this?
-You can explain about your school in your SOP (and also ask your LOR writer to do that).
-Of course, if you're interested in working with Vietnamese, consider  \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{CS programs in the US that have Vietnamese professors}.
+You can explain about your school in your statement (ask your LOR writer to do that too).
+Of course, if you're interested in working with Vietnamese, consider  \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{CS PhD programs in the US that have Vietnamese professors}.
 %It might also be helpful to have your CS dept to put itself on CSRankings so that others know about the dept, its faculty, and research.
@@
 \paragraph{Timeline} The timeline for interviews and notification letters varies.
-Faculty setup interviews based on their (busy) schedule.
+Faculty setup interviews based on their (very busy) schedule.
 Do not be surprised if you get an interview invitation in the last minute.
-Some schools do not do interviews at all.
+Again, some schools do not do interviews at all.
 
@@
 Important factors to consider include the reputation of schools and professors (\autoref{sec:schoolsandprofs}), and funding availability (\autoref{sec:funding}).
-You will have to make your decision by around \emph{April 15} (see \href{https://cgsnet.org/wp-content/uploads/2024/01/CGS_April15_Resolution_Jan312024.pdf}{why?}).
+You will have to make your decision by around \emph{April 15}.
 
@@
   \centering
-  \small
-  \caption{Different types of PhD funding. All cover tuition, insurance, and stipend.}\label{tab:funding}
+  \footnotesize
+  \caption{Different types of PhD funding. All Covered: includes tuition, insurance, and stipend.}\label{tab:funding}
   \begin{tabular}{c|c|c|c}
@@
 
-\textbf{How to get Fellowship?}   Many schools provide fellowships to attract admitted students.
-You likely will not need to do anything and the adcom will recommend such fellowships to admitted students.
-Some schools automatically offer a fellowship to all accepted students, while others only award it to a limited number of admitted students, such as the top percentile.
-
-You likely need to apply for external fellowships including those from the US government and tech companies (e.g., Google, Microsoft, Facebook).
+\textbf{How to get Fellowship?} You need to apply for them.  Many schools provide such fellowships to attract admitted students.
+External fellowships including from the US government and tech companies (e.g., Google, Microsoft, Facebook).
 Fellowships from US government would likely require US citizenship or residency.
 % However, tech companies including Google, Microsoft, Facebook have fellowships that international students can apply for.
-Prestigious external fellowships typically require a clear and good research plan, so it might be a good idea to wait until at least your second year to have research experience and even publication before applying.
+
+Prestigious external fellowships typically require a clear and good research plan, so it is a good idea to wait until at least your second year to have research experience and even publication before applying.
 Remember, you're competing with the top PhD students at top universities worldwide.
@@
 If you want to work with SE \emph{and} AI, you can search for faculty who work in both SE and AI.
-Then you do research about that professor and contact them (\autoref{sec:contact}).
+You can also go to their website, look at their papers and projects.
+In any case, you should contact the professor and talk to them.
+For example, you can (cold) email them and ask about their research and if they are taking new students (\autoref{sec:contact}), and schedule appointments to chat with them (office hrs are often the best time).
 
@@
 \section{Why did I get rejected?}
-Many possible reasons!  Other than having weak profile (e.g., research potentials, GPA, LoRs, SOP), here are some that you might not expect:
+Many possible reasons!  Other than not having very strong profile (e.g., research potentials, GPA, LoRs, SOP), here are some that you might not expect:
 \begin{itemize}
+
+\item \emph{You are not a good fit to the program}.  You could have an excellent profile, but if you are interested in a research area that the program does not have, you might not be admitted.
+Similarly, if there is no faculty who is willing to advise you (e.g., they are on sabbatical, NSF leave, or already have too many students currently), you will not be admitted.
+This is actually good for you, as you don't want to be in a program where no one can advise you.
+
+For example, in the 2024 admission cycle, there is a huge surge in students who are interested in AI and NLP (thanks, ChatGPT!).
+However, our AI/NLP faculty cannot take more students and therefore we cannot admit many applicants, many of whom have excellent profiles (e.g., papers in top conferences, strong LoRs, etc.).
+
 \item  \emph{Your English scores (e.g., IELTS or TOEFL) are not high enough}.
 While professors might not care much about these, there is a certain minimum set by the college or the university that you need to pass to get admitted, especially for TA funding (e.g., you need to be able to communicate well with students as a TA).
-Moreover, while professors are willing to fight to get you admitted, they might be reluctant to go against the standard requirement of the university or college on English scores.
-Note this is mostly for international students, as domestic one do not have to take these tests.
-
-  \item \emph{No faculty available to advise you}.
-You could have an excellent profile, but if there is no faculty who is willing to advise you (e.g., they already have too many students), you will not be admitted.
-This is actually for your own good because you do not want to be in a program where you cannot find an adviser.
+Thus, while professors are willing to argue for your case, they might be reluctant to go against the standard requirement of the university on English scores.
+Note this is mostly for international students, as domestic ones (or those from these \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-GMU#standard-tests-waiver-eligible-countries}{countries}) do not have to take these tests.
+
 \end{itemize}

### 2024-03-03 demystify.tex
Updates from Overleaf

--- 2024-03-03 demystify.tex
@@
 Many programs do not have the tradition of interviewing applicants.
-For example, at GMU, most of our admitted students with TA do not go through any interviews.
+For example, at GMU, most of our admitted students with TA do not go through interviews.
 
@@
 
-Faculty often receive many "cold" e-mails from prospective students.
-Most of the time, we ignore these emails, but in some rare occasions we do answer them.
-So how to write an email that get our attention?
+Faculty often receive "cold" e-mails from prospective students. Most of the time, we ignore these emails, but in some rare occasions we do answer them. So how to write an email that get our attention?
 
@@
 
-So what to do if you want someone to look at your profile and give input?
-You could ask your professors, collaborators, or those who have previously applied.
-For these kind of feedback,  don't ask strangers like random profs., instead ask someone you have personal connection with.
+So how to get someone to look at your profile and give input?
+You could ask your professors, LoR writers, collaborators, or those who have previously applied.
+For these kind of feedback, ask someone you have personal connection with.
 
@@
 \section{Why did I get rejected?}
-Many possible reasons!  Other than having weak profile (e.g., research potentials, GPA, LoRs, SOP), here are some that you might not expect:
+Many possible reasons!  Other than not having a strong profile (e.g., research potentials, GPA, LoRs, SOP), here are some that you might not expect:
 \begin{itemize}

### 2024-03-03 demystify.tex

@@
 \section{Why did I get rejected?}
-Many possible reasons!  Other than not having very strong profile (e.g., research potentials, GPA, LoRs, SOP), here are some that you might not expect:
+Many possible reasons!  Other than not having a strong profile (e.g., research potentials, GPA, LoRs, SOP), here are some that you might not expect:
+
 \begin{itemize}

### 2024-03-03 demystify.tex
Updates from Overleaf

@@
 TA is common in the beginning when you haven't found an adviser who would pay you RA.
+It is also common to sandwich between TA and RA (e.g., when your prof.
+does not have sufficient funding or you want to try get TA experience).
+
+Your TA is paid through the school or department, i.e., they hire you to help teach.
 As a TA, you spend up to 20 hrs/week and help with classes (e.g., grading or teaching labs/recitation).
-Your TA is paid through the school or department, i.e., they hire you to help teach.
-During a semester, a TA might work with several courses and professors (not necessary their adviser).
-TA funding is not typically available during the summer, which has few courses.
-
-\paragraph{How to get TA?}  Unless you have other funding such as RA or Fellowships, TA is typically default for good CS PhD programs.
+During a semester, a TA might work with several courses and professors (not necessary their adviser).  TA funding is not typically available during the summer, which has few courses.
+
+\paragraph{How to get TA?}  Unless you have other funding such as RA or Fellowships, TA is typically default for CS PhD programs.
 When you apply to be a full-time student,  state that you need financial assistant.
@@
 A common scenario is that you first get admitted with TA, and then after a year or two find an adviser to support you with RA.
-See how to contact a prof.
-in \autoref{sec:contact} for research opportunities.
+RA is never guaranteed as it depends on the funding situation of the prof.
+So you should also pay attention to TA, e.g., see if TA is readily available for PhD students in the program.
 
@@
 
-For example, in the 2024 admission cycle, there is a huge surge in students who are interested in AI and NLP (thanks, ChatGPT!).
-However, our AI/NLP faculty cannot take more students and therefore we cannot admit many applicants, many of whom have excellent profiles (e.g., papers in top conferences, strong LoRs, etc.).
+For example, in the 2024 admission cycle, there is a huge surge in students interested in AI and NLP (thanks, ChatGPT!).
+Thus, AI/NLP faculty might be overwhelm with potential students and cannot take additional students, many of whom have excellent profiles.
 

### 2024-03-03 demystify.tex
updates

@@
         \begin{itemize}
-          \item \emph{Yes, definitely}.  CS PhD study in the US is fully funded and admission into good universities is not any harder than non-US schools (\autoref{sec:should}).
+          \item \emph{Yes, definitely}.  CS PhD study in the US is fully funded and admission into good universities is not any harder than in other countries (\autoref{sec:should}).
         \end{itemize}
@@
         \begin{itemize}
-          \item Applications are evaluated by the \emph{PhD Admission} committee and each application is reviewed by typically three faculty (\autoref{sec:evalapps}).
+          \item Applications are evaluated by the \emph{PhD Admission} (adcom) committee and each application is reviewed by typically three faculty (\autoref{sec:evalapps}).
           \item Individually faculty \emph{cannot directly admit} a student---so do not email and ask if you have a chance.
-However, faculty can \emph{advocate} for a student and therefore increase their admission chance---so do contact and introduce yourself (\autoref{sec:contact}).
+However, faculty can \emph{advocate} for a student and therefore increase their admission chance---contact and introduce yourself (\autoref{sec:contact}).
          \end{itemize}
@@
           \item Grades are important, but depend on the reputation of your school (\autoref{sec:grades}).
-          \item Getting an interview is typically a \emph{good sign} as no one wants to interview weak candidates (\autoref{sec:interview}).
+          \item Getting an interview is typically a \emph{good sign}; but no interview does not mean rejection (\autoref{sec:interview}).
+
         \end{itemize}
@@
           \item TA and RA are two main funding sources.
-TA (teaching assistantship) is provided by the department to help profs.
-with classes (e.g., grading).
+TA (teaching assistantship) is provided by the department to help with classes (e.g., grading).
 RA (research assistantship) is provided by profs.
@@
         \begin{itemize}
-          \item Many schools do not offer PhD studies in CS and many CS professors \emph{cannot} formally advise or graduate PhD students  (\autoref{sec:schoolsandprofs}).
+          \item Many schools do not offer PhD studies in CS and many CS professors do not advise or graduate PhD students  (\autoref{sec:schoolsandprofs}).
           \item Contacting a prof. for research opportunities is recommended, but do it \emph{properly} (\autoref{sec:contact}).
         \end{itemize}
+
     \item Miscs and FAQS
         \begin{itemize}
+          \item Why you were rejected (\autoref{sec:why-rejected})
           \item Increasing your admission chance by being unique and standing out (\autoref{sec:improve-your-chance}).
@@
           \item Compared to other countries, CS PhD in the US does not require an MS degree but has longer PhD study time (\autoref{sec:non-us-differences}).
-          \item Your 5--6 year PhD program costs about \$400K in total, but you \emph{do not} pay for it (\autoref{sec:ra-cost}).
+          \item How do you call a professor? (\autoref{sec:address})
+          \item Your 5--6 year PhD program costs about \$400K in total, but you \emph{do not} pay for it (\autoref{sec:ra-cost}
+          \item Don't buy computer equipment and electronics for research, ask your prof. if they can buy them for you (\autoref{sec:buying-equipment}).
           \item Despite some miserable stories on social media, many PhD students have good mentors, supportive lab mates, healthy working environment ... and are happy (\autoref{sec:happy}).
@@
 
-\section{Why did I get rejected?}
+\section{Why did I get rejected?}\label{sec:why-rejected}
 Many possible reasons!  Other than not having a strong profile (e.g., research potentials, GPA, LoRs, SOP), here are some that you might not expect:
@@
 \begin{center}
-  \includegraphics[scale=0.3]{files/c4a.png}
+  \includegraphics[scale=0.5]{files/c4a.png}
 \end{center}

### 2024-03-03 demystify.tex
updates

@@
 that share similar research interests.
-You can learn about faculty and their level of research activity through the faculty's website or even from CSRankings.
 For example, in CSRankings, if you want to work with PL, you can search for those publish in PL conferences.
@@
 Many profs.
-have funding to buy you a computer and other equipments (e.g., keyboards, headphones, tablets, etc).
-Also, do not worry that this is a "waste" of your adviser's money.
-This is part of the funding that your adviser gets to support you and your research.
-So use it to get what you need to do your research (and it's also OK to use them for personal purposes!).
+have funding to buy you a computer and other equipments (e.g., keyboards, headphones, tablets, books, etc).
+
+\section{Can I negotiate my PhD offer?}\label{sec:negotiate}
+For stipend, unlikely.
+TA is set by the university and while RA is set by the adviser, it is often based on the department's standard and it would be unfair to other students if you get more.
+For start date or TA assignment (e.g., TA a specific course), it is uncommon but you can ask for it.
+Also, there is typically no moving allowance for PhD students.
+In short, things that are standard and set by the university or department are unlikely to be negotiable.
+However, you can ask for things such as computer equipments (\autoref{sec:buying-equipment}), books, (and sometimes even furniture) and conference travel budget.
 

### 2024-03-03 demystify.tex
updates

@@
 % It is worth noting that in many cases,  an application could be rejected by most reviewers, however have a "champion", who doesn't even have to be in the Admission comittee, that strongly advocate for your profile.
+
+\section{I am a member of an underrepresented group in CS}\label{sec:urm}
+One question that often comes up is if you should mention, e.g., in SOP, that you are a member of an underrepresented and minority (URM) group.
+This question arises often for international (and even domestic) students.
+The main worry is that (i) they might get discriminated against (as often happen in their countries or personal experiences), and (ii) mentioning this might make them look like they are asking for special treatment.
+
+In my personal opinion, the short answer is \emph{yes}, it can help, especially if you can articulate your experiences and how you can contribute to diversity and inclusion in CS.
+Most universities have a strong commitment to diversity and inclusion and have specific programs to \emph{increase the number of URM students} in their programs.
+Moreover, many universities and industrial organizations (e.g., Google) also have specific scholarships and fellowships for URM students (so you should look for specific scholarships for URM students).
+
+It is completely understandable that some students might feel uncomfortable mentioning this in fear that they lost a chance of working with a prof who might be biased against URM students.
+However, if a prof.
+has such an issue, then you might not want to work with them anyway (and it is likely that the university would not tolerate such behavior).
+In my experience, it is the opposite: many profs.
+are very supportive of URM students and many are actively looking for URM students to work with, and mentioning this in your application can make you stand out.
+It is also worth noting that professors often seek grants from government sponsors such as NSF, which often have broadening participation requirements.
+However, if in doubt, you seek advice from your close professors and collaborators about this.
 
@@
 \section{Can I negotiate my PhD offer?}\label{sec:negotiate}
-For stipend, unlikely.
-TA is set by the university and while RA is set by the adviser, it is often based on the department's standard and it would be unfair to other students if you get more.
+For stipend, unlikely, as TA is set by the university and while RA is set by the adviser, it is often based on the department's standard and it would be unfair to other students if you get more.
+However, your stipend is often increased every year (by a small amount and automatically) and will be more after you pass your qualifying exam.
+
 For start date or TA assignment (e.g., TA a specific course), it is uncommon but you can ask for it.

### 2024-03-03 demystify.tex
updates

--- 2024-03-03 demystify.tex
@@
 % It is worth noting that in many cases,  an application could be rejected by most reviewers, however have a "champion", who doesn't even have to be in the Admission comittee, that strongly advocate for your profile.
+
+\section{I am a member of an underrepresented group in CS}\label{sec:urm}
+One question that often comes up is if you should mention, e.g., in SOP, that you are a member of an underrepresented and minority (URM) group.
+This question arises often for international (and even domestic) students.
+The main worry is that they might get discriminated against, as often happen in their countries or personal experiences.
+%, and (ii) mentioning this might make them look like they are asking for special treatment.
+
+In my personal opinion, the short answer is \emph{yes}, it \emph{does} help, especially if you can articulate your experiences and how you can contribute to diversity and inclusion in CS.
+Most universities have a strong commitment to diversity and inclusion and want to
+\emph{increase the number of URM students} in their programs (e.g., GMU is consistently ranked as one of the most diverse universities in the US).
+Moreover, many universities and industrial organizations also have specific scholarships and fellowships for URM students (so you should look for specific scholarships for URM students).
+
+It is understandable that some students might feel uncomfortable mentioning this in fear that they lost a chance of working with a prof who might be biased against URM students.
+However, if a prof.
+has such an issue, then you might not want to work with them anyway (and it is likely that the university would not tolerate such behavior).
+In my experience, it is the opposite: many profs.
+are actively looking for URM students to work with, and mentioning this in your application can help you stand out.
+
+In addition to being aware of the importance of diversity and inclusion in CS and want to contribute to it, profs.
+have incentives to work with URM students.
+For example, professors often seek grants from government sponsors such as NSF, which strongly encourage broadening participation in STEM (e.g., see \href{https://new.nsf.gov/cise/broadening-participation}{here} and \href{https://www.nsf.gov/pubs/2022/nsf22125/nsf22125.jsp}{here}).
+If in doubt, seek advice from your close professors and collaborators about this.
 

### 2024-03-04 demystify.tex
updates

@@
 
-One main concern for non-CS or non-STEM students is if you have the sufficient technical background, typically obtained through core CS courses.
-You will want to show that you have such knowledge through your coursework, projects, or research.
-For example, if you have taken a course on Algorithms, even online ones like Coursera, you can talk about it in your statement.
-If you have done a project that requires knowledge of OS or have a professional certification (e.g., A+) through work, you can talk about it in your statement.
-If you have done research that requires knowledge of Discrete Maths, you can talk about it in your statement.
+One main concern for non-CS or non-STEM students is if you have the sufficient technical background obtained through core CS courses.
+You need to show that you have such knowledge through your coursework, projects, or research.
+For example, if you have taken a course on Algorithms, even online ones like Coursera, you can talk about it in your SOP.
+If you have done a project that requires knowledge of OS or have a professional certification (e.g., A+) through work, you can talk about it.
+If you have done research that requires knowledge of Discrete Maths, you can talk about it.
 You can also ask your LoR writers to talk about your technical background.
-In summary, in your application, convince us that you have the technical background to do CS PhD research. 
-Note that many online courses teaching about AI does not really help much as these are not core CS knowledge.
+In summary, in your application, convince us that you have the background to do CS PhD research.
 
@@
 \section{Is an MS required for admission to PhD in CS?}\label{sec:msrequirement}
-No. In fact, student with BS can get MS degree ``along the way'' to PhD.  However, MS can help if it gives research experience or is from a more well-known school than your undergrad institution.
+No.
+In fact, student can get MS degree ``along the way'' to PhD, e.g., after finishing course work in the first 2 years.
+However, MS can help admission if it gives research experience or is from a more well-known school than your undergrad institution.
 
@@
 
-\section{How long does it take to complete the CS PhD program?}\label{sec:time}
+\section{How long does to complete the CS PhD program?}\label{sec:time}
 
@@
 
-\paragraph{Standard Tests \autoref{sec:grades}} You do not need to take TOEFL or IELTS because you already did your undergrad (or MS).
-
-\paragraph{Transcripts} You do not need to get your transcripts evaluated by WES or other services (which can be a hassle and inaccurate).
+\paragraph{Standard Tests \autoref{sec:grades}} You do not need to take TOEFL or IELTS because you already did your undergrad (or MS) at a US university.
+You might also be more comfortable communicating in English, e.g., contacting professors (\autoref{sec:contact}).
+
+\paragraph{Transcripts} You do not need to get your transcripts evaluated (which can be a hassle and inaccurate).
 You can just send your official transcripts directly to the school you're applying to.
@@
 For example, you likely have many opportunities to do research as an undergraduates, e.g., through REUs and internships.
-You should highlight these experiences in your application.
-
-\paragraph{Standing out} There are very few domestic applications compared to international ones (i.e., domestic students are minority in the CS PhD application pool).
-That makes you stand out and can really help your case.
+Highlight such experiences in your application.
+
+\paragraph{Standing out} There are very few domestic applications compared to international ones (i.e., domestic students are minority in the CS PhD application pool), and many US universities want to increase the number of domestic students in their programs.
+That makes you stand out and can help your case.
 

### 2024-03-04 demystify.tex
updates

@@
 
-\section{Application Materials}
-
 \paragraph{Fee Waiving} Some school might offer application fee waiver for domestic students.  You should check with the school you're applying to.
@@
 
-\paragraph{Funding \S\ref{sec:funding}} You have more opportunities for funding, e.g., through government scholarships for US citizens and residents.
+\paragraph{Funding \S\ref{sec:funding}} You have more opportunities for funding, e.g., through government scholarships for US citizens and permanent residents.
 You can also apply to specific programs \emph{before} you start your PhD, e.g., NSF Graduate Research Fellowship Program (GRFP) and Hertz Foundation Fellowship.
@@
 
-\paragraph{Standing out} There are very few domestic applications compared to international ones (i.e., domestic students are minority in the CS PhD application pool), and many US universities want to increase the number of domestic students in their programs.
+\paragraph{Standing out} There are \emph{very few} domestic applications compared to international ones (i.e., domestic students are minority in the CS PhD application pool), and many US universities want to increase the number of domestic students in their programs (and as mentioned above, there are specific fellowships for domestic students).
 That makes you stand out and can help your case.

### 2024-03-04 demystify.tex
updates

@@
 
+\paragraph{Follow Up Emails} If you had an interview and have not heard back, you can send an email to ask about the status of your application Don't send this too soon though, perhaps wait at least 2 weeks.
+
+You can also send an email to ask about the status of your application if you have received offers from other schools and need to make a decision soon.
+See \autoref{sec:accept-postpone-decline} for how to accept/postpone/decline offers.
 % \section{Preparing and Tracking Applications}
@@
 
-Students, especially international ones, often get confused on what to do after they get an offer from a professor, e.g., how to postpone or accept/reject offers, and most importantly, how to do so without offending anyone.
+Students, especially international ones, often get confused on what to do after they get an offer from a professor, e.g., is it OK to postpone or accept/reject offers, and most importantly, how to do so without offending anyone.
+Many wonder about accepting an offer and then later rejecting it.
+This section provides some guidance on these issues.
+
 In short, if you are not serious about an offer for any reason (e.g., still waiting for other offers), do not accept it.
-If you are not sure, ask for more time to decide.  But do not accept an offer and then later reject it.  This is considered \emph{rude and unprofessional}, and will likely burn bridges with that prof.
+If you are not sure, ask for more time to decide.
+But do not accept an offer and then later reject it.
+This can be considered \emph{rude and unprofessional}, and will likely burn bridges with that prof.
 

### 2024-03-04 demystify.tex
updates

@@
 
-However, not getting an interview typically means you do not get directly RA support (\autoref{sec:funding}) by an individual faculty (if they want to pay you to do research for them, then likely they want to talk to you first).
-In this case, your application will be reviewed by the adcom and they will make a decision based on your application.
+However, no interviews typically means that you will not likely get an RA (\autoref{sec:funding}), which is offered by an individual faculty (if they want to pay you to do research for them, then likely they want to talk to you first).
+In this case, your application (and TA/fellowship funding) will be reviewed and decided by the adcom.
+
+\paragraph{Follow Up Emails} If you had an interview and have not heard back, you can send an email to ask about the status of your application Don't send this too soon though, perhaps wait at least 2 weeks.
+
+You can also send an email to ask about the status of your application if you have received offers from other schools and need to make a decision soon.
+See \autoref{sec:accept-postpone-decline} for how to accept/postpone/decline offers.
 
@@
 
-In short, you should not worry too much about the timeline. Also do not send emails asking about interviews or status.
+In short, you should not worry too much about the timeline.
+Also do not send emails asking about interviews or status (unless you have interviewed specifically with someone then you can ask that person for status updates and others \autoref{sec:accept-postpone-decline}).
 
@@
 
-\paragraph{Follow Up Emails} If you had an interview and have not heard back, you can send an email to ask about the status of your application Don't send this too soon though, perhaps wait at least 2 weeks.
-
-You can also send an email to ask about the status of your application if you have received offers from other schools and need to make a decision soon.
-See \autoref{sec:accept-postpone-decline} for how to accept/postpone/decline offers.
 % \section{Preparing and Tracking Applications}
@@
 
-Students, especially international ones, often get confused on what to do after they get an offer from a professor, e.g., is it OK to postpone or accept/reject offers, and most importantly, how to do so without offending anyone.
-Many wonder about accepting an offer and then later rejecting it.
-This section provides some guidance on these issues.
+Students, especially international ones, often get confused on what to do after they get an offer from a professor, e.g., is it OK to postpone or accept/reject offers?, and most importantly, how to do so without offending anyone.
+Many also wonder about accepting an offer and then later rejecting it.
 
@@
 But do not ask for too much time, e.g., more than a month.
-You also should not postpone the offers multiple times, which will annoy people.
+You also should not postpone the offer multiple times, which will annoy people.
 

### 2024-03-04 demystify.tex
updates

@@
 
-\paragraph{Why CS depts do not waive application fee?}  Note that some universities do waive application fee, however most do not.
+\paragraph{Why CS depts do not waive application fee?}  Note that some universities do waive application fee (e.g., for domestic students (\autoref{sec:domestic-students})), however most do not.
 In any case, application fee is typically a requirement of the university.
@@
 Important factors to consider include the reputation of schools and professors (\autoref{sec:schoolsandprofs}), and funding availability (\autoref{sec:funding}).
-You will have to make your decision by around \emph{April 15} (see \href{https://cgsnet.org/wp-content/uploads/2024/01/CGS_April15_Resolution_Jan312024.pdf}{why?}).
+You will have to make your decision by around \emph{April 15} (\href{https://cgsnet.org/wp-content/uploads/2024/01/CGS_April15_Resolution_Jan312024.pdf}{why this date?}).
 
@@
 
-\paragraph{Low Stipend?} Students often complain their stipend being too low, but it is actually not bad and you can live comfortably with it.
-In most cases, it is also enough to support your spouse and kids (many CS PhD students have their family with them).
+\paragraph{Low Stipend?} Students often complain their stipend being too low.
+However, in most cases it is actually not bad and you can live comfortably with it.
+It might also be enough to support your spouse and kids (many CS PhD students have their family with them).
 So don't worry too much about stipend.
@@
 
-\textbf{How to get RA?} When a professor recruits you, they might offer you an RA right away (e.g., when you apply).
-A common scenario is that you first get admitted with TA, and then after a year or two find an adviser to support you with RA.
-RA is never guaranteed as it depends on the funding situation of the prof.
-So you should also pay attention to TA, e.g., see if TA is readily available for PhD students in the program.
+\textbf{How to get RA?} When a professor recruits you, they might offer you an RA right away (so you start with an RA).
+However, a more common scenario is that you first get admitted with TA, and then after a year or two find an adviser to support you with RA.
+
+It is important to note that RA is never guaranteed as it depends on the funding situation of the prof.
+So you should also pay attention to TA, which is a good backup plan (remember, typically TA and RA have similar benefits).
+This means you should also ask if TA is readily available for PhD students in the program.
 
@@
 Fellowships are typically competitive and generous, and gives pretty much all benefits tuition/insurance that a TA/RA has.
-Moreover, they often give higher stipend (including summer) and open doors for job opportunities (e.g., internship).
+They might even give higher stipend (including summer) and open doors for job opportunities such as internships.
 
@@
 
-You likely need to apply for external fellowships including those from the US government and tech companies (e.g., Google, Microsoft, Facebook).
+For external fellowships including those from the US government and tech companies (e.g., Google, Microsoft, Facebook), you will need to apply.
 Fellowships from US government would likely require US citizenship or residency.
@@
 
-\subsection{Selecting and Ranking Schools for CS}\label{sec:selecting-ranking-schools}
+\subsection{Selecting and Ranking Schools}\label{sec:selecting-ranking-schools}
 \begin{center}
@@
 
-International students not familiar with US universities often put them into \emph{two} bins:  (i) very top schools that they dream about, and (ii) everything else.
-In many cases, they use resources such as rankings from US News, which are not very transparent and \emph{highly} questionable\footnote{\url{https://cra.org/cra-statement-us-news-world-report-rankings-computer-science-universities/}}.
+International\footnote{Even domestic students who have done their undergrad in the US might not know much about CS PhD programs in the US.} students often put schools into two bins: (i) very top schools that they dream about, and (ii) everything else.
+In many cases, they use rankings from US News, which are not very transparent and highly questionable\footnote{\url{https://cra.org/cra-statement-us-news-world-report-rankings-computer-science-universities/}}.
 Sometimes these students evaluate CS programs using the reputations non-CS programs such as medical, math, or physics.
@@
 However, in general, rankings can be superficial and you need to do more research to be informed and make better decision.
-For example, if you get admissions to several places, you should consider attending Open Houses and contact profs.
+For example, if you get admissions to several places, you should consider attending Open Houses (\autoref{sec:accepted}) and contact profs.
 that you're interested in at those place and talk to them.
@@
 \section{Choosing an Adviser}
-Obviously, there is no one-size-fit-all answer to this question.
-The best adviser is the one that you can work well with.
-This is difficult and takes time.
+Obviously, there is no one-size-fit-all answer to this question. The best adviser is the one that you can work well with.  But how do you find such a person?
+
 Fortunately, unlike many countries that require finding an adviser and research topic before starting the PhD (\autoref{sec:non-us-differences}), CS PhD programs in the US will typically give you a couple of years to search for advisers and research topics.
@@
 Here are some general advice to find professor.
-Assuming you're not familiar with any particular one, then first, search for profs.
+Assuming you're not familiar with any particular one, then first search for profs.
 that share similar research interests.

### 2024-03-07 demystify.tex
updates

@@
           \item Why you were rejected (\autoref{sec:why-rejected})
-          \item Increasing your admission chance by being unique and standing out (\autoref{sec:improve-your-chance}).
+          \item Increasing admission chance by being unique and standing out (\autoref{sec:improve-your-chance}).
+          \item Being in the underepresentative group in CS can help (\autoref{sec:urm}).
           \item You can successfully apply to CS PhD even if you have non-STEM background (\autoref{sec:non-stem}).
+          \item You do not need an MS degree (\autoref{sec:msrequirement}) to apply for PhD in CS, and it takes about 5--7 years for PhD in CS (\autoref{sec:time}).
           \item Compared to other countries, CS PhD in the US does not require an MS degree but has longer PhD study time (\autoref{sec:non-us-differences}).
           \item How do you call a professor? (\autoref{sec:address})
-          \item Your 5--6 year PhD program costs about \$400K in total, but you \emph{do not} pay for it (\autoref{sec:ra-cost}
-          \item Don't buy computer equipment and electronics for research, ask your prof. if they can buy them for you (\autoref{sec:buying-equipment}).
+          \item Your entire PhD program costs about \$400K in total, but you \emph{do not} pay for it (\autoref{sec:ra-cost})
+          \item Ask your prof. if they can buy computer equipments and such for your research (\autoref{sec:buying-equipment}).
           \item Despite some miserable stories on social media, many PhD students have good mentors, supportive lab mates, healthy working environment ... and are happy (\autoref{sec:happy}).
@@
 Given the high number of quality applicants and limited number of spots, other than having a good application profile, you really want to show something that makes you \textbf{stand out}, e.g., do you have a degree or background in \emph{dance} or \emph{music} and want to integrate them with CS?
-are you a female or a minority in CS (research for "URM minority in CS in the US" on Google)?
+are you a female or a minority in CS (research for "URM minority in CS in the US" on Google) (\autoref{sec:urm})?
 Do you participate in outreach activities that help increase diversity and inclusion in CS?
@@
 Most universities have a strong commitment to diversity and inclusion and want to
-\emph{increase the number of URM students} in their programs (e.g., GMU is consistently ranked as one of the most diverse universities in the US).
+\emph{increase the number of URM students} in their programs (e.g., GMU often advertises about being one of the most diverse universities in the US).
 Moreover, many universities and industrial organizations also have specific scholarships and fellowships for URM students (so you should look for specific scholarships for URM students).
@@
 
-In addition to being aware of the importance of diversity and inclusion in CS and want to contribute to it, profs.
+In addition to being aware of the importance of diversity and inclusion and want to contribute to it, profs.
 have incentives to work with URM students.
 For example, professors often seek grants from government sponsors such as NSF, which strongly encourage broadening participation in STEM (e.g., see \href{https://new.nsf.gov/cise/broadening-participation}{here} and \href{https://www.nsf.gov/pubs/2022/nsf22125/nsf22125.jsp}{here}).
+Prof.
+can also apply for grants from organizations and private donors, which have specific programs for URM students.
+Thus, having URM students in their group can help them get these grants.
+
+In the end, this is your choice and you do what make you feel comfortable.
+However, just be aware that URM is often a (huge) plus, not a minus, in CS in the US.
 If in doubt, seek advice from your close professors and collaborators about this.
@@
     \midrule
-    GRA (9-month) & 27K & \\
-    GRA (summer)  &9K	& 3-month, 20hrs/week\\
-    \textbf{Total Salary} &\red{36K}	&\\
+    GRA (9-month) & 29K & \\
+    GRA (summer)  &10	& 3-month, 20hrs/week\\
+    \textbf{Total Salary} &\red{39K}	&\\
     \midrule
@@
     \midrule
-    Total Direct Cost	& \red{57K}	&Salary + Travel + Health + Tuition \\
+    Total Direct Cost	& \red{60}	&Salary + Travel + Health + Tuition \\
     F \& A (MTDC)	& 21K	& Direct Cost - GRA Salary\\
     Total Indirect Cost	& \red{12K}	&58.9\% of MTDC\\
-    \textbf{Total (Direct + Indirect)} &	\red{69K}	& Budget for a GRA\\
+    \textbf{Total (Direct + Indirect)} &	\red{72K}	& Budget for a GRA\\
     \bottomrule
@@
 Private universities may have different numbers.
-For simplicity, in this table I will assume the department has a 9-month stipend of \$27000 (GMU actually pays more) and therefore a 3-month summer of \$9000.
-I will also use GMU tuition rate of about \$15,000/year for full-time study (which is quite cheap compared to private universities, e.g., MIT charges around \$50K) and a 58.9\% rate on \emph{indirect cost}, which is what GMU charges for overhead or administrative costs (yes, after all, universities are businesses!).
+For simplicity, I will assume the department has a 9-month stipend of \$29000 and a 3-month summer of \$10000 (about a third of the 9-month stipend).
+I will also use GMU tuition rate of about \$15,000/year for full-time study (which is quite cheap compared to private universities, e.g., MIT charges around \$50K) and a 58.9\% rate on \emph{indirect cost}, which is a typical rate charged for overhead or administrative costs (yes, after all, universities are businesses!).
 Finally, I assume the student take two conference trips per year, one domestic and one international (conf.
@@
 
-At the end, the total budget comes out to be \$69K/year to support a PhD student.
-\emph{The summary is that over your 5-6 year of your PhD, you cost about \$350K--400K, and while you're paid X, your adviser probably pays 2X for you}.
+At the end, the total budget comes out to be \$72K/year to support a PhD student.
+\emph{The summary is that over your 5--6 years of your PhD, you cost about \$400K, and while your stipend is X, your adviser probably pays 2X for you}.
 
@@
 
-\section{Buying Computer Equipment}\label{sec:buying-equipment} Many students, including myself, get very excited about getting to buy new computer equipment and electronics when they start their PhD.
+\section{Buying Computer Equipment}\label{sec:buying-equipment} Many students get very excited about getting to buy new computer equipment and electronics when they start their PhD.
 However, you should check with your adviser first.
@@
 \section{Can I negotiate my PhD offer?}\label{sec:negotiate}
-For stipend, unlikely, as TA is set by the university and while RA is set by the adviser, it is often based on the department's standard and it would be unfair to other students if you get more.
-However, your stipend is often increased every year (by a small amount and automatically) and will be more after you pass your qualifying exam.
-
-For start date or TA assignment (e.g., TA a specific course), it is uncommon but you can ask for it.
+For stipend, unlikely, as TA is set by the university and while RA is set by the adviser, it is often based on the department's standard, and it also would be unfair to other students if you get more.
+However, your stipend is often increased every year (by a small amount and automatically) and will likely increase after you pass your qualifying exam.
+
+For start date or TA assignment (e.g., TA'ing a specific course), it is uncommon but you can ask for it.
 Also, there is typically no moving allowance for PhD students.
@@
 Most of what is written in this document is applicable to both domestic\footnote{Domestic simply means you did your undergrad (or MS) at a US university.} and international students.
-However, there are some differences that domestic students should be aware of.
-This section lists some of these differences.
+However, there are some differences and benefits that domestic students should be aware of.
+This section lists some of these.
 
@@
 
-\paragraph{Standing out} There are \emph{very few} domestic applications compared to international ones (i.e., domestic students are minority in the CS PhD application pool), and many US universities want to increase the number of domestic students in their programs (and as mentioned above, there are specific fellowships for domestic students).
+\paragraph{Standing out} There are \emph{very few} domestic applications compared to international ones, i.e., domestic students are minority (\autoref{sec:urm}) in the CS PhD application pool.
+Many US universities want to increase the number of domestic students in their programs (and as mentioned above, there are specific fellowships for domestic students).
 That makes you stand out and can help your case.
+
+\paragraph{Open House} (\autoref{sec:accepted}) It is easier for you to attend open house events in person.  This can help you make a better decision on which school to attend.
 

### 2024-03-07 demystify.tex
updates

@@
 
-It is understandable that some students might feel uncomfortable mentioning this in fear that they lost a chance of working with a prof who might be biased against URM students.
-However, if a prof.
-has such an issue, then you might not want to work with them anyway (and it is likely that the university would not tolerate such behavior).
+It is understandable that some students might feel uncomfortable mentioning this in fear that they lost a chance of working with a faculty who might be biased against URM students.
+However, if someone has such an issue, then you might not want to work with them anyway (and it is likely that the university would not tolerate such behavior).
 In my experience, it is the opposite: many profs.
@@
 have incentives to work with URM students.
-For example, professors often seek grants from government sponsors such as NSF, which strongly encourage broadening participation in STEM (e.g., see \href{https://new.nsf.gov/cise/broadening-participation}{here} and \href{https://www.nsf.gov/pubs/2022/nsf22125/nsf22125.jsp}{here}).
-Prof.
-can also apply for grants from organizations and private donors, which have specific programs for URM students.
-Thus, having URM students in their group can help them get these grants.
+For example, we often seek grants from government sponsors such as NSF, which strongly encourage broadening participation in STEM (e.g., see \href{https://new.nsf.gov/cise/broadening-participation}{here} and \href{https://www.nsf.gov/pubs/2022/nsf22125/nsf22125.jsp}{here}).
+Also, sometimes providing this information can make it easier for faculty to apply for certain scholarships or fellowships to support you, e.g., from a private donor that has offered a fellowship only to students from certain background.
 

### 2024-03-07 demystify.tex
updates

@@
 
-\section{I am a member of an underrepresented group in CS}\label{sec:urm}
-One question that often comes up is if you should mention, e.g., in SOP, that you are a member of an underrepresented and minority (URM) group.
-This question arises often for international (and even domestic) students.
+\section{Member of an underrepresented or LBGTQ+ groups}\label{sec:urm}
+A common question is whether you should mention, e.g., in SOP, that you are a member of an underrepresented and minority (URM) group\footnote{URM in CS in the US include African Americans, Hispanic Americans, American Indians/Native Alaskans, Native Hawaiians/Pacific Islanders (excluding Asian Americans), and multiracial students identifying with at least one of the previously listed URM categories.} and LBGTQ+ group.
+This is a strong concern for international (and even domestic) students, who may have encountered discrimination in their countries or personal experiences.
+%This question arises often for international (and even domestic) students.
 The main worry is that they might get discriminated against, as often happen in their countries or personal experiences.
@@
 
-In my personal opinion, the short answer is \emph{yes}, it \emph{does} help, especially if you can articulate your experiences and how you can contribute to diversity and inclusion in CS.
-Most universities have a strong commitment to diversity and inclusion and want to
-\emph{increase the number of URM students} in their programs (e.g., GMU often advertises about being one of the most diverse universities in the US).
-Moreover, many universities and industrial organizations also have specific scholarships and fellowships for URM students (so you should look for specific scholarships for URM students).
-
-It is understandable that some students might feel uncomfortable mentioning this in fear that they lost a chance of working with a faculty who might be biased against URM students.
-However, if someone has such an issue, then you might not want to work with them anyway (and it is likely that the university would not tolerate such behavior).
-In my experience, it is the opposite: many profs.
-are actively looking for URM students to work with, and mentioning this in your application can help you stand out.
-
-In addition to being aware of the importance of diversity and inclusion and want to contribute to it, profs.
-have incentives to work with URM students.
-For example, we often seek grants from government sponsors such as NSF, which strongly encourage broadening participation in STEM (e.g., see \href{https://new.nsf.gov/cise/broadening-participation}{here} and \href{https://www.nsf.gov/pubs/2022/nsf22125/nsf22125.jsp}{here}).
+In my personal opinion, highlighting URM or LBGTQ+ identity can \emph{indeed be beneficial}, especially if you can articulate how your diverse experiences contribute to diversity and inclusion in academic.
+Many universities have a strong commitment to diversity and inclusion and actively
+\emph{recruit students from underepresentative backgrounds} (e.g., some \href{https://cse-climate.engin.umich.edu/reports/climate-dei-reports/cse-climate-and-dei-report-2022-2023/#grad-ethnicity}{stats from UMich CSE} and \href{https://www.gmu.edu/news/2022-09/mason-now-top-10-public-university-diversity-innovation-and-cybersecurity-education-us}{GMU} often promotes itself as one of the most diverse universities in the US).
+Moreover, many scholarships and fellowships specifically for URM and LBGTQ+ students, which applicants should explore if qualify.
+
+It is understandable that some applicants might feel uncomfortable disclosing this, fearing they would lost a chance of working with a faculty with bias.
+However, if someone has such an issue, then you might not want to work with them anyway (and such bias is not acceptable in academic and likely would not be tolerated by the university).
+In my experience, it is the opposite: \emph{many profs.
+actively seek out to work with students from diverse backgrounds and view diversity as an asset to their research group}.
+
+Moreover, promoting diversity aligns with the broadening participation in computing (BPC) goal within academia (e.g., see the \href{https://plans.bpcnet.org/GeorgeMasonUniversity_ComputerScience_DepartmentalBPCPlan.pdf}{BPC plan at GMU CS}) and government funding agencies such as NSF (e.g., see \href{https://new.nsf.gov/cise/broadening-participation}{here} and \href{https://www.nsf.gov/pubs/2022/nsf22125/nsf22125.jsp}{here}).
 Also, sometimes providing this information can make it easier for faculty to apply for certain scholarships or fellowships to support you, e.g., from a private donor that has offered a fellowship only to students from certain background.
 
-In the end, this is your choice and you do what make you feel comfortable.
-However, just be aware that URM is often a (huge) plus, not a minus, in CS in the US.
-If in doubt, seek advice from your close professors and collaborators about this.
+Ultimately, this is a personal choice and you do what make you feel comfortable.
+Seeking guidance from your close professors and collaborators is a good idea.
+However, it is important to recognize that diversity is strongly valued and celebrated within CS programs in the US, and that you should feel empowered to share your unique experiences and perspectives.
 

### 2024-03-07 demystify.tex
updates

@@
 \emph{recruit students from underepresentative backgrounds} (e.g., some \href{https://cse-climate.engin.umich.edu/reports/climate-dei-reports/cse-climate-and-dei-report-2022-2023/#grad-ethnicity}{stats from UMich CSE} and \href{https://www.gmu.edu/news/2022-09/mason-now-top-10-public-university-diversity-innovation-and-cybersecurity-education-us}{GMU} often promotes itself as one of the most diverse universities in the US).
-Moreover, many scholarships and fellowships specifically for URM and LBGTQ+ students, which applicants should explore if qualify.
-
-It is understandable that some applicants might feel uncomfortable disclosing this, fearing they would lost a chance of working with a faculty with bias.
-However, if someone has such an issue, then you might not want to work with them anyway (and such bias is not acceptable in academic and likely would not be tolerated by the university).
+Moreover, many scholarships and fellowships are created specifically for URM and LBGTQ+ students, which you should explore if qualify.
+
+It is understandable that you might feel uncomfortable disclosing this, fearing you would lost a chance of working with a faculty who might have a bias.
+However, if someone has this issue, then you might not want to work with them anyway (and such bias is not acceptable in academic and likely would not be tolerated by the university).
 In my experience, it is the opposite: \emph{many profs.
@@
 
-Moreover, promoting diversity aligns with the broadening participation in computing (BPC) goal within academia (e.g., see the \href{https://plans.bpcnet.org/GeorgeMasonUniversity_ComputerScience_DepartmentalBPCPlan.pdf}{BPC plan at GMU CS}) and government funding agencies such as NSF (e.g., see \href{https://new.nsf.gov/cise/broadening-participation}{here} and \href{https://www.nsf.gov/pubs/2022/nsf22125/nsf22125.jsp}{here}).
+Moreover, promoting diversity aligns with the goal of broadening participation in computing (BPC) that is strongly supported within academia (e.g., see the \href{https://plans.bpcnet.org/GeorgeMasonUniversity_ComputerScience_DepartmentalBPCPlan.pdf}{BPC plan at GMU CS}) and government funding agencies such as NSF (e.g., see \href{https://new.nsf.gov/cise/broadening-participation}{here} and \href{https://www.nsf.gov/pubs/2022/nsf22125/nsf22125.jsp}{here}).
 Also, sometimes providing this information can make it easier for faculty to apply for certain scholarships or fellowships to support you, e.g., from a private donor that has offered a fellowship only to students from certain background.
 
-Ultimately, this is a personal choice and you do what make you feel comfortable.
+Ultimately, this is a personal choice and you do what you feel comfortable.
 Seeking guidance from your close professors and collaborators is a good idea.

### 2024-03-09 demystify.tex
updates

@@
 
-For example, in the 2024 admission cycle, there is a huge surge in students interested in AI and NLP (thanks, ChatGPT!).
-Thus, AI/NLP faculty might be overwhelm with potential students and cannot take additional students, many of whom have excellent profiles.
+For example, in the 2024 admission cycle, there is a huge surge in students interested in AI and NLP (thanks, ChatGPT!). Thus, AI/NLP faculty might be 
+overwhelmed and cannot consider many, even those with excellent profiles.
 
@@
 For example, if you have an open-source project on Github that is used by many people, has lots of stars in Github, do talk about it.
-If you write technical, research-like blogs, talk about them too.
+If you often write technical, research-like blogs with many viewers, talk about them too.
 
@@
 
-It is understandable that you might feel uncomfortable disclosing this, fearing you would lost a chance of working with a faculty who might have a bias.
+It is understandable that you might feel uncomfortable disclosing this, fearing you would lost a chance of working with a faculty due to their bias.
 However, if someone has this issue, then you might not want to work with them anyway (and such bias is not acceptable in academic and likely would not be tolerated by the university).
@@
 However, it is important to recognize that diversity is strongly valued and celebrated within CS programs in the US, and that you should feel empowered to share your unique experiences and perspectives.
+
+%https://www.edsurge.com/news/2018-05-18-how-cornell-university-diversified-its-incoming-phd-computer-science-student-body
+
+%https://academia.stackexchange.com/questions/130591/is-it-standard-for-us-based-universities-to-consider-the-ethnicity-of-an-applica
+
+%https://www.cs.cornell.edu/~bindel/paper/diversity.pdf
+
+%https://cs.brown.edu/degrees/doctoral/applications/helpful-resources-applying-computer-science-phd-programs/
 
@@
 
-\section{How much do \emph{you} cost?}\label{sec:ra-cost}
+\section{How much do \emph{I} cost?}\label{sec:ra-cost}
 PhD students often ask why their salary is so low compared to ludicrous grants their advisers get.
 They also wonder why their offer letters sometimes show that their benefits higher than what they actually receive (i.e., stipend).
-This section aims to shed some light to these questions.
 
@@
 For simplicity, I will assume the department has a 9-month stipend of \$29000 and a 3-month summer of \$10000 (about a third of the 9-month stipend).
-I will also use GMU tuition rate of about \$15,000/year for full-time study (which is quite cheap compared to private universities, e.g., MIT charges around \$50K) and a 58.9\% rate on \emph{indirect cost}, which is a typical rate charged for overhead or administrative costs (yes, after all, universities are businesses!).
+I will also use GMU tuition rate of about \$15,000/year for full-time study (which is quite cheap compared to private universities, e.g., Univ.
+of Chicago is a whooping \$70K) and a 58.9\% rate on \emph{indirect cost}, which is a typical rate charged for overhead or administrative costs (yes, after all, universities are businesses!).
 Finally, I assume the student take two conference trips per year, one domestic and one international (conf.
@@
 Many profs.
-have funding to buy you a computer and other equipments (e.g., keyboards, headphones, tablets, books, etc).
+have funding to buy you computers and other equipments (e.g., software, books, keyboards, headphones, tablets, etc).
 
@@
 For stipend, unlikely, as TA is set by the university and while RA is set by the adviser, it is often based on the department's standard, and it also would be unfair to other students if you get more.
-However, your stipend is often increased every year (by a small amount and automatically) and will likely increase after you pass your qualifying exam.
+However, your stipend is often increased every year (automatically by a small amount) and also after you pass your qualifying exam.
 
@@
 Also, there is typically no moving allowance for PhD students.
-In short, things that are standard and set by the university or department are unlikely to be negotiable.
+In short, standard things set by the university or department are unlikely non-negotiable.
 However, you can ask for things such as computer equipments (\autoref{sec:buying-equipment}), books, (and sometimes even furniture) and conference travel budget.
@@
 
-In short, if you are not serious about an offer for any reason (e.g., still waiting for other offers), do not accept it.
+In short, if you are not serious about an offer for any reason (e.g., still waiting for other offers), \emph{do not accept it}.
 If you are not sure, ask for more time to decide.
@@
 I appreciate your time and consideration.
-I hope we can work together in the future.''}  Professors will understand and wish you good luck.
+I hope we can work together in the future.''}  Professors will understand and wish you luck.
 If you decide to reject an offer, do so quickly.
@@
 Plagiarism and cheating (e.g., exams, assignments) is a BIG no-no in the US.
-If you're caught cheating, you will face very heavy consequence and likely be expelled from the university (e.g., after the second time at GMU).
+If you're caught cheating, you will face serious consequence and likely be expelled from the university (e.g., after the second time at GMU).
 This is quite different from many international countries where cheating is common and often tolerated.
@@
 Imagine your department or adviser being sued for a large sum of money, and it is \emph{you} who caused it.
-If you need to purchase software,  ask your adviser or the department.
+If you need to purchase software,  ask your adviser or the department (\autoref{sec:buying-equipment}).
 

### 2024-03-09 demystify.tex
updates

@@
 
-The most concrete evidence of research ability is having \textbf{papers in reputable international journals or conferences}.
+A concrete evidence of research ability is having \textbf{papers in reputable international journals or conferences}.
 Having published papers, especially at top venues, is a sign that the applicant has successfully involved in research.
@@
 Most PhD programs will require at least \textbf{two LORs}.
-Having a strong letter from an internationally recognized researcher will greatly strengthen your application.
-In fact, a strong letter from a well-known researcher can outweigh the lack of publications.
-
-However, obtaining such letters
+Having a strong letter from an internationally recognized researcher will \emph{greatly strengthen} your application.
+In fact, a strong letter from a well-known researcher can outweigh the lack of publications (or low GPA or unknown school).
+
+However, obtaining letters from famous people 
 can be challenging for international students, who might not have much interactions with such experts.
@@
 
-\paragraph{Generic Letters} When the writers do not know much about the applicants (e.g., just taking some course with them or not making any impression to write about), they might write a generic letter, which is not useful and also considered weak.
+\paragraph{Generic Letters} When the writers do not know much about the applicants (e.g., just taking some course with them or not making any impression to write about), they might write a generic and short letter, which is not useful and also considered weak.
 It might be a good idea to directly ask if the prof.
@@
   \textbf{Vu}: This is an interesting and useful detail that US faculty might not be aware of and students should mention about this in their statements.
-In general, I think it should be fine as long as that person can properly evaluate your research ability.
-\end{commentbox}
-
-\paragraph{Reminding Writers} After you submit your application, you should tell your writers about that and remind them to submit their letters if they haven't done so.
-But don't send too many reminders as that can be annoying to the writers.
+In general, someone with a PhD has been through the research process and therefore can better evaluate your research ability.
+But if you do not have such people, then someone who can properly evaluate your research ability is still OK.
+\end{commentbox}
+
+\paragraph{Reminding Writers} After you submit your application, you should tell your writers about that and let them know they will soon receive an email from the university to submit their letters.
+You should also remind them about the deadline and ask them to submit their letters on time.
+But do not send too many reminders as that can be annoying to the writers.
 
@@
 VEF2.0 follows an interesting model where US faculty members from leading institutions are invited to conduct rigorous interviews with VEF students and subsequently provide reference letters on their behalf.
-Despite the limited interaction between the interviewer and the interviewee (primarily confined to the interview itself), these reference letters are generally effective as they have helped many students getting into top PhD programs in the US.
+Despite the limited interaction between the interviewer and the interviewee (primarily confined to the interview itself), these reference letters are generally effective and have helped many students getting into top PhD programs in the US.
 
@@
 
-While you might not have control over LORs or \hyperref[sec:your-school]{where your go to school}, you do over your
+While you might not have control over LORs (\autoref{sec:lor}) or where your go to school (\autoref{sec:your-school}), you do over your
 statement of purpose (SOP) or personal statement\footnote{Some schools separate these documents and ask you to write both: SOP, which focuses on research experiences, and Personal Statement, which is everything more personal, e.g., why PhD, challenges, diversity, etc}!
@@
 
-You should also talk about things that adcom members might not know about and can help make you different than the other 1000 applicants, e.g.,  your personal Github project with 1K+ stars or your regular contributions to well-known open-source projects (see \autoref{sec:improve-your-chance} for increasing your admission chance).
+You should also talk about things that adcom members might not know about and can help make you different than the other 1000 applicants, e.g., you are member of a minority or LBGTQ+ group (\autoref{sec:urm}), your personal Github project with 1K+ stars or your regular contributions to well-known open-source projects (see \autoref{sec:improve-your-chance} for increasing your admission chance).
 
@@
 Typically, an interview takes about 15--30 minutes, and one important aspect of evaluation is your ability to effectively communicate, including speaking and understanding English. 
+In general treat it as an informal chat, but be prepared to talk about your research experiences and interests (so do have \emph{elevator pitch} about your experiences and research). 
 A professor might want to know more about your research background or even ask you to read a paper (so be sure to do it!).
@@
 Some schools send out admission letters in batches, some do rolling admission, and some do not send anything out (e.g., you're on their waitlist, or they might not send rejection letters).
-However, in general, you should hear back from most schools by around March.
+However, in general, you should hear back from most schools by around March (but some schools might take longer!).
 
@@
 Important factors to consider include the reputation of schools and professors (\autoref{sec:schoolsandprofs}), and funding availability (\autoref{sec:funding}).
-You will have to make your decision by around \emph{April 15} (\href{https://cgsnet.org/wp-content/uploads/2024/01/CGS_April15_Resolution_Jan312024.pdf}{why this date?}).
+You will likely have to make your decision by around \emph{April 15} (\href{https://cgsnet.org/wp-content/uploads/2024/01/CGS_April15_Resolution_Jan312024.pdf}{why this date?}).
 
@@
 This is also a good way to learn about individual faculty (who works on what) and well-known CS conferences\footnote{In CS (and probably only in CS), conferences, not journals, are often the main venue to publish research finding.}.
-\autoref{sec:ranking} gives the top 50 CS programs in the US according to CSRankings.
+%\autoref{sec:ranking} gives the top 50 CS programs in the US according to CSRankings.
 
@@
 
-However, in general, rankings can be superficial and you need to do more research to be informed and make better decision.
+While I find CSrankings useful, it is not perfect, and in general, rankings are superficial and you need to do more research to be informed and make better decision.
 For example, if you get admissions to several places, you should consider attending Open Houses (\autoref{sec:accepted}) and contact profs.
@@
 Also, in the US you also have to pass a series of "exams", e.g., qualifying exam, comprehensive exam, thesis proposal defense\footnote{The word "ABD" (all but dissertation) is used in the US to refer to a PhD candidate who have finished all course work and exams and only need to write and defend their dissertation.}.
-In other countries, you do not have to do these exams or only do a few of them.
+In other countries, you might not have these exams or only have to do a few of them.
 
 \emph{Funding}:  In many countries stipend comes from the university or from the gov't.
-These funding often have a fixed duration (e.g., 3 years).
-In the US, stipend (e.g., RA) comes directly from your adviser (no fixed duration).
+These funding often have a fixed duration (e.g., 3 or 4 years).
+In the US (\autoref{sec:funding}), stipend (e.g., RA) comes directly from your adviser (no fixed duration).
 There are also fewer TA opportunities in the European universities compared to the US.
@@
 This is partly due to the longer PhD program and that US PhD students are often paid through TA, which requires them to do TA in addition to their own research.
-In contrast, PhD students in other countries are often paid through fellowships, which do not require them to do TA.
+In contrast, PhD students in other countries are often paid through fellowships, which might not require doing TA.
 

### 2024-03-09 demystify.tex
updates

@@
 
-The primary focus of the admissions committee is to \textbf{evaluate your background and interest in research} (a PhD in Computer Science is a research degree!), and determine if you would \textbf{fit into} the program.
+The focus of the admissions committee is to \textbf{evaluate your background and interest in research} to see if you can handle a PhD in CS, which is a \emph{research} degree!.
+We also aim to determine if you would \textbf{fit into the program}.
 To evaluate your profile, we consider
@@
 However, they might not push you as hard.
-They might also expect you to figure things out yourself, i.e., so you need to be independent.
+They might also expect you to figure things out yourself, i.e., you need to be independent.
 Some tenured faculty are also no longer active in research and more involved with administrative responsibilities or with their startup companies (this means they will likely not take new students).
@@
 
-In many cases you do not need to do a PhD in CS to work in your area of interested.
-For example, in addition to a traditional CS department, GMU has IST and Cybersecurity departments, which have faculty  with PhD in CS and work on CS topics (e.g., AI, Security, Robotics).
+In many cases you do not need to do a PhD in CS to work in your area of interest.
+For example, in addition to a traditional CS department, GMU has IST and Cybersecurity departments, both of which have faculty  with PhD in CS and work on CS topics (e.g., AI, Security, Robotics).
 So it is totally possible that you still get to do CS research and publish in CS-focused venues even if you're not in a traditional CS program.
@@
 However, if your goal is a PhD in CS, then you likely need to be in the CS dept \emph{and} advised by a CS faculty.
-A non-CS may \emph{co-advise} or be in the PhD dissertation committee, but your main PhD adviser will likely be a tenured or tenure-track faculty in CS (\autoref{sec:faculty-types}).
+A non-CS faculty may \emph{co-advise} or be in the PhD dissertation committee, but your main PhD adviser will likely be a tenured or tenure-track faculty in CS (\autoref{sec:faculty-types}).
 If in doubt, check with the CS department for their requirements.
@@
 
-Most of what is written in this document is applicable to both domestic\footnote{Domestic simply means you did your undergrad (or MS) at a US university.} and international students.
+Most of what is written in this document is applicable to both domestic\footnote{Here domestic means you did your undergrad (or MS) at a US university.} and international students.
 However, there are some differences and benefits that domestic students should be aware of.
@@
 
-\paragraph{School \autoref{sec:your-school}} we already know about your school, which is a plus (see \autoref{sec:your-school}).
+\paragraph{School \autoref{sec:your-school}} We already know about your school, which is a plus (see \autoref{sec:your-school}).
 You are also more familiar with the US education system and culture (so hopefully you are aware of the cultural issues listed in \autoref{sec:cultural}).
@@
 % \end{domesticbox}
+
+\chapter{Research Opportunities for Undergraduates}\label{sec:undergrad-research}
+
+As discussed in \autoref{sec:undergrad-research}, research experience is a crucial component of a strong PhD application.
+It gives you opportunities to try out research, determine what research area you're interested in, publish papers (\autoref{sec:research-ability}), and get strong LoRs (\autoref{sec:lor}).
+
+ This section provides some guidance on how to gain research experience as an undergraduate student.
+It also gives advice to international students or domestic ones from small colleges where research opportunities are limited.
+
+\paragraph{Locally} The most common way to find research is right at your own institution.  
+If you did well or like a class, you can check with the professor of that class for research opportunities.  
+You can also go to the department directory and then professors' websites and see if they are looking for undergraduate researchers.
+Even if they say they are looking for graduate students, you still should contact them and ask.
+
+You can also take honor thesis or independent study courses with a professor.
+This is a good way to get research experience and also get a LoR from the professor.
+You can also ask your academic adviser or other faculty members for suggestions.
+Finally, you can also ask your peers who are already doing research.
+You can use the method described in \autoref{sec:contact} to contact professors
+
+\begin{commentbox}[Vu:] 
+    I enjoy working with undergrads and have advised many (10+).
+    I often recruit undergraduate researchers through my classes, e.g., asking students who did well in my class if they are interested in research.
+Occasionally I was introduced to students by other students or faculty.
+
+    AT GMU we pay undergrads about \$15/hr and about 10 hrs/week. This is a good way to make some money and get research experience.
+  \end{commentbox}
+
+% \paragraph{Online Research Communities and Open Source Contributions} Places such as Github offer many opportunities for research, even though they might not appear so (e.g., they do not explicitly mention research).
+% In many cases, projects are directly related to research (e.g., open-source software from professors' or research labs).
+For example, my research group has a many open-source projects that undergrads can contribute to (see \url{https://github.com/dynaroars/}).
+
+%  By contributing code, fixing bugs, implementing new features, or providing documentation, students can gain practical research experience and interact with experienced developers and researchers in the field.
+Not only you gain research experience, you might be able to get a LoR from the project maintainer and definitely should write about this experience in your SOP (\autoref{sec:research-statement}.)
+
+% \paragraph{Virtual Research Opportunities Beyond Physical Boundaries}: Several universities, organizations, and research institutions offer virtual internships and research programs aimed at providing hands-on research experience.
+These programs often involve working remotely under the guidance of experienced mentors and collaborating with a team of fellow researchers.
+For instance, \href{https://docs.google.com/forms/d/1btIwt4HwjyKMOUk-EMy3rbkfWzFxv2lNrMm_zkd0pA4/viewform?edit_requested=true}{UIUC+ Summer Undergraduate Research in Software Engineering}  offers an unpaid remote internship for software engineering students all over the world to collaborate with mentors from University of Illinois at Urbana Champaign.
+
+% \paragraph{Online Conferences and Workshops}: Attending online conferences and workshops is another way to gain exposure to cutting-edge research and establish connections with experts in the field.
+Many conferences now provide virtual participation options, enabling students to access research talks, poster sessions, and panel discussions and sometimes access designated chat rooms or networking events where participants can engage with researchers, ask questions, and seek potential research collaborations.
+It is beneficial to also create profiles academic collaboration platforms such as ResearchGate.
+By creating profiles, joining relevant research groups, and participating in discussions, students can connect with established researchers, contribute to ongoing projects, and potentially collaborate on publications or research proposals as they provide access
 

### 2024-03-09 demystify.tex
updates

@@
 
+There might be also special programs for undergraduate research from the university, e.g., GMU has the OSCAR program, UNL has UCARE, NSF has REUs (Research Experience for Undergraduates) for US citizen and permanent residents.
+
 You can also take honor thesis or independent study courses with a professor.
@@
 \begin{commentbox}[Vu:] 
-    I enjoy working with undergrads and have advised many (10+).
+    I enjoy working with undergrads (e.g., have mentored over 10 undergrads) and actively recruit them.
     I often recruit undergraduate researchers through my classes, e.g., asking students who did well in my class if they are interested in research.

### 2024-03-10 demystify.tex
updates

@@
-\documentclass[oneside,11pt]{memoir}
-\chapterstyle{pedersen}
+\documentclass[oneside,11pt]{book}
 
@@
 \usepackage[most]{tcolorbox}
-\setsecnumdepth{subsection}
+%\setsecnumdepth{subsection}
 \setcounter{tocdepth}{3}
@@
 \usepackage{booktabs}
-\usepackage{xcolor}
+\usepackage[dvipsnames]{xcolor}
+\usepackage{tikz}
+\usetikzlibrary{calc} 
+\usepackage{anyfontsize}
+\usepackage{sectsty}
+
 \usepackage[makeroom]{cancel}
@@
 
-\title{Demystifying the Computer Science \\
-  PhD Admission in the US\\{\large A Handbook for International \emph{and Domestic} Students}}
-
-\author{\href{https://nguyenthanhvuh.github.io}{ThanhVu (Vu) Nguyen}\\{\small George Mason University, Dept. of Computer Science}}
-
-\makeatletter
-\def\maketitle{%
-  \null
-  \thispagestyle{empty}%
-  \vfill
-  \begin{center}\leavevmode
-    \normalfont
-    {\LARGE\raggedright \textbf{\@title}\par}%
-    \vfill%
-    % \hrulefill\par
-    {\Large \@author\par}%
-    \vfill%
-    {\large\raggedleft \@date\par}%
-  \end{center}%
-  \vfill
-  \null
-  \cleardoublepage
+\begin{document}
+
+\pagestyle{empty}
+\begin{tikzpicture}[overlay,remember picture]
+
+    % Background color
+    \fill[
+    black!2]
+    (current page.south west) rectangle (current page.north east);
+
+    % Rectangles
+    \shade[
+    left color=Dandelion, 
+    right color=Dandelion!40,
+    transform canvas ={rotate around ={45:($(current page.north west)+(0,-6)$)}}] 
+    ($(current page.north west)+(0,-6)$) rectangle ++(9,1.5);
+
+    \shade[
+    left color=lightgray,
+    right color=lightgray!50,
+    rounded corners=0.75cm,
+    transform canvas ={rotate around ={45:($(current page.north west)+(.5,-10)$)}}]
+    ($(current page.north west)+(0.5,-10)$) rectangle ++(15,1.5);
+
+    \shade[
+    left color=lightgray,
+    rounded corners=0.3cm,
+    transform canvas ={rotate around ={45:($(current page.north west)+(.5,-10)$)}}] ($(current page.north west)+(1.5,-9.55)$) rectangle ++(7,.6);
+
+    \shade[
+    left color=orange!80,
+    right color=orange!60,
+    rounded corners=0.4cm,
+    transform canvas ={rotate around ={45:($(current page.north)+(-1.5,-3)$)}}]
+    ($(current page.north)+(-1.5,-3)$) rectangle ++(9,0.8);
+
+    \shade[
+    left color=red!80,
+    right color=red!80,
+    rounded corners=0.9cm,
+    transform canvas ={rotate around ={45:($(current page.north)+(-3,-8)$)}}] ($(current page.north)+(-3,-8)$) rectangle ++(15,1.8);
+
+    \shade[
+    left color=orange,
+    right color=Dandelion,
+    rounded corners=0.9cm,
+    transform canvas ={rotate around ={45:($(current page.north west)+(4,-15.5)$)}}]
+    ($(current page.north west)+(4,-15.5)$) rectangle ++(30,1.8);
+
+    \shade[
+    left color=RoyalBlue,
+    right color=Emerald,
+    rounded corners=0.75cm,
+    transform canvas ={rotate around ={45:($(current page.north west)+(13,-10)$)}}]
+    ($(current page.north west)+(13,-10)$) rectangle ++(15,1.5);
+
+    \shade[
+    left color=ForestGreen,
+    rounded corners=0.3cm,
+    transform canvas ={rotate around ={45:($(current page.north west)+(18,-8)$)}}]
+    ($(current page.north west)+(18,-8)$) rectangle ++(15,0.6);
+
+    \shade[
+    left color=ForestGreen,
+    rounded corners=0.4cm,
+    transform canvas ={rotate around ={45:($(current page.north west)+(19,-5.65)$)}}]
+    ($(current page.north west)+(19,-5.65)$) rectangle ++(15,0.8);
+
+    \shade[
+    left color=OrangeRed,
+    right color=red!80,
+    rounded corners=0.6cm,
+    transform canvas ={rotate around ={45:($(current page.north west)+(20,-9)$)}}] 
+    ($(current page.north west)+(20,-9)$) rectangle ++(14,1.2);
+
+    % Year
+    \draw[ultra thick,gray]
+    ($(current page.center)+(5,2)$) -- ++(0,-3cm) 
+    node[
+    midway,
+    left=0.25cm,
+    text width=5cm,
+    align=right,
+    black!75
+    ]
+    {
+    {\fontsize{25}{30} \selectfont \bf Computer\\[8pt] Science}
     } 
-\makeatother
-
-\begin{document}
-\maketitle
-\frontmatter
+    node[
+    midway,
+    right=0.25cm,
+    text width=6cm,
+    align=left,
+    ForestGreen]
+    {
+    {\fontsize{48}{50} \selectfont PhD}
+    };
+
+    % Title
+    \node[align=center] at ($(current page.center)+(0,-5)$) 
+    {
+    {\fontsize{28}{28} \selectfont {{Demystifying the Computer Science}}} \\[0.15in]
+    {\fontsize{28}{28} \selectfont {{PhD Admission in the US}}} \\[0.2in]    
+    {\fontsize{18}{18} \selectfont {{A Handbook for International \emph{and Domestic} Students}}} \\[0.5in]
+
+    {\fontsize{16}{19.2} \selectfont \textcolor{ForestGreen}{ \bf ThanhVu (Vu) Nguyen}}\\[0.1in]
+    George Mason University, Dept. of Computer Science\\[0.1in]
+    \today{} (latest version available on  \href{https://github.com/nguyenthanhvuh/phd-cs-us}{Github})
+    };
+    \end{tikzpicture}
 
@@
           \item SOP is very important. Write it in such a way that makes you \emph{stand out} (\autoref{sec:research-statement} and \autoref{sec:improve-your-chance})
-          \item GRE \emph{is not} required (\autoref{sec:grades}). Spend your time on something else!
-          \item Grades are important, but depend on the reputation of your school (\autoref{sec:grades}).
+          \item Grades are important, but depend on the reputation of your school (\autoref{sec:your-school}).
+          \item GRE \emph{is not} required (\autoref{sec:standard-tests}). Spend your time on something else!
           \item Getting an interview is typically a \emph{good sign}; but no interview does not mean rejection (\autoref{sec:interview}).
@@
 
-\section{Your School}\label{sec:your-school}
-
-Graduating from top universities \emph{that we recognize} helps.
+\section{Your School and Grades}\label{sec:your-school}
+
+\paragraph{School} Graduating from top universities \emph{that we recognize} helps.
 For example, if you are an international student and your school is well-known, then it is \emph{``top foreign''}, which is definitely a plus.
@@
 
-\section{Grades and GREs}\label{sec:grades}
-
-Having good grades is good, but unless your school is well-known, having top grades or rankings
+\paragraph{Grades} Having good GPA is good. However if your school is not well-known, having top grades or rankings
 usually will not help because we cannot evaluate them.
@@
 
+\section{Standard Tests (GREs, IELTS, etc) and Work Experiences}\label{sec:standard-tests}
+
 \paragraph{GRE} Most good CS PhD programs in the US \textbf{no longer require GREs}, so you don't need to take them.
 However, they might be useful for international students from programs we are not familiar with.
-If you have good GRE scores then you should include them in your application.
-
+If you have good GRE scores then you should include them in your application (but don't expect them to make much difference).
 \paragraph{English Test} Unless your degrees are from specific countries such as \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-GMU#standard-tests-waiver-eligible-countries}{these}, you will need to
-take standardized English test. Just do well enough to pass minimum requirement set by the university, which is required for TA (\autoref{sec:funding}).
+take standardized English test.
+Just do well enough to pass minimum requirement set by the university, which is required for TA (\autoref{sec:funding}).
+Just as with grades, having a high school in these tests might not help, but having too low scores can a \red{red flag} and sometimes results in an automatic rejection (\autoref{sec:why-rejected}), e.g., below the minimum requirement.
 
@@
 
+\paragraph{Work Experiences} Work experiences, especially if they are not research-based, do not carry much weight.
+So do not spend much time talking about them in your SOP (and similarly, LoR from your supervisor for non-research experience will not count much).
+
+However, if you have research-based work experiences, then they can be useful and should be mentioned in your SOP and LORs.
+For example, if you have worked at a research lab (e.g., Microsoft, VinAI in Vietnam, etc), then you should mention that in your SOP and ask your supervisor to write about your research experiences and potentials in their LOR (\autoref{sec:lor}).
+
 \section{CV/Resume}
@@
 
-\paragraph{Interviews} Just like a job application, you might get interviews from schools.
-The most common case is that a professor is interested in working with you and wants to chat to make decision (e.g., to offer RA).
+\paragraph{Interviews} Just like a job application, you likely will get interviews from schools.
+The most common case is that a professor is interested in working with you and wants to chat to make decision e.g., to offer RA (\autoref{sec:ra})).
 In some cases, the interview is done by several professors, e.g., to see if a student fits to their group or to recruit a very strong student to their program.
@@
 Typically, an interview takes about 15--30 minutes, and one important aspect of evaluation is your ability to effectively communicate, including speaking and understanding English. 
-In general treat it as an informal chat, but be prepared to talk about your research experiences and interests (so do have \emph{elevator pitch} about your experiences and research). 
+In general, just treat the interview as an informal chat, but be prepared to talk about your research experiences and interests (so do have \emph{elevator pitch} about your experiences and research). 
 A professor might want to know more about your research background or even ask you to read a paper (so be sure to do it!).
-During the interview, you will have the opportunity to ask questions about the university.
-It's essential to prepare some thoughtful questions, similar to a job interview.
+Some prof.
+can ask you to solve a problem, e.g., a coding problem (so you might want to ask if you need to prepare for that).
+During the interview, you will have the opportunity to ask questions about the university. It's essential to prepare some thoughtful questions, similar to a job interview.
 
@@
 Many programs do not have the tradition of interviewing applicants.
-For example, at GMU, most of our admitted students with TA do not go through interviews.
-
-However, no interviews typically means that you will not likely get an RA (\autoref{sec:funding}), which is offered by an individual faculty (if they want to pay you to do research for them, then likely they want to talk to you first).
+For example, at GMU, most of our admitted students with TA (\autoref{sec:ta}) do not go through interviews.
+
+However, no interviews typically means that you will not likely get an RA (\autoref{sec:ra}), which is offered by an individual faculty (if they want to pay you to do research for them, then likely they want to talk to you first).
 In this case, your application (and TA/fellowship funding) will be reviewed and decided by the adcom.
@@
 
-\section{Graduate Assistantship (TA/RA)}
+\section{Graduate Assistantship (TA/RA)}\label{sec:ta-ra}
 The most common type of funding is \textbf{graduate assistantship}, which is either TA or RA.
@@
 
-\subsection{Research Assistant (RA)}
+\subsection{Research Assistant (RA)}\label{sec:ra}
 RA is provided through a professor through their own funding so you can work on their project.  
@@
 
-\section{Fellowship/Scholarship}
+\section{Fellowships/Scholarships}\label{sec:fellowships}
 
@@
 While this can be confusing due to the large number of universities in the US, a little research, e.g., searching for PhD in CS from the school website, will help you find out.
-All schools listed in \autoref{sec:ranking} have PhD studies in CS, so you can start there.
+All schools listed in \autoref{sec:ranking} have PhD programs in CS, so you can start there.
 
@@
 
-\paragraph{Standard Tests \autoref{sec:grades}} You do not need to take TOEFL or IELTS because you already did your undergrad (or MS) at a US university.
+\paragraph{Standard Tests \autoref{sec:standard-tests}} You do not need to take TOEFL or IELTS because you already did your undergrad (or MS) at a US university.
 You might also be more comfortable communicating in English, e.g., contacting professors (\autoref{sec:contact}).

### 2024-03-10 demystify.tex
mend
add cover page

@@
     {
-    {\fontsize{28}{28} \selectfont {{Demystifying the Computer Science}}} \\[0.15in]
-    {\fontsize{28}{28} \selectfont {{PhD Admission in the US}}} \\[0.2in]    
+    {\fontsize{24}{24} \selectfont {{Demystifying the Computer Science}}} \\[0.15in]
+    {\fontsize{24}{24} \selectfont {{PhD Admission in the US}}} \\[0.2in]    
     {\fontsize{18}{18} \selectfont {{A Handbook for International \emph{and Domestic} Students}}} \\[0.5in]

### 2024-03-10 demystify.tex
mend
updates

@@
 \renewcommand{\subsectionautorefname}{\S\@gobble}
+\renewcommand{\appendixautorefname}{\S\@gobble}
 \makeatother
@@
 
-\chapter{Preface}
+\chapter*{Preface}
 Having been involved in PhD admissions for many years, I've
@@
 
-I wish you the best of luck. And if you follow this guidance, you will at least have a good chance at GMU (see
-\href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-GMU}{why
-  you want to study at here}). Happy school hunting!
+I wish you the best of luck. Happy school hunting!
 
@@
 \newpage
-\tableofcontents*
+\tableofcontents
 
@@
 Each application is assigned to 2--3 faculty members, who will evaluate your profile and reach a consensus.
-Note that while the assigned reviewers are likely the main ones deciding your application, other faculty in the department will likely have access to your application and can provide inputs and opinions on your profile.
+Note that while the assigned reviewers are the main ones deciding your application, other faculty in the department will likely have access to your application and can provide inputs and opinions on your profile.
 
@@
 Some members can be tenured faculty.
-The chair of the committee will likely a senior faculty, but they likely will not review individual applications and instead assign them to committee members.
+The chair of the committee will likely a senior faculty, but they will not review individual applications and instead assign them to committee members.
 The chair will look at various factors such as research interests or mentioning faculty names to assign the applications to appropriate faculty.
 
-At GMU, we usually decide that a full-time PhD candidate is either (i) admit with funding (TA or RA, see \autoref{sec:funding}) or (ii) rejected. In other words, in most cases, we either
+At GMU, we usually decide that a full-time PhD candidate is either (i) admit with funding (\autoref{sec:funding}) or (ii) rejected. In other words, in most cases, we either
 admit you with full funding, or we reject your application. In some rare cases, we admit
@@
 
-\section{Increasing your admission chance}\label{sec:improve-your-chance}
+\section{Increasing my admission chance}\label{sec:improve-your-chance}
 
@@
 \section{Member of an underrepresented or LBGTQ+ groups}\label{sec:urm}
-A common question is whether you should mention, e.g., in SOP, that you are a member of an underrepresented and minority (URM) group\footnote{URM in CS in the US include African Americans, Hispanic Americans, American Indians/Native Alaskans, Native Hawaiians/Pacific Islanders (excluding Asian Americans), and multiracial students identifying with at least one of the previously listed URM categories.} and LBGTQ+ group.
+A common question is whether you should mention that you are a member of an underrepresented and minority (URM) group\footnote{URM in CS in the US include African Americans, Hispanic Americans, American Indians/Native Alaskans, Native Hawaiians/Pacific Islanders (excluding Asian Americans), and multiracial students identifying with at least one of the previously listed URM categories.} and LBGTQ+ group (, e.g., in your SOP).
 This is a strong concern for international (and even domestic) students, who may have encountered discrimination in their countries or personal experiences.
@@
 
-Most of what is written in this document is applicable to both domestic\footnote{Here domestic means you did your undergrad (or MS) at a US university.} and international students.
+Most of what is written in this document is applicable to both domestic\footnote{For simplicity domestic means you did your undergrad (or MS) at a US university.} and international students.
 However, there are some differences and benefits that domestic students should be aware of.
-This section lists some of these.
 
@@
 
-\paragraph{Funding \S\ref{sec:funding}} You have more opportunities for funding, e.g., through government scholarships for US citizens and permanent residents.
+\paragraph{Funding \autoref{sec:funding}} You have more opportunities for funding, e.g., through government scholarships for US citizens and permanent residents.
 You can also apply to specific programs \emph{before} you start your PhD, e.g., NSF Graduate Research Fellowship Program (GRFP) and Hertz Foundation Fellowship.
@@
 
-\paragraph{Research Experiences} You might have more research experiences than international students.
+\paragraph{Research Experiences \autoref{sec:undergrad-research}} You might have more research experiences than international students.
 For example, you likely have many opportunities to do research as an undergraduates, e.g., through REUs and internships.
@@
 
-\paragraph{Standing out} There are \emph{very few} domestic applications compared to international ones, i.e., domestic students are minority (\autoref{sec:urm}) in the CS PhD application pool.
+\paragraph{Standing out \autoref{sec:improve-your-chance}} There are \emph{very few} domestic applications compared to international ones, i.e., domestic students are minority (\autoref{sec:urm}) in the CS PhD application pool.
 Many US universities want to increase the number of domestic students in their programs (and as mentioned above, there are specific fellowships for domestic students).
@@
 
-\paragraph{Open House} (\autoref{sec:accepted}) It is easier for you to attend open house events in person.  This can help you make a better decision on which school to attend.
+\paragraph{Open House \autoref{sec:accepted}} It is easier for you to attend open house events in person.  This can help you make a better decision on which school to attend.
 
@@
 \begin{commentbox}[Vu:] 
-    I enjoy working with undergrads (e.g., have mentored over 10 undergrads) and actively recruit them.
-    I often recruit undergraduate researchers through my classes, e.g., asking students who did well in my class if they are interested in research.
+    I enjoy working with undergrads (e.g., have mentored over 10 undergrads) and always actively recruit.
+    I often get undergraduate researchers through my classes, e.g., asking students who did well in my class if they are interested in research.
 Occasionally I was introduced to students by other students or faculty.
 
-    AT GMU we pay undergrads about \$15/hr and about 10 hrs/week. This is a good way to make some money and get research experience.
+    While most undergrads are understandably not super productive in research, some are and I have published multiple papers with them.
+I also have written strong LoRs for them and have helped them get into PhD programs, including top ones.
+
+    \tcblower
+    At GMU we pay undergrads about \$15/hr and about 10 hrs/week. This is flexible and can be increased as needed.
 \end{commentbox}
@@
 
-\chapter{History and Acknowledgement}\label{sec:ack}
-
-\paragraph{History} This handbook was conceived during a lunch with Craig Yu at GMU.
+\chapter*{History and Acknowledgement}\label{sec:ack}
+
+\paragraph{History} This handbook was conceived during a lunch with a faculty at GMU.
 We talked on about why GMU were not able to attract good Vietnamese and other international students, despite having a much stronger CS program than many schools that these students want to go to (part of the reason is described in \autoref{sec:selecting-ranking-schools}).
@@
 I have been in the PhD admission process at GMU and UNL for many years.
-Currently I serve as the program director of the MS program in Software Engineering at GMU (thus also have some experience with the MS admission process--which is quite different than PhD).
+Currently I serve as the program director of the MS program in Software Engineering at GMU (thus also have some experience with the MS admission process---which is quite different than PhD).
 My personal and lab website is at \href{https://dynaroars.github.io}{dynaroars.github.io}.
 
-Though I'm not an international student, many of my students and collaborators are.
+Though I was not an international student, many of my students and collaborators are/were.
 I also mentor multiple students from Vietnam, and have close colleagues and friends who were once international students.

### 2024-03-10 demystify.tex
mend
updates

@@
 
-There are various guides on writing SOP, e.g.,~\cite{blattman2022writing}, and \href{https://cs-sop.org/}{many example statements} are available.
+There are various guides on writing SOP , and \href{https://cs-sop.org/}{many example statements} are available.
 In short, focus on your research motivation and goal, and convince us that you can achieve it through your experience, e.g., published papers.
@@
 \begin{itemize}
+    \item \href{https://chrisblattman.com/blog/2022/01/11/}{Writing your statement of purpose} by Chris Blattman 
   \item \href{http://www.pl-enthusiast.net/2022/10/03/how-to-write-a-grad-school-personal-statement/}{How to Write a Grad School Personal Statement} by Mike Hicks
@@
 However, if you have research-based work experiences, then they can be useful and should be mentioned in your SOP and LORs.
-For example, if you have worked at a research lab (e.g., Microsoft, VinAI in Vietnam, etc), then you should mention that in your SOP and ask your supervisor to write about your research experiences and potentials in their LOR (\autoref{sec:lor}).
+For example, if you have worked at a research lab (e.g., Microsoft, DoE, VinAI in Vietnam, etc) (\autoref{sec:research-ability}), then you should mention that in your SOP and ask your supervisor to write about your research experiences and potentials in their LOR (\autoref{sec:lor}).
 
@@
 
-\paragraph{Locally} The most common way to find research is right at your own institution.  
+\paragraph{Locally} Start looking for research opportunities at your own institution.  
 If you did well or like a class, you can check with the professor of that class for research opportunities.  
@@
 Finally, you can also ask your peers who are already doing research.
-You can use the method described in \autoref{sec:contact} to contact professors
+Use the method described in \autoref{sec:contact} to contact professors
 
@@
 
-% \paragraph{Online Research Communities and Open Source Contributions} Places such as Github offer many opportunities for research, even though they might not appear so (e.g., they do not explicitly mention research).
-% In many cases, projects are directly related to research (e.g., open-source software from professors' or research labs).
+\paragraph{Online Research Communities and Open Source Contributions} Places such as Github offer many opportunities for research, even though they might not appear so (e.g., they do not explicitly mention research).
+In many cases, professors' or research labs put their research projects on Github.
 For example, my research group has a many open-source projects that undergrads can contribute to (see \url{https://github.com/dynaroars/}).
 
-%  By contributing code, fixing bugs, implementing new features, or providing documentation, students can gain practical research experience and interact with experienced developers and researchers in the field.
+By contributing code, fixing bugs, implementing new features, or providing documentation, students can gain practical research experience and interact with experienced developers and researchers in the field.
 Not only you gain research experience, you might be able to get a LoR from the project maintainer and definitely should write about this experience in your SOP (\autoref{sec:research-statement}.)
 
-% \paragraph{Virtual Research Opportunities Beyond Physical Boundaries}: Several universities, organizations, and research institutions offer virtual internships and research programs aimed at providing hands-on research experience.
+\paragraph{Virtual Research Opportunities}: This is less common but several places offer virtual internships and research programs aimed at providing hands-on research experience.
 These programs often involve working remotely under the guidance of experienced mentors and collaborating with a team of fellow researchers.
@@
 Many conferences now provide virtual participation options, enabling students to access research talks, poster sessions, and panel discussions and sometimes access designated chat rooms or networking events where participants can engage with researchers, ask questions, and seek potential research collaborations.
-It is beneficial to also create profiles academic collaboration platforms such as ResearchGate.
 By creating profiles, joining relevant research groups, and participating in discussions, students can connect with established researchers, contribute to ongoing projects, and potentially collaborate on publications or research proposals as they provide access

### 2024-03-11 demystify.tex
Updates from Overleaf
mend
updates

@@
 \usepackage[most]{tcolorbox}
-%\setsecnumdepth{subsection}
-\setcounter{tocdepth}{3}
+% \setsecnumdepth{subsection}
+% \setcounter{tocdepth}{3}
 \usepackage{enumitem}
@@
     {\fontsize{24}{24} \selectfont {{PhD Admission in the US}}} \\[0.2in]    
-    {\fontsize{18}{18} \selectfont {{A Handbook for International \emph{and Domestic} Students}}} \\[0.5in]
+    {\fontsize{18}{18} \selectfont {{A Handbook for International and Domestic Students}}} \\[0.5in]
 
@@
 
+\begin{mybox}
 This handbook is available at
@@
 If you have questions or comments, feel free to create a \href{https://github.com/nguyenthanhvuh/phd-cs-us/issues}{GitHub issue} for discussion.
+
+\end{mybox}
 
@@
 
-\bibliographystyle{abbrv}
-\bibliography{demystify.bib}
+%\bibliographystyle{abbrv}
+%\bibliography{demystify.bib}
 

### 2024-03-15 demystify.tex
updates

@@
 Furthermore, although many examples are specifics for schools that I and other contributors of this document know about, the information should be generalizable to other R1\footnote{An \href{https://en.wikipedia.org/wiki/List_of_research_universities_in_the_United_States}{R1 institution} in the US is a research-intensive university with a high level of research activity across various disciplines.
-Currently, 146 (out of 4000) universities are classified as R1.} institutions in the US (and universities in other countries).
+Currently, 146 (out of 4000) US universities are classified as R1.} institutions in the US (and universities in other countries).
 
@@
 \begin{mybox}
-This handbook is available at
+This document is available at
 
@@
 \item \emph{You are not a good fit to the program}.  You could have an excellent profile, but if you are interested in a research area that the program does not have, you might not be admitted.
-Similarly, if there is no faculty who is willing to advise you (e.g., they are on sabbatical, NSF leave, or already have too many students currently), you will not be admitted.
+Similarly, if there is no faculty who is willing to advise you (e.g., they are on sabbatical, personal leave, or already have too many students already), you will not be admitted.
 This is actually good for you, as you don't want to be in a program where no one can advise you.
 
-For example, in the 2024 admission cycle, there is a huge surge in students interested in AI and NLP (thanks, ChatGPT!). Thus, AI/NLP faculty might be 
-overwhelmed and cannot consider many, even those with excellent profiles.
+For example, 2024 admission cycle has a huge surge in students interested in AI and NLP (thanks, ChatGPT!).
+Consequently, AI/NLP faculty might be overwhelmed and cannot consider many candidates, even those with excellent profiles.
 
@@
 \section{Member of an underrepresented or LBGTQ+ groups}\label{sec:urm}
-A common question is whether you should mention that you are a member of an underrepresented and minority (URM) group\footnote{URM in CS in the US include African Americans, Hispanic Americans, American Indians/Native Alaskans, Native Hawaiians/Pacific Islanders (excluding Asian Americans), and multiracial students identifying with at least one of the previously listed URM categories.} and LBGTQ+ group (, e.g., in your SOP).
+A common question is whether you should mention that you are a member of an underrepresented and minority (URM) group\footnote{URM in CS in the US include African Americans, Hispanic Americans, American Indians/Native Alaskans, Native Hawaiians/Pacific Islanders (excluding Asian Americans), and multiracial students identifying with at least one of the previously listed URM categories.} or LBGTQ+ group (, e.g., in your SOP).
 This is a strong concern for international (and even domestic) students, who may have encountered discrimination in their countries or personal experiences.
@@
 
-\section{Buying Computer Equipment}\label{sec:buying-equipment} Many students get very excited about getting to buy new computer equipment and electronics when they start their PhD.
+\section{Buying Computer Equipment}\label{sec:buying-equipment} understandably students get very excited about their upcoming PhD journal and want to buy new computer equipment and electronics to prepare.
 However, you should check with your adviser first.
@@
 have funding to buy you computers and other equipments (e.g., software, books, keyboards, headphones, tablets, etc).
+Many CS programs also provide some budget or computer equipments to incoming PhD students (e.g., a laptop).
 
@@
 
-\paragraph{Research Experiences \autoref{sec:undergrad-research}} You might have more research experiences than international students.
-For example, you likely have many opportunities to do research as an undergraduates, e.g., through REUs and internships.
+\paragraph{Research Experiences \autoref{sec:undergrad-research}} You might have many opportunities to do research as an undergraduates, e.g., through REUs and internships at your undergrad university.
 Highlight such experiences in your application.

### 2024-03-16 demystify.tex
Updates from Overleaf

--- 2024-03-10 demystify.tex
@@
 You will be very surprised to learn that a school that you didn't know much about can have very strong research in your interested topic (and vice versa, a school you thought highly about might have no faculty working in the research field you're interested in).
-This is also a good way to learn about individual faculty (who works on what) and well-known CS conferences\footnote{In CS (and probably only in CS), conferences, not journals, are often the main venue to publish research finding.}.
+This is also a good way to learn about individual faculty (who works on what) and well-known CS conferences\footnote{In CS (and probably only in CS), conferences, not journals, are often the main venue to publish research finding (see \href{https://homes.cs.washington.edu/~mernst/advice/conferences-vs-journals.html}{why here}).}.
 %\autoref{sec:ranking} gives the top 50 CS programs in the US according to CSRankings.

### 2024-03-16 demystify.tex
Updates from Overleaf

@@
 \usepackage{booktabs}
-\usepackage[dvipsnames]{xcolor}
+%\usepackage[dvipsnames]{xcolor}
 \usepackage{tikz}
@@
 did your reference writers submit their letters?
-Usually, this screen process is done through a central university system, i.e., not by the CS department.
+Usually, this screen process is done through a central university system, i.e., not by faculty from the CS department.
 If you pass this screen, your application will be forwarded to the CS department for further evaluation.
@@
 
-Then, applications are reviewed by a \emph{PhD admission committee} (\textbf{adcom}) that consists of faculty members in CS (in some cases the committee can involve affiliated faculty from different disciplines).
-These faculty have a wide-range of expertise and background to ensure diverse perspectives in the evaluation process.
+Then, applications are reviewed by a \emph{PhD admission committee} (\textbf{adcom}) that consists of faculty members in CS\footnote{In some cases the committee can involve affiliated faculty from different disciplines.}.
+These adcom members have a wide-range of expertise and background to ensure diverse perspectives in the evaluation process.
 The size and the review load of the adcom depends on the department size.
 At GMU, the PhD adcom typically has 15--20 faculty, and each committee member is assigned to review 25--30 applications.
-Note that many schools, including GMU, have separate adcoms for MS programs.
-
-Each application is assigned to 2--3 faculty members, who will evaluate your profile and reach a consensus.
-Note that while the assigned reviewers are the main ones deciding your application, other faculty in the department will likely have access to your application and can provide inputs and opinions on your profile.
-
-The PhD adcom typically involves assistant professors in the department (see \autoref{sec:faculty-types} for various faculty titles).
+Note that most large schools, including GMU, have separate adcoms for MS programs.
+
+Each application is assigned to about 3 faculty members, who will evaluate your profile and reach a consensus.
+While the assigned reviewers are the main ones deciding your application, other faculty in the department will likely have access to your application and can provide inputs and opinions on your profile.
+
+The PhD adcom typically involves assistant professors in the department (see \autoref{sec:faculty-types} for various types of faculty).
 This provides junior faculty the opportunities to recruit students.
-Some members can be tenured faculty.
+Some members are tenured faculty.
 The chair of the committee will likely a senior faculty, but they will not review individual applications and instead assign them to committee members.
-The chair will look at various factors such as research interests or mentioning faculty names to assign the applications to appropriate faculty.
+The chair will look at various factors such as research interests or mentioning faculty names to assign the applications to appropriate faculty, e.g., I am often assigned to review applicants interested in software engineering.
 
@@
 
-  Another sad reality is that some professors are unwilling to write letters or write weak ones because they do not want (good) students to go abroad or only go to places where they want the students to go to.
+  Several international students mentioned that some professors are unwilling to write letters or write weak ones because they do not want (good) students to go abroad or only go to places where they want the students to go to.
 If you are in this situation, you should find someone else to write for you.
@@
   Sometimes students would go through great length just to get letters from well-known professors in their school (e.g., department head).
-But as mentioned, if these professors do not know you, the letters are generic and carry little value, in fact, \red{red flags}.
-Moreover, a top professor in Vietnam might not be well-known to US faculty (see more details in \autoref{sec:your-school}).
-So save the trouble and just get letters from \emph{any} professors/supervisors who knows you well and can write a good letter about \emph{your} research ability.
+But as mentioned, if these professors do not know you, their letters would likely be generic and carry little value, in fact, \red{red flags}.
+Moreover, a top professor at your university might not be well-known to US faculty (see more details in \autoref{sec:your-school}).
+So save the trouble and  get letters from \emph{any} professors/supervisors who knows you well and can write a good letter about \emph{your} research ability.
 It's better to have a good personalized
@@
 
-What can you do about this?
-You can explain about your school in your SOP (and also ask your LOR writer to do that).
+\textbf{Advice}: You can explain about your school in your SOP (and also ask your LOR writer to do that).
 Of course, if you're interested in working with Vietnamese, consider  \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{CS programs in the US that have Vietnamese professors}.
@@
 usually will not help because we cannot evaluate them.
-
 This can be an issue for students in many top international universities where the competition is so high that very good students can still have low rankings from these schools (and be overlooked by the adcom).
-So what to do with this? As mentioned in \autoref{sec:your-school} you can put a note about this in your statement and ask your LoR writers to talk about it.
-
+
+\textbf{Advice}: As with school reputation, you can put a note about this in your statement and ask your LoR writers to talk about it.
 Note that while having good grades at unknown school might not help,
-having bad grades will be \red{red flag} (unless your LORs or
+having bad grades (e.g., < 3.0) will be \red{red flag} (unless your LORs or
 statement give a proper explanation). This is especially true if you
-have bad grades in relevant courses (e.g., Math and CS).
+have bad grades in relevant courses, e.g., Math and CS.
 
@@
 
-\paragraph{GRE} Most good CS PhD programs in the US \textbf{no longer require GREs}, so you don't need to take them.
-However, they might be useful for international students from programs we are not familiar with.
-If you have good GRE scores then you should include them in your application (but don't expect them to make much difference).
+\paragraph{GRE} Most good CS PhD programs in the US (e.g., R1 universities) \textbf{no longer require GREs}, so you don't need to take them.
+
+\textbf{Advice}: In general, don't waste your time taking GRE.
+However, if you took it and have good scores then you can include and talk about them in your application (but don't expect them to make much difference).
+
 \paragraph{English Test} Unless your degrees are from specific countries such as \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-GMU#standard-tests-waiver-eligible-countries}{these}, you will need to
 take standardized English test.
-Just do well enough to pass minimum requirement set by the university, which is required for TA (\autoref{sec:funding}).
-Just as with grades, having a high school in these tests might not help, but having too low scores can a \red{red flag} and sometimes results in an automatic rejection (\autoref{sec:why-rejected}), e.g., below the minimum requirement.
+
+\textbf{Advice}: Just do well enough to pass minimum requirement set by the university, which is required for TA (\autoref{sec:funding}).
+Just as with grades (\autoref{sec:your-school}) and GRE, having a high scores in these tests might not help, but having too low scores can a \red{red flag} and sometimes results in an automatic rejection (\autoref{sec:why-rejected}), e.g., below the minimum requirement.
 
@@
 
-\paragraph{Work Experiences} Work experiences, especially if they are not research-based, do not carry much weight.
-So do not spend much time talking about them in your SOP (and similarly, LoR from your supervisor for non-research experience will not count much).
-
+\paragraph{Work Experiences} Work experiences, especially if they are not research-based, do not carry much weight. Similarly, LoR from your supervisor for non-research experience will not count much.
+
+\textbf{Advice}: do not spend much time talking about non-research related job in your SOP.  
 However, if you have research-based work experiences, then they can be useful and should be mentioned in your SOP and LORs.
@@
 \section{CV/Resume}
-The CV/Resume should provide a summary of the applicant's achievements.
-It should allow the reviewers to quickly scan to identify major achievements (e.g., Publications, Programming Competition Awards, Teaching Experience).
+The CV/Resume provides a summary of the applicant's achievements.
+
+\textbf{Advice}: Prepare your CV in such away that it allows the reviewers to quickly scan to identify major achievements (e.g., Publications, Programming Competition Awards, Teaching Experience).
 
@@
 Typically, an interview takes about 15--30 minutes, and one important aspect of evaluation is your ability to effectively communicate, including speaking and understanding English.
-In general, just treat the interview as an informal chat, but be prepared to talk about your research experiences and interests (so do have \emph{elevator pitch} about your experiences and research). 
+
+\textbf{Advice}: In general, just treat the interview as an informal chat, but be prepared to talk about your research experiences and interests (so do have \emph{elevator pitch} about your experiences and research).
 A professor might want to know more about your research background or even ask you to read a paper (so be sure to do it!).
 Some prof.
-can ask you to solve a problem, e.g., a coding problem (so you might want to ask if you need to prepare for that).
+can ask you to solve a problem, e.g., a coding problem, so you might want to ask if you need to prepare for that.
 During the interview, you will have the opportunity to ask questions about the university. It's essential to prepare some thoughtful questions, similar to a job interview.
@@
 While this can be confusing due to the large number of universities in the US, a little research, e.g., searching for PhD in CS from the school website, will help you find out.
-All schools listed in \autoref{sec:ranking} have PhD programs in CS, so you can start there.
+Schools listed in \autoref{sec:ranking} have PhD programs in CS, so you can start there.
 
@@
 
-If you want to contact a prof.
+On the other hand, if you want to contact a prof.
 to ask about \emph{research opportunities}, or \emph{GTA/GRA} support, then \emph{yes}, I believe you should---it is \emph{worth it}.
@@
 
-Below is a good example that I would reply to.
+Below is a good example that I would likely reply.
 
@@
 If you really want feedback, ask your professors, collaborators, ref writers, or those who have previously applied.
+
+% \section{How much time did you spend looking at an application?}\label{sec:time-spent}
+% Short answer: less than 5 mins for quick rejection; 15+ mins for an acceptance.  On average, about 10 mins.
+
+% The GMU evaluation system allows us to download a single pdf that consists of a summary page (e.g., school info, GPA, testscores), then transcripts, standard test results, LoRs, CV, SOP, writing samples So I typically read the application in this order.
+Then I make notes on what I've seen, basically evidence that support my evaluation.
+
+% I first look at the summary for school, GPA, est scores (essentially looking for very low GPA or testscores that do not meet the minimum requirements.
+Then I skim on transcripts look for low/failing grades on CS and related courses.
+For these I make comments such as "unknown international school with good GPA, too many low grades on CS and related courses" or "good GPA from top school XYZ".
+I do not look at details on tests (the main scores are already given in the first summary page).
+
+% Then I look at LoRs.
+I read carefully good LoRs and quickly skim over bad or generic ones (very easy to tell, you don't even need much experiences to distinguish between a strong LoRs from the rest).
+Comment examples For LoRs include "very strong LoRs from well-known prof.
+XYZ talking about research experiences" or "generic LoRs, mostly about course work".
+I also note any unusual things in the LoRs.
+
+% Then the SOP, which I read carefully (of course if it is bad then it would a quick read).
+Then my notes for SOP include things such as "SOP mostly on course work, no clear motivation, generic" , or "very enjoyable SOP talking about XYZ, customized for GMU, mentioned profs.
+X,Y,Z".
+If the SOP mentioned about publications then I'll make notes such as "has X papers at top venues" or "has X papers but at unknown places".
+I will not go online search for the papers (unless they upload them as sample papers) .
+
+% I also skim over sampling papers, essentially looking over the abstract to see if reads well.  I would make notes such as "have a paper under submission to unknown venue, but reads well".
+
+%At GMU we use an evaluation system that allows us to download a single, large pdf that consists of everything submitted by the applicant, e.g., GPA, transcripts, test scores, CV, LoRs, SOP, sample writings, etc.
+The first couple of pages show the summary including school, major, GPA, test scores.
 

### 2024-03-17 demystify.tex
updates

@@
-\documentclass[oneside,11pt]{book}
+\documentclass[oneside,11pt,dvipsnames]{book}
 
@@
 \usepackage{booktabs}
-%\usepackage[dvipsnames]{xcolor}
+
 \usepackage{tikz}
 \usetikzlibrary{calc}
+\usepackage{xcolor}
+
 \usepackage{anyfontsize}
@@
         \begin{itemize}
-          \item Applications are evaluated by the \emph{PhD Admission} (adcom) committee and each application is reviewed by typically three faculty (\autoref{sec:evalapps}).
+          \item Applications are evaluated by the \emph{PhD Admission} (adcom) committee and each application is typically reviewed by three faculty (\autoref{sec:evalapps}).
           \item Individually faculty \emph{cannot directly admit} a student---so do not email and ask if you have a chance.
@@
           \item The committee will look at various factors, but the most important ones are research ability, e.g., publications, statement of purpose (SOP), and  letters of recommendation (LORs).
-          \item LORs are very important, but only if they are personalized and talk about your research ability (\autoref{sec:lor}).
-          \item SOP is very important. Write it in such a way that makes you \emph{stand out} (\autoref{sec:research-statement} and \autoref{sec:improve-your-chance})
-          \item Grades are important, but depend on the reputation of your school (\autoref{sec:your-school}).
-          \item GRE \emph{is not} required (\autoref{sec:standard-tests}). Spend your time on something else!
+          \item \emph{LORs are very important}, but only if they are personalized and talk about your research ability (\autoref{sec:lor}).
+          \item \emph{SOP is very important}. Write it in such a way that makes you \emph{stand out} (\autoref{sec:research-statement} and \autoref{sec:improve-your-chance})
+          \item  \emph{Pubication can greatly help}.
+Publications are good venues are concrete evidence that you have successfully engaged in research.
+But they are not required!
+(\autoref{sec:research-ability}).
+
+          \item \emph{Grades might be important}, depends on the reputation of \emph{your school} (\autoref{sec:your-school}).
+While good grades might not help much, bad ones will likley hurt your case.
+          \item \emph{Standard tests are not important} (\autoref{sec:standard-tests}).
+GRE \emph{is not} even required.
+Spend your time on something else!
+For standard English test, just do enough to pass the minimum requirements.
+
+          \item \emph{Contacting a prof. for is recommended}, but do it \emph{properly} (\autoref{sec:contact}).
+
           \item Getting an interview is typically a \emph{good sign}; but no interview does not mean rejection (\autoref{sec:interview}).
@@
           \item Many schools do not offer PhD studies in CS and many CS professors do not advise or graduate PhD students  (\autoref{sec:schoolsandprofs}).
-          \item Contacting a prof. for research opportunities is recommended, but do it \emph{properly} (\autoref{sec:contact}).
         \end{itemize}

### 2024-03-17 demystify.tex
updates

@@
 Spend your time on something else!
-For standard English test, just do enough to pass the minimum requirements.
+For standard English tests (not required for domestic students), just do enough to pass the minimum requirements.
 
@@
 For external fellowships including those from the US government and tech companies (e.g., Google, Microsoft, Facebook), you will need to apply.
-Fellowships from US government would likely require US citizenship or residency.
 % However, tech companies including Google, Microsoft, Facebook have fellowships that international students can apply for.
@@
 
-In general, fellowship, especially competitive ones from the government or company, is prestigious, and you will stand out if you get one.
+In addition to general fellowships that are available to all students, there are also specic fellowships, e.g., for US citizen and permanent residents, for underrepresented minority (URM) students (\autoref{sec:urm}), and for students from specific countries.
+In short, there are many fellowships available, so you should look for them and apply.
+
+In general, fellowships, especially those that are open to everyone, are highly competitive and prestigious, and you will stand out if you get one.
 Every PhD student has pubs, but only few  would have NSF Graduate or Microsoft fellowship.

### 2024-03-17 demystify.tex
updates

@@
 It gives both the professor and student more time, e.g., a whole semester, to work and evaluate the relationship before making any commitment (sounds a bit like a marriage!).
+
 You can also ask if you can do an independent study or research with a prof.
@@
 Ultimately, choose a prof.
-that fits you the most by communicating with them, taking their courses, meeting and asking them questions, and talking to their current students.
+that fits you by communicating with them, taking their courses, meeting and asking them questions, and talking to their current students.
 It will take time and effort, but since you will be working with this person for 5+ years, it is important to try to find the right one.
@@
 
-In many cases you do not need to do a PhD in CS to work in your area of interest.
+You \emph{do not} need to do a PhD in CS to work in your area of interest.
 For example, in addition to a traditional CS department, GMU has IST and Cybersecurity departments, both of which have faculty  with PhD in CS and work on CS topics (e.g., AI, Security, Robotics).

### 2024-03-17 demystify.tex
Updates from Overleaf

@@
         \begin{itemize}
-          \item TA and RA are two main funding sources.
-TA (teaching assistantship) is provided by the department to help with classes (e.g., grading).
-RA (research assistantship) is provided by profs.
-to help with their research (\autoref{sec:funding}).
+          \item TAs, RAs, and fellowships are two main funding sources (\autoref{sec:funding}).
+TAs (teaching assistantships) are provided by the department to help with classes (e.g., grading).
+RAs (research assistantships) are given by profs.
+to help with their research.
+Fellowships, provided by the university, department, or external sources such as government or industry, give move flexibility but are very competitive.
         \end{itemize}
@@
 
-\paragraph{What else to do?} Make decision (\autoref{sec:accept-postpone-decline})!
+\paragraph{What else?} Make decision (\autoref{sec:accept-postpone-decline})!
 accept, reject, or defer the offers (if you have multiple options).
@@
 Ask about computer equipment and software, office space, and other resources; in many case these will be provided for free by your adviser or department (\autoref{sec:buying-equipment}).
+
+Also, do not forget to update and thank your LoR writters and others who have supported you throughout your admission journey.
 

### 2024-03-17 demystify.tex
updates

@@
 Most PhD programs will require at least \textbf{two LORs}.
+A good LoR letter should have detail, enthusiasm, and specific examples of the applicant's achievements, all of which paint a compelling picture of the candidate to the adcom.
+
 Having a strong letter from an internationally recognized researcher will \emph{greatly strengthen} your application.
 In fact, a strong letter from a well-known researcher can outweigh the lack of publications (or low GPA or unknown school).
-
 However, obtaining letters from famous people
@@
 
-\paragraph{Open House} Most schools will have \textbf{Open House} or visit events, which are a great resource to learn about the school, department, faculty, research, living, etc.
-Even if you can't come in person, you should attend virtually and meet with individual faculty.
-During the Open House, you get a chance to talk to individual faculty and current students.
+\paragraph{Open House} Most schools have \emph{Open House} or \emph{Visit Day} events, which are a great resource to learn about the school, department, faculty, research, living, etc.
+\textbf{Advice}: Even if you can't come in person, you should attend virtually and meet with individual faculty.
+During the event, you get a chance to learn more about the program, talk to individual faculty and current students.
 Take notes of faculty who make you excited, count those that are taking in new students (if they meet you, likely they are considering new students!).
@@
 
-Also, do not forget to update and thank your LoR writters and others who have supported you throughout your admission journey.
+Also, do not forget to update and thank your LoR writters and others who have supported you through this process.
 
@@
 Graduate admission can involve randomness and noise.
-Don't bother asking for feedback, you will not likely get any.
 In the meantime, you can work on improving your profile, e.g., get more research experiences, publish more papers, improve your connections for better LoR writers, etc.
+See rejection reasons (\autoref{sec:why-rejected}) and additional advice ()\autoref{sec:improve-your-chance}).
 
@@
 If you do well, you can ask the professor to support you to convert to PhD.
-This is not uncommon for many students.
+
+\paragraph{Should I ask for feedback?}
+No, don't bother.
+You will likely not get any useful feedback.
+We are not willing to reveal your evaluation results or give you advice on how to improve.
+So it is best to just move on.
+If you really want feedback, ask your professors, collaborators, ref writers, or those who have previously applied.
 
@@
 \section{Why did I get rejected?}\label{sec:why-rejected}
-Many possible reasons!  Other than not having a strong profile (e.g., research potentials, GPA, LoRs, SOP), here are some that you might not expect:
-
-\begin{itemize}
-
-\item \emph{You are not a good fit to the program}.  You could have an excellent profile, but if you are interested in a research area that the program does not have, you might not be admitted.
-Similarly, if there is no faculty who is willing to advise you (e.g., they are on sabbatical, personal leave, or already have too many students already), you will not be admitted.
+Many possible reasons!  Other than not having a strong profile (e.g., research potentials, GPA, LoRs, SOP, poor interviews), here are some that you might not expect:
+
+\paragraph{Not a good fit}  You could have an excellent profile, but if you are interested in a research area that the program does not have, you might not be admitted.
+Similarly, if there is no faculty who is willing to advise you (e.g., they are on sabbatical, personal leave, do not have sufficient fundings, or already have too many students already), you will not be admitted.
 This is actually good for you, as you don't want to be in a program where no one can advise you.
@@
 
-\item  \emph{Your English scores (e.g., IELTS or TOEFL) are not high enough}.
-While professors might not care much about these, there is a certain minimum set by the college or the university that you need to pass to get admitted, especially for TA funding (e.g., you need to be able to communicate well with students as a TA).
+\paragraph{Low English exam scores (e.g., IELTS or TOEFL)}  While professors might not care much about these, there might be a certain minimum set by the college or the university that you need to pass to get admitted, especially for TA funding (e.g., you need to communicate well with students as a TA).
 Thus, while professors are willing to argue for your case, they might be reluctant to go against the standard requirement of the university on English scores.
-Note this is mostly for international students, as domestic ones (or those from these \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-GMU#standard-tests-waiver-eligible-countries}{countries}) do not have to take these tests.
-
-\end{itemize}
-
-\paragraph{Should I ask for feedback?}
-No, don't bother.
-You will likely not get any useful feedback.
-We are not willing to reveal your evaluation results or give you advice on how to improve.
-So it is best to just move on.
-If you really want feedback, ask your professors, collaborators, ref writers, or those who have previously applied.
+Note this is for international students, as domestic ones (or those from these \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-GMU#standard-tests-waiver-eligible-countries}{countries}) do not have to take these tests.
+
+\paragraph{Overqualified or Lack Interest}  This might be paradoxical, however if the adcom believes that you are likely to get admitted and go to a better program (i.e., they are just your ``safety'' schools), they might not be so enthusiastic to admit you and want to save the spot for someone else.
+Note though it does happen, for most good schools this is rare and you should not assume this is the case.
+
+A related reason is that you did not show enough interest in the program.
+For example, your SOP did not mention any faculty or research area at the program, or you did not respond to professors interested in interviewing you (which might be considered unprofessional and burn bridges).
+
+\paragraph{Red flags in Your Background} Plagiarism, cheating, or other academic dishonesty are red flags.
+Another one is if you have a history of jumping from one program to another (e.g., you have been in multiple different programs in short time).
+If you think you have red flags, you should address them in your SOP.
+In general, these things are rare, but they do happen and cause concern to the adcom.
+
+  %\item \emph{You did not make a good impression}.
+For example, you might have contacted a professor and they did not like your email.
+Or you might have met a professor (e.g., at a conference) and they did not like your presentation.
+This is rare, but it happens.
 
@@
 
-\section{Buying Computer Equipment}\label{sec:buying-equipment} understandably students get very excited about their upcoming PhD journal and want to buy new computer equipment and electronics to prepare.
+\section{Buying Computer Equipment}\label{sec:buying-equipment}
+
+Understandably students get excited about their upcoming PhD journey and want to buy new computer equipment and electronics to prepare.
 However, you should check with your adviser first.
@@
 
+\paragraph{Standing out \autoref{sec:improve-your-chance}} There are \emph{few} domestic applications compared to international ones, i.e., domestic students are minority (\autoref{sec:urm}) in the CS PhD application pool.
+Many US universities want to increase the number of domestic students in their programs (and as mentioned above, there are specific fellowships for domestic students).
+That makes you stand out and can help your case.
+
 \paragraph{Fee Waiving} Some school might offer application fee waiver for domestic students.  You should check with the school you're applying to.
@@
 
-\paragraph{Transcripts} You do not need to get your transcripts evaluated (which can be a hassle and inaccurate).
+\paragraph{Transcripts} You do not need to get your transcripts evaluated/translated (which can be a hassle and inaccurate).
 You can just send your official transcripts directly to the school you're applying to.
@@
 Highlight such experiences in your application.
-
-\paragraph{Standing out \autoref{sec:improve-your-chance}} There are \emph{very few} domestic applications compared to international ones, i.e., domestic students are minority (\autoref{sec:urm}) in the CS PhD application pool.
-Many US universities want to increase the number of domestic students in their programs (and as mentioned above, there are specific fellowships for domestic students).
-That makes you stand out and can help your case.
 

### 2024-03-20 demystify.tex
updates

--- 2024-03-17 demystify.tex
@@
-\documentclass[oneside,11pt,dvipsnames]{book}
+\documentclass[oneside,11pt]{book}
 
@@
 \usepackage{booktabs}
-
+\usepackage[dvipsnames]{xcolor}
 \usepackage{tikz}
 \usetikzlibrary{calc} 
-\usepackage{xcolor}
-
 \usepackage{anyfontsize}
@@
         \begin{itemize}
-          \item Applications are evaluated by the \emph{PhD Admission} (adcom) committee and each application is typically reviewed by three faculty (\autoref{sec:evalapps}).
+          \item Applications are evaluated by the \emph{PhD Admission} (adcom) committee and each application is reviewed by typically three faculty (\autoref{sec:evalapps}).
           \item Individually faculty \emph{cannot directly admit} a student---so do not email and ask if you have a chance.
@@
           \item The committee will look at various factors, but the most important ones are research ability, e.g., publications, statement of purpose (SOP), and  letters of recommendation (LORs).
-          \item \emph{LORs are very important}, but only if they are personalized and talk about your research ability (\autoref{sec:lor}).
-          \item \emph{SOP is very important}. Write it in such a way that makes you \emph{stand out} (\autoref{sec:research-statement} and \autoref{sec:improve-your-chance})
-          \item  \emph{Pubication can greatly help}.
-Publications are good venues are concrete evidence that you have successfully engaged in research.
-But they are not required!
-(\autoref{sec:research-ability}).
-
-          \item \emph{Grades might be important}, depends on the reputation of \emph{your school} (\autoref{sec:your-school}).
-While good grades might not help much, bad ones will likley hurt your case.
-          \item \emph{Standard tests are not important} (\autoref{sec:standard-tests}).
-GRE \emph{is not} even required.
-Spend your time on something else!
-For standard English tests (not required for domestic students), just do enough to pass the minimum requirements.
-
-          \item \emph{Contacting a prof. for is recommended}, but do it \emph{properly} (\autoref{sec:contact}).
-
+          \item LORs are very important, but only if they are personalized and talk about your research ability (\autoref{sec:lor}).
+          \item SOP is very important. Write it in such a way that makes you \emph{stand out} (\autoref{sec:research-statement} and \autoref{sec:improve-your-chance})
+          \item Grades are important, but depend on the reputation of your school (\autoref{sec:your-school}).
+          \item GRE \emph{is not} required (\autoref{sec:standard-tests}). Spend your time on something else!
           \item Getting an interview is typically a \emph{good sign}; but no interview does not mean rejection (\autoref{sec:interview}).
@@
           \item Many schools do not offer PhD studies in CS and many CS professors do not advise or graduate PhD students  (\autoref{sec:schoolsandprofs}).
+          \item Contacting a prof. for research opportunities is recommended, but do it \emph{properly} (\autoref{sec:contact}).
         \end{itemize}
@@
 did your reference writers submit their letters?
-Usually, this screen process is done through a central university system, i.e., not by faculty from the CS department.
+Usually, this screen process is done through a central university system, i.e., not by the CS department.
 If you pass this screen, your application will be forwarded to the CS department for further evaluation.
@@
 
-Then, applications are reviewed by a \emph{PhD admission committee} (\textbf{adcom}) that consists of faculty members in CS\footnote{In some cases the committee can involve affiliated faculty from different disciplines.}.
-These adcom members have a wide-range of expertise and background to ensure diverse perspectives in the evaluation process.
+Then, applications are reviewed by a \emph{PhD admission committee} (\textbf{adcom}) that consists of faculty members in CS (in some cases the committee can involve affiliated faculty from different disciplines).
+These faculty have a wide-range of expertise and background to ensure diverse perspectives in the evaluation process.
 The size and the review load of the adcom depends on the department size.
 At GMU, the PhD adcom typically has 15--20 faculty, and each committee member is assigned to review 25--30 applications.
-Note that most large schools, including GMU, have separate adcoms for MS programs.
-
-Each application is assigned to about 3 faculty members, who will evaluate your profile and reach a consensus.
-While the assigned reviewers are the main ones deciding your application, other faculty in the department will likely have access to your application and can provide inputs and opinions on your profile.
-
-The PhD adcom typically involves assistant professors in the department (see \autoref{sec:faculty-types} for various types of faculty).
+Note that many schools, including GMU, have separate adcoms for MS programs.
+
+Each application is assigned to 2--3 faculty members, who will evaluate your profile and reach a consensus.
+Note that while the assigned reviewers are the main ones deciding your application, other faculty in the department will likely have access to your application and can provide inputs and opinions on your profile.
+
+The PhD adcom typically involves assistant professors in the department (see \autoref{sec:faculty-types} for various faculty titles).
 This provides junior faculty the opportunities to recruit students.
-Some members are tenured faculty.
+Some members can be tenured faculty.
 The chair of the committee will likely a senior faculty, but they will not review individual applications and instead assign them to committee members.
-The chair will look at various factors such as research interests or mentioning faculty names to assign the applications to appropriate faculty, e.g., I am often assigned to review applicants interested in software engineering.
+The chair will look at various factors such as research interests or mentioning faculty names to assign the applications to appropriate faculty.
 
@@
 
-  Several international students mentioned that some professors are unwilling to write letters or write weak ones because they do not want (good) students to go abroad or only go to places where they want the students to go to.
+  Another sad reality is that some professors are unwilling to write letters or write weak ones because they do not want (good) students to go abroad or only go to places where they want the students to go to.
 If you are in this situation, you should find someone else to write for you.
@@
   Sometimes students would go through great length just to get letters from well-known professors in their school (e.g., department head).
-But as mentioned, if these professors do not know you, their letters would likely be generic and carry little value, in fact, \red{red flags}.
-Moreover, a top professor at your university might not be well-known to US faculty (see more details in \autoref{sec:your-school}).
-So save the trouble and  get letters from \emph{any} professors/supervisors who knows you well and can write a good letter about \emph{your} research ability.
+But as mentioned, if these professors do not know you, the letters are generic and carry little value, in fact, \red{red flags}.
+Moreover, a top professor in Vietnam might not be well-known to US faculty (see more details in \autoref{sec:your-school}).
+So save the trouble and just get letters from \emph{any} professors/supervisors who knows you well and can write a good letter about \emph{your} research ability.
 It's better to have a good personalized
@@
 
-\textbf{Advice}: You can explain about your school in your SOP (and also ask your LOR writer to do that).
+What can you do about this?
+You can explain about your school in your SOP (and also ask your LOR writer to do that).
 Of course, if you're interested in working with Vietnamese, consider  \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{CS programs in the US that have Vietnamese professors}.
@@
 usually will not help because we cannot evaluate them.
+
 This can be an issue for students in many top international universities where the competition is so high that very good students can still have low rankings from these schools (and be overlooked by the adcom).
-
-\textbf{Advice}: As with school reputation, you can put a note about this in your statement and ask your LoR writers to talk about it.
+So what to do with this? As mentioned in \autoref{sec:your-school} you can put a note about this in your statement and ask your LoR writers to talk about it.
+
 Note that while having good grades at unknown school might not help,
-having bad grades (e.g., < 3.0) will be \red{red flag} (unless your LORs or
+having bad grades will be \red{red flag} (unless your LORs or
 statement give a proper explanation). This is especially true if you
-have bad grades in relevant courses, e.g., Math and CS.
+have bad grades in relevant courses (e.g., Math and CS).
 
@@
 
-\paragraph{GRE} Most good CS PhD programs in the US (e.g., R1 universities) \textbf{no longer require GREs}, so you don't need to take them.
-
-\textbf{Advice}: In general, don't waste your time taking GRE.
-However, if you took it and have good scores then you can include and talk about them in your application (but don't expect them to make much difference).
-
+\paragraph{GRE} Most good CS PhD programs in the US \textbf{no longer require GREs}, so you don't need to take them.
+However, they might be useful for international students from programs we are not familiar with.
+If you have good GRE scores then you should include them in your application (but don't expect them to make much difference).
 \paragraph{English Test} Unless your degrees are from specific countries such as \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-GMU#standard-tests-waiver-eligible-countries}{these}, you will need to
 take standardized English test.
-
-\textbf{Advice}: Just do well enough to pass minimum requirement set by the university, which is required for TA (\autoref{sec:funding}).
-Just as with grades (\autoref{sec:your-school}) and GRE, having a high scores in these tests might not help, but having too low scores can a \red{red flag} and sometimes results in an automatic rejection (\autoref{sec:why-rejected}), e.g., below the minimum requirement.
+Just do well enough to pass minimum requirement set by the university, which is required for TA (\autoref{sec:funding}).
+Just as with grades, having a high school in these tests might not help, but having too low scores can a \red{red flag} and sometimes results in an automatic rejection (\autoref{sec:why-rejected}), e.g., below the minimum requirement.
 
@@
 
-\paragraph{Work Experiences} Work experiences, especially if they are not research-based, do not carry much weight. Similarly, LoR from your supervisor for non-research experience will not count much.
-
-\textbf{Advice}: do not spend much time talking about non-research related job in your SOP.
+\paragraph{Work Experiences} Work experiences, especially if they are not research-based, do not carry much weight.
+So do not spend much time talking about them in your SOP (and similarly, LoR from your supervisor for non-research experience will not count much).
+
 However, if you have research-based work experiences, then they can be useful and should be mentioned in your SOP and LORs.
@@
 \section{CV/Resume}
-The CV/Resume provides a summary of the applicant's achievements.
-
-\textbf{Advice}: Prepare your CV in such away that it allows the reviewers to quickly scan to identify major achievements (e.g., Publications, Programming Competition Awards, Teaching Experience).
+The CV/Resume should provide a summary of the applicant's achievements.
+It should allow the reviewers to quickly scan to identify major achievements (e.g., Publications, Programming Competition Awards, Teaching Experience).
 
@@
 Typically, an interview takes about 15--30 minutes, and one important aspect of evaluation is your ability to effectively communicate, including speaking and understanding English. 
-
-\textbf{Advice}: In general, just treat the interview as an informal chat, but be prepared to talk about your research experiences and interests (so do have \emph{elevator pitch} about your experiences and research).
+In general, just treat the interview as an informal chat, but be prepared to talk about your research experiences and interests (so do have \emph{elevator pitch} about your experiences and research). 
 A professor might want to know more about your research background or even ask you to read a paper (so be sure to do it!).
 Some prof.
-can ask you to solve a problem, e.g., a coding problem, so you might want to ask if you need to prepare for that.
+can ask you to solve a problem, e.g., a coding problem (so you might want to ask if you need to prepare for that).
 During the interview, you will have the opportunity to ask questions about the university. It's essential to prepare some thoughtful questions, similar to a job interview.
@@
 For external fellowships including those from the US government and tech companies (e.g., Google, Microsoft, Facebook), you will need to apply.
+Fellowships from US government would likely require US citizenship or residency.
 % However, tech companies including Google, Microsoft, Facebook have fellowships that international students can apply for.
@@
 
-In addition to general fellowships that are available to all students, there are also specic fellowships, e.g., for US citizen and permanent residents, for underrepresented minority (URM) students (\autoref{sec:urm}), and for students from specific countries.
-In short, there are many fellowships available, so you should look for them and apply.
-
-In general, fellowships, especially those that are open to everyone, are highly competitive and prestigious, and you will stand out if you get one.
+In general, fellowship, especially competitive ones from the government or company, is prestigious, and you will stand out if you get one.
 Every PhD student has pubs, but only few  would have NSF Graduate or Microsoft fellowship.
@@
 While this can be confusing due to the large number of universities in the US, a little research, e.g., searching for PhD in CS from the school website, will help you find out.
-Schools listed in \autoref{sec:ranking} have PhD programs in CS, so you can start there.
+All schools listed in \autoref{sec:ranking} have PhD programs in CS, so you can start there.
 
@@
 You will be very surprised to learn that a school that you didn't know much about can have very strong research in your interested topic (and vice versa, a school you thought highly about might have no faculty working in the research field you're interested in).
-This is also a good way to learn about individual faculty (who works on what) and well-known CS conferences\footnote{In CS (and probably only in CS), conferences, not journals, are often the main venue to publish research finding (see \href{https://homes.cs.washington.edu/~mernst/advice/conferences-vs-journals.html}{why here}).}.
+This is also a good way to learn about individual faculty (who works on what) and well-known CS conferences\footnote{In CS (and probably only in CS), conferences, not journals, are often the main venue to publish research finding.}.
 %\autoref{sec:ranking} gives the top 50 CS programs in the US according to CSRankings.
@@
 It gives both the professor and student more time, e.g., a whole semester, to work and evaluate the relationship before making any commitment (sounds a bit like a marriage!).
-
 You can also ask if you can do an independent study or research with a prof.
@@
 Ultimately, choose a prof.
-that fits you by communicating with them, taking their courses, meeting and asking them questions, and talking to their current students.
+that fits you the most by communicating with them, taking their courses, meeting and asking them questions, and talking to their current students.
 It will take time and effort, but since you will be working with this person for 5+ years, it is important to try to find the right one.
@@
 
-On the other hand, if you want to contact a prof.
+If you want to contact a prof.
 to ask about \emph{research opportunities}, or \emph{GTA/GRA} support, then \emph{yes}, I believe you should---it is \emph{worth it}.
@@
 
-Below is a good example that I would likely reply.
+Below is a good example that I would reply to.
 
@@
 
-You \emph{do not} need to do a PhD in CS to work in your area of interest.
+In many cases you do not need to do a PhD in CS to work in your area of interest.
 For example, in addition to a traditional CS department, GMU has IST and Cybersecurity departments, both of which have faculty  with PhD in CS and work on CS topics (e.g., AI, Security, Robotics).
@@
 If you really want feedback, ask your professors, collaborators, ref writers, or those who have previously applied.
-
-% \section{How much time did you spend looking at an application?}\label{sec:time-spent}
-% Short answer: less than 5 mins for quick rejection; 15+ mins for an acceptance.  On average, about 10 mins.
-
-% The GMU evaluation system allows us to download a single pdf that consists of a summary page (e.g., school info, GPA, testscores), then transcripts, standard test results, LoRs, CV, SOP, writing samples So I typically read the application in this order.
-Then I make notes on what I've seen, basically evidence that support my evaluation.
-
-% I first look at the summary for school, GPA, est scores (essentially looking for very low GPA or testscores that do not meet the minimum requirements.
-Then I skim on transcripts look for low/failing grades on CS and related courses.
-For these I make comments such as "unknown international school with good GPA, too many low grades on CS and related courses" or "good GPA from top school XYZ".
-I do not look at details on tests (the main scores are already given in the first summary page).
-
-% Then I look at LoRs.
-I read carefully good LoRs and quickly skim over bad or generic ones (very easy to tell, you don't even need much experiences to distinguish between a strong LoRs from the rest).
-Comment examples For LoRs include "very strong LoRs from well-known prof.
-XYZ talking about research experiences" or "generic LoRs, mostly about course work".
-I also note any unusual things in the LoRs.
-
-% Then the SOP, which I read carefully (of course if it is bad then it would a quick read).
-Then my notes for SOP include things such as "SOP mostly on course work, no clear motivation, generic" , or "very enjoyable SOP talking about XYZ, customized for GMU, mentioned profs.
-X,Y,Z".
-If the SOP mentioned about publications then I'll make notes such as "has X papers at top venues" or "has X papers but at unknown places".
-I will not go online search for the papers (unless they upload them as sample papers) .
-
-% I also skim over sampling papers, essentially looking over the abstract to see if reads well.  I would make notes such as "have a paper under submission to unknown venue, but reads well".
-
-%At GMU we use an evaluation system that allows us to download a single, large pdf that consists of everything submitted by the applicant, e.g., GPA, transcripts, test scores, CV, LoRs, SOP, sample writings, etc.
-The first couple of pages show the summary including school, major, GPA, test scores.
 

### 2024-03-20 demystify.tex

--- 2024-03-17 demystify.tex
@@
 But they are not required!
-(\autoref{sec:research-ability}).
+(\autoref{sec:research-experiences}).
 
@@
 
-\section{Research Ability}\label{sec:research-ability}
-
-A concrete evidence of research ability is having \textbf{papers in reputable international journals or conferences}.
-Having published papers, especially at top venues, is a sign that the applicant has successfully involved in research.
-Students  accepted to top schools often have multiple first-authored papers at top venues.
-
-\begin{commentbox}[Vu:]
-  Many international students mention Scopus Q1, which consists of various journals from IEEE, Elsevier, and many other publishers.
-I don't know/recognize many of journals listed in Scopus Q1.
-This might be something to be mindful of, as \textbf{CS} faculty might not be too familiar with Scopus or journals listed there, so devote some part in your statement to discuss the significance of your papers.
-\end{commentbox}
-
-However, it is understandable that many students do not have the opportunities to publish papers.
-Thus, general writings, even those under submissions or rejected, would still be good (and much better than nothing).
-But be sure to upload your papers with your application and talk about them in your statement (see \autoref{sec:research-statement}).
-Note that local conferences and non-English journals or conferences do
-not carry as much weight since their quality is often unknown to US faculty.
-However, if you have published in such places, you should still upload them, mention them in your statement, and explain why they are good.
-
-% In CS, conferences are important and top conferences are prestigious.
-You can find top CS conferences at places such as \href{https://csrankings.org}{CSRankings}, which ranks CS programs based on how their faculty publish at top \emph{conferences} (see more in \autoref{sec:ranking}).
-
-\begin{commentbox}[Craig:]
-  GMU and many other universities allow you to upload your published papers and other writing samples.
-In many cases, even if the papers were not published at top places, we can still determine their quality by simply skimming over the paper.
-\end{commentbox}
-
-Additionally, \textbf{work experiences at well-known research laboratories}, such as Microsoft Research, can strengthen your
-application.
-Unfortunately, many good research places in your countries, e.g., VinAI in Vietnam, remain relatively unknown to most universities in the US.
-So you or your LoR writers should explicit say something about them in your statement.
-
-\begin{commentbox}[Hung:]
-  The reputation of VinAI has been increasing steadily over the past few years; many of my colleagues heard about VinAI.
-\end{commentbox}
-
-Finally, \textbf{participating internationally recognized competitions} can also demonstrate your research potential.
-For example, participating in Math Olympiads if you want to do theory or  winning ACM programming contests if you want to ``build'' systems, e.g., software analysis.
-
-\begin{commentbox}[Thanh:]
-  Due to academic culture, professors in Vietnam usually aim for (international) journals instead of conferences.
-Could you give some tips on how to know whether a journal is good (CSrankings, unfortunately, only consider conferences)?.
-  \tcblower
-  Vu: One way is looking at what well-known researchers publish at.
-For example, if you are interested in a field X, you can use CSRankings to look at active faculty in X, and then look at their websites to see what journals they publish at.
-\end{commentbox}
-
 \section{Letters of Recommendation (LOR)}\label{sec:lor}
@@
 
+\section{Research Experiences}\label{sec:research-experiences}
+
+\paragraph{Publications} A concrete evidence of research ability is having \emph{papers in reputable international journals or conferences}.
+Having published papers, especially at top venues, is a sign that the applicant has successfully involved in research.
+Students  accepted to top schools often have multiple first-authored papers at top venues.
+
+\begin{commentbox}[Vu:]
+  Many international students mention Scopus Q1, which consists of various journals from IEEE, Elsevier, and many other publishers.
+I don't know/recognize many of journals listed in Scopus Q1.
+This might be something to be mindful of, as \textbf{CS} faculty might not be too familiar with Scopus or journals listed there, so devote some part in your statement to discuss the significance of your papers.
+\end{commentbox}
+
+However, it is understandable that many students do not have the opportunities to publish papers.
+Thus, general writings, even those under submissions or rejected, would still be good (and much better than nothing).
+But be sure to upload your papers with your application and talk about them in your statement (see \autoref{sec:research-statement}).
+Note that local conferences and non-English journals or conferences do
+not carry as much weight since their quality is often unknown to US faculty.
+However, if you have published in such places, you should still upload them, mention them in your statement, and explain why they are good.
+
+% In CS, conferences are important and top conferences are prestigious.
+You can find top CS conferences at places such as \href{https://csrankings.org}{CSRankings}, which ranks CS programs based on how their faculty publish at top \emph{conferences} (see more in \autoref{sec:ranking}).
+
+\begin{commentbox}[Craig:]
+  GMU and many other universities allow you to upload your published papers and other writing samples.
+In many cases, even if the papers were not published at top places, we can still determine their quality by simply skimming over the paper.
+\end{commentbox}
+
+\paragraph{Research Experiences} Work experiences at \emph{well-known research laboratories}, such as Microsoft Research, can strengthen your
+application.
+Unfortunately, many good research places in your countries, e.g., VinAI in Vietnam, remain relatively unknown to most universities in the US.
+So you or your LoR writers should explicit say something about them in your statement.
+
+\begin{commentbox}[Hung:]
+  The reputation of VinAI has been increasing steadily over the past few years; many of my colleagues heard about VinAI.
+\end{commentbox}
+
+\paragraph{Competitions} Winning \emph{internationally recognized competitions} can also demonstrate your research potential.
+For example, participating in Math Olympiads if you want to do theory or  winning ACM programming contests if you want to ``build'' systems, e.g., software analysis.
+
+\begin{commentbox}[Thanh:]
+  Due to academic culture, professors in Vietnam usually aim for (international) journals instead of conferences.
+Could you give some tips on how to know whether a journal is good (CSrankings, unfortunately, only consider conferences)?.
+  \tcblower
+  Vu: One way is looking at what well-known researchers publish at.
+For example, if you are interested in a field X, you can use CSRankings to look at active faculty in X, and then look at their websites to see what journals they publish at.
+\end{commentbox}
+
 \section{Your School and Grades}\label{sec:your-school}
@@
 
-\section{Standard Tests (GREs, IELTS, etc) and Work Experiences}\label{sec:standard-tests}
+\section{Standard Tests (GREs, IELTS, etc), Work Experiences, Online Courses, and Certificates}\label{sec:standard-tests}
 
@@
 
-\paragraph{Work Experiences} Work experiences, especially if they are not research-based, do not carry much weight. Similarly, LoR from your supervisor for non-research experience will not count much.
+\paragraph{Work Experiences} Work experiences, especially if they are not research-based, \emph{do not} carry much weight.
+Similarly, LoR from your supervisor for non-research experience will not count much.
 
@@
 However, if you have research-based work experiences, then they can be useful and should be mentioned in your SOP and LORs.
-For example, if you have worked at a research lab (e.g., Microsoft, DoE, VinAI in Vietnam, etc) (\autoref{sec:research-ability}), then you should mention that in your SOP and ask your supervisor to write about your research experiences and potentials in their LOR (\autoref{sec:lor}).
+For example, if you have worked at a research lab (e.g., Microsoft, DoE, VinAI in Vietnam, etc) (\autoref{sec:research-experiences}), then you should mention that in your SOP and ask your supervisor to write about your research experiences and potentials in their LOR (\autoref{sec:lor}).
+
+\paragraph{Online Courses and Certificates}
+In general, these do not carry much weight as they do not show your research ability.
+We do not care much if you have taken on online Coursera AI course or have a professional certificate from Microsoft.
+However, as mentioned in \autoref{sec:non-stem}, if you do not have a CS background, you can use these to show your ability to handle CS courses.
 
@@
 
-\textbf{Advice}: Prepare your CV in such away that it allows the reviewers to quickly scan to identify major achievements (e.g., Publications, Programming Competition Awards, Teaching Experience).
+\textbf{Advice}: Prepare your CV in such away that it allows the reviewers to quickly scan to identify major achievements, e.g., Publications, Programming Competition Awards, Teaching Experience.
 
@@
 
-\paragraph{Not getting interviews} While it is generally good to get interview, not getting an interview does not mean you're out.
-Many programs do not have the tradition of interviewing applicants.
-For example, at GMU, most of our admitted students with TA (\autoref{sec:ta}) do not go through interviews.
-
-However, no interviews typically means that you will not likely get an RA (\autoref{sec:ra}), which is offered by an individual faculty (if they want to pay you to do research for them, then likely they want to talk to you first).
-In this case, your application (and TA/fellowship funding) will be reviewed and decided by the adcom.
-
 \paragraph{Follow Up Emails} If you had an interview and have not heard back, you can send an email to ask about the status of your application Don't send this too soon though, perhaps wait at least 2 weeks.
@@
 However, if you have new publications or other big achievements, you can send an email asking to update your profile.
+
+\paragraph{Not getting interviews} While it is generally good to get interview, not getting an interview does not mean you're out.
+Many programs do not have the tradition of interviewing applicants.
+For example, at GMU, most of our admitted students with TA (\autoref{sec:ta}) do not go through interviews.
+
+However, no interviews typically means that you will not likely get an RA (\autoref{sec:ra}), which is offered by an individual faculty (if they want to pay you to do research for them, then likely they want to talk to you first).
+In this case, your application (and TA/fellowship funding) will be reviewed and decided by the adcom.
 
@@
 In the meantime, you can work on improving your profile, e.g., get more research experiences, publish more papers, improve your connections for better LoR writers, etc.
-See rejection reasons (\autoref{sec:why-rejected}) and additional advice ()\autoref{sec:improve-your-chance}).
+See rejection reasons (\autoref{sec:why-rejected}) and additional advice (\autoref{sec:improve-your-chance}).
 
@@
 As discussed in \autoref{sec:undergrad-research}, research experience is a crucial component of a strong PhD application.
-It gives you opportunities to try out research, determine what research area you're interested in, publish papers (\autoref{sec:research-ability}), and get strong LoRs (\autoref{sec:lor}).
+It gives you opportunities to try out research, determine what research area you're interested in, publish papers (\autoref{sec:research-experiences}), and get strong LoRs (\autoref{sec:lor}).
 

### 2024-03-20 ms-swe-handbook.tex
Updates from Overleaf

@@
 
-\title{Demystifying the Computer Science \\
-  PhD Admission in the US\\{\large A Guideline for International \emph{and Domestic} Students}}
+\title{GMU MS SWE Handbook\\{\large A Handbook For the Masters of Software Engineering program at George Mason University}}
 
@@
 \begin{itemize}
-\item COMP 501 Computer Programming Foundations I or SWE 510 Object Oriented Programming in Java
-\item COMP 502 Mathematical Foundations of Computing I or INFS 501 Discrete and Logical Structures for Information Systems
-\item COMP 503 Computer Systems Foundations I or INFS 515 Computer Organization
-\item COMP 511 Computer Programming Foundations II or INFS 519 Program Design and Data Structures
+\item \textbf{COMP 501} Computer Programming Foundations I or \textbf{SWE 510} Object Oriented Programming in Java
+\item \textbf{COMP 502} Mathematical Foundations of Computing I or \textbf{INFS 501} Discrete and Logical Structures for Information Systems
+\item \textbf{COMP 503} Computer Systems Foundations I or \textbf{INFS 515} Computer Organization
+\item \textbf{COMP 511} Computer Programming Foundations II or \textbf{INFS 519} Program Design and Data Structures
 \end{itemize}
@@
 
-Students are required to complete 30 credits corresponding to 10 graduate courses.
+Students are required to complete \textbf{30 credits} corresponding to \textbf{10 graduate courses}.
 Students are encouraged to download the course planner spreadsheet, and update it as they proceed through the program.
 
-The course work is divided into three categories: 12 credits of core courses, software engineering related courses of 9 credits, and 9 credits of elective courses.
+The course work is divided into three categories: 12 credits of core courses, 9 credits of SWE related courses, and 9 credits of elective courses.
 
@@
 \begin{enumerate}
-    \item SWE 619 Object-Oriented Software Specification and Construction
-    \item SWE 621 Software Modeling and Architectural Design
-    \item SWE 632 User Interface Design and Development
-    \item SWE 637 Software Testing
+    \item \textbf{SWE 619} Object-Oriented Software Specification and Construction
+    \item \textbf{SWE 621} Software Modeling and Architectural Design
+    \item \textbf{SWE 632} User Interface Design and Development
+    \item \textbf{SWE 637} Software Testing
 \end{enumerate}
@@
 \begin{itemize} 
-\item Software Engineering (SWE)
+\item Software Engineering
 \begin{itemize}
@@
 \end{itemize}
-\item Computer Science (CS)
+\item Computer Science
 
 \begin{itemize}
-    \item CS 540 Language Processors
+    \item CS 540 Language Processors (Compilers)
     \item CS 550 Database Systems
@@
 
-\chapter{Obtaining Funding}
-\section{NSF and other Government Funding}
-
-\paragraph{Other Government Funding}
-
-\section{Internal Funding}
-
-\section{Industrial Gifts}
-
-\chapter{Teaching}
-\section{Teaching Load}
-\section{Teaching Evaluation}
-\section{Teaching Philosophy}
-\section{Teaching Awards}
-
-\chapter{Mentoring}
-\section{Mentoring Undergraduate Students}
-\section{Mentoring Graduate Students}
-\section{Mentoring Postdocs}
-\section{Mentoring Junior Faculty}
-
-\chapter{Services}
-
-\section{Service to your Professional Community}
-\section{Service to the Department}
-\section{Service to the University}
-\section{Service to the Community}
-
 \chapter{History and Acknowledgement}\label{sec:ack}

### 2024-03-21 faculty.tex
Updates from Overleaf

@@
 \section{The 5 jobs of an assistant professor}
-An assistant professor in CS (and likely in other fields) At an R1 university typically has \emph{five} components:
+An tenure-track professor in CS (and likely in other fields) at an R1 university typically has \emph{five} jobs:
 
@@
 
-\chapter{History and Acknowledgement}\label{sec:ack}
-
-\paragraph{History} This document was conceived during a lunch with Craig Yu at GMU.
-We talked on about why GMU were not able to attract good Vietnamese and other international students, despite having a much stronger CS program than many schools that these students want to go to (part of the reason is described in \autoref{sec:selecting-ranking-schools}).
-We wished there were a way for international students to know about the US PhD programs (also for US faculty to understand more about international students and therefore have better chance of recruiting and working with them).
-I was also a member of the large VietPhD group on Facebook and saw many questions from students about PhD programs.
-However, most active participant are students in non-CS fields or not in US, and their answers are unfortunately not always accurate and sometimes leading to more confusion.
-So I thought it would be useful to have a document that is specific to CS PhD programs in the US from an insider prospective.
-
-I started writing this document in May 2023 and have been updating it since then (mostly around deadline time when I tend to procrastinate, i.e., \emph{productive procrastination}!).
-I have put the source code of this document on \href{https://github.com/nguyenthanhvuh/phd-cs-us}{GitHub} so that anyone can contribute to it.
-
-\paragraph{About Me} I am an assistant professor in the CS dept at George Mason University (GMU).
-Prior to GMU, I was at the University of Nebraska-Lincoln (UNL).
-I have been in the PhD admission process at GMU and UNL for many years.
-Currently I serve as the program director of the MS program in Software Engineering at GMU (thus also have some experience with the MS admission process--which is quite different than PhD).
-My personal and lab website is at \href{https://dynaroars.github.io}{dynaroars.github.io}.
-
-Though I'm not an international student, many of my students and collaborators are.
-I also mentor multiple students from Vietnam, and have close colleagues and friends who were once international students.
-I hope to capture the diverse challenges and experiences they've faced in this document, so that it  can be a valuable resource for prospective international students.
-Finally, my upbringing in the US provides a perspective aligned with American culture, allowing me to shed light on various issues, particularly those related to cultural differences (\autoref{sec:cultural}).
-
-\paragraph{Acknowledgement} Many people have contributed to this document.
-Profs. Craig Yu (GMU), Hakan Aydin (GMU), 
-Xiaokuan Zhang (GMU), Hung Le (UMass), and Deepak Kapur (UNM) provided valuable input in the early version.
-Other GMU faculty members also have provided feedback and contributions.
-Many students including Didier (GMU), Thanh (Melbourne), and Dat (Melbourne) have contributed valuable questions and feedback.
-\textbf{Thank you!}
-
 \bibliographystyle{abbrv}

### 2024-03-21 faculty.tex
updates

@@
 \section{Service to your Professional Community}
-\section{Service to the Community}
+
+These are services that you do for your professional community.
+They will help you to get to know people in your field, and to get your name out.
+Common examples include serving on the program committees of conferences, serving on editorial boards of journal, reviewing papers for journals, serving on grant panels.
+
+\paragraph{Serving on Program Committees (PC)} This is a good way to get to know what is hot in your field, and to get to know people in your field.  It is also a good way to get your name out.
+You should try to serve on the program committee of the top conferences in your field.
+
+\textbf{Advice}: You can ask senior colleagues or even your PhD advisor/collaborators to recommend you to serve on a PC.
+You can also directly email the PC chairs and ask to serve on the PC.
+For example, I directly email OOPSLA and PLDI PC chairs and asked to serve on the PC, and they accepted my request (I do have papers at OOPSLA and PLDI when I asked).
+I tried this with SE conferences  e.g., ICSE (and of course also nominate myself through the official form), and it never work, even though I have many papers at ICSE.
+
+\section{Serving on NSF Panels} You should serve on NSF panels.  It is a great way to learn about the grant process: how to distinguish between good and bad proposals and how people review propsoals.
+It is also a good way learn what other people are doing, what is hot, and what is not. In short, it is a good way to learn how to write successful proposals.
+
+Just like serving on a PC, you will review proposals and you will have to argue for why a proposal should be funded or not.
+Unlike PC where discussions can spread over many days (and there are also rebuttals), NSF panels are usually a two-day event.
+You do get paid a bit for serving on NSF panels and you get to travel to NSF headquarters in Virginia (except during the pandemic where everything is remote).
+
+In general, I enjoy serving on NSF panels much more than serving on PC.
+While the work is more intense and takes more time, I found it to be more rewarding and learned a lot from it.
+Since my first NSF panel in 2019, I have served 7 panels (at least once every year since '19) and I look forward to serving more.
+
+\textbf{Advice}: Similarly to serving in PC, you can cold email progarm managers in your area and ask to be on a panel.
+Once you have served, you will likly be invited to serve again.
+Unfortunately, I do not have much experience with non-NSF panels or proposals.
+
 \section{Service to the Department}
 \section{Service to the University}
-\section{Serving on NSF Panels}
+\section{Other Types of Services}
+\paragraph{Writing Recommendation Letters} You will be asked to write a lot of recommendation letters, for both students and colleagues.
 
-You should serve on NSF panels.  It is a great way to learn about the grant process: how to distinguish between good and bad proposals and how people review propsoals.
-It is also a good way learn what other people are doing, what is hot, and what is not. In short, it is a good way to learn how to write successful proposals.
-
-\paragraph{How to get on an NSF panel} You can cold email progarm managers in your area and ask to be on a panel.
-You can also ask your colleagues to make recommend for you.
-Same way you do for getting invited to be on a PC.
-
-\paragraph{How much work is it?} It is a lot of work.
-You will be asked to review a dozen proposals in a month or so.
-It will be similar to a program committee meeting, where all the proposals will be discussed.
-You will be expected to share your view of the paper, and argue for why it should be funded or not.
+\textbf{Advice}: you should have a template for recommendation letters.
+You can find many templates online.
+You should also ask your colleagues to share their templates with you.
+You should also ask your colleagues to share their recommendation letters with you.
+You can learn a lot from reading other people's recommendation letters.
 

### 2024-03-22 demystify.tex
updates

@@
 \chapter*{Preface}
-Having been involved in PhD admissions for many years, I've
-realized that many \textbf{international students} (and also domestic ones), especially those in  smaller countries or less well-known universities, lack a clear understanding of
+Having been involved in PhD admissions for many years, I've realized that many \textbf{international} (and domestic) students, especially those in  smaller countries or less well-known universities, lack a clear understanding of
 the Computer Science PhD admission process at US universities. This confusion not only
@@
   If your grading system is not US standard or you are from a good school unknown outside of your country, you can ask your reference writers to explain about that in their letters.
-For example, "Bach Khoa" are the top universities in Vietnam for STEM studies but few people outside Vietnam know about them.
+For example, ``Bach Khoa'' are the top universities in Vietnam for STEM studies but few people outside Vietnam know about them.
 So if you are from there, you should ask your reference writers to mention that.
@@
 \paragraph{School} Graduating from top universities \emph{that we recognize} helps.
-For example, if you are an international student and your school is well-known, then it is \emph{``top foreign''}, which is definitely a plus.
+For example, if you are an international student and your school is well-known, then it is \emph{``top foreign''}, which is a plus.
 However, if we do not know much about schools in your country, then we are uncertain about the quality of your school and  likely treat your school as \emph{``unknown foreign''}, which can be a minus point.
@@
 Typically, an interview takes about 15--30 minutes, and one important aspect of evaluation is your ability to effectively communicate, including speaking and understanding English. 
-
-\textbf{Advice}: In general, just treat the interview as an informal chat, but be prepared to talk about your research experiences and interests (so do have \emph{elevator pitch} about your experiences and research).
-A professor might want to know more about your research background or even ask you to read a paper (so be sure to do it!).
-Some prof.
-can ask you to solve a problem, e.g., a coding problem, so you might want to ask if you need to prepare for that.
-During the interview, you will have the opportunity to ask questions about the university. It's essential to prepare some thoughtful questions, similar to a job interview.
+You might be asked to talk about your research experiences and interests, and to read a paper and discuss it. In some rare cases you might also be asked to solve a problem, e.g., a coding problem. 
+The interview is also a good opportunity for you to ask questions about the program and the university.
+
+\textbf{Advice}: Treat the interview as an informal chat.
+Have elevator pitches about your research experiences and interests.
+You might also want have some 5-minute presentation about your research.
+If a professor asks you to read a paper, do it and be prepared to discuss it.
+You can also ask if you need to prepare for a coding problem.
+Finally, prepare some thoughtful questions to ask, e.g., about the program, the university, or the professor's research.
 
@@
 
-\epigraph{"Oh...
+\epigraph{``Oh...
 and how is education supposed to make me feel smarter?
 Besides, every time I learn something new, it pushes some old stuff out of my brain.
-Remember when I took that home wine-making course and I forgot how to drive?"}{\textsc{The Simpsons}}
+Remember when I took that home wine-making course and I forgot how to drive?''}{\textsc{The Simpsons}}
 
@@
 Fourth, like most source of income in the US, you will have to \emph{pay tax} on your stipend.
-Finally, private universities might pay more for stipend, but they could have extra "activity" fee or some other hidden ones (e.g., you may need to pay some fees for each credit hour).
+Finally, private universities might pay more for stipend, but they could have extra activity fee or some other hidden ones (e.g., you may need to pay some fees for each credit hour).
 
@@
 But sometimes what is important are things that they (current students) don’t tell you.
-Pay attention to their "level of excitement" being in the group.
+Pay attention to their ``level of excitement'' being in the group.
 \end{commentbox}
@@
 \begin{itemize}
-  \item \href{https://www.cs.columbia.edu/wp-content/uploads/2019/03/Get-Advisor.pdf}{The Definitive "what do I ask/look for" in a PhD Advisor Guide}
+  \item \href{https://www.cs.columbia.edu/wp-content/uploads/2019/03/Get-Advisor.pdf}{The Definitive ``what do I ask/look for'' in a PhD Advisor Guide}
 \end{itemize}
@@
 
-Faculty often receive "cold" e-mails from prospective students. Most of the time, we ignore these emails, but in some rare occasions we do answer them. So how to write an email that get our attention?
+Faculty often receive ``cold'' e-mails from prospective students.
+Most of the time, we ignore these emails, but in some rare occasions we do answer them.
+So how to write an email that get our attention?
 
@@
 \chapter{Miscs and FAQs}
-\epigraph{"I want to share something with you – the three little sentences that will get you through life; number 1: Cover for me, number 2: Oh, good idea, Boss, and number 3: It was like that when I got here."}{\textsc{The Simpsons}}
+\epigraph{``I want to share something with you – the three little sentences that will get you through life; number 1: Cover for me, number 2: Oh, good idea, Boss, and number 3: It was like that when I got here.''}{\textsc{The Simpsons}}
 
 \section{Why did I get rejected?}\label{sec:why-rejected}
-Many possible reasons!  Other than not having a strong profile (e.g., research potentials, GPA, LoRs, SOP, poor interviews), here are some that you might not expect:
+Many possible reasons!  Other than not having a strong profile (e.g., research potentials, GPA, LoRs, SOP, poor interviews), here are some that you might not expect.
 
@@
 
+\paragraph{You aim too high}  You might have applied to schools that are too high for your profile.
+While being ambitious is good, you should also be realistic and apply to a range of schools, including ``safety'' ones.
+I have seen many students who have excellent profiles but only applied to top 10 or even top 20 schools and got rejected from all of them.
+Before applying, you should talk to your professors and ask them to give you an honest opinion on where you should apply
+
 \paragraph{Overqualified or Lack Interest}  This might be paradoxical, however if the adcom believes that you are likely to get admitted and go to a better program (i.e., they are just your ``safety'' schools), they might not be so enthusiastic to admit you and want to save the spot for someone else.
@@
 
-\paragraph{Red flags in Your Background} Plagiarism, cheating, or other academic dishonesty are red flags.
+\paragraph{Red flags in Your Background} Plagiarism, cheating, or other academic dishonesty are \red{red} flags.
 Another one is if you have a history of jumping from one program to another (e.g., you have been in multiple different programs in short time).
@@
 Given the high number of quality applicants and limited number of spots, other than having a good application profile, you really want to show something that makes you \textbf{stand out}, e.g., do you have a degree or background in \emph{dance} or \emph{music} and want to integrate them with CS?
-are you a female or a minority in CS (research for "URM minority in CS in the US" on Google) (\autoref{sec:urm})?
+are you a female or a minority in CS (research for ``URM minority in CS in the US'' on Google) (\autoref{sec:urm})?
 Do you participate in outreach activities that help increase diversity and inclusion in CS?
@@
 However, upon learning that the applicant had led a 100km hike in the Himalayas, he decided to accept the applicant.
-This is a good example of being "stand out", and I would also advocate for that student as this shows they have the persistence and determination required for research.
+This is a good example of being \emph{stand out}, and I would also advocate for that student as this shows they have the persistence and determination required for research.
 
@@
 
-\section{Can I apply to CS PhD if my undergrad was not in CS or related areas?}\label{sec:non-stem}
-
-Yes, \emph{as long as} you can demonstrate you are ready for CS PhD research through research experiences, LoRs, statements, etc as mentioned.
+\section{My undergrad was not in CS or related areas}\label{sec:non-stem}
+
+You still can apply to PhD in CS \emph{as long as} you can demonstrate you are ready for it through your background, research experiences, LoRs, statements, etc.
 You might be even able to leverage this to make your profile stand out as mentioned in \autoref{sec:improve-your-chance}.
 
-One main concern for non-CS or non-STEM students is if you have the sufficient technical background obtained through core CS courses.
-You need to show that you have such knowledge through your coursework, projects, or research.
+A main question adcom has for a non-CS or non-STEM student is if you have the sufficient technical background obtained through core CS courses.
+So you need to show that you have such knowledge through your coursework, projects, or research.
 For example, if you have taken a course on Algorithms, even online ones like Coursera, you can talk about it in your SOP.
@@
 In short, you \emph{do not need} to formally taking CS courses, you just need to show that you have these essential knowledge, e.g., through the mentioned ways.
-Many universities are well-aware that incoming graduate students might not have all the technical background, so they often have a "bridge" courses to help students catch up.
+Many universities are well-aware that incoming graduate students might not have all the technical background, so they often have a \emph{``bridge''} courses to help students catch up.
 For example, GMU has four bridge courses (Data Structures and Algorithms, Computer Systems, Discrete Math, and Programming Foundation) that incoming students can take to catch up on their CS knowledge.
@@
 In fact, student can get MS degree ``along the way'' to PhD, e.g., after finishing course work in the first 2 years.
-However, MS can help admission if it gives research experience or is from a more well-known school than your undergrad institution.
+However, MS can help admission if it gives research experience or is from a more well-known school than your undergrad institution (\autoref{sec:your-school}).
 
@@
 In other countries, you (who already have an MS) start your research right away, e.g., you immediately work on the research project you proposed with the adviser you chose.
-Also, in the US you also have to pass a series of "exams", e.g., qualifying exam, comprehensive exam, thesis proposal defense\footnote{The word "ABD" (all but dissertation) is used in the US to refer to a PhD candidate who have finished all course work and exams and only need to write and defend their dissertation.}.
+Also, in the US you also have to pass a series of exams, e.g., qualifying exam, comprehensive exam, thesis proposal defense\footnote{The word ``ABD'' (all but dissertation) is used in the US to refer to a PhD candidate who have finished all course work and exams and only need to write and defend their dissertation.}.
 In other countries, you might not have these exams or only have to do a few of them.
@@
 In the US, PhD graduates often apply directly for faculty position (postdoc for US graduates is no longer a popular option as it was before).
-%This is also why US PhD programs are longer (remember, 5--7 years).
 
@@
 You can also include their institution if it makes it more precise.
-For example, I can say:  \emph{"I did my postdoc with Jeff Foster at Univ.
-of Maryland"}.
+For example, I can say:  \emph{``I did my postdoc with Jeff Foster at Univ.
+of Maryland''}.
 

### 2024-03-22 demystify.tex
updates

@@
 
-You should also talk about things that adcom members might not know about and can help make you different than the other 1000 applicants, e.g., you are member of a minority or LBGTQ+ group (\autoref{sec:urm}), your personal Github project with 1K+ stars or your regular contributions to well-known open-source projects (see \autoref{sec:improve-your-chance} for increasing your admission chance).
+You should also talk about things that adcom members might not know about and can help make you \textbf{stand out} in the application pool of  thousands of applicants, e.g., you are member of a minority or LGBTQ+ group (\autoref{sec:urm}), your personal Github project with 1K+ stars or your regular contributions to well-known open-source projects (see \autoref{sec:improve-your-chance} for increasing your admission chance).
 
@@
 Having published papers, especially at top venues, is a sign that the applicant has successfully involved in research.
-Students  accepted to top schools often have multiple first-authored papers at top venues.
+Applicants  admitted to top schools often have multiple first-authored papers at top places.
 
@@
   Sometime PhD adcom in the US will share a document such as \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Foreign-Top-Schools}{this one}, which lists the top schools in several countries.
-In some cases we ask other faculty and students if we think they know about the place.
+We also ask other faculty and students if we think they know about the place.
 For example, when I was a postdoc at UMD, members in their CS PhD adcom ask me to evaluate applicants from Vietnam.
@@
 
-Given the high number of quality applicants and limited number of spots, other than having a good application profile, you really want to show something that makes you \textbf{stand out}, e.g., do you have a degree or background in \emph{dance} or \emph{music} and want to integrate them with CS?
-are you a female or a minority in CS (research for ``URM minority in CS in the US'' on Google) (\autoref{sec:urm})?
-Do you participate in outreach activities that help increase diversity and inclusion in CS?
-All of these are unique and would get noticed from reviewers.
-
-Even if you do not have research experience, you can talk about your personal projects, as long as they can help show you can do research.
+Given the high number of quality applicants and limited number of spots, in addition to having a good application profile, you really want to show something that makes you \textbf{stand out}.
+For example, even if you do not have research experience, you can talk about your personal projects, as long as they can help show you can do research.
 For example, if you have an open-source project on Github that is used by many people, has lots of stars in Github, do talk about it.
@@
 
+There are other things you might not think are important but can make you stand out.
+For example, if you have a strong background in a field that is not CS but can be integrated with CS, e.g., you have a degree or background in \emph{dance} or \emph{music} and want to integrate them with CS, do talk about it.
+Are you a female or a minority in CS (\autoref{sec:urm})?
+Do you participate in outreach activities that help increase diversity and inclusion in CS?
+Diversity is highly valued in CS programs in the US, and experiences in these areas can make you stand out and get noticed from reviewers.
+
+\begin{commentbox}[Vu:]
 In his \href{https://matt.might.net/articles/how-to-apply-and-get-in-to-graduate-school-in-science-mathematics-engineering-or-computer-science/}{post}, Matt Might was initially unsure about an application.
@@
 This is a good example of being \emph{stand out}, and I would also advocate for that student as this shows they have the persistence and determination required for research.
+\end{commentbox}
 
@@
 
-\section{Member of an underrepresented or LBGTQ+ groups}\label{sec:urm}
-A common question is whether you should mention that you are a member of an underrepresented and minority (URM) group\footnote{URM in CS in the US include African Americans, Hispanic Americans, American Indians/Native Alaskans, Native Hawaiians/Pacific Islanders (excluding Asian Americans), and multiracial students identifying with at least one of the previously listed URM categories.} or LBGTQ+ group (, e.g., in your SOP).
+\section{Member of an underrepresented or LGBTQ+ groups}\label{sec:urm}
+A common question is whether you should mention that you are a member of an underrepresented and minority (URM) group\footnote{URM in CS in the US include African Americans, Hispanic Americans, American Indians/Native Alaskans, Native Hawaiians/Pacific Islanders (excluding Asian Americans), and multiracial students identifying with at least one of the previously listed URM categories.} or LGBTQ+ group (, e.g., in your SOP).
 This is a strong concern for international (and even domestic) students, who may have encountered discrimination in their countries or personal experiences.
@@
 
-In my personal opinion, highlighting URM or LBGTQ+ identity can \emph{indeed be beneficial}, especially if you can articulate how your diverse experiences contribute to diversity and inclusion in academic.
+In my personal opinion, highlighting URM or LGBTQ+ identity can \emph{indeed be beneficial}, especially if you can articulate how your diverse experiences contribute to diversity and inclusion in academic.
 Many universities have a strong commitment to diversity and inclusion and actively
 \emph{recruit students from underepresentative backgrounds} (e.g., some \href{https://cse-climate.engin.umich.edu/reports/climate-dei-reports/cse-climate-and-dei-report-2022-2023/#grad-ethnicity}{stats from UMich CSE} and \href{https://www.gmu.edu/news/2022-09/mason-now-top-10-public-university-diversity-innovation-and-cybersecurity-education-us}{GMU} often promotes itself as one of the most diverse universities in the US).
-Moreover, many scholarships and fellowships are created specifically for URM and LBGTQ+ students, which you should explore if qualify.
+Moreover, many scholarships and fellowships are created specifically for URM and LGBTQ+ students, which you should explore if qualify.
 
@@
 
-\paragraph{Research Experiences \autoref{sec:undergrad-research}} You might have many opportunities to do research as an undergraduates, e.g., through REUs and internships at your undergrad university.
+\paragraph{Research Experiences \autoref{sec:research-opportunities}} You might have many opportunities to do research as an undergraduates, e.g., through REUs and internships at your undergrad university.
 Highlight such experiences in your application.
@@
 
-\chapter{Research Opportunities for Undergraduates}\label{sec:undergrad-research}
-
-As discussed in \autoref{sec:undergrad-research}, research experience is a crucial component of a strong PhD application.
-It gives you opportunities to try out research, determine what research area you're interested in, publish papers (\autoref{sec:research-experiences}), and get strong LoRs (\autoref{sec:lor}).
-
- This section provides some guidance on how to gain research experience as an undergraduate student.
-It also gives advice to international students or domestic ones from small colleges where research opportunities are limited.
+\chapter{Research Opportunities}\label{sec:research-opportunities}
+
+As discussed in \autoref{sec:research-opportunities}, research experience is a crucial component of a strong PhD application.
+It gives you opportunities to try out research, determine what research area you're interested in, publish papers (\autoref{sec:research-experiences}), connect with researchers and get strong LoRs (\autoref{sec:lor}).
+
+ This section provides some guidance on how to gain research experience, e.g., as an undergraduate student.
+It also gives advice to students from small colleges where research opportunities are limited.
 
@@
 By contributing code, fixing bugs, implementing new features, or providing documentation, students can gain practical research experience and interact with experienced developers and researchers in the field.
-Not only you gain research experience, you might be able to get a LoR from the project maintainer and definitely should write about this experience in your SOP (\autoref{sec:research-statement}.)
-
-\paragraph{Virtual Research Opportunities}: This is less common but several places offer virtual internships and research programs aimed at providing hands-on research experience.
+Not only you gain research experience, you might be able to get a LoR from the project maintainer, and you should write about this experience in your SOP (\autoref{sec:research-statement}.)
+
+\paragraph{Virtual Research Opportunities} This is less common but several places offer virtual internships and research programs aimed at providing hands-on research experience.
 These programs often involve working remotely under the guidance of experienced mentors and collaborating with a team of fellow researchers.

### 2024-03-22 faculty.tex
updates

@@
-\documentclass[oneside,11pt]{memoir}
-\chapterstyle{pedersen}
+\documentclass[oneside,11pt,dvipsnames]{book}
 
@@
 \usepackage[most]{tcolorbox}
-\setsecnumdepth{subsection}
-\setcounter{tocdepth}{3}
+% \setsecnumdepth{subsection}
+% \setcounter{tocdepth}{3}
 \usepackage{enumitem}
@@
 \usepackage{booktabs}
+
+\usepackage{tikz}
+\usetikzlibrary{calc}
 \usepackage{xcolor}
+
+\usepackage{anyfontsize}
+\usepackage{sectsty}
+
 \usepackage[makeroom]{cancel}
@@
 
-\title{A Handbook for Tenured Track Faculty in Computer Science in the US}
+\begin{document}
 
-\author{\href{https://nguyenthanhvuh.github.io}{ThanhVu (Vu) Nguyen}\\{\small George Mason University, Dept. of Computer Science}}
+\pagestyle{empty}
+\begin{tikzpicture}[overlay,remember picture]
 
-\makeatletter
-\def\maketitle{%
-  \null
-  \thispagestyle{empty}%
-  \vfill
-  \begin{center}\leavevmode
-    \normalfont
-    {\LARGE\raggedright \textbf{\@title}\par}%
-    \vfill%
-    % \hrulefill\par
-    {\Large \@author\par}%
-    \vfill%
-    {\large\raggedleft \@date\par}%
-  \end{center}%
-  \vfill
-  \null
-  \cleardoublepage
+    % Background color
+    \fill[
+    black!2]
+    (current page.south west) rectangle (current page.north east);
+
+    % Rectangles
+    \shade[
+    left color=Dandelion, 
+    right color=Dandelion!40,
+    transform canvas ={rotate around ={45:($(current page.north west)+(0,-6)$)}}] 
+    ($(current page.north west)+(0,-6)$) rectangle ++(9,1.5);
+
+    \shade[
+    left color=lightgray,
+    right color=lightgray!50,
+    rounded corners=0.75cm,
+    transform canvas ={rotate around ={45:($(current page.north west)+(.5,-10)$)}}]
+    ($(current page.north west)+(0.5,-10)$) rectangle ++(15,1.5);
+
+    \shade[
+    left color=lightgray,
+    rounded corners=0.3cm,
+    transform canvas ={rotate around ={45:($(current page.north west)+(.5,-10)$)}}] ($(current page.north west)+(1.5,-9.55)$) rectangle ++(7,.6);
+
+    \shade[
+    left color=orange!80,
+    right color=orange!60,
+    rounded corners=0.4cm,
+    transform canvas ={rotate around ={45:($(current page.north)+(-1.5,-3)$)}}]
+    ($(current page.north)+(-1.5,-3)$) rectangle ++(9,0.8);
+
+    \shade[
+    left color=red!80,
+    right color=red!80,
+    rounded corners=0.9cm,
+    transform canvas ={rotate around ={45:($(current page.north)+(-3,-8)$)}}] ($(current page.north)+(-3,-8)$) rectangle ++(15,1.8);
+
+    \shade[
+    left color=orange,
+    right color=Dandelion,
+    rounded corners=0.9cm,
+    transform canvas ={rotate around ={45:($(current page.north west)+(4,-15.5)$)}}]
+    ($(current page.north west)+(4,-15.5)$) rectangle ++(30,1.8);
+
+    \shade[
+    left color=RoyalBlue,
+    right color=Emerald,
+    rounded corners=0.75cm,
+    transform canvas ={rotate around ={45:($(current page.north west)+(13,-10)$)}}]
+    ($(current page.north west)+(13,-10)$) rectangle ++(15,1.5);
+
+    \shade[
+    left color=ForestGreen,
+    rounded corners=0.3cm,
+    transform canvas ={rotate around ={45:($(current page.north west)+(18,-8)$)}}]
+    ($(current page.north west)+(18,-8)$) rectangle ++(15,0.6);
+
+    \shade[
+    left color=ForestGreen,
+    rounded corners=0.4cm,
+    transform canvas ={rotate around ={45:($(current page.north west)+(19,-5.65)$)}}]
+    ($(current page.north west)+(19,-5.65)$) rectangle ++(15,0.8);
+
+    \shade[
+    left color=OrangeRed,
+    right color=red!80,
+    rounded corners=0.6cm,
+    transform canvas ={rotate around ={45:($(current page.north west)+(20,-9)$)}}] 
+    ($(current page.north west)+(20,-9)$) rectangle ++(14,1.2);
+
+    % Year
+    \draw[ultra thick,gray]
+    ($(current page.center)+(5,2)$) -- ++(0,-3cm) 
+    node[
+    midway,
+    left=0.25cm,
+    text width=5cm,
+    align=right,
+    black!75
+    ]
+    {
+    {\fontsize{25}{30} \selectfont \bf Computer\\[8pt] Science}
     } 
-\makeatother
+    node[
+    midway,
+    right=0.25cm,
+    text width=6cm,
+    align=left,
+    ForestGreen]
+    {
+    {\fontsize{48}{50} \selectfont PhD}
+    };
 
-\begin{document}
-\maketitle
-\frontmatter
+    % Title
+    \node[align=center] at ($(current page.center)+(0,-5)$) 
+    {
+    {\fontsize{24}{24} \selectfont {{Demystifying the Computer Science}}} \\[0.15in]
+    {\fontsize{24}{24} \selectfont {{PhD Admission in the US}}} \\[0.2in]    
+    {\fontsize{18}{18} \selectfont {{A Handbook for International and Domestic Students}}} \\[0.5in]
+
+    {\fontsize{16}{19.2} \selectfont \textcolor{ForestGreen}{ \bf ThanhVu (Vu) Nguyen}}\\[0.1in]
+    George Mason University, Dept. of Computer Science\\[0.1in]
+    \today{} (latest version available on  \href{https://github.com/nguyenthanhvuh/phd-cs-us}{Github})
+    };
+    \end{tikzpicture}
 
@@
 
-Chapter 3: Funding
-Let’s talk about what, for many professors, is the least favorite part of the job: making sure you
-have enough money to keep the lights on in the lab. Essentially, a professor applies for grants
-(money from the government, industry, or other charitable foundations) by writing a proposal as
-the Principal Investigator (PI). The funding agency reviews the proposal. If they decide to fund it,
-they send the money to your university. The university then takes its cut of the funding
-(essentially to pay for common services like buildings, admins, etc) and gives you the rest. You
-then spend it on your summer salary, the salaries of your students, equipment, travel,
-registration costs and other expenses associated with your research.
-There are some nuances about the different types of funding, and how you can spend different
-kinds of funding. So let us get into the details. Unfortunately, these details tend to vary a lot by
-country, so everything I’m talking about in this chapter applies only to the US.
-3.1 An example budget
-Let me walk you through an example budget. Let us say you are writing a proposal for \$100,000
-USD (100K USD). The budget will talk about how you plan to spend this money. The following
-table (Table 1) shows the example budget.
-Budget
-Graduate Research Assistant - 1
-Total Salary 40,000
-Fringe Benefits 12,000
-Total Salary and Fringe Benefits 52,000
-Total Materials \& Supplies 0
-International travel to a conference 2500
-Domestic travel to a conference 1481
-Total Travel Costs 3,981
-Tuition for Fall or Spring 4,746
-Tuition for Summer 1,778
-GRA Tuition for 1 year 11,270
-Total Direct Costs 67,251
-Modified Total Direct Cost (MTDC) 55,981
-Indirect Costs 58.5% of MTDC 32,749
-Total Cost (Direct + indirect costs) 100,000
-37
-I will explain the various parts of this example budget. I will be talking about terms such as
-direct costs and indirect costs as defined by the University of Texas – I believe it is similar at
-most other US universities, but there might be differences. Please check with your university to
-make sure.
-Fringe. Let us say you want to support a student on this grant. In addition to paying for the
-student’s salary, you also have to pay for non-wage expenses: things like health insurance, social
-security, and medicare. This is usually expressed as a percentage of the student’s salary. At the
-University of Texas at Austin, this was 30% in 2022 (For a comparison point, it was 32% at
-Princeton). The stipend of a graduate research assistant (GRA) at UT is about \$30K USD for the
-fall and spring semesters ($40K USD including summer). So salary and fringe for a GRA for a
-year is \$52K.
-You also have to pay the tuition of the GRA, and at UT, this is about \$11.2K USD for a year. To
-keep things simple, I am using a single number for the tuition, but this depends on factors such
-as whether you are in-state or out-of-state, your progress in the PhD program (if you are
-all-but-dissertation, your tuition is lower), and whether you are a US citizen or not. It also
-depends on the fraction of time spent as GRA/TA: if this drops below 20 hours a week
-(considered full-time for students), tuition might have to be charged at the out-of-state rate. If
-the student is only doing 5 hours a week as GRA, it might also impact how much of their tuition
-can be covered by the grant; at 5 hours a week, the student might not be eligible for insurance. I
-highly recommend talking to your admin if you are accounting for less than 20 hours a week for
-your student on the project. Some schools also calculate tuition as a % of the stipend.
-So the total cost for supporting a GRA for one year is \$63.2K USD. While you are paying \$63.2K
-for each student for a year, the student's take-home pay is only the stipend part (minus taxes
-that are withheld): \$35.2K assuming a 12% tax rate.
-Note that fringe also applies to you (the PI) when you pay yourself out of your grants for your
-summer months.
-Direct costs. The direct cost part of your budget includes salaries, fringe benefits, travel, and
-equipment and supplies that contribute directly towards performing the proposed research.
-Direct costs are exclusive to each proposal: a piece of equipment can only be a direct cost on
-one proposal, and a GRA can be assigned only to one proposal at a given time. In our example,
-let us assume the only direct costs are supporting the student and paying for the student’s
-travel (including registration, transportation, lodging, and food) for one international and one
-domestic conference.
-Modified Total Direct Costs (MTDC). MTDC is a portion of the total direct costs used for
-calculating the indirect costs. It is the direct costs minus tuition, fellowships, equipment, and
-capital expenditures. Travel also counts as MTDC. In this budget, we are not spending anything
-38
-on materials and supplies in this budget, but if we did, it would be part of MTDC. Check with
-your university to determine exactly what is counted as MTDC – this can be surprisingly tricky.
-For example, you might expect computers to not count towards MTDC as it is “equipment”, but it
-may be counted as a “supply” if the university is unable to amortize its cost. In the example
-above, total direct costs is \$67.2K USD (63.2K for one GRA plus 4K for travel), but MTDC is only
-$56K (leaving out the \$11.2K for tuition).
-Indirect costs. The university takes a portion of all the money you receive, to pay for centralized
-functions such as building costs, utilities, admin, and libraries. This is termed as the “indirect”
-cost, usually expressed as a percentage of the MTDC. At most universities in the US, indirect
-costs are more than 50% of the MTDC. At UT, it is 58.5% (Mike Freedman reports it is 63% at
-Princeton). Thus, in our budget, we would have .585 * 56K = \$32.7K added as indirect costs. As
-a result, it would cost 67.2 + 32.7 = \$100K USD for supporting a single GRA and some travel on
-this grant. Note that even though the 58.5% number is scary, you are only paying 33% of your
-budget for indirect costs; in other words, you don’t lose 58.5% of the total budget to indirect
-costs.
-As you can see from the preceding table, the \$100K USD budget only covers one student for one
-year. At some universities, a GRA costs more, requiring more funding.
-This should give you a picture of how expensive it is to maintain large groups! If you had ten
-students, you would need to pull in a million dollars each year to fund your students.
-It is useful to think of your budget in terms of student-years. \$400K is simply two students for
-two years. A dissertation, roughly five student years, is \$500K. If you get the NSF CAREER grant,
-you would typically be able to support one dissertation. This helps determine the scope of your
-proposal: if you need five students to work on it for two years, a grant that provides \$250K will
-not be a good match.
-Summer months in your budget. PIs at most R1 universities in the US are expected to fund
-themselves via grants in the summer. When you are drawing up your budget, you are expected
-to estimate how much of your summer months you will dedicate to this project. This is taken as
-an indication of PI commitment at the funding agency – you don’t want to write a budget where
-your estimated time involvement is zero. So if you have one NSF grant where you estimate your
-time commitment to be one month, you can only put in one more month for another NSF grant
-proposal (since NSF limits total PI compensation to be two months per year across all NSF
-grants).
-3.2 Types of grants
-Broadly speaking, there are four different kinds of grants or funding, each with different
-restrictions on how you can use the money. The differences come down to what you can spend
-the money on, how long you can keep the money, and what the overhead is.
-39
-We can term the four kinds of funding as grants, contracts, university funding, and gift funding.
-Let me go over each type of funding.
-Grants. The National Science Foundation (NSF), the Defense Advanced Research Projects
-Agency (DARPA), the Army, the Navy, and the Air Force all fund research related to computer
-science. These grants are competitive – NSF has a 25% acceptance rate (though this varies by
-the directorate), and I imagine the other programs are similar.
-This money can typically only be used for funding students, yourself, travel, and equipment
-purchases that you had already put in the budget. For example, if you suddenly want to
-purchase equipment not in the original budget, it needs to be approved by the Office of
-Sponsored Research at your university (which makes sure you are spending the money in the
-right way). You cannot use government funding to pay for things like a meal with your students
-during the semester (meals during travel is allowed).
-Government funding has strict timelines by which the money has to be spent. Typically, the
-money is released on a yearly basis. You have to show that you have used the previous
-installment to get the next round of funding. This sometimes results in professors making last
-minute purchases to exhaust their grant!
-In some cases, you can ask your funding program manager for a no-cost extension to your
-grant. This typically means that you don’t need extra money, but you want extra time to use the
-money. This is usually considered on a case-by-case basis, and I would not recommend
-counting on getting a no-cost extension.
-Indirect costs are levied on government funding (more generally, on all funding that goes
-through the Office of Sponsored Research). The example budget I showed in the previous
-section assumes funding is via sponsored research.
-Grant proposals typically have reporting requirements (for example, you have to submit a report
-once a year for an NSF grant). However, grants typically do not expect any concrete deliverables
-to be provided at the end of the funding period. Essentially, the grant allows you to carry out the
-proposed research: the act of publishing itself is the desired outcome. As a result, grants have a
-strong alignment between the PIs and the funding agency: they both just want to get research
-done.
-Contracts. The second kind of funding is via contracts. Contract funding comes with an explicit
-contract you must fulfill to keep and obtain more funding. For example, there may be quarterly
-or monthly reports, or on-site visits by the sponsor. Contract funding typically also has
-deliverables that are expected at regular intervals during the funding period. One example of
-contract funding is DARPA funding, which comes with a number of deliverables and reporting
-requirements. Contract funding is less aligned with the PIs than grant funding: apart from doing
-the core research, the PI’s students will have to spend time on creating the deliverables, and the
-PI herself has to spend a significant amount of time on reports and meetings.
-40
-NSF vs DARPA. NSF is a good example of grant funding, while DARPA is a good example of
-contract funding. Both these agencies fund a lot of computer science research. DARPA requires
-deliverables and reports to be produced regularly during the grant duration. NSF only requires a
-report to be submitted for each year of the grant duration. You can apply for NSF funds as
-someone on a visa, while DARPA and other similar government funding programs might require
-you to be a US citizen or a permanent resident.
-University funding. Your university, or your college, or your department may choose to give you
-funding. Typically when you start, you get a “startup” package that allows you to kickstart your
-research group. University funding is “internal” funding in a way, and as such has fewer
-restrictions than government funding. But this really depends on the specific university. I know
-universities where startup funding cannot be used for covering summer salary of PIs. Apart
-from startup funding, your university might have internal competitions to award “seed” funding
-(typically under 50K) that you can then use to get preliminary results and later apply for
-traditional grants.
-University funding is also associated with timelines. For example, startup funding usually has to
-be spent before you get tenure. However, university funding deadlines might be a little more
-flexible than grant or contract funding deadlines.
-University funding typically does not have associated indirect costs (as it is already university
-money). You still have to pay fringe support for your students and your own summer salary.
-An interesting sub-category of university funding are Research Centers. A research center is
-funded by sponsors to carry out a certain mission over a span of around 5 years. The research
-center has a program director who is responsible for allocating the funding. PIs at the university
-apply for funding from the research center (a sort of internal grant), and are allocated funding.
-Obtaining funding in this manner is easier than competing for an external grant. Particularly for
-assistant professors who are just starting, research centers can provide seed funding before the
-first grants are obtained. Check out the ADA center at Michigan for a good example.
-Gift funding or unrestricted research funding. Companies or charitable foundations may
-choose to fund your research in the form of “gifts” or unrestricted research grants. This is the
-best kind of funding, for three reasons: indirect costs are typically not levied on these gifts, there
-is no associated timeline by when you should spend them, and you can use them for a broader
-range of things than university and grant/contract funding.
-Indirect costs are not levied because the gift-giver can specify that it cannot be used for indirect
-costs (the university has to honor this). Since it is an unrestricted gift, there are no limits on
-what you can spend them on, or by when you should spend them. This doesn’t mean you can
-spend the money on personal expenses. All expenses still have to be approved by the university,
-and should be used for research-related expenses. But things like taking your research group
-out for a meal can be expensed during gift funding (but not government funding).
-41
-Indirect costs not being levied can make a huge difference. For example, in our budget from
-earlier, if you were using gift funding to fund the student, you don’t have to pay indirect costs,
-and thus get over \$30K back. Considering a student costs \$63K, you get to sponsor the student
-for 1.5 years rather than 1 year!
-How should you use your funding? Let us say that you are lucky enough to have all four kinds of
-funding. And you have an expense that could be assigned to any of these funding (note that this
-is often not the case: you can’t use a grant on topic X to support a student working on Y if X and
-Y are not closely related). What should you use?
-The typical rule is “Use the most restricted funding that will expire the earliest”. Thus, you should
-always use the appropriate grant/contract funding first, then your university funding, and finally
-gift funding. Gift funding is the most flexible, so use it only when you have no other option.
-Always spend your gift funding on expenses that will have overhead (if funded from grants),
-such as students. Do not spend it on things like equipment (be careful to ask what counts as
-equipment at your university) which do not have overhead even when expensed to grants.
-Writing a grant with other PIs. While you could write smaller grants as the sole PI, typically for
-larger grants, you need to collaborate with other PIs and submit as a team. For example, you
-cannot submit NSF Large grants as a solo PI. One thing to remember though: all the funding is
-shared between the PIs. For example, imagine a grant for one million USD. It sounds amazing!
-However, if the grant was awarded to a team of two PIs, and the grant duration is five years,
-each PI can only support one student on it for the grant duration (100K * 2 * 5). Similarly, if a
-grant of 10 million USD was won by a team of five PIs, each PI can only support one student for
-four years (100K * 5 * 4). So in the end, the share of each PI is similar or lower than that of an
-NSF CAREER.
-This is not to say you should not collaborate with other PIs to write grants. Collaboration often
-brings new ideas and research directions that are unlikely or impossible when working alone.
-And bigger grants often provide prestige: for example, landing an NSF Large grant is significant
-and puts you on the map. But it is useful to calculate how much you expect to receive if a grant
-proposal is funded.
-3.3 The grant application process
-So how does one go about submitting a grant? There are several steps in the process:
-1. The funding agency first puts out a call for proposals. For example, check out the NSF
-call for CRII grants, a startup grant meant for new PIs who have not yet received a grant.
-Note that PIs at many R1 universities are not eligible for CRII. This was posted on May 18,
-2022. Each call has a solicitation page with details about the program and how the
-proposal should be structured – you should read this carefully.
-42
-2. In some cases, there are webinars where the program managers explain the grant
-opportunity. CRII had such a webinar on Jul 13 2022. I highly recommend attending
-these webinars if it is your first time submitting for a particular grant opportunity.
-3. In some cases, the funding agency might want you to submit a preliminary proposal.
-They will review the preliminary proposals and inform you if they would like you to go
-ahead with the full proposal. This is not the case for CRII.
-4. There will be a due date by which you need to submit the proposal. For the CRII, this is
-Sep 19 2022. The proposal will have a number of parts, and the solicitation will clearly lay
-out what the proposal should look like.
-5. You write the proposal, and sometime before you submit, you send it to the Office of
-Sponsored Research (OSP or OSR) at your university. They have to approve the proposal
-before you can submit, and they may request to see your proposal a week or 10 days
-before you submit. Please make sure you get a draft to the OSP on time!
-6. Finally, you submit the proposal, and you wait. To be exact, your admin/OSP submits for
-you, so you want to let them know when you are done. You and your admin/OSP will be
-submitting a lot of proposals together; make sure you get to know the folks doing this. I
-only have experience with NSF proposals, but these usually take 6-8 months to be
-reviewed. In some cases, it might take even longer.
-7. If you are lucky enough to get your proposal funded, you might get an email or a call from
-the NSF program manager. They might ask you to slightly modify your budget. For
-example, I was asked to reduce my NSF CAREER budget (and appropriately reduce the
-scope of my tasks as well). You will be asked to provide an abstract that will be displayed
-on the NSF website. For example, check out the abstract for my CAREER award.
-8. Finally, you will get an official email that your grant has been awarded, it will become
-public on the NSF website, and you will get the first installment delivered to your
-university. You can then begin spending it in accordance with what you proposed.
-9. In the unlucky case that your proposal is not funded, you will get back three or more
-reviews from NSF. These reviews tend to be short, and nowhere as detailed as paper
-reviews. As such, it can be hard to figure out what to do after getting a rejection. My
-advice is to go talk to senior faculty in your department if this happens, and they should
-be able to guide you in the right direction.
-3.3 Writing your first grant proposal
-Writing your first grant can be a stressful experience, since it is quite different from writing a
-research paper. There are a number of differences.
-First, the audience for a grant proposal is broader than that of a research paper; it is guaranteed
-that non-experts in your area will be reading and evaluating your proposal (for example, if I write
-43
-a grant proposal about storage research, I can expect it to be read by someone working in
-systems broadly, but not necessarily someone doing storage research). As a result, it has to be
-much broader and more accessible than a research paper. Proposal reviewers might also have a
-smaller amount of time to spend on each proposal – some funding agencies ask reviewers to
-read a dozen proposals in around a month when you serve on their panel. You can imagine how
-much time each proposal would get. Thus, your proposal must be easy to skim and read, must
-get to the major points quickly, and must be accessible to a broad audience.
-Second, in a research paper, you are reporting on something that is already done. The focus is
-on the technical details: how you were able to achieve the result, what are the drawbacks, etc.
-However, in a grant proposal, you are talking about things that you will do in the future. In a grant
-proposal, the most important thing is to be able to inspire the reviewer, and convince them that
-this is a thing worth doing (and funding). In this respect, it is much more similar to a founder
-pitching before an investor, than a professor talking to her peers about her research. While the
-details do matter to some extent, what is much more important is that you paint a big picture of
-how your proposed research will make the world a better place. This is quite different from
-writing a research paper, and takes some practice.
-For writing your first grant proposal, I’d recommend two things. First, talk to folks who have
-gotten the grant you are submitting for, and ask if you could see their grant proposal. Many folks
-are happy to share privately. Some folks, like Jeff Bigham and Gillian Hayes, are publicly sharing
-some of their proposals. Second, write up your draft proposal, and show it to senior folks in your
-department. They would be able to catch errors that could cause rejection.
-When submitting grant proposals, do not be disheartened if it gets rejected. As far as I can tell,
-most proposals, even from experienced faculty, get rejected a few times before getting
-accepted. Grant rejections are a bit more painful because the waiting time is high, you might
-only get to submit to a grant opportunity once per year. But with time and repeated submissions,
-many grant proposals do get funded.
+% Chapter 3: Funding
+% Let’s talk about what, for many professors, is the least favorite part of the job: making sure you
+% have enough money to keep the lights on in the lab. Essentially, a professor applies for grants
+% (money from the government, industry, or other charitable foundations) by writing a proposal as
+% the Principal Investigator (PI). The funding agency reviews the proposal. If they decide to fund it,
+% they send the money to your university. The university then takes its cut of the funding
+% (essentially to pay for common services like buildings, admins, etc) and gives you the rest. You
+% then spend it on your summer salary, the salaries of your students, equipment, travel,
+% registration costs and other expenses associated with your research.
+% There are some nuances about the different types of funding, and how you can spend different
+% kinds of funding. So let us get into the details. Unfortunately, these details tend to vary a lot by
+% country, so everything I’m talking about in this chapter applies only to the US.
+% 3.1 An example budget
+% Let me walk you through an example budget. Let us say you are writing a proposal for \$100,000
+% USD (100K USD). The budget will talk about how you plan to spend this money. The following
+% table (Table 1) shows the example budget.
+% Budget
+% Graduate Research Assistant - 1
+% Total Salary 40,000
+% Fringe Benefits 12,000
+% Total Salary and Fringe Benefits 52,000
+% Total Materials \& Supplies 0
+% International travel to a conference 2500
+% Domestic travel to a conference 1481
+% Total Travel Costs 3,981
+% Tuition for Fall or Spring 4,746
+% Tuition for Summer 1,778
+% GRA Tuition for 1 year 11,270
+% Total Direct Costs 67,251
+% Modified Total Direct Cost (MTDC) 55,981
+% Indirect Costs 58.5% of MTDC 32,749
+% Total Cost (Direct + indirect costs) 100,000
+% 37
+% I will explain the various parts of this example budget. I will be talking about terms such as
+% direct costs and indirect costs as defined by the University of Texas – I believe it is similar at
+% most other US universities, but there might be differences. Please check with your university to
+% make sure.
+% Fringe. Let us say you want to support a student on this grant. In addition to paying for the
+% student’s salary, you also have to pay for non-wage expenses: things like health insurance, social
+% security, and medicare. This is usually expressed as a percentage of the student’s salary. At the
+% University of Texas at Austin, this was 30% in 2022 (For a comparison point, it was 32% at
+% Princeton). The stipend of a graduate research assistant (GRA) at UT is about \$30K USD for the
+% fall and spring semesters ($40K USD including summer). So salary and fringe for a GRA for a
+% year is \$52K.
+% You also have to pay the tuition of the GRA, and at UT, this is about \$11.2K USD for a year. To
+% keep things simple, I am using a single number for the tuition, but this depends on factors such
+% as whether you are in-state or out-of-state, your progress in the PhD program (if you are
+% all-but-dissertation, your tuition is lower), and whether you are a US citizen or not. It also
+% depends on the fraction of time spent as GRA/TA: if this drops below 20 hours a week
+% (considered full-time for students), tuition might have to be charged at the out-of-state rate. If
+% the student is only doing 5 hours a week as GRA, it might also impact how much of their tuition
+% can be covered by the grant; at 5 hours a week, the student might not be eligible for insurance. I
+% highly recommend talking to your admin if you are accounting for less than 20 hours a week for
+% your student on the project. Some schools also calculate tuition as a % of the stipend.
+% So the total cost for supporting a GRA for one year is \$63.2K USD. While you are paying \$63.2K
+% for each student for a year, the student's take-home pay is only the stipend part (minus taxes
+% that are withheld): \$35.2K assuming a 12% tax rate.
+% Note that fringe also applies to you (the PI) when you pay yourself out of your grants for your
+% summer months.
+% Direct costs. The direct cost part of your budget includes salaries, fringe benefits, travel, and
+% equipment and supplies that contribute directly towards performing the proposed research.
+% Direct costs are exclusive to each proposal: a piece of equipment can only be a direct cost on
+% one proposal, and a GRA can be assigned only to one proposal at a given time. In our example,
+% let us assume the only direct costs are supporting the student and paying for the student’s
+% travel (including registration, transportation, lodging, and food) for one international and one
+% domestic conference.
+% Modified Total Direct Costs (MTDC). MTDC is a portion of the total direct costs used for
+% calculating the indirect costs. It is the direct costs minus tuition, fellowships, equipment, and
+% capital expenditures. Travel also counts as MTDC. In this budget, we are not spending anything
+% 38
+% on materials and supplies in this budget, but if we did, it would be part of MTDC. Check with
+% your university to determine exactly what is counted as MTDC – this can be surprisingly tricky.
+% For example, you might expect computers to not count towards MTDC as it is “equipment”, but it
+% may be counted as a “supply” if the university is unable to amortize its cost. In the example
+% above, total direct costs is \$67.2K USD (63.2K for one GRA plus 4K for travel), but MTDC is only
+% $56K (leaving out the \$11.2K for tuition).
+% Indirect costs. The university takes a portion of all the money you receive, to pay for centralized
+% functions such as building costs, utilities, admin, and libraries. This is termed as the “indirect”
+% cost, usually expressed as a percentage of the MTDC. At most universities in the US, indirect
+% costs are more than 50% of the MTDC. At UT, it is 58.5% (Mike Freedman reports it is 63% at
+% Princeton). Thus, in our budget, we would have .585 * 56K = \$32.7K added as indirect costs. As
+% a result, it would cost 67.2 + 32.7 = \$100K USD for supporting a single GRA and some travel on
+% this grant. Note that even though the 58.5% number is scary, you are only paying 33% of your
+% budget for indirect costs; in other words, you don’t lose 58.5% of the total budget to indirect
+% costs.
+% As you can see from the preceding table, the \$100K USD budget only covers one student for one
+% year. At some universities, a GRA costs more, requiring more funding.
+% This should give you a picture of how expensive it is to maintain large groups! If you had ten
+% students, you would need to pull in a million dollars each year to fund your students.
+% It is useful to think of your budget in terms of student-years. \$400K is simply two students for
+% two years. A dissertation, roughly five student years, is \$500K. If you get the NSF CAREER grant,
+% you would typically be able to support one dissertation. This helps determine the scope of your
+% proposal: if you need five students to work on it for two years, a grant that provides \$250K will
+% not be a good match.
+% Summer months in your budget. PIs at most R1 universities in the US are expected to fund
+% themselves via grants in the summer. When you are drawing up your budget, you are expected
+% to estimate how much of your summer months you will dedicate to this project. This is taken as
+% an indication of PI commitment at the funding agency – you don’t want to write a budget where
+% your estimated time involvement is zero. So if you have one NSF grant where you estimate your
+% time commitment to be one month, you can only put in one more month for another NSF grant
+% proposal (since NSF limits total PI compensation to be two months per year across all NSF
+% grants).
+% 3.2 Types of grants
+% Broadly speaking, there are four different kinds of grants or funding, each with different
+% restrictions on how you can use the money. The differences come down to what you can spend
+% the money on, how long you can keep the money, and what the overhead is.
+% 39
+% We can term the four kinds of funding as grants, contracts, university funding, and gift funding.
+% Let me go over each type of funding.
+% Grants. The National Science Foundation (NSF), the Defense Advanced Research Projects
+% Agency (DARPA), the Army, the Navy, and the Air Force all fund research related to computer
+% science. These grants are competitive – NSF has a 25% acceptance rate (though this varies by
+% the directorate), and I imagine the other programs are similar.
+% This money can typically only be used for funding students, yourself, travel, and equipment
+% purchases that you had already put in the budget. For example, if you suddenly want to
+% purchase equipment not in the original budget, it needs to be approved by the Office of
+% Sponsored Research at your university (which makes sure you are spending the money in the
+% right way). You cannot use government funding to pay for things like a meal with your students
+% during the semester (meals during travel is allowed).
+% Government funding has strict timelines by which the money has to be spent. Typically, the
+% money is released on a yearly basis. You have to show that you have used the previous
+% installment to get the next round of funding. This sometimes results in professors making last
+% minute purchases to exhaust their grant!
+% In some cases, you can ask your funding program manager for a no-cost extension to your
+% grant. This typically means that you don’t need extra money, but you want extra time to use the
+% money. This is usually considered on a case-by-case basis, and I would not recommend
+% counting on getting a no-cost extension.
+% Indirect costs are levied on government funding (more generally, on all funding that goes
+% through the Office of Sponsored Research). The example budget I showed in the previous
+% section assumes funding is via sponsored research.
+% Grant proposals typically have reporting requirements (for example, you have to submit a report
+% once a year for an NSF grant). However, grants typically do not expect any concrete deliverables
+% to be provided at the end of the funding period. Essentially, the grant allows you to carry out the
+% proposed research: the act of publishing itself is the desired outcome. As a result, grants have a
+% strong alignment between the PIs and the funding agency: they both just want to get research
+% done.
+% Contracts. The second kind of funding is via contracts. Contract funding comes with an explicit
+% contract you must fulfill to keep and obtain more funding. For example, there may be quarterly
+% or monthly reports, or on-site visits by the sponsor. Contract funding typically also has
+% deliverables that are expected at regular intervals during the funding period. One example of
+% contract funding is DARPA funding, which comes with a number of deliverables and reporting
+% requirements. Contract funding is less aligned with the PIs than grant funding: apart from doing
+% the core research, the PI’s students will have to spend time on creating the deliverables, and the
+% PI herself has to spend a significant amount of time on reports and meetings.
+% 40
+% NSF vs DARPA. NSF is a good example of grant funding, while DARPA is a good example of
+% contract funding. Both these agencies fund a lot of computer science research. DARPA requires
+% deliverables and reports to be produced regularly during the grant duration. NSF only requires a
+% report to be submitted for each year of the grant duration. You can apply for NSF funds as
+% someone on a visa, while DARPA and other similar government funding programs might require
+% you to be a US citizen or a permanent resident.
+% University funding. Your university, or your college, or your department may choose to give you
+% funding. Typically when you start, you get a “startup” package that allows you to kickstart your
+% research group. University funding is “internal” funding in a way, and as such has fewer
+% restrictions than government funding. But this really depends on the specific university. I know
+% universities where startup funding cannot be used for covering summer salary of PIs. Apart
+% from startup funding, your university might have internal competitions to award “seed” funding
+% (typically under 50K) that you can then use to get preliminary results and later apply for
+% traditional grants.
+% University funding is also associated with timelines. For example, startup funding usually has to
+% be spent before you get tenure. However, university funding deadlines might be a little more
+% flexible than grant or contract funding deadlines.
+% University funding typically does not have associated indirect costs (as it is already university
+% money). You still have to pay fringe support for your students and your own summer salary.
+% An interesting sub-category of university funding are Research Centers. A research center is
+% funded by sponsors to carry out a certain mission over a span of around 5 years. The research
+% center has a program director who is responsible for allocating the funding. PIs at the university
+% apply for funding from the research center (a sort of internal grant), and are allocated funding.
+% Obtaining funding in this manner is easier than competing for an external grant. Particularly for
+% assistant professors who are just starting, research centers can provide seed funding before the
+% first grants are obtained. Check out the ADA center at Michigan for a good example.
+% Gift funding or unrestricted research funding. Companies or charitable foundations may
+% choose to fund your research in the form of “gifts” or unrestricted research grants. This is the
+% best kind of funding, for three reasons: indirect costs are typically not levied on these gifts, there
+% is no associated timeline by when you should spend them, and you can use them for a broader
+% range of things than university and grant/contract funding.
+% Indirect costs are not levied because the gift-giver can specify that it cannot be used for indirect
+% costs (the university has to honor this). Since it is an unrestricted gift, there are no limits on
+% what you can spend them on, or by when you should spend them. This doesn’t mean you can
+% spend the money on personal expenses. All expenses still have to be approved by the university,
+% and should be used for research-related expenses. But things like taking your research group
+% out for a meal can be expensed during gift funding (but not government funding).
+% 41
+% Indirect costs not being levied can make a huge difference. For example, in our budget from
+% earlier, if you were using gift funding to fund the student, you don’t have to pay indirect costs,
+% and thus get over \$30K back. Considering a student costs \$63K, you get to sponsor the student
+% for 1.5 years rather than 1 year!
+% How should you use your funding? Let us say that you are lucky enough to have all four kinds of
+% funding. And you have an expense that could be assigned to any of these funding (note that this
+% is often not the case: you can’t use a grant on topic X to support a student working on Y if X and
+% Y are not closely related). What should you use?
+% The typical rule is “Use the most restricted funding that will expire the earliest”. Thus, you should
+% always use the appropriate grant/contract funding first, then your university funding, and finally
+% gift funding. Gift funding is the most flexible, so use it only when you have no other option.
+% Always spend your gift funding on expenses that will have overhead (if funded from grants),
+% such as students. Do not spend it on things like equipment (be careful to ask what counts as
+% equipment at your university) which do not have overhead even when expensed to grants.
+% Writing a grant with other PIs. While you could write smaller grants as the sole PI, typically for
+% larger grants, you need to collaborate with other PIs and submit as a team. For example, you
+% cannot submit NSF Large grants as a solo PI. One thing to remember though: all the funding is
+% shared between the PIs. For example, imagine a grant for one million USD. It sounds amazing!
+% However, if the grant was awarded to a team of two PIs, and the grant duration is five years,
+% each PI can only support one student on it for the grant duration (100K * 2 * 5). Similarly, if a
+% grant of 10 million USD was won by a team of five PIs, each PI can only support one student for
+% four years (100K * 5 * 4). So in the end, the share of each PI is similar or lower than that of an
+% NSF CAREER.
+% This is not to say you should not collaborate with other PIs to write grants. Collaboration often
+% brings new ideas and research directions that are unlikely or impossible when working alone.
+% And bigger grants often provide prestige: for example, landing an NSF Large grant is significant
+% and puts you on the map. But it is useful to calculate how much you expect to receive if a grant
+% proposal is funded.
+% 3.3 The grant application process
+% So how does one go about submitting a grant? There are several steps in the process:
+% 1. The funding agency first puts out a call for proposals. For example, check out the NSF
+% call for CRII grants, a startup grant meant for new PIs who have not yet received a grant.
+% Note that PIs at many R1 universities are not eligible for CRII. This was posted on May 18,
+% 2022. Each call has a solicitation page with details about the program and how the
+% proposal should be structured – you should read this carefully.
+% 42
+% 2. In some cases, there are webinars where the program managers explain the grant
+% opportunity. CRII had such a webinar on Jul 13 2022. I highly recommend attending
+% these webinars if it is your first time submitting for a particular grant opportunity.
+% 3. In some cases, the funding agency might want you to submit a preliminary proposal.
+% They will review the preliminary proposals and inform you if they would like you to go
+% ahead with the full proposal. This is not the case for CRII.
+% 4. There will be a due date by which you need to submit the proposal. For the CRII, this is
+% Sep 19 2022. The proposal will have a number of parts, and the solicitation will clearly lay
+% out what the proposal should look like.
+% 5. You write the proposal, and sometime before you submit, you send it to the Office of
+% Sponsored Research (OSP or OSR) at your university. They have to approve the proposal
+% before you can submit, and they may request to see your proposal a week or 10 days
+% before you submit. Please make sure you get a draft to the OSP on time!
+% 6. Finally, you submit the proposal, and you wait. To be exact, your admin/OSP submits for
+% you, so you want to let them know when you are done. You and your admin/OSP will be
+% submitting a lot of proposals together; make sure you get to know the folks doing this. I
+% only have experience with NSF proposals, but these usually take 6-8 months to be
+% reviewed. In some cases, it might take even longer.
+% 7. If you are lucky enough to get your proposal funded, you might get an email or a call from
+% the NSF program manager. They might ask you to slightly modify your budget. For
+% example, I was asked to reduce my NSF CAREER budget (and appropriately reduce the
+% scope of my tasks as well). You will be asked to provide an abstract that will be displayed
+% on the NSF website. For example, check out the abstract for my CAREER award.
+% 8. Finally, you will get an official email that your grant has been awarded, it will become
+% public on the NSF website, and you will get the first installment delivered to your
+% university. You can then begin spending it in accordance with what you proposed.
+% 9. In the unlucky case that your proposal is not funded, you will get back three or more
+% reviews from NSF. These reviews tend to be short, and nowhere as detailed as paper
+% reviews. As such, it can be hard to figure out what to do after getting a rejection. My
+% advice is to go talk to senior faculty in your department if this happens, and they should
+% be able to guide you in the right direction.
+% 3.3 Writing your first grant proposal
+% Writing your first grant can be a stressful experience, since it is quite different from writing a
+% research paper. There are a number of differences.
+% First, the audience for a grant proposal is broader than that of a research paper; it is guaranteed
+% that non-experts in your area will be reading and evaluating your proposal (for example, if I write
+% 43
+% a grant proposal about storage research, I can expect it to be read by someone working in
+% systems broadly, but not necessarily someone doing storage research). As a result, it has to be
+% much broader and more accessible than a research paper. Proposal reviewers might also have a
+% smaller amount of time to spend on each proposal – some funding agencies ask reviewers to
+% read a dozen proposals in around a month when you serve on their panel. You can imagine how
+% much time each proposal would get. Thus, your proposal must be easy to skim and read, must
+% get to the major points quickly, and must be accessible to a broad audience.
+% Second, in a research paper, you are reporting on something that is already done. The focus is
+% on the technical details: how you were able to achieve the result, what are the drawbacks, etc.
+% However, in a grant proposal, you are talking about things that you will do in the future. In a grant
+% proposal, the most important thing is to be able to inspire the reviewer, and convince them that
+% this is a thing worth doing (and funding). In this respect, it is much more similar to a founder
+% pitching before an investor, than a professor talking to her peers about her research. While the
+% details do matter to some extent, what is much more important is that you paint a big picture of
+% how your proposed research will make the world a better place. This is quite different from
+% writing a research paper, and takes some practice.
+% For writing your first grant proposal, I’d recommend two things. First, talk to folks who have
+% gotten the grant you are submitting for, and ask if you could see their grant proposal. Many folks
+% are happy to share privately. Some folks, like Jeff Bigham and Gillian Hayes, are publicly sharing
+% some of their proposals. Second, write up your draft proposal, and show it to senior folks in your
+% department. They would be able to catch errors that could cause rejection.
+% When submitting grant proposals, do not be disheartened if it gets rejected. As far as I can tell,
+% most proposals, even from experienced faculty, get rejected a few times before getting
+% accepted. Grant rejections are a bit more painful because the waiting time is high, you might
+% only get to submit to a grant opportunity once per year. But with time and repeated submissions,
+% many grant proposals do get funded.
 
-3.4 Serving on grant panels
-44
-3.5 Getting funding from industry
-Since I started at UT, I have been fortunate to receive funding from a number of companies:
-Meta, Google, VMware, and Toyota. How did I get this funding? Unlike with government
-agencies, there is no set process to follow in general (exception being Google with the Google
-Research award).
-Giving talks and networking. The process is more organic: I give talks at various companies
-about my research, and talk to their engineers about how my group’s research can impact them.
-This is useful to gain an understanding of how our work will play out in the real world, and to
-understand the problems they are facing.
-Sometimes after talks like these, the company reaches out to see if they can sponsor our
-research. At other times, the company is looking for academic groups to invest in, and my talk
-puts our group on their radar. Note that you might sometimes receive funding even without
-doing this – good work is always noticed and rewarded (especially if close to the company’s
-interests).
-In rare cases, if you know the folks at the company really well, you can also reach out and say I
-am trying to achieve research goal X, is your company looking to fund anybody in this space.
-But you only do this after building up strong ties with the company over a long period of time.
-You can’t start out by saying “can you give me some money”.
-Networking and building connections always pay out over the long term. So you shouldn’t have
-the mindset of “I gave a talk at company X, I should get some funding from them soon”. Rather,
-you should aim to get your group’s research out there and known by companies in that space.
-The funding will follow later, but time spent networking and giving talks is never wasted.
-3.6 Financial planning for your research group
-Finally, I want to talk a little bit about financial planning for your group. As a professor, you are
-now a manager, and you should ensure that the students working with you are adequately
-funded. If you are unable to fund your graduate students, your department might be able to fund
-them as TAs, though the dept would need some advance notice. For postdocs, this option is not
-there, so you need to make sure you have funding for your postdocs.
-Runway. In general, it is good practice to think six months to 1 year ahead, and make sure you
-have enough funding for existing students/postdocs and folks you are planning to hire. Initially,
-you will have your startup funding, so you start with a full tank of gas. For every
-student/postdoc you hire, you should check how much runway your group has: how long until
-your group runs out of funding. You should ensure you have a runway of 6 months to 1 year.
-45
-This means that if you have 1 year of runway, you need to write grant proposals now (since it
-usually takes about a year to be reviewed and funded). At the start of every semester, review
-your finances with your admin and make sure you have enough of a runway.
-Savings. An important part of financial planning is to always have a little bit extra in your coffers
-than your anticipated expenses. For example, imagine that you got a paper accepted, and there
-is no associated grant you can expense it to. If you don’t have some money left over, you would
-not be able to sponsor your student attending the conference and presenting the paper.
-Summers. Part of this financial planning is also deciding what to do in your summers. I strongly
-advocate ensuring that you can both pay yourself for the summer, and pay your students. This
-might require being conservative in taking on students until you are sure you can pay them.
-There is the temptation to cut back on your own summer salary so that you can pay your
-students; I would recommend avoiding this. While it can be a temporary fix, it is necessary for
-your long-term health, and that of your group, if you maintain your own summer payment.
-If you are able to work at a company for the summer, this can reduce the strain on your group’s
-finances while also helping your research become more grounded. I spent one summer at
-VMware after joining UT Austin, and that time spent was useful for my research. You would
-probably need to reach out much earlier to set this up – the group hosting you would need to
-make sure they have the budget to host you for the summer.
-Coordinating with admins. You want to make sure you and your admin are on the same page
-regarding what funding to use for each expense (sometimes it can be hard to move expenses
-between accounts). Most admins are experienced and get the picture quickly once you describe
-your overall strategy.
-Acknowledgements
-Thanks to Gabriel Parmer, Indranil Gupta, Jeff Bigham, Michael Ekstrand, Michael Freedman,
-Satish Puri, Siddharth Joshi, Solel Pirelli, and Stefan Savage for pointing out errors and providing
-suggestions for the chapter.
-Summary
-In this chapter, we talked about how grants work, what the application process looks like, and
-different sources of funding for research. We went over an example budget that shows all the
-different overheads involved in funding research.
-We have now discussed two of the four main big roles of a professor: managing students and
-managing funding. Next, we will discuss teaching
+% 3.4 Serving on grant panels
+% 44
+% 3.5 Getting funding from industry
+% Since I started at UT, I have been fortunate to receive funding from a number of companies:
+% Meta, Google, VMware, and Toyota. How did I get this funding? Unlike with government
+% agencies, there is no set process to follow in general (exception being Google with the Google
+% Research award).
+% Giving talks and networking. The process is more organic: I give talks at various companies
+% about my research, and talk to their engineers about how my group’s research can impact them.
+% This is useful to gain an understanding of how our work will play out in the real world, and to
+% understand the problems they are facing.
+% Sometimes after talks like these, the company reaches out to see if they can sponsor our
+% research. At other times, the company is looking for academic groups to invest in, and my talk
+% puts our group on their radar. Note that you might sometimes receive funding even without
+% doing this – good work is always noticed and rewarded (especially if close to the company’s
+% interests).
+% In rare cases, if you know the folks at the company really well, you can also reach out and say I
+% am trying to achieve research goal X, is your company looking to fund anybody in this space.
+% But you only do this after building up strong ties with the company over a long period of time.
+% You can’t start out by saying “can you give me some money”.
+% Networking and building connections always pay out over the long term. So you shouldn’t have
+% the mindset of “I gave a talk at company X, I should get some funding from them soon”. Rather,
+% you should aim to get your group’s research out there and known by companies in that space.
+% The funding will follow later, but time spent networking and giving talks is never wasted.
+% 3.6 Financial planning for your research group
+% Finally, I want to talk a little bit about financial planning for your group. As a professor, you are
+% now a manager, and you should ensure that the students working with you are adequately
+% funded. If you are unable to fund your graduate students, your department might be able to fund
+% them as TAs, though the dept would need some advance notice. For postdocs, this option is not
+% there, so you need to make sure you have funding for your postdocs.
+% Runway. In general, it is good practice to think six months to 1 year ahead, and make sure you
+% have enough funding for existing students/postdocs and folks you are planning to hire. Initially,
+% you will have your startup funding, so you start with a full tank of gas. For every
+% student/postdoc you hire, you should check how much runway your group has: how long until
+% your group runs out of funding. You should ensure you have a runway of 6 months to 1 year.
+% 45
+% This means that if you have 1 year of runway, you need to write grant proposals now (since it
+% usually takes about a year to be reviewed and funded). At the start of every semester, review
+% your finances with your admin and make sure you have enough of a runway.
+% Savings. An important part of financial planning is to always have a little bit extra in your coffers
+% than your anticipated expenses. For example, imagine that you got a paper accepted, and there
+% is no associated grant you can expense it to. If you don’t have some money left over, you would
+% not be able to sponsor your student attending the conference and presenting the paper.
+% Summers. Part of this financial planning is also deciding what to do in your summers. I strongly
+% advocate ensuring that you can both pay yourself for the summer, and pay your students. This
+% might require being conservative in taking on students until you are sure you can pay them.
+% There is the temptation to cut back on your own summer salary so that you can pay your
+% students; I would recommend avoiding this. While it can be a temporary fix, it is necessary for
+% your long-term health, and that of your group, if you maintain your own summer payment.
+% If you are able to work at a company for the summer, this can reduce the strain on your group’s
+% finances while also helping your research become more grounded. I spent one summer at
+% VMware after joining UT Austin, and that time spent was useful for my research. You would
+% probably need to reach out much earlier to set this up – the group hosting you would need to
+% make sure they have the budget to host you for the summer.
+% Coordinating with admins. You want to make sure you and your admin are on the same page
+% regarding what funding to use for each expense (sometimes it can be hard to move expenses
+% between accounts). Most admins are experienced and get the picture quickly once you describe
+% your overall strategy.
+% Acknowledgements
+% Thanks to Gabriel Parmer, Indranil Gupta, Jeff Bigham, Michael Ekstrand, Michael Freedman,
+% Satish Puri, Siddharth Joshi, Solel Pirelli, and Stefan Savage for pointing out errors and providing
+% suggestions for the chapter.
+% Summary
+% In this chapter, we talked about how grants work, what the application process looks like, and
+% different sources of funding for research. We went over an example budget that shows all the
+% different overheads involved in funding research.
+% We have now discussed two of the four main big roles of a professor: managing students and
+% managing funding. Next, we will discuss teaching
 
@@
 \chapter{Services}
-These are the services that you are expected to do as a faculty member.
-They are not as important as research, but they are still important.
-They are also a good way to get to know people in your department and in your field.
 
-Moreover, while some services are required (e.g., when your department chair assigns you to some committee), some are not and you can decline (e.g., an invitation to be a PC member).
+You are expected to have certain types of services as a faculty member.
+While some services are required (e.g., when your department chair assigns you to some committee), some are not and you can decline (e.g., an invitation to be a PC member).
 You should be strategic about which services you accept.
@@
 
-\paragraph{Serving on Program Committees (PC)} This is a good way to get to know what is hot in your field, and to get to know people in your field.  It is also a good way to get your name out.
+\subsection{Serving on Program Committees (PC)} This is a good way to get to know what is hot in your field, and to get to know people in your field.  It is also a good way to get your name out.
 You should try to serve on the program committee of the top conferences in your field.
@@
 
-\section{Serving on NSF Panels} You should serve on NSF panels.  It is a great way to learn about the grant process: how to distinguish between good and bad proposals and how people review propsoals.
+\subsection{Serving on NSF Panels} You should serve on NSF panels.
+It is a great way to learn about the grant process: how to distinguish between good and bad proposals and how people review proposals.
 It is also a good way learn what other people are doing, what is hot, and what is not. In short, it is a good way to learn how to write successful proposals.
@@
 
-\textbf{Advice}: Similarly to serving in PC, you can cold email progarm managers in your area and ask to be on a panel.
-Once you have served, you will likly be invited to serve again.
+\textbf{Advice}: Similarly to serving in PC, you can cold email program managers in your area and ask to be on a panel.
+Once you have served, you will likely be invited to serve again.
 Unfortunately, I do not have much experience with non-NSF panels or proposals.
 
+\subsection{Volunteer for more Leadership roles} Once you have served on a few PCs and NSF panels, you should aim for more leadership roles.
+For example, you can volunteer to be a chair of certain sessions/tracks of a large conference, or even a general chair of a workshop or conference.
+These roles are more time consuming, but they are also more rewarding (e.g., more people will know about you) and in general will help your career, especially your tenure case.
+
 \section{Service to the Department}
+Your department chair will assign you to some committees to help run the department.
+Typically for tenure track faculty the services are lighter than tenured ones (e.g., less leadership roles).
+For example, I have served on the \emph{PhD and/or MS admission committee (adcom)}, \emph{faculty search committee}, and \emph{awards committee} when I was at UNL.
+Being a member of the PhD adcom is very common for new assistant professors, as it is a good way to recruit students and build your lab.
+
 \section{Service to the University}
@@
 
-One part of the professor job that doesn’t get a lot of attention is service. However,
-understanding and managing this component is important, both for your own tenure case and
-for the wider community. Let us start by talking about what service is, and why it is important.
-7.1 Defining and managing service
-So what do I mean when I say service? I would loosely define service as optional tasks that your
-community requests from you. You are free to decline these tasks without a direct impact on
-your career (there might be indirect impacts). These tasks are important to keep the community
-running.
-For example, your peer researchers may ask you to serve on the program committee for a
-conference, or review a paper for a journal. You will get a lot of these requests, and you are not
-expected to say yes to all of them. Your department will expect you to do some service each
-year, and there is an expectation that you get invited to serve on the PCs of the top conferences
-in your research area. But beyond that, you don’t get any benefits by doing a lot of service.
-Why should you do service? To put it in a nutshell, doing service is part of being a good citizen,
-both inside your department, and inside your wider academic community. This is because of the
-way academia operates, fully dependent on freely volunteered labor.
-For example, imagine that you want to run a conference. You need someone to organize the
-conference, and someone to review submissions and select papers for the program. One way
-to do this is to recruit reviewers, pay them for their time, and pay other folks to handle the
-logistics of running the conference. Speakers would get paid. All the money would come from a
-combination of sponsors and the fees that attendees provide. This is how industrial
-conferences operate.
-Academia uses a different model. People volunteer their time to act as reviewers or conference
-organizers. This allows the conference organizers to reduce the cost that needs to be paid by
-attendees (even so, the cost is usually several hundreds of US dollars). USENIX uses a slightly
-different model where the conference organizers are paid professionals, but the technical
-reviewing is still done by volunteers. This results in a slightly higher registration fee for
-attendees (compared to when everything is done by volunteers).
-Thus, almost everything in academia runs on volunteer labor. All the events you attend, the
-conferences and journals you publish in, the panels you enjoy, the awards that are given out, all
-depend on people volunteering their time. Since you get a net benefit from the time volunteered
-by others, it is only fair that you do your share of volunteering as well.
-72
-There are some other fringe benefits to volunteering. One is that you get your name out. When
-you are an assistant professor and you are struggling to establish yourself, it can be useful to
-have folks know you and recognize your name. The second benefit is that you get to meet folks
-in the course of doing service and form your network; this is particularly useful when you are
-just starting out.
-Why should you not do service? While it is important to do service, I’ve seen folks go overboard
-on this (and did too much myself in my first year). Doing service feels good: you are doing
-something for your community, and it feels like you are getting your name out. And it is hard to
-say no when folks you respect request you to join a committee.
-But you should always, always remember: you will not get tenure for service. No matter how
-good the service is. The first consideration at R1 universities is always research. If your research
-is not good, having done amazing service will not get you tenure.
-Apart from not helping you with tenure, taking on too many service commitments will make you
-bad at service. If you say yes to too many concurrent PCs, you will have a tough time doing a
-good job at reviewing papers for all of them. Nobody wins in this situation: the conferences and
-authors get crappy reviews, and you get a reputation for writing bad reviews.
-Managing service commitments. As an assistant professor, you should do enough service to be
-a good citizen, but not so much that it affects your other job responsibilities. You should
-conserve your limited energy for research and teaching. Be thoughtful and intentional about the
-service requests you accept – always think about how much time you are committing down the
-line (even if it is several months or a year away). The way to think about it is that you can do two
-to four big service requests each year. A big service request is something like an organizational
-role at a conference, or a program committee assignment where you will be reviewing 10+
-papers. You should really be doing only a few of these each year. Smaller service requests, such
-as being an external reviewer for a conference, you can do more of; but be careful that they don’t
-pile up.
-Saying no. How do you say no politely, especially when it is someone you respect and look up
-to? You should remember that folks who run conferences expect a lot of people to say no. It
-happens all the time, and if you decline, they will not take it personally. You don’t need to worry,
-“If I say no this year, maybe they will not invite me ever again”. I can relate because I had this
-exact worry at the beginning. I’m happy to report that I have declined to review at conferences,
-and still been invited the following year. As long as you do good research and write good
-reviews, you will be invited back. So do not let this consideration make you say yes when you
-don’t have the time to do a good job.
-I usually say something like: “Thank you for inviting me. Unfortunately, I am overloaded on
-service commitments this year, and I don’t think I will have enough time to do a good job.
-Therefore I will have to decline”. As you get a bit more senior, you could recommend a junior
-faculty member they could invite in your place.
-73
-7.2 Serving on program committees and editorial boards
-When you get started as an assistant professor, serving on program committees will form the
-bulk of your service. At first, it can be super flattering to be invited to a program committee: I
-definitely remember my first such invitation. You now have a voice at the table in determining
-the program! You can review papers alongside researchers you looked up to. As a result, you
-pretty much say yes to all the invitations you initially get.
-This is not a great strategy, though: you end up over-committing yourself in your initial years. I
-would recommend saying yes only in the following cases:
-1) The conference is your home venue, where you have been publishing in grad school, and
-where you hope to publish in the future.
-2) It is the top (or one of the top) conferences in your area.
-3) It is a community that you want to join, or an area that you want to publish in.
-4) It is a workshop specializing in your area.
-For example, I publish regularly at systems conferences such as ATC/FAST, so I tend to review
-papers for them. SOSP and OSDI are the top conferences in my area, so I almost always accept
-PC invitations for them. HotOS and HotStorage are the relevant workshops for my area, so I
-review papers for them as well.
-Don’t accept invitations to be on the PC for conferences that you have never published in, and
-are unlikely to do so in the future. Don’t accept invitations just because you know the PC chair if
-you are already committed to other PCs.
-Overall, a good number to aim for is about 1-2 big PCs each year (10+ papers per PC) and 1-2
-smaller ones (5+ papers per PC). Between research and teaching, it will be hard to do a good job
-with more PCs.
-University expectations regarding PCs. Generally, your department or university likes to see that
-you are being invited to serve on the PCs of the top conferences in your area. If you become an
-assistant professor and receive no PC invitations, that can be a bit of a red flag. However,
-beyond expecting to see some service (a few PCs each year), the department/university doesn’t
-really care how much service you do – they assume you will manage your service workload
-efficiently.
-How much time does reviewing a paper take? This really varies by area. I would say anywhere
-between one to four hours per paper is common. Do not spend more than half a day on a single
-paper. You should read actively, trying to get the main ideas and insights, which will enable you
-to write a good review. You should aim to read the paper only once (this gets easier with
-practice), taking notes as you go. Having questions you want to answer as you read will help
-make reviewing easier.
-74
-If you find yourself spending more than 1 day a week on average (or 8 hours per week) for
-reviewing papers, you have accepted too many PC invitations. It can be hard to properly track
-this though, since you have a lot of activity near the reviewing deadlines, and not much on other
-weeks.
-I’ve mostly talked about program committees since conferences are considered the primary
-publication venue in computer science, but much of the advice applies to journals as well. Only
-review for journals where you publish regularly, and track and manage how many submissions
-you accept to review.
-7.3 Serving on dissertation committees
-Another form of service is participating on dissertation committees, where you read the
-dissertation and attend the thesis proposal/defense. This can take up a lot more time, since you
-have to read an entire dissertation. It helps if you already know the students work, since that
-will form a significant portion of the dissertation. Still, serving on a dissertation committee can
-be an entire day’s worth of work.
-Requests for serving on dissertation committees can come from within the department or
-externally. The internal ones are harder to say no, since these are the colleagues you work with
-on a daily basis. Nevertheless, track how much time you are spending on this. Don’t say yes to
-more than one or two dissertation committees per semester/quarter.
-From the university viewpoint, I don’t think there are concrete expectations regarding serving on
-dissertation committees. Not being on dissertation committees early on in your career will not
-be seen as a red flag (to the same extent that not being on PCs would be).
-7.4 Serving as a session chair
-Another form of service is serving as the Session Chair at a conference. The session chair will
-introduce the session (typically 1 to 1.5 hours) at the front, introduce the speakers, and
-moderate the questions. The session chair also kicks off the QnA if there are no questions from
-the audience at first.
-Being a session chair is a fantastic service opportunity, and you should definitely sign up
-whenever you get the chance. It gets you face-time in front of the community, you get your name
-out, and the work required for being a good session chair is modest. Grab session chair
-opportunities with both hands when you are starting out!
-7.5 Organizational service roles
-As you become a bit more experienced, you will be offered organizational roles such as being
-the Publicity Chair, the Web Chair, the Poster Chair, or the Program Committee chair. Of these,
-being the chair of the program committee is most prestigious, since you are responsible for
-75
-delivering a high quality program in coordination with your chosen program committee. So if you
-are invited to be the PC chair of a conference or workshop you normally publish in, you should
-accept the invitation. Your university or department will care about whether you have been
-invited to be the PC chair (once you become a bit more senior).
-However, the other roles involve a lot of work and little recognition. Ideally, these would be done
-by professionals (as in the case of USENIX). However, for non-USENIX conferences, someone
-has to step up and do them. Be aware of the work-reward ratio if you choose to accept an
-assignment like this though.
-During my time as an assistant professor, I co-chaired HotStorage 2020 and SyStor 2021. I was
-also one of the General Chairs for SOSP 21. I enjoyed doing all three roles, but it was definitely a
-lot of work! I would not recommend doing any of these roles (especially being General Chair)
-early on as an assistant professor; it is a better fit when you are close to going up for tenure.
-7.6 Departmental and university service
-Apart from serving on program committees, editorial boards, and organizing conferences and
-workshops, you are also expected to do “local” service for your department and university. This
-can take multiple forms.
-At the department level, this means you will serve on one or more committees. Assistant
-professors are usually assigned to either the PhD admissions committee or the faculty
-recruiting committee. Both committees are extremely important and directly influence the future
-of the department. Other committees include Faculty Evaluation Committees, Budget
-Committees, and Promotion Committees that you can participate on once you become more
-senior. Expect to spend at least a couple of weeks (put together, it will be spread over the
-semester) on departmental committee work each year.
-At the university level, there will be committees deciding overall university policy. Typically, your
-department or college will need to be represented on these committees. As an assistant
-professor, it will be rare for you to be nominated by your department for these committees;
-typically, more senior professors participate in these committees.
-Summary
-The most important thing to remember about service is that it is like chores around the house;
-you need to regularly clean your house, water your plants, etc. You do this because you want a
-nice place to live in; similarly, service provides you a nice community to do research in. However,
-carefully manage the amount of service you do. You will be bombarded with requests, and it is
-all too easy to say yes to too many requests. Don’t accept so many service requests that it
-impacts your other responsibilities as a professor.
-76
+% One part of the professor job that doesn’t get a lot of attention is service. However,
+% understanding and managing this component is important, both for your own tenure case and
+% for the wider community. Let us start by talking about what service is, and why it is important.
+% 7.1 Defining and managing service
+% So what do I mean when I say service? I would loosely define service as optional tasks that your
+% community requests from you. You are free to decline these tasks without a direct impact on
+% your career (there might be indirect impacts). These tasks are important to keep the community
+% running.
+% For example, your peer researchers may ask you to serve on the program committee for a
+% conference, or review a paper for a journal. You will get a lot of these requests, and you are not
+% expected to say yes to all of them. Your department will expect you to do some service each
+% year, and there is an expectation that you get invited to serve on the PCs of the top conferences
+% in your research area. But beyond that, you don’t get any benefits by doing a lot of service.
+% Why should you do service? To put it in a nutshell, doing service is part of being a good citizen,
+% both inside your department, and inside your wider academic community. This is because of the
+% way academia operates, fully dependent on freely volunteered labor.
+% For example, imagine that you want to run a conference. You need someone to organize the
+% conference, and someone to review submissions and select papers for the program. One way
+% to do this is to recruit reviewers, pay them for their time, and pay other folks to handle the
+% logistics of running the conference. Speakers would get paid. All the money would come from a
+% combination of sponsors and the fees that attendees provide. This is how industrial
+% conferences operate.
+% Academia uses a different model. People volunteer their time to act as reviewers or conference
+% organizers. This allows the conference organizers to reduce the cost that needs to be paid by
+% attendees (even so, the cost is usually several hundreds of US dollars). USENIX uses a slightly
+% different model where the conference organizers are paid professionals, but the technical
+% reviewing is still done by volunteers. This results in a slightly higher registration fee for
+% attendees (compared to when everything is done by volunteers).
+% Thus, almost everything in academia runs on volunteer labor. All the events you attend, the
+% conferences and journals you publish in, the panels you enjoy, the awards that are given out, all
+% depend on people volunteering their time. Since you get a net benefit from the time volunteered
+% by others, it is only fair that you do your share of volunteering as well.
+% 72
+% There are some other fringe benefits to volunteering. One is that you get your name out. When
+% you are an assistant professor and you are struggling to establish yourself, it can be useful to
+% have folks know you and recognize your name. The second benefit is that you get to meet folks
+% in the course of doing service and form your network; this is particularly useful when you are
+% just starting out.
+% Why should you not do service? While it is important to do service, I’ve seen folks go overboard
+% on this (and did too much myself in my first year). Doing service feels good: you are doing
+% something for your community, and it feels like you are getting your name out. And it is hard to
+% say no when folks you respect request you to join a committee.
+% But you should always, always remember: you will not get tenure for service. No matter how
+% good the service is. The first consideration at R1 universities is always research. If your research
+% is not good, having done amazing service will not get you tenure.
+% Apart from not helping you with tenure, taking on too many service commitments will make you
+% bad at service. If you say yes to too many concurrent PCs, you will have a tough time doing a
+% good job at reviewing papers for all of them. Nobody wins in this situation: the conferences and
+% authors get crappy reviews, and you get a reputation for writing bad reviews.
+% Managing service commitments. As an assistant professor, you should do enough service to be
+% a good citizen, but not so much that it affects your other job responsibilities. You should
+% conserve your limited energy for research and teaching. Be thoughtful and intentional about the
+% service requests you accept – always think about how much time you are committing down the
+% line (even if it is several months or a year away). The way to think about it is that you can do two
+% to four big service requests each year. A big service request is something like an organizational
+% role at a conference, or a program committee assignment where you will be reviewing 10+
+% papers. You should really be doing only a few of these each year. Smaller service requests, such
+% as being an external reviewer for a conference, you can do more of; but be careful that they don’t
+% pile up.
+% Saying no. How do you say no politely, especially when it is someone you respect and look up
+% to? You should remember that folks who run conferences expect a lot of people to say no. It
+% happens all the time, and if you decline, they will not take it personally. You don’t need to worry,
+% “If I say no this year, maybe they will not invite me ever again”. I can relate because I had this
+% exact worry at the beginning. I’m happy to report that I have declined to review at conferences,
+% and still been invited the following year. As long as you do good research and write good
+% reviews, you will be invited back. So do not let this consideration make you say yes when you
+% don’t have the time to do a good job.
+% I usually say something like: “Thank you for inviting me. Unfortunately, I am overloaded on
+% service commitments this year, and I don’t think I will have enough time to do a good job.
+% Therefore I will have to decline”. As you get a bit more senior, you could recommend a junior
+% faculty member they could invite in your place.
+% 73
+% 7.2 Serving on program committees and editorial boards
+% When you get started as an assistant professor, serving on program committees will form the
+% bulk of your service. At first, it can be super flattering to be invited to a program committee: I
+% definitely remember my first such invitation. You now have a voice at the table in determining
+% the program! You can review papers alongside researchers you looked up to. As a result, you
+% pretty much say yes to all the invitations you initially get.
+% This is not a great strategy, though: you end up over-committing yourself in your initial years. I
+% would recommend saying yes only in the following cases:
+% 1) The conference is your home venue, where you have been publishing in grad school, and
+% where you hope to publish in the future.
+% 2) It is the top (or one of the top) conferences in your area.
+% 3) It is a community that you want to join, or an area that you want to publish in.
+% 4) It is a workshop specializing in your area.
+% For example, I publish regularly at systems conferences such as ATC/FAST, so I tend to review
+% papers for them. SOSP and OSDI are the top conferences in my area, so I almost always accept
+% PC invitations for them. HotOS and HotStorage are the relevant workshops for my area, so I
+% review papers for them as well.
+% Don’t accept invitations to be on the PC for conferences that you have never published in, and
+% are unlikely to do so in the future. Don’t accept invitations just because you know the PC chair if
+% you are already committed to other PCs.
+% Overall, a good number to aim for is about 1-2 big PCs each year (10+ papers per PC) and 1-2
+% smaller ones (5+ papers per PC). Between research and teaching, it will be hard to do a good job
+% with more PCs.
+% University expectations regarding PCs. Generally, your department or university likes to see that
+% you are being invited to serve on the PCs of the top conferences in your area. If you become an
+% assistant professor and receive no PC invitations, that can be a bit of a red flag. However,
+% beyond expecting to see some service (a few PCs each year), the department/university doesn’t
+% really care how much service you do – they assume you will manage your service workload
+% efficiently.
+% How much time does reviewing a paper take? This really varies by area. I would say anywhere
+% between one to four hours per paper is common. Do not spend more than half a day on a single
+% paper. You should read actively, trying to get the main ideas and insights, which will enable you
+% to write a good review. You should aim to read the paper only once (this gets easier with
+% practice), taking notes as you go. Having questions you want to answer as you read will help
+% make reviewing easier.
+% 74
+% If you find yourself spending more than 1 day a week on average (or 8 hours per week) for
+% reviewing papers, you have accepted too many PC invitations. It can be hard to properly track
+% this though, since you have a lot of activity near the reviewing deadlines, and not much on other
+% weeks.
+% I’ve mostly talked about program committees since conferences are considered the primary
+% publication venue in computer science, but much of the advice applies to journals as well. Only
+% review for journals where you publish regularly, and track and manage how many submissions
+% you accept to review.
+% 7.3 Serving on dissertation committees
+% Another form of service is participating on dissertation committees, where you read the
+% dissertation and attend the thesis proposal/defense. This can take up a lot more time, since you
+% have to read an entire dissertation. It helps if you already know the students work, since that
+% will form a significant portion of the dissertation. Still, serving on a dissertation committee can
+% be an entire day’s worth of work.
+% Requests for serving on dissertation committees can come from within the department or
+% externally. The internal ones are harder to say no, since these are the colleagues you work with
+% on a daily basis. Nevertheless, track how much time you are spending on this. Don’t say yes to
+% more than one or two dissertation committees per semester/quarter.
+% From the university viewpoint, I don’t think there are concrete expectations regarding serving on
+% dissertation committees. Not being on dissertation committees early on in your career will not
+% be seen as a red flag (to the same extent that not being on PCs would be).
+% 7.4 Serving as a session chair
+% Another form of service is serving as the Session Chair at a conference. The session chair will
+% introduce the session (typically 1 to 1.5 hours) at the front, introduce the speakers, and
+% moderate the questions. The session chair also kicks off the QnA if there are no questions from
+% the audience at first.
+% Being a session chair is a fantastic service opportunity, and you should definitely sign up
+% whenever you get the chance. It gets you face-time in front of the community, you get your name
+% out, and the work required for being a good session chair is modest. Grab session chair
+% opportunities with both hands when you are starting out!
+% 7.5 Organizational service roles
+% As you become a bit more experienced, you will be offered organizational roles such as being
+% the Publicity Chair, the Web Chair, the Poster Chair, or the Program Committee chair. Of these,
+% being the chair of the program committee is most prestigious, since you are responsible for
+% 75
+% delivering a high quality program in coordination with your chosen program committee. So if you
+% are invited to be the PC chair of a conference or workshop you normally publish in, you should
+% accept the invitation. Your university or department will care about whether you have been
+% invited to be the PC chair (once you become a bit more senior).
+% However, the other roles involve a lot of work and little recognition. Ideally, these would be done
+% by professionals (as in the case of USENIX). However, for non-USENIX conferences, someone
+% has to step up and do them. Be aware of the work-reward ratio if you choose to accept an
+% assignment like this though.
+% During my time as an assistant professor, I co-chaired HotStorage 2020 and SyStor 2021. I was
+% also one of the General Chairs for SOSP 21. I enjoyed doing all three roles, but it was definitely a
+% lot of work! I would not recommend doing any of these roles (especially being General Chair)
+% early on as an assistant professor; it is a better fit when you are close to going up for tenure.
+% 7.6 Departmental and university service
+% Apart from serving on program committees, editorial boards, and organizing conferences and
+% workshops, you are also expected to do “local” service for your department and university. This
+% can take multiple forms.
+% At the department level, this means you will serve on one or more committees. Assistant
+% professors are usually assigned to either the PhD admissions committee or the faculty
+% recruiting committee. Both committees are extremely important and directly influence the future
+% of the department. Other committees include Faculty Evaluation Committees, Budget
+% Committees, and Promotion Committees that you can participate on once you become more
+% senior. Expect to spend at least a couple of weeks (put together, it will be spread over the
+% semester) on departmental committee work each year.
+% At the university level, there will be committees deciding overall university policy. Typically, your
+% department or college will need to be represented on these committees. As an assistant
+% professor, it will be rare for you to be nominated by your department for these committees;
+% typically, more senior professors participate in these committees.
+% Summary
+% The most important thing to remember about service is that it is like chores around the house;
+% you need to regularly clean your house, water your plants, etc. You do this because you want a
+% nice place to live in; similarly, service provides you a nice community to do research in. However,
+% carefully manage the amount of service you do. You will be bombarded with requests, and it is
+% all too easy to say yes to too many requests. Don’t accept so many service requests that it
+% impacts your other responsibilities as a professor.
+% 76
 

### 2024-03-22 demystify.tex
updates

@@
 While you might not have control over LORs (\autoref{sec:lor}) or where your go to school (\autoref{sec:your-school}), you do over your
-statement of purpose (SOP) or personal statement\footnote{Some schools separate these documents and ask you to write both: SOP, which focuses on research experiences, and Personal Statement, which is everything more personal, e.g., why PhD, challenges, diversity, etc}!
+statement of purpose (SOP) or personal statement\footnote{Few schools separate these documents and ask you to write both: SOP, which focuses on research experiences, and personal statement, which is everything more personal, e.g., why PhD, challenges, diversity, etc}!
 So write it well because it could make a big difference.

### 2024-03-22 demystify.tex
updates

@@
 However, it is understandable that many students do not have the opportunities to publish papers.
-Thus, general writings, even those under submissions or rejected, would still be good (and much better than nothing).
-But be sure to upload your papers with your application and talk about them in your statement (see \autoref{sec:research-statement}).
+Thus, general writings, even those under submissions or rejected, would still be good (much better than nothing).
+Be sure to upload your papers with your application and talk about them in your statement (see \autoref{sec:research-statement}).
 Note that local conferences and non-English journals or conferences do
@@
 
-\paragraph{Research Experiences} Work experiences at \emph{well-known research laboratories}, such as Microsoft Research, can strengthen your
+\paragraph{Work Experiences} Work experiences at \emph{well-known research laboratories}, such as Microsoft Research, can strengthen your
 application.
@@
 
-\textbf{Advice}: In general, don't waste your time taking GRE.
+\textbf{Advice}: In general, don't waste time on the GRE.
 However, if you took it and have good scores then you can include and talk about them in your application (but don't expect them to make much difference).
+But if your scores are bad, then you should not include them in your application, which can be a \red{red flag}.
 
@@
 
-\textbf{Advice}: Just do well enough to pass minimum requirement set by the university, which is required for TA (\autoref{sec:funding}).
+\textbf{Advice}: Just do well enough to pass minimum requirement set by the university. %, which is required for TA (\autoref{sec:funding}). 
 Just as with grades (\autoref{sec:your-school}) and GRE, having a high scores in these tests might not help, but having too low scores can a \red{red flag} and sometimes results in an automatic rejection (\autoref{sec:why-rejected}), e.g., below the minimum requirement.
@@
           \begin{itemize}
-            \item TOEF: 88 pts in total AND $\ge 20$ points in each subsection OR
+            \item TOEFL: 88 pts in total AND $\ge 20$ points in each subsection OR
             \item IELTS: $\ge 6.5$ OR
@@
 However, if you have research-based work experiences, then they can be useful and should be mentioned in your SOP and LORs.
-For example, if you have worked at a research lab (e.g., Microsoft, DoE, VinAI in Vietnam, etc) (\autoref{sec:research-experiences}), then you should mention that in your SOP and ask your supervisor to write about your research experiences and potentials in their LOR (\autoref{sec:lor}).
+For example, if you have worked at a research lab, e.g., Microsoft, DoE, VinAI in Vietnam, etc (\autoref{sec:research-experiences}), then you should mention that in your SOP and ask your supervisor to write about your research experiences and potentials in their LOR (\autoref{sec:lor}).
 
@@
 We do not care much if you have taken on online Coursera AI course or have a professional certificate from Microsoft.
-However, as mentioned in \autoref{sec:non-stem}, if you do not have a CS background, you can use these to show your ability to handle CS courses.
+However, as mentioned in \autoref{sec:non-stem}, if you do not have a CS background, you might be able to use these to show your ability to handle CS courses.
 

### 2024-03-23 faculty.tex
Updates from Overleaf

@@
 \section{The 5 jobs of an assistant professor}
-An tenure-track professor in CS (and likely in other fields) at an R1 university typically has \emph{five} jobs:
+An tenure-track professor in CS (and likely in other STEM fields) at an R1 university typically has \emph{five} jobs:
 

### 2024-03-23 faculty.tex
updates

@@
 
-\subsection{Serving on Program Committees (PC)} This is a good way to get to know what is hot in your field, and to get to know people in your field.  It is also a good way to get your name out.
-You should try to serve on the program committee of the top conferences in your field.
+\subsection{Serving on Program Committees (PC)} You should try to serve on the PC  of the \emph{top} conferences in your field.
+This is a good way to get to know what is hot in your field, and to get to know people in your field.
+It is also a good way to get your name out.
 
-\textbf{Advice}: You can ask senior colleagues or even your PhD advisor/collaborators to recommend you to serve on a PC.
-You can also directly email the PC chairs and ask to serve on the PC.
-For example, I directly email OOPSLA and PLDI PC chairs and asked to serve on the PC, and they accepted my request (I do have papers at OOPSLA and PLDI when I asked).
-I tried this with SE conferences  e.g., ICSE (and of course also nominate myself through the official form), and it never work, even though I have many papers at ICSE.
+\textbf{Advice}: You can ask senior colleagues or even your PhD advisor/collaborators to recommend you to serve on a PC. You can also directly email the PC chairs and ask to serve on the PC.
 
@@
 
-In general, I enjoy serving on NSF panels much more than serving on PC.
-While the work is more intense and takes more time, I found it to be more rewarding and learned a lot from it.
-Since my first NSF panel in 2019, I have served 7 panels (at least once every year since '19) and I look forward to serving more.
-
 \textbf{Advice}: Similarly to serving in PC, you can cold email program managers in your area and ask to be on a panel.
@@
 For example, you can volunteer to be a chair of certain sessions/tracks of a large conference, or even a general chair of a workshop or conference.
-These roles are more time consuming, but they are also more rewarding (e.g., more people will know about you) and in general will help your career, especially your tenure case.
+These roles are more time consuming, but they can be more rewarding (e.g., more people will know about you) and in general can help your career, especially your tenure case.
+
+\subsection{Personal Experience} I have served on many PCs and NSF panels.
+The initial services are mostly through ``cold'' emails.
+For example, I directly email OOPSLA and PLDI PC chairs and asked, and they accepted my request (I did have papers at OOPSLA and PLDI when I asked).
+I tried this with SE conferences  e.g., ICSE (and also nominate myself through the official form), and it never works, even though I have many papers at ICSE.
+I also emailed NSF program managers and asked to serve on a panel, and they accepted my request.
+
+I felt that I didn't get much out of serving on PC and didn't really like it much. However, I still do it because I feel I should give back to the community.
+On the otherhand, I throroughly enjoy serving on NSF panels.
+While the work is more intense and takes more time, I found it more rewarding and enjoy the discussions.
+Since my first NSF panel in 2019, I have served 7 panels (at least once every year since '19) and I look forward to serving more.
+
+I have also volunteered for leadership roles, e.g., I directly asked the general chair of ICSE'21 to see if I can take more leadership roles and was then invited to serve as a proceeding co-chair.
+In hindsight this particular role is a lot of work and I didn't really like many aspects of it, but it was an eye-openning experience.
 
@@
 Typically for tenure track faculty the services are lighter than tenured ones (e.g., less leadership roles).
-For example, I have served on the \emph{PhD and/or MS admission committee (adcom)}, \emph{faculty search committee}, and \emph{awards committee} when I was at UNL.
-Being a member of the PhD adcom is very common for new assistant professors, as it is a good way to recruit students and build your lab.
+Common committees include PhD and/or MS admission, faculty search, and colloquium committees.
+
+Depends on your preference and how you are perceived, you can take more leadership roles when you are close to tenure, such as being the chair of a committee or chair/director of a program.
+These roles are more time consuming and might involve some politics, but they are also more rewarding and can help your tenure case.
+
+\textbf{Advice}: You should try to be strategic about which committees you serve on that can help build your lab and tenure case.
+For example, serving in the PhD adcom is very common for new assistant professors, as it is a good way to recruit students for your lab.
+Serving on the faculty search committee is also a good way to recruit new faculty members that you like (and also let you know what the job market is like, just in canse you want to move).
+Serving on the colloquium committee is also a good way to invite speakers in your field that can be your \emph{tenure letter writers}.
+
+\subsubsection{Personal} Since coming to GMU, I have been much more active in serving the department.
+I share the vision of the department and want to help achieve it.
+In addition to the typical services load, I volunteer to serve in additional committees and leadership roles.
+For example, I maintain the CSRankings DB for the department, I help organize the Open House to recruit admitted students, I serve in the Dept Chair renewal committee, etc.
+Even though I am not on the search committee, I speak to every candidate and help recruit strong ones.
+I agree to be director of the MS program, which is a lot of work and typically done by tenured faculty.
+
+Sometimes I feel overwhelmed by the services, but I also feel that I am making a difference in the department and that is rewarding.
+I am also building a good relationship with my colleagues and the department chair, and that is important for my tenure case.
+In addition, I learn a lot from being in various roles (e.g., I learned how class scheduling work, how the deparmtnet is run, how to recruit students, how to recruit faculty, how to run a colloquium, etc) and feel more confident in my leadership skills.
+
+In short, if you are enthusiastic about the department, want to make a difference, and \emph{feel comfortable at handling additional load}, you can volunteer for more services and leadership roles.
 
@@
 \section{Other Types of Services}
-\paragraph{Writing Recommendation Letters} You will be asked to write a lot of recommendation letters, for both students and colleagues.
+\paragraph{Writing Recommendation Letters} You will be asked to write recommendation letters for both students and colleagues.
 
-\textbf{Advice}: you should have a template for recommendation letters.
+\textbf{Advice}: %you should have a template for recommendation letters.
 You can find many templates online.

### 2024-03-24 demystify.tex
updates

@@
 So write it well because it could make a big difference.
-In your statement, you have the opportunity to make your application stand out and unique, which can make you \emph{fit} the CS program you're applying to, even if you don't have very strong research experience.
+
+In your statement, you have the opportunity to connect directly with the adcom and make your application stand out and unique.
+Your SOP can convince the adcom that you would \emph{fit} the program, even if you don't have very strong research experience.
 A well-written SOP also shows that you can communicate, which is very important in research, and that you can effectively teach and communicate with students, which is important for TA funding (see \autoref{sec:funding}).
@@
 
-Finally, you should get your SOP reviewed by LoR writers, professors (especially those who have served in adcom), or even graduate PhD students (because they have been through this process).
+Finally, you should get your SOP reviewed by your LoR writers, professors (especially those who have served in adcom), or even graduate PhD students (because they have been through this process).
 
 \begin{commentbox}[Vu:]
-  I often read the research statement and LORs first. If I am
+  I often read SOP and LORs first. If I am
   persuaded by then, I would skim over other factors and advocate for
@@
   convinced, then I will likely recommend rejection (unless I see
-  something standout in other parts).
+  something really standout in other parts).
   \tcblower
@@
 
-\paragraph{English Test} Unless your degrees are from specific countries such as \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-GMU#standard-tests-waiver-eligible-countries}{these}, you will need to
+\paragraph{English Tests} Unless your degrees are from specific countries such as \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-GMU#standard-tests-waiver-eligible-countries}{these}, you will need to
 take standardized English test.
@@
 \textbf{Advice}: Just do well enough to pass minimum requirement set by the university. %, which is required for TA (\autoref{sec:funding}). 
-Just as with grades (\autoref{sec:your-school}) and GRE, having a high scores in these tests might not help, but having too low scores can a \red{red flag} and sometimes results in an automatic rejection (\autoref{sec:why-rejected}), e.g., below the minimum requirement.
+Just as with grades (\autoref{sec:your-school}) and GRE, having a high scores in these tests might not help, but having too low scores can be a \red{red flag} and sometimes results in an automatic rejection (\autoref{sec:why-rejected}), e.g., below the minimum requirement.
 
@@
 
-However, no interviews typically means that you will not likely get an RA (\autoref{sec:ra}), which is offered by an individual faculty (if they want to pay you to do research for them, then likely they want to talk to you first).
-In this case, your application (and TA/fellowship funding) will be reviewed and decided by the adcom.
+However, no interviews typically means that you will not likely get an RA (\autoref{sec:ra}), which is offered by an individual faculty (if they you to do research for them, then they likely will interview you first).
+In the case of no interviews, your application (and TA/fellowship funding) will be reviewed and decided by the adcom.
 
@@
 
-Fortunately, unlike many countries that require finding an adviser and research topic before starting the PhD (\autoref{sec:non-us-differences}), CS PhD programs in the US will typically give you a couple of years to search for advisers and research topics.
+Fortunately, while some non-US programs require finding an adviser and research topic before starting the PhD (\autoref{sec:non-us-differences}), CS PhD programs in the US will typically give you a couple of years to search for advisers and research topics.
 This is especially true if you're admitted with TA (\autoref{sec:ta}), which gives you time to explore and find adviser.
@@
 Another effective way is taking graduate level courses in the topics you are interested in.
-Professors teaching these \emph{special topics courses} and \emph{research seminars} might be recruiting students---giving you even a higher chance.
+Professors teaching \emph{special topics courses} and \emph{research seminars} might be recruiting students---giving you even a higher chance.
 Do well in the class, answer questions, talk to the prof after classes, etc---being stand out.

### 2024-03-24 faculty.tex
updates

@@
     {
-    {\fontsize{48}{50} \selectfont PhD}
+    {\fontsize{40}{42} \selectfont Faculty}
     };
@@
     {
-    {\fontsize{24}{24} \selectfont {{Demystifying the Computer Science}}} \\[0.15in]
-    {\fontsize{24}{24} \selectfont {{PhD Admission in the US}}} \\[0.2in]    
-    {\fontsize{18}{18} \selectfont {{A Handbook for International and Domestic Students}}} \\[0.5in]
+    {\fontsize{24}{24} \selectfont {{Demystifying Tenure Track:}}} \\[0.15in]
+    {\fontsize{18}{18} \selectfont {{Navigating the Faculty Journey In Computer Science}}} \\[1in]    
+    %{\fontsize{18}{18} \selectfont {{A Handbook for International and Domestic Students}}} \\[0.5in]
 
@@
 
-You are expected to have certain types of services as a faculty member.
-While some services are required (e.g., when your department chair assigns you to some committee), some are not and you can decline (e.g., an invitation to be a PC member).
+While only 10\% of your work is officially for services, they are very important for your tenure and career.
+%You are expected to have certain types of services as a faculty member.  
+Some services are required (e.g., when your department chair assigns you to some committee), but some are not and you can decline (e.g., an invitation to be a PC member).
 You should be strategic about which services you accept.
 
-\section{Service to your Professional Community}
+\section{Service to your Research Community}
 
-These are services that you do for your professional community.
+These are services that you do for your research community.
 They will help you to get to know people in your field, and to get your name out.
-Common examples include serving on the program committees of conferences, serving on editorial boards of journal, reviewing papers for journals, serving on grant panels.
+Common examples include serving on the program committees of conferences, serving on editorial boards of journal, and serving on grant panels.
 
-\subsection{Serving on Program Committees (PC)} You should try to serve on the PC  of the \emph{top} conferences in your field.
-This is a good way to get to know what is hot in your field, and to get to know people in your field.
+\subsection{Serving on Program Committees (PCs) and Editorial Boards} You should try to serve on the PCs  of the \emph{top} conferences in your field.
+This is a good way to know how to write good papers, and get to know the literature, what is hot in your field, and the people (who do what).
 It is also a good way to get your name out.
+Similarly to PCs, you should also try to serve on the editorial boards of top journals in your field.
 
-\textbf{Advice}: You can ask senior colleagues or even your PhD advisor/collaborators to recommend you to serve on a PC. You can also directly email the PC chairs and ask to serve on the PC.
+Serving on PC (or editor boards) essentially means reviewing papers.
+You will have to read papers and write reviews, and then argue for why a paper should be accepted or not.
+You will also have to discuss with other PC members (or editors) and come to a consensus.
+Typically the review process can last for several weeks, and there are also rebuttals and conditionally acceptances (which you should already be familiar with as authors).
 
-\subsection{Serving on NSF Panels} You should serve on NSF panels.
-It is a great way to learn about the grant process: how to distinguish between good and bad proposals and how people review proposals.
+\subsection{Serving on NSF Panels} As mentioned, faculty in CS typically get most of their funding from NSF.
+Thus, you should serve on NSF panels to learn about the grant process: how to distinguish between good and bad proposals and how people review proposals.
 It is also a good way learn what other people are doing, what is hot, and what is not. In short, it is a good way to learn how to write successful proposals.
@@
 
-\textbf{Advice}: Similarly to serving in PC, you can cold email program managers in your area and ask to be on a panel.
-Once you have served, you will likely be invited to serve again.
-Unfortunately, I do not have much experience with non-NSF panels or proposals.
-
 \subsection{Volunteer for more Leadership roles} Once you have served on a few PCs and NSF panels, you should aim for more leadership roles.
@@
 
+\subsection{Advice} To serve, you can directly ask people to recommend you.
+For example, for PC, ask senior colleagues or even your PhD advisor/collaborators to recommend you to the PC chairs.
+You can also directly email the PC chairs and ask to serve.
+Once you have served well, you will likely be invited to serve again.
+Similarly, for NSF Panels, you can cold email program managers in your area and ask to be on their panels.
+Unfortunately, I do not have much experience with non-NSF panels or proposals.
+For leadership roles, you can directly ask the general chair of a conference to see if you can take more leadership roles.
+
 \subsection{Personal Experience} I have served on many PCs and NSF panels.
-The initial services are mostly through ``cold'' emails.
+I mostly cold email people for these opportunities.
 For example, I directly email OOPSLA and PLDI PC chairs and asked, and they accepted my request (I did have papers at OOPSLA and PLDI when I asked).
@@
 I felt that I didn't get much out of serving on PC and didn't really like it much. However, I still do it because I feel I should give back to the community.
-On the otherhand, I throroughly enjoy serving on NSF panels.
+On the other hand, I thoroughly enjoy serving on NSF panels (though of course my proposals still got rejected frequently).
 While the work is more intense and takes more time, I found it more rewarding and enjoy the discussions.
@@
 
-\textbf{Advice}: You should try to be strategic about which committees you serve on that can help build your lab and tenure case.
+\subsection{Advice} You should try to be strategic about which committees you serve on that can help build your lab and tenure case.
 For example, serving in the PhD adcom is very common for new assistant professors, as it is a good way to recruit students for your lab.
-Serving on the faculty search committee is also a good way to recruit new faculty members that you like (and also let you know what the job market is like, just in canse you want to move).
+Serving on the faculty search committee is also a good way to recruit new faculty members that you like (and also let you know what the job market is like, what is hot and not).
 Serving on the colloquium committee is also a good way to invite speakers in your field that can be your \emph{tenure letter writers}.
 
-\subsubsection{Personal} Since coming to GMU, I have been much more active in serving the department.
+\subsection{Personal} Since coming to GMU, I have been very active in serving the department.
 I share the vision of the department and want to help achieve it.
 In addition to the typical services load, I volunteer to serve in additional committees and leadership roles.
-For example, I maintain the CSRankings DB for the department, I help organize the Open House to recruit admitted students, I serve in the Dept Chair renewal committee, etc.
-Even though I am not on the search committee, I speak to every candidate and help recruit strong ones.
-I agree to be director of the MS program, which is a lot of work and typically done by tenured faculty.
+For example, I maintain the CSRankings DB for the department, organize the Open House to recruit admitted students, participate in the Dept Chair renewal committee, etc.
+Even though I am not on the search committee, I meet with every faculty candidate (I talked to 22 people in Spring'23, which was my study break) and help recruit strong ones.
+I also agree to be director of the MS program, which is a lot of work and typically done by tenured faculty.
 
 Sometimes I feel overwhelmed by the services, but I also feel that I am making a difference in the department and that is rewarding.
-I am also building a good relationship with my colleagues and the department chair, and that is important for my tenure case.
-In addition, I learn a lot from being in various roles (e.g., I learned how class scheduling work, how the deparmtnet is run, how to recruit students, how to recruit faculty, how to run a colloquium, etc) and feel more confident in my leadership skills.
+I am also building a good relationship with my colleagues (and also head butt with some), and that is important for my tenure case.
+In addition, I learn a lot from being in various roles (e.g., how class scheduling work, how the department is run, how to recruit students and recruit faculty, how to run a colloquium, etc) and feel more confident in my leadership skills.
 
-In short, if you are enthusiastic about the department, want to make a difference, and \emph{feel comfortable at handling additional load}, you can volunteer for more services and leadership roles.
+In short, if you want to make a difference, and \emph{feel comfortable} at handling additional load, you can volunteer for more services. Otherwise, you can just ``stay low'',
+do the required services and focus on your research and teaching.
 
@@
 
+\chapter{Tenure Package and Application}
+
+\section{The Package}
+
+\subsection{Statements}
+
+\subsection{CV}
+
+\subsection{Letters}
+
+\subsection{Others}
+
 \bibliographystyle{abbrv}

### 2024-03-24 demystify.tex
updates

@@
 \section{How much do \emph{I} cost?}\label{sec:ra-cost}
+
 PhD students often ask why their salary is so low compared to ludicrous grants their advisers get.
@@
 \emph{The summary is that over your 5--6 years of your PhD, you cost about \$400K, and while your stipend is X, your adviser probably pays 2X for you}.
+But of course the nicest thing is that you do not have to pay for any of this!
+You get to gain the knowledge, do research, travel, (and don't even have to buy any research or computer equipments as mentioned in \autoref{sec:buying-equipment}) and also get paid!
 

### 2024-03-24 demystify.tex
updates

@@
 
-First, I want to emphasize that PhD students in Computer Science \emph{do not} need to worry about funding, especially at good R1
-universities in the US. If you are admitted, you will almost certainly
-receive \emph{full funding} to support your study, including tuition,
-health insurance, and stipend (monthly salary). Moreover, depending on the university,
-you may even receive additional benefits such as summer pay, laptops, (conference/workshop) traveling. \autoref{sec:funding} provides more details on funding.
-
-Second, applying to a good US university \emph{should not} be any harder than at schools in other countries.
+Students, especially international ones, often wonder if they should apply to a CS PhD program in the US.
+Specifically, they are worried about the \emph{difficulty} of getting admitted and the \emph{cost} of studying in the US.
+This section addresses these concerns.
+
+First, applying to a good US university \emph{should not} be any harder than at schools in other countries.
 In fact, it might even be more flexible since CS PhD in the US \emph{do not} require having an MS or a topic/proposal/adviser in advance.
@@
 
+Second, PhD students in CS \emph{do not} need to worry about funding, especially at good R1
+universities in the US. If you are admitted, you will almost certainly
+receive \emph{full funding} to support your study, including tuition,
+health insurance, and stipend (i.e., you get paid for your study). Moreover, depending on the university,
+you may even receive additional benefits such as summer pay, laptops, (conference/workshop) traveling. \autoref{sec:funding} provides more details on funding.
+
 \begin{commentbox}[Vu:]
   One of the reasons I create this document is that my colleagues at GMU are interested in recruiting Vietnamese students and are surprised when seeing very few applications in Vietnam (see \autoref{sec:ack}).
-Each year our CS program receives more than 350 PhD applications, most of which are international but only 3--4 are from Vietnam.
+Each year our CS PhD program receives about 400 PhD applications, most of which are international but only 4--5 are from Vietnam.
 In general the number of CS PhD applications from Vietnam to US universities is very low and more would be welcome.
@@
 
-\paragraph{Why CS depts do not waive application fee?}  Note that some universities do waive application fee (e.g., for domestic students (\autoref{sec:domestic-students})), however most do not.
+\paragraph{Why CS depts do not waive application fee?}  Note that some universities do waive, e.g., for domestic students (\autoref{sec:domestic-students}), however most do not.
 In any case, application fee is typically a requirement of the university.
@@
 This is because as someone who just finished your undergraduate study, it is unlikely that you know enough about research to do proposal or picking an adviser.
-This is also why US PhD programs are longer.
+You will have two initial years to explore and find an adviser and research topic.
 
@@
 I started writing this handbook in May 2023 and have been updating it since then (mostly around deadline time when I tend to procrastinate, i.e., \emph{productive procrastination}!).
+The document was initially for international students, but has been generalized to all students.
 I have put the source code of this document on \href{https://github.com/nguyenthanhvuh/phd-cs-us}{GitHub} so that others can contribute to it.
@@
 I have been in the PhD admission process at GMU and UNL for many years.
-Currently I serve as the program director of the MS program in Software Engineering at GMU (thus also have some experience with the MS admission process---which is quite different than PhD).
+Currently I serve as the program director of the MS program in Software Engineering at GMU (thus also have experience with the MS admission process---which is quite different than PhD).
 My personal and lab website is at \href{https://dynaroars.github.io}{dynaroars.github.io}.

### 2024-03-24 demystify.tex
updates

@@
 
-% \section{How much time did you spend looking at an application?}\label{sec:time-spent}
-% Short answer: less than 5 mins for quick rejection; 15+ mins for an acceptance.  On average, about 10 mins.
-
-% The GMU evaluation system allows us to download a single pdf that consists of a summary page (e.g., school info, GPA, testscores), then transcripts, standard test results, LoRs, CV, SOP, writing samples So I typically read the application in this order.
-Then I make notes on what I've seen, basically evidence that support my evaluation.
-
-% I first look at the summary for school, GPA, est scores (essentially looking for very low GPA or testscores that do not meet the minimum requirements.
-Then I skim on transcripts look for low/failing grades on CS and related courses.
-For these I make comments such as "unknown international school with good GPA, too many low grades on CS and related courses" or "good GPA from top school XYZ".
-I do not look at details on tests (the main scores are already given in the first summary page).
-
-% Then I look at LoRs.
+\section{How much time did you spend looking at my application?}\label{sec:time-spent}
+
+Short answer: this is highly dependent on the adcom member.
+For me, it is likely about 10 minutes on average (less than 5 mins for definite rejection; 15+ mins for an acceptance).
+Note that my earlier years, I spent more time, but now I have more experience and can quickly tell if an application is good or bad.
+
+The GMU evaluation system allows us to download a single pdf that consists of a summary page (e.g., school info, GPA, standard test scores), then transcripts, test results, LoRs, CV, SOP, writing samples (if any).
+So I typically scan the application in this order.
+I also make notes on what I've seen, basically evidence that support my evaluation.
+
+I first skim the summary for school, GPA, test scores to looking for red flags like low GPA or test scores that do not meet the minimum requirements.
+Then I move to transcripts to look for low/failing grades on CS and related courses.
+For these I make comments such as ``- unknown international school with good GPA, too many low grades on CS and related courses" or ``+ very good GPA from top school XYZ".
+I do not look at test results as the main scores are already given in the summary page.
+
+Then I look at LoRs.
 I read carefully good LoRs and quickly skim over bad or generic ones (very easy to tell, you don't even need much experiences to distinguish between a strong LoRs from the rest).
-Comment examples For LoRs include "very strong LoRs from well-known prof.
-XYZ talking about research experiences" or "generic LoRs, mostly about course work".
-I also note any unusual things in the LoRs.
-
-% Then the SOP, which I read carefully (of course if it is bad then it would a quick read).
-Then my notes for SOP include things such as "SOP mostly on course work, no clear motivation, generic" , or "very enjoyable SOP talking about XYZ, customized for GMU, mentioned profs.
+My comments for LoRs include ``+ very strong LoRs from well-known prof.
+XYZ talking about research experiences'' or ``- generic LoRs, mostly about course work''.
+I also note any unusual or unique details in the LoRs, e.g., ``+ LoR mentioned about student's significant contribution on well-known open-source project X''.
+Sometimes a LoR came directly from a GMU professor who mentioned they already know the student and intend to offer RA, which is a very strong signal for admission (and my comment would be ``Prof.
+X making RA offer'').
+
+I skim quickly over CV, looking for paper publications, research work experiences, and other relevant achievements such as winning prizes in well-known international programming/math competitions.
+My notes for CV include things such as ``+ has X papers at top venues'' or ``+ gold medal at X''.
+Sometimes I also note about unusual or unique experiences, e.g., ``+ has a patent on X''.
+
+Next is the SOP, which I read carefully (of course if it is bad then it would a quick read).
+My notes for SOP include things such as ``- SOP mostly on course work, no clear motivation, generic" , or ``+ very enjoyable SOP talking about XYZ, customized for GMU, mentioned profs.
 X,Y,Z".
-If the SOP mentioned about publications then I'll make notes such as "has X papers at top venues" or "has X papers but at unknown places".
-I will not go online search for the papers (unless they upload them as sample papers) .
-
-% I also skim over sampling papers, essentially looking over the abstract to see if reads well.  I would make notes such as "have a paper under submission to unknown venue, but reads well".
-
-%At GMU we use an evaluation system that allows us to download a single, large pdf that consists of everything submitted by the applicant, e.g., GPA, transcripts, test scores, CV, LoRs, SOP, sample writings, etc.
-The first couple of pages show the summary including school, major, GPA, test scores.
+If the SOP mentioned about publications then I would have ``+ has X papers at top venues" or ``+/- has X papers but at unknown places''.
+I will not go online search for the papers (unless they upload them as sample papers).
+I also skim over sampling papers, essentially looking over the abstract to see if reads well, and make notes such as ``+ have a paper under submission to unknown venue, but reads well".
+If the student mentioned about a project, I might also look at it (e.g., if it is on Github) and make notes such as ``+ has a project on Github with X stars''.
+
+At this point I have enough to make up my decision, which I enter to the evaluation system.
+We have several options for decision, but typically I just use \emph{reject} and \emph{admit with full funding}.
+For the latter, I also mention funding sources (e.g., RA from prof.
+X or TA).
+I can also recommend very strong students for the University Presidential fellowship.
+Finally, the system has text box where I put my comments to support my decision.
 

### 2024-03-24 demystify.tex
updates

@@
 
-Short answer: this is highly dependent on the adcom member.
-For me, it is likely about 10 minutes on average (less than 5 mins for definite rejection; 15+ mins for an acceptance).
-Note that my earlier years, I spent more time, but now I have more experience and can quickly tell if an application is good or bad.
+Around 15 minutes on average (less than 10 mins for definite rejection; 20+ mins for an acceptance).
+Note I spent more time in my earlier years, but now I have more experience and can quickly tell if an application is good or bad.
 
@@
 For these I make comments such as ``- unknown international school with good GPA, too many low grades on CS and related courses" or ``+ very good GPA from top school XYZ".
-I do not look at test results as the main scores are already given in the summary page.
+I do not look at test details as the summary already gives the main scores.
 
@@
 
-Next is the SOP, which I read carefully (of course if it is bad then it would a quick read).
+Next is the SOP, which I read carefully (but if it is bad then it would a quick read).
 My notes for SOP include things such as ``- SOP mostly on course work, no clear motivation, generic" , or ``+ very enjoyable SOP talking about XYZ, customized for GMU, mentioned profs.
@@
 I also skim over sampling papers, essentially looking over the abstract to see if reads well, and make notes such as ``+ have a paper under submission to unknown venue, but reads well".
-If the student mentioned about a project, I might also look at it (e.g., if it is on Github) and make notes such as ``+ has a project on Github with X stars''.
+If the student mentioned about a project, I might also look at it (e.g., if it is on Github) and make notes such as ``+ has project on Github with X stars''.
 
@@
 Finally, the system has text box where I put my comments to support my decision.
+
+After I submit my evaluation, the adcom chair will look all evaluations for that application and make a decision (almost always following our suggestion).
+If there are large discrepancies such as one vote reject but another full funding acceptance (such cases are rare, but do happen), the chair might ask us to coordinate and re-evaluate the application if necessary.
 

### 2024-03-24 demystify.tex
updates

@@
 
-The GMU evaluation system allows us to download a single pdf that consists of a summary page (e.g., school info, GPA, standard test scores), then transcripts, test results, LoRs, CV, SOP, writing samples (if any).
-So I typically scan the application in this order.
-I also make notes on what I've seen, basically evidence that support my evaluation.
-
-I first skim the summary for school, GPA, test scores to looking for red flags like low GPA or test scores that do not meet the minimum requirements.
-Then I move to transcripts to look for low/failing grades on CS and related courses.
+The evaluation system allows us to download a single pdf that consists of a summary page (e.g., schools, GPA, standard test scores), transcripts, test results, LoRs, CV, SOP, and writing samples (if any).
+I typically review the application in this order.
+I also make notes on what I've seen, using them to support my evaluation.
+
+I first skim the summary for school, GPA, test scores and looking for red flags like low GPA or test scores that do not meet the minimum requirements.
+Then I scan transcripts for low/failing grades on CS and related courses.
 For these I make comments such as ``- unknown international school with good GPA, too many low grades on CS and related courses" or ``+ very good GPA from top school XYZ".

### 2024-03-24 demystify.tex
updates

@@
 
-\section{How much time did you spend looking at my application?}\label{sec:time-spent}
-
-Around 15 minutes on average (less than 10 mins for definite rejection; 20+ mins for an acceptance).
-Note I spent more time in my earlier years, but now I have more experience and can quickly tell if an application is good or bad.
-
-The evaluation system allows us to download a single pdf that consists of a summary page (e.g., schools, GPA, standard test scores), transcripts, test results, LoRs, CV, SOP, and writing samples (if any).
+\section{How much time do you spend looking at an application?}\label{sec:time-spent}
+
+Around 15 minutes on average (less than 10 mins for definite rejection; 20+ mins for acceptance). In earlier years I spent more time, but now I can quickly tell if an application is good or bad.
+
+The evaluation system allows for downloading a single pdf that consists of a summary page (e.g., schools, GPA, standard test scores), transcripts, test results, LoRs, CV, SOP, and writing samples (if any).
 I typically review the application in this order.
@@
 At this point I have enough to make up my decision, which I enter to the evaluation system.
-We have several options for decision, but typically I just use \emph{reject} and \emph{admit with full funding}.
+Typically my decisions are either \emph{reject} or \emph{admit with full funding}.
 For the latter, I also mention funding sources (e.g., RA from prof.
 X or TA).
-I can also recommend very strong students for the University Presidential fellowship.
-Finally, the system has text box where I put my comments to support my decision.
-
-After I submit my evaluation, the adcom chair will look all evaluations for that application and make a decision (almost always following our suggestion).
-If there are large discrepancies such as one vote reject but another full funding acceptance (such cases are rare, but do happen), the chair might ask us to coordinate and re-evaluate the application if necessary.
+I also include my comments to support the decision.
+Finally, I also recommend very strong students for the University Presidential fellowship (\autoref{sec:fellowships}).
+
+The adcom chair will look all evaluations for each application and makes a decision (likely following the decision of the reviewers).
+If there are discrepancies such as one voted reject but another full funding acceptance (such cases are rare, but do happen), the chair might ask the reviewers to coordinate to reach a concensus.
 
@@
 \section{Member of an underrepresented or LGBTQ+ groups}\label{sec:urm}
-A common question is whether you should mention that you are a member of an underrepresented and minority (URM) group\footnote{URM in CS in the US include African Americans, Hispanic Americans, American Indians/Native Alaskans, Native Hawaiians/Pacific Islanders (excluding Asian Americans), and multiracial students identifying with at least one of the previously listed URM categories.} or LGBTQ+ group (, e.g., in your SOP).
+A common question is whether you should mention that you are a member of an underrepresented and minority (URM) group\footnote{URM in CS in the US include African Americans, Hispanic Americans, American Indians/Native Alaskans, Native Hawaiians/Pacific Islanders (excluding Asian Americans), and multiracial students identifying with at least one of the previously listed URM categories.} or LGBTQ+ group (e.g., in your SOP).
 This is a strong concern for international (and even domestic) students, who may have encountered discrimination in their countries or personal experiences.
@@
 Ultimately, this is a personal choice and you do what you feel comfortable.
-Seeking guidance from your close professors and collaborators is a good idea.
-However, it is important to recognize that diversity is strongly valued and celebrated within CS programs in the US, and that you should feel empowered to share your unique experiences and perspectives.
+If you do not feel comfortable disclosing this, you should not feel pressured to do so and you should also ask your LoR writers to not disclose this information.
+Note that you can still talk about your experiences and perspectives without mentioning your identity.
+If uncertain, you can seek additional guidance from your close professors and collaborators However, it is important to recognize that diversity is strongly valued and celebrated within CS programs in the US, and that you should feel empowered to share your unique experiences and perspectives.
 

### 2024-03-24 demystify.tex
updates

@@
 
-\section{How much time do you spend looking at an application?}\label{sec:time-spent}
-
-Around 15 minutes on average (less than 10 mins for definite rejection; 20+ mins for acceptance). In earlier years I spent more time, but now I can quickly tell if an application is good or bad.
-
-The evaluation system allows for downloading a single pdf that consists of a summary page (e.g., schools, GPA, standard test scores), transcripts, test results, LoRs, CV, SOP, and writing samples (if any).
+% \section{How much time do you spend looking at an application?}\label{sec:time-spent}
+
+% Around 15 minutes on average (less than 10 mins for definite rejection; 20+ mins for acceptance). In earlier years I spent more time, but now I can quickly tell if an application is good or bad.
+
+% The evaluation system allows for downloading a single pdf that consists of a summary page (e.g., schools, GPA, standard test scores), transcripts, test results, LoRs, CV, SOP, and writing samples (if any).
 I typically review the application in this order.
@@
 
-I first skim the summary for school, GPA, test scores and looking for red flags like low GPA or test scores that do not meet the minimum requirements.
+% I first skim the summary for school, GPA, test scores and looking for red flags like low GPA or test scores that do not meet the minimum requirements.
 Then I scan transcripts for low/failing grades on CS and related courses.
@@
 
-Then I look at LoRs.
+% Then I look at LoRs.
 I read carefully good LoRs and quickly skim over bad or generic ones (very easy to tell, you don't even need much experiences to distinguish between a strong LoRs from the rest).
@@
 
-I skim quickly over CV, looking for paper publications, research work experiences, and other relevant achievements such as winning prizes in well-known international programming/math competitions.
+% I skim quickly over CV, looking for paper publications, research work experiences, and other relevant achievements such as winning prizes in well-known international programming/math competitions.
 My notes for CV include things such as ``+ has X papers at top venues'' or ``+ gold medal at X''.
@@
 
-Next is the SOP, which I read carefully (but if it is bad then it would a quick read).
+% Next is the SOP, which I read carefully (but if it is bad then it would a quick read).
 My notes for SOP include things such as ``- SOP mostly on course work, no clear motivation, generic" , or ``+ very enjoyable SOP talking about XYZ, customized for GMU, mentioned profs.
@@
 
-At this point I have enough to make up my decision, which I enter to the evaluation system.
+% At this point I have enough to make up my decision, which I enter to the evaluation system.
 Typically my decisions are either \emph{reject} or \emph{admit with full funding}.
@@
 
-The adcom chair will look all evaluations for each application and makes a decision (likely following the decision of the reviewers).
+% The adcom chair will look all evaluations for each application and makes a decision (likely following the decision of the reviewers).
 If there are discrepancies such as one voted reject but another full funding acceptance (such cases are rare, but do happen), the chair might ask the reviewers to coordinate to reach a concensus.

### 2024-03-25 demystify.tex
updates

@@
 \section{Why did I get rejected?}\label{sec:why-rejected}
-Many possible reasons!  Other than not having a strong profile (e.g., research potentials, GPA, LoRs, SOP, poor interviews), here are some that you might not expect.
+Many possible reasons!  Other than not having a strong profile (e.g., research potentials, GPA, LoRs, SOP, poor interviews), here are some other reasons.
+
+\paragraph{You aim too high}  This is perhaps the most common one: you have applied to schools that are too high for your profile.
+While being ambitious is good, you should understand how PhD admission works (e.g., by reading this handbook!
+and realize things such as having a good GPA doesn't mean much) and be realistic!
+
+You should apply to a range of schools, including ``safety'' ones.   Before applying, you should talk to your professors and ask them to give you an honest opinion on where you should apply.
 
 \paragraph{Not a good fit}  You could have an excellent profile, but if you are interested in a research area that the program does not have, you might not be admitted.
-Similarly, if there is no faculty who is willing to advise you (e.g., they are on sabbatical, personal leave, do not have sufficient fundings, or already have too many students already), you will not be admitted.
+Similarly, if there is no faculty who is willing to advise you (e.g., they are on sabbatical, personal leave, do not have sufficient funding, or already have too many students already), you will not be admitted.
 This is actually good for you, as you don't want to be in a program where no one can advise you.
@@
 
-\paragraph{You aim too high}  You might have applied to schools that are too high for your profile.
-While being ambitious is good, you should also be realistic and apply to a range of schools, including ``safety'' ones.
-I have seen many students who have excellent profiles but only applied to top 10 or even top 20 schools and got rejected from all of them.
-Before applying, you should talk to your professors and ask them to give you an honest opinion on where you should apply
-
-\paragraph{Overqualified or Lack Interest}  This might be paradoxical, however if the adcom believes that you are likely to get admitted and go to a better program (i.e., they are just your ``safety'' schools), they might not be so enthusiastic to admit you and want to save the spot for someone else.
+\paragraph{Overqualified or Lack Interest}  This might be paradoxical and is the opposite of aiming too high.
+However if the adcom believes that you are likely to get admitted and go to a better program (i.e., they are just your ``safety'' schools), they might not be so enthusiastic to admit you and want to save the spot for someone else.
 Note though it does happen, for most good schools this is rare and you should not assume this is the case.
@@
 Another one is if you have a history of jumping from one program to another (e.g., you have been in multiple different programs in short time).
-If you think you have red flags, you should address them in your SOP.
+How would the reviewers know?
+They might have contacts in other programs and heard about you, your LoRs might mention it.
+
+If you think you have red flags, you should address them in your SOP or ask your letter writers to do so. 
 In general, these things are rare, but they do happen and cause concern to the adcom.

### 2024-03-25 demystify.tex
updates

--- 2024-03-24 demystify.tex
@@
 
-A related reason is that you did not show enough interest in the program.
+A related reason is that you did not show enough interests in the program.
 For example, your SOP did not mention any faculty or research area at the program, or you did not respond to professors interested in interviewing you (which might be considered unprofessional and burn bridges).
@@
 Another one is if you have a history of jumping from one program to another (e.g., you have been in multiple different programs in short time).
-If you think you have red flags, you should address them in your SOP.
+If you think you have such issues, you should try to address them in your SOP.
 In general, these things are rare, but they do happen and cause concern to the adcom.
@@
 
+\paragraph{Core CS areas} The core CS areas that you should have knowledge in are:
+\begin{itemize}
+  \item Programming Foundation (e.g., programming in C/C++ or Java, OOP)
+  \item Discrete Math (e.g., logic, set theory, proof techniques)
+  \item Data Structures and Algorithms (e.g., linked lists, trees, sorting, searching)
+  \item Computer OS or Systems (e.g., memory management, file systems, processes)
+\end{itemize}
+
 In short, you \emph{do not need} to formally taking CS courses, you just need to show that you have these essential knowledge, e.g., through the mentioned ways.
 Many universities are well-aware that incoming graduate students might not have all the technical background, so they often have a \emph{``bridge''} courses to help students catch up.
-For example, GMU has four bridge courses (Data Structures and Algorithms, Computer Systems, Discrete Math, and Programming Foundation) that incoming students can take to catch up on their CS knowledge.
+For example, GMU has four bridge courses corresponding to the four core areas above that incoming students can take to catch up on their CS knowledge.
 

### 2024-03-25 demystify.tex

--- 2024-03-25 demystify.tex
@@
 
-If you think you have red flags, you should address them in your SOP or ask your letter writers to do so. 
+If you think you have these issues, you should address them in your SOP or ask your letter writers to do so.
 In general, these things are rare, but they do happen and cause concern to the adcom.

### 2024-03-25 demystify.tex
various spelling error

--- 2024-03-25 demystify.tex
@@
         \begin{itemize}
-          \item Applications are evaluated by the \emph{PhD Admission} (adcom) committee and each application is typically reviewed by three faculty (\autoref{sec:evalapps}).
+          \item Applications are evaluated by the \emph{PhD Admission} (adcom) committee and each application is typically reviewed by three faculty members (\autoref{sec:evalapps}).
           \item Individually faculty \emph{cannot directly admit} a student---so do not email and ask if you have a chance.
@@
           \item \emph{SOP is very important}. Write it in such a way that makes you \emph{stand out} (\autoref{sec:research-statement} and \autoref{sec:improve-your-chance})
-          \item  \emph{Pubication can greatly help}.
+          \item  \emph{Publications can greatly help}.
 Publications are good venues are concrete evidence that you have successfully engaged in research.
@@
           \item \emph{Grades might be important}, depends on the reputation of \emph{your school} (\autoref{sec:your-school}).
-While good grades might not help much, bad ones will likley hurt your case.
+While good grades might not help much, bad ones likely will hurt your case.
           \item \emph{Standard tests are not important} (\autoref{sec:standard-tests}).
@@
 
-          \item \emph{Contacting a prof. for is recommended}, but do it \emph{properly} (\autoref{sec:contact}).
+          \item \emph{Contacting a prof. is recommended}, but do it \emph{properly} (\autoref{sec:contact}).
 
@@
           \item Increasing admission chance by being unique and standing out (\autoref{sec:improve-your-chance}).
-          \item Being in the underepresentative group in CS can help (\autoref{sec:urm}).
+          \item Being in the underepresentative groups in CS can help (\autoref{sec:urm}).
           \item You can successfully apply to CS PhD even if you have non-STEM background (\autoref{sec:non-stem}).
@@
   \tcblower
-  \textbf{Vu}: This is an interesting and useful detail that US faculty might not be aware of and students should mention about this in their statements.
+  \textbf{Vu}: This is an interesting and useful detail that US faculty might not be aware of and students should mention this in their statements.
 In general, someone with a PhD has been through the research process and therefore can better evaluate your research ability.
@@
 This shows that you're serious and have done homework on places you're applying to.
-Adcom will look for this part, so do not skip it.
+Adcom will look for this part (\autoref{sec:why-rejected}), so do not skip it.
 
@@
 
-\paragraph{Follow Up Emails} If you had an interview and have not heard back, you can send an email to ask about the status of your application Don't send this too soon though, perhaps wait at least 2 weeks.
-
-You can also send an email to ask about the status of your application if you have received offers from other schools and need to make a decision soon.
+\paragraph{Follow-Up Emails} If you had an interview and have not heard back, you can email to ask about the status of your application Don't send this too soon though, perhaps wait at least 2 weeks.
+
+You can also email to ask about the status of your application if you have received offers from other schools and need to make a decision soon.
 See \autoref{sec:accept-postpone-decline} for how to accept/postpone/decline offers.
@@
 Every PhD student has pubs, but only few  would have NSF Graduate or Microsoft fellowship.
-In fact, these are so prestigious that even if you didn't get it but make it to the final round or even ``honorable mentioning'', school will still mention you on their website and you still should put it on your CV.
+In fact, these are so prestigious that even if you didn't get it but make it to the final round or even ``honorable mentioning'', school will still mention you on their website, and you should put it on your CV.
 
@@
 
-While I find CSrankings useful, it is not perfect, and in general, rankings are superficial and you need to do more research to be informed and make better decision.
+While many find CSRankings useful, it is not perfect, and in general, rankings are superficial, and you need to do more research to be informed and make better decision.
 For example, if you get admissions to several places, you should consider attending Open Houses (\autoref{sec:accepted}) and contact profs.
-that you're interested in at those place and talk to them.
+that you're interested in at those places and talk to them.
 They would be more willing to chat with you now that you have been admitted.
@@
 For example, read their papers, know what they work on, and see if you are interested in their research.
-Then send them an email talking how/why their work would match yours.
+Then email them and talk about how/why their work would match yours.
 In contrast, if you write a generic email that can be sent to multiple professors (e.g., if you just change some names and keywords in the email or copy and paste paper titles), you will not get a response.
@@
 \begin{commentbox}[Vu:]
-  I would strongly advocate for a non-STEM student who shows that they have a strong drive for CS by studying core CS knowledge through various channels (e.g., self-study through online courses, projects, etc).
+  I would strongly advocate for a non-STEM student who shows that they have a strong drive for CS by studying core CS knowledge through various channels (e.g., self-study through online courses, projects, etc.).
 I have seen many students with non-CS background who are very successful in CS PhD.
@@
 In many cases you might take a summer or two off to do internship to get additional research opportunities.
-The PhDComics figure on top shows the ``ambition'' level of a PhD student over their years of study (they miss the 6--7th year where the ambition is \emph{``Just let me graduate''}).
+The PhDComics figure on top shows the ``ambition'' level of a PhD student over their years of study (they miss the 6--7th year when the ambition is \emph{``Just let me graduate''}).
 
@@
 of Chicago is a whooping \$70K) and a 58.9\% rate on \emph{indirect cost}, which is a typical rate charged for overhead or administrative costs (yes, after all, universities are businesses!).
-Finally, I assume the student take two conference trips per year, one domestic and one international (conf.
+Finally, I assume the students take two conference trips per year, one domestic and one international (conf.
 registration, airline tickets, taxi, meals, etc are all included).

### 2024-03-25 demystify.tex
various spelling errors

@@
           \item Increasing admission chance by being unique and standing out (\autoref{sec:improve-your-chance}).
-          \item Being in the underepresentative groups in CS can help (\autoref{sec:urm}).
+          \item Being in an underrepresented group can help can help (\autoref{sec:urm}).
           \item You can successfully apply to CS PhD even if you have non-STEM background (\autoref{sec:non-stem}).
@@
 \begin{commentbox}[Vu:]
-  If your grading system is not US standard or you are from a good school unknown outside of your country, you can ask your reference writers to explain about that in their letters.
+  If your grading system is not US standard or you are from a good school unknown outside your country, you can ask your reference writers to explain about that in their letters.
 For example, ``Bach Khoa'' are the top universities in Vietnam for STEM studies but few people outside Vietnam know about them.
@@
 Moreover, a top professor at your university might not be well-known to US faculty (see more details in \autoref{sec:your-school}).
-So save the trouble and  get letters from \emph{any} professors/supervisors who knows you well and can write a good letter about \emph{your} research ability.
+So save the trouble and  get letters from \emph{any} professors/supervisors who know you well and can write a good letter about \emph{your} research ability.
 It's better to have a good personalized
@@
 
-Reviewers might raise concerns (a red flag) about  a letter that is not waived, e.g., if you do not trust your letter writers, then you should find someone else to write for you.
+If you do not waive your right, then the letter writer might refuse to write for you or write a generic letter that does not help your case.
+Reviewers also might raise concerns about a letter that is not waived, e.g., if you do not trust your letter writers, then you should find someone else to write for you.
 In short, it's a standard practice and a way of keeping things straightforward and respectful in the whole recommendation game.
@@
 Note that most places only have deadlines for the applicant, but are very flexible with the letter writers (in many cases do not even give them any deadline).
-Also, many places do not begin the admission review process right after the deadline and work on application reviews in the next semester (mid January).
-
-\paragraph{VEF 2.0} For Vietnamese students, it's worth mentioning about the \href{https://vef2.org/}{VEF2.0 program}, which has helped many good students in gaining admission to top CS PhD programs.
+Also, many places do not begin the admission review process right after the deadline and work on application reviews in the next semester (mid-January).
+
+\paragraph{VEF 2.0} For Vietnamese students, it's worth mentioning the \href{https://vef2.org/}{VEF2.0 program}, which has helped many good students in gaining admission to top CS PhD programs.
 VEF2.0 follows an interesting model where US faculty members from leading institutions are invited to conduct rigorous interviews with VEF students and subsequently provide reference letters on their behalf.
@@
 \begin{commentbox}[Vu:]
-  At GMU CS, PhD students admitted TA has a 4 years of GTA guaranteed, and in also receive  stipend for the first summer.
+  At GMU CS, PhD students admitted TA has 4 years of GTA guaranteed, and in also receive stipend for the first summer.
 \end{commentbox}
@@
   PhD applicants at GMU are automatically eligible for a \emph{Presidential Fellowship}.
-It is at least as good as GTA but the most important thing is that as a fellowship it is truly free money (i.e., you are not depending on any prof.
+It is at least as good as GTA but the most important thing is that as a fellowship it is truly free money (i.e., you do not depend on any prof.
 or TA).
@@
 
-We will first discuss about universities in the US that offer PhD in CS. Then we will talk about how to rank and select them.
+We will first discuss universities in the US that offer PhD in CS. Then we will talk about how to rank and select them.
 
@@
 However, while many universities offer PhD in their CS programs, many do not.
-These universities might offer just Bachelor degrees (e.g., BS) and no graduate studies (i.e., no MS or Ph.D degrees), or they just offer MS programs (but no PhD).
-For example, Penn State in University Park has Ph.D.
-in CS,  but Penn State in Harrisburg only has BS and MS in CS, and Penn State in York only has BS in CS.
+These universities might offer just Bachelor degrees (e.g., BS) and no graduate studies (i.e., no MS or PhD degrees), or they just offer MS programs (but no PhD).
+For example, Penn State in University Park has PhD in CS,  but Penn State in Harrisburg only has BS and MS in CS, and Penn State in York only has BS in CS.
 On the other hand, multiple locations of the University of Texas, e.g., Austin, Dallas, Arlington, have PhDs in CS.
 
-Thus, if your goal is Ph.D.
-in CS, you have to aim only for schools offering such a degree.
+Thus, if your goal is PhD in CS, you have to aim only for schools offering such a degree.
 %This might sound obvious but many students (even those in the US) get confused due to the large number of universities.
@@
 \begin{commentbox}[Xiaokuan:]
-  Chinese students often only look at USNews rankings when selecting their Ph.D. universities (I did that, too, when I was applying for Ph.D. positions).
+  Chinese students often only look at US News rankings when selecting their PhD universities (I did that, too, when I was applying for PhD positions).
   Now that I am a professor, I find it to be the least promising way.
   %
-  The reason is that USNews does not provide a good metric for evaluating the quality of the Ph.D. program.
+  The reason is that US News does not provide a good metric for evaluating the quality of the PhD program.
   %
@@
   Also,
-  I think Ph.D. is not only about research;
+  I think PhD is not only about research;
   you need to also consider your daily life there, since you will (probably) stay for at least five years.
@@
 
-\paragraph{Tenure-line and teaching faculty} Tenure line faculty, consisting of tenured and tenured-track profs., focus on research, which includes publishing papers, obtaining grants, and mentoring Ph.D students.
+\paragraph{Tenure-line and teaching faculty} Tenure line faculty, consisting of tenured and tenured-track profs., focus on research, which includes publishing papers, obtaining grants, and mentoring PhD students.
 They often have low teaching load (e.g., 1 per semester).
@@
 \emph{Tenure-track} faculty are (more likely young) professors who are on the track to get tenure.
-At most universities, \textbf{only} tenure and tenured track faculty can serve as a formal adviser of Ph.D.
-students.
+At most universities, \textbf{only} tenure and tenured track faculty can serve as a formal adviser of PhD students.
 \autoref{sec:tenure-vs-tenure-track} talks more about choosing between tenure and tenured-track professor as your adviser.
@@
 They typically teach 3--4 classes per semester (which is quite a lot) and \emph{do not} have research responsibilities, e.g., they do not have to worry much about publishing papers or obtaining grants.
-They can mentor Ph.D.
-students but they typically cannot serve as a formal adviser of Ph.D.
-students, i.e., they \emph{cannot graduate} PhD students.
+They can mentor PhD students but they typically cannot serve as a formal adviser of PhD students, i.e., they \emph{cannot graduate} PhD students.
 
@@
 Research faculty (or research scientist) are typically non-tenure line faculty who focus on research.
-Due to their roles, adjunct, emeritus, and research faculty typically do not advise Ph.D.
-students.
+Due to their roles, adjunct, emeritus, and research faculty typically do not advise PhD students.
 
@@
 \begin{commentbox}[Xiaokuan]
-  Applying for Ph.D. and contacting a potential Ph.D. adviser is a classic \textbf{`why me, why you'} problem,
+  Applying for PhD and contacting a potential PhD adviser is a classic \textbf{`why me, why you'} problem,
   similar to looking for a job in a company.
@@
 You just need to observe and follow the conventions at your particular institution.
-Additionally, be aware that not all faculty members may hold a Ph.D., in which case using Prof.
+Additionally, be aware that not all faculty members may hold a PhD, in which case using Prof.
 Lastname is a suitable alternative.
@@
 I drafted the short essay on how to gain research experience remotely in order to improve their research profile and potentially get LOR from experts (hopefully faculty)}\\
-% In the realm of computer science, research experience plays a crucial role in securing admissions to top-tier Ph.D.
-programs.
+% In the realm of computer science, research experience plays a crucial role in securing admissions to top-tier PhD programs.
 However, students from underrepresented small colleges, both in the United States and internationally, often encounter limited research opportunities within their institutions.

### 2024-03-25 demystify.tex
various spelling errors

@@
 \chapter*{Preface}
-Having been involved in PhD admissions for many years, I've realized that many \textbf{international} (and domestic) students, especially those in  smaller countries or less well-known universities, lack a clear understanding of
+Having been involved in PhD admission committees for many years, I've realized that many \textbf{international} (and domestic) students, especially those in smaller countries or less well-known universities, lack a clear understanding of
 the Computer Science PhD admission process at US universities. This confusion not only
@@
 As a TA, you spend up to 20 hrs/week and help with classes (e.g., grading or teaching labs/recitation).
-During a semester, a TA might work with several courses and professors (not necessary their adviser).  TA funding is not typically available during the summer, which has few courses.
+During a semester, a TA might work with several courses and professors (not necessarily their adviser).  TA funding is not typically available during the summer, which has few courses.
 
@@
 \begin{commentbox}[Vu:]
-  At GMU CS, PhD students admitted TA has 4 years of GTA guaranteed, and in also receive stipend for the first summer.
+  At GMU CS, PhD students admitted TA have 4 years of GTA guaranteed, and in also receive a stipend for the first summer.
 \end{commentbox}
@@
 \subsection{Research Assistant (RA)}\label{sec:ra}
-RA is provided through a professor through their own funding so you can work on their project.
+RA is provided through a professor through their funding so you can work on their project.
 You do not need to teach as an RA, so you can focus on your research. Depending on the professor, RA may be available during the summer. \autoref{sec:ra-cost} gives more details on RA budget.
@@
 \begin{commentbox}[Vu:]
-  If you got recruited an offered an RA by a prof., it's very likely you will get admitted.
+  If you got recruited an offered an RA by a prof., you will likely get admitted.
 For example, if a prof., even if not in PhD adcom, wants to fund you, the PhD adcom will respect that decision and admit you (unless your application has many red flags).
@@
 Fellowship is another type of funding that students can get (e.g., from the university, industries, government).
-Fellowships are typically competitive and generous, and gives pretty much all benefits tuition/insurance that a TA/RA has.
+Fellowships are typically competitive and generous and gives pretty much all benefits tuition/insurance that a TA/RA has.
 They might even give higher stipend (including summer) and open doors for job opportunities such as internships.
@@
 
-In addition to general fellowships that are available to all students, there are also specic fellowships, e.g., for US citizen and permanent residents, for underrepresented minority (URM) students (\autoref{sec:urm}), and for students from specific countries.
+In addition to general fellowships that are available to all students, there are also specific fellowships, e.g., for US citizen and permanent residents, for underrepresented minority (URM) students (\autoref{sec:urm}), and for students from specific countries.
 In short, there are many fellowships available, so you should look for them and apply.
@@
 In general, fellowships, especially those that are open to everyone, are highly competitive and prestigious, and you will stand out if you get one.
-Every PhD student has pubs, but only few  would have NSF Graduate or Microsoft fellowship.
+Every PhD student has pubs, but only a few  would have NSF Graduate or Microsoft fellowship.
 In fact, these are so prestigious that even if you didn't get it but make it to the final round or even ``honorable mentioning'', school will still mention you on their website, and you should put it on your CV.
@@
 Choosing a school and an adviser is clearly among the most important things in your mind when you apply and especially when you get admitted.
-This is further complicated due to cultural differences and unfamiliarity of  international students to the US higher education system.
-This section aims to mitigate some confusions and help you make informed decision.
+This is further complicated due to cultural differences and the unfamiliarity of international students with the US higher education system.
+This section aims to mitigate some confusion and help you make informed decisions.
 
@@
 
-Most universities the US have a CS program.
+Most US universities have CS programs.
 However, while many universities offer PhD in their CS programs, many do not.
-These universities might offer just Bachelor degrees (e.g., BS) and no graduate studies (i.e., no MS or PhD degrees), or they just offer MS programs (but no PhD).
+These universities might offer just Bachelor's degrees (e.g., BS) and no graduate studies (i.e., no MS or PhD degrees), or they just offer MS programs (but no PhD).
 For example, Penn State in University Park has PhD in CS,  but Penn State in Harrisburg only has BS and MS in CS, and Penn State in York only has BS in CS.
-On the other hand, multiple locations of the University of Texas, e.g., Austin, Dallas, Arlington, have PhDs in CS.
+On the other hand, multiple locations of the University of Texas, e.g., Austin, Dallas, and Arlington, have PhDs in CS.
 
@@
 
-You can learn about CS programs and research expertise of faculty using resources such as \href{https://csrankings.org}{CSRankings.org}.
-You will be very surprised to learn that a school that you didn't know much about can have very strong research in your interested topic (and vice versa, a school you thought highly about might have no faculty working in the research field you're interested in).
-This is also a good way to learn about individual faculty (who works on what) and well-known CS conferences\footnote{In CS (and probably only in CS), conferences, not journals, are often the main venue to publish research finding (see \href{https://homes.cs.washington.edu/~mernst/advice/conferences-vs-journals.html}{why here}).}.
+You can learn about CS programs and research interests of faculty using resources such as \href{https://csrankings.org}{CSRankings.org}.
+You will be very surprised to learn that a school that you didn't know much about can have very strong research in your topic (and vice versa, a school you thought highly about might have no faculty working in the research field you're interested in).
+This is also a good way to learn about individual faculty (who works on what) and well-known CS conferences\footnote{In CS (and probably only in CS), conferences, not journals, are often the main venue to publish research findings (see \href{https://homes.cs.washington.edu/~mernst/advice/conferences-vs-journals.html}{why here}).}.
 %\autoref{sec:ranking} gives the top 50 CS programs in the US according to CSRankings.
@@
 
-While many find CSRankings useful, it is not perfect, and in general, rankings are superficial, and you need to do more research to be informed and make better decision.
-For example, if you get admissions to several places, you should consider attending Open Houses (\autoref{sec:accepted}) and contact profs.
+While many find CSRankings useful, it is not perfect, and in general, rankings are superficial, and you need to do more research to be informed and make better decisions.
+For example, if you get admission to several places, you should consider attending Open Houses (\autoref{sec:accepted}) and contact profs.
 that you're interested in at those places and talk to them.
 They would be more willing to chat with you now that you have been admitted.
-Ask them questions about their work, how they manage students, their expectations.
+Ask them questions about their work, how they manage students, and their expectations.
 You can even ask to contact their students.
@@
 Fortunately, while some non-US programs require finding an adviser and research topic before starting the PhD (\autoref{sec:non-us-differences}), CS PhD programs in the US will typically give you a couple of years to search for advisers and research topics.
-This is especially true if you're admitted with TA (\autoref{sec:ta}), which gives you time to explore and find adviser.
+This is especially true if you're admitted with TA (\autoref{sec:ta}), which gives you time to explore and find an adviser.
 
@@
 
-Here are some general advice to find professor.
+Here is some general advice to find a professor.
 Assuming you're not familiar with any particular one, then first search for profs.
 that share similar research interests.
-For example, in CSRankings, if you want to work with PL, you can search for those publish in PL conferences.
+For example, in CSRankings, if you want to work with PL, you can search for those published in PL conferences.
 If you want to work with SE \emph{and} AI, you can search for faculty who work in both SE and AI.
-Then you do research about that professor and contact them (\autoref{sec:contact}).
+Then you research that professor and contact them (\autoref{sec:contact}).
 
@@
   which often leads to disagreements or conflicts.
-  IMO, the adviser should be the one who \emph{guides}  students to do research while allowing students to pursue their own interest,
+  IMO, the adviser should be the one who \emph{guides}  students to do research while allowing students to pursue their own interests,
   instead of \emph{dictating} their research.
@@
 
-Another effective way is taking graduate level courses in the topics you are interested in.
+Another effective way is taking graduate-level courses in the topics you are interested in.
 Professors teaching \emph{special topics courses} and \emph{research seminars} might be recruiting students---giving you even a higher chance.
@@
 
-\paragraph{Tenure-line and teaching faculty} Tenure line faculty, consisting of tenured and tenured-track profs., focus on research, which includes publishing papers, obtaining grants, and mentoring PhD students.
+\paragraph{Tenure-line and teaching faculty} Tenure-line faculty, consisting of tenured and tenured-track profs., focus on research, which includes publishing papers, obtaining grants, and mentoring PhD students.
 They often have low teaching load (e.g., 1 per semester).
-\emph{Tenured} faculty are professors who have been promoted to have a permanent position (informally, very hard to fire them).
+\emph{Tenured} faculty are professors who have been promoted to a permanent position (informally, very hard to fire them).
 \emph{Tenure-track} faculty are (more likely young) professors who are on the track to get tenure.
@@
 You could ask your professors, LoR writers, collaborators, or those who have previously applied.
-For these kind of feedback, ask someone you have personal connection with.
+For this kind of feedback, ask someone you have personal connection with.
 
@@
 As discussed in \autoref{sec:research-opportunities}, research experience is a crucial component of a strong PhD application.
-It gives you opportunities to try out research, determine what research area you're interested in, publish papers (\autoref{sec:research-experiences}), connect with researchers and get strong LoRs (\autoref{sec:lor}).
+It gives you opportunities to try out research, determine what research area you're interested in, publish papers (\autoref{sec:research-experiences}), connect with researchers, and get strong LoRs (\autoref{sec:lor}).
 
  This section provides some guidance on how to gain research experience, e.g., as an undergraduate student.
-It also gives advice to students from small colleges where research opportunities are limited.
+It also give advice to students from small colleges where research opportunities are limited.
 
 \paragraph{Locally} Start looking for research opportunities at your own institution.
-If you did well or like a class, you can check with the professor of that class for research opportunities.
+If you did well or liked a class, you can check with the professor of that class for research opportunities.
 You can also go to the department directory and then professors' websites and see if they are looking for undergraduate researchers.
@@
 
-There might be also special programs for undergraduate research from the university, e.g., GMU has the OSCAR program, UNL has UCARE, NSF has REUs (Research Experience for Undergraduates) for US citizen and permanent residents.
+There might be also special programs for undergraduate research from the university, e.g., GMU has the OSCAR program, UNL has UCARE, NSF has REUs (Research Experience for Undergraduates) for US citizens and permanent residents.
 
@@
 \paragraph{Online Research Communities and Open Source Contributions} Places such as Github offer many opportunities for research, even though they might not appear so (e.g., they do not explicitly mention research).
-In many cases, professors' or research labs put their research projects on Github.
+In many cases, professors or research labs put their projects on Github.
 For example, my research group has a many open-source projects that undergrads can contribute to (see \url{https://github.com/dynaroars/}).
@@
 
-Students, especially international ones, often get confused on what to do after they get an offer from a professor, e.g., is it OK to postpone or accept/reject offers?, and most importantly, how to do so without offending anyone.
+Students, especially international ones, often get confused about what to do after they get an offer from a professor, e.g., is it OK to postpone or accept/reject offers?, and most importantly, how to do so without offending anyone.
 Many also wonder about accepting an offer and then later rejecting it.
@@
 
-Plagiarism and cheating (e.g., exams, assignments) is a BIG no-no in the US.
-If you're caught cheating, you will face serious consequence and likely be expelled from the university (e.g., after the second time at GMU).
+Plagiarism and cheating (e.g., exams, assignments) are a BIG no-no in the US.
+If you're caught cheating, you will face serious consequences and likely be expelled from the university (e.g., after the second time at GMU).
 This is quite different from many international countries where cheating is common and often tolerated.
@@
 
-Here is the typically steps: (i) a faculty suspecting a cheating case will report it to the Office of Academic Integrity (OAI) at the university---the report often has supporting evidence and suggested penalty (e.g., a failing grade);  (ii)  OAI will take over and investigate the case; and (iii) OAI will make the final decision.
-It is important to note that after receiving the report from your prof., OAI \emph{completely} takes over and makes decision.
+Here are the typical steps: (i) a faculty suspecting a cheating case will report it to the Office of Academic Integrity (OAI) at the university---the report often has supporting evidence and suggested penalty (e.g., a failing grade);  (ii)  OAI will take over and investigate the case; and (iii) OAI will make the final decision.
+It is important to note that after receiving the report from your prof., OAI \emph{completely} takes over and makes decisions.
 This means begging your professor will not help because they simply are no longer involved in the case.
@@
 \section{Illegal Software} Using illegal/cracked software is very common in many countries (and even in the US).
-However, \emph{do not} install or use them on university computers, even those given to you from your adviser.
+However, \emph{do not} install or use them on university computers, even those given to you by your adviser.
 It is unlikely that the university will track you down, but it is the \emph{software company} that will.
@@
 However, the reality is quite the opposite.
-While people can be straightforward and direct, they are also friendly and informal.
+While people can be straightforward, they are also friendly and informal.
 You can call your professor by their first name (\autoref{sec:address}), disagree with them and argue (and gain respect doing so), seek their help (even on personal matters), come to their houses for parties or gathering, and give them small thoughtful gifts that they put on their desks (\autoref{sec:gifts}).

### 2024-03-27 demystify.tex
updates

@@
 
-\section{Standard Tests (GREs, IELTS, etc), Work Experiences, Online Courses, and Certificates}\label{sec:standard-tests}
-
-\paragraph{GRE} Most good CS PhD programs in the US (e.g., R1 universities) \textbf{no longer require GREs}, so you don't need to take them.
-
-\textbf{Advice}: In general, don't waste time on the GRE.
-However, if you took it and have good scores then you can include and talk about them in your application (but don't expect them to make much difference).
+\section{Standard Tests (GRE, IELTS, etc), Work Experiences, Online Courses, and Certificates}\label{sec:standard-tests}
+
+\paragraph{GRE} While a few schools still require taking the GRE (e.g., UCF), most good CS PhD programs in the US \textbf{no longer} require it.
+
+\textbf{Advice}: If you haven't taken the GRE or have bad scores, then don't waste time (re)taking it.
+However, if you took it and have really good scores then you can include and talk about them in your application (but don't expect them to make much difference).
 But if your scores are bad, then you should not include them in your application, which can be a \red{red flag}.
@@
     \item GPA: $\ge 3.0$ in your undergrad (but we also consider the rank/prestige of your school)
-    \item GRE: not required, though it can help boost your profile
+    \item GRE: not required
     % \item but if you want to use it, then we expect a total (V+Q) of $\ge 311$ (with a $\ge 157$ Q) and A $\ge 3.0-3.5$.

### 2024-03-27 demystify.tex
updates

@@
 universities in the US. If you are admitted, you will almost certainly
-receive \emph{full funding} to support your study, including tuition,
+receive \emph{full funding} (\autoref{sec:funding}) to support your study, including tuition,
 health insurance, and stipend (i.e., you get paid for your study). Moreover, depending on the university,
-you may even receive additional benefits such as summer pay, laptops, (conference/workshop) traveling. \autoref{sec:funding} provides more details on funding.
+you may even receive additional benefits such as summer pay, laptops, (conference/workshop) traveling.
+
+\autoref{sec:summary} provides an overview of the application process, which you can explore in further details in the following chapters.
 
@@
 
-\paragraph{Why CS depts do not waive application fee?}  Note that some universities do waive, e.g., for domestic students (\autoref{sec:domestic-students}), however most do not.
+\paragraph{Why CS depts do not waive application fee?}  Some universities do waive, e.g., Rice for all PhD applicants and many universities for domestic students (\autoref{sec:domestic-students}), however most do not.
 In any case, application fee is typically a requirement of the university.
@@
 In my opinion, requiring applicants to pay the fee helps ensure their seriousness, as it filters out non-serious candidates.
-Also, if the application process were free for everyone, most good CS programs would receive an overwhelming number of applications to review.
+Also, most CS programs already receive many applications, and would be overwhelmed if the application process were free.
 
@@
 \chapter{Research Opportunities}\label{sec:research-opportunities}
+\begin{center}
+    \includegraphics[scale=0.5]{files/phd100404s.png}
+  \end{center}
 
@@
  This section provides some guidance on how to gain research experience, e.g., as an undergraduate student.
-It also give advice to students from small colleges where research opportunities are limited.
+It also gives advice to students from small colleges where research opportunities are limited.
 
@@
 
-There might be also special programs for undergraduate research from the university, e.g., GMU has the OSCAR program, UNL has UCARE, NSF has REUs (Research Experience for Undergraduates) for US citizens and permanent residents.
+There might be also special programs for undergraduate research from the university, e.g., GMU has the OSCAR program, UNL has UCARE, NSF has REUs (for US citizens and permanent residents).
 
@@
 
-\chapter{Cultural Differences and Other Issues}\label{sec:cultural}
-
-This section lists some cultural issues that international students might want to pay attention to. These issues apply in general and not just PhD admission.
+\chapter{Cultural Differences}\label{sec:cultural}
+
+This section lists some cultural issues that students, especially international ones, might want to pay attention to. These issues apply in general and not just PhD admission.
 
@@
 
-Students, especially international ones, often get confused about what to do after they get an offer from a professor, e.g., is it OK to postpone or accept/reject offers?, and most importantly, how to do so without offending anyone.
+Students often ask about what to do after they get an offer from a professor, e.g., is it OK to postpone or accept/reject offers?, and most importantly, how to do so without offending anyone.
 Many also wonder about accepting an offer and then later rejecting it.

### 2024-03-28 demystify.tex
updates

--- 2024-03-27 demystify.tex
@@
 
-\textbf{Advice}: You can explain about your school in your SOP (and also ask your LOR writer to do that).
+\textbf{Advice}: If you think your school is the top in your country but might not be well-known internationally, you should talk about it in your SOP (and also ask your LOR writer to do that).
 Of course, if you're interested in working with Vietnamese, consider  \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{CS programs in the US that have Vietnamese professors}.
@@
 \textbf{Advice}: As with school reputation, you can put a note about this in your statement and ask your LoR writers to talk about it.
-Note that while having good grades at unknown school might not help,
+Note that while having good grades might not help much,
 having bad grades (e.g., < 3.0) will be \red{red flag} (unless your LORs or

### 2024-03-28 demystify.tex
updates

@@
     {\fontsize{24}{24} \selectfont {{PhD Admission in the US}}} \\[0.2in]    
-    {\fontsize{18}{18} \selectfont {{A Handbook for International and Domestic Students}}} \\[0.5in]
+    {\fontsize{18}{18} \selectfont {{A Handbook for International Students}}} \\[0.5in]
 
@@
 \chapter*{Preface}
-Having been involved in PhD admission committees for many years, I've realized that many \textbf{international} (and domestic) students, especially those in smaller countries or less well-known universities, lack a clear understanding of
+Having been involved in PhD admission committees for many years, I've realized that many \textbf{international} students, especially those in smaller countries or less well-known universities, lack a clear understanding of
 the Computer Science PhD admission process at US universities. This confusion not only
@@
 So I want to share some details about the admission process and advice for those who are interested in applying for a \textbf{PhD in Computer Science in the US}.
+Originally, this document was intended for international students, but I have expanded it to include information that might be useful for \emph{domestic students} as well.
 While this handbook is primarily intended for students interested in CS, it might be relevant to students from various disciplines.
@@
         \end{itemize}
+
+    \item Appendices
+    \begin{itemize}
+      \item Specific advice for domestic students (\autoref{sec:domestic-students}).
+      \item How to find research opportunities (\autoref{sec:research-opportunities}).
+      \item Cultural differences between US and other countries (\autoref{sec:cultural}).
+    \end{itemize}
+
 \end{enumerate}
@@
 
-Students, especially international ones, often wonder if they should apply to a CS PhD program in the US.
+Students, especially those from smaller countries or less well-known universities, often wonder if they should apply to a CS PhD program in the US.
 Specifically, they are worried about the \emph{difficulty} of getting admitted and the \emph{cost} of studying in the US.
@@
 \appendix
+
+% \section{Additional Resources}
+% \begin{itemize}
+%   \item \bibentry{nlpphd}
+% \end{itemize}
+
+% \chapter{Additional Resources}
+
+% \chapter{Undergraduate Research Opportunities}
+
+% Prospective PhD students are often unaware of the importance of research experience in their application.
+This provides both concrete research experiences and potential LOR writers that can talk about your research ability.
+% % This is especially true for students from small colleges where research opportunities are limited.
+% This section aims to provide some guidance on how to gain research experience as an undergraduate student.
+%This section is written together Didier, a PhD student at GMU who did his undergraduate study in Rwanda.
+
+% \paragraph{Locally} There are many ways to find research opportunities right in your own institution.
+If you did well or like a class, you can check with the professor of that class for research opportunities.
+Sometimes professors would mention that they look for undergraduates on their websites, so do check those out!
+Even if they say they are looking for graduate students, you still should contact them and ask.
+You can also ask your academic adviser or other faculty members for suggestions.
+Finally, you can also ask your peers who are already doing research.
+You can use the method described in \autoref{sec:contact} to contact professors
+
+% In short, the most important thing is that you need to \emph{ask}!
+Don't be shy, don't be afraid of rejection, and don't be afraid of being ignored or that you're not good enough.
+Just ask!
+The worst thing that can happen is that you get a ``no'' or no response.
+But you will never know if you don't ask.
+And if you get a ``no'', you can always ask again later.
+There are students who got rejected the first time but got accepted the second time.
+So don't give up!
+
+% \paragraph{Online Research Communities and Open Source Contributions}:
+% Places such as Github offer many opportunities for research, even though they might not appear so (e.g., they do not explicitly mention research).
+Students can explore repositories related to their areas of interest, collaborate with other contributors, and actively engage in open source research projects.
+By contributing code, fixing bugs, implementing new features, or providing documentation, students can gain practical research experience and interact with experienced developers and researchers in the field.
+
+% \paragraph{Paid or Unpaid}  travel opportunities, publishing opportunities, etc
+
+% \section{Virtual Research Opportunities Beyond Physical Boundaries (written by Didier)}
+% \didi{In my experience, prospective PhD students (from Rwanda) are often unaware of the information provided in this document throughout their undergraduate.
+For this reason they often find themselves unprepared for PhD admission process by the time they finish undergraduate and have to spend a year or more to improve their application profiles.
+I think it would be beneficial to expand the document to also target students who are not yet ready for the application process to help them improve on all criteria considered during admission.
+I drafted the short essay on how to gain research experience remotely in order to improve their research profile and potentially get LOR from experts (hopefully faculty)}\\
+% In the realm of computer science, research experience plays a crucial role in securing admissions to top-tier PhD programs.
+However, students from underrepresented small colleges, both in the United States and internationally, often encounter limited research opportunities within their institutions.
+Thankfully, there exists an alternative pathway for these aspiring researchers to gain valuable research experience, even without being physically present at a university.
+
+% \textbf{Virtual Research Programs}:
+% Several universities, organizations, and research institutions offer virtual internships and research programs aimed at providing hands-on research experience.
+These programs often involve working remotely under the guidance of experienced mentors and collaborating with a team of fellow researchers.
+For instance, \href{https://docs.google.com/forms/d/1btIwt4HwjyKMOUk-EMy3rbkfWzFxv2lNrMm_zkd0pA4/viewform?edit_requested=true}{UIUC+ Summer Undergraduate Research in Software Engineering}  offers an unpaid remote internship for software engineering students all over the world to collaborate with mentors from University of Illinois at Urbana Champaign.
+% Virtual internships offer opportunities to contribute to ongoing research projects, conduct experiments, analyze data, and write research reports.
+Students can search for such programs through university websites, research institutes, or as of recently chatGPT (or Bard).
+
+% \textbf{Online Conferences and Workshops}:
+% Attending online conferences and workshops is another way to gain exposure to cutting-edge research and establish connections with experts in the field.
+Many conferences now provide virtual participation options, enabling students to access research talks, poster sessions, and panel discussions and sometimes access designated chat rooms or networking events where participants can engage with researchers, ask questions, and seek potential research collaborations.
+It is beneficial to also create profiles academic collaboration platforms such as ResearchGate.
+By creating profiles, joining relevant research groups, and participating in discussions, students can connect with established researchers, contribute to ongoing projects, and potentially collaborate on publications or research proposals as they provide access
+
+% In conclusion, for students lacking research opportunities at their small colleges, the virtual realm offers a wealth of possibilities to gain valuable research experience.
+Engaging with open source projects, joining online research communities, utilizing academic collaboration platforms, participating in virtual internships, and attending online conferences are all avenues to contribute to research, connect with experts, and enhance one's research profile.
+
+\chapter{Domestic Students}\label{sec:domestic-students}
+
+Most of what is written in this document is applicable to both domestic\footnote{For simplicity domestic means you did your undergrad (or MS) at a US university.} and international students.
+However, there are some differences and benefits that domestic students should be aware of.
+
+\paragraph{Standing out \autoref{sec:improve-your-chance}} There are \emph{few} domestic applications compared to international ones, i.e., domestic students are minority (\autoref{sec:urm}) in the CS PhD application pool.
+Many US universities want to increase the number of domestic students in their programs (and as mentioned above, there are specific fellowships for domestic students).
+That makes you stand out and can help your case.
+
+\paragraph{Fee Waiving} Some school might offer application fee waiver for domestic students.  You should check with the school you're applying to.
+
+\paragraph{School \autoref{sec:your-school}} We already know about your school, which is a plus (see \autoref{sec:your-school}).
+You are also more familiar with the US education system and culture (so hopefully you are aware of the cultural issues listed in \autoref{sec:cultural}).
+
+\paragraph{Standard Tests \autoref{sec:standard-tests}} You do not need to take TOEFL or IELTS because you already did your undergrad (or MS) at a US university.
+You might also be more comfortable communicating in English, e.g., contacting professors (\autoref{sec:contact}).
+
+\paragraph{Transcripts} You do not need to get your transcripts evaluated/translated (which can be a hassle and inaccurate).
+You can just send your official transcripts directly to the school you're applying to.
+
+\paragraph{Funding \autoref{sec:funding}} You have more opportunities for funding, e.g., through government scholarships for US citizens and permanent residents.
+You can also apply to specific programs \emph{before} you start your PhD, e.g., NSF Graduate Research Fellowship Program (GRFP) and Hertz Foundation Fellowship.
+These fellowships can be very competitive and will improve your admission chance if you get them.
+
+\paragraph{Research Experiences \autoref{sec:research-opportunities}} You might have many opportunities to do research as an undergraduates, e.g., through REUs and internships at your undergrad university.
+Highlight such experiences in your application.
+
+\paragraph{Open House \autoref{sec:accepted}} It is easier for you to attend open house events in person.  This can help you make a better decision on which school to attend.
+
+% \begin{domesticbox}[Domestic students:] If you're a domestic student, you have several advantages in your application.
+You also have more opportunities for funding, e.g., through government scholarships for US citizens and residents.
+Finally,
+% \end{domesticbox}
+
+\chapter{Research Opportunities}\label{sec:research-opportunities}
+\begin{center}
+    \includegraphics[scale=0.5]{files/phd100404s.png}
+  \end{center}
+
+As discussed in \autoref{sec:research-opportunities}, research experience is a crucial component of a strong PhD application.
+It gives you opportunities to try out research, determine what research area you're interested in, publish papers (\autoref{sec:research-experiences}), connect with researchers, and get strong LoRs (\autoref{sec:lor}).
+
+ This section provides some guidance on how to gain research experience, e.g., as an undergraduate student.
+It also gives advice to students from small colleges where research opportunities are limited.
+
+\paragraph{Locally} Start looking for research opportunities at your own institution.
+If you did well or liked a class, you can check with the professor of that class for research opportunities.
+You can also go to the department directory and then professors' websites and see if they are looking for undergraduate researchers.
+Even if they say they are looking for graduate students, you still should contact them and ask.
+
+There might be also special programs for undergraduate research from the university, e.g., GMU has the OSCAR program, UNL has UCARE, NSF has REUs (for US citizens and permanent residents).
+
+You can also take honor thesis or independent study courses with a professor.
+This is a good way to get research experience and also get a LoR from the professor.
+You can also ask your academic adviser or other faculty members for suggestions.
+Finally, you can also ask your peers who are already doing research.
+Use the method described in \autoref{sec:contact} to contact professors
+
+\begin{commentbox}[Vu:]
+    I enjoy working with undergrads (e.g., have mentored over 10 undergrads) and always actively recruit.
+    I often get undergraduate researchers through my classes, e.g., asking students who did well in my class if they are interested in research.
+Occasionally I was introduced to students by other students or faculty.
+
+    While most undergrads are understandably not super productive in research, some are and I have published multiple papers with them.
+I also have written strong LoRs for them and have helped them get into PhD programs, including top ones.
+
+    \tcblower
+    At GMU we pay undergrads about \$15/hr and about 10 hrs/week. This is flexible and can be increased as needed.
+\end{commentbox}
+
+\paragraph{Online Research Communities and Open Source Contributions} Places such as Github offer many opportunities for research, even though they might not appear so (e.g., they do not explicitly mention research).
+In many cases, professors or research labs put their projects on Github.
+For example, my research group has a many open-source projects that undergrads can contribute to (see \url{https://github.com/dynaroars/}).
+
+By contributing code, fixing bugs, implementing new features, or providing documentation, students can gain practical research experience and interact with experienced developers and researchers in the field.
+Not only you gain research experience, you might be able to get a LoR from the project maintainer, and you should write about this experience in your SOP (\autoref{sec:research-statement}.)
+
+\paragraph{Virtual Research Opportunities} This is less common but several places offer virtual internships and research programs aimed at providing hands-on research experience.
+These programs often involve working remotely under the guidance of experienced mentors and collaborating with a team of fellow researchers.
+For instance, \href{https://docs.google.com/forms/d/1btIwt4HwjyKMOUk-EMy3rbkfWzFxv2lNrMm_zkd0pA4/viewform?edit_requested=true}{UIUC+ Summer Undergraduate Research in Software Engineering}  offers an unpaid remote internship for software engineering students all over the world to collaborate with mentors from University of Illinois at Urbana Champaign.
+
+% \paragraph{Online Conferences and Workshops}: Attending online conferences and workshops is another way to gain exposure to cutting-edge research and establish connections with experts in the field.
+Many conferences now provide virtual participation options, enabling students to access research talks, poster sessions, and panel discussions and sometimes access designated chat rooms or networking events where participants can engage with researchers, ask questions, and seek potential research collaborations.
+By creating profiles, joining relevant research groups, and participating in discussions, students can connect with established researchers, contribute to ongoing projects, and potentially collaborate on publications or research proposals as they provide access
+
+\chapter{Cultural Differences}\label{sec:cultural}
+
+This section lists some cultural issues that students, especially international ones, might want to pay attention to. These issues apply in general and not just PhD admission.
+
+% \paragraph{Diversity} US universities prioritize diversity and inclusion.
+Students need to respect and appreciate varied opinions, backgrounds, and cultures.
+Unlike some countries where certain voices are marginalized, in the US, all perspectives are valued equally (especially at universities).
+Racism or discrimination will have serious consequences, including academic and disciplinary actions.
+
+\section{Checking Status, Accepting, Postponing, and Decline Offers}\label{sec:accept-postpone-decline}
+
+Students often ask about what to do after they get an interview or an offer from a professor, e.g., if they can followup to find out about their status, or is it OK to postpone or accept/reject offers?, and most importantly, how to do so without offending anyone.
+
+\paragraph{Checking your application status} If you have interviewed and not heard back from a professor after a few weeks or especially around the time when universities send out their admission decisions (around late Feb-- mid Mar), you can email to check.
+You can follow up the interview invitation and say: \emph{``Thanks for chatting with me.
+I am very excited about the opportunity to work with you.
+Could you please let me know if you have made a decision or if you need more information from me?''}.
+If you have new updates, e.g., new publications or new fellowship awards, or even new offers from other professors or schools, you can also mention that.
+
+Profs.
+are often very busy, especially during admission time when they have to review many applications, interview many students, and make decisions.
+They might not have time to respond to every email.
+If you do not hear back after a week, you can send another email to check again.
+If you still do not hear back, you can assume that you are not selected.
+
+\paragraph{Accepting an offer} If you decide to accept an offer, you can say: \emph{``Thank you for the offer.
+I would like to accept it and look forward to working with you.
+Could you please send me more details about the offer and what to do next?''}.
+The prof.
+will likely send you more details about the offer and what to do next.
+If you decide to accept an offer, do so quickly.
+
+\paragraph{Postponing an offer} If you need more time to decide, you can ask for more time: \emph{``Thank you for the offer.
+I am very excited about it.
+However, I am still waiting for other offers and need more time to decide.
+Would it be possible to postpone the decision for a few weeks?''}.
+This is perfectly fine and professors will understand and might even appreciate your honesty.
+They will likely give you a few weeks to decide.
+If you need more time, you can ask for more time.
+But do not ask for too much time, e.g., more than a month.
+You also should not postpone the offer multiple times, which will annoy people.
+
+\paragraph{Declining an offer} If you decide to decline or reject an offer, you can say: \emph{``Thank you for the offer.
+However, I have decided to accept another offer.
+I appreciate your time and consideration.
+I hope we can work together in the future.''}  Professors will understand and wish you luck.
+If you decide to reject an offer, do so quickly.
+
+\paragraph{Accepting an offer and later later rejecting it}
+I have seen \emph{many} international students asking about this, e.g., accepting an offer from a prof.
+and then later rejecting it when they get admitted to a better/dream school, which have them on its wait list.
+There are numerous advice, e.g., on Internet forums, that this is OK to do so (most of these advice are from other international students).
+The reasons typically are \emph{``you haven't established much relationship with the prof.
+yet, so it is OK to reject it; no hard feeling, you probably won't see them again anyway''} or that \emph{``it's your life and you should do what's best for you''}.
+
+From my experience, this is \emph{not} a good idea and can be considered \emph{rude and unprofessional}.
+But ultimately, it is a trade-off that you need to decide,  e.g., between your own benefit and the cost to others.
+If you reject an offer, the prof.
+will likely stop looking for other students and might even reject other students; so by rejecting the offer later, you causing a great deal of inconvenience to the prof and also taking away the opportunity from other students.
+This will likely would burn bridges with that prof and they will remember what you did to them.
+Moreover, given that research subfields are typically small, you might run into that professor again in the future, and they will remember what you did to them.
+
+\section{Academic Integrity (Cheating and Plagiarism)}
+
+Plagiarism and cheating (e.g., exams, assignments) are a BIG no-no in the US.
+If you're caught cheating, you will face serious consequences and likely be expelled from the university (e.g., after the second time at GMU).
+This is quite different from many international countries where cheating is common and often tolerated.
+Faculty is extremely good at detecting cheating (we have been dealing with these situations so many times over so many years), and \emph{will} report cheating cases.
+In short, whatever you do, don't cheat---not worth it.
+
+Here are the typical steps: (i) a faculty suspecting a cheating case will report it to the Office of Academic Integrity (OAI) at the university---the report often has supporting evidence and suggested penalty (e.g., a failing grade);  (ii)  OAI will take over and investigate the case; and (iii) OAI will make the final decision.
+It is important to note that after receiving the report from your prof., OAI \emph{completely} takes over and makes decisions.
+This means begging your professor will not help because they simply are no longer involved in the case.
+
+\section{Illegal Software} Using illegal/cracked software is very common in many countries (and even in the US).
+However, \emph{do not} install or use them on university computers, even those given to you by your adviser.
+It is unlikely that the university will track you down, but it is the \emph{software company} that will.
+They have very sophisticated tools to detect illegal software and will sue your university/department.
+Imagine your department or adviser being sued for a large sum of money, and it is \emph{you} who caused it.
+If you need to purchase software,  ask your adviser or the department (\autoref{sec:buying-equipment}).
+
+\section{Gifts}\label{sec:gifts} In many countries, it is customary to give professors gifts, often during holidays.
+These gifts can be costly and profs.
+sometimes expected them.
+In the US, it is uncommon and perfectly OK to not to give gifts.
+\emph{However}, if you'd like to offer small souvenir-like tokens, it's a thoughtful gesture that's appreciated.
+Some professors proudly display their gifts, which can come from students and colleagues (e.g., when they travel to their home countries or conferences).
+In summary, small gifts are fine, but avoid anything that might make your professors uncomfortable.
+
+\section{Relationships with your Profs.}
+There's a misconception that in the US it's all business, with professors as bosses who pay students for their work, and you doing something nice implying you expect something in return.
+However, the reality is quite the opposite.
+While people can be straightforward, they are also friendly and informal.
+You can call your professor by their first name (\autoref{sec:address}), disagree with them and argue (and gain respect doing so), seek their help (even on personal matters), come to their houses for parties or gathering, and give them small thoughtful gifts that they put on their desks (\autoref{sec:gifts}).
+Many students and professors maintain lifelong relationships, staying in touch through cards, emails, and calls, even after their academic journey ends.
+%These connections are genuine and meaningful.
+
+% \begin{commentbox}
+%   [DK:] Here are a few items you may consider addressing: putting international students in touch with other students from their countries, assuring them that they would be picked up from airports upon their arrival and that their initial stay will be taken care of.
+Most universities have other resources for these, but it is worth mentioning that they would be taken care of upon arrival and can get help during the transition phase.
+Learning to cook was a big deal when I arrived over 50 years ago-August 1973.
+But things have changed as one can find many ethnic food places, a big contrast from 1973, when there were two Indian restaurants in Greater Boston area.
+% \end{commentbox}
 
@@
 
-% \section{Additional Resources}
-% \begin{itemize}
-%   \item \bibentry{nlpphd}
-% \end{itemize}
-
-% \chapter{Additional Resources}
-
-% \chapter{Undergraduate Research Opportunities}
-
-% Prospective PhD students are often unaware of the importance of research experience in their application.
-This provides both concrete research experiences and potential LOR writers that can talk about your research ability.
-% % This is especially true for students from small colleges where research opportunities are limited.
-% This section aims to provide some guidance on how to gain research experience as an undergraduate student.
-%This section is written together Didier, a PhD student at GMU who did his undergraduate study in Rwanda.
-
-% \paragraph{Locally} There are many ways to find research opportunities right in your own institution.
-If you did well or like a class, you can check with the professor of that class for research opportunities.
-Sometimes professors would mention that they look for undergraduates on their websites, so do check those out!
-Even if they say they are looking for graduate students, you still should contact them and ask.
-You can also ask your academic adviser or other faculty members for suggestions.
-Finally, you can also ask your peers who are already doing research.
-You can use the method described in \autoref{sec:contact} to contact professors
-
-% In short, the most important thing is that you need to \emph{ask}!
-Don't be shy, don't be afraid of rejection, and don't be afraid of being ignored or that you're not good enough.
-Just ask!
-The worst thing that can happen is that you get a ``no'' or no response.
-But you will never know if you don't ask.
-And if you get a ``no'', you can always ask again later.
-There are students who got rejected the first time but got accepted the second time.
-So don't give up!
-
-% \paragraph{Online Research Communities and Open Source Contributions}:
-% Places such as Github offer many opportunities for research, even though they might not appear so (e.g., they do not explicitly mention research).
-Students can explore repositories related to their areas of interest, collaborate with other contributors, and actively engage in open source research projects.
-By contributing code, fixing bugs, implementing new features, or providing documentation, students can gain practical research experience and interact with experienced developers and researchers in the field.
-
-% \paragraph{Paid or Unpaid}  travel opportunities, publishing opportunities, etc
-
-% \section{Virtual Research Opportunities Beyond Physical Boundaries (written by Didier)}
-% \didi{In my experience, prospective PhD students (from Rwanda) are often unaware of the information provided in this document throughout their undergraduate.
-For this reason they often find themselves unprepared for PhD admission process by the time they finish undergraduate and have to spend a year or more to improve their application profiles.
-I think it would be beneficial to expand the document to also target students who are not yet ready for the application process to help them improve on all criteria considered during admission.
-I drafted the short essay on how to gain research experience remotely in order to improve their research profile and potentially get LOR from experts (hopefully faculty)}\\
-% In the realm of computer science, research experience plays a crucial role in securing admissions to top-tier PhD programs.
-However, students from underrepresented small colleges, both in the United States and internationally, often encounter limited research opportunities within their institutions.
-Thankfully, there exists an alternative pathway for these aspiring researchers to gain valuable research experience, even without being physically present at a university.
-
-% \textbf{Virtual Research Programs}:
-% Several universities, organizations, and research institutions offer virtual internships and research programs aimed at providing hands-on research experience.
-These programs often involve working remotely under the guidance of experienced mentors and collaborating with a team of fellow researchers.
-For instance, \href{https://docs.google.com/forms/d/1btIwt4HwjyKMOUk-EMy3rbkfWzFxv2lNrMm_zkd0pA4/viewform?edit_requested=true}{UIUC+ Summer Undergraduate Research in Software Engineering}  offers an unpaid remote internship for software engineering students all over the world to collaborate with mentors from University of Illinois at Urbana Champaign.
-% Virtual internships offer opportunities to contribute to ongoing research projects, conduct experiments, analyze data, and write research reports.
-Students can search for such programs through university websites, research institutes, or as of recently chatGPT (or Bard).
-
-% \textbf{Online Conferences and Workshops}:
-% Attending online conferences and workshops is another way to gain exposure to cutting-edge research and establish connections with experts in the field.
-Many conferences now provide virtual participation options, enabling students to access research talks, poster sessions, and panel discussions and sometimes access designated chat rooms or networking events where participants can engage with researchers, ask questions, and seek potential research collaborations.
-It is beneficial to also create profiles academic collaboration platforms such as ResearchGate.
-By creating profiles, joining relevant research groups, and participating in discussions, students can connect with established researchers, contribute to ongoing projects, and potentially collaborate on publications or research proposals as they provide access
-
-% In conclusion, for students lacking research opportunities at their small colleges, the virtual realm offers a wealth of possibilities to gain valuable research experience.
-Engaging with open source projects, joining online research communities, utilizing academic collaboration platforms, participating in virtual internships, and attending online conferences are all avenues to contribute to research, connect with experts, and enhance one's research profile.
-
-\chapter{Domestic Students}\label{sec:domestic-students}
-
-Most of what is written in this document is applicable to both domestic\footnote{For simplicity domestic means you did your undergrad (or MS) at a US university.} and international students.
-However, there are some differences and benefits that domestic students should be aware of.
-
-\paragraph{Standing out \autoref{sec:improve-your-chance}} There are \emph{few} domestic applications compared to international ones, i.e., domestic students are minority (\autoref{sec:urm}) in the CS PhD application pool.
-Many US universities want to increase the number of domestic students in their programs (and as mentioned above, there are specific fellowships for domestic students).
-That makes you stand out and can help your case.
-
-\paragraph{Fee Waiving} Some school might offer application fee waiver for domestic students.  You should check with the school you're applying to.
-
-\paragraph{School \autoref{sec:your-school}} We already know about your school, which is a plus (see \autoref{sec:your-school}).
-You are also more familiar with the US education system and culture (so hopefully you are aware of the cultural issues listed in \autoref{sec:cultural}).
-
-\paragraph{Standard Tests \autoref{sec:standard-tests}} You do not need to take TOEFL or IELTS because you already did your undergrad (or MS) at a US university.
-You might also be more comfortable communicating in English, e.g., contacting professors (\autoref{sec:contact}).
-
-\paragraph{Transcripts} You do not need to get your transcripts evaluated/translated (which can be a hassle and inaccurate).
-You can just send your official transcripts directly to the school you're applying to.
-
-\paragraph{Funding \autoref{sec:funding}} You have more opportunities for funding, e.g., through government scholarships for US citizens and permanent residents.
-You can also apply to specific programs \emph{before} you start your PhD, e.g., NSF Graduate Research Fellowship Program (GRFP) and Hertz Foundation Fellowship.
-These fellowships can be very competitive and will improve your admission chance if you get them.
-
-\paragraph{Research Experiences \autoref{sec:research-opportunities}} You might have many opportunities to do research as an undergraduates, e.g., through REUs and internships at your undergrad university.
-Highlight such experiences in your application.
-
-\paragraph{Open House \autoref{sec:accepted}} It is easier for you to attend open house events in person.  This can help you make a better decision on which school to attend.
-
-% \begin{domesticbox}[Domestic students:] If you're a domestic student, you have several advantages in your application.
-You also have more opportunities for funding, e.g., through government scholarships for US citizens and residents.
-Finally,
-% \end{domesticbox}
-
-\chapter{Research Opportunities}\label{sec:research-opportunities}
-\begin{center}
-    \includegraphics[scale=0.5]{files/phd100404s.png}
-  \end{center}
-
-As discussed in \autoref{sec:research-opportunities}, research experience is a crucial component of a strong PhD application.
-It gives you opportunities to try out research, determine what research area you're interested in, publish papers (\autoref{sec:research-experiences}), connect with researchers, and get strong LoRs (\autoref{sec:lor}).
-
- This section provides some guidance on how to gain research experience, e.g., as an undergraduate student.
-It also gives advice to students from small colleges where research opportunities are limited.
-
-\paragraph{Locally} Start looking for research opportunities at your own institution.
-If you did well or liked a class, you can check with the professor of that class for research opportunities.
-You can also go to the department directory and then professors' websites and see if they are looking for undergraduate researchers.
-Even if they say they are looking for graduate students, you still should contact them and ask.
-
-There might be also special programs for undergraduate research from the university, e.g., GMU has the OSCAR program, UNL has UCARE, NSF has REUs (for US citizens and permanent residents).
-
-You can also take honor thesis or independent study courses with a professor.
-This is a good way to get research experience and also get a LoR from the professor.
-You can also ask your academic adviser or other faculty members for suggestions.
-Finally, you can also ask your peers who are already doing research.
-Use the method described in \autoref{sec:contact} to contact professors
-
-\begin{commentbox}[Vu:]
-    I enjoy working with undergrads (e.g., have mentored over 10 undergrads) and always actively recruit.
-    I often get undergraduate researchers through my classes, e.g., asking students who did well in my class if they are interested in research.
-Occasionally I was introduced to students by other students or faculty.
-
-    While most undergrads are understandably not super productive in research, some are and I have published multiple papers with them.
-I also have written strong LoRs for them and have helped them get into PhD programs, including top ones.
-
-    \tcblower
-    At GMU we pay undergrads about \$15/hr and about 10 hrs/week. This is flexible and can be increased as needed.
-\end{commentbox}
-
-\paragraph{Online Research Communities and Open Source Contributions} Places such as Github offer many opportunities for research, even though they might not appear so (e.g., they do not explicitly mention research).
-In many cases, professors or research labs put their projects on Github.
-For example, my research group has a many open-source projects that undergrads can contribute to (see \url{https://github.com/dynaroars/}).
-
-By contributing code, fixing bugs, implementing new features, or providing documentation, students can gain practical research experience and interact with experienced developers and researchers in the field.
-Not only you gain research experience, you might be able to get a LoR from the project maintainer, and you should write about this experience in your SOP (\autoref{sec:research-statement}.)
-
-\paragraph{Virtual Research Opportunities} This is less common but several places offer virtual internships and research programs aimed at providing hands-on research experience.
-These programs often involve working remotely under the guidance of experienced mentors and collaborating with a team of fellow researchers.
-For instance, \href{https://docs.google.com/forms/d/1btIwt4HwjyKMOUk-EMy3rbkfWzFxv2lNrMm_zkd0pA4/viewform?edit_requested=true}{UIUC+ Summer Undergraduate Research in Software Engineering}  offers an unpaid remote internship for software engineering students all over the world to collaborate with mentors from University of Illinois at Urbana Champaign.
-
-% \paragraph{Online Conferences and Workshops}: Attending online conferences and workshops is another way to gain exposure to cutting-edge research and establish connections with experts in the field.
-Many conferences now provide virtual participation options, enabling students to access research talks, poster sessions, and panel discussions and sometimes access designated chat rooms or networking events where participants can engage with researchers, ask questions, and seek potential research collaborations.
-By creating profiles, joining relevant research groups, and participating in discussions, students can connect with established researchers, contribute to ongoing projects, and potentially collaborate on publications or research proposals as they provide access
-
-\chapter{Cultural Differences}\label{sec:cultural}
-
-This section lists some cultural issues that students, especially international ones, might want to pay attention to. These issues apply in general and not just PhD admission.
-
-% \paragraph{Diversity} US universities prioritize diversity and inclusion.
-Students need to respect and appreciate varied opinions, backgrounds, and cultures.
-Unlike some countries where certain voices are marginalized, in the US, all perspectives are valued equally (especially at universities).
-Racism or discrimination will have serious consequences, including academic and disciplinary actions.
-
-\section{Accepting, Postponing, and Decline Offers}\label{sec:accept-postpone-decline}
-
-Students often ask about what to do after they get an offer from a professor, e.g., is it OK to postpone or accept/reject offers?, and most importantly, how to do so without offending anyone.
-Many also wonder about accepting an offer and then later rejecting it.
-
-In short, if you are not serious about an offer for any reason (e.g., still waiting for other offers), \emph{do not accept it}.
-If you are not sure, ask for more time to decide.
-But do not accept an offer and then later reject it.
-This can be considered \emph{rude and unprofessional}, and will likely burn bridges with that prof.
-
-\paragraph{Accepting an offer} If you decide to accept an offer, you can say: \emph{``Thank you for the offer.
-I would like to accept it and look forward to working with you.
-Could you please send me more details about the offer and what to do next?''}.
-The prof.
-will likely send you more details about the offer and what to do next.
-If you decide to accept an offer, do so quickly.
-
-% If you have better offers, you might mention them to the professor and ask if they could match them.
-It is unlikely they can do so as RA might be fixed by the department and not negotiable.
-They might recommend you for departmental fellowships or some additional perks.
-
-\paragraph{Postponing an offer} If you need more time to decide, you can ask for more time: \emph{``Thank you for the offer.
-I am very excited about it.
-However, I am still waiting for other offers and need more time to decide.
-Would it be possible to postpone the decision for a few weeks?''}.
-This is perfectly fine and professors will understand and might even appreciate your honesty.
-They will likely give you a few weeks to decide.
-If you need more time, you can ask for more time.
-But do not ask for too much time, e.g., more than a month.
-You also should not postpone the offer multiple times, which will annoy people.
-
-\paragraph{Declining an offer} If you decide to decline or reject an offer, you can say: \emph{``Thank you for the offer.
-However, I have decided to accept another offer.
-I appreciate your time and consideration.
-I hope we can work together in the future.''}  Professors will understand and wish you luck.
-If you decide to reject an offer, do so quickly.
-
-This is very similar to a professional job offer.
-You should be serious about it and communicate promptly.
-Importantly, do not accept an offer and then later reject it.
-If you do this with a company, you might face legal consequences.
-If you do this with a professor, they will likely not do anything about it (i.e., not worth their time) but they will not want to work with you in the future.
-
-% 1.
-I'm fine if I extend an offer and you request a delay, even if it's because you're exploring other programs.
-No issue here, it's similar to job offerings.
-Of course, I may not be able to wait indefinitely and may extend the offer to another student, but I'll let you know that.
-% 2.
-I'm also okay if you find a program or professor that's a better fit.
-Similar to the first point, no hard feelings at all.
-However, I'd appreciate a prompt notification of your decision so I can extend an offer to another student.
-% 3.
-I'm NOT okay if you accept my offer but then, upon finding something better, tell me of your decision to withdraw Just like in job offers, once you are committed to an offer, there should be a compelling reason to withdraw (and claiming "I found somewhere better" is the weakest excuse).
-% While I won't take any actions against you (a company might), I naturally will remember this and may be cautious in future interactions.
-I'll also reconsider making similar offers to students in similar situations.
-
-\section{Academic Integrity (Cheating and Plagiarism)}
-
-Plagiarism and cheating (e.g., exams, assignments) are a BIG no-no in the US.
-If you're caught cheating, you will face serious consequences and likely be expelled from the university (e.g., after the second time at GMU).
-This is quite different from many international countries where cheating is common and often tolerated.
-Faculty is extremely good at detecting cheating (we have been dealing with these situations so many times over so many years), and \emph{will} report cheating cases.
-In short, whatever you do, don't cheat---not worth it.
-
-Here are the typical steps: (i) a faculty suspecting a cheating case will report it to the Office of Academic Integrity (OAI) at the university---the report often has supporting evidence and suggested penalty (e.g., a failing grade);  (ii)  OAI will take over and investigate the case; and (iii) OAI will make the final decision.
-It is important to note that after receiving the report from your prof., OAI \emph{completely} takes over and makes decisions.
-This means begging your professor will not help because they simply are no longer involved in the case.
-
-\section{Illegal Software} Using illegal/cracked software is very common in many countries (and even in the US).
-However, \emph{do not} install or use them on university computers, even those given to you by your adviser.
-It is unlikely that the university will track you down, but it is the \emph{software company} that will.
-They have very sophisticated tools to detect illegal software and will sue your university/department.
-Imagine your department or adviser being sued for a large sum of money, and it is \emph{you} who caused it.
-If you need to purchase software,  ask your adviser or the department (\autoref{sec:buying-equipment}).
-
-\section{Gifts}\label{sec:gifts} In many countries, it is customary to give professors gifts, often during holidays.
-These gifts can be costly and profs.
-sometimes expected them.
-In the US, it is uncommon and perfectly OK to not to give gifts.
-\emph{However}, if you'd like to offer small souvenir-like tokens, it's a thoughtful gesture that's appreciated.
-Some professors proudly display their gifts, which can come from students and colleagues (e.g., when they travel to their home countries or conferences).
-In summary, small gifts are fine, but avoid anything that might make your professors uncomfortable.
-
-\section{Relationships with your Profs.}
-There's a misconception that in the US it's all business, with professors as bosses who pay students for their work, and you doing something nice implying you expect something in return.
-However, the reality is quite the opposite.
-While people can be straightforward, they are also friendly and informal.
-You can call your professor by their first name (\autoref{sec:address}), disagree with them and argue (and gain respect doing so), seek their help (even on personal matters), come to their houses for parties or gathering, and give them small thoughtful gifts that they put on their desks (\autoref{sec:gifts}).
-Many students and professors maintain lifelong relationships, staying in touch through cards, emails, and calls, even after their academic journey ends.
-%These connections are genuine and meaningful.
-
-% \begin{commentbox}
-%   [DK:] Here are a few items you may consider addressing: putting international students in touch with other students from their countries, assuring them that they would be picked up from airports upon their arrival and that their initial stay will be taken care of.
-Most universities have other resources for these, but it is worth mentioning that they would be taken care of upon arrival and can get help during the transition phase.
-Learning to cook was a big deal when I arrived over 50 years ago-August 1973.
-But things have changed as one can find many ethnic food places, a big contrast from 1973, when there were two Indian restaurants in Greater Boston area.
-% \end{commentbox}
-
 \chapter*{History and Acknowledgement}\label{sec:ack}

### 2024-03-28 demystify.tex
updates

@@
     \textbf{Cover All?} & Yes                      & Yes                             & Yes                                   \\
-    \textbf{Summer?}              & No                       & Maybe                           & Yes                                   \\
+    \textbf{Summer?}              & No                       & Maybe                           & Likely                                   \\
     \midrule
@@
 As a TA, you spend up to 20 hrs/week and help with classes (e.g., grading or teaching labs/recitation).
-During a semester, a TA might work with several courses and professors (not necessarily their adviser).  TA funding is not typically available during the summer, which has few courses.
-
-\paragraph{How to get TA?}  Unless you have other funding such as RA or Fellowships, TA is typically default for CS PhD programs.
+During a semester, a TA might work with several courses and professors (not necessarily their adviser).  TA funding is \emph{not} typically available during the summer, which has few courses.
+
+\paragraph{How to get TA?}  Unless you have other funding such as RA or fellowships, TA is typically default for CS PhD programs.
 When you apply to be a full-time student,  state that you need financial assistant.
@@
 \begin{commentbox}[Vu:]
-  At GMU CS, PhD students admitted TA have 4 years of GTA guaranteed, and in also receive a stipend for the first summer.
+  At GMU CS, admitted PhD students have 4 years of GTA guaranteed, and also receive a stipend for the first summer.
 \end{commentbox}
@@
 
-It is important to note that RA is never guaranteed as it depends on the funding situation of the prof.
-So you should also pay attention to TA, which is a good backup plan (remember, typically TA and RA have similar benefits).
-This means you should also ask if TA is readily available for PhD students in the program.
+It is important to note that RA is \emph{never guaranteed} as it depends on the funding situation of the prof.
+So you should also pay attention to TA, which is a good backup plan (remember, typically TA and RA have quite similar benefits).
+This means you should also check if TA is readily available for PhD students in the program.
 
@@
 Fellowship is another type of funding that students can get (e.g., from the university, industries, government).
-Fellowships are typically competitive and generous and gives pretty much all benefits tuition/insurance that a TA/RA has.
+Fellowships are typically competitive and generous, giving pretty much all benefits tuition/insurance that a TA/RA has.
 They might even give higher stipend (including summer) and open doors for job opportunities such as internships.
 
-\textbf{How to get Fellowship?}   Many schools provide fellowships to attract admitted students.
-You likely will not need to do anything and the adcom will recommend such fellowships to admitted students.
+\textbf{How to get Fellowship?}   Many schools provide fellowships to attract students.
+You likely will not need to do anything and the adcom will recommend such fellowships to strong students.
 Some schools automatically offer a fellowship to all accepted students, while others only award it to a limited number of admitted students, such as the top percentile.
 
-For external fellowships including those from the US government and tech companies (e.g., Google, Microsoft, Facebook), you will need to apply.
+For external fellowships including those from the US government (e.g., NSF, DOD) and tech companies (e.g., Google, Microsoft, Facebook), you will need to apply.
 % However, tech companies including Google, Microsoft, Facebook have fellowships that international students can apply for.
@@
 In addition to general fellowships that are available to all students, there are also specific fellowships, e.g., for US citizen and permanent residents, for underrepresented minority (URM) students (\autoref{sec:urm}), and for students from specific countries.
-In short, there are many fellowships available, so you should look for them and apply.
+There are many fellowships available, so you should look for them and apply.
 
@@
 Every PhD student has pubs, but only a few  would have NSF Graduate or Microsoft fellowship.
-In fact, these are so prestigious that even if you didn't get it but make it to the final round or even ``honorable mentioning'', school will still mention you on their website, and you should put it on your CV.
+In fact, these are so prestigious that even if you didn't get it but make it to the final round or even \emph{``honorable mentioning''}, school will still mention you on their website, and you should put it on your CV.
 

### 2024-03-28 demystify.tex
updates

@@
 If you reject an offer, the prof.
-will likely stop looking for other students and might even reject other students; so by rejecting the offer later, you causing a great deal of inconvenience to the prof and also taking away the opportunity from other students.
+will likely stop looking for other students and might even reject other students; so by rejecting the offer later, causing a great deal of inconvenience to the prof and also taking away the opportunity from other students.
 This will likely would burn bridges with that prof and they will remember what you did to them.
@@
 
-Plagiarism and cheating (e.g., exams, assignments) are a BIG no-no in the US.
+Plagiarism and cheating (e.g., exams and assignments) are a BIG no-no in the US.
 If you're caught cheating, you will face serious consequences and likely be expelled from the university (e.g., after the second time at GMU).
@@
 
-\section{Gifts}\label{sec:gifts} In many countries, it is customary to give professors gifts, often during holidays.
-These gifts can be costly and profs.
-sometimes expected them.
-In the US, it is uncommon and perfectly OK to not to give gifts.
-\emph{However}, if you'd like to offer small souvenir-like tokens, it's a thoughtful gesture that's appreciated.
+\section{Costly Gifts}\label{sec:gifts} In many countries, it is customary to give professors costly gifts (e.g., expensive liquors, jewelry, or even money), often during the holidays, to show respect and appreciation.
+In the US, this is can be considered \emph{inappropriate} and discouraged.
+
+However, if you'd like to offer small souvenir-like tokens, it's a thoughtful gesture that's appreciated.
 Some professors proudly display their gifts, which can come from students and colleagues (e.g., when they travel to their home countries or conferences).
@@
 
-\section{Relationships with your Profs.}
-There's a misconception that in the US it's all business, with professors as bosses who pay students for their work, and you doing something nice implying you expect something in return.
-However, the reality is quite the opposite.
-While people can be straightforward, they are also friendly and informal.
-You can call your professor by their first name (\autoref{sec:address}), disagree with them and argue (and gain respect doing so), seek their help (even on personal matters), come to their houses for parties or gathering, and give them small thoughtful gifts that they put on their desks (\autoref{sec:gifts}).
-Many students and professors maintain lifelong relationships, staying in touch through cards, emails, and calls, even after their academic journey ends.
+\section{Relationships}
+There's a misconception that in the US it's all business, with professors as bosses who pay students for their work and that labmates are just work colleagues; and that doing nice things means expecting something in return.
+
+However, the reality is quite the opposite. While people can be straightforward and appear ``cold'', they are also informal, friendly, and very caring (in ways that might surprise you).
+With labmates and colleagues, you will often work and go to lunch together, confide in each other, help each other navigate the academic journey, and often become lifelong friends.
+With your professor(s), you can call them by their first name (\autoref{sec:address}), disagree with them and argue (and gain respect doing so), seek their help (even on personal matters), come to their houses for parties or gathering, and give them small thoughtful gifts that they put on their desks (\autoref{sec:gifts}).
+Many people maintain lifelong relationships with their professors and colleagues, staying in touch through cards, emails, and calls, even after their academic journey ends.
 %These connections are genuine and meaningful.

### 2024-03-29 demystify.tex
updates

@@
 
-Below we summarize the main points of this guideline. This gives you an overview to decide which specific topics you want to explore more thoroughly.
+This section summarizes the main points of this guideline. This gives you an overview to decide which specific topics you want to explore more thoroughly.
 
@@
         \begin{itemize}
-          \item The committee will look at various factors, but the most important ones are research ability, e.g., publications, statement of purpose (SOP), and  letters of recommendation (LORs).
+          \item The committee will look at various factors, but the most important ones are letters of recommendation (LORs) and research ability, e.g., publications, statement of purpose (SOP).
           \item \emph{LORs are very important}, but only if they are personalized and talk about your research ability (\autoref{sec:lor}).
@@
           \item  \emph{Publications can greatly help}.
-Publications are good venues are concrete evidence that you have successfully engaged in research.
+Publications in good venues are concrete evidence that you have successfully engaged in research.
 But they are not required!
@@
 While good grades might not help much, bad ones likely will hurt your case.
+
           \item \emph{Standard tests are not important} (\autoref{sec:standard-tests}).
-GRE \emph{is not} even required.
-Spend your time on something else!
+GRE typically \emph{is not} required.
 For standard English tests (not required for domestic students), just do enough to pass the minimum requirements.
@@
 
-\textbf{Advice}: As with school reputation, you can put a note about this in your statement and ask your LoR writers to talk about it.
+\textbf{Advice}: As with school reputation, you and your LoR writers can mention about the grading system of your university if you think that is helpful for adcom to evaluate.
 Note that while having good grades might not help much,
@@
 \paragraph{Online Courses and Certificates}
-In general, these do not carry much weight as they do not show your research ability.
-We do not care much if you have taken on online Coursera AI course or have a professional certificate from Microsoft.
-However, as mentioned in \autoref{sec:non-stem}, if you do not have a CS background, you might be able to use these to show your ability to handle CS courses.
+In general, these do not carry much weight as they do not show research ability. We do not care much if you have taken on online Coursera AI course or have a professional certificate from Microsoft.
+However, as mentioned in \autoref{sec:non-stem}, if you do not have a CS background, you might be able to use these to show you have sufficient CS knowledge.
 
@@
 
-\paragraph{Follow-Up Emails} If you had an interview and have not heard back, you can email to ask about the status of your application Don't send this too soon though, perhaps wait at least 2 weeks.
-
-You can also email to ask about the status of your application if you have received offers from other schools and need to make a decision soon.
-See \autoref{sec:accept-postpone-decline} for how to accept/postpone/decline offers.
+\paragraph{Follow-Up Emails} If you had an interview and have not heard back, you can email to ask about the status of your application.
+See \autoref{sec:accept-postpone-decline} for how to check status and follow-up emails.
 
@@
 Some schools send out admission letters in batches, some do rolling admission, and some do not send anything out (e.g., you're on their waitlist, or they might not send rejection letters).
-However, in general, you should hear back from most schools by around March (but some schools might take longer!).
+In general, you should hear back from most schools by around March (but some schools might take longer!).
 
@@
 
-\paragraph{Updating your profile} In general, you should not send emails to update your profile.
-However, if you have new publications or other big achievements, you can send an email asking to update your profile.
-
-\paragraph{Not getting interviews} While it is generally good to get interview, not getting an interview does not mean you're out.
+\paragraph{Updating your profile} You should not send emails to update your profile.
+However, if you have new publications or other big achievements, you can ask to update your application (though no guarantee that they will consider them).
+
+\paragraph{Not getting interviews} While it is generally good to get interview, not getting an interview \emph{does not} mean you're out.
 Many programs do not have the tradition of interviewing applicants.
@@
 However, no interviews typically means that you will not likely get an RA (\autoref{sec:ra}), which is offered by an individual faculty (if they you to do research for them, then they likely will interview you first).
-In the case of no interviews, your application (and TA/fellowship funding) will be reviewed and decided by the adcom.
+In the case of no interviews, your application (and TA/fellowship funding) is decided by the adcom.
 
@@
 
-You can also consider applying to MS programs, which are typically easier to get in; but you need to pay.
+You can consider applying to MS programs, which are typically easier to get in; but you likely need to pay.
 If you get in an MS program at a school of your choice, you can contact professors to work with them.
@@
 
-We will first discuss universities in the US that offer PhD in CS. Then we will talk about how to rank and select them.
+We will first discuss universities in the US that offer PhD in CS. Then we will talk about how to select them.
 
@@
 International\footnote{Even domestic students who have done their undergrad in the US might not know much about CS PhD programs in the US.} students often put schools into two bins: (i) very top schools that they dream about, and (ii) everything else.
-In many cases, they use rankings from US News, which are not very transparent and highly questionable\footnote{\url{https://cra.org/cra-statement-us-news-world-report-rankings-computer-science-universities/}}.
-Sometimes these students evaluate CS programs using the reputations non-CS programs such as medical, math, or physics.
-They even rank universities based on popular states they know in the US, e.g., California and New York.  Clearly, there are so many thing wrong with these methods.
+They commonly use rankings from US News, which is not transparent and questionable\footnote{\url{https://cra.org/cra-statement-us-news-world-report-rankings-computer-science-universities/}}.
+Sometimes they evaluate based on the reputation of the school's undergrad program or the reputation of the school's non-CS programs such as medical, math, or physics.
+They even rank universities based on popular states they know in the US, e.g., California, Texas, and New York.
 
@@
 You will be very surprised to learn that a school that you didn't know much about can have very strong research in your topic (and vice versa, a school you thought highly about might have no faculty working in the research field you're interested in).
-This is also a good way to learn about individual faculty (who works on what) and well-known CS conferences\footnote{In CS (and probably only in CS), conferences, not journals, are often the main venue to publish research findings (see \href{https://homes.cs.washington.edu/~mernst/advice/conferences-vs-journals.html}{why here}).}.
+This is also a good way to learn about individual faculty, e.g., who works on what, and well-known CS conferences\footnote{In CS (and probably only in CS), conferences, not journals, are often the main venue to publish research findings (see \href{https://homes.cs.washington.edu/~mernst/advice/conferences-vs-journals.html}{why here}).}.
 %\autoref{sec:ranking} gives the top 50 CS programs in the US according to CSRankings.
@@
 While many find CSRankings useful, it is not perfect, and in general, rankings are superficial, and you need to do more research to be informed and make better decisions.
-For example, if you get admission to several places, you should consider attending Open Houses (\autoref{sec:accepted}) and contact profs.
+For example, you might prefer places with a large community from your country.
+You might need to think about the cost of living, e.g., living in California is more expensive than in Virginia.
+You might prefer places with high tech companies (Seattle or Silicon Valley), with many outdoor activities (e.g., hiking, skiing), or with better weather (\emph{``PhD can be depressing, so would you rather be depressed in California or in New York?''}).
+
+If you get admission to several places, you should consider attending Open Houses (\autoref{sec:accepted}) and contact profs.
 that you're interested in at those places and talk to them.
 They would be more willing to chat with you now that you have been admitted.
-Ask them questions about their work, how they manage students, and their expectations.
+Ask questions about \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Answers-to-Ph.D-Advisor-Guide}{their work}, how they manage students, and their expectations.
 You can even ask to contact their students.
-See more in \autoref{sec:accepted} on what to do after getting admitted.
 
@@
 
-\paragraph{Checking your application status} If you have interviewed and not heard back from a professor after a few weeks or especially around the time when universities send out their admission decisions (around late Feb-- mid Mar), you can email to check.
+\paragraph{Checking your application status and Follow-up emails} If you have interviewed and not heard back from a professor after a few weeks or especially around the time when universities send out their admission decisions (around late Feb-- mid Mar), you can email to check.
 You can follow up the interview invitation and say: \emph{``Thanks for chatting with me.

### 2024-03-29 demystify.tex
updates

@@
 
-\paragraph{Core CS areas} The core CS areas that you should have knowledge in are:
+\paragraph{Core CS areas} Common CS areas that you should have knowledge in are:
 \begin{itemize}
@@
 However, you should check with your adviser first.
-Many profs.
-have funding to buy you computers and other equipments (e.g., software, books, keyboards, headphones, tablets, etc).
+They might have funding to buy you computers and other equipments (e.g., software, books, keyboards, headphones, tablets, etc).
 Many CS programs also provide some budget or computer equipments to incoming PhD students (e.g., a laptop).

### 2024-03-29 demystify.tex
updates

--- 2024-03-28 demystify.tex
@@
 So I want to share some details about the admission process and advice for those who are interested in applying for a \textbf{PhD in Computer Science in the US}.
-Originally, this document was intended for international students, but I have expanded it to include information that might be useful for \emph{domestic students} as well.
-While this handbook is primarily intended for students interested in CS, it might be relevant to students from various disciplines.
+Originally, this document was intended for international students, but I have expanded it to include information that might also be useful for \emph{domestic students}.
+Moreover, while this handbook is primarily intended for students interested in CS, it might be relevant to students from various disciplines.
 Furthermore, although many examples are specifics for schools that I and other contributors of this document know about, the information should be generalizable to other R1\footnote{An \href{https://en.wikipedia.org/wiki/List_of_research_universities_in_the_United_States}{R1 institution} in the US is a research-intensive university with a high level of research activity across various disciplines.
@@
 \begin{mybox}
-This document is available at
+The latest version of this document is available at
 
@@
 \noindent and its \LaTeX{} source is also on \href{https://github.com/nguyenthanhvuh/phd-cs-us}{Github}.
-If you have questions or comments, feel free to create a \href{https://github.com/nguyenthanhvuh/phd-cs-us/issues}{GitHub issue} for discussion.
+If you have questions or comments, feel free to create new \href{https://github.com/nguyenthanhvuh/phd-cs-us/issues}{GitHub issues} or \href{https://github.com/nguyenthanhvuh/phd-cs-us/discussions}{discussions}.
 
@@
 
-Students, especially those from smaller countries or less well-known universities, often wonder if they should apply to a CS PhD program in the US.
+International students, especially those from less well-known universities and smaller countries, often wonder if they should apply to a CS PhD program in the US.
 Specifically, they are worried about the \emph{difficulty} of getting admitted and the \emph{cost} of studying in the US.
@@
 
-\autoref{sec:summary} provides an overview of the application process, which you can explore in further details in the following chapters.
-
 \begin{commentbox}[Vu:]
@@
 \end{commentbox}
+
+\subsection{Who is this document for?}
+
+While this document can be useful for all students interested in CS PhD in the US, it is primarily intended for international students from \emph{smaller countries and less well-known universities}.
+These students have the disadvantage of not having access to the same resources and information as students from larger countries and well-known universities.
+This document aims to level the playing field by providing information that is not readily available to these students.
 
@@
 \paragraph{History} This handbook was conceived during a lunch with a faculty at GMU.
-We talked on about why GMU were not able to attract good Vietnamese and other international students, despite having a much stronger CS program than many schools that these students want to go to (part of the reason is described in \autoref{sec:selecting-ranking-schools}).
-We wished there were a way for international students to know about the US PhD programs (also for US faculty to understand more about international students and therefore have better chance of recruiting and working with them).
-I was also a member of the large VietPhD group on Facebook and saw many questions from students about PhD programs.
-However, most active participant are students in non-CS fields or not in US, and their answers are unfortunately not always accurate and sometimes leading to more confusion.
+We talked about why GMU was not able to attract good Vietnamese and other international students, despite having a much stronger CS program than many schools that these students want to go to (part of the reason is described in \autoref{sec:selecting-ranking-schools}).
+We wish there were a way for international students to know about the US PhD programs (also for US faculty to understand more about international students and therefore have a better chance of recruiting and working with them).
+
+I was also a member of the large VietPhD group on Facebook and browse Internet forums (e.g., Reddit/gradadmission and GradCafe).
+There I saw many questions from students about PhD programs.
+However, most active participants are students in non-CS fields or not in the US, and their answers are unfortunately not always accurate and sometimes lead to more confusion.
 So I thought it would be useful to have a document that is specific to CS PhD programs in the US from an insider prospective.
@@
 I started writing this handbook in May 2023 and have been updating it since then (mostly around deadline time when I tend to procrastinate, i.e., \emph{productive procrastination}!).
-The document was initially for international students, but has been generalized to all students.
+The document was initially for international students but has been generalized to all students.
 I have put the source code of this document on \href{https://github.com/nguyenthanhvuh/phd-cs-us}{GitHub} so that others can contribute to it.
@@
 Though I was not an international student, many of my students and collaborators are/were.
-I also mentor multiple students from Vietnam, and have close colleagues and friends who were once international students.
-I hope to capture the diverse challenges and experiences they've faced in this handbook, so that it  can be a valuable resource for prospective international students.
+I also mentor multiple students from Vietnam and have close colleagues and friends who were once international students.
+I hope to capture the diverse challenges and experiences they've faced in this document, so that it can be a valuable resource for prospective international students.
 Finally, my upbringing in the US provides a perspective aligned with American culture, allowing me to shed light on various issues, particularly those related to cultural differences (\autoref{sec:cultural}).
@@
 Many students including Didier (GMU), Thanh (Melbourne), and Dat (Melbourne) have contributed valuable questions and feedback.
+
 \textbf{Thank you!}

### 2024-03-29 demystify.tex
updates

@@
 
-\subsection{Who is this document for?}
-
-While this document can be useful for all students interested in CS PhD in the US, it is primarily intended for international students from \emph{smaller countries and less well-known universities}.
-These students have the disadvantage of not having access to the same resources and information as students from larger countries and well-known universities.
+\section{Who is this document for?}
+
+While this document can be useful for all students interested in CS PhD in the US, it is primarily intended for \text bf{international students} from \textbf{smaller countries and less well-known universities} (similarly, for domestic students from universities with no PhD programs or limited research experiences).
 This document aims to level the playing field by providing information that is not readily available to these students.
@@
 While many find CSRankings useful, it is not perfect, and in general, rankings are superficial, and you need to do more research to be informed and make better decisions.
-For example, you might prefer places with a large community from your country.
-You might need to think about the cost of living, e.g., living in California is more expensive than in Virginia.
-You might prefer places with high tech companies (Seattle or Silicon Valley), with many outdoor activities (e.g., hiking, skiing), or with better weather (\emph{``PhD can be depressing, so would you rather be depressed in California or in New York?''}).
+For example, you might prefer places with a large community from your country (e.g., Northern VA has a large population of Vietnamese).
+You might prefer places with high-tech companies (Seattle or Silicon Valley), with many outdoor activities (e.g., hiking, skiing), or with better weather (\emph{``PhD can be depressing, so would you rather be depressed in California or New York?''}).
+You might need to think about the cost of living, e.g., living in California is more expensive than in Virginia, or places with high crime rates (note that while some university might be in a high crime city, but the campus itself is safe).
 
@@
 
-\paragraph{You aim too high}  This is perhaps the most common one: you have applied to schools that are too high for your profile.
-While being ambitious is good, you should understand how PhD admission works (e.g., by reading this handbook!
-and realize things such as having a good GPA doesn't mean much) and be realistic!
-
-You should apply to a range of schools, including ``safety'' ones.   Before applying, you should talk to your professors and ask them to give you an honest opinion on where you should apply.
+\paragraph{You aim too high}  This is a very common reason: you have applied to schools that are too high for your profile.
+While being ambitious is good, you should understand how PhD admission works (e.g., by reading this handbook and realizing things such as having a good GPA doesn't mean much) and be realistic!
+
+Before applying, you should talk to your professors and ask them to give you an honest opinion on where you should apply. You should also apply to a range of schools, including ``safety'' ones.
 
@@
 A related reason is that you did not show enough interests in the program.
-For example, your SOP did not mention any faculty or research area at the program, or you did not respond to professors interested in interviewing you (which might be considered unprofessional and burn bridges).
+For example, your SOP did not mention any faculty or research area at the program, you did not respond to professors interested in interviewing you (which might be considered unprofessional and burn bridges), or during the interview you show little interest or enthusiasm (because they are your safety schools) in the program or profs.
 

### 2024-03-29 demystify.tex
updates

@@
 
-\paragraph{What else?} Make decision (\autoref{sec:accept-postpone-decline})!
-accept, reject, or defer the offers (if you have multiple options).
+\paragraph{What else?} Make decision, accept, reject, or defer the offers  (\autoref{sec:accept-postpone-decline}).
 Ask to meet with potential advisers (e.g., through Open House or separately) and even their students.
@@
 
-Also, do not forget to update and thank your LoR writters and others who have supported you through this process.
-
-\paragraph{If you do not get admitted} If you do not get admitted to any schools or don't want to go to the ones that admit you, try again next time.
+Also, do not forget to update and thank your LoR writers and others who have supported you through this process.
+
+\section{If you do not get admitted}\label{sec:not-accepted}
+
+If you do not get admitted to any schools or don't want to go to the ones that admit you, try again next time.
 Graduate admission can involve randomness and noise.
@@
 
-You can consider applying to MS programs, which are typically easier to get in; but you likely need to pay.
-If you get in an MS program at a school of your choice, you can contact professors to work with them.
-If you do well, you can ask the professor to support you to convert to PhD.
+You can consider applying to MS programs, which are typically easier to get in; but you likely need to pay.  Sometimes, the university that rejected you for PhD might accept you for MS.
+If you get in an MS program at a school of your choice, you can contact professors to work with them. If you do well, you can ask the professor to support you to convert to PhD.
 
@@
 Do not assume that you're young and healthy and ignore insurance.
-At GMU, and at most good CS PhD programs, your GTA or GRA \emph{will always} come with full insurance.
-In fact, at GMU your spouse/children will get significant discount rate for health insurance.
+At GMU, and most good CS PhD programs, your GTA or GRA \emph{will always} come with full insurance.
+In fact, at GMU your spouse/children will get a significant discount rate for health insurance.
 So you will never have to worry much about health issues for you or your family here.
@@
 
-\paragraph{How to get TA?}  Unless you have other funding such as RA or fellowships, TA is typically default for CS PhD programs.
+\paragraph{How to get TA?}  Unless you have other funding such as RA or fellowships, TA is typically the default for CS PhD programs.
 When you apply to be a full-time student,  state that you need financial assistant.
-It is common that the PhD committee will either admit you and give you GTA, or reject you; i.e., we do not admit a student without supporting them.
+Typically, adcom will either admit you and give you GTA, or reject you; i.e., we do not admit a student without supporting them.
 
@@
 
-Even if you have other funding and do not need TA, you still should do TA at least once.
-This allows you to see what teaching is like, which is especially helpful for research career where you often give talks and tell people about your work.
+Even if you have other funding and do not need a TA, you still should do TA at least once.
+This allows you to see what teaching is like, which is especially helpful for a research career where you often give talks and tell people about your work.
 GMU sometimes has classes that a more senior student can teach.
@@
   If you got recruited an offered an RA by a prof., you will likely get admitted.
-For example, if a prof., even if not in PhD adcom, wants to fund you, the PhD adcom will respect that decision and admit you (unless your application has many red flags).
+For example, if a prof., even if not in PhD adcom, wants to fund you, adcom will respect that decision and admit you (unless your application has many red flags).
 \end{commentbox}
@@
 
-Fellowship is another type of funding that students can get (e.g., from the university, industries, government).
+Fellowship is another type of funding that students can get (e.g., from the university, industries, or government).
 Fellowships are typically competitive and generous, giving pretty much all benefits tuition/insurance that a TA/RA has.
-They might even give higher stipend (including summer) and open doors for job opportunities such as internships.
+They might even give higher stipends (including summer) and open doors for job opportunities such as internships.
 
 \textbf{How to get Fellowship?}   Many schools provide fellowships to attract students.
-You likely will not need to do anything and the adcom will recommend such fellowships to strong students.
+You likely will not need to do anything and adcom will recommend such fellowships to strong students.
 Some schools automatically offer a fellowship to all accepted students, while others only award it to a limited number of admitted students, such as the top percentile.
@@
 or TA).
-PhD adcom members nominate applicants for this fellowship and the whole committee will decide.
-\end{commentbox}
-
-In addition to general fellowships that are available to all students, there are also specific fellowships, e.g., for US citizen and permanent residents, for underrepresented minority (URM) students (\autoref{sec:urm}), and for students from specific countries.
+Adcom members nominate applicants for this fellowship and the whole committee will decide.
+\end{commentbox}
+
+In addition to general fellowships that are available to all students, there are also specific fellowships, e.g., for US citizens and permanent residents, for underrepresented minority (URM) students (\autoref{sec:urm}), and students from specific countries.
 There are many fellowships available, so you should look for them and apply.
@@
 
-\section{Choosing a University}
+\section{Choosing a University}\label{sec:choosing-university}
 
@@
 
-You can learn about CS programs and research interests of faculty using resources such as \href{https://csrankings.org}{CSRankings.org}.
+You can learn about CS programs and the research interests of faculty using resources such as \href{https://csrankings.org}{CSRankings.org}.
 You will be very surprised to learn that a school that you didn't know much about can have very strong research in your topic (and vice versa, a school you thought highly about might have no faculty working in the research field you're interested in).
@@
 \section{Choosing an Adviser}
-Obviously, there is no one-size-fit-all answer to this question. The best adviser is the one that you can work well with.  But how do you find such a person?
+There is no one-size-fit-all answer to this question. The best adviser is the one that you can work well with.  But how do you find such a person?
 
@@
 
-Here is some general advice to find a professor.
+Here is some general advice for finding a professor.
 Assuming you're not familiar with any particular one, then first search for profs.
@@
 For example, I do this with multiple students, many of whom are undergrads.
-Many will drop because they find they don't like my research, but some will stay and become my PhD students.
+Many will drop out because they find they don't like my research, but some will stay and become my PhD students.
 
@@
   \textbf{Vu}: \autoref{sec:contact} shows how to contact a professor for research opportunities.
-Many probably will say no (or do not reply) as they do not have the bandwidth to take on random students, but some may say yes if they see potential fit.
+Many probably will say no (or do not reply) as they do not have the bandwidth to take on random students, but some may say yes if they see a potential fit.
 \end{commentbox}
@@
 \paragraph{Tenure-line and teaching faculty} Tenure-line faculty, consisting of tenured and tenured-track profs., focus on research, which includes publishing papers, obtaining grants, and mentoring PhD students.
-They often have low teaching load (e.g., 1 per semester).
+They often have a low teaching load (e.g., 1 per semester).
 \emph{Tenured} faculty are professors who have been promoted to a permanent position (informally, very hard to fire them).
@@
 
-In contrast, \emph{teaching} (or instructional faculty or professor of practice) mainly focus on teaching.
+In contrast, \emph{teaching} (or instructional faculty or professor of practice) mainly focuses on teaching.
 They typically teach 3--4 classes per semester (which is quite a lot) and \emph{do not} have research responsibilities, e.g., they do not have to worry much about publishing papers or obtaining grants.
@@
 
-\paragraph{Titles} Faculty have titles such as assistant, associate, full, regardless if they are tenure line or teaching.
+\paragraph{Titles} Faculty have titles such as assistant, associate, full, regardless if they are tenure-line or teaching.
 Assistant means new faculty, associate means they have been promoted, and full means they are senior.
 A tenure track faculty starts with being an assistant professor and then gets tenure and promoted to associate (typically after 6 years) and then full (time varies, some become full within 3-4 years, some 10+ years, some remain associate).
-A non-tenure track faculty, e.g., teaching faculty or lecturer, also starts with assistant, then promoted to associated, and full.
+A non-tenure track faculty, e.g., teaching faculty or lecturer, also starts with assistant, then promoted to associated, and full prof.
 
@@
 Emeritus means they are retired but still have some affiliation with the university.
-Research faculty (or research scientist) are typically non-tenure line faculty who focus on research.
+Research faculty (or research scientists) are typically non-tenure line faculty who focus on research.
 Due to their roles, adjunct, emeritus, and research faculty typically do not advise PhD students.
@@
 
-The short answer is that tenure-track faculty such as assistant professors are likely young and active in research (they have to, in order to get tenure).
+The short answer is that tenure-track faculty such as assistant professors are likely young and active in research (they have to, to get tenure).
 Thus, they will likely have more time to work with you and push you to do research and publish.
@@
 They might also expect you to figure things out yourself, i.e., you need to be independent.
-Some tenured faculty are also no longer active in research and more involved with administrative responsibilities or with their startup companies (this means they will likely not take new students).
+Some tenured faculty are also no longer active in research and are more involved with administrative responsibilities or with their startup companies (this means they will likely not take new students).
 
@@
 Most of the time, we ignore these emails, but in some rare occasions we do answer them.
-So how to write an email that get our attention?
+So how to write an email that gets our attention?
 
@@
 \paragraph{Overqualified or Lack Interest}  This might be paradoxical and is the opposite of aiming too high.
-However if the adcom believes that you are likely to get admitted and go to a better program (i.e., they are just your ``safety'' schools), they might not be so enthusiastic to admit you and want to save the spot for someone else.
+However if adcom believes that you are likely to get admitted and go to a better program (i.e., they are just your ``safety'' schools), they might not be so enthusiastic to admit you and want to save the spot for someone else.
 Note though it does happen, for most good schools this is rare and you should not assume this is the case.
 
-A related reason is that you did not show enough interests in the program.
-For example, your SOP did not mention any faculty or research area at the program, you did not respond to professors interested in interviewing you (which might be considered unprofessional and burn bridges), or during the interview you show little interest or enthusiasm (because they are your safety schools) in the program or profs.
+A related reason is that you did not show enough interest in the program.
+For example, your SOP did not mention any faculty or research area at the program, you did not respond to professors interested in interviewing you (which might be considered unprofessional and burn bridges), or during the interview you show little knowledge, interest, or enthusiasm (because they are your safety schools) in the program or profs.
 
 \paragraph{Red flags in Your Background} Plagiarism, cheating, or other academic dishonesty are \red{red} flags.
-Another one is if you have a history of jumping from one program to another (e.g., you have been in multiple different programs in short time).
+Another one is if you have a history of jumping from one program to another (e.g., you have been in multiple different programs in a short time).
 How would the reviewers know?

### 2024-03-30 demystify.tex
updates

@@
   \tcblower
-  \textbf{Vu}: This is an interesting and useful detail that US faculty might not be aware of and students should mention this in their statements.
+  \textbf{Vu}: This is an interesting and useful detail that US faculty might not be aware of.
+Students should mention this in their statements.
 In general, someone with a PhD has been through the research process and therefore can better evaluate your research ability.
-But if you do not have such people, then someone who can properly evaluate your research ability is still OK.
+But if you do not have such people, then someone who can properly evaluate your research ability is still OK (and again, explain that in your SOP).
 \end{commentbox}

### 2024-03-30 demystify.tex
updates

@@
 
-  Sometimes students would go through great length just to get letters from well-known professors in their school (e.g., department head).
-But as mentioned, if these professors do not know you, their letters would likely be generic and carry little value, in fact, \red{red flags}.
+Sometimes students would go to great length just to get letters from well-known professors in their school (e.g., department head).
+But as mentioned, if these professors do not know you, their letters would likely be generic and carry little value (sometimes \red{red flags}).
 Moreover, a top professor at your university might not be well-known to US faculty (see more details in \autoref{sec:your-school}).
 So save the trouble and get letters from \emph{any} professors/supervisors who know you well and can write a good letter about \emph{your} research ability.
-It's better to have a good personalized
-  letter about your own research ability from someone who is less
-  well-known than a generic/weak letter from a well-known person.
+It's better to have a good personalized letter about your own research ability from someone who is less well-known than a generic/weak letter from a well-known person.
 
@@
 When you waive your right to see the letter, it adds a layer of trust, showing you're confident in your choice of referees and that you're not trying to twist their words.
-It's also about keeping things open and honest between you and your letter writers, and encourages them to be real about your strengths and qualifications.
+It's also about keeping things open and honest between you and your letter writers and encourages them to be real about your strengths and qualifications.
 Plus, it keeps things private.
@@
 
-Most US universities have CS programs.
-However, while many universities offer PhD in their CS programs, many do not.
+Most US universities have CS study, but many of them do not have a PhD program in CS.
 These universities might offer just Bachelor's degrees (e.g., BS) and no graduate studies (i.e., no MS or PhD degrees), or they just offer MS programs (but no PhD).
@@
 
-International\footnote{Even domestic students who have done their undergrad in the US might not know much about CS PhD programs in the US.} students often put schools into two bins: (i) very top schools that they dream about, and (ii) everything else.
+International students often put schools into two bins: (i) very top schools that they dream about, and (ii) everything else.
 They commonly use rankings from US News, which is not transparent and questionable\footnote{\url{https://cra.org/cra-statement-us-news-world-report-rankings-computer-science-universities/}}.

### 2024-03-31 demystify.tex
updates

@@
 Some schools send out admission letters in batches, some do rolling admission, and some do not send anything out (e.g., you're on their waitlist, or they might not send rejection letters).
-In general, you should hear back from most schools by around March (but some schools might take longer!).
-
-In short, you should not worry too much about the timeline.
+In general, you should hear back from most schools by mid-March (but some schools might take longer!).
+
+In short, there is not much you can do about this other than to be patient and wait.
 Also do not send emails asking about interviews or status (unless you have interviewed specifically with someone then you can ask that person for status updates and others \autoref{sec:accept-postpone-decline}).
@@
 If there are discrepancies such as one voted reject but another full funding acceptance (such cases are rare, but do happen), the chair might ask the reviewers to coordinate to reach a concensus.
+
+%\section{Why are rejection letters sent so late?}\label{sec:late-rejection}
+
+%"Just reject me already!" is a common sentiment among applicants.
+Indeed, school will first send out admission offers to the top candidates.
+They do not send out rejection letters because there is still a chance that some of the top candidates will decline the offer.
+If they do, then the school will go to the next set of candidates and send out offers to them.
+This process continues until all spots are filled.
+This is why you might not hear back until late in the admission cycle.
 

### 2024-04-01 demystify.tex
updates

@@
 \paragraph{Interviews} Just like a job application, you likely will get interviews from schools.
-The most common case is that a professor is interested in working with you and wants to chat to make decision e.g., to offer RA (\autoref{sec:ra})).
-In some cases, the interview is done by several professors, e.g., to see if a student fits to their group or to recruit a very strong student to their program.
+The most common case is that a prof.
+is interested in working with you and wants to chat to make decision, e.g., to offer RA (\autoref{sec:ra})).
+In some cases, the interview is done by several professors, e.g., to see if a student fits in their group or to recruit a very strong student to their program.
 The interview can also be done by adcom members to see if the student fit their program.
@@
 You might also want have some 5-minute presentation about your research.
-If a professor asks you to read a paper, do it and be prepared to discuss it.
+If a prof.
+asks you to read a paper, do it and be prepared to discuss it.
 You can also ask if you need to prepare for a coding problem.
@@
 In general, fellowships, especially those that are open to everyone, are highly competitive and prestigious, and you will stand out if you get one.
-Every PhD student has pubs, but only a few  would have NSF Graduate or Microsoft fellowship.
+Every PhD student has pubs, but only a few would have NSF Graduate (GRFP) or Microsoft fellowships .
 In fact, these are so prestigious that even if you didn't get it but make it to the final round or even \emph{``honorable mentioning''}, school will still mention you on their website, and you should put it on your CV.
@@
   I always encourage the students I admitted to talk with my students and the students of other faculty in other schools who admitted them.
-You will unlikely hear straight-out complaints from current students in a professor’s group.
+You will unlikely hear straight-out complaints from current students in a prof’s group.
 But sometimes what is important are things that they (current students) don’t tell you.
@@
 
-Here is some general advice for finding a professor.
+Here is some general advice for finding a prof.
 Assuming you're not familiar with any particular one, then first search for profs.
@@
 If you want to work with SE \emph{and} AI, you can search for faculty who work in both SE and AI.
-Then you research that professor and contact them (\autoref{sec:contact}).
+Then you research that prof.
+and contact them (\autoref{sec:contact}).
 
@@
 Another effective way is taking graduate-level courses in the topics you are interested in.
-Professors teaching \emph{special topics courses} and \emph{research seminars} might be recruiting students---giving you even a higher chance.
+Profs teaching \emph{special topics courses} and \emph{research seminars} might be recruiting students---giving you even a higher chance.
 Do well in the class, answer questions, talk to the prof after classes, etc---being stand out.
-Many professors, including myself, prefer taking in new students this way.
-It gives both the professor and student more time, e.g., a whole semester, to work and evaluate the relationship before making any commitment (sounds a bit like a marriage!).
+Many profs, including myself, prefer taking in new students this way.
+It gives both the prof and student more time, e.g., a whole semester, to work and evaluate the relationship before making any commitment (sounds a bit like a marriage!).
 
@@
 \paragraph{Not a good fit}  You could have an excellent profile, but if you are interested in a research area that the program does not have, you might not be admitted.
-Similarly, if there is no faculty who is willing to advise you (e.g., they are on sabbatical, personal leave, do not have sufficient funding, or already have too many students already), you will not be admitted.
+Similarly, if there is no faculty who is willing to advise you (e.g., they are on sabbatical or personal leave, do not have sufficient funding, or already have too many students already), you will not be admitted.
 This is actually good for you, as you don't want to be in a program where no one can advise you.
 
-For example, 2024 admission cycle has a huge surge in students interested in AI and NLP (thanks, ChatGPT!).
+For example, the 2024 admission cycle has a huge surge in students interested in AI and NLP (thanks, ChatGPT!).
 Consequently, AI/NLP faculty might be overwhelmed and cannot consider many candidates, even those with excellent profiles.
@@
 However if adcom believes that you are likely to get admitted and go to a better program (i.e., they are just your ``safety'' schools), they might not be so enthusiastic to admit you and want to save the spot for someone else.
-Note though it does happen, for most good schools this is rare and you should not assume this is the case.
 
@@
 
-% \section{How much time do you spend looking at an application?}\label{sec:time-spent}
+%\section{How exactly do you evaluate an application?}\label{sec:ievaluate}
 
@@
 
-% I first skim the summary for school, GPA, test scores and looking for red flags like low GPA or test scores that do not meet the minimum requirements.
+% I first skim the summary for school, GPA, standard tests and look for red flags like low GPA or test scores that do not meet the minimum requirements.
 Then I scan transcripts for low/failing grades on CS and related courses.
@@
 % Then I look at LoRs.
-I read carefully good LoRs and quickly skim over bad or generic ones (very easy to tell, you don't even need much experiences to distinguish between a strong LoRs from the rest).
+I read carefully good LoRs and quickly skim over bad or generic ones (very easy to tell, you don't even need much experience to distinguish between strong LoRs from the rest).
 My comments for LoRs include ``+ very strong LoRs from well-known prof.
-XYZ talking about research experiences'' or ``- generic LoRs, mostly about course work''.
+XYZ talks about research experiences'' or ``- generic LoRs, mostly about course work''.
 I also note any unusual or unique details in the LoRs, e.g., ``+ LoR mentioned about student's significant contribution on well-known open-source project X''.
@@
 
-% I skim quickly over CV, looking for paper publications, research work experiences, and other relevant achievements such as winning prizes in well-known international programming/math competitions.
+% I skim over CV, looking for paper publications, research work experiences, and other relevant achievements such as winning prizes in well-known international programming/math competitions.
 My notes for CV include things such as ``+ has X papers at top venues'' or ``+ gold medal at X''.
@@
 
-% Next is the SOP, which I read carefully (but if it is bad then it would a quick read).
+% Next is the SOP, which I read carefully (but if it is bad then it would be a quick read).
 My notes for SOP include things such as ``- SOP mostly on course work, no clear motivation, generic" , or ``+ very enjoyable SOP talking about XYZ, customized for GMU, mentioned profs.
 X,Y,Z".
-If the SOP mentioned about publications then I would have ``+ has X papers at top venues" or ``+/- has X papers but at unknown places''.
-I will not go online search for the papers (unless they upload them as sample papers).
+If the SOP mentioned publications then I would have ``+ has X papers at top venues" or ``+/- has X papers but at unknown places''.
+I will not go online to search for the papers (unless they upload them as sample papers).
 I also skim over sampling papers, essentially looking over the abstract to see if reads well, and make notes such as ``+ have a paper under submission to unknown venue, but reads well".
-If the student mentioned about a project, I might also look at it (e.g., if it is on Github) and make notes such as ``+ has project on Github with X stars''.
-
-% At this point I have enough to make up my decision, which I enter to the evaluation system.
+If the student mentioned a project, I might also look at it (e.g., if it is on Github) and make notes such as ``+ has project on Github with X stars''.
+
+% At this point I have enough to make up my decision, which I enter into the evaluation system.
 Typically my decisions are either \emph{reject} or \emph{admit with full funding}.
@@
 
-% The adcom chair will look all evaluations for each application and makes a decision (likely following the decision of the reviewers).
-If there are discrepancies such as one voted reject but another full funding acceptance (such cases are rare, but do happen), the chair might ask the reviewers to coordinate to reach a concensus.
+% The adcom chair will look at all evaluations for each application and make a decision (likely following the decision of the reviewers).
+If there are discrepancies such as one voting reject but another full funding acceptance (such cases are rare, but do happen), the chair might ask the reviewers to coordinate to reach a consensus.
 
@@
 There are other things you might not think are important but can make you stand out.
-For example, if you have a strong background in a field that is not CS but can be integrated with CS, e.g., you have a degree or background in \emph{dance} or \emph{music} and want to integrate them with CS, do talk about it.
+For example, if you have a strong background in a non-CS field but can be integrated with CS, e.g., you have a degree or background in \emph{dance} or \emph{music} and want to integrate them with CS, do talk about it.
 Are you a female or a minority in CS (\autoref{sec:urm})?
@@
 
-\section{PhDs in the US vs. Other Countries}\label{sec:non-us-differences}
+\section{PhD in the US vs. Other Countries}\label{sec:non-us-differences}
 
@@
 
-\emph{MS Degree requirement}:  as mentioned in \autoref{sec:msrequirement} and \autoref{sec:time}, CS PhD programs in the US do not require MS degree.
-In contrast, many other countries do require MS degree before joining a PhD program.
+\emph{MS requirement and PhD duration}:  as mentioned in \autoref{sec:msrequirement} and \autoref{sec:time}, CS PhD programs in the US do not require MS degree.
+In contrast, many other countries require having an MS degree before joining a PhD program.
 This means that US PhD programs are longer (5--7 years, 2 of which are coursework) than other countries (3--4 years, no coursework).
@@
 \emph{Project proposal}: in many countries, you have to choose a project and adviser \emph{during} the application process (e.g., you write a proposal to a potential adviser).
+You start your research right from the beginning.
+
 In the US, you often start your PhD without an adviser or project and find them later.
-This is because as someone who just finished your undergraduate study, it is unlikely that you know enough about research to do proposal or picking an adviser.
 You will have two initial years to explore and find an adviser and research topic.
-
-\emph{Course work}: in the US you will spend the first couple of years taking classes and explore potential adviser and research topics.
-In other countries, you (who already have an MS) start your research right away, e.g., you immediately work on the research project you proposed with the adviser you chose.
-Also, in the US you also have to pass a series of exams, e.g., qualifying exam, comprehensive exam, thesis proposal defense\footnote{The word ``ABD'' (all but dissertation) is used in the US to refer to a PhD candidate who have finished all course work and exams and only need to write and defend their dissertation.}.
-In other countries, you might not have these exams or only have to do a few of them.
-
-\emph{Funding}:  In many countries stipend comes from the university or from the gov't.
-These funding often have a fixed duration (e.g., 3 or 4 years).
-In the US (\autoref{sec:funding}), stipend (e.g., RA) comes directly from your adviser (no fixed duration).
-There are also fewer TA opportunities in the European universities compared to the US.
-
-\emph{Faculty position after PhD}: In other countries, PhD graduates interested academia typically apply for additional research appointments, i.e., postdocs, and then consider faculty position.
-In the US, PhD graduates often apply directly for faculty position (postdoc for US graduates is no longer a popular option as it was before).
+And then you often explore projects with that adviser before settling on one (can take a couple of years).
+
+\emph{Course work}: as mentioned above in the US you will spend the first couple of years taking classes and exploring potential adviser and research topics. 
+After that, you also have to pass a series of exams during your PhD, e.g., qualifying exam, comprehensive exam, thesis proposal defense\footnote{The word ``ABD'' (all but dissertation) is used in the US to refer to a PhD candidate who have finished all course work and exams and only need to write and defend their dissertation.}.
+
+In other countries, you start your research right away, e.g., you immediately work on the research project you proposed with the adviser you chose.
+Moreover, you might not have exams like those in the US or only have to do a few of them.
+
+\emph{Funding}:  In many countries funding comes from the university or the gov't.
+This funding often has a fixed duration (e.g., 3 or 4 years).
+In the US (\autoref{sec:funding}), funding (e.g., RA) comes directly from your adviser (no fixed duration).
+There are also fewer TA opportunities in European universities compared to the US.
+
+\emph{Academic position after PhD}: In other countries, PhD graduates interested in academia typically apply for additional research appointments, i.e., postdocs, and then consider faculty positions.
+
+In the US, PhD graduates often apply directly for faculty positions (postdoc for US graduates is no longer a popular option as it was before).
 

### 2024-04-01 demystify.tex
updates

@@
 
-Also note that professors in a non-CS department unlikely can serve as formal PhD advisers for CS students (see \autoref{sec:related-fields}).
+Also note that non-CS profs unlikely can serve as formal PhD advisers for CS students (see \autoref{sec:related-fields}).
 
@@
 However, if your goal is a PhD in CS, then you likely need to be in the CS dept \emph{and} advised by a CS faculty.
-A non-CS faculty may \emph{co-advise} or be in the PhD dissertation committee, but your main PhD adviser will likely be a tenured or tenure-track faculty in CS (\autoref{sec:faculty-types}).
+A non-CS faculty can serve in the PhD dissertation committee (common) or \emph{co-advise} (somewhat rare), but your main PhD adviser will likely be a tenured or tenure-track faculty in CS (\autoref{sec:faculty-types}).
+For example, a prof.
+in Stats or Math dept might be able to serve as a co-adviser, but not as a sole adviser of a student in a CS PhD program.
+Similarly, a prof.
+in CS cannot serve as a formal adviser for a PhD student in a Math or Stats.
 If in doubt, check with the CS department for their requirements.

### 2024-04-03 demystify.tex
updates

@@
 
-In your statement, you have the opportunity to connect directly with the adcom and make your application stand out and unique.
-Your SOP can convince the adcom that you would \emph{fit} the program, even if you don't have very strong research experience.
+In your statement, you have the opportunity to connect directly with us and make your application stand out and unique.
+Your SOP can convince adcom members that you would \emph{fit} the program, even if you don't have very strong research experience.
 A well-written SOP also shows that you can communicate, which is very important in research, and that you can effectively teach and communicate with students, which is important for TA funding (see \autoref{sec:funding}).
@@
 There are various guides on writing SOP , and \href{https://cs-sop.org/}{many example statements} are available.
-In short, focus on your research motivation and goal, and convince us that you can achieve it through your experience, e.g., published papers.
+In short, focus on your research motivation and convince us that you can achieve it through your experience, e.g., published papers.
 You should also back up your claims with concrete evidences.
@@
 
-You should also talk about things that adcom members might not know about and can help make you \textbf{stand out} in the application pool of  thousands of applicants, e.g., you are member of a minority or LGBTQ+ group (\autoref{sec:urm}), your personal Github project with 1K+ stars or your regular contributions to well-known open-source projects (see \autoref{sec:improve-your-chance} for increasing your admission chance).
+You should talk about things that adcom members might not know about and can help make you \textbf{stand out} in the application pool of  thousands of applicants, e.g., you are member of a minority or LGBTQ+ group (\autoref{sec:urm}), your personal Github project with hundreds of stars or your regular contributions to well-known open-source projects (see \autoref{sec:improve-your-chance} for increasing your admission chance).
 
@@
 
-Finally, you should get your SOP reviewed by your LoR writers, professors (especially those who have served in adcom), or even graduate PhD students (because they have been through this process).
+Finally, you should get your SOP reviewed by your LoR writers, professors (especially those who have served in adcom), or even postdocs or PhD students (because they have been through this process).
 
 \begin{commentbox}[Vu:]
-  I often read SOP and LORs first. If I am
+  I often read LORs and SOP first. If I am
   persuaded by then, I would skim over other factors and advocate for
@@
 
+This section discusses various research experiences that can strengthen your application.
+\autoref{sec:research-opportunities} provides more information on how to find research opportunities (e.g., during your undergrad study).
+
 \paragraph{Publications} A concrete evidence of research ability is having \emph{papers in reputable international journals or conferences}.
@@
 
-\textbf{Advice}: If you think your school is the top in your country but might not be well-known internationally, you should talk about it in your SOP (and also ask your LOR writer to do that).
+\textbf{Advice}: If your school is the top in your country but you think it might not be well-known internationally, you should talk about it in your SOP (and also ask your LOR writer to do that).
 Of course, if you're interested in working with Vietnamese, consider  \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{CS programs in the US that have Vietnamese professors}.
@@
 
-\textbf{Advice}: As with school reputation, you and your LoR writers can mention about the grading system of your university if you think that is helpful for adcom to evaluate.
+\textbf{Advice}: As with school reputation, you and your LoR writers can mention the grading system of your university if you think that is helpful for adcom to evaluate.
 Note that while having good grades might not help much,
@@
 \textbf{Advice}: Even if you can't come in person, you should attend virtually and meet with individual faculty.
-During the event, you get a chance to learn more about the program, talk to individual faculty and current students.
-Take notes of faculty who make you excited, count those that are taking in new students (if they meet you, likely they are considering new students!).
-Talk to students about their advisers, the dept, the area, funding situation etc.
+During the event, you get a chance to learn more about the program, and talk to individual faculty and current students.
+Take notes of faculty who make you excited, and count those that are taking in new students (if they meet you, likely they are considering new students!).
+Talk to students about their advisers, the dept, the area, and funding situation.
 Ask about anything you want to determine that they deserve \emph{you}.
@@
 
-\paragraph{What else?} Make decision, accept, reject, or defer the offers  (\autoref{sec:accept-postpone-decline}).
+\paragraph{What else?} Make a decision, accept, reject, or defer the offers  (\autoref{sec:accept-postpone-decline}).
 Ask to meet with potential advisers (e.g., through Open House or separately) and even their students.
-Ask about computer equipment and software, office space, and other resources; in many case these will be provided for free by your adviser or department (\autoref{sec:buying-equipment}).
+Ask about computer equipment and software, office space, and other resources; in many cases these will be provided for free by your adviser or department (\autoref{sec:buying-equipment}).
 
@@
 However, you should check with your adviser first.
-They might have funding to buy you computers and other equipments (e.g., software, books, keyboards, headphones, tablets, etc).
-Many CS programs also provide some budget or computer equipments to incoming PhD students (e.g., a laptop).
-
-\section{Can I negotiate my PhD offer?}\label{sec:negotiate}
-For stipend, unlikely, as TA is set by the university and while RA is set by the adviser, it is often based on the department's standard, and it also would be unfair to other students if you get more.
-However, your stipend is often increased every year (automatically by a small amount) and also after you pass your qualifying exam.
+They might have funding to buy you computers and other equipment (e.g., software, books, keyboards, headphones, tablets, etc).
+Many CS programs also provide some budget or computer equipment to incoming PhD students (e.g., a laptop).
+
+\section{Can I negotiate my PhD offer (e.g., if I have multiple offers)?}\label{sec:negotiate}
+
+For stipend, unlikely for TA, which is set by the university.
+Moreover, while RA is set by the adviser, it is often based on the department's standard, and it also would be unfair to other students if you get more.
+However, your stipend is often increased every year (automatically by a small amount) and also after you pass your qualifier exams.
 
@@
 In short, standard things set by the university or department are unlikely non-negotiable.
-However, you can ask for things such as computer equipments (\autoref{sec:buying-equipment}), books, (and sometimes even furniture) and conference travel budget.
+However, you can ask for things such as computer equipment (\autoref{sec:buying-equipment}), books, (and sometimes even furniture), and a conference travel budget.
 

### 2024-04-04 demystify.tex
updates

@@
 
-While this document can be useful for all students interested in CS PhD in the US, it is primarily intended for \text bf{international students} from \textbf{smaller countries and less well-known universities} (similarly, for domestic students from universities with no PhD programs or limited research experiences).
-This document aims to level the playing field by providing information that is not readily available to these students.
+While the information presented here can be useful for all students interested in CS PhD in the US, this guideline is primarily intended for \textbf{international students} from \textbf{smaller countries and less well-known universities} (similarly, for domestic students from universities with no PhD programs or limited research experiences).
+My goal is to level the playing field by providing information that is not readily available to such students.
 
@@
 
-\paragraph{Grades} Having good GPA is good. However if your school is not well-known, having top grades or rankings
+\paragraph{Grades} Having a good GPA is good. However if your school is not well-known, having top grades or rankings
 usually will not help because we cannot evaluate them.
-This can be an issue for students in many top international universities where the competition is so high that very good students can still have low rankings from these schools (and be overlooked by the adcom).
+This can be an issue for students in many top international universities where the competition is so high that very good students can still have low rankings from these schools (and be overlooked by adcom).
 

### 2024-04-05 demystify.tex
updates

@@
 
-While the information presented here can be useful for all students interested in CS PhD in the US, this guideline is primarily intended for \textbf{international students} from \textbf{smaller countries and less well-known universities} (similarly, for domestic students from universities with no PhD programs or limited research experiences).
-My goal is to level the playing field by providing information that is not readily available to such students.
+While the information presented here can be useful for all students interested in applying to a CS PhD program in the US, this guideline is primarily intended for \textbf{international students} from \textbf{smaller countries and less well-known universities} (similarly, for domestic students from universities with no PhD programs or limited research experiences).
+Students from top schools with strong research cultural and experiences might already know most of the information presented here.
+My goal is to level the playing field by providing information that is not readily available to less privileged students.
+I also hope to encourage more students from these countries to apply to CS PhD programs in the US.
 
@@
 If you say you work on project X, then show some results (paper submitted (or even rejected), achieved X\% performance improvement over ..).
+Also, do not use more than 2 pages (you can have references in the third page).
 

### 2024-04-05 faculty.tex
updates

@@
 \chapter{Obtaining Funding}
+Getting funding is arguably the most important part of getting tenure.
+Most new professors struggle with grant writing, which is very different than writing research papers.
+However, without funding, you can't pay your students, travel to conferences, or even pay your own salary.
+In short, the lack of funding cause lost of stress and anxiety, and threaten your tenure.
+In this chapter, we will discuss the various ways to obtain funding.
+
+\section{Most people suck at grant writing (at first)} Most new assistant professors have no experience in grant writing, and struggle with it in the beginning.
+Then eventually they will get used to it and learn to like it (or at least tolerate it).
+The key is to keep trying, and to not give up.
+The best researchers at the best universities get rejected all the time.
+It's part of the process.
+So don't get discouraged.
+Keep trying, and you will get better over time.
+
+\paragraph{It is very different than writing research papers} Grant writing is a very different skill than writing research papers.
+In research papers, you describe what you \emph{did}.
+You convince people the merits of your work by showing the technical details, e.g., the algorithm,math, and proofs, and presenting experiments and results supporting your claim.
+You don't need to tell people who you are (often the authors are anonymous), and you don't need to convince people that you are the right person to do the work.
+Your reviewers are typically experts in your field and know a bit about the paper topic, and they are judging your work based on the technical merits.
+
+In grant writing, you are trying to convince people to give you money to do something you \emph{haven't done yet}.
+You need to convince people at a much higher level that your idea is important (impactful) and that you are the right person (with the right collaboration, resource, and plan) to do it.
+Getting too much into technical details can be a turn off.
+Your reviewers might not also experts in your field, and they are judging your work based on the \emph{big picture}.
+
+However, it's a skill that can be learned.
+The best way to learn is to read other people's grants, and to write your own.
+You will likely get rejected many times, but that's normal.
+You will get better over time.
+The key is to keep trying, and to not give up.
+If you are not getting rejected, you are not aiming high enough.
+The best researchers in the world get rejected all the time.
+It's part of the process.
+So don't get discouraged.
+Keep trying, and you will get better over time.
+
+It's a very different skill than writing research papers.
+However, it's a skill that can be learned.
+The best way to learn is to read other people's grants, and to write your own.
+You will likely get rejected many times, but that's normal.
+You will get better over time.
+The key is to keep trying, and to not give up.
+If you are not getting rejected, you are not aiming high enough.
+The best researchers in the world get rejected all the time.
+It's part of the process.
+So don't get discouraged.
+Keep trying, and you will get better over time.
 \section{NSF and other Government Funding}
@@
 
+\appendix
+\chapter{NSF Panel Reviews}
+
 \bibliographystyle{abbrv}

### 2024-04-07 demystify.tex
updates
typos

@@
 \begin{commentbox}[Vu:]
-  One of the reasons I create this document is that my colleagues at GMU are interested in recruiting Vietnamese students and are surprised when seeing very few applications in Vietnam (see \autoref{sec:ack}).
+  One of the reasons I created this document is that my colleagues at GMU are interested in recruiting Vietnamese students and are surprised when seeing very few applications in Vietnam (see \autoref{sec:ack}).
 Each year our CS PhD program receives about 400 PhD applications, most of which are international but only 4--5 are from Vietnam.
@@
 While the information presented here can be useful for all students interested in applying to a CS PhD program in the US, this guideline is primarily intended for \textbf{international students} from \textbf{smaller countries and less well-known universities} (similarly, for domestic students from universities with no PhD programs or limited research experiences).
-Students from top schools with strong research cultural and experiences might already know most of the information presented here.
+Students from top schools with strong research cultures and experiences might already know most of the information presented here.
 My goal is to level the playing field by providing information that is not readily available to less privileged students.
@@
 did your reference writers submit their letters?
-Usually, this screen process is done through a central university system, i.e., not by faculty from the CS department.
+Usually, this screening process is done through a central university system, i.e., not by faculty from the CS department.
 If you pass this screen, your application will be forwarded to the CS department for further evaluation.
@@
 These adcom members have a wide-range of expertise and background to ensure diverse perspectives in the evaluation process.
-The size and the review load of the adcom depends on the department size.
+The size and the review load of the adcom depend on the department size.
 At GMU, the PhD adcom typically has 15--20 faculty, and each committee member is assigned to review 25--30 applications.
@@
 Some members are tenured faculty.
-The chair of the committee will likely a senior faculty, but they will not review individual applications and instead assign them to committee members.
+The chair of the committee will likely be a senior faculty, but they will not review individual applications and instead assign them to committee members.
 The chair will look at various factors such as research interests or mentioning faculty names to assign the applications to appropriate faculty, e.g., I am often assigned to review applicants interested in software engineering.
 
-At GMU, we usually decide that a full-time PhD candidate is either (i) admit with funding (\autoref{sec:funding}) or (ii) rejected. In other words, in most cases, we either
-admit you with full funding, or we reject your application. In some rare cases, we admit
+At GMU, we usually decide that a full-time PhD candidate is either (i) admitted with funding (\autoref{sec:funding}) or (ii) rejected. In other words, in most cases, we either
+admit you with full funding or reject your application. In some rare cases, we admit
 without funding because you have funding on your own, e.g.,
@@
 
-\paragraph{Why CS depts do not waive application fee?}  Some universities do waive, e.g., Rice for all PhD applicants and many universities for domestic students (\autoref{sec:domestic-students}), however most do not.
-In any case, application fee is typically a requirement of the university.
+\paragraph{Why CS depts do not waive the application fee?}  Some universities do waive, e.g., Rice for all PhD applicants and many universities for domestic students (\autoref{sec:domestic-students}), however most do not.
+In any case, the application fee is typically a requirement of the university.
 Individual departments and programs do not have the flexibility to waive the application fee, even if they want to.
@@
 Having a strong letter from an internationally recognized researcher will \emph{greatly strengthen} your application.
-In fact, a strong letter from a well-known researcher can outweigh the lack of publications (or low GPA or unknown school).
+A strong letter from a well-known researcher can outweigh the lack of publications (or low GPA or unknown school).
 However, obtaining letters from famous people
-can be challenging for international students, who might not have much interactions with such experts.
-So it is fine to have a letter from people (e.g., profs, researchers, postdocs who mentored you) that know you well enough to talk about \emph{your research experience and capabilities}.
-Many students get letters from supervisors from companies where they did internships or are working at.
-It is OK as long as it is a research-based personalized letter (once again, we are talking about PhD applications, not MS).
-Again, the emphasis here is \emph{research}, i.e., the letters should describe your research experiences and potentials.
-Letters focusing your course work or non-research projects at a company won't carry much weight.
+can be challenging for international students, who might not have many interactions with such experts.
+So it is fine to have a letter from people (e.g., profs, researchers, postdocs who mentored you) who know you well enough to talk about \emph{your research experience and capabilities}.
+
+Many students get letters from supervisors from companies where they did internships or are working.
+It is OK as long as it is a research-based personalized letter.
+Again, the emphasis here is \emph{research}, i.e., the letters should describe your research experiences and potential.
+Letters focusing on your course work or non-research projects at a company won't carry much weight.
 
 \begin{commentbox}[Vu:]
-  If your grading system is not US standard or you are from a good school unknown outside your country, you can ask your reference writers to explain about that in their letters.
+  If your grading system is not US standard or you are from a good school unknown outside your country, you can ask your reference writers to explain that in their letters.
 For example, ``Bach Khoa'' are the top universities in Vietnam for STEM studies but few people outside Vietnam know about them.
@@
 \paragraph{How to help your LOR writer} Your LoR writer can ask you to provide additional information to help them write your letter.
-For example, they might ask your GPA, research and work experiences, papers (if any), or anything you want them to mention.
+For example, they might ask about GPA, research and work experiences, papers (if any), or anything you want them to mention.
 You can also provide them with a draft of your SOP so that they can see what you are saying about yourself and can complement that with their own perspective.
-It might also be a good idea to provide them such information even when they do not ask, especially if you have not interacted with them much or have not done much research with them.
+It might also be a good idea to provide them with such information even when they do not ask, especially if you have not interacted with them much or have not done much research with them.
 
 \paragraph{Self-written Letters} Many international students write letters themselves, typically due to the requests of their profs., and have them signed by their profs.
-Such letters have little values and are considered weak by reviewers---why can you not even find someone who cares or knows enough about you to write a candid personal reference letter?.
+Such letters have little value and are considered weak by reviewers---why can you not even find someone who cares or knows enough about you to write a candid personal reference letter?
 Instead of the reference writer talking about you, in this case it is you who write about yourself (and they just sign the letter).
@@
 If not, then you should ask someone else.
-For example, if a student I don't know well ask me to write letter for them, I will explicitly tell them I don't know them that well to write much about them, and such a short, generic, and weak letter will not help their case.
+For example, if a student I don't know well asks me to write a letter for them, I will explicitly tell them I don't know them that well to write much about them, and such a short, generic, and weak letter will not help their case.
 
@@
 In short, focus on your research motivation and convince us that you can achieve it through your experience, e.g., published papers.
-You should also back up your claims with concrete evidences.
+You should also back up your claims with concrete evidence.
 For example, if you say you have experience with teaching, then show what you did, e.g., undergrad TA or mentoring someone.
 If you say you work on project X, then show some results (paper submitted (or even rejected), achieved X\% performance improvement over ..).
-Also, do not use more than 2 pages (you can have references in the third page).
+Also, do not use more than 2 pages (you can have references on the third page).
 
@@
   Do careful research on professors, don't mention \emph{emeritus} or adjunct faculty (see more about various types of faculty in \autoref{sec:faculty-types}).
-  Also, be careful not to send statements to wrong schools or mixing
-  facts (e.g., talking about school X but mentioned about working with
-  profs. at school Y; and definitely do not talk about George Washington when applying to George Mason). I have seen such statements more time that I
-  should.
+  Also, be careful not to send statements to the wrong schools or mix
+  facts (e.g., talking about school X but mentioning working with
+  profs. at school Y; and do not talk about George Washington when applying to George Mason). I have seen such statements more times than I should.
 \end{commentbox}
@@
 \paragraph{Publications} A concrete evidence of research ability is having \emph{papers in reputable international journals or conferences}.
-Having published papers, especially at top venues, is a sign that the applicant has successfully involved in research.
+Having published papers, especially at top venues, is a sign that the applicant has been successfully involved in research.
 Applicants admitted to top schools often have multiple first-authored papers at top places.
@@
   Many international students mention Scopus Q1, which consists of various journals from IEEE, Elsevier, and many other publishers.
-I don't know/recognize many of journals listed in Scopus Q1.
+I don't know/recognize many of the journals listed in Scopus Q1.
 This might be something to be mindful of, as \textbf{CS} faculty might not be too familiar with Scopus or journals listed there, so devote some part in your statement to discuss the significance of your papers.
@@
 
-However, it is understandable that many students do not have the opportunities to publish papers.
+However, it is understandable that many students do not have the opportunity to publish papers.
 Thus, general writings, even those under submissions or rejected, would still be good (much better than nothing).
@@
 Unfortunately, many good research places in your countries, e.g., VinAI in Vietnam, remain relatively unknown to most universities in the US.
-So you or your LoR writers should explicit say something about them in your statement.
+So you or your LoR writers should explicitly say something about them in your statement.
 
@@
   Due to academic culture, professors in Vietnam usually aim for (international) journals instead of conferences.
-Could you give some tips on how to know whether a journal is good (CSrankings, unfortunately, only consider conferences)?.
+Could you give some tips on how to know whether a journal is good (CSrankings, unfortunately, only consider conferences)?
   \tcblower
-  Vu: One way is looking at what well-known researchers publish at.
+  Vu: One way is looking at what well-known researchers publish.
 For example, if you are interested in a field X, you can use CSRankings to look at active faculty in X, and then look at their websites to see what journals they publish at.
@@
 \begin{commentbox}[Vu:]
-  Sometime PhD adcom in the US will share a document such as \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Foreign-Top-Schools}{this one}, which lists the top schools in several countries.
+  Sometimes PhD adcom in the US will share a document such as \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Foreign-Top-Schools}{this one}, which lists the top schools in several countries.
 We also ask other faculty and students if we think they know about the place.
-For example, when I was a postdoc at UMD, members in their CS PhD adcom ask me to evaluate applicants from Vietnam.
-During my time at UNL and now here at GMU, I have looked at Vietnamese applications (whether they are assigned to me or not) and provide input to their reviewers, e.g., X is the top tech school in Vietnam and so it should be \emph{top} instead of \emph{unknown} foreign, which makes a huge difference.
+For example, when I was a postdoc at UMD, members of their CS PhD adcom asked me to evaluate applicants from Vietnam.
+During my time at UNL and now here at GMU, I have looked at Vietnamese applications (whether they are assigned to me or not) and provided input to their reviewers, e.g., X is the top tech school in Vietnam and so it should be \emph{top} instead of \emph{unknown} foreign, which makes a huge difference.
 \end{commentbox}
@@
 However, these programs often set higher requirements for students, including more demanding tests and assignments, resulting in lower GPAs and overall rankings.
-For example, an 3.5 GPA students from such talented programs are typically much better than a 4.0 GPA students not in those programs.
+For example, a 3.5 GPA students from such talented programs are typically much better than a 4.0 GPA students not in those programs.
 Similarly, variations in GPA standards exist among different universities, with technical universities generally having lower GPAs than economical universities.
@@
 \paragraph{English Tests} Unless your degrees are from specific countries such as \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-GMU#standard-tests-waiver-eligible-countries}{these}, you will need to
-take standardized English test.
-
-\textbf{Advice}: Just do well enough to pass minimum requirement set by the university. %, which is required for TA (\autoref{sec:funding}). 
-Just as with grades (\autoref{sec:your-school}) and GRE, having a high scores in these tests might not help, but having too low scores can be a \red{red flag} and sometimes results in an automatic rejection (\autoref{sec:why-rejected}), e.g., below the minimum requirement.
+take standardized English tests.
+
+\textbf{Advice}: Just do well enough to pass the minimum requirement set by the university. %, which is required for TA (\autoref{sec:funding}). 
+Just as with grades (\autoref{sec:your-school}) and GRE, having high scores in these tests might not help, but having too low scores can be a \red{red flag} and sometimes results in an automatic rejection (\autoref{sec:why-rejected}), e.g., below the minimum requirement.
 
@@
 
-\textbf{Advice}: do not spend much time talking about non-research related job in your SOP.
+\textbf{Advice}: do not spend much time talking about non-research-related job in your SOP.
 However, if you have research-based work experiences, then they can be useful and should be mentioned in your SOP and LORs.
@@
 
-\textbf{Advice}: Prepare your CV in such away that it allows the reviewers to quickly scan to identify major achievements, e.g., Publications, Programming Competition Awards, Teaching Experience.
+\textbf{Advice}: Prepare your CV in such away that it allows the reviewers to quickly scan to identify major achievements, e.g., Publications, Programming Competition Awards, and Teaching Experience.
 
@@
 
-After you submitted your applications, the waiting game begins! For many students, this is a very stressful time. This section provides some information and tips to help you get through this time.
+After you submit your applications, the waiting game begins! For many students, this is a very stressful time. This section provides some information and tips to help you get through this time.
 
@@
 You can also apply to specific programs \emph{before} you start your PhD, e.g., NSF Graduate Research Fellowship Program (GRFP) and Hertz Foundation Fellowship.
-These fellowships can be very competitive and will improve your admission chance if you get them.
+These fellowships are very competitive and can significantly improve your admission chance if you get them.
 

### 2024-04-07 demystify.tex
typos

@@
 However, if you have research-based work experiences, then they can be useful and should be mentioned in your SOP and LORs.
-For example, if you have worked at a research lab, e.g., Microsoft, DoE, VinAI in Vietnam, etc (\autoref{sec:research-experiences}), then you should mention that in your SOP and ask your supervisor to write about your research experiences and potentials in their LOR (\autoref{sec:lor}).
+For example, if you have worked at a research lab, e.g., Microsoft, DoE, VinAI in Vietnam, etc (\autoref{sec:research-experiences}), then you should mention that in your SOP and ask your supervisor to write about your research experiences and potential in their LOR (\autoref{sec:lor}).
 
@@
 The most common case is that a prof.
-is interested in working with you and wants to chat to make decision, e.g., to offer RA (\autoref{sec:ra})).
+is interested in working with you and wants to chat to make a decision, e.g., to offer RA (\autoref{sec:ra})).
 In some cases, the interview is done by several professors, e.g., to see if a student fits in their group or to recruit a very strong student to their program.
-The interview can also be done by adcom members to see if the student fit their program.
+The interview can also be done by adcom members to see if the student fits their program.
 %The professor might include one of their current students as part of the interview.
@@
 Typically, an interview takes about 15--30 minutes, and one important aspect of evaluation is your ability to effectively communicate, including speaking and understanding English. 
-You might be asked to talk about your research experiences and interests, and to read a paper and discuss it. In some rare cases you might also be asked to solve a problem, e.g., a coding problem. 
+You might be asked to talk about your research experiences and interests and to read a paper and discuss it. In some rare cases you might also be asked to solve a problem, e.g., a coding problem. 
 The interview is also a good opportunity for you to ask questions about the program and the university.
@@
 Have elevator pitches about your research experiences and interests.
-You might also want have some 5-minute presentation about your research.
+You might also want to have a 5-minute presentation about your research.
 If a prof.
@@
 For very strong candidates, the interview is actually to recruit them.
-In some cases a faculty interviews a candidate that they see potentials and want to advocate for their admission.
+In some cases a faculty interviews a candidate that they see potential and want to advocate for their admission.
 Without the interview, such applications may be more likely to be rejected.
@@
 
-\paragraph{Not getting interviews} While it is generally good to get interview, not getting an interview \emph{does not} mean you're out.
+\paragraph{Not getting interviews} While it is generally good to get an interview, not getting one \emph{does not} mean you're out.
 Many programs do not have the tradition of interviewing applicants.
@@
 \paragraph{Low Stipend?} Students often complain their stipend being too low.
-However, in most cases it is actually not bad and you can live comfortably with it.
+However, in most cases it is actually not bad, and you can live comfortably with it.
 It might also be enough to support your spouse and kids (many CS PhD students have their family with them).
@@
 A good school would know that it has to be competitive to attract students.
-For example, at GMU, every year we discuss about improving the benefits, and especially stipend, for our graduate students.
+For example, at GMU, every year we discuss improving the benefits, and especially stipend, for our graduate students.
 
@@
 \section{Why did I get rejected?}\label{sec:why-rejected}
-Many possible reasons!  Other than not having a strong profile (e.g., research potentials, GPA, LoRs, SOP, poor interviews), here are some other reasons.
+Many possible reasons!  Other than not having a strong profile (e.g., research potential, GPA, LoRs, SOP, poor interviews), here are some other reasons.
 

### 2024-04-07 demystify.tex
updates

--- 2024-04-05 demystify.tex
@@
 \paragraph{You aim too high}  This is a very common reason: you have applied to schools that are too high for your profile.
-While being ambitious is good, you should understand how PhD admission works (e.g., by reading this handbook and realizing things such as having a good GPA doesn't mean much) and be realistic!
+While being ambitious is good, you should understand how PhD admission works (e.g., by reading this document and realizing things such as having a good GPA doesn't mean much to top programs) and be realistic!
 
@@
 
-% I first skim the summary for school, GPA, standard tests and look for red flags like low GPA or test scores that do not meet the minimum requirements.
+% I first skim the summary for school, GPA, standard tests and look for red flags like low GPA or test scores that do not meet the minimum requirements set by the university.
 Then I scan transcripts for low/failing grades on CS and related courses.

### 2024-04-07 faculty.tex
updates

@@
 \chapter{Tenure Package and Application}
+After 5 or 6 years, it is time for you to go up (or apply) for tenure.
+You will need to prepare a tenure package and submit it to your department.
+The department will then review your package and vote on whether to recommend you for tenure.
+If the department votes in your favor, the package will then go to the college, and then to the university.
+The university will then make the final decision on whether to grant you tenure.
+
+This document shows my experience in preparing my tenure package and applying for tenure.
+I will share my timeline, the package (statements, letter writer sections), and the process.
+I will also share some advice on this process.
+
+\section{Timeline}
+Typically you will first mention to your department chair that you are planning to go up for tenure.
+The department chair then assign you a mentor who will help you with the process.
+You will apply for tenure in the Fall semester, and the decision will be made in the Spring semester.
+
+My own timeline is as follows:
+
+\begin{itemize}
+    \item Fall 2021: Moved to GMU (I started at UNL in 2016)
+    \item (late) Fall 2022: I told my department chair that I plan to go up for tenure in 2023.
+    \item Spring 2023: I started preparing my tenure package, which includes writing statements and think reference letters.
+I also took a semester off because GMU has a pre-tenure study break policy.
+    \item Summer 2023: I submitted my tenure package.
+    \item Fall 2024: I gave a tenure talk. Department voted in my favor, and then the college voted in my favor, and the decision goes up to the university.
+    \item Spring 2024: In March, I got the official notification that I got tenure and promoted to Associate Prof., beginning in Fall 2024.
+\end{itemize}
 
@@
 \subsection{Letters}
+This is a huge part of the tenure package. You will need to ask for letters from senior people in your area.  Someone who can vouch for your research, teaching, and services.
+
+You will need 4-6 letters, however you will not be the one asking for these letters, but your department will.
+However you can suggest who to ask.
+Typically, there will be two list of letter writers: one list consists of people suggested by you and the other list has people suggested by the department (e.g., people who work in your field in your dept).
+
+Originally I asked several senior people in my area to write letters for me.
+Then I found out that there is a rule at GMU that I cannot ask people whom have collaboration with me in the last 7 years!!!
+That essentially make most of my letter writers invalid.
+Moreover, GMU (and likely most other universities) prefer letter writers not from my list (because people on my list are likely to be biased), and has a rule that at least 60\% of the letter writers must be from the department's list.
+Also, there's a good chance that a letter writer will decline the invitation, so there should be more than 4-6 names in the lists.
+
+After talking to various people in my dept and outside, I decide to put just 2 names in my list and let the department decides the rest.
+I even thought about leaving my list empty but then that seems to be too weird and I don't want to give the impression that I don't have any friends in the world.
+Essentially this means that likely none or 1 of
+
+The reason is because I feel I don't have too many "enemies" out there that would write bad things about me.
+Those that don't know me well will likely just decline the letter invitation.
+Turned out to be a good decision because later I learn that all my letters wrote very positive things about me.
+I will never know who wrote the letters, but I am very grateful to them.
 

### 2024-04-07 demystify.tex
updates

@@
 
-In contrast, \emph{teaching} (or instructional faculty or professor of practice) mainly focuses on teaching.
+In contrast, \emph{teaching} (sometimes called instructional faculty or professor of practice) mainly focuses on teaching.
 They typically teach 3--4 classes per semester (which is quite a lot) and \emph{do not} have research responsibilities, e.g., they do not have to worry much about publishing papers or obtaining grants.
-They can mentor PhD students but they typically cannot serve as a formal adviser of PhD students, i.e., they \emph{cannot graduate} PhD students.
+As such they typically do not have funding (e.g., RA~\autoref{sec:funding}) to support PhD students.
+Depends on the university, they might be able to \emph{co-advise} and mentor PhD students with a tenure-line faculty.
 
@@
 Emeritus means they are retired but still have some affiliation with the university.
-Research faculty (or research scientists) are typically non-tenure line faculty who focus on research.
 Due to their roles, adjunct, emeritus, and research faculty typically do not advise PhD students.
+
+\emph{Research faculty} or research scientists are typically non-tenure line faculty who focus on research.
+While they cannot serve as a formal adviser, they often co-advise and mentor students with tenure-line faculty.
+They can recruit students and provide RA funding for them as they typically have their own funding.
 

### 2024-04-07 demystify.tex
updates

@@
 
-\emph{Research faculty} or research scientists are typically non-tenure line faculty who focus on research.
+\emph{Research faculty} or research scientists are typically non-tenure line faculty who focus on research (e.g., publishing papers and getting grants).
 While they cannot serve as a formal adviser, they often co-advise and mentor students with tenure-line faculty.
-They can recruit students and provide RA funding for them as they typically have their own funding.
+They can recruit students and provide RA funding for them.
 

### 2024-04-07 demystify.tex
updates

@@
 \emph{Tenure-track} faculty are (more likely young) professors who are on the track to get tenure.
-At most universities, \textbf{only} tenure and tenured track faculty can serve as a formal adviser of PhD students.
+At most universities, \textbf{only tenure and tenured track} faculty can serve as a formal adviser of PhD students.
 \autoref{sec:tenure-vs-tenure-track} talks more about choosing between tenure and tenured-track professor as your adviser.
@@
 
-\emph{Research faculty} or research scientists are typically non-tenure line faculty who focus on research (e.g., publishing papers and getting grants).
-While they cannot serve as a formal adviser, they often co-advise and mentor students with tenure-line faculty.
+Some universities have \emph{research faculty} or research scientists, who are typically non-tenure line faculty who focus on research (e.g., publishing papers and getting grants).
+While they cannot solely graduate PhD students, they often co-advise and mentor students with tenure-line faculty.
 They can recruit students and provide RA funding for them.

### 2024-04-07 demystify.tex
updates

@@
 
-Some universities have \emph{research faculty} or research scientists, who are typically non-tenure line faculty who focus on research (e.g., publishing papers and getting grants).
+Some universities have \emph{research faculty} or research scientists, who are typically non-tenure line faculty and focus on research (e.g., publishing papers and getting grants).
 While they cannot solely graduate PhD students, they often co-advise and mentor students with tenure-line faculty.
 They can recruit students and provide RA funding for them.
+They can also teach classes, but typically have a lower teaching load than teaching faculty.
 

### 2024-04-08 ms-swe-handbook.tex
updates

@@
 
+% https://cs.gmu.edu/current-students/ms-students/ms-in-software-engineering/
+
 The MS in Software Engineering (MS-SWE) program prepares students to become leaders in engineering high quality, large scale, computing solutions to real life problems.
-The MS-SWE program prioritizes working professionals—as such, all SWE courses are taught \emph{late afternoon} or \emph{early evening}.
+The MS-SWE program prioritizes working professionals--as such, all SWE courses are taught \textbf{late afternoon} or \textbf{early evening}.
 
@@
 Students are required to complete \textbf{30 credits} corresponding to \textbf{10 graduate courses}.
-Students are encouraged to download the course planner spreadsheet, and update it as they proceed through the program.
+Students are encouraged to download the \href{https://cs.gmu.edu/media/uploads/programs/graduate/masters/ms-swe-courseplanner.xlsx}{course planner spreadsheet}, and update it as they proceed through the program.
 
@@
 \begin{enumerate}
-    \item \textbf{SWE 619} Object-Oriented Software Specification and Construction
-    \item \textbf{SWE 621} Software Modeling and Architectural Design
-    \item \textbf{SWE 632} User Interface Design and Development
-    \item \textbf{SWE 637} Software Testing
+    \item \href{https://catalog.gmu.edu/search/?P=SWE%20619}{\textbf{SWE 619}} Object-Oriented Software Specification and Construction
+    \item \href{https://catalog.gmu.edu/search/?P=SWE%20621}{\textbf{SWE 621}} Software Modeling and Architectural Design
+    \item \href{https://catalog.gmu.edu/search/?P=SWE%20632}{\textbf{SWE 632}} User Interface Design and Development
+    \item \href{https://catalog.gmu.edu/search/?P=SWE%20637}{\textbf{SWE 637}} Software Testing
 \end{enumerate}
@@
 \begin{itemize}
-    \item CS 540 Language Processors (Compilers)
-    \item CS 550 Database Systems
-    \item CS 555 Computer Communications and Networking
-    \item CS 571 Operating Systems
-    \item CS 584 Theory and Applications of Data Mining
-    \item CS 675 Distributed Systems
-    \item CS 678 Advanced Natural Language Processing
+    \item \href{https://catalog.gmu.edu/search/?P=CS%20540}{CS 540} Language Processors (Compilers)
+    \item \href{https://catalog.gmu.edu/search/?P=CS%20550}{CS 550} Database Systems
+    \item \href{https://catalog.gmu.edu/search/?P=CS%20555}{CS 555} Computer Communications and Networking
+    \item \href{https://catalog.gmu.edu/search/?P=CS%20571}{CS 571} Operating Systems
+    \item \href{https://catalog.gmu.edu/search/?P=CS%20584}{CS 584} Theory and Applications of Data Mining
+    \item \href{https://catalog.gmu.edu/search/?P=CS%20675}{CS 675} Distributed Systems
+    \item \href{https://catalog.gmu.edu/search/?P=CS%20678}{CS 678} Advanced Natural Language Processing
 \end{itemize}
@@
 \begin{itemize}
-    \item ISA 562 Information Security Theory and Practice
-    \item ISA 650 Security Policy
-    \item ISA 673 Operating Systems Security
+    \item \href{https://catalog.gmu.edu/search/?P=ISA%20562}{ISA 562} Information Security Theory and Practice
+    \item \href{https://catalog.gmu.edu/search/?P=ISA%20650}{ISA 650} Security Policy
+    \item \href{https://catalog.gmu.edu/search/?P=ISA%20673}{ISA 673} Operating Systems Security
 \end{itemize}
@@
 \begin{itemize}
-    \item INFS 740 Database Programming for the World Wide Web
+    \item \href{https://catalog.gmu.edu/search/?P=INFS%20740}{INFS} 740 Database Programming for the World Wide Web
 \end{itemize}
@@
 \begin{itemize}
-    \item OR 542 Operations Research: Stochastic Models
+    \item \href{https://catalog.gmu.edu/search/?P=OR%20542}{OR 542} Operations Research: Stochastic Models
 \end{itemize}

### 2024-04-08 faculty.tex
updates

@@
 \subsection{Letters}
-This is a huge part of the tenure package. You will need to ask for letters from senior people in your area.  Someone who can vouch for your research, teaching, and services.
-
-You will need 4-6 letters, however you will not be the one asking for these letters, but your department will.
+This is a huge part of the tenure package. You will need letters from senior people in your area.  Someone who can vouch for your research, teaching, and services.
+
+You will need 4--6 letters, however you will not be the one asking for these letters, but your department will.
 However you can suggest who to ask.

### 2024-04-08 faculty.tex
updates

--- 2024-04-07 faculty.tex
@@
 \subsection{Letters}
-This is a huge part of the tenure package. You will need to ask for letters from senior people in your area.  Someone who can vouch for your research, teaching, and services.
-
-You will need 4-6 letters, however you will not be the one asking for these letters, but your department will.
-However you can suggest who to ask.
-Typically, there will be two list of letter writers: one list consists of people suggested by you and the other list has people suggested by the department (e.g., people who work in your field in your dept).
-
-Originally I asked several senior people in my area to write letters for me.
+This is a huge part of your tenure package.  You will need to ask for letters from senior people in your area.  Someone who can vouch for your research, teaching, and services.
+You will need 4-6 letters. However, you will not be the one asking for these letters, but your department will. You can suggest who to ask.
+
+Typically, there are two lists of letter writers: one consisting of people suggested by you and the other consisting of people suggested by the department (e.g., people who work in your field in your dept).
+There can also be another list, the ``blacklist'' where you list people that you don't want to write letters for you.
+
+Originally I asked several senior people to write my letters.
 Then I found out that there is a rule at GMU that I cannot ask people whom have collaboration with me in the last 7 years!!!
-That essentially make most of my letter writers invalid.
-Moreover, GMU (and likely most other universities) prefer letter writers not from my list (because people on my list are likely to be biased), and has a rule that at least 60\% of the letter writers must be from the department's list.
-Also, there's a good chance that a letter writer will decline the invitation, so there should be more than 4-6 names in the lists.
-
-After talking to various people in my dept and outside, I decide to put just 2 names in my list and let the department decides the rest.
-I even thought about leaving my list empty but then that seems to be too weird and I don't want to give the impression that I don't have any friends in the world.
-Essentially this means that likely none or 1 of
-
-The reason is because I feel I don't have too many "enemies" out there that would write bad things about me.
+That invalidates most of my letter writers.
+Moreover, GMU (and likely most other universities) would prefer writers not from my list (because people on my list are likely to be biased), and has a rule that at least 60\% of the letter writers must be from the department's list.
+Moreover, there's a good chance that a letter writer will decline the invitation, so there should be more names.
+
+After talking to various people in my dept and outside, I decide to put just 2 names in my list (0 on blacklist) and let the department decides the rest.
+I also thought about leaving my list empty but then that seems to be too weird, and I don't want to give the impression that I don't have any friends in the world.
+The reason is that I feel I don't have too many "enemies" out there that would write bad things about me.
 Those that don't know me well will likely just decline the letter invitation.
-Turned out to be a good decision because later I learn that all my letters wrote very positive things about me.
+Turned out to be a good decision because in the end the department selected 1 of the 2 names on my list and 4 more names from the department's list.
+Moreover, it looks like no one declines so my letter writers are all set fairly quickly in the process.
+All of my letters were good and give many positive comments about me as summarized in my evaluation letter (sent to me after the tenure decision).
 I will never know who wrote the letters, but I am very grateful to them.

### 2024-04-08 demystify.tex
updates

--- 2024-04-07 demystify.tex
@@
 Important factors to consider include the reputation of schools and professors (\autoref{sec:schoolsandprofs}), and funding availability (\autoref{sec:funding}).
-You will likely have to make your decision by around \emph{April 15} (\href{https://cgsnet.org/wp-content/uploads/2024/01/CGS_April15_Resolution_Jan312024.pdf}{why this date?}).
+You will likely have to make your decision (\autoref{sec:accept-postpone-decline}) by around \emph{April 15}, which is the deadline for most schools
+(\href{https://cgsnet.org/wp-content/uploads/2024/01/CGS_April15_Resolution_Jan312024.pdf}{why this date?}).
 
@@
 
-\paragraph{Tenure-line and teaching faculty} Tenure-line faculty, consisting of tenured and tenured-track profs., focus on research, which includes publishing papers, obtaining grants, and mentoring PhD students.
+\paragraph{Types of Faculty} There are typically three types of faculty: tenure-line, teaching, and research.
+The main difference is their focus: research, teaching, or both.
+The type of faculty can affect their ability to serve as your adviser.
+
+\textbf{Tenure-line faculty}, consisting of tenured and tenured-track profs., focus on research, which includes publishing papers, obtaining grants, and mentoring PhD students.
 They often have a low teaching load (e.g., 1 per semester).
@@
 \emph{Tenure-track} faculty are (more likely young) professors who are on the track to get tenure.
-At most universities, \textbf{only} tenure and tenured track faculty can serve as a formal adviser of PhD students.
+For most schools, tenure-line is the main type of faculty that can serve as a formal adviser of PhD students.
 \autoref{sec:tenure-vs-tenure-track} talks more about choosing between tenure and tenured-track professor as your adviser.
 
-In contrast, \emph{teaching} (sometimes called instructional faculty or professor of practice) mainly focuses on teaching.
+\textbf{Teaching faculty} (sometimes called instructional faculty or professor of practice) mainly focuses on teaching.
 They typically teach 3--4 classes per semester (which is quite a lot) and \emph{do not} have research responsibilities, e.g., they do not have to worry much about publishing papers or obtaining grants.
 As such they typically do not have funding (e.g., RA~\autoref{sec:funding}) to support PhD students.
-Depends on the university, they might be able to \emph{co-advise} and mentor PhD students with a tenure-line faculty.
-
-Also note that non-CS profs unlikely can serve as formal PhD advisers for CS students (see \autoref{sec:related-fields}).
-
-\paragraph{Titles} Faculty have titles such as assistant, associate, full, regardless if they are tenure-line or teaching.
+Depends on the university, they might be able to \emph{co-advise} and mentor PhD students with a tenure-line faculty, but typically do not formally advise and graduate PhD students.
+
+Note that non-CS profs (even in tenure-line ones) often cannot solely advise and graduate PhD students (see \autoref{sec:related-fields}).
+
+Some universities have \textbf{research faculty} or research scientists, who are typically non-tenure line faculty and focus on research (e.g., publishing papers and getting grants). 
+They can recruit students and provide RA funding for them. 
+They can also teach classes, but typically have a lower teaching load.
+For some universities, they \emph{can} solely advise and graduate PhD students, but for others, they cannot.
+
+\textbf{Adjunct faculty} is not full-time, e.g., they might be working in industry and teach a class or two for fun.
+\textbf{Emeritus faculty} means they are retired from the university, e.g., they could move to the industry, but still have some affiliation with the university.
+Due to their roles, these profs.
+typically do not advise PhD students.
+
+\paragraph{Ranks} Faculty have ranks such as assistant, associate, full, regardless if they are tenure-line or teaching.
 Assistant means new faculty, associate means they have been promoted, and full means they are senior.
 A tenure track faculty starts with being an assistant professor and then gets tenure and promoted to associate (typically after 6 years) and then full (time varies, some become full within 3-4 years, some 10+ years, some remain associate).
-A non-tenure track faculty, e.g., teaching faculty or lecturer, also starts with assistant, then promoted to associated, and full prof.
-
-Adjunct faculty is not full-time, e.g., they might be working in industry and teach a class or two for fun.
-Emeritus means they are retired but still have some affiliation with the university.
-Due to their roles, adjunct, emeritus, and research faculty typically do not advise PhD students.
-
-\emph{Research faculty} or research scientists are typically non-tenure line faculty who focus on research (e.g., publishing papers and getting grants).
-While they cannot serve as a formal adviser, they often co-advise and mentor students with tenure-line faculty.
-They can recruit students and provide RA funding for them.
+A non-tenure track faculty, e.g., teaching or research faculty, also starts with assistant, then promoted to associated, and full prof.
 
@@
 
-\paragraph{Accepting an offer and later later rejecting it}
-I have seen \emph{many} international students asking about this, e.g., accepting an offer from a prof.
-and then later rejecting it when they get admitted to a better/dream school, which have them on its wait list.
-There are numerous advice, e.g., on Internet forums, that this is OK to do so (most of these advice are from other international students).
-The reasons typically are \emph{``you haven't established much relationship with the prof.
-yet, so it is OK to reject it; no hard feeling, you probably won't see them again anyway''} or that \emph{``it's your life and you should do what's best for you''}.
-
-From my experience, this is \emph{not} a good idea and can be considered \emph{rude and unprofessional}.
-But ultimately, it is a trade-off that you need to decide,  e.g., between your own benefit and the cost to others.
-If you reject an offer, the prof.
-will likely stop looking for other students and might even reject other students; so by rejecting the offer later, causing a great deal of inconvenience to the prof and also taking away the opportunity from other students.
-This will likely would burn bridges with that prof and they will remember what you did to them.
-Moreover, given that research subfields are typically small, you might run into that professor again in the future, and they will remember what you did to them.
+\paragraph{Accepting an offer and later rejecting it}
+
+I've noticed many students, especially international, face a dilemma when they commit to a graduate offer but then receive a better one, often discussed in online forums.
+Advice often suggests it's okay to switch, using reasons like not having a strong relationship with the prof.
+or prioritizing personal benefit.
+
+In my opinion, these reasons are not strong enough to justify retracting an acceptance.
+A more valid reason is using the \href{https://cgsnet.org/wp-content/uploads/2024/01/CGS_April15_Resolution_Jan312024.pdf}{April 15 resolution}, which many universities participate.
+Among various things, this resolution states that students are free to accept a new offer from a different institution until 4/15, even if they have already accepted an offer elsewhere.
+
+However, regardless of the resolution, retracting an acceptance can have ethical implications.
+When you accept an offer, you are making a commitment to work with that prof, who then might stop looking for other students.
+So by retracting your offer, you are breaking your commitment and also causing a great deal of inconvenience to the prof and also takes away the opportunity from other students.
+Ultimately, this choice is personal, involving a balance between personal benefit and the ethical impact on others.
 
@@
 \section{Relationships}
-There's a misconception that in the US it's all business, with professors as bosses who pay students for their work and that labmates are just work colleagues; and that doing nice things means expecting something in return.
+There's a misconception that in the US it's all business, with professors as bosses who pay students for their work and that lab mates are just work colleagues; and that doing nice things means expecting something in return.
 
 However, the reality is quite the opposite. While people can be straightforward and appear ``cold'', they are also informal, friendly, and very caring (in ways that might surprise you).
-With labmates and colleagues, you will often work and go to lunch together, confide in each other, help each other navigate the academic journey, and often become lifelong friends.
+With lab mates and colleagues, you will often work and go to lunch together, confide in each other, help each other navigate the academic journey, and often become lifelong friends.
 With your professor(s), you can call them by their first name (\autoref{sec:address}), disagree with them and argue (and gain respect doing so), seek their help (even on personal matters), come to their houses for parties or gathering, and give them small thoughtful gifts that they put on their desks (\autoref{sec:gifts}).
@@
 
-\paragraph{History} This handbook was conceived during a lunch with a faculty at GMU.
+\paragraph{History} This document was conceived during a lunch with a faculty at GMU.
 We talked about why GMU was not able to attract good Vietnamese and other international students, despite having a much stronger CS program than many schools that these students want to go to (part of the reason is described in \autoref{sec:selecting-ranking-schools}).
@@
 
-I was also a member of the large VietPhD group on Facebook and browse Internet forums (e.g., Reddit/gradadmission and GradCafe).
+I was also a member of the large VietPhD group on Facebook and browsed Internet forums (e.g., \texttt{Reddit/gradadmission} and \texttt{GradCafe}).
 There I saw many questions from students about PhD programs.
@@
 
-I started writing this handbook in May 2023 and have been updating it since then (mostly around deadline time when I tend to procrastinate, i.e., \emph{productive procrastination}!).
+I started writing this document in May 2023 and have been updating it since then (mostly around deadline time when I tend to procrastinate, i.e., \emph{productive procrastination}!).
 The document was initially for international students but has been generalized to all students.
@@
 \paragraph{About Me} I am an assistant professor in the CS dept at George Mason University (GMU).
-Prior to GMU, I was at the University of Nebraska-Lincoln (UNL).
+Before GMU, I was at the University of Nebraska-Lincoln (UNL).
 I have been in the PhD admission process at GMU and UNL for many years.
@@
 I also mentor multiple students from Vietnam and have close colleagues and friends who were once international students.
-I hope to capture the diverse challenges and experiences they've faced in this document, so that it can be a valuable resource for prospective international students.
+I hope to capture the diverse challenges and experiences they've faced in this document so that it can be a valuable resource for prospective international students.
 Finally, my upbringing in the US provides a perspective aligned with American culture, allowing me to shed light on various issues, particularly those related to cultural differences (\autoref{sec:cultural}).

### 2024-04-08 demystify.tex
updates

@@
 For example, in addition to a traditional CS department, GMU has IST and Cybersecurity departments, both of which have faculty with PhD in CS and work on CS topics (e.g., AI, Security, Robotics).
-So it is totally possible that you still get to do CS research and publish in CS-focused venues even if you're not in a traditional CS program.
-It is common to see faculty with CS PhD in a non-CS department as well as faculty with non-CS PhD in CS department.
+So it is possible that you still get to do CS research and publish in CS-focused venues even if you're not in a traditional CS program.
+It is common to see faculty with CS PhD in a non-CS department as well as faculty with non-CS PhD in a CS department.
 
 However, if your goal is a PhD in CS, then you likely need to be in the CS dept \emph{and} advised by a CS faculty.
-A non-CS faculty can serve in the PhD dissertation committee (common) or \emph{co-advise} (somewhat rare), but your main PhD adviser will likely be a tenured or tenure-track faculty in CS (\autoref{sec:faculty-types}).
+A non-CS faculty can serve in the PhD dissertation committee (common) or \emph{co-advise} (somewhat rare), but your main PhD adviser will likely be a tenure-line faculty in CS (\autoref{sec:faculty-types}).
 For example, a prof.
@@
 Similarly, a prof.
-in CS cannot serve as a formal adviser for a PhD student in a Math or Stats.
+in CS cannot be the sole adviser and graduate a PhD student in a Math or Stats.
 If in doubt, check with the CS department for their requirements.
 
-For this specific reason,  CSRankings includes only tenure line faculty who can advise CS PhD students.
+For this specific reason,  CSRankings includes only faculty who can advise CS PhD students.
 I also have compiled a \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{list} of Vietnamese faculty who can advise PhD students.

### 2024-04-14 demystify.tex
updates

@@
 
-\section{Who is this document for?}
-
-While the information presented here can be useful for all students interested in applying to a CS PhD program in the US, this guideline is primarily intended for \textbf{international students} from \textbf{smaller countries and less well-known universities} (similarly, for domestic students from universities with no PhD programs or limited research experiences).
+\section*{Who is this document for?}
+
+The information presented here can be useful for all students, but this guideline is primarily intended for \textbf{international students} from \textbf{smaller countries and less well-known universities} (similarly, for domestic students from universities with no PhD programs or limited research experiences).
 Students from top schools with strong research cultures and experiences might already know most of the information presented here.
-My goal is to level the playing field by providing information that is not readily available to less privileged students.
-I also hope to encourage more students from these countries to apply to CS PhD programs in the US.
+My goal is thus to level the playing field by providing information that is not readily available to less privileged students. 
+I also hope to encourage more students with such backgrounds to apply to CS PhD programs in the US.
+
+The document might also not sufficient for students aiming for top-10 schools, which might have different requirements\footnote{e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU}.
+However, the information here should be useful for students aiming for other good schools, e.g., R1 institutions in the US.
+It is definitely useful for students aiming for GMU :)
 
@@
 Then, applications are reviewed by a \emph{PhD admission committee} (\textbf{adcom}) that consists of faculty members in CS\footnote{In some cases the committee can involve affiliated faculty from different disciplines.}.
-These adcom members have a wide-range of expertise and background to ensure diverse perspectives in the evaluation process.
+These adcom members have a wide range of expertise and background to ensure diverse perspectives in the evaluation process.
 The size and the review load of the adcom depend on the department size.
@@
 
-\paragraph{Why CS depts do not waive the application fee?}  Some universities do waive, e.g., Rice for all PhD applicants and many universities for domestic students (\autoref{sec:domestic-students}), however most do not.
+\paragraph{Why CS depts do not waive the application fee?}  Some universities do waive, e.g., Rice does that for PhD applicants and many universities for domestic students (\autoref{sec:domestic-students}), however most do not.
 In any case, the application fee is typically a requirement of the university.
@@
 A strong letter from a well-known researcher can outweigh the lack of publications (or low GPA or unknown school).
+
 However, obtaining letters from famous people
@@
 
-If you do not waive your right, then the letter writer might refuse to write for you or write a generic letter that does not help your case.
+If you do not waive your right,  the letter writer might refuse to write for you or write a generic letter that does not help your case.
 Reviewers also might raise concerns about a letter that is not waived, e.g., if you do not trust your letter writers, then you should find someone else to write for you.
@@
 The most common type of funding is \textbf{graduate assistantship}, which is either TA or RA.
-Both TA and RA come with tuition waiving (you don't have to pay tuition), health insurance (this takes care of your insurance, which is a must have in the US), and most importantly, your stipend (i.e., your salary).
-Some universities also give significant discount or pay insurance for spouse/children.
+Both TA and RA come with tuition waiving (you don't have to pay tuition), health insurance (this takes care of your insurance, which is a must-have in the US), and most importantly, your stipend (i.e., your salary).
+Some universities also give significant discounts or pay insurance for spouses/children.
 
 Several things about stipends.
-First, the amount of stipend \emph{varies} and depends on factors such as location (e.g., a stipend in Washington DC is likely higher than in Lincoln, Nebraska due to higher living cost).
-Second, an academic year (AY)  year is typically \emph{9-month} in the US, so stipend is for 9 months (so divide by 9 for each month).
+First, the amount of stipend \emph{varies} and depends on factors such as location (e.g., a stipend in Washington DC is likely higher than in Lincoln, Nebraska due to higher living costs).
+Second, an academic year (AY)  year is typically \emph{9-month} in the US, so the stipend is for 9 months (so divide by 9 for each month).
 Third, you might get paid over the summer through funding from your professor or fellowship (typically no TA over the summer).
-Fourth, like most source of income in the US, you will have to \emph{pay tax} on your stipend.
-Finally, private universities might pay more for stipend, but they could have extra activity fee or some other hidden ones (e.g., you may need to pay some fees for each credit hour).
+Fourth, like most sources of income in the US, you will have to \emph{pay tax} on your stipend.
+Finally, private universities might pay more for stipend, but they could have extra activity fees or some other hidden ones (e.g., you may need to pay some fees for each credit hour).
 
@@
 However, in most cases it is actually not bad, and you can live comfortably with it.
-It might also be enough to support your spouse and kids (many CS PhD students have their family with them).
+It might also be enough to support your spouse and kids (many CS PhD students have their families with them).
 So don't worry too much about stipend.
@@
   TA and RA at GMU have similar benefits in tuition waiving and insurance.
-The college and department will set the rate for 9-month graduate assistant stipend.
-TA, which is paid by the department, will likely be that amount but RA might be higher depending on the stage of the student (1st year vs ABD\footnote{All but dissertation: really close to graduate.}) and the prof.
+The college and department will set the rate for a 9-month graduate assistant stipend.
+TA, which is paid by the department, will likely be that amount but RA might be higher depending on the stage of the student (1st year vs ABD\footnote{All but dissertation: close to graduate.}) and the prof.
   \tcblower
@@
 It is also common to sandwich between TA and RA (e.g., when your prof.
-does not have sufficient funding or you want to try get TA experience).
+does not have sufficient funding or you want to try the TA experience).
 
@@
 \begin{commentbox}[Vu:]
-  I would strongly advocate for a non-STEM student who shows that they have a strong drive for CS by studying core CS knowledge through various channels (e.g., self-study through online courses, projects, etc.).
+  I would advocate for a non-STEM student who shows that they have a strong drive for CS by studying core CS knowledge through various channels (e.g., self-study through online courses, projects, etc.).
 I have seen many students with non-CS background who are very successful in CS PhD.
@@
 
-If you have an MS then some course work \emph{might be} transferred for course credits, which save a bit of time. But overall don't count on this, especially if your MS is not from the US.
+If you have an MS then some course work \emph{might be} transferred for course credits, which saves a bit of time. But overall don't count on this, especially if your MS is not from the US.
 
@@
 The last 1-2 years you continue to publish, write and defend your dissertation, and look for job.
-In many cases you might take a summer or two off to do internship to get additional research opportunities.
+In many cases you might take a summer or two off to do an internship to get additional research opportunities.
 The PhDComics figure on top shows the ``ambition'' level of a PhD student over their years of study (they miss the 6--7th year when the ambition is \emph{``Just let me graduate''}).
@@
 
-\emph{MS requirement and PhD duration}:  as mentioned in \autoref{sec:msrequirement} and \autoref{sec:time}, CS PhD programs in the US do not require MS degree.
+\emph{MS requirement and PhD duration}:  as mentioned in \autoref{sec:msrequirement} and \autoref{sec:time}, CS PhD programs in the US do not require an MS degree.
 In contrast, many other countries require having an MS degree before joining a PhD program.
@@
 As you become more familiar with your prof and depending on their preferences, you may transition to addressing them by their first name.
-For example, I prefer that my students and colleague call me Vu. Some students call me \emph{Dr. Vu}, which I find a bit amusing but am totally fine with it.
+For example, I prefer that my students and colleagues call me Vu. Some students call me \emph{Dr. Vu}, which I find a bit amusing but am totally fine with it.
 
@@
 PhD students often ask why their salary is so low compared to ludicrous grants their advisers get.
-They also wonder why their offer letters sometimes show that their benefits higher than what they actually receive (i.e., stipend).
+They also wonder why their offer letters sometimes show that their benefits are higher than what they receive (i.e., stipend).
 
@@
 
-At the end, the total budget comes out to be \$72K/year to support a PhD student.
+In the end, the total budget comes out to be \$72K/year to support a PhD student.
 \emph{The summary is that over your 5--6 years of your PhD, you cost about \$400K, and while your stipend is X, your adviser probably pays 2X for you}.
-But of course the nicest thing is that you do not have to pay for any of this!
-You get to gain the knowledge, do research, travel, (and don't even have to buy any research or computer equipments as mentioned in \autoref{sec:buying-equipment}) and also get paid!
+But of course, the nicest thing is that you do not have to pay for any of this!
+You get to gain the knowledge, do research, travel, (and don't even have to buy any research or computer equipment as mentioned in \autoref{sec:buying-equipment}) and also get paid!
 
@@
 
-For start date or TA assignment (e.g., TA'ing a specific course), it is uncommon but you can ask for it.
+For start date or TA assignment (e.g., TA'ing a specific course), it is uncommon, but you can ask for it.
 Also, there is typically no moving allowance for PhD students.
@@
 \section{Will I be miserable during my a PhD?}\label{sec:happy}
-There are many stories on how students are mistreated, stressed, and miserable.
-Issues including bad relationships with professors, conflicts with co-authors and lab mates, feeling discriminated (e.g., because you're an international student), etc \emph{do} exist, and it is good to be aware of those.
-However, in reality there are many good mentors, supportive lab mates and department, and so on.
+There are many stories of how students are mistreated, stressed, and miserable.
+Issues including bad relationships with professors, conflicts with co-authors and lab mates, feeling discriminated against (e.g., because you're an international student), \emph{do} exist, and it is good to be aware of those.
+However, in reality, there are many good mentors, supportive lab mates and department, and so on.
 So don't let social media make you feel pessimistic and deter your quest to advance knowledge.
@@
 
-Most of what is written in this document is applicable to both domestic\footnote{For simplicity domestic means you did your undergrad (or MS) at a US university.} and international students.
+Most of what is written in this document applies to both domestic\footnote{For simplicity domestic means you did your undergrad (or MS) at a US university.} and international students.
 However, there are some differences and benefits that domestic students should be aware of.
 
-\paragraph{Standing out \autoref{sec:improve-your-chance}} There are \emph{few} domestic applications compared to international ones, i.e., domestic students are minority (\autoref{sec:urm}) in the CS PhD application pool.
+\paragraph{Standing out \autoref{sec:improve-your-chance}} There are \emph{few} domestic applications compared to international ones, i.e., domestic students are the minority (\autoref{sec:urm}) in the CS PhD application pool.
 Many US universities want to increase the number of domestic students in their programs (and as mentioned above, there are specific fellowships for domestic students).
@@
 
-\paragraph{Fee Waiving} Some school might offer application fee waiver for domestic students.  You should check with the school you're applying to.
+\paragraph{Fee Waiving} Some schools might offer application fee waivers for domestic students.  You should check with the school you're applying to.
 
@@
 
- This section provides some guidance on how to gain research experience, e.g., as an undergraduate student.
-It also gives advice to students from small colleges where research opportunities are limited.
-
-\paragraph{Locally} Start looking for research opportunities at your own institution.
+ This section provides some guidance on how to gain research experience, e.g., as an undergraduate student.  It also advises students from small colleges where research opportunities are limited.
+
+\paragraph{Locally} Start looking for research opportunities at your institution.
 If you did well or liked a class, you can check with the professor of that class for research opportunities.
@@
 
-\paragraph{Online Research Communities and Open Source Contributions} Places such as Github offer many opportunities for research, even though they might not appear so (e.g., they do not explicitly mention research).
-In many cases, professors or research labs put their projects on Github.
-For example, my research group has a many open-source projects that undergrads can contribute to (see \url{https://github.com/dynaroars/}).
+\paragraph{Online Research Communities and Open Source Contributions} Places such as GitHub offer many research opportunities, even though they might not appear so (e.g., they do not explicitly mention research).
+In many cases, professors or research labs put their projects on GitHub.
+For example, my research group has many open-source projects that undergrads can contribute to (see \url{https://github.com/dynaroars/}).
 
 By contributing code, fixing bugs, implementing new features, or providing documentation, students can gain practical research experience and interact with experienced developers and researchers in the field.
-Not only you gain research experience, you might be able to get a LoR from the project maintainer, and you should write about this experience in your SOP (\autoref{sec:research-statement}.)
+Not only you gain research experience, but you might be able to get a LoR from the project maintainer, and you should write about this experience in your SOP (\autoref{sec:research-statement}.)
 
@@
 
-\paragraph{Checking your application status and Follow-up emails} If you have interviewed and not heard back from a professor after a few weeks or especially around the time when universities send out their admission decisions (around late Feb-- mid Mar), you can email to check.
+\paragraph{Checking your application status and Follow-up emails} If you have interviewed and not heard back from a professor after a few weeks or especially around the time when universities send out their admission decisions (around late Feb-- mid-Mar), you can email to check.
 You can follow up the interview invitation and say: \emph{``Thanks for chatting with me.
@@
 Profs.
-are often very busy, especially during admission time when they have to review many applications, interview many students, and make decisions.
+are often very busy, especially during admission time when they have to many reviews and interviews.
 They might not have time to respond to every email.

### 2024-04-19 demystify.tex
Updates from Overleaf

@@
 
-\noindent and its \LaTeX{} source is also on \href{https://github.com/nguyenthanhvuh/phd-cs-us}{Github}.
+\noindent and its \LaTeX{} source is also on \href{https://github.com/nguyenthanhvuh/phd-cs-us}{GitHub}.
 If you have questions or comments, feel free to create new \href{https://github.com/nguyenthanhvuh/phd-cs-us/issues}{GitHub issues} or \href{https://github.com/nguyenthanhvuh/phd-cs-us/discussions}{discussions}.
@@
   One of the reasons I created this document is that my colleagues at GMU are interested in recruiting Vietnamese students and are surprised when seeing very few applications in Vietnam (see \autoref{sec:ack}).
-Each year our CS PhD program receives about 400 PhD applications, most of which are international but only 4--5 are from Vietnam.
+Each year our CS PhD program receives 500+ PhD applications, most of which are international but only 5--6 are from Vietnam.
 In general the number of CS PhD applications from Vietnam to US universities is very low and more would be welcome.
@@
 
-The document might also not sufficient for students aiming for top-10 schools, which might have different requirements\footnote{e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU}.
+The document might not be sufficient for students aiming for top-10 schools, which typically have different requirements\footnote{e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU}.
 However, the information here should be useful for students aiming for other good schools, e.g., R1 institutions in the US.

### 2024-04-19 demystify.tex
updates

@@
 
-The focus of the admissions committee is to \textbf{evaluate your background and interest in research} to see if you can handle a PhD in CS, which is a \emph{research} degree!.
-We also aim to determine if you would \textbf{fit into the program}.
-To evaluate your profile, we consider
-the following key indicators.
+The focus of the admissions committee is to evaluate your background and interest in research to see if you \textbf{fit} into our program and can handle a PhD in CS, which is a \emph{research} degree!
+To evaluate your profile, we consider the following key indicators.
 
@@
 \section{Why did I get rejected?}\label{sec:why-rejected}
+
+Many students lament that they have a good profile, but still got no interviews or rejected, and that the process seems random.
+However, while it is true that the process is not perfect, it is not random.
+The short answer is that you do not "fit" the program or the prof.
+Why?
+
 Many possible reasons!  Other than not having a strong profile (e.g., research potential, GPA, LoRs, SOP, poor interviews), here are some other reasons.
@@
 \paragraph{Not a good fit}  You could have an excellent profile, but if you are interested in a research area that the program does not have, you might not be admitted.
-Similarly, if there is no faculty who is willing to advise you (e.g., they are on sabbatical or personal leave, do not have sufficient funding, or already have too many students already), you will not be admitted.
+Similarly, if no faculty who is willing to advise you (e.g., they are on sabbatical or personal leave, do not have sufficient funding, or already have too many students already), you will not be admitted.
 This is actually good for you, as you don't want to be in a program where no one can advise you.

### 2024-04-19 demystify.tex
updates

@@
 Similarly, if no faculty who is willing to advise you (e.g., they are on sabbatical or personal leave, do not have sufficient funding, or already have too many students already), you will not be admitted.
-This is actually good for you, as you don't want to be in a program where no one can advise you.
+This is good for you, as you don't want to be in a program where no one can advise you.
 
@@
 
-\paragraph{Low English exam scores (e.g., IELTS or TOEFL)}  While professors might not care much about these, there might be a certain minimum set by the college or the university that you need to pass to get admitted, especially for TA funding (e.g., you need to communicate well with students as a TA).
+\paragraph{Low English exam scores (e.g., IELTS or TOEFL)}  While profs.
+might not care much about these, there might be a certain minimum set by the college or the university that you need to pass to get admitted, especially for TA funding (e.g., you need to communicate well with students as a TA).
 Thus, while professors are willing to argue for your case, they might be reluctant to go against the standard requirement of the university on English scores.

### 2024-04-19 demystify.tex
updates

@@
 
-Many students lament that they have a good profile, but still got no interviews or rejected, and that the process seems random.
-However, while it is true that the process is not perfect, it is not random.
-The short answer is that you do not "fit" the program or the prof.
-Why?
-
-Many possible reasons!  Other than not having a strong profile (e.g., research potential, GPA, LoRs, SOP, poor interviews), here are some other reasons.
+Many students lament that they get no interviews or are rejected (with strong profiles) and that the process seems random.  However, while it is true that the process is not perfect, it is not random.
+There are many possible reasons!  Other than not having a strong profile (e.g., research potential, GPA, LoRs, SOP, poor interviews), here are some other reasons.
 

### 2024-04-20 demystify.tex
updates

@@
 
-Many students lament that they get no interviews or are rejected (with strong profiles) and that the process seems random.  However, while it is true that the process is not perfect, it is not random.
-There are many possible reasons!  Other than not having a strong profile (e.g., research potential, GPA, LoRs, SOP, poor interviews), here are some other reasons.
+Many students lament that they get no interviews or are rejected (with strong profiles) and that the admission process seems random.
+However, while it is true that the process is not perfect, it is not random.
+There are many possible reasons for rejections!  Other than not having a strong profile (e.g., research potential, GPA, LoRs, SOP, poor interviews), here are some other reasons.
 
@@
 
-\paragraph{Red flags in Your Background} Plagiarism, cheating, or other academic dishonesty are \red{red} flags.
-Another one is if you have a history of jumping from one program to another (e.g., you have been in multiple different programs in a short time).
+\paragraph{Red flags} Various types of red flags can cause concern to the adcom (e.g., plagiarism, cheating, or other academic dishonesty).
+Another one is that you have a history of jumping from one program to another (e.g., you have been in multiple different programs in a short time).
 How would the reviewers know?
@@
 
-%\section{How exactly do you evaluate an application?}\label{sec:ievaluate}
-
-% Around 15 minutes on average (less than 10 mins for definite rejection; 20+ mins for acceptance). In earlier years I spent more time, but now I can quickly tell if an application is good or bad.
-
-% The evaluation system allows for downloading a single pdf that consists of a summary page (e.g., schools, GPA, standard test scores), transcripts, test results, LoRs, CV, SOP, and writing samples (if any).
+\section{How exactly do you evaluate an application?}\label{sec:ievaluate}
+
+Around 15 minutes on average (less than 10 mins for definite rejection; 20+ mins for acceptance). In earlier years I spent more time, but now I can quickly tell if an application is good or bad.
+
+The evaluation system allows for downloading a single pdf that consists of a summary page (e.g., schools, GPA, standard test scores), transcripts, test results, LoRs, CV, SOP, and writing samples (if any).
 I typically review the application in this order.
@@
 
-% I first skim the summary for school, GPA, standard tests and look for red flags like low GPA or test scores that do not meet the minimum requirements set by the university.
+I first skim the summary for school, GPA, standard tests and look for red flags like low GPA or test scores that do not meet the minimum requirements set by the university.
 Then I scan transcripts for low/failing grades on CS and related courses.
@@
 
-% Then I look at LoRs.
+Then I look at LoRs.
 I read carefully good LoRs and quickly skim over bad or generic ones (very easy to tell, you don't even need much experience to distinguish between strong LoRs from the rest).
@@
 
-% I skim over CV, looking for paper publications, research work experiences, and other relevant achievements such as winning prizes in well-known international programming/math competitions.
+I skim over CV, looking for paper publications, research work experiences, and other relevant achievements such as winning prizes in well-known international programming/math competitions.
 My notes for CV include things such as ``+ has X papers at top venues'' or ``+ gold medal at X''.
@@
 
-% Next is the SOP, which I read carefully (but if it is bad then it would be a quick read).
+Next is the SOP, which I read carefully (but if it is bad then it would be a quick read).
 My notes for SOP include things such as ``- SOP mostly on course work, no clear motivation, generic" , or ``+ very enjoyable SOP talking about XYZ, customized for GMU, mentioned profs.
@@
 
-% At this point I have enough to make up my decision, which I enter into the evaluation system.
+At this point I have enough to make up my decision, which I enter into the evaluation system.
 Typically my decisions are either \emph{reject} or \emph{admit with full funding}.
@@
 
-% The adcom chair will look at all evaluations for each application and make a decision (likely following the decision of the reviewers).
+The adcom chair will look at all evaluations for each application and make a decision (likely following the decision of the reviewers).
 If there are discrepancies such as one voting reject but another full funding acceptance (such cases are rare, but do happen), the chair might ask the reviewers to coordinate to reach a consensus.

### 2024-04-20 demystify.tex
updates

--- 2024-04-19 demystify.tex
@@
 For example, a prof.
-in Stats or Math dept might be able to serve as a co-adviser, but not as a sole adviser of a student in a CS PhD program.
+in Stats or Math might be able to serve as a co-adviser, but not as a sole adviser of a student in a CS PhD program.
 Similarly, a prof.
-in CS cannot be the sole adviser and graduate a PhD student in a Math or Stats.
+in CS cannot be the sole adviser and graduate a PhD student in Math or Stats.
 If in doubt, check with the CS department for their requirements.
@@
 A related reason is that you did not show enough interest in the program.
-For example, your SOP did not mention any faculty or research area at the program, you did not respond to professors interested in interviewing you (which might be considered unprofessional and burn bridges), or during the interview you show little knowledge, interest, or enthusiasm (because they are your safety schools) in the program or profs.
-
-\paragraph{Red flags in Your Background} Plagiarism, cheating, or other academic dishonesty are \red{red} flags.
-Another one is if you have a history of jumping from one program to another (e.g., you have been in multiple different programs in a short time).
+For example, your SOP did not mention any faculty or research area at the program, you did not respond to professors interested in interviewing you (which might be considered unprofessional and burn bridges), or during the interview you showed little knowledge, interest, or enthusiasm (because they are your safety schools) in the program or profs.
+
+\paragraph{Red flags} Various types of red flags can cause concern to the adcom (e.g., plagiarism, cheating, or other academic dishonesty).
+Another one is that you have a history of jumping from one program to another (e.g., you have been in multiple different programs in a short time).
 How would the reviewers know?
@@
 
-%\section{How exactly do you evaluate an application?}\label{sec:ievaluate}
+% \section{How long do you evaluate an application?}\label{sec:ievaluate}
+
+% On average, I spend about 15 minutes reviewing each application (less for clear rejections and more for potential acceptances).
+In earlier years I spent more time but now  I’ve become more efficient in evaluating..
+
+% The review system compiles a single PDF per applicant, including a summary page, transcripts, test scores, LoRs, a CV, SOP, and any writing samples.
+I review these documents in order, taking notes to support my evaluations.
+
+% I start with the summary, checking for low GPAs or test scores below the university minumum.
+I then skim the transcripts for low grades in relevant courses, noting issues like "- many low grades in main courses and - unknown international school with good GPA".
+
+% I pay close attention to strong LoRs and skim weaker ones, noting either strong endorsements from well-known professors or generic content.
+For the CV, I look for publications, research experiences, and notable achievements, noting things like "+ published papers in top venues" or "+ gold medal in an international competition."
+
+% The SOP is read thoroughly unless it's poorly written; a brief skim suffices in those cases. I note whether the SOP is engaging and well-tailored to our program or if it’s generic and uninspired.
+
+% Finally, I make a decision—usually either a direct rejection or an offer of admission with full funding, mentioning potential funding sources.
+Exceptionally strong candidates are recommended for the University Presidential fellowship.
+The admissions committee chair reviews all evaluations, typically following the recommendations, but discrepancies are resolved through discussion among reviewers to reach a consensus.
 

### 2024-04-20 demystify.tex
updates

@@
         \begin{itemize}
-          \item The committee will look at various factors, but the most important ones are letters of recommendation (LORs) and research ability, e.g., publications, statement of purpose (SOP).
+          \item The committee will look at various factors, but the most important ones are letters of recommendation (LORs),  statements of purpose (SOP), and research experience and ability, e.g., publications.
           \item \emph{LORs are very important}, but only if they are personalized and talk about your research ability (\autoref{sec:lor}).
@@
           \item Your entire PhD program costs about \$400K in total, but you \emph{do not} pay for it (\autoref{sec:ra-cost})
-          \item Ask your prof. if they can buy computer equipments and such for your research (\autoref{sec:buying-equipment}).
+          \item Ask your prof. if they can buy computer equipment and such for your research (\autoref{sec:buying-equipment}).
           \item Despite some miserable stories on social media, many PhD students have good mentors, supportive lab mates, healthy working environment ... and are happy (\autoref{sec:happy}).
@@
 VEF2.0 follows an interesting model where US faculty members from leading institutions are invited to conduct rigorous interviews with VEF students and subsequently provide reference letters on their behalf.
-Despite the limited interaction between the interviewer and the interviewee (primarily confined to the interview itself), these reference letters are generally effective and have helped many students getting into top PhD programs in the US.
+Despite the limited interaction between the interviewer and the interviewee (primarily confined to the interview itself), these reference letters are generally effective and have helped many students get into top PhD programs in the US.
 
@@
   convinced, then I will likely recommend rejection (unless I see
-  something really standout in other parts).
+  something really stands out in other parts).
   \tcblower
@@
 
-\paragraph{Publications} A concrete evidence of research ability is having \emph{papers in reputable international journals or conferences}.
+\paragraph{Publications} Concrete evidence of research ability is having \emph{papers in reputable international journals or conferences}.
 Having published papers, especially at top venues, is a sign that the applicant has been successfully involved in research.
@@
 However, these programs often set higher requirements for students, including more demanding tests and assignments, resulting in lower GPAs and overall rankings.
-For example, a 3.5 GPA students from such talented programs are typically much better than a 4.0 GPA students not in those programs.
+For example, 3.5 GPA students from such talented programs are typically much better than 4.0 GPA students not in those programs.
 Similarly, variations in GPA standards exist among different universities, with technical universities generally having lower GPAs than economical universities.
@@
   \textbf{Vu}: Vietnamese students and even faculty often lament how this grading system hurts Vietnamese students applying abroad.
-One way to mitigate this is making these issues known in your SOP.
+One way to mitigate this is to make these issues known in your SOP.
 \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{Universities with Vietnamese profs} are probably aware of them, but in general your letter writers \emph{and} you can explicitly mention these in their letters and your statement.
@@
 \paragraph{Online Courses and Certificates}
-In general, these do not carry much weight as they do not show research ability. We do not care much if you have taken on online Coursera AI course or have a professional certificate from Microsoft.
+In general, these do not carry much weight as they do not show research ability. We do not care much if you have taken an online Coursera AI course or have a professional certificate from Microsoft.
 However, as mentioned in \autoref{sec:non-stem}, if you do not have a CS background, you might be able to use these to show you have sufficient CS knowledge.
@@
 
-\textbf{Advice}: Prepare your CV in such away that it allows the reviewers to quickly scan to identify major achievements, e.g., Publications, Programming Competition Awards, and Teaching Experience.
+\textbf{Advice}: Prepare your CV in such a way that it allows the reviewers to quickly scan to identify major achievements, e.g., Publications, Programming Competition Awards, and Teaching Experience.
 
@@
 \paragraph{Timeline} The timeline for interviews and notification letters varies.
-Faculty setup interviews based on their (busy) schedule.
-Do not be surprised if you get an interview invitation in the last minute.
+Faculty set up interviews based on their (busy) schedule.
+Do not be surprised if you get an interview invitation at the last minute.
 Some schools do not do interviews at all.
@@
 \paragraph{Updating your profile} You should not send emails to update your profile.
-However, if you have new publications or other big achievements, you can ask to update your application (though no guarantee that they will consider them).
+However, if you have new publications or other big achievements, you can ask them to update your application (though no guarantee that they will consider them).
 
@@
 
-However, no interviews typically means that you will not likely get an RA (\autoref{sec:ra}), which is offered by an individual faculty (if they you to do research for them, then they likely will interview you first).
+However, no interviews typically mean that you will not likely get an RA (\autoref{sec:ra}), which is offered by an individual faculty (if they want you to do research for them, then they likely will interview you first).
 In the case of no interviews, your application (and TA/fellowship funding) is decided by the adcom.
@@
 During the event, you get a chance to learn more about the program, and talk to individual faculty and current students.
-Take notes of faculty who make you excited, and count those that are taking in new students (if they meet you, likely they are considering new students!).
-Talk to students about their advisers, the dept, the area, and funding situation.
+Take notes of faculty who make you excited, and count those taking in new students (if they meet you, likely they are considering new students!).
+Talk to students about their advisers, the dept, the area, and the funding situation.
 Ask about anything you want to determine that they deserve \emph{you}.
@@
 You can consider applying to MS programs, which are typically easier to get in; but you likely need to pay.  Sometimes, the university that rejected you for PhD might accept you for MS.
-If you get in an MS program at a school of your choice, you can contact professors to work with them. If you do well, you can ask the professor to support you to convert to PhD.
+If you get into an MS program at a school of your choice, you can contact professors to work with them. If you do well, you can ask the professor to support you to convert to PhD.
 
@@
 Fourth, like most sources of income in the US, you will have to \emph{pay tax} on your stipend.
-Finally, private universities might pay more for stipend, but they could have extra activity fees or some other hidden ones (e.g., you may need to pay some fees for each credit hour).
-
-\paragraph{Low Stipend?} Students often complain their stipend being too low.
-However, in most cases it is actually not bad, and you can live comfortably with it.
+Finally, private universities might pay more for stipends, but they could have extra activity fees or some other hidden ones (e.g., you may need to pay some fees for each credit hour).
+
+\paragraph{Low Stipend?} Students often complain their stipend is too low.
+However, in most cases it is not bad, and you can live comfortably with it.
 It might also be enough to support your spouse and kids (many CS PhD students have their families with them).
-So don't worry too much about stipend.
+So don't worry too much about the stipend.
 If you're admitted to a good CS PhD program, you will be fine.
@@
 In general, fellowships, especially those that are open to everyone, are highly competitive and prestigious, and you will stand out if you get one.
-Every PhD student has pubs, but only a few would have NSF Graduate (GRFP) or Microsoft fellowships .
-In fact, these are so prestigious that even if you didn't get it but make it to the final round or even \emph{``honorable mentioning''}, school will still mention you on their website, and you should put it on your CV.
+Every PhD student has pubs, but only a few would have NSF Graduate (GRFP) or Microsoft fellowships.
+In fact, these are so prestigious that even if you didn't get it but make it to the final round or even \emph{``honorable mentioning''}, you school will still mention you on their website, and you should put it on your CV.
 
@@
 
-Most US universities have CS study, but many of them do not have a PhD program in CS.
+Most US universities have CS studies, but many of them do not have a PhD program in CS.
 These universities might offer just Bachelor's degrees (e.g., BS) and no graduate studies (i.e., no MS or PhD degrees), or they just offer MS programs (but no PhD).
@@
 \begin{commentbox}[Dat:] Most Vietnamese students, including those from top schools, \textbf{do not know} about CSRankings.
-May be applicants who worked at top research places such as VinAI would know about it.
+Maybe applicants who worked at top research places such as VinAI would know about it.
 \end{commentbox}
@@
 You might prefer places with high-tech companies (Seattle or Silicon Valley), with many outdoor activities (e.g., hiking, skiing), or with better weather (\emph{``PhD can be depressing, so would you rather be depressed in California or New York?''}).
-You might need to think about the cost of living, e.g., living in California is more expensive than in Virginia, or places with high crime rates (note that while some university might be in a high crime city, but the campus itself is safe).
+You might need to think about the cost of living, e.g., living in California is more expensive than in Virginia, or places with high crime rates (note that while some universities might be in a high-crime city but the campus itself is safe).
 
@@
   I think PhD is not only about research;
-  you need to also consider your daily life there, since you will (probably) stay for at least five years.
+  you need to also consider your daily life there since you will (probably) stay for at least five years.
   %
@@
 \section{Choosing an Adviser}
-There is no one-size-fit-all answer to this question. The best adviser is the one that you can work well with.  But how do you find such a person?
+There is no one-size-fits-all answer to this question. The best adviser is the one that you can work well with.  But how do you find such a person?
 
@@
 Profs teaching \emph{special topics courses} and \emph{research seminars} might be recruiting students---giving you even a higher chance.
-Do well in the class, answer questions, talk to the prof after classes, etc---being stand out.
+Do well in the class, answer questions, talk to the prof.
+after classes, etc---being stand out.
 Many profs, including myself, prefer taking in new students this way.
@@
 For example, you might hear about tenured,  tenure track, and teaching faculty.
-You might also hear about assistant, associate, full, adjunct, emeritus, teaching, research professors.
+You might also hear about assistant, associate, full, adjunct, emeritus, teaching, and research professors.
 Here is a primer on these terminologies.
@@
 
-\textbf{Tenure-line faculty}, consisting of tenured and tenured-track profs., focus on research, which includes publishing papers, obtaining grants, and mentoring PhD students.
+\textbf{Tenure-line faculty}, consisting of tenured and tenure-track profs., focus on research, which includes publishing papers, obtaining grants, and mentoring PhD students.
 They often have a low teaching load (e.g., 1 per semester).
-\emph{Tenured} faculty are professors who have been promoted to a permanent position (informally, very hard to fire them).
+\emph{Tenured} faculty are professors who have been promoted to permanent positions (informally, very hard to fire them).
 \emph{Tenure-track} faculty are (more likely young) professors who are on the track to get tenure.
@@
 As such they typically do not have funding (e.g., RA~\autoref{sec:funding}) to support PhD students.
-Depends on the university, they might be able to \emph{co-advise} and mentor PhD students with a tenure-line faculty, but typically do not formally advise and graduate PhD students.
+Depending on the university, they might be able to \emph{co-advise} and mentor PhD students with a tenure-line faculty but typically do not formally advise and graduate PhD students.
 
@@
 They can recruit students and provide RA funding for them. 
-They can also teach classes, but typically have a lower teaching load.
+They can also teach classes but typically have a lower teaching load.
 For some universities, they \emph{can} solely advise and graduate PhD students, but for others, they cannot.
@@
 
-\paragraph{Ranks} Faculty have ranks such as assistant, associate, full, regardless if they are tenure-line or teaching.
+\paragraph{Ranks} Faculty have ranks such as assistant, associate, and full, regardless if they are tenure-line or teaching.
 Assistant means new faculty, associate means they have been promoted, and full means they are senior.
 A tenure track faculty starts with being an assistant professor and then gets tenure and promoted to associate (typically after 6 years) and then full (time varies, some become full within 3-4 years, some 10+ years, some remain associate).
-A non-tenure track faculty, e.g., teaching or research faculty, also starts with assistant, then promoted to associated, and full prof.
+A non-tenure track faculty, e.g., teaching or research faculty, also starts with assistant, then is promoted to associated, and full prof.
 
@@
 Faculty often receive ``cold'' e-mails from prospective students.
-Most of the time, we ignore these emails, but in some rare occasions we do answer them.
+Most of the time, we ignore these emails, but on some rare occasions, we do answer them.
 So how to write an email that gets our attention?
@@
 We don't know and can't answer because as explained in \autoref{sec:evalapps}, we don't make individual decisions and might not even be assigned to evaluate your application.
-It is the same as sending a paper draft to a journal editor and ask them if your paper has a chance.
+It is the same as sending a paper draft to a journal editor and asking them if your paper has a chance.
 
@@
 You could ask your professors, LoR writers, collaborators, or those who have previously applied.
-For this kind of feedback, ask someone you have personal connection with.
+For this kind of feedback, ask someone you have a personal connection with.
 
@@
 
-First, read the prof's website, see if they say something about contacting them.
+First, read the prof's website, and see if they say something about contacting them.
 Many profs.
@@
 
-  This is a good example because it is clearly written just for me.  It shows that the student knows about my work on invariant generation and has related  background (paper C and project G).
+  This is a good example because it is written just for me.  It shows that the student knows about my work on invariant generation and has a related background (paper C and project G).
 \end{commentbox}

### 2024-04-20 demystify.tex
updates

@@
 
-The focus of the admissions committee is to evaluate your background and interest in research to see if you \textbf{fit} into our program and can handle a PhD in CS, which is a \emph{research} degree!
+The focus of adcom is to evaluate your background and interest in research to see if you \textbf{fit} into our program and can handle a PhD in CS, which is a \emph{research} degree!
 To evaluate your profile, we consider the following key indicators.
@@
 \begin{commentbox}[Vu:]
-  If your grading system is not US standard or you are from a good school unknown outside your country, you can ask your reference writers to explain that in their letters.
+  If your grading system is not US standard, or you are from a good school unknown outside your country, you can ask your reference writers to explain that in their letters.
 For example, ``Bach Khoa'' are the top universities in Vietnam for STEM studies but few people outside Vietnam know about them.
@@
 
-\paragraph{How to help your LOR writer} Your LoR writer can ask you to provide additional information to help them write your letter.
+\paragraph{How to help your LOR writer} Your writer can ask you to provide additional information to help them write your letter.
 For example, they might ask about GPA, research and work experiences, papers (if any), or anything you want them to mention.
@@
 
-There are various guides on writing SOP , and \href{https://cs-sop.org/}{many example statements} are available.
-In short, focus on your research motivation and convince us that you can achieve it through your experience, e.g., published papers.
+In your SOP, focus on research motivation and convince us that you can achieve it through your experience, e.g., published papers.
 You should also back up your claims with concrete evidence.
@@
 If you say you work on project X, then show some results (paper submitted (or even rejected), achieved X\% performance improvement over ..).
-Also, do not use more than 2 pages (you can have references on the third page).
+Also, do not use more than two pages (references can be on the third page).
+\href{https://cs-sop.org/}{Many SOP samples} are available.
 
@@
 
-This is something easy to do, but is neglected by many
-applicants: \textbf{customize the statement} for the school you're applying to,
+This is a simple task often overlooked by many applicants: \textbf{tailor your SOP} to the institution you're applying for,
 e.g., why do you apply \emph{here}?
-provide names of professors who you're interested in (if they are not already in the adcom, your application might get forwarded to them for evaluation; and they might be interested in interviewing and recruiting you).
+Provide names of professors who you're interested in (if they are not already in the adcom, your application might get forwarded to them for evaluation; and they might be interested in interviewing and recruiting you).
 This shows that you're serious and have done homework on places you're applying to.
-Adcom will look for this part (\autoref{sec:why-rejected}), so do not skip it.
+Adcom will look for this part (\autoref{sec:why-rejected}).
 

### 2024-04-21 demystify.tex
updates

@@
 
-The document might not be sufficient for students aiming for top-10 schools, which typically have different requirements\footnote{e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU}.
-However, the information here should be useful for students aiming for other good schools, e.g., R1 institutions in the US.
-It is definitely useful for students aiming for GMU :)
+While this document may not fully meet the needs of students aiming for top-10 schools, which typically have more stringent requirements (e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU), it should be valuable for those targeting other reputable institutions, such as R1 universities in the US.
+It is certainly useful for students aiming for GMU!
 

### 2024-04-21 demystify.tex
updates

@@
 Again, the emphasis here is \emph{research}, i.e., the letters should describe your research experiences and potential.
-Letters focusing on your course work or non-research projects at a company won't carry much weight.
+Letters focusing on your coursework or non-research projects at a company won't carry much weight.
 
@@
 TA is paid by the department for you to help with teaching.
-Finally, fellowship is an independent funding that can come from a school, a company, or an organization.
+Finally, fellowship is independent funding that can come from a school, a company, or an organization.
 \autoref{tab:funding} summarizes the differences.
@@
 This can be longer than CS PhDs in other countries, which might require MS first (recall that CS PhD programs in the US do not require MS and you can get MS along the way to PhD).
-Within these 5--7 years, CS PhD students often take a ``leave of absence'' for 1--2 semesters to do internship at companies and research labs.
-
-The first 2 years you typically take coursework (somewhat equivalent to an MS study), find an adviser, learn how to do research.
-The next 2--3 years you focus on your research, form dissertation topic, and get results published.
-The last 1-2 years you continue to publish, write and defend your dissertation, and look for job.
+Within these 5--7 years, CS PhD students often take a ``leave of absence'' for 1--2 semesters to do internships at companies and research labs.
+
+The first 2 years you typically take coursework (somewhat equivalent to an MS study), find an adviser, and learn how to do research.
+The next 2--3 years you focus on your research, form a dissertation topic, and get results published.
+The last 1-2 years you continue to publish, write and defend your dissertation, and look for a job.
 In many cases you might take a summer or two off to do an internship to get additional research opportunities.
@@
 \emph{Course work}: as mentioned above in the US you will spend the first couple of years taking classes and exploring potential adviser and research topics. 
-After that, you also have to pass a series of exams during your PhD, e.g., qualifying exam, comprehensive exam, thesis proposal defense\footnote{The word ``ABD'' (all but dissertation) is used in the US to refer to a PhD candidate who have finished all course work and exams and only need to write and defend their dissertation.}.
+After that, you also have to pass a series of exams during your PhD, e.g., qualifying exam, comprehensive exam, thesis proposal defense\footnote{The word ``ABD'' (all but dissertation) is used in the US to refer to a PhD candidate who has finished all course work and exams and only needs to write and defend their dissertation.}.
 
@@
 
-PhD students often ask why their salary is so low compared to ludicrous grants their advisers get.
+PhD students often ask why their salary is so low compared to the large grants their advisers get.
 They also wonder why their offer letters sometimes show that their benefits are higher than what they receive (i.e., stipend).
@@
 Issues including bad relationships with professors, conflicts with co-authors and lab mates, feeling discriminated against (e.g., because you're an international student), \emph{do} exist, and it is good to be aware of those.
-However, in reality, there are many good mentors, supportive lab mates and department, and so on.
+However, in reality, there are many good mentors, supportive lab mates and departments, and so on.
 So don't let social media make you feel pessimistic and deter your quest to advance knowledge.
@@
 You can also apply to specific programs \emph{before} you start your PhD, e.g., NSF Graduate Research Fellowship Program (GRFP) and Hertz Foundation Fellowship.
-These fellowships are very competitive and can significantly improve your admission chance if you get them.
+These fellowships are very competitive and can significantly improve your admission chances if you get them.
 

### 2024-04-26 demystify.tex
updates

@@
 
-\paragraph{Interviews} Just like a job application, you likely will get interviews from schools.
+\subsection{Interviews}
+
+Just like a job application, you likely will get interviews from schools.
 The most common case is that a prof.
-is interested in working with you and wants to chat to make a decision, e.g., to offer RA (\autoref{sec:ra})).
+is interested in working with you and wants to chat to make a decision, e.g., to offer RA (\autoref{sec:ra}).
 In some cases, the interview is done by several professors, e.g., to see if a student fits in their group or to recruit a very strong student to their program.
@@
 
-\paragraph{Timeline} The timeline for interviews and notification letters varies.
+\subsection{Timeline}
+
+The timeline for interviews and notification letters varies.
 Faculty set up interviews based on their (busy) schedule.

### 2024-04-26 demystify.tex
updates

@@
 
-For stipend, unlikely for TA, which is set by the university.
-Moreover, while RA is set by the adviser, it is often based on the department's standard, and it also would be unfair to other students if you get more.
-However, your stipend is often increased every year (automatically by a small amount) and also after you pass your qualifier exams.
-
-For start date or TA assignment (e.g., TA'ing a specific course), it is uncommon, but you can ask for it.
+It is unlike that you can negotiate TA stipend, which is set by the university.
+Moreover, while RA is set by the adviser, it is often based on the department's standard, and it also would be unfair to other students if you get more (especially since you're new and have not proven yourself yet).
+However, your stipend is often increased every year by the university (automatically by a small amount).
+
+For start date or TA assignment (e.g., TA'ing a specific course), you can ask for it.
 Also, there is typically no moving allowance for PhD students.
 In short, standard things set by the university or department are unlikely non-negotiable.
-However, you can ask for things such as computer equipment (\autoref{sec:buying-equipment}), books, (and sometimes even furniture), and a conference travel budget.
-
-\section{Will I be miserable during my a PhD?}\label{sec:happy}
-There are many stories of how students are mistreated, stressed, and miserable.
+However, you can ask for things such as computer equipment (\autoref{sec:buying-equipment}), books, (and maybe even furniture), and a conference travel budget.
+
+% \section{Will I be miserable during my a PhD?}\label{sec:happy}
+% There are many stories of how students are mistreated, stressed, and miserable.
 Issues including bad relationships with professors, conflicts with co-authors and lab mates, feeling discriminated against (e.g., because you're an international student), \emph{do} exist, and it is good to be aware of those.

### 2024-04-27 demystify.tex
updates

@@
 
-The focus of adcom is to evaluate your background and interest in research to see if you \textbf{fit} into our program and can handle a PhD in CS, which is a \emph{research} degree!
-To evaluate your profile, we consider the following key indicators.
+The focus of adcom is to evaluate your background and interest in research to see if you \textbf{fit into our PhD program}!
+The emphasis here is \emph{fitting}, which varies from school to school, faculty to faculty, and even from year to year.
+However, in general, to evaluate your profile, we consider the following key indicators.
 
@@
 
-\paragraph{VEF 2.0} For Vietnamese students, it's worth mentioning the \href{https://vef2.org/}{VEF2.0 program}, which has helped many good students in gaining admission to top CS PhD programs.
-VEF2.0 follows an interesting model where US faculty members from leading institutions are invited to conduct rigorous interviews with VEF students and subsequently provide reference letters on their behalf.
-Despite the limited interaction between the interviewer and the interviewee (primarily confined to the interview itself), these reference letters are generally effective and have helped many students get into top PhD programs in the US.
+\paragraph{VEF 2.0} For Vietnamese students, it's worth mentioning the \href{https://vef2.org/}{VEF2.0 program}, which has helped many students in gaining admission to top CS PhD programs.
+VEF2.0 invites US faculty members from leading institutions to conduct rigorous interviews with VEF students and subsequently provide reference letters on their behalf.
+Despite the limited interaction between the interviewer and the interviewee (primarily just the interview), these reference letters are generally effective and have helped many students get into top PhD programs in the US.
 
@@
 Also, do not use more than two pages (references can be on the third page).
-\href{https://cs-sop.org/}{Many SOP samples} are available.
+Many SOP samples are \href{https://cs-sop.org/}{available here}.
 
@@
 
-In short, you \emph{do not need} to formally taking CS courses, you just need to show that you have these essential knowledge, e.g., through the mentioned ways.
+In short, you \emph{do not need} to formally taking CS courses, you just need to show that you have this essential knowledge, e.g., through the mentioned ways.
 Many universities are well-aware that incoming graduate students might not have all the technical background, so they often have a \emph{``bridge''} courses to help students catch up.
@@
 
-If you have an MS then some course work \emph{might be} transferred for course credits, which saves a bit of time. But overall don't count on this, especially if your MS is not from the US.
+If you have an MS then some course work \emph{might be} transferred for course credits, which saves a bit of time. But don't count on this, especially if your MS is not from the US.
+
+\begin{commentbox}[Vu:]
+    I start my PhD with an MS in CS from a US university.
+I found that the MS helped me a lot in research, but not with course work and I had to retake many courses.
+In general, don't worry if you don't have an MS.
+But also don't feel that you wasted your time if you have an MS, as it can help you in research.
+  \end{commentbox}
 
@@
 
-The first 2 years you typically take coursework (somewhat equivalent to an MS study), find an adviser, and learn how to do research.
+The first two years you typically take coursework (somewhat equivalent to an MS study), find an adviser, and learn how to do research.
 The next 2--3 years you focus on your research, form a dissertation topic, and get results published.
@@
 The PhDComics figure on top shows the ``ambition'' level of a PhD student over their years of study (they miss the 6--7th year when the ambition is \emph{``Just let me graduate''}).
+
+\begin{commentbox}[Vu:]
+    I start my PhD with an MS, and it took me 7 years (Fall'2007-- Fall'14). I did spend half a year doing internship at Naval Research Lab and 1 year doing internship at Google
+  \end{commentbox}
 

### 2024-04-28 demystify.tex
updates

@@
 Second, an academic year (AY)  year is typically \emph{9-month} in the US, so the stipend is for 9 months (so divide by 9 for each month).
-Third, you might get paid over the summer through funding from your professor or fellowship (typically no TA over the summer).
+Third, you might get paid over the summer (\autoref{sec:summer-funding}) through funding from your professor or fellowship (typically no TA over the summer).
 Fourth, like most sources of income in the US, you will have to \emph{pay tax} on your stipend.
@@
 As a TA, you spend up to 20 hrs/week and help with classes (e.g., grading or teaching labs/recitation).
-During a semester, a TA might work with several courses and professors (not necessarily their adviser).  TA funding is \emph{not} typically available during the summer, which has few courses.
+During a semester, a TA might work with several courses and professors (not necessarily their adviser).
+TA funding is \emph{not} typically available during the summer (\autoref{sec:summer-funding}), which has few courses.
 
@@
 In fact, these are so prestigious that even if you didn't get it but make it to the final round or even \emph{``honorable mentioning''}, you school will still mention you on their website, and you should put it on your CV.
+
+\section{Funding In the Summer}\label{sec:summer-funding}
+
+One confusion students often have is whether they will get paid over the summer.  The answer is that it depends on the funding  source.
+First, remember that an academic year (AY), which consists of Fall and Spring semesters, is typically 9 months, so your stipend is for 9 months (and many places allow you to spread it over 12 months).
+
+If your funding source is TA (\autoref{sec:ta}), you typically do not get paid over the summer as there are few courses.
+Some CS departments offer some summer funding, but it is typically not guaranteed and might not be a lot.
+For example, at GMU, we offer summer funding for 1st-year PhD TA students.
+The amount over the 3-month summer is similar to their monthly stipend (i.e., their 9-month stipend divided by 3 for the 3 summer months).
+
+For RA (\autoref{sec:ta}), it depends on your prof.
+and their funding.
+When writing grant proposals, profs.
+typically include summer funding for their students (\autoref{sec:ra-cost}).
+However, funding is never guaranteed, e.g., the prof might not get the grant.
+For my students, I have been fortunate to have funding to support them over the summer.
+Over the 3-summer months, I typically pay them 1/3 of their 9-month stipend.
+I prioritize summer funding for my students because GMU has very good TA resources so they never have to worry about funding during the AY.
+
+Finally, for fellowships (\autoref{sec:fellowships}) you might get paid over the summer depending on your fellowship (\autoref{sec:fellowships}).
+Typically good ones, e.g., from NSF, Google, Microsoft, will pay you the whole year.
 
@@
 No.
-In fact, student can get MS degree ``along the way'' to PhD, e.g., after finishing course work in the first 2 years.
+In fact, student can get MS degree ``along the way'' to PhD, e.g., after finishing the 2-year course work.
 However, MS can help admission if it gives research experience or is from a more well-known school than your undergrad institution (\autoref{sec:your-school}).
 
-If you have an MS then some course work \emph{might be} transferred for course credits, which saves a bit of time. But don't count on this, especially if your MS is not from the US.
+If you have an MS then some course work \emph{might be} transferred for course credits, which \emph{might} save some time. But don't count on this.
 
@@
     \midrule
-    GRA (9-month) & 29K & \\
+    GRA (9-month) & 30K & \\
     GRA (summer)  &10	& 3-month, 20hrs/week\\
-    \textbf{Total Salary} &\red{39K}	&\\
+    \textbf{Total Salary} &\red{40K}	&\\
     \midrule
@@
     \midrule
-    Total Direct Cost	& \red{60}	&Salary + Travel + Health + Tuition \\
+    Total Direct Cost	& \red{61K}	&Salary + Travel + Health + Tuition \\
     F \& A (MTDC)	& 21K	& Direct Cost - GRA Salary\\
     Total Indirect Cost	& \red{12K}	&58.9\% of MTDC\\
-    \textbf{Total (Direct + Indirect)} &	\red{72K}	& Budget for a GRA\\
+    \textbf{Total (Direct + Indirect)} &	\red{73K}	& Budget for a GRA\\
     \bottomrule
@@
 Private universities may have different numbers.
-For simplicity, I will assume the department has a 9-month stipend of \$29000 and a 3-month summer of \$10000 (about a third of the 9-month stipend).
+For simplicity, I will assume the department has a 9-month stipend of \$30000 and a 3-month summer of \$10000 (a third of the 9-month stipend).
 I will also use GMU tuition rate of about \$15,000/year for full-time study (which is quite cheap compared to private universities, e.g., Univ.
@@
 \paragraph{Standing out \autoref{sec:improve-your-chance}} There are \emph{few} domestic applications compared to international ones, i.e., domestic students are the minority (\autoref{sec:urm}) in the CS PhD application pool.
-Many US universities want to increase the number of domestic students in their programs (and as mentioned above, there are specific fellowships for domestic students).
+Many US universities want to increase the number of domestic students in their programs (and as mentioned later, there are specific fellowships for domestic students).
 That makes you stand out and can help your case.

### 2024-05-01 demystify.tex
updates

@@
           \item Increasing admission chance by being unique and standing out (\autoref{sec:improve-your-chance}).
-          \item Being in an underrepresented group can help can help (\autoref{sec:urm}).
+          \item Being in an underrepresented group can help (\autoref{sec:urm}).
           \item You can successfully apply to CS PhD even if you have non-STEM background (\autoref{sec:non-stem}).
@@
 \begin{commentbox}[Vu:]
-  At GMU CS, admitted PhD students have 4 years of GTA guaranteed, and also receive a stipend for the first summer.
+  At GMU CS, admitted PhD students have 4 years of GTA guaranteed, and also receive a stipend for the first summer (\autoref{sec:summer-funding}).
 \end{commentbox}
@@
 \section{Member of an underrepresented or LGBTQ+ groups}\label{sec:urm}
+
 A common question is whether you should mention that you are a member of an underrepresented and minority (URM) group\footnote{URM in CS in the US include African Americans, Hispanic Americans, American Indians/Native Alaskans, Native Hawaiians/Pacific Islanders (excluding Asian Americans), and multiracial students identifying with at least one of the previously listed URM categories.} or LGBTQ+ group (e.g., in your SOP).
@@
 
-It is understandable that you might feel uncomfortable disclosing this, fearing you would lost a chance of working with a faculty due to their bias.
-However, if someone has this issue, then you might not want to work with them anyway (and such bias is not acceptable in academic and likely would not be tolerated by the university).
+It is understandable that you might feel uncomfortable disclosing this, fearing you would lose a chance of working with a faculty due to their bias.
+However, if someone has this issue, then you might not want to work with them anyway (and such bias is not acceptable in academics and likely would not be tolerated by the university).
 In my experience, it is the opposite: \emph{many profs.
-actively seek out to work with students from diverse backgrounds and view diversity as an asset to their research group}.
+actively seek to work with students from diverse backgrounds and view diversity as an asset to their research group}.
 
 Moreover, promoting diversity aligns with the goal of broadening participation in computing (BPC) that is strongly supported within academia (e.g., see the \href{https://plans.bpcnet.org/GeorgeMasonUniversity_ComputerScience_DepartmentalBPCPlan.pdf}{BPC plan at GMU CS}) and government funding agencies such as NSF (e.g., see \href{https://new.nsf.gov/cise/broadening-participation}{here} and \href{https://www.nsf.gov/pubs/2022/nsf22125/nsf22125.jsp}{here}).
-Also, sometimes providing this information can make it easier for faculty to apply for certain scholarships or fellowships to support you, e.g., from a private donor that has offered a fellowship only to students from certain background.
+Also, sometimes providing this information can make it easier for faculty to apply for certain scholarships or fellowships to support you, e.g., from a private donor that has offered a fellowship only to students from certain backgrounds.
 
@@
 
-\paragraph{Core CS areas} Common CS areas that you should have knowledge in are:
+\paragraph{Core CS areas} Common CS areas that you should know are:
 \begin{itemize}
@@
 
-In short, you \emph{do not need} to formally taking CS courses, you just need to show that you have this essential knowledge, e.g., through the mentioned ways.
-Many universities are well-aware that incoming graduate students might not have all the technical background, so they often have a \emph{``bridge''} courses to help students catch up.
+In short, you \emph{do not need} to formally take CS courses, you just need to show that you have this essential knowledge, e.g., through the mentioned ways.
+Many universities are well aware that incoming graduate students might not have all the technical background, so they often have a \emph{``bridge''} courses to help students catch up.
 For example, GMU has four bridge courses corresponding to the four core areas above that incoming students can take to catch up on their CS knowledge.
@@
   I would advocate for a non-STEM student who shows that they have a strong drive for CS by studying core CS knowledge through various channels (e.g., self-study through online courses, projects, etc.).
-I have seen many students with non-CS background who are very successful in CS PhD.
+I have seen many students with non-CS backgrounds who are very successful in CS PhD.
 I also have seen many students with CS background who are not successful in CS PhD.
@@
 No.
-In fact, student can get MS degree ``along the way'' to PhD, e.g., after finishing the 2-year course work.
+In fact, students can get MS degree ``along the way'' to PhD, e.g., after finishing the 2-year course work.
 However, MS can help admission if it gives research experience or is from a more well-known school than your undergrad institution (\autoref{sec:your-school}).
@@
 
-\begin{commentbox}[Vu:]
-    I start my PhD with an MS, and it took me 7 years (Fall'2007-- Fall'14). I did spend half a year doing internship at Naval Research Lab and 1 year doing internship at Google
-  \end{commentbox}
+% \begin{commentbox}[Vu:]
+%     I start my PhD with an MS, and it took me 7 years (Fall'2007-- Fall'14). I spent half a year doing an internship at the Naval Research Lab.
+%   \end{commentbox}
 
@@
 
-\emph{Funding}:  In many countries funding comes from the university or the gov't.
+\emph{Funding}:  In many countries, funding comes from the university or the gov't.
 This funding often has a fixed duration (e.g., 3 or 4 years).

### 2024-05-01 demystify.tex
updates

@@
 Finally, for fellowships (\autoref{sec:fellowships}) you might get paid over the summer depending on your fellowship (\autoref{sec:fellowships}).
-Typically good ones, e.g., from NSF, Google, Microsoft, will pay you the whole year.
+Good ones, e.g., from NSF, Google, and Microsoft, will pay you the whole year.
 
@@
 Assistant means new faculty, associate means they have been promoted, and full means they are senior.
-A tenure track faculty starts with being an assistant professor and then gets tenure and promoted to associate (typically after 6 years) and then full (time varies, some become full within 3-4 years, some 10+ years, some remain associate).
+Typically, a tenure-track faculty starts with being an assistant professor and then gets tenure and promoted to associate (typically after 6 years) and then full (time varies, some become full within 3--4 years, some 10+ years, some remain associate).
 A non-tenure track faculty, e.g., teaching or research faculty, also starts with assistant, then is promoted to associated, and full prof.
 
-\subsection{Tenured or tenure-track faculty? Who do you choose?}\label{sec:tenure-vs-tenure-track}
+\subsection{Tenured or tenure-track faculty?}\label{sec:tenure-vs-tenure-track}
 
@@
 
-The short answer is that tenure-track faculty such as assistant professors are likely young and active in research (they have to, to get tenure).
+Tenure-track faculty such as assistant professors are likely young and very active in research (they have to, to get tenure).
 Thus, they will likely have more time to work with you and push you to do research and publish.
@@
 Tenured faculty, e.g., associate and full profs., are likely older, more well-known, and have more experience in managing students.
-However, they might not push you as hard.
-They might also expect you to figure things out yourself, i.e., you need to be independent.
+However, they might not push you as hard and expect you to figure things out yourself, i.e., you need to be independent.
 Some tenured faculty are also no longer active in research and are more involved with administrative responsibilities or with their startup companies (this means they will likely not take new students).
+\autoref{sec:faculty-types} gives more details on faculty types.
 
@@
 Many profs.
-explicitly indicate how prospective students should (or should not) contact them (e.g., using specific email subjects).
+explicitly state how prospective students should (or should not) contact them, e.g., using specific email subjects.
 In general, the best way to catch the prof.'s attention is to \emph{customize your email} for them.
@@
 A non-CS faculty can serve in the PhD dissertation committee (common) or \emph{co-advise} (somewhat rare), but your main PhD adviser will likely be a tenure-line faculty in CS (\autoref{sec:faculty-types}).
-For example, a prof.
-in Stats or Math might be able to serve as a co-adviser, but not as a sole adviser of a student in a CS PhD program.
-Similarly, a prof.
-in CS cannot be the sole adviser and graduate a PhD student in Math or Stats.
+For example, a prof. in Stats or Math might be able to serve as a co-adviser, but not as a sole adviser of a student in a CS PhD program. 
 If in doubt, check with the CS department for their requirements.
@@
 might not care much about these, there might be a certain minimum set by the college or the university that you need to pass to get admitted, especially for TA funding (e.g., you need to communicate well with students as a TA).
-Thus, while professors are willing to argue for your case, they might be reluctant to go against the standard requirement of the university on English scores.
-Note this is for international students, as domestic ones (or those from these \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-GMU#standard-tests-waiver-eligible-countries}{countries}) do not have to take these tests.
+Thus, while profs.
+are willing to argue for your case, they might be reluctant to go against the requirement of the university on standard exams.
+
+Note this is for international students, as domestic ones (and likely those from these \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-GMU#standard-tests-waiver-eligible-countries}{countries}) do not have to take these tests.
 
 \paragraph{Overqualified or Lack Interest}  This might be paradoxical and is the opposite of aiming too high.
-However if adcom believes that you are likely to get admitted and go to a better program (i.e., they are just your ``safety'' schools), they might not be so enthusiastic to admit you and want to save the spot for someone else.
+However if adcom believes that you are likely to get admitted and go to a better program, they might not be so enthusiastic to admit you and want to save the spot for someone else.
 
@@
 Understandably students get excited about their upcoming PhD journey and want to buy new computer equipment and electronics to prepare.
-However, you should check with your adviser first.
+However, you should check with your prof.
+first.
 They might have funding to buy you computers and other equipment (e.g., software, books, keyboards, headphones, tablets, etc).
-Many CS programs also provide some budget or computer equipment to incoming PhD students (e.g., a laptop).
+Many CS programs also provide budget or computer equipment to incoming PhD students (e.g., a laptop).
 

### 2024-05-23 faculty.tex
updates

@@
 Getting funding is arguably the most important part of getting tenure.
-Most new professors struggle with grant writing, which is very different than writing research papers.
-However, without funding, you can't pay your students, travel to conferences, or even pay your own salary.
-In short, the lack of funding cause lost of stress and anxiety, and threaten your tenure.
+Most new professors struggle with grant writing, which is very different compared to research papers.
+However, without funding, you can't pay your students, conference registration and travel, or even pay your own summer salary.
+In short, the lack of funding causes lost of stress and anxiety, and threaten your tenure.
 In this chapter, we will discuss the various ways to obtain funding.
@@
 For example, you can volunteer to be a chair of certain sessions/tracks of a large conference, or even a general chair of a workshop or conference.
-These roles are more time consuming, but they can be more rewarding (e.g., more people will know about you) and in general can help your career, especially your tenure case.
+These roles are more time-consuming, but they can be more rewarding (e.g., more people will know about you) and in general can help your career, especially your tenure case.
 
@@
 
-\subsection{Personal Experience} I have served on many PCs and NSF panels.
-I mostly cold email people for these opportunities.
+\subsection{Personal Experience} In my case I mostly cold email people for opportunities.
 For example, I directly email OOPSLA and PLDI PC chairs and asked, and they accepted my request (I did have papers at OOPSLA and PLDI when I asked).
 I tried this with SE conferences  e.g., ICSE (and also nominate myself through the official form), and it never works, even though I have many papers at ICSE.
-I also emailed NSF program managers and asked to serve on a panel, and they accepted my request.
-
-I felt that I didn't get much out of serving on PC and didn't really like it much. However, I still do it because I feel I should give back to the community.
+I also got invited to serve on a PC because a senior colleague recommended me.
+For NSF, I emailed program managers and asked to serve on their panels.
+
+I felt that I didn't get much out of serving on PCs and didn't really like the review process much. However, I still do it because I feel it is my responsibility give back to the community.
+
 On the other hand, I thoroughly enjoy serving on NSF panels (though of course my proposals still got rejected frequently).
 While the work is more intense and takes more time, I found it more rewarding and enjoy the discussions.
-Since my first NSF panel in 2019, I have served 7 panels (at least once every year since '19) and I look forward to serving more.
+Since my first NSF panel in 2019, I have served 8 panels (at least once every year since '19) and I look forward to serving more.
 
@@
 
-\subsection{Personal} Since coming to GMU, I have been very active in serving the department.
+A good department will try to match you with committees that can help junior faculty build their lab and tenure case (and shield you from committees that are not as useful).
+
+\subsection{Personal Experience} Since coming to GMU, I have been very active in serving the department.
 I share the vision of the department and want to help achieve it.
@@
 Even though I am not on the search committee, I meet with every faculty candidate (I talked to 22 people in Spring'23, which was my study break) and help recruit strong ones.
-I also agree to be director of the MS program, which is a lot of work and typically done by tenured faculty.
-
-Sometimes I feel overwhelmed by the services, but I also feel that I am making a difference in the department and that is rewarding.
+I also agree to be director of the MS SWE program, which is a lot of work and typically done by tenured faculty.
+
+Occasionally I feel overwhelmed by having many services, but I also feel that I am making a difference in the department and that is rewarding.
 I am also building a good relationship with my colleagues (and also head butt with some), and that is important for my tenure case.
-In addition, I learn a lot from being in various roles (e.g., how class scheduling work, how the department is run, how to recruit students and recruit faculty, how to run a colloquium, etc) and feel more confident in my leadership skills.
+In addition, I learn a lot from being in various roles (e.g., how class scheduling work, how the department is run, how to recruit students and recruit faculty, how to run a colloquium, etc).
 
@@
 
-\chapter{Tenure Package and Application}
-After 5 or 6 years, it is time for you to go up (or apply) for tenure.
+\chapter{Renewal and Tenure Application}
+
+Toward the end of your 3rd year, you will need to prepare a renewal package and submit it to your department.
+The department will then review your package and vote on whether to renew you for your 4th (and perhaps also 5th year).
+Your next renewal is essentially your tenure application.
+The department will vote on whether to recommend your renewal.
+
+Similarly, after 5 or 6 years, it is time for you to go up (or apply) for tenure.
 You will need to prepare a tenure package and submit it to your department.
 The department will then review your package and vote on whether to recommend you for tenure.
-If the department votes in your favor, the package will then go to the college, and then to the university.
+If the department votes in your favor, the package will then go to the college, and then to the university (e.g., provost, president, board of visitors, etc).
 The university will then make the final decision on whether to grant you tenure.
 
-This document shows my experience in preparing my tenure package and applying for tenure.
+%I will focus more on the tenure application process, as the renewal process is similar but less  intense.
+
+%are so important, you should start preparing for them early.
+You should also talk to your department chair and mentor about the process and what is expected of you.
+You should also talk to your colleagues who have gone through the process and ask for their advice.
+
+%This document shows my experience in preparing my tenure package and applying for tenure.
 I will share my timeline, the package (statements, letter writer sections), and the process.
@@
 \begin{itemize}
-    \item Fall 2021: Moved to GMU (I started at UNL in 2016)
-    \item (late) Fall 2022: I told my department chair that I plan to go up for tenure in 2023.
+    \item Fall 2021: Moved to GMU with accelerated tenure clock (meaning I don't have to restart from the beginning as I have already been at UNL since 2016).
+    \item (late) Fall 2022: I told my department chair that I plan to apply for tenure in 2023 (which means if things go well then I get tenured in Fall'24).
     \item Spring 2023: I started preparing my tenure package, which includes writing statements and think reference letters.
@@
     \item Fall 2024: I gave a tenure talk. Department voted in my favor, and then the college voted in my favor, and the decision goes up to the university.
-    \item Spring 2024: In March, I got the official notification that I got tenure and promoted to Associate Prof., beginning in Fall 2024.
+    \item Spring 2024: In May, I got the official notification that I got tenured and promoted to Associate Prof., beginning in Fall 2024.
 \end{itemize}
@@
 \subsection{Letters}
-This is a huge part of your tenure package.  You will need to ask for letters from senior people in your area.  Someone who can vouch for your research, teaching, and services.
-You will need 4-6 letters. However, you will not be the one asking for these letters, but your department will. You can suggest who to ask.
-
-Typically, there are two lists of letter writers: one consisting of people suggested by you and the other consisting of people suggested by the department (e.g., people who work in your field in your dept).
+This is a huge part of your tenure package.  You will need letters from senior people in your area.  Someone who can vouch for you.
+You will need 4--6 letters. However, you will not be the one asking for these letters, but your department will. You can suggest who to ask.
+
+Typically, there are \emph{two lists} of letter writers: one consisting of people suggested by you and the other consisting of people suggested by the department (i.e., your dept will come up with a list of people they think are good letter writers for you).
 There can also be another list, the ``blacklist'' where you list people that you don't want to write letters for you.
 
+\subsubsection{Personal Experience}
 Originally I asked several senior people to write my letters.
@@
 
+\paragraph{Tenure Visits} Many faculty will visit other universities to give talks and meet people.
+This is a good way to get your name out and to get letters.
+For example, if you plan to put person X on your list or you think person X will likely be selected by the department, you can ask X to visit their university and give a talk.
+
+\paragraph{Inviting speakers to colloquium} Many departments have colloquium series where they invite speakers to give talks.  You can invite potential letter writers to your department's colloquium.
+
+In general, these are a good way to get to know people in your field and to get your name out. Many of these, especially tenure visits, are common practices for junior faculty applying for tenure.
+
+\section{Renewal Package}
+
 \appendix

### 2024-05-23 demystify.tex
Updates from Overleaf

@@
 
-The information presented here can be useful for all students, but this guideline is primarily intended for \textbf{international students} from \textbf{smaller countries and less well-known universities} (similarly, for domestic students from universities with no PhD programs or limited research experiences).
+The guideline can be useful for all students, but it is primarily intended for \textbf{international students} from \textbf{smaller countries and less well-known universities} (similarly, for domestic students from universities with no PhD programs or limited research experiences).
 Students from top schools with strong research cultures and experiences might already know most of the information presented here.
@@
 
-While this document may not fully meet the needs of students aiming for top-10 schools, which typically have more stringent requirements (e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU), it should be valuable for those targeting other reputable institutions, such as R1 universities in the US.
+While this document may not fully meet the needs of students aiming for top-10 schools, which typically have more stringent requirements (e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU), it should be valuable for those targeting other R1 reputable institutions.
 It is certainly useful for students aiming for GMU!

### 2024-06-07 demystify.tex
updates

@@
 
-By around March you should hear back from most PhD programs you applied to. If you haven't heard back, reach out via email and ask about the status of your application.
+By around March you should hear back from most PhD programs that want to admit you.  In many cases you will not hear back from schools that do not want to admit you.
 If you receive offers, congratulations!
@@
 \paragraph{Open House} Most schools have \emph{Open House} or \emph{Visit Day} events, which are a great resource to learn about the school, department, faculty, research, living, etc.
-\textbf{Advice}: Even if you can't come in person, you should attend virtually and meet with individual faculty.
+
+Even if you can't come in person, you should attend virtually and meet with individual faculty.
 During the event, you get a chance to learn more about the program, and talk to individual faculty and current students.
@@
 \epigraph{Bart, with \$10,000, we’d be millionaires! We could buy all kinds of useful things like…love!’}{\textsc{The Simpsons}}
-
-As mentioned, if you're admitted to a \emph{good} CS PhD program, you should not have to worry about funding!
+If you're admitted to a \emph{good} CS PhD program, you should not have to worry about funding!
 In the US, the common types of funding for PhDs are \emph{graduate teaching assistant} (GTA or TA), \emph{graduate research assistant} (GRA or RA), and \emph{Fellowship}.
@@
 for you to do their research.
-TA is paid by the department for you to help with teaching.
+TA is paid by the dept.
+for you to help with teaching.
 Finally, fellowship is independent funding that can come from a school, a company, or an organization.
@@
 Note that funding is typically more available for PhD students than
-Masters.
+MS.
 
@@
 Several things about stipends.
-First, the amount of stipend \emph{varies} and depends on factors such as location (e.g., a stipend in Washington DC is likely higher than in Lincoln, Nebraska due to higher living costs).
+First, the amount of stipend \emph{varies} and depends on factors such as location (e.g., a stipend in the DC area is likely higher than in Lincoln, Nebraska due to higher living costs).
 Second, an academic year (AY)  year is typically \emph{9-month} in the US, so the stipend is for 9 months (so divide by 9 for each month).

### 2024-06-07 faculty.tex
updates

@@
 Most new professors struggle with grant writing, which is very different compared to research papers.
-However, without funding, you can't pay your students, conference registration and travel, or even pay your own summer salary.
-In short, the lack of funding causes lost of stress and anxiety, and threaten your tenure.
+However, without funding, you can't pay your students, conference registration and travel, or even your own summer salary.
+In short, the lack of funding causes lost of stress, uncertainty, and threaten your career.
 In this chapter, we will discuss the various ways to obtain funding.
@@
 Typically you will first mention to your department chair that you are planning to go up for tenure.
-The department chair then assign you a mentor who will help you with the process.
-You will apply for tenure in the Fall semester, and the decision will be made in the Spring semester.
-
-My own timeline is as follows:
+The chair might then assign you a mentor who will help you with the process.
+You will apply for tenure in the Fall semester, the decision will be made in the next Spring semester, and you will get tenure in the Fall semester after that.
+
+My timeline is as follows:
 
@@
   \item Fall 2021: Moved to GMU with accelerated tenure clock (meaning I don't have to restart from the beginning as I have already been at UNL since 2016).
-    \item (late) Fall 2022: I told my department chair that I plan to apply for tenure in 2023 (which means if things go well then I get tenured in Fall'24).
-    \item Spring 2023: I started preparing my tenure package, which includes writing statements and think reference letters.
-I also took a semester off because GMU has a pre-tenure study break policy.
+  \item (late) Fall 2022: I told my department chair that I plan to apply for tenure in 2023 (which means if things go well then I get tenure in Fall'24).
+  \item Spring 2023: I started preparing my tenure package, which includes writing statements and think reference letters.  I also took a semester off because GMU has a pre-tenure study break policy.
   \item Summer 2023: I submitted my tenure package.
   \item Fall 2024: I gave a tenure talk. Department voted in my favor, and then the college voted in my favor, and the decision goes up to the university.
-    \item Spring 2024: In May, I got the official notification that I got tenured and promoted to Associate Prof., beginning in Fall 2024.
+  \item Spring 2024: In May, I got the official notification that I got tenure and promoted to Associate Prof., beginning in Fall 2024.
 \end{itemize}
@@
 
+\subsection{CV}
+
+GMU (and likely other universities) has a very strict and specific format for the CV (e.g., what to include, how to format, etc).
+Fortunately I have been using this format since I joined GMU and also kept my CV up-to-date, so I didn't have to do much.
+I did have to add a few more sections (e.g., teaching evaluation and number of students in courses taught).
+I also have to make sure that the CV is consistent with my statements and other parts of the package (e.g., funding amount must match with what the university has in their system).
+
+I also looked at CVs of other people at GMU who recently got tenure. I also sent my CV to  my mentors and got some feedback.
+
 \subsection{Statements}
 
-\subsection{CV}
+Here is my \href{https://dynaroars.github.io/people/nguyenthanhvuh/files/pt/statement-annotated.pdf}{tenure application statement}, annotated with comments.
+At GMU, the statement is limited to 8 pages.
+My statement follows the standard structure and talks about my achievements and plans in (i) research and funding, (ii) advising and mentoring, (iii) teaching, and (iv) services.
+The main unique part of my statement is the first page where I highlight and summarize my achievements (e.g., papers, fundings, students, etc).
+I get most of these information and numbers from my CV.
+
+While preparing the statement was the most time-consuming part of my tenure package, it didn't take as long (may be just a couple of days).
+I get statements from several colleague at GMU who recently got tenure and a couple of people at other universities to get some ideas on how it should be structured.
+I was able to reuse lots of text and contents from my job application statements (when I applied to GMU in 2021).
+I did not send it for feedback and also did not revise it much.
 
@@
 \paragraph{Tenure Visits} Many faculty will visit other universities to give talks and meet people.
-This is a good way to get your name out and to get letters.
-For example, if you plan to put person X on your list or you think person X will likely be selected by the department, you can ask X to visit their university and give a talk.
+This is a good way to get your name out and to get letters (e.g., from the person that host you).
+Physical visits were not possible during the pandemic, but virtual visits are possible and in fact easier.
+However,  I did not do any tenure visits or give talks in the year leading up to my tenure application.
+This was because I am not a travel person and also I felt that it would not help much in my case.
 

### 2024-06-13 demystify.tex
updates

@@
 Originally, this document was intended for international students, but I have expanded it to include information that might also be useful for \emph{domestic students}.
-Moreover, while this handbook is primarily intended for students interested in CS, it might be relevant to students from various disciplines.
+Moreover, while this handbook is primarily intended for students interested in CS, it might be relevant to students from various STEM (Science, Technologies, Engineering, and Mathematics) disciplines.
 Furthermore, although many examples are specifics for schools that I and other contributors of this document know about, the information should be generalizable to other R1\footnote{An \href{https://en.wikipedia.org/wiki/List_of_research_universities_in_the_United_States}{R1 institution} in the US is a research-intensive university with a high level of research activity across various disciplines.
@@
 
-\textbf{Advice}: If you haven't taken the GRE or have bad scores, then don't waste time (re)taking it.
-However, if you took it and have really good scores then you can include and talk about them in your application (but don't expect them to make much difference).
+\textbf{Advice}: If you haven't taken the GRE or have bad scores, then don't waste time (re)taking it. It does not help much!
+
+However, if you took it and have really good scores then it might be worth it to include and talk about them in your application (but don't expect them to make much difference).
 But if your scores are bad, then you should not include them in your application, which can be a \red{red flag}.
 
-\paragraph{English Tests} Unless your degrees are from specific countries such as \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-GMU#standard-tests-waiver-eligible-countries}{these}, you will need to
+\paragraph{English Tests (IELTS, TOEFL)} Unless your degrees are from specific countries such as \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-GMU#standard-tests-waiver-eligible-countries}{these}, you will need to
 take standardized English tests.
@@
 \section{CV/Resume}
-The CV/Resume provides a summary of the applicant's achievements.
+The CV/Resume provides a summary of the applicant's achievements. Most schools require you to upload your CV with your application.
 
@@
 Many programs do not have the tradition of interviewing applicants.
-For example, at GMU, most of our admitted students with TA (\autoref{sec:ta}) do not go through interviews.
+For example, at GMU, most admitted students with TA (\autoref{sec:ta}) do not go through interviews.
 
@@
 
-Not every faculty can serve as your official PhD adviser.
+\emph{Not} every faculty can serve as your official PhD adviser.
 Let's try to understand the different types of faculty and their roles.
@@
 
-\paragraph{Types of Faculty} There are typically three types of faculty: tenure-line, teaching, and research.
+\paragraph{Types of Faculty} There are typically two types of faculty: tenure-line and teaching.
 The main difference is their focus: research, teaching, or both.
@@
 
-Note that non-CS profs (even in tenure-line ones) often cannot solely advise and graduate PhD students (see \autoref{sec:related-fields}).
+Note that at many universities do not allow \emph{non-CS profs}, even in tenure-line ones, to solely advise and graduate PhD students in CS (see \autoref{sec:related-fields}).
 
@@
 
-A common question is whether you should mention that you are a member of an underrepresented and minority (URM) group\footnote{URM in CS in the US include African Americans, Hispanic Americans, American Indians/Native Alaskans, Native Hawaiians/Pacific Islanders (excluding Asian Americans), and multiracial students identifying with at least one of the previously listed URM categories.} or LGBTQ+ group (e.g., in your SOP).
+A common question is whether you should mention, e.g., in your SOP, that you are a member of an underrepresented and minority (URM) group\footnote{URM in CS in the US include African Americans, Hispanic Americans, American Indians/Native Alaskans, Native Hawaiians/Pacific Islanders (excluding Asian Americans), and multiracial students identifying with at least one of the previously listed URM categories.} or LGBTQ+ group.
 This is a strong concern for international (and even domestic) students, who may have encountered discrimination in their countries or personal experiences.
@@
 
-In my personal opinion, highlighting URM or LGBTQ+ identity can \emph{indeed be beneficial}, especially if you can articulate how your diverse experiences contribute to diversity and inclusion in academic.
+In my personal opinion, highlighting URM or LGBTQ+ identity can \emph{indeed be beneficial}, especially if you can articulate how your diverse experiences contribute to diversity and inclusion in academics.
 Many universities have a strong commitment to diversity and inclusion and actively
@@
 
-It is understandable that you might feel uncomfortable disclosing this, fearing you would lose a chance of working with a faculty due to their bias.
+Understandably, you might feel uncomfortable disclosing this, fearing you would lose a chance of working with a faculty due to their bias.
 However, if someone has this issue, then you might not want to work with them anyway (and such bias is not acceptable in academics and likely would not be tolerated by the university).
@@
 Private universities may have different numbers.
-For simplicity, I will assume the department has a 9-month stipend of \$30000 and a 3-month summer of \$10000 (a third of the 9-month stipend).
-I will also use GMU tuition rate of about \$15,000/year for full-time study (which is quite cheap compared to private universities, e.g., Univ.
+For simplicity, I will assume the department has a 9-month stipend of \$30K and a 3-month summer of \$10K (a third of the 9-month stipend).
+I will also use GMU tuition rate of about \$15K/year for full-time study (which is quite cheap compared to private universities, e.g., Univ.
 of Chicago is a whooping \$70K) and a 58.9\% rate on \emph{indirect cost}, which is a typical rate charged for overhead or administrative costs (yes, after all, universities are businesses!).
@@
 
-In the end, the total budget comes out to be \$72K/year to support a PhD student.
+In the end, the total budget comes out to be \$73K/year to support a PhD student.
 \emph{The summary is that over your 5--6 years of your PhD, you cost about \$400K, and while your stipend is X, your adviser probably pays 2X for you}.
@@
 
-For start date or TA assignment (e.g., TA'ing a specific course), you can ask for it.
+For a specific start date or TA assignment (e.g., TA'ing a particular course), you can ask for it.
 Also, there is typically no moving allowance for PhD students.
-In short, standard things set by the university or department are unlikely non-negotiable.
+In short, standard things set by the university or department are unlikely negotiable.
 However, you can ask for things such as computer equipment (\autoref{sec:buying-equipment}), books, (and maybe even furniture), and a conference travel budget.
@@
 
-I've noticed many students, especially international, face a dilemma when they commit to a graduate offer but then receive a better one, often discussed in online forums.
-Advice often suggests it's okay to switch, using reasons like not having a strong relationship with the prof.
-or prioritizing personal benefit.
+I've seen many students, especially international, face a dilemma when they \emph{commit} to a graduate offer but then receive a better one.
+Advice given in online forums is often along the line that it's okay to switch, using reasons like you haven't yet had a strong relationship with the prof.
+or you should prioritize personal benefit.
 
@@
 %These connections are genuine and meaningful.
+
+\begin{commentbox}[Vu:]
+    My lab has various activities together (see \href{https://photos.app.goo.gl/LFtbqQUuznq9eiL7A}{here}).
+For example, we always have a Thanksgiving get-together at my house.
+Many faculty I know have similar activities with their lab members.
+  \end{commentbox}
 
@@
 
-I was also a member of the large VietPhD group on Facebook and browsed Internet forums (e.g., \texttt{Reddit/gradadmission} and \texttt{GradCafe}).
+I was also a member of the large VietPhD group on Facebook and often browse Internet forums (e.g., \texttt{Reddit/gradadmission} and \texttt{GradCafe}).
 There I saw many questions from students about PhD programs.
 However, most active participants are students in non-CS fields or not in the US, and their answers are unfortunately not always accurate and sometimes lead to more confusion.
-So I thought it would be useful to have a document that is specific to CS PhD programs in the US from an insider prospective.
+So I thought it would be useful to have a document that is specific to CS PhD programs in the US from an insider perspective.
 
 I started writing this document in May 2023 and have been updating it since then (mostly around deadline time when I tend to procrastinate, i.e., \emph{productive procrastination}!).
-The document was initially for international students but has been generalized to all students.
+The document was initially intended for international students but has been generalized to all students.
 I have put the source code of this document on \href{https://github.com/nguyenthanhvuh/phd-cs-us}{GitHub} so that others can contribute to it.
 
-\paragraph{About Me} I am an assistant professor in the CS dept at George Mason University (GMU).
+\paragraph{About Me} I am an associate professor in the CS dept at George Mason University (GMU).
 Before GMU, I was at the University of Nebraska-Lincoln (UNL).

### 2024-06-18 demystify.tex
updates

@@
 
-% \section{How long do you evaluate an application?}\label{sec:ievaluate}
-
-% On average, I spend about 15 minutes reviewing each application (less for clear rejections and more for potential acceptances).
-In earlier years I spent more time but now  I’ve become more efficient in evaluating..
-
-% The review system compiles a single PDF per applicant, including a summary page, transcripts, test scores, LoRs, a CV, SOP, and any writing samples.
-I review these documents in order, taking notes to support my evaluations.
-
-% I start with the summary, checking for low GPAs or test scores below the university minumum.
-I then skim the transcripts for low grades in relevant courses, noting issues like "- many low grades in main courses and - unknown international school with good GPA".
-
-% I pay close attention to strong LoRs and skim weaker ones, noting either strong endorsements from well-known professors or generic content.
-For the CV, I look for publications, research experiences, and notable achievements, noting things like "+ published papers in top venues" or "+ gold medal in an international competition."
-
-% The SOP is read thoroughly unless it's poorly written; a brief skim suffices in those cases. I note whether the SOP is engaging and well-tailored to our program or if it’s generic and uninspired.
-
-% Finally, I make a decision—usually either a direct rejection or an offer of admission with full funding, mentioning potential funding sources.
-Exceptionally strong candidates are recommended for the University Presidential fellowship.
-The admissions committee chair reviews all evaluations, typically following the recommendations, but discrepancies are resolved through discussion among reviewers to reach a consensus.
+\section{How long do you evaluate an application?}\label{sec:ievaluate}
+
+%
+
+On average, I spend about 10 minutes reviewing each application (less for clear rejections and more for potential acceptances).
+While this seems short, it is not that difficult to tell if an application is good or bad.
+In fact, this time
+is twice what other faculty spend on average, e.g., Philip Guo spent \href{https://pg.ucsd.edu/PhD-application-tips.htm}{3--5 minutes} per application.
+
+For each application, the GMU application system compiles a single PDF including a summary, transcripts, test scores, LoRs, a CV, SOP, and writing samples.
+I usually start with the \emph{summary}, checking for low GPAs or test scores below the university minimum.
+I then skim the \emph{transcripts} for low grades in relevant courses, noting issues like ``many low grades in main courses or unknown international school with good GPA''.
+These are not as important as LORs or SOP but I read them first because the review system has questions about them (e.g., ``is GPA good?
+is IELTS score sufficient?'')
+
+I pay close attention to \emph{strong LoRs} and skim weaker ones, noting either strong letters from well-known professors/researchers or generic content.
+For the \emph{CV}, I look for publications, research experiences, and notable achievements, noting things like ``published papers in top venues'' or ``gold medal in an international competition.''
+I read \emph{SOP} thoroughly (unless it's poorly written then I just skim).
+I note whether the SOP is exciting, research-oriented, and tailored to our program.
+All these notes are entered into the evaluation system.
+
+Finally, I make a decision, which is usually either a rejection or an offer of admission with full funding (e.g., from TA).
+I also recommend very strong candidates for the University Presidential fellowship.
+The adcom chair then reviews all evaluations, looking at entered notes, and ask reviewers to discuss and resolve discrepancies to reach a consensus.
+Typically there is no involvement from the adcom chair, department chair, or anyone else (i.e., the decision is made by the reviewers).
 
@@
 In many cases, professors or research labs put their projects on GitHub.
-For example, my research group has many open-source projects that undergrads can contribute to (see \url{https://github.com/dynaroars/}).
+For example, my research group, \href{https://github.com/dynaroars/}{Dynaroars}, has many open-source projects that undergrads can contribute to.
 
@@
 In my opinion, these reasons are not strong enough to justify retracting an acceptance.
-A more valid reason is using the \href{https://cgsnet.org/wp-content/uploads/2024/01/CGS_April15_Resolution_Jan312024.pdf}{April 15 resolution}, which many universities participate.
+A more valid reason is using the \href{https://cgsnet.org/wp-content/uploads/2024/01/CGS_April15_Resolution_Jan312024.pdf}{April 15 resolution}, in which many universities participate.
 Among various things, this resolution states that students are free to accept a new offer from a different institution until 4/15, even if they have already accepted an offer elsewhere.
@@
 However, regardless of the resolution, retracting an acceptance can have ethical implications.
-When you accept an offer, you are making a commitment to work with that prof, who then might stop looking for other students.
-So by retracting your offer, you are breaking your commitment and also causing a great deal of inconvenience to the prof and also takes away the opportunity from other students.
+When you accept an offer, you are committing to work with that prof, who then might stop looking for other students.
+So by retracting your offer, you are breaking your commitment and also causing a great deal of inconvenience to the prof and also taking away the opportunity from other students.
 Ultimately, this choice is personal, involving a balance between personal benefit and the ethical impact on others.

### 2024-06-18 faculty.tex
updates

@@
 Getting funding is arguably the most important part of getting tenure.
-Most new professors struggle with grant writing, which is very different compared to research papers.
+Writing grant propsal is very different from writing research papers.
 However, without funding, you can't pay your students, conference registration and travel, or even your own summer salary.
-In short, the lack of funding causes lost of stress, uncertainty, and threaten your career.
-In this chapter, we will discuss the various ways to obtain funding.
+%In short, the lack of funding causes lost of stress, uncertainty, and threaten your career.
+Most new professors struggle with grant writing at first, but eventually get better at it.  In this chapter, we will discuss the various ways to obtain funding.
+
+%Like most new professors, I struggled big time grant writing at first, but eventually got better at it.
+
+%, which is very different compared to research papers.   In this chapter, we will discuss the various ways to obtain funding.
 
 \section{Most people suck at grant writing (at first)} Most new assistant professors have no experience in grant writing, and struggle with it in the beginning.
-Then eventually they will get used to it and learn to like it (or at least tolerate it).
-The key is to keep trying, and to not give up.
-The best researchers at the best universities get rejected all the time.
-It's part of the process.
+Then eventually they will get used to it and learn to like it (or at least do not hate it as much).
 So don't get discouraged.
@@
 Keep trying, and you will get better over time.
-\section{NSF and other Government Funding}
+
+\section{Types of Funding}
+\subsection{NSF and other Government Funding}
 
@@
 
-\section{Internal Funding}
-
-\section{Industrial Gifts}
+\subsection{Internal Funding}
+
+\subsection{Industrial Gifts}
+
+\section{My Story}
 

### 2024-06-20 demystify.tex
updates
Updates from Overleaf

@@
 
-Here is some general advice for finding a prof.
+Here is some advice for finding a prof.
 Assuming you're not familiar with any particular one, then first search for profs.
@@
 
-Another effective way is taking graduate-level courses in the topics you are interested in.
-Profs teaching \emph{special topics courses} and \emph{research seminars} might be recruiting students---giving you even a higher chance.
+Another way is taking graduate-level courses in the topics you are interested in.
+Many profs teach \emph{special topics courses} and \emph{research seminars}, and they might be recruiting students.
 Do well in the class, answer questions, talk to the prof.
@@
 Many profs, including myself, prefer taking in new students this way.
-It gives both the prof and student more time, e.g., a whole semester, to work and evaluate the relationship before making any commitment (sounds a bit like a marriage!).
-
-You can also ask if you can do an independent study or research with a prof.
+It gives both the prof and student more time, e.g., a whole semester, to work and evaluate the relationship before making any commitment (sounds like a marriage!).
+
+You can also ask to do an independent study or research with a prof.
 This can be informal (no credit) and takes place during the summer or winter break.
-For example, I do this with multiple students, many of whom are undergrads.
-Many will drop out because they find they don't like my research, but some will stay and become my PhD students.
+For example, I do this with several students, some of whom are undergrads.
+Many will drop out because they find they don't like my research, but some find that they like the work.
 
@@
 Another one is that you have a history of jumping from one program to another (e.g., you have been in multiple different programs in a short time).
-How would the reviewers know?
-They might have contacts in other programs and heard about you, your LoRs might mention it.
+Adcom members might have contacts in other programs and heard about your stories, or your LoRs might mention these.
 
@@
 
-PhD students often ask why their salary is so low compared to the large grants their advisers get.
+PhD students often ask why their salary is low compared to the large grants their advisers get.
 They also wonder why their offer letters sometimes show that their benefits are higher than what they receive (i.e., stipend).
@@
 \paragraph{Standing out \autoref{sec:improve-your-chance}} There are \emph{few} domestic applications compared to international ones, i.e., domestic students are the minority (\autoref{sec:urm}) in the CS PhD application pool.
-Many US universities want to increase the number of domestic students in their programs (and as mentioned later, there are specific fellowships for domestic students).
+Many US universities want to increase the number of domestic students in their programs (and as mentioned later, there are specific fellowships and funding for domestic students).
 That makes you stand out and can help your case.
@@
 
-\paragraph{School \autoref{sec:your-school}} We already know about your school, which is a plus (see \autoref{sec:your-school}).
+\paragraph{School \autoref{sec:your-school}} Adcom already knows about your school, which is a plus (see \autoref{sec:your-school}).
 You are also more familiar with the US education system and culture (so hopefully you are aware of the cultural issues listed in \autoref{sec:cultural}).
@@
 You can also apply to specific programs \emph{before} you start your PhD, e.g., NSF Graduate Research Fellowship Program (GRFP) and Hertz Foundation Fellowship.
-These fellowships are very competitive and can significantly improve your admission chances if you get them.
+These fellowships are very competitive and can significantly improve your admission chances.
 
@@
 Advice given in online forums is often along the line that it's okay to switch, using reasons like you haven't yet had a strong relationship with the prof.
-or you should prioritize personal benefit.
+or you should prioritize your personal benefit.
 
@@
 
-However, regardless of the resolution, retracting an acceptance can have ethical implications.
+However, in general, retracting an acceptance can have ethical implications.
 When you accept an offer, you are committing to work with that prof, who then might stop looking for other students.
@@
 In the US, this is can be considered \emph{inappropriate} and discouraged.
+Most international students are aware of this, however, they might take it to the extreme and
 

### 2024-06-20 demystify.tex
updates

@@
 In the US, this is can be considered \emph{inappropriate} and discouraged.
-Most international students are aware of this, however, they might take it to the extreme and
 
@@
 With lab mates and colleagues, you will often work and go to lunch together, confide in each other, help each other navigate the academic journey, and often become lifelong friends.
-With your professor(s), you can call them by their first name (\autoref{sec:address}), disagree with them and argue (and gain respect doing so), seek their help (even on personal matters), come to their houses for parties or gathering, and give them small thoughtful gifts that they put on their desks (\autoref{sec:gifts}).
-Many people maintain lifelong relationships with their professors and colleagues, staying in touch through cards, emails, and calls, even after their academic journey ends.
-%These connections are genuine and meaningful.
+With your professor(s), you can call them by their first name (\autoref{sec:address}), disagree with them and argue (and gain respect doing so), seek their help (even on personal matters), come to their houses for parties or gathering (e.g., my lab always come to my house for \href{https://photos.app.goo.gl/LFtbqQUuznq9eiL7A}{Thanksgiving}), and give them small thoughtful gifts that they put on their desks (\autoref{sec:gifts}).
+Many people maintain lifelong relationships with their professors and colleagues, staying in touch through cards, emails, and visits, even after they nolonger work together.
 
 \begin{commentbox}[Vu:]
-    My lab has various activities together (see \href{https://photos.app.goo.gl/LFtbqQUuznq9eiL7A}{here}).
-For example, we always have a Thanksgiving get-together at my house.
-Many faculty I know have similar activities with their lab members.
+    I maintain a close relationship with my former professors and mentors.
+When there is a new event in my life (or theirs), I often send them an email or call them, e.g., when I get married, have a new baby, new job, etc.
+When I was having issues at UNL and considering leaving, I talked to them.
+I don't think this bothers them a bit; I think they are actually interested in these ``dramas'' in my life and would be offended if I didn't come to them.
+
+    I also visit my former professors when I am in their area.
+I meet Thang Bui (my MS adviser) at least once a year when I come back to Harrisburg to visit my parents.
+When Steph was in DC for a meeting, I invited her to Mason to give a talk on her work.
+I have also collaborated with them after I graduated e.g., I am working with Deepak on an NSF proposal on invariant generation.
+
+    In short, while I am a bit closer to my former advisers and mentors than most people (e.g., I still keep in touch with my middle school teacher in Hawaii), it is always a good idea to maintain a good relationship with people who have helped or worked well with you.
+A simple, short email or text once in a while (e.g., a \emph{``Hi X, I heard you just got promoted ...
+Congrats!''}) would suffice.
+They will appreciate it and you never know when you might need their help.
+
   \end{commentbox}

### 2024-06-21 demystify.tex
Updates from Overleaf
updates

@@
 So I want to share some details about the admission process and advice for those who are interested in applying for a \textbf{PhD in Computer Science in the US}.
-Originally, this document was intended for international students, but I have expanded it to include information that might also be useful for \emph{domestic students}.
-Moreover, while this handbook is primarily intended for students interested in CS, it might be relevant to students from various STEM (Science, Technologies, Engineering, and Mathematics) disciplines.
+Originally, this document was intended for international students, but I have expanded it to include information that might also be useful for \emph{US domestic students}.
+Moreover, while this is primarily intended for students interested in CS, it might be relevant to students from various STEM (Science, Technologies, Engineering, and Mathematics) disciplines.
 Furthermore, although many examples are specifics for schools that I and other contributors of this document know about, the information should be generalizable to other R1\footnote{An \href{https://en.wikipedia.org/wiki/List_of_research_universities_in_the_United_States}{R1 institution} in the US is a research-intensive university with a high level of research activity across various disciplines.
-Currently, 146 (out of 4000) US universities are classified as R1.} institutions in the US (and universities in other countries).
-
-In addition, this handbook can help \textbf{US faculty and admission committee} gain a better understanding of international students and their cultural differences.
+Currently, 146 (out of 4000) US universities are classified as R1.} institutions in the US.
+
+This information can also help \textbf{US faculty and admission committee} gain a better understanding of international students and their cultural differences.
 By recognizing and leveraging these differences, CS programs in the US can attract larger and more competitive application pools from international students.
@@
           \item Ask your prof. if they can buy computer equipment and such for your research (\autoref{sec:buying-equipment}).
-          \item Despite some miserable stories on social media, many PhD students have good mentors, supportive lab mates, healthy working environment ... and are happy (\autoref{sec:happy}).
+          %\item Despite some miserable stories on social media, many PhD students have good mentors, supportive lab mates, healthy working environment ... and are happy (\autoref{sec:happy}).
         \end{itemize}
@@
 
-The guideline can be useful for all students, but it is primarily intended for \textbf{international students} from \textbf{smaller countries and less well-known universities} (similarly, for domestic students from universities with no PhD programs or limited research experiences).
-Students from top schools with strong research cultures and experiences might already know most of the information presented here.
+The guideline can be useful for all students, but it is primarily intended for \textbf{international students} from \textbf{smaller countries and less well-known universities} (similarly, for domestic students from US universities with no PhD programs or limited research experiences).
+Students from top schools with strong research cultures and experiences might already know most of the information given here.
 My goal is thus to level the playing field by providing information that is not readily available to less privileged students. 
@@
 
-While this document may not fully meet the needs of students aiming for top-10 schools, which typically have more stringent requirements (e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU), it should be valuable for those targeting other R1 reputable institutions.
+While this document may not fully meet the needs of students aiming for top-10 schools, which typically have more stringent requirements (e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU), it should be valuable for those targeting other reputable R1 institutions.
 It is certainly useful for students aiming for GMU!
@@
 
-After you submit your PhD application (usually around December), it will first be screened
+After you submit your PhD application (usually around December of year $X-1$ for the Fall admission of year $X$), it will first be screened
 for general requirements, e.g., did you submit your transcripts and standard scores?
 did your reference writers submit their letters?
-Usually, this screening process is done through a central university system, i.e., not by faculty from the CS department.
+Usually, this screening process is done through a central university system, i.e., not by CS faculty.
 If you pass this screen, your application will be forwarded to the CS department for further evaluation.
@@
 
-\paragraph{Why CS depts do not waive the application fee?}  Some universities do waive, e.g., Rice does that for PhD applicants and many universities for domestic students (\autoref{sec:domestic-students}), however most do not.
+\paragraph{Why CS depts do not waive the application fee?}  Some universities do waive, e.g., Rice does for all PhD applicants and many universities for domestic students (\autoref{sec:domestic-students}), however most do not.
 In any case, the application fee is typically a requirement of the university.
@@
 Also, most CS programs already receive many applications, and would be overwhelmed if the application process were free.
-
-% \begin{commentbox}[Hung]
-%   The application fee varies between schools, and is about \$70--\$100.
-I found this number pretty ridiculous.
-In most cases, there is no proper justification for that amount.
-Yes, the application fee could act as a filter, but a nominal amount would do the job.
-While a blanket waiver might not be feasible, universities should extend the eligibility of fee waiver applications to more countries.
-% \end{commentbox}
 
@@
 Most PhD programs will require at least \textbf{two LORs}.
-A good LoR letter should have detail, enthusiasm, and specific examples of the applicant's achievements, all of which paint a compelling picture of the candidate to the adcom.
+A good LoR letter should give details of the applicant's research background and achievements, all of which paint a compelling picture of the candidate to the adcom.
 
@@
 When I was having issues at UNL and considering leaving, I talked to them.
-I don't think this bothers them a bit; I think they are actually interested in these ``dramas'' in my life and would be offended if I didn't come to them.
+I think this does not bother them a bit and that they are actually interested in these ``dramas'' in my life.
 

### 2024-06-21 demystify.tex
Updates from Overleaf

--- 2024-06-20 demystify.tex
@@
 
-In my personal opinion, highlighting URM or LGBTQ+ identity can \emph{indeed be beneficial}, especially if you can articulate how your diverse experiences contribute to diversity and inclusion in academics.
+In my personal opinion, highlighting URM or LGBTQ+ identity can \emph{be beneficial}, especially if you can articulate how your diverse experiences contribute to diversity and inclusion in academics.
 Many universities have a strong commitment to diversity and inclusion and actively
@@
 
+Typically, it takes 5--7 years for CS PhD in the US.
+This can be longer than CS PhDs in other countries, which might require MS first (recall from \autoref{sec:msrequirement} that CS PhD programs in the US do not require MS and you can get MS along the way to PhD).
+Within these 5--7 years, CS PhD students often take a ``leave of absence'' for 1--2 semesters to do internships at companies and research labs.
+
 \begin{center}
@@
 
-Typically, it takes 5--7 years for CS PhD in the US.
-This can be longer than CS PhDs in other countries, which might require MS first (recall that CS PhD programs in the US do not require MS and you can get MS along the way to PhD).
-Within these 5--7 years, CS PhD students often take a ``leave of absence'' for 1--2 semesters to do internships at companies and research labs.
-
 The first two years you typically take coursework (somewhat equivalent to an MS study), find an adviser, and learn how to do research.
 The next 2--3 years you focus on your research, form a dissertation topic, and get results published.
-The last 1-2 years you continue to publish, write and defend your dissertation, and look for a job.
+The last 1--2 years you continue to publish, write and defend your dissertation, and look for a job.
 In many cases you might take a summer or two off to do an internship to get additional research opportunities.
-The PhDComics figure on top shows the ``ambition'' level of a PhD student over their years of study (they miss the 6--7th year when the ambition is \emph{``Just let me graduate''}).
+The PhDComics figure above shows the ``ambition'' level of a PhD student over their years of study (they miss the 6--7th year when the ambition is \emph{``Just let me graduate''}).
 
@@
 
-\emph{MS requirement and PhD duration}:  as mentioned in \autoref{sec:msrequirement} and \autoref{sec:time}, CS PhD programs in the US do not require an MS degree.
+\emph{MS requirement and PhD duration}:  CS PhD programs in the US do not require an MS degree (as mentioned in \autoref{sec:msrequirement} and \autoref{sec:time}).
 In contrast, many other countries require having an MS degree before joining a PhD program.
@@
 \emph{Project proposal}: in many countries, you have to choose a project and adviser \emph{during} the application process (e.g., you write a proposal to a potential adviser).
-You start your research right from the beginning.
+But this allows you to start your research right from the beginning.
 
@@
 You will have two initial years to explore and find an adviser and research topic.
-And then you often explore projects with that adviser before settling on one (can take a couple of years).
+After that you explore projects with that adviser before settling on one (can take a couple of years).
 
 \emph{Course work}: as mentioned above in the US you will spend the first couple of years taking classes and exploring potential adviser and research topics. 
-After that, you also have to pass a series of exams during your PhD, e.g., qualifying exam, comprehensive exam, thesis proposal defense\footnote{The word ``ABD'' (all but dissertation) is used in the US to refer to a PhD candidate who has finished all course work and exams and only needs to write and defend their dissertation.}.
+After that, you have to pass a series of exams during your PhD, e.g., qualifying exam, comprehensive exam, thesis proposal defense\footnote{The word ``ABD'' (all but dissertation) is used in the US to refer to a PhD candidate who has finished all course work and exams and only needs to write and defend their dissertation.}.
 
@@
 
-In the US, PhD graduates often apply directly for faculty positions (postdoc for US graduates is no longer a popular option as it was before).
+In the US, PhD graduates often apply directly for faculty positions. Postdoc for US graduates is no longer a popular option as it was before.
 
@@
 
-Understandably students get excited about their upcoming PhD journey and want to buy new computer equipment and electronics to prepare.
+Students understandbly get excited about their upcoming PhD journey and want to buy new computer equipment and electronics to prepare.
 However, you should check with your prof.
@@
 You can also go to the department directory and then professors' websites and see if they are looking for undergraduate researchers.
-Even if they say they are looking for graduate students, you still should contact them and ask.
-
-There might be also special programs for undergraduate research from the university, e.g., GMU has the OSCAR program, UNL has UCARE, NSF has REUs (for US citizens and permanent residents).
+Even if they say they are looking for graduate students, you can still contact them and ask.
+
+There might be also special programs for undergraduate research from the university, e.g., Mason has the OSCAR program, UNL has UCARE, and NSF has REUs (for US citizens and permanent residents).
 
@@
 When there is a new event in my life (or theirs), I often send them an email or call them, e.g., when I get married, have a new baby, new job, etc.
-When I was having issues at UNL and considering leaving, I talked to them.
-I don't think this bothers them a bit; I think they are actually interested in these ``dramas'' in my life and would be offended if I didn't come to them.
+When I was having issues at UNL and considering leaving, I talked to my PhD advisers Deepak and Steph.
+I think this does not bother them a bit and that they are actually interested in these ``dramas'' in my life.
 
@@
 
-\paragraph{About Me} I am an associate professor in the CS dept at George Mason University (GMU).
-Before GMU, I was at the University of Nebraska-Lincoln (UNL).
-I have been in the PhD admission process at GMU and UNL for many years.
-Currently I serve as the program director of the MS program in Software Engineering at GMU (thus also have experience with the MS admission process---which is quite different than PhD).
+\paragraph{About Me} I am an associate professor in the CS dept at George Mason University.
+Before Mason, I was at the University of Nebraska-Lincoln (UNL).
+I have been in the PhD admission process at Mason and UNL for many years.
+Currently I serve as the program director of the MS program in Software Engineering at Mason (thus also have experience with the MS admission process---which is quite different than PhD).
 My personal and lab website is at \href{https://dynaroars.github.io}{dynaroars.github.io}.

### 2024-06-21 demystify.tex

--- 2024-06-21 demystify.tex
@@
 When there is a new event in my life (or theirs), I often send them an email or call them, e.g., when I get married, have a new baby, new job, etc.
-When I was having issues at UNL and considering leaving, I talked to them.
+When I was having issues at UNL and considering leaving, I talked to my PhD advisers Deepak and Steph.
 I think this does not bother them a bit and that they are actually interested in these ``dramas'' in my life.

### 2024-06-21 demystify.tex
updates

@@
 International students, especially those from less well-known universities and smaller countries, often wonder if they should apply to a CS PhD program in the US.
-Specifically, they are worried about the \emph{difficulty} of getting admitted and the \emph{cost} of studying in the US.
+Specifically, they are worried about the \textbf{difficulty of getting admitted} and the \textbf{cost of graduate study} in the US.
 This section addresses these concerns.
@@
 First, applying to a good US university \emph{should not} be any harder than at schools in other countries.
-In fact, it might even be more flexible since CS PhD in the US \emph{do not} require having an MS or a topic/proposal/adviser in advance.
+In fact, it might even be more flexible since CS PhD in the US \emph{do not} require having an MS or a research topic/proposal/adviser in advance.
 If you believe you have a chance in other countries, e.g., South Korea, Singapore, Germany, UK, Japan, and Australia, then you will surely have a chance in the US as well.
@@
 \begin{commentbox}[Vu:]
-  One of the reasons I created this document is that my colleagues at GMU are interested in recruiting Vietnamese students and are surprised when seeing very few applications in Vietnam (see \autoref{sec:ack}).
+  One of the reasons I created this document is that my colleagues at Mason are interested in recruiting Vietnamese students and are surprised when seeing very few applications in Vietnam (see \autoref{sec:ack}).
 Each year our CS PhD program receives 500+ PhD applications, most of which are international but only 5--6 are from Vietnam.
@@
 
-The guideline can be useful for all students, but it is primarily intended for \textbf{international students} from \textbf{smaller countries and less well-known universities} (similarly, for domestic students from US universities with no PhD programs or limited research experiences).
+The guideline can be useful for all students, but it is primarily intended for \textbf{international students} from \textbf{smaller countries and less well-known universities} (or domestic students from US universities with no PhD programs or limited research experiences).
 Students from top schools with strong research cultures and experiences might already know most of the information given here.
@@
 While this document may not fully meet the needs of students aiming for top-10 schools, which typically have more stringent requirements (e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU), it should be valuable for those targeting other reputable R1 institutions.
-It is certainly useful for students aiming for GMU!
+It is certainly useful for students aiming for Mason!
 
@@
   \item  \href{https://pg.ucsd.edu/PhD-application-tips.htm}{Blog post on PhD Admission} by Philip Guo
-  \item   \href{https://talkingtorobots.com/yonatanbisk.html}{PhD Admission post} by Yonatan Bisk (customized for CMU admission)
+  \item   \href{https://talkingtorobots.com/FAQ.html}{PhD Admission post} by Yonatan Bisk (customized for CMU admission)
 \end{itemize}
@@
 
-After you submit your PhD application (usually around December of year $X-1$ for the Fall admission of year $X$), it will first be screened
+After you submit your PhD application (e.g., around December of year $X$ for the Fall admission of year $X+1$), it will first be screened
 for general requirements, e.g., did you submit your transcripts and standard scores?
@@
 \begin{commentbox}[Hakan:]
-  At GMU, for full consideration, students should make sure to submit \textbf{ALL} required documents by the application deadline, and should never assume that some required documents (such as official TOEFL scores or official diplomas/transcripts) will be waived by the admissions office.
+  At Mason, for full consideration, students should make sure to submit \textbf{ALL} required documents by the application deadline, and should never assume that some required documents (such as official TOEFL scores or official diplomas/transcripts) will be waived by the admissions office.
 If something is listed and not marked as ``optional'', it is mandatory and they should plan for submitting all those.
@@
 The size and the review load of the adcom depend on the department size.
-At GMU, the PhD adcom typically has 15--20 faculty, and each committee member is assigned to review 25--30 applications.
-Note that most large schools, including GMU, have separate adcoms for MS programs.
+At Mason, the PhD adcom typically has 15--20 faculty, and each committee member is assigned to review 25--30 applications.
+Note that most large schools, including Mason, have separate adcoms for MS programs.
 
@@
 
-At GMU, we usually decide that a full-time PhD candidate is either (i) admitted with funding (\autoref{sec:funding}) or (ii) rejected. In other words, in most cases, we either
+At Mason, we usually decide that a full-time PhD candidate is either (i) admitted with funding (\autoref{sec:funding}) or (ii) rejected. In other words, in most cases, we either
 admit you with full funding or reject your application. In some rare cases, we admit
@@
 \begin{commentbox}[Craig:]
-  GMU and many other universities allow you to upload your published papers and other writing samples.
+  Mason and many other universities allow you to upload your published papers and other writing samples.
 In many cases, even if the papers were not published at top places, we can still determine their quality by simply skimming over the paper.
@@
 For example, when I was a postdoc at UMD, members of their CS PhD adcom asked me to evaluate applicants from Vietnam.
-During my time at UNL and now here at GMU, I have looked at Vietnamese applications (whether they are assigned to me or not) and provided input to their reviewers, e.g., X is the top tech school in Vietnam and so it should be \emph{top} instead of \emph{unknown} foreign, which makes a huge difference.
+During my time at UNL and now here at Mason, I have looked at Vietnamese applications (whether they are assigned to me or not) and provided input to their reviewers, e.g., X is the top tech school in Vietnam and so it should be \emph{top} instead of \emph{unknown} foreign, which makes a huge difference.
 \end{commentbox}
@@
 
-\paragraph{English Tests (IELTS, TOEFL)} Unless your degrees are from specific countries such as \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-GMU#standard-tests-waiver-eligible-countries}{these}, you will need to
+\paragraph{English Tests (IELTS, TOEFL)} Unless your degrees are from specific countries such as \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-Mason#standard-tests-waiver-eligible-countries}{these}, you will need to
 take standardized English tests.
@@
 \begin{commentbox}[Vu:]
-  The minimum for GMU (being above this might not mean much, but below is a \red{red flag}).
+  The minimum for Mason (being above this might not mean much, but below is a \red{red flag}).
   \begin{itemize}
@@
 \begin{commentbox}[Vu:]
-  At GMU, faculty are encouraged to interview candidates.
+  At Mason, faculty are encouraged to interview candidates.
 For very strong candidates, the interview is actually to recruit them.
@@
 Many programs do not have the tradition of interviewing applicants.
-For example, at GMU, most admitted students with TA (\autoref{sec:ta}) do not go through interviews.
+For example, at Mason, most admitted students with TA (\autoref{sec:ta}) do not go through interviews.
 
@@
 \begin{commentbox}[Vu:]
-  GMU has \emph{Virtual} Open House (VOH), e.g., \url{https://cs-gmu.github.io/cs-phd-voh-s23/}.
-We invite all admitted PhD students to the VOH through Zoom to learn about the CS program, the department, GMU, and the DC area in general.
+  Mason has \emph{Virtual} Open House (VOH), e.g., \url{https://cs-Mason.github.io/cs-phd-voh-s23/}.
+We invite all admitted PhD students to the VOH through Zoom to learn about the CS program, the department, Mason, and the DC area in general.
 Students also get opportunities to chat with professors and current students.
@@
 A good school would know that it has to be competitive to attract students.
-For example, at GMU, every year we discuss improving the benefits, and especially stipend, for our graduate students.
+For example, at Mason, every year we discuss improving the benefits, and especially stipend, for our graduate students.
 
@@
 \begin{commentbox}[Vu:]
-  TA and RA at GMU have similar benefits in tuition waiving and insurance.
+  TA and RA at Mason have similar benefits in tuition waiving and insurance.
 The college and department will set the rate for a 9-month graduate assistant stipend.
@@
 Do not assume that you're young and healthy and ignore insurance.
-At GMU, and most good CS PhD programs, your GTA or GRA \emph{will always} come with full insurance.
-In fact, at GMU your spouse/children will get a significant discount rate for health insurance.
+At Mason, and most good CS PhD programs, your GTA or GRA \emph{will always} come with full insurance.
+In fact, at Mason your spouse/children will get a significant discount rate for health insurance.
 So you will never have to worry much about health issues for you or your family here.
@@
 \begin{commentbox}[Vu:]
-  At GMU CS, admitted PhD students have 4 years of GTA guaranteed, and also receive a stipend for the first summer (\autoref{sec:summer-funding}).
+  At Mason CS, admitted PhD students have 4 years of GTA guaranteed, and also receive a stipend for the first summer (\autoref{sec:summer-funding}).
 \end{commentbox}
@@
 This allows you to see what teaching is like, which is especially helpful for a research career where you often give talks and tell people about your work.
-GMU sometimes has classes that a more senior student can teach.
+Mason sometimes has classes that a more senior student can teach.
 In that case, you will be paid as GTA or even sometimes as a lecturer.
@@
 \begin{commentbox}[Vu:]
-  PhD applicants at GMU are automatically eligible for a \emph{Presidential Fellowship}.
+  PhD applicants at Mason are automatically eligible for a \emph{Presidential Fellowship}.
 It is at least as good as GTA but the most important thing is that as a fellowship it is truly free money (i.e., you do not depend on any prof.
@@
 Some CS departments offer some summer funding, but it is typically not guaranteed and might not be a lot.
-For example, at GMU, we offer summer funding for 1st-year PhD TA students.
+For example, at Mason, we offer summer funding for 1st-year PhD TA students.
 The amount over the 3-month summer is similar to their monthly stipend (i.e., their 9-month stipend divided by 3 for the 3 summer months).
@@
 Over the 3-summer months, I typically pay them 1/3 of their 9-month stipend.
-I prioritize summer funding for my students because GMU has very good TA resources so they never have to worry about funding during the AY.
+I prioritize summer funding for my students because Mason has very good TA resources so they never have to worry about funding during the AY.
 
@@
 
-  I am writing to inquire about potential research opportunities as a GRA in your group at GMU. Currently I am an undergraduate student in Computer Science at UNIV and plan to graduate in May 2023.
+  I am writing to inquire about potential research opportunities as a GRA in your group at Mason. Currently I am an undergraduate student in Computer Science at UNIV and plan to graduate in May 2023.
 
@@
 You \emph{do not} need to do a PhD in CS to work in your area of interest.
-For example, in addition to a traditional CS department, GMU has IST and Cybersecurity departments, both of which have faculty with PhD in CS and work on CS topics (e.g., AI, Security, Robotics).
+For example, in addition to a traditional CS department, Mason has IST and Cybersecurity departments, both of which have faculty with PhD in CS and work on CS topics (e.g., AI, Security, Robotics).
 So it is possible that you still get to do CS research and publish in CS-focused venues even if you're not in a traditional CS program.
@@
 
-\chapter{Miscs and FAQs}
+\chapter{Miscs and FAQs}\label{sec:faqs}
+
 \epigraph{``I want to share something with you – the three little sentences that will get you through life; number 1: Cover for me, number 2: Oh, good idea, Boss, and number 3: It was like that when I got here.''}{\textsc{The Simpsons}}
@@
 
+\paragraph{Not a good fit and Bad luck}  You could have an excellent profile, but if you are interested in a research area that the program does not have, you might not be admitted.
+Similarly, if no faculty who is willing to advise you (e.g., they are on sabbatical or personal leave, do not have sufficient funding, or already have too many students already), you will not be admitted.
+This is good for you, as you don't want to be in a program where no one can advise you.
+
+Related to this is that you just have bad luck and apply at the wrong time.
+For example, since 2024 there has been a huge surge in students interested in AI and NLP (thanks, ChatGPT!).
+Consequently, AI/NLP faculty might be overwhelmed and cannot consider many candidates, even those with excellent profiles.
+
 \paragraph{You aim too high} This is a very common reason: you have applied to schools that are too high for your profile.
@@
 
-Before applying, you should talk to your professors and ask them to give you an honest opinion on where you should apply. You should also apply to a range of schools, including ``safety'' ones.
-
-\paragraph{Not a good fit}  You could have an excellent profile, but if you are interested in a research area that the program does not have, you might not be admitted.
-Similarly, if no faculty who is willing to advise you (e.g., they are on sabbatical or personal leave, do not have sufficient funding, or already have too many students already), you will not be admitted.
-This is good for you, as you don't want to be in a program where no one can advise you.
-
-For example, the 2024 admission cycle has a huge surge in students interested in AI and NLP (thanks, ChatGPT!).
-Consequently, AI/NLP faculty might be overwhelmed and cannot consider many candidates, even those with excellent profiles.
-
-\paragraph{Low English exam scores (e.g., IELTS or TOEFL)}  While profs.
-might not care much about these, there might be a certain minimum set by the college or the university that you need to pass to get admitted, especially for TA funding (e.g., you need to communicate well with students as a TA).
+Before applying, you should talk to your professors and ask them to give you an honest opinion on where you should apply.
+To increase their chances, many students apply to a range of schools, including ``safety'' ones.
+
+\paragraph{Overqualified or Lack Interest}  This might be paradoxical and is the opposite of aiming too high.
+However if adcom believes that you are likely to get admitted and go to a better program, they might not be so enthusiastic to admit you and want to save the spot for someone else.
+
+A related reason is that you did not show enough interest in the program.
+For example, your SOP did not mention any faculty or research area at the program, you did not respond to professors interested in interviewing you (which might be considered unprofessional and burn bridges), or during the interview you showed little knowledge, interest, or enthusiasm (because they are your safety schools) in the program or profs.
+
+\paragraph{Low English exam scores (e.g., IELTS or TOEFL)}  Profs.
+might not care much about these, but the college or the university often sets a minimum that you need to pass to be considered, especially for TA funding (e.g., you need to communicate well with students as a TA).
 Thus, while profs.
@@
 
-Note this is for international students, as domestic ones (and likely those from these \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-GMU#standard-tests-waiver-eligible-countries}{countries}) do not have to take these tests.
-
-\paragraph{Overqualified or Lack Interest}  This might be paradoxical and is the opposite of aiming too high.
-However if adcom believes that you are likely to get admitted and go to a better program, they might not be so enthusiastic to admit you and want to save the spot for someone else.
-
-A related reason is that you did not show enough interest in the program.
-For example, your SOP did not mention any faculty or research area at the program, you did not respond to professors interested in interviewing you (which might be considered unprofessional and burn bridges), or during the interview you showed little knowledge, interest, or enthusiasm (because they are your safety schools) in the program or profs.
+Note that English exams are not required for all students, e.g., domestic students and those from certain countries are waived from taking these exams.
+For example, at Mason also waive the English exams for students from these \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-Mason#standard-tests-waiver-eligible-countries}{countries}.
 
@@
 
-  %\item \emph{You did not make a good impression}.
-For example, you might have contacted a professor and they did not like your email.
-Or you might have met a professor (e.g., at a conference) and they did not like your presentation.
-This is rare, but it happens.
-
-\section{How long do you evaluate an application?}\label{sec:ievaluate}
-
-%
-
-On average, I spend about 10 minutes reviewing each application (less for clear rejections and more for potential acceptances).
-While this seems short, it is not that difficult to tell if an application is good or bad.
-In fact, this time
-is twice what other faculty spend on average, e.g., Philip Guo spent \href{https://pg.ucsd.edu/PhD-application-tips.htm}{3--5 minutes} per application.
-
-For each application, the GMU application system compiles a single PDF including a summary, transcripts, test scores, LoRs, a CV, SOP, and writing samples.
-I usually start with the \emph{summary}, checking for low GPAs or test scores below the university minimum.
-I then skim the \emph{transcripts} for low grades in relevant courses, noting issues like ``many low grades in main courses or unknown international school with good GPA''.
-These are not as important as LORs or SOP but I read them first because the review system has questions about them (e.g., ``is GPA good?
-is IELTS score sufficient?'')
-
-I pay close attention to \emph{strong LoRs} and skim weaker ones, noting either strong letters from well-known professors/researchers or generic content.
-For the \emph{CV}, I look for publications, research experiences, and notable achievements, noting things like ``published papers in top venues'' or ``gold medal in an international competition.''
-I read \emph{SOP} thoroughly (unless it's poorly written then I just skim).
-I note whether the SOP is exciting, research-oriented, and tailored to our program.
-All these notes are entered into the evaluation system.
-
-Finally, I make a decision, which is usually either a rejection or an offer of admission with full funding (e.g., from TA).
-I also recommend very strong candidates for the University Presidential fellowship.
-The adcom chair then reviews all evaluations, looking at entered notes, and ask reviewers to discuss and resolve discrepancies to reach a consensus.
-Typically there is no involvement from the adcom chair, department chair, or anyone else (i.e., the decision is made by the reviewers).
-
-% Around 15 minutes on average (less than 10 mins for definite rejection; 20+ mins for acceptance). In earlier years I spent more time, but now I can quickly tell if an application is good or bad.
-
-% The evaluation system allows for downloading a single pdf that consists of a summary page (e.g., schools, GPA, standard test scores), transcripts, test results, LoRs, CV, SOP, and writing samples (if any).
-I typically review the application in this order.
-I also make notes on what I've seen, using them to support my evaluation.
-
-% I first skim the summary for school, GPA, standard tests and look for red flags like low GPA or test scores that do not meet the minimum requirements set by the university.
-Then I scan transcripts for low/failing grades on CS and related courses.
-For these I make comments such as ``- unknown international school with good GPA, too many low grades on CS and related courses" or ``+ very good GPA from top school XYZ".
-I do not look at test details as the summary already gives the main scores.
-
-% Then I look at LoRs.
-I read carefully good LoRs and quickly skim over bad or generic ones (very easy to tell, you don't even need much experience to distinguish between strong LoRs from the rest).
-My comments for LoRs include ``+ very strong LoRs from well-known prof.
-XYZ talks about research experiences'' or ``- generic LoRs, mostly about course work''.
-I also note any unusual or unique details in the LoRs, e.g., ``+ LoR mentioned about student's significant contribution on well-known open-source project X''.
-Sometimes a LoR came directly from a GMU professor who mentioned they already know the student and intend to offer RA, which is a very strong signal for admission (and my comment would be ``Prof.
-X making RA offer'').
-
-% I skim over CV, looking for paper publications, research work experiences, and other relevant achievements such as winning prizes in well-known international programming/math competitions.
-My notes for CV include things such as ``+ has X papers at top venues'' or ``+ gold medal at X''.
-Sometimes I also note about unusual or unique experiences, e.g., ``+ has a patent on X''.
-
-% Next is the SOP, which I read carefully (but if it is bad then it would be a quick read).
-My notes for SOP include things such as ``- SOP mostly on course work, no clear motivation, generic" , or ``+ very enjoyable SOP talking about XYZ, customized for GMU, mentioned profs.
-X,Y,Z".
-If the SOP mentioned publications then I would have ``+ has X papers at top venues" or ``+/- has X papers but at unknown places''.
-I will not go online to search for the papers (unless they upload them as sample papers).
-I also skim over sampling papers, essentially looking over the abstract to see if reads well, and make notes such as ``+ have a paper under submission to unknown venue, but reads well".
-If the student mentioned a project, I might also look at it (e.g., if it is on Github) and make notes such as ``+ has project on Github with X stars''.
-
-% At this point I have enough to make up my decision, which I enter into the evaluation system.
-Typically my decisions are either \emph{reject} or \emph{admit with full funding}.
-For the latter, I also mention funding sources (e.g., RA from prof.
-X or TA).
-I also include my comments to support the decision.
-Finally, I also recommend very strong students for the University Presidential fellowship (\autoref{sec:fellowships}).
-
-% The adcom chair will look at all evaluations for each application and make a decision (likely following the decision of the reviewers).
-If there are discrepancies such as one voting reject but another full funding acceptance (such cases are rare, but do happen), the chair might ask the reviewers to coordinate to reach a consensus.
-
-%\section{Why are rejection letters sent so late?}\label{sec:late-rejection}
+\section{Timeline: Why are rejection letters sent so late?}\label{sec:late-rejection}
 
@@
 
+\paragraph{Acceptance Letters} Universities typically prioritize sending out acceptance letters first.
+This allows the admitted students to start planning and preparing for their studies.
+These letters go to the top candidates who are offered admission directly without being placed on a waitlist.
+Some universities have rolling admissions, meaning they send out acceptance letters as they review applications (e.g., some universities sent out letters as soon as early February).
+Others have a specific date when they send out the first round of acceptance letters.
+
+\paragraph{Response Deadlines}: Accepted students are usually given a strict deadline to respond to their offers (e.g., whether they will accept the offer), often around \textbf{April 15th} in the US.
+This is a standard deadline for PhD programs due to the Council of Graduate Schools (CGS) Resolution.
+After this date, programs can gauge how many slots remain unfilled.
+
+\paragraph{Waitlist}: If the CS program has X slots, then many strong candidates will be waitlisted.
+If accepted students decline the offer, then offers are sent to students on the waitlist.
+So if you see people getting accepted, that does not mean you are out yet.
+
+\paragraph{Rejection Letters} Typically, schools start sending out rejection letters to remaining applicants after they have finalized their admissions decisions.
+Thus, rejection letters are often sent out late (e.g., after April 15th or even much later).
+
 \section{Increasing my admission chance}\label{sec:improve-your-chance}
@@
 
-Given the high number of quality applicants and limited number of spots, in addition to having a good application profile, you really want to show something that makes you \textbf{stand out}.
+Given the high number of quality applicants and a limited number of spots, in addition to having a good application profile, you really want to show something that makes you \textbf{stand out}.
 For example, even if you do not have research experience, you can talk about your personal projects, as long as they can help show you can do research.
-For example, if you have an open-source project on Github that is used by many people, has lots of stars in Github, do talk about it.
+For example, if you have an open-source project that has lots of stars in Github, then you should mention it.
 If you often write technical, research-like blogs with many viewers, talk about them too.
@@
 
+\section{How long do you evaluate an application?}\label{sec:ievaluate}
+
+On average, I spend about 10 minutes reviewing each application (less for clear rejections and more for potential acceptances).
+While this seems short, it is not that difficult to tell if an application is good or bad.
+In fact, this time
+is twice what other faculty spend on average, e.g., Philip Guo spent \href{https://pg.ucsd.edu/PhD-application-tips.htm}{3--5 minutes} per application.
+
+For each application, the Mason application system compiles a single PDF including a summary, transcripts, test scores, LoRs, a CV, SOP, and writing samples.
+I usually start with the \emph{summary}, checking for low GPAs or test scores below the university minimum (\autoref{sec:standard-tests}).
+I then skim the \emph{transcripts} for low grades in relevant courses, noting issues like ``many low grades in main courses or unknown international school with good GPA''.
+These are not as important as LORs or SOP but I read them first because the review system has questions about them (e.g., ``is GPA good?
+is IELTS sufficient?'')
+
+I read carefully \emph{strong LoRs} and skim weaker ones, noting either strong letters from well-known professors/researchers or generic content.
+For the \emph{CV}, I look for publications, research experiences, and notable achievements, noting things like ``published papers in top venues'' or ``gold medal in an international competition.''
+I read strong \emph{SOP} thoroughly (and skim weak ones). I note whether the SOP is exciting, research-oriented, and tailored to our program. All these notes are entered into the evaluation system.
+
+Finally, I make a decision, which is usually either a rejection or an offer of admission with full funding (e.g., from TA).
+I also recommend very strong candidates for the University Presidential fellowship.
+The adcom chair then reviews all evaluations, looking at entered notes, and ask reviewers to discuss and resolve discrepancies to reach a consensus.
+Typically there is no involvement from the adcom chair, department chair, or anyone else (i.e., the decision is made by the reviewers).
+
 \section{Member of an underrepresented or LGBTQ+ groups}\label{sec:urm}
@@
 Many universities have a strong commitment to diversity and inclusion and actively
-\emph{recruit students from underepresentative backgrounds} (e.g., some \href{https://cse-climate.engin.umich.edu/reports/climate-dei-reports/cse-climate-and-dei-report-2022-2023/#grad-ethnicity}{stats from UMich CSE} and \href{https://www.gmu.edu/news/2022-09/mason-now-top-10-public-university-diversity-innovation-and-cybersecurity-education-us}{GMU} often promotes itself as one of the most diverse universities in the US).
+\emph{recruit students from underepresentative backgrounds} (e.g., some \href{https://cse-climate.engin.umich.edu/reports/climate-dei-reports/cse-climate-and-dei-report-2022-2023/#grad-ethnicity}{stats from UMich CSE} and \href{https://www.Mason.edu/news/2022-09/mason-now-top-10-public-university-diversity-innovation-and-cybersecurity-education-us}{Mason} often promotes itself as one of the most diverse universities in the US).
 Moreover, many scholarships and fellowships are created specifically for URM and LGBTQ+ students, which you should explore if qualify.
@@
 
-Moreover, promoting diversity aligns with the goal of broadening participation in computing (BPC) that is strongly supported within academia (e.g., see the \href{https://plans.bpcnet.org/GeorgeMasonUniversity_ComputerScience_DepartmentalBPCPlan.pdf}{BPC plan at GMU CS}) and government funding agencies such as NSF (e.g., see \href{https://new.nsf.gov/cise/broadening-participation}{here} and \href{https://www.nsf.gov/pubs/2022/nsf22125/nsf22125.jsp}{here}).
+Moreover, promoting diversity aligns with the goal of broadening participation in computing (BPC) that is strongly supported within academia (e.g., see the \href{https://plans.bpcnet.org/GeorgeMasonUniversity_ComputerScience_DepartmentalBPCPlan.pdf}{BPC plan at Mason CS}) and government funding agencies such as NSF (e.g., see \href{https://new.nsf.gov/cise/broadening-participation}{here} and \href{https://www.nsf.gov/pubs/2022/nsf22125/nsf22125.jsp}{here}).
 Also, sometimes providing this information can make it easier for faculty to apply for certain scholarships or fellowships to support you, e.g., from a private donor that has offered a fellowship only to students from certain backgrounds.
@@
 Many universities are well aware that incoming graduate students might not have all the technical background, so they often have a \emph{``bridge''} courses to help students catch up.
-For example, GMU has four bridge courses corresponding to the four core areas above that incoming students can take to catch up on their CS knowledge.
+For example, Mason has four bridge courses corresponding to the four core areas above that incoming students can take to catch up on their CS knowledge.
 
@@
 For simplicity, I will assume the department has a 9-month stipend of \$30K and a 3-month summer of \$10K (a third of the 9-month stipend).
-I will also use GMU tuition rate of about \$15K/year for full-time study (which is quite cheap compared to private universities, e.g., Univ.
+I will also use Mason tuition rate of about \$15K/year for full-time study (which is quite cheap compared to private universities, e.g., Univ.
 of Chicago is a whooping \$70K) and a 58.9\% rate on \emph{indirect cost}, which is a typical rate charged for overhead or administrative costs (yes, after all, universities are businesses!).
@@
 % This section aims to provide some guidance on how to gain research experience as an undergraduate student.
-%This section is written together Didier, a PhD student at GMU who did his undergraduate study in Rwanda.
+%This section is written together Didier, a PhD student at Mason who did his undergraduate study in Rwanda.
 
@@
     \tcblower
-    At GMU we pay undergrads about \$15/hr and about 10 hrs/week. This is flexible and can be increased as needed.
+    At Mason we pay undergrads about \$15/hr and about 10 hrs/week. This is flexible and can be increased as needed.
 \end{commentbox}
@@
 Plagiarism and cheating (e.g., exams and assignments) are a BIG no-no in the US.
-If you're caught cheating, you will face serious consequences and likely be expelled from the university (e.g., after the second time at GMU).
+If you're caught cheating, you will face serious consequences and likely be expelled from the university (e.g., after the second time at Mason).
 This is quite different from many international countries where cheating is common and often tolerated.
@@
 With your professor(s), you can call them by their first name (\autoref{sec:address}), disagree with them and argue (and gain respect doing so), seek their help (even on personal matters), come to their houses for parties or gathering (e.g., my lab always come to my house for \href{https://photos.app.goo.gl/LFtbqQUuznq9eiL7A}{Thanksgiving}), and give them small thoughtful gifts that they put on their desks (\autoref{sec:gifts}).
-Many people maintain lifelong relationships with their professors and colleagues, staying in touch through cards, emails, and visits, even after they nolonger work together.
+Many people maintain lifelong relationships with their professors and colleagues, staying in touch through cards, emails, and visits, even after they no longer work together.
 
@@
     I maintain a close relationship with my former professors and mentors.
-When there is a new event in my life (or theirs), I often send them an email or call them, e.g., when I get married, have a new baby, new job, etc.
+When there is a new event in my life (or theirs), I often email them or call them, e.g., when I get married, have a new baby, new job, etc.
 When I was having issues at UNL and considering leaving, I talked to my PhD advisers Deepak and Steph.
-I think this does not bother them a bit and that they are actually interested in these ``dramas'' in my life.
+I think this does not bother them a bit; they are genuinely interested in knowing and helping solve these ``dramas'' in my life.
 
@@
 Congrats!''}) would suffice.
-They will appreciate it and you never know when you might need their help.
+They will appreciate it, and you never know when you might need their help.
 
@@
 
-\paragraph{History} This document was conceived during a lunch with a faculty at GMU.
-We talked about why GMU was not able to attract good Vietnamese and other international students, despite having a much stronger CS program than many schools that these students want to go to (part of the reason is described in \autoref{sec:selecting-ranking-schools}).
+\paragraph{History} This document was conceived during a lunch with a faculty at Mason.
+We talked about why Mason was not able to attract good Vietnamese and other international students, despite having a much stronger CS program than many schools that these students want to go to (part of the reason is described in \autoref{sec:selecting-ranking-schools}).
 We wish there were a way for international students to know about the US PhD programs (also for US faculty to understand more about international students and therefore have a better chance of recruiting and working with them).
@@
 \paragraph{Acknowledgement} Many people have contributed to this work.
-Profs. Craig Yu (GMU), Hakan Aydin (GMU), 
-Xiaokuan Zhang (GMU), Hung Le (UMass), and Deepak Kapur (UNM) provided valuable input in the early version.
-Other GMU faculty members also have provided feedback and contributions.
-Many students including Didier (GMU), Thanh (Melbourne), and Dat (Melbourne) have contributed valuable questions and feedback.
+Profs. Craig Yu (Mason), Hakan Aydin (Mason), 
+Xiaokuan Zhang (Mason), Hung Le (UMass), and Deepak Kapur (UNM) provided valuable input in the early version.
+Other Mason faculty members also have provided feedback and contributions.
+Many students including Didier (Mason), Thanh (Melbourne), and Dat (Melbourne) have contributed valuable questions and feedback.
 

### 2024-06-24 demystify.tex
updates

@@
 
-\paragraph{VEF 2.0} For Vietnamese students, it's worth mentioning the \href{https://vef2.org/}{VEF2.0 program}, which has helped many students in gaining admission to top CS PhD programs.
+\paragraph{VEF 2.0} For Vietnamese students, it's worth mentioning the \href{https://vef2.org/}{VEF2.0 program}, which has helped many students gain admission to top CS PhD programs in the US.
 VEF2.0 invites US faculty members from leading institutions to conduct rigorous interviews with VEF students and subsequently provide reference letters on their behalf.
-Despite the limited interaction between the interviewer and the interviewee (primarily just the interview), these reference letters are generally effective and have helped many students get into top PhD programs in the US.
+Despite the limited interaction between the interviewer and the interviewee (primarily just the interview), these reference letters are generally effective and have helped many students get into good PhD programs.
 
@@
 
-\textbf{Advice}: If you haven't taken the GRE or have bad scores, then don't waste time (re)taking it. It does not help much!
+Thus, if you haven't taken the GRE or have bad scores, then don't waste time (re)taking it. It does not help much!
 
@@
 
-\textbf{Advice}: Just do well enough to pass the minimum requirement set by the university. %, which is required for TA (\autoref{sec:funding}). 
+You should just do well enough to pass the minimum requirement set by the university. %, which is required for TA (\autoref{sec:funding}). 
 Just as with grades (\autoref{sec:your-school}) and GRE, having high scores in these tests might not help, but having too low scores can be a \red{red flag} and sometimes results in an automatic rejection (\autoref{sec:why-rejected}), e.g., below the minimum requirement.
@@
 
-\textbf{Advice}: do not spend much time talking about non-research-related job in your SOP.
+Do not spend much time talking about non-research-related job in your SOP.
 However, if you have research-based work experiences, then they can be useful and should be mentioned in your SOP and LORs.
@@
 
-\textbf{Advice}: Prepare your CV in such a way that it allows the reviewers to quickly scan to identify major achievements, e.g., Publications, Programming Competition Awards, and Teaching Experience.
+Prepare your CV in such a way that it allows the reviewers to quickly scan to identify major achievements, e.g., Publications, Programming Competition Awards, and Teaching Experience.
 
@@
 
-\textbf{Advice}: Treat the interview as an informal chat.
-Have elevator pitches about your research experiences and interests.
+You should treat the interview as an informal chat.
+Have ``elevator pitches'' about your research experiences and interests.
 You might also want to have a 5-minute presentation about your research.
@@
 
-By around March you should hear back from most PhD programs that want to admit you.  In many cases you will not hear back from schools that do not want to admit you.
+By around March you should hear back from most PhD programs that want to admit you.  In many cases you will not hear back from schools that do not want to admit you (see \autoref{sec:late-rejection}).
 If you receive offers, congratulations!
@@
 First, the amount of stipend \emph{varies} and depends on factors such as location (e.g., a stipend in the DC area is likely higher than in Lincoln, Nebraska due to higher living costs).
-Second, an academic year (AY)  year is typically \emph{9-month} in the US, so the stipend is for 9 months (so divide by 9 for each month).
+Second, an academic year (AY)  year is typically \emph{9-month} in the US, so the stipend is for 9 months.
 Third, you might get paid over the summer (\autoref{sec:summer-funding}) through funding from your professor or fellowship (typically no TA over the summer).
@@
 \paragraph{How to get TA?}  Unless you have other funding such as RA or fellowships, TA is typically the default for CS PhD programs.
-When you apply to be a full-time student,  state that you need financial assistant.
-Typically, adcom will either admit you and give you GTA, or reject you; i.e., we do not admit a student without supporting them.
+In your application, indicate that you need financial assistant.
+Typically, adcom will either admit you and give you TA, or reject you.
+We do not admit a student without supporting them.
 

### 2024-06-24 demystify.tex
updates

@@
 
-\paragraph{Acceptance Letters} Universities typically prioritize sending out acceptance letters first.
-This allows the admitted students to start planning and preparing for their studies.
+\paragraph{Acceptance Letters} Universities prioritize sending out acceptance letters first.
+This allows the admitted students to make decision and plan for their studies.
 These letters go to the top candidates who are offered admission directly without being placed on a waitlist.
+
 Some universities have rolling admissions, meaning they send out acceptance letters as they review applications (e.g., some universities sent out letters as soon as early February).
@@
 
-\paragraph{Response Deadlines}: Accepted students are usually given a strict deadline to respond to their offers (e.g., whether they will accept the offer), often around \textbf{April 15th} in the US.
+\paragraph{Response Deadlines} Accepted students are usually given a strict deadline to make decision on their offers, often around \textbf{April 15th} in the US.
 This is a standard deadline for PhD programs due to the Council of Graduate Schools (CGS) Resolution.
@@
 
-\paragraph{Waitlist}: If the CS program has X slots, then many strong candidates will be waitlisted.
+\paragraph{Waitlist} Most CS programs have a limited number of slots for PhD students, and thus put many students on a waitlist.
 If accepted students decline the offer, then offers are sent to students on the waitlist.

### 2024-06-24 demystify.tex
updates

@@
 
-\paragraph{Acceptance Letters} Universities prioritize sending out acceptance letters first.
-This allows the admitted students to make decision and plan for their studies.
-These letters go to the top candidates who are offered admission directly without being placed on a waitlist.
+\paragraph{Acceptance Letters} Universities prioritize sending out acceptance letters first. This allows the admitted students to make decision and plan for their studies.
 
@@
 
-\paragraph{Response Deadlines} Accepted students are usually given a strict deadline to make decision on their offers, often around \textbf{April 15th} in the US.
-This is a standard deadline for PhD programs due to the Council of Graduate Schools (CGS) Resolution.
-After this date, programs can gauge how many slots remain unfilled.
+\paragraph{Response Deadlines} Accepted students are usually given a deadline to make decision on their offers, often around around \textbf{April 15th} in the US.
+This is a standard deadline for PhD programs due to the Council of Graduate Schools (CGS) Resolution (\href{https://cgsnet.org/wp-content/uploads/2024/01/CGS_April15_Resolution_Jan312024.pdf}{see more).
+After this date, CS programs can gauge how many slots remain unfilled.
 

### 2024-06-29 demystify.tex
updates

@@
 
-\paragraph{How to help your LOR writer} Your writer can ask you to provide additional information to help them write your letter.
-For example, they might ask about GPA, research and work experiences, papers (if any), or anything you want them to mention.
-You can also provide them with a draft of your SOP so that they can see what you are saying about yourself and can complement that with their own perspective.
-It might also be a good idea to provide them with such information even when they do not ask, especially if you have not interacted with them much or have not done much research with them.
-
 \paragraph{Self-written Letters} Many international students write letters themselves, typically due to the requests of their profs., and have them signed by their profs.
@@
 Instead of the reference writer talking about you, in this case it is you who write about yourself (and they just sign the letter).
+Moreover, experienced reviewers can easily spot (or suspect) such letters, which can be a \red{red flag}.
+
+\paragraph{How to help your LOR writer} As mentioned above, do not write your own letter.
+Instead, you can help your LOR writer by providing details or unique things about yourself.
+For example, let them know about your GPA, research and work experiences, papers (if any), or anything you want them to mention.
+You can also provide them with a draft of your SOP so that they can see what you are saying about yourself and can complement that with their own perspective.
+Sometimes your writer will explicit ask you for such information, but if not, you should provide it anyway (especially if you have not interacted with them much or have not done much research with them).
 
@@
 \paragraph{Response Deadlines} Accepted students are usually given a deadline to make decision on their offers, often around around \textbf{April 15th} in the US.
-This is a standard deadline for PhD programs due to the Council of Graduate Schools (CGS) Resolution (\href{https://cgsnet.org/wp-content/uploads/2024/01/CGS_April15_Resolution_Jan312024.pdf}{see more).
+This is a standard deadline for PhD programs due to the Council of Graduate Schools (CGS) Resolution (\href{https://cgsnet.org/wp-content/uploads/2024/01/CGS_April15_Resolution_Jan312024.pdf}{see more}).
 After this date, CS programs can gauge how many slots remain unfilled.
 
-\paragraph{Waitlist} Most CS programs have a limited number of slots for PhD students, and thus put many students on a waitlist.
+\paragraph{Waitlist} Most CS programs have a limited number of slots for PhD students, and thus put many good students on a waitlist.
 If accepted students decline the offer, then offers are sent to students on the waitlist.
@@
 Thus, rejection letters are often sent out late (e.g., after April 15th or even much later).
+Not much you can do here, you can try to contact the school to ask about your status, but they might not reply, they might say they are still reviewing applications, or give you inaccurate information (e.g., you will hear in two weeks -- which might not happen).
+In short, you just have to wait.
 
@@
 
-For each application, the Mason application system compiles a single PDF including a summary, transcripts, test scores, LoRs, a CV, SOP, and writing samples.
+For each application, the Mason application system compiles a single PDF file, which consists of a summary, transcripts, test scores, LoRs, a CV, SOP, and writing samples.
 I usually start with the \emph{summary}, checking for low GPAs or test scores below the university minimum (\autoref{sec:standard-tests}).
@@
 I also recommend very strong candidates for the University Presidential fellowship.
-The adcom chair then reviews all evaluations, looking at entered notes, and ask reviewers to discuss and resolve discrepancies to reach a consensus.
-Typically there is no involvement from the adcom chair, department chair, or anyone else (i.e., the decision is made by the reviewers).
+The adcom chair then reviews all evaluations, looking at entered notes, and ask reviewers to discuss and resolve discrepancies to reach a consensus (e.g., a reviewer wants to accept but the other wants to reject).
+Typically, the decision is made entirely by the reviewers; there is no involvement from the adcom chair, department chair, or others.
 
@@
   \tcblower
-  \textbf{Vu}: Yes, for some reason I enjoy addressing you as ``Prof.'' (without appending a last or first name).  The use of Prof. Vu may have foreshadowed my future in academia.
+  \textbf{Vu}: Yes, for some reason I enjoy calling you ``Professor'' (without appending a last or first name).  The use of Prof. Vu may have foreshadowed my future in academia.
 \end{commentbox}
@@
 
-Here are the typical steps: (i) a faculty suspecting a cheating case will report it to the Office of Academic Integrity (OAI) at the university---the report often has supporting evidence and suggested penalty (e.g., a failing grade);  (ii)  OAI will take over and investigate the case; and (iii) OAI will make the final decision.
-It is important to note that after receiving the report from your prof., OAI \emph{completely} takes over and makes decisions.
-This means begging your professor will not help because they simply are no longer involved in the case.
+Here are the typical steps: (i) a faculty suspecting a cheating case will report it to the Office of Academic Integrity (OAI) at the university---the report often has supporting evidence and suggested penalty (e.g., a failing grade);  (ii)  OAI will completely take over and investigate the case independently (the prof.
+is not involved); and (iii) OAI will make the final decision.
+It is important to note that after receiving the report from your prof., OAI \emph{completely} takes over and makes its own decision.
+This means begging your professor will not help because they simply are no longer involved in the case and cannot do much.
 
@@
 I meet Thang Bui (my MS adviser) at least once a year when I come back to Harrisburg to visit my parents.
-When Steph was in DC for a meeting, I invited her to Mason to give a talk on her work.
+When Steph was in DC for a meeting, I invited her to to give a research talk at Mason.
 I have also collaborated with them after I graduated e.g., I am working with Deepak on an NSF proposal on invariant generation.

### 2024-07-07 demystify.tex
updates

--- 2024-06-24 demystify.tex
@@
 % However, tech companies including Google, Microsoft, Facebook have fellowships that international students can apply for.
-Prestigious external fellowships typically require a clear and good research plan, so it might be a good idea to wait until at least your second year to have research experience and even publication before applying.
+Prestigious external fellowships typically require a clear and good research plan (the GRFP also requires  outreach activities and plans).
+So it might be a good idea to wait until your second year to have research experience and even publication before applying.
 Remember, you're competing with the top PhD students at top universities worldwide.
@@
 In general, fellowships, especially those that are open to everyone, are highly competitive and prestigious, and you will stand out if you get one.
-Every PhD student has pubs, but only a few would have NSF Graduate (GRFP) or Microsoft fellowships.
-In fact, these are so prestigious that even if you didn't get it but make it to the final round or even \emph{``honorable mentioning''}, you school will still mention you on their website, and you should put it on your CV.
+Every PhD student has pubs, but only a few would have NSF Graduate (GRFP\footnote{\url{https://www.alexhunterlang.com/nsf-fellowship} is a good starting place for the GRFP with lots of proposal examples.}) or Microsoft fellowships.
+In fact, these are so prestigious that even if you didn't get it but make it to the final round or even \emph{``honorable mentioning''}, your school will still mention you on their website, and you should put it on your CV.
 
@@
 \paragraph{Response Deadlines} Accepted students are usually given a deadline to make decision on their offers, often around around \textbf{April 15th} in the US.
-This is a standard deadline for PhD programs due to the Council of Graduate Schools (CGS) Resolution (\href{https://cgsnet.org/wp-content/uploads/2024/01/CGS_April15_Resolution_Jan312024.pdf}{see more).
+This is a standard deadline for PhD programs due to the Council of Graduate Schools (CGS) Resolution (\href{https://cgsnet.org/wp-content/uploads/2024/01/CGS_April15_Resolution_Jan312024.pdf}{see more}.
 After this date, CS programs can gauge how many slots remain unfilled.

### 2024-07-07 faculty.tex
updates

@@
 Some services are required (e.g., when your department chair assigns you to some committee), but some are not and you can decline (e.g., an invitation to be a PC member).
-You should be strategic about which services you accept.
-
-\section{Service to your Research Community}
+
+\section{Service to Research Community}
 
@@
 They will help you to get to know people in your field, and to get your name out.
-Common examples include serving on the program committees of conferences, serving on editorial boards of journal, and serving on grant panels.
+Common examples include serving on the program committees (PC) of conferences, serving on editorial boards of journal, and serving on grant panels (e.g., from NSF).
 
@@
 
-\subsection{Volunteer for more Leadership roles} Once you have served on a few PCs and NSF panels, you should aim for more leadership roles.
+\subsection{Volunteer for Leadership roles} Once you have served on a few PCs and NSF panels, you should aim for more leadership roles.
 For example, you can volunteer to be a chair of certain sessions/tracks of a large conference, or even a general chair of a workshop or conference.
@@
 \subsection{Personal Experience} In my case I mostly cold email people for opportunities.
-For example, I directly email OOPSLA and PLDI PC chairs and asked, and they accepted my request (I did have papers at OOPSLA and PLDI when I asked).
+For example, I directly email OOPSLA and PLDI PC chairs and asked, and they accepted my request.
 I tried this with SE conferences  e.g., ICSE (and also nominate myself through the official form), and it never works, even though I have many papers at ICSE.
@@
 
-I felt that I didn't get much out of serving on PCs and didn't really like the review process much. However, I still do it because I feel it is my responsibility give back to the community.
+I felt that I didn't get much out of serving on PCs and didn't really like the review process much.
+However, I still do it because I feel it is my responsibility, i.e., people spent time reviewing my papers, so I should also volunteer my time to review theirs.
 
@@
 I have also volunteered for leadership roles, e.g., I directly asked the general chair of ICSE'21 to see if I can take more leadership roles and was then invited to serve as a proceeding co-chair.
-In hindsight this particular role is a lot of work and I didn't really like many aspects of it, but it was an eye-openning experience.
+In hindsight this particular role is a lot of work, and I didn't really like many aspects of it, but it was an eye-openning experience.
 
@@
 Your department chair will assign you to some committees to help run the department.
-Typically for tenure track faculty the services are lighter than tenured ones (e.g., less leadership roles).
+For tenure-track faculty the services are often lighter than tenured ones (e.g., less leadership roles).
 Common committees include PhD and/or MS admission, faculty search, and colloquium committees.
 
-Depends on your preference and how you are perceived, you can take more leadership roles when you are close to tenure, such as being the chair of a committee or chair/director of a program.
-These roles are more time consuming and might involve some politics, but they are also more rewarding and can help your tenure case.
+You can also take more leadership roles when you are close to tenure, such as being the chair of a committee or chair/director of a program.
+These are more time-consuming and might involve some politics, but they are also more rewarding and can help your tenure case.
 
@@
 
-A good department will try to match you with committees that can help junior faculty build their lab and tenure case (and shield you from committees that are not as useful).
+A good department/chair will try to match you with committees that can help junior faculty build their lab and tenure case (and shield you from committees that are not as useful).
 
 \subsection{Personal Experience} Since coming to GMU, I have been very active in serving the department.
-I share the vision of the department and want to help achieve it.
 In addition to the typical services load, I volunteer to serve in additional committees and leadership roles.
 For example, I maintain the CSRankings DB for the department, organize the Open House to recruit admitted students, participate in the Dept Chair renewal committee, etc.
-Even though I am not on the search committee, I meet with every faculty candidate (I talked to 22 people in Spring'23, which was my study break) and help recruit strong ones.
+Even though I am not on the search committee, I meet with every faculty candidate (I talked to 22 people in Spring'23, which was during my study break).
 I also agree to be director of the MS SWE program, which is a lot of work and typically done by tenured faculty.
@@
 Occasionally I feel overwhelmed by having many services, but I also feel that I am making a difference in the department and that is rewarding.
-I am also building a good relationship with my colleagues (and also head butt with some), and that is important for my tenure case.
-In addition, I learn a lot from being in various roles (e.g., how class scheduling work, how the department is run, how to recruit students and recruit faculty, how to run a colloquium, etc).
-
+I am also building a good relationship with my colleagues (and also head-butt with some), and that is important for my tenure case.
+But of course I don't do this because of my tenure case, I do it because I enjoy it and believe I can make a difference.
+In addition, I learn a lot from being in various roles (e.g., how class scheduling work, how the department is run, how to recruit students and recruit faculty, how to deal with difficult people and situation, etc).
+
+In general, I was way more active than most junior faculty in the department.
+I was also a lot more involved in the department than I was in my previous department at UNL.
+Here I share the vision of the department and want to help achieve it.
 In short, if you want to make a difference, and \emph{feel comfortable} at handling additional load, you can volunteer for more services. Otherwise, you can just ``stay low'',
-do the required services and focus on your research and teaching.
+do the required services and focus on your research and teaching. This is probably the best strategy for most junior faculty.
 
 \section{Service to the University}
+In most cases, these are optional services and assistant professors often do not take any of them.
+Common examples include serving the faculty senate, scholarship committee, etc.
+The university often sends out emails asking for volunteers (so somewhat generic and you can safely ignore), and sometimes people (e.g., your dept chair or dean) will directly ask you to serve (in which case you will want to consider it).
+
+\subsection{Advice and Personal Experience} I only have limited experience with these services.
+I served in a couple of grants and scholarship committees, in which I review applications for scholarship and grants for students.
+I also help represent the college in serving as judge of the KimmyDuong foundation scholarship.
+These are in general time-consuming (e.g., the KimmyDuong scholarship reviews and interview process happen during weekends!), but they are also rewarding as you get to know students and help them.
+In general, I would recommend doing these services only if you have the time.
+However, sometimes they can be rewarding and fun, so you might want to try them out.
+
 \section{Other Types of Services}

### 2024-07-07 demystify.tex

--- 2024-06-29 demystify.tex
@@
 
-\paragraph{Response Deadlines} Accepted students are usually given a deadline to make decision on their offers, often around around \textbf{April 15th} in the US.
-This is a standard deadline for PhD programs due to the Council of Graduate Schools (CGS) Resolution (\href{https://cgsnet.org/wp-content/uploads/2024/01/CGS_April15_Resolution_Jan312024.pdf}{see more}).
+\paragraph{Response Deadlines} Accepted students are usually given a deadline to make decision on their offers, often around \textbf{April 15th} in the US.
+This is a standard deadline for PhD programs due to the Council of Graduate Schools (CGS) Resolution (\href{https://cgsnet.org/wp-content/uploads/2024/01/CGS_April15_Resolution_Jan312024.pdf}{see more}.
 After this date, CS programs can gauge how many slots remain unfilled.
@@
 Thus, rejection letters are often sent out late (e.g., after April 15th or even much later).
-Not much you can do here, you can try to contact the school to ask about your status, but they might not reply, they might say they are still reviewing applications, or give you inaccurate information (e.g., you will hear in two weeks -- which might not happen).
-In short, you just have to wait.
+Not much you can do here, you can try to contact the school to ask about your status, but they might not reply, they might say they are still reviewing applications, or give you inaccurate information (e.g., you will hear in two weeks).
+In short, you just have to be patient and wait.
 

### 2024-07-07 demystify.tex
updates

@@
 \paragraph{GRE} While a few schools still require taking the GRE (e.g., UCF), most good CS PhD programs in the US \textbf{no longer} require it.
-
-Thus, if you haven't taken the GRE or have bad scores, then don't waste time (re)taking it. It does not help much!
+Thus, if you haven't taken the GRE or have bad scores, then don't waste time (re)taking it.
+It does not help much!
 
@@
 
-\paragraph{English Tests (IELTS, TOEFL)} Unless your degrees are from specific countries such as \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-Mason#standard-tests-waiver-eligible-countries}{these}, you will need to
-take standardized English tests.
+\paragraph{English Tests (IELTS, TOEFL)} Unless your degrees are from specific countries such as \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-Mason#standard-tests-waiver-eligible-countries}{these}, you will need to take standardized English tests.
 
@@
 \begin{commentbox}[Vu:]
-  If you got recruited an offered an RA by a prof., you will likely get admitted.
-For example, if a prof., even if not in PhD adcom, wants to fund you, adcom will respect that decision and admit you (unless your application has many red flags).
+  If you got recruited and offered an RA by a prof., you will likely get admitted.
+For example, if a prof., even if not in PhD adcom, wants to fund you, adcom will likely respect that decision and admit you.
 \end{commentbox}

### 2024-07-08 faculty.tex
services

@@
 \chapter{Preface}
-Having been involved in PhD admissions for many years, I've
-realized that many \textbf{international students} (and also domestic ones), especially those in  smaller countries or less well-known universities, lack a clear understanding of
-the Computer Science PhD admission process at US universities. This confusion not only
-discourages students from applying but also creates the perception that
-getting admitted to a CS PhD program in the US is difficult compared to other countries.
-
-% though \emph{very} top schools could be very selective, e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU
-So I want to share some details about the admission process and advice for those who are interested in applying for a \textbf{PhD in Computer Science in the US}.
-While this document is primarily intended for students interested in CS, it might be relevant to students from various disciplines.
-Furthermore, although many examples are specifics for schools that I and other contributors of this document know about, the information should be generalizable to other R1\footnote{An \href{https://en.wikipedia.org/wiki/List_of_research_universities_in_the_United_States}{R1 institution} in the US is a research-intensive university with a high level of research activity across various disciplines.
-Currently, 146 (out of 4000) universities are classified as R1.} institutions in the US (and universities in other countries).
-
-In addition, this document can help \textbf{US faculty and admission committee} gain a better understanding of international students and their cultural differences.
-By recognizing and leveraging these differences, CS programs in the US can attract larger and more competitive application pools from international students.
-
-I wish you the best of luck. And if you follow this guidance, you will at least have a good chance at GMU (see
-\href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-GMU}{why
-  you want to study at here}). Happy school hunting!
+This documents my tenure track journey in computer science and the lessons I have learned along the way.
+It is a way for me to reflect on my experiences and share them with others who may be interested in pursuing a similar path.
+I hope that it will be helpful to those who are in their early stages of academia (e.g., tenure-track faculty), especially in computer science.
 
@@
 \begin{center}
-  \href{https://nguyenthanhvuh.github.io/phd-cs-us/demystify.pdf}{nguyenthanhvuh.github.io/phd-cs-us/demystify.pdf},
+  \href{https://nguyenthanhvuh.github.io/phd-cs-us/faculty.pdf}{nguyenthanhvuh.github.io/phd-cs-us/faculty.pdf},
 \end{center}
@@
 
-\subsection{Advice} To serve, you can directly ask people to recommend you.
+\subsection{How to get invited for these services?}
+
+You can directly ask people to recommend you.
 For example, for PC, ask senior colleagues or even your PhD advisor/collaborators to recommend you to the PC chairs.
@@
 
-\subsection{Personal Experience} In my case I mostly cold email people for opportunities.
-For example, I directly email OOPSLA and PLDI PC chairs and asked, and they accepted my request.
-I tried this with SE conferences  e.g., ICSE (and also nominate myself through the official form), and it never works, even though I have many papers at ICSE.
+\subsection{Personal Experience} I mostly cold email people for opportunities.
+For example, I directly email OOPSLA and PLDI PC chairs and asked, and they invited.
+I tried this with ICSE (and also nominate myself through the official form), and it never works, even though I have many papers at ICSE (that leaves a bad taste in my mouth about ICSE, but that is another story).
 I also got invited to serve on a PC because a senior colleague recommended me.
-For NSF, I emailed program managers and asked to serve on their panels.
-
-I felt that I didn't get much out of serving on PCs and didn't really like the review process much.
+For NSF, I just emailed program managers and asked to serve on their panels, and they invited me.
+
+I felt that I didn't get much out of serving on conf. PCs and didn't really like the review process much. 
 However, I still do it because I feel it is my responsibility, i.e., people spent time reviewing my papers, so I should also volunteer my time to review theirs.
 
-On the other hand, I thoroughly enjoy serving on NSF panels (though of course my proposals still got rejected frequently).
-While the work is more intense and takes more time, I found it more rewarding and enjoy the discussions.
-Since my first NSF panel in 2019, I have served 8 panels (at least once every year since '19) and I look forward to serving more.
+On the other hand, I thoroughly enjoy serving on NSF panels (though of course my proposals still got rejected very often).
+To this date I have served on 8 NSF panels (at least once every year since 2019) and I look forward to serving more.
+While NSF review is more intense and takes more time (compared to conf.
+reviews), I found it more rewarding and enjoy the discussions.
 
 I have also volunteered for leadership roles, e.g., I directly asked the general chair of ICSE'21 to see if I can take more leadership roles and was then invited to serve as a proceeding co-chair.
-In hindsight this particular role is a lot of work, and I didn't really like many aspects of it, but it was an eye-openning experience.
+In hindsight this particular role is a lot of work, and I didn't really like some aspects of it, but it was an eye-openning experience.
 
 \section{Service to the Department}
-Your department chair will assign you to some committees to help run the department.
-For tenure-track faculty the services are often lighter than tenured ones (e.g., less leadership roles).
+Your department chair will assign you to some committees to help run the department.  For tenure-track faculty the services are often lighter than tenured ones (e.g., less leadership roles).
+
 Common committees include PhD and/or MS admission, faculty search, and colloquium committees.
+Serving in the PhD adcom is very common for new assistant professors, as it is a good way to recruit students for your lab.
+Serving on the faculty search committee is also a good way to recruit new faculty members that you like (and also let you know what the job market is like, what is hot and not).
+Finally, being on the colloquium committee is a good way to invite speakers in your field that can be your \emph{tenure letter writers}.
 
@@
 These are more time-consuming and might involve some politics, but they are also more rewarding and can help your tenure case.
-
-\subsection{Advice} You should try to be strategic about which committees you serve on that can help build your lab and tenure case.
-For example, serving in the PhD adcom is very common for new assistant professors, as it is a good way to recruit students for your lab.
-Serving on the faculty search committee is also a good way to recruit new faculty members that you like (and also let you know what the job market is like, what is hot and not).
-Serving on the colloquium committee is also a good way to invite speakers in your field that can be your \emph{tenure letter writers}.
 
@@
 \section{Service to the University}
-In most cases, these are optional services and assistant professors often do not take any of them.
+You can also serve the university in various committees.
 Common examples include serving the faculty senate, scholarship committee, etc.
@@
 
-\subsection{Advice and Personal Experience} I only have limited experience with these services.
+\subsection{Personal Experience} I have limited experience with these services.
 I served in a couple of grants and scholarship committees, in which I review applications for scholarship and grants for students.
-I also help represent the college in serving as judge of the KimmyDuong foundation scholarship.
-These are in general time-consuming (e.g., the KimmyDuong scholarship reviews and interview process happen during weekends!), but they are also rewarding as you get to know students and help them.
-In general, I would recommend doing these services only if you have the time.
-However, sometimes they can be rewarding and fun, so you might want to try them out.
+I also help represent the college in serving as judge of the Kimmy-Duong foundation scholarship.
+These are time-consuming (e.g., the KimmyDuong scholarship reviews and interview process happen during weekends!), but they are also rewarding as you get to know students and help them (many have unique and interesting stories).
+
+In general, I would recommend doing these services only if you have the time and have interest in them. However, sometimes they can be rewarding and fun, so you might want to try them out.
 
 \section{Other Types of Services}
-\paragraph{Writing Recommendation Letters} You will be asked to write recommendation letters for both students and colleagues.
-
-\textbf{Advice}: %you should have a template for recommendation letters.
-You can find many templates online.
-You should also ask your colleagues to share their templates with you.
-You should also ask your colleagues to share their recommendation letters with you.
-You can learn a lot from reading other people's recommendation letters.
+
+\paragraph{Writing Recommendation Letters} You will be asked to write recommendation letters for both students and sometimes even colleagues.
+For example, I write on average ten reference letters for students each year, including letters for grad schools, jobs, or scholarships.
+I view writing letters as my responsibility as a faculty member, and so I usually do not say no to any student who asks for a letter.
+However, I would tell a student I do not know well that I can only write a generic  (i.e., weak) letter for them.
+
+\section{Conclusion}
+
+Many would advise about being strategic about your services.I think it is a good advice as you can get more out of your services.
+However, I do not follow this, and in many cases I would volunteer for tasks that no one wants to do because I feel that someone gotta do it and I can help out.
 

### 2024-07-08 faculty.tex
updates

@@
 
-Many would advise about being strategic about your services.I think it is a good advice as you can get more out of your services.
-However, I do not follow this, and in many cases I would volunteer for tasks that no one wants to do because I feel that someone gotta do it and I can help out.
+Many would (rightly) advise about being strategic about your services.
+I do not really generally this and just volunteer for services that I think it is my responsibility (e.g., being in PCs, though I never liked them much) and those that I enjoy (NSF panels, services to the depts).
+In many cases I would also volunteer for tasks that no one wants to do because I feel that someone gotta do it and I can help out, e.g., various misc tasks in the dept.
 
@@
 
-Toward the end of your 3rd year, you will need to prepare a renewal package and submit it to your department.
+Toward the end of your 3rd year (or early 4th year), you will need to prepare and submit a renewal package.
 The department will then review your package and vote on whether to renew you for your 4th (and perhaps also 5th year).
-Your next renewal is essentially your tenure application.
-The department will vote on whether to recommend your renewal.
-
-Similarly, after 5 or 6 years, it is time for you to go up (or apply) for tenure.
-You will need to prepare a tenure package and submit it to your department.
+Many faculty members actually do not get renewed, so do take this process seriously.
+
+Then, after 5 or 6 years, it is time for you to go up (or apply) for tenure.
+You will need to submit a tenure package.
 The department will then review your package and vote on whether to recommend you for tenure.
 If the department votes in your favor, the package will then go to the college, and then to the university (e.g., provost, president, board of visitors, etc).
-The university will then make the final decision on whether to grant you tenure.
+The university will then make the final decision on whether to grant you tenure (typically they follow the department's recommendation, but not always).
+
+I will focus on the tenure application process, as the renewal process is similar but less intense (e.g., no external letters).
 
@@
 \section{Timeline}
-Typically you will first mention to your department chair that you are planning to go up for tenure.
+
+If you plan to apply for tenure in year X (and get tenure in year X+1), you will first mention your intention to the dept chair (at GMU this is at least a semester before, i.e., in December of year X-1).
 The chair might then assign you a mentor who will help you with the process.
-You will apply for tenure in the Fall semester, the decision will be made in the next Spring semester, and you will get tenure in the Fall semester after that.
+You will apply for tenure in the summer or early Fall of year X, the decision will be made in the next Spring semester of year X+1, and if everything goes well, you will get tenure in the Fall semester in year X+1.
 
@@
   \item Fall 2021: Moved to GMU with accelerated tenure clock (meaning I don't have to restart from the beginning as I have already been at UNL since 2016).
-  \item (late) Fall 2022: I told my department chair that I plan to apply for tenure in 2023 (which means if things go well then I get tenure in Fall'24).
-  \item Spring 2023: I started preparing my tenure package, which includes writing statements and think reference letters.  I also took a semester off because GMU has a pre-tenure study break policy.
-  \item Summer 2023: I submitted my tenure package.
-  \item Fall 2024: I gave a tenure talk. Department voted in my favor, and then the college voted in my favor, and the decision goes up to the university.
-  \item Spring 2024: In May, I got the official notification that I got tenure and promoted to Associate Prof., beginning in Fall 2024.
+  \item (late) Fall 2022: Told my dept. chair that I plan to apply for tenure in 2023.
+  \item Spring 2023: Prepared my tenure package, which includes writing statements and think reference letters.
+Also took this semester off (no teaching no services) because GMU has a pre-tenure study break policy.
+  \item Summer 2023: Submitted my tenure package.
+  \item Fall 2024: Gave a tenure talk. Department voted in my favor, and then the college voted in favor, and the decision goes up to the university.
+  \item Spring 2024: In May, got the official notification that I got tenure and promoted to Associate Prof., beginning in Fall 2024.
 \end{itemize}
@@
 
-GMU (and likely other universities) has a very strict and specific format for the CV (e.g., what to include, how to format, etc).
-Fortunately I have been using this format since I joined GMU and also kept my CV up-to-date, so I didn't have to do much.
-I did have to add a few more sections (e.g., teaching evaluation and number of students in courses taught).
-I also have to make sure that the CV is consistent with my statements and other parts of the package (e.g., funding amount must match with what the university has in their system).
-
-I also looked at CVs of other people at GMU who recently got tenure. I also sent my CV to  my mentors and got some feedback.
+Both GMU and UNL ( and likely other universities) have a very strict and specific format for the CV, e.g., what to include, how to format, order of sections and subsections, etc.  
+Fortunately I have always been using the university CV format and also kept my CV up-to-date, so I didn't have to do much. 
+I did have to add more \emph{personal} details (e.g., teaching evaluation and funding amounts that match what the university has in their system).
+
+I also looked at CVs of other people in the dept.
+who recently got tenure.
+I also sent my CV to  my mentors and got many feedback.
+In fact, I also asked a Division Dean in the College who is in charge of P\&T procedure to review my CV for formatting and other issues.
+In short, the CV is one of the easiest part of the tenure package, and so I want to make sure mine has no silly mistakes.
 
@@
 At GMU, the statement is limited to 8 pages.
-My statement follows the standard structure and talks about my achievements and plans in (i) research and funding, (ii) advising and mentoring, (iii) teaching, and (iv) services.
-The main unique part of my statement is the first page where I highlight and summarize my achievements (e.g., papers, fundings, students, etc).
-I get most of these information and numbers from my CV.
-
-While preparing the statement was the most time-consuming part of my tenure package, it didn't take as long (may be just a couple of days).
-I get statements from several colleague at GMU who recently got tenure and a couple of people at other universities to get some ideas on how it should be structured.
+My statement talks about my achievements and plans in (i) research and funding, (ii) advising and mentoring, (iii) teaching, and (iv) services.
+This organization is somewhat standard, and the main unique part of my statement is the first page where I highlight and summarize my achievements (e.g., papers, fundings, students, etc).
+I got these information and numbers from my CV.
+
+While preparing the statement was the most time-consuming part of my tenure package, it wasn't too bad and may be just took a couple of days.
 I was able to reuse lots of text and contents from my job application statements (when I applied to GMU in 2021).
-I did not send it for feedback and also did not revise it much.
+I also got statements from colleagues at GMU who recently got tenure and a couple of people at other universities to get some ideas on what to write.
+I did not send my statement for feedback because I felt that it was good enough and I didn't want to bother people (reading an 8-page statement is a lot more work).
 

### 2024-07-08 faculty.tex
updates

@@
 In fact, I also asked a Division Dean in the College who is in charge of P\&T procedure to review my CV for formatting and other issues.
-In short, the CV is one of the easiest part of the tenure package, and so I want to make sure mine has no silly mistakes.
+In short, the CV is the easiest part of my tenure package, and so I want to make sure mine has no silly mistakes.
 
@@
 
+You will need to prepare a statement for your tenure package.
+Typically, the statement talks about your achievements and plans in (i) research and funding, (ii) advising and mentoring, (iii) teaching, and (iv) services.
+This order is somewhat standard, but you can change it if you want.
+In a sense, this statement is similar to your job application statement, but it is more focused on what you have done (and less on what you will do).
+However, unlike job application, tenure statement has a page limit (e.g., 8 pages at GMU).
+
 Here is my \href{https://dynaroars.github.io/people/nguyenthanhvuh/files/pt/statement-annotated.pdf}{tenure application statement}, annotated with comments.
-At GMU, the statement is limited to 8 pages.
-My statement talks about my achievements and plans in (i) research and funding, (ii) advising and mentoring, (iii) teaching, and (iv) services.
-This organization is somewhat standard, and the main unique part of my statement is the first page where I highlight and summarize my achievements (e.g., papers, fundings, students, etc).
+The unique part of my statement is the first page where I highlight and summarize my achievements (e.g., papers, fundings, students, etc).
 I got these information and numbers from my CV.
-
-While preparing the statement was the most time-consuming part of my tenure package, it wasn't too bad and may be just took a couple of days.
+While preparing the statement was the most time-consuming part of my tenure package, it wasn't too bad and just took a couple of days.
 I was able to reuse lots of text and contents from my job application statements (when I applied to GMU in 2021).
@@
 \subsection{Letters}
-This is a huge part of your tenure package.  You will need letters from senior people in your area.  Someone who can vouch for you.
-You will need 4--6 letters. However, you will not be the one asking for these letters, but your department will. You can suggest who to ask.
-
-Typically, there are \emph{two lists} of letter writers: one consisting of people suggested by you and the other consisting of people suggested by the department (i.e., your dept will come up with a list of people they think are good letter writers for you).
+You will need letters from senior people in your area.  Someone who can vouch for you.
+You will need 4--6 letters.
+However, you will not be the one asking for these letters, but your department will.
+You can suggest who to ask.
+This process doesn't take much of your time (because you don't write the letters) but it might cause you a lot of stress (because you don't have full control over it).
+
+Typically, your department will compile \emph{two lists} of letter writers: one consisting of people suggested by you and the other consisting of people suggested by the department (i.e., your dept will come up with a list of people they think are good letter writers for you).
+The people on the departmental list often "weight" more than those on your list.
 There can also be another list, the ``blacklist'' where you list people that you don't want to write letters for you.
 
-\subsubsection{Personal Experience}
+\begin{commentbox}[Vu:]
 Originally I asked several senior people to write my letters.
@@
 Moreover, GMU (and likely most other universities) would prefer writers not from my list (because people on my list are likely to be biased), and has a rule that at least 60\% of the letter writers must be from the department's list.
-Moreover, there's a good chance that a letter writer will decline the invitation, so there should be more names.
-
-After talking to various people in my dept and outside, I decide to put just 2 names in my list (0 on blacklist) and let the department decides the rest.
-I also thought about leaving my list empty but then that seems to be too weird, and I don't want to give the impression that I don't have any friends in the world.
-The reason is that I feel I don't have too many "enemies" out there that would write bad things about me.
+There's also a good chance that a letter writer will decline the invitation, so there should be more names.
+\end{commentbox}
+
+In my case I put just 2 names in my list (0 on blacklist) and let the department decide the rest.
+The reason that I only put 2 names is that I feel I don't have too many "enemies" out there that would write bad things about me.
 Those that don't know me well will likely just decline the letter invitation.
-Turned out to be a good decision because in the end the department selected 1 of the 2 names on my list and 4 more names from the department's list.
-Moreover, it looks like no one declines so my letter writers are all set fairly quickly in the process.
-All of my letters were good and give many positive comments about me as summarized in my evaluation letter (sent to me after the tenure decision).
+This also gives the department more choices on their list.
+Note I also thought about leaving my list empty, but then that seems to be too weird, and I don't want to give the impression that I don't have any "friends" in academic.
+
+The department selected 1 of the 2 names on my list and 4 more names from the department's list.
+Moreover, it appears that no one declines so my letter writers are all set fairly quickly in the process.
+All of my letters were very good and give many positive comments about me as summarized in my evaluation letter (sent to me after the tenure decision).
 I will never know who wrote the letters, but I am very grateful to them.
 
+\subsection{Tenure Talk}
+After you submit your tenure package, you might be asked to give a tenure talk.
+This is a talk where you summarize your research, teaching, and services, and future plans.
+The talk is typically 1 hour long, and there will be a Q\&A session afterward.
+The talk is open to the public.
+
 \subsection{Others}
 
-\paragraph{Tenure Visits} Many faculty will visit other universities to give talks and meet people.
+\paragraph{Tenure Visits} Many people will visit other universities to give talks and meet people.
 This is a good way to get your name out and to get letters (e.g., from the person that host you).

### 2024-07-08 faculty.tex
updates

@@
 
-Toward the end of your 3rd year (or early 4th year), you will need to prepare and submit a renewal package.
-The department will then review your package and vote on whether to renew you for your 4th (and perhaps also 5th year).
-Many faculty members actually do not get renewed, so do take this process seriously.
+Toward the end of your 3rd year (or early 4th year), you will need to submit a \emph{renewal} application.
+The department will then review your application and vote on whether to renew you for your 4th (and perhaps also 5th year).
+At some university this renewal is almost guaranteed, but at some other universities, it is not and people do get denied.
 
@@
 
-Both GMU and UNL ( and likely other universities) have a very strict and specific format for the CV, e.g., what to include, how to format, order of sections and subsections, etc.  
-Fortunately I have always been using the university CV format and also kept my CV up-to-date, so I didn't have to do much. 
+For P\&T (Promotion and Tenure), universities often have a very strict and specific format for the CV, e.g., what to include, how to format, order of sections and subsections, etc.  
+Fortunately I have always been using the university CV format and also kept my CV up-to-date, so I didn't have to do much (here's my most \href{https://dynaroars.github.io/people/nguyenthanhvuh/latex-cv/cv-nguyen.pdf}{recent CV}).
 I did have to add more \emph{personal} details (e.g., teaching evaluation and funding amounts that match what the university has in their system).
@@
 \subsection{Letters}
-You will need letters from senior people in your area.  Someone who can vouch for you.
-You will need 4--6 letters.
-However, you will not be the one asking for these letters, but your department will.
+You will need 4--6 letters from senior people in your area to vouch for your work.
+Your department will be ask people for these letters on your behalf.
 You can suggest who to ask.
-This process doesn't take much of your time (because you don't write the letters) but it might cause you a lot of stress (because you don't have full control over it).
+This process doesn't take much of your time (because you don't write the letters) but it might cause you a lot of stress (because you don't have much control over it).
 
@@
 
-In my case I put just 2 names in my list (0 on blacklist) and let the department decide the rest.
-The reason that I only put 2 names is that I feel I don't have too many "enemies" out there that would write bad things about me.
-Those that don't know me well will likely just decline the letter invitation.
-This also gives the department more choices on their list.
-Note I also thought about leaving my list empty, but then that seems to be too weird, and I don't want to give the impression that I don't have any "friends" in academic.
-
-The department selected 1 of the 2 names on my list and 4 more names from the department's list.
-Moreover, it appears that no one declines so my letter writers are all set fairly quickly in the process.
-All of my letters were very good and give many positive comments about me as summarized in my evaluation letter (sent to me after the tenure decision).
+I had just 2 names on my list (0 on blacklist) and let the department decide the rest.
+I feel I don't have too many ``enemies'' out there that would write bad things about me ( those that don't know me well will likely just decline the letter invitation instead of writing a bad letter).
+This also gives the department more choices on their list (recall that people on the dept.
+list often weight more).
+Note I also thought about leaving my list empty, but then that seems to be too weird. I don't want to give the impression that I don't have any ``friends'' in academic.
+
+The department selected 1 of the 2 names on my list and 4 names from the department's list.
+It appears that no one declines, so my letter writers were all set fairly quickly.
+All of my letters were very good and give positive comments about me as summarized in my evaluation letter (sent to me after the tenure decision).
 I will never know who wrote the letters, but I am very grateful to them.
 
+\section{Others}
 \subsection{Tenure Talk}
-After you submit your tenure package, you might be asked to give a tenure talk.
-This is a talk where you summarize your research, teaching, and services, and future plans.
-The talk is typically 1 hour long, and there will be a Q\&A session afterward.
-The talk is open to the public.
-
+After you submit your tenure package, you might be asked to give a tenure talk, which summarizes your research, teaching, and services, and future plans.
+The talk is open to the public, about 1 hour long, and has Q\&A session afterward.
+
+Here is my \href{https://dynaroars.github.io/people/nguyenthanhvuh/files/pt/slides.pdf}{talk}.
+I focus on my recent research direction on neural network verification and just briefly on my past work.
+Overall, I didn't have to spend too much time on preparing because I reused slides from my job and research talk and just added at the end a ``stats'' slide on my funding, pubs, teaching, services, etc.
+
+\subsection{Your ``Champion'' in the P\&T Meeting}
+
+You will need a senior faculty member (or ``champion'') from your department to present your case in the department P\&T meeting.
+This person will present your achievements and argue for you.
+Typically, this person is your mentor, but it can be anyone in the department that you trust and who knows your work well.
+A good champion won't likely be able to help a bad case, but a bad champion can hurt a good case.  So do pay attention to this.
+
+At GMU I was assigned a mentor, Paul Ammann, right when I joined.
+For my tenure case, he put together my ``stats'' page and asked me to review it.
+I also discussed with him about letter writers, etc.
+Overall, I think my case is relatively strong and so the whole thing was very ``smooth sail'' according to Paul.
+However, I did get much advice and guidance from him in other things e.g., how to run the MS SWE program (more on that later).
+
+\begin{commentbox}[Vu:]
+    While I have had not experienced with an ineffective mentor, I have heard about cases where the mentor did not do a good job, which can hurt strong candidates during the 4-th year renewal process.
+For example, a candidate might have few, but impactful high-quality publications, and the mentor did not emphasize this enough.
+In some case, an ineffective mentor actually turn a significant accomplishment into a negative, e.g., getting a 10-year Most Influential paper award is a big deal in certain fields, but the mentor did not emphasize this or allow it being downplayed by others (and allow idiotic comments from other faculty such as the work has been done 10 year ago).
+So do pay attention to this and make sure you have a good mentor who can defend you.
+If necessary, you can ask your chair to assign you a different mentor.
+
+\end{commentbox}
 \subsection{Others}

### 2024-07-08 faculty.tex
updates

@@
 It is also a good way to get your name out.
-Similarly to PCs, you should also try to serve on the editorial boards of top journals in your field.
-
-Serving on PC (or editor boards) essentially means reviewing papers.
-You will have to read papers and write reviews, and then argue for why a paper should be accepted or not.
-You will also have to discuss with other PC members (or editors) and come to a consensus.
-Typically the review process can last for several weeks, and there are also rebuttals and conditionally acceptances (which you should already be familiar with as authors).
+
+You should also try to serve on the editorial boards of top journals in your field.
+Similar to being in conf.
+PCs, this essentially means reviewing paper submissions.
+However, unlike PC reviews, you typically will not have to discuss and reach a consensus with other reviewers.
+The journal editor will make the decision based on the reviews.
 
@@
 
-\subsection{Personal Experience} I mostly cold email people for opportunities.
+\paragraph{Personal Experience} I mostly cold email people for opportunities.
 For example, I directly email OOPSLA and PLDI PC chairs and asked, and they invited.
@@
 
-\section{Service to the Department}
+\section{Service to the Department}\label{sec:services-dept}
 Your department chair will assign you to some committees to help run the department.  For tenure-track faculty the services are often lighter than tenured ones (e.g., less leadership roles).
@@
 
-\subsection{Personal Experience} Since coming to GMU, I have been very active in serving the department.
+\paragraph{Personal Experience} Since coming to GMU, I have been very active in serving the department.
 In addition to the typical services load, I volunteer to serve in additional committees and leadership roles.
@@
 Even though I am not on the search committee, I meet with every faculty candidate (I talked to 22 people in Spring'23, which was during my study break).
-I also agree to be director of the MS SWE program, which is a lot of work and typically done by tenured faculty.
-
-Occasionally I feel overwhelmed by having many services, but I also feel that I am making a difference in the department and that is rewarding.
+I also serve in the Chair Renewal committee and as the director of the MS SWE program, which is a lot of work and typically done by tenured faculty.
+
+Occasionally I do feel overwhelmed by having many services, but I share the vision of the department and want to help achieve it.
+I learn a lot from being in various roles (e.g., how class scheduling work, how the department is run, how to recruit students and recruit faculty, how to deal with difficult people and situation, etc).
 I am also building a good relationship with my colleagues (and also head-butt with some), and that is important for my tenure case.
-But of course I don't do this because of my tenure case, I do it because I enjoy it and believe I can make a difference.
-In addition, I learn a lot from being in various roles (e.g., how class scheduling work, how the department is run, how to recruit students and recruit faculty, how to deal with difficult people and situation, etc).
-
-In general, I was way more active than most junior faculty in the department.
-I was also a lot more involved in the department than I was in my previous department at UNL.
-Here I share the vision of the department and want to help achieve it.
+
+In general, I was way more active than most junior faculty in the department.  I was also a lot more involved in the department than I was in my previous department at UNL.
 In short, if you want to make a difference, and \emph{feel comfortable} at handling additional load, you can volunteer for more services. Otherwise, you can just ``stay low'',
@@
 
-\subsection{Personal Experience} I have limited experience with these services.
+\paragraph{Personal Experience} I have limited experience with these services.
 I served in a couple of grants and scholarship committees, in which I review applications for scholarship and grants for students.
@@
 
-\chapter{Renewal and Tenure Application}
-
-Toward the end of your 3rd year (or early 4th year), you will need to submit a \emph{renewal} application.
-The department will then review your application and vote on whether to renew you for your 4th (and perhaps also 5th year).
-At some university this renewal is almost guaranteed, but at some other universities, it is not and people do get denied.
-
-Then, after 5 or 6 years, it is time for you to go up (or apply) for tenure.
+\chapter{Tenure Application}
+
+After 5 or 6 years, it is time for you to go up (or apply) for tenure.
 You will need to submit a tenure package.
@@
 
-I will focus on the tenure application process, as the renewal process is similar but less intense (e.g., no external letters).
-
 %I will focus more on the tenure application process, as the renewal process is similar but less  intense.
@@
   \item Fall 2021: Moved to GMU with accelerated tenure clock (meaning I don't have to restart from the beginning as I have already been at UNL since 2016).
-  \item (late) Fall 2022: Told my dept. chair that I plan to apply for tenure in 2023.
-  \item Spring 2023: Prepared my tenure package, which includes writing statements and think reference letters.
+  \item (late) Fall 2022: Told my dept. chair that I plan to apply for tenure in Summer 2023.
+  \item Spring 2023: Prepared my tenure package, which includes writing statements and thinking about reference letter writers.
 Also took this semester off (no teaching no services) because GMU has a pre-tenure study break policy.
@@
 
-Here is my \href{https://dynaroars.github.io/people/nguyenthanhvuh/files/pt/slides.pdf}{talk}.
+Here is my \href{https://dynaroars.github.io/people/nguyenthanhvuh/files/pt/slides.pdf}{talk}, which took place in Fall'23 (after I submit my tenure package in Summer'23).
 I focus on my recent research direction on neural network verification and just briefly on my past work.
@@
 
-You will need a senior faculty member (or ``champion'') from your department to present your case in the department P\&T meeting.
-This person will present your achievements and argue for you.
-Typically, this person is your mentor, but it can be anyone in the department that you trust and who knows your work well.
+You will need a senior faculty member from your department to present your case in the department P\&T meeting.
+This person (i.e., your ``champion'') will present your achievements and argue for your case.
+Typically, the champion is your mentor, but it can be anyone in the department that you trust and who knows your work well.
 A good champion won't likely be able to help a bad case, but a bad champion can hurt a good case.  So do pay attention to this.
 
-At GMU I was assigned a mentor, Paul Ammann, right when I joined.
-For my tenure case, he put together my ``stats'' page and asked me to review it.
-I also discussed with him about letter writers, etc.
-Overall, I think my case is relatively strong and so the whole thing was very ``smooth sail'' according to Paul.
-However, I did get much advice and guidance from him in other things e.g., how to run the MS SWE program (more on that later).
+At GMU I was assigned a mentor when I joined.
+I discussed with him about letter writers, and he put together my ``stats'' page and asked me to review it.
+Overall, I think my case is relatively strong and so the whole thing was a ``smooth sail'' according my mentor.
 
 \begin{commentbox}[Vu:]
-    While I have had not experienced with an ineffective mentor, I have heard about cases where the mentor did not do a good job, which can hurt strong candidates during the 4-th year renewal process.
+    If a case is not strong enough (e.g., lack of funding) then the department chair and mentor will likely advise the candidate against applying for tenure and instead seek for extension to improve.
+I think this is a good practice to  ``save face'' for the candidate and avoid a negative tenure decision, in which case they likely have to find a new job.
+\end{commentbox}
+
+\subsection{Tenure Visits and Colloqium} Close to the tenure application (e.g., during the Spring), some tenure candidates will visit other universities to give talks and meet people. 
+Sometimes the candidate is assigned to be the chair of the colloquium in the department, and they can invite their potential tenure letter writers to give talks.
+These are all good ways to get your name out and to get letters.
+The traditional way is a physical visit, but since the pandemic, many of these visits are done remotely (which are more convenient for many people).
+
+I did not do any tenure visits or host anyone for the purpose of getting tenure letters.
+This was partly because I am not a travel person, and also I felt that it would not help much in my case; I believe that I have strong record and very good relationship with my colleagues and the dept.
+Instead, during the Spring'23 semester before my tenure application, I put lots of effort in helping the university in recruiting faculty (by meeting with 20+ faculty candidates), even though I am not in the search committee and really didn't have service requirements as I am on study break that semester.
+As mentioned in \autoref{sec:services-dept}, I didn't really do any of these for my tenure case, but because I enjoy the working env at GMU and want to help out.
+
+\section{3rd year Renewal Package}
+
+%As mentioned, a good dept chair or mentor will likely advise against applying for tenure unless they think your case is strong enough.
+This is to avoid a direct tenure deny, which can be very hard and bad for the candidate.
+%On the other hand, 4-th year renewal is unavoidable and often used as a ``warning'' to the candidate that they need to improve.
+
+Toward the end of your 3rd year (or early 4th year), you will need to submit a ``midterm'' \emph{renewal} application to renew your contract for the 4th year (and perhaps also 5th year).
+While somewhat similar to tenure evaluation, this is less intense and  does not require reference letters or presentation.
+%The department will then review your application and vote on whether to renew you for your 4th (and perhaps also 5th year). 
+Your mentor presents and argues for your case in the dept.
+meeting, but the decision is made within the dept.
+and usually does not go up to the college or university.
+It is still signed by the Dean, who often just follows the dept's recommendation.
+
+While the renewal is straightforward for most people, it can be surprise and shock to some.
+
+%I didn't have a good experience with the 4-th year renewal process at UNL, which is one of the reasons why I had to leave UNL.
+
+% \begin{commentbox}[Vu:]
+%     While I have not experienced with an ineffective mentor, I have heard about cases where the mentor did not do a good job, which can hurt strong candidates during the 4-th year renewal process.
 For example, a candidate might have few, but impactful high-quality publications, and the mentor did not emphasize this enough.
@@
 
-\end{commentbox}
-\subsection{Others}
-
-\paragraph{Tenure Visits} Many people will visit other universities to give talks and meet people.
-This is a good way to get your name out and to get letters (e.g., from the person that host you).
-Physical visits were not possible during the pandemic, but virtual visits are possible and in fact easier.
-However,  I did not do any tenure visits or give talks in the year leading up to my tenure application.
-This was because I am not a travel person and also I felt that it would not help much in my case.
-
-\paragraph{Inviting speakers to colloquium} Many departments have colloquium series where they invite speakers to give talks.  You can invite potential letter writers to your department's colloquium.
-
-In general, these are a good way to get to know people in your field and to get your name out. Many of these, especially tenure visits, are common practices for junior faculty applying for tenure.
-
-\section{Renewal Package}
+% \end{commentbox}
 

### 2024-07-08 faculty.tex
updates

@@
 Getting funding is arguably the most important part of getting tenure.
-Writing grant propsal is very different from writing research papers.
-However, without funding, you can't pay your students, conference registration and travel, or even your own summer salary.
-%In short, the lack of funding causes lost of stress, uncertainty, and threaten your career.
-Most new professors struggle with grant writing at first, but eventually get better at it.  In this chapter, we will discuss the various ways to obtain funding.
+without funding, you can't pay your students, conference registration and travel, or even your own summer salary.
+However, writing grant proposals is very different from writing research papers.
+Thus, most new professors struggle with grant writing at first, but eventually get better at it.
 
@@
 \section{Teaching Evaluation}
+
+At the end of each semester, students are asked to evaluate the instructor.
+The results are shared anonymously with the instructor to help them improve their teaching and course.
+Teaching evaluation is also commonly used in annual and tenure evaluation.
+
+One main issue with teaching evaluation is that not many students respond, despite multiple reminders and urging from the instructor.
+Moreover, the few that respond are usually the students who are unhappy and thus will give highly negative evaluation.
+This is the reason why many instructors do not like teaching evaluations, especially when it is used to evaluate their performance.
+
+A strategy that I found effective is giving the entire class a small bonus point if everyone completes the evaluation.
+The bonus is computed as the percentage of responses * the bonus (X).
+So if X is 0.5 and everyone responds, everyone gets 0.5 points.
+If 50\% respond, each gets 0.25 points.
+While X is a tiny number, I emphasize that it can make a difference for borderline grades.
+I don’t curve grades, so this is the only way for students to bump.
+I also emphasize that in my previous classes many students benefit from this.
+So that really triggers their interest.
+And also because evaluations are before final exams, everyone is incentivized to participate because they don’t know their final grades and don’t want to miss this opportunity.
+Not only the students do it, they also remind their classmates/friends to do it.
+
+I only employ this method recently, and response rate has been very good so far.
+For Spr’24, it was 90\% (25/28 students) and for F’23, it was 100\% (16/16 students).
+Before this, my response rate is about 50\%.
+
 \section{Teaching Philosophy}
@@
 
-\subsection{Tenure Visits and Colloqium} Close to the tenure application (e.g., during the Spring), some tenure candidates will visit other universities to give talks and meet people. 
+\subsection{Tenure Visits and Colloquium} Closer to the tenure application (e.g., during the Spring), some candidates will visit other universities to give talks and meet people. 
 Sometimes the candidate is assigned to be the chair of the colloquium in the department, and they can invite their potential tenure letter writers to give talks.
-These are all good ways to get your name out and to get letters.
-The traditional way is a physical visit, but since the pandemic, many of these visits are done remotely (which are more convenient for many people).
-
-I did not do any tenure visits or host anyone for the purpose of getting tenure letters.
-This was partly because I am not a travel person, and also I felt that it would not help much in my case; I believe that I have strong record and very good relationship with my colleagues and the dept.
-Instead, during the Spring'23 semester before my tenure application, I put lots of effort in helping the university in recruiting faculty (by meeting with 20+ faculty candidates), even though I am not in the search committee and really didn't have service requirements as I am on study break that semester.
-As mentioned in \autoref{sec:services-dept}, I didn't really do any of these for my tenure case, but because I enjoy the working env at GMU and want to help out.
+Both are good ways to get your name out and to get letters.
+The traditional way is a physical visit, but since the pandemic, many of these visits are done remotely, which is more convenient for many people.
+
+I did not do these visits, partly because I am not a travel person, and also I felt that I would not have problems finding letter writers.
+Instead, during the Spring'23 semester before my tenure application, I put lots of effort in helping the university in recruiting faculty (by meeting with 20+ faculty candidates), even though I am not in the search committee and really didn't have service requirements during this pre-tenure study break.
+This helps me build a good relationship with my colleagues in the dept.
 

### 2024-07-10 demystify.tex
updates

@@
 
-This section summarizes the main points of this guideline. This gives you an overview to decide which specific topics you want to explore more thoroughly.
+This section summarizes the main points of this guideline. It also gives you an overview to decide which specific topics you want to explore more thoroughly.
 
 \begin{enumerate}
-  \item Should you apply?
+  \item Should you apply to a PhD CS program in the US?
         \begin{itemize}
@@
           \item Applications are evaluated by the \emph{PhD Admission} (adcom) committee and each application is typically reviewed by three faculty members (\autoref{sec:evalapps}).
-          \item Individually faculty \emph{cannot directly admit} a student---so do not email and ask if you have a chance.
-However, faculty can \emph{advocate} for a student and therefore increase their admission chance---contact and introduce yourself (\autoref{sec:contact}).
+          \item Individually faculty member \emph{cannot directly admit} a student---so do not email and ask if you have a chance.
+However, faculty can \emph{advocate} for a student and therefore increase their admission chance---so do contact and introduce yourself (\autoref{sec:contact}).
          \end{itemize}
@@
         \begin{itemize}
-          \item The committee will look at various factors, but the most important ones are letters of recommendation (LORs),  statements of purpose (SOP), and research experience and ability, e.g., publications.
+          \item The committee will look at various factors, but the most important ones are letters of recommendation (LORs),  statements of purpose (SOP), and research experience and potential, e.g., publications.
           \item \emph{LORs are very important}, but only if they are personalized and talk about your research ability (\autoref{sec:lor}).
@@
           \item  \emph{Publications can greatly help}.
-Publications in good venues are concrete evidence that you have successfully engaged in research.
+Papers in good venues are concrete evidence that you have successfully engaged in research.
 But they are not required!
@@
 
-          \item \emph{Grades might be important}, depends on the reputation of \emph{your school} (\autoref{sec:your-school}).
-While good grades might not help much, bad ones likely will hurt your case.
+          \item \emph{High grades probably won't help much} (unless you're from a well-known school), but bad ones likely will raise concerns (\autoref{sec:your-school}).
 
@@
 
-          \item \emph{Contacting a prof. is recommended}, but do it \emph{properly} (\autoref{sec:contact}).
+          \item \emph{Contacting a prof. is recommended}, but do it \emph{properly} to get a reply (\autoref{sec:contact}).
 
@@
         \begin{itemize}
-          \item TAs, RAs, and fellowships are two main funding sources (\autoref{sec:funding}).
+          \item TAs, RAs, and fellowships are main funding sources for PhDs (\autoref{sec:funding}).
 TAs (teaching assistantships) are provided by the department to help with classes (e.g., grading).
@@
         \begin{itemize}
-          \item Many schools do not offer PhD studies in CS and many CS professors do not advise or graduate PhD students  (\autoref{sec:schoolsandprofs}).
+          \item Not every professor, even those in CS, can advise or graduate CS PhD students (\autoref{sec:schoolsandprofs}).
         \end{itemize}
@@
         \begin{itemize}
-          \item Why you were rejected (\autoref{sec:why-rejected})
+          \item Why you were rejected (\autoref{sec:why-rejected}). One potential reason is because you're applying to AI/ML, a super competitive field in recent years with many applicants.
+          \item Why rejection letters are sent so late?
+(\autoref{sec:late-rejection}).
+Grad programs often wait for the accepted students to make their decisions, typically by April 15, before sending out rejection letters.
+
           \item Increasing admission chance by being unique and standing out (\autoref{sec:improve-your-chance}).
-          \item Being in an underrepresented group can help (\autoref{sec:urm}).
-          \item You can successfully apply to CS PhD even if you have non-STEM background (\autoref{sec:non-stem}).
-          \item You do not need an MS degree (\autoref{sec:msrequirement}) to apply for PhD in CS, and it takes about 5--7 years for PhD in CS (\autoref{sec:time}).
-          \item Compared to other countries, CS PhD in the US does not require an MS degree but has longer PhD study time (\autoref{sec:non-us-differences}).
-          \item How do you call a professor? (\autoref{sec:address})
+          \item It take about 10 minutes to review an application (\autoref{sec:ievaluate}).
+          \item Being a minority can help (\autoref{sec:urm}).
+          \item You can successfully apply to CS PhD even if you have non-CS background (\autoref{sec:non-stem}).
+          \item You do not need an MS degree (\autoref{sec:msrequirement}), and CS PhD takes about 5--7 years (\autoref{sec:time}).
+Compared to other countries, CS PhD in the US does not require an MS degree but has a longer PhD study time (\autoref{sec:non-us-differences}).
+          \item How do you call your professors? In many cases, just by first names! (\autoref{sec:address})
           \item Your entire PhD program costs about \$400K in total, but you \emph{do not} pay for it (\autoref{sec:ra-cost})
@@
 
-\paragraph{Acceptance Letters} Universities prioritize sending out acceptance letters first. This allows the admitted students to make decision and plan for their studies.
+\paragraph{Acceptance Letters} Universities prioritize sending out acceptance letters first. This allows the admitted students to make decisions and plan for their studies.
 
@@
 
-\paragraph{Response Deadlines} Accepted students are usually given a deadline to make decision on their offers, often around \textbf{April 15th} in the US.
+\paragraph{Response Deadlines} Accepted students are usually given a deadline to make decisions on their offers, often around \textbf{April 15th} in the US.
 This is a standard deadline for PhD programs due to the Council of Graduate Schools (CGS) Resolution (\href{https://cgsnet.org/wp-content/uploads/2024/01/CGS_April15_Resolution_Jan312024.pdf}{see more}.
@@
 
-Given the high number of quality applicants and a limited number of spots, in addition to having a good application profile, you really want to show something that makes you \textbf{stand out}.
+Given the high number of quality applicants and a limited number of spots, in addition to having a good application profile, you want to show something that makes you \textbf{stand out}.
 For example, even if you do not have research experience, you can talk about your personal projects, as long as they can help show you can do research.
@@
 I also recommend very strong candidates for the University Presidential fellowship.
-The adcom chair then reviews all evaluations, looking at entered notes, and ask reviewers to discuss and resolve discrepancies to reach a consensus (e.g., a reviewer wants to accept but the other wants to reject).
+The adcom chair then reviews all evaluations, looks at entered notes, and asks reviewers to discuss and resolve discrepancies to reach a consensus (e.g., a reviewer wants to accept but the other wants to reject).
 Typically, the decision is made entirely by the reviewers; there is no involvement from the adcom chair, department chair, or others.

### 2024-07-10 faculty.tex
updates

@@
 Getting funding is arguably the most important part of getting tenure.
-without funding, you can't pay your students, conference registration and travel, or even your own summer salary.
+Without funding, you can't pay your students, conference registration and travel, or even your own summer salary.
 However, writing grant proposals is very different from writing research papers.
@@
 
-A strategy that I found effective is giving the entire class a small bonus point if everyone completes the evaluation.
-The bonus is computed as the percentage of responses * the bonus (X).
-So if X is 0.5 and everyone responds, everyone gets 0.5 points.
-If 50\% respond, each gets 0.25 points.
-While X is a tiny number, I emphasize that it can make a difference for borderline grades.
-I don’t curve grades, so this is the only way for students to bump.
-I also emphasize that in my previous classes many students benefit from this.
-So that really triggers their interest.
-And also because evaluations are before final exams, everyone is incentivized to participate because they don’t know their final grades and don’t want to miss this opportunity.
-Not only the students do it, they also remind their classmates/friends to do it.
-
-I only employ this method recently, and response rate has been very good so far.
-For Spr’24, it was 90\% (25/28 students) and for F’23, it was 100\% (16/16 students).
-Before this, my response rate is about 50\%.
+A strategy that I found effective is that the whole class gets a small bonus point if everyone completes the evaluation.
+The bonus is calculated as the percentage of responses multiplied by a factor (X).
+So, if X is 0.5 and everyone participates, they get 0.5 points each.
+If only half respond, it's 0.25 points each.
+It might seem small, but it can make a difference for borderline grades.
+Since I don't curve grades, this is their only chance to boost their scores.
+Many students have benefited from this in my previous classes, which really grabs their attention.
+Plus, since evaluations happen before finals and grades aren't finalized, everyone's motivated to participate—they don't want to miss out.
+Students not only do it themselves but also remind their friends.
+It's been working well for me: Spring '24 I had a 90\% response rate (25 out of 28 students), and Fall '23 was a perfect 100\% (16 out of 16 students).
+Before, my response rate was about 50\%, which is considered higher than average in the department.
 
@@
 
-\chapter{Services}
-
-While only 10\% of your work is officially for services, they are very important for your tenure and career.
+\chapter{Services}\label{sec:services}
+
+Officially services are only counted for 10\% of your work, however they are very important for your tenure and career.
 %You are expected to have certain types of services as a faculty member.  
-Some services are required (e.g., when your department chair assigns you to some committee), but some are not and you can decline (e.g., an invitation to be a PC member).
+Some services are required (e.g., when your department chair assigns you to some committees), but some are not and can be declined (e.g., an invitation to be a PC member).
 
@@
 
-Common committees include PhD and/or MS admission, faculty search, and colloquium committees.
+Common ones include PhD and/or MS admission, faculty search,  colloquium committees.
 Serving in the PhD adcom is very common for new assistant professors, as it is a good way to recruit students for your lab.
@@
 
-You can also take more leadership roles when you are close to tenure, such as being the chair of a committee or chair/director of a program.
-These are more time-consuming and might involve some politics, but they are also more rewarding and can help your tenure case.
+Other types of services include leadership roles (e.g., being the director of a program), and representing the dept in various university committees.
+These are not assigned by the dept, but rather you volunteer for them.
+Taking on a leadership role or representing the dept in certain university committees, which is less common for tenure-track profs., help the dept/chair in certain areas and can make you a good ``citizen'' which helps in your tenure case.
 
@@
 
+\paragraph{PhD Dissertation Committee} Being in a PhD committee is a common task to help your colleagues (e.g., the adviser of the student asks you to be in student dissertation committee) and the students (e.g., you are involved or are interested in the research of the student).
+You can serve in the committee of students in your dept, in other dept, or even in a different university (external committee member).
+This is a good way to build your network (and also help your students build their network).
+
 \paragraph{Writing Recommendation Letters} You will be asked to write recommendation letters for both students and sometimes even colleagues.
@@
 I do not really generally this and just volunteer for services that I think it is my responsibility (e.g., being in PCs, though I never liked them much) and those that I enjoy (NSF panels, services to the depts).
-In many cases I would also volunteer for tasks that no one wants to do because I feel that someone gotta do it and I can help out, e.g., various misc tasks in the dept.
+I enjoy services to my dept.
+and in many cases would take on tasks that no one wants to do because I feel that someone gotta do it and I can help out, e.g., various misc tasks in the dept.
 

### 2024-07-10 demystify.tex
updates

@@
           \item Why you were rejected (\autoref{sec:why-rejected}). One potential reason is because you're applying to AI/ML, a super competitive field in recent years with many applicants.
+
           \item Why rejection letters are sent so late?
@@
           \item You can successfully apply to CS PhD even if you have non-CS background (\autoref{sec:non-stem}).
-          \item You do not need an MS degree (\autoref{sec:msrequirement}), and CS PhD takes about 5--7 years (\autoref{sec:time}).
+          \item You do not need an MS (\autoref{sec:msrequirement}), and CS PhD takes about 5--7 years (\autoref{sec:time}).
 Compared to other countries, CS PhD in the US does not require an MS degree but has a longer PhD study time (\autoref{sec:non-us-differences}).
-          \item How do you call your professors? In many cases, just by first names! (\autoref{sec:address})
+          \item What should you call your profs.? (\autoref{sec:address}) Many possibilities (e.g., prof., Dr., and even their firstnames), but probably not Mr. or Mrs.
           \item Your entire PhD program costs about \$400K in total, but you \emph{do not} pay for it (\autoref{sec:ra-cost})

### 2024-07-10 demystify.tex
updates

@@
         \begin{itemize}
-          \item Why you were rejected (\autoref{sec:why-rejected}). One potential reason is because you're applying to AI/ML, a super competitive field in recent years with many applicants.
+          \item One reason why you were rejected (\autoref{sec:why-rejected}) is because you applied to AI/ML, a super competitive field in recent years with many applicants.
 

### 2024-07-11 faculty.tex
updates

@@
 \section{Types of Funding}
-\subsection{NSF and other Government Funding}
-
-\paragraph{Other Government Funding}
+\paragraph{NSF} NSF is the obvious funding source.
+Among many good things about NSF, two stand out: (transparency) proposals are evaluated by an independent panel and are based on merit rather than personal connections; (autonomy) the flexibility and freedom given to focus on pure research without the pressure of producing concrete deliverables.
+Because of these reasons, NSF funding is often considered the prestigious and competitive, even for very experienced senior researchers from top places (which don't mean anything).
+Note that a PI can use at most 2 months of summer salary from NSF grants (if you need the 3rd summer month, you need to find other sources of funding).
+
+\paragraph{DARPA, ONR, and other defense sponsors} DARPA, ONR, Army, Airforce, etc are like a contract, with deliverables required at regular intervals.
+Contracts are awarded in phases to multiple teams, and funding is cut for underperforming teams.
+This puts a lot of pressure on both the PIs and their students, requiring them to spend significant time on reports, meetings, and product development (e.g., creating hacker-proof software rather than research prototypes).
+Additionally, in many cases funding often depends on existing connections or being affiliated with institutions that frequently receive such awards.
+
+However, these sponsors are good sources of funding and can provide the additional resources (e.g., summer months for faculty).
+Note that some defense programs, e.g., Young Investigator Award, are more like NSF grants, with less stringent requirements and no deliverables required.
 
@@
 \subsection{Industrial Gifts}
+As its name suggests, industrial gifts are unrestricted research funding from companies such as Amazon, Microsoft, Facebook, and Google.
+These are highly competitive and prestigious (e.g., the solicitation is open worldwide, and you compete with top researchers from top places).
+They are the best kind of funding for three reasons: (1) indirect costs or overheads are typically not levied on these gifts (so you get most of the money), (2) there is no associated timeline by when you should spend them (they give it to you and forget about it), and (3) you can use them for a broader range of things than university and grant/contract funding (e.g., fun events for your research group).
 
@@
 \chapter{Mentoring}
+
+Successfully mentoring students is a key and most rewarding part of your job as a faculty member. 
+There will be times when you feel discouraged and frustrated, but there will also be times when you feel proud and happy.  
+You will be expected to mentor undergraduate students and graduate students. You will likely need to graduate at least one PhD student to get tenure.
+
+\section{Mentoring Graduate Students}
+
 \section{Mentoring Undergraduate Students}
-\section{Mentoring Graduate Students}
+Surprisingly, some of the best research projects I have seen come from undergraduate students.
+
+I have worked with many undergrads (15+ since 2016). My success rate (publish a research paper with me) with undergrads
+
 \section{Mentoring Postdocs}
@@
 
-Officially services are only counted for 10\% of your work, however they are very important for your tenure and career.
+Officially services are counted for only 10\% of your work. However, they are very important for your tenure and career.
 %You are expected to have certain types of services as a faculty member.  
@@
 %The department will then review your application and vote on whether to renew you for your 4th (and perhaps also 5th year). 
-Your mentor presents and argues for your case in the dept.
+Your mentor will also present and argue for your case in the dept.
 meeting, but the decision is made within the dept.
@@
 While the renewal is straightforward for most people, it can be surprise and shock to some.
+Ideally, it should not be a surprise, as your annual evaluation should give you some idea on how you are doing, e.g., if you are not doing well, the annual evaluation should tell you that you need to improve.
+However, it can be a surprise due to various reasons such as new dept.
+chair or administrative changes who have different expectations.
 

### 2024-07-13 demystify.tex
updates

@@
 
-\section{How long does to complete the CS PhD program?}\label{sec:time}
+\section{How long to complete the CS PhD program?}\label{sec:time}
 
@@
 
-Furthermore, avoid using Mr.
-or Mrs., or the professor's first name if you're not acquainted with them yet.
+Furthermore, avoid using Mr., Mrs., Miss, or the professor's first name if you're not acquainted with them yet.
 As you become more familiar with your prof and depending on their preferences, you may transition to addressing them by their first name.
@@
 Lastname is preferred over Prof.
-You just need to observe and follow the conventions at your particular institution.
-Additionally, be aware that not all faculty members may hold a PhD, in which case using Prof.
+Moreover, be aware that not all faculty members may hold a PhD, in which case using Prof.
 Lastname is a suitable alternative.
+	You just need to observe and follow the conventions at your particular institution.One way to determine how to address a professor is to observe how they sign their emails or how they introduce themselves in class.
+For example, I introduce myself as ``Vu'' in class.
 
@@
 
-Students understandbly get excited about their upcoming PhD journey and want to buy new computer equipment and electronics to prepare.
+Students understandably get excited about their upcoming PhD journey and want to buy new computer equipment and electronics to prepare.
 However, you should check with your prof.
@@
 
-\section{Can I negotiate my PhD offer (e.g., if I have multiple offers)?}\label{sec:negotiate}
+\section{Negotiating PhD offer (e.g., having multiple offers)?}\label{sec:negotiate}
 
@@
 
-\begin{table}
+\begin{table}[h]
   \centering
   \small
-  \caption{Top 50 CS PhD programs in the U.S.
+  \caption{The top 50 CS programs in the US from \href{https://www.csrankings.org}{CSRankings.org}, a ranking system  based on faculty publications at top CS conferences.
 (\href{https://csrankings.org}{CSRankings}, Jan.
@@
 % Virginia Tech\red{$^*$}  \\
-\autoref{tab:ranking} lists the top 50 CS programs in the US from \href{https://www.csrankings.org}{CSRankings.org}, a ranking system  based on faculty publications at top CS conferences.
+
+%\autoref{tab:ranking} lists .
 
@@
 We talked about why Mason was not able to attract good Vietnamese and other international students, despite having a much stronger CS program than many schools that these students want to go to (part of the reason is described in \autoref{sec:selecting-ranking-schools}).
-We wish there were a way for international students to know about the US PhD programs (also for US faculty to understand more about international students and therefore have a better chance of recruiting and working with them).
+We wish there were a way for international students to know about PhD programs in the US.
 
@@
 There I saw many questions from students about PhD programs.
-However, most active participants are students in non-CS fields or not in the US, and their answers are unfortunately not always accurate and sometimes lead to more confusion.
+However, most participants are students in non-CS fields or not in the US, and their answers are not always accurate and can lead to more confusion.
 So I thought it would be useful to have a document that is specific to CS PhD programs in the US from an insider perspective.
@@
 
-\paragraph{About Me} I am an associate professor in the CS dept at George Mason University.
+\paragraph{About Me} I am an associate professor in the CS dept at George Mason University (Mason).
 Before Mason, I was at the University of Nebraska-Lincoln (UNL).
 I have been in the PhD admission process at Mason and UNL for many years.
-Currently I serve as the program director of the MS program in Software Engineering at Mason (thus also have experience with the MS admission process---which is quite different than PhD).
+I also serve as the program director of the MS program in Software Engineering at Mason (thus also have experience with the MS admission process---which is quite different than PhD).
 My personal and lab website is at \href{https://dynaroars.github.io}{dynaroars.github.io}.

### 2024-07-13 faculty.tex
updates

@@
 
+\section{Types of Funding}
+\paragraph{NSF} NSF is the obvious funding source for CS. 
+Two things stand out for NSF grants: (transparency) proposals are evaluated by an independent panel and are based on merit rather than personal connections; (autonomy) the flexibility and freedom given to focus on pure research without the pressure of producing concrete deliverables.
+Because of these reasons, NSF funding is often considered the prestigious and competitive, even for very experienced senior researchers from top places (which don't mean anything).
+Note that a PI can use at most 2 months of summer salary from NSF grants (if you need the 3rd summer month, you need to find other sources of funding).
+
+\paragraph{DARPA, ONR, and other defense sponsors} Most of these awards are like a \emph{contract}, with deliverables required at regular intervals.
+Contracts are awarded in phases to multiple teams, and funding is cut for underperforming teams.
+This puts a lot of pressure on both the PIs and their students, requiring them to spend significant time on reports, meetings, and product development (e.g., creating hacker-proof software rather than research prototypes).
+Additionally, in many cases funding often depends on existing connections or being affiliated with institutions that frequently receive such awards.
+
+However, these sponsors are good sources of funding and can provide the additional resources (e.g., summer months for faculty).
+Note that some defense programs, e.g., Young Investigator Award, are more like NSF grants, with less stringent requirements and no deliverables required.
+
+\subsection{Internal Funding}
+These are funds from your university, city, or state (e.g., Virginia has the CCI program that provide ``seed'' grants for researchers from universities in Virginia).
+Also, your ``startup'' package is also an internal funding to kickstart your research.
+These funding are less competitive than NSF, but they are also less prestigious (e.g., they aren't much and usually don't count as much for tenure).
+However, they are good sources of funding to get preliminary results and later apply for traditional grants (and many internal funding programs explicitly encourage this).
+
+\subsection{Industrial Gifts}
+Industrial gifts are \emph{unrestricted funding} from companies such as Amazon, Microsoft, Facebook, and Google.
+These are highly competitive and prestigious (e.g., the solicitation is open worldwide, and you compete with top researchers from top places).
+They are the best kind of funding for three reasons: (1) indirect costs or overheads are typically not levied on these gifts (so you get most of the money), (2) there is no associated timeline by when you should spend them (they give it to you and forget about it), and (3) you can use them for a broader range of things than university and grant/contract funding (e.g., summary salary, equipments, or even bowling events for your group).
+
+\paragraph{Personal Experiences}
+
 \section{Most people suck at grant writing (at first)} Most new assistant professors have no experience in grant writing, and struggle with it in the beginning.
@@
 Keep trying, and you will get better over time.
-
-\section{Types of Funding}
-\paragraph{NSF} NSF is the obvious funding source.
-Among many good things about NSF, two stand out: (transparency) proposals are evaluated by an independent panel and are based on merit rather than personal connections; (autonomy) the flexibility and freedom given to focus on pure research without the pressure of producing concrete deliverables.
-Because of these reasons, NSF funding is often considered the prestigious and competitive, even for very experienced senior researchers from top places (which don't mean anything).
-Note that a PI can use at most 2 months of summer salary from NSF grants (if you need the 3rd summer month, you need to find other sources of funding).
-
-\paragraph{DARPA, ONR, and other defense sponsors} DARPA, ONR, Army, Airforce, etc are like a contract, with deliverables required at regular intervals.
-Contracts are awarded in phases to multiple teams, and funding is cut for underperforming teams.
-This puts a lot of pressure on both the PIs and their students, requiring them to spend significant time on reports, meetings, and product development (e.g., creating hacker-proof software rather than research prototypes).
-Additionally, in many cases funding often depends on existing connections or being affiliated with institutions that frequently receive such awards.
-
-However, these sponsors are good sources of funding and can provide the additional resources (e.g., summer months for faculty).
-Note that some defense programs, e.g., Young Investigator Award, are more like NSF grants, with less stringent requirements and no deliverables required.
-
-\subsection{Internal Funding}
-
-\subsection{Industrial Gifts}
-As its name suggests, industrial gifts are unrestricted research funding from companies such as Amazon, Microsoft, Facebook, and Google.
-These are highly competitive and prestigious (e.g., the solicitation is open worldwide, and you compete with top researchers from top places).
-They are the best kind of funding for three reasons: (1) indirect costs or overheads are typically not levied on these gifts (so you get most of the money), (2) there is no associated timeline by when you should spend them (they give it to you and forget about it), and (3) you can use them for a broader range of things than university and grant/contract funding (e.g., fun events for your research group).
-
-\section{My Story}
 
@@
 There will be times when you feel discouraged and frustrated, but there will also be times when you feel proud and happy.  
-You will be expected to mentor undergraduate students and graduate students. You will likely need to graduate at least one PhD student to get tenure.
+For tenure you will be expected to mentor undergraduate students and graduate students. You will likely need to graduate at least one PhD student to get tenure.
 
@@
 \paragraph{PhD Dissertation Committee} Being in a PhD committee is a common task to help your colleagues (e.g., the adviser of the student asks you to be in student dissertation committee) and the students (e.g., you are involved or are interested in the research of the student).
-You can serve in the committee of students in your dept, in other dept, or even in a different university (external committee member).
-This is a good way to build your network (and also help your students build their network).
+You can serve in the PhD or MS committee of students in your dept, in other dept, or even in a different university (external committee member).
 
@@
 
-I did not do these visits, partly because I am not a travel person, and also I felt that I would not have problems finding letter writers.
-Instead, during the Spring'23 semester before my tenure application, I put lots of effort in helping the university in recruiting faculty (by meeting with 20+ faculty candidates), even though I am not in the search committee and really didn't have service requirements during this pre-tenure study break.
+I did not do these visits, partly because I am not a travel person, and also I felt that I would have enough good letter writers.
+However, during the Spring'23 semester before my tenure application, I did participate more in the dept.
+activities, e.g., in recruiting by meeting with 20+ faculty candidates, even though I am not in the search committee and really didn't have service requirements during this pre-tenure study break.
 This helps me build a good relationship with my colleagues in the dept.
 
-\section{3rd year Renewal Package}
+\section{Midterm Review}
 
@@
 
-Toward the end of your 3rd year (or early 4th year), you will need to submit a ``midterm'' \emph{renewal} application to renew your contract for the 4th year (and perhaps also 5th year).
-While somewhat similar to tenure evaluation, this is less intense and  does not require reference letters or presentation.
+Midterm review happens midway through the tenure period, i.e., the 3rd year in a 6-year tenure track.
+You will need to submit a midterm \emph{renewal} application to renew your contract (and your next renewal will be the tenure application).
+While somewhat similar to tenure evaluation described above, midterm review is less intense and  does not require reference letters or presentation.
 %The department will then review your application and vote on whether to renew you for your 4th (and perhaps also 5th year). 
@@
 
-While the renewal is straightforward for most people, it can be surprise and shock to some.
-Ideally, it should not be a surprise, as your annual evaluation should give you some idea on how you are doing, e.g., if you are not doing well, the annual evaluation should tell you that you need to improve.
-However, it can be a surprise due to various reasons such as new dept.
+While the renewal is straightforward for most people, it can be a surprise and shock to some.
+Ideally, it should not be a surprise, as your annual evaluation or dept.
+chair and mentor should tell you if you're not doing well and what to improve.
+However, things can become complicated due to various reasons such as new dept.
 chair or administrative changes who have different expectations.

### 2024-07-13 demystify.tex
updates

@@
 
+% \section{Timeline}
+
+% \begin{table}[h!]
+% \centering
+% \begin{tabular}{l|l}
+%     \toprule
+% \textbf{Month} & \textbf{Task} \\
+% \midrule
+% January-March & Research programs and prepare for tests \\
+% April-June & Take standardized tests (GRE, TOEFL, etc.) \\
+% July-September & Prepare application materials (SOP, LORs, etc.) \\
+% October-December & Submit applications \\
+% January-March & Interviews and follow-ups \\
+% April & Admission decisions \\
+% \bottomrule
+% \end{tabular}
+% \caption{Application Timeline}
+% \end{table}
+
 \section*{Who is this document for?}
@@
 
+\begin{commentbox}{Vu:}
 At Mason, we usually decide that a full-time PhD candidate is either (i) admitted with funding (\autoref{sec:funding}) or (ii) rejected. In other words, in most cases, we either
@@
 supported by your government or having external fellowships. We justify
-our decision with a summary of your application, where we list
+our decision (\autoref{sec:ievaluate}) with a summary of your application, where we list
 strengths, e.g., a well-known school, and weaknesses, e.g., weak
 LORs.
+\end{commentbox}
 
@@
 
-\paragraph{How to help your LOR writer} As mentioned above, do not write your own letter.
+\paragraph{Helping your LOR writer} As mentioned above, do not write your own letter.
 Instead, you can help your LOR writer by providing details or unique things about yourself.
 For example, let them know about your GPA, research and work experiences, papers (if any), or anything you want them to mention.
-You can also provide them with a draft of your SOP so that they can see what you are saying about yourself and can complement that with their own perspective.
+You can also provide them with a draft of your SOP so that they can see what you are saying about yourself and complement that with their own perspective.
 Sometimes your writer will explicit ask you for such information, but if not, you should provide it anyway (especially if you have not interacted with them much or have not done much research with them).
@@
 Graduate admission can involve randomness and noise.
-In the meantime, you can work on improving your profile, e.g., get more research experiences, publish more papers, improve your connections for better LoR writers, etc.
-See rejection reasons (\autoref{sec:why-rejected}) and additional advice (\autoref{sec:improve-your-chance}).
+In the meantime, you can work on improving your profile, e.g., get more research experiences, publish more papers, improve your connections for stronger LoRs, etc.
+See rejection reasons (\autoref{sec:why-rejected}), timeline (\autoref{sec:late-rejection}), and additional advice (\autoref{sec:improve-your-chance}).
 
@@
 
-While many find CSRankings useful, it is not perfect, and in general, rankings are superficial, and you need to do more research to be informed and make better decisions.
+\paragraph{What matters to you?} While many find CSRankings useful, it is not perfect, and in general, rankings are superficial, and you need to do more research to be informed and make better decisions.
 For example, you might prefer places with a large community from your country (e.g., Northern VA has a large population of Vietnamese).
@@
 
-This summarizes the main differences between CS PhD in the US and other countries. % This is based on my experience and what I heard from others who did PhD in other countries.
+\begin{table}[h!]
+\centering
+\begin{tabular}{|c|c|c|}
+\toprule
+\textbf{Aspect} & \textbf{US PhD Programs} & \textbf{Other Countries} \\
+\midrule
+Duration & 5-7 years & 3-5 years \\
+MS Required & Not required & Often required \\
+Coursework Required & Yes (first 2 years) & No \\
+Research Proposal Required & No & Yes (in some countries) \\
+Academic (Faculty) Job & Direct & Postdoc \\
+Work Life Balance & Less & More \\
+\bottomrule
+\end{tabular}
+\caption{Comparison of US and Other PhD Programs}\label{tab:us-vs-other}
+\end{table}
+
+\autoref{tab:us-vs-other} summarizes the main differences between CS PhD in the US and other countries. % This is based on my experience and what I heard from others who did PhD in other countries.
 

### 2024-07-14 demystify.tex
updates

@@
 
-After you submit your PhD application (e.g., around December of year $X$ for the Fall admission of year $X+1$), it will first be screened
+After you submit your PhD application (e.g., in late Fall of year $X$ for the Fall admission of year $X+1$), it will first be screened
 for general requirements, e.g., did you submit your transcripts and standard scores?
@@
 Usually, this screening process is done through a central university system, i.e., not by CS faculty.
-If you pass this screen, your application will be forwarded to the CS department for further evaluation.
+If you pass this, your application will be forwarded to the CS department for further evaluation.
 If you don't,  the system will tell you what is missing and what you need to do.
@@
 \begin{commentbox}[Vu:]
-  I often read LORs and SOP first. If I am
+  I often read SOP and LORs first. If I am
   persuaded by then, I would skim over other factors and advocate for
@@
 
-Some schools send out admission letters in batches, some do rolling admission, and some do not send anything out (e.g., you're on their waitlist, or they might not send rejection letters).
+Some schools send out admission letters in batches, some do rolling admission, and some do not send anything out (e.g., you're on their waitlist).
 In general, you should hear back from most schools by mid-March (but some schools might take longer!).
+Note that rejection letters might come very late (see \autoref{sec:late-rejection}).
 
@@
 In the meantime, you can work on improving your profile, e.g., get more research experiences, publish more papers, improve your connections for stronger LoRs, etc.
-See rejection reasons (\autoref{sec:why-rejected}), timeline (\autoref{sec:late-rejection}), and additional advice (\autoref{sec:improve-your-chance}).
+See rejection reasons (\autoref{sec:why-rejected}) and timeline (\autoref{sec:late-rejection}), and additional advice to improve your chance next time (\autoref{sec:improve-your-chance}).
 
@@
 You will likely not get any useful feedback.
-We are not willing to reveal your evaluation results or give you advice on how to improve.
+We are not willing to reveal your evaluation results or give you advice on how to improve your profile.
 So it is best to just move on.

### 2024-07-16 faculty.tex
services

@@
 These are funds from your university, city, or state (e.g., Virginia has the CCI program that provide ``seed'' grants for researchers from universities in Virginia).
-Also, your ``startup'' package is also an internal funding to kickstart your research.
+Your ``startup'' package is also be considered as a large funding from an internal source (your university) to kickstart your research.
 These funding are less competitive than NSF, but they are also less prestigious (e.g., they aren't much and usually don't count as much for tenure).
@@
 \paragraph{Personal Experiences}
+As with most tenure track professors, I focus mostly on NSF grants and most of my funding comes from NSF.
+I only have one grant from Army Research Office (ARO) in which I am a co-PI and didn't have to write the proposal.
+I have tried to get other DoD funding, but without much success, I found that DoD program managers are very picky and often have their own agenda.
+I have also received internal funding from my university and industrial gifts from Amazon,and Facebook.
+I have found that industrial gifts are the best kind of funding, as they are unrestricted and have no associated timeline by when you should spend them.
+They are also highly competitive and prestigious, and are a great way to get funding for your research.
 
@@
 \chapter{NSF Panel Reviews}
+\chapter{Issues}
+
+\section{Change in Departmental Leadership. Moving from UNL to GMU}
+
+I joined UNL in 2016 and was very happy there until 2020, which is when the dept. changed its leadership and organization and lead to my decision to be on the job market again.
+
+I do not want to get into details but below summarizes several things that happened
+
+\begin{itemize}
+\item 2019: four senior professors (in SE) left.  One of these 4 had became the chair of the dept and then left just after 2 months.  
+\item 2020:  an even split among the faculty on moving to completely to the College of Engr.
+Multiple senior people warn about this decision.
+The new chair candidate only agrees to accept the job if the faculty moved to CoE, putting pressure on the move.
+\item Salary inversion:  new chair offers new AP salary that is much higher than current AP, leading to jokes that current AP should just reapply to UNL.
+This became a serious issue and many people at UNL use this as a reason on why they move.
+
+\end{itemize}
+
+Note that even after leaving Lincoln, my family and I still have a soft spot for the city and its people.
+It was a great place to live and raise a family, but unfortunately not a good place to work for me.
 

### 2024-07-20 demystify.tex
services

@@
 I have been in the PhD admission process at Mason and UNL for many years.
-I also serve as the program director of the MS program in Software Engineering at Mason (thus also have experience with the MS admission process---which is quite different than PhD).
+This allows me to have a good understanding of the process, the challenges students face, and what faculty are looking for.
+I also serve as the program director of the MS program in Software Engineering at Mason (thus also have experience with MS admission---which is quite different than PhD).
 My personal and lab website is at \href{https://dynaroars.github.io}{dynaroars.github.io}.

### 2024-07-20 faculty.tex
services

@@
 \chapter{Research}
+Research here focuses on paper publication, which is a major component in getting your name out there and of course in your departmental evaluation.
+
 \chapter{Obtaining Funding}
@@
 Many would (rightly) advise about being strategic about your services.
-I do not really generally this and just volunteer for services that I think it is my responsibility (e.g., being in PCs, though I never liked them much) and those that I enjoy (NSF panels, services to the depts).
+I didn't really follow this and just volunteer for services that I think it is my responsibility (e.g., being in PCs, though I never liked them much) and those that I enjoy (NSF panels, services to the depts).
 I enjoy services to my dept.
@@
 
-\section{Change in Departmental Leadership. Moving from UNL to GMU}
-
-I joined UNL in 2016 and was very happy there until 2020, which is when the dept. changed its leadership and organization and lead to my decision to be on the job market again.
-
-I do not want to get into details but below summarizes several things that happened
+\section{Change in Department. Moving from UNL to GMU}
+
+I joined UNL in 2016 and was very happy there until 2020, when people start leaving and the dept.
+changed its leadership and organization.
+All of these uncertainties and other issues lead to my decision to be on the job market again.
+
+I will skip the details and just summarize several things that I didn't like.
 
 \begin{itemize}
-\item 2019: four senior professors (in SE) left.  One of these 4 had became the chair of the dept and then left just after 2 months.  
-\item 2020:  an even split among the faculty on moving to completely to the College of Engr.
-Multiple senior people warn about this decision.
+\item 2019: four senior professors (in SE) left.
+One of these 4 was the dept chair (and he left just after 2 months of being chair).
+There were issues of upper admin didn't work well with the dept chair (and this issue has been going for years).
+\item 2020: College of Engr pressure dept to move entirely to it.
 The new chair candidate only agrees to accept the job if the faculty moved to CoE, putting pressure on the move.
+Faculty are split on the decision and eventually moved with lots of mistrusts and concerns under CoE.
 \item Salary inversion:  new chair offers new AP salary that is much higher than current AP, leading to jokes that current AP should just reapply to UNL.
 This became a serious issue and many people at UNL use this as a reason on why they move.
+\item Senior faculty fighting on mailing list.
 

### 2024-07-20 demystify.tex
services

@@
           \item Increasing admission chance by being unique and standing out (\autoref{sec:improve-your-chance}).
-          \item It take about 10 minutes to review an application (\autoref{sec:ievaluate}).
+          \item It takes about 10 minutes to review an application (\autoref{sec:ievaluate}).
+          \item Sometimes adcom members do dicuss applicants, but in most cases they make independent decisions (\autoref{sec:adcom-discuss}).           
           \item Being a minority can help (\autoref{sec:urm}).
@@
 While this seems short, it is not that difficult to tell if an application is good or bad.
-In fact, this time
+In fact, this
 is twice what other faculty spend on average, e.g., Philip Guo spent \href{https://pg.ucsd.edu/PhD-application-tips.htm}{3--5 minutes} per application.
@@
 Typically, the decision is made entirely by the reviewers; there is no involvement from the adcom chair, department chair, or others.
+
+\section{Do adcom members talk to each other?}\label{sec:adcom-discuss}
+
+In most cases we make decision independently and do not talk to each other.
+However, we do discuss when there are discrepancies in evaluations, in which case the adcom chair will ask reviewers discuss the application to reach a consensus.
+We also chat to each other for interesting or strong applications that we think should be admitted, e.g., how to recruit this student or who should be the adviser.
+If the student mentioned a faculty member in their SOP, we might ask that faculty member if they are interested in the student.
+In short, while we do not talk to each other for most applications, we do for some.
 

### 2024-07-20 demystify.tex
services

@@
           \item It takes about 10 minutes to review an application (\autoref{sec:ievaluate}).
-          \item Sometimes adcom members do dicuss applicants, but in most cases they make independent decisions (\autoref{sec:adcom-discuss}).           
-          \item Being a minority can help (\autoref{sec:urm}).
+          \item Sometimes adcom members dicuss applicants, but in most cases they make independent decisions (\autoref{sec:adcom-discuss}).           
+          \item Adcom members only review applications assigned to them (typically matching their expertise) and rarely get involved in other applications (\autoref{sec:read-all}).
+
+          \item Being a minority can significantly help your application (\autoref{sec:urm}).
           \item You can successfully apply to CS PhD even if you have non-CS background (\autoref{sec:non-stem}).
@@
 
-In most cases we make decision independently and do not talk to each other.
-However, we do discuss when there are discrepancies in evaluations, in which case the adcom chair will ask reviewers discuss the application to reach a consensus.
-We also chat to each other for interesting or strong applications that we think should be admitted, e.g., how to recruit this student or who should be the adviser.
+In most cases adcom members, even those reviewing the same application, make decision independently and do not talk to each other.
+However, in some rare cases we do talk to each other.
+For example, when there are discrepancies in evaluations, in which case the adcom chair will ask reviewers discuss the application to reach a consensus.
+We also talk to each other for interesting or strong applications, e.g., how to recruit this student or who should be the adviser.
 If the student mentioned a faculty member in their SOP, we might ask that faculty member if they are interested in the student.
-In short, while we do not talk to each other for most applications, we do for some.
+
+\section{Do adcom members read all applications?}\label{sec:read-all}
+
+We only review those that are assigned to us. Adcom chair will assign applications to reviewers based on their expertise (e.g., if a student says they want to do SE or interested in working with me).
+
+For example, I only review those that are assigned to me, which is typically around 50--70 applications for the Fall admission. 
+I do not look at any other applicants, unless I know them (e.g., they contacted me) or have some special circumstances (e.g., they are a student from Vietnam).
+But usually, even if I look at them, I do not get involved in their evaluation.
 

### 2024-07-21 demystify.tex
Updates from Overleaf
updates

@@
 \end{commentbox}
+
+\subsection{Common Pitfalls to Avoid}\label{sec:sop-pitfalls}
+
+SOP is crucial for making a strong impression on adcom reviewers. This outlines common pitfalls to avoid in your SOP.
+
+\begin{enumerate}
+    \item \textbf{Generic Statements}: Using vague or general statements that could apply to any applicant.
+For example, saying blank statements such as “I am passionate about computer science” without providing specific examples or evidence.
+    \item \textbf{Not Customized for the University you're applying to}: Failing to mention why this univ is a good fit for you and which faculty members you want to work with.
+    \item \textbf{Too Long}: Keep it nore more than 2 pages.  Adcom members have to read many applications and a long SOP can be a turn-off.
+
+    \item \textbf{Lack of Passion or Motivation}: Failing to convey genuine enthusiasm for research and the specific program.
+    \item \textbf{Grammatical and Typographical Errors}: Submitting an SOP filled with grammatical mistakes, typos, and poorly constructed sentences.
+    \item \textbf{Excessive Self-Praise}: Overly boasting about achievements without providing concrete evidence or examples to back them up.
+    \item \textbf{Irrelevant Information}: Including personal anecdotes or experiences that do not relate to the field of study or research interests.
+    \item \textbf{Unrealistic Goals}: Stating research goals that are not feasible or that show a lack of understanding of the field’s current state.
+    \item \textbf{Negative Tone}: Criticizing previous institutions, professors, or colleagues.
+    \item \textbf{Failure to Address Fit}: Not explaining why you are a good fit for the program and how the program is a good fit for your goals.
+    \item \textbf{Generic Opening and Closing Statements}: Starting or ending the SOP with clichés or unoriginal statements that do not add value.
+    \item \textbf{Lack of Customization}: Submitting the same SOP to multiple programs without tailoring it to each specific program's strengths and faculty.
+\end{enumerate}
 

### 2024-07-23 demystify.tex
updates
services

--- 2024-07-20 demystify.tex
@@
           \item It takes about 10 minutes to review an application (\autoref{sec:ievaluate}).
-          \item Sometimes adcom members dicuss applicants, but in most cases they make independent decisions (\autoref{sec:adcom-discuss}).           
+          \item Sometimes adcom members discuss applicants, but in most cases they make independent decisions (\autoref{sec:adcom-discuss}).           
           \item Adcom members only review applications assigned to them (typically matching their expertise) and rarely get involved in other applications (\autoref{sec:read-all}).
@@
 For example, if you say you have experience with teaching, then show what you did, e.g., undergrad TA or mentoring someone.
-If you say you work on project X, then show some results (paper submitted (or even rejected), achieved X\% performance improvement over ..).
+If you say you work on a project, then show some results, e.g., paper submitted (or even rejected), achieved X\% performance improvement over ...
 Also, do not use more than two pages (references can be on the third page).
@@
   convinced, then I will likely recommend rejection (unless I see
-  something really stands out in other parts).
+  something stand out in other parts).
   \tcblower
@@
 For example, 3.5 GPA students from such talented programs are typically much better than 4.0 GPA students not in those programs.
-Similarly, variations in GPA standards exist among different universities, with technical universities generally having lower GPAs than economical universities.
+Similarly, variations in GPA standards exist among different universities, with technical universities generally having lower GPAs than economic universities.
 These make gaining admission in the US difficult as US faculty are not familiar with these issues.
@@
 \paragraph{How to get TA?}  Unless you have other funding such as RA or fellowships, TA is typically the default for CS PhD programs.
-In your application, indicate that you need financial assistant.
+In your application, indicate that you need financial assistance.
 Typically, adcom will either admit you and give you TA, or reject you.
@@
 Typically, a tenure-track faculty starts with being an assistant professor and then gets tenure and promoted to associate (typically after 6 years) and then full (time varies, some become full within 3--4 years, some 10+ years, some remain associate).
-A non-tenure track faculty, e.g., teaching or research faculty, also starts with assistant, then is promoted to associated, and full prof.
+A non-tenure track faculty, e.g., teaching or research faculty, also starts with Assistant, then is promoted to Associated, and Full prof.
 

### 2024-07-24 demystify.tex
updates

@@
 
+% \subsection{Dos and Don'ts of SOP}
+% \paragraph{Do's}
+% \begin{enumerate}
+%     \item \textbf{Be Specific}: Clearly articulate your research interests and why you are drawn to the program.
+%     \item \textbf{Show Enthusiasm}: Demonstrate genuine passion for your field of study.
+%     \item \textbf{Proofread}: Ensure your statement is free of grammatical and typographical errors.
+%     \item \textbf{Be Personal}: Include relevant personal stories that highlight your journey and motivations.
+%     \item \textbf{Tailor the Statement}: Customize your statement for each program you apply to.
+% \end{enumerate}
+
+% \paragraph{Don'ts}
+% \begin{enumerate}
+%     \item \textbf{Avoid Generic Statements}: Don’t use vague or general phrases.
+%     \item \textbf{Don't Repeat Your Resume}: Provide new insights or context beyond what’s listed in your CV.
+%     \item \textbf{Steer Clear of Jargon}: Avoid overusing technical jargon that might confuse readers.
+%     \item \textbf{Don’t Exaggerate}: Avoid excessive self-praise without evidence.
+%     \item \textbf{Avoid Negative Tones}: Don’t criticize past institutions or colleagues.
+% \end{enumerate}
+%\subsection{Common Pitfalls to Avoid}\label{sec:sop-pitfalls}
+
+%SOP is crucial for making a strong impression on adcom reviewers. This outlines common pitfalls to avoid in your SOP.
+
+%\begin{enumerate}
+%m \textbf{Generic Statements}: Using vague or general statements that could apply to any applicant.
+For example, saying blank statements such as “I am passionate about computer science” without providing specific examples or evidence.
+%    \item \textbf{Not Customized}: Failing to mention why you chose this univ and faculty members you want to work with.
+If you can submit the same SOP to multiple programs with few changes, it is not customized enough.
+%    \item \textbf{Too Long}: Adcom members have to read many applications and a long SOP can be a turn-off. Keep it under 2 pages.
+%    \item \textbf{Lack of Passion or Motivation}: Failing to convey genuine enthusiasm for research and the specific program.
+%    \item \textbf{Grammatical and Typographical Errors}: Submitting an SOP filled with grammatical mistakes, typos, and poorly constructed sentences.
+%    \item \textbf{Excessive Self-Praise}: Overly boasting about achievements without providing concrete evidence or examples to back them up.
+%    \item \textbf{Irrelevant Information}: Including personal anecdotes or experiences that do not relate to the field of study or research interests.
+%    \item \textbf{Unrealistic Goals}: Stating research goals that are not feasible or that show a lack of understanding of the field’s current state.
+%    \item \textbf{Negative Tone}: Criticizing previous institutions, professors, or colleagues.
+%    \item \textbf{Failure to Address Fit}: Not explaining why you are a good fit for the program and how the program is a good fit for your goals.
+%    \item \textbf{Generic Opening and Closing Statements}: Starting or ending the SOP with clichés or unoriginal statements that do not add value.
+%    \item \textbf{Lack of Customization}: Submitting the same SOP to multiple programs without tailoring it to each specific program's strengths and faculty.
+%\end{enumerate}
+
 \subsection*{Additional Resources}
@@
   3) why do you want to apply for university X, not Y?
-  If you don't want to spend time to do your homework,
+  If you don't want to spend time doing your homework,
   the chance of getting a reply is close to zero.
@@
   In my view, cold emails are not welcome by most faculty members and should be avoided.
-However, if one is already admitted to a program in some department, by all means, send an email to the faculty you may be interested in working with, but do mention right at the beginning that you are already admitted to the program as well as several other universities.
+However, if one is already admitted to a program in some department, by all means, email the faculty you may be interested in working with, but do mention right at the beginning that you are already admitted to the program as well as several other universities.
 State specific areas (preferably specific topics-ML, robotics instead of AI).

### 2024-07-24 demystify.tex
updates

@@
 My goal is thus to level the playing field by providing information that is not readily available to less privileged students. 
-I also hope to encourage more students with such backgrounds to apply to CS PhD programs in the US.
+I hope to encourage more students with such backgrounds to apply to CS PhD programs in the US.
 
@@
 \paragraph{Not a good fit and Bad luck}  You could have an excellent profile, but if you are interested in a research area that the program does not have, you might not be admitted.
-Similarly, if no faculty who is willing to advise you (e.g., they are on sabbatical or personal leave, do not have sufficient funding, or already have too many students already), you will not be admitted.
+Similarly, if no faculty who is willing to advise you (e.g., they are on sabbatical or personal leave, do not have sufficient funding, or already have too many students), you will not be admitted.
 This is good for you, as you don't want to be in a program where no one can advise you.
@@
 
-In most cases adcom members, even those reviewing the same application, make decision independently and do not talk to each other.
+In most cases adcom members, even those reviewing the same application, make decisions independently and do not talk to each other.
+It's not a strict rule but a common practice to avoid bias and ensure fairness (and also we don't have time to talk to each other).
+
 However, in some rare cases we do talk to each other.
-For example, when there are discrepancies in evaluations, in which case the adcom chair will ask reviewers discuss the application to reach a consensus.
+For example, when there are discrepancies in evaluations, in which case the adcom chair will ask reviewers to discuss the application to reach a consensus.
 We also talk to each other for interesting or strong applications, e.g., how to recruit this student or who should be the adviser.
@@
 
-We only review those that are assigned to us. Adcom chair will assign applications to reviewers based on their expertise (e.g., if a student says they want to do SE or interested in working with me).
+We only review those that are assigned to us, which are already too many.
+Adcom chair will assign applications to reviewers based on their expertise (e.g., if a student says they want to do SE or interested in working with me).
 
@@
 I do not look at any other applicants, unless I know them (e.g., they contacted me) or have some special circumstances (e.g., they are a student from Vietnam).
-But usually, even if I look at them, I do not get involved in their evaluation.
+However, even if I look at them, I usually do not get involved in their evaluation.
 
@@
 
+Keep in mind that these computers would be university property, which might be monitored and have certain restrictions, e.g., do not install illegal software on them (\autoref{sec:illegal-software}).
+You might also have to return them when you graduate.
+
 \section{Negotiating PhD offer (e.g., having multiple offers)?}\label{sec:negotiate}
@@
 
-\section{Illegal Software} Using illegal/cracked software is very common in many countries (and even in the US).
+\section{Illegal Software}\label{sec:illegal-software}
+ Using illegal/cracked software is very common in many countries (and even in the US).
 However, \emph{do not} install or use them on university computers, even those given to you by your adviser.

### 2024-07-25 demystify.tex
updates

@@
 The size and the review load of the adcom depend on the department size.
-At Mason, the PhD adcom typically has 15--20 faculty, and each committee member is assigned to review 25--30 applications.
+At Mason, the PhD adcom typically has 15--20 faculty, and each committee member is assigned to review about 30 applications.
 Note that most large schools, including Mason, have separate adcoms for MS programs.
 
-Each application is assigned to about 3 faculty members, who will evaluate your profile and reach a consensus.
+Each application is assigned to about 3 adcom members, who will evaluate your profile and reach a consensus.
 While the assigned reviewers are the main ones deciding your application, other faculty in the department will likely have access to your application and can provide inputs and opinions on your profile.
@@
 
-For example, I only review those that are assigned to me, which is typically around 50--70 applications for the Fall admission. 
+For example, I only review those that are assigned to me.
 I do not look at any other applicants, unless I know them (e.g., they contacted me) or have some special circumstances (e.g., they are a student from Vietnam).
@@
 
-\paragraph{Core CS areas} Common CS areas that you should know are:
+\paragraph{Core CS topics} Common CS knowledge that you should know are:
 \begin{itemize}
-  \item Programming Foundation (e.g., programming in C/C++ or Java, OOP)
+  \item Programming Foundation (e.g., programming concepts in some modern languages)
   \item Discrete Math (e.g., logic, set theory, proof techniques)

### 2024-07-29 demystify.tex
updates

@@
 However, while it is true that the process is not perfect, it is not random.
-There are many possible reasons for rejections!  Other than not having a strong profile (e.g., research potential, GPA, LoRs, SOP, poor interviews), here are some other reasons.
+Other than not having a strong profile (e.g., research potential, GPA, LoRs, SOP, poor interviews), here are some other reasons for rejections.
 
@@
 
-We only review those that are assigned to us, which are already too many.
-Adcom chair will assign applications to reviewers based on their expertise (e.g., if a student says they want to do SE or interested in working with me).
-
-For example, I only review those that are assigned to me.
-I do not look at any other applicants, unless I know them (e.g., they contacted me) or have some special circumstances (e.g., they are a student from Vietnam).
-However, even if I look at them, I usually do not get involved in their evaluation.
+While adcom members (in fact, any faculty) can view any application, we only review those that are assigned to us, which are already too many.
+Adcom chair will assign applications to reviewers based on their expertise (e.g., if a student says they want to do SE or interested in working with me), and reviewers will only review those applications.
+Occasionally we might look at other applications, e.g., if we know the student or have some special circumstances (e.g., if that student contacted me, I know that student, they are from Vietnam, etc).
+However, even if we look at them, we usually do not get involved in their evaluation.
 

### 2024-07-30 demystify.tex
updates

--- 2024-07-25 demystify.tex
@@
 \usepackage{graphicx}
+\usepackage{pdfpages}
 \usepackage{hyperref}
@@
 
-% \section{Additional Resources}
-% \begin{itemize}
-%   \item \bibentry{nlpphd}
-% \end{itemize}
-
-% \chapter{Additional Resources}
-
-% \chapter{Undergraduate Research Opportunities}
-
-% Prospective PhD students are often unaware of the importance of research experience in their application.
-This provides both concrete research experiences and potential LOR writers that can talk about your research ability.
-% % This is especially true for students from small colleges where research opportunities are limited.
-% This section aims to provide some guidance on how to gain research experience as an undergraduate student.
-%This section is written together Didier, a PhD student at Mason who did his undergraduate study in Rwanda.
-
-% \paragraph{Locally} There are many ways to find research opportunities right in your own institution.
-If you did well or like a class, you can check with the professor of that class for research opportunities.
-Sometimes professors would mention that they look for undergraduates on their websites, so do check those out!
-Even if they say they are looking for graduate students, you still should contact them and ask.
-You can also ask your academic adviser or other faculty members for suggestions.
-Finally, you can also ask your peers who are already doing research.
-You can use the method described in \autoref{sec:contact} to contact professors
-
-% In short, the most important thing is that you need to \emph{ask}!
-Don't be shy, don't be afraid of rejection, and don't be afraid of being ignored or that you're not good enough.
-Just ask!
-The worst thing that can happen is that you get a ``no'' or no response.
-But you will never know if you don't ask.
-And if you get a ``no'', you can always ask again later.
-There are students who got rejected the first time but got accepted the second time.
-So don't give up!
-
-% \paragraph{Online Research Communities and Open Source Contributions}:
-% Places such as Github offer many opportunities for research, even though they might not appear so (e.g., they do not explicitly mention research).
-Students can explore repositories related to their areas of interest, collaborate with other contributors, and actively engage in open source research projects.
-By contributing code, fixing bugs, implementing new features, or providing documentation, students can gain practical research experience and interact with experienced developers and researchers in the field.
-
-% \paragraph{Paid or Unpaid}  travel opportunities, publishing opportunities, etc
-
-% \section{Virtual Research Opportunities Beyond Physical Boundaries (written by Didier)}
-% \didi{In my experience, prospective PhD students (from Rwanda) are often unaware of the information provided in this document throughout their undergraduate.
-For this reason they often find themselves unprepared for PhD admission process by the time they finish undergraduate and have to spend a year or more to improve their application profiles.
-I think it would be beneficial to expand the document to also target students who are not yet ready for the application process to help them improve on all criteria considered during admission.
-I drafted the short essay on how to gain research experience remotely in order to improve their research profile and potentially get LOR from experts (hopefully faculty)}\\
-% In the realm of computer science, research experience plays a crucial role in securing admissions to top-tier PhD programs.
-However, students from underrepresented small colleges, both in the United States and internationally, often encounter limited research opportunities within their institutions.
-Thankfully, there exists an alternative pathway for these aspiring researchers to gain valuable research experience, even without being physically present at a university.
-
-% \textbf{Virtual Research Programs}:
-% Several universities, organizations, and research institutions offer virtual internships and research programs aimed at providing hands-on research experience.
-These programs often involve working remotely under the guidance of experienced mentors and collaborating with a team of fellow researchers.
-For instance, \href{https://docs.google.com/forms/d/1btIwt4HwjyKMOUk-EMy3rbkfWzFxv2lNrMm_zkd0pA4/viewform?edit_requested=true}{UIUC+ Summer Undergraduate Research in Software Engineering}  offers an unpaid remote internship for software engineering students all over the world to collaborate with mentors from University of Illinois at Urbana Champaign.
-% Virtual internships offer opportunities to contribute to ongoing research projects, conduct experiments, analyze data, and write research reports.
-Students can search for such programs through university websites, research institutes, or as of recently chatGPT (or Bard).
-
-% \textbf{Online Conferences and Workshops}:
-% Attending online conferences and workshops is another way to gain exposure to cutting-edge research and establish connections with experts in the field.
-Many conferences now provide virtual participation options, enabling students to access research talks, poster sessions, and panel discussions and sometimes access designated chat rooms or networking events where participants can engage with researchers, ask questions, and seek potential research collaborations.
-It is beneficial to also create profiles academic collaboration platforms such as ResearchGate.
-By creating profiles, joining relevant research groups, and participating in discussions, students can connect with established researchers, contribute to ongoing projects, and potentially collaborate on publications or research proposals as they provide access
-
-% In conclusion, for students lacking research opportunities at their small colleges, the virtual realm offers a wealth of possibilities to gain valuable research experience.
-Engaging with open source projects, joining online research communities, utilizing academic collaboration platforms, participating in virtual internships, and attending online conferences are all avenues to contribute to research, connect with experts, and enhance one's research profile.
+\chapter{Visa and all that for International Students}\label{sec:visa}
+
+% \begin{center}
+%     \includegraphics[scale=0.8]{files/visa.pdf}
+% \end{center}
+
+For international students pursuing a PhD in the US, \textbf{F-1} is the main visa that allows a student to study full-time at an accredited institution.  Here are some key points about the F-1 visa:
+
+\begin{itemize}
+\item \textbf{Employment:} You are allowed to work on-campus for up to \textbf{20 hours per week during the academic year} (because you still need to take classes) and \textbf{full-time} (typically 40 hrs) during official school breaks (e.g., summer and winter breaks).
+Off-campus employment requires authorization, which can be obtained through programs like Curricular Practical Training and Optional Practical Training described below.
+
+\item \textbf{Curricular Practical Training (CPT):} This program allows you to participate in internships or practical training that is an integral part of their academic curriculum.
+CPT must be related to the student's field of study and can be full-time or part-time.
+It requires prior authorization from the Designated School Official (DSO) and must be completed before graduation.
+
+\item \textbf{Optional Practical Training (OPT):} OPT provides up to 12 months of work authorization for students before or after completing their degree.
+For students in STEM fields, there is an additional 24-month extension available.
+OPT requires prior application and approval from the United States Citizenship and Immigration Services (USCIS).
+
+\item \textbf{Full-time Enrollment:} You must maintain full-time enrollment status during the academic year.
+This means taking a minimum number of credits each semester, as defined by your program.
+Dropping below full-time status can result in loss of visa status.
+\end{itemize}
+
+\includepdf[pages=-]{files/visa.pdf}
+
+\section{For Spouses and Children}
+
+The spouses and children of F-1 visa holders can enter the US under the F-2 visa status. The F-2 visa allows family members to live in the US with the following conditions:
+
+For Spouse: 
+\begin{itemize}
+\item \textbf{Work Restrictions:} F-2 spouses are not permitted to work in the US.
+\item \textbf{Education:} They can study part-time but cannot enroll in full-time degree programs.
+\end{itemize}
+For Children:
+\begin{itemize}
+\item \textbf{Education:} F-2 children can attend K-12 schools but cannot pursue higher education full-time.
+\item \textbf{Work Restrictions:} Like F-2 spouses, children are not allowed to work under any circumstances.
+\end{itemize}
+F-2 visa holders must leave the US if the primary F-1 student visa holder loses status or completes their program.
 

### 2024-07-30 demystify.tex
updates

@@
 \begin{mybox}
-The latest version of this document is available at
+This document will be updated regularly to reflect the latest information and updates in the admission process. Its latest version is available at
 
@@
 Students understandably get excited about their upcoming PhD journey and want to buy new computer equipment and electronics to prepare.
-However, you should check with your prof.
-first.
+However, you should first check with your professor.
 They might have funding to buy you computers and other equipment (e.g., software, books, keyboards, headphones, tablets, etc).
@@
 
-\chapter{Visa and all that for International Students}\label{sec:visa}
+\chapter{Visa for International Students}\label{sec:visa}
 
@@
 
-\item \textbf{Curricular Practical Training (CPT):} This program allows you to participate in internships or practical training that is an integral part of their academic curriculum.
-CPT must be related to the student's field of study and can be full-time or part-time.
-It requires prior authorization from the Designated School Official (DSO) and must be completed before graduation.
+\item \textbf{Curricular Practical Training (CPT):} CPT allows you to participate in internships or practical training that is an integral part of their academic curriculum.
+CPT must be related to your field of study and can be full-time or part-time.
+It requires prior authorization from your university and must be completed before graduation.
 
@@
 For students in STEM fields, there is an additional 24-month extension available.
-OPT requires prior application and approval from the United States Citizenship and Immigration Services (USCIS).
+OPT requires prior application and approval from USCIS.
 

### 2024-07-30 demystify.tex
updates

@@
 \begin{center}
-  \includegraphics[scale=0.4]{files/c2.png}
+  \includegraphics[scale=0.5]{files/c2.png}
 \end{center}
@@
 For example, if you say you have experience with teaching, then show what you did, e.g., undergrad TA or mentoring someone.
-If you say you work on a project, then show some results, e.g., paper submitted (or even rejected), achieved X\% performance improvement over ...
+If you say you work on a project, then show some results, e.g., paper submitted (or even rejected), achieved certain performance improvement over the state of the art.
 Also, do not use more than two pages (references can be on the third page).
@@
 \chapter{Choosing Schools and Professors}\label{sec:schoolsandprofs}
+
+\epigraph{\vspace{-0.2in} The answer to life's problems aren't at the bottom of a bottle. Heh heh! They're on TV!}{\textsc{The Simpsons}}
 
@@
 
+\epigraph{\vspace{-0.2in} I didn’t do it. Nobody saw me do it. You can’t prove anything!}{\textsc{The Simpsons}}
+
 % \begin{center}
@@
 
+\epigraph{\vspace{-0.2in} I’m not a bad guy! I work hard, and I love my kids. So why should I spend half my Sunday hearing about how I’m going to Hell?}{\textsc{The Simpsons}}
+
 Most of what is written in this document applies to both domestic\footnote{For simplicity domestic means you did your undergrad (or MS) at a US university.} and international students.
-However, there are some differences and benefits that domestic students should be aware of.
+However, there are some differences and benefits that domestic students should be aware of and can leverage to improve their chances of admission.
 
@@
 
-\paragraph{Research Experiences \autoref{sec:research-opportunities}} You might have many opportunities to do research as an undergraduates, e.g., through REUs and internships at your undergrad university.
+\paragraph{Research Experiences \autoref{sec:research-opportunities}} You might have many opportunities to do research as an undergraduate, e.g., through REUs and internships at your undergrad university.
 Highlight such experiences in your application.
@@
 \chapter{Research Opportunities}\label{sec:research-opportunities}
+
+\epigraph{\vspace{-0.2in} Kids, you tried your best and you failed miserably. The lesson is: never try.}{\textsc{The Simpsons}}
+
 \begin{center}
@@
 \chapter{Cultural Differences}\label{sec:cultural}
+
+\epigraph{\vspace{-0.2in} Ay caramba!}{\textsc{The Simpsons}}
 
@@
 \chapter{Rankings of CS PhD programs}\label{sec:ranking}
+
+\epigraph{\vspace{-0.2in} The whole damn system is wrong!}{\textsc{The Simpsons}}
 
@@
 
+Also thanks to NSF its flexibility and allows faculty to be creative in research and education, which allows me to have time to write this document.
+
 \textbf{Thank you!}

### 2024-08-01 demystify.tex
updates

@@
 \begin{center}
-  \includegraphics[scale=0.5]{files/c1.png}
+  \includegraphics[scale=0.6]{files/c1.png}
 \end{center}
 
-International students often put schools into two bins: (i) very top schools that they dream about, and (ii) everything else.
+International students often put schools into two bins: (i) top schools that they dream about, and (ii) everything else.
 They commonly use rankings from US News, which is not transparent and questionable\footnote{\url{https://cra.org/cra-statement-us-news-world-report-rankings-computer-science-universities/}}.
 Sometimes they evaluate based on the reputation of the school's undergrad program or the reputation of the school's non-CS programs such as medical, math, or physics.
-They even rank universities based on popular states they know in the US, e.g., California, Texas, and New York.
-
-You can learn about CS programs and the research interests of faculty using resources such as \href{https://csrankings.org}{CSRankings.org}.
+They even rank universities based on popular places they know in the US, e.g., California, Texas, and New York.
+
+Instead of these superficial criteria, you should consider the CS program and the research interests of faculty members.
+You can learn about these using resources such as \href{https://csrankings.org}{CSRankings.org}.
 You will be very surprised to learn that a school that you didn't know much about can have very strong research in your topic (and vice versa, a school you thought highly about might have no faculty working in the research field you're interested in).
@@
 
+\subsection{PhD in other related fields: CE, IST, Cybersecurity}\label{sec:related-fields}
+
+You \emph{do not} need to do a PhD in CS to work in your area of interest.
+For example, in addition to a traditional CS department, Mason has IST and Cybersecurity departments, both of which have faculty with PhD in CS and work on CS topics (e.g., AI, Security, Robotics).
+So it is possible that you still get to do CS research and publish in CS-focused venues even if you're not in a traditional CS program.
+It is common to see faculty with CS PhD in a non-CS department as well as faculty with non-CS PhD in a CS department.
+
+However, if your goal is a PhD in CS, then you likely need to be in the CS dept \emph{and} advised by a CS faculty.
+A non-CS faculty can serve in the PhD dissertation committee (common) or \emph{co-advise} (somewhat rare), but your main PhD adviser will likely be a tenure-line faculty in CS (\autoref{sec:faculty-types}).
+For example, a prof. in Stats or Math might be able to serve as a co-adviser, but not as a sole adviser of a student in a CS PhD program. 
+If in doubt, check with the CS department for their requirements.
+
+For this specific reason,  CSRankings includes only faculty who can advise CS PhD students.
+I also have compiled a \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{list} of Vietnamese faculty who can advise PhD students.
+\autoref{sec:faculty-types} talks more about who can serve as your PhD adviser.
+
 \section{Choosing an Adviser}
@@
 \end{itemize}
-\subsection{Types of Faculty: Who can serve as a PhD adviser?}\label{sec:faculty-types}
+\subsection{Who can serve as a PhD adviser?}\label{sec:faculty-types}
 
@@
 
-Tenure-track faculty such as assistant professors are likely young and very active in research (they have to, to get tenure).
-Thus, they will likely have more time to work with you and push you to do research and publish.
+Now that you know a bit about tenured and tenure-track faculty (\autoref{sec:faculty-types}).
+Which one should you choose as your adviser?
+The short answer is that either can be a good adviser, but they have different strengths and weaknesses.
+
+Tenure-track faculty, e.g., assistant professors, are likely young and active in research (they have to, to get tenure).
+Thus, they will devote more time to work with you and push you to do research and publish.
 However, they may not have as much experience in managing students and may not have as much funding (yet).
@@
 
-\subsection{PhD in other related fields: CE, IST, Cybersecurity}\label{sec:related-fields}
-
-You \emph{do not} need to do a PhD in CS to work in your area of interest.
-For example, in addition to a traditional CS department, Mason has IST and Cybersecurity departments, both of which have faculty with PhD in CS and work on CS topics (e.g., AI, Security, Robotics).
-So it is possible that you still get to do CS research and publish in CS-focused venues even if you're not in a traditional CS program.
-It is common to see faculty with CS PhD in a non-CS department as well as faculty with non-CS PhD in a CS department.
-
-However, if your goal is a PhD in CS, then you likely need to be in the CS dept \emph{and} advised by a CS faculty.
-A non-CS faculty can serve in the PhD dissertation committee (common) or \emph{co-advise} (somewhat rare), but your main PhD adviser will likely be a tenure-line faculty in CS (\autoref{sec:faculty-types}).
-For example, a prof. in Stats or Math might be able to serve as a co-adviser, but not as a sole adviser of a student in a CS PhD program. 
-If in doubt, check with the CS department for their requirements.
-
-For this specific reason,  CSRankings includes only faculty who can advise CS PhD students.
-I also have compiled a \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{list} of Vietnamese faculty who can advise PhD students.
-\autoref{sec:faculty-types} talks more about who can serve as your PhD adviser.
+\subsection{Things to avoid when cold emailing a professor}
+
+\begin{itemize}
+
+    \item \textbf{Not customizing the email}.
+Many students write a generic email and send it to multiple professors.
+Tthese emails typically talk about the student and not why they are interested in the professor's work.
+See writing tips in \autoref{sec:contact}
+    \item \textbf{Sending the same email to multiple professors}.
+Some students send to multiple emails using BCC or even CC (where we can see all addresses the email being sent to).
+This is a big turn-off for professors and will result in no reply or a very harsh reply on how unprofessional you are.
+    \item \textbf{Too much information}. Keep your email short and to the point. If we are interested, we will ask for more information.
+    \item \textbf{Not being professional}, e.g., do not call the professor by their first name in the first email.  Use Prof. Lastname or Dr. Lastname (\autoref{sec:address}).
+    \item \textbf{Not reading the professor's website} and asking questions that are already answered there.
+\end{itemize}
 
@@
 
-Also thanks to NSF its flexibility and allows faculty to be creative in research and education, which allows me to have time to write this document.
+Also thanks to NSF for encouraging faculty to be creative in research and education, which allows me to write this document.
 

### 2024-08-02 demystify.tex
updates

@@
 A good LoR letter should give details of the applicant's research background and achievements, all of which paint a compelling picture of the candidate to the adcom.
-
 Having a strong letter from an internationally recognized researcher will \emph{greatly strengthen} your application.
@@
 
-\paragraph{Self-written Letters} Many international students write letters themselves, typically due to the requests of their profs., and have them signed by their profs.
-Such letters have little value and are considered weak by reviewers---why can you not even find someone who cares or knows enough about you to write a candid personal reference letter?
+\paragraph{Why Self-written Letters are Bad?} Many international students write letters themselves, typically due to the requests of their profs., and have them signed by their profs.
+Such letters have \textbf{very little value} and are considered weak by reviewers---why can you not even find someone who cares or knows enough about you to write a candid personal reference letter?
 Instead of the reference writer talking about you, in this case it is you who write about yourself (and they just sign the letter).
-Moreover, experienced reviewers can easily spot (or suspect) such letters, which can be a \red{red flag}.
-
-\paragraph{Helping your LOR writer} As mentioned above, do not write your own letter.
-Instead, you can help your LOR writer by providing details or unique things about yourself.
-For example, let them know about your GPA, research and work experiences, papers (if any), or anything you want them to mention.
-You can also provide them with a draft of your SOP so that they can see what you are saying about yourself and complement that with their own perspective.
-Sometimes your writer will explicit ask you for such information, but if not, you should provide it anyway (especially if you have not interacted with them much or have not done much research with them).
-
-\paragraph{Generic Letters} When the writers do not know much about the applicants (e.g., just taking some course with them or not making any impression to write about), they might write a generic and short letter, which is not useful and also considered weak.
-It might be a good idea to directly ask if the prof.
+Experienced reviewers can easily spot (or suspect) such letters, which can be a \red{red flag}.
+
+\paragraph{Why Generic Letters are Bad?} When the writers do not know much about the applicants (e.g., just taking some course with them or not making any impression to write about), they might write a generic and short letter, which is not useful and also considered weak.
+This does not mean the professor is not good or does not care about you, but they just do not know you well enough to write a strong letter.
+So it might be a good idea to directly ask if the prof.
 is willing to write a \emph{strong} letter for you.
@@
 
-Sometimes students would go to great length just to get letters from well-known professors in their school (e.g., department head).
+Sometimes students would go to great lengths just to get letters from well-known professors in their school (e.g., department head).
 But as mentioned, if these professors do not know you, their letters would likely be generic and carry little value (sometimes \red{red flags}).
@@
 
-\paragraph{Waiving Your Right}  Choosing not to look at a reference letter is pretty standard in school and job applications.
+\paragraph{Why You Should Waive Your Right}  Choosing not to look at a reference letter is pretty standard in school and job applications.
 When you waive your right to see the letter, it adds a layer of trust, showing you're confident in your choice of referees and that you're not trying to twist their words.
@@
 
-\paragraph{Reminding Writers} After you submit your application, you should tell your writers about that and let them know they will soon receive an email from the university to submit their letters.
-You should also remind them about the deadline and ask them to submit their letters on time.
-But do not send too many reminders as that can be annoying to the writers.
-
+\paragraph{Why You Want to Help your LOR Writer} As mentioned above, do not write your own letter and generic letters do not give much value.
+One thing you can do to help your writer to write a strong, customized LoR for you is providing details or unique things about yourself.
+For example, let them know about your GPA, research and work experiences, papers (if any), or anything you want them to mention.
+You can also provide them with a draft of your SOP so that they can see what you are saying about yourself and complement that with their own perspective.
+Sometimes your writer will explicitly ask you for such information, but if not, you should provide it anyway (especially if you have not interacted with them much or have not done much research with them).
+
+\paragraph{Why You Need to Remind Your Writers} After you submit your application, you should tell your writers about that and let them know they will soon receive an email from the university to submit their letters.
+You should also remind them when you submit your application and ask them to submit their letters on time if they haven't done so.
 Note that most places only have deadlines for the applicant, but are very flexible with the letter writers (in many cases do not even give them any deadline).
 Also, many places do not begin the admission review process right after the deadline and work on application reviews in the next semester (mid-January).
+
+Thus you do want to send reminders because professors can be quite busy and might forget to submit their letters, especially when there is no explicit deadline.
+However, do not send too many reminders as that can be annoying to the writers.
 
@@
 So it is possible that you still get to do CS research and publish in CS-focused venues even if you're not in a traditional CS program.
-It is common to see faculty with CS PhD in a non-CS department as well as faculty with non-CS PhD in a CS department.
+It is common to see faculty with PhD in CS in a non-CS department as well as faculty with non-CS PhD in a CS department.
 
@@
 
-    \item \textbf{Not customizing the email}.
-Many students write a generic email and send it to multiple professors.
-Tthese emails typically talk about the student and not why they are interested in the professor's work.
-See writing tips in \autoref{sec:contact}
+    \item \textbf{Not customizing the email}. These emails typically talk about the student and not why they are interested in the professor's work. See writing tips in \autoref{sec:contact}
     \item \textbf{Sending the same email to multiple professors}.
-Some students send to multiple emails using BCC or even CC (where we can see all addresses the email being sent to).
-This is a big turn-off for professors and will result in no reply or a very harsh reply on how unprofessional you are.
-    \item \textbf{Too much information}. Keep your email short and to the point. If we are interested, we will ask for more information.
+Sending emails to multiple profs.
+using BCC or even CC (where we can see all addresses the email is being sent to) will result in no reply or a very harsh one on how unprofessional this is.
+    \item \textbf{Too much information}.
+Keep your email short and to the point.
+If we are interested, we will ask for more information.
+2--3 paragraphs are enough.
+Personally I also do not want to see course transcript or CV in the first email.
     \item \textbf{Not being professional}, e.g., do not call the professor by their first name in the first email.  Use Prof. Lastname or Dr. Lastname (\autoref{sec:address}).

### 2024-08-04 demystify.tex
updates

@@
 In my opinion, requiring applicants to pay the fee helps ensure their seriousness, as it filters out non-serious candidates.
-Also, most CS programs already receive many applications, and would be overwhelmed if the application process were free.
+Also, most CS programs already receive many applications and would be overwhelmed if the application process were free.
 
@@
 
-\paragraph{Why Self-written Letters are Bad?} Many international students write letters themselves, typically due to the requests of their profs., and have them signed by their profs.
+\paragraph{Self-written letters are bad} Many international students write letters themselves, typically due to the requests of their profs., and have them signed by their profs.
 Such letters have \textbf{very little value} and are considered weak by reviewers---why can you not even find someone who cares or knows enough about you to write a candid personal reference letter?
@@
 
-\paragraph{Why Generic Letters are Bad?} When the writers do not know much about the applicants (e.g., just taking some course with them or not making any impression to write about), they might write a generic and short letter, which is not useful and also considered weak.
+\paragraph{Generic letters are bad} When the writers do not know much about the applicants (e.g., just taking some course with them or not making any impression to write about), they might write a generic and short letter, which is not useful and also considered weak.
 This does not mean the professor is not good or does not care about you, but they just do not know you well enough to write a strong letter.
@@
 
-\paragraph{Why You Should Waive Your Right}  Choosing not to look at a reference letter is pretty standard in school and job applications.
+\paragraph{Waiving Your Right}  Choosing not to look at a reference letter is pretty standard in school and job applications.
 When you waive your right to see the letter, it adds a layer of trust, showing you're confident in your choice of referees and that you're not trying to twist their words.
@@
 
-\paragraph{Why You Want to Help your LOR Writer} As mentioned above, do not write your own letter and generic letters do not give much value.
-One thing you can do to help your writer to write a strong, customized LoR for you is providing details or unique things about yourself.
+\paragraph{Help your LOR Writer} As mentioned above, do not write your own letter and generic letters do not give much value.
+One thing you can do to help your writer to write a strong, customized LoR for you is to provide details or unique things about yourself.
 For example, let them know about your GPA, research and work experiences, papers (if any), or anything you want them to mention.
@@
 
-\paragraph{Why You Need to Remind Your Writers} After you submit your application, you should tell your writers about that and let them know they will soon receive an email from the university to submit their letters.
+\paragraph{Remind Your Writers} After you submit your application, you should tell your writers about that and let them know they will soon receive an email from the university to submit their letters.
 You should also remind them when you submit your application and ask them to submit their letters on time if they haven't done so.
@@
 
-In your statement, you have the opportunity to connect directly with us and make your application stand out and unique.
+In your statement, you have the opportunity to connect directly with adcom members and make your application stand out and unique.
 Your SOP can convince adcom members that you would \emph{fit} the program, even if you don't have very strong research experience.
@@
 
-\textbf{Advice}: If your school is the top in your country but you think it might not be well-known internationally, you should talk about it in your SOP (and also ask your LOR writer to do that).
+If your school is the top in your country but you think it might not be well-known internationally, you should talk about it in your SOP (and also ask your LOR writer to do that).
 Of course, if you're interested in working with Vietnamese, consider  \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{CS programs in the US that have Vietnamese professors}.
@@
 
-\textbf{Advice}: As with school reputation, you and your LoR writers can mention the grading system of your university if you think that is helpful for adcom to evaluate.
+As with school reputation, you and your LoR writers can mention the grading system of your university if you think that is helpful for adcom to evaluate.
 Note that while having good grades might not help much,
@@
 
-\paragraph{GRE} While a few schools still require taking the GRE (e.g., UCF), most good CS PhD programs in the US \textbf{no longer} require it.
+\paragraph{GREs Do Not Matter} While a few schools still require taking the GRE (e.g., UCF), most good CS PhD programs in the US \textbf{no longer} require it.
 Thus, if you haven't taken the GRE or have bad scores, then don't waste time (re)taking it.
@@
 Note that English exams are not required for all students, e.g., domestic students and those from certain countries are waived from taking these exams.
-For example, at Mason also waive the English exams for students from these \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-Mason#standard-tests-waiver-eligible-countries}{countries}.
-
-\paragraph{Red flags} Various types of red flags can cause concern to the adcom (e.g., plagiarism, cheating, or other academic dishonesty).
+For example, Mason waives English exams for students from these \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-Mason#standard-tests-waiver-eligible-countries}{countries}.
+
+\paragraph{Red flags} Various types of red flags can cause concern to the adcom (e.g., too many STEM courses with low or withdraw grades, plagiarism, cheating, or other academic dishonesty).
 Another one is that you have a history of jumping from one program to another (e.g., you have been in multiple different programs in a short time).
@@
 
-\paragraph{Rejection Letters} Typically, schools start sending out rejection letters to remaining applicants after they have finalized their admissions decisions.
+\paragraph{Rejection Letters} Schools typically start sending out rejection letters to remaining applicants \emph{after they have finalized their admissions decisions}.
 Thus, rejection letters are often sent out late (e.g., after April 15th or even much later).
 Not much you can do here, you can try to contact the school to ask about your status, but they might not reply, they might say they are still reviewing applications, or give you inaccurate information (e.g., you will hear in two weeks).
-In short, you just have to be patient and wait.
+In short, you just have to be patient and wait (however, beware that some schools do not send out rejection letters at all).
 

### 2024-08-04 faculty.tex
updates

@@
     {\fontsize{24}{24} \selectfont {{Demystifying Tenure Track:}}} \\[0.15in]
-    {\fontsize{18}{18} \selectfont {{Navigating the Faculty Journey In Computer Science}}} \\[1in]    
+    {\fontsize{18}{18} \selectfont {{My Academic Journey in Computer Science}}} \\[1in]    
     %{\fontsize{18}{18} \selectfont {{A Handbook for International and Domestic Students}}} \\[0.5in]
@@
 \chapter{Research}
-Research here focuses on paper publication, which is a major component in getting your name out there and of course in your departmental evaluation.
+This section focuses on paper publication, which is a major component in getting your name out there and of course in your departmental evaluation.
+
+\section{What is needed for tenure?}
+
+Publications are a major component of tenure evaluation.
+There are no specific numbers of papers or which conferences/journals are required, but typically you need to have a good number of papers in top venues to get tenure.
+Also work that won awards (e.g., best paper awards or 10-year most influential awards) also count a lot in tenure evaluation.
+Citations are not required but work with very high citations would help.
+
+% 5.1 Evaluating Research from the Point of View of Tenure
+
+% 	•	Key Components: Publication record and expert letters.
+% 	•	Consistency & Impact: Steady publication in top venues, with standout work defining one’s reputation.
+% 	•	Research Independence: Importance of showing independent work, separate from past advisors.
+
+% 5.2 Coming up with a Research Agenda
+
+% 	•	Research Agenda vs. Project: A broad goal guiding multiple projects.
+% 	•	Criteria: Inspiring, technically challenging, and doable work, ideally tied to the professor’s expertise.
+
+% 5.3 Picking Research Problems
+
+% 	•	Importance of Selection: Prioritizing meaningful, high-impact problems.
+% 	•	Criteria: Should be either intellectually elegant or practically useful, balancing theory and practical applications.
+
+% 5.4 Knowing When to Give Up
+
+% 	•	Decision Making: Assessing progress and potential, setting early criteria for success, and being willing to pivot if necessary.
+
+% 5.5 Managing Collaborations
+
+% 	•	Balancing Collaborations: Limiting them to maintain focus on primary research.
+% 	•	Industrial Collaborations: Handling agreements, IP, and publication protocols.
+
+% Summary
+
+% The chapter covers developing a research agenda, selecting impactful problems, knowing when to pivot, and managing collaborations, emphasizing focus and strategic decision-making.
 
@@
 
+\section{From a Tenure Evaluation Point of View}
+Without a doubt, grants are one of the most, if not the most, important component of tenure evaluation.
+The most common reason for not getting tenure is not having enough funding.
+In an ideal world, research results, e.g., publications, would be the more important; but in reality, funding is often considered more important for tenure evaluation at most R1 universities.
+The amount of funding you have is a direct reflection of your research impact and productivity.
+It is also a reflection of your ability to write proposals, manage projects, and work with collaborators.
+More realistically, grants are needed to pay for your students, travel conference, research equipments, and even your own summer salary.
+
+In general, the more funding you have, the better.
+However, the quality of the funding is also important.
+For example, external NSF grants are more prestigious than an internal grant.
+Also, funding from industry, while has a much smaller amount, is often highly competitive, prestigious, and can help show that your research is of interest to industry.
+In general, you should aim to get funding from a variety of sources, including NSF, Defense, industry, and internal grants.
+This will show that you are able to work with different types of sponsors and that your research is of interest to a wide range of people.
+
 \section{Types of Funding}
@@
 These funding are less competitive than NSF, but they are also less prestigious (e.g., they aren't much and usually don't count as much for tenure).
-However, they are good sources of funding to get preliminary results and later apply for traditional grants (and many internal funding programs explicitly encourage this).
+However, they are good sources of funding to get preliminary results and later apply for traditional grants.
 
@@
 Industrial gifts are \emph{unrestricted funding} from companies such as Amazon, Microsoft, Facebook, and Google.
+While some are connection-based (e.g., for your ongoing work with a collaborator at Facebook), most are based on the merits of your research.
 These are highly competitive and prestigious (e.g., the solicitation is open worldwide, and you compete with top researchers from top places).  
-They are the best kind of funding for three reasons: (1) indirect costs or overheads are typically not levied on these gifts (so you get most of the money), (2) there is no associated timeline by when you should spend them (they give it to you and forget about it), and (3) you can use them for a broader range of things than university and grant/contract funding (e.g., summary salary, equipments, or even bowling events for your group).
+They also show that your research has values and impact beyond academia and is interested to industry.
+Practically, they are the best kind of funding for three reasons: (1) indirect costs or overheads are typically not levied on these gifts (so you get most of the money), (2) there is no associated timeline by when you should spend them (they give it to you and forget about it), and (3) you can use them for a broader range of things than university and grant/contract funding (e.g., summary salary, equipments, or even bowling events for your group).
 
@@
 I have tried to get other DoD funding, but without much success, I found that DoD program managers are very picky and often have their own agenda.
-I have also received internal funding from my university and industrial gifts from Amazon,and Facebook.
+I have also received internal funding from my university and industrial gifts from Amazon, and Facebook.
 I have found that industrial gifts are the best kind of funding, as they are unrestricted and have no associated timeline by when you should spend them.
-They are also highly competitive and prestigious, and are a great way to get funding for your research.
+They are also highly competitive and prestigious, and are a great way to show research impact.
 
@@
 \chapter{Teaching}
+
+\section{From a P\&T Point of View}
+
+While teaching should take about 40\% of your time and your salary is mainly for teaching, teaching is not as important as research for tenure.
+Moreover, for most new faculty, teaching is not as difficult as getting grants.
+Thus, tenure-track faculty often do not worry much about teaching.
+
 \section{Teaching Load}

### 2024-08-06 demystify.tex
updates

@@
 You can also ask if you need to prepare for a coding problem.
-Finally, prepare some thoughtful questions to ask, e.g., about the program, the university, or the professor's research.
+Finally, prepare some thoughtful questions to ask, e.g., about the program, the university, or the professor's research (see \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Answers-to-Ph.D-Advisor-Guide}{some questions} you can ask a potential adviser).
 
@@
 
-\paragraph{Red flags} Various types of red flags can cause concern to the adcom (e.g., too many STEM courses with low or withdraw grades, plagiarism, cheating, or other academic dishonesty).
+\paragraph{Red flags} Various types of red flags can cause concern to the adcom (e.g., too many STEM courses with low or withdrawn grades, plagiarism, cheating, or other academic dishonesty).
 Another one is that you have a history of jumping from one program to another (e.g., you have been in multiple different programs in a short time).

### 2024-08-06 faculty.tex
updates

@@
 
-\section{From a Tenure Evaluation Point of View}
-Without a doubt, grants are one of the most, if not the most, important component of tenure evaluation.
+\section{For P\&T}
+Without a doubt, grants are one of the most, if not \emph{the} most, important factors in P\&T (or any kind of performance) evaluation.
 The most common reason for not getting tenure is not having enough funding.
@@
 In general, the more funding you have, the better.
-However, the quality of the funding is also important.
-For example, external NSF grants are more prestigious than an internal grant.
+However, the \emph{quality} of the funding is also important.
+For example, external NSF grants are more prestigious than an internal grant because they give more money and are also more competitive.
 Also, funding from industry, while has a much smaller amount, is often highly competitive, prestigious, and can help show that your research is of interest to industry.
-In general, you should aim to get funding from a variety of sources, including NSF, Defense, industry, and internal grants.
+Finally, you should aim to get funding from a variety of sources, including NSF, Defense, industry, and internal grants.
 This will show that you are able to work with different types of sponsors and that your research is of interest to a wide range of people.
@@
 \paragraph{NSF} NSF is the obvious funding source for CS. 
-Two things stand out for NSF grants: (transparency) proposals are evaluated by an independent panel and are based on merit rather than personal connections; (autonomy) the flexibility and freedom given to focus on pure research without the pressure of producing concrete deliverables.
+Two things stand out for NSF grants are (i.
+\emph{transparency}) proposals are evaluated by an independent panel and are based on merit rather than personal connections, and (ii.
+\emph{autonomy}) the flexibility and freedom given to focus on pure research without the pressure of producing concrete deliverables.
+
 Because of these reasons, NSF funding is often considered the prestigious and competitive, even for very experienced senior researchers from top places (which don't mean anything).
@@
 
-However, these sponsors are good sources of funding and can provide the additional resources (e.g., summer months for faculty).
+However, these sponsors are good sources of funding and can provide the additional resources (e.g., your "3rd" non-NSF summer month).
 Note that some defense programs, e.g., Young Investigator Award, are more like NSF grants, with less stringent requirements and no deliverables required.
@@
 \section{Most people suck at grant writing (at first)} Most new assistant professors have no experience in grant writing, and struggle with it in the beginning.
-Then eventually they will get used to it and learn to like it (or at least do not hate it as much).
-So don't get discouraged.
-Keep trying, and you will get better over time.
-
-\paragraph{It is very different than writing research papers} Grant writing is a very different skill than writing research papers.
-In research papers, you describe what you \emph{did}.
-You convince people the merits of your work by showing the technical details, e.g., the algorithm,math, and proofs, and presenting experiments and results supporting your claim.
-You don't need to tell people who you are (often the authors are anonymous), and you don't need to convince people that you are the right person to do the work.
-Your reviewers are typically experts in your field and know a bit about the paper topic, and they are judging your work based on the technical merits.
-
-In grant writing, you are trying to convince people to give you money to do something you \emph{haven't done yet}.
-You need to convince people at a much higher level that your idea is important (impactful) and that you are the right person (with the right collaboration, resource, and plan) to do it.
-Getting too much into technical details can be a turn off.
-Your reviewers might not also experts in your field, and they are judging your work based on the \emph{big picture}.
-
-However, it's a skill that can be learned.
-The best way to learn is to read other people's grants, and to write your own.
-You will likely get rejected many times, but that's normal.
-You will get better over time.
+It becomes especially stressful starting your 3rd year when departmental evaluation starts to give you more pressure and warning if you don't have enough funding.
+By your midterm or fourth-year review, you are expected to have at least some external funding, and can be in serious trouble if you don't have any.
+So start early, and don't wait until the last minute to start writing grants.
+It's a skill that can be learned, and you will get better over time.
 The key is to keep trying, and to not give up.
@@
 
-It's a very different skill than writing research papers.
+\paragraph{It is very different than writing research papers} Grant writing is a very different skill than writing research papers.
+In research papers, you describe what you \emph{did}.
+You convince people the merits of your work by showing the technical details, e.g., the algorithm,math, and proofs, and presenting experiments and results supporting your claim.
+You don't need to tell people who you are (often the authors are anonymous), and you don't need to convince people that you are the right person to do the work.
+Your reviewers are typically experts in your field and know a bit about the paper topic, and they are judging your work based on the technical merits.
+
+In grant writing, you are trying to convince people to give you money to do something you \emph{haven't done yet}.
+You need to convince people at a much higher level that your idea is important (impactful) and that you are the right person (with the right collaboration, resource, and plan) to do it.
+Getting too much into technical details can be a turn off.
+Your reviewers might not also experts in your field, and they are judging your work based on the \emph{big picture}.
+
 However, it's a skill that can be learned.
@@
 
-While teaching should take about 40\% of your time and your salary is mainly for teaching, teaching is not as important as research for tenure.
-Moreover, for most new faculty, teaching is not as difficult as getting grants.
+While teaching should take about 40\% of your time (equivalent to research) and your salary is mainly for teaching, teaching is not as nearly as important as research for tenure.
+Moreover, for most new faculty, teaching is not as difficult as getting grants or publishing papers.
 Thus, tenure-track faculty often do not worry much about teaching.
@@
 There will be times when you feel discouraged and frustrated, but there will also be times when you feel proud and happy.  
-For tenure you will be expected to mentor undergraduate students and graduate students. You will likely need to graduate at least one PhD student to get tenure.
+For P\&T you will be expected to have successes in mentoring students.  For most good universities, eou will likely need to graduate at least one PhD student to get tenure.
 

### 2024-08-07 nnv.tex
create NNV book

--- 2024-08-06 demystify.tex
@@
 
-\paragraph{School} Graduating from top universities \emph{that we recognize} helps.
+\subsection{School} Graduating from top universities \emph{that we recognize} helps.
 For example, if you are an international student and your school is well-known, then it is \emph{``top foreign''}, which is a plus.
@@
 
-\paragraph{Grades} Having a good GPA is good. However if your school is not well-known, having top grades or rankings
-usually will not help because we cannot evaluate them.
+\subsection{Grades} Grades generally do not matter much for CS PhD admission.
+PhD in CS is a research degree and doing well in undergrad courses does not necessarily mean you can do research (other factors such as research experiences, pubs, LoRs are more important).
+
+Nonetheless, if you are from a well-known school, having good grades do help (not a lot though), e.g., adcom members often note details such as "good GPA from well-known school (\autoref{sec:ievaluate}).
+However if your school is not well-known, having top grades or rankings usually will not help because we cannot evaluate them (e.g., we don't know how hard it is to get a 4.0 or A's at your school).
 This can be an issue for students in many top international universities where the competition is so high that very good students can still have low rankings from these schools (and be overlooked by adcom).
@@
 As with school reputation, you and your LoR writers can mention the grading system of your university if you think that is helpful for adcom to evaluate.
-Note that while having good grades might not help much,
-having bad grades (e.g., < 3.0) will be \red{red flag} (unless your LORs or
-statement give a proper explanation). This is especially true if you
-have bad grades in relevant courses, e.g., Math and CS.
 
@@
   \end{commentbox}
+
+\paragraph{Bad Grades} 
+While having good GPA might not help much (again, because of other much more important factors),
+having bad GPA can be a \red{red flag}. In fact, many universities have a minimum GPA requirement (e.g., > 3.0) and will automatically reject applications with lower GPAs.  
+If you have bad grades, you should explain them in your SOP or have your LOR writers explain them in their letters.
+
+Moreover, having bad grades in relevant courses, e.g., Math and CS, can be a \red{red flag}.
+Adcom members often scan through transcripts looking for C and lower grades in Math and CS courses and might raise concerns if they see several of them.
+Note that bad grades in non-relevant courses, e.g., e.g., about politics or history, are not as concerning.
+
+A common question is whether you should be concern and explain bad grades in relevant courses in your SOP.
+If you have just a few, then they do not matter much, so don't spend much time explaining them (let's face it, many of us, including faculty, have bad grades in some courses, including relevant ones).
+But if you have many bad grades for an entire semester or year due to some specific reasons, then you may be you should explain them in your SOP.
 

### 2024-08-07 nnv.tex
nnv bib

--- 2024-08-04 demystify.tex
@@
 
-    % Year
-    \draw[ultra thick,gray]
-    ($(current page.center)+(5,2)$) -- ++(0,-3cm) 
-    node[
-    midway,
-    left=0.25cm,
-    text width=5cm,
-    align=right,
-    black!75
-    ]
-    {
-    {\fontsize{25}{30} \selectfont \bf Computer\\[8pt] Science}
-    } 
-    node[
-    midway,
-    right=0.25cm,
-    text width=6cm,
-    align=left,
-    ForestGreen]
-    {
-    {\fontsize{48}{50} \selectfont PhD}
-    };
-
     % Title
@@
     {
-    {\fontsize{24}{24} \selectfont {{Demystifying the Computer Science}}} \\[0.15in]
-    {\fontsize{24}{24} \selectfont {{PhD Admission in the US}}} \\[0.2in]    
-    {\fontsize{18}{18} \selectfont {{A Handbook for International Students}}} \\[0.5in]
+    {\fontsize{38}{1} \selectfont {{Neural Network Verification}}}\\[0.5in]
 
-    {\fontsize{16}{19.2} \selectfont \textcolor{ForestGreen}{ \bf ThanhVu (Vu) Nguyen}}\\[0.1in]
-    George Mason University, Dept. of Computer Science\\[0.1in]
+    {\fontsize{14}{19.2} \selectfont \textcolor{ForestGreen}{ \bf ThanhVu (Vu) Nguyen}}\\[0.1in]
     \today{} (latest version available on  \href{https://github.com/nguyenthanhvuh/phd-cs-us}{Github})
@@
 
-This section summarizes the main points of this guideline. It also gives you an overview to decide which specific topics you want to explore more thoroughly.
-
-\begin{enumerate}
-  \item Should you apply to a PhD CS program in the US?
-        \begin{itemize}
-          \item \emph{Yes, definitely}.  CS PhD study in the US is fully funded and admission into good universities is not any harder than in other countries (\autoref{sec:should}).
-        \end{itemize}
-  \item How is your application evaluated?
-        \begin{itemize}
-          \item Applications are evaluated by the \emph{PhD Admission} (adcom) committee and each application is typically reviewed by three faculty members (\autoref{sec:evalapps}).
-          \item Individually faculty member \emph{cannot directly admit} a student---so do not email and ask if you have a chance.
-However, faculty can \emph{advocate} for a student and therefore increase their admission chance---so do contact and introduce yourself (\autoref{sec:contact}).
-         \end{itemize}
-  \item Application Materials
-        \begin{itemize}
-          \item The committee will look at various factors, but the most important ones are letters of recommendation (LORs),  statements of purpose (SOP), and research experience and potential, e.g., publications.
-          \item \emph{LORs are very important}, but only if they are personalized and talk about your research ability (\autoref{sec:lor}).
-          \item \emph{SOP is very important}. Write it in such a way that makes you \emph{stand out} (\autoref{sec:research-statement} and \autoref{sec:improve-your-chance})
-          \item  \emph{Publications can greatly help}.
-Papers in good venues are concrete evidence that you have successfully engaged in research.
-But they are not required!
-(\autoref{sec:research-experiences}).
-
-          \item \emph{High grades probably won't help much} (unless you're from a well-known school), but bad ones likely will raise concerns (\autoref{sec:your-school}).
-
-          \item \emph{Standard tests are not important} (\autoref{sec:standard-tests}).
-GRE typically \emph{is not} required.
-For standard English tests (not required for domestic students), just do enough to pass the minimum requirements.
-
-          \item \emph{Contacting a prof. is recommended}, but do it \emph{properly} to get a reply (\autoref{sec:contact}).
-
-          \item Getting an interview is typically a \emph{good sign}; but no interview does not mean rejection (\autoref{sec:interview}).
-
-        \end{itemize}
-  \item What to do after getting admitted?
-        \begin{itemize}
-                \item  Celebrating! Now it is your turn to evaluate the school!
-          \item Attend \emph{Open House} to learn more about the place and \emph{interview} profs---they would be much more willing to talk to you now (\autoref{sec:accepted}).
-        \end{itemize}
-  \item Funding
-        \begin{itemize}
-          \item TAs, RAs, and fellowships are main funding sources for PhDs (\autoref{sec:funding}).
-TAs (teaching assistantships) are provided by the department to help with classes (e.g., grading).
-RAs (research assistantships) are given by profs.
-to help with their research.
-Fellowships, provided by the university, department, or external sources such as government or industry, give move flexibility but are very competitive.
-        \end{itemize}
-  \item Choosing School and Professors
-        \begin{itemize}
-          \item Not every professor, even those in CS, can advise or graduate CS PhD students (\autoref{sec:schoolsandprofs}).
-        \end{itemize}
-
-    \item Miscs and FAQS
-        \begin{itemize}
-          \item One reason why you were rejected (\autoref{sec:why-rejected}) is because you applied to AI/ML, a super competitive field in recent years with many applicants.
-
-          \item Why rejection letters are sent so late?
-(\autoref{sec:late-rejection}).
-Grad programs often wait for the accepted students to make their decisions, typically by April 15, before sending out rejection letters.
-
-          \item Increasing admission chance by being unique and standing out (\autoref{sec:improve-your-chance}).
-          \item It takes about 10 minutes to review an application (\autoref{sec:ievaluate}).
-          \item Sometimes adcom members discuss applicants, but in most cases they make independent decisions (\autoref{sec:adcom-discuss}).           
-          \item Adcom members only review applications assigned to them (typically matching their expertise) and rarely get involved in other applications (\autoref{sec:read-all}).
-
-          \item Being a minority can significantly help your application (\autoref{sec:urm}).
-          \item You can successfully apply to CS PhD even if you have non-CS background (\autoref{sec:non-stem}).
-          \item You do not need an MS (\autoref{sec:msrequirement}), and CS PhD takes about 5--7 years (\autoref{sec:time}).
-Compared to other countries, CS PhD in the US does not require an MS degree but has a longer PhD study time (\autoref{sec:non-us-differences}).
-          \item What should you call your profs.? (\autoref{sec:address}) Many possibilities (e.g., prof., Dr., and even their firstnames), but probably not Mr. or Mrs.
-          \item Your entire PhD program costs about \$400K in total, but you \emph{do not} pay for it (\autoref{sec:ra-cost})
-          \item Ask your prof. if they can buy computer equipment and such for your research (\autoref{sec:buying-equipment}).
-          %\item Despite some miserable stories on social media, many PhD students have good mentors, supportive lab mates, healthy working environment ... and are happy (\autoref{sec:happy}).
-        \end{itemize}
-
-    \item Appendices
-    \begin{itemize}
-      \item Specific advice for domestic students (\autoref{sec:domestic-students}).
-      \item How to find research opportunities (\autoref{sec:research-opportunities}).
-      \item Cultural differences between US and other countries (\autoref{sec:cultural}).
-    \end{itemize}
-
-\end{enumerate}
-
-\mainmatter
-\chapter{Should You Apply to a CS PhD Program in the US?}\label{sec:should}
-
-\epigraph{\vspace{-0.2in} Don't make fun of graduate students. They just made a terrible life choice.}{\textsc{The Simpsons}}
-
-International students, especially those from less well-known universities and smaller countries, often wonder if they should apply to a CS PhD program in the US.
-Specifically, they are worried about the \textbf{difficulty of getting admitted} and the \textbf{cost of graduate study} in the US.
-This section addresses these concerns.
-
-First, applying to a good US university \emph{should not} be any harder than at schools in other countries.
-In fact, it might even be more flexible since CS PhD in the US \emph{do not} require having an MS or a research topic/proposal/adviser in advance.
-If you believe you have a chance in other countries, e.g., South Korea, Singapore, Germany, UK, Japan, and Australia, then you will surely have a chance in the US as well.
-\autoref{sec:non-us-differences} compares CS PhD study in the US to other countries.
-
-Second, PhD students in CS \emph{do not} need to worry about funding, especially at good R1
-universities in the US. If you are admitted, you will almost certainly
-receive \emph{full funding} (\autoref{sec:funding}) to support your study, including tuition,
-health insurance, and stipend (i.e., you get paid for your study). Moreover, depending on the university,
-you may even receive additional benefits such as summer pay, laptops, (conference/workshop) traveling.
-
-\begin{commentbox}[Vu:]
-  One of the reasons I created this document is that my colleagues at Mason are interested in recruiting Vietnamese students and are surprised when seeing very few applications in Vietnam (see \autoref{sec:ack}).
-Each year our CS PhD program receives 500+ PhD applications, most of which are international but only 5--6 are from Vietnam.
-In general the number of CS PhD applications from Vietnam to US universities is very low and more would be welcome.
-\end{commentbox}
-
-% \section{Timeline}
-
-% \begin{table}[h!]
-% \centering
-% \begin{tabular}{l|l}
-%     \toprule
-% \textbf{Month} & \textbf{Task} \\
-% \midrule
-% January-March & Research programs and prepare for tests \\
-% April-June & Take standardized tests (GRE, TOEFL, etc.) \\
-% July-September & Prepare application materials (SOP, LORs, etc.) \\
-% October-December & Submit applications \\
-% January-March & Interviews and follow-ups \\
-% April & Admission decisions \\
-% \bottomrule
-% \end{tabular}
-% \caption{Application Timeline}
-% \end{table}
-
-\section*{Who is this document for?}
-
-The guideline can be useful for all students, but it is primarily intended for \textbf{international students} from \textbf{smaller countries and less well-known universities} (or domestic students from US universities with no PhD programs or limited research experiences).
-Students from top schools with strong research cultures and experiences might already know most of the information given here.
-My goal is thus to level the playing field by providing information that is not readily available to less privileged students. 
-I hope to encourage more students with such backgrounds to apply to CS PhD programs in the US.
-
-While this document may not fully meet the needs of students aiming for top-10 schools, which typically have more stringent requirements (e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU), it should be valuable for those targeting other reputable R1 institutions.
-It is certainly useful for students aiming for Mason!
-
-%% Vu: *what's a PhD?*  This [[https://matt.might.net/articles/phd-school-in-pictures/][series of pictures]] from [[https://matt.might.net][Matt Might]] illustrates what a PhD means.
-%% \end{commentbox}
-
-\section*{Additional Resources}
-\begin{itemize}
-  \item \href{https://parentheticallyspeaking.org/articles/us-cs-phd-faq/}{Getting a Computer Science PhD in the USA} by Shriram Krishnamurthi
-  \item  \href{https://pg.ucsd.edu/PhD-application-tips.htm}{Blog post on PhD Admission} by Philip Guo
-  \item   \href{https://talkingtorobots.com/FAQ.html}{PhD Admission post} by Yonatan Bisk (customized for CMU admission)
-\end{itemize}
-
-\chapter{How is Your Application Evaluated?}\label{sec:evalapps}
-
-\epigraph{How is education supposed to make me feel smarter?
-Besides, every time I learn something new, it pushes some of the old stuff out of my brain.
-Remember when I took that home wine making course, and I forgot how to drive?}{\textsc{The Simpsons}}
-
-After you submit your PhD application (e.g., in late Fall of year $X$ for the Fall admission of year $X+1$), it will first be screened
-for general requirements, e.g., did you submit your transcripts and standard scores?
-did your reference writers submit their letters?
-Usually, this screening process is done through a central university system, i.e., not by CS faculty.
-If you pass this, your application will be forwarded to the CS department for further evaluation.
-If you don't,  the system will tell you what is missing and what you need to do.
-So pay attention to your email and check your application status regularly.
-
-\begin{commentbox}[Hakan:]
-  At Mason, for full consideration, students should make sure to submit \textbf{ALL} required documents by the application deadline, and should never assume that some required documents (such as official TOEFL scores or official diplomas/transcripts) will be waived by the admissions office.
-If something is listed and not marked as ``optional'', it is mandatory and they should plan for submitting all those.
-\end{commentbox}
-
-Then, applications are reviewed by a \emph{PhD admission committee} (\textbf{adcom}) that consists of faculty members in CS\footnote{In some cases the committee can involve affiliated faculty from different disciplines.}.
-These adcom members have a wide range of expertise and background to ensure diverse perspectives in the evaluation process.
-The size and the review load of the adcom depend on the department size.
-At Mason, the PhD adcom typically has 15--20 faculty, and each committee member is assigned to review about 30 applications.
-Note that most large schools, including Mason, have separate adcoms for MS programs.
-
-Each application is assigned to about 3 adcom members, who will evaluate your profile and reach a consensus.
-While the assigned reviewers are the main ones deciding your application, other faculty in the department will likely have access to your application and can provide inputs and opinions on your profile.
-
-The PhD adcom typically involves assistant professors in the department (see \autoref{sec:faculty-types} for various types of faculty).
-This provides junior faculty the opportunities to recruit students.
-Some members are tenured faculty.
-The chair of the committee will likely be a senior faculty, but they will not review individual applications and instead assign them to committee members.
-The chair will look at various factors such as research interests or mentioning faculty names to assign the applications to appropriate faculty, e.g., I am often assigned to review applicants interested in software engineering.
-
-\begin{commentbox}{Vu:}
-At Mason, we usually decide that a full-time PhD candidate is either (i) admitted with funding (\autoref{sec:funding}) or (ii) rejected. In other words, in most cases, we either
-admit you with full funding or reject your application. In some rare cases, we admit
-without funding because you have funding on your own, e.g.,
-supported by your government or having external fellowships. We justify
-our decision (\autoref{sec:ievaluate}) with a summary of your application, where we list
-strengths, e.g., a well-known school, and weaknesses, e.g., weak
-LORs.
-\end{commentbox}
-
-% \didi{Is there more information on typical strengths and common weaknesses of applications.
-This is especially useful to sophomore and junior students as they still have time to work on those strengths.}
-% tvn: the main thing is research experiences
-
-\paragraph{Why CS depts do not waive the application fee?}  Some universities do waive, e.g., Rice does for all PhD applicants and many universities for domestic students (\autoref{sec:domestic-students}), however most do not.
-In any case, the application fee is typically a requirement of the university.
-Individual departments and programs do not have the flexibility to waive the application fee, even if they want to.
-
-In my opinion, requiring applicants to pay the fee helps ensure their seriousness, as it filters out non-serious candidates.
-Also, most CS programs already receive many applications and would be overwhelmed if the application process were free.
-
-\chapter{Your Application}\label{sec:application}
-
-\epigraph{Son, if you really want something in this life, you have to work for it. Now quiet! They’re about to announce the lottery numbers.}{\textsc{The Simpsons}}
-
-The focus of adcom is to evaluate your background and interest in research to see if you \textbf{fit into our PhD program}!
-The emphasis here is \emph{fitting}, which varies from school to school, faculty to faculty, and even from year to year.
-However, in general, to evaluate your profile, we consider the following key indicators.
-
-\section{Letters of Recommendation (LOR)}\label{sec:lor}
-
-\begin{center}
-  \includegraphics[width=0.6\linewidth]{files/c6.png}
-\end{center}
-
-Most PhD programs will require at least \textbf{two LORs}.
-A good LoR letter should give details of the applicant's research background and achievements, all of which paint a compelling picture of the candidate to the adcom.
-Having a strong letter from an internationally recognized researcher will \emph{greatly strengthen} your application.
-A strong letter from a well-known researcher can outweigh the lack of publications (or low GPA or unknown school).
-
-However, obtaining letters from famous people
-can be challenging for international students, who might not have many interactions with such experts.
-So it is fine to have a letter from people (e.g., profs, researchers, postdocs who mentored you) who know you well enough to talk about \emph{your research experience and capabilities}.
-
-Many students get letters from supervisors from companies where they did internships or are working.
-It is OK as long as it is a research-based personalized letter.
-Again, the emphasis here is \emph{research}, i.e., the letters should describe your research experiences and potential.
-Letters focusing on your coursework or non-research projects at a company won't carry much weight.
-
-\begin{commentbox}[Vu:]
-  If your grading system is not US standard, or you are from a good school unknown outside your country, you can ask your reference writers to explain that in their letters.
-For example, ``Bach Khoa'' are the top universities in Vietnam for STEM studies but few people outside Vietnam know about them.
-So if you are from there, you should ask your reference writers to mention that.
-\end{commentbox}
-
-\paragraph{Self-written letters are bad} Many international students write letters themselves, typically due to the requests of their profs., and have them signed by their profs.
-Such letters have \textbf{very little value} and are considered weak by reviewers---why can you not even find someone who cares or knows enough about you to write a candid personal reference letter?
-Instead of the reference writer talking about you, in this case it is you who write about yourself (and they just sign the letter).
-Experienced reviewers can easily spot (or suspect) such letters, which can be a \red{red flag}.
-
-\paragraph{Generic letters are bad} When the writers do not know much about the applicants (e.g., just taking some course with them or not making any impression to write about), they might write a generic and short letter, which is not useful and also considered weak.
-This does not mean the professor is not good or does not care about you, but they just do not know you well enough to write a strong letter.
-So it might be a good idea to directly ask if the prof.
-is willing to write a \emph{strong} letter for you.
-If not, then you should ask someone else.
-For example, if a student I don't know well asks me to write a letter for them, I will explicitly tell them I don't know them that well to write much about them, and such a short, generic, and weak letter will not help their case.
-
-\begin{commentbox}[Hung:]
-  A sad reality is that most professors in Vietnam \textbf{DO NOT} know how to write a good letter, or are lazy in writing letters hence delegate the writing to the students.
-Unfortunately, there is no easy solution to this problem.
-\end{commentbox}
-
-\begin{commentbox}[Vu:]
-
-  Several international students mentioned that some professors are unwilling to write letters or write weak ones because they do not want (good) students to go abroad or only go to places where they want the students to go to.
-If you are in this situation, you should find someone else to write for you.
-  \tcblower
-
-Sometimes students would go to great lengths just to get letters from well-known professors in their school (e.g., department head).
-But as mentioned, if these professors do not know you, their letters would likely be generic and carry little value (sometimes \red{red flags}).
-Moreover, a top professor at your university might not be well-known to US faculty (see more details in \autoref{sec:your-school}).
-So save the trouble and get letters from \emph{any} professors/supervisors who know you well and can write a good letter about \emph{your} research ability.
-It's better to have a good personalized letter about your own research ability from someone who is less well-known than a generic/weak letter from a well-known person.
-
-\end{commentbox}
-
-\paragraph{Waiving Your Right}  Choosing not to look at a reference letter is pretty standard in school and job applications.
-When you waive your right to see the letter, it adds a layer of trust, showing you're confident in your choice of referees and that you're not trying to twist their words.
-It's also about keeping things open and honest between you and your letter writers and encourages them to be real about your strengths and qualifications.
-Plus, it keeps things private.
-
-If you do not waive your right,  the letter writer might refuse to write for you or write a generic letter that does not help your case.
-Reviewers also might raise concerns about a letter that is not waived, e.g., if you do not trust your letter writers, then you should find someone else to write for you.
-In short, it's a standard practice and a way of keeping things straightforward and respectful in the whole recommendation game.
-
-\begin{commentbox}[Didier]
-  \emph{Should letter writers have PhDs?}  In Rwanda, a lot of students interact more with teaching faculty who might not have PhD.
-  \tcblower
-  \textbf{Vu}: This is an interesting and useful detail that US faculty might not be aware of.
-Students should mention this in their statements.
-In general, someone with a PhD has been through the research process and therefore can better evaluate your research ability.
-But if you do not have such people, then someone who can properly evaluate your research ability is still OK (and again, explain that in your SOP).
-\end{commentbox}
-
-\paragraph{Help your LOR Writer} As mentioned above, do not write your own letter and generic letters do not give much value.
-One thing you can do to help your writer to write a strong, customized LoR for you is to provide details or unique things about yourself.
-For example, let them know about your GPA, research and work experiences, papers (if any), or anything you want them to mention.
-You can also provide them with a draft of your SOP so that they can see what you are saying about yourself and complement that with their own perspective.
-Sometimes your writer will explicitly ask you for such information, but if not, you should provide it anyway (especially if you have not interacted with them much or have not done much research with them).
-
-\paragraph{Remind Your Writers} After you submit your application, you should tell your writers about that and let them know they will soon receive an email from the university to submit their letters.
-You should also remind them when you submit your application and ask them to submit their letters on time if they haven't done so.
-Note that most places only have deadlines for the applicant, but are very flexible with the letter writers (in many cases do not even give them any deadline).
-Also, many places do not begin the admission review process right after the deadline and work on application reviews in the next semester (mid-January).
-
-Thus you do want to send reminders because professors can be quite busy and might forget to submit their letters, especially when there is no explicit deadline.
-However, do not send too many reminders as that can be annoying to the writers.
-
-\paragraph{VEF 2.0} For Vietnamese students, it's worth mentioning the \href{https://vef2.org/}{VEF2.0 program}, which has helped many students gain admission to top CS PhD programs in the US.
-VEF2.0 invites US faculty members from leading institutions to conduct rigorous interviews with VEF students and subsequently provide reference letters on their behalf.
-Despite the limited interaction between the interviewer and the interviewee (primarily just the interview), these reference letters are generally effective and have helped many students get into good PhD programs.
-
-\section{Statement of Purpose (SOP)}\label{sec:research-statement}
-
-\begin{center}
-  \includegraphics[scale=0.5]{files/c2.png}
-\end{center}
-
-While you might not have control over LORs (\autoref{sec:lor}) or where your go to school (\autoref{sec:your-school}), you do over your
-statement of purpose (SOP) or personal statement\footnote{Few schools separate these documents and ask you to write both: SOP, which focuses on research experiences, and personal statement, which is everything more personal, e.g., why PhD, challenges, diversity, etc}!
-So write it well because it could make a big difference.
-
-In your statement, you have the opportunity to connect directly with adcom members and make your application stand out and unique.
-Your SOP can convince adcom members that you would \emph{fit} the program, even if you don't have very strong research experience.
-A well-written SOP also shows that you can communicate, which is very important in research, and that you can effectively teach and communicate with students, which is important for TA funding (see \autoref{sec:funding}).
-% is important because if you need (GTA) funding, it will provide evidence
-% that you can teach and communicate with students.
-
-In your SOP, focus on research motivation and convince us that you can achieve it through your experience, e.g., published papers.
-You should also back up your claims with concrete evidence.
-For example, if you say you have experience with teaching, then show what you did, e.g., undergrad TA or mentoring someone.
-If you say you work on a project, then show some results, e.g., paper submitted (or even rejected), achieved certain performance improvement over the state of the art.
-Also, do not use more than two pages (references can be on the third page).
-Many SOP samples are \href{https://cs-sop.org/}{available here}.
-
-You should talk about things that adcom members might not know about and can help make you \textbf{stand out} in the application pool of thousands of applicants, e.g., you are member of a minority or LGBTQ+ group (\autoref{sec:urm}), your personal Github project with hundreds of stars or your regular contributions to well-known open-source projects (see \autoref{sec:improve-your-chance} for increasing your admission chance).
-
-This is a simple task often overlooked by many applicants: \textbf{tailor your SOP} to the institution you're applying for,
-e.g., why do you apply \emph{here}?
-Provide names of professors who you're interested in (if they are not already in the adcom, your application might get forwarded to them for evaluation; and they might be interested in interviewing and recruiting you).
-This shows that you're serious and have done homework on places you're applying to.
-Adcom will look for this part (\autoref{sec:why-rejected}).
-
-Finally, you should get your SOP reviewed by your LoR writers, professors (especially those who have served in adcom), or even postdocs or PhD students (because they have been through this process).
-
-\begin{commentbox}[Vu:]
-  I often read SOP and LORs first. If I am
-  persuaded by then, I would skim over other factors and advocate for
-  admission (unless I see red flags in other parts). However, if I am not
-  convinced, then I will likely recommend rejection (unless I see
-  something stand out in other parts).
-  \tcblower
-  Do careful research on professors, don't mention \emph{emeritus} or adjunct faculty (see more about various types of faculty in \autoref{sec:faculty-types}).
-  Also, be careful not to send statements to the wrong schools or mix
-  facts (e.g., talking about school X but mentioning working with
-  profs. at school Y; and do not talk about George Washington when applying to George Mason). I have seen such statements more times than I should.
-\end{commentbox}
-
-% \subsection{Dos and Don'ts of SOP}
-% \paragraph{Do's}
-% \begin{enumerate}
-%     \item \textbf{Be Specific}: Clearly articulate your research interests and why you are drawn to the program.
-%     \item \textbf{Show Enthusiasm}: Demonstrate genuine passion for your field of study.
-%     \item \textbf{Proofread}: Ensure your statement is free of grammatical and typographical errors.
-%     \item \textbf{Be Personal}: Include relevant personal stories that highlight your journey and motivations.
-%     \item \textbf{Tailor the Statement}: Customize your statement for each program you apply to.
-% \end{enumerate}
-
-% \paragraph{Don'ts}
-% \begin{enumerate}
-%     \item \textbf{Avoid Generic Statements}: Don’t use vague or general phrases.
-%     \item \textbf{Don't Repeat Your Resume}: Provide new insights or context beyond what’s listed in your CV.
-%     \item \textbf{Steer Clear of Jargon}: Avoid overusing technical jargon that might confuse readers.
-%     \item \textbf{Don’t Exaggerate}: Avoid excessive self-praise without evidence.
-%     \item \textbf{Avoid Negative Tones}: Don’t criticize past institutions or colleagues.
-% \end{enumerate}
-%\subsection{Common Pitfalls to Avoid}\label{sec:sop-pitfalls}
-
-%SOP is crucial for making a strong impression on adcom reviewers. This outlines common pitfalls to avoid in your SOP.
-
-%\begin{enumerate}
-%m \textbf{Generic Statements}: Using vague or general statements that could apply to any applicant.
-For example, saying blank statements such as “I am passionate about computer science” without providing specific examples or evidence.
-%    \item \textbf{Not Customized}: Failing to mention why you chose this univ and faculty members you want to work with.
-If you can submit the same SOP to multiple programs with few changes, it is not customized enough.
-%    \item \textbf{Too Long}: Adcom members have to read many applications and a long SOP can be a turn-off. Keep it under 2 pages.
-%    \item \textbf{Lack of Passion or Motivation}: Failing to convey genuine enthusiasm for research and the specific program.
-%    \item \textbf{Grammatical and Typographical Errors}: Submitting an SOP filled with grammatical mistakes, typos, and poorly constructed sentences.
-%    \item \textbf{Excessive Self-Praise}: Overly boasting about achievements without providing concrete evidence or examples to back them up.
-%    \item \textbf{Irrelevant Information}: Including personal anecdotes or experiences that do not relate to the field of study or research interests.
-%    \item \textbf{Unrealistic Goals}: Stating research goals that are not feasible or that show a lack of understanding of the field’s current state.
-%    \item \textbf{Negative Tone}: Criticizing previous institutions, professors, or colleagues.
-%    \item \textbf{Failure to Address Fit}: Not explaining why you are a good fit for the program and how the program is a good fit for your goals.
-%    \item \textbf{Generic Opening and Closing Statements}: Starting or ending the SOP with clichés or unoriginal statements that do not add value.
-%    \item \textbf{Lack of Customization}: Submitting the same SOP to multiple programs without tailoring it to each specific program's strengths and faculty.
-%\end{enumerate}
-
-\subsection*{Additional Resources}
-\begin{itemize}
-    \item \href{https://chrisblattman.com/blog/2022/01/11/}{Writing your statement of purpose} by Chris Blattman
-  \item \href{http://www.pl-enthusiast.net/2022/10/03/how-to-write-a-grad-school-personal-statement/}{How to Write a Grad School Personal Statement} by Mike Hicks
-  \item     \href{https://cs-sop.notion.site/cs-sop/CS-PhD-Statements-of-Purpose-df39955313834889b7ac5411c37b958d?p=f5d5980a71524ebaa4e6ae57266b847c&pm=s}{CS PhD SOP database} by cs-sop.org
-\end{itemize}
-% \subsection{Outline of a LOR}\label{sec:lor-outline}
-
-\section{Research Experiences}\label{sec:research-experiences}
-
-This section discusses various research experiences that can strengthen your application.
-\autoref{sec:research-opportunities} provides more information on how to find research opportunities (e.g., during your undergrad study).
-
-\paragraph{Publications} Concrete evidence of research ability is having \emph{papers in reputable international journals or conferences}.
-Having published papers, especially at top venues, is a sign that the applicant has been successfully involved in research.
-Applicants admitted to top schools often have multiple first-authored papers at top places.
-
-\begin{commentbox}[Vu:]
-  Many international students mention Scopus Q1, which consists of various journals from IEEE, Elsevier, and many other publishers.
-I don't know/recognize many of the journals listed in Scopus Q1.
-This might be something to be mindful of, as \textbf{CS} faculty might not be too familiar with Scopus or journals listed there, so devote some part in your statement to discuss the significance of your papers.
-\end{commentbox}
-
-However, it is understandable that many students do not have the opportunity to publish papers.
-Thus, general writings, even those under submissions or rejected, would still be good (much better than nothing).
-Be sure to upload your papers with your application and talk about them in your statement (see \autoref{sec:research-statement}).
-Note that local conferences and non-English journals or conferences do
-not carry as much weight since their quality is often unknown to US faculty.
-However, if you have published in such places, you should still upload them, mention them in your statement, and explain why they are good.
-
-% In CS, conferences are important and top conferences are prestigious.
-You can find top CS conferences at places such as \href{https://csrankings.org}{CSRankings}, which ranks CS programs based on how their faculty publish at top \emph{conferences} (see more in \autoref{sec:ranking}).
-
-\begin{commentbox}[Craig:]
-  Mason and many other universities allow you to upload your published papers and other writing samples.
-In many cases, even if the papers were not published at top places, we can still determine their quality by simply skimming over the paper.
-\end{commentbox}
-
-\paragraph{Work Experiences} Work experiences at \emph{well-known research laboratories}, such as Microsoft Research, can strengthen your
-application.
-Unfortunately, many good research places in your countries, e.g., VinAI in Vietnam, remain relatively unknown to most universities in the US.
-So you or your LoR writers should explicitly say something about them in your statement.
-
-\begin{commentbox}[Hung:]
-  The reputation of VinAI has been increasing steadily over the past few years; many of my colleagues heard about VinAI.
-\end{commentbox}
-
-\paragraph{Competitions} Winning \emph{internationally recognized competitions} can also demonstrate your research potential.
-For example, participating in Math Olympiads if you want to do theory or winning ACM programming contests if you want to ``build'' systems, e.g., software analysis.
-
-\begin{commentbox}[Thanh:]
-  Due to academic culture, professors in Vietnam usually aim for (international) journals instead of conferences.
-Could you give some tips on how to know whether a journal is good (CSrankings, unfortunately, only consider conferences)?
-  \tcblower
-  Vu: One way is looking at what well-known researchers publish.
-For example, if you are interested in a field X, you can use CSRankings to look at active faculty in X, and then look at their websites to see what journals they publish at.
-\end{commentbox}
-
-\section{Your School and Grades}\label{sec:your-school}
-
-\paragraph{School} Graduating from top universities \emph{that we recognize} helps.
-For example, if you are an international student and your school is well-known, then it is \emph{``top foreign''}, which is a plus.
-However, if we do not know much about schools in your country, then we are uncertain about the quality of your school and likely treat your school as \emph{``unknown foreign''}, which can be a minus point.
-
-If your school is the top in your country but you think it might not be well-known internationally, you should talk about it in your SOP (and also ask your LOR writer to do that).
-Of course, if you're interested in working with Vietnamese, consider  \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{CS programs in the US that have Vietnamese professors}.
-%It might also be helpful to have your CS dept to put itself on CSRankings so that others know about the dept, its faculty, and research.
-
-\begin{commentbox}[Vu:]
-  Sometimes PhD adcom in the US will share a document such as \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Foreign-Top-Schools}{this one}, which lists the top schools in several countries.
-We also ask other faculty and students if we think they know about the place.
-For example, when I was a postdoc at UMD, members of their CS PhD adcom asked me to evaluate applicants from Vietnam.
-During my time at UNL and now here at Mason, I have looked at Vietnamese applications (whether they are assigned to me or not) and provided input to their reviewers, e.g., X is the top tech school in Vietnam and so it should be \emph{top} instead of \emph{unknown} foreign, which makes a huge difference.
-\end{commentbox}
-\begin{commentbox}[Deepak:]
-  If an applicant is anxious about their school not being known outside their country, they can provide information about their school and department, with independent sources where such information can be verified.
-\end{commentbox}
-
-\paragraph{Grades} Having a good GPA is good. However if your school is not well-known, having top grades or rankings
-usually will not help because we cannot evaluate them.
-This can be an issue for students in many top international universities where the competition is so high that very good students can still have low rankings from these schools (and be overlooked by adcom).
-
-As with school reputation, you and your LoR writers can mention the grading system of your university if you think that is helpful for adcom to evaluate.
-Note that while having good grades might not help much,
-having bad grades (e.g., < 3.0) will be \red{red flag} (unless your LORs or
-statement give a proper explanation). This is especially true if you
-have bad grades in relevant courses, e.g., Math and CS.
-
-\begin{commentbox}[Thanh:]
-  Vietnamese universities typically offer specialized programs, such as the talented engineer program at HUST, that have highly competitive entrance exams and a limited number of available slots (e.g., 30 per year).
-However, these programs often set higher requirements for students, including more demanding tests and assignments, resulting in lower GPAs and overall rankings.
-For example, 3.5 GPA students from such talented programs are typically much better than 4.0 GPA students not in those programs.
-Similarly, variations in GPA standards exist among different universities, with technical universities generally having lower GPAs than economic universities.
-These make gaining admission in the US difficult as US faculty are not familiar with these issues.
-  \tcblower
-  \textbf{Vu}: Vietnamese students and even faculty often lament how this grading system hurts Vietnamese students applying abroad.
-One way to mitigate this is to make these issues known in your SOP.
-\href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{Universities with Vietnamese profs} are probably aware of them, but in general your letter writers \emph{and} you can explicitly mention these in their letters and your statement.
-\end{commentbox}
-
-\section{Standard Tests (GRE, IELTS, etc), Work Experiences, Online Courses, and Certificates}\label{sec:standard-tests}
-
-\paragraph{GREs Do Not Matter} While a few schools still require taking the GRE (e.g., UCF), most good CS PhD programs in the US \textbf{no longer} require it.
-Thus, if you haven't taken the GRE or have bad scores, then don't waste time (re)taking it.
-It does not help much!
-
-However, if you took it and have really good scores then it might be worth it to include and talk about them in your application (but don't expect them to make much difference).
-But if your scores are bad, then you should not include them in your application, which can be a \red{red flag}.
-
-\paragraph{English Tests (IELTS, TOEFL)} Unless your degrees are from specific countries such as \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-Mason#standard-tests-waiver-eligible-countries}{these}, you will need to take standardized English tests.
-
-You should just do well enough to pass the minimum requirement set by the university. %, which is required for TA (\autoref{sec:funding}). 
-Just as with grades (\autoref{sec:your-school}) and GRE, having high scores in these tests might not help, but having too low scores can be a \red{red flag} and sometimes results in an automatic rejection (\autoref{sec:why-rejected}), e.g., below the minimum requirement.
-
-\begin{commentbox}[Vu:]
-  The minimum for Mason (being above this might not mean much, but below is a \red{red flag}).
-  \begin{itemize}
-    \item GPA: $\ge 3.0$ in your undergrad (but we also consider the rank/prestige of your school)
-    \item GRE: not required
-    % \item but if you want to use it, then we expect a total (V+Q) of $\ge 311$ (with a $\ge 157$ Q) and A $\ge 3.0-3.5$.
-    \item English requirement tests (one of the below)
-          \begin{itemize}
-            \item TOEFL: 88 pts in total AND $\ge 20$ points in each subsection OR
-            \item IELTS: $\ge 6.5$ OR
-            \item DuoLingo Graduate English: $\ge 120$ OR
-            \item Pearson Test of Academic English: $\ge 67$
-          \end{itemize}
-  \end{itemize}
-\end{commentbox}
-
-\paragraph{Work Experiences} Work experiences, especially if they are not research-based, \emph{do not} carry much weight.
-Similarly, LoR from your supervisor for non-research experience will not count much.
-
-Do not spend much time talking about non-research-related job in your SOP.
-However, if you have research-based work experiences, then they can be useful and should be mentioned in your SOP and LORs.
-For example, if you have worked at a research lab, e.g., Microsoft, DoE, VinAI in Vietnam, etc (\autoref{sec:research-experiences}), then you should mention that in your SOP and ask your supervisor to write about your research experiences and potential in their LOR (\autoref{sec:lor}).
-
-\paragraph{Online Courses and Certificates}
-In general, these do not carry much weight as they do not show research ability. We do not care much if you have taken an online Coursera AI course or have a professional certificate from Microsoft.
-However, as mentioned in \autoref{sec:non-stem}, if you do not have a CS background, you might be able to use these to show you have sufficient CS knowledge.
-
-\section{CV/Resume}
-The CV/Resume provides a summary of the applicant's achievements. Most schools require you to upload your CV with your application.
-
-Prepare your CV in such a way that it allows the reviewers to quickly scan to identify major achievements, e.g., Publications, Programming Competition Awards, and Teaching Experience.
-
-\section{Interview and the Waiting Game}\label{sec:interview}
-
-After you submit your applications, the waiting game begins! For many students, this is a very stressful time. This section provides some information and tips to help you get through this time.
-
-\subsection{Interviews}
-
-Just like a job application, you likely will get interviews from schools.
-The most common case is that a prof.
-is interested in working with you and wants to chat to make a decision, e.g., to offer RA (\autoref{sec:ra}).
-In some cases, the interview is done by several professors, e.g., to see if a student fits in their group or to recruit a very strong student to their program.
-The interview can also be done by adcom members to see if the student fits their program.
-%The professor might include one of their current students as part of the interview.
-
-Typically, an interview takes about 15--30 minutes, and one important aspect of evaluation is your ability to effectively communicate, including speaking and understanding English. 
-You might be asked to talk about your research experiences and interests and to read a paper and discuss it. In some rare cases you might also be asked to solve a problem, e.g., a coding problem. 
-The interview is also a good opportunity for you to ask questions about the program and the university.
-
-You should treat the interview as an informal chat.
-Have ``elevator pitches'' about your research experiences and interests.
-You might also want to have a 5-minute presentation about your research.
-If a prof.
-asks you to read a paper, do it and be prepared to discuss it.
-You can also ask if you need to prepare for a coding problem.
-Finally, prepare some thoughtful questions to ask, e.g., about the program, the university, or the professor's research.
-
-\begin{commentbox}[Vu:]
-  At Mason, faculty are encouraged to interview candidates.
-For very strong candidates, the interview is actually to recruit them.
-In some cases a faculty interviews a candidate that they see potential and want to advocate for their admission.
-Without the interview, such applications may be more likely to be rejected.
-  \tcblower
-  In short, getting an interview is a good sign; it means that someone is considering you. If we are not interested in your application, we will not proceed with an interview.
-\end{commentbox}
-
-\paragraph{Follow-Up Emails} If you had an interview and have not heard back, you can email to ask about the status of your application.
-See \autoref{sec:accept-postpone-decline} for how to check status and follow-up emails.
-
-\subsection{Timeline}
-
-The timeline for interviews and notification letters varies.
-Faculty set up interviews based on their (busy) schedule.
-Do not be surprised if you get an interview invitation at the last minute.
-Some schools do not do interviews at all.
-
-Some schools send out admission letters in batches, some do rolling admission, and some do not send anything out (e.g., you're on their waitlist).
-In general, you should hear back from most schools by mid-March (but some schools might take longer!).
-Note that rejection letters might come very late (see \autoref{sec:late-rejection}).
-
-In short, there is not much you can do about this other than to be patient and wait.
-Also do not send emails asking about interviews or status (unless you have interviewed specifically with someone then you can ask that person for status updates and others \autoref{sec:accept-postpone-decline}).
-
-\paragraph{Updating your profile} You should not send emails to update your profile.
-However, if you have new publications or other big achievements, you can ask them to update your application (though no guarantee that they will consider them).
-
-\paragraph{Not getting interviews} While it is generally good to get an interview, not getting one \emph{does not} mean you're out.
-Many programs do not have the tradition of interviewing applicants.
-For example, at Mason, most admitted students with TA (\autoref{sec:ta}) do not go through interviews.
-
-However, no interviews typically mean that you will not likely get an RA (\autoref{sec:ra}), which is offered by an individual faculty (if they want you to do research for them, then they likely will interview you first).
-In the case of no interviews, your application (and TA/fellowship funding) is decided by the adcom.
-
-% \section{Preparing and Tracking Applications}
-
-% \paragraph{Shared Spreadsheet} Many students use a (shared) spreadsheet, e.g., Google Sheets/Docs, to help them and their letter writers to keep track of their applications.
-Here are some information to put on the sheet.
-% \begin{itemize}
-%   \item \textbf{Your info}: Full name, email, phone, link to website/CV.  This is helpful for the writers just in case they want to quickly get some information about you.
-
-%   \item \textbf{Applications details}: University, Dept, Application System URL, Submission Deadlines, Application Status (e.g., submitted,  rejected, wait-listed, accepted)
-
-%   \item \textbf{LoRs}: Writer 1, Writer 2, Writer 3.
-If this is a shared document, you might want to omit names and just use writer X.
-Under each is their status, e.g., sent/not yet/reminder needed, etc.
-
-% \end{itemize}
-
-% You can then share this sheet with your reference writers and remind them to submit LoRs on your behalf (see \autoref{sec:lor}).
-Also, you should periodically update your writers with your status; a simple note such as \emph{``I am getting interviewed by prof.
-X at school Y.
-Do you have any advice?''} can get you a lot of help.
-
-% \paragraph{Communicating with LoR writers and People Who Support You}
-
-% stressfull time, rely on support from others
-% ask ppl to review your SOPs
-% provide information for LORs, ..
-%
-
-\chapter{Getting Admitted}\label{sec:accepted}
-
-\epigraph{``Oh...
-and how is education supposed to make me feel smarter?
-Besides, every time I learn something new, it pushes some old stuff out of my brain.
-Remember when I took that home wine-making course and I forgot how to drive?''}{\textsc{The Simpsons}}
-
-By around March you should hear back from most PhD programs that want to admit you.  In many cases you will not hear back from schools that do not want to admit you (see \autoref{sec:late-rejection}).
-If you receive offers, congratulations!
-Now you're at a different game because the schools that have admitted you will try to get you to accept them!
-Important factors to consider include the reputation of schools and professors (\autoref{sec:schoolsandprofs}), and funding availability (\autoref{sec:funding}).
-You will likely have to make your decision (\autoref{sec:accept-postpone-decline}) by around \emph{April 15}, which is the deadline for most schools
-(\href{https://cgsnet.org/wp-content/uploads/2024/01/CGS_April15_Resolution_Jan312024.pdf}{why this date?}).
-
-\paragraph{Open House} Most schools have \emph{Open House} or \emph{Visit Day} events, which are a great resource to learn about the school, department, faculty, research, living, etc.
-
-Even if you can't come in person, you should attend virtually and meet with individual faculty.
-During the event, you get a chance to learn more about the program, and talk to individual faculty and current students.
-Take notes of faculty who make you excited, and count those taking in new students (if they meet you, likely they are considering new students!).
-Talk to students about their advisers, the dept, the area, and the funding situation.
-Ask about anything you want to determine that they deserve \emph{you}.
-
-\begin{commentbox}[Vu:]
-  Mason has \emph{Virtual} Open House (VOH), e.g., \url{https://cs-Mason.github.io/cs-phd-voh-s23/}.
-We invite all admitted PhD students to the VOH through Zoom to learn about the CS program, the department, Mason, and the DC area in general.
-Students also get opportunities to chat with professors and current students.
-\end{commentbox}
-
-\paragraph{What else?} Make a decision, accept, reject, or defer the offers  (\autoref{sec:accept-postpone-decline}).
-Ask to meet with potential advisers (e.g., through Open House or separately) and even their students.
-Ask about computer equipment and software, office space, and other resources; in many cases these will be provided for free by your adviser or department (\autoref{sec:buying-equipment}).
-
-Also, do not forget to update and thank your LoR writers and others who have supported you through this process.
-
-\section{If you do not get admitted}\label{sec:not-accepted}
-
-If you do not get admitted to any schools or don't want to go to the ones that admit you, try again next time.
-Graduate admission can involve randomness and noise.
-In the meantime, you can work on improving your profile, e.g., get more research experiences, publish more papers, improve your connections for stronger LoRs, etc.
-See rejection reasons (\autoref{sec:why-rejected}) and timeline (\autoref{sec:late-rejection}), and additional advice to improve your chance next time (\autoref{sec:improve-your-chance}).
-
-You can consider applying to MS programs, which are typically easier to get in; but you likely need to pay.  Sometimes, the university that rejected you for PhD might accept you for MS.
-If you get into an MS program at a school of your choice, you can contact professors to work with them. If you do well, you can ask the professor to support you to convert to PhD.
-
-\paragraph{Should I ask for feedback?}
-No, don't bother.
-You will likely not get any useful feedback.
-We are not willing to reveal your evaluation results or give you advice on how to improve your profile.
-So it is best to just move on.
-If you really want feedback, ask your professors, collaborators, ref writers, or those who have previously applied.
-
-\chapter{Funding}\label{sec:funding}
-
-\epigraph{Bart, with \$10,000, we’d be millionaires! We could buy all kinds of useful things like…love!’}{\textsc{The Simpsons}}
-If you're admitted to a \emph{good} CS PhD program, you should not have to worry about funding!
-In the US, the common types of funding for PhDs are \emph{graduate teaching assistant} (GTA or TA), \emph{graduate research assistant} (GRA or RA), and \emph{Fellowship}.
-RA is paid by a prof.
-for you to do their research.
-TA is paid by the dept.
-for you to help with teaching.
-Finally, fellowship is independent funding that can come from a school, a company, or an organization.
-\autoref{tab:funding} summarizes the differences.
-Note that funding is typically more available for PhD students than
-MS.
-
-\begin{table}
-  \centering
-  \small
-  \caption{Different types of PhD funding. All cover tuition, insurance, and stipend.}\label{tab:funding}
-  \begin{tabular}{c|c|c|c}
-    \toprule
-    &\textbf{TA}&\textbf{RA}&\textbf{Fellowship}\\
-    \midrule
-    \textbf{From} & School & Profs. & School/External\\
-    \textbf{For}                  & Teaching Assist.       & Research                        & Research                              \\
-    \textbf{Cover All?} & Yes                      & Yes                             & Yes                                   \\
-    \textbf{Summer?}              & No                       & Maybe                           & Likely                                   \\
-    \midrule
-    \textbf{Pros}                 & Research Freedom         & Get to do research              & Research Freedom                      \\
-    \textbf{Cons}                 & Teaching Duties           & Research Restriction & Competitive, Limited             \\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
-\section{Graduate Assistantship (TA/RA)}\label{sec:ta-ra}
-The most common type of funding is \textbf{graduate assistantship}, which is either TA or RA.
-Both TA and RA come with tuition waiving (you don't have to pay tuition), health insurance (this takes care of your insurance, which is a must-have in the US), and most importantly, your stipend (i.e., your salary).
-Some universities also give significant discounts or pay insurance for spouses/children.
-
-Several things about stipends.
-First, the amount of stipend \emph{varies} and depends on factors such as location (e.g., a stipend in the DC area is likely higher than in Lincoln, Nebraska due to higher living costs).
-Second, an academic year (AY)  year is typically \emph{9-month} in the US, so the stipend is for 9 months.
-Third, you might get paid over the summer (\autoref{sec:summer-funding}) through funding from your professor or fellowship (typically no TA over the summer).
-Fourth, like most sources of income in the US, you will have to \emph{pay tax} on your stipend.
-Finally, private universities might pay more for stipends, but they could have extra activity fees or some other hidden ones (e.g., you may need to pay some fees for each credit hour).
-
-\paragraph{Low Stipend?} Students often complain their stipend is too low.
-However, in most cases it is not bad, and you can live comfortably with it.
-It might also be enough to support your spouse and kids (many CS PhD students have their families with them).
-So don't worry too much about the stipend.
-If you're admitted to a good CS PhD program, you will be fine.
-A good school would know that it has to be competitive to attract students.
-For example, at Mason, every year we discuss improving the benefits, and especially stipend, for our graduate students.
-
-For a full breakdown of how much a graduate student costs, see \autoref{sec:ra-cost}.
-
-\begin{commentbox}[Vu:]
-  TA and RA at Mason have similar benefits in tuition waiving and insurance.
-The college and department will set the rate for a 9-month graduate assistant stipend.
-TA, which is paid by the department, will likely be that amount but RA might be higher depending on the stage of the student (1st year vs ABD\footnote{All but dissertation: close to graduate.}) and the prof.
-  \tcblower
-  Having health insurance is required at many US universities.
-Do not assume that you're young and healthy and ignore insurance.
-At Mason, and most good CS PhD programs, your GTA or GRA \emph{will always} come with full insurance.
-In fact, at Mason your spouse/children will get a significant discount rate for health insurance.
-So you will never have to worry much about health issues for you or your family here.
-\end{commentbox}
-
-\subsection{Teaching Assistant (TA)}\label{sec:ta}
-
-TA is common in the beginning when you haven't found an adviser who would pay you RA.
-It is also common to sandwich between TA and RA (e.g., when your prof.
-does not have sufficient funding or you want to try the TA experience).
-
-Your TA is paid through the school or department, i.e., they hire you to help teach.
-As a TA, you spend up to 20 hrs/week and help with classes (e.g., grading or teaching labs/recitation).
-During a semester, a TA might work with several courses and professors (not necessarily their adviser).
-TA funding is \emph{not} typically available during the summer (\autoref{sec:summer-funding}), which has few courses.
-
-\paragraph{How to get TA?}  Unless you have other funding such as RA or fellowships, TA is typically the default for CS PhD programs.
-In your application, indicate that you need financial assistance.
-Typically, adcom will either admit you and give you TA, or reject you.
-We do not admit a student without supporting them.
-
-\begin{commentbox}[Vu:]
-  At Mason CS, admitted PhD students have 4 years of GTA guaranteed, and also receive a stipend for the first summer (\autoref{sec:summer-funding}).
-\end{commentbox}
-
-Even if you have other funding and do not need a TA, you still should do TA at least once.
-This allows you to see what teaching is like, which is especially helpful for a research career where you often give talks and tell people about your work.
-Mason sometimes has classes that a more senior student can teach.
-In that case, you will be paid as GTA or even sometimes as a lecturer.
-This is a good opportunity for students to get teaching experience and also get paid more (as a lecturer).
-
-\subsection{Research Assistant (RA)}\label{sec:ra}
-RA is provided through a professor through their funding so you can work on their project.
-You do not need to teach as an RA, so you can focus on your research. Depending on the professor, RA may be available during the summer. \autoref{sec:ra-cost} gives more details on RA budget.
-
-\textbf{How to get RA?} When a professor recruits you, they might offer you an RA right away (so you start with an RA).
-However, a more common scenario is that you first get admitted with TA, and then after a year or two find an adviser to support you with RA.
-
-It is important to note that RA is \emph{never guaranteed} as it depends on the funding situation of the prof.
-So you should also pay attention to TA, which is a good backup plan (remember, typically TA and RA have quite similar benefits).
-This means you should also check if TA is readily available for PhD students in the program.
-
-\begin{commentbox}[Vu:]
-  If you got recruited and offered an RA by a prof., you will likely get admitted.
-For example, if a prof., even if not in PhD adcom, wants to fund you, adcom will likely respect that decision and admit you.
-\end{commentbox}
-
-\section{Fellowships/Scholarships}\label{sec:fellowships}
-
-Fellowship is another type of funding that students can get (e.g., from the university, industries, or government).
-Fellowships are typically competitive and generous, giving pretty much all benefits tuition/insurance that a TA/RA has.
-They might even give higher stipends (including summer) and open doors for job opportunities such as internships.
-
-\textbf{How to get Fellowship?}   Many schools provide fellowships to attract students.
-You likely will not need to do anything and adcom will recommend such fellowships to strong students.
-Some schools automatically offer a fellowship to all accepted students, while others only award it to a limited number of admitted students, such as the top percentile.
-
-For external fellowships including those from the US government (e.g., NSF, DOD) and tech companies (e.g., Google, Microsoft, Facebook), you will need to apply.
-% However, tech companies including Google, Microsoft, Facebook have fellowships that international students can apply for.
-Prestigious external fellowships typically require a clear and good research plan (the GRFP also requires outreach activities and plans).
-So it might be a good idea to wait until your second year to have research experience and even publication before applying.
-Remember, you're competing with the top PhD students at top universities worldwide.
-
-\begin{commentbox}[Vu:]
-  PhD applicants at Mason are automatically eligible for a \emph{Presidential Fellowship}.
-It is at least as good as GTA but the most important thing is that as a fellowship it is truly free money (i.e., you do not depend on any prof.
-or TA).
-Adcom members nominate applicants for this fellowship and the whole committee will decide.
-\end{commentbox}
-
-In addition to general fellowships that are available to all students, there are also specific fellowships, e.g., for US citizens and permanent residents, for underrepresented minority (URM) students (\autoref{sec:urm}), and students from specific countries.
-There are many fellowships available, so you should look for them and apply.
-
-In general, fellowships, especially those that are open to everyone, are highly competitive and prestigious, and you will stand out if you get one.
-Every PhD student has pubs, but only a few would have NSF Graduate (GRFP\footnote{\url{https://www.alexhunterlang.com/nsf-fellowship} is a good starting place for the GRFP with lots of proposal examples.}) or Microsoft fellowships.
-In fact, these are so prestigious that even if you didn't get it but make it to the final round or even \emph{``honorable mentioning''}, your school will still mention you on their website, and you should put it on your CV.
-
-\section{Funding In the Summer}\label{sec:summer-funding}
-
-One confusion students often have is whether they will get paid over the summer.  The answer is that it depends on the funding source.
-First, remember that an academic year (AY), which consists of Fall and Spring semesters, is typically 9 months, so your stipend is for 9 months (and many places allow you to spread it over 12 months).
-
-If your funding source is TA (\autoref{sec:ta}), you typically do not get paid over the summer as there are few courses.
-Some CS departments offer some summer funding, but it is typically not guaranteed and might not be a lot.
-For example, at Mason, we offer summer funding for 1st-year PhD TA students.
-The amount over the 3-month summer is similar to their monthly stipend (i.e., their 9-month stipend divided by 3 for the 3 summer months).
-
-For RA (\autoref{sec:ta}), it depends on your prof.
-and their funding.
-When writing grant proposals, profs.
-typically include summer funding for their students (\autoref{sec:ra-cost}).
-However, funding is never guaranteed, e.g., the prof might not get the grant.
-For my students, I have been fortunate to have funding to support them over the summer.
-Over the 3-summer months, I typically pay them 1/3 of their 9-month stipend.
-I prioritize summer funding for my students because Mason has very good TA resources so they never have to worry about funding during the AY.
-
-Finally, for fellowships (\autoref{sec:fellowships}) you might get paid over the summer depending on your fellowship (\autoref{sec:fellowships}).
-Good ones, e.g., from NSF, Google, and Microsoft, will pay you the whole year.
-
-\chapter{Choosing Schools and Professors}\label{sec:schoolsandprofs}
-
-\epigraph{\vspace{-0.2in} The answer to life's problems aren't at the bottom of a bottle. Heh heh! They're on TV!}{\textsc{The Simpsons}}
-
-Choosing a school and an adviser is clearly among the most important things in your mind when you apply and especially when you get admitted.
-This is further complicated due to cultural differences and the unfamiliarity of international students with the US higher education system.
-This section aims to mitigate some confusion and help you make informed decisions.
-
-\section{Choosing a University}\label{sec:choosing-university}
-
-We will first discuss universities in the US that offer PhD in CS. Then we will talk about how to select them.
-
-\subsection{Schools that offer PhD in CS}
-
-Most US universities have CS studies, but many of them do not have a PhD program in CS.
-These universities might offer just Bachelor's degrees (e.g., BS) and no graduate studies (i.e., no MS or PhD degrees), or they just offer MS programs (but no PhD).
-For example, Penn State in University Park has PhD in CS,  but Penn State in Harrisburg only has BS and MS in CS, and Penn State in York only has BS in CS.
-On the other hand, multiple locations of the University of Texas, e.g., Austin, Dallas, and Arlington, have PhDs in CS.
-
-Thus, if your goal is PhD in CS, you have to aim only for schools offering such a degree.
-%This might sound obvious but many students (even those in the US) get confused due to the large number of universities.
-While this can be confusing due to the large number of universities in the US, a little research, e.g., searching for PhD in CS from the school website, will help you find out.
-Schools listed in \autoref{sec:ranking} have PhD programs in CS, so you can start there.
-
-% \subsection{R1, R2, ...}
-
-\subsection{Selecting and Ranking Schools}\label{sec:selecting-ranking-schools}
-\begin{center}
-  \includegraphics[scale=0.6]{files/c1.png}
-\end{center}
-
-International students often put schools into two bins: (i) top schools that they dream about, and (ii) everything else.
-They commonly use rankings from US News, which is not transparent and questionable\footnote{\url{https://cra.org/cra-statement-us-news-world-report-rankings-computer-science-universities/}}.
-Sometimes they evaluate based on the reputation of the school's undergrad program or the reputation of the school's non-CS programs such as medical, math, or physics.
-They even rank universities based on popular places they know in the US, e.g., California, Texas, and New York.
-
-Instead of these superficial criteria, you should consider the CS program and the research interests of faculty members.
-You can learn about these using resources such as \href{https://csrankings.org}{CSRankings.org}.
-You will be very surprised to learn that a school that you didn't know much about can have very strong research in your topic (and vice versa, a school you thought highly about might have no faculty working in the research field you're interested in).
-This is also a good way to learn about individual faculty, e.g., who works on what, and well-known CS conferences\footnote{In CS (and probably only in CS), conferences, not journals, are often the main venue to publish research findings (see \href{https://homes.cs.washington.edu/~mernst/advice/conferences-vs-journals.html}{why here}).}.
-%\autoref{sec:ranking} gives the top 50 CS programs in the US according to CSRankings.
-
-\begin{commentbox}[Dat:] Most Vietnamese students, including those from top schools, \textbf{do not know} about CSRankings.
-Maybe applicants who worked at top research places such as VinAI would know about it.
-\end{commentbox}
-
-\paragraph{What matters to you?} While many find CSRankings useful, it is not perfect, and in general, rankings are superficial, and you need to do more research to be informed and make better decisions.
-For example, you might prefer places with a large community from your country (e.g., Northern VA has a large population of Vietnamese).
-You might prefer places with high-tech companies (Seattle or Silicon Valley), with many outdoor activities (e.g., hiking, skiing), or with better weather (\emph{``PhD can be depressing, so would you rather be depressed in California or New York?''}).
-You might need to think about the cost of living, e.g., living in California is more expensive than in Virginia, or places with high crime rates (note that while some universities might be in a high-crime city but the campus itself is safe).
-
-If you get admission to several places, you should consider attending Open Houses (\autoref{sec:accepted}) and contact profs.
-that you're interested in at those places and talk to them.
-They would be more willing to chat with you now that you have been admitted.
-Ask questions about \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Answers-to-Ph.D-Advisor-Guide}{their work}, how they manage students, and their expectations.
-You can even ask to contact their students.
-
-\begin{commentbox}[Hung:]
-  I always encourage the students I admitted to talk with my students and the students of other faculty in other schools who admitted them.
-You will unlikely hear straight-out complaints from current students in a prof’s group.
-But sometimes what is important are things that they (current students) don’t tell you.
-Pay attention to their ``level of excitement'' being in the group.
-\end{commentbox}
-
-\begin{commentbox}[Xiaokuan:]
-  Chinese students often only look at US News rankings when selecting their PhD universities (I did that, too, when I was applying for PhD positions).
-  Now that I am a professor, I find it to be the least promising way.
-  %
-  The reason is that US News does not provide a good metric for evaluating the quality of the PhD program.
-  %
-  If you want to do great research, CSRankings is the best way to find good and active professors (which did not exist when I was applying),
-  since it solely focuses on publications at top-tier CS conferences.
-  %
-  Also,
-  I think PhD is not only about research;
-  you need to also consider your daily life there since you will (probably) stay for at least five years.
-  %
-  You might regret it if you did not consider this seriously before applying.
-  %
-\end{commentbox}
-
-\subsection{PhD in other related fields: CE, IST, Cybersecurity}\label{sec:related-fields}
-
-You \emph{do not} need to do a PhD in CS to work in your area of interest.
-For example, in addition to a traditional CS department, Mason has IST and Cybersecurity departments, both of which have faculty with PhD in CS and work on CS topics (e.g., AI, Security, Robotics).
-So it is possible that you still get to do CS research and publish in CS-focused venues even if you're not in a traditional CS program.
-It is common to see faculty with PhD in CS in a non-CS department as well as faculty with non-CS PhD in a CS department.
-
-However, if your goal is a PhD in CS, then you likely need to be in the CS dept \emph{and} advised by a CS faculty.
-A non-CS faculty can serve in the PhD dissertation committee (common) or \emph{co-advise} (somewhat rare), but your main PhD adviser will likely be a tenure-line faculty in CS (\autoref{sec:faculty-types}).
-For example, a prof. in Stats or Math might be able to serve as a co-adviser, but not as a sole adviser of a student in a CS PhD program. 
-If in doubt, check with the CS department for their requirements.
-
-For this specific reason,  CSRankings includes only faculty who can advise CS PhD students.
-I also have compiled a \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{list} of Vietnamese faculty who can advise PhD students.
-\autoref{sec:faculty-types} talks more about who can serve as your PhD adviser.
-
-\section{Choosing an Adviser}
-There is no one-size-fits-all answer to this question. The best adviser is the one that you can work well with.  But how do you find such a person?
-
-Fortunately, while some non-US programs require finding an adviser and research topic before starting the PhD (\autoref{sec:non-us-differences}), CS PhD programs in the US will typically give you a couple of years to search for advisers and research topics.
-This is especially true if you're admitted with TA (\autoref{sec:ta}), which gives you time to explore and find an adviser.
-
-\subsection{Finding the right adviser}
-
-Here is some advice for finding a prof.
-Assuming you're not familiar with any particular one, then first search for profs.
-that share similar research interests.
-For example, in CSRankings, if you want to work with PL, you can search for those published in PL conferences.
-If you want to work with SE \emph{and} AI, you can search for faculty who work in both SE and AI.
-Then you research that prof.
-and contact them (\autoref{sec:contact}).
-
-\begin{commentbox}[Xiaokuan:]
-  Whether the student's research interest matches that of the adviser is very important;
-  if there is a mismatch,
-  either the student or the adviser has to make compromises,
-  which often leads to disagreements or conflicts.
-  IMO, the adviser should be the one who \emph{guides}  students to do research while allowing students to pursue their own interests,
-  instead of \emph{dictating} their research.
-\end{commentbox}
-
-Another way is taking graduate-level courses in the topics you are interested in.
-Many profs teach \emph{special topics courses} and \emph{research seminars}, and they might be recruiting students.
-Do well in the class, answer questions, talk to the prof.
-after classes, etc---being stand out.
-Many profs, including myself, prefer taking in new students this way.
-It gives both the prof and student more time, e.g., a whole semester, to work and evaluate the relationship before making any commitment (sounds like a marriage!).
-
-You can also ask to do an independent study or research with a prof.
-This can be informal (no credit) and takes place during the summer or winter break.
-For example, I do this with several students, some of whom are undergrads.
-Many will drop out because they find they don't like my research, but some find that they like the work.
-
-Ultimately, choose a prof.
-that fits you by communicating with them, taking their courses, meeting and asking them questions, and talking to their current students.
-It will take time and effort, but since you will be working with this person for 5+ years, it is important to try to find the right one.
-
-\begin{commentbox}[Thanh:]
-  In my opinion, having a well-suited adviser is crucial for a successful PhD and research career.
-One effective approach to finding a suitable professor is by working with a professor during your undergraduate studies.
-An exemplary instance is VinAI's residency program, where residents collaborate with professors from the US for two years before applying to PhD programs.
-Many VinAI residents have achieved remarkable results and gained admission to prestigious US universities.
-Unfortunately, VinAI's resident program is limited to AI research.
-
-  In other fields, e.g.
-Software Engineering, Vietnamese students face challenges in reaching US professors.
-Do you have any tips for Vietnamese students who want to connect with US professors and work as research assistants?
-  \tcblower
-  \textbf{Vu}: \autoref{sec:contact} shows how to contact a professor for research opportunities.
-Many probably will say no (or do not reply) as they do not have the bandwidth to take on random students, but some may say yes if they see a potential fit.
-\end{commentbox}
-
-\subsubsection*{Additional Resources}
-\begin{itemize}
-  \item \href{https://www.cs.columbia.edu/wp-content/uploads/2019/03/Get-Advisor.pdf}{The Definitive ``what do I ask/look for'' in a PhD Advisor Guide}
-\end{itemize}
-\subsection{Who can serve as a PhD adviser?}\label{sec:faculty-types}
-
-\emph{Not} every faculty can serve as your official PhD adviser.
-Let's try to understand the different types of faculty and their roles.
-For example, you might hear about tenured,  tenure track, and teaching faculty.
-You might also hear about assistant, associate, full, adjunct, emeritus, teaching, and research professors.
-Here is a primer on these terminologies.
-
-\paragraph{Types of Faculty} There are typically two types of faculty: tenure-line and teaching.
-The main difference is their focus: research, teaching, or both.
-The type of faculty can affect their ability to serve as your adviser.
-
-\textbf{Tenure-line faculty}, consisting of tenured and tenure-track profs., focus on research, which includes publishing papers, obtaining grants, and mentoring PhD students.
-They often have a low teaching load (e.g., 1 per semester).
-\emph{Tenured} faculty are professors who have been promoted to permanent positions (informally, very hard to fire them).
-\emph{Tenure-track} faculty are (more likely young) professors who are on the track to get tenure.
-For most schools, tenure-line is the main type of faculty that can serve as a formal adviser of PhD students.
-\autoref{sec:tenure-vs-tenure-track} talks more about choosing between tenure and tenured-track professor as your adviser.
-
-\textbf{Teaching faculty} (sometimes called instructional faculty or professor of practice) mainly focuses on teaching.
-They typically teach 3--4 classes per semester (which is quite a lot) and \emph{do not} have research responsibilities, e.g., they do not have to worry much about publishing papers or obtaining grants.
-As such they typically do not have funding (e.g., RA~\autoref{sec:funding}) to support PhD students.
-Depending on the university, they might be able to \emph{co-advise} and mentor PhD students with a tenure-line faculty but typically do not formally advise and graduate PhD students.
-
-Note that at many universities do not allow \emph{non-CS profs}, even in tenure-line ones, to solely advise and graduate PhD students in CS (see \autoref{sec:related-fields}).
-
-Some universities have \textbf{research faculty} or research scientists, who are typically non-tenure line faculty and focus on research (e.g., publishing papers and getting grants). 
-They can recruit students and provide RA funding for them. 
-They can also teach classes but typically have a lower teaching load.
-For some universities, they \emph{can} solely advise and graduate PhD students, but for others, they cannot.
-
-\textbf{Adjunct faculty} is not full-time, e.g., they might be working in industry and teach a class or two for fun.
-\textbf{Emeritus faculty} means they are retired from the university, e.g., they could move to the industry, but still have some affiliation with the university.
-Due to their roles, these profs.
-typically do not advise PhD students.
-
-\paragraph{Ranks} Faculty have ranks such as assistant, associate, and full, regardless if they are tenure-line or teaching.
-Assistant means new faculty, associate means they have been promoted, and full means they are senior.
-Typically, a tenure-track faculty starts with being an assistant professor and then gets tenure and promoted to associate (typically after 6 years) and then full (time varies, some become full within 3--4 years, some 10+ years, some remain associate).
-A non-tenure track faculty, e.g., teaching or research faculty, also starts with Assistant, then is promoted to Associated, and Full prof.
-
-\subsection{Tenured or tenure-track faculty?}\label{sec:tenure-vs-tenure-track}
-
-\begin{center}
-  \includegraphics[scale=0.4]{files/c8.png}
-\end{center}
-
-Now that you know a bit about tenured and tenure-track faculty (\autoref{sec:faculty-types}).
-Which one should you choose as your adviser?
-The short answer is that either can be a good adviser, but they have different strengths and weaknesses.
-
-Tenure-track faculty, e.g., assistant professors, are likely young and active in research (they have to, to get tenure).
-Thus, they will devote more time to work with you and push you to do research and publish.
-However, they may not have as much experience in managing students and may not have as much funding (yet).
-
-Tenured faculty, e.g., associate and full profs., are likely older, more well-known, and have more experience in managing students.
-However, they might not push you as hard and expect you to figure things out yourself, i.e., you need to be independent.
-Some tenured faculty are also no longer active in research and are more involved with administrative responsibilities or with their startup companies (this means they will likely not take new students).
-\autoref{sec:faculty-types} gives more details on faculty types.
-
-% \subsection{Strong Faculty vs High-ranked School}
-% \begin{commentbox}[Thanh:]
-%   When considering PhD programs, we often wonder if we should prioritize a high-ranking university or a professor with a strong reputation?
-Of course, both are great, but it is hard to achieve both.
-I believe receiving some guidance from you  would be incredibly valuable.
-% \end{commentbox}
-
-% This is a good question and there doesn't seem to be any best answer or specific algorithm that you can use.
-But I'll throw out something for you to think about.
-To me, there are many good schools in the US and so you have some leeway in choosing.
-For example, I don't believe there's much difference among schools in top 5 (e.g., CMU, UIUC, UCSD, MIT, GTech are in the same equivalent class), top 10 (e.g., Michigan,  Stanford, UWash, UC-Berkeley, Cornell is another equivalent class), top 20, top 30, top 40, and so on (though I would say schools above 100 are pretty much in the same equivalent class).
-It's hard to say a top X is much much better than a top X+20 (e.g.,  Purdue at 15 is stronger than Yale at 35, but not \emph{that much} stronger as you would think).
-So keep that in mind, while ranking does matter (e.g., more active faculty, stronger students), the lower ranked universities are still pretty good, might have sub fields that are extremely strong (NCState is 40th, but in Software Engineering it is easily top 5), and many unique things.
-This is in contrast with universities in Vietnam where there are large gaps among the top 5, 10, and so on.
-
-\subsection{Should you contact a professor? How to get a desired reply?}\label{sec:contact}
-
-\begin{center}
-  \includegraphics[scale=0.6]{files/emails.png}
-\end{center}
-
-Faculty often receive ``cold'' e-mails from prospective students.
-Most of the time, we ignore these emails, but on some rare occasions, we do answer them.
-So how to write an email that gets our attention?
-
-First, if you want to contact a prof.
-to \emph{ask about your admission chance}, please \textbf{don't}.
-We don't know and can't answer because as explained in \autoref{sec:evalapps}, we don't make individual decisions and might not even be assigned to evaluate your application.
-It is the same as sending a paper draft to a journal editor and asking them if your paper has a chance.
-
-So how to get someone to look at your profile and give input?
-You could ask your professors, LoR writers, collaborators, or those who have previously applied.
-For this kind of feedback, ask someone you have a personal connection with.
-
-On the other hand, if you want to contact a prof.
-to ask about \emph{research opportunities}, or \emph{GTA/GRA} support, then \emph{yes}, I believe you should---it is \emph{worth it}.
-However, you really need to put effort into it and do it the \emph{right way}.
-
-First, read the prof's website, and see if they say something about contacting them.
-Many profs.
-explicitly state how prospective students should (or should not) contact them, e.g., using specific email subjects.
-In general, the best way to catch the prof.'s attention is to \emph{customize your email} for them.
-For example, read their papers, know what they work on, and see if you are interested in their research.
-Then email them and talk about how/why their work would match yours.
-In contrast, if you write a generic email that can be sent to multiple professors (e.g., if you just change some names and keywords in the email or copy and paste paper titles), you will not get a response.
-
-Below is a good example that I would likely reply.
-
-\begin{commentbox}[Dear Prof. Nguyen,]
-
-  I am writing to inquire about potential research opportunities as a GRA in your group at Mason. Currently I am an undergraduate student in Computer Science at UNIV and plan to graduate in May 2023.
-
-  I have read your TSE'21 paper on numerical invariant generation, and I am interested in this line of dynamic invariant research.
-I have worked (optional: with prof.
-Y at Z) on static program analysis and I think it could be used to tackle the spurious issues mentioned in your paper.
-I have a short paper at conference/workshop C and a project on symbolic execution (Github repo G).
-
-  ...
-
-  This is a good example because it is written just for me.  It shows that the student knows about my work on invariant generation and has a related background (paper C and project G).
-\end{commentbox}
-
-Finally, profs. are very busy so don't take it personally if you don't get anything from them (though I would be very surprised if such thoughtful emails get no replies!).
-
-\begin{commentbox}[Xiaokuan]
-  Applying for PhD and contacting a potential PhD adviser is a classic \textbf{`why me, why you'} problem,
-  similar to looking for a job in a company.
-  On a high level,
-  you need to show that you have done your \emph{homework}
-  regarding the professor and the university,
-  and clearly explain:
-  1) why do you think you are a good fit in professor A's group?
-  2) why do you want to be advised by professor A, not B?
-  3) why do you want to apply for university X, not Y?
-  If you don't want to spend time doing your homework,
-  the chance of getting a reply is close to zero.
-\end{commentbox}
-
-\begin{commentbox}
-  [Deepak:]
-  In my view, cold emails are not welcome by most faculty members and should be avoided.
-However, if one is already admitted to a program in some department, by all means, email the faculty you may be interested in working with, but do mention right at the beginning that you are already admitted to the program as well as several other universities.
-State specific areas (preferably specific topics-ML, robotics instead of AI).
-\end{commentbox}
-
-\subsubsection*{Additional Resources}
-\begin{itemize}
-  \item \href{https://yonatanbisk.com/emailing_professors.html}{A Note about Emailing Professors} by Yonatan Bisk
-  \item \autoref{sec:accept-postpone-decline} How to accept/postpone/decline PhD offers (and do it gracefully)
-\end{itemize}
-
-\subsection{Things to avoid when cold emailing a professor}
-
-\begin{itemize}
-
-    \item \textbf{Not customizing the email}. These emails typically talk about the student and not why they are interested in the professor's work. See writing tips in \autoref{sec:contact}
-    \item \textbf{Sending the same email to multiple professors}.
-Sending emails to multiple profs.
-using BCC or even CC (where we can see all addresses the email is being sent to) will result in no reply or a very harsh one on how unprofessional this is.
-    \item \textbf{Too much information}.
-Keep your email short and to the point.
-If we are interested, we will ask for more information.
-2--3 paragraphs are enough.
-Personally I also do not want to see course transcript or CV in the first email.
-    \item \textbf{Not being professional}, e.g., do not call the professor by their first name in the first email.  Use Prof. Lastname or Dr. Lastname (\autoref{sec:address}).
-    \item \textbf{Not reading the professor's website} and asking questions that are already answered there.
-\end{itemize}
-
-\chapter{Miscs and FAQs}\label{sec:faqs}
-
-\epigraph{``I want to share something with you – the three little sentences that will get you through life; number 1: Cover for me, number 2: Oh, good idea, Boss, and number 3: It was like that when I got here.''}{\textsc{The Simpsons}}
-
-\section{Why did I get rejected?}\label{sec:why-rejected}
-
-Many students lament that they get no interviews or are rejected (with strong profiles) and that the admission process seems random.
-However, while it is true that the process is not perfect, it is not random.
-Other than not having a strong profile (e.g., research potential, GPA, LoRs, SOP, poor interviews), here are some other reasons for rejections.
-
-\paragraph{Not a good fit and Bad luck}  You could have an excellent profile, but if you are interested in a research area that the program does not have, you might not be admitted.
-Similarly, if no faculty who is willing to advise you (e.g., they are on sabbatical or personal leave, do not have sufficient funding, or already have too many students), you will not be admitted.
-This is good for you, as you don't want to be in a program where no one can advise you.
-
-Related to this is that you just have bad luck and apply at the wrong time.
-For example, since 2024 there has been a huge surge in students interested in AI and NLP (thanks, ChatGPT!).
-Consequently, AI/NLP faculty might be overwhelmed and cannot consider many candidates, even those with excellent profiles.
-
-\paragraph{You aim too high} This is a very common reason: you have applied to schools that are too high for your profile.
-While being ambitious is good, you should understand how PhD admission works (e.g., by reading this document and realizing things such as having a good GPA doesn't mean much to top programs) and be realistic!
-
-Before applying, you should talk to your professors and ask them to give you an honest opinion on where you should apply.
-To increase their chances, many students apply to a range of schools, including ``safety'' ones.
-
-\paragraph{Overqualified or Lack Interest}  This might be paradoxical and is the opposite of aiming too high.
-However if adcom believes that you are likely to get admitted and go to a better program, they might not be so enthusiastic to admit you and want to save the spot for someone else.
-
-A related reason is that you did not show enough interest in the program.
-For example, your SOP did not mention any faculty or research area at the program, you did not respond to professors interested in interviewing you (which might be considered unprofessional and burn bridges), or during the interview you showed little knowledge, interest, or enthusiasm (because they are your safety schools) in the program or profs.
-
-\paragraph{Low English exam scores (e.g., IELTS or TOEFL)}  Profs.
-might not care much about these, but the college or the university often sets a minimum that you need to pass to be considered, especially for TA funding (e.g., you need to communicate well with students as a TA).
-Thus, while profs.
-are willing to argue for your case, they might be reluctant to go against the requirement of the university on standard exams.
-
-Note that English exams are not required for all students, e.g., domestic students and those from certain countries are waived from taking these exams.
-For example, Mason waives English exams for students from these \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-Mason#standard-tests-waiver-eligible-countries}{countries}.
-
-\paragraph{Red flags} Various types of red flags can cause concern to the adcom (e.g., too many STEM courses with low or withdraw grades, plagiarism, cheating, or other academic dishonesty).
-Another one is that you have a history of jumping from one program to another (e.g., you have been in multiple different programs in a short time).
-Adcom members might have contacts in other programs and heard about your stories, or your LoRs might mention these.
-
-If you think you have these issues, you should address them in your SOP or ask your letter writers to do so.
-In general, these things are rare, but they do happen and cause concern to the adcom.
-
-\section{Timeline: Why are rejection letters sent so late?}\label{sec:late-rejection}
-
-%"Just reject me already!" is a common sentiment among applicants.
-Indeed, school will first send out admission offers to the top candidates.
-They do not send out rejection letters because there is still a chance that some of the top candidates will decline the offer.
-If they do, then the school will go to the next set of candidates and send out offers to them.
-This process continues until all spots are filled.
-This is why you might not hear back until late in the admission cycle.
-
-\paragraph{Acceptance Letters} Universities prioritize sending out acceptance letters first. This allows the admitted students to make decisions and plan for their studies.
-
-Some universities have rolling admissions, meaning they send out acceptance letters as they review applications (e.g., some universities sent out letters as soon as early February).
-Others have a specific date when they send out the first round of acceptance letters.
-
-\paragraph{Response Deadlines} Accepted students are usually given a deadline to make decisions on their offers, often around \textbf{April 15th} in the US.
-This is a standard deadline for PhD programs due to the Council of Graduate Schools (CGS) Resolution (\href{https://cgsnet.org/wp-content/uploads/2024/01/CGS_April15_Resolution_Jan312024.pdf}{see more}.
-After this date, CS programs can gauge how many slots remain unfilled.
-
-\paragraph{Waitlist} Most CS programs have a limited number of slots for PhD students, and thus put many good students on a waitlist.
-If accepted students decline the offer, then offers are sent to students on the waitlist.
-So if you see people getting accepted, that does not mean you are out yet.
-
-\paragraph{Rejection Letters} Schools typically start sending out rejection letters to remaining applicants \emph{after they have finalized their admissions decisions}.
-Thus, rejection letters are often sent out late (e.g., after April 15th or even much later).
-Not much you can do here, you can try to contact the school to ask about your status, but they might not reply, they might say they are still reviewing applications, or give you inaccurate information (e.g., you will hear in two weeks).
-In short, you just have to be patient and wait (however, beware that some schools do not send out rejection letters at all).
-
-\section{Increasing my admission chance}\label{sec:improve-your-chance}
-
-\begin{center}
-  \includegraphics[scale=0.2]{files/alpinist-climbing-peak-mountain-comic-hand-drawn-vector-illustration.jpg}
-\end{center}
-
-Given the high number of quality applicants and a limited number of spots, in addition to having a good application profile, you want to show something that makes you \textbf{stand out}.
-For example, even if you do not have research experience, you can talk about your personal projects, as long as they can help show you can do research.
-For example, if you have an open-source project that has lots of stars in Github, then you should mention it.
-If you often write technical, research-like blogs with many viewers, talk about them too.
-
-There are other things you might not think are important but can make you stand out.
-For example, if you have a strong background in a non-CS field but can be integrated with CS, e.g., you have a degree or background in \emph{dance} or \emph{music} and want to integrate them with CS, do talk about it.
-Are you a female or a minority in CS (\autoref{sec:urm})?
-Do you participate in outreach activities that help increase diversity and inclusion in CS?
-Diversity is highly valued in CS programs in the US, and experiences in these areas can make you stand out and get noticed from reviewers.
-
-\begin{commentbox}[Vu:]
-In his \href{https://matt.might.net/articles/how-to-apply-and-get-in-to-graduate-school-in-science-mathematics-engineering-or-computer-science/}{post}, Matt Might was initially unsure about an application.
-However, upon learning that the applicant had led a 100km hike in the Himalayas, he decided to accept the applicant.
-This is a good example of being \emph{stand out}, and I would also advocate for that student as this shows they have the persistence and determination required for research.
-\end{commentbox}
-
-% It is worth noting that in many cases,  an application could be rejected by most reviewers, however have a "champion", who doesn't even have to be in the Admission comittee, that strongly advocate for your profile.
-
-\section{How long do you evaluate an application?}\label{sec:ievaluate}
-
-On average, I spend about 10 minutes reviewing each application (less for clear rejections and more for potential acceptances).
-While this seems short, it is not that difficult to tell if an application is good or bad.
-In fact, this
-is twice what other faculty spend on average, e.g., Philip Guo spent \href{https://pg.ucsd.edu/PhD-application-tips.htm}{3--5 minutes} per application.
-
-For each application, the Mason application system compiles a single PDF file, which consists of a summary, transcripts, test scores, LoRs, a CV, SOP, and writing samples.
-I usually start with the \emph{summary}, checking for low GPAs or test scores below the university minimum (\autoref{sec:standard-tests}).
-I then skim the \emph{transcripts} for low grades in relevant courses, noting issues like ``many low grades in main courses or unknown international school with good GPA''.
-These are not as important as LORs or SOP but I read them first because the review system has questions about them (e.g., ``is GPA good?
-is IELTS sufficient?'')
-
-I read carefully \emph{strong LoRs} and skim weaker ones, noting either strong letters from well-known professors/researchers or generic content.
-For the \emph{CV}, I look for publications, research experiences, and notable achievements, noting things like ``published papers in top venues'' or ``gold medal in an international competition.''
-I read strong \emph{SOP} thoroughly (and skim weak ones). I note whether the SOP is exciting, research-oriented, and tailored to our program. All these notes are entered into the evaluation system.
-
-Finally, I make a decision, which is usually either a rejection or an offer of admission with full funding (e.g., from TA).
-I also recommend very strong candidates for the University Presidential fellowship.
-The adcom chair then reviews all evaluations, looks at entered notes, and asks reviewers to discuss and resolve discrepancies to reach a consensus (e.g., a reviewer wants to accept but the other wants to reject).
-Typically, the decision is made entirely by the reviewers; there is no involvement from the adcom chair, department chair, or others.
-
-\section{Do adcom members talk to each other?}\label{sec:adcom-discuss}
-
-In most cases adcom members, even those reviewing the same application, make decisions independently and do not talk to each other.
-It's not a strict rule but a common practice to avoid bias and ensure fairness (and also we don't have time to talk to each other).
-
-However, in some rare cases we do talk to each other.
-For example, when there are discrepancies in evaluations, in which case the adcom chair will ask reviewers to discuss the application to reach a consensus.
-We also talk to each other for interesting or strong applications, e.g., how to recruit this student or who should be the adviser.
-If the student mentioned a faculty member in their SOP, we might ask that faculty member if they are interested in the student.
-
-\section{Do adcom members read all applications?}\label{sec:read-all}
-
-While adcom members (in fact, any faculty) can view any application, we only review those that are assigned to us, which are already too many.
-Adcom chair will assign applications to reviewers based on their expertise (e.g., if a student says they want to do SE or interested in working with me), and reviewers will only review those applications.
-Occasionally we might look at other applications, e.g., if we know the student or have some special circumstances (e.g., if that student contacted me, I know that student, they are from Vietnam, etc).
-However, even if we look at them, we usually do not get involved in their evaluation.
-
-\section{Member of an underrepresented or LGBTQ+ groups}\label{sec:urm}
-
-A common question is whether you should mention, e.g., in your SOP, that you are a member of an underrepresented and minority (URM) group\footnote{URM in CS in the US include African Americans, Hispanic Americans, American Indians/Native Alaskans, Native Hawaiians/Pacific Islanders (excluding Asian Americans), and multiracial students identifying with at least one of the previously listed URM categories.} or LGBTQ+ group.
-This is a strong concern for international (and even domestic) students, who may have encountered discrimination in their countries or personal experiences.
-%This question arises often for international (and even domestic) students.
-The main worry is that they might get discriminated against, as often happen in their countries or personal experiences.
-%, and (ii) mentioning this might make them look like they are asking for special treatment.
-
-In my personal opinion, highlighting URM or LGBTQ+ identity can \emph{be beneficial}, especially if you can articulate how your diverse experiences contribute to diversity and inclusion in academics.
-Many universities have a strong commitment to diversity and inclusion and actively
-\emph{recruit students from underepresentative backgrounds} (e.g., some \href{https://cse-climate.engin.umich.edu/reports/climate-dei-reports/cse-climate-and-dei-report-2022-2023/#grad-ethnicity}{stats from UMich CSE} and \href{https://www.Mason.edu/news/2022-09/mason-now-top-10-public-university-diversity-innovation-and-cybersecurity-education-us}{Mason} often promotes itself as one of the most diverse universities in the US).
-Moreover, many scholarships and fellowships are created specifically for URM and LGBTQ+ students, which you should explore if qualify.
-
-Understandably, you might feel uncomfortable disclosing this, fearing you would lose a chance of working with a faculty due to their bias.
-However, if someone has this issue, then you might not want to work with them anyway (and such bias is not acceptable in academics and likely would not be tolerated by the university).
-In my experience, it is the opposite: \emph{many profs.
-actively seek to work with students from diverse backgrounds and view diversity as an asset to their research group}.
-
-Moreover, promoting diversity aligns with the goal of broadening participation in computing (BPC) that is strongly supported within academia (e.g., see the \href{https://plans.bpcnet.org/GeorgeMasonUniversity_ComputerScience_DepartmentalBPCPlan.pdf}{BPC plan at Mason CS}) and government funding agencies such as NSF (e.g., see \href{https://new.nsf.gov/cise/broadening-participation}{here} and \href{https://www.nsf.gov/pubs/2022/nsf22125/nsf22125.jsp}{here}).
-Also, sometimes providing this information can make it easier for faculty to apply for certain scholarships or fellowships to support you, e.g., from a private donor that has offered a fellowship only to students from certain backgrounds.
-
-Ultimately, this is a personal choice and you do what you feel comfortable.
-If you do not feel comfortable disclosing this, you should not feel pressured to do so and you should also ask your LoR writers to not disclose this information.
-Note that you can still talk about your experiences and perspectives without mentioning your identity.
-If uncertain, you can seek additional guidance from your close professors and collaborators However, it is important to recognize that diversity is strongly valued and celebrated within CS programs in the US, and that you should feel empowered to share your unique experiences and perspectives.
-
-%https://www.edsurge.com/news/2018-05-18-how-cornell-university-diversified-its-incoming-phd-computer-science-student-body
-
-%https://academia.stackexchange.com/questions/130591/is-it-standard-for-us-based-universities-to-consider-the-ethnicity-of-an-applica
-
-%https://www.cs.cornell.edu/~bindel/paper/diversity.pdf
-
-%https://cs.brown.edu/degrees/doctoral/applications/helpful-resources-applying-computer-science-phd-programs/
-
-\section{My undergrad was not in CS or related areas}\label{sec:non-stem}
-
-You still can apply to PhD in CS \emph{as long as} you can demonstrate you are ready for it through your background, research experiences, LoRs, statements, etc.
-You might be even able to leverage this to make your profile stand out as mentioned in \autoref{sec:improve-your-chance}.
-
-A main question adcom has for a non-CS or non-STEM student is if you have the sufficient technical background obtained through core CS courses.
-So you need to show that you have such knowledge through your coursework, projects, or research.
-For example, if you have taken a course on Algorithms, even online ones like Coursera, you can talk about it in your SOP.
-If you have done a project that requires knowledge of OS or have a professional certification (e.g., A+) through work, you can talk about it.
-If you have done research that requires knowledge of Discrete Maths, you can talk about it.
-You can also ask your LoR writers to talk about your technical background.
-In summary, in your application, convince us that you have the background to do CS PhD research.
-
-\paragraph{Core CS topics} Common CS knowledge that you should know are:
-\begin{itemize}
-  \item Programming Foundation (e.g., programming concepts in some modern languages)
-  \item Discrete Math (e.g., logic, set theory, proof techniques)
-  \item Data Structures and Algorithms (e.g., linked lists, trees, sorting, searching)
-  \item Computer OS or Systems (e.g., memory management, file systems, processes)
-\end{itemize}
-
-In short, you \emph{do not need} to formally take CS courses, you just need to show that you have this essential knowledge, e.g., through the mentioned ways.
-Many universities are well aware that incoming graduate students might not have all the technical background, so they often have a \emph{``bridge''} courses to help students catch up.
-For example, Mason has four bridge courses corresponding to the four core areas above that incoming students can take to catch up on their CS knowledge.
-
-\begin{commentbox}[Vu:]
-  I would advocate for a non-STEM student who shows that they have a strong drive for CS by studying core CS knowledge through various channels (e.g., self-study through online courses, projects, etc.).
-I have seen many students with non-CS backgrounds who are very successful in CS PhD.
-I also have seen many students with CS background who are not successful in CS PhD.
-So it is not about your background, it is about your drive and passion for CS research.
-\end{commentbox}
-
-\section{Is an MS required for admission to PhD in CS?}\label{sec:msrequirement}
-No.
-In fact, students can get MS degree ``along the way'' to PhD, e.g., after finishing the 2-year course work.
-However, MS can help admission if it gives research experience or is from a more well-known school than your undergrad institution (\autoref{sec:your-school}).
-
-If you have an MS then some course work \emph{might be} transferred for course credits, which \emph{might} save some time. But don't count on this.
-
-\begin{commentbox}[Vu:]
-    I start my PhD with an MS in CS from a US university.
-I found that the MS helped me a lot in research, but not with course work and I had to retake many courses.
-In general, don't worry if you don't have an MS.
-But also don't feel that you wasted your time if you have an MS, as it can help you in research.
-  \end{commentbox}
-
-\section{How long to complete the CS PhD program?}\label{sec:time}
-
-Typically, it takes 5--7 years for CS PhD in the US.
-This can be longer than CS PhDs in other countries, which might require MS first (recall from \autoref{sec:msrequirement} that CS PhD programs in the US do not require MS and you can get MS along the way to PhD).
-Within these 5--7 years, CS PhD students often take a ``leave of absence'' for 1--2 semesters to do internships at companies and research labs.
-
-\begin{center}
-  \includegraphics[scale=0.5]{files/c4a.png}
-\end{center}
-
-The first two years you typically take coursework (somewhat equivalent to an MS study), find an adviser, and learn how to do research.
-The next 2--3 years you focus on your research, form a dissertation topic, and get results published.
-The last 1--2 years you continue to publish, write and defend your dissertation, and look for a job.
-In many cases you might take a summer or two off to do an internship to get additional research opportunities.
-The PhDComics figure above shows the ``ambition'' level of a PhD student over their years of study (they miss the 6--7th year when the ambition is \emph{``Just let me graduate''}).
-
-% \begin{commentbox}[Vu:]
-%     I start my PhD with an MS, and it took me 7 years (Fall'2007-- Fall'14). I spent half a year doing an internship at the Naval Research Lab.
-%   \end{commentbox}
-
-\section{PhD in the US vs. Other Countries}\label{sec:non-us-differences}
-
-\begin{table}[h!]
-\centering
-\begin{tabular}{|c|c|c|}
-\toprule
-\textbf{Aspect} & \textbf{US PhD Programs} & \textbf{Other Countries} \\
-\midrule
-Duration & 5-7 years & 3-5 years \\
-MS Required & Not required & Often required \\
-Coursework Required & Yes (first 2 years) & No \\
-Research Proposal Required & No & Yes (in some countries) \\
-Academic (Faculty) Job & Direct & Postdoc \\
-Work Life Balance & Less & More \\
-\bottomrule
-\end{tabular}
-\caption{Comparison of US and Other PhD Programs}\label{tab:us-vs-other}
-\end{table}
-
-\autoref{tab:us-vs-other} summarizes the main differences between CS PhD in the US and other countries. % This is based on my experience and what I heard from others who did PhD in other countries.
-
-\emph{MS requirement and PhD duration}:  CS PhD programs in the US do not require an MS degree (as mentioned in \autoref{sec:msrequirement} and \autoref{sec:time}).
-In contrast, many other countries require having an MS degree before joining a PhD program.
-This means that US PhD programs are longer (5--7 years, 2 of which are coursework) than other countries (3--4 years, no coursework).
-
-\emph{Project proposal}: in many countries, you have to choose a project and adviser \emph{during} the application process (e.g., you write a proposal to a potential adviser).
-But this allows you to start your research right from the beginning.
-
-In the US, you often start your PhD without an adviser or project and find them later.
-You will have two initial years to explore and find an adviser and research topic.
-After that you explore projects with that adviser before settling on one (can take a couple of years).
-
-\emph{Course work}: as mentioned above in the US you will spend the first couple of years taking classes and exploring potential adviser and research topics. 
-After that, you have to pass a series of exams during your PhD, e.g., qualifying exam, comprehensive exam, thesis proposal defense\footnote{The word ``ABD'' (all but dissertation) is used in the US to refer to a PhD candidate who has finished all course work and exams and only needs to write and defend their dissertation.}.
-
-In other countries, you start your research right away, e.g., you immediately work on the research project you proposed with the adviser you chose.
-Moreover, you might not have exams like those in the US or only have to do a few of them.
-
-\emph{Funding}:  In many countries, funding comes from the university or the gov't.
-This funding often has a fixed duration (e.g., 3 or 4 years).
-In the US (\autoref{sec:funding}), funding (e.g., RA) comes directly from your adviser (no fixed duration).
-There are also fewer TA opportunities in European universities compared to the US.
-
-\emph{Academic position after PhD}: In other countries, PhD graduates interested in academia typically apply for additional research appointments, i.e., postdocs, and then consider faculty positions.
-
-In the US, PhD graduates often apply directly for faculty positions. Postdoc for US graduates is no longer a popular option as it was before.
-
-\emph{Work-life balance}: PhD students in the US are often said to be overworked compared to other countries, e.g., in Europe.
-This is partly due to the longer PhD program and that US PhD students are often paid through TA, which requires them to do TA in addition to their own research.
-In contrast, PhD students in other countries are often paid through fellowships, which might not require doing TA.
-
-\section{How do I call or address a professor?}\label{sec:address}
-
-\begin{center}
-  \includegraphics[scale=0.5]{files/c5.png}
-\end{center}
-
-If you're reaching out to a professor for the first time,  address them as Prof.
-or Dr.
-Lastname.
-Many international students use Prof.
-or Dr.
-FirstName LastName, but this can come across as if you're simply copying and pasting names.
-It's not necessary, so stick with Prof.
-or Dr.
-Lastname.
-
-Furthermore, avoid using Mr., Mrs., Miss, or the professor's first name if you're not acquainted with them yet.
-As you become more familiar with your prof and depending on their preferences, you may transition to addressing them by their first name.
-For example, I prefer that my students and colleagues call me Vu. Some students call me \emph{Dr. Vu}, which I find a bit amusing but am totally fine with it.
-
-\begin{commentbox}[DK:]
-  I was amused to read this as if I recall correctly, you never called me by my first name when you were at UNM. You always called me Prof. And, many times, I would jokingly call you back as Prof. Vu.
-  \tcblower
-  \textbf{Vu}: Yes, for some reason I enjoy calling you ``Professor'' (without appending a last or first name).  The use of Prof. Vu may have foreshadowed my future in academia.
-\end{commentbox}
-
-Note that in some universities the formal title Dr.
-Lastname is preferred over Prof.
-Moreover, be aware that not all faculty members may hold a PhD, in which case using Prof.
-Lastname is a suitable alternative.
-	You just need to observe and follow the conventions at your particular institution.One way to determine how to address a professor is to observe how they sign their emails or how they introduce themselves in class.
-For example, I introduce myself as ``Vu'' in class.
-
-\paragraph{Referring to professors you know} Because you are already familiar with these individuals, you can just informally use their names if they are OK with it as mentioned above (or Dr./Prof., if you want to be formal).
-You can also include their institution if it makes it more precise.
-For example, I can say:  \emph{``I did my postdoc with Jeff Foster at Univ.
-of Maryland''}.
-
-Do not include ranking (e.g., Assistant, Associate, Scientist, ...) when referring to someone.
-I see many international students include a lengthy title of people they know, e.g., \emph{I am advised by Asst.
-Prof.
-X, and I also collaborate with Distinguished Scientist Y}.
-This is \emph{not necessary} and makes it look like you're trying to show off your connections.
-These nuances represent some cultural differences that you may encounter and will gradually adapt to.
-More on cultural differences in \autoref{sec:cultural}.
-
-\section{How much do \emph{I} cost?}\label{sec:ra-cost}
-
-PhD students often ask why their salary is low compared to the large grants their advisers get.
-They also wonder why their offer letters sometimes show that their benefits are higher than what they receive (i.e., stipend).
-
-\begin{center}
-  \includegraphics[scale=0.3]{files/c6.png}
-\end{center}
-
-\begin{table}
-  \centering
-  \small
-  \caption{GRA cost breakdown. F \& A is Facilities \& Administrative Cost Base and
-    MTDC is Modified Total Direct Cost. These are things that the university can charge overhead to.}\label{tab:cost}
-  \begin{tabular}{rcl}
-    \textbf{Budget} & \textbf{Cost \$} & \textbf{Notes} \\
-    \midrule
-    GRA (9-month) & 30K & \\
-    GRA (summer)  &10	& 3-month, 20hrs/week\\
-    \textbf{Total Salary} &\red{40K}	&\\
-    \midrule
-    Health Insurance	&3K	& full year\\
-    Tuition (In-State) &	15K	& (\$680/ Credit + \$150/Student Fee/ Credit)* 9 credits = \\
-                    &&\$7470 (\$6120 + \$1350) per semester\\
-    \textbf{Total Tuition \& Insurance}	&\red{18K}	&Full year tuition + insurance\\
-    \midrule
-    Conference Registration	& 500 & \\
-    International Travel &	1800& \\
-    Domestic Travel	& 700	& \\
-    \textbf{Total Travel}&	\red{3K}	\\
-    \midrule
-    Total Direct Cost	& \red{61K}	&Salary + Travel + Health + Tuition \\
-    F \& A (MTDC)	& 21K	& Direct Cost - GRA Salary\\
-    Total Indirect Cost	& \red{12K}	&58.9\% of MTDC\\
-    \textbf{Total (Direct + Indirect)} &	\red{73K}	& Budget for a GRA\\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
-\autoref{tab:cost} shows the budget breakdown for a GRA per year.
-These numbers are based on my experience at public universities in the US.
-Private universities may have different numbers.
-For simplicity, I will assume the department has a 9-month stipend of \$30K and a 3-month summer of \$10K (a third of the 9-month stipend).
-I will also use Mason tuition rate of about \$15K/year for full-time study (which is quite cheap compared to private universities, e.g., Univ.
-of Chicago is a whooping \$70K) and a 58.9\% rate on \emph{indirect cost}, which is a typical rate charged for overhead or administrative costs (yes, after all, universities are businesses!).
-Finally, I assume the students take two conference trips per year, one domestic and one international (conf.
-registration, airline tickets, taxi, meals, etc are all included).
-
-In the end, the total budget comes out to be \$73K/year to support a PhD student.
-\emph{The summary is that over your 5--6 years of your PhD, you cost about \$400K, and while your stipend is X, your adviser probably pays 2X for you}.
-But of course, the nicest thing is that you do not have to pay for any of this!
-You get to gain the knowledge, do research, travel, (and don't even have to buy any research or computer equipment as mentioned in \autoref{sec:buying-equipment}) and also get paid!
-
-% \section{Having fun during a PhD?}
-% PhD students \emph{and faculty} probably find it amusing about the notion that students, especially international ones, can genuinely enjoy their PhD studies.
-In fact, after reading posts after posts on VietPhD.org on how PhD students are commonly mistreated, stressed, it seems being miserable is a norm during a PhD study.
-
-% There are many advice on surviving PhD that you can follow. But here I just list a few that works for me and what I advice my students to do.\tvn{TODO}
-
-\section{Buying Computer Equipment}\label{sec:buying-equipment}
-
-Students understandably get excited about their upcoming PhD journey and want to buy new computer equipment and electronics to prepare.
-However, you should first check with your professor.
-They might have funding to buy you computers and other equipment (e.g., software, books, keyboards, headphones, tablets, etc).
-Many CS programs also provide budget or computer equipment to incoming PhD students (e.g., a laptop).
-
-Keep in mind that these computers would be university property, which might be monitored and have certain restrictions, e.g., do not install illegal software on them (\autoref{sec:illegal-software}).
-You might also have to return them when you graduate.
-
-\section{Negotiating PhD offer (e.g., having multiple offers)?}\label{sec:negotiate}
-
-It is unlike that you can negotiate TA stipend, which is set by the university.
-Moreover, while RA is set by the adviser, it is often based on the department's standard, and it also would be unfair to other students if you get more (especially since you're new and have not proven yourself yet).
-However, your stipend is often increased every year by the university (automatically by a small amount).
-
-For a specific start date or TA assignment (e.g., TA'ing a particular course), you can ask for it.
-Also, there is typically no moving allowance for PhD students.
-In short, standard things set by the university or department are unlikely negotiable.
-However, you can ask for things such as computer equipment (\autoref{sec:buying-equipment}), books, (and maybe even furniture), and a conference travel budget.
-
-% \section{Will I be miserable during my a PhD?}\label{sec:happy}
-% There are many stories of how students are mistreated, stressed, and miserable.
-Issues including bad relationships with professors, conflicts with co-authors and lab mates, feeling discriminated against (e.g., because you're an international student), \emph{do} exist, and it is good to be aware of those.
-However, in reality, there are many good mentors, supportive lab mates and departments, and so on.
-So don't let social media make you feel pessimistic and deter your quest to advance knowledge.
-
-\appendix
-
-\chapter{Visa for International Students}\label{sec:visa}
-
-\epigraph{\vspace{-0.2in} I didn’t do it. Nobody saw me do it. You can’t prove anything!}{\textsc{The Simpsons}}
-
-% \begin{center}
-%     \includegraphics[scale=0.8]{files/visa.pdf}
-% \end{center}
-
-For international students pursuing a PhD in the US, \textbf{F-1} is the main visa that allows a student to study full-time at an accredited institution.  Here are some key points about the F-1 visa:
-
-\begin{itemize}
-\item \textbf{Employment:} You are allowed to work on-campus for up to \textbf{20 hours per week during the academic year} (because you still need to take classes) and \textbf{full-time} (typically 40 hrs) during official school breaks (e.g., summer and winter breaks).
-Off-campus employment requires authorization, which can be obtained through programs like Curricular Practical Training and Optional Practical Training described below.
-
-\item \textbf{Curricular Practical Training (CPT):} CPT allows you to participate in internships or practical training that is an integral part of their academic curriculum.
-CPT must be related to your field of study and can be full-time or part-time.
-It requires prior authorization from your university and must be completed before graduation.
-
-\item \textbf{Optional Practical Training (OPT):} OPT provides up to 12 months of work authorization for students before or after completing their degree.
-For students in STEM fields, there is an additional 24-month extension available.
-OPT requires prior application and approval from USCIS.
-
-\item \textbf{Full-time Enrollment:} You must maintain full-time enrollment status during the academic year.
-This means taking a minimum number of credits each semester, as defined by your program.
-Dropping below full-time status can result in loss of visa status.
-\end{itemize}
-
-\includepdf[pages=-]{files/visa.pdf}
-
-\section{For Spouses and Children}
-
-The spouses and children of F-1 visa holders can enter the US under the F-2 visa status. The F-2 visa allows family members to live in the US with the following conditions:
-
-For Spouse: 
-\begin{itemize}
-\item \textbf{Work Restrictions:} F-2 spouses are not permitted to work in the US.
-\item \textbf{Education:} They can study part-time but cannot enroll in full-time degree programs.
-\end{itemize}
-For Children:
-\begin{itemize}
-\item \textbf{Education:} F-2 children can attend K-12 schools but cannot pursue higher education full-time.
-\item \textbf{Work Restrictions:} Like F-2 spouses, children are not allowed to work under any circumstances.
-\end{itemize}
-F-2 visa holders must leave the US if the primary F-1 student visa holder loses status or completes their program.
-
-\chapter{Domestic Students}\label{sec:domestic-students}
-
-\epigraph{\vspace{-0.2in} I’m not a bad guy! I work hard, and I love my kids. So why should I spend half my Sunday hearing about how I’m going to Hell?}{\textsc{The Simpsons}}
-
-Most of what is written in this document applies to both domestic\footnote{For simplicity domestic means you did your undergrad (or MS) at a US university.} and international students.
-However, there are some differences and benefits that domestic students should be aware of and can leverage to improve their chances of admission.
-
-\paragraph{Standing out \autoref{sec:improve-your-chance}} There are \emph{few} domestic applications compared to international ones, i.e., domestic students are the minority (\autoref{sec:urm}) in the CS PhD application pool.
-Many US universities want to increase the number of domestic students in their programs (and as mentioned later, there are specific fellowships and funding for domestic students).
-That makes you stand out and can help your case.
-
-\paragraph{Fee Waiving} Some schools might offer application fee waivers for domestic students.  You should check with the school you're applying to.
-
-\paragraph{School \autoref{sec:your-school}} Adcom already knows about your school, which is a plus (see \autoref{sec:your-school}).
-You are also more familiar with the US education system and culture (so hopefully you are aware of the cultural issues listed in \autoref{sec:cultural}).
-
-\paragraph{Standard Tests \autoref{sec:standard-tests}} You do not need to take TOEFL or IELTS because you already did your undergrad (or MS) at a US university.
-You might also be more comfortable communicating in English, e.g., contacting professors (\autoref{sec:contact}).
-
-\paragraph{Transcripts} You do not need to get your transcripts evaluated/translated (which can be a hassle and inaccurate).
-You can just send your official transcripts directly to the school you're applying to.
-
-\paragraph{Funding \autoref{sec:funding}} You have more opportunities for funding, e.g., through government scholarships for US citizens and permanent residents.
-You can also apply to specific programs \emph{before} you start your PhD, e.g., NSF Graduate Research Fellowship Program (GRFP) and Hertz Foundation Fellowship.
-These fellowships are very competitive and can significantly improve your admission chances.
-
-\paragraph{Research Experiences \autoref{sec:research-opportunities}} You might have many opportunities to do research as an undergraduate, e.g., through REUs and internships at your undergrad university.
-Highlight such experiences in your application.
-
-\paragraph{Open House \autoref{sec:accepted}} It is easier for you to attend open house events in person.  This can help you make a better decision on which school to attend.
-
-% \begin{domesticbox}[Domestic students:] If you're a domestic student, you have several advantages in your application.
-You also have more opportunities for funding, e.g., through government scholarships for US citizens and residents.
-Finally,
-% \end{domesticbox}
-
-\chapter{Research Opportunities}\label{sec:research-opportunities}
-
-\epigraph{\vspace{-0.2in} Kids, you tried your best and you failed miserably. The lesson is: never try.}{\textsc{The Simpsons}}
-
-\begin{center}
-    \includegraphics[scale=0.5]{files/phd100404s.png}
-  \end{center}
-
-As discussed in \autoref{sec:research-opportunities}, research experience is a crucial component of a strong PhD application.
-It gives you opportunities to try out research, determine what research area you're interested in, publish papers (\autoref{sec:research-experiences}), connect with researchers, and get strong LoRs (\autoref{sec:lor}).
-
- This section provides some guidance on how to gain research experience, e.g., as an undergraduate student.  It also advises students from small colleges where research opportunities are limited.
-
-\paragraph{Locally} Start looking for research opportunities at your institution.
-If you did well or liked a class, you can check with the professor of that class for research opportunities.
-You can also go to the department directory and then professors' websites and see if they are looking for undergraduate researchers.
-Even if they say they are looking for graduate students, you can still contact them and ask.
-
-There might be also special programs for undergraduate research from the university, e.g., Mason has the OSCAR program, UNL has UCARE, and NSF has REUs (for US citizens and permanent residents).
-
-You can also take honor thesis or independent study courses with a professor.
-This is a good way to get research experience and also get a LoR from the professor.
-You can also ask your academic adviser or other faculty members for suggestions.
-Finally, you can also ask your peers who are already doing research.
-Use the method described in \autoref{sec:contact} to contact professors
-
-\begin{commentbox}[Vu:]
-    I enjoy working with undergrads (e.g., have mentored over 10 undergrads) and always actively recruit.
-    I often get undergraduate researchers through my classes, e.g., asking students who did well in my class if they are interested in research.
-Occasionally I was introduced to students by other students or faculty.
-
-    While most undergrads are understandably not super productive in research, some are and I have published multiple papers with them.
-I also have written strong LoRs for them and have helped them get into PhD programs, including top ones.
-
-    \tcblower
-    At Mason we pay undergrads about \$15/hr and about 10 hrs/week. This is flexible and can be increased as needed.
-\end{commentbox}
-
-\paragraph{Online Research Communities and Open Source Contributions} Places such as GitHub offer many research opportunities, even though they might not appear so (e.g., they do not explicitly mention research).
-In many cases, professors or research labs put their projects on GitHub.
-For example, my research group, \href{https://github.com/dynaroars/}{Dynaroars}, has many open-source projects that undergrads can contribute to.
-
-By contributing code, fixing bugs, implementing new features, or providing documentation, students can gain practical research experience and interact with experienced developers and researchers in the field.
-Not only you gain research experience, but you might be able to get a LoR from the project maintainer, and you should write about this experience in your SOP (\autoref{sec:research-statement}.)
-
-\paragraph{Virtual Research Opportunities} This is less common but several places offer virtual internships and research programs aimed at providing hands-on research experience.
-These programs often involve working remotely under the guidance of experienced mentors and collaborating with a team of fellow researchers.
-For instance, \href{https://docs.google.com/forms/d/1btIwt4HwjyKMOUk-EMy3rbkfWzFxv2lNrMm_zkd0pA4/viewform?edit_requested=true}{UIUC+ Summer Undergraduate Research in Software Engineering}  offers an unpaid remote internship for software engineering students all over the world to collaborate with mentors from University of Illinois at Urbana Champaign.
-
-% \paragraph{Online Conferences and Workshops}: Attending online conferences and workshops is another way to gain exposure to cutting-edge research and establish connections with experts in the field.
-Many conferences now provide virtual participation options, enabling students to access research talks, poster sessions, and panel discussions and sometimes access designated chat rooms or networking events where participants can engage with researchers, ask questions, and seek potential research collaborations.
-By creating profiles, joining relevant research groups, and participating in discussions, students can connect with established researchers, contribute to ongoing projects, and potentially collaborate on publications or research proposals as they provide access
-
-\chapter{Cultural Differences}\label{sec:cultural}
-
-\epigraph{\vspace{-0.2in} Ay caramba!}{\textsc{The Simpsons}}
-
-This section lists some cultural issues that students, especially international ones, might want to pay attention to. These issues apply in general and not just PhD admission.
-
-% \paragraph{Diversity} US universities prioritize diversity and inclusion.
-Students need to respect and appreciate varied opinions, backgrounds, and cultures.
-Unlike some countries where certain voices are marginalized, in the US, all perspectives are valued equally (especially at universities).
-Racism or discrimination will have serious consequences, including academic and disciplinary actions.
-
-\section{Checking Status, Accepting, Postponing, and Decline Offers}\label{sec:accept-postpone-decline}
-
-Students often ask about what to do after they get an interview or an offer from a professor, e.g., if they can followup to find out about their status, or is it OK to postpone or accept/reject offers?, and most importantly, how to do so without offending anyone.
-
-\paragraph{Checking your application status and Follow-up emails} If you have interviewed and not heard back from a professor after a few weeks or especially around the time when universities send out their admission decisions (around late Feb-- mid-Mar), you can email to check.
-You can follow up the interview invitation and say: \emph{``Thanks for chatting with me.
-I am very excited about the opportunity to work with you.
-Could you please let me know if you have made a decision or if you need more information from me?''}.
-If you have new updates, e.g., new publications or new fellowship awards, or even new offers from other professors or schools, you can also mention that.
-
-Profs.
-are often very busy, especially during admission time when they have to many reviews and interviews.
-They might not have time to respond to every email.
-If you do not hear back after a week, you can send another email to check again.
-If you still do not hear back, you can assume that you are not selected.
-
-\paragraph{Accepting an offer} If you decide to accept an offer, you can say: \emph{``Thank you for the offer.
-I would like to accept it and look forward to working with you.
-Could you please send me more details about the offer and what to do next?''}.
-The prof.
-will likely send you more details about the offer and what to do next.
-If you decide to accept an offer, do so quickly.
-
-\paragraph{Postponing an offer} If you need more time to decide, you can ask for more time: \emph{``Thank you for the offer.
-I am very excited about it.
-However, I am still waiting for other offers and need more time to decide.
-Would it be possible to postpone the decision for a few weeks?''}.
-This is perfectly fine and professors will understand and might even appreciate your honesty.
-They will likely give you a few weeks to decide.
-If you need more time, you can ask for more time.
-But do not ask for too much time, e.g., more than a month.
-You also should not postpone the offer multiple times, which will annoy people.
-
-\paragraph{Declining an offer} If you decide to decline or reject an offer, you can say: \emph{``Thank you for the offer.
-However, I have decided to accept another offer.
-I appreciate your time and consideration.
-I hope we can work together in the future.''}  Professors will understand and wish you luck.
-If you decide to reject an offer, do so quickly.
-
-\paragraph{Accepting an offer and later rejecting it}
-
-I've seen many students, especially international, face a dilemma when they \emph{commit} to a graduate offer but then receive a better one.
-Advice given in online forums is often along the line that it's okay to switch, using reasons like you haven't yet had a strong relationship with the prof.
-or you should prioritize your personal benefit.
-
-In my opinion, these reasons are not strong enough to justify retracting an acceptance.
-A more valid reason is using the \href{https://cgsnet.org/wp-content/uploads/2024/01/CGS_April15_Resolution_Jan312024.pdf}{April 15 resolution}, in which many universities participate.
-Among various things, this resolution states that students are free to accept a new offer from a different institution until 4/15, even if they have already accepted an offer elsewhere.
-
-However, in general, retracting an acceptance can have ethical implications.
-When you accept an offer, you are committing to work with that prof, who then might stop looking for other students.
-So by retracting your offer, you are breaking your commitment and also causing a great deal of inconvenience to the prof and also taking away the opportunity from other students.
-Ultimately, this choice is personal, involving a balance between personal benefit and the ethical impact on others.
-
-\section{Academic Integrity (Cheating and Plagiarism)}
-
-Plagiarism and cheating (e.g., exams and assignments) are a BIG no-no in the US.
-If you're caught cheating, you will face serious consequences and likely be expelled from the university (e.g., after the second time at Mason).
-This is quite different from many international countries where cheating is common and often tolerated.
-Faculty is extremely good at detecting cheating (we have been dealing with these situations so many times over so many years), and \emph{will} report cheating cases.
-In short, whatever you do, don't cheat---not worth it.
-
-Here are the typical steps: (i) a faculty suspecting a cheating case will report it to the Office of Academic Integrity (OAI) at the university---the report often has supporting evidence and suggested penalty (e.g., a failing grade);  (ii)  OAI will completely take over and investigate the case independently (the prof.
-is not involved); and (iii) OAI will make the final decision.
-It is important to note that after receiving the report from your prof., OAI \emph{completely} takes over and makes its own decision.
-This means begging your professor will not help because they simply are no longer involved in the case and cannot do much.
-
-\section{Illegal Software}\label{sec:illegal-software}
- Using illegal/cracked software is very common in many countries (and even in the US).
-However, \emph{do not} install or use them on university computers, even those given to you by your adviser.
-It is unlikely that the university will track you down, but it is the \emph{software company} that will.
-They have very sophisticated tools to detect illegal software and will sue your university/department.
-Imagine your department or adviser being sued for a large sum of money, and it is \emph{you} who caused it.
-If you need to purchase software,  ask your adviser or the department (\autoref{sec:buying-equipment}).
-
-\section{Costly Gifts}\label{sec:gifts} In many countries, it is customary to give professors costly gifts (e.g., expensive liquors, jewelry, or even money), often during the holidays, to show respect and appreciation.
-In the US, this is can be considered \emph{inappropriate} and discouraged.
-
-However, if you'd like to offer small souvenir-like tokens, it's a thoughtful gesture that's appreciated.
-Some professors proudly display their gifts, which can come from students and colleagues (e.g., when they travel to their home countries or conferences).
-In summary, small gifts are fine, but avoid anything that might make your professors uncomfortable.
-
-\section{Relationships}
-There's a misconception that in the US it's all business, with professors as bosses who pay students for their work and that lab mates are just work colleagues; and that doing nice things means expecting something in return.
-
-However, the reality is quite the opposite. While people can be straightforward and appear ``cold'', they are also informal, friendly, and very caring (in ways that might surprise you).
-With lab mates and colleagues, you will often work and go to lunch together, confide in each other, help each other navigate the academic journey, and often become lifelong friends.
-With your professor(s), you can call them by their first name (\autoref{sec:address}), disagree with them and argue (and gain respect doing so), seek their help (even on personal matters), come to their houses for parties or gathering (e.g., my lab always come to my house for \href{https://photos.app.goo.gl/LFtbqQUuznq9eiL7A}{Thanksgiving}), and give them small thoughtful gifts that they put on their desks (\autoref{sec:gifts}).
-Many people maintain lifelong relationships with their professors and colleagues, staying in touch through cards, emails, and visits, even after they no longer work together.
-
-\begin{commentbox}[Vu:]
-    I maintain a close relationship with my former professors and mentors.
-When there is a new event in my life (or theirs), I often email them or call them, e.g., when I get married, have a new baby, new job, etc.
-When I was having issues at UNL and considering leaving, I talked to my PhD advisers Deepak and Steph.
-I think this does not bother them a bit; they are genuinely interested in knowing and helping solve these ``dramas'' in my life.
-
-    I also visit my former professors when I am in their area.
-I meet Thang Bui (my MS adviser) at least once a year when I come back to Harrisburg to visit my parents.
-When Steph was in DC for a meeting, I invited her to to give a research talk at Mason.
-I have also collaborated with them after I graduated e.g., I am working with Deepak on an NSF proposal on invariant generation.
-
-    In short, while I am a bit closer to my former advisers and mentors than most people (e.g., I still keep in touch with my middle school teacher in Hawaii), it is always a good idea to maintain a good relationship with people who have helped or worked well with you.
-A simple, short email or text once in a while (e.g., a \emph{``Hi X, I heard you just got promoted ...
-Congrats!''}) would suffice.
-They will appreciate it, and you never know when you might need their help.
-
-  \end{commentbox}
-
-% \begin{commentbox}
-%   [DK:] Here are a few items you may consider addressing: putting international students in touch with other students from their countries, assuring them that they would be picked up from airports upon their arrival and that their initial stay will be taken care of.
-Most universities have other resources for these, but it is worth mentioning that they would be taken care of upon arrival and can get help during the transition phase.
-Learning to cook was a big deal when I arrived over 50 years ago-August 1973.
-But things have changed as one can find many ethnic food places, a big contrast from 1973, when there were two Indian restaurants in Greater Boston area.
-% \end{commentbox}
-
-\chapter{Rankings of CS PhD programs}\label{sec:ranking}
-
-\epigraph{\vspace{-0.2in} The whole damn system is wrong!}{\textsc{The Simpsons}}
-
-\begin{table}[h]
-  \centering
-  \small
-  \caption{The top 50 CS programs in the US from \href{https://www.csrankings.org}{CSRankings.org}, a ranking system  based on faculty publications at top CS conferences.
-(\href{https://csrankings.org}{CSRankings}, Jan.
-2024).
-\red{$^*$} indicates that the university has \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{Vietnamese prof.} that can advise CS PhD students.}\label{tab:ranking}
-  \begin{tabular}{rl|rl}
-    \toprule
-    1 & Carnegie Mellon & 26 & Duke University \\
-    2 & Univ. of Illinois at Urbana-Champaign\red{$^*$}  & 27 & Univ. of California - Santa Barbara \\
-    3 & Univ. of California-San Diego & 28 & Rutgers University\red{$^*$} \\
-    4 & Georgia Institute of Technology & 29 & Univ. of California - Riverside\\
-    5 & MIT                            & 30 & Pennsylvania State University  \\
-    6 & Univ. of California - Berkeley& 31 & Northwestern University\\
-    7 & University of Michigan - Ann Arbor\red{$^*$}   & 32& George Mason University\red{$^*$}\\
-    8 & University of Washington      &33 &  Texas A\&M University\red{$^*$} \\
-    9 &  Stanford University  &34&  University of Utah \\
-    10 & Cornell University  & 35 &  North Carolina State University\\\
-    11 & University of Maryland - College Park &  36& Ohio State \\
-    12 & Northeastern University\red{$^*$} &37& University of Virginia  \\
-    13 & Purdue University &38& Yale University \\
-    14 & University of Wisconsin - Madison\red{$^*$} &39 & Univ. of California - Santa Cruz \\
-    15 & University of Texas at Austin &40& Brown University \\
-    16 & University of Pennsylvania &41 & Harvard University \\
-    17 & Columbia University\red{$^*$} &42 & Boston University  \\
-    18 & Princeton University\red{$^*$}  & 43& Rice University\\
-    19 & New York University  & 44&  University at Buffalo\red{$^*$}\\
-    20 &  University of Massachusetts-Amherst\red{$^*$} &45& University of Colorado-Boulder \\
-    21 & Univ. of California - Los Angeles &46& University of Illinois at Chicago  \\
-    22 & University of Southern California &47& University of Minnesota \\
-    23 & University of Chicago &48& University of North Carolina\red{$^*$} \\
-    24 & Stony Brook University\red{$^*$} &49& Arizona State University\red{$^*$} \\
-    25 &  Univ. of California - Irvine&50& Univ. Of California - Davis \\
-    \bottomrule
-  \end{tabular}
-\end{table}
-% Virginia Tech\red{$^*$}  \\
-
-%\autoref{tab:ranking} lists .
-
-\chapter*{History and Acknowledgement}\label{sec:ack}
-
-\paragraph{History} This document was conceived during a lunch with a faculty at Mason.
-We talked about why Mason was not able to attract good Vietnamese and other international students, despite having a much stronger CS program than many schools that these students want to go to (part of the reason is described in \autoref{sec:selecting-ranking-schools}).
-We wish there were a way for international students to know about PhD programs in the US.
-
-I was also a member of the large VietPhD group on Facebook and often browse Internet forums (e.g., \texttt{Reddit/gradadmission} and \texttt{GradCafe}).
-There I saw many questions from students about PhD programs.
-However, most participants are students in non-CS fields or not in the US, and their answers are not always accurate and can lead to more confusion.
-So I thought it would be useful to have a document that is specific to CS PhD programs in the US from an insider perspective.
-
-I started writing this document in May 2023 and have been updating it since then (mostly around deadline time when I tend to procrastinate, i.e., \emph{productive procrastination}!).
-The document was initially intended for international students but has been generalized to all students.
-I have put the source code of this document on \href{https://github.com/nguyenthanhvuh/phd-cs-us}{GitHub} so that others can contribute to it.
-
-\paragraph{About Me} I am an associate professor in the CS dept at George Mason University (Mason).
-Before Mason, I was at the University of Nebraska-Lincoln (UNL).
-I have been in the PhD admission process at Mason and UNL for many years.
-This allows me to have a good understanding of the process, the challenges students face, and what faculty are looking for.
-I also serve as the program director of the MS program in Software Engineering at Mason (thus also have experience with MS admission---which is quite different than PhD).
-My personal and lab website is at \href{https://dynaroars.github.io}{dynaroars.github.io}.
-
-Though I was not an international student, many of my students and collaborators are/were.
-I also mentor multiple students from Vietnam and have close colleagues and friends who were once international students.
-I hope to capture the diverse challenges and experiences they've faced in this document so that it can be a valuable resource for prospective international students.
-Finally, my upbringing in the US provides a perspective aligned with American culture, allowing me to shed light on various issues, particularly those related to cultural differences (\autoref{sec:cultural}).
-
-\paragraph{Acknowledgement} Many people have contributed to this work.
-Profs. Craig Yu (Mason), Hakan Aydin (Mason), 
-Xiaokuan Zhang (Mason), Hung Le (UMass), and Deepak Kapur (UNM) provided valuable input in the early version.
-Other Mason faculty members also have provided feedback and contributions.
-Many students including Didier (Mason), Thanh (Melbourne), and Dat (Melbourne) have contributed valuable questions and feedback.
-
-Also thanks to NSF for encouraging faculty to be creative in research and education, which allows me to write this document.
-
-\textbf{Thank you!}
-
-%\bibliographystyle{abbrv}
-%\bibliography{demystify.bib}
+\bibliographystyle{abbrv}
+\bibliography{demystify.bib}
 

### 2024-08-10 faculty.tex
updates

@@
 
-\chapter{Research}
-This section focuses on paper publication, which is a major component in getting your name out there and of course in your departmental evaluation.
-
-\section{What is needed for tenure?}
-
-Publications are a major component of tenure evaluation.
-There are no specific numbers of papers or which conferences/journals are required, but typically you need to have a good number of papers in top venues to get tenure.
-Also work that won awards (e.g., best paper awards or 10-year most influential awards) also count a lot in tenure evaluation.
-Citations are not required but work with very high citations would help.
+\chapter{Research}\label{chap:research}
+This section focuses on paper publication, which is a major component of P\&T and of course in getting your name out there.
+P\&T evaluation requires no specific conferences/journals or numbers of papers, but typically you will want to have a good number of papers in top venues.
+P\&T members or letter writers (obviously) are familiar with top places in your field, and they will definitely look for those in your profile.
+Work that won awards, e.g., paper awards, can help a lot in tenure evaluation.
+Papers with very high citations also help.
+Research areas that you started and are now popular also count.
+Finally, new research directions that can help get grants are also good.
+
+\section{Venues}
+
+Depending on your research area, you might focus on publishing at top conferences or journals.
+You probably should know what's popular in your area already (if not, ask your senior colleagues).
+If you're at an R1 university, you will want to publish at top venues, and not just any venue.
+You should also aim to publish at a variety of venues, and not just one.
+This will show that your research is of interest to a wide range of people.
+
+\paragraph{Conferences}
+In most CS fields, top conferences are more common and competitive than journals.
+Each field also has several of such top conferences yearly (e.g., in Software Engineering, ICSE, FSE, ASE, and ISSTA are top four).
+They are often scheduled so that if you miss (or get rejected) from one, you can submit to another.
+
+You probably will want to get at least about 2 papers at top places each year.
+This is an average and so you might have 3 one year and 1 the next.
+Papers with your students are likely more preferred than papers with your peers (and their students) because it shows that you are mentoring your students well and that the project is lead by your group.
+
+\paragraph{Artifact Evaluation}
+
+\paragraph{Journals}
+In some journals are more common and prestigious.
+Moreover, many areas have top journals in additions to well-known conferences (e.g., Trans.
+of Software Engr.
+or TSE is a very well-known journal in Software Engineering, which has 4 major confs.).
+Typically, or at least in SE, the common approach is publishing new work to conference and followup work to journals (e.g., so the journal paper is an substantial extension of the conf.
+paper with new experiments, studies, theorems, etc).
+In some sense, journals are for more mature work, and conferences are for new ideas.
+
+\paragraph{Short papers, workshops, etc}
+
+\section{Collaboration and co-authorship}
+While it's general true that profs.
+are busy, but for me and probably many others, it's not the main reason.
+I do skim every single emails from prospective students, but I ignore most of them because they were quite generic (likely written for dozen of other profs.).
+Sometimes they try to "customize" it by mentioning the names of my papers, but
+
+.  So even if I am not busy (e.g., spending time on reddit), I still would not reply to such emails.
+
+\paragraph{Papers with your adviser(s)}
 
@@
 
-\chapter{Obtaining Funding}
-Getting funding is arguably the most important part of getting tenure.
-Without funding, you can't pay your students, conference registration and travel, or even your own summer salary.
-However, writing grant proposals is very different from writing research papers.
-Thus, most new professors struggle with grant writing at first, but eventually get better at it.
-
-%Like most new professors, I struggled big time grant writing at first, but eventually got better at it.
-
-%, which is very different compared to research papers.   In this chapter, we will discuss the various ways to obtain funding.
-
-\section{For P\&T}
+\chapter{Obtaining Funding}\label{chap}
 Without a doubt, grants are one of the most, if not \emph{the} most, important factors in P\&T (or any kind of performance) evaluation.
-The most common reason for not getting tenure is not having enough funding.
-In an ideal world, research results, e.g., publications, would be the more important; but in reality, funding is often considered more important for tenure evaluation at most R1 universities.
+The most common reason for not getting tenure is the lack of funding.
+Ideally research results, e.g., publications (\autoref{chap:research}), would be the more important thing for a researcher; but in reality, funding is often considered more important for tenure evaluation at most R1 universities.
 The amount of funding you have is a direct reflection of your research impact and productivity.
@@
 Finally, you should aim to get funding from a variety of sources, including NSF, Defense, industry, and internal grants.
-This will show that you are able to work with different types of sponsors and that your research is of interest to a wide range of people.
-
-\section{Types of Funding}
+This shows that you can work with different types of sponsors and that your research is of interest to a wide range of people.
+
+\section{Types of Funding}\label{sec:fuding-types}
+There are several types of funding sources, including external funding (e.g., NSF, DARPA), internal funding (e.g., from your university), and industrial gifts (e.g., from Amazon or Google).
+Each type of funding has its own pros and cons, and you should aim to get funding from a variety of sources.
+\subsection{External Funding}\label{sec:external}
 \paragraph{NSF} NSF is the obvious funding source for CS. 
@@
 
-However, these sponsors are good sources of funding and can provide the additional resources (e.g., your "3rd" non-NSF summer month).
+However, these sponsors are good sources of funding and can provide the additional resources (e.g., your ``3rd'' non-NSF summer month).
 Note that some defense programs, e.g., Young Investigator Award, are more like NSF grants, with less stringent requirements and no deliverables required.
 
-\subsection{Internal Funding}
-These are funds from your university, city, or state (e.g., Virginia has the CCI program that provide ``seed'' grants for researchers from universities in Virginia).
+\subsection{Internal Funding}\label{sec:internal}
+These are funds from your university, city, or state (e.g., Virginia has the CCI program that provide ``seed'' grants for researchers from public universities in Virginia).
 Your ``startup'' package is also be considered as a large funding from an internal source (your university) to kickstart your research.
@@
 While some are connection-based (e.g., for your ongoing work with a collaborator at Facebook), most are based on the merits of your research.
-These are highly competitive and prestigious (e.g., the solicitation is open worldwide, and you compete with top researchers from top places).  
+While such unrestricted gifts are not a lot (e.g., \$100K), they are highly competitive and prestigious (e.g., the solicitation is open worldwide, and you compete with top researchers from top places).
 They also show that your research has values and impact beyond academia and is interested to industry. 
-Practically, they are the best kind of funding for three reasons: (1) indirect costs or overheads are typically not levied on these gifts (so you get most of the money), (2) there is no associated timeline by when you should spend them (they give it to you and forget about it), and (3) you can use them for a broader range of things than university and grant/contract funding (e.g., summary salary, equipments, or even bowling events for your group).
+They are also very flexible (hence the term unrestricted):  (1) indirect costs or overheads are typically not levied on these gifts (so you get most of the money), (2) there is no associated timeline by when you should spend them (they give it to you and forget about it), and (3) you can use them for a broader range of things than university and grant/contract funding (e.g., summary salary, equipments, or even bowling events for your group).
 

### 2024-08-10 nnv.tex
updates

@@
 
-\chapter{Summary}\label{sec:summary}
+\chapter{Basic of Neural Network}\label{sec:basic}
+
+A \emph{neural network} (\textbf{NN})~\cite{Goodfellow-et-al-2016} consists of an input layer, multiple hidden layers, and an output layer.
+Each layer has a number of neurons, each connected to neurons from previous layers through a predefined set of weights (derived by training the network with data).
+A \textbf{DNN} is an NN with at least two hidden layers.
+
+%A deep neural network consists of three types of layers: an input layer, multiple hidden layers, and an output layer.
+Each layer consists of a number of neurons, each connected to neurons from another layers through a predefined set of weights (derived by training the network with data).
+A network is called a \emph{fully connected feed-forward} neural network (FNN) if it connects every neuron in a layer (i.e., all weights are non-zero) to a neuron in the next layer.
+
+The output of a DNN is obtained by iteratively computing  the  values  of  neurons  in  each  layer.
+The value of a neuron in the input layer is the input data.
+The value of a neuron in the hidden layers is computed by applying an \emph{affine transformation} to values of neurons in the previous layers, then followed by an \emph{activation function} such as the popular Rectified Linear Unit (ReLU) activation.
+
+For this activation, the value of a hidden neuron \(y\) is 
+$ReLU(w_1v_1 + \dots{} + w_nv_n + b)$, where \(b\) is the bias parameter of \(y\), \(w_i, \dots, w_n\) are the weights of \(y\), \(v_1,\dots,v_n\) are the neuron values of preceding layer, \(w_1v_1 + \dots + w_nv_n+b\) is the affine transformation, and \(ReLU(x) = \max(x,0)\) is the ReLU activation.
+The values of a neuron in the output layer is evaluated similarly but it may skip the activation function.
+A ReLU activated neuron is said to be \emph{active} if its input value is greater than
+zero and \emph{inactive} otherwise.
+
+\paragraph{DNN Verification} Given a DNN \(N\) and a property $\phi$, the \emph{DNN verification problem} asks if $\phi$ is a valid property of $N$.
+Typically, $\phi$ is a formula of the form $\phi_{in} \Rightarrow \phi_{out}$, where $\phi_{in}$ is a property over the inputs of $N$ and $\phi_{out}$ is a property over the outputs of $N$.
+%This form of properties has been used to encode safety and security requirements of DNNs, e.g., safety specifications to avoid collision in unmanned aircraft~\cite{kochenderfer2012next} and \emph{adversarial robustness}~\cite{katz2017towards} properties desired by all DNNs, in which a small input perturbation does not cause major spikes in the DNN's outputs.
+A DNN verifier attempts to find a \emph{counterexample} input to $N$ that satisfies $\phi_{in}$ but violates $\phi_{out}$.
+If no such counterexample exists, $\phi$ is a valid property of $N$.
+Otherwise, $\phi$ is not valid and the counterexample can be used to retrain or debug the DNN~\cite{huang2017safety}.
+
+% Verification tool such as Marabou and nnenum are then applied to the network to prove that the network is safe or identifier counterexample representing small input differences causing large output changes.
+
+% \footnote{This is encoded as the differences of the inputs being within a certain small range  ($\phi_{in}$) implies the differences of the outputs still fall within a certain range in $\phi_{out}$)}.
+
+\begin{wrapfigure}{r}{0.3\textwidth}
+    \centering
+    \includegraphics[width=1\linewidth]{figure/dnn.pdf}
+    \caption{\label{fig:dnn} An FNN with ReLU.}
+  \end{wrapfigure}
+\paragraph{Example} Fig.~\ref{fig:dnn} shows a simple DNN with two inputs $x_1,x_2$, two hidden neurons $x_3,x_4$, and one output $x_5$.
+The weights of a neuron are shown on its incoming edges , and the bias is shown above or below each neuron.
+The outputs of the hidden neurons  are computed the affine transformation and ReLU, e.g., $x_3 = ReLU(-0.5x_1+0.5x_2+1.0)$.
+The output neuron is computed with just the affine transformation, i.e., $x_5=-x_3+x_4-1$.
+
+A valid property for this DNN is that the output is $x_5 \le 0$ for any inputs $x_1 \in [-1,1], x_2\in[-2,2]$. An invalid property for this network is that $x_5 > 0$ for those similar inputs.
+A counterexample showing this property violation is $\{x_1=-1, x_2=2\}$, from which the network evaluates to $x_5=-3.5$.
+Such properties can capture \emph{safety requirements} (e.g., a rule in an  collision avoidance system in~\cite{kochenderfer2012next,katz2017reluplex} is ``if the intruder is distant and significantly slower than us, then we stay below a certain threshold'') or \emph{local robustness}~\cite{katz2017towards} conditions (a form of adversarial robustness stating that small perturbations of a given input all yield the same output).
+
+%Fig.~\ref{fig:dnn}b shows the same network but with each hidden neuron $x$ split into two neurons $x'$ and $x''$ representing the result of affine transformation on $x$ and ReLU activation on $x'$, respectively, e.g.,  $x_3'=-x_1-0.5x_2-1.0$ and  $x_3'' = ReLU(x_3')$.
+This ReLU-slitting representation is adopted by \tool{} and other DNN analyses (e.g., ~\cite{katz2017reluplex,wang2018efficient,henriksen2020efficient})
+% including Reluplex~\cite{katz2017reluplex}, Neurify~\cite{wang2018efficient}, and VeriNet~\cite{henriksen2020efficient},
+%because it does not change the semantics or complexity of the problem and is easier to reason about as we will show in \S\ref{sec:overview}.
+
+\section{Affine Transformation}
+
+\section{Activation Function}
+
+\section{Properties of Neural Networks}
+
+\subsection{Robustness}
+
+\subsection{Safety}
+
+\chapter{Verification of Neural Networks}\label{sec:verification}
+
+\section{Complexity}
+
+\chapter{Search Algorithms}
+
+\chapter{Constraint Solving}
+\section{SMT}
+\section{MILP}
+\chapter{Abstraction}
+
+\section{Interval}
+\section{Zotope}
+\section{Polytope}
+
+\chapter{Popular Techniques and Tools}
+\chapter{Verifying the Verifiers}
+\chapter{Conclusion}
 

### 2024-08-10 demystify.tex
updates

--- 2024-08-02 demystify.tex
@@
 \usepackage{graphicx}
-\usepackage{pdfpages}
 \usepackage{hyperref}
@@
 \begin{mybox}
-This document will be updated regularly to reflect the latest information and updates in the admission process. Its latest version is available at
+The latest version of this document is available at
 
@@
 A good LoR letter should give details of the applicant's research background and achievements, all of which paint a compelling picture of the candidate to the adcom.
+
 Having a strong letter from an internationally recognized researcher will \emph{greatly strengthen} your application.
@@
 
-\paragraph{Why Self-written Letters are Bad?} Many international students write letters themselves, typically due to the requests of their profs., and have them signed by their profs.
-Such letters have \textbf{very little value} and are considered weak by reviewers---why can you not even find someone who cares or knows enough about you to write a candid personal reference letter?
+\paragraph{Self-written Letters} Many international students write letters themselves, typically due to the requests of their profs., and have them signed by their profs.
+Such letters have little value and are considered weak by reviewers---why can you not even find someone who cares or knows enough about you to write a candid personal reference letter?
 Instead of the reference writer talking about you, in this case it is you who write about yourself (and they just sign the letter).
-Experienced reviewers can easily spot (or suspect) such letters, which can be a \red{red flag}.
-
-\paragraph{Why Generic Letters are Bad?} When the writers do not know much about the applicants (e.g., just taking some course with them or not making any impression to write about), they might write a generic and short letter, which is not useful and also considered weak.
-This does not mean the professor is not good or does not care about you, but they just do not know you well enough to write a strong letter.
-So it might be a good idea to directly ask if the prof.
+Moreover, experienced reviewers can easily spot (or suspect) such letters, which can be a \red{red flag}.
+
+\paragraph{Helping your LOR writer} As mentioned above, do not write your own letter.
+Instead, you can help your LOR writer by providing details or unique things about yourself.
+For example, let them know about your GPA, research and work experiences, papers (if any), or anything you want them to mention.
+You can also provide them with a draft of your SOP so that they can see what you are saying about yourself and complement that with their own perspective.
+Sometimes your writer will explicit ask you for such information, but if not, you should provide it anyway (especially if you have not interacted with them much or have not done much research with them).
+
+\paragraph{Generic Letters} When the writers do not know much about the applicants (e.g., just taking some course with them or not making any impression to write about), they might write a generic and short letter, which is not useful and also considered weak.
+It might be a good idea to directly ask if the prof.
 is willing to write a \emph{strong} letter for you.
@@
 
-Sometimes students would go to great lengths just to get letters from well-known professors in their school (e.g., department head).
+Sometimes students would go to great length just to get letters from well-known professors in their school (e.g., department head).
 But as mentioned, if these professors do not know you, their letters would likely be generic and carry little value (sometimes \red{red flags}).
@@
 
-\paragraph{Why You Want to Help your LOR Writer} As mentioned above, do not write your own letter and generic letters do not give much value.
-One thing you can do to help your writer to write a strong, customized LoR for you is providing details or unique things about yourself.
-For example, let them know about your GPA, research and work experiences, papers (if any), or anything you want them to mention.
-You can also provide them with a draft of your SOP so that they can see what you are saying about yourself and complement that with their own perspective.
-Sometimes your writer will explicitly ask you for such information, but if not, you should provide it anyway (especially if you have not interacted with them much or have not done much research with them).
-
-\paragraph{Why You Need to Remind Your Writers} After you submit your application, you should tell your writers about that and let them know they will soon receive an email from the university to submit their letters.
-You should also remind them when you submit your application and ask them to submit their letters on time if they haven't done so.
+\paragraph{Reminding Writers} After you submit your application, you should tell your writers about that and let them know they will soon receive an email from the university to submit their letters.
+You should also remind them about the deadline and ask them to submit their letters on time.
+But do not send too many reminders as that can be annoying to the writers.
+
 Note that most places only have deadlines for the applicant, but are very flexible with the letter writers (in many cases do not even give them any deadline).
 Also, many places do not begin the admission review process right after the deadline and work on application reviews in the next semester (mid-January).
-
-Thus you do want to send reminders because professors can be quite busy and might forget to submit their letters, especially when there is no explicit deadline.
-However, do not send too many reminders as that can be annoying to the writers.
 
@@
 \begin{center}
-  \includegraphics[scale=0.5]{files/c2.png}
+  \includegraphics[scale=0.4]{files/c2.png}
 \end{center}
@@
 For example, if you say you have experience with teaching, then show what you did, e.g., undergrad TA or mentoring someone.
-If you say you work on a project, then show some results, e.g., paper submitted (or even rejected), achieved certain performance improvement over the state of the art.
+If you say you work on a project, then show some results, e.g., paper submitted (or even rejected), achieved X\% performance improvement over ...
 Also, do not use more than two pages (references can be on the third page).
@@
 
-\epigraph{\vspace{-0.2in} The answer to life's problems aren't at the bottom of a bottle. Heh heh! They're on TV!}{\textsc{The Simpsons}}
-
 Choosing a school and an adviser is clearly among the most important things in your mind when you apply and especially when you get admitted.
@@
 \begin{center}
-  \includegraphics[scale=0.6]{files/c1.png}
+  \includegraphics[scale=0.5]{files/c1.png}
 \end{center}
 
-International students often put schools into two bins: (i) top schools that they dream about, and (ii) everything else.
+International students often put schools into two bins: (i) very top schools that they dream about, and (ii) everything else.
 They commonly use rankings from US News, which is not transparent and questionable\footnote{\url{https://cra.org/cra-statement-us-news-world-report-rankings-computer-science-universities/}}.
 Sometimes they evaluate based on the reputation of the school's undergrad program or the reputation of the school's non-CS programs such as medical, math, or physics.
-They even rank universities based on popular places they know in the US, e.g., California, Texas, and New York.
-
-Instead of these superficial criteria, you should consider the CS program and the research interests of faculty members.
-You can learn about these using resources such as \href{https://csrankings.org}{CSRankings.org}.
+They even rank universities based on popular states they know in the US, e.g., California, Texas, and New York.
+
+You can learn about CS programs and the research interests of faculty using resources such as \href{https://csrankings.org}{CSRankings.org}.
 You will be very surprised to learn that a school that you didn't know much about can have very strong research in your topic (and vice versa, a school you thought highly about might have no faculty working in the research field you're interested in).
@@
 
-\subsection{PhD in other related fields: CE, IST, Cybersecurity}\label{sec:related-fields}
-
-You \emph{do not} need to do a PhD in CS to work in your area of interest.
-For example, in addition to a traditional CS department, Mason has IST and Cybersecurity departments, both of which have faculty with PhD in CS and work on CS topics (e.g., AI, Security, Robotics).
-So it is possible that you still get to do CS research and publish in CS-focused venues even if you're not in a traditional CS program.
-It is common to see faculty with PhD in CS in a non-CS department as well as faculty with non-CS PhD in a CS department.
-
-However, if your goal is a PhD in CS, then you likely need to be in the CS dept \emph{and} advised by a CS faculty.
-A non-CS faculty can serve in the PhD dissertation committee (common) or \emph{co-advise} (somewhat rare), but your main PhD adviser will likely be a tenure-line faculty in CS (\autoref{sec:faculty-types}).
-For example, a prof. in Stats or Math might be able to serve as a co-adviser, but not as a sole adviser of a student in a CS PhD program. 
-If in doubt, check with the CS department for their requirements.
-
-For this specific reason,  CSRankings includes only faculty who can advise CS PhD students.
-I also have compiled a \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{list} of Vietnamese faculty who can advise PhD students.
-\autoref{sec:faculty-types} talks more about who can serve as your PhD adviser.
-
 \section{Choosing an Adviser}
@@
 \end{itemize}
-\subsection{Who can serve as a PhD adviser?}\label{sec:faculty-types}
+\subsection{Types of Faculty: Who can serve as a PhD adviser?}\label{sec:faculty-types}
 
@@
 
-Now that you know a bit about tenured and tenure-track faculty (\autoref{sec:faculty-types}).
-Which one should you choose as your adviser?
-The short answer is that either can be a good adviser, but they have different strengths and weaknesses.
-
-Tenure-track faculty, e.g., assistant professors, are likely young and active in research (they have to, to get tenure).
-Thus, they will devote more time to work with you and push you to do research and publish.
+Tenure-track faculty such as assistant professors are likely young and very active in research (they have to, to get tenure).
+Thus, they will likely have more time to work with you and push you to do research and publish.
 However, they may not have as much experience in managing students and may not have as much funding (yet).
@@
 
-\subsection{Things to avoid when cold emailing a professor}
-
-\begin{itemize}
-
-    \item \textbf{Not customizing the email}. These emails typically talk about the student and not why they are interested in the professor's work. See writing tips in \autoref{sec:contact}
-    \item \textbf{Sending the same email to multiple professors}.
-Sending emails to multiple profs.
-using BCC or even CC (where we can see all addresses the email is being sent to) will result in no reply or a very harsh one on how unprofessional this is.
-    \item \textbf{Too much information}.
-Keep your email short and to the point.
-If we are interested, we will ask for more information.
-2--3 paragraphs are enough.
-Personally I also do not want to see course transcript or CV in the first email.
-    \item \textbf{Not being professional}, e.g., do not call the professor by their first name in the first email.  Use Prof. Lastname or Dr. Lastname (\autoref{sec:address}).
-    \item \textbf{Not reading the professor's website} and asking questions that are already answered there.
-\end{itemize}
+\subsection{PhD in other related fields: CE, IST, Cybersecurity}\label{sec:related-fields}
+
+You \emph{do not} need to do a PhD in CS to work in your area of interest.
+For example, in addition to a traditional CS department, Mason has IST and Cybersecurity departments, both of which have faculty with PhD in CS and work on CS topics (e.g., AI, Security, Robotics).
+So it is possible that you still get to do CS research and publish in CS-focused venues even if you're not in a traditional CS program.
+It is common to see faculty with CS PhD in a non-CS department as well as faculty with non-CS PhD in a CS department.
+
+However, if your goal is a PhD in CS, then you likely need to be in the CS dept \emph{and} advised by a CS faculty.
+A non-CS faculty can serve in the PhD dissertation committee (common) or \emph{co-advise} (somewhat rare), but your main PhD adviser will likely be a tenure-line faculty in CS (\autoref{sec:faculty-types}).
+For example, a prof. in Stats or Math might be able to serve as a co-adviser, but not as a sole adviser of a student in a CS PhD program. 
+If in doubt, check with the CS department for their requirements.
+
+For this specific reason,  CSRankings includes only faculty who can advise CS PhD students.
+I also have compiled a \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{list} of Vietnamese faculty who can advise PhD students.
+\autoref{sec:faculty-types} talks more about who can serve as your PhD adviser.
 
@@
 
-\paragraph{Not a good fit and Bad luck}  You could have an excellent profile, but if you are interested in a research area that the program does not have, you might not be admitted.
-Similarly, if no faculty who is willing to advise you (e.g., they are on sabbatical or personal leave, do not have sufficient funding, or already have too many students), you will not be admitted.
-This is good for you, as you don't want to be in a program where no one can advise you.
+\paragraph{Not a Good fit and Bad luck}  You could have an excellent profile, but if you are interested in a research area that the program does not have, you will not be admitted.
+Similarly, if no faculty is willing to advise you (e.g., they are on sabbatical or personal leave, do not have sufficient funding, or already have too many students), you will not be admitted.
+This is actually good for you, as you don't want to be in a program where no one can advise you.
 
@@
 Students understandably get excited about their upcoming PhD journey and want to buy new computer equipment and electronics to prepare.
-However, you should first check with your professor.
+However, you should check with your prof.
+first.
 They might have funding to buy you computers and other equipment (e.g., software, books, keyboards, headphones, tablets, etc).
@@
 
-\chapter{Visa for International Students}\label{sec:visa}
-
-\epigraph{\vspace{-0.2in} I didn’t do it. Nobody saw me do it. You can’t prove anything!}{\textsc{The Simpsons}}
-
-% \begin{center}
-%     \includegraphics[scale=0.8]{files/visa.pdf}
-% \end{center}
-
-For international students pursuing a PhD in the US, \textbf{F-1} is the main visa that allows a student to study full-time at an accredited institution.  Here are some key points about the F-1 visa:
-
-\begin{itemize}
-\item \textbf{Employment:} You are allowed to work on-campus for up to \textbf{20 hours per week during the academic year} (because you still need to take classes) and \textbf{full-time} (typically 40 hrs) during official school breaks (e.g., summer and winter breaks).
-Off-campus employment requires authorization, which can be obtained through programs like Curricular Practical Training and Optional Practical Training described below.
-
-\item \textbf{Curricular Practical Training (CPT):} CPT allows you to participate in internships or practical training that is an integral part of their academic curriculum.
-CPT must be related to your field of study and can be full-time or part-time.
-It requires prior authorization from your university and must be completed before graduation.
-
-\item \textbf{Optional Practical Training (OPT):} OPT provides up to 12 months of work authorization for students before or after completing their degree.
-For students in STEM fields, there is an additional 24-month extension available.
-OPT requires prior application and approval from USCIS.
-
-\item \textbf{Full-time Enrollment:} You must maintain full-time enrollment status during the academic year.
-This means taking a minimum number of credits each semester, as defined by your program.
-Dropping below full-time status can result in loss of visa status.
-\end{itemize}
-
-\includepdf[pages=-]{files/visa.pdf}
-
-\section{For Spouses and Children}
-
-The spouses and children of F-1 visa holders can enter the US under the F-2 visa status. The F-2 visa allows family members to live in the US with the following conditions:
-
-For Spouse: 
-\begin{itemize}
-\item \textbf{Work Restrictions:} F-2 spouses are not permitted to work in the US.
-\item \textbf{Education:} They can study part-time but cannot enroll in full-time degree programs.
-\end{itemize}
-For Children:
-\begin{itemize}
-\item \textbf{Education:} F-2 children can attend K-12 schools but cannot pursue higher education full-time.
-\item \textbf{Work Restrictions:} Like F-2 spouses, children are not allowed to work under any circumstances.
-\end{itemize}
-F-2 visa holders must leave the US if the primary F-1 student visa holder loses status or completes their program.
+% \section{Additional Resources}
+% \begin{itemize}
+%   \item \bibentry{nlpphd}
+% \end{itemize}
+
+% \chapter{Additional Resources}
+
+% \chapter{Undergraduate Research Opportunities}
+
+% Prospective PhD students are often unaware of the importance of research experience in their application.
+This provides both concrete research experiences and potential LOR writers that can talk about your research ability.
+% % This is especially true for students from small colleges where research opportunities are limited.
+% This section aims to provide some guidance on how to gain research experience as an undergraduate student.
+%This section is written together Didier, a PhD student at Mason who did his undergraduate study in Rwanda.
+
+% \paragraph{Locally} There are many ways to find research opportunities right in your own institution.
+If you did well or like a class, you can check with the professor of that class for research opportunities.
+Sometimes professors would mention that they look for undergraduates on their websites, so do check those out!
+Even if they say they are looking for graduate students, you still should contact them and ask.
+You can also ask your academic adviser or other faculty members for suggestions.
+Finally, you can also ask your peers who are already doing research.
+You can use the method described in \autoref{sec:contact} to contact professors
+
+% In short, the most important thing is that you need to \emph{ask}!
+Don't be shy, don't be afraid of rejection, and don't be afraid of being ignored or that you're not good enough.
+Just ask!
+The worst thing that can happen is that you get a ``no'' or no response.
+But you will never know if you don't ask.
+And if you get a ``no'', you can always ask again later.
+There are students who got rejected the first time but got accepted the second time.
+So don't give up!
+
+% \paragraph{Online Research Communities and Open Source Contributions}:
+% Places such as Github offer many opportunities for research, even though they might not appear so (e.g., they do not explicitly mention research).
+Students can explore repositories related to their areas of interest, collaborate with other contributors, and actively engage in open source research projects.
+By contributing code, fixing bugs, implementing new features, or providing documentation, students can gain practical research experience and interact with experienced developers and researchers in the field.
+
+% \paragraph{Paid or Unpaid}  travel opportunities, publishing opportunities, etc
+
+% \section{Virtual Research Opportunities Beyond Physical Boundaries (written by Didier)}
+% \didi{In my experience, prospective PhD students (from Rwanda) are often unaware of the information provided in this document throughout their undergraduate.
+For this reason they often find themselves unprepared for PhD admission process by the time they finish undergraduate and have to spend a year or more to improve their application profiles.
+I think it would be beneficial to expand the document to also target students who are not yet ready for the application process to help them improve on all criteria considered during admission.
+I drafted the short essay on how to gain research experience remotely in order to improve their research profile and potentially get LOR from experts (hopefully faculty)}\\
+% In the realm of computer science, research experience plays a crucial role in securing admissions to top-tier PhD programs.
+However, students from underrepresented small colleges, both in the United States and internationally, often encounter limited research opportunities within their institutions.
+Thankfully, there exists an alternative pathway for these aspiring researchers to gain valuable research experience, even without being physically present at a university.
+
+% \textbf{Virtual Research Programs}:
+% Several universities, organizations, and research institutions offer virtual internships and research programs aimed at providing hands-on research experience.
+These programs often involve working remotely under the guidance of experienced mentors and collaborating with a team of fellow researchers.
+For instance, \href{https://docs.google.com/forms/d/1btIwt4HwjyKMOUk-EMy3rbkfWzFxv2lNrMm_zkd0pA4/viewform?edit_requested=true}{UIUC+ Summer Undergraduate Research in Software Engineering}  offers an unpaid remote internship for software engineering students all over the world to collaborate with mentors from University of Illinois at Urbana Champaign.
+% Virtual internships offer opportunities to contribute to ongoing research projects, conduct experiments, analyze data, and write research reports.
+Students can search for such programs through university websites, research institutes, or as of recently chatGPT (or Bard).
+
+% \textbf{Online Conferences and Workshops}:
+% Attending online conferences and workshops is another way to gain exposure to cutting-edge research and establish connections with experts in the field.
+Many conferences now provide virtual participation options, enabling students to access research talks, poster sessions, and panel discussions and sometimes access designated chat rooms or networking events where participants can engage with researchers, ask questions, and seek potential research collaborations.
+It is beneficial to also create profiles academic collaboration platforms such as ResearchGate.
+By creating profiles, joining relevant research groups, and participating in discussions, students can connect with established researchers, contribute to ongoing projects, and potentially collaborate on publications or research proposals as they provide access
+
+% In conclusion, for students lacking research opportunities at their small colleges, the virtual realm offers a wealth of possibilities to gain valuable research experience.
+Engaging with open source projects, joining online research communities, utilizing academic collaboration platforms, participating in virtual internships, and attending online conferences are all avenues to contribute to research, connect with experts, and enhance one's research profile.
 
@@
 
-\epigraph{\vspace{-0.2in} I’m not a bad guy! I work hard, and I love my kids. So why should I spend half my Sunday hearing about how I’m going to Hell?}{\textsc{The Simpsons}}
-
 Most of what is written in this document applies to both domestic\footnote{For simplicity domestic means you did your undergrad (or MS) at a US university.} and international students.
-However, there are some differences and benefits that domestic students should be aware of and can leverage to improve their chances of admission.
+However, there are some differences and benefits that domestic students should be aware of.
 
@@
 
-\paragraph{Research Experiences \autoref{sec:research-opportunities}} You might have many opportunities to do research as an undergraduate, e.g., through REUs and internships at your undergrad university.
+\paragraph{Research Experiences \autoref{sec:research-opportunities}} You might have many opportunities to do research as an undergraduates, e.g., through REUs and internships at your undergrad university.
 Highlight such experiences in your application.
@@
 \chapter{Research Opportunities}\label{sec:research-opportunities}
-
-\epigraph{\vspace{-0.2in} Kids, you tried your best and you failed miserably. The lesson is: never try.}{\textsc{The Simpsons}}
-
 \begin{center}
@@
 \chapter{Cultural Differences}\label{sec:cultural}
-
-\epigraph{\vspace{-0.2in} Ay caramba!}{\textsc{The Simpsons}}
 
@@
 \chapter{Rankings of CS PhD programs}\label{sec:ranking}
-
-\epigraph{\vspace{-0.2in} The whole damn system is wrong!}{\textsc{The Simpsons}}
 
@@
 
-Also thanks to NSF for encouraging faculty to be creative in research and education, which allows me to write this document.
-
 \textbf{Thank you!}

### 2024-08-10 faculty.tex

@@
 I do skim every single emails from prospective students, but I ignore most of them because they were quite generic (likely written for dozen of other profs.).
-Sometimes they try to "customize" it by mentioning the names of my papers, but
-
-.  So even if I am not busy (e.g., spending time on reddit), I still would not reply to such emails.
+Sometimes they try to "customize" it by mentioning the names of my papers, but the contents are still quite irrelevant.
+So even if I am not busy (e.g.,  spending time on reddit), I still would not reply to such emails.
+
+For the OP: you can follow up with an email now since it has been almost a month,  or do it later.
+I wouldn't care if you send now or later --  if it's generic I'll ignore it,  but if it is personalized and especially relevant to my work, I'll likely reply.
+If I am not, then you can assume I am not the right fit and should move on.
 

### 2024-08-10 demystify.tex
updates

@@
 
-\paragraph{Self-written letters are bad} Many international students write letters themselves, typically due to the requests of their profs., and have them signed by their profs.
-Such letters have \textbf{very little value} and are considered weak by reviewers---why can you not even find someone who cares or knows enough about you to write a candid personal reference letter?
-Instead of the reference writer talking about you, in this case it is you who write about yourself (and they just sign the letter).
-Experienced reviewers can easily spot (or suspect) such letters, which can be a \red{red flag}.
-
-\paragraph{Generic letters are bad} When the writers do not know much about the applicants (e.g., just taking some course with them or not making any impression to write about), they might write a generic and short letter, which is not useful and also considered weak.
+\subsection{Generic Letters are Bad} When the writers do not know much about the applicants (e.g., just taking some course with them or not making any impression to write about), they might write a generic and short letter, which is not useful and also considered weak.
+
 This does not mean the professor is not good or does not care about you, but they just do not know you well enough to write a strong letter.
@@
 
-\paragraph{Waiving Your Right}  Choosing not to look at a reference letter is pretty standard in school and job applications.
+\subsection{Self-written Letters are Bad} Many international students write letters themselves, typically due to the requests of their profs., and have them signed by their profs.
+Such letters have \textbf{little value} and are considered weak by reviewers---why can you not even find someone who cares or knows enough about you to write a candid personal reference letter?
+Instead of the reference writer talking about you, in this case it is you who write about yourself (and they just sign the letter).
+Experienced reviewers can easily spot (or suspect) such letters, which can be a \red{red flag}.
+
+Most well-known profs.
+would not ask you to write your own letter (in fact, even not well-known ones wouldn't do this).
+However, many international students do not know this (e.g., normal at their university) and might be in a situation they do not have a choice as they need the letter.
+However, think about this: if a prof.
+does this often, then they are likely not a good English writer or relatively unknown in the field.
+In either case, their letters won't worth much.
+So if you are in this situation, you should find someone else to write for you.
+
+\subsubsection{Letters from Admin people}  
+The ``department head'', ``dean'' or ``director'' at your university might not be well-known to adcom members (sorry, we likely recognized researchers, not admin people), so it's fine to to have "less well-known" people write for you as long as they can write a good letter about your research ability).
+
+\subsection{Waiving Your Right}  Choosing not to look at a reference letter is pretty standard in school and job applications.
 When you waive your right to see the letter, it adds a layer of trust, showing you're confident in your choice of referees and that you're not trying to twist their words.
@@
 
-\paragraph{Publications} Concrete evidence of research ability is having \emph{papers in reputable international journals or conferences}.
-Having published papers, especially at top venues, is a sign that the applicant has been successfully involved in research.
-Applicants admitted to top schools often have multiple first-authored papers at top places.
+\paragraph{Publications} The most concrete evidence of research ability is having \emph{papers in reputable international journals or conferences}.
+Having published papers, especially at top venues, is a sign that the you has been successfully involved in research.
+Publications are never required for application, however given the competitiveness of the process, they can significantly strengthen your application and are becoming the norm, e.g., those admitted to top schools often have multiple first-authored papers at top places.
 

### 2024-08-11 demystify.tex
updates

@@
 But they are not required!
-(\autoref{sec:research-experiences}).
+(\autoref{sec:research-experience}).
 
@@
 
-\paragraph{Help your LOR Writer} As mentioned above, do not write your own letter and generic letters do not give much value.
+\subsection{Help your LOR Writer} As mentioned above, do not write your own letter and generic letters do not give much value.
 One thing you can do to help your writer to write a strong, customized LoR for you is to provide details or unique things about yourself.
-For example, let them know about your GPA, research and work experiences, papers (if any), or anything you want them to mention.
+For example, let them know about your GPA, research and work experience, papers (if any), or anything you want them to mention.
 You can also provide them with a draft of your SOP so that they can see what you are saying about yourself and complement that with their own perspective.
@@
 
-\paragraph{Remind Your Writers} After you submit your application, you should tell your writers about that and let them know they will soon receive an email from the university to submit their letters.
+\subsection{Remind Your Writers} After you submit your application, you should tell your writers about that and let them know they will soon receive an email from the university to submit their letters.
 You should also remind them when you submit your application and ask them to submit their letters on time if they haven't done so.
@@
 
-\paragraph{VEF 2.0} For Vietnamese students, it's worth mentioning the \href{https://vef2.org/}{VEF2.0 program}, which has helped many students gain admission to top CS PhD programs in the US.
+\subsection{VEF 2.0} For Vietnamese students, it's worth mentioning the \href{https://vef2.org/}{VEF2.0 program}, which has helped many students gain admission to top CS PhD programs in the US.
 VEF2.0 invites US faculty members from leading institutions to conduct rigorous interviews with VEF students and subsequently provide reference letters on their behalf.
@@
 This is a simple task often overlooked by many applicants: \textbf{tailor your SOP} to the institution you're applying for,
-e.g., why do you apply \emph{here}?
+e.g., why do you apply \emph{here}? who do you want to work with?
 Provide names of professors who you're interested in (if they are not already in the adcom, your application might get forwarded to them for evaluation; and they might be interested in interviewing and recruiting you).
@@
 
-\section{Research Experiences}\label{sec:research-experiences}
+\section{Research Experience}\label{sec:research-experience}
 
@@
 
-\paragraph{Publications} The most concrete evidence of research ability is having \emph{papers in reputable international journals or conferences}.
+\subsection{Publications} The most concrete evidence of research ability is having \emph{papers in reputable international journals or conferences}.
 Having published papers, especially at top venues, is a sign that the you has been successfully involved in research.
@@
 
-\paragraph{Work Experiences} Work experiences at \emph{well-known research laboratories}, such as Microsoft Research, can strengthen your
+\subsection{Work Experience} Work experience at \emph{well-known research laboratories}, such as Microsoft Research, can strengthen your
 application.
@@
 
-\paragraph{Competitions} Winning \emph{internationally recognized competitions} can also demonstrate your research potential.
+\subsection{Competitions} Winning \emph{internationally recognized competitions} can demonstrate your research potential.
 For example, participating in Math Olympiads if you want to do theory or winning ACM programming contests if you want to ``build'' systems, e.g., software analysis.
@@
 
-\paragraph{GREs Do Not Matter} While a few schools still require taking the GRE (e.g., UCF), most good CS PhD programs in the US \textbf{no longer} require it.
+\subsection{GREs Do Not Matter Much} While a few schools still require taking the GRE (e.g., UCF), most good CS PhD programs in the US \textbf{no longer} require it.
 Thus, if you haven't taken the GRE or have bad scores, then don't waste time (re)taking it.
@@
 
-\paragraph{English Tests (IELTS, TOEFL)} Unless your degrees are from specific countries such as \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-Mason#standard-tests-waiver-eligible-countries}{these}, you will need to take standardized English tests.
+\subsection{English Tests (IELTS, TOEFL)} Unless your degrees are from specific countries such as \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-Mason#standard-tests-waiver-eligible-countries}{these}, you will need to take standardized English tests.
 
@@
 
-\paragraph{Work Experiences} Work experiences, especially if they are not research-based, \emph{do not} carry much weight.
+\subsection{Work Experience} Work experience, especially if they are \emph{not} research-based, \emph{do not} carry much weight for PhD (they can help for MS application though).
 Similarly, LoR from your supervisor for non-research experience will not count much.
@@
 However, if you have research-based work experiences, then they can be useful and should be mentioned in your SOP and LORs.
-For example, if you have worked at a research lab, e.g., Microsoft, DoE, VinAI in Vietnam, etc (\autoref{sec:research-experiences}), then you should mention that in your SOP and ask your supervisor to write about your research experiences and potential in their LOR (\autoref{sec:lor}).
-
-\paragraph{Online Courses and Certificates}
+For example, if you have worked at a research lab, e.g., Microsoft, DoE, VinAI in Vietnam, etc (\autoref{sec:research-experience}), then you should mention that in your SOP and ask your supervisor to write about your research experiences and potential in their LOR (\autoref{sec:lor}).
+
+\subsection{Online Courses and Certificates}
 In general, these do not carry much weight as they do not show research ability. We do not care much if you have taken an online Coursera AI course or have a professional certificate from Microsoft.
@@
 
-You still can apply to PhD in CS \emph{as long as} you can demonstrate you are ready for it through your background, research experiences, LoRs, statements, etc.
+You still can apply to PhD in CS \emph{as long as} you can demonstrate you are ready for it through your background, research experience, LoRs, statements, etc.
 You might be even able to leverage this to make your profile stand out as mentioned in \autoref{sec:improve-your-chance}.
@@
 
-\paragraph{Research Experiences \autoref{sec:research-opportunities}} You might have many opportunities to do research as an undergraduate, e.g., through REUs and internships at your undergrad university.
-Highlight such experiences in your application.
+\paragraph{Research Experience \autoref{sec:research-opportunities}} You might have many opportunities to do research as an undergraduate, e.g., through REUs and internships at your undergrad university.
+Highlight such experience in your application.
 
@@
 As discussed in \autoref{sec:research-opportunities}, research experience is a crucial component of a strong PhD application.
-It gives you opportunities to try out research, determine what research area you're interested in, publish papers (\autoref{sec:research-experiences}), connect with researchers, and get strong LoRs (\autoref{sec:lor}).
+It gives you opportunities to try out research, determine what research area you're interested in, publish papers (\autoref{sec:research-experience}), connect with researchers, and get strong LoRs (\autoref{sec:lor}).
 

### 2024-08-11 demystify.tex
updates

@@
 % \end{table}
-
-\section*{Who is this document for?}
+\section{Why the US and not other countries?}
+TODO
+
+\section{Who is this document for?}
 
@@
 
+\section{What's the difference between PhD and MS in CS?}\label{sec:phd-vs-ms}
+
+While both are graduate degrees, they are different in terms of objective, admission requirements, course requirements, duration, and funding as summarized in \autoref{tab:phd-vs-ms}.
+
+\begin{table}
+  \caption{PhD vs. MS}\label{tab:phd-vs-ms}
+  \begin{tabular}{c|c|c}
+  \toprule
+  \textbf{} & \textbf{PhD} & \textbf{MS} \\
+  \midrule
+  Objective & Industry & Research \\
+  Admission Req & Research experience & No research experience \\
+  Coursework Req & Yes & Yes (research is much more important)  \\
+  Duration & 2 years & 5--7 years \\
+  Adviser Req & No & Yes \\
+  Funding & No & Yes \\
+  \bottomrule
+  \end{tabular}
+\end{table}
+
+\begin{itemize}
+
+  \item Objective: an MS is typically to prepare you for industry, while a PhD is to prepare you for research.
+Some MS has thesis option but the research is much lighter compared to a PhD, which is research-focused.
+
+  \item Admission requirements: MS also requires a good GPA, LoRs, SOP, and test scores, but research experience is not required.
+PhD requires all of these, but research experience is a must.
+Moreover, MS programs are typically much less competitive than PhD programs.
+In fact, many MS program are viewed as a revenue source (\emph{cash cow}) for the department, as they are not funded and students pay tuition.
+
+  \item Course requirements: MS has a specific number of courses (typically can be done in 2 years).
+You graduate with an MS when you're done with the courses.
+PhD also has coursework requirements (typically taken in your first 2 years) but focuses on research after that.
+You graduate with a PhD when you have done enough research and written a dissertation (which is typically way harder and takes much longer time than coursework).
+
+  \item Duration: an MS typically takes 2 years while a PhD takes 5--7 years (or even longer).  Many students get an MS along the way to a PhD, e.g., after finishing the 2-year course work.
+
+  \item Adviser: MS students typically do not have an adviser (if you do thesis option then you will have one), while PhD students need an adviser who guides them in their research.
+
+  \item Funding: MS is typically \emph{not funded}, while PhD is typically funded (e.g., through TA or RA as explained in \autoref{sec:funding}).
+
+\end{itemize}
+
 \section{PhD in the US vs. Other Countries}\label{sec:non-us-differences}
 
-\begin{table}[h!]
-\centering
-\begin{tabular}{|c|c|c|}
+\begin{table}
+\caption{Comparison of US and Other PhD Programs}\label{tab:us-vs-other}
+\begin{tabular}{c|c|c}
 \toprule
@@
 \end{tabular}
-\caption{Comparison of US and Other PhD Programs}\label{tab:us-vs-other}
 \end{table}

### 2024-08-13 demystify.tex
updates

@@
 % \end{table}
-\section{Why the US and not other countries?}
-TODO
+% \section{Why the US and not other countries?}
+% TODO
 
@@
 
-Sometimes students would go to great lengths just to get letters from well-known professors in their school (e.g., department head).
+Sometimes students would go to great lengths just to get letters from well-known professors in their school (e.g., department head or dean).
 But as mentioned, if these professors do not know you, their letters would likely be generic and carry little value (sometimes \red{red flags}).
-Moreover, a top professor at your university might not be well-known to US faculty (see more details in \autoref{sec:your-school}).
+Moreover, a top professor at your university might not be well-known to US faculty (see more details in \autoref{sec:admin-letters} and \autoref{sec:your-school}).
 So save the trouble and get letters from \emph{any} professors/supervisors who know you well and can write a good letter about \emph{your} research ability.
@@
 \subsection{Self-written Letters are Bad} Many international students write letters themselves, typically due to the requests of their profs., and have them signed by their profs.
-Such letters have \textbf{little value} and are considered weak by reviewers---why can you not even find someone who cares or knows enough about you to write a candid personal reference letter?
-Instead of the reference writer talking about you, in this case it is you who write about yourself (and they just sign the letter).
-Experienced reviewers can easily spot (or suspect) such letters, which can be a \red{red flag}.
-
-Most well-known profs.
+Such letters have \emph{little value} and are considered weak by reviewers---why can you not even find someone who cares or knows enough about you to write a candid personal reference letter?
+Instead of the ref.
+writer talking about you, in this case it is you who write about yourself (and they just sign the letter).
+Experienced reviewers can easily spot  (because the student often has little experience in writing LoRs) and  \red{red flag} such letters.
+
+Most well-known and well-respected profs.
 would not ask you to write your own letter (in fact, even not well-known ones wouldn't do this).
-However, many international students do not know this (e.g., normal at their university) and might be in a situation they do not have a choice as they need the letter.
+However, many international students do not know this (e.g., it is a common practice at their university) and might be in a situation they do not have a choice as they need the letter.
 However, think about this: if a prof.
-does this often, then they are likely not a good English writer or relatively unknown in the field.
-In either case, their letters won't worth much.
+does this often, then they are likely not a good English writer, relatively unknown in the field, or simply do not know/care enough about you.
+In any of these cases, their letters won't worth much.
 So if you are in this situation, you should find someone else to write for you.
 
-\subsubsection{Letters from Admin people}  
-The ``department head'', ``dean'' or ``director'' at your university might not be well-known to adcom members (sorry, we likely recognized researchers, not admin people), so it's fine to to have "less well-known" people write for you as long as they can write a good letter about your research ability).
+\subsection{Letters from Admin people}\label{sec:admin-letters}
+
+Many international students try to get letters from high-ranking admin people in their universities, e.g., department head, dean, or director.
+These people, while well-known and respected in your university, might not be too active in research (e.g., they focus on administration) and thus are not recognized by adcom members, who often recognize good researchers (even if they are not in their field'', but not admin people.
+So it's fine to to have `less well-known'' people write for you as long as they can write a good letter about your research ability.
+
+\begin{commentbox}[Vu:]
+  In my experience in reviewing applications, letters from admin people are often generic and do not provide much value.  If you are in this situation, you should find someone else to write for you.
+\end{commentbox}
 
@@
 \begin{commentbox}[Vu:]
-  I often read SOP and LORs first. If I am
+  I often read SOP and LORs first (\autoref{sec:ievaluate}). If I am
   persuaded by then, I would skim over other factors and advocate for
@@
 So you or your LoR writers should explicitly say something about them in your statement.
+In general, if you did some good research work at such places, then you should mention that in your SOP and ask your supervisor to write about your research experiences and potential in their LORs (\autoref{sec:lor}).
 
@@
 \end{commentbox}
+
+On the other hand, if non-research work (e.g., software development) \emph{do not} carry much weight for PhD (they can help for MS application though).
+Similarly, LoR from your supervisor for non-research experience will not count much.
+So do not spend much time talking about non-research-related job in your SOP.
 
@@
 
-\section{Standard Tests (GRE, IELTS, etc), Work Experiences, Online Courses, and Certificates}\label{sec:standard-tests}
+\section{Standard Tests (GRE, IELTS, etc), Online Courses, and Certificates}\label{sec:standard-tests}
 
@@
 \end{commentbox}
-
-\subsection{Work Experience} Work experience, especially if they are \emph{not} research-based, \emph{do not} carry much weight for PhD (they can help for MS application though).
-Similarly, LoR from your supervisor for non-research experience will not count much.
-
-Do not spend much time talking about non-research-related job in your SOP.
-However, if you have research-based work experiences, then they can be useful and should be mentioned in your SOP and LORs.
-For example, if you have worked at a research lab, e.g., Microsoft, DoE, VinAI in Vietnam, etc (\autoref{sec:research-experience}), then you should mention that in your SOP and ask your supervisor to write about your research experiences and potential in their LOR (\autoref{sec:lor}).
 
@@
 
-For each application, the Mason application system compiles a single PDF file, which consists of a summary, transcripts, test scores, LoRs, a CV, SOP, and writing samples.
-I usually start with the \emph{summary}, checking for low GPAs or test scores below the university minimum (\autoref{sec:standard-tests}).
+For each application, our system compiles a single PDF file, which consists of a summary (degrees, GPAs, etc), transcripts, test scores, LoRs, a CV, SOP, and writing samples.
+I usually read in this order.
+I start with the \emph{summary}, checking for low GPAs or test scores below the university minimum (\autoref{sec:standard-tests}).
 I then skim the \emph{transcripts} for low grades in relevant courses, noting issues like ``many low grades in main courses or unknown international school with good GPA''.
@@
 I read carefully \emph{strong LoRs} and skim weaker ones, noting either strong letters from well-known professors/researchers or generic content.
-For the \emph{CV}, I look for publications, research experiences, and notable achievements, noting things like ``published papers in top venues'' or ``gold medal in an international competition.''
-I read strong \emph{SOP} thoroughly (and skim weak ones). I note whether the SOP is exciting, research-oriented, and tailored to our program. All these notes are entered into the evaluation system.
-
-Finally, I make a decision, which is usually either a rejection or an offer of admission with full funding (e.g., from TA).
+I skim over \emph{CV} and look for publications, research experiences, and notable achievements, noting things like ``published papers in top venues'' or ``gold medal in an international competition.''
+I read strong \emph{SOPs} thoroughly (and skim weak ones).
+I note whether the SOP is exciting, research-oriented, standing out, e.g., applicant is minority (\autoref{sec:urm}), and tailored to our program (e.g., if they are familiar with the work of some faculty or have talked to them).
+These notes are entered into the evaluation system.
+
+Finally, I enter my decision, which is usually either a rejection or an offer of admission with full funding (e.g., from TA).
 I also recommend very strong candidates for the University Presidential fellowship.
-The adcom chair then reviews all evaluations, looks at entered notes, and asks reviewers to discuss and resolve discrepancies to reach a consensus (e.g., a reviewer wants to accept but the other wants to reject).
+Note that while the system has other decision options, eg., admit without funding, provisional admission (e.g., if they need to take some courses), I rarely use them, simply because we already have too many good students and do not have the resources to support them.
+
+\subsection{How Decisions are Made}\label{sec:how-decisions}
+
+After all reviewers have evaluated an application, adcom chair will review all evaluations, look at entered notes, and ask reviewers to discuss and resolve discrepancies to reach a consensus (e.g., a reviewer wants to accept but the other wants to reject).
 Typically, the decision is made entirely by the reviewers; there is no involvement from the adcom chair, department chair, or others.
+In most cases adcom members, even those reviewing the same application, make decisions independently and do not talk to each other (just a common practice to avoid biasing).
+In some rare cases we might (\autoref{sec:adcom-discuss}).
+
+Even if \emph{all} reviewers recommend acceptance, the application is not automatically accepted, especially if no faculty is willing to advise the student.
+For example, if the student is interested in a research area that no faculty is working on, or an area where no faculty is taking new students (e.g., AI/ML where faculty likely already has many students) then the student will not be admitted.
+This is increasingly common as the number of applicants increases significantly in recent years in CS.
+
+However, if the student has contacted a faculty member and that faculty member is interested in the student and has made this known to the admission committee, then the student is likely to be admitted, \emph{even if the faculty does not review the application or has funding} (thus the importance of contacting faculty).
+Similarly, if the student mentioned a faculty member in their SOP, adcom might ask that faculty member to look at their application and if they are interested in the student.
+Even if the student has a weak profile (but still passes the minimum requirement from the university), they might be admitted if a faculty member is willing to take them.
+Adcom members, especially in the US, are very reluctant to go against the faculty's decision (e.g., if a faculty wants to admit a student, we are not going to reject them).
 
 \section{Do adcom members talk to each other?}\label{sec:adcom-discuss}
-
-In most cases adcom members, even those reviewing the same application, make decisions independently and do not talk to each other.
-It's not a strict rule but a common practice to avoid bias and ensure fairness (and also we don't have time to talk to each other).
 

### 2024-08-13 demystify.tex
updates

@@
 Moreover, having bad grades in relevant courses, e.g., Math and CS, can be a \red{red flag}.
-Adcom members often scan through transcripts looking for C and lower grades in Math and CS courses and might raise concerns if they see several of them.
+Adcom members often scan through transcripts (\autoref{sec:ievaluate}) looking for C and lower grades in Math and CS courses and might raise concerns if they see several of them.
 Note that bad grades in non-relevant courses, e.g., e.g., about politics or history, are not as concerning.
@@
 A common question is whether you should be concern and explain bad grades in relevant courses in your SOP.
-If you have just a few, then they do not matter much, so don't spend much time explaining them (let's face it, many of us, including faculty, have bad grades in some courses, including relevant ones).
+If you have just a few, then they do not matter much, so don't spend much time explaining them (many adcom reviewers, like myself, have bad grades in relevent courses!).
 But if you have many bad grades for an entire semester or year due to some specific reasons, then you may be you should explain them in your SOP.
 
-\section{Standard Tests (GRE, IELTS, etc), Online Courses, and Certificates}\label{sec:standard-tests}
+\section{Standard Tests (e.g., GRE and IELTS), Online Courses, and Certificates}\label{sec:standard-tests}
 
@@
 
-\paragraph{Not getting interviews} While it is generally good to get an interview, not getting one \emph{does not} mean you're out.
+\subsection{Not Getting Interviews} While it is generally good to get an interview, not getting one \emph{does not} mean you're out.
 Many programs do not have the tradition of interviewing applicants.
@@
 
-\paragraph{What else?} Make a decision, accept, reject, or defer the offers  (\autoref{sec:accept-postpone-decline}).
+\paragraph{What next?} Make a decision, accept, reject, or defer the offers  (\autoref{sec:accept-postpone-decline}).
 Ask to meet with potential advisers (e.g., through Open House or separately) and even their students.
@@
 You will likely not get any useful feedback.
-We are not willing to reveal your evaluation results or give you advice on how to improve your profile.
-So it is best to just move on.
+We are not willing (allowed) to reveal your evaluation results or give you advice on how to improve your profile.
+So just move on.
 If you really want feedback, ask your professors, collaborators, ref writers, or those who have previously applied.

### 2024-08-13 demystify.tex
updates

@@
 You will likely not get any useful feedback.
-We are not willing (allowed) to reveal your evaluation results or give you advice on how to improve your profile.
-So just move on.
-If you really want feedback, ask your professors, collaborators, ref writers, or those who have previously applied.
+We are not willing (somtimes allowed) to reveal your evaluation results or give you feedback on how to improve your profile.
+\emph{So just move on}.
+If you really want advice, ask your professors, collaborators, ref writers, or those who have previously applied.
 

### 2024-08-14 demystify.tex
updates

@@
 A common question is whether you should be concern and explain bad grades in relevant courses in your SOP.
-If you have just a few, then they do not matter much, so don't spend much time explaining them (many adcom reviewers, like myself, have bad grades in relevent courses!).
+If you have just a few, then they do not matter much, so don't spend much time explaining them (many adcom reviewers themselves have bad grades in relevent courses!).
 But if you have many bad grades for an entire semester or year due to some specific reasons, then you may be you should explain them in your SOP.
@@
 Thus, if you haven't taken the GRE or have bad scores, then don't waste time (re)taking it.
-It does not help much!
-
-However, if you took it and have really good scores then it might be worth it to include and talk about them in your application (but don't expect them to make much difference).
+Being optional really means optional, and not taking it will not hurt your application.
+
+However, if you took it and have really good scores then it might be worth it to include and talk about them in your application, but don't expect them to make much difference (many adcom reviewers never took it).
 But if your scores are bad, then you should not include them in your application, which can be a \red{red flag}.
@@
 \subsection{English Tests (IELTS, TOEFL)} Unless your degrees are from specific countries such as \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-Mason#standard-tests-waiver-eligible-countries}{these}, you will need to take standardized English tests.
-
+On one hand you will need to show some level of English proficiency, but on the other hand, you do not need to have very good scores in these tests (many adcom members themselves were once international students and struggled with English).
 You should just do well enough to pass the minimum requirement set by the university. %, which is required for TA (\autoref{sec:funding}).
@@
 The CV/Resume provides a summary of the applicant's achievements. Most schools require you to upload your CV with your application.
-
 Prepare your CV in such a way that it allows the reviewers to quickly scan to identify major achievements, e.g., Publications, Programming Competition Awards, and Teaching Experience.
@@
 
-Just like a job application, you likely will get interviews from schools.
+After you apply, you might get interviews.
 The most common case is that a prof.
-is interested in working with you and wants to chat to make a decision, e.g., to offer RA (\autoref{sec:ra}).
+is interested in working with you and wants to chat with you, e.g., to offer RA (\autoref{sec:ra}).
 In some cases, the interview is done by several professors, e.g., to see if a student fits in their group or to recruit a very strong student to their program.
 The interview can also be done by adcom members to see if the student fits their program.
-%The professor might include one of their current students as part of the interview.
 
 Typically, an interview takes about 15--30 minutes, and one important aspect of evaluation is your ability to effectively communicate, including speaking and understanding English. 
-You might be asked to talk about your research experiences and interests and to read a paper and discuss it. In some rare cases you might also be asked to solve a problem, e.g., a coding problem. 
-The interview is also a good opportunity for you to ask questions about the program and the university.
+You might be asked to talk about your research experience and interests and to read a paper and discuss it.
+In some rare cases you might also be asked to solve a problem, e.g., a coding problem (one of my colleagues at Mason often does this).
+The interview is also a good opportunity for you to ask questions, e.g., about the prof, their group, the CS PhD program, and the university.
 
 You should treat the interview as an informal chat.
-Have ``elevator pitches'' about your research experiences and interests.
+Have ``elevator pitches'' about your research experience and interests.
 You might also want to have a 5-minute presentation about your research.
@@
 The timeline for interviews and notification letters varies.
-Faculty set up interviews based on their (busy) schedule.
+Faculty set up interviews based on their (busy and erratic) schedule.
 Do not be surprised if you get an interview invitation at the last minute.
-Some schools do not do interviews at all.
+Some schools do not do interviews at all (\autoref{sec:no-interview}).
 
@@
 
-\subsection{Not Getting Interviews} While it is generally good to get an interview, not getting one \emph{does not} mean you're out.
+\subsection{Not Getting Interviews}\label{sec:no-interview}
+While it is generally good to get an interview, not getting one \emph{does not} mean you're out.
 Many programs do not have the tradition of interviewing applicants.
@@
 
-However, no interviews typically mean that you will not likely get an RA (\autoref{sec:ra}), which is offered by an individual faculty (if they want you to do research for them, then they likely will interview you first).
+However, no interviews could mean that you will not likely get an RA (\autoref{sec:ra}), which is offered by an individual faculty (if they want you to do research for them, then they likely will interview you first).
 In the case of no interviews, your application (and TA/fellowship funding) is decided by the adcom.
@@
 Do not assume that you're young and healthy and ignore insurance.
-At Mason, and most good CS PhD programs, your GTA or GRA \emph{will always} come with full insurance.
+At Mason, and most CS PhD programs, your GTA or GRA comes with full insurance.
 In fact, at Mason your spouse/children will get a significant discount rate for health insurance.
@@
 For example, in addition to a traditional CS department, Mason has IST and Cybersecurity departments, both of which have faculty with PhD in CS and work on CS topics (e.g., AI, Security, Robotics).
-So it is possible that you still get to do CS research and publish in CS-focused venues even if you're not in a traditional CS program.
+So you still can do CS research and publish in CS-focused venues even if you're not in a traditional CS program.
 It is common to see faculty with PhD in CS in a non-CS department as well as faculty with non-CS PhD in a CS department.
 
-However, if your goal is a PhD in CS, then you likely need to be in the CS dept \emph{and} advised by a CS faculty.
+However, if your goal is a PhD in CS, then you need to be in the CS dept \emph{and} advised by a CS faculty.
 A non-CS faculty can serve in the PhD dissertation committee (common) or \emph{co-advise} (somewhat rare), but your main PhD adviser will likely be a tenure-line faculty in CS (\autoref{sec:faculty-types}).

### 2024-08-15 faculty.tex
updates

--- 2024-08-10 faculty.tex
@@
 
-\chapter{Research}\label{chap:research}
-This section focuses on paper publication, which is a major component of P\&T and of course in getting your name out there.
+\chapter{Research and Publications}\label{chap:research}
+
+This section focuses on research publication, which is a major component of P\&T and of course in getting your name out there.
 P\&T evaluation requires no specific conferences/journals or numbers of papers, but typically you will want to have a good number of papers in top venues.
-P\&T members or letter writers (obviously) are familiar with top places in your field, and they will definitely look for those in your profile.
-Work that won awards, e.g., paper awards, can help a lot in tenure evaluation.
+%P\&T members or letter writers (obviously) are familiar with top places in your field, and they will definitely look for those in your profile.
+Work that won awards, e.g., best paper awards, can help a lot in P\&T.
 Papers with very high citations also help.
@@
 
-Depending on your research area, you might focus on publishing at top conferences or journals.
+Depending on your research area, you might focus on publishing at top conferences or journal.
 You probably should know what's popular in your area already (if not, ask your senior colleagues).
@@
 
-\paragraph{Conferences}
-In most CS fields, top conferences are more common and competitive than journals.
-Each field also has several of such top conferences yearly (e.g., in Software Engineering, ICSE, FSE, ASE, and ISSTA are top four).
+\subsection{Conferences}
+In most CS fields, conferences are more common and competitive than journals.
+Each field has several of top conferences every year (e.g., in Software Engineering, ICSE, FSE, ASE, and ISSTA are top four).
 They are often scheduled so that if you miss (or get rejected) from one, you can submit to another.
@@
 
-\paragraph{Artifact Evaluation}
-
-\paragraph{Journals}
-In some journals are more common and prestigious.
-Moreover, many areas have top journals in additions to well-known conferences (e.g., Trans.
+\paragraph{Artifact Evaluation} This is probably more well-known in SE conferences, and is intended to promote the reproducibility of research results.
+However, I never liked it because it's a lot of work and many of my experiments require heavy computation that the reviewers do not have.
+I also don't think it's very useful as I have always make my code, benchmarks, and results open source (through Github).
+I found it strange that many SE conference have recently required this as part of the submission process.
+It makes the submission process much more complicated and much more time-consuming.
+
+\subsection{Journals}
+In some fields, journals are more common and prestigious.
+Moreover, many areas have top journals in additions to well-known conferences (e.g., in addition to the main fource confs as mentioned above, SE also has Trans.
 of Software Engr.
-or TSE is a very well-known journal in Software Engineering, which has 4 major confs.).
+or TSE, which is a very well-known and high-quality journal).
 Typically, or at least in SE, the common approach is publishing new work to conference and followup work to journals (e.g., so the journal paper is an substantial extension of the conf.
@@
 
-\paragraph{Short papers, workshops, etc}
-
-\section{Collaboration and co-authorship}
-While it's general true that profs.
-are busy, but for me and probably many others, it's not the main reason.
-I do skim every single emails from prospective students, but I ignore most of them because they were quite generic (likely written for dozen of other profs.).
-Sometimes they try to "customize" it by mentioning the names of my papers, but
-
-.  So even if I am not busy (e.g., spending time on reddit), I still would not reply to such emails.
-
-\paragraph{Papers with your adviser(s)}
+While you are not expected to publish regularly in journals, having a few papers in top journals by your tenure evaluation shows the diversity of your research and that you can publish in different venues.
+
+\subsection{Short papers, workshops, etc} Short papers in workshop or non-research tracks at a conference (e.g., tool, demo, vision, etc) do not count as much.
+They describe work that might not be as mature (e.g., new idea) or as complete (e.g., a tool or a demo) as a full paper, and also have higher acceptance rate.
+Many people use them as a way to get feedback on their work before submitting to a full paper venue.
+Most new professors do not focus on these kinds of papers.
+Many P\&T committees do not count or care much about these and having too many of them can be a negative signal.
+
+Personally, I find that such papers are very useful as preliminary work for writing grant proposals.
+They show that you have a new idea and good preliminary results that are worth pursuing further.
+Sometimes a full paper is not as good because the research is relatively complete with strong results, which bring the concern that you already did all the cool work and the proposal is minor and not exciting.
+
+\section{FAQs}
+
+\subsection{Your university does not recognize conference value} For most non-CS engineering fields journal are more prestigious than conferences.
+Places with a strong CS department are well aware of the value of top confs., but many other places are not (especially smaller universities or CS depts run by engineering people).
+These places also do not know much about the quality of the venue, and so in many cases put a workshop paper in the same class as a top conference paper and also cannot differentiate between a good vs bad journals in CS (because they do not know CS well).
+Also,  they often value quantity over quality because they cannot properly evaluate paper quality.
+
+This is a cultural issue and is very difficult to change.
+Many good researchers have left such places because of this.
+However, some were able to leverage this and publish at easy places and do really well within their department.
+However, while they thrive at such places, they will likely not get funding and recognition in their field.
+This is a trade-off that you need to consider.
+
+\subsection{Papers with your adviser(s)}
+One common question is whether you should publish papers with your adviser(s) after you start your own lab and become a professor.
+The straight answer is that if you collaborate too much with your adviser(s), it's not good and brings concern about you not being an independent researcher (you are already a professor and so are expected to be able to do research on your own).
+I have seen many cases where a new professor has published many papers with their adviser(s) and then had trouble getting tenure (even renewal).
+
+However, this does not mean you should not work with your adviser(s).
+You can absolutely continue do research and write grant proposals with them, but you should also work and write papers with other people, and especially your students.
 

### 2024-08-18 demystify.tex
updates

@@
 
-\paragraph{Low Stipend?} Students often complain their stipend is too low.
+\paragraph{Low Stipend?} Students sometimes complain about their stipend being low.
 However, in most cases it is not bad, and you can live comfortably with it.
@@
 A good school would know that it has to be competitive to attract students.
-For example, at Mason, every year we discuss improving the benefits, and especially stipend, for our graduate students.
+For example, at Mason, every year we try to improve the benefits, and especially stipend, for our graduate students.
 
@@
   Having health insurance is required at many US universities.
-Do not assume that you're young and healthy and ignore insurance.
+Do not assume that you're young and healthy and ignore insurance (\autoref{sec:cultural-misc}).
 At Mason, and most CS PhD programs, your GTA or GRA comes with full insurance.
@@
 
-In addition to general fellowships that are available to all students, there are also specific fellowships, e.g., for US citizens and permanent residents, for underrepresented minority (URM) students (\autoref{sec:urm}), and students from specific countries.
+In addition to general fellowships that are available to all students, there are also specific fellowships, e.g., for US citizens and permanent residents (\autoref{sec:domestic-students}), for underrepresented minority (URM) students (\autoref{sec:urm}), and students from specific countries.
 There are many fellowships available, so you should look for them and apply.
@@
 
-\section{Funding In the Summer}\label{sec:summer-funding}
+\section{Miscs}
+
+\subsection{Funding In the Summer}\label{sec:summer-funding}
 
@@
 
-\chapter{Choosing Schools and Professors}\label{sec:schoolsandprofs}
-
-\epigraph{\vspace{-0.2in} The answer to life's problems aren't at the bottom of a bottle. Heh heh! They're on TV!}{\textsc{The Simpsons}}
-
-Choosing a school and an adviser is clearly among the most important things in your mind when you apply and especially when you get admitted.
-This is further complicated due to cultural differences and the unfamiliarity of international students with the US higher education system.
-This section aims to mitigate some confusion and help you make informed decisions.
-
-\section{Choosing a University}\label{sec:choosing-university}
-
-We will first discuss universities in the US that offer PhD in CS. Then we will talk about how to select them.
-
-\subsection{Schools that offer PhD in CS}
-
-Most US universities have CS studies, but many of them do not have a PhD program in CS.
-These universities might offer just Bachelor's degrees (e.g., BS) and no graduate studies (i.e., no MS or PhD degrees), or they just offer MS programs (but no PhD).
-For example, Penn State in University Park has PhD in CS,  but Penn State in Harrisburg only has BS and MS in CS, and Penn State in York only has BS in CS.
-On the other hand, multiple locations of the University of Texas, e.g., Austin, Dallas, and Arlington, have PhDs in CS.
-
-Thus, if your goal is PhD in CS, you have to aim only for schools offering such a degree.
-%This might sound obvious but many students (even those in the US) get confused due to the large number of universities.
-While this can be confusing due to the large number of universities in the US, a little research, e.g., searching for PhD in CS from the school website, will help you find out.
-Schools listed in \autoref{sec:ranking} have PhD programs in CS, so you can start there.
-
-% \subsection{R1, R2, ...}
-
-\subsection{Selecting and Ranking Schools}\label{sec:selecting-ranking-schools}
-\begin{center}
-  \includegraphics[scale=0.6]{files/c1.png}
-\end{center}
-
-International students often put schools into two bins: (i) top schools that they dream about, and (ii) everything else.
-They commonly use rankings from US News, which is not transparent and questionable\footnote{\url{https://cra.org/cra-statement-us-news-world-report-rankings-computer-science-universities/}}.
-Sometimes they evaluate based on the reputation of the school's undergrad program or the reputation of the school's non-CS programs such as medical, math, or physics.
-They even rank universities based on popular places they know in the US, e.g., California, Texas, and New York.
-
-Instead of these superficial criteria, you should consider the CS program and the research interests of faculty members.
-You can learn about these using resources such as \href{https://csrankings.org}{CSRankings.org}.
-You will be very surprised to learn that a school that you didn't know much about can have very strong research in your topic (and vice versa, a school you thought highly about might have no faculty working in the research field you're interested in).
-This is also a good way to learn about individual faculty, e.g., who works on what, and well-known CS conferences\footnote{In CS (and probably only in CS), conferences, not journals, are often the main venue to publish research findings (see \href{https://homes.cs.washington.edu/~mernst/advice/conferences-vs-journals.html}{why here}).}.
-%\autoref{sec:ranking} gives the top 50 CS programs in the US according to CSRankings.
-
-\begin{commentbox}[Dat:] Most Vietnamese students, including those from top schools, \textbf{do not know} about CSRankings.
-Maybe applicants who worked at top research places such as VinAI would know about it.
-\end{commentbox}
-
-\paragraph{What matters to you?} While many find CSRankings useful, it is not perfect, and in general, rankings are superficial, and you need to do more research to be informed and make better decisions.
-For example, you might prefer places with a large community from your country (e.g., Northern VA has a large population of Vietnamese).
-You might prefer places with high-tech companies (Seattle or Silicon Valley), with many outdoor activities (e.g., hiking, skiing), or with better weather (\emph{``PhD can be depressing, so would you rather be depressed in California or New York?''}).
-You might need to think about the cost of living, e.g., living in California is more expensive than in Virginia, or places with high crime rates (note that while some universities might be in a high-crime city but the campus itself is safe).
-
-If you get admission to several places, you should consider attending Open Houses (\autoref{sec:accepted}) and contact profs.
-that you're interested in at those places and talk to them.
-They would be more willing to chat with you now that you have been admitted.
-Ask questions about \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Answers-to-Ph.D-Advisor-Guide}{their work}, how they manage students, and their expectations.
-You can even ask to contact their students.
-
-\begin{commentbox}[Hung:]
-  I always encourage the students I admitted to talk with my students and the students of other faculty in other schools who admitted them.
-You will unlikely hear straight-out complaints from current students in a prof’s group.
-But sometimes what is important are things that they (current students) don’t tell you.
-Pay attention to their ``level of excitement'' being in the group.
-\end{commentbox}
-
-\begin{commentbox}[Xiaokuan:]
-  Chinese students often only look at US News rankings when selecting their PhD universities (I did that, too, when I was applying for PhD positions).
-  Now that I am a professor, I find it to be the least promising way.
-  %
-  The reason is that US News does not provide a good metric for evaluating the quality of the PhD program.
-  %
-  If you want to do great research, CSRankings is the best way to find good and active professors (which did not exist when I was applying),
-  since it solely focuses on publications at top-tier CS conferences.
-  %
-  Also,
-  I think PhD is not only about research;
-  you need to also consider your daily life there since you will (probably) stay for at least five years.
-  %
-  You might regret it if you did not consider this seriously before applying.
-  %
-\end{commentbox}
-
-\subsection{PhD in other related fields: CE, IST, Cybersecurity}\label{sec:related-fields}
-
-You \emph{do not} need to do a PhD in CS to work in your area of interest.
-For example, in addition to a traditional CS department, Mason has IST and Cybersecurity departments, both of which have faculty with PhD in CS and work on CS topics (e.g., AI, Security, Robotics).
-So you still can do CS research and publish in CS-focused venues even if you're not in a traditional CS program.
-It is common to see faculty with PhD in CS in a non-CS department as well as faculty with non-CS PhD in a CS department.
-
-However, if your goal is a PhD in CS, then you need to be in the CS dept \emph{and} advised by a CS faculty.
-A non-CS faculty can serve in the PhD dissertation committee (common) or \emph{co-advise} (somewhat rare), but your main PhD adviser will likely be a tenure-line faculty in CS (\autoref{sec:faculty-types}).
-For example, a prof. in Stats or Math might be able to serve as a co-adviser, but not as a sole adviser of a student in a CS PhD program. 
-If in doubt, check with the CS department for their requirements.
-
-For this specific reason,  CSRankings includes only faculty who can advise CS PhD students.
-I also have compiled a \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{list} of Vietnamese faculty who can advise PhD students.
-\autoref{sec:faculty-types} talks more about who can serve as your PhD adviser.
-
-\section{Choosing an Adviser}
-There is no one-size-fits-all answer to this question. The best adviser is the one that you can work well with.  But how do you find such a person?
-
-Fortunately, while some non-US programs require finding an adviser and research topic before starting the PhD (\autoref{sec:non-us-differences}), CS PhD programs in the US will typically give you a couple of years to search for advisers and research topics.
-This is especially true if you're admitted with TA (\autoref{sec:ta}), which gives you time to explore and find an adviser.
-
-\subsection{Finding the right adviser}
-
-Here is some advice for finding a prof.
-Assuming you're not familiar with any particular one, then first search for profs.
-that share similar research interests.
-For example, in CSRankings, if you want to work with PL, you can search for those published in PL conferences.
-If you want to work with SE \emph{and} AI, you can search for faculty who work in both SE and AI.
-Then you research that prof.
-and contact them (\autoref{sec:contact}).
-
-\begin{commentbox}[Xiaokuan:]
-  Whether the student's research interest matches that of the adviser is very important;
-  if there is a mismatch,
-  either the student or the adviser has to make compromises,
-  which often leads to disagreements or conflicts.
-  IMO, the adviser should be the one who \emph{guides}  students to do research while allowing students to pursue their own interests,
-  instead of \emph{dictating} their research.
-\end{commentbox}
-
-Another way is taking graduate-level courses in the topics you are interested in.
-Many profs teach \emph{special topics courses} and \emph{research seminars}, and they might be recruiting students.
-Do well in the class, answer questions, talk to the prof.
-after classes, etc---being stand out.
-Many profs, including myself, prefer taking in new students this way.
-It gives both the prof and student more time, e.g., a whole semester, to work and evaluate the relationship before making any commitment (sounds like a marriage!).
-
-You can also ask to do an independent study or research with a prof.
-This can be informal (no credit) and takes place during the summer or winter break.
-For example, I do this with several students, some of whom are undergrads.
-Many will drop out because they find they don't like my research, but some find that they like the work.
-
-Ultimately, choose a prof.
-that fits you by communicating with them, taking their courses, meeting and asking them questions, and talking to their current students.
-It will take time and effort, but since you will be working with this person for 5+ years, it is important to try to find the right one.
-
-\begin{commentbox}[Thanh:]
-  In my opinion, having a well-suited adviser is crucial for a successful PhD and research career.
-One effective approach to finding a suitable professor is by working with a professor during your undergraduate studies.
-An exemplary instance is VinAI's residency program, where residents collaborate with professors from the US for two years before applying to PhD programs.
-Many VinAI residents have achieved remarkable results and gained admission to prestigious US universities.
-Unfortunately, VinAI's resident program is limited to AI research.
-
-  In other fields, e.g.
-Software Engineering, Vietnamese students face challenges in reaching US professors.
-Do you have any tips for Vietnamese students who want to connect with US professors and work as research assistants?
-  \tcblower
-  \textbf{Vu}: \autoref{sec:contact} shows how to contact a professor for research opportunities.
-Many probably will say no (or do not reply) as they do not have the bandwidth to take on random students, but some may say yes if they see a potential fit.
-\end{commentbox}
-
-\subsubsection*{Additional Resources}
-\begin{itemize}
-  \item \href{https://www.cs.columbia.edu/wp-content/uploads/2019/03/Get-Advisor.pdf}{The Definitive ``what do I ask/look for'' in a PhD Advisor Guide}
-\end{itemize}
-\subsection{Who can serve as a PhD adviser?}\label{sec:faculty-types}
-
-\emph{Not} every faculty can serve as your official PhD adviser.
-Let's try to understand the different types of faculty and their roles.
-For example, you might hear about tenured,  tenure track, and teaching faculty.
-You might also hear about assistant, associate, full, adjunct, emeritus, teaching, and research professors.
-Here is a primer on these terminologies.
-
-\paragraph{Types of Faculty} There are typically two types of faculty: tenure-line and teaching.
-The main difference is their focus: research, teaching, or both.
-The type of faculty can affect their ability to serve as your adviser.
-
-\textbf{Tenure-line faculty}, consisting of tenured and tenure-track profs., focus on research, which includes publishing papers, obtaining grants, and mentoring PhD students.
-They often have a low teaching load (e.g., 1 per semester).
-\emph{Tenured} faculty are professors who have been promoted to permanent positions (informally, very hard to fire them).
-\emph{Tenure-track} faculty are (more likely young) professors who are on the track to get tenure.
-For most schools, tenure-line is the main type of faculty that can serve as a formal adviser of PhD students.
-\autoref{sec:tenure-vs-tenure-track} talks more about choosing between tenure and tenured-track professor as your adviser.
-
-\textbf{Teaching faculty} (sometimes called instructional faculty or professor of practice) mainly focuses on teaching.
-They typically teach 3--4 classes per semester (which is quite a lot) and \emph{do not} have research responsibilities, e.g., they do not have to worry much about publishing papers or obtaining grants.
-As such they typically do not have funding (e.g., RA~\autoref{sec:funding}) to support PhD students.
-Depending on the university, they might be able to \emph{co-advise} and mentor PhD students with a tenure-line faculty but typically do not formally advise and graduate PhD students.
-
-Note that at many universities do not allow \emph{non-CS profs}, even in tenure-line ones, to solely advise and graduate PhD students in CS (see \autoref{sec:related-fields}).
-
-Some universities have \textbf{research faculty} or research scientists, who are typically non-tenure line faculty and focus on research (e.g., publishing papers and getting grants). 
-They can recruit students and provide RA funding for them. 
-They can also teach classes but typically have a lower teaching load.
-For some universities, they \emph{can} solely advise and graduate PhD students, but for others, they cannot.
-
-\textbf{Adjunct faculty} is not full-time, e.g., they might be working in industry and teach a class or two for fun.
-\textbf{Emeritus faculty} means they are retired from the university, e.g., they could move to the industry, but still have some affiliation with the university.
-Due to their roles, these profs.
-typically do not advise PhD students.
-
-\paragraph{Ranks} Faculty have ranks such as assistant, associate, and full, regardless if they are tenure-line or teaching.
-Assistant means new faculty, associate means they have been promoted, and full means they are senior.
-Typically, a tenure-track faculty starts with being an assistant professor and then gets tenure and promoted to associate (typically after 6 years) and then full (time varies, some become full within 3--4 years, some 10+ years, some remain associate).
-A non-tenure track faculty, e.g., teaching or research faculty, also starts with Assistant, then is promoted to Associated, and Full prof.
-
-\subsection{Tenured or tenure-track faculty?}\label{sec:tenure-vs-tenure-track}
-
-\begin{center}
-  \includegraphics[scale=0.4]{files/c8.png}
-\end{center}
-
-Now that you know a bit about tenured and tenure-track faculty (\autoref{sec:faculty-types}).
-Which one should you choose as your adviser?
-The short answer is that either can be a good adviser, but they have different strengths and weaknesses.
-
-Tenure-track faculty, e.g., assistant professors, are likely young and active in research (they have to, to get tenure).
-Thus, they will devote more time to work with you and push you to do research and publish.
-However, they may not have as much experience in managing students and may not have as much funding (yet).
-
-Tenured faculty, e.g., associate and full profs., are likely older, more well-known, and have more experience in managing students.
-However, they might not push you as hard and expect you to figure things out yourself, i.e., you need to be independent.
-Some tenured faculty are also no longer active in research and are more involved with administrative responsibilities or with their startup companies (this means they will likely not take new students).
-\autoref{sec:faculty-types} gives more details on faculty types.
-
-% \subsection{Strong Faculty vs High-ranked School}
-% \begin{commentbox}[Thanh:]
-%   When considering PhD programs, we often wonder if we should prioritize a high-ranking university or a professor with a strong reputation?
-Of course, both are great, but it is hard to achieve both.
-I believe receiving some guidance from you  would be incredibly valuable.
-% \end{commentbox}
-
-% This is a good question and there doesn't seem to be any best answer or specific algorithm that you can use.
-But I'll throw out something for you to think about.
-To me, there are many good schools in the US and so you have some leeway in choosing.
-For example, I don't believe there's much difference among schools in top 5 (e.g., CMU, UIUC, UCSD, MIT, GTech are in the same equivalent class), top 10 (e.g., Michigan,  Stanford, UWash, UC-Berkeley, Cornell is another equivalent class), top 20, top 30, top 40, and so on (though I would say schools above 100 are pretty much in the same equivalent class).
-It's hard to say a top X is much much better than a top X+20 (e.g.,  Purdue at 15 is stronger than Yale at 35, but not \emph{that much} stronger as you would think).
-So keep that in mind, while ranking does matter (e.g., more active faculty, stronger students), the lower ranked universities are still pretty good, might have sub fields that are extremely strong (NCState is 40th, but in Software Engineering it is easily top 5), and many unique things.
-This is in contrast with universities in Vietnam where there are large gaps among the top 5, 10, and so on.
-
-\subsection{Should you contact a professor? How to get a desired reply?}\label{sec:contact}
-
-\begin{center}
-  \includegraphics[scale=0.6]{files/emails.png}
-\end{center}
-
-Faculty often receive ``cold'' e-mails from prospective students.
-Most of the time, we ignore these emails, but on some rare occasions, we do answer them.
-So how to write an email that gets our attention?
-
-First, if you want to contact a prof.
-to \emph{ask about your admission chance}, please \textbf{don't}.
-We don't know and can't answer because as explained in \autoref{sec:evalapps}, we don't make individual decisions and might not even be assigned to evaluate your application.
-It is the same as sending a paper draft to a journal editor and asking them if your paper has a chance.
-
-So how to get someone to look at your profile and give input?
-You could ask your professors, LoR writers, collaborators, or those who have previously applied.
-For this kind of feedback, ask someone you have a personal connection with.
-
-On the other hand, if you want to contact a prof.
-to ask about \emph{research opportunities}, or \emph{GTA/GRA} support, then \emph{yes}, I believe you should---it is \emph{worth it}.
-However, you really need to put effort into it and do it the \emph{right way}.
-
-First, read the prof's website, and see if they say something about contacting them.
-Many profs.
-explicitly state how prospective students should (or should not) contact them, e.g., using specific email subjects.
-In general, the best way to catch the prof.'s attention is to \emph{customize your email} for them.
-For example, read their papers, know what they work on, and see if you are interested in their research.
-Then email them and talk about how/why their work would match yours.
-In contrast, if you write a generic email that can be sent to multiple professors (e.g., if you just change some names and keywords in the email or copy and paste paper titles), you will not get a response.
-
-Below is a good example that I would likely reply.
-
-\begin{commentbox}[Dear Prof. Nguyen,]
-
-  I am writing to inquire about potential research opportunities as a GRA in your group at Mason. Currently I am an undergraduate student in Computer Science at UNIV and plan to graduate in May 2023.
-
-  I have read your TSE'21 paper on numerical invariant generation, and I am interested in this line of dynamic invariant research.
-I have worked (optional: with prof.
-Y at Z) on static program analysis and I think it could be used to tackle the spurious issues mentioned in your paper.
-I have a short paper at conference/workshop C and a project on symbolic execution (Github repo G).
-
-  ...
-
-  This is a good example because it is written just for me.  It shows that the student knows about my work on invariant generation and has a related background (paper C and project G).
-\end{commentbox}
-
-Finally, profs. are very busy so don't take it personally if you don't get anything from them (though I would be very surprised if such thoughtful emails get no replies!).
-
-\begin{commentbox}[Xiaokuan]
-  Applying for PhD and contacting a potential PhD adviser is a classic \textbf{`why me, why you'} problem,
-  similar to looking for a job in a company.
-  On a high level,
-  you need to show that you have done your \emph{homework}
-  regarding the professor and the university,
-  and clearly explain:
-  1) why do you think you are a good fit in professor A's group?
-  2) why do you want to be advised by professor A, not B?
-  3) why do you want to apply for university X, not Y?
-  If you don't want to spend time doing your homework,
-  the chance of getting a reply is close to zero.
-\end{commentbox}
-
-\begin{commentbox}
-  [Deepak:]
-  In my view, cold emails are not welcome by most faculty members and should be avoided.
-However, if one is already admitted to a program in some department, by all means, email the faculty you may be interested in working with, but do mention right at the beginning that you are already admitted to the program as well as several other universities.
-State specific areas (preferably specific topics-ML, robotics instead of AI).
-\end{commentbox}
-
-\subsubsection*{Additional Resources}
-\begin{itemize}
-  \item \href{https://yonatanbisk.com/emailing_professors.html}{A Note about Emailing Professors} by Yonatan Bisk
-  \item \autoref{sec:accept-postpone-decline} How to accept/postpone/decline PhD offers (and do it gracefully)
-\end{itemize}
-
-\subsection{Things to avoid when cold emailing a professor}
-
-\begin{itemize}
-
-    \item \textbf{Not customizing the email}. These emails typically talk about the student and not why they are interested in the professor's work. See writing tips in \autoref{sec:contact}
-    \item \textbf{Sending the same email to multiple professors}.
-Sending emails to multiple profs.
-using BCC or even CC (where we can see all addresses the email is being sent to) will result in no reply or a very harsh one on how unprofessional this is.
-    \item \textbf{Too much information}.
-Keep your email short and to the point.
-If we are interested, we will ask for more information.
-2--3 paragraphs are enough.
-Personally I also do not want to see course transcript or CV in the first email.
-    \item \textbf{Not being professional}, e.g., do not call the professor by their first name in the first email.  Use Prof. Lastname or Dr. Lastname (\autoref{sec:address}).
-    \item \textbf{Not reading the professor's website} and asking questions that are already answered there.
-\end{itemize}
-
-\chapter{Miscs and FAQs}\label{sec:faqs}
-
-\epigraph{``I want to share something with you – the three little sentences that will get you through life; number 1: Cover for me, number 2: Oh, good idea, Boss, and number 3: It was like that when I got here.''}{\textsc{The Simpsons}}
-
-\section{Why did I get rejected?}\label{sec:why-rejected}
-
-Many students lament that they get no interviews or are rejected (with strong profiles) and that the admission process seems random.
-However, while it is true that the process is not perfect, it is not random.
-Other than not having a strong profile (e.g., research potential, GPA, LoRs, SOP, poor interviews), here are some other reasons for rejections.
-
-\paragraph{Not a Good fit and Bad luck}  You could have an excellent profile, but if you are interested in a research area that the program does not have, you will not be admitted.
-Similarly, if no faculty is willing to advise you (e.g., they are on sabbatical or personal leave, do not have sufficient funding, or already have too many students), you will not be admitted.
-This is actually good for you, as you don't want to be in a program where no one can advise you.
-
-Related to this is that you just have bad luck and apply at the wrong time.
-For example, since 2024 there has been a huge surge in students interested in AI and NLP (thanks, ChatGPT!).
-Consequently, AI/NLP faculty might be overwhelmed and cannot consider many candidates, even those with excellent profiles.
-
-\paragraph{You aim too high} This is a very common reason: you have applied to schools that are too high for your profile.
-While being ambitious is good, you should understand how PhD admission works (e.g., by reading this document and realizing things such as having a good GPA doesn't mean much to top programs) and be realistic!
-
-Before applying, you should talk to your professors and ask them to give you an honest opinion on where you should apply.
-To increase their chances, many students apply to a range of schools, including ``safety'' ones.
-
-\paragraph{Overqualified or Lack Interest}  This might be paradoxical and is the opposite of aiming too high.
-However if adcom believes that you are likely to get admitted and go to a better program, they might not be so enthusiastic to admit you and want to save the spot for someone else.
-
-A related reason is that you did not show enough interest in the program.
-For example, your SOP did not mention any faculty or research area at the program, you did not respond to professors interested in interviewing you (which might be considered unprofessional and burn bridges), or during the interview you showed little knowledge, interest, or enthusiasm (because they are your safety schools) in the program or profs.
-
-\paragraph{Low English exam scores (e.g., IELTS or TOEFL)}  Profs.
-might not care much about these, but the college or the university often sets a minimum that you need to pass to be considered, especially for TA funding (e.g., you need to communicate well with students as a TA).
-Thus, while profs.
-are willing to argue for your case, they might be reluctant to go against the requirement of the university on standard exams.
-
-Note that English exams are not required for all students, e.g., domestic students and those from certain countries are waived from taking these exams.
-For example, Mason waives English exams for students from these \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-Mason#standard-tests-waiver-eligible-countries}{countries}.
-
-\paragraph{Red flags} Various types of red flags can cause concern to the adcom (e.g., too many STEM courses with low or withdrawn grades, plagiarism, cheating, or other academic dishonesty).
-Another one is that you have a history of jumping from one program to another (e.g., you have been in multiple different programs in a short time).
-Adcom members might have contacts in other programs and heard about your stories, or your LoRs might mention these.
-
-If you think you have these issues, you should address them in your SOP or ask your letter writers to do so.
-In general, these things are rare, but they do happen and cause concern to the adcom.
-
-\section{Timeline: Why are rejection letters sent so late?}\label{sec:late-rejection}
-
-%"Just reject me already!" is a common sentiment among applicants.
-Indeed, school will first send out admission offers to the top candidates.
-They do not send out rejection letters because there is still a chance that some of the top candidates will decline the offer.
-If they do, then the school will go to the next set of candidates and send out offers to them.
-This process continues until all spots are filled.
-This is why you might not hear back until late in the admission cycle.
-
-\paragraph{Acceptance Letters} Universities prioritize sending out acceptance letters first. This allows the admitted students to make decisions and plan for their studies.
-
-Some universities have rolling admissions, meaning they send out acceptance letters as they review applications (e.g., some universities sent out letters as soon as early February).
-Others have a specific date when they send out the first round of acceptance letters.
-
-\paragraph{Response Deadlines} Accepted students are usually given a deadline to make decisions on their offers, often around \textbf{April 15th} in the US.
-This is a standard deadline for PhD programs due to the Council of Graduate Schools (CGS) Resolution (\href{https://cgsnet.org/wp-content/uploads/2024/01/CGS_April15_Resolution_Jan312024.pdf}{see more}.
-After this date, CS programs can gauge how many slots remain unfilled.
-
-\paragraph{Waitlist} Most CS programs have a limited number of slots for PhD students, and thus put many good students on a waitlist.
-If accepted students decline the offer, then offers are sent to students on the waitlist.
-So if you see people getting accepted, that does not mean you are out yet.
-
-\paragraph{Rejection Letters} Schools typically start sending out rejection letters to remaining applicants \emph{after they have finalized their admissions decisions}.
-Thus, rejection letters are often sent out late (e.g., after April 15th or even much later).
-Not much you can do here, you can try to contact the school to ask about your status, but they might not reply, they might say they are still reviewing applications, or give you inaccurate information (e.g., you will hear in two weeks).
-In short, you just have to be patient and wait (however, beware that some schools do not send out rejection letters at all).
-
-\section{Increasing my admission chance}\label{sec:improve-your-chance}
-
-\begin{center}
-  \includegraphics[scale=0.2]{files/alpinist-climbing-peak-mountain-comic-hand-drawn-vector-illustration.jpg}
-\end{center}
-
-Given the high number of quality applicants and a limited number of spots, in addition to having a good application profile, you want to show something that makes you \textbf{stand out}.
-For example, even if you do not have research experience, you can talk about your personal projects, as long as they can help show you can do research.
-For example, if you have an open-source project that has lots of stars in Github, then you should mention it.
-If you often write technical, research-like blogs with many viewers, talk about them too.
-
-There are other things you might not think are important but can make you stand out.
-For example, if you have a strong background in a non-CS field but can be integrated with CS, e.g., you have a degree or background in \emph{dance} or \emph{music} and want to integrate them with CS, do talk about it.
-Are you a female or a minority in CS (\autoref{sec:urm})?
-Do you participate in outreach activities that help increase diversity and inclusion in CS?
-Diversity is highly valued in CS programs in the US, and experiences in these areas can make you stand out and get noticed from reviewers.
-
-\begin{commentbox}[Vu:]
-In his \href{https://matt.might.net/articles/how-to-apply-and-get-in-to-graduate-school-in-science-mathematics-engineering-or-computer-science/}{post}, Matt Might was initially unsure about an application.
-However, upon learning that the applicant had led a 100km hike in the Himalayas, he decided to accept the applicant.
-This is a good example of being \emph{stand out}, and I would also advocate for that student as this shows they have the persistence and determination required for research.
-\end{commentbox}
-
-% It is worth noting that in many cases,  an application could be rejected by most reviewers, however have a "champion", who doesn't even have to be in the Admission comittee, that strongly advocate for your profile.
-
-\section{How long do you evaluate an application?}\label{sec:ievaluate}
-
-On average, I spend about 10 minutes reviewing each application (less for clear rejections and more for potential acceptances).
-While this seems short, it is not that difficult to tell if an application is good or bad.
-In fact, this
-is twice what other faculty spend on average, e.g., Philip Guo spent \href{https://pg.ucsd.edu/PhD-application-tips.htm}{3--5 minutes} per application.
-
-For each application, our system compiles a single PDF file, which consists of a summary (degrees, GPAs, etc), transcripts, test scores, LoRs, a CV, SOP, and writing samples.
-I usually read in this order.
-I start with the \emph{summary}, checking for low GPAs or test scores below the university minimum (\autoref{sec:standard-tests}).
-I then skim the \emph{transcripts} for low grades in relevant courses, noting issues like ``many low grades in main courses or unknown international school with good GPA''.
-These are not as important as LORs or SOP but I read them first because the review system has questions about them (e.g., ``is GPA good?
-is IELTS sufficient?'')
-
-I read carefully \emph{strong LoRs} and skim weaker ones, noting either strong letters from well-known professors/researchers or generic content.
-I skim over \emph{CV} and look for publications, research experiences, and notable achievements, noting things like ``published papers in top venues'' or ``gold medal in an international competition.''
-I read strong \emph{SOPs} thoroughly (and skim weak ones).
-I note whether the SOP is exciting, research-oriented, standing out, e.g., applicant is minority (\autoref{sec:urm}), and tailored to our program (e.g., if they are familiar with the work of some faculty or have talked to them).
-These notes are entered into the evaluation system.
-
-Finally, I enter my decision, which is usually either a rejection or an offer of admission with full funding (e.g., from TA).
-I also recommend very strong candidates for the University Presidential fellowship.
-Note that while the system has other decision options, eg., admit without funding, provisional admission (e.g., if they need to take some courses), I rarely use them, simply because we already have too many good students and do not have the resources to support them.
-
-\subsection{How Decisions are Made}\label{sec:how-decisions}
-
-After all reviewers have evaluated an application, adcom chair will review all evaluations, look at entered notes, and ask reviewers to discuss and resolve discrepancies to reach a consensus (e.g., a reviewer wants to accept but the other wants to reject).
-Typically, the decision is made entirely by the reviewers; there is no involvement from the adcom chair, department chair, or others.
-In most cases adcom members, even those reviewing the same application, make decisions independently and do not talk to each other (just a common practice to avoid biasing).
-In some rare cases we might (\autoref{sec:adcom-discuss}).
-
-Even if \emph{all} reviewers recommend acceptance, the application is not automatically accepted, especially if no faculty is willing to advise the student.
-For example, if the student is interested in a research area that no faculty is working on, or an area where no faculty is taking new students (e.g., AI/ML where faculty likely already has many students) then the student will not be admitted.
-This is increasingly common as the number of applicants increases significantly in recent years in CS.
-
-However, if the student has contacted a faculty member and that faculty member is interested in the student and has made this known to the admission committee, then the student is likely to be admitted, \emph{even if the faculty does not review the application or has funding} (thus the importance of contacting faculty).
-Similarly, if the student mentioned a faculty member in their SOP, adcom might ask that faculty member to look at their application and if they are interested in the student.
-Even if the student has a weak profile (but still passes the minimum requirement from the university), they might be admitted if a faculty member is willing to take them.
-Adcom members, especially in the US, are very reluctant to go against the faculty's decision (e.g., if a faculty wants to admit a student, we are not going to reject them).
-
-\section{Do adcom members talk to each other?}\label{sec:adcom-discuss}
-
-However, in some rare cases we do talk to each other.
-For example, when there are discrepancies in evaluations, in which case the adcom chair will ask reviewers to discuss the application to reach a consensus.
-We also talk to each other for interesting or strong applications, e.g., how to recruit this student or who should be the adviser.
-If the student mentioned a faculty member in their SOP, we might ask that faculty member if they are interested in the student.
-
-\section{Do adcom members read all applications?}\label{sec:read-all}
-
-While adcom members (in fact, any faculty) can view any application, we only review those that are assigned to us, which are already too many.
-Adcom chair will assign applications to reviewers based on their expertise (e.g., if a student says they want to do SE or interested in working with me), and reviewers will only review those applications.
-Occasionally we might look at other applications, e.g., if we know the student or have some special circumstances (e.g., if that student contacted me, I know that student, they are from Vietnam, etc).
-However, even if we look at them, we usually do not get involved in their evaluation.
-
-\section{Member of an underrepresented or LGBTQ+ groups}\label{sec:urm}
-
-A common question is whether you should mention, e.g., in your SOP, that you are a member of an underrepresented and minority (URM) group\footnote{URM in CS in the US include African Americans, Hispanic Americans, American Indians/Native Alaskans, Native Hawaiians/Pacific Islanders (excluding Asian Americans), and multiracial students identifying with at least one of the previously listed URM categories.} or LGBTQ+ group.
-This is a strong concern for international (and even domestic) students, who may have encountered discrimination in their countries or personal experiences.
-%This question arises often for international (and even domestic) students.
-The main worry is that they might get discriminated against, as often happen in their countries or personal experiences.
-%, and (ii) mentioning this might make them look like they are asking for special treatment.
-
-In my personal opinion, highlighting URM or LGBTQ+ identity can \emph{be beneficial}, especially if you can articulate how your diverse experiences contribute to diversity and inclusion in academics.
-Many universities have a strong commitment to diversity and inclusion and actively
-\emph{recruit students from underepresentative backgrounds} (e.g., some \href{https://cse-climate.engin.umich.edu/reports/climate-dei-reports/cse-climate-and-dei-report-2022-2023/#grad-ethnicity}{stats from UMich CSE} and \href{https://www.Mason.edu/news/2022-09/mason-now-top-10-public-university-diversity-innovation-and-cybersecurity-education-us}{Mason} often promotes itself as one of the most diverse universities in the US).
-Moreover, many scholarships and fellowships are created specifically for URM and LGBTQ+ students, which you should explore if qualify.
-
-Understandably, you might feel uncomfortable disclosing this, fearing you would lose a chance of working with a faculty due to their bias.
-However, if someone has this issue, then you might not want to work with them anyway (and such bias is not acceptable in academics and likely would not be tolerated by the university).
-In my experience, it is the opposite: \emph{many profs.
-actively seek to work with students from diverse backgrounds and view diversity as an asset to their research group}.
-
-Moreover, promoting diversity aligns with the goal of broadening participation in computing (BPC) that is strongly supported within academia (e.g., see the \href{https://plans.bpcnet.org/GeorgeMasonUniversity_ComputerScience_DepartmentalBPCPlan.pdf}{BPC plan at Mason CS}) and government funding agencies such as NSF (e.g., see \href{https://new.nsf.gov/cise/broadening-participation}{here} and \href{https://www.nsf.gov/pubs/2022/nsf22125/nsf22125.jsp}{here}).
-Also, sometimes providing this information can make it easier for faculty to apply for certain scholarships or fellowships to support you, e.g., from a private donor that has offered a fellowship only to students from certain backgrounds.
-
-Ultimately, this is a personal choice and you do what you feel comfortable.
-If you do not feel comfortable disclosing this, you should not feel pressured to do so and you should also ask your LoR writers to not disclose this information.
-Note that you can still talk about your experiences and perspectives without mentioning your identity.
-If uncertain, you can seek additional guidance from your close professors and collaborators However, it is important to recognize that diversity is strongly valued and celebrated within CS programs in the US, and that you should feel empowered to share your unique experiences and perspectives.
-
-%https://www.edsurge.com/news/2018-05-18-how-cornell-university-diversified-its-incoming-phd-computer-science-student-body
-
-%https://academia.stackexchange.com/questions/130591/is-it-standard-for-us-based-universities-to-consider-the-ethnicity-of-an-applica
-
-%https://www.cs.cornell.edu/~bindel/paper/diversity.pdf
-
-%https://cs.brown.edu/degrees/doctoral/applications/helpful-resources-applying-computer-science-phd-programs/
-
-\section{My undergrad was not in CS or related areas}\label{sec:non-stem}
-
-You still can apply to PhD in CS \emph{as long as} you can demonstrate you are ready for it through your background, research experience, LoRs, statements, etc.
-You might be even able to leverage this to make your profile stand out as mentioned in \autoref{sec:improve-your-chance}.
-
-A main question adcom has for a non-CS or non-STEM student is if you have the sufficient technical background obtained through core CS courses.
-So you need to show that you have such knowledge through your coursework, projects, or research.
-For example, if you have taken a course on Algorithms, even online ones like Coursera, you can talk about it in your SOP.
-If you have done a project that requires knowledge of OS or have a professional certification (e.g., A+) through work, you can talk about it.
-If you have done research that requires knowledge of Discrete Maths, you can talk about it.
-You can also ask your LoR writers to talk about your technical background.
-In summary, in your application, convince us that you have the background to do CS PhD research.
-
-\paragraph{Core CS topics} Common CS knowledge that you should know are:
-\begin{itemize}
-  \item Programming Foundation (e.g., programming concepts in some modern languages)
-  \item Discrete Math (e.g., logic, set theory, proof techniques)
-  \item Data Structures and Algorithms (e.g., linked lists, trees, sorting, searching)
-  \item Computer OS or Systems (e.g., memory management, file systems, processes)
-\end{itemize}
-
-In short, you \emph{do not need} to formally take CS courses, you just need to show that you have this essential knowledge, e.g., through the mentioned ways.
-Many universities are well aware that incoming graduate students might not have all the technical background, so they often have a \emph{``bridge''} courses to help students catch up.
-For example, Mason has four bridge courses corresponding to the four core areas above that incoming students can take to catch up on their CS knowledge.
-
-\begin{commentbox}[Vu:]
-  I would advocate for a non-STEM student who shows that they have a strong drive for CS by studying core CS knowledge through various channels (e.g., self-study through online courses, projects, etc.).
-I have seen many students with non-CS backgrounds who are very successful in CS PhD.
-I also have seen many students with CS background who are not successful in CS PhD.
-So it is not about your background, it is about your drive and passion for CS research.
-\end{commentbox}
-
-\section{Is an MS required for admission to PhD in CS?}\label{sec:msrequirement}
-No.
-In fact, students can get MS degree ``along the way'' to PhD, e.g., after finishing the 2-year course work.
-However, MS can help admission if it gives research experience or is from a more well-known school than your undergrad institution (\autoref{sec:your-school}).
-
-If you have an MS then some course work \emph{might be} transferred for course credits, which \emph{might} save some time. But don't count on this.
-
-\begin{commentbox}[Vu:]
-    I start my PhD with an MS in CS from a US university.
-I found that the MS helped me a lot in research, but not with course work and I had to retake many courses.
-In general, don't worry if you don't have an MS.
-But also don't feel that you wasted your time if you have an MS, as it can help you in research.
-  \end{commentbox}
-
-\section{How long to complete the CS PhD program?}\label{sec:time}
-
-Typically, it takes 5--7 years for CS PhD in the US.
-This can be longer than CS PhDs in other countries, which might require MS first (recall from \autoref{sec:msrequirement} that CS PhD programs in the US do not require MS and you can get MS along the way to PhD).
-Within these 5--7 years, CS PhD students often take a ``leave of absence'' for 1--2 semesters to do internships at companies and research labs.
-
-\begin{center}
-  \includegraphics[scale=0.5]{files/c4a.png}
-\end{center}
-
-The first two years you typically take coursework (somewhat equivalent to an MS study), find an adviser, and learn how to do research.
-The next 2--3 years you focus on your research, form a dissertation topic, and get results published.
-The last 1--2 years you continue to publish, write and defend your dissertation, and look for a job.
-In many cases you might take a summer or two off to do an internship to get additional research opportunities.
-The PhDComics figure above shows the ``ambition'' level of a PhD student over their years of study (they miss the 6--7th year when the ambition is \emph{``Just let me graduate''}).
-
-% \begin{commentbox}[Vu:]
-%     I start my PhD with an MS, and it took me 7 years (Fall'2007-- Fall'14). I spent half a year doing an internship at the Naval Research Lab.
-%   \end{commentbox}
-
-\section{What's the difference between PhD and MS in CS?}\label{sec:phd-vs-ms}
-
-While both are graduate degrees, they are different in terms of objective, admission requirements, course requirements, duration, and funding as summarized in \autoref{tab:phd-vs-ms}.
-
-\begin{table}
-  \caption{PhD vs. MS}\label{tab:phd-vs-ms}
-  \begin{tabular}{c|c|c}
-  \toprule
-  \textbf{} & \textbf{PhD} & \textbf{MS} \\
-  \midrule
-  Objective & Industry & Research \\
-  Admission Req & Research experience & No research experience \\
-  Coursework Req & Yes & Yes (research is much more important)  \\
-  Duration & 2 years & 5--7 years \\
-  Adviser Req & No & Yes \\
-  Funding & No & Yes \\
-  \bottomrule
-  \end{tabular}
-\end{table}
-
-\begin{itemize}
-
-  \item Objective: an MS is typically to prepare you for industry, while a PhD is to prepare you for research.
-Some MS has thesis option but the research is much lighter compared to a PhD, which is research-focused.
-
-  \item Admission requirements: MS also requires a good GPA, LoRs, SOP, and test scores, but research experience is not required.
-PhD requires all of these, but research experience is a must.
-Moreover, MS programs are typically much less competitive than PhD programs.
-In fact, many MS program are viewed as a revenue source (\emph{cash cow}) for the department, as they are not funded and students pay tuition.
-
-  \item Course requirements: MS has a specific number of courses (typically can be done in 2 years).
-You graduate with an MS when you're done with the courses.
-PhD also has coursework requirements (typically taken in your first 2 years) but focuses on research after that.
-You graduate with a PhD when you have done enough research and written a dissertation (which is typically way harder and takes much longer time than coursework).
-
-  \item Duration: an MS typically takes 2 years while a PhD takes 5--7 years (or even longer).  Many students get an MS along the way to a PhD, e.g., after finishing the 2-year course work.
-
-  \item Adviser: MS students typically do not have an adviser (if you do thesis option then you will have one), while PhD students need an adviser who guides them in their research.
-
-  \item Funding: MS is typically \emph{not funded}, while PhD is typically funded (e.g., through TA or RA as explained in \autoref{sec:funding}).
-
-\end{itemize}
-
-\section{PhD in the US vs. Other Countries}\label{sec:non-us-differences}
-
-\begin{table}
-\caption{Comparison of US and Other PhD Programs}\label{tab:us-vs-other}
-\begin{tabular}{c|c|c}
-\toprule
-\textbf{Aspect} & \textbf{US PhD Programs} & \textbf{Other Countries} \\
-\midrule
-Duration & 5-7 years & 3-5 years \\
-MS Required & Not required & Often required \\
-Coursework Required & Yes (first 2 years) & No \\
-Research Proposal Required & No & Yes (in some countries) \\
-Academic (Faculty) Job & Direct & Postdoc \\
-Work Life Balance & Less & More \\
-\bottomrule
-\end{tabular}
-\end{table}
-
-\autoref{tab:us-vs-other} summarizes the main differences between CS PhD in the US and other countries. % This is based on my experience and what I heard from others who did PhD in other countries.
-
-\emph{MS requirement and PhD duration}:  CS PhD programs in the US do not require an MS degree (as mentioned in \autoref{sec:msrequirement} and \autoref{sec:time}).
-In contrast, many other countries require having an MS degree before joining a PhD program.
-This means that US PhD programs are longer (5--7 years, 2 of which are coursework) than other countries (3--4 years, no coursework).
-
-\emph{Project proposal}: in many countries, you have to choose a project and adviser \emph{during} the application process (e.g., you write a proposal to a potential adviser).
-But this allows you to start your research right from the beginning.
-
-In the US, you often start your PhD without an adviser or project and find them later.
-You will have two initial years to explore and find an adviser and research topic.
-After that you explore projects with that adviser before settling on one (can take a couple of years).
-
-\emph{Course work}: as mentioned above in the US you will spend the first couple of years taking classes and exploring potential adviser and research topics. 
-After that, you have to pass a series of exams during your PhD, e.g., qualifying exam, comprehensive exam, thesis proposal defense\footnote{The word ``ABD'' (all but dissertation) is used in the US to refer to a PhD candidate who has finished all course work and exams and only needs to write and defend their dissertation.}.
-
-In other countries, you start your research right away, e.g., you immediately work on the research project you proposed with the adviser you chose.
-Moreover, you might not have exams like those in the US or only have to do a few of them.
-
-\emph{Funding}:  In many countries, funding comes from the university or the gov't.
-This funding often has a fixed duration (e.g., 3 or 4 years).
-In the US (\autoref{sec:funding}), funding (e.g., RA) comes directly from your adviser (no fixed duration).
-There are also fewer TA opportunities in European universities compared to the US.
-
-\emph{Academic position after PhD}: In other countries, PhD graduates interested in academia typically apply for additional research appointments, i.e., postdocs, and then consider faculty positions.
-
-In the US, PhD graduates often apply directly for faculty positions. Postdoc for US graduates is no longer a popular option as it was before.
-
-\emph{Work-life balance}: PhD students in the US are often said to be overworked compared to other countries, e.g., in Europe.
-This is partly due to the longer PhD program and that US PhD students are often paid through TA, which requires them to do TA in addition to their own research.
-In contrast, PhD students in other countries are often paid through fellowships, which might not require doing TA.
-
-\section{How do I call or address a professor?}\label{sec:address}
-
-\begin{center}
-  \includegraphics[scale=0.5]{files/c5.png}
-\end{center}
-
-If you're reaching out to a professor for the first time,  address them as Prof.
-or Dr.
-Lastname.
-Many international students use Prof.
-or Dr.
-FirstName LastName, but this can come across as if you're simply copying and pasting names.
-It's not necessary, so stick with Prof.
-or Dr.
-Lastname.
-
-Furthermore, avoid using Mr., Mrs., Miss, or the professor's first name if you're not acquainted with them yet.
-As you become more familiar with your prof and depending on their preferences, you may transition to addressing them by their first name.
-For example, I prefer that my students and colleagues call me Vu. Some students call me \emph{Dr. Vu}, which I find a bit amusing but am totally fine with it.
-
-\begin{commentbox}[DK:]
-  I was amused to read this as if I recall correctly, you never called me by my first name when you were at UNM. You always called me Prof. And, many times, I would jokingly call you back as Prof. Vu.
-  \tcblower
-  \textbf{Vu}: Yes, for some reason I enjoy calling you ``Professor'' (without appending a last or first name).  The use of Prof. Vu may have foreshadowed my future in academia.
-\end{commentbox}
-
-Note that in some universities the formal title Dr.
-Lastname is preferred over Prof.
-Moreover, be aware that not all faculty members may hold a PhD, in which case using Prof.
-Lastname is a suitable alternative.
-	You just need to observe and follow the conventions at your particular institution.One way to determine how to address a professor is to observe how they sign their emails or how they introduce themselves in class.
-For example, I introduce myself as ``Vu'' in class.
-
-\paragraph{Referring to professors you know} Because you are already familiar with these individuals, you can just informally use their names if they are OK with it as mentioned above (or Dr./Prof., if you want to be formal).
-You can also include their institution if it makes it more precise.
-For example, I can say:  \emph{``I did my postdoc with Jeff Foster at Univ.
-of Maryland''}.
-
-Do not include ranking (e.g., Assistant, Associate, Scientist, ...) when referring to someone.
-I see many international students include a lengthy title of people they know, e.g., \emph{I am advised by Asst.
-Prof.
-X, and I also collaborate with Distinguished Scientist Y}.
-This is \emph{not necessary} and makes it look like you're trying to show off your connections.
-These nuances represent some cultural differences that you may encounter and will gradually adapt to.
-More on cultural differences in \autoref{sec:cultural}.
-
-\section{How much do \emph{I} cost?}\label{sec:ra-cost}
+\subsection{How much do \emph{I} cost?}\label{sec:ra-cost}
 
@@
 
+\chapter{Choosing Schools and Professors}\label{sec:schoolsandprofs}
+
+\epigraph{\vspace{-0.2in} The answer to life's problems aren't at the bottom of a bottle. Heh heh! They're on TV!}{\textsc{The Simpsons}}
+
+Choosing a school and an adviser is clearly among the most important things in your mind when you apply and especially when you get admitted.
+This is further complicated due to cultural differences and the unfamiliarity of international students with the US higher education system.
+This section aims to mitigate some confusion and help you make informed decisions.
+
+\section{Choosing a University}\label{sec:choosing-university}
+
+We will first discuss universities in the US that offer PhD in CS. Then we will talk about how to select them.
+
+\subsection{Schools that offer PhD in CS}
+
+Most US universities have CS studies, but many of them do not have a PhD program in CS.
+These universities might offer just Bachelor's degrees (e.g., BS) and no graduate studies (i.e., no MS or PhD degrees), or they just offer MS programs (but no PhD).
+For example, Penn State in University Park has PhD in CS,  but Penn State in Harrisburg only has BS and MS in CS, and Penn State in York only has BS in CS.
+On the other hand, multiple locations of the University of Texas, e.g., Austin, Dallas, and Arlington, have PhDs in CS.
+
+Thus, if your goal is PhD in CS, you have to aim only for schools offering such a degree.
+%This might sound obvious but many students (even those in the US) get confused due to the large number of universities.
+While this can be confusing due to the large number of universities in the US, a little research, e.g., searching for PhD in CS from the school website, will help you find out.
+Schools listed in \autoref{sec:ranking} have PhD programs in CS, so you can start there.
+
+% \subsection{R1, R2, ...}
+
+\subsection{Selecting and Ranking Schools}\label{sec:selecting-ranking-schools}
+\begin{center}
+  \includegraphics[scale=0.6]{files/c1.png}
+\end{center}
+
+International students often put schools into two bins: (i) top schools that they dream about, and (ii) everything else.
+They commonly use rankings from US News, which is not transparent and questionable\footnote{\url{https://cra.org/cra-statement-us-news-world-report-rankings-computer-science-universities/}}.
+Sometimes they evaluate based on the reputation of the school's undergrad program or the reputation of the school's non-CS programs such as medical, math, or physics.
+They even rank universities based on popular places they know in the US, e.g., California, Texas, and New York.
+
+Instead of these superficial criteria, you should consider the CS program and the research interests of faculty members.
+You can learn about these using resources such as \href{https://csrankings.org}{CSRankings.org}.
+You will be very surprised to learn that a school that you didn't know much about can have very strong research in your topic (and vice versa, a school you thought highly about might have no faculty working in the research field you're interested in).
+This is also a good way to learn about individual faculty, e.g., who works on what, and well-known CS conferences\footnote{In CS (and probably only in CS), conferences, not journals, are often the main venue to publish research findings (see \href{https://homes.cs.washington.edu/~mernst/advice/conferences-vs-journals.html}{why here}).}.
+%\autoref{sec:ranking} gives the top 50 CS programs in the US according to CSRankings.
+
+\begin{commentbox}[Dat:] Most Vietnamese students, including those from top schools, \textbf{do not know} about CSRankings.
+Maybe applicants who worked at top research places such as VinAI would know about it.
+\end{commentbox}
+
+\paragraph{What matters to you?} While many find CSRankings useful, it is not perfect, and in general, rankings are superficial, and you need to do more research to be informed and make better decisions.
+For example, you might prefer places with a large community from your country (e.g., Northern VA has a large population of Vietnamese).
+You might prefer places with high-tech companies (Seattle or Silicon Valley), with many outdoor activities (e.g., hiking, skiing), or with better weather (\emph{``PhD can be depressing, so would you rather be depressed in California or New York?''}).
+You might need to think about the cost of living, e.g., living in California is more expensive than in Virginia, or places with high crime rates (note that while some universities might be in a high-crime city but the campus itself is safe).
+
+If you get admission to several places, you should consider attending Open Houses (\autoref{sec:accepted}) and contact profs.
+that you're interested in at those places and talk to them.
+They would be more willing to chat with you now that you have been admitted.
+Ask questions about \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Answers-to-Ph.D-Advisor-Guide}{their work}, how they manage students, and their expectations.
+You can even ask to contact their students.
+
+\begin{commentbox}[Hung:]
+  I always encourage the students I admitted to talk with my students and the students of other faculty in other schools who admitted them.
+You will unlikely hear straight-out complaints from current students in a prof’s group.
+But sometimes what is important are things that they (current students) don’t tell you.
+Pay attention to their ``level of excitement'' being in the group.
+\end{commentbox}
+
+\begin{commentbox}[Xiaokuan:]
+  Chinese students often only look at US News rankings when selecting their PhD universities (I did that, too, when I was applying for PhD positions).
+  Now that I am a professor, I find it to be the least promising way.
+  %
+  The reason is that US News does not provide a good metric for evaluating the quality of the PhD program.
+  %
+  If you want to do great research, CSRankings is the best way to find good and active professors (which did not exist when I was applying),
+  since it solely focuses on publications at top-tier CS conferences.
+  %
+  Also,
+  I think PhD is not only about research;
+  you need to also consider your daily life there since you will (probably) stay for at least five years.
+  %
+  You might regret it if you did not consider this seriously before applying.
+  %
+\end{commentbox}
+
+\subsection{PhD in other related fields: CE, IST, Cybersecurity}\label{sec:related-fields}
+
+You \emph{do not} need to do a PhD in CS to work in CS areas.
+For example, in addition to a traditional CS department, Mason has IST and Cybersecurity departments, both of which have faculty with PhD in CS and work on CS topics (e.g., AI, Security, Robotics).
+So you still can do CS research and publish in CS-focused venues even if you're not in a traditional CS program.
+It is common to see faculty with PhD in CS in a non-CS department as well as faculty with non-CS PhD in a CS department.
+
+However, if your goal is a PhD in CS, then you need to be in the CS dept \emph{and} advised by a CS faculty.
+A non-CS faculty can serve in the PhD dissertation committee (common) or \emph{co-advise} (somewhat rare), but your main PhD adviser will likely be a tenure-line faculty in CS (\autoref{sec:faculty-types}).
+For example, a prof. in Stats or Math might be able to serve as a co-adviser, but not as a sole adviser of a student in a CS PhD program. 
+If in doubt, check with the CS department for their requirements.
+
+For this specific reason,  CSRankings includes only faculty who can advise CS PhD students.
+I also have compiled a \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{list} of Vietnamese faculty who can advise PhD students.
+\autoref{sec:faculty-types} talks more about who can serve as your PhD adviser.
+
+\section{Choosing an Adviser}
+There is no one-size-fits-all answer to this question. The best adviser is the one that you can work well with.  But how do you find such a person?
+
+Fortunately, while some non-US programs require finding an adviser and research topic before starting the PhD (\autoref{sec:non-us-differences}), CS PhD programs in the US will typically give you a couple of years to ``shop'' for advisers and research topics.
+This is especially true if you're admitted with TA (\autoref{sec:ta}), which gives you time to explore and find an adviser.
+
+\subsection{Finding the right adviser}
+
+Here is some advice for finding a prof.
+Assuming you're not familiar with any particular one, then first search for profs.
+that share similar research interests.
+For example, in CSRankings, if you want to work with PL, you can search for those published in PL conferences.
+If you want to work with SE \emph{and} AI, you can search for faculty who work in both SE and AI.
+Then you research about that prof.
+(e.g., go to their website) and contact them (\autoref{sec:contact}).
+
+\begin{commentbox}[Xiaokuan:]
+  Whether the student's research interest matches that of the adviser is very important;
+  if there is a mismatch,
+  either the student or the adviser has to make compromises,
+  which often leads to disagreements or conflicts.
+  IMO, the adviser should be the one who \emph{guides}  students to do research while allowing students to pursue their own interests,
+  instead of \emph{dictating} their research.
+\end{commentbox}
+
+Another way is taking graduate-level courses in the topics you are interested in.
+Many profs teach \emph{special topics courses} and \emph{research seminars}, and they might be recruiting students.
+Do well in the class, answer questions, talk to the prof.
+after classes, etc---being stand out.
+Many profs, including myself, prefer taking in new students this way.
+It gives both the prof and student more time, e.g., a whole semester, to work and evaluate the relationship before making any commitment (sounds like a marriage!).
+
+You can also ask to do an independent study or research with a prof.
+This can be informal (no credit) and takes place during the summer or winter break.
+For example, I do this with several students, some of whom are undergrads.
+Many will drop out because they find they don't like my research, but some find that they like the work.
+
+Ultimately, choose a prof.
+that fits you by communicating with them, taking their courses, meeting and asking them questions, and talking to their current students.
+It will take time and effort, but since you will be working with this person for 5+ years, it is important to try to find the right one.
+
+\begin{commentbox}[Thanh:]
+  In my opinion, having a well-suited adviser is crucial for a successful PhD and research career.
+One effective approach to finding a suitable professor is by working with a professor during your undergraduate studies.
+An exemplary instance is VinAI's residency program, where residents collaborate with professors from the US for two years before applying to PhD programs.
+Many VinAI residents have achieved remarkable results and gained admission to prestigious US universities.
+Unfortunately, VinAI's resident program is limited to AI research.
+
+  In other fields, e.g.
+Software Engineering, Vietnamese students face challenges in reaching US professors.
+Do you have any tips for Vietnamese students who want to connect with US professors and work as research assistants?
+  \tcblower
+  \textbf{Vu}: \autoref{sec:contact} shows how to contact a professor for research opportunities.
+Many probably will say no (or do not reply) as they do not have the bandwidth to take on random students, but some may say yes if they see a potential fit.
+\end{commentbox}
+
+\subsubsection*{Additional Resources}
+\begin{itemize}
+  \item \href{https://www.cs.columbia.edu/wp-content/uploads/2019/03/Get-Advisor.pdf}{The Definitive ``what do I ask/look for'' in a PhD Advisor Guide}
+\end{itemize}
+\subsection{Who can serve as a PhD adviser?}\label{sec:faculty-types}
+
+\emph{Not} every faculty can serve as your official PhD adviser.
+Let's try to understand the different types of faculty and their roles.
+For example, you might hear about tenured,  tenure track, and teaching faculty.
+You might also hear about assistant, associate, full, adjunct, emeritus, teaching, and research professors.
+Here is a primer on these terminologies.
+
+\paragraph{Types of Faculty} There are typically two types of faculty: tenure-line and teaching.
+The main difference is their focus: research, teaching, or both.
+The type of faculty can affect their ability to serve as your adviser.
+
+\textbf{Tenure-line faculty}, consisting of tenured and tenure-track profs., focus on research, which includes publishing papers, obtaining grants, and mentoring PhD students.
+They often have a low teaching load (e.g., 1 per semester).
+\emph{Tenured} faculty are professors who have been promoted to permanent positions (informally, very hard to fire them).
+\emph{Tenure-track} faculty are (more likely young) professors who are on the track to get tenure.
+For most schools, tenure-line is the main type of faculty that can serve as a formal adviser of PhD students.
+\autoref{sec:tenure-vs-tenure-track} talks more about choosing between tenure and tenured-track professor as your adviser.
+
+\textbf{Teaching faculty} (sometimes called instructional faculty or professor of practice) mainly focuses on teaching.
+They typically teach 3--4 classes per semester (which is quite a lot) and \emph{do not} have research responsibilities, e.g., they do not have to worry much about publishing papers or obtaining grants.
+As such they typically do not have funding (e.g., RA~\autoref{sec:funding}) to support PhD students.
+Depending on the university, they might be able to \emph{co-advise} and mentor PhD students with a tenure-line faculty but typically do not formally advise and graduate PhD students.
+
+Note that at many universities do not allow \emph{non-CS profs}, even in tenure-line ones, to solely advise and graduate PhD students in CS (see \autoref{sec:related-fields}).
+
+Some universities have \textbf{research faculty} or research scientists, who are typically non-tenure line faculty and focus on research (e.g., publishing papers and getting grants). 
+They can recruit students and provide RA funding for them. 
+They can also teach classes but typically have a lower teaching load.
+For some universities, they \emph{can} solely advise and graduate PhD students, but for others, they cannot.
+
+\textbf{Adjunct faculty} is not full-time, e.g., they might be working in industry and teach a class or two for fun.
+\textbf{Emeritus faculty} means they are retired from the university, e.g., they could move to the industry, but still have some affiliation with the university.
+Due to their roles, these profs.
+typically do not advise PhD students.
+
+\paragraph{Ranks} Faculty have ranks such as assistant, associate, and full, regardless if they are tenure-line or teaching.
+Assistant means new faculty, associate means they have been promoted, and full means they are senior.
+Typically, a tenure-track faculty starts with being an assistant professor and then gets tenure and promoted to associate (typically after 6 years) and then full (time varies, some become full within 3--4 years, some 10+ years, some remain associate).
+A non-tenure track faculty, e.g., teaching or research faculty, also starts with Assistant, then is promoted to Associated, and Full prof.
+
+\subsection{Tenured or tenure-track faculty?}\label{sec:tenure-vs-tenure-track}
+
+\begin{center}
+  \includegraphics[scale=0.4]{files/c8.png}
+\end{center}
+
+Now that you know a bit about tenured and tenure-track faculty (\autoref{sec:faculty-types}).
+Which one should you choose as your adviser?
+The short answer is that either can be a good adviser, but they have different strengths and weaknesses.
+
+Tenure-track faculty, e.g., assistant professors, are likely young and active in research (they have to, to get tenure).
+Thus, they will devote more time to work with you and push you to do research and publish.
+However, they may not have as much experience in managing students and may not have as much funding (yet).
+
+Tenured faculty, e.g., associate and full profs., are likely older, more well-known, and have more experience in managing students.
+However, they might not push you as hard and expect you to figure things out yourself, i.e., you need to be independent.
+Some tenured faculty are also no longer active in research and are more involved with administrative responsibilities or with their startup companies (this means they will likely not take new students).
+\autoref{sec:faculty-types} gives more details on faculty types.
+
+% \subsection{Strong Faculty vs High-ranked School}
+% \begin{commentbox}[Thanh:]
+%   When considering PhD programs, we often wonder if we should prioritize a high-ranking university or a professor with a strong reputation?
+Of course, both are great, but it is hard to achieve both.
+I believe receiving some guidance from you  would be incredibly valuable.
+% \end{commentbox}
+
+% This is a good question and there doesn't seem to be any best answer or specific algorithm that you can use.
+But I'll throw out something for you to think about.
+To me, there are many good schools in the US and so you have some leeway in choosing.
+For example, I don't believe there's much difference among schools in top 5 (e.g., CMU, UIUC, UCSD, MIT, GTech are in the same equivalent class), top 10 (e.g., Michigan,  Stanford, UWash, UC-Berkeley, Cornell is another equivalent class), top 20, top 30, top 40, and so on (though I would say schools above 100 are pretty much in the same equivalent class).
+It's hard to say a top X is much much better than a top X+20 (e.g.,  Purdue at 15 is stronger than Yale at 35, but not \emph{that much} stronger as you would think).
+So keep that in mind, while ranking does matter (e.g., more active faculty, stronger students), the lower ranked universities are still pretty good, might have sub fields that are extremely strong (NCState is 40th, but in Software Engineering it is easily top 5), and many unique things.
+This is in contrast with universities in Vietnam where there are large gaps among the top 5, 10, and so on.
+
+\subsection{Should you contact a professor? How to get a desired reply?}\label{sec:contact}
+
+\begin{center}
+  \includegraphics[scale=0.6]{files/emails.png}
+\end{center}
+
+Faculty often receive ``cold'' e-mails from prospective students.
+Most of the time, we ignore these emails, but on some rare occasions, we do answer them.
+So how to write an email that gets our attention?
+
+First, if you want to contact a prof.
+to \emph{ask about your admission chance}, please \textbf{don't}.
+We don't know and can't answer because as explained in \autoref{sec:evalapps}, we don't make individual decisions and might not even be assigned to evaluate your application.
+It is the same as sending a paper draft to a journal editor and asking them if your paper has a chance.
+
+So how to get someone to look at your profile and give input?
+You could ask your professors, LoR writers, collaborators, or those who have previously applied.
+For this kind of feedback, ask someone you have a personal connection with.
+
+On the other hand, if you want to contact a prof.
+to ask about \emph{research opportunities}, or \emph{GTA/GRA} support, then \emph{yes}, I believe you should---it is \emph{worth it}.
+However, you really need to put effort into it and do it the \emph{right way}.
+
+First, read the prof's website, and see if they say something about contacting them.
+Many profs.
+explicitly state how prospective students should (or should not) contact them, e.g., using specific email subjects.
+In general, the best way to catch the prof.'s attention is to \emph{customize your email} for them.
+For example, read their papers, know what they work on, and see if you are interested in their research.
+Then email them and talk about how/why their work would match yours.
+In contrast, if you write a generic email that can be sent to multiple professors (e.g., if you just change some names and keywords in the email or copy and paste paper titles), you will not get a response.
+
+Below is a good example that I would likely reply.
+
+\begin{commentbox}[Dear Prof. Nguyen,]
+
+  I am writing to inquire about potential research opportunities as a GRA in your group at Mason. Currently I am an undergraduate student in Computer Science at UNIV and plan to graduate in May 2023.
+
+  I have read your TSE'21 paper on numerical invariant generation, and I am interested in this line of dynamic invariant research.
+I have worked (optional: with prof.
+Y at Z) on static program analysis and I think it could be used to tackle the spurious issues mentioned in your paper.
+I have a short paper at conference/workshop C and a project on symbolic execution (Github repo G).
+
+  ...
+
+  This is a good example because it is written just for me.  It shows that the student knows about my work on invariant generation and has a related background (paper C and project G).
+\end{commentbox}
+
+Finally, profs. are very busy so don't take it personally if you don't get anything from them (though I would be very surprised if such thoughtful emails get no replies!).
+
+\begin{commentbox}[Xiaokuan]
+  Applying for PhD and contacting a potential PhD adviser is a classic \textbf{`why me, why you'} problem,
+  similar to looking for a job in a company.
+  On a high level,
+  you need to show that you have done your \emph{homework}
+  regarding the professor and the university,
+  and clearly explain:
+  1) why do you think you are a good fit in professor A's group?
+  2) why do you want to be advised by professor A, not B?
+  3) why do you want to apply for university X, not Y?
+  If you don't want to spend time doing your homework,
+  the chance of getting a reply is close to zero.
+\end{commentbox}
+
+\begin{commentbox}
+  [Deepak:]
+  In my view, cold emails are not welcome by most faculty members and should be avoided.
+However, if one is already admitted to a program in some department, by all means, email the faculty you may be interested in working with, but do mention right at the beginning that you are already admitted to the program as well as several other universities.
+State specific areas (preferably specific topics-ML, robotics instead of AI).
+\end{commentbox}
+
+\subsubsection*{Additional Resources}
+\begin{itemize}
+  \item \href{https://yonatanbisk.com/emailing_professors.html}{A Note about Emailing Professors} by Yonatan Bisk
+  \item \autoref{sec:accept-postpone-decline} How to accept/postpone/decline PhD offers (and do it gracefully)
+\end{itemize}
+
+\subsection{Things to avoid when cold emailing a professor}
+
+\begin{itemize}
+
+    \item \textbf{Not customizing the email}. These emails typically talk about the student and not why they are interested in the professor's work. See writing tips in \autoref{sec:contact}
+    \item \textbf{Sending the same email to multiple professors}.
+Sending emails to multiple profs.
+using BCC or even CC (where we can see all addresses the email is being sent to) will result in no reply or a very harsh one on how unprofessional this is.
+    \item \textbf{Too much information}.
+Keep your email short and to the point.
+If we are interested, we will ask for more information.
+2--3 paragraphs are enough.
+Personally I also do not want to see course transcript or CV in the first email.
+    \item \textbf{Not being professional}, e.g., do not call the professor by their first name in the first email.  Use Prof. Lastname or Dr. Lastname (\autoref{sec:address}).
+    \item \textbf{Not reading the professor's website} and asking questions that are already answered there.
+\end{itemize}
+
+\chapter{Miscs and FAQs}\label{sec:faqs}
+
+\epigraph{``I want to share something with you – the three little sentences that will get you through life; number 1: Cover for me, number 2: Oh, good idea, Boss, and number 3: It was like that when I got here.''}{\textsc{The Simpsons}}
+
+\section{Why did I get rejected?}\label{sec:why-rejected}
+
+Many students lament that they get no interviews or are rejected (with strong profiles) and that the admission process seems random.
+However, while it is true that the process is not perfect, it is not random.
+Other than not having a strong profile (e.g., research potential, GPA, LoRs, SOP, poor interviews), here are some other reasons for rejections.
+
+\paragraph{Not a Good fit and Bad luck}  You could have an excellent profile, but if you are interested in a research area that the program does not have, you will not be admitted.
+Similarly, if no faculty is willing to advise you (e.g., they are on sabbatical or personal leave, do not have sufficient funding, or already have too many students), you will not be admitted.
+This is actually good for you, as you don't want to be in a program where no one can advise you.
+
+Related to this is that you just have bad luck and apply at the wrong time.
+For example, since 2024 there has been a huge surge in students interested in AI and NLP (thanks, ChatGPT!).
+Consequently, AI/NLP faculty might be overwhelmed and cannot consider many candidates, even those with excellent profiles.
+
+\paragraph{You aim too high} This is a very common reason: you have applied to schools that are too high for your profile.
+While being ambitious is good, you should understand how PhD admission works (e.g., by reading this document and realizing things such as having a good GPA doesn't mean much to top programs) and be realistic!
+
+Before applying, you should talk to your professors and ask them to give you an honest opinion on where you should apply.
+To increase their chances, many students apply to a range of schools, including ``safety'' ones.
+
+\paragraph{Overqualified or Lack Interest}  This might be paradoxical and is the opposite of aiming too high.
+However if adcom believes that you are likely to get admitted and go to a better program, they might not be so enthusiastic to admit you and want to save the spot for someone else.
+
+A related reason is that you did not show enough interest in the program.
+For example, your SOP did not mention any faculty or research area at the program, you did not respond to professors interested in interviewing you (which might be considered unprofessional and burn bridges), or during the interview you showed little knowledge, interest, or enthusiasm (because they are your safety schools) in the program or profs.
+
+\paragraph{Low English exam scores (e.g., IELTS or TOEFL)}  Profs.
+might not care much about these, but the college or the university often sets a minimum that you need to pass to be considered, especially for TA funding (e.g., you need to communicate well with students as a TA).
+Thus, while profs.
+are willing to argue for your case, they might be reluctant to go against the requirement of the university on standard exams.
+
+Note that English exams are not required for all students, e.g., domestic students and those from certain countries are waived from taking these exams.
+For example, Mason waives English exams for students from these \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-Mason#standard-tests-waiver-eligible-countries}{countries}.
+
+\paragraph{Red flags} Various types of red flags can cause concern to the adcom (e.g., too many STEM courses with low or withdrawn grades, plagiarism, cheating, or other academic dishonesty).
+Another one is that you have a history of jumping from one program to another (e.g., you have been in multiple different programs in a short time).
+Adcom members might have contacts in other programs and heard about your stories, or your LoRs might mention these.
+
+If you think you have these issues, you should address them in your SOP or ask your letter writers to do so.
+In general, these things are rare, but they do happen and cause concern to the adcom.
+
+\section{Timeline: Why are rejection letters sent so late?}\label{sec:late-rejection}
+
+%"Just reject me already!" is a common sentiment among applicants.
+Indeed, school will first send out admission offers to the top candidates.
+They do not send out rejection letters because there is still a chance that some of the top candidates will decline the offer.
+If they do, then the school will go to the next set of candidates and send out offers to them.
+This process continues until all spots are filled.
+This is why you might not hear back until late in the admission cycle.
+
+\paragraph{Acceptance Letters} Universities prioritize sending out acceptance letters first. This allows the admitted students to make decisions and plan for their studies.
+
+Some universities have rolling admissions, meaning they send out acceptance letters as they review applications (e.g., some universities sent out letters as soon as early February).
+Others have a specific date when they send out the first round of acceptance letters.
+
+\paragraph{Response Deadlines} Accepted students are usually given a deadline to make decisions on their offers, often around \textbf{April 15th} in the US.
+This is a standard deadline for PhD programs due to the Council of Graduate Schools (CGS) Resolution (\href{https://cgsnet.org/wp-content/uploads/2024/01/CGS_April15_Resolution_Jan312024.pdf}{see more}.
+After this date, CS programs can gauge how many slots remain unfilled.
+
+\paragraph{Waitlist} Most CS programs have a limited number of slots for PhD students, and thus put many good students on a waitlist.
+If accepted students decline the offer, then offers are sent to students on the waitlist.
+So if you see people getting accepted, that does not mean you are out yet.
+
+\paragraph{Rejection Letters} Schools typically start sending out rejection letters to remaining applicants \emph{after they have finalized their admissions decisions}.
+Thus, rejection letters are often sent out late (e.g., after April 15th or even much later).
+Not much you can do here, you can try to contact the school to ask about your status, but they might not reply, they might say they are still reviewing applications, or give you inaccurate information (e.g., you will hear in two weeks).
+In short, you just have to be patient and wait (however, beware that some schools do not send out rejection letters at all).
+
+\section{Increasing my admission chance}\label{sec:improve-your-chance}
+
+\begin{center}
+  \includegraphics[scale=0.2]{files/alpinist-climbing-peak-mountain-comic-hand-drawn-vector-illustration.jpg}
+\end{center}
+
+Given the high number of quality applicants and a limited number of spots, in addition to having a good application profile, you want to show something that makes you \textbf{stand out}.
+For example, even if you do not have research experience, you can talk about your personal projects, as long as they can help show you can do research.
+For example, if you have an open-source project that has lots of stars in Github, then you should mention it.
+If you often write technical, research-like blogs with many viewers, talk about them too.
+
+There are other things you might not think are important but can make you stand out.
+For example, if you have a strong background in a non-CS field but can be integrated with CS, e.g., you have a degree or background in \emph{dance} or \emph{music} and want to integrate them with CS, do talk about it.
+Are you a female or a minority in CS (\autoref{sec:urm})?
+Do you participate in outreach activities that help increase diversity and inclusion in CS?
+Diversity is highly valued in CS programs in the US, and experiences in these areas can make you stand out and get noticed from reviewers.
+
+\begin{commentbox}[Vu:]
+In his \href{https://matt.might.net/articles/how-to-apply-and-get-in-to-graduate-school-in-science-mathematics-engineering-or-computer-science/}{post}, Matt Might was initially unsure about an application.
+However, upon learning that the applicant had led a 100km hike in the Himalayas, he decided to accept the applicant.
+This is a good example of being \emph{stand out}, and I would also advocate for that student as this shows they have the persistence and determination required for research.
+\end{commentbox}
+
+% It is worth noting that in many cases,  an application could be rejected by most reviewers, however have a "champion", who doesn't even have to be in the Admission comittee, that strongly advocate for your profile.
+
+\section{How long do you evaluate an application?}\label{sec:ievaluate}
+
+On average, I spend about 10 minutes reviewing each application (less for clear rejections and more for potential acceptances).
+While this seems short, it is not that difficult to tell if an application is good or bad.
+In fact, this
+is twice what other faculty spend on average, e.g., Philip Guo spent \href{https://pg.ucsd.edu/PhD-application-tips.htm}{3--5 minutes} per application.
+
+For each application, our system compiles a single PDF file, which consists of a summary (degrees, GPAs, etc), transcripts, test scores, LoRs, a CV, SOP, and writing samples.
+I usually read in this order.
+I start with the \emph{summary}, checking for low GPAs or test scores below the university minimum (\autoref{sec:standard-tests}).
+I then skim the \emph{transcripts} for low grades in relevant courses, noting issues like ``many low grades in main courses or unknown international school with good GPA''.
+These are not as important as LORs or SOP but I read them first because the review system has questions about them (e.g., ``is GPA good?
+is IELTS sufficient?'')
+
+I read carefully \emph{strong LoRs} and skim weaker ones, noting either strong letters from well-known professors/researchers or generic content.
+I skim over \emph{CV} and look for publications, research experiences, and notable achievements, noting things like ``published papers in top venues'' or ``gold medal in an international competition.''
+I read strong \emph{SOPs} thoroughly (and skim weak ones).
+I note whether the SOP is exciting, research-oriented, standing out, e.g., applicant is minority (\autoref{sec:urm}), and tailored to our program (e.g., if they are familiar with the work of some faculty or have talked to them).
+These notes are entered into the evaluation system.
+
+Finally, I enter my decision, which is usually either a rejection or an offer of admission with full funding (e.g., from TA).
+I also recommend very strong candidates for the University Presidential fellowship.
+Note that while the system has other decision options, eg., admit without funding, provisional admission (e.g., if they need to take some courses), I rarely use them, simply because we already have too many good students and do not have the resources to support them.
+
+\subsection{How Decisions are Made}\label{sec:how-decisions}
+
+After reviewers have evaluated an application, adcom chair will review all evaluations, look at entered notes, and ask reviewers to discuss and resolve discrepancies to reach a consensus (e.g., a reviewer wants to accept but the other wants to reject).
+Typically, the decision is made entirely by the reviewers; there is no involvement from the adcom chair, department chair, or others.
+In most cases adcom members, even those reviewing the same application, make decisions independently and do not talk to each other (just a common practice to avoid biasing).
+In some rare cases we might (\autoref{sec:adcom-discuss}).
+
+Even if \emph{all} reviewers recommend acceptance, the application is not automatically accepted, especially if no faculty is willing to advise the student.
+For example, if the student is interested in a research area that no faculty is working on, or an area where no faculty is taking new students (e.g., AI/ML where faculty likely already has many students) then the student will not be admitted.
+This is increasingly common as the number of applicants increases significantly in recent years in CS.
+
+However, if the student has contacted a faculty member and that faculty member is interested in the student and has made this known to the admission committee, then the student is likely to be admitted, \emph{even if the faculty does not review the application or has funding} (thus the importance of contacting faculty).
+Similarly, if the student mentioned a faculty member in their SOP, adcom might ask that faculty member to look at their application and if they are interested in the student.
+Even if the student has a weak profile (but still passes the minimum requirement from the university), they might be admitted if a faculty member is willing to take them.
+Adcom members, especially in the US, are very reluctant to go against the faculty's decision (e.g., if a faculty wants to admit a student, we are not going to reject them).
+
+\section{Do adcom members talk to each other?}\label{sec:adcom-discuss}
+
+However, in some rare cases we do talk to each other.
+For example, when there are discrepancies in evaluations, in which case the adcom chair will ask reviewers to discuss the application to reach a consensus.
+We also talk to each other for interesting or strong applications, e.g., how to recruit this student or who should be the adviser.
+If the student mentioned a faculty member in their SOP, we might ask that faculty member if they are interested in the student.
+
+\section{Do adcom members read all applications?}\label{sec:read-all}
+
+While adcom members (in fact, any faculty) can view any application, we only review those that are assigned to us, which are already too many.
+Adcom chair will assign applications to reviewers based on their expertise (e.g., if a student says they want to do SE or interested in working with me), and reviewers will only review those applications.
+Occasionally we might look at other applications, e.g., if we know the student or have some special circumstances (e.g., if that student contacted me, I know that student, they are from Vietnam, etc).
+However, even if we look at them, we usually do not get involved in their evaluation.
+
+\section{Member of an underrepresented or LGBTQ+ groups}\label{sec:urm}
+
+A common question is whether you should mention, e.g., in your SOP, that you are a member of an underrepresented and minority (URM) group\footnote{URM in CS in the US include African Americans, Hispanic Americans, American Indians/Native Alaskans, Native Hawaiians/Pacific Islanders (excluding Asian Americans), and multiracial students identifying with at least one of the previously listed URM categories.} or LGBTQ+ group.
+This is a strong concern for international (and even domestic) students, who may have encountered discrimination in their countries or personal experiences.
+%This question arises often for international (and even domestic) students.
+The main worry is that they might get discriminated against, as often happen in their countries or personal experiences.
+%, and (ii) mentioning this might make them look like they are asking for special treatment.
+
+In my personal opinion, highlighting URM or LGBTQ+ identity can \emph{be beneficial}, especially if you can articulate how your diverse experiences contribute to diversity and inclusion in academics.
+Many universities have a strong commitment to diversity and inclusion and actively
+\emph{recruit students from underepresentative backgrounds} (e.g., some \href{https://cse-climate.engin.umich.edu/reports/climate-dei-reports/cse-climate-and-dei-report-2022-2023/#grad-ethnicity}{stats from UMich CSE} and \href{https://www.Mason.edu/news/2022-09/mason-now-top-10-public-university-diversity-innovation-and-cybersecurity-education-us}{Mason} often promotes itself as one of the most diverse universities in the US).
+Moreover, many scholarships and fellowships are created specifically for URM and LGBTQ+ students, which you should explore if qualify.
+
+Understandably, you might feel uncomfortable disclosing this, fearing you would lose a chance of working with a faculty due to their bias.
+However, if someone has this issue, then you might not want to work with them anyway (and such bias is not acceptable in academics and likely would not be tolerated by the university).
+In my experience, it is the opposite: \emph{many profs.
+actively seek to work with students from diverse backgrounds and view diversity as an asset to their research group}.
+
+Moreover, promoting diversity aligns with the goal of broadening participation in computing (BPC) that is strongly supported within academia (e.g., see the \href{https://plans.bpcnet.org/GeorgeMasonUniversity_ComputerScience_DepartmentalBPCPlan.pdf}{BPC plan at Mason CS}) and government funding agencies such as NSF (e.g., see \href{https://new.nsf.gov/cise/broadening-participation}{here} and \href{https://www.nsf.gov/pubs/2022/nsf22125/nsf22125.jsp}{here}).
+Also, sometimes providing this information can make it easier for faculty to apply for certain scholarships or fellowships to support you, e.g., from a private donor that has offered a fellowship only to students from certain backgrounds.
+
+Ultimately, this is a personal choice and you do what you feel comfortable.
+If you do not feel comfortable disclosing this, you should not feel pressured to do so and you should also ask your LoR writers to not disclose this information.
+Note that you can still talk about your experiences and perspectives without mentioning your identity.
+If uncertain, you can seek additional guidance from your close professors and collaborators However, it is important to recognize that diversity is strongly valued and celebrated within CS programs in the US, and that you should feel empowered to share your unique experiences and perspectives.
+
+%https://www.edsurge.com/news/2018-05-18-how-cornell-university-diversified-its-incoming-phd-computer-science-student-body
+
+%https://academia.stackexchange.com/questions/130591/is-it-standard-for-us-based-universities-to-consider-the-ethnicity-of-an-applica
+
+%https://www.cs.cornell.edu/~bindel/paper/diversity.pdf
+
+%https://cs.brown.edu/degrees/doctoral/applications/helpful-resources-applying-computer-science-phd-programs/
+
+\section{My undergrad was not in CS or related areas}\label{sec:non-stem}
+
+You still can apply to PhD in CS \emph{as long as} you can demonstrate you are ready for it through your background, research experience, LoRs, statements, etc.
+You might be even able to leverage this to make your profile stand out as mentioned in \autoref{sec:improve-your-chance}.
+
+A main question adcom has for a non-CS or non-STEM student is if you have the sufficient technical background obtained through core CS courses.
+So you need to show that you have such knowledge through your coursework, projects, or research.
+For example, if you have taken a course on Algorithms, even online ones like Coursera, you can talk about it in your SOP.
+If you have done a project that requires knowledge of OS or have a professional certification (e.g., A+) through work, you can talk about it.
+If you have done research that requires knowledge of Discrete Maths, you can talk about it.
+You can also ask your LoR writers to talk about your technical background.
+In summary, in your application, convince us that you have the background to do CS PhD research.
+
+\paragraph{Core CS topics} Common CS knowledge that you should know are:
+\begin{itemize}
+  \item Programming Foundation (e.g., programming concepts in some modern languages)
+  \item Discrete Math (e.g., logic, set theory, proof techniques)
+  \item Data Structures and Algorithms (e.g., linked lists, trees, sorting, searching)
+  \item Computer OS or Systems (e.g., memory management, file systems, processes)
+\end{itemize}
+
+In short, you \emph{do not need} to formally take CS courses, you just need to show that you have this essential knowledge, e.g., through the mentioned ways.
+Many universities are well aware that incoming graduate students might not have all the technical background, so they often have a \emph{``bridge''} courses to help students catch up.
+For example, Mason has four bridge courses corresponding to the four core areas above that incoming students can take to catch up on their CS knowledge.
+
+\begin{commentbox}[Vu:]
+  I would advocate for a non-STEM student who shows that they have a strong drive for CS by studying core CS knowledge through various channels (e.g., self-study through online courses, projects, etc.).
+I have seen many students with non-CS backgrounds who are very successful in CS PhD.
+I also have seen many students with CS background who are not successful in CS PhD.
+So it is not about your background, it is about your drive and passion for CS research.
+\end{commentbox}
+
+\section{Is an MS required for admission to PhD in CS?}\label{sec:msrequirement}
+No.
+In fact, students can get MS degree ``along the way'' to PhD, e.g., after finishing the 2-year course work.
+However, MS can help admission if it gives research experience or is from a more well-known school than your undergrad institution (\autoref{sec:your-school}).
+
+If you have an MS then some course work \emph{might be} transferred for course credits, which \emph{might} save some time. But don't count on this.
+
+\begin{commentbox}[Vu:]
+    I start my PhD with an MS in CS from a US university.
+I found that the MS helped me a lot in research, but not with course work and I had to retake many courses.
+In general, don't worry if you don't have an MS.
+But also don't feel that you wasted your time if you have an MS, as it can help you in research.
+  \end{commentbox}
+
+\section{How long to complete the CS PhD program?}\label{sec:time}
+
+Typically, it takes 5--7 years for CS PhD in the US.
+This can be longer than CS PhDs in other countries, which might require MS first (recall from \autoref{sec:msrequirement} that CS PhD programs in the US do not require MS and you can get MS along the way to PhD).
+Within these 5--7 years, CS PhD students often take a ``leave of absence'' for 1--2 semesters to do internships at companies and research labs.
+
+\begin{center}
+  \includegraphics[scale=0.5]{files/c4a.png}
+\end{center}
+
+The first two years you typically take coursework (somewhat equivalent to an MS study), find an adviser, and learn how to do research.
+The next 2--3 years you focus on your research, form a dissertation topic, and get results published.
+The last 1--2 years you continue to publish, write and defend your dissertation, and look for a job.
+In many cases you might take a summer or two off to do an internship to get additional research opportunities.
+The PhDComics figure above shows the ``ambition'' level of a PhD student over their years of study (they miss the 6--7th year when the ambition is \emph{``Just let me graduate''}).
+
+% \begin{commentbox}[Vu:]
+%     I start my PhD with an MS, and it took me 7 years (Fall'2007-- Fall'14). I spent half a year doing an internship at the Naval Research Lab.
+%   \end{commentbox}
+
+\section{What's the difference between PhD and MS in CS?}\label{sec:phd-vs-ms}
+
+While both are graduate degrees, they are different in terms of objective, admission requirements, course requirements, duration, and funding as summarized in \autoref{tab:phd-vs-ms}.
+
+\begin{table}
+  \caption{PhD vs. MS}\label{tab:phd-vs-ms}
+  \begin{tabular}{c|c|c}
+  \toprule
+  \textbf{} & \textbf{PhD} & \textbf{MS} \\
+  \midrule
+  Objective & Industry & Research \\
+  Admission Req & Research experience & No research experience \\
+  Coursework Req & Yes & Yes (research is much more important)  \\
+  Duration & 2 years & 5--7 years \\
+  Adviser Req & No & Yes \\
+  Funding & No & Yes \\
+  \bottomrule
+  \end{tabular}
+\end{table}
+
+\begin{itemize}
+
+  \item Objective: an MS is typically to prepare you for industry, while a PhD is to prepare you for research.
+Some MS has thesis option but the research is much lighter compared to a PhD, which is research-focused.
+
+  \item Admission requirements: MS also requires a good GPA, LoRs, SOP, and test scores, but research experience is not required.
+PhD requires all of these, but research experience is a must.
+Moreover, MS programs are typically much less competitive than PhD programs.
+In fact, many MS program are viewed as a revenue source (\emph{cash cow}) for the department, as they are not funded and students pay tuition.
+
+  \item Course requirements: MS has a specific number of courses (typically can be done in 2 years).
+You graduate with an MS when you're done with the courses.
+PhD also has coursework requirements (typically taken in your first 2 years) but focuses on research after that.
+You graduate with a PhD when you have done enough research and written a dissertation (which is typically way harder and takes much longer time than coursework).
+
+  \item Duration: an MS typically takes 2 years while a PhD takes 5--7 years (or even longer).  Many students get an MS along the way to a PhD, e.g., after finishing the 2-year course work.
+
+  \item Adviser: MS students typically do not have an adviser (if you do thesis option then you will have one), while PhD students need an adviser who guides them in their research.
+
+  \item Funding: MS is typically \emph{not funded}, while PhD is typically funded (e.g., through TA or RA as explained in \autoref{sec:funding}).
+
+\end{itemize}
+
+\section{PhD in the US vs. Other Countries}\label{sec:non-us-differences}
+
+\begin{table}
+\caption{Comparison of US and Other PhD Programs}\label{tab:us-vs-other}
+\begin{tabular}{c|c|c}
+\toprule
+\textbf{Aspect} & \textbf{US PhD Programs} & \textbf{Other Countries} \\
+\midrule
+Duration & 5-7 years & 3-5 years \\
+MS Required & Not required & Often required \\
+Coursework Required & Yes (first 2 years) & No \\
+Research Proposal Required & No & Yes (in some countries) \\
+Academic (Faculty) Job & Direct & Postdoc \\
+Work Life Balance & Less & More \\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\autoref{tab:us-vs-other} summarizes the main differences between CS PhD in the US and other countries. % This is based on my experience and what I heard from others who did PhD in other countries.
+
+\emph{MS requirement and PhD duration}:  CS PhD programs in the US do not require an MS degree (as mentioned in \autoref{sec:msrequirement} and \autoref{sec:time}).
+In contrast, many other countries require having an MS degree before joining a PhD program.
+This means that US PhD programs are longer (5--7 years, 2 of which are coursework) than other countries (3--4 years, no coursework).
+
+\emph{Project proposal}: in many countries, you have to choose a project and adviser \emph{during} the application process (e.g., you write a proposal to a potential adviser).
+But this allows you to start your research right from the beginning.
+
+In the US, you often start your PhD without an adviser or project and find them later.
+You will have two initial years to explore and find an adviser and research topic.
+After that you explore projects with that adviser before settling on one (can take a couple of years).
+
+\emph{Course work}: as mentioned above in the US you will spend the first couple of years taking classes and exploring potential adviser and research topics. 
+After that, you have to pass a series of exams during your PhD, e.g., qualifying exam, comprehensive exam, thesis proposal defense\footnote{The word ``ABD'' (all but dissertation) is used in the US to refer to a PhD candidate who has finished all course work and exams and only needs to write and defend their dissertation.}.
+
+In other countries, you start your research right away, e.g., you immediately work on the research project you proposed with the adviser you chose.
+Moreover, you might not have exams like those in the US or only have to do a few of them.
+
+\emph{Funding}:  In many countries, funding comes from the university or the gov't.
+This funding often has a fixed duration (e.g., 3 or 4 years).
+In the US (\autoref{sec:funding}), funding (e.g., RA) comes directly from your adviser (no fixed duration).
+There are also fewer TA opportunities in European universities compared to the US.
+
+\emph{Academic position after PhD}: In other countries, PhD graduates interested in academia typically apply for additional research appointments, i.e., postdocs, and then consider faculty positions.
+
+In the US, PhD graduates often apply directly for faculty positions. Postdoc for US graduates is no longer a popular option as it was before.
+
+\emph{Work-life balance}: PhD students in the US are often said to be overworked compared to other countries, e.g., in Europe.
+This is partly due to the longer PhD program and that US PhD students are often paid through TA, which requires them to do TA in addition to their own research.
+In contrast, PhD students in other countries are often paid through fellowships, which might not require doing TA.
+
+\section{How do I call or address a professor?}\label{sec:address}
+
+\begin{center}
+  \includegraphics[scale=0.5]{files/c5.png}
+\end{center}
+
+If you're reaching out to a professor for the first time,  address them as Prof.
+or Dr.
+Lastname (if they do not have a PhD, then use Prof.).
+Many international students use Prof.
+or Dr.
+FirstName LastName, but this can come across as if you're simply copying and pasting names.
+So just stick with Prof.
+or Dr.
+Lastname.
+Using Prof.
+is generally the safest option.
+
+Furthermore, do not use Mr., Mrs., Ms, or Miss.
+This happens rarely, but I have seen  new students (e.g., undergraduate freshmen in the US) sometimes use these, which are used in K-12 schools but not in higher education.
+
+Moreover, do not call the prof. by their first name at first.  As you become more familiar with your prof and depending on their preferences, you may transition to addressing them by their first name.
+For example, I prefer that my students and colleagues call me Vu. Some students call me \emph{Dr. Vu}, which I find a bit amusing but am totally fine with it.
+
+\begin{commentbox}[DK:]
+  I was amused to read this as if I recall correctly, you never called me by my first name when you were at UNM. You always called me Prof. And, many times, I would jokingly call you back as Prof. Vu.
+  \tcblower
+  \textbf{Vu}: Yes, for some reason I enjoy calling you ``Professor'' (without appending a last or first name).  The use of Prof. Vu may have foreshadowed my future in academia.
+\end{commentbox}
+
+Note that in some universities the formal title Dr.
+Lastname is preferred over Prof.
+Moreover, be aware that not all faculty members may hold a PhD, in which case using Prof.
+Lastname is a suitable alternative.
+	You just need to observe and follow the conventions at your particular institution.
+One way to determine how to address a professor is to observe how they sign their emails or how they introduce themselves in class.
+For example, I introduce myself as ``Vu'' in class.
+
+\paragraph{Referring to professors you know} When referring or talking about a prof (e.g., your mentor) that you know, you can just informally use their names if they are OK with it as mentioned above (or Dr./Prof., if you want to be formal).
+You can also include their institution if it makes it more precise.
+For example, I can say:  \emph{``I did my postdoc with Jeff Foster at Univ.
+of Maryland''}.
+
+Do not include ranking (e.g., Assistant, Associate, Scientist, ...) when referring to someone.
+I see many international students include a lengthy title of people they know, e.g., \emph{I am advised by Asst.
+Prof.
+X, and I also collaborate with Distinguished Scientist Y}.
+This is \emph{not necessary} and makes it look like you're trying to show off your connections.
+These nuances represent some cultural differences that you may encounter and will gradually adapt to.
+More on cultural differences in \autoref{sec:cultural}.
+
 % \section{Having fun during a PhD?}
@@
 
+\chapter{Glossary}\label{sec:glossary}
+
+\begin{description}
+
+  \item[ABD] All But Dissertation. A PhD candidate who has finished all course work and exams and only needs to write and defend their dissertation.
+
+  \item[Adcom] Admission Committee. The group of faculty members who review applications and make admission decisions.
+  \item [Adcom members] Faculty members who are part of the admission committee. These are the people who review applications and make admission decisions.
+  \item[Adcom chair] The faculty member who leads the admission committee.
+This person is often not involved in individual admission decisions but oversees the process (e.g., resolve disputes, ensure fairness).
+  \item[Advisor/Supervisor]: A faculty member who guides and mentors a PhD student throughout their research.
+This person typically plays a crucial role in a student's academic journey.
+In the US, the term ``adviser'' is more commonly used than ``supervisor''.
+  \item[Cohort] A group of students who start a program at the same time and take classes together. You should get to know your cohort as they will be your colleagues and friends during your PhD.
+  \item  [Deferred Admission] An option allowing admitted students to postpone their start date, typically by one year.
+  \item [Diversity Statement] A document sometimes required in applications where applicants describe their contributions to diversity and inclusion, or how their background adds to the diversity of the academic community.
+This document is often required for faculty positions, but some universities also require it for graduate admissions.
+  \item[Domestic Student] A student who is typically a US citizen or permanent resident (``greencard holder'').
+        But sometimes it can also refer to students who did their undergrad (or MS) at a US university.
+  \item[Field or Area] A specific research area, e.g., Machine Learning, Computer Vision, Software Engineering, etc.
+  \item [In-state vs.
+Out-of-state] In-state tuition is the tuition rate for students who are residents of the state where the university is located.
+Out-of-state tuition is the tuition rate for students who are not residents of that state.
+In-state tuition is typically much lower than out-of-state tuition.
+PhD students typically do not have to pay tuition as it is covered by their funding, but this might be important for MS students.
+  \item[LoR] Letter of Recommendation. A letter written by a professor or supervisor that assesses your qualifications and potential for graduate study (\autoref{sec:lor}).
+  \item[SOP] Statement of Purpose. A document written by yourself to explains your research interests, background, and reasons for applying to a PhD program (\autoref{sec:research-statement}).
+
+  \item[RA] Research Assistantship. A form of funding where you work as a research assistant for a professor.
+  \item[TA] Teaching Assistantship. A form of funding where you work as a teaching assistant for a professor.
+  \item[TOEFL] Test of English as a Foreign Language. A standardized test to measure the English language ability of non-native speakers.
+  \item[IELTS] International English Language Testing System. A standardized test to measure the English language ability of non-native speakers.
+  \item[GRE] Graduate Record Examination.
+A standardized test that is an admissions requirement for many graduate schools in the US.
+However, it is not required for most CS PhD programs (\autoref{sec:standard-tests}).
+
+  \item [Lab] Research Lab.
+A group of researchers (profs.
+and students) working on a common research area, e.g., a Software Engineering Lab.
+Typically a lab is led by a single professor, but sometimes it can be co-led by multiple professors.
+For example, my lab is DynaROARS, which focuses on my own research area (software verification), but I am also part of the more general SWE Lab at Mason with other faculty in Software Engineering.
+
+  \item[MS] Master of Science. A graduate degree that typically requires 2 years of study.
+  \item [PI]  Principal Investigator.
+A lead researcher on a grant or research project, e.g., a PI on an NSF grant.
+Prospective students often use this term to refer to a professor they are interested in working with.
+While this term is used frequently in online forums among students, we do not use it this way in academia, e.g., ``who is your PI?'' might only make sense to fellow Reddit applicants, but might not be understood by others, ``who is your adviser?'' is what you want to use.
+    \item[STEM] Science, Technology, Engineering, and Mathematics. A term used to group together these academic disciplines. CS is considered a STEM field.
+  \item[URM] Underrepresented Minority. Refers to groups that are underrepresented in higher education, e.g., Black, Hispanic, and Native American students.
+  \item[NSF] National Science Foundation.
+A US government agency that supports research and education in science and engineering.
+NSF is a major source of funding for RAs in CS research.
+There are other agencies that support RAs but they might require US citizenship or permanent residency.
+        \item [R1 vs.
+R2 Universities] Research 1 (R1) universities are universities with the highest level of research activity.
+Research 2 (R2) universities have a high level of research activity but not as high as R1 universities.
+R1 universities are typically more prestigious and have more funding and resources for research.
+R2 universities are also good but might have fewer resources and funding.
+  \item [Rolling Admission] A process where applications are reviewed as they are received, and decisions are made throughout the admission cycle.
+This is in contrast to a process where all applications are reviewed after the deadline and decisions are made at the same time.
+Rolling admission is more common for MS programs and less common for PhD programs (e.g., at Mason, MS is rolling and PhD is not).
+  \item [Top-tier conferences and journals] These are the most prestigious venues for publishing research in CS.
+For example, in Software Engineering, top-tier conferences include ICSE, FSE, and ASE and top-tier journals include TSE and TOSEM.
+Publishing in these venues is highly competitive and prestigious, and can significantly improve your chances of getting admitted to a good PhD program.
+
+        \item [Stipend] A fixed regular sum paid to students as part of their funding package (i.e., salary).
+This is typically paid monthly or bi-weekly.
+Stipend is typically paid to PhD students and not MS students.
+  \item[REU] Research Experience for Undergraduates. A program funded by the NSF to provide research opportunities for undergraduate students.
+    \item[OPT] Optional Practical Training. A program that allows F-1 students to work in the US for up to 12 months after graduation.
+  \item[CPT] Curricular Practical Training. A program that allows F-1 students to participate in internships or practical training that is an integral part of their academic curriculum.
+  \item[F-1] Student Visa. A visa that allows international students to study full-time at an accredited institution in the US.
+  \item[F-2] Dependent Visa. A visa that allows the spouse and children of an F-1 visa holder to live in the US.
+\end{description}
+
 \chapter{Cultural Differences}\label{sec:cultural}
@@
 
-\section{Relationships}
+\section{Maintaining Good Relationships with your Profs.}
 There's a misconception that in the US it's all business, with professors as bosses who pay students for their work and that lab mates are just work colleagues; and that doing nice things means expecting something in return.
@@
   \end{commentbox}
+
+  \section{Miscs}\label{sec:cultural-misc}
+  Several other common surprises for international students (and foreigners in general) in the US:
+
+  \paragraph{Healthcare System}  You (and your spouse) will need health insurance!
+Otherwise you will be charged a lot for healthcare services.
+However, as mentioned in \autoref{sec:funding}, your  TA/RA  will cover health insurance (and your spouse/children also get insurance or discount for it).
+However, be aware that heathcare services can still be expensive even with insurance.
+So  know what your insurance covers and be prepared for unexpected costs.
+Moreover, healthcare system has many confusing jargons such as HMO, PPO, deductibles, copays, and coinsurance.
+Get yourself familiar with them and ask for help when needed.
+
+  \paragraph{Tipping Culture} Unlike many other countries, tipping is expected for various services, especially in restaurant.
+So adding an extra 15--20\% to your bills is common, especially in restaurants.
+  You should also tip for other services, e.g., Uber, taxi, haircuts, and hotel services.
+The minimum wage for tipped employees is lower than the standard minimum wage, so tips are an important part of their income.
+If you do not tip, it is considered rude and disrespectful.
+
+  \paragraph{Car Dependency} Most of the US is highly car-dependent.
+If you do not have a car, you will need to rely on friends, Uber, or public transportation, which can be inconvenient and time-consuming.
+Most international students end up getting a driver's license and buying a car.
+
+	%\paragraph{Individualism} The strong culture of individualism in the US can be both liberating and isolating.
+People are more focused on their personal rights and tend to be less concerned with what others think.
+
+  % \paragraph{Racism and Segregation} While some find the US to be less racist than expected, others note that racism and segregation are still prevalent, often manifesting subtly in daily life or visibly in segregated neighborhoods.
+
+	%\paragraph{Litigiousness} The US is known for its litigious culture, where lawsuits are common, and people can be sued over seemingly minor issues.
+This can be surprising for newcomers used to different legal systems.
+
+  % \paragraph{Consumerism and Scale} The US is often associated with large portions, big cars, and massive stores like Costco. The emphasis on consumption and abundance is notable.
+
+	% \paragraph{Religious Influence} Religion, particularly Christianity, plays a more visible role in public and political life in the US compared to many other countries.
 

### 2024-08-18 faculty.tex
updates

@@
 
-Depending on your research area, you might focus on publishing at top conferences or journal.
+Depending on your field, you might focus on publishing at top conferences or journal.
 You probably should know what's popular in your area already (if not, ask your senior colleagues).
@@
 \subsection{Short papers, workshops, etc} Short papers in workshop or non-research tracks at a conference (e.g., tool, demo, vision, etc) do not count as much.
-They describe work that might not be as mature (e.g., new idea) or as complete (e.g., a tool or a demo) as a full paper, and also have higher acceptance rate.
+They describe work that might not be as mature (e.g., new idea) or as complete (e.g., a tool or a demo), and also have higher acceptance rate.
 Many people use them as a way to get feedback on their work before submitting to a full paper venue.
-Most new professors do not focus on these kinds of papers.
 Many P\&T committees do not count or care much about these and having too many of them can be a negative signal.
 
-Personally, I find that such papers are very useful as preliminary work for writing grant proposals.
+While many faculty do not value short papers, I find that they are are very useful for writing grant proposals.
 They show that you have a new idea and good preliminary results that are worth pursuing further.
@@
 
-\subsection{Your university does not recognize conference value} For most non-CS engineering fields journal are more prestigious than conferences.
-Places with a strong CS department are well aware of the value of top confs., but many other places are not (especially smaller universities or CS depts run by engineering people).
-These places also do not know much about the quality of the venue, and so in many cases put a workshop paper in the same class as a top conference paper and also cannot differentiate between a good vs bad journals in CS (because they do not know CS well).
+\subsection{Your university does not recognize conference value}
+
+Journals, not conferences, are the main venue for publication for most non-CS engineering fields.
+While places with a strong CS department are well aware of the value of top confs.
+in CS, many other places are not (especially smaller universities or CS depts run by engineering people).
+In many cases such places put a workshop paper in the same class as a top conference paper and also cannot differentiate between a good vs bad journals in CS (because they do not know CS well).
 Also,  they often value quantity over quality because they cannot properly evaluate paper quality.
@@
 The most common reason for not getting tenure is the lack of funding.
-Ideally research results, e.g., publications (\autoref{chap:research}), would be the more important thing for a researcher; but in reality, funding is often considered more important for tenure evaluation at most R1 universities.
+Ideally research results, e.g., publications (\autoref{chap:research}), would be the more important thing for a researcher; but in reality, funding is often considered more important for P\&T at most R1 universities.
 The amount of funding you have is a direct reflection of your research impact and productivity.
-It is also a reflection of your ability to write proposals, manage projects, and work with collaborators.
-More realistically, grants are needed to pay for your students, travel conference, research equipments, and even your own summer salary.
+More practically, grants are needed to pay for your students, travel conference, research equipments, and even your own summer salary.
 
@@
 \paragraph{Personal Experiences}
-As with most tenure track professors, I focus mostly on NSF grants and most of my funding comes from NSF.
+As with most TT professors, I focus mostly on NSF grants.
 I only have one grant from Army Research Office (ARO) in which I am a co-PI and didn't have to write the proposal.
-I have tried to get other DoD funding, but without much success, I found that DoD program managers are very picky and often have their own agenda.
+I have tried to get other DoD funding, but without much success, I find that DoD program managers difficult to make contact with and to get feedback from.
 I have also received internal funding from my university and industrial gifts from Amazon, and Facebook.

### 2024-08-19 demystify.tex
updates

@@
 \renewcommand{\appendixautorefname}{\S\@gobble}
+
+\newcommand*\l@chapterinfo{\@nodottedtocline{0}{0.0em}{1.5em}}
+\newcommand*\l@sectioninfo{\@nodottedtocline{1}{1.5em}{2.3em}}
+\newcommand*\l@subsectioninfo{\@nodottedtocline{2}{3.8em}{3.2em}}
+\newcommand*\l@subsubsectioninfo{\@nodottedtocline{3}{7.0em}{4.1em}}
+\newcommand*\l@paragraphinfo{\@nodottedtocline{4}{10em}{5em}}
+%\newcommand*\l@subparagraphinfo{\@nodottedtocline{5}{12em}{6em}}line{5}{12em}{6em}}
+
+% Copied from the book class macro `\@dottedtocline`. Removed the dots and page number
+\def\@nodottedtocline#1#2#3#4#5{%
+  \ifnum #1>\c@tocdepth \else
+    \vskip \z@ \@plus.2\p@
+    {\leftskip #2\relax \rightskip \@tocrmarg \parfillskip -\rightskip
+     \parindent #2\relax\@afterindenttrue
+     \interlinepenalty\@M
+     \leavevmode
+     \@tempdima #3\relax
+     \advance\leftskip \@tempdima \null\nobreak\hskip -\leftskip
+     {#4}\nobreak
+     \leaders\hbox{$\m@th
+        \mkern \@dotsep mu\hbox{\,}\mkern \@dotsep
+        mu$}\hfill
+     \nobreak
+     \hb@xt@\@pnumwidth{\hfil\normalfont \normalcolor }%
+     \par}%
+  \fi}
+
 \makeatother
+
+\def\chapterinfo#1{%
+  \addcontentsline{toc}{chapterinfo}{%
+    \noexpand\numberline{}\color{black}{#1}}%
+}
+\def\sectioninfo#1{%
+  \addcontentsline{toc}{sectioninfo}{%
+    \noexpand\numberline{}\color{black}{#1}}%
+}
+\def\subsectioninfo#1{%
+    \addcontentsline{toc}{subsectioninfo}{%
+    \noexpand\numberline{}\color{black}{#1}}%
+}
 
@@
 \newcommand{\tl}[1]{\mycomment{ThanhLe}{#1}}{}
+\newcommand{\xz}[1]{\mycomment{Xiaokuan}{[#1]}}{}
 \newcommand{\red}[1]{{\color{red}{#1}}}
-\newcommand{\xz}[1]{\mycomment{Xiaokuan}{[#1]}}{}
 
@@
 
-\chapter{Summary}\label{sec:summary}
-
-This section summarizes the main points of this guideline. It also gives you an overview to decide which specific topics you want to explore more thoroughly.
-
-\begin{enumerate}
-  \item Should you apply to a PhD CS program in the US?
-        \begin{itemize}
-          \item \emph{Yes, definitely}.  CS PhD study in the US is fully funded and admission into good universities is not any harder than in other countries (\autoref{sec:should}).
-        \end{itemize}
-  \item How is your application evaluated?
-        \begin{itemize}
-          \item Applications are evaluated by the \emph{PhD Admission} (adcom) committee and each application is typically reviewed by three faculty members (\autoref{sec:evalapps}).
-          \item Individually faculty member \emph{cannot directly admit} a student---so do not email and ask if you have a chance.
-However, faculty can \emph{advocate} for a student and therefore increase their admission chance---so do contact and introduce yourself (\autoref{sec:contact}).
-         \end{itemize}
-  \item Application Materials
-        \begin{itemize}
-          \item The committee will look at various factors, but the most important ones are letters of recommendation (LORs),  statements of purpose (SOP), and research experience and potential, e.g., publications.
-          \item \emph{LORs are very important}, but only if they are personalized and talk about your research ability (\autoref{sec:lor}).
-          \item \emph{SOP is very important}. Write it in such a way that makes you \emph{stand out} (\autoref{sec:research-statement} and \autoref{sec:improve-your-chance})
-          \item  \emph{Publications can greatly help}.
-Papers in good venues are concrete evidence that you have successfully engaged in research.
-But they are not required!
-(\autoref{sec:research-experience}).
-
-          \item \emph{High grades probably won't help much} (unless you're from a well-known school), but bad ones likely will raise concerns (\autoref{sec:your-school}).
-
-          \item \emph{Standard tests are not important} (\autoref{sec:standard-tests}).
-GRE typically \emph{is not} required.
-For standard English tests (not required for domestic students), just do enough to pass the minimum requirements.
-
-          \item \emph{Contacting a prof. is recommended}, but do it \emph{properly} to get a reply (\autoref{sec:contact}).
-
-          \item Getting an interview is typically a \emph{good sign}; but no interview does not mean rejection (\autoref{sec:interview}).
-
-        \end{itemize}
-  \item What to do after getting admitted?
-        \begin{itemize}
-                \item  Celebrating! Now it is your turn to evaluate the school!
-          \item Attend \emph{Open House} to learn more about the place and \emph{interview} profs---they would be much more willing to talk to you now (\autoref{sec:accepted}).
-        \end{itemize}
-  \item Funding
-        \begin{itemize}
-          \item TAs, RAs, and fellowships are main funding sources for PhDs (\autoref{sec:funding}).
-TAs (teaching assistantships) are provided by the department to help with classes (e.g., grading).
-RAs (research assistantships) are given by profs.
-to help with their research.
-Fellowships, provided by the university, department, or external sources such as government or industry, give move flexibility but are very competitive.
-        \end{itemize}
-  \item Choosing School and Professors
-        \begin{itemize}
-          \item Not every professor, even those in CS, can advise or graduate CS PhD students (\autoref{sec:schoolsandprofs}).
-        \end{itemize}
-
-    \item Miscs and FAQS
-        \begin{itemize}
-          \item One reason why you were rejected (\autoref{sec:why-rejected}) is because you applied to AI/ML, a super competitive field in recent years with many applicants.
-
-          \item Why rejection letters are sent so late?
-(\autoref{sec:late-rejection}).
-Grad programs often wait for the accepted students to make their decisions, typically by April 15, before sending out rejection letters.
-
-          \item Increasing admission chance by being unique and standing out (\autoref{sec:improve-your-chance}).
-          \item It takes about 10 minutes to review an application (\autoref{sec:ievaluate}).
-          \item Sometimes adcom members discuss applicants, but in most cases they make independent decisions (\autoref{sec:adcom-discuss}).           
-          \item Adcom members only review applications assigned to them (typically matching their expertise) and rarely get involved in other applications (\autoref{sec:read-all}).
-
-          \item Being a minority can significantly help your application (\autoref{sec:urm}).
-          \item You can successfully apply to CS PhD even if you have non-CS background (\autoref{sec:non-stem}).
-          \item You do not need an MS (\autoref{sec:msrequirement}), and CS PhD takes about 5--7 years (\autoref{sec:time}).
+% \chapter{Summary}\label{sec:summary}
+
+% This section summarizes the main points of this guideline. It also gives you an overview to decide which specific topics you want to explore more thoroughly.
+
+  %\item Should you apply to a PhD CS program in the US?
+
+          %\item \emph{Contacting a prof. is recommended}, but do it \emph{properly} to get a reply (\autoref{sec:contact}).
+
+          %\item  (\autoref{sec:interview}).
+
+    % \item Miscs and FAQS
+    %     \begin{itemize}
+
+    %       \item (\autoref{sec:read-all}).
+
+    %       \item
+    %       \item
+    %       \item You do not need an MS (\autoref{sec:msrequirement}), and CS PhD takes about 5--7 years (\autoref{sec:time}).
 Compared to other countries, CS PhD in the US does not require an MS degree but has a longer PhD study time (\autoref{sec:non-us-differences}).
-          \item What should you call your profs.? (\autoref{sec:address}) Many possibilities (e.g., prof., Dr., and even their firstnames), but probably not Mr. or Mrs.
-          \item Your entire PhD program costs about \$400K in total, but you \emph{do not} pay for it (\autoref{sec:ra-cost})
-          \item Ask your prof. if they can buy computer equipment and such for your research (\autoref{sec:buying-equipment}).
-          %\item Despite some miserable stories on social media, many PhD students have good mentors, supportive lab mates, healthy working environment ... and are happy (\autoref{sec:happy}).
-        \end{itemize}
-
-    \item Appendices
-    \begin{itemize}
-      \item Specific advice for domestic students (\autoref{sec:domestic-students}).
-      \item How to find research opportunities (\autoref{sec:research-opportunities}).
-      \item Cultural differences between US and other countries (\autoref{sec:cultural}).
-    \end{itemize}
-
-\end{enumerate}
+    %       \item
+    %       \item
+    %       \item
+    %       %\item Despite some miserable stories on social media, many PhD students have good mentors, supportive lab mates, healthy working environment ... and are happy (\autoref{sec:happy}).
+    %     \end{itemize}
+
+    % \item Appendices
+    % \begin{itemize}
+    %   \item Specific advice for domestic students (\autoref{sec:domestic-students}).
+    %   \item How to find research opportunities (\autoref{sec:research-opportunities}).
+    %   \item Cultural differences between US and other countries (\autoref{sec:cultural}).
+    % \end{itemize}
 
@@
 \chapter{Should You Apply to a CS PhD Program in the US?}\label{sec:should}
+
+\chapterinfo{\emph{Yes, definitely}.  CS PhD study in the US is fully funded and admission into good universities is not any harder than in other countries.}
 
@@
 
+\chapterinfo{Applications are evaluated by the \emph{PhD Admission} (adcom) committee and each application is typically reviewed by three faculty members.}
+
 \epigraph{How is education supposed to make me feel smarter?
@@
 
-After you submit your PhD application (e.g., in late Fall of year $X$ for the Fall admission of year $X+1$), it will first be screened
+After you submit your PhD application (e.g., in late Fall of year $X$ for the Fall admission of year $X+1$), it will first be ``screened''
 for general requirements, e.g., did you submit your transcripts and standard scores?
@@
 
-Then, applications are reviewed by a \emph{PhD admission committee} (\textbf{adcom}) that consists of faculty members in CS\footnote{In some cases the committee can involve affiliated faculty from different disciplines.}.
+\paragraph{Adcom} After screening applications are reviewed by a \emph{PhD admission committee} (\textbf{adcom}) that consists of faculty members in CS\footnote{In some cases the committee can involve affiliated faculty from different disciplines.}.
 These adcom members have a wide range of expertise and background to ensure diverse perspectives in the evaluation process.
@@
 
+\chapterinfo{The committee will look at various factors, but the most important ones are letters of recommendation (LORs),  statements of purpose (SOP), and research experience and potential, e.g., publications.}
+
 \epigraph{Son, if you really want something in this life, you have to work for it. Now quiet! They’re about to announce the lottery numbers.}{\textsc{The Simpsons}}
 
-The focus of adcom is to evaluate your background and interest in research to see if you \textbf{fit into our PhD program}!
+The goal of adcom is to evaluate your research experience, potential, and interest to see if you \textbf{fit into our PhD program}!
 The emphasis here is \emph{fitting}, which varies from school to school, faculty to faculty, and even from year to year.
@@
 \section{Letters of Recommendation (LOR)}\label{sec:lor}
+\sectioninfo{\emph{LORs are very important}, but only if they are personalized and talk about your research ability.}
 
@@
 
-\section{Statement of Purpose (SOP)}\label{sec:research-statement}
+\section{Statement of Purpose (SOP)}\label{sec:sop}
+\sectioninfo{\\emph{SOP is very important}. Write it in such a way that makes you \emph{stand out}.}
 
@@
 
+\sectioninfo{\emph{Publications can greatly help}. Papers in good venues are concrete evidence that you have successfully engaged in research. But they are not required!}
+
 This section discusses various research experiences that can strengthen your application.
@@
 Thus, general writings, even those under submissions or rejected, would still be good (much better than nothing).
-Be sure to upload your papers with your application and talk about them in your statement (see \autoref{sec:research-statement}).
+Be sure to upload your papers with your application and talk about them in your statement (see \autoref{sec:sop}).
 Note that local conferences and non-English journals or conferences do
@@
 \section{Your School and Grades}\label{sec:your-school}
+\sectioninfo{\emph{High grades probably won't help much} (unless you're from a well-known school), but bad ones likely will raise concerns.}
 
@@
 \section{Standard Tests (e.g., GRE and IELTS), Online Courses, and Certificates}\label{sec:standard-tests}
-
-\subsection{GREs Do Not Matter Much} While a few schools still require taking the GRE (e.g., UCF), most good CS PhD programs in the US \textbf{no longer} require it.
+\sectioninfo{\emph{Standard tests are not important}.
+GRE typically \emph{is not} required.
+For standard English tests (not required for domestic students), just do enough to pass the minimum requirements.}
+
+\subsection{GREs Are Optional and Do Not Matter} While a few schools still require taking the GRE (e.g., UCF), most good CS PhD programs in the US \textbf{no longer} require it.
 Thus, if you haven't taken the GRE or have bad scores, then don't waste time (re)taking it.
@@
 \section{Interview and the Waiting Game}\label{sec:interview}
+\sectioninfo{Getting an interview is typically a \emph{good sign}; but no interview does not mean rejection}
 
@@
 
+\chapterinfo{Celebrating! Now it is your turn to evaluate the school!
+Attend \emph{Open House} to learn more about the place and \emph{interview} profs---they would be much more willing to talk to you now.}
+
 \epigraph{``Oh...
@@
 \chapter{Funding}\label{sec:funding}
+
+\chapterinfo{TAs, RAs, and fellowships are main funding sources for PhDs.
+TAs (teaching assistantships) are provided by the department to help with classes (e.g., grading).
+RAs (research assistantships) are given by profs.
+to help with their research.
+Fellowships, provided by the university, department, or external sources such as government or industry, give move flexibility but are very competitive.}
 
@@
 For a full breakdown of how much a graduate student costs, see \autoref{sec:ra-cost}.
+\sectioninfo{Your entire PhD program costs about \$400K in total, but you \emph{do not} pay for it.}
 
@@
 
+\chapterinfo{Not every professor, even those in CS, can advise or graduate CS PhD students.}
+
 \epigraph{\vspace{-0.2in} The answer to life's problems aren't at the bottom of a bottle. Heh heh! They're on TV!}{\textsc{The Simpsons}}
@@
 
+\subsectioninfo{Individually faculty member \emph{cannot directly admit} a student---so do not email and ask if you have a chance.
+However, faculty can \emph{advocate} for a student and therefore increase their admission chance---so do contact and introduce yourself.}
+
 \begin{center}
@@
 
+\sectioninfo{One reason why you were rejected is because you applied to AI/ML, a super competitive field in recent years with many applicants.}
+
 Many students lament that they get no interviews or are rejected (with strong profiles) and that the admission process seems random.
@@
 \section{Timeline: Why are rejection letters sent so late?}\label{sec:late-rejection}
+\sectioninfo{Why rejection letters are sent so late? Grad programs often wait for the accepted students to make their decisions, typically by April 15, before sending out rejection letters.}
 
@@
 
+\sectioninfo{Increasing admission chance by being unique and standing out.}
+
 \begin{center}
@@
 \section{How long do you evaluate an application?}\label{sec:ievaluate}
+\sectioninfo{It takes about 10 minutes to review an application.}
 
@@
 \section{Do adcom members talk to each other?}\label{sec:adcom-discuss}
+\sectioninfo{Sometimes adcom members discuss applicants, but in most cases they make independent decisions.}
 
@@
 \section{Do adcom members read all applications?}\label{sec:read-all}
+\sectioninfo{Adcom members only review applications assigned to them (typically matching their expertise) and rarely get involved in other applications}
 
@@
 \section{Member of an underrepresented or LGBTQ+ groups}\label{sec:urm}
+\sectioninfo{Being a minority can significantly help your application.}
 
@@
 \section{My undergrad was not in CS or related areas}\label{sec:non-stem}
+\sectioninfo{You can successfully apply to CS PhD even if you have non-CS background.}
 
@@
 \section{How do I call or address a professor?}\label{sec:address}
+\sectioninfo{What should you call your profs.? Many possibilities (e.g., prof., Dr., and even their firstnames), but not Mr. or Mrs.}
 
@@
 \section{Buying Computer Equipment}\label{sec:buying-equipment}
+\sectioninfo{Ask your prof. if they can buy computer equipment and such for your research.}
 
@@
 By contributing code, fixing bugs, implementing new features, or providing documentation, students can gain practical research experience and interact with experienced developers and researchers in the field.
-Not only you gain research experience, but you might be able to get a LoR from the project maintainer, and you should write about this experience in your SOP (\autoref{sec:research-statement}.)
+Not only you gain research experience, but you might be able to get a LoR from the project maintainer, and you should write about this experience in your SOP (\autoref{sec:sop}.)
 
@@
   \item[LoR] Letter of Recommendation. A letter written by a professor or supervisor that assesses your qualifications and potential for graduate study (\autoref{sec:lor}).
-  \item[SOP] Statement of Purpose. A document written by yourself to explains your research interests, background, and reasons for applying to a PhD program (\autoref{sec:research-statement}).
+  \item[SOP] Statement of Purpose. A document written by yourself to explains your research interests, background, and reasons for applying to a PhD program (\autoref{sec:sop}).
 

### 2024-08-20 demystify.tex
updates

@@
     %       \item
-    %       \item You do not need an MS (\autoref{sec:msrequirement}), and CS PhD takes about 5--7 years (\autoref{sec:time}).
-Compared to other countries, CS PhD in the US does not require an MS degree but has a longer PhD study time (\autoref{sec:non-us-differences}).
+    %       \item (\autoref{sec:msrequirement}), and CS PhD takes abou (\autoref{sec:time}). (\autoref{sec:non-us-differences}).
     %       \item
@@
 
-\chapterinfo{\emph{Yes, definitely}.  CS PhD study in the US is fully funded and admission into good universities is not any harder than in other countries.}
+\chapterinfo{\emph{Yes, definitely}.  CS PhD study in the US is fully funded and admission into good universities in the US is not any harder than in other countries.}
 
@@
 
-\section{Who is this document for?}
-
-The guideline can be useful for all students, but it is primarily intended for \textbf{international students} from \textbf{smaller countries and less well-known universities} (or domestic students from US universities with no PhD programs or limited research experiences).
-Students from top schools with strong research cultures and experiences might already know most of the information given here.
-My goal is thus to level the playing field by providing information that is not readily available to less privileged students. 
-I hope to encourage more students with such backgrounds to apply to CS PhD programs in the US.
-
-While this document may not fully meet the needs of students aiming for top-10 schools, which typically have more stringent requirements (e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU), it should be valuable for those targeting other reputable R1 institutions.
-It is certainly useful for students aiming for Mason!
-
-%% Vu: *what's a PhD?*  This [[https://matt.might.net/articles/phd-school-in-pictures/][series of pictures]] from [[https://matt.might.net][Matt Might]] illustrates what a PhD means.
-%% \end{commentbox}
-
-\section*{Additional Resources}
-\begin{itemize}
-  \item \href{https://parentheticallyspeaking.org/articles/us-cs-phd-faq/}{Getting a Computer Science PhD in the USA} by Shriram Krishnamurthi
-  \item  \href{https://pg.ucsd.edu/PhD-application-tips.htm}{Blog post on PhD Admission} by Philip Guo
-  \item   \href{https://talkingtorobots.com/FAQ.html}{PhD Admission post} by Yonatan Bisk (customized for CMU admission)
-\end{itemize}
-
-\chapter{How is Your Application Evaluated?}\label{sec:evalapps}
-
-\chapterinfo{Applications are evaluated by the \emph{PhD Admission} (adcom) committee and each application is typically reviewed by three faculty members.}
+\section{How is Your Application Evaluated?}\label{sec:evalapps}
+
+\sectioninfo{Applications are evaluated by the \emph{PhD Admission} (adcom) committee and each application is typically reviewed by three faculty members.}
 
@@
 
-Each application is assigned to about 3 adcom members, who will evaluate your profile and reach a consensus.
+Each application is assigned to about three adcom members, who will evaluate your profile and reach a consensus.
 While the assigned reviewers are the main ones deciding your application, other faculty in the department will likely have access to your application and can provide inputs and opinions on your profile.
@@
 
+\section{Who is this document for?}
+\sectioninfo{This document is mainly intended for international students from smaller countries and domestic students from less well-known universities.}
+
+The guideline can be useful for all students, but it is primarily intended for \textbf{international students} from \textbf{smaller countries and less well-known universities} (or domestic students from US universities with no PhD programs or limited research experiences).
+Students from top schools with strong research cultures and experiences might already know most of the information given here.
+My goal is thus to level the playing field by providing information that is not readily available to less privileged students. 
+I hope to encourage more students with such backgrounds to apply to CS PhD programs in the US.
+
+While this document may not fully meet the needs of students aiming for top-10 schools, which typically have more stringent requirements (e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU), it should be valuable for those targeting other reputable R1 institutions.
+It is certainly useful for students aiming for Mason!
+
+%% Vu: *what's a PhD?*  This [[https://matt.might.net/articles/phd-school-in-pictures/][series of pictures]] from [[https://matt.might.net][Matt Might]] illustrates what a PhD means.
+%% \end{commentbox}
+
+\section*{Additional Resources}
+\begin{itemize}
+  \item \href{https://parentheticallyspeaking.org/articles/us-cs-phd-faq/}{Getting a Computer Science PhD in the USA} by Shriram Krishnamurthi
+  \item  \href{https://pg.ucsd.edu/PhD-application-tips.htm}{Blog post on PhD Admission} by Philip Guo
+  \item   \href{https://talkingtorobots.com/FAQ.html}{PhD Admission post} by Yonatan Bisk (customized for CMU admission)
+\end{itemize}
+
 \chapter{Your Application}\label{sec:application}
@@
 \section{Statement of Purpose (SOP)}\label{sec:sop}
-\sectioninfo{\\emph{SOP is very important}. Write it in such a way that makes you \emph{stand out}.}
+\sectioninfo{\emph{SOP is very important}. Write it in such a way that makes you \emph{stand out}.}
 
@@
 \section{Your School and Grades}\label{sec:your-school}
-\sectioninfo{\emph{High grades probably won't help much} (unless you're from a well-known school), but bad ones likely will raise concerns.}
+\sectioninfo{\emph{High grades probably won't help much} (unless you're from a very top and well-known school), but bad ones likely will raise concerns.}
 
@@
 \section{CV/Resume}
+\sectioninfo{Highlight and summarize major achievements, e.g., Publications, Programming Competition Awards, and Teaching Experience.}
+
 The CV/Resume provides a summary of the applicant's achievements. Most schools require you to upload your CV with your application.
@@
 
-\chapterinfo{Celebrating! Now it is your turn to evaluate the school!
+\chapterinfo{Celebrate! Now it is your turn to evaluate the school!
 Attend \emph{Open House} to learn more about the place and \emph{interview} profs---they would be much more willing to talk to you now.}
@@
 \chapterinfo{TAs, RAs, and fellowships are main funding sources for PhDs.
-TAs (teaching assistantships) are provided by the department to help with classes (e.g., grading).
-RAs (research assistantships) are given by profs.
+TAs are provided by the department to help with classes (e.g., grading).
+RAs are given by profs.
 to help with their research.
-Fellowships, provided by the university, department, or external sources such as government or industry, give move flexibility but are very competitive.}
+Fellowships, provided by the university, department, or external sources such as government or industry, give move flexibility but can be very competitive.}
 
@@
 For a full breakdown of how much a graduate student costs, see \autoref{sec:ra-cost}.
-\sectioninfo{Your entire PhD program costs about \$400K in total, but you \emph{do not} pay for it.}
 
@@
 \subsection{How much do \emph{I} cost?}\label{sec:ra-cost}
+\sectioninfo{Your entire PhD program costs about \$400K in total, but you \emph{do not} pay for it.}
 
@@
 
-\sectioninfo{One reason why you were rejected is because you applied to AI/ML, a super competitive field in recent years with many applicants.}
+\sectioninfo{You aim too high, are overqualified, or even because you applied to AI/ML, a super competitive field in recent years with many applicants.}
 
@@
 \section{Is an MS required for admission to PhD in CS?}\label{sec:msrequirement}
+\sectioninfo{You do not need an MS to do PhD in CS.}
+
 No.
@@
 \section{How long to complete the CS PhD program?}\label{sec:time}
+\sectioninfo{About 5--7 years in the US}
 
@@
 \section{PhD in the US vs. Other Countries}\label{sec:non-us-differences}
+\sectioninfo{Compared to other countries, CS PhD in the US does not require an MS degree but has a longer PhD study time.}
 

### 2024-08-20 demystify.tex
updates

@@
     %       \item
-    %       \item (\autoref{sec:msrequirement}), and CS PhD takes abou (\autoref{sec:time}). (\autoref{sec:non-us-differences}).
+    %       \item (\autoref{sec:msrequirement}), and CS PhD takes abou (\autoref{sec:time}). Pu(\autoref{sec:non-us-differences}).
     %       \item
@@
 
-\subsection{How much do \emph{I} cost?}\label{sec:ra-cost}
+\subsection{How much do you cost?}\label{sec:ra-cost}
 \sectioninfo{Your entire PhD program costs about \$400K in total, but you \emph{do not} pay for it.}
@@
 
-\section{Why did I get rejected?}\label{sec:why-rejected}
+\section{Why did you get rejected?}\label{sec:why-rejected}
 
@@
 
-\section{Increasing my admission chance}\label{sec:improve-your-chance}
+\section{Increasing your admission chance}\label{sec:improve-your-chance}
 
@@
 
-\section{How long do you evaluate an application?}\label{sec:ievaluate}
-\sectioninfo{It takes about 10 minutes to review an application.}
+\section{How long does it take to evaluate an application?}\label{sec:ievaluate}
+\sectioninfo{It takes me about 10 minutes to review an application.}
 
@@
 
-\subsection{How Decisions are Made}\label{sec:how-decisions}
+\section{How are decisions made?}\label{sec:how-decisions}
+\sectioninfo{Even if all adcom reviewers recommend acceptance, the application can still be rejected. Vice versa, if all reviewers think the application is weak, the student might still be admitted.}
 
@@
 \section{Member of an underrepresented or LGBTQ+ groups}\label{sec:urm}
-\sectioninfo{Being a minority can significantly help your application.}
+\sectioninfo{Being a minority can make you stand out and help your application.}
 
@@
 \section{PhD in the US vs. Other Countries}\label{sec:non-us-differences}
-\sectioninfo{Compared to other countries, CS PhD in the US does not require an MS degree but has a longer PhD study time.}
+\sectioninfo{Among several differences, CS PhD in the US does not require an MS degree but has a longer PhD study time.}
 
@@
 
-\section{How do I call or address a professor?}\label{sec:address}
+\section{How to call or address a professor?}\label{sec:address}
 \sectioninfo{What should you call your profs.? Many possibilities (e.g., prof., Dr., and even their firstnames), but not Mr. or Mrs.}
@@
 \appendix
-
-\chapter{Visa for International Students}\label{sec:visa}
-
-\epigraph{\vspace{-0.2in} I didn’t do it. Nobody saw me do it. You can’t prove anything!}{\textsc{The Simpsons}}
-
-% \begin{center}
-%     \includegraphics[scale=0.8]{files/visa.pdf}
-% \end{center}
-
-For international students pursuing a PhD in the US, \textbf{F-1} is the main visa that allows a student to study full-time at an accredited institution.  Here are some key points about the F-1 visa:
-
-\begin{itemize}
-\item \textbf{Employment:} You are allowed to work on-campus for up to \textbf{20 hours per week during the academic year} (because you still need to take classes) and \textbf{full-time} (typically 40 hrs) during official school breaks (e.g., summer and winter breaks).
-Off-campus employment requires authorization, which can be obtained through programs like Curricular Practical Training and Optional Practical Training described below.
-
-\item \textbf{Curricular Practical Training (CPT):} CPT allows you to participate in internships or practical training that is an integral part of their academic curriculum.
-CPT must be related to your field of study and can be full-time or part-time.
-It requires prior authorization from your university and must be completed before graduation.
-
-\item \textbf{Optional Practical Training (OPT):} OPT provides up to 12 months of work authorization for students before or after completing their degree.
-For students in STEM fields, there is an additional 24-month extension available.
-OPT requires prior application and approval from USCIS.
-
-\item \textbf{Full-time Enrollment:} You must maintain full-time enrollment status during the academic year.
-This means taking a minimum number of credits each semester, as defined by your program.
-Dropping below full-time status can result in loss of visa status.
-\end{itemize}
-
-\includepdf[pages=-]{files/visa.pdf}
-
-\section{For Spouses and Children}
-
-The spouses and children of F-1 visa holders can enter the US under the F-2 visa status. The F-2 visa allows family members to live in the US with the following conditions:
-
-For Spouse: 
-\begin{itemize}
-\item \textbf{Work Restrictions:} F-2 spouses are not permitted to work in the US.
-\item \textbf{Education:} They can study part-time but cannot enroll in full-time degree programs.
-\end{itemize}
-For Children:
-\begin{itemize}
-\item \textbf{Education:} F-2 children can attend K-12 schools but cannot pursue higher education full-time.
-\item \textbf{Work Restrictions:} Like F-2 spouses, children are not allowed to work under any circumstances.
-\end{itemize}
-F-2 visa holders must leave the US if the primary F-1 student visa holder loses status or completes their program.
-
-\chapter{Domestic Students}\label{sec:domestic-students}
-
-\epigraph{\vspace{-0.2in} I’m not a bad guy! I work hard, and I love my kids. So why should I spend half my Sunday hearing about how I’m going to Hell?}{\textsc{The Simpsons}}
-
-Most of what is written in this document applies to both domestic\footnote{For simplicity domestic means you did your undergrad (or MS) at a US university.} and international students.
-However, there are some differences and benefits that domestic students should be aware of and can leverage to improve their chances of admission.
-
-\paragraph{Standing out \autoref{sec:improve-your-chance}} There are \emph{few} domestic applications compared to international ones, i.e., domestic students are the minority (\autoref{sec:urm}) in the CS PhD application pool.
-Many US universities want to increase the number of domestic students in their programs (and as mentioned later, there are specific fellowships and funding for domestic students).
-That makes you stand out and can help your case.
-
-\paragraph{Fee Waiving} Some schools might offer application fee waivers for domestic students.  You should check with the school you're applying to.
-
-\paragraph{School \autoref{sec:your-school}} Adcom already knows about your school, which is a plus (see \autoref{sec:your-school}).
-You are also more familiar with the US education system and culture (so hopefully you are aware of the cultural issues listed in \autoref{sec:cultural}).
-
-\paragraph{Standard Tests \autoref{sec:standard-tests}} You do not need to take TOEFL or IELTS because you already did your undergrad (or MS) at a US university.
-You might also be more comfortable communicating in English, e.g., contacting professors (\autoref{sec:contact}).
-
-\paragraph{Transcripts} You do not need to get your transcripts evaluated/translated (which can be a hassle and inaccurate).
-You can just send your official transcripts directly to the school you're applying to.
-
-\paragraph{Funding \autoref{sec:funding}} You have more opportunities for funding, e.g., through government scholarships for US citizens and permanent residents.
-You can also apply to specific programs \emph{before} you start your PhD, e.g., NSF Graduate Research Fellowship Program (GRFP) and Hertz Foundation Fellowship.
-These fellowships are very competitive and can significantly improve your admission chances.
-
-\paragraph{Research Experience \autoref{sec:research-opportunities}} You might have many opportunities to do research as an undergraduate, e.g., through REUs and internships at your undergrad university.
-Highlight such experience in your application.
-
-\paragraph{Open House \autoref{sec:accepted}} It is easier for you to attend open house events in person.  This can help you make a better decision on which school to attend.
-
-% \begin{domesticbox}[Domestic students:] If you're a domestic student, you have several advantages in your application.
-You also have more opportunities for funding, e.g., through government scholarships for US citizens and residents.
-Finally,
-% \end{domesticbox}
-
-\chapter{Research Opportunities}\label{sec:research-opportunities}
-
-\epigraph{\vspace{-0.2in} Kids, you tried your best and you failed miserably. The lesson is: never try.}{\textsc{The Simpsons}}
-
-\begin{center}
-    \includegraphics[scale=0.5]{files/phd100404s.png}
-  \end{center}
-
-As discussed in \autoref{sec:research-opportunities}, research experience is a crucial component of a strong PhD application.
-It gives you opportunities to try out research, determine what research area you're interested in, publish papers (\autoref{sec:research-experience}), connect with researchers, and get strong LoRs (\autoref{sec:lor}).
-
- This section provides some guidance on how to gain research experience, e.g., as an undergraduate student.  It also advises students from small colleges where research opportunities are limited.
-
-\paragraph{Locally} Start looking for research opportunities at your institution.
-If you did well or liked a class, you can check with the professor of that class for research opportunities.
-You can also go to the department directory and then professors' websites and see if they are looking for undergraduate researchers.
-Even if they say they are looking for graduate students, you can still contact them and ask.
-
-There might be also special programs for undergraduate research from the university, e.g., Mason has the OSCAR program, UNL has UCARE, and NSF has REUs (for US citizens and permanent residents).
-
-You can also take honor thesis or independent study courses with a professor.
-This is a good way to get research experience and also get a LoR from the professor.
-You can also ask your academic adviser or other faculty members for suggestions.
-Finally, you can also ask your peers who are already doing research.
-Use the method described in \autoref{sec:contact} to contact professors
-
-\begin{commentbox}[Vu:]
-    I enjoy working with undergrads (e.g., have mentored over 10 undergrads) and always actively recruit.
-    I often get undergraduate researchers through my classes, e.g., asking students who did well in my class if they are interested in research.
-Occasionally I was introduced to students by other students or faculty.
-
-    While most undergrads are understandably not super productive in research, some are and I have published multiple papers with them.
-I also have written strong LoRs for them and have helped them get into PhD programs, including top ones.
-
-    \tcblower
-    At Mason we pay undergrads about \$15/hr and about 10 hrs/week. This is flexible and can be increased as needed.
-\end{commentbox}
-
-\paragraph{Online Research Communities and Open Source Contributions} Places such as GitHub offer many research opportunities, even though they might not appear so (e.g., they do not explicitly mention research).
-In many cases, professors or research labs put their projects on GitHub.
-For example, my research group, \href{https://github.com/dynaroars/}{Dynaroars}, has many open-source projects that undergrads can contribute to.
-
-By contributing code, fixing bugs, implementing new features, or providing documentation, students can gain practical research experience and interact with experienced developers and researchers in the field.
-Not only you gain research experience, but you might be able to get a LoR from the project maintainer, and you should write about this experience in your SOP (\autoref{sec:sop}.)
-
-\paragraph{Virtual Research Opportunities} This is less common but several places offer virtual internships and research programs aimed at providing hands-on research experience.
-These programs often involve working remotely under the guidance of experienced mentors and collaborating with a team of fellow researchers.
-For instance, \href{https://docs.google.com/forms/d/1btIwt4HwjyKMOUk-EMy3rbkfWzFxv2lNrMm_zkd0pA4/viewform?edit_requested=true}{UIUC+ Summer Undergraduate Research in Software Engineering}  offers an unpaid remote internship for software engineering students all over the world to collaborate with mentors from University of Illinois at Urbana Champaign.
-
-% \paragraph{Online Conferences and Workshops}: Attending online conferences and workshops is another way to gain exposure to cutting-edge research and establish connections with experts in the field.
-Many conferences now provide virtual participation options, enabling students to access research talks, poster sessions, and panel discussions and sometimes access designated chat rooms or networking events where participants can engage with researchers, ask questions, and seek potential research collaborations.
-By creating profiles, joining relevant research groups, and participating in discussions, students can connect with established researchers, contribute to ongoing projects, and potentially collaborate on publications or research proposals as they provide access
-
 \chapter{Glossary}\label{sec:glossary}
@@
 PhD students typically do not have to pay tuition as it is covered by their funding, but this might be important for MS students.
+  \item [Lab] Research Lab.
+A group of researchers (profs.
+and students) working on a common research area, e.g., a Software Engineering Lab.
+Typically a lab is led by a single professor, but sometimes it can be co-led by multiple professors.
+For example, my lab is DynaROARS, which focuses on my own research area (software verification), but I am also part of the more general SWE Lab at Mason with other faculty in Software Engineering.
+
   \item[LoR] Letter of Recommendation. A letter written by a professor or supervisor that assesses your qualifications and potential for graduate study (\autoref{sec:lor}).
+  \item[MS] Master of Science. A graduate degree that typically requires 2 years of study.
+  \item[Open House] An event where admitted students visit the university to meet faculty, students, and staff, and learn more about the program.
+This is a great opportunity to get a feel for the university and the department and decide if it is a good fit for you.
+
   \item[SOP] Statement of Purpose. A document written by yourself to explains your research interests, background, and reasons for applying to a PhD program (\autoref{sec:sop}).
@@
 
-  \item [Lab] Research Lab.
-A group of researchers (profs.
-and students) working on a common research area, e.g., a Software Engineering Lab.
-Typically a lab is led by a single professor, but sometimes it can be co-led by multiple professors.
-For example, my lab is DynaROARS, which focuses on my own research area (software verification), but I am also part of the more general SWE Lab at Mason with other faculty in Software Engineering.
-
-  \item[MS] Master of Science. A graduate degree that typically requires 2 years of study.
   \item [PI]  Principal Investigator.
@@
 \end{description}
+
+\chapter{Visa for International Students}\label{sec:visa}
+
+\epigraph{\vspace{-0.2in} I didn’t do it. Nobody saw me do it. You can’t prove anything!}{\textsc{The Simpsons}}
+
+% \begin{center}
+%     \includegraphics[scale=0.8]{files/visa.pdf}
+% \end{center}
+
+For international students pursuing a PhD in the US, \textbf{F-1} is the main visa that allows a student to study full-time at an accredited institution.  Here are some key points about the F-1 visa:
+
+\begin{itemize}
+\item \textbf{Employment:} You are allowed to work on-campus for up to \textbf{20 hours per week during the academic year} (because you still need to take classes) and \textbf{full-time} (typically 40 hrs) during official school breaks (e.g., summer and winter breaks).
+Off-campus employment requires authorization, which can be obtained through programs like Curricular Practical Training and Optional Practical Training described below.
+
+\item \textbf{Curricular Practical Training (CPT):} CPT allows you to participate in internships or practical training that is an integral part of their academic curriculum.
+CPT must be related to your field of study and can be full-time or part-time.
+It requires prior authorization from your university and must be completed before graduation.
+
+\item \textbf{Optional Practical Training (OPT):} OPT provides up to 12 months of work authorization for students before or after completing their degree.
+For students in STEM fields, there is an additional 24-month extension available.
+OPT requires prior application and approval from USCIS.
+
+\item \textbf{Full-time Enrollment:} You must maintain full-time enrollment status during the academic year.
+This means taking a minimum number of credits each semester, as defined by your program.
+Dropping below full-time status can result in loss of visa status.
+\end{itemize}
+
+\includepdf[pages=-]{files/visa.pdf}
+
+\section{For Spouses and Children}
+
+The spouses and children of F-1 visa holders can enter the US under the F-2 visa status. The F-2 visa allows family members to live in the US with the following conditions:
+
+For Spouse: 
+\begin{itemize}
+\item \textbf{Work Restrictions:} F-2 spouses are not permitted to work in the US.
+\item \textbf{Education:} They can study part-time but cannot enroll in full-time degree programs.
+\end{itemize}
+For Children:
+\begin{itemize}
+\item \textbf{Education:} F-2 children can attend K-12 schools but cannot pursue higher education full-time.
+\item \textbf{Work Restrictions:} Like F-2 spouses, children are not allowed to work under any circumstances.
+\end{itemize}
+F-2 visa holders must leave the US if the primary F-1 student visa holder loses status or completes their program.
+
+\chapter{Domestic Students}\label{sec:domestic-students}
+\chapterinfo{Specific benefits and opportunities for domestic students applying to CS PhD programs.}
+
+\epigraph{\vspace{-0.2in} I’m not a bad guy! I work hard, and I love my kids. So why should I spend half my Sunday hearing about how I’m going to Hell?}{\textsc{The Simpsons}}
+
+Most of what is written in this document applies to both domestic\footnote{For simplicity domestic means you did your undergrad (or MS) at a US university.} and international students.
+However, there are some differences and benefits that domestic students should be aware of and can leverage to improve their chances of admission.
+
+\paragraph{Standing out \autoref{sec:improve-your-chance}} There are \emph{few} domestic applications compared to international ones, i.e., domestic students are the minority (\autoref{sec:urm}) in the CS PhD application pool.
+Many US universities want to increase the number of domestic students in their programs (and as mentioned later, there are specific fellowships and funding for domestic students).
+That makes you stand out and can help your case.
+
+\paragraph{Fee Waiving} Some schools might offer application fee waivers for domestic students.  You should check with the school you're applying to.
+
+\paragraph{School \autoref{sec:your-school}} Adcom already knows about your school, which is a plus (see \autoref{sec:your-school}).
+You are also more familiar with the US education system and culture (so hopefully you are aware of the cultural issues listed in \autoref{sec:cultural}).
+
+\paragraph{Standard Tests \autoref{sec:standard-tests}} You do not need to take TOEFL or IELTS because you already did your undergrad (or MS) at a US university.
+You might also be more comfortable communicating in English, e.g., contacting professors (\autoref{sec:contact}).
+
+\paragraph{Transcripts} You do not need to get your transcripts evaluated/translated (which can be a hassle and inaccurate).
+You can just send your official transcripts directly to the school you're applying to.
+
+\paragraph{Funding \autoref{sec:funding}} You have more opportunities for funding, e.g., through government scholarships for US citizens and permanent residents.
+You can also apply to specific programs \emph{before} you start your PhD, e.g., NSF Graduate Research Fellowship Program (GRFP) and Hertz Foundation Fellowship.
+These fellowships are very competitive and can significantly improve your admission chances.
+
+\paragraph{Research Experience \autoref{sec:research-opportunities}} You might have many opportunities to do research as an undergraduate, e.g., through REUs and internships at your undergrad university.
+Highlight such experience in your application.
+
+\paragraph{Open House \autoref{sec:accepted}} It is easier for you to attend open house events in person.  This can help you make a better decision on which school to attend.
+
+% \begin{domesticbox}[Domestic students:] If you're a domestic student, you have several advantages in your application.
+You also have more opportunities for funding, e.g., through government scholarships for US citizens and residents.
+Finally,
+% \end{domesticbox}
+
+\chapter{Research Opportunities}\label{sec:research-opportunities}
+\chapterinfo{How to get research experience as an undergrad.}
+
+\epigraph{\vspace{-0.2in} Kids, you tried your best and you failed miserably. The lesson is: never try.}{\textsc{The Simpsons}}
+
+\begin{center}
+    \includegraphics[scale=0.5]{files/phd100404s.png}
+  \end{center}
+
+As discussed in \autoref{sec:research-opportunities}, research experience is a crucial component of a strong PhD application.
+It gives you opportunities to try out research, determine what research area you're interested in, publish papers (\autoref{sec:research-experience}), connect with researchers, and get strong LoRs (\autoref{sec:lor}).
+
+ This section provides some guidance on how to gain research experience, e.g., as an undergraduate student.  It also advises students from small colleges where research opportunities are limited.
+
+\paragraph{Locally} Start looking for research opportunities at your institution.
+If you did well or liked a class, you can check with the professor of that class for research opportunities.
+You can also go to the department directory and then professors' websites and see if they are looking for undergraduate researchers.
+Even if they say they are looking for graduate students, you can still contact them and ask.
+
+There might be also special programs for undergraduate research from the university, e.g., Mason has the OSCAR program, UNL has UCARE, and NSF has REUs (for US citizens and permanent residents).
+
+You can also take honor thesis or independent study courses with a professor.
+This is a good way to get research experience and also get a LoR from the professor.
+You can also ask your academic adviser or other faculty members for suggestions.
+Finally, you can also ask your peers who are already doing research.
+Use the method described in \autoref{sec:contact} to contact professors
+
+\begin{commentbox}[Vu:]
+    I enjoy working with undergrads (e.g., have mentored over 10 undergrads) and always actively recruit.
+    I often get undergraduate researchers through my classes, e.g., asking students who did well in my class if they are interested in research.
+Occasionally I was introduced to students by other students or faculty.
+
+    While most undergrads are understandably not super productive in research, some are and I have published multiple papers with them.
+I also have written strong LoRs for them and have helped them get into PhD programs, including top ones.
+
+    \tcblower
+    At Mason we pay undergrads about \$15/hr and about 10 hrs/week. This is flexible and can be increased as needed.
+\end{commentbox}
+
+\paragraph{Online Research Communities and Open Source Contributions} Places such as GitHub offer many research opportunities, even though they might not appear so (e.g., they do not explicitly mention research).
+In many cases, professors or research labs put their projects on GitHub.
+For example, my research group, \href{https://github.com/dynaroars/}{Dynaroars}, has many open-source projects that undergrads can contribute to.
+
+By contributing code, fixing bugs, implementing new features, or providing documentation, students can gain practical research experience and interact with experienced developers and researchers in the field.
+Not only you gain research experience, but you might be able to get a LoR from the project maintainer, and you should write about this experience in your SOP (\autoref{sec:sop}.)
+
+\paragraph{Virtual Research Opportunities} This is less common but several places offer virtual internships and research programs aimed at providing hands-on research experience.
+These programs often involve working remotely under the guidance of experienced mentors and collaborating with a team of fellow researchers.
+For instance, \href{https://docs.google.com/forms/d/1btIwt4HwjyKMOUk-EMy3rbkfWzFxv2lNrMm_zkd0pA4/viewform?edit_requested=true}{UIUC+ Summer Undergraduate Research in Software Engineering}  offers an unpaid remote internship for software engineering students all over the world to collaborate with mentors from University of Illinois at Urbana Champaign.
+
+% \paragraph{Online Conferences and Workshops}: Attending online conferences and workshops is another way to gain exposure to cutting-edge research and establish connections with experts in the field.
+Many conferences now provide virtual participation options, enabling students to access research talks, poster sessions, and panel discussions and sometimes access designated chat rooms or networking events where participants can engage with researchers, ask questions, and seek potential research collaborations.
+By creating profiles, joining relevant research groups, and participating in discussions, students can connect with established researchers, contribute to ongoing projects, and potentially collaborate on publications or research proposals as they provide access
 

### 2024-08-20 demystify.tex
updates

@@
 International students, especially those from less well-known universities and smaller countries, often wonder if they should apply to a CS PhD program in the US.
-Specifically, they are worried about the \textbf{difficulty of getting admitted} and the \textbf{cost of graduate study} in the US.
+Specifically, they are worried about (i) the \textbf{difficulty of getting admitted} and (ii) the \textbf{cost of graduate study} in the US.
 This section addresses these concerns.
@@
 health insurance, and stipend (i.e., you get paid for your study). Moreover, depending on the university,
-you may even receive additional benefits such as summer pay, laptops, (conference/workshop) traveling.
+you may even receive additional benefits such as summer pay, laptops, and traveling to conferences and workshops.
 
@@
 
-After you submit your PhD application (e.g., in late Fall of year $X$ for the Fall admission of year $X+1$), it will first be ``screened''
+After you submit your PhD application (e.g., in late Fall of year $X$ for the Fall admission of year $X+1$), it will first be checked
 for general requirements, e.g., did you submit your transcripts and standard scores?
@@
 Usually, this screening process is done through a central university system, i.e., not by CS faculty.
+
 If you pass this, your application will be forwarded to the CS department for further evaluation.
@@
 
-\paragraph{Adcom} After screening applications are reviewed by a \emph{PhD admission committee} (\textbf{adcom}) that consists of faculty members in CS\footnote{In some cases the committee can involve affiliated faculty from different disciplines.}.
+\paragraph{Adcom} After screening applications are reviewed by a PhD admission committee or \emph{adcom} that consists of faculty members in CS\footnote{In some cases the committee can involve affiliated faculty from different disciplines.}.
 These adcom members have a wide range of expertise and background to ensure diverse perspectives in the evaluation process.
@@
 PhD students typically do not have to pay tuition as it is covered by their funding, but this might be important for MS students.
+
   \item [Lab] Research Lab.
@@
   \item[LoR] Letter of Recommendation. A letter written by a professor or supervisor that assesses your qualifications and potential for graduate study (\autoref{sec:lor}).
+
+  \item[Major] A student's primary field of study, e.g., Computer Science, Electrical Engineering, Mathematics, etc.
+Often used in the context of undergraduate studies (e.g., ``what was your major in undergrad?'').
+
   \item[MS] Master of Science. A graduate degree that typically requires 2 years of study.
@@
 
-  \item[RA] Research Assistantship. A form of funding where you work as a research assistant for a professor.
-  \item[TA] Teaching Assistantship. A form of funding where you work as a teaching assistant for a professor.
-  \item[TOEFL] Test of English as a Foreign Language. A standardized test to measure the English language ability of non-native speakers.
-  \item[IELTS] International English Language Testing System. A standardized test to measure the English language ability of non-native speakers.
+  \item[RA and TA] Research Assistantship (RA) is form of funding where you work on a research project for a professor.
+Teaching Assistantship (TA) is a form of funding where you help a a professor with teaching (e.g., grading assignments).
+
+  \item[TOEFL/IELTS] Test of English as a Foreign Language. A standardized test to measure the English language ability of non-native speakers.
   \item[GRE] Graduate Record Examination.
@@
 There are other agencies that support RAs but they might require US citizenship or permanent residency.
-        \item [R1 vs.
-R2 Universities] Research 1 (R1) universities are universities with the highest level of research activity.
+        \item [R1/R2 Universities] Research 1 (R1) universities are universities with the highest level of research activity.
 Research 2 (R2) universities have a high level of research activity but not as high as R1 universities.

### 2024-08-20 demystify.tex
updates

@@
 PhD students typically do not have to pay tuition as it is covered by their funding, but this might be important for MS students.
+  \item [Ivy League] A group of eight private universities in the US known for their academic excellence and social prestige.
+The Ivy League schools are Brown, Columbia, Cornell, Dartmouth, Harvard, the UPenn, Princeton, and Yale.
+Note that these schools are typically known for their undergraduate programs more than their graduate programs.
+Moverover, most top CS programs such as CMU, UIUC, UCSD, UWash, MIT, Stanford, are not in the Ivy League.
 

### 2024-08-24 demystify.tex
updates

@@
 \section{How long to complete the CS PhD program?}\label{sec:time}
-\sectioninfo{About 5--7 years in the US}
+\sectioninfo{About 5--7 years in the US.}
 
@@
 \section{What's the difference between PhD and MS in CS?}\label{sec:phd-vs-ms}
+\sectioninfo{MS focuses on coursework and prepares you for industry, while PhD focuses on research and prepares you for academia or research labs.}
 
@@
 Many CS programs also provide budget or computer equipment to incoming PhD students (e.g., a laptop).
+
+However, note that the prof.
+might not have the funding to buy you things.
+More importantly, they are not obligated to do so.
+It might also a need-based decision, e.g., if you need a powerful computer for your research, then the prof.
+might buy it for you (the lab might already has powerful servers and you can just use them for your work).
+But if you just want a new computer because others have it, then the prof.
+might not buy it for you.
 

### 2024-08-24 demystify.tex
updates

@@
 \section{CV/Resume}
-\sectioninfo{Highlight and summarize major achievements, e.g., Publications, Programming Competition Awards, and Teaching Experience.}
+\sectioninfo{Highlight and summarize major achievements, e.g., Publications, Programming Awards}
 
@@
 
-\chapterinfo{Celebrate! Now it is your turn to evaluate the school!
+\chapterinfo{Congrats! Now it is your turn to evaluate the school!
 Attend \emph{Open House} to learn more about the place and \emph{interview} profs---they would be much more willing to talk to you now.}
@@
 
+\section{Checking Status, Accepting, Postponing, and Decline Offers}\label{sec:accept-postpone-decline}
+
+Students often ask about what to do after they get an interview or an offer from a professor, e.g., if they can followup to find out about their status, or is it OK to postpone or accept/reject offers?, and most importantly, how to do so without offending anyone.
+
+\paragraph{Checking your application status and Follow-up emails} If you have interviewed and not heard back from a professor after a few weeks or especially around the time when universities send out their admission decisions (around late Feb-- mid-Mar), you can email to check.
+You can follow up the interview invitation and say: \emph{``Thanks for chatting with me.
+I am very excited about the opportunity to work with you.
+Could you please let me know if you have made a decision or if you need more information from me?''}.
+If you have new updates, e.g., new publications or new fellowship awards, or even new offers from other professors or schools, you can also mention that.
+
+Profs.
+are often very busy, especially during admission time when they have to many reviews and interviews.
+They might not have time to respond to every email.
+If you do not hear back after a week, you can send another email to check again.
+If you still do not hear back, you can assume that you are not selected.
+
+\paragraph{Accepting an offer} If you decide to accept an offer, you can say: \emph{``Thank you for the offer.
+I would like to accept it and look forward to working with you.
+Could you please send me more details about the offer and what to do next?''}.
+The prof.
+will likely send you more details about the offer and what to do next.
+If you decide to accept an offer, do so quickly.
+
+\paragraph{Postponing an offer} If you need more time to decide, you can ask for more time: \emph{``Thank you for the offer.
+I am very excited about it.
+However, I am still waiting for other offers and need more time to decide.
+Would it be possible to postpone the decision for a few weeks?''}.
+This is perfectly fine and professors will understand and might even appreciate your honesty.
+They will likely give you a few weeks to decide.
+If you need more time, you can ask for more time.
+But do not ask for too much time, e.g., more than a month.
+You also should not postpone the offer multiple times, which will annoy people.
+
+\paragraph{Declining an offer} If you decide to decline or reject an offer, you can say: \emph{``Thank you for the offer.
+However, I have decided to accept another offer.
+I appreciate your time and consideration.
+I hope we can work together in the future.''}  Professors will understand and wish you luck.
+If you decide to reject an offer, do so quickly.
+
+\paragraph{Accepting an offer and later rejecting it}
+
+I've seen many students, especially international, face a dilemma when they \emph{commit} to a graduate offer but then receive a better one.
+Advice given in online forums is often along the line that it's okay to switch, using reasons like you haven't yet had a strong relationship with the prof.
+or you should prioritize your personal benefit.
+
+In my opinion, these reasons are not strong enough to justify retracting an acceptance.
+A more valid reason is using the \href{https://cgsnet.org/wp-content/uploads/2024/01/CGS_April15_Resolution_Jan312024.pdf}{April 15 resolution}, in which many universities participate.
+Among various things, this resolution states that students are free to accept a new offer from a different institution until 4/15, even if they have already accepted an offer elsewhere.
+
+However, in general, retracting an acceptance can have ethical implications.
+When you accept an offer, you are committing to work with that prof, who then might stop looking for other students.
+So by retracting your offer, you are breaking your commitment and also causing a great deal of inconvenience to the prof and also taking away the opportunity from other students.
+Ultimately, this choice is personal, involving a balance between personal benefit and the ethical impact on others.
+
 \section{If you do not get admitted}\label{sec:not-accepted}
@@
 You can consider applying to MS programs, which are typically easier to get in; but you likely need to pay.  Sometimes, the university that rejected you for PhD might accept you for MS.
-If you get into an MS program at a school of your choice, you can contact professors to work with them. If you do well, you can ask the professor to support you to convert to PhD.
-
-\paragraph{Should I ask for feedback?}
+If you get into an MS program at a school of your choice, you can contact professors to work with them.
+If you do well, you can ask the professor to support you to convert to PhD.
+You will likely need to apply again, but you will have a much better chance because you have the direct support of a professor there.
+
+\subsection{Why did you get rejected?}\label{sec:why-rejected}
+
+\sectioninfo{You aim too high, are overqualified, or even because you applied to AI/ML, a super competitive field in recent years with many applicants.}
+
+Many students lament that they get no interviews or are rejected (with strong profiles) and that the admission process seems random.
+However, while it is true that the process is not perfect, it is not random.
+Other than not having a strong profile (e.g., research potential, GPA, LoRs, SOP, poor interviews), here are some other reasons for rejections.
+
+\paragraph{Not a Good fit and Bad luck}  You could have an excellent profile, but if you are interested in a research area that the program does not have, you will not be admitted.
+Similarly, if no faculty is willing to advise you (e.g., they are on sabbatical or personal leave, do not have sufficient funding, or already have too many students), you will not be admitted.
+This is actually good for you, as you don't want to be in a program where no one can advise you.
+
+Related to this is that you just have bad luck and apply at the wrong time.
+For example, since 2024 there has been a huge surge in students interested in AI and NLP (thanks, ChatGPT!).
+Consequently, AI/NLP faculty might be overwhelmed and cannot consider many candidates, even those with excellent profiles.
+
+\paragraph{You aim too high} This is a very common reason: you have applied to schools that are too high for your profile.
+While being ambitious is good, you should understand how PhD admission works (e.g., by reading this document and realizing things such as having a good GPA doesn't mean much to top programs) and be realistic!
+
+Before applying, you should talk to your professors and ask them to give you an honest opinion on where you should apply.
+To increase their chances, many students apply to a range of schools, including ``safety'' ones.
+
+\paragraph{Overqualified or Lack Interest}  This might be paradoxical and is the opposite of aiming too high.
+However if adcom believes that you are likely to get admitted and go to a better program, they might not be so enthusiastic to admit you and want to save the spot for someone else.
+
+A related reason is that you did not show enough interest in the program.
+For example, your SOP did not mention any faculty or research area at the program, you did not respond to professors interested in interviewing you (which might be considered unprofessional and burn bridges), or during the interview you showed little knowledge, interest, or enthusiasm (because they are your safety schools) in the program or profs.
+
+\paragraph{Low English exam scores (e.g., IELTS or TOEFL)}  Profs.
+might not care much about these, but the college or the university often sets a minimum that you need to pass to be considered, especially for TA funding (e.g., you need to communicate well with students as a TA).
+Thus, while profs.
+are willing to argue for your case, they might be reluctant to go against the requirement of the university on standard exams.
+
+Note that English exams are not required for all students, e.g., domestic students and those from certain countries are waived from taking these exams.
+For example, Mason waives English exams for students from these \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-Mason#standard-tests-waiver-eligible-countries}{countries}.
+
+\paragraph{Red flags} Various types of red flags can cause concern to the adcom (e.g., too many STEM courses with low or withdrawn grades, plagiarism, cheating, or other academic dishonesty).
+Another one is that you have a history of jumping from one program to another (e.g., you have been in multiple different programs in a short time).
+Adcom members might have contacts in other programs and heard about your stories, or your LoRs might mention these.
+
+If you think you have these issues, you should address them in your SOP or ask your letter writers to do so.
+In general, these things are rare, but they do happen and cause concern to the adcom.
+
+\subsection{Timeline: Why are rejection letters sent so late?}\label{sec:late-rejection}
+\sectioninfo{Why rejection letters are sent so late? Grad programs often wait for the accepted students to make their decisions, typically by April 15, before sending out rejection letters.}
+
+%"Just reject me already!" is a common sentiment among applicants.
+Indeed, school will first send out admission offers to the top candidates.
+They do not send out rejection letters because there is still a chance that some of the top candidates will decline the offer.
+If they do, then the school will go to the next set of candidates and send out offers to them.
+This process continues until all spots are filled.
+This is why you might not hear back until late in the admission cycle.
+
+\paragraph{Acceptance Letters} Universities prioritize sending out acceptance letters first. This allows the admitted students to make decisions and plan for their studies.
+
+Some universities have rolling admissions, meaning they send out acceptance letters as they review applications (e.g., some universities sent out letters as soon as early February).
+Others have a specific date when they send out the first round of acceptance letters.
+
+\paragraph{Response Deadlines} Accepted students are usually given a deadline to make decisions on their offers, often around \textbf{April 15th} in the US.
+This is a standard deadline for PhD programs due to the Council of Graduate Schools (CGS) Resolution (\href{https://cgsnet.org/wp-content/uploads/2024/01/CGS_April15_Resolution_Jan312024.pdf}{see more}.
+After this date, CS programs can gauge how many slots remain unfilled.
+
+\paragraph{Waitlist} Most CS programs have a limited number of slots for PhD students, and thus put many good students on a waitlist.
+If accepted students decline the offer, then offers are sent to students on the waitlist.
+So if you see people getting accepted, that does not mean you are out yet.
+
+\paragraph{Rejection Letters} Schools typically start sending out rejection letters to remaining applicants \emph{after they have finalized their admissions decisions}.
+Thus, rejection letters are often sent out late (e.g., after April 15th or even much later).
+Not much you can do here, you can try to contact the school to ask about your status, but they might not reply, they might say they are still reviewing applications, or give you inaccurate information (e.g., you will hear in two weeks).
+In short, you just have to be patient and wait (however, beware that some schools do not send out rejection letters at all).
+
+\subsubsection{Should I ask for feedback?}
 No, don't bother.
 You will likely not get any useful feedback.
-We are not willing (somtimes allowed) to reveal your evaluation results or give you feedback on how to improve your profile.
+We are not willing (sometimes not allowed) to reveal your evaluation results or give you feedback on how to improve your profile.
 \emph{So just move on}.
@@
 \chapterinfo{TAs, RAs, and fellowships are main funding sources for PhDs.
-TAs are provided by the department to help with classes (e.g., grading).
+TAs are provided by the department to help with classes.
 RAs are given by profs.
@@
 
-\chapterinfo{Not every professor, even those in CS, can advise or graduate CS PhD students.}
+\chapterinfo{Not every university has PhD progarms in CS. Not every professor, even those in CS, can advise or graduate CS PhD students.}
 
@@
 \section{Choosing a University}\label{sec:choosing-university}
+\sectioninfo{Select schools based on their CS program and faculty research interests.}
 
@@
 \section{Choosing an Adviser}
+\sectioninfo{The best adviser is the one that you can work well with. But you do not know that until you start working with them.}
+
 There is no one-size-fits-all answer to this question. The best adviser is the one that you can work well with.  But how do you find such a person?
@@
 \epigraph{``I want to share something with you – the three little sentences that will get you through life; number 1: Cover for me, number 2: Oh, good idea, Boss, and number 3: It was like that when I got here.''}{\textsc{The Simpsons}}
-
-\section{Why did you get rejected?}\label{sec:why-rejected}
-
-\sectioninfo{You aim too high, are overqualified, or even because you applied to AI/ML, a super competitive field in recent years with many applicants.}
-
-Many students lament that they get no interviews or are rejected (with strong profiles) and that the admission process seems random.
-However, while it is true that the process is not perfect, it is not random.
-Other than not having a strong profile (e.g., research potential, GPA, LoRs, SOP, poor interviews), here are some other reasons for rejections.
-
-\paragraph{Not a Good fit and Bad luck}  You could have an excellent profile, but if you are interested in a research area that the program does not have, you will not be admitted.
-Similarly, if no faculty is willing to advise you (e.g., they are on sabbatical or personal leave, do not have sufficient funding, or already have too many students), you will not be admitted.
-This is actually good for you, as you don't want to be in a program where no one can advise you.
-
-Related to this is that you just have bad luck and apply at the wrong time.
-For example, since 2024 there has been a huge surge in students interested in AI and NLP (thanks, ChatGPT!).
-Consequently, AI/NLP faculty might be overwhelmed and cannot consider many candidates, even those with excellent profiles.
-
-\paragraph{You aim too high} This is a very common reason: you have applied to schools that are too high for your profile.
-While being ambitious is good, you should understand how PhD admission works (e.g., by reading this document and realizing things such as having a good GPA doesn't mean much to top programs) and be realistic!
-
-Before applying, you should talk to your professors and ask them to give you an honest opinion on where you should apply.
-To increase their chances, many students apply to a range of schools, including ``safety'' ones.
-
-\paragraph{Overqualified or Lack Interest}  This might be paradoxical and is the opposite of aiming too high.
-However if adcom believes that you are likely to get admitted and go to a better program, they might not be so enthusiastic to admit you and want to save the spot for someone else.
-
-A related reason is that you did not show enough interest in the program.
-For example, your SOP did not mention any faculty or research area at the program, you did not respond to professors interested in interviewing you (which might be considered unprofessional and burn bridges), or during the interview you showed little knowledge, interest, or enthusiasm (because they are your safety schools) in the program or profs.
-
-\paragraph{Low English exam scores (e.g., IELTS or TOEFL)}  Profs.
-might not care much about these, but the college or the university often sets a minimum that you need to pass to be considered, especially for TA funding (e.g., you need to communicate well with students as a TA).
-Thus, while profs.
-are willing to argue for your case, they might be reluctant to go against the requirement of the university on standard exams.
-
-Note that English exams are not required for all students, e.g., domestic students and those from certain countries are waived from taking these exams.
-For example, Mason waives English exams for students from these \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-Mason#standard-tests-waiver-eligible-countries}{countries}.
-
-\paragraph{Red flags} Various types of red flags can cause concern to the adcom (e.g., too many STEM courses with low or withdrawn grades, plagiarism, cheating, or other academic dishonesty).
-Another one is that you have a history of jumping from one program to another (e.g., you have been in multiple different programs in a short time).
-Adcom members might have contacts in other programs and heard about your stories, or your LoRs might mention these.
-
-If you think you have these issues, you should address them in your SOP or ask your letter writers to do so.
-In general, these things are rare, but they do happen and cause concern to the adcom.
-
-\section{Timeline: Why are rejection letters sent so late?}\label{sec:late-rejection}
-\sectioninfo{Why rejection letters are sent so late? Grad programs often wait for the accepted students to make their decisions, typically by April 15, before sending out rejection letters.}
-
-%"Just reject me already!" is a common sentiment among applicants.
-Indeed, school will first send out admission offers to the top candidates.
-They do not send out rejection letters because there is still a chance that some of the top candidates will decline the offer.
-If they do, then the school will go to the next set of candidates and send out offers to them.
-This process continues until all spots are filled.
-This is why you might not hear back until late in the admission cycle.
-
-\paragraph{Acceptance Letters} Universities prioritize sending out acceptance letters first. This allows the admitted students to make decisions and plan for their studies.
-
-Some universities have rolling admissions, meaning they send out acceptance letters as they review applications (e.g., some universities sent out letters as soon as early February).
-Others have a specific date when they send out the first round of acceptance letters.
-
-\paragraph{Response Deadlines} Accepted students are usually given a deadline to make decisions on their offers, often around \textbf{April 15th} in the US.
-This is a standard deadline for PhD programs due to the Council of Graduate Schools (CGS) Resolution (\href{https://cgsnet.org/wp-content/uploads/2024/01/CGS_April15_Resolution_Jan312024.pdf}{see more}.
-After this date, CS programs can gauge how many slots remain unfilled.
-
-\paragraph{Waitlist} Most CS programs have a limited number of slots for PhD students, and thus put many good students on a waitlist.
-If accepted students decline the offer, then offers are sent to students on the waitlist.
-So if you see people getting accepted, that does not mean you are out yet.
-
-\paragraph{Rejection Letters} Schools typically start sending out rejection letters to remaining applicants \emph{after they have finalized their admissions decisions}.
-Thus, rejection letters are often sent out late (e.g., after April 15th or even much later).
-Not much you can do here, you can try to contact the school to ask about your status, but they might not reply, they might say they are still reviewing applications, or give you inaccurate information (e.g., you will hear in two weeks).
-In short, you just have to be patient and wait (however, beware that some schools do not send out rejection letters at all).
 
@@
 
-However, note that the prof.
-might not have the funding to buy you things.
-More importantly, they are not obligated to do so.
+However, note this is a privilege and a courtesy and not a right.
+This means that the prof.
+is not obligated to buy you things, and that you should not expect these are automatically provided.
+Also, while the prof.
+might be able to do so in the past, they might not be able to do so now due to budget constraints.
 It might also a need-based decision, e.g., if you need a powerful computer for your research, then the prof.
 might buy it for you (the lab might already has powerful servers and you can just use them for your work).
-But if you just want a new computer because others have it, then the prof.
-might not buy it for you.
-
-Keep in mind that these computers would be university property, which might be monitored and have certain restrictions, e.g., do not install illegal software on them (\autoref{sec:illegal-software}).
-You might also have to return them when you graduate.
+Finally, the prof.
+might have a policy on what they can buy for you, e.g., they prefer to get you desktop instead of a laptop (to encourage you to come to the office and work).
+
+Keep in mind that these computers and equipment would be university property, which might be monitored and have certain restrictions, e.g., do not install illegal software on them (\autoref{sec:illegal-software}).
+You will likely need to return them when you graduate.
 
 \section{Negotiating PhD offer (e.g., having multiple offers)?}\label{sec:negotiate}
+\sectioninfo{You will not be able to negotiate stipend, but you can ask for specific start date, TA assignment, and conference travel budget.}
 
@@
 
-\section{Checking Status, Accepting, Postponing, and Decline Offers}\label{sec:accept-postpone-decline}
-
-Students often ask about what to do after they get an interview or an offer from a professor, e.g., if they can followup to find out about their status, or is it OK to postpone or accept/reject offers?, and most importantly, how to do so without offending anyone.
-
-\paragraph{Checking your application status and Follow-up emails} If you have interviewed and not heard back from a professor after a few weeks or especially around the time when universities send out their admission decisions (around late Feb-- mid-Mar), you can email to check.
-You can follow up the interview invitation and say: \emph{``Thanks for chatting with me.
-I am very excited about the opportunity to work with you.
-Could you please let me know if you have made a decision or if you need more information from me?''}.
-If you have new updates, e.g., new publications or new fellowship awards, or even new offers from other professors or schools, you can also mention that.
-
-Profs.
-are often very busy, especially during admission time when they have to many reviews and interviews.
-They might not have time to respond to every email.
-If you do not hear back after a week, you can send another email to check again.
-If you still do not hear back, you can assume that you are not selected.
-
-\paragraph{Accepting an offer} If you decide to accept an offer, you can say: \emph{``Thank you for the offer.
-I would like to accept it and look forward to working with you.
-Could you please send me more details about the offer and what to do next?''}.
-The prof.
-will likely send you more details about the offer and what to do next.
-If you decide to accept an offer, do so quickly.
-
-\paragraph{Postponing an offer} If you need more time to decide, you can ask for more time: \emph{``Thank you for the offer.
-I am very excited about it.
-However, I am still waiting for other offers and need more time to decide.
-Would it be possible to postpone the decision for a few weeks?''}.
-This is perfectly fine and professors will understand and might even appreciate your honesty.
-They will likely give you a few weeks to decide.
-If you need more time, you can ask for more time.
-But do not ask for too much time, e.g., more than a month.
-You also should not postpone the offer multiple times, which will annoy people.
-
-\paragraph{Declining an offer} If you decide to decline or reject an offer, you can say: \emph{``Thank you for the offer.
-However, I have decided to accept another offer.
-I appreciate your time and consideration.
-I hope we can work together in the future.''}  Professors will understand and wish you luck.
-If you decide to reject an offer, do so quickly.
-
-\paragraph{Accepting an offer and later rejecting it}
-
-I've seen many students, especially international, face a dilemma when they \emph{commit} to a graduate offer but then receive a better one.
-Advice given in online forums is often along the line that it's okay to switch, using reasons like you haven't yet had a strong relationship with the prof.
-or you should prioritize your personal benefit.
-
-In my opinion, these reasons are not strong enough to justify retracting an acceptance.
-A more valid reason is using the \href{https://cgsnet.org/wp-content/uploads/2024/01/CGS_April15_Resolution_Jan312024.pdf}{April 15 resolution}, in which many universities participate.
-Among various things, this resolution states that students are free to accept a new offer from a different institution until 4/15, even if they have already accepted an offer elsewhere.
-
-However, in general, retracting an acceptance can have ethical implications.
-When you accept an offer, you are committing to work with that prof, who then might stop looking for other students.
-So by retracting your offer, you are breaking your commitment and also causing a great deal of inconvenience to the prof and also taking away the opportunity from other students.
-Ultimately, this choice is personal, involving a balance between personal benefit and the ethical impact on others.
-
 \section{Academic Integrity (Cheating and Plagiarism)}
@@
 
-\chapter{Rankings of CS PhD programs}\label{sec:ranking}
+\chapter{CSRankings: Rankings of CS PhD programs}\label{sec:ranking}
+\chapterinfo{CSRankings.org is a ranking system based on faculty publications at top CS conferences.}
 

### 2024-08-26 demystify.tex
updates

--- 2024-08-20 demystify.tex
@@
     \draw[ultra thick,gray]
-    ($(current page.center)+(5,2)$) -- ++(0,-3cm) 
+    ($(current page.center)+(5,3.5)$) -- ++(0,-3cm)
     node[
@@
     % Title
-    \node[align=center] at ($(current page.center)+(0,-5)$) 
+    \node[align=center] at ($(current page.center)+(0,-7)$)
     {
@@
     {\fontsize{24}{24} \selectfont {{PhD Admission in the US}}} \\[0.2in]    
-    {\fontsize{18}{18} \selectfont {{A Handbook for International Students}}} \\[0.5in]
+    {\fontsize{18}{18} \selectfont {{A Handbook for International Students}}} \\[1.in]
 

### 2024-08-26 faculty.tex
updates

@@
 P\&T evaluation requires no specific conferences/journals or numbers of papers, but typically you will want to have a good number of papers in top venues.
-%P\&T members or letter writers (obviously) are familiar with top places in your field, and they will definitely look for those in your profile.
 Work that won awards, e.g., best paper awards, can help a lot in P\&T.
@@
 Papers with your students are likely more preferred than papers with your peers (and their students) because it shows that you are mentoring your students well and that the project is lead by your group.
-
-\paragraph{Artifact Evaluation} This is probably more well-known in SE conferences, and is intended to promote the reproducibility of research results.
-However, I never liked it because it's a lot of work and many of my experiments require heavy computation that the reviewers do not have.
-I also don't think it's very useful as I have always make my code, benchmarks, and results open source (through Github).
-I found it strange that many SE conference have recently required this as part of the submission process.
-It makes the submission process much more complicated and much more time-consuming.
 
@@
 Sometimes a full paper is not as good because the research is relatively complete with strong results, which bring the concern that you already did all the cool work and the proposal is minor and not exciting.
+I have used results from short papers in many of my grants, e.g., NIER (New Idea and Emerging Results) papers for my NSF CAREER and Tool Demo and SEIP (SE in Practice) paper for my Facebook gift.
+
+\paragraph{Artifact Evaluation (AE)} AE is is probably more well-known in SE conferences, but it allows authors of accepted papers to submit their tools to be evaluated by reviewers.
+The intention is to show that your research is reproducible.
+However, I never liked it because it's a lot of work and many of my experiments require heavy computation that the reviewers do not have.
+In many cases reviewers are students and have no technical expertise how to run things (even though it is already in Docker) or insists on using their own environment (e.g., Mac) while the developed tool is explicitly for Linux.
+Finally, I also don't think it's very useful as I have always make my code, benchmarks, and results open source (through Github).
+I found it strange that many SE conference have recently required this as part of the submission process.
+It makes the submission process much more complicated and time-consuming.
 

### 2024-08-27 demystify.tex
updates

--- 2024-08-24 demystify.tex
@@
 
-  \item[SOP] Statement of Purpose. A document written by yourself to explains your research interests, background, and reasons for applying to a PhD program (\autoref{sec:sop}).
+  \item[Q1 journal] Commonly mentioned by international students to refer to top-tier journals in CS.
+This is not a standard in CS in the US (which often focuses on confs.
+rather than journals), and many CS faculty in the US might not know what this means.
+If you want to refer to top-tier journals, you should mention the specific journals, e.g., TSE in Software Engineering.
+
+  \item[SOP] Statement of Purpose is a document written by yourself to explains your research interests, background, and reasons for applying to a PhD program (\autoref{sec:sop}).
 
@@
 
-  \item [PI]  Principal Investigator.
-A lead researcher on a grant or research project, e.g., a PI on an NSF grant.
+  \item [PI]  Principal Investigator is lead researcher on a grant or research project, e.g., a PI on an NSF grant.
 Prospective students often use this term to refer to a professor they are interested in working with.
-While this term is used frequently in online forums among students, we do not use it this way in academia, e.g., ``who is your PI?'' might only make sense to fellow Reddit applicants, but might not be understood by others, ``who is your adviser?'' is what you want to use.
+However, we do not use it this way in academia, e.g., ``who is your PI?'' might only make sense to fellow Reddit applicants, but might not be understood by others, ``who is your adviser?'' is what you want to use.
     \item[STEM] Science, Technology, Engineering, and Mathematics. A term used to group together these academic disciplines. CS is considered a STEM field.
@@
 Research 2 (R2) universities have a high level of research activity but not as high as R1 universities.
-R1 universities are typically more prestigious and have more funding and resources for research.
+R1 universities are typically larger, and have more funding and resources for research.
 R2 universities are also good but might have fewer resources and funding.
-  \item [Rolling Admission] A process where applications are reviewed as they are received, and decisions are made throughout the admission cycle.
-This is in contrast to a process where all applications are reviewed after the deadline and decisions are made at the same time.
+
+        \item [Rolling Admission] Applications are reviewed as they are received (instead of all at once after the deadline), and decisions are made throughout the admission cycle.
 Rolling admission is more common for MS programs and less common for PhD programs (e.g., at Mason, MS is rolling and PhD is not).
+
   \item [Top-tier conferences and journals] These are the most prestigious venues for publishing research in CS.
@@
   \section{Miscs}\label{sec:cultural-misc}
-  Several other common surprises for international students (and foreigners in general) in the US:
+
+  Several other common surprises for international students (and foreigners in general) in the US.
+Note that I skip topics involving politics, religion, tax, and racism as these happen in many countries and are not unique to the US.
+
+  \paragraph{Small talks} People often engage in small talks, e.g., about the weather, sports, or weekend plans.  This is a way to start a conversation and how social interaction starts in the US.
+
+  Moreover, avoid asking personal questions, e.g., about salary, age, relationship status, or health, as these are considered private (talking about kids' activities or schools are OK).
+Also, do not talk about politics or religion in general (in fact, we often do not talk about these with our family and friends to avoid conflicts).
+Sometimes foreigners are surprised by how Americans do not talk about their personal lives, e.g., sharing details about their families, health, or relationships, and that their conversations are often not very ``deep'' or ``mind-provoking''.
+This is just a cultural norm about privacy and personal space.
 
@@
 However, as mentioned in \autoref{sec:funding}, your  TA/RA  will cover health insurance (and your spouse/children also get insurance or discount for it).
+
   However, be aware that heathcare services can still be expensive even with insurance.
 So  know what your insurance covers and be prepared for unexpected costs.
-Moreover, healthcare system has many confusing jargons such as HMO, PPO, deductibles, copays, and coinsurance.
-Get yourself familiar with them and ask for help when needed.
+Moreover, healthcare system has many confusing jargons such as HMO, PPO, deductibles, copays, and coinsurance (try take a look at the Explanation of Benefits or EoB statement you received from your insurance company).
+It's arguably the most complicated system in the US and even Americans often do not understand it (and politicians often exploit this to their advantage).
+So do not hesitate to ask your HR or the insurance company for help.
 
@@
 The minimum wage for tipped employees is lower than the standard minimum wage, so tips are an important part of their income.
-If you do not tip, it is considered rude and disrespectful.
 
@@
 	%\paragraph{Individualism} The strong culture of individualism in the US can be both liberating and isolating.
-People are more focused on their personal rights and tend to be less concerned with what others think.
+For example, Americans are encouraged to express their opinions and pursue their goals, and are more focused on their personal rights (e.g., they do not want people do things in front of their house).
 

### 2024-08-27 faculty.tex
updates

--- 2024-08-18 faculty.tex
@@
 Many good researchers have left such places because of this.
-However, some were able to leverage this and publish at easy places and do really well within their department.
+Some were able to leverage this and publish at easy places and do really well within their department.
 However, while they thrive at such places, they will likely not get funding and recognition in their field.
@@
 
-\chapter{Teaching}
-
-\section{From a P\&T Point of View}
-
-While teaching should take about 40\% of your time (equivalent to research) and your salary is mainly for teaching, teaching is not as nearly as important as research for tenure.
-Moreover, for most new faculty, teaching is not as difficult as getting grants or publishing papers.
-Thus, tenure-track faculty often do not worry much about teaching.
+\chapter{Teaching}\label{chapter}
+
+While teaching should take about 40\% of your time (equivalent to research) and your salary is mainly for teaching, it is not as nearly as important as research for tenure.
+Moreover, for most new faculty, teaching is not as difficult as getting grants or publishing papers, and thus is often taken lightly by new faculty.
+However, teaching is still important for your tenure case, and being a good teacher can be very rewarding.
+Moreover, innovative integration of research and teaching can help you win good students and even NSF grants.
 
 \section{Teaching Load}
+New TT faculty often have a lighter teaching load in their first several years or even throughout their tenure track.
+This is to give you time to establish your research program.
+However, you should expect to teach at least one course per semester.
+The teaching load is usually higher at teaching universities, and lower at research universities.
+At some universities, you might be able to buy out of teaching by getting a grant.
+
 \section{Teaching Evaluation}

### 2024-08-31 faculty.tex
updates

--- 2024-08-26 faculty.tex
@@
 In some fields, journals are more common and prestigious.
-Moreover, many areas have top journals in additions to well-known conferences (e.g., in addition to the main fource confs as mentioned above, SE also has Trans.
+Moreover, many areas have top journals in additions to well-known conferences (e.g., in addition to the main four confs as mentioned above, SE also has Trans.
 of Software Engr.
@@
 However, this does not mean you should not work with your adviser(s).
-You can absolutely continue do research and write grant proposals with them, but you should also work and write papers with other people, and especially your students.
+You can continue doing research and write grant proposals with them, but you should also work and write papers with other people, and especially your students.
 
@@
 
-\chapter{Obtaining Funding}\label{chap}
+\chapter{Funding and Grant Proposals}\label{chap}
 Without a doubt, grants are one of the most, if not \emph{the} most, important factors in P\&T (or any kind of performance) evaluation.
@@
 
-\section{From a P\&T Point of View}
-
 While teaching should take about 40\% of your time (equivalent to research) and your salary is mainly for teaching, teaching is not as nearly as important as research for tenure.
@@
 At the end of each semester, students are asked to evaluate the instructor.
-The results are shared anonymously with the instructor to help them improve their teaching and course.
+The results are shared anonymously with the instructor to help them improve their teaching.
 Teaching evaluation is also commonly used in annual and tenure evaluation.
@@
 There will be times when you feel discouraged and frustrated, but there will also be times when you feel proud and happy.  
-For P\&T you will be expected to have successes in mentoring students.  For most good universities, eou will likely need to graduate at least one PhD student to get tenure.
+For P\&T you will need to successfully mentoring students.  Many universities expect you to need to graduate at least one PhD student to get tenure.
 
@@
 
-Officially services are counted for only 10\% of your work. However, they are very important for your tenure and career.
+Officially services are counted for only 10\% of your work. However, they are very important for your P\&T and career in general.
 %You are expected to have certain types of services as a faculty member.  
@@
 \paragraph{Personal Experience} I mostly cold email people for opportunities.
-For example, I directly email OOPSLA and PLDI PC chairs and asked, and they invited.
-I tried this with ICSE (and also nominate myself through the official form), and it never works, even though I have many papers at ICSE (that leaves a bad taste in my mouth about ICSE, but that is another story).
-I also got invited to serve on a PC because a senior colleague recommended me.
+For example, I directly email OOPSLA and PLDI PC chairs and asked, and they invited me.
+I tried this with ICSE (and also nominate myself through the official form), and it never works, even though I have many papers at ICSE.
 For NSF, I just emailed program managers and asked to serve on their panels, and they invited me.
@@
 
-I will skip the details and just summarize several things that I didn't like.
+I will skip the details and just summarize several important and disappointing events.
 

### 2024-09-01 demystify.tex

@@
 
-You can consider applying to MS programs, which are typically easier to get in; but you likely need to pay.  Sometimes, the university that rejected you for PhD might accept you for MS.
+You can consider applying to MS programs, which are typically easier to get in; but you likely need to pay (\autoref{sec:ms-admission}).
+A university that rejected you for PhD might accept you for MS.
 If you get into an MS program at a school of your choice, you can contact professors to work with them.
@@
 
-\section{What's the difference between PhD and MS in CS?}\label{sec:phd-vs-ms}
-\sectioninfo{MS focuses on coursework and prepares you for industry, while PhD focuses on research and prepares you for academia or research labs.}
-
-While both are graduate degrees, they are different in terms of objective, admission requirements, course requirements, duration, and funding as summarized in \autoref{tab:phd-vs-ms}.
-
-\begin{table}
-  \caption{PhD vs. MS}\label{tab:phd-vs-ms}
-  \begin{tabular}{c|c|c}
-  \toprule
-  \textbf{} & \textbf{PhD} & \textbf{MS} \\
-  \midrule
-  Objective & Industry & Research \\
-  Admission Req & Research experience & No research experience \\
-  Coursework Req & Yes & Yes (research is much more important)  \\
-  Duration & 2 years & 5--7 years \\
-  Adviser Req & No & Yes \\
-  Funding & No & Yes \\
-  \bottomrule
-  \end{tabular}
-\end{table}
-
-\begin{itemize}
-
-  \item Objective: an MS is typically to prepare you for industry, while a PhD is to prepare you for research.
-Some MS has thesis option but the research is much lighter compared to a PhD, which is research-focused.
-
-  \item Admission requirements: MS also requires a good GPA, LoRs, SOP, and test scores, but research experience is not required.
-PhD requires all of these, but research experience is a must.
-Moreover, MS programs are typically much less competitive than PhD programs.
-In fact, many MS program are viewed as a revenue source (\emph{cash cow}) for the department, as they are not funded and students pay tuition.
-
-  \item Course requirements: MS has a specific number of courses (typically can be done in 2 years).
-You graduate with an MS when you're done with the courses.
-PhD also has coursework requirements (typically taken in your first 2 years) but focuses on research after that.
-You graduate with a PhD when you have done enough research and written a dissertation (which is typically way harder and takes much longer time than coursework).
-
-  \item Duration: an MS typically takes 2 years while a PhD takes 5--7 years (or even longer).  Many students get an MS along the way to a PhD, e.g., after finishing the 2-year course work.
-
-  \item Adviser: MS students typically do not have an adviser (if you do thesis option then you will have one), while PhD students need an adviser who guides them in their research.
-
-  \item Funding: MS is typically \emph{not funded}, while PhD is typically funded (e.g., through TA or RA as explained in \autoref{sec:funding}).
-
-\end{itemize}
-
 \section{PhD in the US vs. Other Countries}\label{sec:non-us-differences}
@@
 % \end{domesticbox}
+
+\chapter{MS Admission}
+\chapterinfo{MS focuses on coursework and prepares you for industry, while PhD focuses on research and prepares you for academia or research labs.
+This section provides guidance on applying to MS programs.}
+
+While both MS and PhD programs are graduate degrees, they are \emph{very} different in terms of objective, admission requirements, course requirements, duration, and funding.
+This section provides guidance on applying to MS programs in CS.
+
+\section{Differences between PhD and MS}
+
+\begin{table}
+    \caption{PhD vs. MS}\label{tab:phd-vs-ms}
+    \begin{tabular}{c|c|c}
+    \toprule
+    \textbf{} & \textbf{PhD} & \textbf{MS} \\
+    \midrule
+    Objective & Industry & Research \\
+    Admission Req & Research experience & No research experience \\
+    Coursework Req & Yes & Yes (research is much more important)  \\
+    Duration & 2 years & 5--7 years \\
+    Adviser Req & No & Yes \\
+    Funding & No & Yes \\
+    \bottomrule
+    \end{tabular}
+  \end{table}
+
+\autoref{tab:phd-vs-ms} summarizes the main differences between PhD and MS programs:
+\begin{itemize}
+
+  \item Objective: an MS is typically to prepare you for industry, while a PhD is to prepare you for research. Some MS has thesis option but the research is much lighter compared to a PhD.
+
+  \item Admission requirements (\autoref{sec:ms-admission}): MS also requires a good GPA, LoRs, SOP, and test scores, but research experience is not required.
+PhD requires all of these, but research experience is a must.
+
+  \item Course requirements: MS has a specific number of courses (typically can be done in 2 years).
+You graduate with an MS when you're done with the courses.
+PhD also has coursework requirements (typically taken in your first 2 years) but focuses on research after that.
+You graduate with a PhD when you have done enough research and written a dissertation (which is typically way harder and takes much longer time than coursework).
+
+  \item Duration: an MS typically takes 2 years while a PhD takes 5--7 years (or even longer).  Many students get an MS along the way to a PhD, e.g., after finishing the 2-year course work.
+
+  \item Adviser: MS students typically do not have an adviser (if you do thesis option then you will have one), while PhD students need an adviser who guides them in their research.
+
+  \item Funding: MS is typically \emph{not funded}, while PhD is typically funded (\autoref{sec:funding}).
+
+\end{itemize}
+
+\section{MS Admission}\label{sec:ms-admission}
+
+Except for a very few highly selective programs, MS programs are typically much less competitive than PhD programs.
+Some refer to MS programs as a \emph{cash cow} for the department, as MS students are often not funded and have to pay tuition (i.e., you're likely to get in if you can afford it).
+
+While admission requirements are similar to PhD programs (e.g., GPA, SOP, LORs), research is not a focus in MS programs.
+In most cases, the main requirements are that you have sufficient background in CS (e.g., through your undergrad degree).
+
+\subsection{Admission Committee}
+MS admission also involves an adcom that reviews applications and makes admission decisions.
+However, MS admission is typically rolling (i.e., they review applications as they come in and make decisions throughout the admission cycle).
+Moreover, unlike PhD that has multiple reviewers for each application, an MS application typically involves only one reviewer and does not review much time to review compared to a PhD application (\autoref{sec:ievaluate}).
+
+\begin{commentbox}[Vu:]
+    Mason has a very large the number of MS students in CS (1000+ MS vs 200 PhDs).
+In contrast, other similar size universities often have much smaller MS CS programs (or none).
+Location plays an important role as GMU is close to DC with many developer professionals who want to get an MS, which are often covered by their employers.
+Our MS program is geared towards working professionals, e.g., all of our MS courses are offered in the evening and online.
+
+    The CS department has three separate committees for MS admission: MS in CS (the traditional one, which is the largest), MS in Software Engineering (SWE), and MS in Information System. 
+    Each committee has its own chair and members. For example, I am often in the MS SWE adcom, which has about 3 adcom members (including the chair who also does the review). 
+\end{commentbox}
+
+%Third, adcom members can involve both teaching and tenure-line faculty (\autoref{sec:faculty-types}).
+
+\subsection{Application Materials}
+
+In an MS application, you typically need to submit the similar materials as in a PhD application, e.g., transcripts, LoRs, SOP, and test scores.
+However, research experience is not required, and LoRs can be from anyone who can speak about your academic or working abilities (e.g., professors, employers, etc).
+Moreover, your SOP you should focus on your academic and work (e.g., software development) background, why you want to get an MS, and how the program fits your goals.
+As with PhDs, GREs are often not required.
+Moreover, some MS programs do not even require LoRs or SOPs.
+
+\paragraph{Transcripts and Undergrad Background} Since research is not required, your undergrad background is more important in an MS application.
+You should have a strong background in CS, e.g., through your undergrad degree.
+If you do not have a CS degree, you might need to take some prerequisite courses before you can start the MS program.
+GPA is also important, as it is often used as a filter for MS applications.
+
+\paragraph{SOP} Your SOP should explain why you want to get an MS, how the program fits your goals, and why you're a good fit for the program.
+You don't need to mention about working with specific professors unless you want to do a thesis option.
+You still need to customize it for each school (e.g., you picked GMU due to its strong SWE program or the DC area has many job opportunities).
+
+\paragraph{LoRs} LoRs can be from professors, employers, or anyone who can speak about your academic or working abilities.
+Unlike PhDs, your LoRs do not need to be from professors and talk about research experience.
+Many schools do not even require LoRs for MS.
+
+\paragraph{Test Scores} GREs are often not required for MS applications.
+However, some schools might still require them, and you should check with the school you're applying to.
+If you have a low GPA, a high GRE score can help offset that.
 

### 2024-09-01 faculty.tex

@@
 
-\chapter{Teaching}
-
-While teaching should take about 40\% of your time (equivalent to research) and your salary is mainly for teaching, teaching is not as nearly as important as research for tenure.
-Moreover, for most new faculty, teaching is not as difficult as getting grants or publishing papers.
-Thus, tenure-track faculty often do not worry much about teaching.
+\chapter{Teaching}\label{chapter}
+
+While teaching should take about 40\% of your time (equivalent to research) and your salary is mainly for teaching, it is not as nearly as important as research for tenure.
+Moreover, for most new faculty, teaching is not as difficult as getting grants or publishing papers, and thus is often taken lightly by new faculty.
+However, teaching is still important for your tenure case, and being a good teacher can be very rewarding.
+Moreover, innovative integration of research and teaching can help you win good students and even NSF grants.
 
 \section{Teaching Load}
+New TT faculty often have a lighter teaching load in their first several years or even throughout their tenure track.
+This is to give you time to establish your research program.
+However, you should expect to teach at least one course per semester.
+The teaching load is usually higher at teaching universities, and lower at research universities.
+At some universities, you might be able to buy out of teaching by getting a grant.
+
 \section{Teaching Evaluation}

### 2024-09-01 demystify.tex
updates

@@
 \chapter{MS Admission}
-\chapterinfo{MS focuses on coursework and prepares you for industry, while PhD focuses on research and prepares you for academia or research labs.
-This section provides guidance on applying to MS programs.}
+\chapterinfo{MS focuses on coursework and prepares you for industry, while PhD focuses on research and prepares you for academia or research.}
 
@@
 
-In an MS application, you typically need to submit the similar materials as in a PhD application, e.g., transcripts, LoRs, SOP, and test scores.
-However, research experience is not required, and LoRs can be from anyone who can speak about your academic or working abilities (e.g., professors, employers, etc).
-Moreover, your SOP you should focus on your academic and work (e.g., software development) background, why you want to get an MS, and how the program fits your goals.
+You typically need to submit the similar materials as in a PhD application, e.g., transcripts, LoRs, SOP, and test scores.
+However, research experience is not required, and LoRs can be from anyone who can speak about your academic or working abilities.
+Moreover, your SOP you should focus on your academic and work background, why you want to get an MS, and how the program fits your goals.
 As with PhDs, GREs are often not required.
-Moreover, some MS programs do not even require LoRs or SOPs.
-
-\paragraph{Transcripts and Undergrad Background} Since research is not required, your undergrad background is more important in an MS application.
+Some MS programs do not even require LoRs or SOPs.
+
+\paragraph{Undergrad Background and GPA} Since research is not required, your undergrad background is more important in an MS application.
 You should have a strong background in CS, e.g., through your undergrad degree.
-If you do not have a CS degree, you might need to take some prerequisite courses before you can start the MS program.
 GPA is also important, as it is often used as a filter for MS applications.
@@
 
-\paragraph{LoRs} LoRs can be from professors, employers, or anyone who can speak about your academic or working abilities.
-Unlike PhDs, your LoRs do not need to be from professors and talk about research experience.
-Many schools do not even require LoRs for MS.
+\paragraph{LoRs} Unlike PhDs, your LoRs do not need to be from professors or talk about research experience. Many MS programs do not even require LoRs.
 
@@
 If you have a low GPA, a high GRE score can help offset that.
+As with PhD, English proficiency tests (TOEFL/IELTS) are required for international students.
+
+\subsection{Funding}
+Unlike PhD programs, which typically guarantee funding (\autoref{sec:funding}), MS programs are typically not funded.
+You have to pay tuition, which can be expensive.
+It is difficult for an MS student to get funding, e.g., through TA or RA positions, because these are typically reserved for PhD students.
+\begin{itemize}
+\item (RA) Profs.
+are not willing to take MS students as RAs because they are not around long enough to be productive.
+It can take a while for a student to get up to speed and start being productive, and by the time they are, they are about to graduate.
+Moreover, the goal of most MS students is to get a job, not to do research, so they are not as motivated to do research.
+\item (TA) MS students are typically not given TA positions because PhDs are given priority.
+Many depts.
+even have shortage of TA positions for PhD students and so they cannot afford to give them to MS students.
+\end{itemize}
+
+However, there are always exceptions, e.g., if you have a strong background and can demonstrate that you can be productive, then a prof.
+might be willing to take you as an RA.
+Moreover, some schools do have TAs for MS students, and you can apply for these positions.
+For example, Mason CS dept has TAs for MS students (of course not for all of our 1000+ MS students, but quite a few); my courses in the past two years have MS TAs.
 

### 2024-09-12 faculty.tex
updates

@@
 
-While teaching should take about 40\% of your time (equivalent to research) and your salary is mainly for teaching, it is not as nearly as important as research for tenure.
+While teaching takes about 40\% of your time, it is not as nearly as important as research for tenure.
 Moreover, for most new faculty, teaching is not as difficult as getting grants or publishing papers, and thus is often taken lightly by new faculty.
 However, teaching is still important for your tenure case, and being a good teacher can be very rewarding.
-Moreover, innovative integration of research and teaching can help you win good students and even NSF grants.
+Moreover, innovative integration of research and teaching can help you get good students and even NSF grants.
 
@@
 The teaching load is usually higher at teaching universities, and lower at research universities.
-At some universities, you might be able to buy out of teaching by getting a grant.
+
+The desirable teaching load for TT at most R1 universities is 1+1 (1 course / semester) during the tenure track.
+While some univs might give you a lighter load in your first year, most univs expect you to teach 1+1 from your second year.
+In many cases, 1+1 is only for the first 2 or 3 years and after that it increases to 1+2 (or 2+1).
+It is rare that a TT in CS at R1 universities would have to teach 2+2 during the tenure track.
+
+In many cases, your contract might list 1+1 for the first X year and 1+2 after until tenure, but the dept unofficial gives TT faculty 1+1 until tenure.
+This is because the dept chair can reduce the teaching load based on performance.
+There are also other ways to reduce teaching load such as \emph{buying out} of teaching with a grant, leading a very large grant, or have taught many large classes.
+
+TT faculty will teach both undergrad and grad courses.
+Grad courses include seminars, special topics, and core courses.
+Most new faculty use seminars and special topics to recruit students for their research group.
+Undergrad courses are typically higher level courses in your area of expertise (e.g., 400 level course on Software Testing).
+It is rare that you have to teach 100 level courses (e.g., Intro to Programming) because there are just too many students in these courses.
+
+\begin{commentbox}[Vu]
+
+  At UNL, TT faculty is 1+1 for the first 3 years, and 1+2 after that.
+At GMU, it is 1+1 for the first 3 years, but I haven't seen any TT faculty having to teach more than that during their tenure track.
+In addition, GMU grants all TT faculty a semester off from teaching before tenure.
+Some faculty use this to work on their NSF CAREER proposal while others use it to prepare their tenure application.
+
+\end{commentbox}
 
@@
 
+\begin{commentbox}[Vu]
 A strategy that I found effective is that the whole class gets a small bonus point if everyone completes the evaluation.
@@
 It's been working well for me: Spring '24 I had a 90\% response rate (25 out of 28 students), and Fall '23 was a perfect 100\% (16 out of 16 students).
-Before, my response rate was about 50\%, which is considered higher than average in the department.
+Before, my response rate was about 50\%, which is already higher than the average.
+\end{commentbox}
 

### 2024-09-12 faculty.tex
updates

@@
 
-\chapter{Teaching}\label{chapter}
+\chapter{Teaching}\label{chapter:teaching}
 
@@
 This is to give you time to establish your research program.
-However, you should expect to teach at least one course per semester.
 The teaching load is usually higher at teaching universities, and lower at research universities.
 
-The desirable teaching load for TT at most R1 universities is 1+1 (1 course / semester) during the tenure track.
+The desirable teaching load for TT at most R1 universities is 1+1 (1 course / semester).
 While some univs might give you a lighter load in your first year, most univs expect you to teach 1+1 from your second year.
@@
 In many cases, your contract might list 1+1 for the first X year and 1+2 after until tenure, but the dept unofficial gives TT faculty 1+1 until tenure.
-This is because the dept chair can reduce the teaching load based on performance.
-There are also other ways to reduce teaching load such as \emph{buying out} of teaching with a grant, leading a very large grant, or have taught many large classes.
-
-TT faculty will teach both undergrad and grad courses.
+Some other ways to reduce teaching load such as \emph{buying out} of teaching with a grant, leading a very large grant, or have taught many large classes.
+
+\begin{commentbox}[Vu]
+
+    At UNL, TT faculty is 1+1 for the first 3 years, and 1+2 after that.
+At GMU, it is 1+1 for the first 3 years, but I haven't seen any TT having to teach more than that.
+In addition, GMU grants all TT faculty a semester off from teaching before tenure.
+Some faculty use this to work on their NSF CAREER proposal while others use it to prepare their tenure application.
+
+  \end{commentbox}
+
+\section{Courses}TT faculty will teach both undergrad and grad courses.
 Grad courses include seminars, special topics, and core courses.
@@
 
+Most depts.
+would allow faculty to pick their favorite courses to teach.
+In some cases, they might ask you to teach an existing core undergraduate courses (e.g., Theory, Operating Systems) and develop your own grad course (e.g., Software Verification).
+It is likely that you will repeat the same courses.
+This significant reduces the amount of time you need to prepare for the course.
+However, you should not teach the same course every semester, and you will want to alternate between grad and undergrad courses.
+
 \begin{commentbox}[Vu]
 
-  At UNL, TT faculty is 1+1 for the first 3 years, and 1+2 after that.
-At GMU, it is 1+1 for the first 3 years, but I haven't seen any TT faculty having to teach more than that during their tenure track.
-In addition, GMU grants all TT faculty a semester off from teaching before tenure.
-Some faculty use this to work on their NSF CAREER proposal while others use it to prepare their tenure application.
-
+    At GMU, I teach the same Software Specification and Construction course at both grad (SWE-619) and undergrad (SWE-419) levels.
+While it is a bit boring to teach the same materials every semester, it is also nice because I don't have to prepare much for the course.
+I did try to teach an AI Assurance seminar course once, but never liked it much because it only has MS students, who were not very interested in research.
+
+    In contrast, I teach various courses at UNL.
+I teach the several undergraduate courses including Theory and design my own two courses on Compilers and Software Analysis.
+I also create a grad seminar on Software Verification.
   \end{commentbox}
@@
 
-\section{Teaching Philosophy}
-\section{Teaching Awards}
-
-\chapter{Mentoring}
+\section{Recruiting Students for Research} Many TT faculty use their teaching to recruit students for their research group.
+Seminar and special topics courses are designed to introduce grad students to your research area.
+Similarly, you might be able to recruit undergrads from your 400 level courses.
+Taking students from your class has the benefits that you and the students already know each other, and you know their strengths and weaknesses, e.g., through their programming assignments in your class.
+
+\chapter{Recruiting and Mentoring}
 

### 2024-09-14 demystify.tex
updates
Updates from Overleaf

@@
 would not ask you to write your own letter (in fact, even not well-known ones wouldn't do this).
-However, many international students do not know this (e.g., it is a common practice at their university) and might be in a situation they do not have a choice as they need the letter.
+This might be a common practice at their university or the students do not have a choice as they need the letter.
 However, think about this: if a prof.
@@
   \tcblower
-  Vu: One way is looking at what well-known researchers publish.
+  \textbf{Vu}: One way is looking at what well-known researchers publish.
 For example, if you are interested in a field X, you can use CSRankings to look at active faculty in X, and then look at their websites to see what journals they publish at.
@@
 In some cases a faculty interviews a candidate that they see potential and want to advocate for their admission.
-Without the interview, such applications may be more likely to be rejected.
-  \tcblower
+Without the interview, such applications may be more likely to be rejected.\\
+
   In short, getting an interview is a good sign; it means that someone is considering you. If we are not interested in your application, we will not proceed with an interview.
@@
   \textbf{Vu}: \autoref{sec:contact} shows how to contact a professor for research opportunities.
-Many probably will say no (or do not reply) as they do not have the bandwidth to take on random students, but some may say yes if they see a potential fit.
+Many will say no (or do not reply) as they do not have the bandwidth to take on random students, but some may say yes if they see a potential fit.
 \end{commentbox}
@@
 
-\begin{commentbox}[Dear Prof. Nguyen,]
+\begin{commentbox}[Dear Prof. Nguyen,]\\
 
   I am writing to inquire about potential research opportunities as a GRA in your group at Mason. Currently I am an undergraduate student in Computer Science at UNIV and plan to graduate in May 2023.
+    \\
 
@@
 
+ \tcblower
   This is a good example because it is written just for me.  It shows that the student knows about my work on invariant generation and has a related background (paper C and project G).
@@
         \item [R1/R2 Universities] Research 1 (R1) universities are universities with the highest level of research activity.
-Research 2 (R2) universities have a high level of research activity but not as high as R1 universities.
+R2 universities have a high level of research activity but not as high as R1 universities.
 R1 universities are typically larger, and have more funding and resources for research.
@@
 Our MS program is geared towards working professionals, e.g., all of our MS courses are offered in the evening and online.
+    \\
 
@@
 \subsection{Funding}
-Unlike PhD programs, which typically guarantee funding (\autoref{sec:funding}), MS programs are typically not funded.
-You have to pay tuition, which can be expensive.
-It is difficult for an MS student to get funding, e.g., through TA or RA positions, because these are typically reserved for PhD students.
+Unlike PhD programs, which often have funding (\autoref{sec:funding}), MS students are typically not funded through either RA or TA.
 \begin{itemize}
@@
 Moreover, some schools do have TAs for MS students, and you can apply for these positions.
-For example, Mason CS dept has TAs for MS students (of course not for all of our 1000+ MS students, but quite a few); my courses in the past two years have MS TAs.
+For example, Mason CS dept has quite a few TAs for MS students; my courses in the past two years have MS TAs.
 
@@
 I think this does not bother them a bit; they are genuinely interested in knowing and helping solve these ``dramas'' in my life.
+    \\
 
@@
 When Steph was in DC for a meeting, I invited her to to give a research talk at Mason.
-I have also collaborated with them after I graduated e.g., I am working with Deepak on an NSF proposal on invariant generation.
+I have also collaborated with them after I graduated e.g., I recently got an NSF grant with Deepak.
+    \\
 

### 2024-09-14 faculty.tex
Updates from Overleaf

@@
 While teaching takes about 40\% of your time, it is not as nearly as important as research for tenure.
-Moreover, for most new faculty, teaching is not as difficult as getting grants or publishing papers, and thus is often taken lightly by new faculty.
+Moreover, most new faculty do not prioritize teaching, e.g., compared to writing grants or recruiting students for research.
 However, teaching is still important for your tenure case, and being a good teacher can be very rewarding.
-Moreover, innovative integration of research and teaching can help you get good students and even NSF grants.
+Innovative integration of research and teaching can also help you get good students and even NSF grants.
 
 \section{Teaching Load}
-New TT faculty often have a lighter teaching load in their first several years or even throughout their tenure track.
-This is to give you time to establish your research program.
-The teaching load is usually higher at teaching universities, and lower at research universities.
-
-The desirable teaching load for TT at most R1 universities is 1+1 (1 course / semester).
-While some univs might give you a lighter load in your first year, most univs expect you to teach 1+1 from your second year.
+New TT faculty often have a lighter teaching load in their first several years or until they get tenure.  This is to give you time to establish your research program.
+
+At most large R1 universities the teaching load for TT faculty is 1+1 (1 course / semester).
+While some univs might give you a lighter load in your first year, most univs expect you to teach 1+1 starting your second year.
 In many cases, 1+1 is only for the first 2 or 3 years and after that it increases to 1+2 (or 2+1).
@@
 
-In many cases, your contract might list 1+1 for the first X year and 1+2 after until tenure, but the dept unofficial gives TT faculty 1+1 until tenure.
-Some other ways to reduce teaching load such as \emph{buying out} of teaching with a grant, leading a very large grant, or have taught many large classes.
+In some cases your contract might list 1+1 for the first X year and 1+2 after until tenure, but the dept unofficial gives TT faculty 1+1 until tenure.
+Some other ways to reduce teaching load include \emph{buying out} of teaching with a grant, leading a very large grant, and teaching many large classes.
 
 \begin{commentbox}[Vu]
-
     At UNL, TT faculty is 1+1 for the first 3 years, and 1+2 after that.
-At GMU, it is 1+1 for the first 3 years, but I haven't seen any TT having to teach more than that.
+At GMU, it is 1+1 for the first 3 years, but most TT faculty would continue this teaching load until they get tenure.
 In addition, GMU grants all TT faculty a semester off from teaching before tenure.
 Some faculty use this to work on their NSF CAREER proposal while others use it to prepare their tenure application.
-
   \end{commentbox}
 
-\section{Courses}TT faculty will teach both undergrad and grad courses.
-Grad courses include seminars, special topics, and core courses.
-Most new faculty use seminars and special topics to recruit students for their research group.
-Undergrad courses are typically higher level courses in your area of expertise (e.g., 400 level course on Software Testing).
-It is rare that you have to teach 100 level courses (e.g., Intro to Programming) because there are just too many students in these courses.
+\section{Courses}
+
+In most cases you will teach both undergrad and grad courses.
+New faculty often create graduate seminars and special topic courses to recruit students for their research group.
+Undergrad courses typically include higher level courses in your area of expertise (e.g., 400 level course on Software Testing).
+It is rare that you have to teach 100 level courses (e.g., Intro to Programming) because they are very large and thus time-consuming.
 
@@
 In some cases, they might ask you to teach an existing core undergraduate courses (e.g., Theory, Operating Systems) and develop your own grad course (e.g., Software Verification).
-It is likely that you will repeat the same courses.
+It is likely that you get to repeat the same courses.
 This significant reduces the amount of time you need to prepare for the course.
@@
 \begin{commentbox}[Vu]
-
     At GMU, I teach the same Software Specification and Construction course at both grad (SWE-619) and undergrad (SWE-419) levels.
-While it is a bit boring to teach the same materials every semester, it is also nice because I don't have to prepare much for the course.
-I did try to teach an AI Assurance seminar course once, but never liked it much because it only has MS students, who were not very interested in research.
+This is very nice because I don't have to prepare much for the course.
+I did try to teach an AI Assurance graduate seminar once, but never liked it much because it only has MS students, who were not very interested in research.
 
     In contrast, I teach various courses at UNL.
-I teach the several undergraduate courses including Theory and design my own two courses on Compilers and Software Analysis.
+I teach the several undergraduate courses including Theory, Software Engineering, and design my own two courses on Compilers and Software Analysis.
 I also create a grad seminar on Software Verification.
+
+    I feel that UNL is a bit more flexible in creating new classes compared to GMU, where I am pretty much the main instructor for SWE-619/SWE-419.
   \end{commentbox}

### 2024-09-25 faculty.tex
updates

--- 2024-09-12 faculty.tex
@@
 \section{Mentoring Undergraduate Students}
+I have mentored a dozen of undergrad students. Most of them did not last long, but some of them did great work. 
 Surprisingly, some of the best research projects I have seen come from undergraduate students.

### 2024-09-25 demystify.tex
updates

@@
 
+The rest of this document will provide you with more details on the CS PhD admission process in the US, including how to prepare your application, what to expect during process, what to do with acceptance (and how to deal with rejection), and so on.
+The goal is to help you understand the process and make the best decisions for your future.
+
 \begin{commentbox}[Vu:]
@@
 At Mason, the PhD adcom typically has 15--20 faculty, and each committee member is assigned to review about 30 applications.
-Note that most large schools, including Mason, have separate adcoms for MS programs.
+Note that most large schools, including Mason, have separate adcoms for MS programs (\autoref{chap:ms}).
 
 Each application is assigned to about three adcom members, who will evaluate your profile and reach a consensus.
-While the assigned reviewers are the main ones deciding your application, other faculty in the department will likely have access to your application and can provide inputs and opinions on your profile.
+While the assigned reviewers are the main ones deciding your application, other faculty in the department can also have access to your application and provide inputs and opinions on your profile.
 
@@
 
-\paragraph{Why CS depts do not waive the application fee?}  Some universities do waive, e.g., Rice does for all PhD applicants and many universities for domestic students (\autoref{sec:domestic-students}), however most do not.
+\paragraph{Why CS depts do not waive the application fee?}  Some universities do waive, e.g., Rice does for all PhD applicants and many universities for domestic students (\autoref{sec:domestic-students}), however most do not for international students.
 In any case, the application fee is typically a requirement of the university.
@@
 Also, most CS programs already receive many applications and would be overwhelmed if the application process were free.
+Note that if you have financial difficulties, you can ask the department for a waiver, but this is typically only granted in exceptional cases.
 
@@
 Most PhD programs will require at least \textbf{two LORs}. 
-A good LoR letter should give details of the applicant's research background and achievements, all of which paint a compelling picture of the candidate to the adcom.
+The goal of the LoR is to provide the adcom with an assessment of your research ability and potential. 
+The letter should 
+focus on your research background, achievements, and potentials, all of which paint a compelling picture of the candidate to the adcom.
+
+\paragraph{LoR writers} LoR writers should be someone who (i) can talk in depth about your research experience and potential and (ii) have the credibility to evaluate your research ability.
+
+A LoR writer should be someone who knows you well, so that they can provide a strong-supporting and personalized letter for you.  
+LoR writers should also have the credibility to convince adcom members.
+Such a person is typically someone who has a PhD (so they know what a PhD is about) and is active in research (so they can evaluate your research ability).
+LoR writers are often your research advisors, professors, or supervisors who have worked with you on research projects.
+
 Having a strong letter from an internationally recognized researcher will \emph{greatly strengthen} your application.
@@
 
-\chapter{MS Admission}
+\chapter{MS Admission}\label{chap:ms}
 \chapterinfo{MS focuses on coursework and prepares you for industry, while PhD focuses on research and prepares you for academia or research.}

### 2024-09-26 demystify.tex
updates

@@
 
-\subsection{Self-written Letters are Bad} Many international students write letters themselves, typically due to the requests of their profs., and have them signed by their profs.
+\subsection{Self-written Letters are Bad}\label{sec:self-letters}
+ Many international students write letters themselves, typically due to the requests of their profs., and have them signed by their profs.
 Such letters have \emph{little value} and are considered weak by reviewers---why can you not even find someone who cares or knows enough about you to write a candid personal reference letter?
@@
 
-Most well-known and well-respected profs.
-would not ask you to write your own letter (in fact, even not well-known ones wouldn't do this).
+Most well-known and well-respected profs would not ask you to write your own letter (in fact, even not well-known ones wouldn't do this).
 This might be a common practice at their university or the students do not have a choice as they need the letter.
@@
 
-Many international students try to get letters from high-ranking admin people in their universities, e.g., department head, dean, or director.
-These people, while well-known and respected in your university, might not be too active in research (e.g., they focus on administration) and thus are not recognized by adcom members, who often recognize good researchers (even if they are not in their field'', but not admin people.
-So it's fine to to have `less well-known'' people write for you as long as they can write a good letter about your research ability.
+Many international students try to get letters from high-ranking admin people in their universities such as department head, dean, or director.
+The students often mistakenly believe that these letters are more valuable because such people are important in their universities.
+However, these people, while well-known and respected in your university, might not be active in research (e.g., they focus on administration) and thus are not recognized by adcom members.
+In fact, to adcom members, such a letter has little value if the writer does not know you well enough (e.g., did not work with you) to talk about your research ability.
 
 \begin{commentbox}[Vu:]
-  In my experience in reviewing applications, letters from admin people are often generic and do not provide much value.  If you are in this situation, you should find someone else to write for you.
+  In my experience in reviewing applications, letters from admin people are often generic and do not provide much value. 
+  In many case, the letter reads like it was written by a student (\autoref{sec:self-letters}), and thus is a red flag.  So if you are in this situation, you should find someone else to write for you.
+    So if you are in this situation, find someone else to write for you.
 \end{commentbox}
@@
 \subsection{Help your LOR Writer} As mentioned above, do not write your own letter and generic letters do not give much value.
-One thing you can do to help your writer to write a strong, customized LoR for you is to provide details or unique things about yourself.
+However, you can help your writer to write a strong, customized LoR for you by providing details or unique things about yourself.
 For example, let them know about your GPA, research and work experience, papers (if any), or anything you want them to mention.
 You can also provide them with a draft of your SOP so that they can see what you are saying about yourself and complement that with their own perspective.
+
 Sometimes your writer will explicitly ask you for such information, but if not, you should provide it anyway (especially if you have not interacted with them much or have not done much research with them).
@@
 So write it well because it could make a big difference.
-
-In your statement, you have the opportunity to connect directly with adcom members and make your application stand out and unique.
-Your SOP can convince adcom members that you would \emph{fit} the program, even if you don't have very strong research experience.
 A well-written SOP also shows that you can communicate, which is very important in research, and that you can effectively teach and communicate with students, which is important for TA funding (see \autoref{sec:funding}).
+
+%In your statement, you have the opportunity to connect directly with adcom members and make your application stand out and unique.
+Your SOP can convince adcom members that you would \emph{fit} their program, even if you don't have very strong research experience.
 % is important because if you need (GTA) funding, it will provide evidence
@@
 
-In your SOP, focus on research motivation and convince us that you can achieve it through your experience, e.g., published papers.
+In your SOP, focus on research potential and convince us through your experience, e.g., published papers.
 You should also back up your claims with concrete evidence.
@@
 
-Finally, you should get your SOP reviewed by your LoR writers, professors (especially those who have served in adcom), or even postdocs or PhD students (because they have been through this process).
+Finally, you should get your SOP reviewed by your LoR writers and professors, especially those who have served in adcom, or even postdocs or PhD students as they have been through this process.
 
@@
 
-\subsection{Work Experience} Work experience at \emph{well-known research laboratories}, such as Microsoft Research, can strengthen your
-application.
-Unfortunately, many good research places in your countries, e.g., VinAI in Vietnam, remain relatively unknown to most universities in the US.
-So you or your LoR writers should explicitly say something about them in your statement.
-In general, if you did some good research work at such places, then you should mention that in your SOP and ask your supervisor to write about your research experiences and potential in their LORs (\autoref{sec:lor}).
-
-\begin{commentbox}[Hung:]
-  The reputation of VinAI has been increasing steadily over the past few years; many of my colleagues heard about VinAI.
-\end{commentbox}
-
-On the other hand, if non-research work (e.g., software development) \emph{do not} carry much weight for PhD (they can help for MS application though).
-Similarly, LoR from your supervisor for non-research experience will not count much.
-So do not spend much time talking about non-research-related job in your SOP.
-
-\subsection{Competitions} Winning \emph{internationally recognized competitions} can demonstrate your research potential.
-For example, participating in Math Olympiads if you want to do theory or winning ACM programming contests if you want to ``build'' systems, e.g., software analysis.
-
 \begin{commentbox}[Thanh:]
@@
   \end{commentbox}
+
+\subsection{Work Experience} Work experience at \emph{well-known research laboratories}, such as Microsoft Research, can strengthen your
+application.
+Unfortunately, many good research places in your countries, e.g., VinAI in Vietnam, remain relatively unknown to most universities in the US.
+So you or your LoR writers should explicitly say something about them in your statement.
+In general, if you did some good research work at such places, then you should mention that in your SOP and ask your supervisor to write about your research experiences and potential in their LORs (\autoref{sec:lor}).
+
+\begin{commentbox}[Hung:]
+  The reputation of VinAI has been increasing steadily over the past few years; many of my colleagues heard about VinAI.
+\end{commentbox}
+
+On the other hand, if non-research work (e.g., software development) \emph{do not} carry much weight for PhD (they can help for MS application though).
+Similarly, LoR from your supervisor for non-research experience will not count much.
+So do not spend much time talking about non-research-related job in your SOP.
+
+\subsection{Competitions} Winning \emph{internationally recognized competitions} can demonstrate your research potential.
+For example, participating in Math Olympiads if you want to do theory or winning ACM programming contests if you want to ``build'' systems, e.g., software analysis.
 
@@
 If you have just a few, then they do not matter much, so don't spend much time explaining them (many adcom reviewers themselves have bad grades in relevent courses!).
-But if you have many bad grades for an entire semester or year due to some specific reasons, then you may be you should explain them in your SOP.
+But if you have many bad grades for an entire semester or year due to some specific reasons, then you should explain them in your SOP.
 
@@
 
+\begin{commentbox}[Vu:]
+    Unlike a job application, CV is not as important for PhD admission.
+We care more about your LOR, SOP, etc.
+So do not spend too much time on your CV, just make sure it is easy to skim through and well-organized around research activities and achievements.
+\end{commentbox}
+
 \section{Interview and the Waiting Game}\label{sec:interview}
@@
 In some cases, the interview is done by several professors, e.g., to see if a student fits in their group or to recruit a very strong student to their program.
-The interview can also be done by adcom members to see if the student fits their program.
 
@@
 You might be asked to talk about your research experience and interests and to read a paper and discuss it.
-In some rare cases you might also be asked to solve a problem, e.g., a coding problem (one of my colleagues at Mason often does this).
+In some rare cases you might also be asked to solve a problem (one of my colleagues at Mason often does coding interview).
 The interview is also a good opportunity for you to ask questions, e.g., about the prof, their group, the CS PhD program, and the university.
@@
 So  know what your insurance covers and be prepared for unexpected costs.
-Moreover, healthcare system has many confusing jargons such as HMO, PPO, deductibles, copays, and coinsurance (try take a look at the Explanation of Benefits or EoB statement you received from your insurance company).
+Moreover, healthcare system has many confusing jargons such as HMO, PPO, deductibles, co-pays, and coinsurance (try take a look at the Explanation of Benefits or EoB statement you received from your insurance company).
 It's arguably the most complicated system in the US and even Americans often do not understand it (and politicians often exploit this to their advantage).

### 2024-09-30 faculty.tex
updates

--- 2024-09-14 faculty.tex
@@
 
-\section{Most people suck at grant writing (at first)} Most new assistant professors have no experience in grant writing, and struggle with it in the beginning.
-It becomes especially stressful starting your 3rd year when departmental evaluation starts to give you more pressure and warning if you don't have enough funding.
-By your midterm or fourth-year review, you are expected to have at least some external funding, and can be in serious trouble if you don't have any.
-So start early, and don't wait until the last minute to start writing grants.
-It's a skill that can be learned, and you will get better over time.
-The key is to keep trying, and to not give up.
-If you are not getting rejected, you are not aiming high enough.
+\section{Most people suck at grant writing (at first)} Most new assistant professors have no experience in grant writing, and struggle with it.
+It becomes especially stressful starting your 3rd year when departmental evaluation starts to look at your funding.
+By your midterm or fourth-year review, you are expected to have at least some external funding, and can be in trouble if you don't have any.
+However, despite initial struggles, most people get better at grant writing over time.
 The best researchers in the world get rejected all the time.
-It's part of the process.
 So don't get discouraged.
@@
 would allow faculty to pick their favorite courses to teach.
-In some cases, they might ask you to teach an existing core undergraduate courses (e.g., Theory, Operating Systems) and develop your own grad course (e.g., Software Verification).
+In some cases, they let  you develop your own grad course (e.g., Software Verification) but ask you to teach an existing core undergraduate courses (e.g., Theory, Operating Systems).
 It is likely that you get to repeat the same courses.
@@
 I did try to teach an AI Assurance graduate seminar once, but never liked it much because it only has MS students, who were not very interested in research.
+    \\
 
@@
 I teach the several undergraduate courses including Theory, Software Engineering, and design my own two courses on Compilers and Software Analysis.
-I also create a grad seminar on Software Verification.
-
-    I feel that UNL is a bit more flexible in creating new classes compared to GMU, where I am pretty much the main instructor for SWE-619/SWE-419.
+I also create a grad seminar on Software Verification.\\
+
+    I feel that UNL is a bit more flexible in creating new classes compared to GMU.  However, I spent much less time prepare for courses at GMU.
   \end{commentbox}
 
-\section{Teaching Evaluation}
-
+\paragraph{Co-Teaching}
+Some places encourage new TT faculty to co-teach with another faculty.
+This is a way to share the workload and to learn from your more experienced colleague.
+It has many benefits, but it can also be challenging if you and your co-teacher have different teaching styles.
+
+\begin{commentbox}[Vu]
+I co-taught an undergrad Software Engineering course once at UNL with a teaching faculty.
+I didn't like it because the teaching faculty was very experienced and good at teaching, and I felt that I was not contributing much.
+I never co-taught again after that.
+\end{commentbox}
+
+\section{Teaching Evaluation}\label{sec:teaching-evaluation}
+
+Teaching evaluations are a critical component of your annual review and P\&T. It provides an analysis on how students and peers perceive your teaching. 
+You pay attention to evaluations, and make changes to your teaching based on the feedback, and show that you continuously strive to improve your teaching.
+
+\paragraph{Student Evaluation}
 At the end of each semester, students are asked to evaluate the instructor.
+The evaluation often includes both quantitative (e.g., ratings on a scale of 1-5) and qualitative (e.g., comments) feedback.
+In addition to default questions, you can add your own questions.
 The results are shared anonymously with the instructor to help them improve their teaching.
-Teaching evaluation is also commonly used in annual and tenure evaluation.
 
@@
 Students not only do it themselves but also remind their friends.
+  \\
+
   It's been working well for me: Spring '24 I had a 90\% response rate (25 out of 28 students), and Fall '23 was a perfect 100\% (16 out of 16 students).
@@
   \end{commentbox}
+
+\paragraph{Peer Evaluations}
+
+In addition to student feedback, peer evaluations can offer valuable insights on your teaching. At many places, peer evaluations are required for P\&T.
+Request evaluations from experienced colleagues who can observe your class and provide constructive criticism.
+Like with student evaluation, incorporating these peer feedback into your teaching and describe your changes and improvements.
+Peer evaluation letters can also help you get Teaching Awards and other recognitions.
 
@@
 
+\begin{commentbox}[Vu]
+  I have recruited my first PhD student Guolong from my seminar and many undergrads from my classes.
+However, other than Guolong, the undergrads recruited from my class did not work out well.
+I found that undergrads recruited from my class are not as motivated as those who come to me on their own.
+So, I stop recruiting students from my class and only take those who come to me on their own.
+More about recruiting students in \autoref{sec:recruiting}
+\end{commentbox}
+
 \chapter{Recruiting and Mentoring}
@@
 There will be times when you feel discouraged and frustrated, but there will also be times when you feel proud and happy.
+
 For P\&T you will need to successfully mentoring students.  Many universities expect you to need to graduate at least one PhD student to get tenure.
+
+\section{Recruiting}\label{sec:recruiting}
 
@@
 
-I have worked with many undergrads (15+ since 2016). My success rate (publish a research paper with me) with undergrads
+I have worked with many undergrads (15+ since 2016).
 
 \section{Mentoring Postdocs}
+
+I never formally mentored a postdoc, but I have worked with several postdocs.
+They are more like colleague than students as they are more mature and independent, e.g., writing own papers and leading projects.
+
 \section{Mentoring Junior Faculty}
@@
 
-After 5 or 6 years, it is time for you to go up (or apply) for tenure.
+After 5 or 6 years, it is time for you to go up (apply) for tenure.
 You will need to submit a tenure package.
@@
 
-\section{Change in Department. Moving from UNL to GMU}
-
-I joined UNL in 2016 and was very happy there until 2020, when people start leaving and the dept.
+% \section{Change in Department. Moving from UNL to GMU}
+
+% I joined UNL in 2016 and was very happy there until 2020, when people start leaving and the dept.
 changed its leadership and organization.
@@
 
-I will skip the details and just summarize several important and disappointing events.
-
-\begin{itemize}
-\item 2019: four senior professors (in SE) left.
+% I will skip the details and just summarize several important and disappointing events.
+
+% \begin{itemize}
+% \item 2019: four senior professors (in SE) left.
 One of these 4 was the dept chair (and he left just after 2 months of being chair).
 There were issues of upper admin didn't work well with the dept chair (and this issue has been going for years).
-\item 2020: College of Engr pressure dept to move entirely to it.
+% \item 2020: College of Engr pressure dept to move entirely to it.
 The new chair candidate only agrees to accept the job if the faculty moved to CoE, putting pressure on the move.
 Faculty are split on the decision and eventually moved with lots of mistrusts and concerns under CoE.
-\item Salary inversion:  new chair offers new AP salary that is much higher than current AP, leading to jokes that current AP should just reapply to UNL.
+% \item Salary inversion:  new chair offers new AP salary that is much higher than current AP, leading to jokes that current AP should just reapply to UNL.
 This became a serious issue and many people at UNL use this as a reason on why they move.
-\item Senior faculty fighting on mailing list.
-
-\end{itemize}
-
-Note that even after leaving Lincoln, my family and I still have a soft spot for the city and its people.
+% \item Senior faculty fighting on mailing list.
+
+% \end{itemize}
+
+% Note that even after leaving Lincoln, my family and I still have a soft spot for the city and its people.
 It was a great place to live and raise a family, but unfortunately not a good place to work for me.

### 2024-09-30 demystify.tex
updates

@@
 
-After you submit your PhD application (e.g., in late Fall of year $X$ for the Fall admission of year $X+1$), it will first be checked
-for general requirements, e.g., did you submit your transcripts and standard scores?
+After you submit your PhD application (e.g., in late Fall of year $X$ for the Fall admission of year $X+1$), it will first be checked for general requirements, e.g., did you submit your transcripts and standard scores?
 did your reference writers submit their letters?
@@
 
-If you pass this, your application will be forwarded to the CS department for further evaluation.
-If you don't,  the system will tell you what is missing and what you need to do.
+After screening, your application is complete and forwarded to the CS department for further evaluation.
+If you don't pass screening,  the system will tell you what is missing and what you need to do.
 So pay attention to your email and check your application status regularly.
@@
 \subsection{Self-written Letters are Bad}\label{sec:self-letters}
- Many international students write letters themselves, typically due to the requests of their profs., and have them signed by their profs.
+ Many international students write letters themselves, typically due to the requests of their profs.
+or letter writers, and have them signed by their profs.
 Such letters have \emph{little value} and are considered weak by reviewers---why can you not even find someone who cares or knows enough about you to write a candid personal reference letter?
@@
 
-Most well-known and well-respected profs would not ask you to write your own letter (in fact, even not well-known ones wouldn't do this).
-This might be a common practice at their university or the students do not have a choice as they need the letter.
+Most well-known and well-respected profs would \emph{not} ask you to write your own letter (in fact, even not well-known ones wouldn't do this).
+This might be a common practice at specific universities and the students do not have a choice as they need the letter.
 However, think about this: if a prof.
-does this often, then they are likely not a good English writer, relatively unknown in the field, or simply do not know/care enough about you.
-In any of these cases, their letters won't worth much.
+does this often, then they either do not know/care enough about you or don't know how to write a good letter.
+In any case, such letters are not useful and might even hurt your application.
 So if you are in this situation, you should find someone else to write for you.
 
-\subsection{Letters from Admin people}\label{sec:admin-letters}
-
-Many international students try to get letters from high-ranking admin people in their universities such as department head, dean, or director.
-The students often mistakenly believe that these letters are more valuable because such people are important in their universities.
+\subsection{Letters from Admin People}\label{sec:admin-letters}
+
+Many international students try to get letters from high-ranking admin people in their universities such as department chair/head, dean, or director.
+The students often believe that these letters are more valuable because such people are important in their universities.
 However, these people, while well-known and respected in your university, might not be active in research (e.g., they focus on administration) and thus are not recognized by adcom members.
-In fact, to adcom members, such a letter has little value if the writer does not know you well enough (e.g., did not work with you) to talk about your research ability.
+Such a letter has little value if the writer does not know you well enough (e.g., did not work with you) to talk about your research ability.
 
@@
 
-  \item[Adcom] Admission Committee. The group of faculty members who review applications and make admission decisions.
-  \item [Adcom members] Faculty members who are part of the admission committee. These are the people who review applications and make admission decisions.
-  \item[Adcom chair] The faculty member who leads the admission committee.
+  \item[Adcom \autoref{sec:evalapps}] Admission Committee. The group of faculty members who review applications and make admission decisions.
+
+  \item [Adcom members \autoref{sec:evalapps}] Faculty members who are part of the admission committee. These are the people who review applications and make admission decisions.
+
+  \item[Adcom chair \autoref{sec:evalapps}] The faculty member who leads the admission committee.
 This person is often not involved in individual admission decisions but oversees the process (e.g., resolve disputes, ensure fairness).
+
   \item[Advisor/Supervisor]: A faculty member who guides and mentors a PhD student throughout their research.
@@
 In the US, the term ``adviser'' is more commonly used than ``supervisor''.
+
   \item[Cohort] A group of students who start a program at the same time and take classes together. You should get to know your cohort as they will be your colleagues and friends during your PhD.
+
   \item [Deferred Admission] An option allowing admitted students to postpone their start date, typically by one year.
@@
 This document is often required for faculty positions, but some universities also require it for graduate admissions.
+Note that you do not need to be from an underrepresented group to write a diversity statement.
+
   \item[Domestic Student] A student who is typically a US citizen or permanent resident (``greencard holder'').
         But sometimes it can also refer to students who did their undergrad (or MS) at a US university.
+
   \item[Field or Area] A specific research area, e.g., Machine Learning, Computer Vision, Software Engineering, etc.
+
   \item [In-state vs.
@@
 PhD students typically do not have to pay tuition as it is covered by their funding, but this might be important for MS students.
+
   \item [Ivy League] A group of eight private universities in the US known for their academic excellence and social prestige.
 The Ivy League schools are Brown, Columbia, Cornell, Dartmouth, Harvard, the UPenn, Princeton, and Yale.
-Note that these schools are typically known for their undergraduate programs more than their graduate programs.
-Moverover, most top CS programs such as CMU, UIUC, UCSD, UWash, MIT, Stanford, are not in the Ivy League.
+These schools are typically known for their undergraduate programs more than their graduate programs.
+Moreover, most top CS programs such as CMU, UIUC, UCSD, UWash, MIT, Stanford, are not in the Ivy League.
+See selecting and ranking schools in \autoref{sec:selecting-ranking-schools}.
 
@@
 Typically a lab is led by a single professor, but sometimes it can be co-led by multiple professors.
-For example, my lab is DynaROARS, which focuses on my own research area (software verification), but I am also part of the more general SWE Lab at Mason with other faculty in Software Engineering.
+For example, my lab \href{https://dynaroars.cs.gmu.edu}{DynaROARS} focuses on my own research area (software verification), but I am also part of the more general SWE Lab at Mason with other faculty in Software Engineering.
 
@@
 
-  \item[MS] Master of Science. A graduate degree that typically requires 2 years of study.
-  \item[Open House] An event where admitted students visit the university to meet faculty, students, and staff, and learn more about the program.
+  \item[MS \autoref{sec:msrequirement}] Master of Science. A graduate degree that typically requires 2 years of study.
+
+  \item[Open House \autoref{sec:accepted}] An event where admitted students visit the university to meet faculty, students, and staff, and learn more about the program.
 This is a great opportunity to get a feel for the university and the department and decide if it is a good fit for you.
 
-  \item[Q1 journal] Commonly mentioned by international students to refer to top-tier journals in CS.
+  \item[Q1 Journal] Commonly mentioned by international students to refer to top-tier journals in CS.
 This is not a standard in CS in the US (which often focuses on confs.
@@
 \begin{table}
-    \caption{PhD vs. MS}\label{tab:phd-vs-ms}
+    \caption{MS vs. PhD}\label{tab:phd-vs-ms}
     \begin{tabular}{c|c|c}
     \toprule
-    \textbf{} & \textbf{PhD} & \textbf{MS} \\
+    \textbf{} & \textbf{MS} & \textbf{PhD} \\
     \midrule
@@
 
-\autoref{tab:phd-vs-ms} summarizes the main differences between PhD and MS programs:
+\autoref{tab:phd-vs-ms} summarizes the main differences between MS and PhD programs:
 \begin{itemize}
@@
 
-\subsection{Admission Committee}
+\subsection{Admission Committee}\label{sec:ms-adcom}
 MS admission also involves an adcom that reviews applications and makes admission decisions.

### 2024-09-30 demystify.tex
updates

@@
     \item \textbf{Not reading the professor's website} and asking questions that are already answered there.
+    \item \textbf{Not using plain text}. Fancy format, colors, or fonts are not professional. CS people like plain text.
 \end{itemize}
@@
 \item \textbf{Employment:} You are allowed to work on-campus for up to \textbf{20 hours per week during the academic year} (because you still need to take classes) and \textbf{full-time} (typically 40 hrs) during official school breaks (e.g., summer and winter breaks).
-Off-campus employment requires authorization, which can be obtained through programs like Curricular Practical Training and Optional Practical Training described below.
+Off-campus employment requires authorization, which can be obtained through CPT and OPT programs described below.
 
@@
 PhD also has coursework requirements (typically taken in your first 2 years) but focuses on research after that.
-You graduate with a PhD when you have done enough research and written a dissertation (which is typically way harder and takes much longer time than coursework).
+You graduate with a PhD when you have done enough research and written a dissertation, which usually takes much longer time than coursework.
 
@@
 
-  \item Funding: MS is typically \emph{not funded}, while PhD is typically funded (\autoref{sec:funding}).
+  \item Funding: MS is typically \emph{not funded}, while PhD is typically funded (\autoref{sec:funding}). See \autoref{sec:ms-funding} for more details on MS funding.
 
@@
 
-\subsection{Funding}
+\subsection{Funding}\label{sec:ms-funding}
 Unlike PhD programs, which often have funding (\autoref{sec:funding}), MS students are typically not funded through either RA or TA.
@@
 
-However, there are always exceptions, e.g., if you have a strong background and can demonstrate that you can be productive, then a prof.
-might be willing to take you as an RA.
-Moreover, some schools do have TAs for MS students, and you can apply for these positions.
-For example, Mason CS dept has quite a few TAs for MS students; my courses in the past two years have MS TAs.
+However, there are always exceptions, e.g., if you have a strong background and can demonstrate that you can be productive, then a prof. might be willing to take you as their RA.
+Moreover, some schools have TAs for MS students, and you can apply for these positions. For example, Mason CS dept has quite a few TAs for MS students; my courses in the past two years have MS TAs.
 
@@
 This allows me to have a good understanding of the process, the challenges students face, and what faculty are looking for.
-I also serve as the program director of the MS program in Software Engineering at Mason (thus also have experience with MS admission---which is quite different than PhD).
+I also serve as the program director of the MS program in Software Engineering at Mason (thus also have experience with MS admission---which is quite different than PhD as discussed in \autoref{chap:ms}).
 My personal and lab website is at \href{https://dynaroars.github.io}{dynaroars.github.io}.

### 2024-10-02 demystify.tex
updates

@@
 
-After you submit your PhD application (e.g., in late Fall of year $X$ for the Fall admission of year $X+1$), it will first be checked for general requirements, e.g., did you submit your transcripts and standard scores?
+After you submit your PhD application, it will first be checked for general requirements, e.g., did you submit your transcripts and standard scores?
 did your reference writers submit their letters?
@@
 
-The guideline can be useful for all students, but it is primarily intended for \textbf{international students} from \textbf{smaller countries and less well-known universities} (or domestic students from US universities with no PhD programs or limited research experiences).
+The guideline can be useful for all students, but it is primarily intended for \textbf{international students} from \textbf{smaller countries and less well-known universities} (or domestic students from US universities with no PhD programs or limited research opportunities).
+
 Students from top schools with strong research cultures and experiences might already know most of the information given here.
@@
 
-While this document may not fully meet the needs of students aiming for top-10 schools, which typically have more stringent requirements (e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU), it should be valuable for those targeting other reputable R1 institutions.
+While this document may not fully meet the needs of students aiming for top-10 schools (\autoref{sec:ranking}), which typically have more stringent requirements (e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU), it should be valuable for those targeting other reputable R1 institutions.
 It is certainly useful for students aiming for Mason!
@@
 
-\subsection{Remind Your Writers} After you submit your application, you should tell your writers about that and let them know they will soon receive an email from the university to submit their letters.
+\subsection{Remind Your Writers}
+
+After you submit your application, you should tell your writers about that and let them know they will soon receive an email from the university to submit their letters.
 You should also remind them when you submit your application and ask them to submit their letters on time if they haven't done so.
@@
 
-\subsection{VEF 2.0} For Vietnamese students, it's worth mentioning the \href{https://vef2.org/}{VEF2.0 program}, which has helped many students gain admission to top CS PhD programs in the US.
+\begin{commentbox}[Vu:]
+For Vietnamese students, it's worth mentioning the \href{https://vef2.org/}{VEF2.0 program}, which has helped many students gain admission to top CS PhD programs in the US.
 VEF2.0 invites US faculty members from leading institutions to conduct rigorous interviews with VEF students and subsequently provide reference letters on their behalf.
 Despite the limited interaction between the interviewer and the interviewee (primarily just the interview), these reference letters are generally effective and have helped many students get into good PhD programs.
+\end{commentbox}
 
@@
 
-A common question is whether you should be concern and explain bad grades in relevant courses in your SOP.
-If you have just a few, then they do not matter much, so don't spend much time explaining them (many adcom reviewers themselves have bad grades in relevent courses!).
+Should you explain bad grades in relevant courses in your SOP?
+If you have just a few,  they do not matter much, so don't spend much time explaining them (many adcom reviewers themselves have bad grades in relevant courses!).
 But if you have many bad grades for an entire semester or year due to some specific reasons, then you should explain them in your SOP.
@@
 You can also ask to do an independent study or research with a prof.
-This can be informal (no credit) and takes place during the summer or winter break.
+This can be informal (no credits) and takes place during the summer or winter break.
 For example, I do this with several students, some of whom are undergrads.
@@
 \textbf{Tenure-line faculty}, consisting of tenured and tenure-track profs., focus on research, which includes publishing papers, obtaining grants, and mentoring PhD students.
-They often have a low teaching load (e.g., 1 per semester).
+They often have a low teaching load, e.g., 1 course per semester.
 \emph{Tenured} faculty are professors who have been promoted to permanent positions (informally, very hard to fire them).
-\emph{Tenure-track} faculty are (more likely young) professors who are on the track to get tenure.
+\emph{Tenure-track} faculty are professors who are on the track to get tenure.
 For most schools, tenure-line is the main type of faculty that can serve as a formal adviser of PhD students.
@@
 Assistant means new faculty, associate means they have been promoted, and full means they are senior.
-Typically, a tenure-track faculty starts with being an assistant professor and then gets tenure and promoted to associate (typically after 6 years) and then full (time varies, some become full within 3--4 years, some 10+ years, some remain associate).
-A non-tenure track faculty, e.g., teaching or research faculty, also starts with Assistant, then is promoted to Associated, and Full prof.
+Typically, a tenure-track faculty starts with being an assistant professor and then gets tenure and promoted to associate (around 6 years) and then full (time varies, some become full within 3--4 years, some 10+ years, some remain associate).
+A non-tenure track faculty, e.g., teaching or research faculty, also starts with Assistant, then is promoted to Associated, and finally Full.
 

### 2024-10-02 demystify.tex
updates

@@
 
-\paragraph{Adcom} After screening applications are reviewed by a PhD admission committee or \emph{adcom} that consists of faculty members in CS\footnote{In some cases the committee can involve affiliated faculty from different disciplines.}.
+\subsection*{Admission Committee (adcom)}\label{sec:adcom}
+
+Your applications are reviewed by a PhD admission committee or \emph{adcom} that consists of faculty members in CS\footnote{In some cases the committee can involve affiliated faculty from different disciplines.}.
 These adcom members have a wide range of expertise and background to ensure diverse perspectives in the evaluation process.
@@
 This provides junior faculty the opportunities to recruit students.
-Some members are tenured faculty.
 The chair of the committee will likely be a senior faculty, but they will not review individual applications and instead assign them to committee members.
@@
 
-\paragraph{Why CS depts do not waive the application fee?}  Some universities do waive, e.g., Rice does for all PhD applicants and many universities for domestic students (\autoref{sec:domestic-students}), however most do not for international students.
-In any case, the application fee is typically a requirement of the university.
-Individual departments and programs do not have the flexibility to waive the application fee, even if they want to.
-
-In my opinion, requiring applicants to pay the fee helps ensure their seriousness, as it filters out non-serious candidates.
-Also, most CS programs already receive many applications and would be overwhelmed if the application process were free.
-Note that if you have financial difficulties, you can ask the department for a waiver, but this is typically only granted in exceptional cases.
-
 \section{Who is this document for?}
@@
 \end{itemize}
+
+\section{Miscs}
+
+\subsubsection{Why CS depts do not waive the application fee?}
+
+Some universities do waive, e.g., Rice does for all PhD applicants and many universities for domestic students (\autoref{sec:domestic-students}), however most do not for international students.
+In any case, the application fee is typically a requirement of the university.
+Individual departments and programs do not have the flexibility to waive the application fee, even if they want to.
+
+\begin{commentbox}{Vu:}
+In my opinion, requiring applicants to pay the fee helps ensure their seriousness, as it filters out non-serious candidates.
+Also, most CS programs already receive many applications and would be overwhelmed if the application process were free.
+Note that if you have financial difficulties, you can ask the department for a waiver, but this is typically only granted in exceptional cases.
+\end{commentbox}
+\subsection{How are decisions made?}\label{sec:how-decisions}
+\subsectioninfo{Even if all adcom reviewers recommend acceptance, the application can still be rejected.
+Vice versa, if all reviewers think the application is weak, the student might still be admitted.}
+
+After reviewers have evaluated an application, adcom chair will review all evaluations, look at entered notes, and ask reviewers to discuss and resolve discrepancies to reach a consensus (e.g., a reviewer wants to accept but the other wants to reject).
+Typically, the decision is made entirely by the reviewers; there is no involvement from the adcom chair, department chair, or others.
+In most cases adcom members, even those reviewing the same application, make decisions independently and do not talk to each other (just a common practice to avoid biasing).
+In some rare cases we might (\autoref{sec:adcom-discuss}).
+
+Even if \emph{all} reviewers recommend acceptance, the application is not automatically accepted, especially if no faculty is willing to advise the student.
+For example, if the student is interested in a research area that no faculty is working on, or an area where no faculty is taking new students (e.g., AI/ML where faculty likely already has many students) then the student will not be admitted.
+This is increasingly common as the number of applicants increases significantly in recent years in CS.
+
+However, if the student has contacted a faculty member and that faculty member is interested in the student and has made this known to the admission committee, then the student is likely to be admitted, \emph{even if the faculty does not review the application or has funding} (thus the importance of contacting faculty).
+Similarly, if the student mentioned a faculty member in their SOP, adcom might ask that faculty member to look at their application and if they are interested in the student.
+Even if the student has a weak profile (but still passes the minimum requirement from the university), they might be admitted if a faculty member is willing to take them.
+Adcom members, especially in the US, are very reluctant to go against the faculty's decision (e.g., if a faculty wants to admit a student, we are not going to reject them).
+
+\subsection{Do adcom members talk to each other?}\label{sec:adcom-discuss}
+\subsectioninfo{Sometimes adcom members discuss applicants, but in most cases they make independent decisions.}
+
+However, in some rare cases we do talk to each other.
+For example, when there are discrepancies in evaluations, in which case the adcom chair will ask reviewers to discuss the application to reach a consensus.
+We also talk to each other for interesting or strong applications, e.g., how to recruit this student or who should be the adviser.
+If the student mentioned a faculty member in their SOP, we might ask that faculty member if they are interested in the student.
+
+\subsection{Do adcom members read all applications?}\label{sec:read-all}
+\subsectioninfo{Adcom members only review applications assigned to them (typically matching their expertise) and rarely get involved in other applications}
+
+While adcom members (in fact, any faculty) can view any application, we only review those that are assigned to us, which are already too many.
+Adcom chair will assign applications to reviewers based on their expertise (e.g., if a student says they want to do SE or interested in working with me), and reviewers will only review those applications.
+Occasionally we might look at other applications, e.g., if we know the student or have some special circumstances (e.g., if that student contacted me, I know that student, they are from Vietnam, etc).
+However, even if we look at them, we usually do not get involved in their evaluation.
 
@@
 \subsection{Why did you get rejected?}\label{sec:why-rejected}
-
-\sectioninfo{You aim too high, are overqualified, or even because you applied to AI/ML, a super competitive field in recent years with many applicants.}
+\subsectioninfo{You aim too high, are overqualified, or even because you applied to AI/ML, a super competitive field in recent years with many applicants.}
 
@@
 \subsection{Timeline: Why are rejection letters sent so late?}\label{sec:late-rejection}
-\sectioninfo{Why rejection letters are sent so late? Grad programs often wait for the accepted students to make their decisions, typically by April 15, before sending out rejection letters.}
+\subsectioninfo{Grad programs often wait for the accepted students to make their decisions, typically by April 15, before sending out rejection letters.}
 
@@
 If you really want advice, ask your professors, collaborators, ref writers, or those who have previously applied.
+
+\subsection{Increasing your admission chance}\label{sec:improve-your-chance}
+\subsectioninfo{You can improve your profile by being unique and standing out.}
+
+\begin{center}
+  \includegraphics[scale=0.2]{files/alpinist-climbing-peak-mountain-comic-hand-drawn-vector-illustration.jpg}
+\end{center}
+
+Given the high number of quality applicants and a limited number of spots, in addition to having a good application profile, you want to show something that makes you \textbf{stand out}.
+For example, even if you do not have research experience, you can talk about your personal projects, as long as they can help show you can do research.
+For example, if you have an open-source project that has lots of stars in Github, then you should mention it.
+If you often write technical, research-like blogs with many viewers, talk about them too.
+
+There are other things you might not think are important but can make you stand out.
+For example, if you have a strong background in a non-CS field but can be integrated with CS, e.g., you have a degree or background in \emph{dance} or \emph{music} and want to integrate them with CS, do talk about it.
+Are you a female or a minority in CS (\autoref{sec:urm})?
+Do you participate in outreach activities that help increase diversity and inclusion in CS?
+Diversity is highly valued in CS programs in the US, and experiences in these areas can make you stand out and get noticed from reviewers.
+
+\begin{commentbox}[Vu:]
+In his \href{https://matt.might.net/articles/how-to-apply-and-get-in-to-graduate-school-in-science-mathematics-engineering-or-computer-science/}{post}, Matt Might was initially unsure about an application.
+However, upon learning that the applicant had led a 100km hike in the Himalayas, he decided to accept the applicant.
+This is a good example of being \emph{stand out}, and I would also advocate for that student as this shows they have the persistence and determination required for research.
+\end{commentbox}
+
+% It is worth noting that in many cases,  an application could be rejected by most reviewers, however have a "champion", who doesn't even have to be in the Admission comittee, that strongly advocate for your profile.
 
@@
 
-One confusion students often have is whether they will get paid over the summer.  The answer is that it depends on the funding source.
-First, remember that an academic year (AY), which consists of Fall and Spring semesters, is typically 9 months, so your stipend is for 9 months (and many places allow you to spread it over 12 months).
-
-If your funding source is TA (\autoref{sec:ta}), you typically do not get paid over the summer as there are few courses.
-Some CS departments offer some summer funding, but it is typically not guaranteed and might not be a lot.
+One confusion students often have is whether they will get paid over the summer.  This depends on the funding source.
+First, recall that an academic year (AY), which consists of Fall and Spring semesters, is typically 9 months, so your stipend is for 9 months (and many places allow you to spread it over 12 months).
+
+If your funding source is TA (\autoref{sec:ta}), you typically do not get paid over the summer, which has few courses.
+Some CS departments offer summer funding, but it is not guaranteed and might not be a lot.
 For example, at Mason, we offer summer funding for 1st-year PhD TA students.
@@
 However, funding is never guaranteed, e.g., the prof might not get the grant.
+
+\begin{commentbox}[Vu:]
 For my students, I have been fortunate to have funding to support them over the summer.
@@
 I prioritize summer funding for my students because Mason has very good TA resources so they never have to worry about funding during the AY.
+\end{commentbox}
 
@@
 
-\section{Increasing your admission chance}\label{sec:improve-your-chance}
-
-\sectioninfo{Increasing admission chance by being unique and standing out.}
-
-\begin{center}
-  \includegraphics[scale=0.2]{files/alpinist-climbing-peak-mountain-comic-hand-drawn-vector-illustration.jpg}
-\end{center}
-
-Given the high number of quality applicants and a limited number of spots, in addition to having a good application profile, you want to show something that makes you \textbf{stand out}.
-For example, even if you do not have research experience, you can talk about your personal projects, as long as they can help show you can do research.
-For example, if you have an open-source project that has lots of stars in Github, then you should mention it.
-If you often write technical, research-like blogs with many viewers, talk about them too.
-
-There are other things you might not think are important but can make you stand out.
-For example, if you have a strong background in a non-CS field but can be integrated with CS, e.g., you have a degree or background in \emph{dance} or \emph{music} and want to integrate them with CS, do talk about it.
-Are you a female or a minority in CS (\autoref{sec:urm})?
-Do you participate in outreach activities that help increase diversity and inclusion in CS?
-Diversity is highly valued in CS programs in the US, and experiences in these areas can make you stand out and get noticed from reviewers.
-
-\begin{commentbox}[Vu:]
-In his \href{https://matt.might.net/articles/how-to-apply-and-get-in-to-graduate-school-in-science-mathematics-engineering-or-computer-science/}{post}, Matt Might was initially unsure about an application.
-However, upon learning that the applicant had led a 100km hike in the Himalayas, he decided to accept the applicant.
-This is a good example of being \emph{stand out}, and I would also advocate for that student as this shows they have the persistence and determination required for research.
-\end{commentbox}
-
-% It is worth noting that in many cases,  an application could be rejected by most reviewers, however have a "champion", who doesn't even have to be in the Admission comittee, that strongly advocate for your profile.
-
 \section{How long does it take to evaluate an application?}\label{sec:ievaluate}
@@
 Note that while the system has other decision options, eg., admit without funding, provisional admission (e.g., if they need to take some courses), I rarely use them, simply because we already have too many good students and do not have the resources to support them.
-
-\section{How are decisions made?}\label{sec:how-decisions}
-\sectioninfo{Even if all adcom reviewers recommend acceptance, the application can still be rejected. Vice versa, if all reviewers think the application is weak, the student might still be admitted.}
-
-After reviewers have evaluated an application, adcom chair will review all evaluations, look at entered notes, and ask reviewers to discuss and resolve discrepancies to reach a consensus (e.g., a reviewer wants to accept but the other wants to reject).
-Typically, the decision is made entirely by the reviewers; there is no involvement from the adcom chair, department chair, or others.
-In most cases adcom members, even those reviewing the same application, make decisions independently and do not talk to each other (just a common practice to avoid biasing).
-In some rare cases we might (\autoref{sec:adcom-discuss}).
-
-Even if \emph{all} reviewers recommend acceptance, the application is not automatically accepted, especially if no faculty is willing to advise the student.
-For example, if the student is interested in a research area that no faculty is working on, or an area where no faculty is taking new students (e.g., AI/ML where faculty likely already has many students) then the student will not be admitted.
-This is increasingly common as the number of applicants increases significantly in recent years in CS.
-
-However, if the student has contacted a faculty member and that faculty member is interested in the student and has made this known to the admission committee, then the student is likely to be admitted, \emph{even if the faculty does not review the application or has funding} (thus the importance of contacting faculty).
-Similarly, if the student mentioned a faculty member in their SOP, adcom might ask that faculty member to look at their application and if they are interested in the student.
-Even if the student has a weak profile (but still passes the minimum requirement from the university), they might be admitted if a faculty member is willing to take them.
-Adcom members, especially in the US, are very reluctant to go against the faculty's decision (e.g., if a faculty wants to admit a student, we are not going to reject them).
-
-\section{Do adcom members talk to each other?}\label{sec:adcom-discuss}
-\sectioninfo{Sometimes adcom members discuss applicants, but in most cases they make independent decisions.}
-
-However, in some rare cases we do talk to each other.
-For example, when there are discrepancies in evaluations, in which case the adcom chair will ask reviewers to discuss the application to reach a consensus.
-We also talk to each other for interesting or strong applications, e.g., how to recruit this student or who should be the adviser.
-If the student mentioned a faculty member in their SOP, we might ask that faculty member if they are interested in the student.
-
-\section{Do adcom members read all applications?}\label{sec:read-all}
-\sectioninfo{Adcom members only review applications assigned to them (typically matching their expertise) and rarely get involved in other applications}
-
-While adcom members (in fact, any faculty) can view any application, we only review those that are assigned to us, which are already too many.
-Adcom chair will assign applications to reviewers based on their expertise (e.g., if a student says they want to do SE or interested in working with me), and reviewers will only review those applications.
-Occasionally we might look at other applications, e.g., if we know the student or have some special circumstances (e.g., if that student contacted me, I know that student, they are from Vietnam, etc).
-However, even if we look at them, we usually do not get involved in their evaluation.
 
@@
 %\autoref{tab:ranking} lists .
-
-\chapter*{History and Acknowledgement}\label{sec:ack}
-
+\chapter{History and Acknowledgement}\label{sec:ack}
 \paragraph{History} This document was conceived during a lunch with a faculty at Mason.

### 2024-10-02 demystify.tex
reognizing

@@
 
+\section{Who is this document for?}
+\sectioninfo{This document is intended for international students from smaller countries and domestic students from less well-known universities.}
+
+The guideline can be useful for all students, but it is primarily intended for \textbf{international students} from \textbf{smaller countries and less well-known universities} (or domestic students from US universities with no PhD programs or limited research opportunities).
+
+Students from top schools with strong research cultures and experiences might already know most of the information given here.
+My goal is thus to level the playing field by providing information that is not readily available to less privileged students. 
+I hope to encourage more students with such backgrounds to apply to CS PhD programs in the US.
+
+While this document may not fully meet the needs of students aiming for top-10 schools (\autoref{sec:ranking}), which typically have more stringent requirements (e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU), it should be valuable for those targeting other reputable R1 institutions.
+It is certainly useful for students aiming for Mason!
+
+\section{How long to complete the CS PhD program?}\label{sec:time}
+\sectioninfo{About 5--7 years in the US.}
+
+Typically, it takes 5--7 years for CS PhD in the US.
+This can be longer than CS PhDs in other countries, which might require MS first (recall from \autoref{sec:msrequirement} that CS PhD programs in the US do not require MS and you can get MS along the way to PhD).
+Within these 5--7 years, CS PhD students often take a ``leave of absence'' for 1--2 semesters to do internships at companies and research labs.
+
+\begin{center}
+  \includegraphics[scale=0.5]{files/c4a.png}
+\end{center}
+
+The first two years you typically take coursework (somewhat equivalent to an MS study), find an adviser, and learn how to do research.
+The next 2--3 years you focus on your research, form a dissertation topic, and get results published.
+The last 1--2 years you continue to publish, write and defend your dissertation, and look for a job.
+In many cases you might take a summer or two off to do an internship to get additional research opportunities.
+The PhDComics figure above shows the ``ambition'' level of a PhD student over their years of study (they miss the 6--7th year when the ambition is \emph{``Just let me graduate''}).
+
+% \begin{commentbox}[Vu:]
+%     I start my PhD with an MS, and it took me 7 years (Fall'2007-- Fall'14). I spent half a year doing an internship at the Naval Research Lab.
+%   \end{commentbox}
+
+\section{My undergrad was not in CS or related areas}\label{sec:non-stem}
+\sectioninfo{You can successfully apply to CS PhD even if you have non-CS background.}
+
+You still can apply to PhD in CS \emph{as long as} you can demonstrate you are ready for it through your background, research experience, LoRs, statements, etc.
+You might be even able to leverage this to make your profile stand out as mentioned in \autoref{sec:improve-your-chance}.
+
+A main question adcom has for a non-CS or non-STEM student is if you have the sufficient technical background obtained through core CS courses.
+So you need to show that you have such knowledge through your coursework, projects, or research.
+For example, if you have taken a course on Algorithms, even online ones like Coursera, you can talk about it in your SOP.
+If you have done a project that requires knowledge of OS or have a professional certification (e.g., A+) through work, you can talk about it.
+If you have done research that requires knowledge of Discrete Maths, you can talk about it.
+You can also ask your LoR writers to talk about your technical background.
+In summary, in your application, convince us that you have the background to do CS PhD research.
+
+\paragraph{Core CS topics} Common CS knowledge that you should know are:
+\begin{itemize}
+  \item Programming Foundation (e.g., programming concepts in some modern languages)
+  \item Discrete Math (e.g., logic, set theory, proof techniques)
+  \item Data Structures and Algorithms (e.g., linked lists, trees, sorting, searching)
+  \item Computer OS or Systems (e.g., memory management, file systems, processes)
+\end{itemize}
+
+In short, you \emph{do not need} to formally take CS courses, you just need to show that you have this essential knowledge, e.g., through the mentioned ways.
+Many universities are well aware that incoming graduate students might not have all the technical background, so they often have a \emph{``bridge''} courses to help students catch up.
+For example, Mason has four bridge courses corresponding to the four core areas above that incoming students can take to catch up on their CS knowledge.
+
+\begin{commentbox}[Vu:]
+  I would advocate for a non-STEM student who shows that they have a strong drive for CS by studying core CS knowledge through various channels (e.g., self-study through online courses, projects, etc.).
+I have seen many students with non-CS backgrounds who are very successful in CS PhD.
+I also have seen many students with CS background who are not successful in CS PhD.
+So it is not about your background, it is about your drive and passion for CS research.
+\end{commentbox}
+
+\section{Is an MS required for admission to PhD in CS?}\label{sec:msrequirement}
+\sectioninfo{You do not need an MS to do PhD in CS.}
+
+No.
+In fact, students can get MS degree ``along the way'' to PhD, e.g., after finishing the 2-year course work.
+However, MS can help admission if it gives research experience or is from a more well-known school than your undergrad institution (\autoref{sec:your-school}).
+
+If you have an MS then some course work \emph{might be} transferred for course credits, which \emph{might} save some time. But don't count on this.
+
+\begin{commentbox}[Vu:]
+    I start my PhD with an MS in CS from a US university.
+I found that the MS helped me a lot in research, but not with course work and I had to retake many courses.
+In general, don't worry if you don't have an MS.
+But also don't feel that you wasted your time if you have an MS, as it can help you in research.
+  \end{commentbox}
+
+\section{PhD in the US vs. Other Countries}\label{sec:non-us-differences}
+\sectioninfo{Among several differences, CS PhD in the US does not require an MS degree but has a longer PhD study time.}
+
+\begin{table}
+\caption{Comparison of US and Other PhD Programs}\label{tab:us-vs-other}
+\begin{tabular}{c|c|c}
+\toprule
+\textbf{Aspect} & \textbf{US PhD Programs} & \textbf{Other Countries} \\
+\midrule
+Duration & 5-7 years & 3-5 years \\
+MS Required & Not required & Often required \\
+Coursework Required & Yes (first 2 years) & No \\
+Research Proposal Required & No & Yes (in some countries) \\
+Academic (Faculty) Job & Direct & Postdoc \\
+Work Life Balance & Less & More \\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\autoref{tab:us-vs-other} summarizes the main differences between CS PhD in the US and other countries. % This is based on my experience and what I heard from others who did PhD in other countries.
+
+\emph{MS requirement and PhD duration}:  CS PhD programs in the US do not require an MS degree (as mentioned in \autoref{sec:msrequirement} and \autoref{sec:time}).
+In contrast, many other countries require having an MS degree before joining a PhD program.
+This means that US PhD programs are longer (5--7 years, 2 of which are coursework) than other countries (3--4 years, no coursework).
+
+\emph{Project proposal}: in many countries, you have to choose a project and adviser \emph{during} the application process (e.g., you write a proposal to a potential adviser).
+But this allows you to start your research right from the beginning.
+
+In the US, you often start your PhD without an adviser or project and find them later.
+You will have two initial years to explore and find an adviser and research topic.
+After that you explore projects with that adviser before settling on one (can take a couple of years).
+
+\emph{Course work}: as mentioned above in the US you will spend the first couple of years taking classes and exploring potential adviser and research topics. 
+After that, you have to pass a series of exams during your PhD, e.g., qualifying exam, comprehensive exam, thesis proposal defense\footnote{The word ``ABD'' (all but dissertation) is used in the US to refer to a PhD candidate who has finished all course work and exams and only needs to write and defend their dissertation.}.
+
+In other countries, you start your research right away, e.g., you immediately work on the research project you proposed with the adviser you chose.
+Moreover, you might not have exams like those in the US or only have to do a few of them.
+
+\emph{Funding}:  In many countries, funding comes from the university or the gov't.
+This funding often has a fixed duration (e.g., 3 or 4 years).
+In the US (\autoref{sec:funding}), funding (e.g., RA) comes directly from your adviser (no fixed duration).
+There are also fewer TA opportunities in European universities compared to the US.
+
+\emph{Academic position after PhD}: In other countries, PhD graduates interested in academia typically apply for additional research appointments, i.e., postdocs, and then consider faculty positions.
+
+In the US, PhD graduates often apply directly for faculty positions. Postdoc for US graduates is no longer a popular option as it was before.
+
+\emph{Work-life balance}: PhD students in the US are often said to be overworked compared to other countries, e.g., in Europe.
+This is partly due to the longer PhD program and that US PhD students are often paid through TA, which requires them to do TA in addition to their own research.
+In contrast, PhD students in other countries are often paid through fellowships, which might not require doing TA.
+
+%% Vu: *what's a PhD?*  This [[https://matt.might.net/articles/phd-school-in-pictures/][series of pictures]] from [[https://matt.might.net][Matt Might]] illustrates what a PhD means.
+%% \end{commentbox}
+
 % \section{Timeline}
@@
 % TODO
-
-\section{How is Your Application Evaluated?}\label{sec:evalapps}
+\subsubsection{Why CS depts do not waive the application fee?}
+
+Some universities do waive, e.g., Rice does for all PhD applicants and many universities for domestic students (\autoref{sec:domestic-students}), however most do not for international students.
+In any case, the application fee is typically a requirement of the university.
+Individual departments and programs do not have the flexibility to waive the application fee, even if they want to.
+
+\begin{commentbox}{Vu:}
+In my opinion, requiring applicants to pay the fee helps ensure their seriousness, as it filters out non-serious candidates.
+Also, most CS programs already receive many applications and would be overwhelmed if the application process were free.
+Note that if you have financial difficulties, you can ask the department for a waiver, but this is typically only granted in exceptional cases.
+\end{commentbox}
+
+\chapter{How is Your Application Evaluated?}\label{sec:evalapps}
 
@@
 
-\subsection*{Admission Committee (adcom)}\label{sec:adcom}
+\section{Admission Committee (adcom)}\label{sec:adcom}
 
@@
 
-\section{Who is this document for?}
-\sectioninfo{This document is mainly intended for international students from smaller countries and domestic students from less well-known universities.}
-
-The guideline can be useful for all students, but it is primarily intended for \textbf{international students} from \textbf{smaller countries and less well-known universities} (or domestic students from US universities with no PhD programs or limited research opportunities).
-
-Students from top schools with strong research cultures and experiences might already know most of the information given here.
-My goal is thus to level the playing field by providing information that is not readily available to less privileged students. 
-I hope to encourage more students with such backgrounds to apply to CS PhD programs in the US.
-
-While this document may not fully meet the needs of students aiming for top-10 schools (\autoref{sec:ranking}), which typically have more stringent requirements (e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU), it should be valuable for those targeting other reputable R1 institutions.
-It is certainly useful for students aiming for Mason!
-
-%% Vu: *what's a PhD?*  This [[https://matt.might.net/articles/phd-school-in-pictures/][series of pictures]] from [[https://matt.might.net][Matt Might]] illustrates what a PhD means.
-%% \end{commentbox}
-
 \section*{Additional Resources}
@@
 
-\section{Miscs}
-
-\subsubsection{Why CS depts do not waive the application fee?}
-
-Some universities do waive, e.g., Rice does for all PhD applicants and many universities for domestic students (\autoref{sec:domestic-students}), however most do not for international students.
-In any case, the application fee is typically a requirement of the university.
-Individual departments and programs do not have the flexibility to waive the application fee, even if they want to.
-
-\begin{commentbox}{Vu:}
-In my opinion, requiring applicants to pay the fee helps ensure their seriousness, as it filters out non-serious candidates.
-Also, most CS programs already receive many applications and would be overwhelmed if the application process were free.
-Note that if you have financial difficulties, you can ask the department for a waiver, but this is typically only granted in exceptional cases.
-\end{commentbox}
-\subsection{How are decisions made?}\label{sec:how-decisions}
-\subsectioninfo{Even if all adcom reviewers recommend acceptance, the application can still be rejected.
-Vice versa, if all reviewers think the application is weak, the student might still be admitted.}
+\section{How are decisions made?}\label{sec:how-decisions}
+\sectioninfo{Even if all adcom reviewers recommend acceptance, the application can still be rejected. Vice versa, if all reviewers think the application is weak, the student might still be admitted.}
 
@@
 
-\subsection{Do adcom members talk to each other?}\label{sec:adcom-discuss}
-\subsectioninfo{Sometimes adcom members discuss applicants, but in most cases they make independent decisions.}
+\section{Do adcom members talk to each other?}\label{sec:adcom-discuss}
+\sectioninfo{Sometimes adcom members discuss applicants, but in most cases they make independent decisions.}
 
@@
 
-\subsection{Do adcom members read all applications?}\label{sec:read-all}
-\subsectioninfo{Adcom members only review applications assigned to them (typically matching their expertise) and rarely get involved in other applications}
+\section{Do adcom members read all applications?}\label{sec:read-all}
+\sectioninfo{Adcom members only review applications assigned to them (typically matching their expertise) and rarely get involved in other applications}
 
@@
 
-\subsection{Timeline}
-
-The timeline for interviews and notification letters varies.
+\paragraph{When do interviews happen?}
+
+The timeline for interviews varies.
 Faculty set up interviews based on their (busy and erratic) schedule.
@@
 Some schools do not do interviews at all (\autoref{sec:no-interview}).
-
-Some schools send out admission letters in batches, some do rolling admission, and some do not send anything out (e.g., you're on their waitlist).
-In general, you should hear back from most schools by mid-March (but some schools might take longer!).
-Note that rejection letters might come very late (see \autoref{sec:late-rejection}).
-
-In short, there is not much you can do about this other than to be patient and wait.
-Also do not send emails asking about interviews or status (unless you have interviewed specifically with someone then you can ask that person for status updates and others \autoref{sec:accept-postpone-decline}).
 
@@
 
+\subsection{Notification Timeline: Why rejection letters are sent so late?}\label{sec:late-rejection}
+\subsectioninfo{Grad programs often wait for the accepted students to make their decisions, typically by April 15, before sending out rejection letters.}
+
+%"Just reject me already!" is a common sentiment among applicants.
+Indeed, school will first send out admission offers to the top candidates.
+They do not send out rejection letters because there is still a chance that some of the top candidates will decline the offer.
+If they do, then the school will go to the next set of candidates and send out offers to them.
+This process continues until all spots are filled.
+This is why you might not hear back until late in the admission cycle.
+
+Some schools send out admission letters in batches, some do rolling admission, and some do not send anything out (e.g., you're on their waitlist).
+You should hear back from most schools by mid-March (though rejection can come out very late).
+
+Not much you can do other than to be patient and wait.
+Do not send emails asking about interviews or status (unless you have interviewed specifically with someone (\autoref{sec:interview}) then you can ask that person for status updates and other questions (\autoref{sec:accept-postpone-decline}).
+
+\paragraph{Acceptance Letters} Universities prioritize sending out acceptance letters first. This allows the admitted students to make decisions and plan for their studies.
+
+Some universities have rolling admissions, meaning they send out acceptance letters as they review applications (e.g., some universities sent out letters as soon as early February).
+Others have a specific date when they send out the first round of acceptance letters.
+
+\paragraph{Response Deadlines} Accepted students are usually given a deadline to make decisions on their offers, often around \textbf{April 15th} in the US.
+This is a standard deadline for PhD programs due to the Council of Graduate Schools (CGS) Resolution (\href{https://cgsnet.org/wp-content/uploads/2024/01/CGS_April15_Resolution_Jan312024.pdf}{see more}.
+After this date, CS programs can gauge how many slots remain unfilled.
+
+\paragraph{Waitlist} Most CS programs have a limited number of slots for PhD students, and thus put many good students on a waitlist.
+If accepted students decline the offer, then offers are sent to students on the waitlist.
+So if you see people getting accepted, that does not mean you are out yet.
+
+\paragraph{Rejection Letters} Schools typically start sending out rejection letters to remaining applicants \emph{after they have finalized their admissions decisions}.
+Thus, rejection letters are often sent out late (e.g., after April 15th or even much later).
+Not much you can do here.
+You can try to contact the school to ask about your status, but they might not reply, they might say they are still reviewing applications, or give you inaccurate information (e.g., you will hear in two weeks).
+In short, you just have to be patient and wait (however, beware that some schools do not send out rejection letters at all).
+
 % \section{Preparing and Tracking Applications}
@@
 \chapter{Getting Admitted}\label{sec:accepted}
-
 \chapterinfo{Congrats! Now it is your turn to evaluate the school!
@@
 In the meantime, you can work on improving your profile, e.g., get more research experiences, publish more papers, improve your connections for stronger LoRs, etc.
-See rejection reasons (\autoref{sec:why-rejected}) and timeline (\autoref{sec:late-rejection}), and additional advice to improve your chance next time (\autoref{sec:improve-your-chance}).
+See rejection reasons (\autoref{sec:why-rejected}) and additional advice to improve your chance next time (\autoref{sec:improve-your-chance}).
 
@@
 
-\subsection{Why did you get rejected?}\label{sec:why-rejected}
-\subsectioninfo{You aim too high, are overqualified, or even because you applied to AI/ML, a super competitive field in recent years with many applicants.}
-
-Many students lament that they get no interviews or are rejected (with strong profiles) and that the admission process seems random.
-However, while it is true that the process is not perfect, it is not random.
-Other than not having a strong profile (e.g., research potential, GPA, LoRs, SOP, poor interviews), here are some other reasons for rejections.
-
-\paragraph{Not a Good fit and Bad luck}  You could have an excellent profile, but if you are interested in a research area that the program does not have, you will not be admitted.
-Similarly, if no faculty is willing to advise you (e.g., they are on sabbatical or personal leave, do not have sufficient funding, or already have too many students), you will not be admitted.
-This is actually good for you, as you don't want to be in a program where no one can advise you.
-
-Related to this is that you just have bad luck and apply at the wrong time.
-For example, since 2024 there has been a huge surge in students interested in AI and NLP (thanks, ChatGPT!).
-Consequently, AI/NLP faculty might be overwhelmed and cannot consider many candidates, even those with excellent profiles.
-
-\paragraph{You aim too high} This is a very common reason: you have applied to schools that are too high for your profile.
-While being ambitious is good, you should understand how PhD admission works (e.g., by reading this document and realizing things such as having a good GPA doesn't mean much to top programs) and be realistic!
-
-Before applying, you should talk to your professors and ask them to give you an honest opinion on where you should apply.
-To increase their chances, many students apply to a range of schools, including ``safety'' ones.
-
-\paragraph{Overqualified or Lack Interest}  This might be paradoxical and is the opposite of aiming too high.
-However if adcom believes that you are likely to get admitted and go to a better program, they might not be so enthusiastic to admit you and want to save the spot for someone else.
-
-A related reason is that you did not show enough interest in the program.
-For example, your SOP did not mention any faculty or research area at the program, you did not respond to professors interested in interviewing you (which might be considered unprofessional and burn bridges), or during the interview you showed little knowledge, interest, or enthusiasm (because they are your safety schools) in the program or profs.
-
-\paragraph{Low English exam scores (e.g., IELTS or TOEFL)}  Profs.
-might not care much about these, but the college or the university often sets a minimum that you need to pass to be considered, especially for TA funding (e.g., you need to communicate well with students as a TA).
-Thus, while profs.
-are willing to argue for your case, they might be reluctant to go against the requirement of the university on standard exams.
-
-Note that English exams are not required for all students, e.g., domestic students and those from certain countries are waived from taking these exams.
-For example, Mason waives English exams for students from these \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-Mason#standard-tests-waiver-eligible-countries}{countries}.
-
-\paragraph{Red flags} Various types of red flags can cause concern to the adcom (e.g., too many STEM courses with low or withdrawn grades, plagiarism, cheating, or other academic dishonesty).
-Another one is that you have a history of jumping from one program to another (e.g., you have been in multiple different programs in a short time).
-Adcom members might have contacts in other programs and heard about your stories, or your LoRs might mention these.
-
-If you think you have these issues, you should address them in your SOP or ask your letter writers to do so.
-In general, these things are rare, but they do happen and cause concern to the adcom.
-
-\subsection{Timeline: Why are rejection letters sent so late?}\label{sec:late-rejection}
-\subsectioninfo{Grad programs often wait for the accepted students to make their decisions, typically by April 15, before sending out rejection letters.}
-
-%"Just reject me already!" is a common sentiment among applicants.
-Indeed, school will first send out admission offers to the top candidates.
-They do not send out rejection letters because there is still a chance that some of the top candidates will decline the offer.
-If they do, then the school will go to the next set of candidates and send out offers to them.
-This process continues until all spots are filled.
-This is why you might not hear back until late in the admission cycle.
-
-\paragraph{Acceptance Letters} Universities prioritize sending out acceptance letters first. This allows the admitted students to make decisions and plan for their studies.
-
-Some universities have rolling admissions, meaning they send out acceptance letters as they review applications (e.g., some universities sent out letters as soon as early February).
-Others have a specific date when they send out the first round of acceptance letters.
-
-\paragraph{Response Deadlines} Accepted students are usually given a deadline to make decisions on their offers, often around \textbf{April 15th} in the US.
-This is a standard deadline for PhD programs due to the Council of Graduate Schools (CGS) Resolution (\href{https://cgsnet.org/wp-content/uploads/2024/01/CGS_April15_Resolution_Jan312024.pdf}{see more}.
-After this date, CS programs can gauge how many slots remain unfilled.
-
-\paragraph{Waitlist} Most CS programs have a limited number of slots for PhD students, and thus put many good students on a waitlist.
-If accepted students decline the offer, then offers are sent to students on the waitlist.
-So if you see people getting accepted, that does not mean you are out yet.
-
-\paragraph{Rejection Letters} Schools typically start sending out rejection letters to remaining applicants \emph{after they have finalized their admissions decisions}.
-Thus, rejection letters are often sent out late (e.g., after April 15th or even much later).
-Not much you can do here, you can try to contact the school to ask about your status, but they might not reply, they might say they are still reviewing applications, or give you inaccurate information (e.g., you will hear in two weeks).
-In short, you just have to be patient and wait (however, beware that some schools do not send out rejection letters at all).
-
-\subsubsection{Should I ask for feedback?}
+\paragraph{Should I ask for feedback?}
 No, don't bother.
@@
 
+\subsection{Why did you get rejected?}\label{sec:why-rejected}
+\subsectioninfo{You aim too high, are overqualified, or even because you applied to AI/ML, a super competitive field in recent years with many applicants.}
+
+Many students lament that they get no interviews or are rejected (with strong profiles) and that the admission process seems random.
+However, while it is true that the process is not perfect, it is not random.
+Other than not having a strong profile (e.g., research potential, GPA, LoRs, SOP, poor interviews), here are some other reasons for rejections.
+
+\paragraph{Not a Good fit and Bad luck}  You could have an excellent profile, but if you are interested in a research area that the program does not have, you will not be admitted.
+Similarly, if no faculty is willing to advise you (e.g., they are on sabbatical or personal leave, do not have sufficient funding, or already have too many students), you will not be admitted.
+This is actually good for you, as you don't want to be in a program where no one can advise you.
+
+Related to this is that you just have bad luck and apply at the wrong time.
+For example, since 2024 there has been a huge surge in students interested in AI and NLP (thanks, ChatGPT!).
+Consequently, AI/NLP faculty might be overwhelmed and cannot consider many candidates, even those with excellent profiles.
+
+\paragraph{You aim too high} This is a very common reason: you have applied to schools that are too high for your profile.
+While being ambitious is good, you should understand how PhD admission works (e.g., by reading this document and realizing things such as having a good GPA doesn't mean much to top programs) and be realistic!
+
+Before applying, you should talk to your professors and ask them to give you an honest opinion on where you should apply.
+To increase their chances, many students apply to a range of schools, including ``safety'' ones.
+
+\paragraph{Overqualified or Lack Interest}  This might be paradoxical and is the opposite of aiming too high.
+However if adcom believes that you are likely to get admitted and go to a better program, they might not be so enthusiastic to admit you and want to save the spot for someone else.
+
+A related reason is that you did not show enough interest in the program.
+For example, your SOP did not mention any faculty or research area at the program, you did not respond to professors interested in interviewing you (which might be considered unprofessional and burn bridges), or during the interview you showed little knowledge, interest, or enthusiasm (because they are your safety schools) in the program or profs.
+
+\paragraph{Low English exam scores (e.g., IELTS or TOEFL)}  Profs.
+might not care much about these, but the college or the university often sets a minimum that you need to pass to be considered, especially for TA funding (e.g., you need to communicate well with students as a TA).
+Thus, while profs.
+are willing to argue for your case, they might be reluctant to go against the requirement of the university.
+
+Note that English exams are not required for all students, e.g., domestic students and those from certain countries are waived from taking these exams.
+For example, Mason waives English exams for students from these \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-Mason#standard-tests-waiver-eligible-countries}{countries}.
+
+\paragraph{Red flags} Various types of red flags can cause concern to the adcom (e.g., too many STEM courses with low or withdrawn grades, plagiarism, cheating, or other academic dishonesty).
+Another one is that you have a history of jumping from one program to another (e.g., you have been in multiple different programs in a short time).
+Adcom members might have contacts in other programs and heard about your stories, or your LoRs might mention these.
+
+If you think you have these issues, you should address them in your SOP or ask your letter writers to do so.
+In general, these things are rare, but they do happen and cause concern to the adcom.
+
 \subsection{Increasing your admission chance}\label{sec:improve-your-chance}
@@
 \chapter{Funding}\label{sec:funding}
-
 \chapterinfo{TAs, RAs, and fellowships are main funding sources for PhDs.
@@
 %https://cs.brown.edu/degrees/doctoral/applications/helpful-resources-applying-computer-science-phd-programs/
-
-\section{My undergrad was not in CS or related areas}\label{sec:non-stem}
-\sectioninfo{You can successfully apply to CS PhD even if you have non-CS background.}
-
-You still can apply to PhD in CS \emph{as long as} you can demonstrate you are ready for it through your background, research experience, LoRs, statements, etc.
-You might be even able to leverage this to make your profile stand out as mentioned in \autoref{sec:improve-your-chance}.
-
-A main question adcom has for a non-CS or non-STEM student is if you have the sufficient technical background obtained through core CS courses.
-So you need to show that you have such knowledge through your coursework, projects, or research.
-For example, if you have taken a course on Algorithms, even online ones like Coursera, you can talk about it in your SOP.
-If you have done a project that requires knowledge of OS or have a professional certification (e.g., A+) through work, you can talk about it.
-If you have done research that requires knowledge of Discrete Maths, you can talk about it.
-You can also ask your LoR writers to talk about your technical background.
-In summary, in your application, convince us that you have the background to do CS PhD research.
-
-\paragraph{Core CS topics} Common CS knowledge that you should know are:
-\begin{itemize}
-  \item Programming Foundation (e.g., programming concepts in some modern languages)
-  \item Discrete Math (e.g., logic, set theory, proof techniques)
-  \item Data Structures and Algorithms (e.g., linked lists, trees, sorting, searching)
-  \item Computer OS or Systems (e.g., memory management, file systems, processes)
-\end{itemize}
-
-In short, you \emph{do not need} to formally take CS courses, you just need to show that you have this essential knowledge, e.g., through the mentioned ways.
-Many universities are well aware that incoming graduate students might not have all the technical background, so they often have a \emph{``bridge''} courses to help students catch up.
-For example, Mason has four bridge courses corresponding to the four core areas above that incoming students can take to catch up on their CS knowledge.
-
-\begin{commentbox}[Vu:]
-  I would advocate for a non-STEM student who shows that they have a strong drive for CS by studying core CS knowledge through various channels (e.g., self-study through online courses, projects, etc.).
-I have seen many students with non-CS backgrounds who are very successful in CS PhD.
-I also have seen many students with CS background who are not successful in CS PhD.
-So it is not about your background, it is about your drive and passion for CS research.
-\end{commentbox}
-
-\section{Is an MS required for admission to PhD in CS?}\label{sec:msrequirement}
-\sectioninfo{You do not need an MS to do PhD in CS.}
-
-No.
-In fact, students can get MS degree ``along the way'' to PhD, e.g., after finishing the 2-year course work.
-However, MS can help admission if it gives research experience or is from a more well-known school than your undergrad institution (\autoref{sec:your-school}).
-
-If you have an MS then some course work \emph{might be} transferred for course credits, which \emph{might} save some time. But don't count on this.
-
-\begin{commentbox}[Vu:]
-    I start my PhD with an MS in CS from a US university.
-I found that the MS helped me a lot in research, but not with course work and I had to retake many courses.
-In general, don't worry if you don't have an MS.
-But also don't feel that you wasted your time if you have an MS, as it can help you in research.
-  \end{commentbox}
-
-\section{How long to complete the CS PhD program?}\label{sec:time}
-\sectioninfo{About 5--7 years in the US.}
-
-Typically, it takes 5--7 years for CS PhD in the US.
-This can be longer than CS PhDs in other countries, which might require MS first (recall from \autoref{sec:msrequirement} that CS PhD programs in the US do not require MS and you can get MS along the way to PhD).
-Within these 5--7 years, CS PhD students often take a ``leave of absence'' for 1--2 semesters to do internships at companies and research labs.
-
-\begin{center}
-  \includegraphics[scale=0.5]{files/c4a.png}
-\end{center}
-
-The first two years you typically take coursework (somewhat equivalent to an MS study), find an adviser, and learn how to do research.
-The next 2--3 years you focus on your research, form a dissertation topic, and get results published.
-The last 1--2 years you continue to publish, write and defend your dissertation, and look for a job.
-In many cases you might take a summer or two off to do an internship to get additional research opportunities.
-The PhDComics figure above shows the ``ambition'' level of a PhD student over their years of study (they miss the 6--7th year when the ambition is \emph{``Just let me graduate''}).
-
-% \begin{commentbox}[Vu:]
-%     I start my PhD with an MS, and it took me 7 years (Fall'2007-- Fall'14). I spent half a year doing an internship at the Naval Research Lab.
-%   \end{commentbox}
-
-\section{PhD in the US vs. Other Countries}\label{sec:non-us-differences}
-\sectioninfo{Among several differences, CS PhD in the US does not require an MS degree but has a longer PhD study time.}
-
-\begin{table}
-\caption{Comparison of US and Other PhD Programs}\label{tab:us-vs-other}
-\begin{tabular}{c|c|c}
-\toprule
-\textbf{Aspect} & \textbf{US PhD Programs} & \textbf{Other Countries} \\
-\midrule
-Duration & 5-7 years & 3-5 years \\
-MS Required & Not required & Often required \\
-Coursework Required & Yes (first 2 years) & No \\
-Research Proposal Required & No & Yes (in some countries) \\
-Academic (Faculty) Job & Direct & Postdoc \\
-Work Life Balance & Less & More \\
-\bottomrule
-\end{tabular}
-\end{table}
-
-\autoref{tab:us-vs-other} summarizes the main differences between CS PhD in the US and other countries. % This is based on my experience and what I heard from others who did PhD in other countries.
-
-\emph{MS requirement and PhD duration}:  CS PhD programs in the US do not require an MS degree (as mentioned in \autoref{sec:msrequirement} and \autoref{sec:time}).
-In contrast, many other countries require having an MS degree before joining a PhD program.
-This means that US PhD programs are longer (5--7 years, 2 of which are coursework) than other countries (3--4 years, no coursework).
-
-\emph{Project proposal}: in many countries, you have to choose a project and adviser \emph{during} the application process (e.g., you write a proposal to a potential adviser).
-But this allows you to start your research right from the beginning.
-
-In the US, you often start your PhD without an adviser or project and find them later.
-You will have two initial years to explore and find an adviser and research topic.
-After that you explore projects with that adviser before settling on one (can take a couple of years).
-
-\emph{Course work}: as mentioned above in the US you will spend the first couple of years taking classes and exploring potential adviser and research topics. 
-After that, you have to pass a series of exams during your PhD, e.g., qualifying exam, comprehensive exam, thesis proposal defense\footnote{The word ``ABD'' (all but dissertation) is used in the US to refer to a PhD candidate who has finished all course work and exams and only needs to write and defend their dissertation.}.
-
-In other countries, you start your research right away, e.g., you immediately work on the research project you proposed with the adviser you chose.
-Moreover, you might not have exams like those in the US or only have to do a few of them.
-
-\emph{Funding}:  In many countries, funding comes from the university or the gov't.
-This funding often has a fixed duration (e.g., 3 or 4 years).
-In the US (\autoref{sec:funding}), funding (e.g., RA) comes directly from your adviser (no fixed duration).
-There are also fewer TA opportunities in European universities compared to the US.
-
-\emph{Academic position after PhD}: In other countries, PhD graduates interested in academia typically apply for additional research appointments, i.e., postdocs, and then consider faculty positions.
-
-In the US, PhD graduates often apply directly for faculty positions. Postdoc for US graduates is no longer a popular option as it was before.
-
-\emph{Work-life balance}: PhD students in the US are often said to be overworked compared to other countries, e.g., in Europe.
-This is partly due to the longer PhD program and that US PhD students are often paid through TA, which requires them to do TA in addition to their own research.
-In contrast, PhD students in other countries are often paid through fellowships, which might not require doing TA.
 

### 2024-10-06 demystify.tex
updates

@@
     {\fontsize{24}{24} \selectfont {{Demystifying the Computer Science}}} \\[0.15in]
-    {\fontsize{24}{24} \selectfont {{PhD Admission in the US}}} \\[0.2in]    
-    {\fontsize{18}{18} \selectfont {{A Handbook for International Students}}} \\[1.in]
+    {\fontsize{24}{24} \selectfont {{PhD Admission in the US}}} \\[1in]    
+    %{\fontsize{18}{18} \selectfont {{A Handbook for International Students}}} \\[1.in]
 
@@
 the Computer Science PhD admission process at US universities. This confusion not only
-discourages students from applying but also creates the perception that
-getting admitted to a CS PhD program in the US is difficult compared to other countries.
+discourages students from applying but also creates the perception that getting admitted to a CS PhD program in the US is difficult compared to other countries.
 
@@
 So I want to share some details about the admission process and advice for those who are interested in applying for a \textbf{PhD in Computer Science in the US}.
-Originally, this document was intended for international students, but I have expanded it to include information that might also be useful for \emph{US domestic students}.
-Moreover, while this is primarily intended for students interested in CS, it might be relevant to students from various STEM (Science, Technologies, Engineering, and Mathematics) disciplines.
-Furthermore, although many examples are specifics for schools that I and other contributors of this document know about, the information should be generalizable to other R1\footnote{An \href{https://en.wikipedia.org/wiki/List_of_research_universities_in_the_United_States}{R1 institution} in the US is a research-intensive university with a high level of research activity across various disciplines.
+Originally, this document was intended for \emph{international students}, but I have expanded it to include information that might also be useful for \emph{US domestic students}.
+Moreover, while this primarily aims at students interested in CS, it might be relevant to students from various STEM (Science, Technologies, Engineering, and Mathematics) disciplines.
+Furthermore, although many examples are specifics for schools that I and other contributors of this document know about, the information should be generalizable to other good R1\footnote{An \href{https://en.wikipedia.org/wiki/List_of_research_universities_in_the_United_States}{R1 institution} in the US is a research-intensive university with a high level of research activity across various disciplines.
 Currently, 146 (out of 4000) US universities are classified as R1.} institutions in the US.
@@
 
-I wish you the best of luck. Happy school hunting!
+I wish you the best of luck.
 
 \begin{mybox}
-This document will be updated regularly to reflect the latest information and updates in the admission process. Its latest version is available at
+This document is open source and updated regularly. Its latest version is available at
 
@@
 First, applying to a good US university \emph{should not} be any harder than at schools in other countries.
-In fact, it might even be more flexible since CS PhD in the US \emph{do not} require having an MS or a research topic/proposal/adviser in advance.
+In fact, it might even be more flexible since CS PhD in the US \emph{do not} require having an MS or a research topic, proposal, or adviser in advance (\autoref{sec:non-us-differences} compares CS PhD study in the US to other countries).
 If you believe you have a chance in other countries, e.g., South Korea, Singapore, Germany, UK, Japan, and Australia, then you will surely have a chance in the US as well.
-\autoref{sec:non-us-differences} compares CS PhD study in the US to other countries.
 
@@
 universities in the US. If you are admitted, you will almost certainly
-receive \emph{full funding} (\autoref{sec:funding}) to support your study, including tuition,
+receive \emph{full funding} (\autoref{sec:funding}) to support your study. This includes tuition,
 health insurance, and stipend (i.e., you get paid for your study). Moreover, depending on the university,
-you may even receive additional benefits such as summer pay, laptops, and traveling to conferences and workshops.
+you may receive additional benefits such as summer pay, laptops, and traveling to conferences and workshops.
 
@@
 The goal is to help you understand the process and make the best decisions for your future.
+
+\section{Who is this document for?}
+\sectioninfo{This document is intended for international students from smaller countries and domestic students from less well-known universities.}
+
+The guideline can be useful for all students, but it is primarily intended for \textbf{international students} from \textbf{smaller countries and less well-known universities} (or domestic students from US universities with no PhD programs or limited research opportunities).
+
+Students from top schools with strong research cultures and experience might already know most of the information given here.
+My goal is thus to level the playing field by providing information that is not readily available to less privileged students. 
+I hope to encourage more students with such backgrounds to apply to CS PhD programs in the US.
+
+While this document may not fully meet the needs of students aiming for top-10 schools (\autoref{sec:ranking}), which typically have more stringent requirements (e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU), it should be valuable for those targeting other reputable R1 institutions.
+It is certainly useful for students aiming for Mason!
 
@@
 
-\section{Who is this document for?}
-\sectioninfo{This document is intended for international students from smaller countries and domestic students from less well-known universities.}
-
-The guideline can be useful for all students, but it is primarily intended for \textbf{international students} from \textbf{smaller countries and less well-known universities} (or domestic students from US universities with no PhD programs or limited research opportunities).
-
-Students from top schools with strong research cultures and experiences might already know most of the information given here.
-My goal is thus to level the playing field by providing information that is not readily available to less privileged students. 
-I hope to encourage more students with such backgrounds to apply to CS PhD programs in the US.
-
-While this document may not fully meet the needs of students aiming for top-10 schools (\autoref{sec:ranking}), which typically have more stringent requirements (e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU), it should be valuable for those targeting other reputable R1 institutions.
-It is certainly useful for students aiming for Mason!
-
 \section{How long to complete the CS PhD program?}\label{sec:time}
@@
 Typically, it takes 5--7 years for CS PhD in the US.
-This can be longer than CS PhDs in other countries, which might require MS first (recall from \autoref{sec:msrequirement} that CS PhD programs in the US do not require MS and you can get MS along the way to PhD).
+This can be longer than CS PhDs in other countries (\autoref{sec:msrequirement}), which might require MS first.
 Within these 5--7 years, CS PhD students often take a ``leave of absence'' for 1--2 semesters to do internships at companies and research labs.
@@
 % TODO
-\subsubsection{Why CS depts do not waive the application fee?}
-
-Some universities do waive, e.g., Rice does for all PhD applicants and many universities for domestic students (\autoref{sec:domestic-students}), however most do not for international students.
-In any case, the application fee is typically a requirement of the university.
-Individual departments and programs do not have the flexibility to waive the application fee, even if they want to.
-
-\begin{commentbox}{Vu:}
-In my opinion, requiring applicants to pay the fee helps ensure their seriousness, as it filters out non-serious candidates.
-Also, most CS programs already receive many applications and would be overwhelmed if the application process were free.
-Note that if you have financial difficulties, you can ask the department for a waiver, but this is typically only granted in exceptional cases.
-\end{commentbox}
 
@@
 However, even if we look at them, we usually do not get involved in their evaluation.
+
+\subsection{Why CS depts do not waive the application fee?}
+
+Some universities do waive, e.g., Rice does for all PhD applicants and many universities for domestic students (\autoref{sec:domestic-students}), however most do not for international students.
+In any case, the application fee is typically a requirement of the university.
+Individual departments and programs do not have the flexibility to waive the application fee, even if they want to.
+
+\begin{commentbox}{Vu:}
+In my opinion, requiring applicants to pay the fee helps ensure their seriousness, as it filters out non-serious candidates.
+Also, most CS programs already receive many applications and would be overwhelmed if the application process were free.
+Note that if you have financial difficulties, you can ask the department for a waiver, but this is typically only granted in exceptional cases.
+\end{commentbox}
 

### 2024-10-14 demystify.tex
update

@@
 Specifically, they are worried about (i) the \textbf{difficulty of getting admitted} and (ii) the \textbf{cost of graduate study} in the US.
-This section addresses these concerns.
+Part of the reasons is the lack of information and guidance on the admission process in the US (let's face it, public forums like Facebook communities and Reddit are full of confusion and concerns from students).
+This section and the rest of this document aim to address these concerns.
 
 First, applying to a good US university \emph{should not} be any harder than at schools in other countries.
-In fact, it might even be more flexible since CS PhD in the US \emph{do not} require having an MS or a research topic, proposal, or adviser in advance (\autoref{sec:non-us-differences} compares CS PhD study in the US to other countries).
+It might even be more flexible since CS PhD in the US \emph{do not} require having an MS or a research topic, proposal, or adviser in advance (\autoref{sec:non-us-differences} compares CS PhD study in the US to other countries).
 If you believe you have a chance in other countries, e.g., South Korea, Singapore, Germany, UK, Japan, and Australia, then you will surely have a chance in the US as well.
@@
 Second, PhD students in CS \emph{do not} need to worry about funding, especially at good R1
-universities in the US. If you are admitted, you will almost certainly
+universities in the US. 
+If you are admitted, you will almost certainly
 receive \emph{full funding} (\autoref{sec:funding}) to support your study. This includes tuition,
-health insurance, and stipend (i.e., you get paid for your study). Moreover, depending on the university,
-you may receive additional benefits such as summer pay, laptops, and traveling to conferences and workshops.
-
-The rest of this document will provide you with more details on the CS PhD admission process in the US, including how to prepare your application, what to expect during process, what to do with acceptance (and how to deal with rejection), and so on.
-The goal is to help you understand the process and make the best decisions for your future.
+health insurance, and stipend (i.e., you get paid for your study). 
+In fact, full funding for CS PhD students is the norm in the US, and I'd go as far as to say that if you are not admitted with full funding, you should be cautious about the program and might want to not accept the offer.
+Moreover, depending on the university, you may receive additional benefits such as summer pay, laptops, and traveling to conferences and workshops.
 
@@
 I hope to encourage more students with such backgrounds to apply to CS PhD programs in the US.
+
+Another goal is to explain the \emph{``why''} behind the admission process, e.g., why LoRs are important, why you should not draft your own LoR, why you should contact professors, etc.
 
@@
 
-Typically, it takes 5--7 years for CS PhD in the US.
-This can be longer than CS PhDs in other countries (\autoref{sec:msrequirement}), which might require MS first.
-Within these 5--7 years, CS PhD students often take a ``leave of absence'' for 1--2 semesters to do internships at companies and research labs.
+Typically, it takes 5--7 years for CS PhD in the US.  This can be longer than CS PhDs in other countries (\autoref{sec:msrequirement}), which might require MS first.
 
@@
 The last 1--2 years you continue to publish, write and defend your dissertation, and look for a job.
-In many cases you might take a summer or two off to do an internship to get additional research opportunities.
+Within these 5--7 years, CS PhD students often take a ``leave of absence'' for 1--2 semesters or summer to do internships at companies and research labs.
+
 The PhDComics figure above shows the ``ambition'' level of a PhD student over their years of study (they miss the 6--7th year when the ambition is \emph{``Just let me graduate''}).
 
-% \begin{commentbox}[Vu:]
-%     I start my PhD with an MS, and it took me 7 years (Fall'2007-- Fall'14). I spent half a year doing an internship at the Naval Research Lab.
-%   \end{commentbox}
+\begin{commentbox}[Vu:]
+    I start my PhD with an MS, and it took me 7 years (Fall'07-- Fall'14).
+I spent half a year doing an internship at the Naval Research Lab.
+My PhD did take a bit longer than usual, but allows me to explore various research areas and topics.
+  \end{commentbox}
 

### 2024-10-15 demystify.tex
update

@@
 
-If you have an MS then some course work \emph{might be} transferred for course credits, which \emph{might} save some time. But don't count on this.
+If you have an MS then some course work \emph{might be} transferred for course credits, which \emph{might} save some time. But in general don't count finishing earlier just because you have an MS.
 
@@
     I start my PhD with an MS in CS from a US university.
-I found that the MS helped me a lot in research, but not with course work and I had to retake many courses.
-In general, don't worry if you don't have an MS.
-But also don't feel that you wasted your time if you have an MS, as it can help you in research.
+I found that the MS helped me in gaining research experience, but I still had to retake courses because I did my MS at a different university.
+So in the end, I did not save any time because of the MS.
+
+    In general, don't worry if you don't have an MS. But also don't feel that you wasted your time if you have an MS, as it can help you in research.
   \end{commentbox}

### 2024-10-15 ennv.tex
update

--- 2024-08-10 nnv.tex
@@
     {
-    {\fontsize{38}{1} \selectfont {{Neural Network Verification}}}\\[0.5in]
+        {\fontsize{24}{24} \selectfont {{Engineering A Verifier}}} \\[0.15in]
+        {\fontsize{24}{24} \selectfont {{for Deep Neural Networks}}} \\[1in]    
+        %{\fontsize{18}{18} \selectfont {{A Handbook for International Students}}} \\[1.in]
 
@@
 
-\begin{wrapfigure}{r}{0.3\textwidth}
-    \centering
-    \includegraphics[width=1\linewidth]{figure/dnn.pdf}
-    \caption{\label{fig:dnn} An FNN with ReLU.}
-  \end{wrapfigure}
+%\begin{wrapfigure}{r}{0.3\textwidth}
+%    \centering
+%    \includegraphics[width=1\linewidth]{figure/dnn.pdf}
+%    \caption{\label{fig:dnn} An FNN with ReLU.}
+%  \end{wrapfigure}
 \paragraph{Example} Fig.~\ref{fig:dnn} shows a simple DNN with two inputs $x_1,x_2$, two hidden neurons $x_3,x_4$, and one output $x_5$.
@@
 \bibliographystyle{abbrv}
-\bibliography{demystify.bib}
+\bibliography{nnv.bib}
 

### 2024-10-15 faculty.tex
update

@@
 
-This section focuses on research publication, which is a major component of P\&T and of course in getting your name out there.
+Research, in particular publications, is a major component of P\&T and of course in getting your name out there.
 P\&T evaluation requires no specific conferences/journals or numbers of papers, but typically you will want to have a good number of papers in top venues.
-Work that won awards, e.g., best paper awards, can help a lot in P\&T.
+Papers and work that won awards, e.g., best paper awards, can help a lot in P\&T.
 Papers with very high citations also help.
@@
 
-Successfully mentoring students is a key and most rewarding part of your job as a faculty member. 
+Successfully mentoring students is import for P\&T process.  Many universities expect you to graduate at least one PhD student to get tenure.  
 There will be times when you feel discouraged and frustrated, but there will also be times when you feel proud and happy.
-
-For P\&T you will need to successfully mentoring students.  Many universities expect you to need to graduate at least one PhD student to get tenure.
+In short, mentoring students is a key and most rewarding part of your job as a faculty member.
 
@@
 \section{Mentoring Graduate Students}
+You will need to mentor PhD students and help them find research topics, do research, publish papers, and become independent researchers.
+This is a long and challenging process, but it is also very rewarding.
 

### 2024-10-15 ennv.tex
update

@@
 Each layer has a number of neurons, each connected to neurons from previous layers through a predefined set of weights (derived by training the network with data).
-A \textbf{DNN} is an NN with at least two hidden layers.
+A \emph{Deep Neural Network} (\textbf{DNN}) is an NN with at least two hidden layers.
 
@@
 \section{Affine Transformation}
+The affine transformation (AF) of a neuron is the sum of the products of the weights of the incoming edges and the values of the neurons in the previous layer, plus the bias of the neuron.
+More specifically, the AF of a neuron \(y\) with weights \(w_1, \dots, w_n\) and bias \(b\) and the values of neurons in the previous layer \(v_1, \dots, v_n\) is \(w_1v_1 + \dots + w_nv_n + b\).
 
-\section{Activation Function}
+For example, the AF of a neuron \(y\) with weights \(-0.5, 0.5\) and bias \(1.0\) and the values of neurons in the previous layer \(x_1, x_2\) is \(-0.5x_1 + 0.5x_2 + 1.0\).
+
+For DNN verification, AF is straightforward to reason about because it is a linear function.
+However, AFs are often followed by non-linear activation functions, described next in \autoref{sec:activation}, which make the verification problem more challenging.
+
+\section{Activation Functions}\label{sec:activation}
+Several popular activation functions used in DNNs include ReLU, Sigmoid, Tanh, and Softmax.
+All of these are non-linear\footnote{Non-linear means that the output of the function is not a linear combination of its inputs.} functions that introduce non-linearity to the network, allowing it to learn complex patterns in the data.
+
+\begin{itemize}
+\item ReLU (Rectified Linear Unit): ReLU is the most popular activation function in DNNs.
+It returns 0 if the input is less than zero, and the input itself otherwise.
+It is often used in hidden layers.\\
+    \begin{center}
+        $ReLU(x) = \max(x,0)$
+    \end{center}
+\item Sigmoid: Sigmoid is a smooth function that maps any real value to the range (0,1). It is often used in the output layer of a binary classification problem.\\
+    \begin{center}
+        $Sigmoid(x) = \frac{1}{1+e^{-x}}$
+    \end{center}
+\item Tanh: Tanh is similar to the sigmoid function but maps any real value to the range (-1,1). It is often used in the output layer of a multi-class classification problem.\\
+    \begin{center}
+        $Tanh(x) = \frac{e^x-e^{-x}}{e^x+e^{-x}}$
+    \end{center}
+\item Softmax: Softmax is a generalization of the sigmoid function that maps any real value to the range (0,1) and ensures that the sum of the output values is 1.
+It is often used in the output layer of a multi-class classification problem.\\
+    \begin{center}
+    $Softmax(x)_i = \frac{e^{x_i}}{\sum_{j=1}^{n}e^{x_j}}$
+    \end{center}
+\end{itemize}
+
+For DNN verification, these non-linear activation functions make verification difficult because it introduces multiple possible outcomes for any input, making it hard to reason about the output of the network.
+For example, ReLU has two possible outputs for any input: 0 if the input is less than zero, and the input itself otherwise, and Sigmoid has a smooth curve with infinite possible outputs for any input.
 
@@
 \bibliographystyle{abbrv}
-\bibliography{nnv.bib}
+\bibliography{ennv.bib}
 

### 2024-10-15 ennv.tex
update

@@
 The value of a neuron in the input layer is the input data.
-The value of a neuron in the hidden layers is computed by applying an \emph{affine transformation} to values of neurons in the previous layers, then followed by an \emph{activation function} such as the popular Rectified Linear Unit (ReLU) activation.
-
-For this activation, the value of a hidden neuron \(y\) is 
-$ReLU(w_1v_1 + \dots{} + w_nv_n + b)$, where \(b\) is the bias parameter of \(y\), \(w_i, \dots, w_n\) are the weights of \(y\), \(v_1,\dots,v_n\) are the neuron values of preceding layer, \(w_1v_1 + \dots + w_nv_n+b\) is the affine transformation, and \(ReLU(x) = \max(x,0)\) is the ReLU activation.
-The values of a neuron in the output layer is evaluated similarly but it may skip the activation function.
-A ReLU activated neuron is said to be \emph{active} if its input value is greater than
-zero and \emph{inactive} otherwise.
-
-\paragraph{DNN Verification} Given a DNN \(N\) and a property $\phi$, the \emph{DNN verification problem} asks if $\phi$ is a valid property of $N$.
-Typically, $\phi$ is a formula of the form $\phi_{in} \Rightarrow \phi_{out}$, where $\phi_{in}$ is a property over the inputs of $N$ and $\phi_{out}$ is a property over the outputs of $N$.
-%This form of properties has been used to encode safety and security requirements of DNNs, e.g., safety specifications to avoid collision in unmanned aircraft~\cite{kochenderfer2012next} and \emph{adversarial robustness}~\cite{katz2017towards} properties desired by all DNNs, in which a small input perturbation does not cause major spikes in the DNN's outputs.
-A DNN verifier attempts to find a \emph{counterexample} input to $N$ that satisfies $\phi_{in}$ but violates $\phi_{out}$.
-If no such counterexample exists, $\phi$ is a valid property of $N$.
-Otherwise, $\phi$ is not valid and the counterexample can be used to retrain or debug the DNN~\cite{huang2017safety}.
-
-% Verification tool such as Marabou and nnenum are then applied to the network to prove that the network is safe or identifier counterexample representing small input differences causing large output changes.
-
-% \footnote{This is encoded as the differences of the inputs being within a certain small range  ($\phi_{in}$) implies the differences of the outputs still fall within a certain range in $\phi_{out}$)}.
-
-%\begin{wrapfigure}{r}{0.3\textwidth}
-%    \centering
-%    \includegraphics[width=1\linewidth]{figure/dnn.pdf}
-%    \caption{\label{fig:dnn} An FNN with ReLU.}
-%  \end{wrapfigure}
-\paragraph{Example} Fig.~\ref{fig:dnn} shows a simple DNN with two inputs $x_1,x_2$, two hidden neurons $x_3,x_4$, and one output $x_5$.
-The weights of a neuron are shown on its incoming edges , and the bias is shown above or below each neuron.
-The outputs of the hidden neurons  are computed the affine transformation and ReLU, e.g., $x_3 = ReLU(-0.5x_1+0.5x_2+1.0)$.
-The output neuron is computed with just the affine transformation, i.e., $x_5=-x_3+x_4-1$.
-
-A valid property for this DNN is that the output is $x_5 \le 0$ for any inputs $x_1 \in [-1,1], x_2\in[-2,2]$. An invalid property for this network is that $x_5 > 0$ for those similar inputs.
-A counterexample showing this property violation is $\{x_1=-1, x_2=2\}$, from which the network evaluates to $x_5=-3.5$.
-Such properties can capture \emph{safety requirements} (e.g., a rule in an  collision avoidance system in~\cite{kochenderfer2012next,katz2017reluplex} is ``if the intruder is distant and significantly slower than us, then we stay below a certain threshold'') or \emph{local robustness}~\cite{katz2017towards} conditions (a form of adversarial robustness stating that small perturbations of a given input all yield the same output).
+The value of a neuron in the hidden layers is computed by applying an \emph{affine transformation} (\autoref{sec:affine}) to values of neurons in the previous layers, then followed by an \emph{activation function} (\autoref{sec:activation}) such as the popular Rectified Linear Unit (ReLU) activation.
 
@@
 
-\section{Affine Transformation}
+\section{Affine Transformation}\label{sec:affine}
 The affine transformation (AF) of a neuron is the sum of the products of the weights of the incoming edges and the values of the neurons in the previous layer, plus the bias of the neuron.
@@
 
-For example, the AF of a neuron \(y\) with weights \(-0.5, 0.5\) and bias \(1.0\) and the values of neurons in the previous layer \(x_1, x_2\) is \(-0.5x_1 + 0.5x_2 + 1.0\).
+%AF is often applied to hidden layers of a DNN to compute the value of a neuron in the hidden layer.
+%The values of a neuron in the output layer is evaluated similarly but it may skip the activation function.
+
+For example, the AF of a neuron \(x_3\) in \autoref{fig:dnn} with (incoming arrows) weights \(-0.5, 0.5\) and bias \(1.0\) and the values of neurons in the previous layer \(x_1, x_2\) is \(-0.5x_1 + 0.5x_2 + 1.0\).
 
@@
 It returns 0 if the input is less than zero, and the input itself otherwise.
-It is often used in hidden layers.\\
+It is often used in hidden layers and skipped in the output layer.
+A ReLU activated neuron is said to be \emph{active} if its input value is greater than
+zero and \emph{inactive} otherwise.\\
     \begin{center}
@@
     \end{center}
+
 \item Sigmoid: Sigmoid is a smooth function that maps any real value to the range (0,1). It is often used in the output layer of a binary classification problem.\\
@@
 
+\section{Example} 
+Fig.~\ref{fig:dnn} shows a simple DNN with two inputs $x_1,x_2$, two hidden neurons $x_3,x_4$, and one output $x_5$.
+The weights of a neuron are shown on its incoming edges , and the bias is shown above or below each neuron.
+The outputs of the hidden neurons  are computed the affine transformation and ReLU, e.g., $x_3 = ReLU(-0.5x_1+0.5x_2+1.0)$.
+The output neuron is computed with just the affine transformation, i.e., $x_5=-x_3+x_4-1$.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.5\linewidth]{figure/dnn.pdf}
+    \caption{\label{fig:dnn} An FNN with ReLU.}
+\end{figure}
+
 \section{Properties of Neural Networks}
@@
 \chapter{Verification of Neural Networks}\label{sec:verification}
+\paragraph{DNN Verification} Given a DNN \(N\) and a property $\phi$, the \emph{DNN verification problem} asks if $\phi$ is a valid property of $N$.
+Typically, $\phi$ is a formula of the form $\phi_{in} \Rightarrow \phi_{out}$, where $\phi_{in}$ is a property over the inputs of $N$ and $\phi_{out}$ is a property over the outputs of $N$.
+%This form of properties has been used to encode safety and security requirements of DNNs, e.g., safety specifications to avoid collision in unmanned aircraft~\cite{kochenderfer2012next} and \emph{adversarial robustness}~\cite{katz2017towards} properties desired by all DNNs, in which a small input perturbation does not cause major spikes in the DNN's outputs.
+A DNN verifier attempts to find a \emph{counterexample} input to $N$ that satisfies $\phi_{in}$ but violates $\phi_{out}$.
+If no such counterexample exists, $\phi$ is a valid property of $N$.
+Otherwise, $\phi$ is not valid and the counterexample can be used to retrain or debug the DNN~\cite{huang2017safety}.
+
+% Verification tool such as Marabou and nnenum are then applied to the network to prove that the network is safe or identifier counterexample representing small input differences causing large output changes.
+
+% \footnote{This is encoded as the differences of the inputs being within a certain small range  ($\phi_{in}$) implies the differences of the outputs still fall within a certain range in $\phi_{out}$)}.
+
+\paragraph{Example} A valid property for the DNN in \autoref{fig:dnn} is that the output is $x_5 \le 0$ for any inputs $x_1 \in [-1,1], x_2\in[-2,2]$.
+An invalid property for this network is that $x_5 > 0$ for those similar inputs.
+A counterexample showing this property violation is $\{x_1=-1, x_2=2\}$, from which the network evaluates to $x_5=-3.5$.
+Such properties can capture \emph{safety requirements} (e.g., a rule in an  collision avoidance system in~\cite{kochenderfer2012next,katz2017reluplex} is ``if the intruder is distant and significantly slower than us, then we stay below a certain threshold'') or \emph{local robustness}~\cite{katz2017towards} conditions (a form of adversarial robustness stating that small perturbations of a given input all yield the same output).
 

### 2024-10-15 ennv.tex
update

@@
 
+\section{Types of Neural Networks}
+
+\paragraph{Convolutional Neural Networks} Convolutional Neural Networks (CNNs) are a type of neural network that is often used in image recognition and classification.
+CNNs are designed to automatically and adaptively learn spatial hierarchies of features from input images.
+CNNs are made up of neurons that have learnable weights and biases.
+Each neuron receives several inputs, takes a weighted sum over them, passes it through an activation function, and responds with an output.
+
+\paragraph{Recurrent Neural Networks} Recurrent Neural Networks (RNNs) are a type of neural network that is often used in natural language processing and speech recognition.
+RNNs are designed to recognize patterns in sequences of data.
+RNNs have loops in them, allowing information to persist.
+This loop allows information to be passed from one step of the network to the next.
+
+\paragraph{Generative Adversarial Networks} Generative Adversarial Networks (GANs) are a type of neural network that is often used in generating new data.
+GANs are made up of two networks, a generator and a discriminator.
+The generator creates new data instances, while the discriminator evaluates them for authenticity.
+The goal of the generator is to generate data that is indistinguishable from real data, while the goal of the discriminator is to correctly classify real and generated data.
+
+\paragraph{Autoencoders} Autoencoders are a type of neural network that is often used in unsupervised learning.
+Autoencoders are designed to learn efficient representations of data.
+Autoencoders consist of an encoder and a decoder.
+The encoder compresses the input data into a lower-dimensional representation, while the decoder reconstructs the input data from the lower-dimensional representation.
+
+\paragraph{Residual Networks} Residual Networks (ResNets) are a type of neural network that is often used in image recognition and classification.
+ResNets are designed to address the vanishing gradient problem that occurs in deep neural networks.
+ResNets introduce skip connections that allow the gradient to flow directly through the network, making it easier to train deep networks.
+
 \section{Properties of Neural Networks}

### 2024-10-15 ennv.tex
update

@@
 
+\paragraph{Feed-Forward Network} A Feed-Forward Network (FFN) is a type of neural network where the connections between the neurons do not form a cycle.
+In an FFN, the information flows in one direction, from the input layer to the output layer.
+FFNs are the simplest type of neural network and are often used in classification and regression tasks.
+A fully connected feed-forward neural network (FNN) is a type of FFN where every neuron in a layer is connected to every neuron in the next layer.
+\begin{center}
+\begin{tikzpicture}[scale=0.8, transform shape]
+% Input Layer
+\node at (0,4) {\textbf{Input Layer}};
+\foreach \x in {1,2,3,4} {
+    \node[circle, draw=black, fill=white, minimum size=0.6cm] (I\x) at (0,4-\x) {};
+}
+
+% Hidden Layer 1
+\node at (3,4) {\textbf{Hidden Layer 1}};
+\foreach \x in {1,2,3,4,5} {
+    \node[circle, draw=black, fill=white, minimum size=0.6cm] (H1\x) at (3,4.5-\x) {};
+}
+
+% Hidden Layer 2
+\node at (6,4) {\textbf{Hidden Layer 2}};
+\foreach \x in {1,2,3,4,5} {
+    \node[circle, draw=black, fill=white, minimum size=0.6cm] (H2\x) at (6,4.5-\x) {};
+}
+
+% Output Layer
+\node at (9,3.5) {\textbf{Output Layer}};
+\foreach \x in {1,2,3} {
+    \node[circle, draw=black, fill=white, minimum size=0.6cm] (O\x) at (9,3.5-\x) {};
+}
+
+% Connections
+\foreach \i in {1,2,3,4}
+    \foreach \j in {1,2,3,4,5}
+        \draw[->] (I\i) -- (H1\j);
+
+\foreach \i in {1,2,3,4,5}
+    \foreach \j in {1,2,3,4,5}
+        \draw[->] (H1\i) -- (H2\j);
+
+\foreach \i in {1,2,3,4,5}
+    \foreach \j in {1,2,3}
+        \draw[->] (H2\i) -- (O\j);
+
+\end{tikzpicture}
+\end{center}
+
 \paragraph{Convolutional Neural Networks} Convolutional Neural Networks (CNNs) are a type of neural network that is often used in image recognition and classification.

### 2024-10-16 demystify.tex
updates

--- 2024-10-06 demystify.tex
@@
     {\fontsize{24}{24} \selectfont {{Demystifying the Computer Science}}} \\[0.15in]
-    {\fontsize{24}{24} \selectfont {{PhD Admission in the US}}} \\[1in]    
-    %{\fontsize{18}{18} \selectfont {{A Handbook for International Students}}} \\[1.in]
+    {\fontsize{24}{24} \selectfont {{PhD Admission in the US}}} \\[0.2in]    
+    {\fontsize{18}{18} \selectfont {{A Handbook for International Students}}} \\[1.in]
 
@@
 the Computer Science PhD admission process at US universities. This confusion not only
-discourages students from applying but also creates the perception that getting admitted to a CS PhD program in the US is difficult compared to other countries.
+discourages students from applying but also creates the perception that
+getting admitted to a CS PhD program in the US is difficult compared to other countries.
 
@@
 So I want to share some details about the admission process and advice for those who are interested in applying for a \textbf{PhD in Computer Science in the US}.
-Originally, this document was intended for \emph{international students}, but I have expanded it to include information that might also be useful for \emph{US domestic students}.
-Moreover, while this primarily aims at students interested in CS, it might be relevant to students from various STEM (Science, Technologies, Engineering, and Mathematics) disciplines.
-Furthermore, although many examples are specifics for schools that I and other contributors of this document know about, the information should be generalizable to other good R1\footnote{An \href{https://en.wikipedia.org/wiki/List_of_research_universities_in_the_United_States}{R1 institution} in the US is a research-intensive university with a high level of research activity across various disciplines.
+Originally, this document was intended for international students, but I have expanded it to include information that might also be useful for \emph{US domestic students}.
+Moreover, while this is primarily intended for students interested in CS, it might be relevant to students from various STEM (Science, Technologies, Engineering, and Mathematics) disciplines.
+Furthermore, although many examples are specifics for schools that I and other contributors of this document know about, the information should be generalizable to other R1\footnote{An \href{https://en.wikipedia.org/wiki/List_of_research_universities_in_the_United_States}{R1 institution} in the US is a research-intensive university with a high level of research activity across various disciplines.
 Currently, 146 (out of 4000) US universities are classified as R1.} institutions in the US.
@@
 
-I wish you the best of luck.
+I wish you the best of luck. Happy school hunting!
 
 \begin{mybox}
-This document is open source and updated regularly. Its latest version is available at
+This document will be updated regularly to reflect the latest information and updates in the admission process. Its latest version is available at
 
@@
 First, applying to a good US university \emph{should not} be any harder than at schools in other countries.
-In fact, it might even be more flexible since CS PhD in the US \emph{do not} require having an MS or a research topic, proposal, or adviser in advance (\autoref{sec:non-us-differences} compares CS PhD study in the US to other countries).
+In fact, it might even be more flexible since CS PhD in the US \emph{do not} require having an MS or a research topic/proposal/adviser in advance.
 If you believe you have a chance in other countries, e.g., South Korea, Singapore, Germany, UK, Japan, and Australia, then you will surely have a chance in the US as well.
+\autoref{sec:non-us-differences} compares CS PhD study in the US to other countries.
 
@@
 universities in the US. If you are admitted, you will almost certainly
-receive \emph{full funding} (\autoref{sec:funding}) to support your study. This includes tuition,
+receive \emph{full funding} (\autoref{sec:funding}) to support your study, including tuition,
 health insurance, and stipend (i.e., you get paid for your study). Moreover, depending on the university,
-you may receive additional benefits such as summer pay, laptops, and traveling to conferences and workshops.
+you may even receive additional benefits such as summer pay, laptops, and traveling to conferences and workshops.
 
@@
 The goal is to help you understand the process and make the best decisions for your future.
-
-\section{Who is this document for?}
-\sectioninfo{This document is intended for international students from smaller countries and domestic students from less well-known universities.}
-
-The guideline can be useful for all students, but it is primarily intended for \textbf{international students} from \textbf{smaller countries and less well-known universities} (or domestic students from US universities with no PhD programs or limited research opportunities).
-
-Students from top schools with strong research cultures and experience might already know most of the information given here.
-My goal is thus to level the playing field by providing information that is not readily available to less privileged students. 
-I hope to encourage more students with such backgrounds to apply to CS PhD programs in the US.
-
-While this document may not fully meet the needs of students aiming for top-10 schools (\autoref{sec:ranking}), which typically have more stringent requirements (e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU), it should be valuable for those targeting other reputable R1 institutions.
-It is certainly useful for students aiming for Mason!
 
@@
 
+\section{Who is this document for?}
+\sectioninfo{This document is intended for international students from smaller countries and domestic students from less well-known universities.}
+
+The guideline can be useful for all students, but it is primarily intended for \textbf{international students} from \textbf{smaller countries and less well-known universities} (or domestic students from US universities with no PhD programs or limited research opportunities).
+
+Students from top schools with strong research cultures and experiences might already know most of the information given here.
+My goal is thus to level the playing field by providing information that is not readily available to less privileged students. 
+I hope to encourage more students with such backgrounds to apply to CS PhD programs in the US.
+
+While this document may not fully meet the needs of students aiming for top-10 schools (\autoref{sec:ranking}), which typically have more stringent requirements (e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU), it should be valuable for those targeting other reputable R1 institutions.
+It is certainly useful for students aiming for Mason!
+
 \section{How long to complete the CS PhD program?}\label{sec:time}
@@
 Typically, it takes 5--7 years for CS PhD in the US.
-This can be longer than CS PhDs in other countries (\autoref{sec:msrequirement}), which might require MS first.
+This can be longer than CS PhDs in other countries, which might require MS first (recall from \autoref{sec:msrequirement} that CS PhD programs in the US do not require MS and you can get MS along the way to PhD).
 Within these 5--7 years, CS PhD students often take a ``leave of absence'' for 1--2 semesters to do internships at companies and research labs.
@@
 % TODO
+\subsubsection{Why CS depts do not waive the application fee?}
+
+Some universities do waive, e.g., Rice does for all PhD applicants and many universities for domestic students (\autoref{sec:domestic-students}), however most do not for international students.
+In any case, the application fee is typically a requirement of the university.
+Individual departments and programs do not have the flexibility to waive the application fee, even if they want to.
+
+\begin{commentbox}{Vu:}
+In my opinion, requiring applicants to pay the fee helps ensure their seriousness, as it filters out non-serious candidates.
+Also, most CS programs already receive many applications and would be overwhelmed if the application process were free.
+Note that if you have financial difficulties, you can ask the department for a waiver, but this is typically only granted in exceptional cases.
+\end{commentbox}
 
@@
 
-\section*{Additional Resources}
-\begin{itemize}
-  \item \href{https://parentheticallyspeaking.org/articles/us-cs-phd-faq/}{Getting a Computer Science PhD in the USA} by Shriram Krishnamurthi
-  \item  \href{https://pg.ucsd.edu/PhD-application-tips.htm}{Blog post on PhD Admission} by Philip Guo
-  \item   \href{https://talkingtorobots.com/FAQ.html}{PhD Admission post} by Yonatan Bisk (customized for CMU admission)
-\end{itemize}
-
 \section{How are decisions made?}\label{sec:how-decisions}
@@
 However, even if we look at them, we usually do not get involved in their evaluation.
-
-\subsection{Why CS depts do not waive the application fee?}
-
-Some universities do waive, e.g., Rice does for all PhD applicants and many universities for domestic students (\autoref{sec:domestic-students}), however most do not for international students.
-In any case, the application fee is typically a requirement of the university.
-Individual departments and programs do not have the flexibility to waive the application fee, even if they want to.
-
-\begin{commentbox}{Vu:}
-In my opinion, requiring applicants to pay the fee helps ensure their seriousness, as it filters out non-serious candidates.
-Also, most CS programs already receive many applications and would be overwhelmed if the application process were free.
-Note that if you have financial difficulties, you can ask the department for a waiver, but this is typically only granted in exceptional cases.
-\end{commentbox}
 

### 2024-10-16 faculty.tex
updates

--- 2024-09-30 faculty.tex
@@
 While teaching takes about 40\% of your time, it is not as nearly as important as research for tenure.
-Moreover, most new faculty do not prioritize teaching, e.g., compared to writing grants or recruiting students for research.
+Moreover, most TT faculty do not prioritize teaching compared to writing grants or working with students on papers.
 However, teaching is still important for your tenure case, and being a good teacher can be very rewarding.
@@
 
-\section{Teaching Load}
-New TT faculty often have a lighter teaching load in their first several years or until they get tenure.  This is to give you time to establish your research program.
-
-At most large R1 universities the teaching load for TT faculty is 1+1 (1 course / semester).
+\section{Courses Taught}\label{sec:teaching-load}
+
+% At most large R1 universities the teaching load for TT faculty is 1+1 (1 course / semester).
 While some univs might give you a lighter load in your first year, most univs expect you to teach 1+1 starting your second year.
@@
 It is rare that a TT in CS at R1 universities would have to teach 2+2 during the tenure track.
-
-In some cases your contract might list 1+1 for the first X year and 1+2 after until tenure, but the dept unofficial gives TT faculty 1+1 until tenure.
+% In some cases your contract might list 1+1 for the first X year and 1+2 after until tenure, but the dept unofficial gives TT faculty 1+1 until tenure.
 Some other ways to reduce teaching load include \emph{buying out} of teaching with a grant, leading a very large grant, and teaching many large classes.
 
+Most places want to see that you have taught both undergrad and grad courses by the time you go up for tenure.
+A good dept.
+would allow and encourage faculty to pick their favorite courses to teach.
+In some cases, they let you develop your own grad course (e.g., Software Verification) but ask you to teach an existing core undergraduate one (e.g., Theory, Operating Systems).
+A good dept.
+also "protects" you from teaching intro courses (e.g., Intro to Programming), which are usually very large and the most time-consuming to teach.
+
+It is likely that you get to repeat the same courses, which significant reduces the amount of time you need to prepare for the course.
+However, you should not teach the same course every semester, and you will want to alternate between grad and undergrad courses.
+
+\begin{commentbox}[Vu]
+  At GMU, I teach the same Software Specification and Construction course at both grad (SWE-619) and undergrad (SWE-419) levels.
+This is very nice because I don't have to prepare much for the course.
+I did try to teach an AI Assurance graduate seminar once, but never liked it much because it only has MS students, who were not very interested in research.
+  \\
+
+  In contrast, I teach various courses at UNL.
+I teach the several undergraduate courses including Theory, Software Engineering, and design my own two courses on Compilers and Software Analysis.
+I also create a grad seminar on Software Verification.\\
+
+ THus UNL is a bit more flexible in creating new classes compared to GMU.  However, I spent much less time prepare for courses at GMU.
+\end{commentbox}
+
+\paragraph{Teaching load} New TT faculty often have a lighter teaching load in their first several years or until they get tenure.  This is to give you time to establish your research program.
 \begin{commentbox}[Vu]
@@
 
-\section{Courses}
-
-In most cases you will teach both undergrad and grad courses.
-New faculty often create graduate seminars and special topic courses to recruit students for their research group.
-Undergrad courses typically include higher level courses in your area of expertise (e.g., 400 level course on Software Testing).
-It is rare that you have to teach 100 level courses (e.g., Intro to Programming) because they are very large and thus time-consuming.
-
-Most depts.
-would allow faculty to pick their favorite courses to teach.
-In some cases, they let  you develop your own grad course (e.g., Software Verification) but ask you to teach an existing core undergraduate courses (e.g., Theory, Operating Systems).
-It is likely that you get to repeat the same courses.
-This significant reduces the amount of time you need to prepare for the course.
-However, you should not teach the same course every semester, and you will want to alternate between grad and undergrad courses.
-
-\begin{commentbox}[Vu]
-    At GMU, I teach the same Software Specification and Construction course at both grad (SWE-619) and undergrad (SWE-419) levels.
-This is very nice because I don't have to prepare much for the course.
-I did try to teach an AI Assurance graduate seminar once, but never liked it much because it only has MS students, who were not very interested in research.
-    \\
-
-    In contrast, I teach various courses at UNL.
-I teach the several undergraduate courses including Theory, Software Engineering, and design my own two courses on Compilers and Software Analysis.
-I also create a grad seminar on Software Verification.\\
-
-    I feel that UNL is a bit more flexible in creating new classes compared to GMU.  However, I spent much less time prepare for courses at GMU.
-  \end{commentbox}
-
 \paragraph{Co-Teaching}
@@
 
-Teaching evaluations are a critical component of your annual review and P\&T. It provides an analysis on how students and peers perceive your teaching. 
-You pay attention to evaluations, and make changes to your teaching based on the feedback, and show that you continuously strive to improve your teaching.
+Teaching evaluations are a critical component of P\&T, and you will include these in your dossier when you go up for tenure. 
+It provides a quantitative measure on how students and peers perceive your teaching.
+You should pay attention to evaluations, and make changes to your teaching based on the feedback, and show that you continuously strive to improve your teaching.
 
@@
 The results are shared anonymously with the instructor to help them improve their teaching.
-
-One main issue with teaching evaluation is that not many students respond, despite multiple reminders and urging from the instructor.
-Moreover, the few that respond are usually the students who are unhappy and thus will give highly negative evaluation.
+You will want to get good evaluations, ``above'' the average of the dept or college (which might be challenging because you are compared to all faculty, including teaching faculty who are very good at teaching).
+
+One issue with teaching evaluation is that not many students respond, and the few that do are usually the students who are unhappy and thus will give highly negative evaluation.
 This is the reason why many instructors do not like teaching evaluations, especially when it is used to evaluate their performance.
@@
 \begin{commentbox}[Vu]
-  A strategy that I found effective is that the whole class gets a small bonus point if everyone completes the evaluation.
+  In recent years I encourage student to do teaching evaluation by giving the whole class a small bonus point if everyone completes the evaluation.
 The bonus is calculated as the percentage of responses multiplied by a factor (X).
@@
 Since I don't curve grades, this is their only chance to boost their scores.
-Many students have benefited from this in my previous classes, which really grabs their attention.
-Plus, since evaluations happen before finals and grades aren't finalized, everyone's motivated to participate—they don't want to miss out.
+Moreover, since evaluations happen before finals and grades aren't finalized, everyone's motivated to participate as they don't want to miss out.
 Students not only do it themselves but also remind their friends.
@@
 
-  It's been working well for me: Spring '24 I had a 90\% response rate (25 out of 28 students), and Fall '23 was a perfect 100\% (16 out of 16 students).
+  This works well for me: Spring '24 I had a 90\% response rate (25 out of 28 students), and Fall '23 was a perfect 100\% (16 out of 16 students).
 Before, my response rate was about 50\%, which is already higher than the average.
@@
 
-In addition to student feedback, peer evaluations can offer valuable insights on your teaching. At many places, peer evaluations are required for P\&T.
-Request evaluations from experienced colleagues who can observe your class and provide constructive criticism.
-Like with student evaluation, incorporating these peer feedback into your teaching and describe your changes and improvements.
+In addition to student feedback, there is peer evaluation where a more experienced faculty member observes your class and provides feedback.
+This is usually required for P\&T (e.g., both GMU and UNL require it).
+Like with student evaluation, incorporating feedback from peer evaluation into your teaching and describe your changes and improvements.
 Peer evaluation letters can also help you get Teaching Awards and other recognitions.

### 2024-10-16 demystify.tex

--- 2024-10-15 demystify.tex
@@
 % tvn: the main thing is research experiences
-
-\section*{Additional Resources}
-\begin{itemize}
-  \item \href{https://parentheticallyspeaking.org/articles/us-cs-phd-faq/}{Getting a Computer Science PhD in the USA} by Shriram Krishnamurthi
-  \item  \href{https://pg.ucsd.edu/PhD-application-tips.htm}{Blog post on PhD Admission} by Philip Guo
-  \item   \href{https://talkingtorobots.com/FAQ.html}{PhD Admission post} by Yonatan Bisk (customized for CMU admission)
-\end{itemize}
 

### 2024-10-16 ennv.tex

--- 2024-10-15 ennv.tex
@@
 
-\begin{figure}
-    \centering
-    \includegraphics[width=0.5\linewidth]{figure/dnn.pdf}
-    \caption{\label{fig:dnn} An FNN with ReLU.}
-\end{figure}
+% \begin{figure}
+%     \centering
+%     \includegraphics[width=0.5\linewidth]{figure/dnn.pdf}
+%     \caption{\label{fig:dnn} An FNN with ReLU.}
+% \end{figure}
+
+\section{Types of Neural Networks}
+
+\paragraph{Feed-Forward Network (FFN)} In an FFN information flows in one direction, from the input layer to hidden layers to the output layer (and thus no cycle).
+
+A fully connected feed-forward neural network (FNN), shown below, is an FFN where every neuron in a layer is connected to every neuron in the next layer.
+\begin{center}
+\begin{tikzpicture}[scale=0.8, transform shape]
+% Input Layer
+\node at (0,4) {\textbf{Input Layer}};
+\foreach \x in {1,2,3,4} {
+    \node[circle, draw=black, fill=white, minimum size=0.6cm] (I\x) at (0,4-\x) {};
+}
+
+% Hidden Layer 1
+\node at (3,4) {\textbf{Hidden Layer 1}};
+\foreach \x in {1,2,3,4,5} {
+    \node[circle, draw=black, fill=white, minimum size=0.6cm] (H1\x) at (3,4.5-\x) {};
+}
+
+% Hidden Layer 2
+\node at (6,4) {\textbf{Hidden Layer 2}};
+\foreach \x in {1,2,3,4,5} {
+    \node[circle, draw=black, fill=white, minimum size=0.6cm] (H2\x) at (6,4.5-\x) {};
+}
+
+% Output Layer
+\node at (9,3.5) {\textbf{Output Layer}};
+\foreach \x in {1,2,3} {
+    \node[circle, draw=black, fill=white, minimum size=0.6cm] (O\x) at (9,3.5-\x) {};
+}
+
+% Connections
+\foreach \i in {1,2,3,4}
+    \foreach \j in {1,2,3,4,5}
+        \draw[->] (I\i) -- (H1\j);
+
+\foreach \i in {1,2,3,4,5}
+    \foreach \j in {1,2,3,4,5}
+        \draw[->] (H1\i) -- (H2\j);
+
+\foreach \i in {1,2,3,4,5}
+    \foreach \j in {1,2,3}
+        \draw[->] (H2\i) -- (O\j);
+
+\end{tikzpicture}
+\end{center}
+
+\paragraph{Convolutional Neural Networks (CNNs)} are a type of neural network that is often used in image recognition and classification.
+CNNs consist of of neurons that have learnable weights and biases.
+Each neuron receives several inputs, takes a weighted sum over them, passes it through an activation function, and responds with an output.
+
+\paragraph{Recurrent Neural Networks} Recurrent Neural Networks (RNNs) are a type of neural network that is often used in natural language processing and speech recognition.
+RNNs are designed to recognize patterns in sequences of data.
+RNNs have loops in them, allowing information to persist.
+This loop allows information to be passed from one step of the network to the next.
+
+\paragraph{Residual Networks} Residual Networks (ResNets) are a type of neural network that is often used in image recognition and classification.
+ResNets introduce skip connections that allow the gradient to flow directly through the network, making it easier to train deep networks.
 

### 2024-10-16 ennv.tex
update

@@
 The value of a neuron in the input layer is the input data.
-The value of a neuron in the hidden layers is computed by applying an \emph{affine transformation} (\autoref{sec:affine}) to values of neurons in the previous layers, then followed by an \emph{activation function} (\autoref{sec:activation}) such as the popular Rectified Linear Unit (ReLU) activation.
+The value of a neuron in the hidden layers is computed by applying an \emph{affine transformation} (\autoref{sec:affine}) to values of neurons in the previous layers, then followed by an \emph{activation function} (\autoref{sec:activation}) such as ReLU and Sigmoid.
+The value of a neuron in the output layer is computed similarly but may skip the activation function.
 
@@
 
-% \begin{figure}
-%     \centering
-%     \includegraphics[width=0.5\linewidth]{figure/dnn.pdf}
-%     \caption{\label{fig:dnn} An FNN with ReLU.}
-% \end{figure}
+\begin{figure}
+    \centering
+    \includegraphics[width=0.5\linewidth]{figure/dnn.pdf}
+    \caption{\label{fig:dnn} An FNN with ReLU.}
+\end{figure}
 
@@
 \section{Properties of Neural Networks}
+Similar to software programs, neural networks have desirable properties to ensure the network behaves as expected.
+These could be specific to the applications modeled by the network, e.g., safety properties in a collision avoidance system or general properties that are desired by all networks, e.g., robustness to adversarial attacks.
 
-\subsection{Robustness}
+\paragraph{Robustness Properties}
+\emph{Robustness}, a desirable property for all networks, ensures that small perturbations in the input data do not cause major changes in the output of the network.
+For example, if a few pixels in an image are changed, the network should still classify the image correctly.
+\emph{Adversarial attacks} are a common way to test the robustness of a neural network.
+In an adversarial attack, an attacker makes small changes to the input data to cause the network to misclassify the data.
 
-\subsection{Safety}
+\emph{Local} robustness refers to robustness of a neural network within a \emph{small neighborhood or region} of the input data.
+In contrast, \emph{global} robustness refers to robustness of a network across the \emph{entire input space}.
+Global robustness is harder to achieve than local robustness, as it requires the network to be robust to all possible inputs.
+
+\paragraph{$\epsilon$-robustness} A neural network is $\epsilon$-robust if the difference between any two inputs $x$ and $x'$ is within a small range $\epsilon$, the output $f$ of the network does not change significantly (or remain the same), i.e., $\|x-x'\| \leq \epsilon \implies f(x) \approx f(x')$.
+
+\paragraph{Safety Properties} Safety properties are specific to the application modeled by the network.
+For example, a safety property in a collision avoidance system might be that if the intruder is distant and significantly slower than us, then we stay below a certain threshold, i.e., $d_{intruder} > d_{threshold} \land v_{intruder} < v_{threshold} \implies v_{us} < v_{threshold}$.
+
+\subsection{Challenges}
+
+\paragraph{Formalization}
+
+\paragraph{Expressiveness}
 

### 2024-10-16 ennv.tex
update

@@
 
+\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
+
+\SetAlgoCaptionSeparator{.}
+\SetAlCapFnt{\footnotesize\rm}
+\SetAlCapNameFnt{\footnotesize\rm}
+
+\newcommand{\functiontextformat}[1]{\textrm{\texttt{#1}}}
+
+% \newtoggle{usecomment}
+% \settoggle{usecomment}{true}
+
+% \newcommand{\tcomment}[1]{\iftoggle{usecomment}{#1}{}}
+% \newcommand{\tvn}[1]{\iftoggle{usecomment}{{\color{red}{[TVN]: #1}}}{}}
+
+% \newcommand{\mbd}[1]{\iftoggle{usecomment}{{\color{magenta}{[MBD]: #1}}}{}}
+% \newcommand{\matt}[1]{\iftoggle{usecomment}{{\color{magenta}{[MBD]: #1}}}{}}
+
+% \newcommand{\hd}[1]{\iftoggle{usecomment}{{\color{blue}{[HD]: #1}}}{}}
+
+% \newcommand{\mynew}[1]{\iftoggle{usecomment}{{\color{red}{[NEW]: #1}}}{}}
+
+%\newcommand{\ignore}[1]{}
+
 \newtcolorbox{mybox}{
@@
 
+% for captions
+\def\Section{\S}
 \renewcommand{\figurename}{Fig.}
 \renewcommand{\tablename}{Tab.}
-\def\Section{\S}
+\renewcommand{\algorithmcfname}{Alg.}
+
+% for cross references
+% \renewcommand{\algorithmcflinename}{L.}
+\renewcommand{\algorithmautorefname}{Alg.}
 \renewcommand{\figureautorefname}{Fig.}
 \renewcommand{\tableautorefname}{Tab.}
-\makeatletter
+\renewcommand{\equationautorefname}{Eq.}
 \renewcommand{\chapterautorefname}{\S\@gobble}
@@
 \renewcommand{\appendixautorefname}{\S\@gobble}
+
+\newcommand{\nnproofgen}{\texttt{APTPgen}}
+\newcommand{\nnproofcheck}{\texttt{APTPcheck}}
+\newcommand{\nnproofchecker}{\texttt{APTPchecker}}
+\newcommand{\nnproofformat}{\texttt{APTP}}
+%\newcommand{\tool}{\texttt{NNProofChecker}}
+\newcommand{\veristable}{\texttt{VeriStable}}
+\newcommand{\crown}{\texttt{$\alpha\beta$-CROWN}}
+\newcommand{\nnenum}{\texttt{nnenum}}
+\newcommand{\marabou}{\texttt{Marabou}}
+\newcommand{\eran}{\texttt{ERAN}}
+\newcommand{\reluplex}{\texttt{Reluplex}}
+\newcommand{\reluval}{\texttt{Reluval}}
+\newcommand{\neurify}{\texttt{Neurify}}
+\newcommand{\nnv}{\texttt{NNV}}
+\newcommand{\ovalnnv}{\texttt{OVAL}}
+\newcommand{\dnnv}{\texttt{DNNV}}
+\newcommand{\verinet}{\texttt{VeriNet}}
+\newcommand{\mnbab}{\texttt{MN-BaB}}
+% \newcommand{\vnncomp}{VNN-COMP'23}
+\newcommand{\planet}{\texttt{Planet}}
+\newcommand{\dd}{\texttt{BaB$_{\text{NV}}$}}
+\newcommand{\neuralsat}{\texttt{NeuralSAT}}
+\newcommand{\mipverify}{\texttt{MIPVerify}}
+\newcommand{\crowndefault}{\texttt{$\alpha\beta$-CROWN} (default)}
+
+\makeatletter
 \makeatother
@@
 \chapter{Search Algorithms}
+\autoref{fig:alg} shows \dd{}, a reference architecture~\cite{nakagawa2014consolidating} 
+for modern DNN verifiers that we use to illustrate our observations.  
+\dd{} takes as input a ReLU-based DNN $\mathcal{N}$ and a formulae $\phi_{in}\Rightarrow \phi_{out}$ representing the property of interest.
+
+\dd{} iterates between two components: \texttt{Decide} (branching, \autoref{line:decide}), which decides (assigns) an activation status value for a neuron, and \texttt{Deduce} (bounding, \autoref{line:deduce}), which checks the feasibility of the current activation pattern.
+To add proof generation capability, \dd{} is instrumented with a proof tree (\texttt{proof}) variable (\autoref{line:prooftree}) to record these branching decisions.
+The proof is represented as a binary tree structure, where each node represents a neuron and its left and right edges represent its activation decision (active or inactive).
+%The proof tree is then used to generate a proof in the \nnproofformat{} format (\autoref{sec:proof-format}).
+
+\begin{figure}[t]
+    \begin{minipage}[b]{\linewidth}
+        \centering
+        \begin{minipage}[t]{0.48\textwidth}
+            \centering  
+            \includegraphics[width=\linewidth]{figure/proof_net.pdf}
+            \caption*{(a)}
+        \end{minipage}
+        \begin{minipage}[t]{0.48\textwidth}
+            \centering
+            \includegraphics[width=\linewidth]{figure/proof_tree.pdf}
+            \caption*{(b)}
+        \end{minipage}
+        \caption{(a) A simple DNN.  (b) A proof tree produced verifying the property $(x_1, x_2) \in [-2.0, 2.0] \times [-1.0, 1,0] \Rightarrow (y_1 > y_2)$.}
+        \label{fig:example}
+    \end{minipage}
+\end{figure}
+%White nodes correspond to branching nodes where \neuralsat{} makes decisions to split ReLU neurons.
+
+\paragraph{Example} \autoref{fig:example}a illustrates a DNN and how \dd{} determines unsatisfiability (i.e., verifies the problem) and generates the unsat proof in \autoref{fig:example}b.
+
+First, \dd{} initializes the activation pattern set \functiontextformat{ActPatterns} with an empty activation pattern $\emptyset$.
+Then \dd{} enters a loop (\autoref{line:dpllstart}-\autoref{line:dpllend}) to search for a satisfying assignment or a proof of unsatisfiability.
+In the first iteration, \dd{} selects the only available activation pattern $\emptyset \in \functiontextformat{ActPatterns}$.
+It calls \functiontextformat{Deduce} to check the feasibility of the problem based on the current activation pattern.
+\functiontextformat{Deduce} uses abstraction to approximate that from the input constraints the output values are feasible for the given network.
+Since \functiontextformat{Deduce} cannot decide infeasibility, \dd{} randomly selects a neuron to split (\functiontextformat{Decide}).
+Let us assume that it chooses $v_4$ to split, which essentially means the problem is split into two independent subproblems: one with $v_4$ active and the other with $v_4$ inactive.
+\dd{} then adds $v_4$ and $\overline{v_4}$ to \functiontextformat{ActPatterns}.
+
+In the second iteration, \dd{} has two subproblems (that can be processed in parallel).
+For the first subproblem with $v_4$, \functiontextformat{Deduce} cannot decide infeasibility, so it selects $v_2$ to split.
+It then conjoins $v_4$ with $v_2$ and then with $\overline{v_2}$ and adds both conjuncts to \texttt{ActPatterns}.
+For the second subproblem with $\overline{v_4}$ inactive, \functiontextformat{Deduce} determines that the problem is unsatisfiable and \dd{} saves the node $v_4$ to the proof tree, as node 3, to indicate one unsatisfiable pattern, i.e., whenever the network has $v_4$ being inactive, the problem is unsatisfiable.
+
+In the third iteration, \dd{} has two subproblems for $v_4 \land v_2$ and $v_4 \land \overline{v_2}$.
+For the first subproblem, \functiontextformat{Deduce} cannot decide infeasibility, so it selects $v_1$ to split.
+It then conjoins $v_1$ and then $\overline{v_1}$ to the current activation pattern and adds them to \functiontextformat{ActPatterns}.
+For the second subproblem, \functiontextformat{Deduce} determines that the problem is unsatisfiable and \dd{} saves the node $v_4 \land \overline{v_2}$ to the proof tree, as node 5.
+
+In the fourth iteration, \dd{} has two subproblems for $v_4 \land v_2 \land v_1$ and $v_4 \land v_2 \land \overline{v_1}$.
+Both subproblems are determined to be unsatisfiable, and \dd{} saves them to the proof tree as nodes 6 and 7, respectively.
+
+Finally, \dd{} has an empty \texttt{ActPatterns}, stops the search, and returns \texttt{unsat} and the proof tree.
 

### 2024-10-16 ennv.tex
update

@@
 \begin{itemize}
-\item ReLU (Rectified Linear Unit): ReLU is the most popular activation function in DNNs.
-It returns 0 if the input is less than zero, and the input itself otherwise.
-It is often used in hidden layers and skipped in the output layer.
-A ReLU activated neuron is said to be \emph{active} if its input value is greater than
-zero and \emph{inactive} otherwise.\\
+\item ReLU (Rectified Linear Unit):  ReLU returns 0 if the input is less than zero, and the input itself otherwise. 
+A ReLU activated neuron is said to be \emph{active} if its input value is greater than zero and \emph{inactive} otherwise.
+ReLU is the most popular activation function in DNNs.\\
     \begin{center}
@@
 \chapter{Search Algorithms}
+
+\begin{algorithm}[t]
+    \SetKwFor{Parfor}{parfor}{do}{endparfor}
+
+    \SetKwInOut{Input}{input}
+    \SetKwInOut{Output}{output}
+
+    \SetKw{Break}{break}
+    \SetKw{Continue}{continue}
+    \SetKw{In}{in}
+
+    \SetKwData{model}{model}
+    \SetKwData{layer}{layer}
+    \SetKwData{nextlayer}{layer$_{i+1}$}
+    \SetKwData{prooftree}{proof}
+    \SetKwData{leaf}{node}
+    \SetKwData{status}{status}
+    \SetKwData{minimum}{objval}
+    \SetKwData{mynull}{null}
+
+    \SetKwFunction{InputMILP}{AddInputConstrs}
+    \SetKwFunction{GetUnstableNeurons}{GetUnstableNeurons}
+    \SetKwFunction{PiecewiseLinearMILP}{AddConstrsPWL}
+    \SetKwFunction{LinearMILP}{AddConstrsLinear}
+    \SetKwFunction{Maximize}{Maximize}
+    \SetKwFunction{Minimize}{Minimize}
+    \SetKwFunction{Feasible}{CheckFeasibility}
+    \SetKwFunction{Optimize}{Optimize}
+    \SetKwFunction{isPiecewiseLinear}{isPiecewiseLinear}
+    \SetKwFunction{CreateStabilizedMILP}{CreateStabilizedMILP}
+    \SetKwFunction{GetLeafNodes}{GetLeafNodes}
+    \SetKwFunction{AddConstrs}{AddConstrs}
+    \SetKwFunction{RemoveConstrs}{RemoveConstrs}
+    \SetKwFunction{AddObjective}{AddObjectives}
+    \SetKwFunction{ShortenSplitConstrs}{ShortenSplitConstrs}
+    \SetKwFunction{RemoveLeafNodes}{RemoveLeaves}
+    \SetKwFunction{StoppingConditions}{StoppingConditions}
+    \SetKwFunction{RepOK}{RepOK}
+    \SetKwFunction{RaiseError}{RaiseError}
+
+    \SetKwInOut{Input}{input}
+    \SetKwInOut{Output}{output}
+    \SetKw{Break}{break}
+    \SetKw{Continue}{continue}
+    \SetKwFunction{Backtrack}{Backtrack}
+    \SetKwFunction{Select}{Select}
+    \SetKwFunction{Decide}{Decide}
+    \SetKwFunction{BCP}{BCP}
+    \SetKwFunction{Deduce}{Deduce}
+    \SetKwFunction{AnalyzeConflict}{AnalyzeConflict}
+    \SetKwFunction{BooleanAbstraction}{BooleanAbstraction}
+    \SetKwFunction{AddClause}{AddClause}
+    \SetKwFunction{isTotal}{isTotal}
+    \SetKwFunction{randomattack}{RandomAttack}
+    \SetKwFunction{pgd}{PGDAttack}
+
+    \SetKwFunction{DPLLT}{DPLLT}
+    \SetKwFunction{isValid}{isValid}
+    \SetKwFunction{isEmpty}{isEmpty}
+    \SetKwFunction{LPSolver}{LPSolver}
+    \SetKwFunction{Solve}{Solve}
+    \SetKwFunction{FindLayerNodes}{FindLayerNodes}
+    \SetKwFunction{TightenInputBounds}{TightenInputBounds}
+    \SetKwFunction{Abstract}{Abstract}
+    \SetKwFunction{Check}{Check}
+    \SetKwFunction{Decide}{Decide}
+    \SetKwFunction{Imply}{Imply}
+    \SetKwFunction{Lower}{LowerBound}
+    \SetKwFunction{Upper}{UpperBound}
+    \SetKwFunction{GetInputBounds}{GetInputBounds}
+    \SetKwFunction{GetInputs}{GetInputs}
+    \SetKwFunction{GetNumInputs}{GetNumInputs}
+    \SetKwFunction{CurrentConflictClause}{CurrentConflictClause}
+    \SetKwFunction{StopCriterion}{StopCriterion}
+    \SetKwFunction{LastAssignedLiteral}{LastLiteral}
+    \SetKwFunction{LiteralToVariable}{LiteralToVariable}
+    \SetKwFunction{Antecedent}{Antecedent}
+    \SetKwFunction{BinRes}{BinRes}
+    \SetKwFunction{BacktrackLevel}{BacktrackLevel}
+    \SetKwFunction{AddClause}{AddClause}
+    \SetKwFunction{ActivationStatus}{ActivationStatus}
+    \SetKwFunction{Backjump}{Backjump}
+    \SetKwFunction{EstimateBounds}{EstimateBounds}
+
+    \SetKwData{problems}{ActPatterns}
+    \SetKwData{implicationgraph}{igraph}
+    \SetKwData{literal}{lit}
+    \SetKwData{variable}{var}
+    \SetKwData{antecedent}{ante}
+    \SetKwData{conflict}{conflict}
+    \SetKwData{none}{none}
+    \SetKwData{layerid}{lid}
+    \SetKwData{hiddenbounds}{hidden\_bounds}
+    \SetKwData{layerbounds}{layer\_bounds}
+    \SetKwData{inputs}{inputs}
+    \SetKwData{inputbounds}{input\_bounds}
+    \SetKwData{outputbounds}{output\_bounds}
+    \SetKwData{infeasible}{INFEASIBLE}
+    \SetKwData{unreachable}{UNREACHABLE}
+    \SetKwData{maxinputs}{MAX\_NUM\_INPUT}
+    \SetKwData{assignment}{$\sigma$}
+    \SetKwData{dl}{dl}
+    \SetKwData{lpmodel}{solver}
+    \SetKwData{clauses}{clauses}
+    \SetKwData{conflict}{conflict}
+    \SetKwData{clause}{clause}
+    \SetKwData{igraph}{igraph}
+    \SetKwData{cex}{cex}
+    \SetKwData{sat}{sat}
+    \SetKwData{unsat}{unsat}
+    \SetKwData{certified}{certified}
+    \SetKwData{uncertified}{uncertified}
+    \SetKwData{submodel}{sub\_model}
+
+    \SetKwData{true}{true}
+    \SetKwData{false}{false}
+    \DontPrintSemicolon
+
+    \SetKwFunction{Restart}{Restart}
+
+    \SetKwData{counterexample}{cex}
+    \SetKwData{conflictclause}{conflict\_clause}
+    \SetKwData{isconflict}{is\_conflict}
+
+    \footnotesize
+    \DontPrintSemicolon
+
+    \Input{DNN $\mathcal{N}$, property $\phi_{in} \Rightarrow \phi_{out}$}
+    Output{($\unsat, \prooftree$)}
+    \Output{($\unsat, \prooftree$) if property is valid, otherwise ($\sat, \counterexample$)}
+    \BlankLine
+
+    $\problems \leftarrow \{ \emptyset \}$ \tcp{initialize verification problems} 
+    $\prooftree \gets \{ ~ \}$ \tcp{initialize proof tree}\label{line:prooftree}
+
+    \While(\tcp*[h]{main loop}){$\problems$}{\label{line:dpllstart}
+        % \tcp{$\sigma_i$ is the activation pattern of problem $i$-th}
+        $\sigma_i \gets \Select(\problems)$ \tcp{process problem $i$-th}
+        % \Parfor(\tcp*[h]{process in parallel}){$\sigma_i ~\In~ \problems$}{ \label{line:parfor}
+            \If{\Deduce{$\mathcal{N}, \phi_{in}, \phi_{out}, \sigma_i$}}{\label{line:deduce}
+                $(\counterexample, v_i) \leftarrow \Decide(\mathcal{N}, \phi_{in}, \phi_{out}, \sigma_i)$ \\ \label{line:decide}
+                \If(\tcp*[h]{found a valid counter-example}){$\counterexample$}{
+                    \Return{$(\sat, \counterexample)$} 
+                }
+                \tcp{create new activation patterns}
+                $\problems \leftarrow \problems \cup \{ \sigma_i \land v_i ~;~ \sigma_i \land \overline{v_i} \}$ \;
+            }
+            \Else(\tcp*[h]{detect a conflict}){
+                % $\clauses \leftarrow \clauses \cup \AnalyzeConflict(\igraph_i)$ \\ \label{line:analyze_conflict}
+                $\prooftree \leftarrow \prooftree \cup \{ \sigma_i \}$ \tcp{build proof tree} \label{line:record_proof}
+            }
+        % }
+        % \If(\tcp*[h]{no more problems}){$\isEmpty(\problems)$}{
+        % }
+
+    }\label{line:dpllend}
+    \Return{$(\unsat, \prooftree)$}
+
+    \caption{The \dd{} algorithm with proof generation.}\label{fig:alg}
+\end{algorithm}
+
 \autoref{fig:alg} shows \dd{}, a reference architecture~\cite{nakagawa2014consolidating} 

### 2024-10-21 demystify.tex
updates

@@
 Specifically, they are worried about (i) the \textbf{difficulty of getting admitted} and (ii) the \textbf{cost of graduate study} in the US.
-Part of the reasons is the lack of information and guidance on the admission process in the US (let's face it, public forums like Facebook communities and Reddit are full of confusion and concerns from students).
-This section and the rest of this document aim to address these concerns.
+Part of the reasons is the lack of information and guidance on the admission process in the US (let's face it, social places like Facebook and Reddit are full of confusion and contradictory information).
+This document aims to address these concerns.
 
@@
 It might even be more flexible since CS PhD in the US \emph{do not} require having an MS or a research topic, proposal, or adviser in advance (\autoref{sec:non-us-differences} compares CS PhD study in the US to other countries).
+It doesn't even require having a CS background (\autoref{sec:non-stem}).
 If you believe you have a chance in other countries, e.g., South Korea, Singapore, Germany, UK, Japan, and Australia, then you will surely have a chance in the US as well.
@@
 health insurance, and stipend (i.e., you get paid for your study). 
-In fact, full funding for CS PhD students is the norm in the US, and I'd go as far as to say that if you are not admitted with full funding, you should be cautious about the program and might want to not accept the offer.
+For many students, stipends are not only sufficient for them to live but also for their families (e.g., spouse and children) to live comfortably.
+
 Moreover, depending on the university, you may receive additional benefits such as summer pay, laptops, and traveling to conferences and workshops. 
+Full funding for CS PhD students is the norm in the US, and I'd go as far as to say that if you are not admitted with full funding, you might want to not accept the offer.
 
@@
 I hope to encourage more students with such backgrounds to apply to CS PhD programs in the US.
-
-Another goal is to explain the \emph{``why''} behind the admission process, e.g., why LoRs are important, why you should not draft your own LoR, why you should contact professors, etc.
+Another goal of this document is to explain the \emph{``why''} behind the admission process, e.g., why LoRs are important, why you should not draft your own LoR, why you should contact professors, etc.
 
@@
 
-\paragraph{What next?} Make a decision, accept, reject, or defer the offers  (\autoref{sec:accept-postpone-decline}).
+\paragraph{What's next?} Make a decision, accept, reject, or defer the offers  (\autoref{sec:accept-postpone-decline}).
 Ask to meet with potential advisers (e.g., through Open House or separately) and even their students.
@@
 Ultimately, this choice is personal, involving a balance between personal benefit and the ethical impact on others.
+
+\section{Negotiating PhD offer (e.g., having multiple offers)?}\label{sec:negotiate}
+\sectioninfo{You will not be able to negotiate stipend, but you can ask for specific start date, TA assignment, and conference travel budget.}
+
+It is unlike that you can negotiate TA stipend, which is set by the university.
+Moreover, while RA is set by the adviser, it is often based on the department's standard, and it also would be unfair to other students if you get more (especially since you're new and have not proven yourself yet).
+However, your stipend is often increased every year by the university (automatically by a small amount).
+
+For a specific start date or TA assignment (e.g., TA'ing a particular course), you can ask for it.
+Also, there is typically no moving allowance for PhD students.
+In short, standard things set by the university or department are unlikely negotiable.
+However, you can ask for things such as computer equipment (\autoref{sec:buying-equipment}), books, (and maybe even furniture), and a conference travel budget.
 
@@
 
-\paragraph{Types of Faculty} There are typically two types of faculty: tenure-line and teaching.
+\paragraph{Types of Faculty} There are typically two types of faculty at an R1 research university: tenure-line and teaching.
 The main difference is their focus: research, teaching, or both.
-The type of faculty can affect their ability to serve as your adviser.
-
-\textbf{Tenure-line faculty}, consisting of tenured and tenure-track profs., focus on research, which includes publishing papers, obtaining grants, and mentoring PhD students.
+Tenure-line faculty are typically more research-focused and can mentor PhD students, while teaching faculty are more teaching-focused and typically do not advise PhD students.
+Thus when contacting a prof.
+for research opportunities (\autoref{sec:contact}), you should focus on tenure-line faculty (and not adjunct or emeritus :p).
+
+\textbf{Tenure-line faculty}, consisting of tenured and tenure-track profs., focuses on research, which includes publishing papers, obtaining grants, and mentoring PhD students.
 They often have a low teaching load, e.g., 1 course per semester.
@@
 
-\textbf{Teaching faculty} (sometimes called instructional faculty or professor of practice) mainly focuses on teaching.
+\textbf{Teaching faculty}, sometimes called instructional faculty or professor of practice, mainly focuses on teaching.
 They typically teach 3--4 classes per semester (which is quite a lot) and \emph{do not} have research responsibilities, e.g., they do not have to worry much about publishing papers or obtaining grants.
 As such they typically do not have funding (e.g., RA~\autoref{sec:funding}) to support PhD students.
-Depending on the university, they might be able to \emph{co-advise} and mentor PhD students with a tenure-line faculty but typically do not formally advise and graduate PhD students.
-
-Note that at many universities do not allow \emph{non-CS profs}, even in tenure-line ones, to solely advise and graduate PhD students in CS (see \autoref{sec:related-fields}).
+Depending on the university, they might be able to \emph{co-advise} and mentor PhD students with a tenure-line faculty, but they typically do not formally advise and graduate PhD students.
+
+Note that at many universities do not allow \emph{non-CS profs}, even tenure-line ones, to solely advise and graduate PhD students in CS (see \autoref{sec:related-fields}).
 
@@
 
+%\section{"Chance me" questions}\label{sec:chance-me}
+
+%Reddit and other grad admission forums are full of ``Chance me'' and ``roast my profile'' questions.
+These are questions where students post their profiles and ask others to evaluate their chances of getting admitted to their choice of schools.
+
+%These questions are not useful and should be avoided.
+First, no one knows!
+Even adcom members can't tell if you would get admitted to their programs or not.
+As mentioned throughout this guide, the admission process is complex and involves many factors that are not part of your profile, e.g., whether the faculty have funding or are interested in your background.
+Second, answers to this questions are often not helpful to you.
+Faculty and people knowledgeable about the process will not answer these questions, and those who do are often not knowledgeable and might give you false hope or despair.
+OK, someone say your GRE is low (which does not even matter for most schools (\autoref{sec:standard-tests}) , are you going to retake it?
+Or someone says your GPA is low, are you going to apply to a different school?
+No, you should apply to the schools you want to go to and see what happens.
+If you don't get in, then you can ask for feedback and improve your profile for the next round.
+
 \section{How long does it take to evaluate an application?}\label{sec:ievaluate}
-\sectioninfo{It takes me about 10 minutes to review an application.}
-
-On average, I spend about 10 minutes reviewing each application (less for clear rejections and more for potential acceptances).
+\sectioninfo{It takes me about 10--15 minutes to review an application.}
+
+On average, I spend about 10--15 minutes reviewing each application (less for clear rejections and more for potential acceptances).
 While this seems short, it is not that difficult to tell if an application is good or bad.
@@
 I also recommend very strong candidates for the University Presidential fellowship.
-Note that while the system has other decision options, eg., admit without funding, provisional admission (e.g., if they need to take some courses), I rarely use them, simply because we already have too many good students and do not have the resources to support them.
+Note that while the system has other decision options, eg., admit without funding, provisional admission (e.g., if they need to take some courses), I rarely use them, simply because we want to fully fund our PhD students.
 
@@
 Many universities have a strong commitment to diversity and inclusion and actively
-\emph{recruit students from underepresentative backgrounds} (e.g., some \href{https://cse-climate.engin.umich.edu/reports/climate-dei-reports/cse-climate-and-dei-report-2022-2023/#grad-ethnicity}{stats from UMich CSE} and \href{https://www.Mason.edu/news/2022-09/mason-now-top-10-public-university-diversity-innovation-and-cybersecurity-education-us}{Mason} often promotes itself as one of the most diverse universities in the US).
-Moreover, many scholarships and fellowships are created specifically for URM and LGBTQ+ students, which you should explore if qualify.
+\emph{recruit students from underepresentative backgrounds} (e.g., some \href{https://cse-climate.engin.umich.edu/reports/climate-dei-reports/cse-climate-and-dei-report-2022-2023/#grad-ethnicity}{stats from UMich CSE} and \href{https://www.Mason.edu/news/2022-09/mason-now-top-10-public-university-diversity-innovation-and-cybersecurity-education-us}{Mason} often tout
+itself as one of the most diverse universities in the US). Moreover, many scholarships and fellowships are created specifically for URM and LGBTQ+ students, which you should explore if qualify.
 
@@
 Moreover, promoting diversity aligns with the goal of broadening participation in computing (BPC) that is strongly supported within academia (e.g., see the \href{https://plans.bpcnet.org/GeorgeMasonUniversity_ComputerScience_DepartmentalBPCPlan.pdf}{BPC plan at Mason CS}) and government funding agencies such as NSF (e.g., see \href{https://new.nsf.gov/cise/broadening-participation}{here} and \href{https://www.nsf.gov/pubs/2022/nsf22125/nsf22125.jsp}{here}).
-Also, sometimes providing this information can make it easier for faculty to apply for certain scholarships or fellowships to support you, e.g., from a private donor that has offered a fellowship only to students from certain backgrounds.
+Providing this information can also make it easier for faculty to apply for certain scholarships or fellowships to support you, e.g., from a private donor that has offered a fellowship only to students from certain backgrounds.
 
 Ultimately, this is a personal choice and you do what you feel comfortable.
-If you do not feel comfortable disclosing this, you should not feel pressured to do so and you should also ask your LoR writers to not disclose this information.
-Note that you can still talk about your experiences and perspectives without mentioning your identity.
-If uncertain, you can seek additional guidance from your close professors and collaborators However, it is important to recognize that diversity is strongly valued and celebrated within CS programs in the US, and that you should feel empowered to share your unique experiences and perspectives.
+If you do not feel comfortable disclosing this, you should not feel pressured to do so.
+However, if you feel comfortable but are unsure if it is accepted in higher education in the US, then I hope this section helps you understand that it is not only accepted but also valued and celebrated.
+
+%If uncertain, you can seek additional guidance from your close professors and collaborators.
 
@@
 You will likely need to return them when you graduate.
-
-\section{Negotiating PhD offer (e.g., having multiple offers)?}\label{sec:negotiate}
-\sectioninfo{You will not be able to negotiate stipend, but you can ask for specific start date, TA assignment, and conference travel budget.}
-
-It is unlike that you can negotiate TA stipend, which is set by the university.
-Moreover, while RA is set by the adviser, it is often based on the department's standard, and it also would be unfair to other students if you get more (especially since you're new and have not proven yourself yet).
-However, your stipend is often increased every year by the university (automatically by a small amount).
-
-For a specific start date or TA assignment (e.g., TA'ing a particular course), you can ask for it.
-Also, there is typically no moving allowance for PhD students.
-In short, standard things set by the university or department are unlikely negotiable.
-However, you can ask for things such as computer equipment (\autoref{sec:buying-equipment}), books, (and maybe even furniture), and a conference travel budget.
 

### 2024-10-21 ennv.tex
updates

@@
 \usepackage[margin=1.5in]{geometry}
+\usepackage{amsmath}
 \usepackage{soul}
@@
 
+\makeatletter
 % for cross references
@@
 \renewcommand{\appendixautorefname}{\S\@gobble}
+
+\makeatother
 
@@
 \newcommand{\crowndefault}{\texttt{$\alpha\beta$-CROWN} (default)}
-
-\makeatletter
-\makeatother
 
@@
 A \emph{neural network} (\textbf{NN})~\cite{Goodfellow-et-al-2016} consists of an input layer, multiple hidden layers, and an output layer.
-Each layer has a number of neurons, each connected to neurons from previous layers through a predefined set of weights (derived by training the network with data).
-A \emph{Deep Neural Network} (\textbf{DNN}) is an NN with at least two hidden layers.
+Each layer has a number of neurons, each connected to neurons in the next layer through a predefined set of weights (derived by training the network with data).
+A \emph{Deep Neural Network} (\textbf{DNN}) is an NN with two or more hidden layers.
 
@@
 
-The output of a DNN is obtained by iteratively computing  the  values  of  neurons  in  each  layer.
+The output of an NN is obtained by iteratively computing  the  values  of  neurons  in  each  layer.
 The value of a neuron in the input layer is the input data.
@@
 \chapter{Verification of Neural Networks}\label{sec:verification}
+
+\section{The Neural Network Verification Problem and Its Formalization}
 \paragraph{DNN Verification} Given a DNN \(N\) and a property $\phi$, the \emph{DNN verification problem} asks if $\phi$ is a valid property of $N$.
@@
 
+\section{Satisfiability and Activation Pattern Search} As with traditional software, DNN verification is often represented as a satisfiability problem.
+We thus need to define a formula to represent the network.
+Typically this formula is a conjunction of constraints representing the affine transformation and activation function of each neuron in the network.
+For a network with $L$ layers, $N$ neurons per layer, and ReLU activations this formula is:
+\begin{align*}
+\alpha = \bigwedge_{\begin{smallmatrix}i \in [1,L]\\ j \in [1,N]\end{smallmatrix}} v_{i,j} = \max \Big( \sum_{k \in [1,N]} (w_{i-1,j,k} \cdot v_{i-1,j}) + b_{i,j}, 0 \Big)
+\end{align*}
+With this definition DNN verification can be formulated as checking the satisfiability of:
+\begin{equation}\label{eq:prob}
+  \alpha \land \phi_{in} \land \neg \phi_{out}
+\end{equation}
+If \autoref{eq:prob} is unsatisfiable, $\phi$ is a valid property of $\mathcal{N}$.
+Otherwise, $\phi$ is not valid (and the counterexample of the original problem is any satisfying assignment that makes \autoref{eq:prob} true).
+
+For the widely-used ReLU activation problem, this problem becomes a search for \emph{activation patterns}, i.e., boolean assignments representing activation status of neurons, that lead to satisfaction the formula in \autoref{eq:prob}.
+Modern DNN verification techniques~\cite{bunel2020branch,wang2021beta,ferrari2022complete,duong2024harnessing,duong2023dpllt,ovalbab,katz2019marabou,bak2021nnenum} all adopt this idea and search for satisfying activation patterns.
+
 \section{Complexity}
 
-\chapter{Search Algorithms}
+\chapter{Common Search Algorithms}
+
+\section{Symbolic Search and Constraint solving}
+\section{Branch-and-Bound (BnB)}
 

### 2024-10-21 demystify.tex
updates

@@
 A common question is whether you should mention, e.g., in your SOP, that you are a member of an underrepresented and minority (URM) group\footnote{URM in CS in the US include African Americans, Hispanic Americans, American Indians/Native Alaskans, Native Hawaiians/Pacific Islanders (excluding Asian Americans), and multiracial students identifying with at least one of the previously listed URM categories.} or LGBTQ+ group.
-This is a strong concern for international (and even domestic) students, who may have encountered discrimination in their countries or personal experiences.
+The main concern is that mentioning this might make you look like you are asking for special treatment or that you might get discriminated against.
+
+%This is a strong concern for international (and even domestic) students, who may have encountered discrimination in their countries or personal experiences. 
 %This question arises often for international (and even domestic) students.
@@
 Many universities have a strong commitment to diversity and inclusion and actively
-\emph{recruit students from underepresentative backgrounds} (e.g., some \href{https://cse-climate.engin.umich.edu/reports/climate-dei-reports/cse-climate-and-dei-report-2022-2023/#grad-ethnicity}{stats from UMich CSE} and \href{https://www.Mason.edu/news/2022-09/mason-now-top-10-public-university-diversity-innovation-and-cybersecurity-education-us}{Mason} often tout
-itself as one of the most diverse universities in the US). Moreover, many scholarships and fellowships are created specifically for URM and LGBTQ+ students, which you should explore if qualify.
+\emph{recruit students from underepresentative backgrounds} (e.g., some \href{https://cse-climate.engin.umich.edu/reports/climate-dei-reports/cse-climate-and-dei-report-2022-2023/#grad-ethnicity}{stats from UMich CSE} and GMU often touts itself as one of the most diverse universities in the US).
+Moreover, many scholarships and fellowships are created specifically for URM and LGBTQ+ students, which you should explore if qualify.
 
@@
 If you do not feel comfortable disclosing this, you should not feel pressured to do so.
-However, if you feel comfortable but are unsure if it is accepted in higher education in the US, then I hope this section helps you understand that it is not only accepted but also valued and celebrated.
+However, if you feel comfortable but are unsure if it is accepted in higher education in the US, then I hope this section helps you understand that it is not only accepted but also valued and appreciated.
 

### 2024-10-21 demystify.tex
updates

@@
 If you do not feel comfortable disclosing this, you should not feel pressured to do so.
-However, if you feel comfortable but are unsure if it is accepted in higher education in the US, then I hope this section helps you understand that it is not only accepted but also valued and appreciated.
+However, if you feel comfortable but are unsure if it is accepted in higher education in the US, then I hope this helps you understand that it is not only accepted but also valued and appreciated.
 

### 2024-10-22 demystify.tex
updates

@@
 A common question is whether you should mention, e.g., in your SOP, that you are a member of an underrepresented and minority (URM) group\footnote{URM in CS in the US include African Americans, Hispanic Americans, American Indians/Native Alaskans, Native Hawaiians/Pacific Islanders (excluding Asian Americans), and multiracial students identifying with at least one of the previously listed URM categories.} or LGBTQ+ group.
-The main concern is that mentioning this might make you look like you are asking for special treatment or that you might get discriminated against.
+The main concern is that you can be discriminated against or that you are asking for special treatment.
 
@@
 
-  \item[RA and TA] Research Assistantship (RA) is form of funding where you work on a research project for a professor.
-Teaching Assistantship (TA) is a form of funding where you help a a professor with teaching (e.g., grading assignments).
+  \item[RA and TA] Research Assistantship (RA, \autoref{sec:ra}) is a funding where you work on a research project for a professor.
+Teaching Assistantship (TA, \autoref{sec:ta}) is a funding where you help a a professor with teaching (e.g., grading assignments).
 

### 2024-10-22 demystify.tex
updates

@@
 There are other agencies that support RAs but they might require US citizenship or permanent residency.
+
   \item [R1/R2 Universities] Research 1 (R1) universities are universities with the highest level of research activity.
@@
 
-While both MS and PhD programs are graduate degrees, they are \emph{very} different in terms of objective, admission requirements, course requirements, duration, and funding.
+While both MS and PhD programs are graduate degrees, they are \emph{very different} in terms of objective, admission requirements, course requirements, duration, and funding.
 This section provides guidance on applying to MS programs in CS.
@@
     \caption{MS vs. PhD}\label{tab:phd-vs-ms}
+    \centering
+    \small
     \begin{tabular}{c|c|c}
@@
 
-Except for a very few highly selective programs, MS programs are typically much less competitive than PhD programs.
-Some refer to MS programs as a \emph{cash cow} for the department, as MS students are often not funded and have to pay tuition (i.e., you're likely to get in if you can afford it).
+Except for a very few highly selective programs, MS programs are typically much less competitive than PhD programs, i.e., you're likely to get in if you can afford it.
+Some refer to MS programs as a \emph{cash cow} for the department, as MS students are often not funded and have to pay tuition.
 
 While admission requirements are similar to PhD programs (e.g., GPA, SOP, LORs), research is not a focus in MS programs.
-In most cases, the main requirements are that you have sufficient background in CS (e.g., through your undergrad degree).
+In most cases, the main requirements are that you have sufficient background in CS, e.g., through your undergrad degree.
 
@@
 However, MS admission is typically rolling (i.e., they review applications as they come in and make decisions throughout the admission cycle).
-Moreover, unlike PhD that has multiple reviewers for each application, an MS application typically involves only one reviewer and does not review much time to review compared to a PhD application (\autoref{sec:ievaluate}).
+Moreover, unlike PhD that has multiple reviewers for each application, an MS application typically involves only one reviewer and does require much time to review compared to a PhD application (\autoref{sec:ievaluate}).
 
@@
 \subsection{Funding}\label{sec:ms-funding}
-Unlike PhD programs, which often have funding (\autoref{sec:funding}), MS students are typically not funded through either RA or TA.
+Unlike PhD programs, which often have funding (\autoref{sec:funding}), MS students are typically not funded.
+The reason is because MS students do not focus on research and thus are not funded through RA and PhD students have priority for TA positions.
 \begin{itemize}
@@
 
-This section lists some cultural issues that students, especially international ones, might want to pay attention to. These issues apply in general and not just PhD admission.
+This section lists some general cultural issues that students, especially international ones, might want to pay attention to.
 
@@
 
-  Moreover, avoid asking personal questions, e.g., about salary, age, relationship status, or health, as these are considered private (talking about kids' activities or schools are OK).
-Also, do not talk about politics or religion in general (in fact, we often do not talk about these with our family and friends to avoid conflicts).
+Moreover, avoid asking personal questions, e.g., about salary, age, relationship status, or health, as these are considered private.
+Talking about kids' activities or schools are OK.
+Also, do not talk about politics or religion.
+In fact, we often do not talk about these taboo subjects with our family and friends to avoid conflicts.
+
 Sometimes foreigners are surprised by how Americans do not talk about their personal lives, e.g., sharing details about their families, health, or relationships, and that their conversations are often not very ``deep'' or ``mind-provoking''.
@@
 
-  \paragraph{Car Dependency} Most of the US is highly car-dependent.
+  \paragraph{Car Dependency} Most places in the US are highly car-dependent.
 If you do not have a car, you will need to rely on friends, Uber, or public transportation, which can be inconvenient and time-consuming.
-Most international students end up getting a driver's license and buying a car.
-
+Many international students end up getting a driver's license, which is highly convenient and replaces many documents (e.g., ID, passport) and buying a car.
+
+  %\paragraphP{Conversation} Americans are generally friendly and open.  However, there are certain things that we rarely talk about, e.g., salary, relationship status, or health. 
+  These are considered private and should not be asked.
+Also, do not talk about politics or religion in general (in fact, we do not even talk about these with our family and friends to avoid conflicts).
+
+  %Sometimes foreigners are surprised by how Americans do not talk about their personal lives, e.g., sharing details about their families, health, or relationships, and that their conversations are often not very ``deep'' or ``mind-provoking''.
+This is just a cultural norm about privacy and personal space.
+
+  %This can be surprising for people from countries where strangers do not typically talk to each other.
+Americans are also generally informal and use first names, even with people they have just met.
+This can be surprising for people from countries where titles and last names are used, even with close friends.
 	%\paragraph{Individualism} The strong culture of individualism in the US can be both liberating and isolating.

### 2024-10-22 demystify.tex
updates

@@
 \begin{commentbox}[Vu:]
-  I often read SOP and LORs first (\autoref{sec:ievaluate}). If I am
+  I often read LORs and SOP first (\autoref{sec:ievaluate}). If I am
   persuaded by then, I would skim over other factors and advocate for
@@
 
-% \subsection{Dos and Don'ts of SOP}
-% \paragraph{Do's}
-% \begin{enumerate}
-%     \item \textbf{Be Specific}: Clearly articulate your research interests and why you are drawn to the program.
-%     \item \textbf{Show Enthusiasm}: Demonstrate genuine passion for your field of study.
-%     \item \textbf{Proofread}: Ensure your statement is free of grammatical and typographical errors.
-%     \item \textbf{Be Personal}: Include relevant personal stories that highlight your journey and motivations.
-%     \item \textbf{Tailor the Statement}: Customize your statement for each program you apply to.
-% \end{enumerate}
-
-% \paragraph{Don'ts}
-% \begin{enumerate}
-%     \item \textbf{Avoid Generic Statements}: Don’t use vague or general phrases.
-%     \item \textbf{Don't Repeat Your Resume}: Provide new insights or context beyond what’s listed in your CV.
-%     \item \textbf{Steer Clear of Jargon}: Avoid overusing technical jargon that might confuse readers.
-%     \item \textbf{Don’t Exaggerate}: Avoid excessive self-praise without evidence.
-%     \item \textbf{Avoid Negative Tones}: Don’t criticize past institutions or colleagues.
-% \end{enumerate}
-%\subsection{Common Pitfalls to Avoid}\label{sec:sop-pitfalls}
-
-%SOP is crucial for making a strong impression on adcom reviewers. This outlines common pitfalls to avoid in your SOP.
-
-%\begin{enumerate}
-%m \textbf{Generic Statements}: Using vague or general statements that could apply to any applicant.
-For example, saying blank statements such as “I am passionate about computer science” without providing specific examples or evidence.
-%    \item \textbf{Not Customized}: Failing to mention why you chose this univ and faculty members you want to work with.
-If you can submit the same SOP to multiple programs with few changes, it is not customized enough.
-%    \item \textbf{Too Long}: Adcom members have to read many applications and a long SOP can be a turn-off. Keep it under 2 pages.
-%    \item \textbf{Lack of Passion or Motivation}: Failing to convey genuine enthusiasm for research and the specific program.
-%    \item \textbf{Grammatical and Typographical Errors}: Submitting an SOP filled with grammatical mistakes, typos, and poorly constructed sentences.
-%    \item \textbf{Excessive Self-Praise}: Overly boasting about achievements without providing concrete evidence or examples to back them up.
-%    \item \textbf{Irrelevant Information}: Including personal anecdotes or experiences that do not relate to the field of study or research interests.
-%    \item \textbf{Unrealistic Goals}: Stating research goals that are not feasible or that show a lack of understanding of the field’s current state.
-%    \item \textbf{Negative Tone}: Criticizing previous institutions, professors, or colleagues.
-%    \item \textbf{Failure to Address Fit}: Not explaining why you are a good fit for the program and how the program is a good fit for your goals.
-%    \item \textbf{Generic Opening and Closing Statements}: Starting or ending the SOP with clichés or unoriginal statements that do not add value.
-%    \item \textbf{Lack of Customization}: Submitting the same SOP to multiple programs without tailoring it to each specific program's strengths and faculty.
-%\end{enumerate}
+\subsection{Kiss of Death}
+
+\begin{itemize}
+\item \textbf{Too personal} Don't talk about your personal issues, e.g., family, health, relationship, etc.
+This can raise concerns about your ability to handle the PhD.
+Don't talk about religious or political beliefs (\autoref{sec:cultural}).
+\item \textbf{Criticizing} your current or previous institution or professors: just like in a job interview, don't badmouth your current or previous employer. 
+\item \textbf{No concrete evidence to back up claims}: e.g., saying you are passionate about a research topic without showing any experience.
+This is where specific names (work submitted at conf.
+X), numbers (outperformed SOTA by Y \%), and examples (worked on project Z) can help.
+\item \textbf{Use flowery and AI-like language}: don't use AI to write your SOP. Not that hard to raise suspicion.
+\item \textbf{Not customized to the program}: If your SOP can be sent to multiple programs with few changes, it is too generic.
+Do some homework and mention why you want to spend the next 7 years there.
+\item \textbf{Mentioning wrong professors}: do not mention emeritus professors or those who have left.
+Do your homework to see if the professors you mention are still active in research.
+Teaching and adjunct faculty are often not active in mentoring PhD students (\autoref{sec:faculty-types}).
+\item \textbf{Too Long and fancy format}: As mentioned, keep it under 2 pages.
+Don't use too much coloring or fancy fonts (like those in Words).
+CS academics like using LaTeX (common way to write our papers and other documents), so write your SOP using LaTeX.
+\end{itemize}
 

### 2024-10-22 demystify.tex
updates

@@
 
-This section discusses various research experiences that can strengthen your application.
+This section discusses publications and other research experiences that can strengthen your application.
 \autoref{sec:research-opportunities} provides more information on how to find research opportunities (e.g., during your undergrad study).
@@
 \subsection{Publications} The most concrete evidence of research ability is having \emph{papers in reputable international journals or conferences}.
-Having published papers, especially at top venues, is a sign that the you has been successfully involved in research.
-Publications are never required for application, however given the competitiveness of the process, they can significantly strengthen your application and are becoming the norm, e.g., those admitted to top schools often have multiple first-authored papers at top places.
+Having published papers, especially at top venues, is a sign that you have been successfully involved in research.
+
+Publications are never required for application, however given the competitiveness of CS admission, they can significantly strengthen your application and are becoming the norm.
+Applicants admitted to top schools, especially in popular fields such as ML and NLP, often have multiple first-authored papers at top places.
 
@@
 This might be something to be mindful of, as \textbf{CS} faculty might not be too familiar with Scopus or journals listed there, so devote some part in your statement to discuss the significance of your papers.
-\end{commentbox}
-
-However, it is understandable that many students do not have the opportunity to publish papers.
-Thus, general writings, even those under submissions or rejected, would still be good (much better than nothing).
-Be sure to upload your papers with your application and talk about them in your statement (see \autoref{sec:sop}).
-Note that local conferences and non-English journals or conferences do
-not carry as much weight since their quality is often unknown to US faculty.
-However, if you have published in such places, you should still upload them, mention them in your statement, and explain why they are good.
-
-% In CS, conferences are important and top conferences are prestigious.
-You can find top CS conferences at places such as \href{https://csrankings.org}{CSRankings}, which ranks CS programs based on how their faculty publish at top \emph{conferences} (see more in \autoref{sec:ranking}).
-
-\begin{commentbox}[Craig:]
-  Mason and many other universities allow you to upload your published papers and other writing samples.
-In many cases, even if the papers were not published at top places, we can still determine their quality by simply skimming over the paper.
 \end{commentbox}
@@
 For example, if you are interested in a field X, you can use CSRankings to look at active faculty in X, and then look at their websites to see what journals they publish at.
+\end{commentbox}
+\paragraph{What if I don't have any pubs} It is understandable that many students do not have the opportunity to publish papers.
+Thus, other writings, even those under submissions or rejected, would still be good and better than nothing.
+Be sure to upload your papers with your application and talk about them in your statement (see \autoref{sec:sop}).
+
+Note that local conferences and non-English journals or conferences do
+not carry as much weight since their quality is often unknown to US faculty.
+However, if you have published in such places, you should still upload them, mention them in your statement, and explain why they are good.
+
+% In CS, conferences are important and top conferences are prestigious.
+You can find top CS conferences at places such as \href{https://csrankings.org}{CSRankings}, which ranks CS programs based on how their faculty publish at top \emph{conferences} (see more in \autoref{sec:ranking}).
+
+\paragraph{I am not the first author} Being the first author is good because it indicates you own the work.
+However, it's perfectly OK to be second or third or last.
+It is difficult to publish a good paper and so being a co-author is still a good sign about your research experience.
+In any case (and especially in the case when you're not the first author), you should explain the work and your contribution.
+Better yet, have your LoR writers talk about your contribution in their letters.
+
+\begin{commentbox}[Craig:]
+  Mason and many other universities allow you to upload your published papers and other writing samples.
+In many cases, even if the papers were not published at top places, we can still determine their quality by simply skimming over the paper.
 \end{commentbox}

### 2024-10-22 demystify.tex
updates

@@
 
-Having a strong letter from an internationally recognized researcher will \emph{greatly strengthen} your application.
-A strong letter from a well-known researcher can outweigh the lack of publications (or low GPA or unknown school).
-
-However, obtaining letters from famous people
-can be challenging for international students, who might not have many interactions with such experts.
-So it is fine to have a letter from people (e.g., profs, researchers, postdocs who mentored you) who know you well enough to talk about \emph{your research experience and capabilities}.
+\paragraph{LoR from Famous Ppl?} Having a strong letter from an internationally recognized researcher will \emph{greatly strengthen} your application.
+A strong LOR from such a researcher can outweigh the lack of publications (or low GPA or unknown school).
+
+%However, obtaining letters from famous people
+%can be challenging for international students, who might not have many interactions with such experts.
+
+However, let's face it.  Most of us don't have an opportunity to interact or work with well-known researchers.  
+So what to do?
+It is perfectly fine to have a letter from people who know you well enough to talk about \emph{your research experience and capabilities}.
+These include profs., researchers, postdocs who have mentored and worked with you.
+In fact, a personal, customized, and strong letter from even a senior PhD grad student who mentors and knows you well is much better than a generic letter from a well-known superstar who does not know you well.
+
+It is also worth noting that a LoR from a ``full'' senior professor who has not been active in research might not be as strong as a LoR from a junior faculty who has been actively mentoring you in research.
+This applies to administrative (e.g., chairs, directors, and deans) also (\autoref{sec:admin-letters}.) The fame of the writer is less important than how well they know you and can write about your research ability.
 

### 2024-10-22 demystify.tex
updates

@@
  Many international students write letters themselves, typically due to the requests of their profs.
-or letter writers, and have them signed by their profs.
+or letter writers.
 Such letters have \emph{little value} and are considered weak by reviewers---why can you not even find someone who cares or knows enough about you to write a candid personal reference letter?
@@
 
-Most well-known and well-respected profs would \emph{not} ask you to write your own letter (in fact, even not well-known ones wouldn't do this).
+Most well-known and well-respected profs would \emph{not} ask you to write your own letter (in fact, even not well-known ones wouldn't do this to students they care about).
 This might be a common practice at specific universities and the students do not have a choice as they need the letter.
 However, think about this: if a prof.
-does this often, then they either do not know/care enough about you or don't know how to write a good letter.
-In any case, such letters are not useful and might even hurt your application.
+does this often, then they either don't know how to write a LOR (more common than you would think) or simply do not know or care enough about you.
+In any case, such LoRs are not useful and might even hurt your application.
 So if you are in this situation, you should find someone else to write for you.
 
-\subsection{Letters from Admin People}\label{sec:admin-letters}
-
-Many international students try to get letters from high-ranking admin people in their universities such as department chair/head, dean, or director.
-The students often believe that these letters are more valuable because such people are important in their universities.
-However, these people, while well-known and respected in your university, might not be active in research (e.g., they focus on administration) and thus are not recognized by adcom members.
-Such a letter has little value if the writer does not know you well enough (e.g., did not work with you) to talk about your research ability.
+\subsection{LoRs from Admin People}\label{sec:admin-letters}
+
+Many international students try to get LoRs from high-ranking administrators in their universities such as department chair/head, dean, or director.
+The students never worked with these people (they might take a class or so with these profs), but mistakenly believe that these LoRs are valuable due to the writer's high position in the university.
+However, as emphasized, a generic LoR has little value because the writer does not know you well and can talk in depth about your research ability.
+
+Moreover, while being well-known and respected in your local university, these writers might not be very active in research (e.g., they haven't published in recent years).
+Thus they might not be well-known and recognized by adcom members.
 
@@
   In my experience in reviewing applications, letters from admin people are often generic and do not provide much value. 
-  In many case, the letter reads like it was written by a student (\autoref{sec:self-letters}), and thus is a red flag.  So if you are in this situation, you should find someone else to write for you.
-    So if you are in this situation, find someone else to write for you.
+  In many case, the letter reads like it was written by a student (\autoref{sec:self-letters}), and thus is a red flag.
+So if you are in this situation, you should find someone else to write for you as mentioned in this section.
 \end{commentbox}

### 2024-10-24 demystify.tex
updates

@@
 
-\subsection{Generic Letters are Bad} When the writers do not know much about the applicants (e.g., just taking some course with them or not making any impression to write about), they might write a generic and short letter, which is not useful and also considered weak.
+\subsection{Generic Letters are Bad}\label{sec:generic-letters} 
+When the writers do not know much about the applicants (e.g., just taking some course with them or not making any impression to write about), they might write a generic and short letter, which is not useful and also considered weak.
 
@@
 
-\subsection{Help your LOR Writer} As mentioned above, do not write your own letter and generic letters do not give much value.
-However, you can help your writer to write a strong, customized LoR for you by providing details or unique things about yourself.
+\subsection{Help your LOR Writer}]\label{sec:help-your-LOR-writers}
+
+As mentioned above in \autoref{sec:self-letters} and \autoref{sec:generic-letters}, do not write your own letter and generic letters do not give much value.
+Thus, to help your writer to write a strong, customized LoR, you should provide them details or unique things about yourself.
 For example, let them know about your GPA, research and work experience, papers (if any), or anything you want them to mention.
+If the GPA in your program is highly competitive (\autoref{sec:gpa}) and they know that, remind them to talk about it in the LOR.
 You can also provide them with a draft of your SOP so that they can see what you are saying about yourself and complement that with their own perspective.
@@
 
-\subsection{Kiss of Death}
+\subsection{Kiss of Death in SOP} \label{sec:kiss-of-death-sop}
 
@@
 \item \textbf{Not customized to the program}: If your SOP can be sent to multiple programs with few changes, it is too generic.
-Do some homework and mention why you want to spend the next 7 years there.
+Do some homework and mention why you want to spend the next 5--7 years there.
 \item \textbf{Mentioning wrong professors}: do not mention emeritus professors or those who have left.
@@
 \end{commentbox}
-
-\subsection{Grades} Grades generally do not matter much for CS PhD admission.
+\subsection{Grades}\label{sec:gpa}
+
+Grades generally do not matter much for CS PhD admission.
 PhD in CS is a research degree and doing well in undergrad courses does not necessarily mean you can do research (other factors such as research experiences, pubs, LoRs are more important).
@@
 
-As with school reputation, you and your LoR writers can mention the grading system of your university if you think that is helpful for adcom to evaluate.
+As with school reputation, you and your LoR writers can mention the grading system of your university if you think that is helpful for adcom to evaluate (\autoref{sec:help-your-LOR-writers}).
 
@@
 \begin{commentbox}[Vu:]
-  The minimum for Mason (being above this might not mean much, but below is a \red{red flag}).
+  Here is the minimum requirements at Mason. 
+  Being above this might not mean much, but below is a \red{red flag}.
   \begin{itemize}
@@
     % \item but if you want to use it, then we expect a total (V+Q) of $\ge 311$ (with a $\ge 157$ Q) and A $\ge 3.0-3.5$.
-    \item English requirement tests (one of the below)
+    \item \href{https://www.gmu.edu/international/english-language-requirements}{English proficiency requirements} (one of the below)
           \begin{itemize}
-            \item TOEFL: 88 pts in total AND $\ge 20$ points in each subsection OR
+            \item TOEFL: 80 OR
             \item IELTS: $\ge 6.5$ OR
@@
 \subsection{Online Courses and Certificates}
-In general, these do not carry much weight as they do not show research ability. We do not care much if you have taken an online Coursera AI course or have a professional certificate from Microsoft.
+
+These do not carry much weight as they do not show research ability. We do not care much if you have taken an online Coursera AI course or have a professional certificate from Microsoft.
 However, as mentioned in \autoref{sec:non-stem}, if you do not have a CS background, you might be able to use these to show you have sufficient CS knowledge.
@@
 
-By around March you should hear back from most PhD programs that want to admit you.  In many cases you will not hear back from schools that do not want to admit you (see \autoref{sec:late-rejection}).
+By around March you should hear back from most PhD programs that want to admit you. 
+But you likely won't hear back from schools that do not want to admit you (\autoref{sec:late-rejection}).
+
 If you receive offers, congratulations!
@@
 
-Many students lament that they get no interviews or are rejected (with strong profiles) and that the admission process seems random.
-However, while it is true that the process is not perfect, it is not random.
+Many students lament that they get no interviews or are rejected and that the admission process seems random.  However, while it is true that the process is not perfect, it is not random.
 Other than not having a strong profile (e.g., research potential, GPA, LoRs, SOP, poor interviews), here are some other reasons for rejections.
@@
 A related reason is that you did not show enough interest in the program.
-For example, your SOP did not mention any faculty or research area at the program, you did not respond to professors interested in interviewing you (which might be considered unprofessional and burn bridges), or during the interview you showed little knowledge, interest, or enthusiasm (because they are your safety schools) in the program or profs.
+For example, you did not respond to professors interested in interviewing you (which might be considered unprofessional and burn bridges), or during the interview you showed little knowledge, interest, or enthusiasm in the profs or program (because it is your safety).
 
@@
 \paragraph{How to get TA?}  Unless you have other funding such as RA or fellowships, TA is typically the default for CS PhD programs.
-In your application, indicate that you need financial assistance.
+In your application, just simply indicate that you need financial assistance.
 Typically, adcom will either admit you and give you TA, or reject you.
-We do not admit a student without supporting them.
+We do not admit a student without supporting them (\autoref{sec:ievaluate}).
 
@@
 In that case, you will be paid as GTA or even sometimes as a lecturer.
-This is a good opportunity for students to get teaching experience and also get paid more (as a lecturer).
+This is a good opportunity for students to get teaching experience and you might even get paid a bit more.
 
@@
 
-Fellowship is another type of funding that students can get (e.g., from the university, industries, or government).
+Fellowship is another type of funding that students can get from the university, industries, or government.
 Fellowships are typically competitive and generous, giving pretty much all benefits tuition/insurance that a TA/RA has.
@@
 \begin{center}
-  \includegraphics[scale=0.3]{files/c6.png}
+  \includegraphics[scale=0.5]{files/c6.png}
 \end{center}
@@
 
-\subsection{Finding the right adviser}
-
-Here is some advice for finding a prof.
+\subsection{Finding an Adviser}
+
 Assuming you're not familiar with any particular one, then first search for profs.
@@
 
-\paragraph{Types of Faculty} There are typically two types of faculty at an R1 research university: tenure-line and teaching.
+\subsubsection{Types of Faculty} There are typically two types of faculty at an R1 research university: tenure-line and teaching.
 The main difference is their focus: research, teaching, or both.
@@
 
-\paragraph{Ranks} Faculty have ranks such as assistant, associate, and full, regardless if they are tenure-line or teaching.
-Assistant means new faculty, associate means they have been promoted, and full means they are senior.
-Typically, a tenure-track faculty starts with being an assistant professor and then gets tenure and promoted to associate (around 6 years) and then full (time varies, some become full within 3--4 years, some 10+ years, some remain associate).
+\subsubsection{Ranks} Faculty have ranks such as assistant, associate, and full, regardless if they are tenure-line or teaching.
+Assistant means new faculty, associate means they have been promoted (or was hired after being in the industry for a while), and full means they are senior.
+
+Typically, a tenure-track faculty starts with being an assistant professor and then gets tenure and promoted to associate (around 6 years).
+Note that getting tenure is a huge deal, and it might change the way they work with students, e.g., they might be more relaxed and less stressed (\autoref{sec:tenure-vs-tenure-track}).
+
+After that they go up for full prof.
+(time varies, some become full within 3--4 years, some 10+ years, some remain associate).
 A non-tenure track faculty, e.g., teaching or research faculty, also starts with Assistant, then is promoted to Associated, and finally Full.
@@
 
-Now that you know a bit about tenured and tenure-track faculty (\autoref{sec:faculty-types}).
-Which one should you choose as your adviser?
-The short answer is that either can be a good adviser, but they have different strengths and weaknesses.
+Now that you know a bit about tenured and tenure-track faculty (\autoref{sec:faculty-types}).  Which one should you choose as your adviser? 
+Either can be a good adviser, but they have different strengths and weaknesses.
 
@@
 Thus, they will devote more time to work with you and push you to do research and publish.
-However, they may not have as much experience in managing students and may not have as much funding (yet).
+However, you might not be too independent when you graduate because they have been too hands-on with you.
+Also, they may not have as much experience in managing students and may not have as much funding (yet).
 
@@
 
-\subsection{Things to avoid when cold emailing a professor}
+\subsection{Kiss of Death in Emails}\label{sec:kiss-of-death-emails}
 
 \begin{itemize}
-
-    \item \textbf{Not customizing the email}. These emails typically talk about the student and not why they are interested in the professor's work. See writing tips in \autoref{sec:contact}
-    \item \textbf{Sending the same email to multiple professors}.
+    \item \textbf{Not customizing the email, too generic}: a copy and paste kind of email or those that can be sent to multiple professors with very little modifications.
+This shows the lack of interest and will be treated as spam.
+    \item \textbf{Self-focus} You focus too much about you and your achievements but not why you are interested in the prof.'s work (\autoref{sec:contact}).
+Why their work is interesting and how your background fits in.
+    \item \textbf{Too long}. Keep it to about 3--4 paragraphs. Seriously, less is more and long emails are often not read and discarded.\\
+    Personally I do not want to see course transcript or CV in the first email.  If we are interested, we will ask for more information.
+    \item \textbf{Flowery greetings and language}. Don't use ``Dear esteemed professor'', ``kindly consider my application'', etc. And of course, don't ``hey there''. 
+    \item \textbf{Sending the similar mail to multiple profs}.
 Sending emails to multiple profs.
 using BCC or even CC (where we can see all addresses the email is being sent to) will result in no reply or a very harsh one on how unprofessional this is.
-    \item \textbf{Too much information}.
-Keep your email short and to the point.
-If we are interested, we will ask for more information.
-2--3 paragraphs are enough.
-Personally I also do not want to see course transcript or CV in the first email.
-    \item \textbf{Not being professional}, e.g., do not call the professor by their first name in the first email.  Use Prof. Lastname or Dr. Lastname (\autoref{sec:address}).
-    \item \textbf{Not reading the professor's website} and asking questions that are already answered there.
-    \item \textbf{Not using plain text}. Fancy format, colors, or fonts are not professional. CS people like plain text.
+    \item \textbf{Not being professional}, e.g., do not call the professor by their first name in the first email (some don't care but you don't want to take the risk -- you don't know them that well yet).
+Do not use Mr., Mrs., etc.
+Use Prof.
+Lastname or Dr.
+Lastname (\autoref{sec:address}).
+    \item \textbf{Ignoring the Prof's guideline} and asking questions that are already answered on their website (most CS profs have personal websites).
+For example, on my website, I put very specific information on how to contact me (e.g., email subject).
+    \item \textbf{Fancy format}. CS people like plain text.  Do not use colors, fonts, or fancy formats.
 \end{itemize}

### 2024-10-24 demystify.tex
updates

@@
     \item \textbf{Flowery greetings and language}. Don't use ``Dear esteemed professor'', ``kindly consider my application'', etc. And of course, don't ``hey there''. 
-    \item \textbf{Sending the similar mail to multiple profs}.
-Sending emails to multiple profs.
-using BCC or even CC (where we can see all addresses the email is being sent to) will result in no reply or a very harsh one on how unprofessional this is.
     \item \textbf{Not being professional}, e.g., do not call the professor by their first name in the first email (some don't care but you don't want to take the risk -- you don't know them that well yet).
@@
     \item \textbf{Fancy format}. CS people like plain text.  Do not use colors, fonts, or fancy formats.
+    \item \textbf{Mass emails}.
+This is too obvious but I've seen it many times when a student mass emails all profs in a department (e.g., through CC or even BCC---we know because faculty talk to each other).
+This will result in no reply or a very harsh one on how unprofessional you are.
 \end{itemize}

### 2024-10-24 demystify.tex
updates

@@
 \begin{itemize}
+    \item \textbf{Send emails to the wrong prof.}.
+While this is not mentioned, it is the number one reason why you don't get a reply: you send email to profs.
+who do not / cannot advise CS PhD students (e.g., adjunct, emeritus, teaching, non-CS, etc.).
+See \autoref{sec:faculty-types} for details.
     \item \textbf{Not customizing the email, too generic}: a copy and paste kind of email or those that can be sent to multiple professors with very little modifications.
@@
 Why their work is interesting and how your background fits in.
-    \item \textbf{Too long}. Keep it to about 3--4 paragraphs. Seriously, less is more and long emails are often not read and discarded.\\
+    \item \textbf{Too long}. Keep it to about 3--4 paragraphs. Seriously, less is more and too long emails are often not read and discarded.\\
     Personally I do not want to see course transcript or CV in the first email.  If we are interested, we will ask for more information.
-    \item \textbf{Flowery greetings and language}. Don't use ``Dear esteemed professor'', ``kindly consider my application'', etc. And of course, don't ``hey there''. 
-    \item \textbf{Not being professional}, e.g., do not call the professor by their first name in the first email (some don't care but you don't want to take the risk -- you don't know them that well yet).
+    \item \textbf{Flowery greetings and language}.
+Don't use ``Dear esteemed professor'', don't ``hey there''.
+Do not call the prof.
+by their first name in the first email (some don't care but you don't want to take the risk -- you don't know them that well yet).
 Do not use Mr., Mrs., etc.
-Use Prof.
+To be safe, use Prof.
 Lastname or Dr.
@@
 For example, on my website, I put very specific information on how to contact me (e.g., email subject).
-    \item \textbf{Fancy format}. CS people like plain text.  Do not use colors, fonts, or fancy formats.
+Following this helps you stand out, and I will likely respond to you than others.
+    \item \textbf{Fancy format}. CS (and probably many other fields) likes plain email.  Do not use colors, fonts, or fancy formats.
     \item \textbf{Mass emails}.
-This is too obvious but I've seen it many times when a student mass emails all profs in a department (e.g., through CC or even BCC---we know because faculty talk to each other).
+I've seen it many times when a student mass emails all profs in a department (e.g., through CC or even BCC---we know because faculty talk to each other).
 This will result in no reply or a very harsh one on how unprofessional you are.

### 2024-10-24 demystify.tex
kiss of death cold email

@@
 See \autoref{sec:faculty-types} for details.
-    \item \textbf{Not customizing the email, too generic}: a copy and paste kind of email or those that can be sent to multiple professors with very little modifications.
-This shows the lack of interest and will be treated as spam.
-    \item \textbf{Self-focus} You focus too much about you and your achievements but not why you are interested in the prof.'s work (\autoref{sec:contact}).
-Why their work is interesting and how your background fits in.
-    \item \textbf{Too long}. Keep it to about 3--4 paragraphs. Seriously, less is more and too long emails are often not read and discarded.\\
-    Personally I do not want to see course transcript or CV in the first email.  If we are interested, we will ask for more information.
+    \item \textbf{Generic}. A copy and paste kind of email or those that can be sent to multiple ppl with very little modifications. This shows the lack of interest and will be treated as spam.  
+    \item \textbf{Self-focus}.
+Focusing too much about you and your achievements but not why you are interested in the prof.'s work (\autoref{sec:contact}).
+Mention why you're interested in their work and how your background can contribute.
+    \item \textbf{Too long}. Keep it to about 3--4 short paragraphs. Less is more and too long emails are often not read and discarded.\\
+    Personally I do not want to see course transcript or test scores in the first email.
+If I am interested I'll ask for them.
+Attaching CV is OK.
+Sample papers and links to your Arxiv papers or GitHub projects are also OK if they are relevant to my work.
     \item \textbf{Flowery greetings and language}.
@@
 Lastname (\autoref{sec:address}).
-    \item \textbf{Ignoring the Prof's guideline} and asking questions that are already answered on their website (most CS profs have personal websites).
-For example, on my website, I put very specific information on how to contact me (e.g., email subject).
-Following this helps you stand out, and I will likely respond to you than others.
-    \item \textbf{Fancy format}. CS (and probably many other fields) likes plain email.  Do not use colors, fonts, or fancy formats.
+    \item \textbf{Ignoring the Prof's guideline} and asking questions that are already answered on their website.
+Most CS profs have websites and many put very specific information on how to contact them (e.g., email subject, what to include).
+Following this helps you stand out and increase your chance of getting a reply.
+    \item \textbf{Fancy format}. CS (and probably many other fields) prefers plain email.  Do not use colors, fonts, or fancy formats.
     \item \textbf{Mass emails}.

### 2024-10-26 ennv.tex
updates

@@
 
-\paragraph{Convolutional Neural Networks (CNNs)} are a type of neural network that is often used in image recognition and classification.
-CNNs consist of of neurons that have learnable weights and biases.
+\paragraph{Convolutional Neural Networks (CNNs)} CNNs, often used in image recognition and classification, consist of neurons that have learnable weights and biases.
 Each neuron receives several inputs, takes a weighted sum over them, passes it through an activation function, and responds with an output.
 
-\paragraph{Recurrent Neural Networks} Recurrent Neural Networks (RNNs) are a type of neural network that is often used in natural language processing and speech recognition.
-RNNs are designed to recognize patterns in sequences of data.
-RNNs have loops in them, allowing information to persist.
-This loop allows information to be passed from one step of the network to the next.
+\paragraph{Recurrent Neural Networks (RNNs)} RNNs, often used in natural language process and speech recognition, are designed to recognize patterns in sequences of data.
+RNNs have \emph{loops} in them, allowing information to be sent forward and backward.
 

### 2024-10-26 demystify.tex
updates

@@
 
-\subsection{Should you contact a professor? How to get a desired reply?}\label{sec:contact}
+\section{Contacting a prof. How to get a desired reply?}\label{sec:contact}
 
@@
 
-\subsection{Kiss of Death in Emails}\label{sec:kiss-of-death-emails}
+\section{Kiss of Death in Emails}\label{sec:kiss-of-death-emails}
 
@@
     \item \textbf{Send emails to the wrong prof.}.
-While this is not mentioned, it is the number one reason why you don't get a reply: you send email to profs.
+Most students often do not pay attention or know about this, but a very common reason why you don't get a reply is that you send email to profs.
 who do not / cannot advise CS PhD students (e.g., adjunct, emeritus, teaching, non-CS, etc.).
 See \autoref{sec:faculty-types} for details.
-    \item \textbf{Generic}. A copy and paste kind of email or those that can be sent to multiple ppl with very little modifications. This shows the lack of interest and will be treated as spam.  
+While these profs.
+might be able to co-advise, they typically do not have the bandwidth, funding, and the desire to mentor students (they are already overloaded with teaching and other service duties).
+So do some research before sending emails, e.g., most CS profs.
+who are active in research will have a website with their work (publications) and students; some also have lab website dedicated to their group.
+Most teaching and adjunct profs.
+do not have such a research page.
+
+    \item \textbf{Generic}.
+You should already know this!
+A copy and paste kind of email or those that can be sent to multiple ppl with very little modifications show the lack of interest and will be treated as spam.
+Most likely we will not reply to these emails.
+
     \item \textbf{Self-focus}.
@@
 Mention why you're interested in their work and how your background can contribute.
+
     \item \textbf{Too long}. Keep it to about 3--4 short paragraphs. Less is more and too long emails are often not read and discarded.\\
@@
 Lastname (\autoref{sec:address}).
+
     \item \textbf{Ignoring the Prof's guideline} and asking questions that are already answered on their website.
-Most CS profs have websites and many put very specific information on how to contact them (e.g., email subject, what to include).
+Many profs.
+put very specific information on how to contact them on their website (e.g., email subject, what to include).
 Following this helps you stand out and increase your chance of getting a reply.
-    \item \textbf{Fancy format}. CS (and probably many other fields) prefers plain email.  Do not use colors, fonts, or fancy formats.
+
+    \item \textbf{Fancy format}.
+Do not use colors, fancy fonts or formats.
+While not really a kiss of death, it is very annoying, especially for ppl in CS (and probably many other fields) who often prefer plain messages.
+
     \item \textbf{Mass emails}.
@@
 This will result in no reply or a very harsh one on how unprofessional you are.
+
+    \item \textbf{Don't CALL}.
+Not related to email but sometimes students get desperate and call the prof.
+This is a big no-no, especially for CS ppl who often prefer email over phone calls.
+Seriously though, don't call any prof.
+unless they ask you to.
 \end{itemize}
+
+Even if you do everything right, you will likely get a generic (\autoref{sec:interpreting-response}) or no response from a prof.
+Don't take it personally.
+They might not have the bandwidth to take on new students (very common), or they might not be interested in your background.
+
+\section{Interpreting Response}\label{sec:interpreting-response}
+
+Some generic responses are:
+\begin{itemize}
+\item \textbf{No reply}. Very common. You might try again in a few weeks or months, but don't expect a reply. And after a couple of tries, you should move on.  It simply means they are not interested.
+\item \textbf{Not taking students but encourage you to apply}. Polite way of saying not interested and referring you to the admission process.
+\item \textbf{Not taken student this year (but encourage you to apply next year)}.
+Polite but generic response.
+But If you do apply next year, you can mention that you contacted them last year and they encouraged you to apply.
+\item \textbf{Come talk to me after you're admitted}. Generic. Refer you to the admission process. But if you get admitted then you can reply to them and say you're admitted and would like to talk.
+\item \textbf{Cannot admit you directly. Need to go through admission process}. Generic. They are not interested and refer you to the admission process.
+\item \textbf{I am not taking students but I co-advise/can serve on your committee} While this might sound good, it's generic because it says once you're admitted and have an adviser, then contact me again.
+\item \textbf{I am not taking student but Prof.
+X might be}.
+Not common practice as most profs.
+will not refer you to their colleagues.
+However, this is better than the previous responses.
+While they cannot take you, they think you are a good fit for X.
+So follow up with a thank you and say you'll contact X.
+And then contact X and say that Y referred you to them.
+\end{itemize}
+
+In short, all of these replies mean the prof.
+is not interested.
+The best positive response is that they want more information from you, e.g., your CV, transcript, paper, or a chat (like an interview).
 

### 2024-10-27 demystify.tex
updates

@@
 
-\pagestyle{empty}
+%\pagestyle{empty}
+
 \begin{tikzpicture}[overlay,remember picture]
@@
 
-          %\item \emph{Contacting a prof. is recommended}, but do it \emph{properly} to get a reply (\autoref{sec:contact}).
+          %\item \emph{Contacting a prof. is recommended}, but do it \emph{properly} to get a reply (\autoref{chap:contact}).
 
@@
 Then you research about that prof.
-(e.g., go to their website) and contact them (\autoref{sec:contact}).
+(e.g., go to their website) and contact them (\autoref{chap:contact}).
 
@@
   \tcblower
-  \textbf{Vu}: \autoref{sec:contact} shows how to contact a professor for research opportunities.
+  \textbf{Vu}: \autoref{chap:contact} shows how to contact a professor for research opportunities.
 Many will say no (or do not reply) as they do not have the bandwidth to take on random students, but some may say yes if they see a potential fit.
@@
 Thus when contacting a prof.
-for research opportunities (\autoref{sec:contact}), you should focus on tenure-line faculty (and not adjunct or emeritus :p).
+for research opportunities (\autoref{chap:contact}), you should focus on tenure-line faculty (and not adjunct or emeritus :p).
 
@@
 
-\section{Contacting a prof. How to get a desired reply?}\label{sec:contact}
-
-\subsectioninfo{Individually faculty member \emph{cannot directly admit} a student---so do not email and ask if you have a chance.
+\chapter{Contacting a prof. How to get a desired reply?}\label{chap:contact}
+
+\chapterinfo{Individually faculty member \emph{cannot directly admit} a student---so do not email and ask if you have a chance.
 However, faculty can \emph{advocate} for a student and therefore increase their admission chance---so do contact and introduce yourself.}
@@
 Finally, profs. are very busy so don't take it personally if you don't get anything from them (though I would be very surprised if such thoughtful emails get no replies!).
+See~\autoref{sec:kiss-of-death-emails} for common mistakes in emails and~\autoref{sec:interpreting-response} for interpreting replies.
 
@@
 
-\subsubsection*{Additional Resources}
+\subsection*{Additional Resources}
 \begin{itemize}
@@
     \item \textbf{Self-focus}.
-Focusing too much about you and your achievements but not why you are interested in the prof.'s work (\autoref{sec:contact}).
+Focusing too much about you and your achievements but not why you are interested in the prof.'s work (\autoref{chap:contact}).
 Mention why you're interested in their work and how your background can contribute.
 
-    \item \textbf{Too long}. Keep it to about 3--4 short paragraphs. Less is more and too long emails are often not read and discarded.\\
-    Personally I do not want to see course transcript or test scores in the first email.
-If I am interested I'll ask for them.
+    \item \textbf{Too long}.
+Keep it to about 3--4 short paragraphs.
+Less is more and too long emails are often not read and discarded.
+Don't attach course transcript or test scores in the first email.
+If they are  interested they will ask for them.
 Attaching CV is OK.
-Sample papers and links to your Arxiv papers or GitHub projects are also OK if they are relevant to my work.
+Sample papers and links to your Arxiv papers or GitHub projects are also OK if they are relevant.
+
     \item \textbf{Flowery greetings and language}.
@@
 \paragraph{Standard Tests \autoref{sec:standard-tests}} You do not need to take TOEFL or IELTS because you already did your undergrad (or MS) at a US university.
-You might also be more comfortable communicating in English, e.g., contacting professors (\autoref{sec:contact}).
+You might also be more comfortable communicating in English, e.g., contacting professors (\autoref{chap:contact}).
 
@@
 Finally, you can also ask your peers who are already doing research.
-Use the method described in \autoref{sec:contact} to contact professors
+Use the method described in \autoref{chap:contact} to contact professors
 

### 2024-10-27 demystify.tex
updates

--- 2024-10-26 demystify.tex
@@
 
-However, no interviews could mean that you will not likely get an RA (\autoref{sec:ra}), which is offered by an individual faculty (if they want you to do research for them, then they likely will interview you first).
-In the case of no interviews, your application (and TA/fellowship funding) is decided by the adcom.
+However, no interviews mean that you will not likely get an RA (\autoref{sec:ra}), which is offered by an individual faculty (if they want you to do research for them, then they likely will interview you first).
+If you have no interviews, your application (and TA/fellowship funding) is decided by the adcom.
 
@@
 You can try to contact the school to ask about your status, but they might not reply, they might say they are still reviewing applications, or give you inaccurate information (e.g., you will hear in two weeks).
-In short, you just have to be patient and wait (however, beware that some schools do not send out rejection letters at all).
+In short, you just have to be patient and wait (and also beware that some schools do not send out rejection letters at all).
 
@@
 However, you can ask for things such as computer equipment (\autoref{sec:buying-equipment}), books, (and maybe even furniture), and a conference travel budget.
+
+\section{Buying Computer Equipment}\label{sec:buying-equipment}
+\sectioninfo{Ask your prof. if they can buy computer equipment and such for your research.}
+
+Students understandably get excited about their upcoming PhD journey and want to buy new computer equipment and electronics to prepare.
+However, you should first check with your professor.
+They might have funding to buy you computers and other equipment (e.g., software, books, keyboards, headphones, tablets, etc).
+Many CS programs also provide budget or computer equipment to incoming PhD students (e.g., a laptop).
+
+However, note this is a privilege and a courtesy and not a right.
+This means that the prof.
+is not obligated to buy you things, and that you should not expect these are automatically provided.
+Also, while the prof.
+might be able to do so in the past, they might not be able to do so now due to budget constraints.
+It might also a need-based decision, e.g., if you need a powerful computer for your research, then the prof.
+might buy it for you (the lab might already has powerful servers and you can just use them for your work).
+Finally, the prof.
+might have a policy on what they can buy for you, e.g., they prefer to get you desktop instead of a laptop (to encourage you to come to the office and work).
+
+Keep in mind that these computers and equipment would be university property, which might be monitored and have certain restrictions, e.g., do not install illegal software on them (\autoref{sec:illegal-software}).
+You will likely need to return them when you graduate.
 
@@
 
-\subsectioninfo{Individually faculty member \emph{cannot directly admit} a student---so do not email and ask if you have a chance.
+\sectioninfo{Individually faculty member \emph{cannot directly admit} a student---so do not email and ask if you have a chance.
 However, faculty can \emph{advocate} for a student and therefore increase their admission chance---so do contact and introduce yourself.}
@@
 
-\section{Kiss of Death in Emails}\label{sec:kiss-of-death-emails}
+\subsection{Kiss of Death in Emails}\label{sec:kiss-of-death-emails}
 
@@
 
-    \item \textbf{Too long}. Keep it to about 3--4 short paragraphs. Less is more and too long emails are often not read and discarded.\\
-    Personally I do not want to see course transcript or test scores in the first email.
-If I am interested I'll ask for them.
+    \item \textbf{Too long}.
+Keep it to about 3--4 short paragraphs.
+Less is more and too long emails are often not read and discarded.
+Don't attach course transcript or test scores in the first email.
+If they are  interested they will ask for them.
 Attaching CV is OK.
-Sample papers and links to your Arxiv papers or GitHub projects are also OK if they are relevant to my work.
+Sample papers and links to your Arxiv papers or GitHub projects are also OK if they are relevant.
+
     \item \textbf{Flowery greetings and language}.
@@
 
-\section{Interpreting Response}\label{sec:interpreting-response}
+\subsection{Interpreting Response}\label{sec:interpreting-response}
 
@@
 
-\section{Buying Computer Equipment}\label{sec:buying-equipment}
-\sectioninfo{Ask your prof. if they can buy computer equipment and such for your research.}
-
-Students understandably get excited about their upcoming PhD journey and want to buy new computer equipment and electronics to prepare.
-However, you should first check with your professor.
-They might have funding to buy you computers and other equipment (e.g., software, books, keyboards, headphones, tablets, etc).
-Many CS programs also provide budget or computer equipment to incoming PhD students (e.g., a laptop).
-
-However, note this is a privilege and a courtesy and not a right.
-This means that the prof.
-is not obligated to buy you things, and that you should not expect these are automatically provided.
-Also, while the prof.
-might be able to do so in the past, they might not be able to do so now due to budget constraints.
-It might also a need-based decision, e.g., if you need a powerful computer for your research, then the prof.
-might buy it for you (the lab might already has powerful servers and you can just use them for your work).
-Finally, the prof.
-might have a policy on what they can buy for you, e.g., they prefer to get you desktop instead of a laptop (to encourage you to come to the office and work).
-
-Keep in mind that these computers and equipment would be university property, which might be monitored and have certain restrictions, e.g., do not install illegal software on them (\autoref{sec:illegal-software}).
-You will likely need to return them when you graduate.
-
 % \section{Will I be miserable during my a PhD?}\label{sec:happy}
@@
 \chapter{MS Admission}\label{chap:ms}
-\chapterinfo{MS focuses on coursework and prepares you for industry, while PhD focuses on research and prepares you for academia or research.}
+\chapterinfo{MS focuses on coursework and prepares you for \emph{industry}, while PhD focuses on research and prepares you for academia or research.}
 
@@
 
-  \item Objective: an MS is typically to prepare you for industry, while a PhD is to prepare you for research. Some MS has thesis option but the research is much lighter compared to a PhD.
+  \item Objective: an MS is typically to prepare you for \emph{industry}, while a PhD is to prepare you for research and academia.
+Some MS has thesis option but the research is much lighter compared to a PhD.
 
@@
 You graduate with an MS when you're done with the courses.
-PhD also has coursework requirements (typically taken in your first 2 years) but focuses on research after that.
+PhD also has coursework requirements (typically taken in your first 2 years as mentioned in \autoref{sec:time}) but focuses on research after that.
 You graduate with a PhD when you have done enough research and written a dissertation, which usually takes much longer time than coursework.
@@
 
-  \item Funding: MS is typically \emph{not funded}, while PhD is typically funded (\autoref{sec:funding}). See \autoref{sec:ms-funding} for more details on MS funding.
+  \item Funding: MS is typically \emph{not funded}, while PhD is (\autoref{sec:funding}). See \autoref{sec:ms-funding} for more details on MS funding.
 
@@
 
-Except for a very few highly selective programs, MS programs are typically much less competitive than PhD programs, i.e., you're likely to get in if you can afford it.
+MS programs are much less competitive than PhD programs, i.e., you're likely to get in if you can afford it.
 Some refer to MS programs as a \emph{cash cow} for the department, as MS students are often not funded and have to pay tuition.
@@
 In most cases, the main requirements are that you have sufficient background in CS, e.g., through your undergrad degree.
+This does not mean MS programs are easy to get admitted, but the requirements are much lower compared to PhD programs.
 
@@
 MS admission also involves an adcom that reviews applications and makes admission decisions.
-However, MS admission is typically rolling (i.e., they review applications as they come in and make decisions throughout the admission cycle).
+However, MS admission is typically rolling, i.e., they review applications as they come in and make decisions throughout the admission cycle.
 Moreover, unlike PhD that has multiple reviewers for each application, an MS application typically involves only one reviewer and does require much time to review compared to a PhD application (\autoref{sec:ievaluate}).
@@
 However, research experience is not required, and LoRs can be from anyone who can speak about your academic or working abilities.
-Moreover, your SOP you should focus on your academic and work background, why you want to get an MS, and how the program fits your goals.
+Moreover, your SOP should focus on your academic and work background, why you want to get an MS, and how the program fits your goals.
 As with PhDs, GREs are often not required.
@@
 are not willing to take MS students as RAs because they are not around long enough to be productive.
-It can take a while for a student to get up to speed and start being productive, and by the time they are, they are about to graduate.
+It can take a while for a student to get up to speed and start being productive, and by that time, they are about to graduate.
 Moreover, the goal of most MS students is to get a job, not to do research, so they are not as motivated to do research.
-\item (TA) MS students are typically not given TA positions because PhDs are given priority.
-Many depts.
-even have shortage of TA positions for PhD students and so they cannot afford to give them to MS students.
+
+\item (TA) MS students are typically not given TA positions because PhDs are given priority. Many depts. do not have enough TAs for PhD students and so they cannot afford to give them to MS students.
 \end{itemize}
 
-However, there are always exceptions, e.g., if you have a strong background and can demonstrate that you can be productive, then a prof. might be willing to take you as their RA.
-Moreover, some schools have TAs for MS students, and you can apply for these positions. For example, Mason CS dept has quite a few TAs for MS students; my courses in the past two years have MS TAs.
+However, there are always exceptions, e.g., if you have a strong background and can demonstrate that you can be productive in research, then a prof. might be willing to take you as their RA.
+Moreover, some schools have TAs for MS students, and you can apply for these positions. For example, Mason CS has quite a few TAs for MS students; my courses in the past two years have MS TAs.
+
+\section{Selecting and Ranking Schools}\label{sec:selecting-ranking-schools-ms}
+%\sectioninfo{How to select and rank MS programs in CS.}
+
+Because of the differences between MS and PhD programs, you should consider different factors when selecting and ranking MS programs.
+For example, in PhD programs, you should focus on research areas and advisers (\autoref{sec:schoolsandprofs}), while in MS programs, you might want to consider factors such as location, industry connections, and job placement.
+
+\begin{itemize}
+
+\item \textbf{Location:} In addition to living in a place you like (e.g., warm weather), you should also consider the job opportunities in the area.
+If you want to work in the tech industry, you might want to consider MS programs in tech hubs like the Bay Area, Seattle, or Boston.
+For government jobs, you might want to consider schools in the DC DMV area.
+Many students at Mason work for the government or defense contractors in the DC area and take classes in the evening or online.
+
+\item \textbf{Industry Connections:} Universities often have strong connections with local companies and can help you get internships and jobs.
+For example, schools in Washington area often have their students working with Microsoft and Amazon while those in the Bay Area have students working with Google and Facebook.
+Mason naturally has strong connections with the government and defense contractors and Amazon, whose 2nd HQ is in Arlington.
+
+\item \textbf{Ranking:} Unlike PhD programs where the adviser and research are perhaps the most important factor, in MS programs the ranking of the school is more important as you do not need to spend 5--7 years doing research.
+Thus, you should consider the ranking of the school in CS, especially in the specific area you're interested in.
+For example, if you're interested in software engineering, you might want to consider schools with strong SWE programs.
+
+\item \textbf{Living Cost:} As MS students are typically not funded, living cost can be important.
+You should consider the tuition and living expenses of the area.
+However, I find that in the beginning living costs, e.g., renting, seems scary but students often find way to make it work, e.g., by sharing an apartment with other students.
+
+\end{itemize}
 

### 2024-10-27 demystify.tex
updates

@@
 Such a person is typically someone who has a PhD (so they know what a PhD is about) and is active in research (so they can evaluate your research ability).
-LoR writers are often your research advisors, professors, or supervisors who have worked with you on research projects.
+LoR writers are often your research advisers, professors, or supervisors who have worked with you on research projects.
 
@@
 Because of the differences between MS and PhD programs, you should consider different factors when selecting and ranking MS programs.
-For example, in PhD programs, you should focus on research areas and advisers (\autoref{sec:schoolsandprofs}), while in MS programs, you might want to consider factors such as location, industry connections, and job placement.
+For example, you might want to consider factors such as location, industry connections, and job placement instead of research areas and advisers (\autoref{sec:schoolsandprofs}).
 
@@
 
-\item \textbf{Location:} In addition to living in a place you like (e.g., warm weather), you should also consider the job opportunities in the area.
-If you want to work in the tech industry, you might want to consider MS programs in tech hubs like the Bay Area, Seattle, or Boston.
-For government jobs, you might want to consider schools in the DC DMV area.
-Many students at Mason work for the government or defense contractors in the DC area and take classes in the evening or online.
+\item \textbf{Location:} In addition to living in a place you like (e.g., warm weather, historical city, etc.), you should also consider the job opportunities in the area.
+For tech industry, consider MS programs in tech hubs like the Bay Area and Seattle.
+For government jobs, look at schools in the DC area.
+For example, many students at Mason work for the government or defense contractors in the DC area and take classes in the evening or online.
 
 \item \textbf{Industry Connections:} Universities often have strong connections with local companies and can help you get internships and jobs.
-For example, schools in Washington area often have their students working with Microsoft and Amazon while those in the Bay Area have students working with Google and Facebook.
-Mason naturally has strong connections with the government and defense contractors and Amazon, whose 2nd HQ is in Arlington.
-
-\item \textbf{Ranking:} Unlike PhD programs where the adviser and research are perhaps the most important factor, in MS programs the ranking of the school is more important as you do not need to spend 5--7 years doing research.
+For example, Microsoft and Amazon look for students in the Washington area while  Google and Facebook recruit those in the Bay Area.
+Mason students natural get jobs from the government, defense contractors, and Amazon, whose 2nd HQ is in Arlington.
+
+\item \textbf{Ranking:} Unlike PhD programs where the adviser and research are perhaps the most important factor, in MS programs the ranking of the school is more important.
 Thus, you should consider the ranking of the school in CS, especially in the specific area you're interested in.
@@
 
-\item \textbf{Living Cost:} As MS students are typically not funded, living cost can be important.
+\item \textbf{Living Cost:} MS students are typically not funded, thus living cost can be a big hassle.
 You should consider the tuition and living expenses of the area.
-However, I find that in the beginning living costs, e.g., renting, seems scary but students often find way to make it work, e.g., by sharing an apartment with other students.
+Note that sometimes living costs, e.g., renting, seems scary at first but students often find way to make it work, e.g., by sharing an apartment with other students.
 
@@
 
-As discussed in \autoref{sec:research-opportunities}, research experience is a crucial component of a strong PhD application.
-It gives you opportunities to try out research, determine what research area you're interested in, publish papers (\autoref{sec:research-experience}), connect with researchers, and get strong LoRs (\autoref{sec:lor}).
+As discussed in \autoref{sec:research-experience}, having a successful research experience can greatly strengthen a PhD application.
+Research experience gives you opportunities to try out research, determine what research area you're interested in, publish papers (\autoref{sec:research-experience}), connect with researchers, and get strong LoRs (\autoref{sec:lor}).
 
@@
 \begin{commentbox}[Vu:]
-    I enjoy working with undergrads (e.g., have mentored over 10 undergrads) and always actively recruit.
+    I enjoy working with undergrads (e.g., I have mentored over 15 undergrads) and always actively recruit.
     I often get undergraduate researchers through my classes, e.g., asking students who did well in my class if they are interested in research.
-Occasionally I was introduced to students by other students or faculty.
+Occasionally I was introduced to students by other students or faculty.\\
 
     While most undergrads are understandably not super productive in research, some are and I have published multiple papers with them.
-I also have written strong LoRs for them and have helped them get into PhD programs, including top ones.
+I also have written strong LoRs for them and have helped them get into good PhD programs.
 

### 2024-10-28 demystify.tex
updates

@@
 
-\section{Do adcom members read all applications?}\label{sec:read-all}
+\section{How applications are assigned to adcom members?}\label{sec:applications-assigned}
 \sectioninfo{Adcom members only review applications assigned to them (typically matching their expertise) and rarely get involved in other applications}
@@
 Some universities do waive, e.g., Rice does for all PhD applicants and many universities for domestic students (\autoref{sec:domestic-students}), however most do not for international students.
-In any case, the application fee is typically a requirement of the university.
-Individual departments and programs do not have the flexibility to waive the application fee, even if they want to.
+Some require proof of financial difficulties, e.g., a statement from a financial adviser or a bank statement.
+
+In gneral, application fee is typically a requirement of the university. Individual departments and programs do not have the flexibility to waive the application fee, even if they want to.
 
@@
 In my opinion, requiring applicants to pay the fee helps ensure their seriousness, as it filters out non-serious candidates.
-Also, most CS programs already receive many applications and would be overwhelmed if the application process were free.
+Most CS programs already receive many applications and would be overwhelmed if the application process were free.
 Note that if you have financial difficulties, you can ask the department for a waiver, but this is typically only granted in exceptional cases.
@@
 
-\paragraph{LoR from Famous Ppl?} Having a strong letter from an internationally recognized researcher will \emph{greatly strengthen} your application.
+\paragraph{LoR from Famous People} Having a strong letter from an internationally recognized researcher will \emph{greatly strengthen} your application.
 A strong LOR from such a researcher can outweigh the lack of publications (or low GPA or unknown school).
@@
 
-\begin{commentbox}[Vu:]
-For Vietnamese students, it's worth mentioning the \href{https://vef2.org/}{VEF2.0 program}, which has helped many students gain admission to top CS PhD programs in the US.
+% \begin{commentbox}[Vu:]
+% For Vietnamese students, it's worth mentioning the \href{https://vef2.org/}{VEF2.0 program}, which has helped many students gain admission to top CS PhD programs in the US.
 VEF2.0 invites US faculty members from leading institutions to conduct rigorous interviews with VEF students and subsequently provide reference letters on their behalf.
 Despite the limited interaction between the interviewer and the interviewee (primarily just the interview), these reference letters are generally effective and have helped many students get into good PhD programs.
-\end{commentbox}
+% \end{commentbox}
 
 \section{Statement of Purpose (SOP)}\label{sec:sop}
-\sectioninfo{\emph{SOP is very important}. Write it in such a way that makes you \emph{stand out}.}
+\sectioninfo{\emph{SOP is important}. Write it in such a way that makes you \emph{stand out}.}
 
@@
 statement of purpose (SOP) or personal statement\footnote{Few schools separate these documents and ask you to write both: SOP, which focuses on research experiences, and personal statement, which is everything more personal, e.g., why PhD, challenges, diversity, etc}!
-So write it well because it could make a big difference.
 A well-written SOP also shows that you can communicate, which is very important in research, and that you can effectively teach and communicate with students, which is important for TA funding (see \autoref{sec:funding}).
+Many SOP samples for CS are \href{https://cs-sop.org/}{available here}.
 
@@
 In your SOP, focus on research potential and convince us through your experience, e.g., published papers.
-You should also back up your claims with concrete evidence.
+Back up your claims with \textbf{concrete evidence}.
 For example, if you say you have experience with teaching, then show what you did, e.g., undergrad TA or mentoring someone.
 If you say you work on a project, then show some results, e.g., paper submitted (or even rejected), achieved certain performance improvement over the state of the art.
-Also, do not use more than two pages (references can be on the third page).
-Many SOP samples are \href{https://cs-sop.org/}{available here}.
 
@@
 
-Finally, you should get your SOP reviewed by your LoR writers and professors, especially those who have served in adcom, or even postdocs or PhD students as they have been through this process.
+Finally, your SOP reviewed by your LoR writers and professors, especially those who have served in adcom, or even postdocs or PhD students as they have been through this process.
 
@@
 Research experience gives you opportunities to try out research, determine what research area you're interested in, publish papers (\autoref{sec:research-experience}), connect with researchers, and get strong LoRs (\autoref{sec:lor}).
-
- This section provides some guidance on how to gain research experience, e.g., as an undergraduate student.  It also advises students from small colleges where research opportunities are limited.
+This section provides some guidance on how to gain research experience as an undergraduate student or student at a smaller college where research opportunities might be limited.
 

### 2024-10-28 demystify.tex
updates

@@
 So by retracting your offer, you are breaking your commitment and also causing a great deal of inconvenience to the prof and also taking away the opportunity from other students.
-Ultimately, this choice is personal, involving a balance between personal benefit and the ethical impact on others.
+Ultimately, this choice is personal and involves a balance between personal benefit and ethical considerations.
 
@@
 
-It is unlike that you can negotiate TA stipend, which is set by the university.
-Moreover, while RA is set by the adviser, it is often based on the department's standard, and it also would be unfair to other students if you get more (especially since you're new and have not proven yourself yet).
+It is unlike that you can negotiate TA stipend (\autoref{sec:ta}), which is set by the university.
+Moreover, while RA is set by the adviser (\autoref{sec:ra}), it is often based on the department's standard, and it also would be unfair to other students if you get more (especially since you're new and have not proven yourself yet).
 However, your stipend is often increased every year by the university (automatically by a small amount).
@@
 However, note this is a privilege and a courtesy and not a right.
-This means that the prof.
+Your prof.
 is not obligated to buy you things, and that you should not expect these are automatically provided.
@@
 Finally, the prof.
-might have a policy on what they can buy for you, e.g., they prefer to get you desktop instead of a laptop (to encourage you to come to the office and work).
+might have a policy on what they can buy for you, e.g., they prefer to get you desktop instead of a laptop so that you come and work in the lab.
 
@@
 \begin{itemize}
-  \item \href{https://www.cs.columbia.edu/wp-content/uploads/2019/03/Get-Advisor.pdf}{The Definitive ``what do I ask/look for'' in a PhD Advisor Guide}
+  \item \href{https://www.cs.columbia.edu/wp-content/uploads/2019/03/Get-Advisor.pdf}{The Definitive ``what do I ask/look for'' in a PhD Adviser Guide}
 \end{itemize}
@@
 
+\section{Are profs. very busy (and therefore ghost me)?}\label{sec:busy}
+
+Yes, profs.
+are busy.
+They have many deadlines, meetings, and emails, many of which are from prospective students looking for research opportunities.
+Tenure-line profs.
+(\autoref{sec:faculty-types}) typically have 5 jobs (research, teaching, service, grant writing, and mentoring students).
+They also have a life outside of work, e.g., family, hobbies, etc.
+So yes, they are super busy.
+
+However, this is not why they ``ghost'' you.
+The main reason is that they are not interested.
+They might not have funding, the bandwidth to take on new students, or they might not be interested in your background.
+So don't take it personally if you don't get a reply.
+If you want to increase your chance of getting a reply, see \autoref{sec:contact}.
+
+\begin{commentbox}{Vu}
+I scan through \emph{every single email} sent by students for research opportunities (many of which are from undergrads and highschool students).
+However, I only reply to those that I think are a good fit and ask them to chat with me.
+I tend to be more responsive to undergrads and highschool students , especially those in the local area, to encourage them to pursue research.
+\end{commentbox}
+
 \section{How long does it take to evaluate an application?}\label{sec:ievaluate}
@@
 
-  \item[Advisor/Supervisor]: A faculty member who guides and mentors a PhD student throughout their research.
+  \item[Adviser/Superviser]: A faculty member who guides and mentors a PhD student throughout their research.
 This person typically plays a crucial role in a student's academic journey.

### 2024-10-29 demystify.tex
updates

@@
 Most CS programs already receive many applications and would be overwhelmed if the application process were free.
+Even with application fee the competition is already very tough, imagine if the application is free and the number of applications doubles or triples.\\
+
 Note that if you have financial difficulties, you can ask the department for a waiver, but this is typically only granted in exceptional cases.
@@
 
-Most well-known and well-respected profs would \emph{not} ask you to write your own letter (in fact, even not well-known ones wouldn't do this to students they care about).
+\begin{commentbox}[Vu:]
+ Well-known and well-respected profs would \emph{not} ask you to write your own letter (in fact, even not well-known ones wouldn't do this to students they care about).
 This might be a common practice at specific universities and the students do not have a choice as they need the letter.
@@
 So if you are in this situation, you should find someone else to write for you.
+\end{commentbox}
 
@@
 
-\subsection{Help your LOR Writer}]\label{sec:help-your-LOR-writers}
+\subsection{Help your LOR Writer}\label{sec:help-your-LOR-writers}
 
@@
 
-\subsection{Remind Your Writers}
+\subsection{Remind Your Writers}\label{sec:remind-writers}
 
@@
 However, do not send too many reminders as that can be annoying to the writers.
+
+\subsection{Thank Your Writers}\label{sec:thank-writers}
+
+After your application is submitted and your writers have submitted their letters (i.e., the wait begins), you should send a quick thank you note, which serves both as an acknowledgement that you know they have submitted the letters and as an appreciate for their help.
+You should also let them know the outcome of your application, regardless of whether you are admitted or not.
+In addition to being a common courtesy, this can also help maintain a good relationship with your writers, which can be useful in the future (e.g., if you need another letter for another round of applications or for a job reference).
 
@@
 % \end{commentbox}
+
+\subsection{How do you write LoR? Is it lots of work?}\label{sec:lor-writing}
+
+If a student asks me to write letter for them, I will agree as I believe it is part of my job.
+However, I will let them know if I won't be able to write a strong letter for them (e.g., they just take my class and I don't know them well enough or they have not done much research with me), and suggest they find someone else.
+If they insist, then I will write the letter, but it will likely be a weak one.
+However, I will still try to write something positive about the student, e.g., they are hardworking, or that I heard they are working on a research project, etc.
+I usually ask the student to provide me with some details about themselves, e.g., their GPA and transcript,  and use that to write the letter.
+Regardless, the letters will still be short and weak (\autoref{sec:generic-letters}) and usually take about 5-10 minutes to write.
+
+For students who I know well, i.e., strong letters, then it will take a lot longer as it will be personalized.
+While I have a general template of such strong letters, it still can take me an hour or more to write such a letter.
+I also often ask the students to provide information (\autoref{sec:help-your-LOR-writers}) and what they think I should highlight in the letter.
+If they already have a draft of their SOP, they can also provide that to me so I can complement what they say with my own perspective.
+I also offer to review their SOP and provide feedback.
+
+%I do ask the students to waive their right to see the letter, as it is a standard practice and a way of keeping things straightforward and respectful in the whole recommendation game.
+I also let them know that I will not show them the letter, even if they do not waive their right.
+I also let them know that I will not write a letter for them if they do not waive their right, as I believe that such letters are not useful and might even hurt their application.
+
+%\subsection{Will it annoy them if they have to submit many letters? }\label{sec:many-letters}
+%Profs. are used to write letters for their students
 
@@
 
-If your school is the top in your country but you think it might not be well-known internationally, you should talk about it in your SOP (and also ask your LOR writer to do that).
+Many international students mistakenly believe that their school is well-known, but in fact, it is not (e.g., many Vietnamese students believe that ``Bach Khoa'' is well-known internationally, but it is not).
+So you should explicitly mention that in your SOP and ask your LOR writers to do that as well.
 Of course, if you're interested in working with Vietnamese, consider  \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{CS programs in the US that have Vietnamese professors}.
@@
 While having good GPA might not help much (again, because of other much more important factors),
-having bad GPA can be a \red{red flag}. In fact, many universities have a minimum GPA requirement (e.g., > 3.0) and will automatically reject applications with lower GPAs.  
-If you have bad grades, you should explain them in your SOP or have your LOR writers explain them in their letters.
+having bad GPA can hurt your application.  Many universities have a minimum GPA requirement (e.g., $> 3.0$) and will automatically reject applications with lower GPAs.  
+If you have bad grades, you should explain them in your SOP or better yet, have your LOR writers explain them in their letters if they know the reasons.
 
@@
 \begin{center}
-  \includegraphics[scale=0.6]{files/c1.png}
+  \includegraphics[scale=0.5]{files/c1.png}
 \end{center}
 
-International students often put schools into two bins: (i) top schools that they dream about, and (ii) everything else.
+Many students just put schools into two bins: (i) top schools that they dream about, and (ii) everything else.
 They commonly use rankings from US News, which is not transparent and questionable\footnote{\url{https://cra.org/cra-statement-us-news-world-report-rankings-computer-science-universities/}}.
 Sometimes they evaluate based on the reputation of the school's undergrad program or the reputation of the school's non-CS programs such as medical, math, or physics.
-They even rank universities based on popular places they know in the US, e.g., California, Texas, and New York.
-
-Instead of these superficial criteria, you should consider the CS program and the research interests of faculty members.
+Many international students rank universities based on popular places they know in the US, e.g., California, Texas, and New York.
+
+Instead of these superficial criteria, you should specifically consider the CS program and the research interests of faculty members.
 You can learn about these using resources such as \href{https://csrankings.org}{CSRankings.org}.
@@
 
-\paragraph{What matters to you?} While many find CSRankings useful, it is not perfect, and in general, rankings are superficial, and you need to do more research to be informed and make better decisions.
+\paragraph{What matters to you?} While many find CSRankings useful, it is still superficial as every other ranking.
+You should consider other factors that matter to you.
 For example, you might prefer places with a large community from your country (e.g., Northern VA has a large population of Vietnamese).
@@
 I read carefully \emph{strong LoRs} and skim weaker ones, noting either strong letters from well-known professors/researchers or generic content.
-I skim over \emph{CV} and look for publications, research experiences, and notable achievements, noting things like ``published papers in top venues'' or ``gold medal in an international competition.''
-I read strong \emph{SOPs} thoroughly (and skim weak ones).
+I skim over \emph{CV} and look for publications, research experiences, and notable achievements.
+I take notes of things like ``published papers in top venues'' or ``gold medal in an international competition''.
+
+I skim weak \emph{SOPs} but read strong ones thoroughly.
 I note whether the SOP is exciting, research-oriented, standing out, e.g., applicant is minority (\autoref{sec:urm}), and tailored to our program (e.g., if they are familiar with the work of some faculty or have talked to them).
@@
 
-Most of what is written in this document applies to both domestic\footnote{For simplicity domestic means you did your undergrad (or MS) at a US university.} and international students.
+Most of what is written in this document applies to both domestic\footnote{For simplicity domestic means you did your undergrad (or MS) at a US university.
+Different universities might have different definitions, e.g., permanent residents or US citizens.} and international students.
 However, there are some differences and benefits that domestic students should be aware of and can leverage to improve their chances of admission.
-
 \paragraph{Standing out \autoref{sec:improve-your-chance}} There are \emph{few} domestic applications compared to international ones, i.e., domestic students are the minority (\autoref{sec:urm}) in the CS PhD application pool.
@@
 While both MS and PhD programs are graduate degrees, they are \emph{very different} in terms of objective, admission requirements, course requirements, duration, and funding.
-This section provides guidance on applying to MS programs in CS.
+This section discusses the differences and provides guidance on applying to MS programs.
 
@@
 
-MS programs are much less competitive than PhD programs, i.e., you're likely to get in if you can afford it.
-Some refer to MS programs as a \emph{cash cow} for the department, as MS students are often not funded and have to pay tuition.
+In most cases MS CS programs are much less competitive than PhD programs, i.e., you're likely to get in if you can afford it.
+Many think of MS programs as a \emph{cash cow} because students are often not funded and have to pay tuition.
 
@@
 \begin{itemize}
-\item (RA) Profs.
+\item[\textbf{RA}] Profs.
 are not willing to take MS students as RAs because they are not around long enough to be productive.
@@
 
-\item (TA) MS students are typically not given TA positions because PhDs are given priority. Many depts. do not have enough TAs for PhD students and so they cannot afford to give them to MS students.
+\item[\textbf{TA}] MS students are typically not given TA positions because PhDs are given priority.
+Many depts.
+do not have enough TAs for PhD students and so they cannot afford to give them to MS students.
 \end{itemize}
@@
 
-\item \textbf{Living Cost:} MS students are typically not funded, thus living cost can be a big hassle.
+\item \textbf{Living Cost:} MS students are typically not funded, so living cost can be a big factor.
 You should consider the tuition and living expenses of the area.
-Note that sometimes living costs, e.g., renting, seems scary at first but students often find way to make it work, e.g., by sharing an apartment with other students.
+Note that sometimes living costs, e.g., renting, seem scary at first but students often find way to make it work, e.g., by sharing an apartment with other students.
 

### 2024-10-30 demystify.tex
updates

@@
 
+\subsection{Interpreting Response}\label{sec:interpreting-response}
 Even if you do everything right, you will likely get a generic (\autoref{sec:interpreting-response}) or no response from a prof.
@@
 
-\subsection{Interpreting Response}\label{sec:interpreting-response}
-
 Some generic responses are:
 \begin{itemize}
-\item \textbf{No reply}. Very common. You might try again in a few weeks or months, but don't expect a reply. And after a couple of tries, you should move on.  It simply means they are not interested.
-\item \textbf{Not taking students but encourage you to apply}. Polite way of saying not interested and referring you to the admission process.
+\item \textbf{No reply}. This is by far the most common response (see why in \autoref{sec:busy}).  It means they are not interested.
+You might try again in a few weeks or months, but don't expect a reply. And after a couple of tries, you should move on.  It simply means they are not interested.
+\item \textbf{Not taking students but encourage you to apply}.
+Polite way of saying not interested and referring you to the admission process.
+Note that this does not in any way mean that they think you have a good chance of getting admitted.
+
 \item \textbf{Not taken student this year (but encourage you to apply next year)}.
 Polite but generic response.
-But If you do apply next year, you can mention that you contacted them last year and they encouraged you to apply.
+And like the previous one, encouragement to apply does not mean they think you have a good chance of getting admitted.
+
 \item \textbf{Come talk to me after you're admitted}. Generic. Refer you to the admission process. But if you get admitted then you can reply to them and say you're admitted and would like to talk.
+
 \item \textbf{Cannot admit you directly. Need to go through admission process}. Generic. They are not interested and refer you to the admission process.
+
 \item \textbf{I am not taking students but I co-advise/can serve on your committee} While this might sound good, it's generic because it says once you're admitted and have an adviser, then contact me again.
+
 \item \textbf{I am not taking student but Prof.
 X might be}.
-Not common practice as most profs.
+Not common as most profs.
 will not refer you to their colleagues.
@@
 The best positive response is that they want more information from you, e.g., your CV, transcript, paper, or a chat (like an interview).
+
+\section{Are profs. so busy that they completely ignore emails?} \label{sec:busy}
+
+Profs.
+are busy.
+They have many deadlines, meetings, and emails, many of which are from prospective students looking for research opportunities.
+They also have a life outside of work, e.g., family, hobbies, etc.
+
+But this is not why they ``ghost'' you and provide no response.
+The main reason is that a response, no matter how generic, might be misinterpreted by the student.
+For example, if they say they are not taking students but still (strongly) encourage you to apply, the student might take it as a positive sign of admission.
+Or if they say they are not interested, the student might take it as a personal rejection.
+
+So it is safer to not reply at all.
+This is why you should not take it personally if you do not get a reply.
+It is not because profs.
+are busy, but because they are trying to avoid misunderstandings.
+
+\begin{commentbox}{Vu: }
+    I do scan through \emph{every single email} sent by students for admission and research opportunities (many of which are from undergrads and highschool students).
+However, I only reply to those that I think are a good fit and ask them to chat with me.
+    \end{commentbox}
 
@@
 If you don't get in, then you can ask for feedback and improve your profile for the next round.
-
-\section{Are profs. very busy (and therefore ghost me)?}\label{sec:busy}
-
-Yes, profs.
-are busy.
-They have many deadlines, meetings, and emails, many of which are from prospective students looking for research opportunities.
-Tenure-line profs.
-(\autoref{sec:faculty-types}) typically have 5 jobs (research, teaching, service, grant writing, and mentoring students).
-They also have a life outside of work, e.g., family, hobbies, etc.
-So yes, they are super busy.
-
-However, this is not why they ``ghost'' you.
-The main reason is that they are not interested.
-They might not have funding, the bandwidth to take on new students, or they might not be interested in your background.
-So don't take it personally if you don't get a reply.
-If you want to increase your chance of getting a reply, see \autoref{sec:contact}.
-
-\begin{commentbox}{Vu}
-I scan through \emph{every single email} sent by students for research opportunities (many of which are from undergrads and highschool students).
-However, I only reply to those that I think are a good fit and ask them to chat with me.
-I tend to be more responsive to undergrads and highschool students , especially those in the local area, to encourage them to pursue research.
-\end{commentbox}
 

### 2024-10-30 demystify.tex
updates

@@
 But this is not why they ``ghost'' you and provide no response.
-The main reason is that a response, no matter how generic, might be misinterpreted by the student.
+It is because they are trying to avoid misunderstandings.
+A response, no matter how generic, might be misinterpreted by the student.
 For example, if they say they are not taking students but still (strongly) encourage you to apply, the student might take it as a positive sign of admission.
 Or if they say they are not interested, the student might take it as a personal rejection.
-
-So it is safer to not reply at all.
-This is why you should not take it personally if you do not get a reply.
-It is not because profs.
-are busy, but because they are trying to avoid misunderstandings.
+So it is safer to not reply at all to avoid misunderstandings.
 
 \begin{commentbox}{Vu: }
-    I do scan through \emph{every single email} sent by students for admission and research opportunities (many of which are from undergrads and highschool students).
+    I skim through \emph{every email} sent by students for admission and research opportunities (many of which are from undergrads and highschool students).
+So that means I do read your email, and it doesn't go to my spam/trash folder.
     However, I only reply to those that I think are a good fit and ask them to chat with me.

### 2024-10-30 demystify.tex
updates

@@
 \subsection{Interpreting Response}\label{sec:interpreting-response}
-Even if you do everything right, you will likely get a generic (\autoref{sec:interpreting-response}) or no response from a prof.
-Don't take it personally.
-They might not have the bandwidth to take on new students (very common), or they might not be interested in your background.
+Even if you avoid the kiss of death emails (\autoref{sec:kiss-of-death-emails}), you might still not get a reply.
+This is very common and you should not take it personally.
+There are many reasons why you might not get a reply, e.g., the prof.
+is not interested, they are too busy (\autoref{sec:busy}), they are not taking students, they are on vacation, etc.
+Here are some common responses and what they mean:
 
@@
 are busy.
-They have many deadlines, meetings, and emails, many of which are from prospective students looking for research opportunities.
-They also have a life outside of work, e.g., family, hobbies, etc.
-
-But this is not why they ``ghost'' you and provide no response.
-It is because they are trying to avoid misunderstandings.
-A response, no matter how generic, might be misinterpreted by the student.
-For example, if they say they are not taking students but still (strongly) encourage you to apply, the student might take it as a positive sign of admission.
-Or if they say they are not interested, the student might take it as a personal rejection.
-So it is safer to not reply at all to avoid misunderstandings.
+We have many deadlines, meetings, and emails, many of which are from prospective students looking for research opportunities.
+we also have a life outside of work, e.g., family, hobbies, etc.
+
+However, this is not why we ``ghost'' you and provide no response.
+It is because we want to avoid misunderstandings.
+A response (\autoref{sec:interpreting-response}), no matter how clear we think it is, might be misinterpreted by the student.
+For example, if we say we are not taking students but still (strongly) encourage to apply, then the student might take it as a positive sign of admission.
+Or if we say we are not interested, the student might take it as a personal rejection and not apply at all.
+So it is not because we are too busy to reply, but because we do not want misunderstandings and have to deal with them later.
 

### 2024-10-30 ennv.tex
updates

@@
 
-For the widely-used ReLU activation problem, this problem becomes a search for \emph{activation patterns}, i.e., boolean assignments representing activation status of neurons, that lead to satisfaction the formula in \autoref{eq:prob}.
+\section{Activation Pattern Search} For the widely-used ReLU activation problem, the DNN verification problem becomes a search for \emph{activation patterns}, i.e., boolean assignments representing activation status of neurons, that lead to satisfaction the formula in \autoref{eq:prob}.
+
 Modern DNN verification techniques~\cite{bunel2020branch,wang2021beta,ferrari2022complete,duong2024harnessing,duong2023dpllt,ovalbab,katz2019marabou,bak2021nnenum} all adopt this idea and search for satisfying activation patterns.
@@
     \Input{DNN $\mathcal{N}$, property $\phi_{in} \Rightarrow \phi_{out}$}
-    Output{($\unsat, \prooftree$)}
-    \Output{($\unsat, \prooftree$) if property is valid, otherwise ($\sat, \counterexample$)}
+    \Output{$\unsat$ if property is valid, otherwise ($\sat, \counterexample$)}
     \BlankLine
@@
             }
-            \Else(\tcp*[h]{detect a conflict}){
+            %\Else(\tcp*[h]{detect a conflict}){
                 % $\clauses \leftarrow \clauses \cup \AnalyzeConflict(\igraph_i)$ \\ \label{line:analyze_conflict}
-                $\prooftree \leftarrow \prooftree \cup \{ \sigma_i \}$ \tcp{build proof tree} \label{line:record_proof}
-            }
+            %    $\prooftree \leftarrow \prooftree \cup \{ \sigma_i \}$ \tcp{build proof tree} \label{line:record_proof}
+            %}
         % }
@@
     }\label{line:dpllend}
-    \Return{$(\unsat, \prooftree)$}
+    \Return{\unsat}
 
@@
 
-\autoref{fig:alg} shows \dd{}, a reference architecture~\cite{nakagawa2014consolidating} 
-for modern DNN verifiers that we use to illustrate our observations.  
+Modern DNN verifiers often adopt the branch-and-bound (BnB) approach to search for activation patterns that satisfy the DNN verification problem.
+A BnB search consists of two main components: (branch) splitting into the problem smaller subproblems
+by using \emph{neuron splitting}, which decides boolean values representing neuron activation status, 
+and (bound) using abstraction and LP solving to approximate bounds on neuron values to determine 
+the satisfiability of the partial activation pattern formed by the split.
+
+\autoref{fig:alg} shows \dd{}, a reference BnB architecture~\cite{nakagawa2014consolidating} for modern DNN verifiers.
 \dd{} takes as input a ReLU-based DNN $\mathcal{N}$ and a formulae $\phi_{in}\Rightarrow \phi_{out}$ representing the property of interest.
-
-\dd{} iterates between two components: \texttt{Decide} (branching, \autoref{line:decide}), which decides (assigns) an activation status value for a neuron, and \texttt{Deduce} (bounding, \autoref{line:deduce}), which checks the feasibility of the current activation pattern.
-To add proof generation capability, \dd{} is instrumented with a proof tree (\texttt{proof}) variable (\autoref{line:prooftree}) to record these branching decisions.
+\dd{} iterates between (i) branching ()\autoref{line:decide}), which \emph{decides} (assigns) an activation status value for a neuron, and (ii) bounding (\autoref{line:deduce}), which \emph{deduces} or checks the feasibility of the current activation pattern.
+%To add proof generation capability, \dd{} is instrumented with a proof tree (\texttt{proof}) variable (\autoref{line:prooftree}) to record these branching decisions.
 The proof is represented as a binary tree structure, where each node represents a neuron and its left and right edges represent its activation decision (active or inactive).

### 2024-10-30 demystify.tex
updates

@@
 If a student asks me to write letter for them, I will agree as I believe it is part of my job.
-However, I will let them know if I won't be able to write a strong letter for them (e.g., they just take my class and I don't know them well enough or they have not done much research with me), and suggest they find someone else.
-If they insist, then I will write the letter, but it will likely be a weak one.
-However, I will still try to write something positive about the student, e.g., they are hardworking, or that I heard they are working on a research project, etc.
-I usually ask the student to provide me with some details about themselves, e.g., their GPA and transcript,  and use that to write the letter.
-Regardless, the letters will still be short and weak (\autoref{sec:generic-letters}) and usually take about 5-10 minutes to write.
-
-For students who I know well, i.e., strong letters, then it will take a lot longer as it will be personalized.
+However, I will let them know if I won't be able to write a strong letter for them (e.g., they just take my class and I don't know them well enough), and suggest they find someone else.
+If they insist, then I will write the letter.
+It will be a weak and generic letter though I still try to have something positive, e.g., the student is hardworking, or that I heard they are working on a research project, etc.
+Regardless, the letters will still be short and weak (\autoref{sec:generic-letters}) and usually take about 5--10 minutes to write.
+
+For students who I know well, i.e., strong letters, then writing will take a lot longer as it will be personalized.
 While I have a general template of such strong letters, it still can take me an hour or more to write such a letter.
-I also often ask the students to provide information (\autoref{sec:help-your-LOR-writers}) and what they think I should highlight in the letter.
-If they already have a draft of their SOP, they can also provide that to me so I can complement what they say with my own perspective.
-I also offer to review their SOP and provide feedback.
-
-%I do ask the students to waive their right to see the letter, as it is a standard practice and a way of keeping things straightforward and respectful in the whole recommendation game.
-I also let them know that I will not show them the letter, even if they do not waive their right.
-I also let them know that I will not write a letter for them if they do not waive their right, as I believe that such letters are not useful and might even hurt their application.
+I often ask the students to provide information (\autoref{sec:help-your-LOR-writers}) and what they think I should highlight in the letter.
+They can also provide me their SOP (if they already have written one) so I can complement what they say with my own perspective.
 

### 2024-10-30 demystify.tex
updates

@@
 
-\subsection{Waiving Your Right}  Choosing not to look at a reference letter is pretty standard in school and job applications.
+\subsection{Waiving Your Right}\label{sec:waive-right}
+
+When you ask someone to write a letter for you, you will be asked if you want to waive your right to see the letter.  You should \emph{always} waive your right.
+
+Choosing not to look at a reference letter is pretty standard in school and job applications.
 When you waive your right to see the letter, it adds a layer of trust, showing you're confident in your choice of referees and that you're not trying to twist their words.
@@
 If a student asks me to write letter for them, I will agree as I believe it is part of my job.
-However, I will let them know if I won't be able to write a strong letter for them (e.g., they just take my class and I don't know them well enough), and suggest they find someone else.
-If they insist, then I will write the letter.
-It will be a weak and generic letter though I still try to have something positive, e.g., the student is hardworking, or that I heard they are working on a research project, etc.
-Regardless, the letters will still be short and weak (\autoref{sec:generic-letters}) and usually take about 5--10 minutes to write.
+However, I will ask them to waive their right to see the letter (\autoref{sec:waive-right}), and will not write if they do not do so.
+
+I will also let the student know if I can write a strong letter for them.
+If I can't, I will tell them and suggest they find someone else.
+If they insist, then I will write for them.
+While I try to say something positive, e.g., the student is hardworking, or that I heard they are working on a research project, etc, the letter will still be short and weak (\autoref{sec:generic-letters}).
+Usually it takes me about 5--10 minutes to write such a letter.
 

### 2024-10-31 ennv.tex
updates

@@
 
-Modern DNN verifiers often adopt the branch-and-bound (BnB) approach to search for activation patterns that satisfy the DNN verification problem.
+Most modern DNN verifiers adopt the branch-and-bound (BnB) approach to search for activation patterns for DNN verification.
 A BnB search consists of two main components: (branch) splitting into the problem smaller subproblems
@@
 \dd{} takes as input a ReLU-based DNN $\mathcal{N}$ and a formulae $\phi_{in}\Rightarrow \phi_{out}$ representing the property of interest.
-\dd{} iterates between (i) branching ()\autoref{line:decide}), which \emph{decides} (assigns) an activation status value for a neuron, and (ii) bounding (\autoref{line:deduce}), which \emph{deduces} or checks the feasibility of the current activation pattern.
+\dd{} iterates between (i) branching (\autoref{line:decide}), which \emph{decides} (assigns) an activation status value for a neuron, and (ii) bounding (\autoref{line:deduce}), which \emph{deduces} or checks the feasibility of the current activation pattern.
 %To add proof generation capability, \dd{} is instrumented with a proof tree (\texttt{proof}) variable (\autoref{line:prooftree}) to record these branching decisions.

### 2024-10-31 ennv.tex
updates

@@
 Let us assume that it chooses $v_4$ to split, which essentially means the problem is split into two independent subproblems: one with $v_4$ active and the other with $v_4$ inactive.
-\dd{} then adds $v_4$ and $\overline{v_4}$ to \functiontextformat{ActPatterns}.
+\dd{} then adds $\{v_4\}$ and $\{\overline{v_4}\}$ to \functiontextformat{ActPatterns}.
 
@@
 \section{Interval}
+
+An interval abstraction is a simple and widely-used abstraction technique in DNN verification.
+It approximates the range of values that a neuron can take by using a lower and upper bound $[l, u]$.
+The interval abstraction is efficient and easy to implement, but it can be imprecise, especially for non-linear activation functions like ReLU.
+
+\paragraph{Example}: For the DNN in \autoref{fig:dnn}, the interval abstraction for the input layer is $x_1 \in [-1, 1]$ and $x_2 \in [-2, 2]$.
+The interval abstraction for the hidden layer is $x_3 \in [0, 2]$ because $x_3 = [ReLU(-0.5x_1 + 0.5x_2 + 1.0), ReLU(-0.5x_1 + 0.5x_2 + 1.0)] = [ReLU(-0.5 \times -1 + 0.5 \times -2 + 1.0), ReLU(-0.5 \times 1 + 0.5 \times 2 + 1.0)] = [0, 2]$.
+Similarly, $x_4 \in [0, 2]$.
+The interval abstraction for the output layer is $x_5 \in [-3, 3]$.
+
 \section{Zotope}

### 2024-11-02 demystify.tex
updates

@@
 \chapter*{Preface}
-Having been involved in PhD admission committees for many years, I've realized that many \textbf{international} students, especially those in smaller countries or less well-known universities, lack a clear understanding of
+Having been involved in PhD admission committees for many years, I've realized that \textbf{international} students, especially those in smaller countries or less well-known universities, often lack a clear understanding of
 the Computer Science PhD admission process at US universities. This confusion not only
@@
 
-\paragraph{Not a Good fit and Bad luck}  You could have an excellent profile, but if you are interested in a research area that the program does not have, you will not be admitted.
+\paragraph{You aim too high} 
+You have applied to schools that are \emph{way too high} for your profile (\autoref{sec:selecting-ranking-schools}).
+Many students simply just go for very top schools (e.g., top 10) and are surprised when they are rejected to all of them, in multiple years, and completely get shut out.
+This is very obvious but many students still do this.
+An analogy is a person who has never hike wanting to climb Everest, which btw if you could, you might have a better admission chance (\autoref{sec:improve-your-chance}).
+
+While being ambitious and aiming high is good, you should understand how PhD admission works (e.g., by reading \emph{this} document and realizing things such as having a good GPA or GRE doesn't mean much to top programs), that in the US there are many good schools, and just be realistic.
+
+\paragraph{Not a Good fit and Bad luck}  You could have an excellent profile (e.g., strong research and LoRs), but if you are interested in a research area that the program does not have, you will not be admitted.
 Similarly, if no faculty is willing to advise you (e.g., they are on sabbatical or personal leave, do not have sufficient funding, or already have too many students), you will not be admitted.
@@
 
-\paragraph{You aim too high} This is a very common reason: you have applied to schools that are too high for your profile.
-While being ambitious is good, you should understand how PhD admission works (e.g., by reading this document and realizing things such as having a good GPA doesn't mean much to top programs) and be realistic!
-
 Before applying, you should talk to your professors and ask them to give you an honest opinion on where you should apply.
@@
 \paragraph{Low English exam scores (e.g., IELTS or TOEFL)}  Profs.
-might not care much about these, but the college or the university often sets a minimum that you need to pass to be considered, especially for TA funding (e.g., you need to communicate well with students as a TA).
+might not care much about these, but the college or the university often sets a minimum that you need to pass to be considered, especially for TA funding (a low English score causes concern that you might not be able to communicate well with students as a TA).
 Thus, while profs.
@@
 
-\paragraph{Red flags} Various types of red flags can cause concern to the adcom (e.g., too many STEM courses with low or withdrawn grades, plagiarism, cheating, or other academic dishonesty).
-Another one is that you have a history of jumping from one program to another (e.g., you have been in multiple different programs in a short time).
+\paragraph{Red flags} Various types of red flags can cause concern to the adcom.
+Common examples include too many STEM courses with low or withdrawn grades, plagiarism, cheating, or other academic dishonesty.
+Another one is that you have a history of jumping from one program to another.
 Adcom members might have contacts in other programs and heard about your stories, or your LoRs might mention these.
@@
 While these profs.
-might be able to co-advise, they typically do not have the bandwidth, funding, and the desire to mentor students (they are already overloaded with teaching and other service duties).
-So do some research before sending emails, e.g., most CS profs.
+might be able to co-advise, they typically do not have the bandwidth, funding, and the desire to mentor students for research (they are already overloaded with teaching and other service duties).
+So do some homework before sending emails, e.g., most CS profs.
 who are active in research will have a website with their work (publications) and students; some also have lab website dedicated to their group.

### 2024-11-02 demystify.tex
updates

--- 2024-10-30 demystify.tex
@@
 
-Students from top schools with strong research cultures and experience might already know most of the information given here.
-My goal is thus to level the playing field by providing information that is not readily available to less privileged students. 
+Students from top schools with strong research programs and experience might already know most of the information given here.
+My goal is thus to level the playing field by providing info that is not readily available to less privileged students. 
 I hope to encourage more students with such backgrounds to apply to CS PhD programs in the US.
+
 Another goal of this document is to explain the \emph{``why''} behind the admission process, e.g., why LoRs are important, why you should not draft your own LoR, why you should contact professors, etc.
-
-While this document may not fully meet the needs of students aiming for top-10 schools (\autoref{sec:ranking}), which typically have more stringent requirements (e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU), it should be valuable for those targeting other reputable R1 institutions.
+Understanding the reason and mindset of the admission committee and profs.
+can help you prepare better.
+
+While this document may not fully meet the needs of students aiming for top-10 schools (\autoref{sec:selecting-ranking-schools} and \autoref{sec:ranking}), which typically have more stringent requirements (e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU), it should be valuable for those targeting other reputable R1 institutions.
 It is certainly useful for students aiming for Mason!
 
-\begin{commentbox}[Vu:]
-    One of the reasons I created this document is that my colleagues at Mason are interested in recruiting Vietnamese students and are surprised when seeing very few applications in Vietnam (see \autoref{sec:ack}).
+%\begin{commentbox}[Vu:]
+%    One of the reasons I created this document is that my colleagues at Mason are interested in recruiting Vietnamese students and are surprised when seeing very few applications in Vietnam (see \autoref{sec:ack}).
 Each year our CS PhD program receives 500+ PhD applications, most of which are international but only 5--6 are from Vietnam.
-In general the number of CS PhD applications from Vietnam to US universities is very low and more would be welcome.
-  \end{commentbox}
+%  \end{commentbox}
 
@@
 
-Typically, it takes 5--7 years for CS PhD in the US.  This can be longer than CS PhDs in other countries (\autoref{sec:msrequirement}), which might require MS first.
+Typically, it takes 5--7 years for CS PhD in the US.  This can be longer than CS PhDs in other countries (\autoref{sec:non-us-differences}), which might require MS (\autoref{sec:phd-vs-ms}) first.
 
@@
 
-\section{How applications are assigned to adcom members?}\label{sec:applications-assigned}
+\section{Do adcom members read all applications?}\label{sec:read-all}
 \sectioninfo{Adcom members only review applications assigned to them (typically matching their expertise) and rarely get involved in other applications}
@@
 Some universities do waive, e.g., Rice does for all PhD applicants and many universities for domestic students (\autoref{sec:domestic-students}), however most do not for international students.
-Some require proof of financial difficulties, e.g., a statement from a financial adviser or a bank statement.
-
-In gneral, application fee is typically a requirement of the university. Individual departments and programs do not have the flexibility to waive the application fee, even if they want to.
+In any case, the application fee is typically a requirement of the university.
+Individual departments and programs do not have the flexibility to waive the application fee, even if they want to.
 
@@
 In my opinion, requiring applicants to pay the fee helps ensure their seriousness, as it filters out non-serious candidates.
-Most CS programs already receive many applications and would be overwhelmed if the application process were free.
-Even with application fee the competition is already very tough, imagine if the application is free and the number of applications doubles or triples.\\
-
+Also, most CS programs already receive many applications and would be overwhelmed if the application process were free.
 Note that if you have financial difficulties, you can ask the department for a waiver, but this is typically only granted in exceptional cases.
@@
 Such a person is typically someone who has a PhD (so they know what a PhD is about) and is active in research (so they can evaluate your research ability).
-LoR writers are often your research advisers, professors, or supervisors who have worked with you on research projects.
-
-\paragraph{LoR from Famous People} Having a strong letter from an internationally recognized researcher will \emph{greatly strengthen} your application.
+LoR writers are often your research advisors, professors, or supervisors who have worked with you on research projects.
+
+\paragraph{LoR from Famous Ppl?} Having a strong letter from an internationally recognized researcher will \emph{greatly strengthen} your application.
 A strong LOR from such a researcher can outweigh the lack of publications (or low GPA or unknown school).
@@
 
-\begin{commentbox}[Vu:]
- Well-known and well-respected profs would \emph{not} ask you to write your own letter (in fact, even not well-known ones wouldn't do this to students they care about).
+Most well-known and well-respected profs would \emph{not} ask you to write your own letter (in fact, even not well-known ones wouldn't do this to students they care about).
 This might be a common practice at specific universities and the students do not have a choice as they need the letter.
@@
 So if you are in this situation, you should find someone else to write for you.
-\end{commentbox}
 
@@
 
-\subsection{Help your LOR Writer}\label{sec:help-your-LOR-writers}
+\subsection{Help your LOR Writer}]\label{sec:help-your-LOR-writers}
 
@@
 
-\subsection{Remind Your Writers}\label{sec:remind-writers}
+\subsection{Remind Your Writers}
 
@@
 
-\subsection{Thank Your Writers}\label{sec:thank-writers}
-
-After your application is submitted and your writers have submitted their letters (i.e., the wait begins), you should send a quick thank you note, which serves both as an acknowledgement that you know they have submitted the letters and as an appreciate for their help.
-You should also let them know the outcome of your application, regardless of whether you are admitted or not.
-In addition to being a common courtesy, this can also help maintain a good relationship with your writers, which can be useful in the future (e.g., if you need another letter for another round of applications or for a job reference).
-
-% \begin{commentbox}[Vu:]
-% For Vietnamese students, it's worth mentioning the \href{https://vef2.org/}{VEF2.0 program}, which has helped many students gain admission to top CS PhD programs in the US.
+\begin{commentbox}[Vu:]
+For Vietnamese students, it's worth mentioning the \href{https://vef2.org/}{VEF2.0 program}, which has helped many students gain admission to top CS PhD programs in the US.
 VEF2.0 invites US faculty members from leading institutions to conduct rigorous interviews with VEF students and subsequently provide reference letters on their behalf.
 Despite the limited interaction between the interviewer and the interviewee (primarily just the interview), these reference letters are generally effective and have helped many students get into good PhD programs.
-% \end{commentbox}
-
-\subsection{How do you write LoR? Is it lots of work?}\label{sec:lor-writing}
-
-If a student asks me to write letter for them, I will agree as I believe it is part of my job.
-However, I will let them know if I won't be able to write a strong letter for them (e.g., they just take my class and I don't know them well enough), and suggest they find someone else.
-If they insist, then I will write the letter.
-It will be a weak and generic letter though I still try to have something positive, e.g., the student is hardworking, or that I heard they are working on a research project, etc.
-Regardless, the letters will still be short and weak (\autoref{sec:generic-letters}) and usually take about 5--10 minutes to write.
-
-For students who I know well, i.e., strong letters, then writing will take a lot longer as it will be personalized.
-While I have a general template of such strong letters, it still can take me an hour or more to write such a letter.
-I often ask the students to provide information (\autoref{sec:help-your-LOR-writers}) and what they think I should highlight in the letter.
-They can also provide me their SOP (if they already have written one) so I can complement what they say with my own perspective.
-
-%\subsection{Will it annoy them if they have to submit many letters? }\label{sec:many-letters}
-%Profs. are used to write letters for their students
+\end{commentbox}
 
 \section{Statement of Purpose (SOP)}\label{sec:sop}
-\sectioninfo{\emph{SOP is important}. Write it in such a way that makes you \emph{stand out}.}
+\sectioninfo{\emph{SOP is very important}. Write it in such a way that makes you \emph{stand out}.}
 
@@
 statement of purpose (SOP) or personal statement\footnote{Few schools separate these documents and ask you to write both: SOP, which focuses on research experiences, and personal statement, which is everything more personal, e.g., why PhD, challenges, diversity, etc}!
+So write it well because it could make a big difference.
 A well-written SOP also shows that you can communicate, which is very important in research, and that you can effectively teach and communicate with students, which is important for TA funding (see \autoref{sec:funding}).
-Many SOP samples for CS are \href{https://cs-sop.org/}{available here}.
 
@@
 In your SOP, focus on research potential and convince us through your experience, e.g., published papers.
-Back up your claims with \textbf{concrete evidence}.
+You should also back up your claims with concrete evidence.
 For example, if you say you have experience with teaching, then show what you did, e.g., undergrad TA or mentoring someone.
 If you say you work on a project, then show some results, e.g., paper submitted (or even rejected), achieved certain performance improvement over the state of the art.
+Also, do not use more than two pages (references can be on the third page).
+Many SOP samples are \href{https://cs-sop.org/}{available here}.
 
@@
 
-Finally, your SOP reviewed by your LoR writers and professors, especially those who have served in adcom, or even postdocs or PhD students as they have been through this process.
+Finally, you should get your SOP reviewed by your LoR writers and professors, especially those who have served in adcom, or even postdocs or PhD students as they have been through this process.
 
@@
 
-Many international students mistakenly believe that their school is well-known, but in fact, it is not (e.g., many Vietnamese students believe that ``Bach Khoa'' is well-known internationally, but it is not).
-So you should explicitly mention that in your SOP and ask your LOR writers to do that as well.
+If your school is the top in your country but you think it might not be well-known internationally, you should talk about it in your SOP (and also ask your LOR writer to do that).
 Of course, if you're interested in working with Vietnamese, consider  \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{CS programs in the US that have Vietnamese professors}.
@@
 While having good GPA might not help much (again, because of other much more important factors),
-having bad GPA can hurt your application.  Many universities have a minimum GPA requirement (e.g., $> 3.0$) and will automatically reject applications with lower GPAs.  
-If you have bad grades, you should explain them in your SOP or better yet, have your LOR writers explain them in their letters if they know the reasons.
+having bad GPA can be a \red{red flag}. In fact, many universities have a minimum GPA requirement (e.g., > 3.0) and will automatically reject applications with lower GPAs.  
+If you have bad grades, you should explain them in your SOP or have your LOR writers explain them in their letters.
 
@@
 
-However, no interviews mean that you will not likely get an RA (\autoref{sec:ra}), which is offered by an individual faculty (if they want you to do research for them, then they likely will interview you first).
-If you have no interviews, your application (and TA/fellowship funding) is decided by the adcom.
+However, no interviews could mean that you will not likely get an RA (\autoref{sec:ra}), which is offered by an individual faculty (if they want you to do research for them, then they likely will interview you first).
+In the case of no interviews, your application (and TA/fellowship funding) is decided by the adcom.
 
@@
 You can try to contact the school to ask about your status, but they might not reply, they might say they are still reviewing applications, or give you inaccurate information (e.g., you will hear in two weeks).
-In short, you just have to be patient and wait (and also beware that some schools do not send out rejection letters at all).
+In short, you just have to be patient and wait (however, beware that some schools do not send out rejection letters at all).
 
@@
 So by retracting your offer, you are breaking your commitment and also causing a great deal of inconvenience to the prof and also taking away the opportunity from other students.
-Ultimately, this choice is personal and involves a balance between personal benefit and ethical considerations.
+Ultimately, this choice is personal, involving a balance between personal benefit and the ethical impact on others.
 
@@
 
-It is unlike that you can negotiate TA stipend (\autoref{sec:ta}), which is set by the university.
-Moreover, while RA is set by the adviser (\autoref{sec:ra}), it is often based on the department's standard, and it also would be unfair to other students if you get more (especially since you're new and have not proven yourself yet).
+It is unlike that you can negotiate TA stipend, which is set by the university.
+Moreover, while RA is set by the adviser, it is often based on the department's standard, and it also would be unfair to other students if you get more (especially since you're new and have not proven yourself yet).
 However, your stipend is often increased every year by the university (automatically by a small amount).
@@
 However, you can ask for things such as computer equipment (\autoref{sec:buying-equipment}), books, (and maybe even furniture), and a conference travel budget.
-
-\section{Buying Computer Equipment}\label{sec:buying-equipment}
-\sectioninfo{Ask your prof. if they can buy computer equipment and such for your research.}
-
-Students understandably get excited about their upcoming PhD journey and want to buy new computer equipment and electronics to prepare.
-However, you should first check with your professor.
-They might have funding to buy you computers and other equipment (e.g., software, books, keyboards, headphones, tablets, etc).
-Many CS programs also provide budget or computer equipment to incoming PhD students (e.g., a laptop).
-
-However, note this is a privilege and a courtesy and not a right.
-Your prof.
-is not obligated to buy you things, and that you should not expect these are automatically provided.
-Also, while the prof.
-might be able to do so in the past, they might not be able to do so now due to budget constraints.
-It might also a need-based decision, e.g., if you need a powerful computer for your research, then the prof.
-might buy it for you (the lab might already has powerful servers and you can just use them for your work).
-Finally, the prof.
-might have a policy on what they can buy for you, e.g., they prefer to get you desktop instead of a laptop so that you come and work in the lab.
-
-Keep in mind that these computers and equipment would be university property, which might be monitored and have certain restrictions, e.g., do not install illegal software on them (\autoref{sec:illegal-software}).
-You will likely need to return them when you graduate.
 
@@
 \begin{center}
-  \includegraphics[scale=0.5]{files/c1.png}
+  \includegraphics[scale=0.6]{files/c1.png}
 \end{center}
 
-Many students just put schools into two bins: (i) top schools that they dream about, and (ii) everything else.
+International students often put schools into two bins: (i) top schools that they dream about, and (ii) everything else.
 They commonly use rankings from US News, which is not transparent and questionable\footnote{\url{https://cra.org/cra-statement-us-news-world-report-rankings-computer-science-universities/}}.
 Sometimes they evaluate based on the reputation of the school's undergrad program or the reputation of the school's non-CS programs such as medical, math, or physics.
-Many international students rank universities based on popular places they know in the US, e.g., California, Texas, and New York.
-
-Instead of these superficial criteria, you should specifically consider the CS program and the research interests of faculty members.
+They even rank universities based on popular places they know in the US, e.g., California, Texas, and New York.
+
+Instead of these superficial criteria, you should consider the CS program and the research interests of faculty members.
 You can learn about these using resources such as \href{https://csrankings.org}{CSRankings.org}.
@@
 
-\paragraph{What matters to you?} While many find CSRankings useful, it is still superficial as every other ranking.
-You should consider other factors that matter to you.
+\paragraph{What matters to you?} While many find CSRankings useful, it is not perfect, and in general, rankings are superficial, and you need to do more research to be informed and make better decisions.
 For example, you might prefer places with a large community from your country (e.g., Northern VA has a large population of Vietnamese).
@@
 \begin{itemize}
-  \item \href{https://www.cs.columbia.edu/wp-content/uploads/2019/03/Get-Advisor.pdf}{The Definitive ``what do I ask/look for'' in a PhD Adviser Guide}
+  \item \href{https://www.cs.columbia.edu/wp-content/uploads/2019/03/Get-Advisor.pdf}{The Definitive ``what do I ask/look for'' in a PhD Advisor Guide}
 \end{itemize}
@@
 
-\section{Contacting a prof. How to get a desired reply?}\label{sec:contact}
-
-\sectioninfo{Individually faculty member \emph{cannot directly admit} a student---so do not email and ask if you have a chance.
+\chapter{Contacting a prof. How to get a desired reply?}\label{chap:contact}
+
+\chapterinfo{Individually faculty member \emph{cannot directly admit} a student---so do not email and ask if you have a chance.
 However, faculty can \emph{advocate} for a student and therefore increase their admission chance---so do contact and introduce yourself.}
@@
     \item \textbf{Self-focus}.
-Focusing too much about you and your achievements but not why you are interested in the prof.'s work (\autoref{sec:contact}).
+Focusing too much about you and your achievements but not why you are interested in the prof.'s work (\autoref{chap:contact}).
 Mention why you're interested in their work and how your background can contribute.
@@
 
-\subsection{Interpreting Response}\label{sec:interpreting-response}
-Even if you avoid the kiss of death emails (\autoref{sec:kiss-of-death-emails}), you might still not get a reply.
-This is very common and you should not take it personally.
-There are many reasons why you might not get a reply, e.g., the prof.
-is not interested, they are too busy (\autoref{sec:busy}), they are not taking students, they are on vacation, etc.
-Here are some common responses and what they mean:
+Even if you do everything right, you will likely get a generic (\autoref{sec:interpreting-response}) or no response from a prof.
+Don't take it personally.
+They might not have the bandwidth to take on new students (very common), or they might not be interested in your background.
+
+\section{Interpreting Response}\label{sec:interpreting-response}
 
@@
 \begin{itemize}
-\item \textbf{No reply}. This is by far the most common response (see why in \autoref{sec:busy}).  It means they are not interested.
-You might try again in a few weeks or months, but don't expect a reply. And after a couple of tries, you should move on.  It simply means they are not interested.
-\item \textbf{Not taking students but encourage you to apply}.
-Polite way of saying not interested and referring you to the admission process.
-Note that this does not in any way mean that they think you have a good chance of getting admitted.
-
+\item \textbf{No reply}. Very common. You might try again in a few weeks or months, but don't expect a reply. And after a couple of tries, you should move on.  It simply means they are not interested.
+\item \textbf{Not taking students but encourage you to apply}. Polite way of saying not interested and referring you to the admission process.
 \item \textbf{Not taken student this year (but encourage you to apply next year)}.
 Polite but generic response.
-And like the previous one, encouragement to apply does not mean they think you have a good chance of getting admitted.
-
+But If you do apply next year, you can mention that you contacted them last year and they encouraged you to apply.
 \item \textbf{Come talk to me after you're admitted}. Generic. Refer you to the admission process. But if you get admitted then you can reply to them and say you're admitted and would like to talk.
-
 \item \textbf{Cannot admit you directly. Need to go through admission process}. Generic. They are not interested and refer you to the admission process.
-
 \item \textbf{I am not taking students but I co-advise/can serve on your committee} While this might sound good, it's generic because it says once you're admitted and have an adviser, then contact me again.
-
 \item \textbf{I am not taking student but Prof.
 X might be}.
-Not common as most profs.
+Not common practice as most profs.
 will not refer you to their colleagues.
@@
 The best positive response is that they want more information from you, e.g., your CV, transcript, paper, or a chat (like an interview).
-
-\section{Are profs. so busy that they completely ignore emails?} \label{sec:busy}
-
-Profs.
-are busy.
-We have many deadlines, meetings, and emails, many of which are from prospective students looking for research opportunities.
-we also have a life outside of work, e.g., family, hobbies, etc.
-
-However, this is not why we ``ghost'' you and provide no response.
-It is because we want to avoid misunderstandings.
-A response (\autoref{sec:interpreting-response}), no matter how clear we think it is, might be misinterpreted by the student.
-For example, if we say we are not taking students but still (strongly) encourage to apply, then the student might take it as a positive sign of admission.
-Or if we say we are not interested, the student might take it as a personal rejection and not apply at all.
-So it is not because we are too busy to reply, but because we do not want misunderstandings and have to deal with them later.
-
-\begin{commentbox}{Vu: }
-    I skim through \emph{every email} sent by students for admission and research opportunities (many of which are from undergrads and highschool students).
-So that means I do read your email, and it doesn't go to my spam/trash folder.
-    However, I only reply to those that I think are a good fit and ask them to chat with me.
-    \end{commentbox}
 
@@
 I read carefully \emph{strong LoRs} and skim weaker ones, noting either strong letters from well-known professors/researchers or generic content.
-I skim over \emph{CV} and look for publications, research experiences, and notable achievements.
-I take notes of things like ``published papers in top venues'' or ``gold medal in an international competition''.
-
-I skim weak \emph{SOPs} but read strong ones thoroughly.
+I skim over \emph{CV} and look for publications, research experiences, and notable achievements, noting things like ``published papers in top venues'' or ``gold medal in an international competition.''
+I read strong \emph{SOPs} thoroughly (and skim weak ones).
 I note whether the SOP is exciting, research-oriented, standing out, e.g., applicant is minority (\autoref{sec:urm}), and tailored to our program (e.g., if they are familiar with the work of some faculty or have talked to them).
@@
 
+\section{Buying Computer Equipment}\label{sec:buying-equipment}
+\sectioninfo{Ask your prof. if they can buy computer equipment and such for your research.}
+
+Students understandably get excited about their upcoming PhD journey and want to buy new computer equipment and electronics to prepare.
+However, you should first check with your professor.
+They might have funding to buy you computers and other equipment (e.g., software, books, keyboards, headphones, tablets, etc).
+Many CS programs also provide budget or computer equipment to incoming PhD students (e.g., a laptop).
+
+However, note this is a privilege and a courtesy and not a right.
+This means that the prof.
+is not obligated to buy you things, and that you should not expect these are automatically provided.
+Also, while the prof.
+might be able to do so in the past, they might not be able to do so now due to budget constraints.
+It might also a need-based decision, e.g., if you need a powerful computer for your research, then the prof.
+might buy it for you (the lab might already has powerful servers and you can just use them for your work).
+Finally, the prof.
+might have a policy on what they can buy for you, e.g., they prefer to get you desktop instead of a laptop (to encourage you to come to the office and work).
+
+Keep in mind that these computers and equipment would be university property, which might be monitored and have certain restrictions, e.g., do not install illegal software on them (\autoref{sec:illegal-software}).
+You will likely need to return them when you graduate.
+
 % \section{Will I be miserable during my a PhD?}\label{sec:happy}
@@
 
-  \item[Adviser/Superviser]: A faculty member who guides and mentors a PhD student throughout their research.
+  \item[Advisor/Supervisor]: A faculty member who guides and mentors a PhD student throughout their research.
 This person typically plays a crucial role in a student's academic journey.
@@
 
-Most of what is written in this document applies to both domestic\footnote{For simplicity domestic means you did your undergrad (or MS) at a US university.
-Different universities might have different definitions, e.g., permanent residents or US citizens.} and international students.
+Most of what is written in this document applies to both domestic\footnote{For simplicity domestic means you did your undergrad (or MS) at a US university.} and international students.
 However, there are some differences and benefits that domestic students should be aware of and can leverage to improve their chances of admission.
+
 \paragraph{Standing out \autoref{sec:improve-your-chance}} There are \emph{few} domestic applications compared to international ones, i.e., domestic students are the minority (\autoref{sec:urm}) in the CS PhD application pool.
@@
 \chapter{MS Admission}\label{chap:ms}
-\chapterinfo{MS focuses on coursework and prepares you for \emph{industry}, while PhD focuses on research and prepares you for academia or research.}
+\chapterinfo{MS focuses on coursework and prepares you for industry, while PhD focuses on research and prepares you for academia or research.}
 
 While both MS and PhD programs are graduate degrees, they are \emph{very different} in terms of objective, admission requirements, course requirements, duration, and funding.
-This section discusses the differences and provides guidance on applying to MS programs.
-
-\section{Differences between PhD and MS}
+This section provides guidance on applying to MS programs in CS.
+
+\section{Differences between PhD and MS}\label{sec:phd-vs-ms}
 
@@
 
-  \item Objective: an MS is typically to prepare you for \emph{industry}, while a PhD is to prepare you for research and academia.
-Some MS has thesis option but the research is much lighter compared to a PhD.
+  \item Objective: an MS is typically to prepare you for industry, while a PhD is to prepare you for research. Some MS has thesis option but the research is much lighter compared to a PhD.
 
@@
 You graduate with an MS when you're done with the courses.
-PhD also has coursework requirements (typically taken in your first 2 years as mentioned in \autoref{sec:time}) but focuses on research after that.
+PhD also has coursework requirements (typically taken in your first 2 years) but focuses on research after that.
 You graduate with a PhD when you have done enough research and written a dissertation, which usually takes much longer time than coursework.
@@
 
-  \item Funding: MS is typically \emph{not funded}, while PhD is (\autoref{sec:funding}). See \autoref{sec:ms-funding} for more details on MS funding.
+  \item Funding: MS is typically \emph{not funded}, while PhD is typically funded (\autoref{sec:funding}). See \autoref{sec:ms-funding} for more details on MS funding.
 
@@
 
-In most cases MS CS programs are much less competitive than PhD programs, i.e., you're likely to get in if you can afford it.
-Many think of MS programs as a \emph{cash cow} because students are often not funded and have to pay tuition.
+Except for a very few highly selective programs, MS programs are typically much less competitive than PhD programs, i.e., you're likely to get in if you can afford it.
+Some refer to MS programs as a \emph{cash cow} for the department, as MS students are often not funded and have to pay tuition.
 
@@
 In most cases, the main requirements are that you have sufficient background in CS, e.g., through your undergrad degree.
-This does not mean MS programs are easy to get admitted, but the requirements are much lower compared to PhD programs.
 
@@
 MS admission also involves an adcom that reviews applications and makes admission decisions.
-However, MS admission is typically rolling, i.e., they review applications as they come in and make decisions throughout the admission cycle.
+However, MS admission is typically rolling (i.e., they review applications as they come in and make decisions throughout the admission cycle).
 Moreover, unlike PhD that has multiple reviewers for each application, an MS application typically involves only one reviewer and does require much time to review compared to a PhD application (\autoref{sec:ievaluate}).
@@
 However, research experience is not required, and LoRs can be from anyone who can speak about your academic or working abilities.
-Moreover, your SOP should focus on your academic and work background, why you want to get an MS, and how the program fits your goals.
+Moreover, your SOP you should focus on your academic and work background, why you want to get an MS, and how the program fits your goals.
 As with PhDs, GREs are often not required.
@@
 \begin{itemize}
-\item[\textbf{RA}] Profs.
+\item (RA) Profs.
 are not willing to take MS students as RAs because they are not around long enough to be productive.
-It can take a while for a student to get up to speed and start being productive, and by that time, they are about to graduate.
+It can take a while for a student to get up to speed and start being productive, and by the time they are, they are about to graduate.
 Moreover, the goal of most MS students is to get a job, not to do research, so they are not as motivated to do research.
-
-\item[\textbf{TA}] MS students are typically not given TA positions because PhDs are given priority.
+\item (TA) MS students are typically not given TA positions because PhDs are given priority.
 Many depts.
-do not have enough TAs for PhD students and so they cannot afford to give them to MS students.
+even have shortage of TA positions for PhD students and so they cannot afford to give them to MS students.
 \end{itemize}
 
-However, there are always exceptions, e.g., if you have a strong background and can demonstrate that you can be productive in research, then a prof. might be willing to take you as their RA.
-Moreover, some schools have TAs for MS students, and you can apply for these positions. For example, Mason CS has quite a few TAs for MS students; my courses in the past two years have MS TAs.
-
-\section{Selecting and Ranking Schools}\label{sec:selecting-ranking-schools-ms}
-%\sectioninfo{How to select and rank MS programs in CS.}
-
-Because of the differences between MS and PhD programs, you should consider different factors when selecting and ranking MS programs.
-For example, you might want to consider factors such as location, industry connections, and job placement instead of research areas and advisers (\autoref{sec:schoolsandprofs}).
-
-\begin{itemize}
-
-\item \textbf{Location:} In addition to living in a place you like (e.g., warm weather, historical city, etc.), you should also consider the job opportunities in the area.
-For tech industry, consider MS programs in tech hubs like the Bay Area and Seattle.
-For government jobs, look at schools in the DC area.
-For example, many students at Mason work for the government or defense contractors in the DC area and take classes in the evening or online.
-
-\item \textbf{Industry Connections:} Universities often have strong connections with local companies and can help you get internships and jobs.
-For example, Microsoft and Amazon look for students in the Washington area while  Google and Facebook recruit those in the Bay Area.
-Mason students natural get jobs from the government, defense contractors, and Amazon, whose 2nd HQ is in Arlington.
-
-\item \textbf{Ranking:} Unlike PhD programs where the adviser and research are perhaps the most important factor, in MS programs the ranking of the school is more important.
-Thus, you should consider the ranking of the school in CS, especially in the specific area you're interested in.
-For example, if you're interested in software engineering, you might want to consider schools with strong SWE programs.
-
-\item \textbf{Living Cost:} MS students are typically not funded, so living cost can be a big factor.
-You should consider the tuition and living expenses of the area.
-Note that sometimes living costs, e.g., renting, seem scary at first but students often find way to make it work, e.g., by sharing an apartment with other students.
-
-\end{itemize}
+However, there are always exceptions, e.g., if you have a strong background and can demonstrate that you can be productive, then a prof. might be willing to take you as their RA.
+Moreover, some schools have TAs for MS students, and you can apply for these positions. For example, Mason CS dept has quite a few TAs for MS students; my courses in the past two years have MS TAs.
 
@@
 
-As discussed in \autoref{sec:research-experience}, having a successful research experience can greatly strengthen a PhD application.
-Research experience gives you opportunities to try out research, determine what research area you're interested in, publish papers (\autoref{sec:research-experience}), connect with researchers, and get strong LoRs (\autoref{sec:lor}).
-This section provides some guidance on how to gain research experience as an undergraduate student or student at a smaller college where research opportunities might be limited.
+As discussed in \autoref{sec:research-opportunities}, research experience is a crucial component of a strong PhD application.
+It gives you opportunities to try out research, determine what research area you're interested in, publish papers (\autoref{sec:research-experience}), connect with researchers, and get strong LoRs (\autoref{sec:lor}).
+
+ This section provides some guidance on how to gain research experience, e.g., as an undergraduate student.  It also advises students from small colleges where research opportunities are limited.
 
@@
 Finally, you can also ask your peers who are already doing research.
-Use the method described in \autoref{sec:contact} to contact professors
+Use the method described in \autoref{chap:contact} to contact professors
 
 \begin{commentbox}[Vu:]
-    I enjoy working with undergrads (e.g., I have mentored over 15 undergrads) and always actively recruit.
+    I enjoy working with undergrads (e.g., have mentored over 10 undergrads) and always actively recruit.
     I often get undergraduate researchers through my classes, e.g., asking students who did well in my class if they are interested in research.
-Occasionally I was introduced to students by other students or faculty.\\
+Occasionally I was introduced to students by other students or faculty.
 
     While most undergrads are understandably not super productive in research, some are and I have published multiple papers with them.
-I also have written strong LoRs for them and have helped them get into good PhD programs.
+I also have written strong LoRs for them and have helped them get into PhD programs, including top ones.
 

### 2024-11-02 demystify.tex

--- 2024-11-02 demystify.tex
@@
 
-\section{Differences between PhD and MS}
+\section{Differences between PhD and MS}\label{sec:phd-vs-ms}
 

### 2024-11-03 demystify.tex
updates

@@
 Faculty often receive ``cold'' e-mails from prospective students.
-Most of the time, we ignore these emails, but on some rare occasions, we do answer them.
+Most of the time, we ignore these emails (\autoref{sec:busy}), but on some rare occasions, we do answer them.
 So how to write an email that gets our attention?
@@
 
-However, this is not why we ``ghost'' you and provide no response.
+However, this is not why we ``ghost'' you and provide no response (\autoref{sec:interpreting-response}).
 It is because we want to avoid misunderstandings.

### 2024-11-03 demystify.tex
updates

--- 2024-11-02 demystify.tex
@@
 
-Students from top schools with strong research programs and experience might already know most of the information given here.
-My goal is thus to level the playing field by providing info that is not readily available to less privileged students. 
+Students from top schools with strong research cultures and experience might already know most of the information given here.
+My goal is thus to level the playing field by providing information that is not readily available to less privileged students. 
 I hope to encourage more students with such backgrounds to apply to CS PhD programs in the US.
-
 Another goal of this document is to explain the \emph{``why''} behind the admission process, e.g., why LoRs are important, why you should not draft your own LoR, why you should contact professors, etc.
-Understanding the reason and mindset of the admission committee and profs.
-can help you prepare better.
-
-While this document may not fully meet the needs of students aiming for top-10 schools (\autoref{sec:selecting-ranking-schools} and \autoref{sec:ranking}), which typically have more stringent requirements (e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU), it should be valuable for those targeting other reputable R1 institutions.
+
+While this document may not fully meet the needs of students aiming for top-10 schools (\autoref{sec:ranking}), which typically have more stringent requirements (e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU), it should be valuable for those targeting other reputable R1 institutions.
 It is certainly useful for students aiming for Mason!
 
-%\begin{commentbox}[Vu:]
-%    One of the reasons I created this document is that my colleagues at Mason are interested in recruiting Vietnamese students and are surprised when seeing very few applications in Vietnam (see \autoref{sec:ack}).
+\begin{commentbox}[Vu:]
+    One of the reasons I created this document is that my colleagues at Mason are interested in recruiting Vietnamese students and are surprised when seeing very few applications in Vietnam (see \autoref{sec:ack}).
 Each year our CS PhD program receives 500+ PhD applications, most of which are international but only 5--6 are from Vietnam.
-%  \end{commentbox}
+In general the number of CS PhD applications from Vietnam to US universities is very low and more would be welcome.
+  \end{commentbox}
 
@@
 
-Typically, it takes 5--7 years for CS PhD in the US.  This can be longer than CS PhDs in other countries (\autoref{sec:non-us-differences}), which might require MS (\autoref{sec:phd-vs-ms}) first.
+Typically, it takes 5--7 years for CS PhD in the US.  This can be longer than CS PhDs in other countries (\autoref{sec:msrequirement}), which might require MS first.
 
@@
 Given the high number of quality applicants and a limited number of spots, in addition to having a good application profile, you want to show something that makes you \textbf{stand out}.
-For example, even if you do not have research experience, you can talk about your personal projects, as long as they can help show you can do research.
-For example, if you have an open-source project that has lots of stars in Github, then you should mention it.
+For example, even if you do not have research experience, you can talk about your personal projects, as long as they show you can do research.
+For example, if you have an open-source project that has lots of stars in Github, mention it.
 If you often write technical, research-like blogs with many viewers, talk about them too.
-
-There are other things you might not think are important but can make you stand out.
-For example, if you have a strong background in a non-CS field but can be integrated with CS, e.g., you have a degree or background in \emph{dance} or \emph{music} and want to integrate them with CS, do talk about it.
-Are you a female or a minority in CS (\autoref{sec:urm})?
-Do you participate in outreach activities that help increase diversity and inclusion in CS?
-Diversity is highly valued in CS programs in the US, and experiences in these areas can make you stand out and get noticed from reviewers.
 
@@
 % It is worth noting that in many cases,  an application could be rejected by most reviewers, however have a "champion", who doesn't even have to be in the Admission comittee, that strongly advocate for your profile.
+
+There are other things you might not think are important but can make you stand out.
+For example, if you have a strong background in a non-CS field but can be integrated with CS, e.g., you have a degree or background in \emph{dance} or \emph{music} and want to integrate them with CS, do talk about it.
+Are you a female or a minority in CS?
+Do you participate in outreach activities that help increase diversity and inclusion in CS?
+Diversity is highly valued in CS programs in the US (\autoref{sec:urm}), and experiences in these areas can make you stand out and get noticed from reviewers.
 
@@
 
-\section{Differences between PhD and MS}\label{sec:phd-vs-ms}
+\section{Differences between PhD and MS}
 

### 2024-11-07 demystify.tex
updates

@@
 
-\subsection{English Tests (IELTS, TOEFL)} Unless your degrees are from specific countries such as \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-Mason#standard-tests-waiver-eligible-countries}{these}, you will need to take standardized English tests.
+\subsection{English Tests (IELTS, TOEFL)} Unless your degrees are from specific countries such as \href{https://www.gmu.edu/international/english-language-requirements}{these}, you will need to take standardized English tests.
 On one hand you will need to show some level of English proficiency, but on the other hand, you do not need to have very good scores in these tests (many adcom members themselves were once international students and struggled with English).
@@
 
-One confusion students often have is whether they will get paid over the summer.  This depends on the funding source.
+Students often get confused about summer funding, e.g., if they will get paid during the summer.
+This is a good question because PhD students have no classes over the summer and so are often more productive and produce good research.
+However, summer salary is largely depending on your funding source.
+
 First, recall that an academic year (AY), which consists of Fall and Spring semesters, is typically 9 months, so your stipend is for 9 months (and many places allow you to spread it over 12 months).

### 2024-11-07 demystify.tex
updates

@@
 \epigraph{Bart, with \$10,000, we’d be millionaires! We could buy all kinds of useful things like…love!’}{\textsc{The Simpsons}}
+
 If you're admitted to a \emph{good} CS PhD program, you should not have to worry about funding!
@@
 Finally, fellowship is independent funding that can come from a school, a company, or an organization.
+
 \autoref{tab:funding} summarizes the differences.
-Note that funding is typically more available for PhD students than
-MS.
+Note that funding is typically more available for PhD students than MS (\autoref{sec:ms-funding}).
 
@@
 The college and department will set the rate for a 9-month graduate assistant stipend.
-TA, which is paid by the department, will likely be that amount but RA might be higher depending on the stage of the student (1st year vs ABD\footnote{All but dissertation: close to graduate.}) and the prof.
+TA, which is paid by the department, will likely be that amount but RA might be higher depending on  the prof and the stage of the student (1st year vs ABD\footnote{All but dissertation: close to graduate.}).\\
+
+  At Mason, CS TA/RA is paid the highest.
+First, CS is under the College of Engineering and Computing (CEC), which already has a higher TA stipend than other colleges.
+Second, CS is the largest department in the college, and has the higher TA stipend than what CEC suggests.
+Finally, it is a tradition that CS faculty pay their RAs more than the department's TA rate.
+So a winning combination for CS PhD students.
+
   \tcblower
@@
 
-\subsection{Application Materials}
+\subsection{Application Materials}\label{sec:ms-application}
 

### 2024-11-10 demystify.tex
updates

@@
 
-    {\fontsize{16}{19.2} \selectfont \textcolor{ForestGreen}{ \bf ThanhVu (Vu) Nguyen}}\\[0.1in]
+    {\fontsize{16}{19.2} \selectfont \textcolor{ForestGreen}{\bf ThanhVu (Vu) Nguyen}}\\[0.1in]
     George Mason University, Dept. of Computer Science\\[0.1in]
@@
 So I want to share some details about the admission process and advice for those who are interested in applying for a \textbf{PhD in Computer Science in the US}.
-Originally, this document was intended for \emph{international students}, but I have expanded it to include information that might also be useful for \emph{US domestic students}.
+Originally, this document was intended for \emph{international students}, but has been expanded to include information for \emph{US domestic students} (\autoref{sec:domestic-students}).
 Moreover, while this primarily aims at students interested in CS, it might be relevant to students from various STEM (Science, Technologies, Engineering, and Mathematics) disciplines.
@@
 \sectioninfo{Individually faculty member \emph{cannot directly admit} a student---so do not email and ask if you have a chance.
-However, faculty can \emph{advocate} for a student and therefore increase their admission chance---so do contact and introduce yourself.}
+However, faculty can \emph{advocate} for a student and therefore increase their admission chance---so contact and introduce yourself.}
 
@@
 \section{How to call or address a professor?}\label{sec:address}
-\sectioninfo{What should you call your profs.? Many possibilities (e.g., prof., Dr., and even their firstnames), but not Mr. or Mrs.}
+\sectioninfo{What should you call your profs.? Many possibilities (e.g., prof., Dr., and even their firstnames), but not Mr., Mrs., or their firstname in the first email.}
 
@@
 \chapter{MS Admission}\label{chap:ms}
-\chapterinfo{MS focuses on coursework and prepares you for \emph{industry}, while PhD focuses on research and prepares you for academia or research.}
+\chapterinfo{MS focuses on coursework and prepares you for \emph{industry}, while PhD focuses on research and prepares you for academia or research. So admission requirements are different.}
 
@@
 Before Mason, I was at the University of Nebraska-Lincoln (UNL).
+Both UNL and Mason are large R1 research universities in the US and have many international students in their CS graduate programs.
 I have been in the PhD admission process at Mason and UNL for many years.
 This allows me to have a good understanding of the process, the challenges students face, and what faculty are looking for.
-I also serve as the program director of the MS program in Software Engineering at Mason (thus also have experience with MS admission---which is quite different than PhD as discussed in \autoref{chap:ms}).
+Currently I serve as the program director of the MS program in Software Engineering at Mason (thus also have experience with MS admission---which is quite different than PhD as discussed in \autoref{chap:ms}).
 My personal and lab website is at \href{https://dynaroars.github.io}{dynaroars.github.io}.
@@
 
-\textbf{Thank you!}
-
 %\bibliographystyle{abbrv}

### 2024-11-16 demystify.tex
updates

@@
 
-\section{Is an MS required for admission to PhD in CS?}\label{sec:msrequirement}
+\section{Is an MS required for admission to PhD in CS in the US?}\label{sec:msrequirement}
 \sectioninfo{You do not need an MS to do PhD in CS.}
 
-No.
-In fact, students can get MS degree ``along the way'' to PhD, e.g., after finishing the 2-year course work.
+No, not in the US (other countries often do though~\autoref{sec:non-us-differences}).
+In the US, it is common to apply for a PhD program directly after your undergrad.
+Most CS PhD programs are designed so that students can get MS degree ``along the way'' to PhD, e.g., after finishing the 2-year course work.
+
 However, MS can help admission if it gives research experience or is from a more well-known school than your undergrad institution (\autoref{sec:your-school}).
-
-If you have an MS then some course work \emph{might be} transferred for course credits, which \emph{might} save some time. But in general don't count finishing earlier just because you have an MS.
+Moreover, if you have an MS then some course work \emph{might be} transferred for course credits, which \emph{might} save some time.
+But in general don't count finishing earlier just because you have an MS.
 
@@
 
-\emph{MS requirement and PhD duration}:  CS PhD programs in the US do not require an MS degree (as mentioned in \autoref{sec:msrequirement} and \autoref{sec:time}).
+\emph{MS requirement and PhD duration}:  CS PhD programs in the US do not require an MS degree (as mentioned in \autoref{sec:time} and \autoref{sec:msrequirement}).
 In contrast, many other countries require having an MS degree before joining a PhD program.

### 2024-11-16 demystify.tex
updates

@@
 
-\subsubsection{Types of Faculty} There are typically two types of faculty at an R1 research university: tenure-line and teaching.
+\paragraph{Faculty} There are typically two types of faculty at an R1 research university: tenure-line and teaching.
 The main difference is their focus: research, teaching, or both.
@@
 
-\textbf{Tenure-line faculty}, consisting of tenured and tenure-track profs., focuses on research, which includes publishing papers, obtaining grants, and mentoring PhD students.
+\emph{Tenure-line faculty}, consisting of tenured and tenure-track profs., focuses on research, which includes publishing papers, obtaining grants, and mentoring PhD students.
 They often have a low teaching load, e.g., 1 course per semester.
@@
 
-\textbf{Teaching faculty}, sometimes called instructional faculty or professor of practice, mainly focuses on teaching.
+\emph{Teaching faculty}, sometimes called instructional faculty or professor of practice, mainly focuses on teaching.
 They typically teach 3--4 classes per semester (which is quite a lot) and \emph{do not} have research responsibilities, e.g., they do not have to worry much about publishing papers or obtaining grants.
@@
 
-Some universities have \textbf{research faculty} or research scientists, who are typically non-tenure line faculty and focus on research (e.g., publishing papers and getting grants). 
+Some universities have \emph{research faculty} or research scientists, who are typically non-tenure line faculty and focus on research (e.g., publishing papers and getting grants). 
 They can recruit students and provide RA funding for them. 
@@
 
-\textbf{Adjunct faculty} is not full-time, e.g., they might be working in industry and teach a class or two for fun.
+\emph{Adjunct faculty} is not full-time, e.g., they might be working in industry and teach a class or two for fun.
 \textbf{Emeritus faculty} means they are retired from the university, e.g., they could move to the industry, but still have some affiliation with the university.
@@
 
-\subsubsection{Ranks} Faculty have ranks such as assistant, associate, and full, regardless if they are tenure-line or teaching.
+\paragraph{Ranks} Faculty have ranks such as assistant, associate, and full, regardless if they are tenure-line or teaching.
 Assistant means new faculty, associate means they have been promoted (or was hired after being in the industry for a while), and full means they are senior.
@@
 Typically, a tenure-track faculty starts with being an assistant professor and then gets tenure and promoted to associate (around 6 years).
-Note that getting tenure is a huge deal, and it might change the way they work with students, e.g., they might be more relaxed and less stressed (\autoref{sec:tenure-vs-tenure-track}).
+Note that getting tenure is a huge deal, and it might change the way they work with students (\autoref{sec:tenure-vs-tenure-track}).
 

### 2024-11-18 demystify.tex
updates

@@
 
-\subsection{Why CS depts do not waive the application fee?}
+\subsection{Why CS depts do not waive the application fee?}\label{sec:fee-waive}
 
@@
 
-In gneral, application fee is typically a requirement of the university. Individual departments and programs do not have the flexibility to waive the application fee, even if they want to.
+In general, application fee is typically a requirement of the university. Individual departments and programs do not have the flexibility to waive the application fee, even if they want to.
 
@@
 % \subsection{Outline of a LOR}\label{sec:lor-outline}
+
+\subsection{Diversity Statement}\label{sec:diversity-statement}
+
+In recent years many universities start requiring a statement about diversity, equity, and inclusion (DEI) as part of the application (also see \autoref{sec:urm}).
+This has caused confusion among many applicants, especially international students.
+They often wonder what to write about, especially if they are not from a minority group, have not experienced anything related the topic, or even not sure what it is about.
+
+A DEI statement is about showing your \emph{awareness of DEI}, your understanding of the challenges others face, and actions you've taken to contribute.
+Customize what you write to your own background and experiences.
+Here are some tips:
+
+\begin{itemize}
+\item \textbf{You don't have to be from a minority group.} You just need to understand what DEI is, e.g., be aware of challenges faced by minority groups in studying CS, STEM fields, academia, workplace.
+Show that you understand the importance of DEI and how it can benefit the CS field and society.
+
+\item \textbf{Show your awareness, customized to your own experiences}.
+For example, if you come from a different culture or country, talk about DEI here, e.g., many countries completely lack diversity in CS programs, talk about it, what does minority group face in your country?
+How do you think DEI can benefit?
+
+\item \textbf{Highlight your contributions}.
+You do not need to be in any group or organization specific to DEI.
+You can talk about your contributions to your community, e.g., mentoring students, organizing events, or even just helping people in needs.
+If you don't believe you have done anything much, then talk about what you plan to do.
+
+\item \textbf{Customized it to your application} Can your PhD study or research contribute to DEI?
+Can you help or mentor students from underrepresented groups from your country?
+These are very doable and can be very convincing.
+
+\end{itemize}
+
+\begin{commentbox}[Vu:]
+One common pitfall is to write about how you are from a poor family, a small town, or a developing country.
+These are not related to DEI and raise concerns about you not understanding or even bother to do some basic research about the topic.
+I have seen people consider themselves as minority because they are Asian or international students studying in the US (have you seen how many Asian or international students are in CS programs in the US?).\\
+
+Moreover, do not make up stories or exaggerate your experiences.
+Just show your awareness and understanding of the topic and what you did or how you plan to contribute to it (e.g., mentoring students, helping others).
+One of the best DEI essay I have read start the acknowledgement that the applicant is not a minority, but a white male in a STEM field (i.e., that make them the majority), and that they do not have much experience in DEI, but show that they are aware of the challenges faced by minority groups and how they can contribute to it.
+\end{commentbox}
 
@@
 \begin{commentbox}[Hung:]
-  I always encourage the students I admitted to talk with my students and the students of other faculty in other schools who admitted them.
+  I always encourage the students I admit to talk with my students and the students of other faculty in other schools who admitted them.
 You will unlikely hear straight-out complaints from current students in a prof’s group.
@@
 
-However, this is not why we ``ghost'' you and provide no response (\autoref{sec:interpreting-response}).
+However, this is \emph{not} why we ``ghost'' you and provide no response (\autoref{sec:interpreting-response}).
 It is because we want to avoid misunderstandings.
@@
 
-  \item[Adcom \autoref{sec:evalapps}] Admission Committee. The group of faculty members who review applications and make admission decisions.
-
-  \item [Adcom members \autoref{sec:evalapps}] Faculty members who are part of the admission committee. These are the people who review applications and make admission decisions.
-
-  \item[Adcom chair \autoref{sec:evalapps}] The faculty member who leads the admission committee.
+  \item[Adcom~\autoref{sec:evalapps}] Admission Committee. The group of faculty members who review applications and make admission decisions.
+
+  \item [Adcom members~\autoref{sec:evalapps}] Faculty members who are part of the admission committee. These are the people who review applications and make admission decisions.
+
+  \item[Adcom chair~\autoref{sec:evalapps}] The faculty member who leads the admission committee.
 This person is often not involved in individual admission decisions but oversees the process (e.g., resolve disputes, ensure fairness).
@@
   \item [Deferred Admission] An option allowing admitted students to postpone their start date, typically by one year.
-  \item [Diversity Statement] A document sometimes required in applications where applicants describe their contributions to diversity and inclusion, or how their background adds to the diversity of the academic community.
-This document is often required for faculty positions, but some universities also require it for graduate admissions.
+
+  \item [Diversity Statement~\autoref{sec:diversity-statement}] A document sometimes required in applications where applicants describe their contributions to diversity and inclusion, or how their background adds to the diversity of the academic community.
+This document is often required for faculty positions, but more universities are requiring it for graduate admissions.
 Note that you do not need to be from an underrepresented group to write a diversity statement.
@@
 
-\paragraph{Fee Waiving} Some schools might offer application fee waivers for domestic students.  You should check with the school you're applying to.
+\paragraph{Fee Waiving~\autoref{sec:fee-waive}} Some schools might offer application fee waivers for domestic students.  You should check with the school you're applying to.
 

### 2024-11-18 demystify.tex
updates

@@
 In recent years many universities start requiring a statement about diversity, equity, and inclusion (DEI) as part of the application (also see \autoref{sec:urm}).
-This has caused confusion among many applicants, especially international students.
-They often wonder what to write about, especially if they are not from a minority group, have not experienced anything related the topic, or even not sure what it is about.
-
-A DEI statement is about showing your \emph{awareness of DEI}, your understanding of the challenges others face, and actions you've taken to contribute.
-Customize what you write to your own background and experiences.
-Here are some tips:
+Students, especially international ones, often wonder what to write about, especially if they are not from a minority group, have not experienced anything related the topic, or even not sure what it is about.
+
+A DEI statement is about showing your \emph{awareness} of DEI, \emph{understanding} of the challenges others face, and \emph{potential things} you can do to contribute.
 
@@
 \item \textbf{Show your awareness, customized to your own experiences}.
-For example, if you come from a different culture or country, talk about DEI here, e.g., many countries completely lack diversity in CS programs, talk about it, what does minority group face in your country?
+If you come from a different culture or country, talk about DEI there.
+Many countries, especially developing countries, do not have/care about DEI or even have laws against it, so does minority group face in your country?
 How do you think DEI can benefit?
+Remember, you would have a unique perspective on this topic that we might not know about, so use that to your advantage.
 
@@
 Just show your awareness and understanding of the topic and what you did or how you plan to contribute to it (e.g., mentoring students, helping others).
-One of the best DEI essay I have read start the acknowledgement that the applicant is not a minority, but a white male in a STEM field (i.e., that make them the majority), and that they do not have much experience in DEI, but show that they are aware of the challenges faced by minority groups and how they can contribute to it.
+One of the best DEI statements I have read starts the acknowledgement that the applicant is not a minority, but a white male in a STEM field (i.e., that make them the majority), and that they do not have much experience in DEI, but show that they are aware of the challenges faced by minority groups and how they can contribute to it.
 \end{commentbox}

### 2024-11-19 demystify.tex
updates

@@
 
-In recent years many universities start requiring a statement about diversity, equity, and inclusion (DEI) as part of the application (also see \autoref{sec:urm}).
-Students, especially international ones, often wonder what to write about, especially if they are not from a minority group, have not experienced anything related the topic, or even not sure what it is about.
+Many universities now require a diversity, equity, and inclusion (DEI) statement as part of the application (also see \autoref{sec:urm}).
+Students, especially international ones, often struggle with what to write, especially when they are not from a minority group or do not have much experience in DEI.
 

### 2024-11-30 demystify.tex
updates

@@
 
-\section{Is an MS required for admission to PhD in CS in the US?}\label{sec:msrequirement}
+\section{Is MS required for admission to PhD in CS?}\label{sec:msrequirement}
 \sectioninfo{You do not need an MS to do PhD in CS.}
@@
 If you don't get in, then you can ask for feedback and improve your profile for the next round.
+
+\section{``Chance me'' on getting into MIT or Stanford or Ivy League schools}\label{sec:chance-me}
+
+Many students, especially international students, aim for very top CS PhD programs such as Stanford and MIT, and Ivy Leagues.
+Every year Reddit and other forums have numerious posts asking for evaluation of their chances of getting into these schools (the so called \emph{``chance me''} or \emph{``roast my CV/profile''} posts), and then later posts on being \emph{``rejected everywhere''}.
+
+\paragraph{You are unlikely to get in these schools}  While being ambitious is always good, it is important to be realistic, and the harsh reality is that you are very likely NOT getting in any of these.
+Why?
+people who do get in these schools often were explicitly encouraged by their mentors and LoR writers, who themselves might be MIT-alum or are academic celebrities who've sent many of their students there.
+If you are not in this category (doubt you are, because you're asking random people on the internet instead of your trusted LoR writers), then you're likely not going to get in, and MIT and Stanford would glady take your application money.
+
+\paragraph{Top in US News might not really top in CS} Morever, your ``top'' schools might not be top in CS.
+The Ivies?
+yes good for many things but usually not coding marathon and typically MIA at top CS confs (ok, they are still good, but a school somewhere in Maryland, Wisconsin, or Urbana Champain cornfield would probably beat them by far).
+MIT and Stanford?
+yes, they are top, but many are equivalent, and there could be CS areas where they are not as strong as other schools.
+Caltech and John Hopkins?
+...
+are you confused CS with Physics and Medicine?
+
+You're missing out some serious places: UCSD, UW, UMD, Michigan, Purdue, Stony brook.
+These all have super strong CS programs but rarely get mentioned.
+Utah, Rutgers, Buffalo, Penn State, UVA, Ohio, UC Riverside, George Mason ...
+any of them rings a bell?
+These might not ranked very high on QS or US News and are not household names to you, your friends, mom and dad, but they are excellent CS schools where graduating students from your dream schools often apply for faculty positions (e.g., see \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-GMU}{GMU stats}).
+You can totally go to these places and be superstar there.
+
+\paragraph{You don't need to be top}  It is understandable top schools will open doors to more opportunities (you are likely from a very good schools in your country to aim for a PhD in CS in the US).
+In countries like Vietnam, there are usually only a few top schools, and if you don't get into them, you're doomed (or so you think).
+But that is not the case in the US.
+There are simply too many good schools, and it is not a binary choice between top X or nothing.
+In fact, ``equity'' is emphasized in the US and many ``smaller'' schools actually are given more resources and attention to help students succeed.
+
+In short, don't aim for just the moon and the stars, also try the sky, clouds, or even just the trees.  You might be happier there and it might catapult you to the sun and beyond.
 

### 2024-11-30 demystify.tex
updates

--- 2024-11-19 demystify.tex
@@
 
+\newtcolorbox{myemphasisbox}{
+  enhanced,
+  boxrule=1pt, frame hidden,
+  %borderline = {1pt}{0pt}{blue!75!black},
+  colback=gray!20!white,
+  sharp corners
+}
+
+\newenvironment{emphasisbox}[1][]{
+  \small
+  \begin{myemphasisbox}
+    {\small \textbf{#1}}
+  }{
+  \end{myemphasisbox}
+}
+
 \renewcommand{\figurename}{Fig.}
@@
 If you don't get in, then you can ask for feedback and improve your profile for the next round.
+
+\section{``Chance me'' on getting admitted to Top 10}\label{sec:chance-me}
+
+Many students, especially international students, aim for very top CS PhD programs such as Stanford and MIT, and Ivy Leagues.
+Every year Reddit and other forums have numerous students asking for evaluation of their chances of getting into these schools (the so called \emph{``chance me''} or \emph{``roast my CV/profile''} posts) and then later posts on being \emph{``rejected everywhere''}.
+
+\paragraph{You are unlikely to get in these schools}  While being ambitious is always good, it is important to be realistic, and the harsh reality is that it is very unlikely that you will get in MIT, or Harvard.
+Why?
+people who do get in these schools often were explicitly encouraged by their mentors and LoR writers, who themselves might be alumni of these schools or are academic celebrities who've sent many of their students there.
+If you are not in this category (doubt you are, because you're asking random people on the internet instead of your trusted LoR writers), then you're likely not going to get in, and MIT and Stanford would glady take your application money.
+
+\paragraph{``Your'' top might not really top in CS} Harvard and the Ivies?
+yes good for many things but usually not coding marathon and typically MIA at top CS confs (ok, they are still good, but a school somewhere in Maryland, Wisconsin, or Urbana Champaign cornfield would likely beat them by far in CS).
+MIT and Stanford?
+yes, they are top, but many are equivalent, and there could be CS areas where they are not as strong as other schools.
+Caltech and John Hopkins?
+...
+are you confused CS with Physics and Medicine?
+
+You're missing out some serious places: UCSD, UW, UMD, Michigan, Purdue, Stony brook.
+These all have super strong CS programs but rarely get mentioned.
+Utah, Rutgers, Buffalo, Penn State, UVA, Ohio, UC Riverside, George Mason ...
+any of them rings a bell?
+These might not ranked very high on QS or US News and are not household names to you, your friends, mom and dad, but they are excellent CS schools where graduating students from your dream schools often apply for faculty positions (e.g., see \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-GMU}{GMU stats}).
+You can go to these places and be superstar there.
+
+\paragraph{You don't need to be top}  Top schools will open doors to more opportunities.
+In countries like Vietnam, there are usually only a few top schools, and if you don't get into them, you are doomed  (or so you think).
+But that is not the case in the US.
+There are simply too many good schools, and it is not a binary choice between top X or nothing.
+In fact, ``equity'' is emphasized in the US and ``smaller'' schools are often given more resources and attention to help students succeed.
+
+\begin{emphasisbox}
+In short, don't aim for just the moon and the stars, also try the sky, clouds, or even just the trees.
+You might find much happiness and success there, which then can catapult you places beyond the moon and the stars.
+This is the American dream, which is strongly embodied in the its higher education system.
+\end{emphasisbox}
 

### 2024-11-30 demystify.tex
updates

@@
 
+\section{Research Experience}\label{sec:research-experience}
+
+\sectioninfo{\emph{Publications can greatly help}. Papers in good venues are concrete evidence that you have successfully engaged in research. But they are not required!}
+
+This section discusses publications and other research experiences that can strengthen your application.
+\autoref{sec:research-opportunities} provides more information on how to find research opportunities (e.g., during your undergrad study).
+
+\subsection{Publications} The most concrete evidence of research ability is having \emph{papers in reputable international journals or conferences}.
+Having published papers, especially at top venues, is a sign that you have been successfully involved in research.
+
+Publications are never required for application, however given the competitiveness of CS admission, they can significantly strengthen your application and are becoming the norm.
+Applicants admitted to top schools, especially in popular fields such as ML and NLP, often have multiple first-authored papers at top places.
+
+\begin{commentbox}[Vu:]
+  Many international students mention Scopus Q1, which consists of various journals from IEEE, Elsevier, and many other publishers.
+I don't know/recognize many of the journals listed in Scopus Q1.
+This might be something to be mindful of, as \textbf{CS} faculty might not be too familiar with Scopus or journals listed there, so devote some part in your statement to discuss the significance of your papers.
+\end{commentbox}
+
+\begin{commentbox}[Thanh:]
+    Due to academic culture, professors in Vietnam usually aim for (international) journals instead of conferences.
+Could you give some tips on how to know whether a journal is good (CSrankings, unfortunately, only consider conferences)?
+    \tcblower
+    \textbf{Vu}: One way is looking at what well-known researchers publish.
+For example, if you are interested in a field X, you can use CSRankings to look at active faculty in X, and then look at their websites to see what journals they publish at.
+\end{commentbox}
+\paragraph{What if I don't have any pubs} It is understandable that many students do not have the opportunity to publish papers.
+Thus, other writings, even those under submissions or rejected, would still be good and better than nothing.
+Be sure to upload your papers with your application and talk about them in your statement (see \autoref{sec:sop}).
+
+Note that local conferences and non-English journals or conferences do
+not carry as much weight since their quality is often unknown to US faculty.
+However, if you have published in such places, you should still upload them, mention them in your statement, and explain why they are good.
+
+% In CS, conferences are important and top conferences are prestigious.
+You can find top CS conferences at places such as \href{https://csrankings.org}{CSRankings}, which ranks CS programs based on how their faculty publish at top \emph{conferences} (see more in \autoref{sec:ranking}).
+
+\paragraph{I am not the first author} Being the first author is good because it indicates you own the work.
+However, it's perfectly OK to be second or third or last.
+It is difficult to publish a good paper and so being a co-author is still a good sign about your research experience.
+In any case (and especially in the case when you're not the first author), you should explain the work and your contribution.
+Better yet, have your LoR writers talk about your contribution in their letters.
+
+\begin{commentbox}[Craig:]
+  Mason and many other universities allow you to upload your published papers and other writing samples.
+In many cases, even if the papers were not published at top places, we can still determine their quality by simply skimming over the paper.
+\end{commentbox}
+
+\subsection{Work Experience} Work experience at \emph{well-known research laboratories}, such as Microsoft Research, can strengthen your
+application.
+Unfortunately, many good research places in your countries, e.g., VinAI in Vietnam, remain relatively unknown to most universities in the US.
+So you or your LoR writers should explicitly say something about them in your statement.
+In general, if you did some good research work at such places, then you should mention that in your SOP and ask your supervisor to write about your research experiences and potential in their LORs (\autoref{sec:lor}).
+
+\begin{commentbox}[Hung:]
+  The reputation of VinAI has been increasing steadily over the past few years; many of my colleagues heard about VinAI.
+\end{commentbox}
+
+On the other hand, if non-research work (e.g., software development) \emph{do not} carry much weight for PhD (they can help for MS application though).
+Similarly, LoR from your supervisor for non-research experience will not count much.
+So do not spend much time talking about non-research-related job in your SOP.
+
+\subsection{Competitions} Winning \emph{internationally recognized competitions} can demonstrate your research potential.
+For example, participating in Math Olympiads if you want to do theory or winning ACM programming contests if you want to ``build'' systems, e.g., software analysis.
+
 \section{Statement of Purpose (SOP)}\label{sec:sop}
@@
 
-\section{Research Experience}\label{sec:research-experience}
-
-\sectioninfo{\emph{Publications can greatly help}. Papers in good venues are concrete evidence that you have successfully engaged in research. But they are not required!}
-
-This section discusses publications and other research experiences that can strengthen your application.
-\autoref{sec:research-opportunities} provides more information on how to find research opportunities (e.g., during your undergrad study).
-
-\subsection{Publications} The most concrete evidence of research ability is having \emph{papers in reputable international journals or conferences}.
-Having published papers, especially at top venues, is a sign that you have been successfully involved in research.
-
-Publications are never required for application, however given the competitiveness of CS admission, they can significantly strengthen your application and are becoming the norm.
-Applicants admitted to top schools, especially in popular fields such as ML and NLP, often have multiple first-authored papers at top places.
-
-\begin{commentbox}[Vu:]
-  Many international students mention Scopus Q1, which consists of various journals from IEEE, Elsevier, and many other publishers.
-I don't know/recognize many of the journals listed in Scopus Q1.
-This might be something to be mindful of, as \textbf{CS} faculty might not be too familiar with Scopus or journals listed there, so devote some part in your statement to discuss the significance of your papers.
-\end{commentbox}
-
-\begin{commentbox}[Thanh:]
-    Due to academic culture, professors in Vietnam usually aim for (international) journals instead of conferences.
-Could you give some tips on how to know whether a journal is good (CSrankings, unfortunately, only consider conferences)?
-    \tcblower
-    \textbf{Vu}: One way is looking at what well-known researchers publish.
-For example, if you are interested in a field X, you can use CSRankings to look at active faculty in X, and then look at their websites to see what journals they publish at.
-\end{commentbox}
-\paragraph{What if I don't have any pubs} It is understandable that many students do not have the opportunity to publish papers.
-Thus, other writings, even those under submissions or rejected, would still be good and better than nothing.
-Be sure to upload your papers with your application and talk about them in your statement (see \autoref{sec:sop}).
-
-Note that local conferences and non-English journals or conferences do
-not carry as much weight since their quality is often unknown to US faculty.
-However, if you have published in such places, you should still upload them, mention them in your statement, and explain why they are good.
-
-% In CS, conferences are important and top conferences are prestigious.
-You can find top CS conferences at places such as \href{https://csrankings.org}{CSRankings}, which ranks CS programs based on how their faculty publish at top \emph{conferences} (see more in \autoref{sec:ranking}).
-
-\paragraph{I am not the first author} Being the first author is good because it indicates you own the work.
-However, it's perfectly OK to be second or third or last.
-It is difficult to publish a good paper and so being a co-author is still a good sign about your research experience.
-In any case (and especially in the case when you're not the first author), you should explain the work and your contribution.
-Better yet, have your LoR writers talk about your contribution in their letters.
-
-\begin{commentbox}[Craig:]
-  Mason and many other universities allow you to upload your published papers and other writing samples.
-In many cases, even if the papers were not published at top places, we can still determine their quality by simply skimming over the paper.
-\end{commentbox}
-
-\subsection{Work Experience} Work experience at \emph{well-known research laboratories}, such as Microsoft Research, can strengthen your
-application.
-Unfortunately, many good research places in your countries, e.g., VinAI in Vietnam, remain relatively unknown to most universities in the US.
-So you or your LoR writers should explicitly say something about them in your statement.
-In general, if you did some good research work at such places, then you should mention that in your SOP and ask your supervisor to write about your research experiences and potential in their LORs (\autoref{sec:lor}).
-
-\begin{commentbox}[Hung:]
-  The reputation of VinAI has been increasing steadily over the past few years; many of my colleagues heard about VinAI.
-\end{commentbox}
-
-On the other hand, if non-research work (e.g., software development) \emph{do not} carry much weight for PhD (they can help for MS application though).
-Similarly, LoR from your supervisor for non-research experience will not count much.
-So do not spend much time talking about non-research-related job in your SOP.
-
-\subsection{Competitions} Winning \emph{internationally recognized competitions} can demonstrate your research potential.
-For example, participating in Math Olympiads if you want to do theory or winning ACM programming contests if you want to ``build'' systems, e.g., software analysis.
-
 \section{Your School and Grades}\label{sec:your-school}
@@
 
-Grades generally do not matter much for CS PhD admission.
+Compared to other factors such as LoRs (\autoref{sec:lor}) and research experiences (\autoref{sec:research-experience}), grades generally do not matter much for CS PhD admission.
 PhD in CS is a research degree and doing well in undergrad courses does not necessarily mean you can do research (other factors such as research experiences, pubs, LoRs are more important).
-
-Nonetheless, if you are from a well-known school, having good grades do help (not a lot though), e.g., adcom members often note details such as "good GPA from well-known school (\autoref{sec:ievaluate}).
+In fact, many CS faculty members themselves have bad grades in undergrad courses (and some were proud of that!).
+
+Nonetheless, if you are from a well-known school (\autoref{sec:your-school}), having good grades do help (not a lot though), e.g., adcom members often note details such as "good GPA from well-known school (\autoref{sec:ievaluate}).
 However if your school is not well-known, having top grades or rankings usually will not help because we cannot evaluate them (e.g., we don't know how hard it is to get a 4.0 or A's at your school).

### 2024-12-01 demystify.tex
updates

@@
 Don't talk about religious or political beliefs (\autoref{sec:cultural}).
-\item \textbf{Criticizing} your current or previous institution or professors: just like in a job interview, don't badmouth your current or previous employer. 
+\item \textbf{Criticizing} your current or previous institution or professors. Just like in a job interview, don't badmouth your current or previous employer. 
 \item \textbf{No concrete evidence to back up claims}: e.g., saying you are passionate about a research topic without showing any experience.
@@
 X), numbers (outperformed SOTA by Y \%), and examples (worked on project Z) can help.
-\item \textbf{Use flowery and AI-like language}: don't use AI to write your SOP. Not that hard to raise suspicion.
+\item \textbf{Use flowery and AI-like language}. Don't use AI to write your SOP. Not that hard to raise suspicion. Though you can ask AI to check your grammar and spelling.
 \item \textbf{Not customized to the program}: If your SOP can be sent to multiple programs with few changes, it is too generic.
-Do some homework and mention why you want to spend the next 5--7 years there.
+Do some research and mention why you want to spend the next 5--7 years there.
 \item \textbf{Mentioning wrong professors}: do not mention emeritus professors or those who have left.
-Do your homework to see if the professors you mention are still active in research.
 Teaching and adjunct faculty are often not active in mentoring PhD students (\autoref{sec:faculty-types}).
+Do your homework and mention profs who are till active in research.
 \item \textbf{Too Long and fancy format}: As mentioned, keep it under 2 pages.
 Don't use too much coloring or fancy fonts (like those in Words).
-CS academics like using LaTeX (common way to write our papers and other documents), so write your SOP using LaTeX.
+CS academics like using LaTeX (common way to write our papers and other documents), so write your SOP using LaTeX (with Times or default font, 11pt, and 1-inch margin).
 \end{itemize}

### 2024-12-01 demystify.tex
updates

--- 2024-11-30 demystify.tex
@@
 
-\newtcolorbox{myemphasisbox}{
-  enhanced,
-  boxrule=1pt, frame hidden,
-  %borderline = {1pt}{0pt}{blue!75!black},
-  colback=gray!20!white,
-  sharp corners
-}
-
-\newenvironment{emphasisbox}[1][]{
-  \small
-  \begin{myemphasisbox}
-    {\small \textbf{#1}}
-  }{
-  \end{myemphasisbox}
-}
-
 \renewcommand{\figurename}{Fig.}
@@
 
-\section{Is MS required for admission to PhD in CS?}\label{sec:msrequirement}
+\section{Is an MS required for admission to PhD in CS in the US?}\label{sec:msrequirement}
 \sectioninfo{You do not need an MS to do PhD in CS.}
@@
 \paragraph{Red flags} Various types of red flags can cause concern to the adcom.
-Common examples include too many STEM courses with low or withdrawn grades, plagiarism, cheating, or other academic dishonesty.
+Common examples include many STEM courses with low or withdrawn grades, plagiarism, cheating, or other academic dishonesty.
 Another one is that you have a history of jumping from one program to another.
-Adcom members might have contacts in other programs and heard about your stories, or your LoRs might mention these.
+Adcom members might have contacts in other programs and heard about your stories, or your LoRs might mention them.
 
@@
 If you don't get in, then you can ask for feedback and improve your profile for the next round.
-
-\section{``Chance me'' on getting admitted to Top 10}\label{sec:chance-me}
-
-Many students, especially international students, aim for very top CS PhD programs such as Stanford and MIT, and Ivy Leagues.
-Every year Reddit and other forums have numerous students asking for evaluation of their chances of getting into these schools (the so called \emph{``chance me''} or \emph{``roast my CV/profile''} posts) and then later posts on being \emph{``rejected everywhere''}.
-
-\paragraph{You are unlikely to get in these schools}  While being ambitious is always good, it is important to be realistic, and the harsh reality is that it is very unlikely that you will get in MIT, or Harvard.
-Why?
-people who do get in these schools often were explicitly encouraged by their mentors and LoR writers, who themselves might be alumni of these schools or are academic celebrities who've sent many of their students there.
-If you are not in this category (doubt you are, because you're asking random people on the internet instead of your trusted LoR writers), then you're likely not going to get in, and MIT and Stanford would glady take your application money.
-
-\paragraph{``Your'' top might not really top in CS} Harvard and the Ivies?
-yes good for many things but usually not coding marathon and typically MIA at top CS confs (ok, they are still good, but a school somewhere in Maryland, Wisconsin, or Urbana Champaign cornfield would likely beat them by far in CS).
-MIT and Stanford?
-yes, they are top, but many are equivalent, and there could be CS areas where they are not as strong as other schools.
-Caltech and John Hopkins?
-...
-are you confused CS with Physics and Medicine?
-
-You're missing out some serious places: UCSD, UW, UMD, Michigan, Purdue, Stony brook.
-These all have super strong CS programs but rarely get mentioned.
-Utah, Rutgers, Buffalo, Penn State, UVA, Ohio, UC Riverside, George Mason ...
-any of them rings a bell?
-These might not ranked very high on QS or US News and are not household names to you, your friends, mom and dad, but they are excellent CS schools where graduating students from your dream schools often apply for faculty positions (e.g., see \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-GMU}{GMU stats}).
-You can go to these places and be superstar there.
-
-\paragraph{You don't need to be top}  Top schools will open doors to more opportunities.
-In countries like Vietnam, there are usually only a few top schools, and if you don't get into them, you are doomed  (or so you think).
-But that is not the case in the US.
-There are simply too many good schools, and it is not a binary choice between top X or nothing.
-In fact, ``equity'' is emphasized in the US and ``smaller'' schools are often given more resources and attention to help students succeed.
-
-\begin{emphasisbox}
-In short, don't aim for just the moon and the stars, also try the sky, clouds, or even just the trees.
-You might find much happiness and success there, which then can catapult you places beyond the moon and the stars.
-This is the American dream, which is strongly embodied in the its higher education system.
-\end{emphasisbox}
 

### 2024-12-01 demystify.tex
updates

@@
 
-Typically, it takes 5--7 years for CS PhD in the US.  This can be longer than CS PhDs in other countries (\autoref{sec:non-us-differences}), which might require MS (\autoref{sec:phd-vs-ms}) first.
+Typically, it takes 5--7 years for CS PhD in the US.  This can be longer compared to other countries (\autoref{sec:non-us-differences}), which might require MS (\autoref{sec:phd-vs-ms}) first.
 
@@
 \caption{Comparison of US and Other PhD Programs}\label{tab:us-vs-other}
+\center
+\small
 \begin{tabular}{c|c|c}
@@
 
-\paragraph{LoR from Famous People} Having a strong letter from an internationally recognized researcher will \emph{greatly strengthen} your application.
+\subsection{LoR from Famous People}
+
+Having a strong letter from an internationally recognized researcher will \emph{greatly strengthen} your application.
 A strong LOR from such a researcher can outweigh the lack of publications (or low GPA or unknown school).
@@
 It is also worth noting that a LoR from a ``full'' senior professor who has not been active in research might not be as strong as a LoR from a junior faculty who has been actively mentoring you in research.
-This applies to administrative (e.g., chairs, directors, and deans) also (\autoref{sec:admin-letters}.) The fame of the writer is less important than how well they know you and can write about your research ability.
+This also applies to administrative, e.g., chairs, directors, and deans (\autoref{sec:admin-letters}.) The fame of the writer is less important than how well they know you and can write about your research ability.
 
@@
 So if you are from there, you should ask your reference writers to mention that.
+\end{commentbox}
+
+\begin{commentbox}[Didier]
+    \emph{Should letter writers have PhDs?}  In Rwanda, a lot of students interact more with teaching faculty who might not have PhD.
+    \tcblower
+    \textbf{Vu}: This is an interesting and useful detail that US faculty might not be aware of.
+Students should mention this in their statements.
+In general, someone with a PhD has been through the research process and therefore can better evaluate your research ability.
+But if you do not have such people, then someone who can properly evaluate your research ability is still OK (and again, explain that in your SOP).
   \end{commentbox}
@@
 
-\subsection{Waiving Your Right}\label{sec:waive-right}
-
-When you ask someone to write a letter for you, you will be asked if you want to waive your right to see the letter.  You should \emph{always} waive your right.
+\subsection{What can you do?}\label{sec:lor-what-to-do}
+
+While you cannot control what your letter writers write, there are multiple things you can do to help them write a strong letter for you.
+
+\paragraph{Waiving Your Right} When you ask someone to write a letter for you, you will be asked if you want to waive your right to see the letter.  You should \emph{always} waive your right.
 
@@
 
-\begin{commentbox}[Didier]
-  \emph{Should letter writers have PhDs?}  In Rwanda, a lot of students interact more with teaching faculty who might not have PhD.
-  \tcblower
-  \textbf{Vu}: This is an interesting and useful detail that US faculty might not be aware of.
-Students should mention this in their statements.
-In general, someone with a PhD has been through the research process and therefore can better evaluate your research ability.
-But if you do not have such people, then someone who can properly evaluate your research ability is still OK (and again, explain that in your SOP).
-\end{commentbox}
-
-\subsection{Help your LOR Writer}\label{sec:help-your-LOR-writers}
-
-As mentioned above in \autoref{sec:self-letters} and \autoref{sec:generic-letters}, do not write your own letter and generic letters do not give much value.
+\paragraph{Help your LOR Writer} As mentioned above in \autoref{sec:self-letters} and \autoref{sec:generic-letters}, do not write your own letter and generic letters do not give much value.
 Thus, to help your writer to write a strong, customized LoR, you should provide them details or unique things about yourself.
@@
 
-\subsection{Remind Your Writers}\label{sec:remind-writers}
-
-After you submit your application, you should tell your writers about that and let them know they will soon receive an email from the university to submit their letters.
+\paragraph{Remind Your Writers} After you submit your application, you should tell your writers about that and let them know they will soon receive an email from the university to submit their letters.
 You should also remind them when you submit your application and ask them to submit their letters on time if they haven't done so.
-Note that most places only have deadlines for the applicant, but are very flexible with the letter writers (in many cases do not even give them any deadline).
+
+Note that most places only have deadlines for the applicant, but are very flexible with the letter writers.
+In many cases your LOR writers are not even given any deadline.
 Also, many places do not begin the admission review process right after the deadline and work on application reviews in the next semester (mid-January).
@@
 
-\subsection{Thank Your Writers}\label{sec:thank-writers}
-
-After your application is submitted and your writers have submitted their letters (i.e., the wait begins), you should send a quick thank you note, which serves both as an acknowledgement that you know they have submitted the letters and as an appreciate for their help.
+\paragraph{Thank Your Writers} After your application is submitted and your writers have submitted their letters (i.e., the wait begins), you should send a quick thank you note, which serves both as an acknowledgement that you know they have submitted the letters and as an appreciate for their help.
 You should also let them know the outcome of your application, regardless of whether you are admitted or not.
@@
 If a student asks me to write letter for them, I will agree as I believe it is part of my job.
-However, I will ask them to waive their right to see the letter (\autoref{sec:waive-right}), and will not write if they do not do so.
+However, I will ask them to waive their right to see the letter (\autoref{sec:lor-what-to-do}), and will not write if they do not do so.
 
@@
 While I have a general template of such strong letters, it still can take me an hour or more to write such a letter.
-I often ask the students to provide information (\autoref{sec:help-your-LOR-writers}) and what they think I should highlight in the letter.
+I often ask the students to provide information (\autoref{sec:lor-what-to-do}) and what they think I should highlight in the letter.
 They can also provide me their SOP (if they already have written one) so I can complement what they say with my own perspective.

### 2024-12-01 demystify.tex
updates

@@
 
-\paragraph{Remind Your Writers} After you submit your application, you should tell your writers about that and let them know they will soon receive an email from the university to submit their letters.
+\paragraph{Remind Your Writers} After entering your writers' information in the application system, you should tell your writers about that and let them know they will soon receive an email from the university to submit their letters.
 You should also remind them when you submit your application and ask them to submit their letters on time if they haven't done so.
@@
 Note that most places only have deadlines for the applicant, but are very flexible with the letter writers.
-In many cases your LOR writers are not even given any deadline.
+In many cases your LOR writers \emph{are not given any deadline}.
+For example, the following list several LOR invitation emails I received from various universities in the past few years:
+
+\begin{emphasisbox}
+    \footnotesize
+    \paragraph{No deadline given} Examples with no deadline given:
+
+    \textbf{Sample 1:}
+
+    .. is applying for admission into the Computer Science (Ph.D.) program in the Graduate School at the University of Massachusetts Amherst and has listed you as a reference.
+
+    \textbf{They have waived their right of access to see your reference.}
+
+    You may submit your reference online via the Graduate Reference Center, which is located at .. Please use the email address and code shown below to log in to submit this reference.
+
+    A timely response is important for this applicant to be favorably considered. Please be aware that the applicant's admission could be contingent on your prompt response.
+
+    \textbf{Sample 2:}
+
+    Dear ..
+
+    ..
+has requested that you write a letter of recommendation for their Illinois graduate application.
+We greatly appreciate your feedback on X's ability to succeed in their graduate studies at Illinois.
+
+    In an effort to make this process as easy as possible for you, we offer the ability to upload your recommendation letter online through our secure website.
+To submit your recommendation, please use the link below.
+This link is unique to this recommendation and should not be shared or forwarded.
+
+    Thank you!
+
+    \textbf{Sample 3:}
+
+    .. is applying to the University of Nebraska-Lincoln and has named you as a recommender. \textbf{This applicant has waived the right to view your recommendation.}
+
+    Please complete our brief recommendation form (or if necessary, decline to recommend).
+
+    Should you experience any technical difficulties or require assistance with your account, please contact CollegeNET Support.
+
+    Thank you.
+
+    \paragraph{Deadline given} Several examples with a deadline or some date mentioned:
+
+    \textbf{Sample 1:}
+
+    Dear Dr ..
+
+    ..
+has listed you as a reference in an application to the EECS graduate program at MIT.
+If you are not .., please ask ..
+for the correct letter submission information.
+Please submit your letter here: ..
+
+    The deadline for the applicant is \textbf{Dec 15}; we start the admissions process immediately after that date, so we would appreciate receiving your recommendation by then.
+The applicant can see if you have submitted a letter, and may remind you.
+
+    Even though the applicant knows the link for submitting a letter, the applicant is unable to read your submitted letter.
+The https URL ensures that you connect to MIT's graduate admissions server.
+That server accepts letter submissions, but will not reveal a submitted letter.
+The link ..
+provides a bit more information on MIT's approach to collecting letters of requests.
+
+    Thank you!
+
+    \textbf{Sample 2:}
+
+    Dear Recommender,
+
+    The applicant listed below has applied to a graduate program in the School of Computer Science at Carnegie Mellon University and has requested a letter of recommendation from you.
+
+    Applicant: ..
+
+    Click Here if your mail system provides html content or use the URL below to enter your letter of recommendation in PDF form only.
+
+    Your letter of recommendation is due by \textbf{12 p.m. (Noon) EST on December 10, 20XX}. Thank you for taking the time to respond.
+\end{emphasisbox}
+
 Also, many places do not begin the admission review process right after the deadline and work on application reviews in the next semester (mid-January).

### 2024-12-01 demystify.tex
updates

@@
 
-A LoR writer should be someone who knows you well, so that they can provide a strong-supporting and personalized letter for you.  
-LoR writers should also have the credibility to convince adcom members.
+A LoR writer should be someone who \emph{knows you well}, so that they can provide a strong-supporting and personalized letter for you.
+LoR writers should also have the credibility to convince adcom members on your research ability and potential.
 Such a person is typically someone who has a PhD (so they know what a PhD is about) and is active in research (so they can evaluate your research ability).
@@
 
-\subsection{LoR from Famous People}
-
-Having a strong letter from an internationally recognized researcher will \emph{greatly strengthen} your application.
+\paragraph{LoR from Famous People}  Having a strong letter from an internationally recognized researcher will \emph{greatly strengthen} your application.
 A strong LOR from such a researcher can outweigh the lack of publications (or low GPA or unknown school).
+The reason is simple: adcom members trust people that they know and respect.
+If a well-known researcher enthusiastically recommends you, then the adcom members will likely trust that recommendation.
 
@@
 
-However, let's face it.  Most of us don't have an opportunity to interact or work with well-known researchers.  
+However, let's face it.  Most students don't have an opportunity to interact or work with well-known researchers.  
 So what to do?
@@
 
-This does not mean the professor is not good or does not care about you, but they just do not know you well enough to write a strong letter.
+\emph{This does not mean the professor is not good or does not care about you}, but they just do not know you well enough to write a strong letter.
 So it might be a good idea to directly ask if the prof.
@@
 \subsection{Self-written Letters are Bad}\label{sec:self-letters}
- Many international students write letters themselves, typically due to the requests of their profs.
+Many students write letters themselves, typically due to the requests of their profs.
 or letter writers.
@@
 writer talking about you, in this case it is you who write about yourself (and they just sign the letter).
-Experienced reviewers can easily spot  (because the student often has little experience in writing LoRs) and  \red{red flag} such letters.
+
+Adcom members are just too experienced (having reviews hundreds of those letters) to spot and suspect such letters.
+There is a very difference between a letter written by an experienced professor and a student.
+For example, they can provide convincing and concrete examples based on their experience and compare you to their own students, and of course the writing style of a professor is very different from an undergraduate student.
+Worse yet, if we suspect that the student wrote the letter, it is a \red{red flag} as we will question both the student's integrity and the letter writer's credibility.
 
@@
 In many cases your LOR writers \emph{are not given any deadline}.
-For example, the following list several LOR invitation emails I received from various universities in the past few years:
+The following lists several LOR invitation emails I received from various universities in the past few years:
 
@@
 application.
-Unfortunately, many good research places in your countries, e.g., VinAI in Vietnam, remain relatively unknown to most universities in the US.
+Note that the emphasis here is \emph{research} places, not software development or non-research work.
+For example, working at a FAANG (Facebook, Amazon, Apple, Netflix, Google) company as a software engineer is not research experience, though it can be helpful for MS applications discussed in \autoref{chap:ms}.
+Similarly, a LoR from your supervisor for non-research experience might not count much.
+So do not spend much time talking about non-research job in your SOP.
+
+Note that adcom reviewers might not be familiar with all research labs, especially those outside the US For example, while VinAI is  well-known in Vietnam and potentially in Asia, it might not be very well-known in the US.
 So you or your LoR writers should explicitly say something about them in your statement.
-In general, if you did some good research work at such places, then you should mention that in your SOP and ask your supervisor to write about your research experiences and potential in their LORs (\autoref{sec:lor}).
-
-\begin{commentbox}[Hung:]
-  The reputation of VinAI has been increasing steadily over the past few years; many of my colleagues heard about VinAI.
-\end{commentbox}
-
-On the other hand, if non-research work (e.g., software development) \emph{do not} carry much weight for PhD (they can help for MS application though).
-Similarly, LoR from your supervisor for non-research experience will not count much.
-So do not spend much time talking about non-research-related job in your SOP.
+In general, if you did some good research work, then you should mention that in your SOP and ask your supervisor to write about it in their LOR (\autoref{sec:lor}).
+
+% \begin{commentbox}[Hung:]
+%   The reputation of VinAI has been increasing steadily over the past few years; many of my colleagues heard about VinAI.
+% \end{commentbox}
 

### 2024-12-01 demystify.tex
updates

@@
 
-\subsection{School} Graduating from top universities \emph{that we recognize} helps.
+\subsection{School} Graduating from top universities \emph{that adcom members recognize helps}.
 For example, if you are an international student and your school is well-known, then it is \emph{``top foreign''}, which is a plus.
@@
 
-Many international students mistakenly believe that their school is well-known, but in fact, it is not (e.g., many Vietnamese students believe that ``Bach Khoa'' is well-known internationally, but it is not).
-So you should explicitly mention that in your SOP and ask your LOR writers to do that as well.
+Many international students mistakenly believe that their school is well-known, but in fact, it is not, e.g., many Vietnamese students believe that ``Bach Khoa'' is well-known, but it is not (and their confusing acronyms HUST and HCMUT only make it worse).
+So if you think your school is well-known, mention it (and back up with some evidence like rankings or awards) in your SOP.
+You can ask your LOR writers to talk about your school as well (\autoref{sec:lor-what-to-do}).
 Of course, if you're interested in working with Vietnamese, consider  \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{CS programs in the US that have Vietnamese professors}.

### 2024-12-11 demystify.tex
updates

@@
 \mainmatter
+
+\part{Introduction To CS PhD Admission in the US}\label{part:intro}
+
 \chapter{Should You Apply to a CS PhD Program in the US?}\label{sec:should}
@@
 
-However, MS can help admission if it gives research experience or is from a more well-known school than your undergrad institution (\autoref{sec:your-school}).
+However, MS can help admission if it gives research experience or is from a more well-known school than your undergrad institution (\autoref{chap:your-school}).
 Moreover, if you have an MS then some course work \emph{might be} transferred for course credits, which \emph{might} save some time.
@@
 
-\chapter{Your Application}\label{sec:application}
-
-\chapterinfo{The committee will look at various factors, but the most important ones are letters of recommendation (LORs),  statements of purpose (SOP), and research experience and potential, e.g., publications.}
+\part{Your Application Materials}\label{part:application}
+
+%\partinfo{.}
 
@@
 The emphasis here is \emph{fitting}, which varies from school to school, faculty to faculty, and even from year to year.
-However, in general, to evaluate your profile, we consider the following key indicators.
-
-\section{Letters of Recommendation (LOR)}\label{sec:lor}
-\sectioninfo{\emph{LORs are very important}, but only if they are personalized and talk about your research ability.}
+The committee will look at various factors, but the most important ones are letters of recommendation (LORs),  statements of purpose (SOP), and research experience and potential, e.g., publications.
+
+\chapter{Letters of Recommendation (LOR)}\label{chap:lor}
+\chapterinfo{\emph{LORs are very important}, but only if they are personalized and talk about your research ability.}
 
@@
 The letter should 
-focus on your research background, achievements, and potentials, all of which paint a compelling picture of the candidate to the adcom.
-
-\paragraph{LoR writers} LoR writers should be someone who (i) can talk in depth about your research experience and potential and (ii) have the credibility to evaluate your research ability.
+focus on your research background, achievements, and potentials, all of which paint a compelling picture of you as a researcher.
+
+\section{LoR writers}\label{sec:lor-writers}
+
+LoR writers should be someone who (i) can talk in depth about your research experience and potential and (ii) have the credibility to evaluate your research ability.
 
 A LoR writer should be someone who \emph{knows you well}, so that they can provide a strong-supporting and personalized letter for you.
-LoR writers should also have the credibility to convince adcom members on your research ability and potential.
+LoR writers should also \emph{have the credibility} to convince adcom members on your research ability and potential.
 Such a person is typically someone who has a PhD (so they know what a PhD is about) and is active in research (so they can evaluate your research ability).
@@
 
-\paragraph{LoR from Famous People}  Having a strong letter from an internationally recognized researcher will \emph{greatly strengthen} your application.
+\subsection{LoR from Famous People}  Having a strong letter from an internationally recognized researcher will \emph{greatly strengthen} your application.
 A strong LOR from such a researcher can outweigh the lack of publications (or low GPA or unknown school).
@@
 
-However, let's face it.  Most students don't have an opportunity to interact or work with well-known researchers.  
+However, most students don't have an opportunity to interact or work with well-known researchers.  
 So what to do?
@@
 
+  \subsection{LoRs from Admin People}\label{sec:admin-letters}
+
+  Many international students try to get LoRs from high-ranking administrators in their universities such as department chair/head, dean, or director.
+The students never worked with these people (they might take a class or so with these profs), but mistakenly believe that these LoRs are valuable due to the writer's high position in the university.
+  However, as emphasized, a generic LoR has little value because the writer does not know you well and can talk in depth about your research ability.
+
+  Moreover, while being well-known and respected in your local university, these writers might not be very active in research (e.g., they haven't published in recent years).
+Thus they might not be well-known and recognized by adcom members.
+
+  \begin{commentbox}[Vu:]
+    In my experience in reviewing applications, letters from admin people are often generic and do not provide much value. 
+    In many case, the letter reads like it was written by a student (\autoref{sec:self-letters}), and thus is a red flag.
+So if you are in this situation, you should find someone else to write for you as mentioned in this section.
+  \end{commentbox}
+
 \subsection{Generic Letters are Bad}\label{sec:generic-letters} 
@@
 But as mentioned, if these professors do not know you, their letters would likely be generic and carry little value (sometimes \red{red flags}).
-Moreover, a top professor at your university might not be well-known to US faculty (see more details in \autoref{sec:admin-letters} and \autoref{sec:your-school}).
+Moreover, a top professor at your university might not be well-known to US faculty (see more details in \autoref{sec:admin-letters} and \autoref{chap:your-school}).
 So save the trouble and get letters from \emph{any} professors/supervisors who know you well and can write a good letter about \emph{your} research ability.
@@
 Instead of the ref.
-writer talking about you, in this case it is you who write about yourself (and they just sign the letter).
+writer talking about you, in it is you who write about yourself (and they just sign the letter).
 
@@
 
-\subsection{LoRs from Admin People}\label{sec:admin-letters}
-
-Many international students try to get LoRs from high-ranking administrators in their universities such as department chair/head, dean, or director.
-The students never worked with these people (they might take a class or so with these profs), but mistakenly believe that these LoRs are valuable due to the writer's high position in the university.
-However, as emphasized, a generic LoR has little value because the writer does not know you well and can talk in depth about your research ability.
-
-Moreover, while being well-known and respected in your local university, these writers might not be very active in research (e.g., they haven't published in recent years).
-Thus they might not be well-known and recognized by adcom members.
-
-\begin{commentbox}[Vu:]
-  In my experience in reviewing applications, letters from admin people are often generic and do not provide much value. 
-  In many case, the letter reads like it was written by a student (\autoref{sec:self-letters}), and thus is a red flag.
-So if you are in this situation, you should find someone else to write for you as mentioned in this section.
-\end{commentbox}
-
-\subsection{What can you do?}\label{sec:lor-what-to-do}
-
-While you cannot control what your letter writers write, there are multiple things you can do to help them write a strong letter for you.
-
-\paragraph{Waiving Your Right} When you ask someone to write a letter for you, you will be asked if you want to waive your right to see the letter.  You should \emph{always} waive your right.
+\section{Asking for LORs}\label{sec:asking-lor}
+
+As mentioned in~\autoref{sec:lor-writers}, LoR writers should be someone who knows you well and has the credibility to evaluate your research ability.
+You can explicitly ask if they are willing to write a strong letter for you.
+If they are not willing (you should be thankful that they are honest with you), then you should ask someone else.
+
+% Below are some tips to approach your LoR writers:
+% \begin{itemize}
+% \item \textbf{Ask in advance}: You should ask for LoRs \emph{at least a month} before the deadline.
+People are busy with their own work and likely have many other students asking for LoRs.
+Also, a strong LoR takes time to write, so you should give them enough time to write a strong letter for you.
+
+% \item \textbf{Help your writers}: You should provide your writers with details about the programs you are applying to, the deadlines, and any other information they need to know.
+You should also provide them with details about your research experience and potential so that they can write a strong letter for you.
+% \item \textbf{Follow up}: After you ask for LoRs, you should follow up with your writers to make sure they have submitted the letters on time.
+You should also thank them for writing the letters for you.
+% \item \textbf{Waive your right}: When you ask someone to write a letter for you, you will be asked if you want to waive your right to see the letter.
+You should always waive your right.
+This shows that you trust your writers and that you are not trying to twist their words.
+% \item \textbf{Remind your writers}: After you enter your writers' information in the application system, you should tell your writers about that and let them know they will soon receive an email from the university to submit their letters.
+You should also remind them when you submit your application and ask them to submit their letters on time if they haven't done so.
+% \item \textbf{Help your writers}: To help your writers write a strong letter for you, you should provide them with details about your research experience and potential.
+You should also provide them with a draft of your SOP so that they can see what you are saying about yourself and complement that with their own perspective.
+% \item \textbf{Thank your writers}: After your writers have submitted their letters, you should thank them for writing the letters for you.
+You should also let them know the outcome of your applications.
+% \item \textbf{Ask for feedback}: After you have been admitted or rejected, you should ask your writers for feedback on your application.
+This will help you improve your application for future applications.
+% \item \textbf{Stay in touch}: After you have been admitted or rejected, you should stay in touch with your writers.
+This will help you build a relationship with them and will help you in the future.
+% \end{itemize}
+
+\subsection{Waiving Your Right} When you ask someone to write a letter for you, you will be asked if you want to waive your right to see the letter.  You should \emph{always} waive your right.
 
@@
 
-\paragraph{Help your LOR Writer} As mentioned above in \autoref{sec:self-letters} and \autoref{sec:generic-letters}, do not write your own letter and generic letters do not give much value.
+\section{What can you do?}\label{sec:lor-what-to-do}
+
+While you cannot control what your letter writers write, there are multiple things you can do to help them write a strong letter for you.
+
+\subsection{Help your LOR Writer}\label{sec:help-your-LOR-writers}
+
+ As mentioned above in \autoref{sec:self-letters} and \autoref{sec:generic-letters}, do not write your own letter and generic letters do not give much value.
 Thus, to help your writer to write a strong, customized LoR, you should provide them details or unique things about yourself.
@@
 
-\section{Research Experience}\label{sec:research-experience}
+\chapter{Research Experience}\label{chap:research-experience}
 
@@
 
-This section discusses publications and other research experiences that can strengthen your application.
-\autoref{sec:research-opportunities} provides more information on how to find research opportunities (e.g., during your undergrad study).
-
-\subsection{Publications} The most concrete evidence of research ability is having \emph{papers in reputable international journals or conferences}.
+Here we look at publications and other research experiences that can strengthen your application.
+\autoref{chap:research-opportunities} provides more information on how to find research opportunities (e.g., during your undergrad study).
+
+\section{Publications} The most concrete evidence of research ability is having \emph{papers in reputable international journals or conferences}.
 Having published papers, especially at top venues, is a sign that you have been successfully involved in research.
@@
 Thus, other writings, even those under submissions or rejected, would still be good and better than nothing.
-Be sure to upload your papers with your application and talk about them in your statement (see \autoref{sec:sop}).
+Be sure to upload your papers with your application and talk about them in your statement (see \autoref{chap:sop}).
 
@@
 
-\subsection{Work Experience} Work experience at \emph{well-known research laboratories}, such as Microsoft Research, can strengthen your
+\section{Work Experience} Work experience at \emph{well-known research laboratories}, such as Microsoft Research, can strengthen your
 application.
@@
 So you or your LoR writers should explicitly say something about them in your statement.
-In general, if you did some good research work, then you should mention that in your SOP and ask your supervisor to write about it in their LOR (\autoref{sec:lor}).
+In general, if you did some good research work, then you should mention that in your SOP and ask your supervisor to write about it in their LOR (\autoref{chap:lor}).
 
@@
 
-\subsection{Competitions} Winning \emph{internationally recognized competitions} can demonstrate your research potential.
+\section{Competitions} Winning \emph{internationally recognized competitions} can demonstrate your research potential.
 For example, participating in Math Olympiads if you want to do theory or winning ACM programming contests if you want to ``build'' systems, e.g., software analysis.
 
-\section{Statement of Purpose (SOP)}\label{sec:sop}
-\sectioninfo{\emph{SOP is important}. Write it in such a way that makes you \emph{stand out}.}
+\chapter{Statement of Purpose (SOP)}\label{chap:sop}
+\chapterinfo{\emph{SOP is important}. Write it in such a way that makes you \emph{stand out}.}
 
@@
 
-While you might not have control over LORs (\autoref{sec:lor}) or where your go to school (\autoref{sec:your-school}), you do over your
+While you might not have control over LORs (\autoref{chap:lor}) or where your go to school (\autoref{chap:your-school}), you do over your
 statement of purpose (SOP) or personal statement\footnote{Few schools separate these documents and ask you to write both: SOP, which focuses on research experiences, and personal statement, which is everything more personal, e.g., why PhD, challenges, diversity, etc}!
@@
 
-\subsection{Kiss of Death in SOP} \label{sec:kiss-of-death-sop}
+\section{Kiss of Death in SOP} \label{sec:kiss-of-death-sop}
 
@@
 
-\subsection{Diversity Statement}\label{sec:diversity-statement}
+\section{Diversity Statement}\label{sec:diversity-statement}
 
@@
 
-\section{Your School and Grades}\label{sec:your-school}
-\sectioninfo{\emph{High grades probably won't help much} (unless you're from a very top and well-known school), but bad ones likely will raise concerns.}
-
-\subsection{School} Graduating from top universities \emph{that adcom members recognize helps}.
+\chapter{Your School and Grades}\label{chap:your-school}
+\chapterinfo{\emph{High grades probably won't help much} (unless you're from a very top and well-known school), but bad ones likely will raise concerns.}
+
+\section{School} Graduating from top universities \emph{that adcom members recognize helps}.
 For example, if you are an international student and your school is well-known, then it is \emph{``top foreign''}, which is a plus.
@@
 \end{commentbox}
-\subsection{Grades}\label{sec:gpa}
-
-Compared to other factors such as LoRs (\autoref{sec:lor}) and research experiences (\autoref{sec:research-experience}), grades generally do not matter much for CS PhD admission.
+\section{Grades}\label{sec:gpa}
+
+Compared to other factors such as LoRs (\autoref{chap:lor}) and research experiences (\autoref{chap:research-experience}), grades generally do not matter much for CS PhD admission.
 PhD in CS is a research degree and doing well in undergrad courses does not necessarily mean you can do research (other factors such as research experiences, pubs, LoRs are more important).
@@
 
-Nonetheless, if you are from a well-known school (\autoref{sec:your-school}), having good grades do help (not a lot though), e.g., adcom members often note details such as "good GPA from well-known school (\autoref{sec:ievaluate}).
+Nonetheless, if you are from a well-known school (\autoref{chap:your-school}), having good grades do help (not a lot though), e.g., adcom members often note details such as "good GPA from well-known school (\autoref{sec:ievaluate}).
 However if your school is not well-known, having top grades or rankings usually will not help because we cannot evaluate them (e.g., we don't know how hard it is to get a 4.0 or A's at your school).
@@
 
-\section{Standard Tests (e.g., GRE and IELTS), Online Courses, and Certificates}\label{sec:standard-tests}
-\sectioninfo{\emph{Standard tests are not important}.
+\chapter{Standard Tests (GRE and IELTS)}\label{chap:standard-tests}
+\chapterinfo{\emph{Standard tests are not important}.
 GRE typically \emph{is not} required.
@@
 
-\subsection{GREs Are Optional and Do Not Matter} While a few schools still require taking the GRE (e.g., UCF), most good CS PhD programs in the US \textbf{no longer} require it.
+\section{GREs Are Optional and Do Not Matter} While a few schools still require taking the GRE (e.g., UCF), most good CS PhD programs in the US \textbf{no longer} require it.
 Thus, if you haven't taken the GRE or have bad scores, then don't waste time (re)taking it.
@@
 
-\subsection{English Tests (IELTS, TOEFL)} Unless your degrees are from specific countries such as \href{https://www.gmu.edu/international/english-language-requirements}{these}, you will need to take standardized English tests.
+\section{English Tests (IELTS, TOEFL)} Unless your degrees are from specific countries such as \href{https://www.gmu.edu/international/english-language-requirements}{these}, you will need to take standardized English tests.
 On one hand you will need to show some level of English proficiency, but on the other hand, you do not need to have very good scores in these tests (many adcom members themselves were once international students and struggled with English).
 You should just do well enough to pass the minimum requirement set by the university. %, which is required for TA (\autoref{sec:funding}).
-Just as with grades (\autoref{sec:your-school}) and GRE, having high scores in these tests might not help, but having too low scores can be a \red{red flag} and sometimes results in an automatic rejection (\autoref{sec:why-rejected}), e.g., below the minimum requirement.
+Just as with grades (\autoref{chap:your-school}) and GRE, having high scores in these tests might not help, but having too low scores can be a \red{red flag} and sometimes results in an automatic rejection (\autoref{sec:why-rejected}), e.g., below the minimum requirement.
 
@@
 
-\subsection{Online Courses and Certificates}
+\chapter{Miscellaneous}\label{chap:misc}
+\section{CV/Resume}
+\sectioninfo{Highlight and summarize major achievements, e.g., Publications, Programming Awards}
+
+The CV/Resume provides a summary of the applicant's achievements. Most schools require you to upload your CV with your application.
+Prepare your CV in such a way that it allows the reviewers to quickly scan to identify major achievements, e.g., Publications, Programming Competition Awards, and Teaching Experience.
+
+\begin{commentbox}[Vu:]
+    Unlike a job application, CV is not as important for PhD admission because we do not screen applicants based on their CV.
+Moreover, we care and pay attention more about your LOR, SOP, etc.
+So do not spend too much time on your CV, just make sure it is easy to skim through and well-organized around research activities and achievements.
+\end{commentbox}
+
+\section{Writing Sample}\label{sec:writing-sample}
+TBD TODO
+
+\section{Online Courses and Certificates}
 
@@
 
-\section{CV/Resume}
-\sectioninfo{Highlight and summarize major achievements, e.g., Publications, Programming Awards}
-
-The CV/Resume provides a summary of the applicant's achievements. Most schools require you to upload your CV with your application.
-Prepare your CV in such a way that it allows the reviewers to quickly scan to identify major achievements, e.g., Publications, Programming Competition Awards, and Teaching Experience.
-
-\begin{commentbox}[Vu:]
-    Unlike a job application, CV is not as important for PhD admission.
-We care more about your LOR, SOP, etc.
-So do not spend too much time on your CV, just make sure it is easy to skim through and well-organized around research activities and achievements.
-\end{commentbox}
-
-\section{Interview and the Waiting Game}\label{sec:interview}
-\sectioninfo{Getting an interview is typically a \emph{good sign}; but no interview does not mean rejection}
+\chapter{Interview and the Waiting Game}\label{chap:interview}
+\chapterinfo{Getting an interview is typically a \emph{good sign}; but no interview does not mean rejection}
 
@@
 
-\subsection{Interviews}
+\section{Interviews}
 
@@
 
-\subsection{Not Getting Interviews}\label{sec:no-interview}
+\section{Not Getting Interviews}\label{sec:no-interview}
 While it is generally good to get an interview, not getting one \emph{does not} mean you're out.
@@
 
-\subsection{Notification Timeline: Why rejection letters are sent so late?}\label{sec:late-rejection}
-\subsectioninfo{Grad programs often wait for the accepted students to make their decisions, typically by April 15, before sending out rejection letters.}
+\section{Notification Timeline: Why rejection letters are sent so late?}\label{sec:late-rejection}
+\sectioninfo{Grad programs often wait for the accepted students to make their decisions, typically by April 15, before sending out rejection letters.}
 
@@
 Not much you can do other than to be patient and wait.
-Do not send emails asking about interviews or status (unless you have interviewed specifically with someone (\autoref{sec:interview}) then you can ask that person for status updates and other questions (\autoref{sec:accept-postpone-decline}).
+Do not send emails asking about interviews or status (unless you have interviewed specifically with someone then you can ask that person for status updates and other questions (\autoref{sec:accept-postpone-decline}).
 
@@
 
-% You can then share this sheet with your reference writers and remind them to submit LoRs on your behalf (see \autoref{sec:lor}).
+% You can then share this sheet with your reference writers and remind them to submit LoRs on your behalf (see \autoref{chap:lor}).
 Also, you should periodically update your writers with your status; a simple note such as \emph{``I am getting interviewed by prof.
@@
 
+\part{Funding}\label{part:funding}
 \chapter{Funding}\label{sec:funding}
@@
 Faculty and people knowledgeable about the process will not answer these questions, and those who do are often not knowledgeable and might give you false hope or despair.
-OK, someone say your GRE is low (which does not even matter for most schools (\autoref{sec:standard-tests}) , are you going to retake it?
+OK, someone say your GRE is low (which does not even matter for most schools (\autoref{chap:standard-tests}) , are you going to retake it?
 Or someone says your GPA is low, are you going to apply to a different school?
@@
 I usually read in this order.
-I start with the \emph{summary}, checking for low GPAs or test scores below the university minimum (\autoref{sec:standard-tests}).
+I start with the \emph{summary}, checking for low GPAs or test scores below the university minimum (\autoref{chap:standard-tests}).
 I then skim the \emph{transcripts} for low grades in relevant courses, noting issues like ``many low grades in main courses or unknown international school with good GPA''.
@@
 
-  \item[LoR] Letter of Recommendation. A letter written by a professor or supervisor that assesses your qualifications and potential for graduate study (\autoref{sec:lor}).
+  \item[LoR] Letter of Recommendation. A letter written by a professor or supervisor that assesses your qualifications and potential for graduate study (\autoref{chap:lor}).
 
@@
 
-  \item[SOP] Statement of Purpose is a document written by yourself to explains your research interests, background, and reasons for applying to a PhD program (\autoref{sec:sop}).
+  \item[SOP] Statement of Purpose is a document written by yourself to explains your research interests, background, and reasons for applying to a PhD program (\autoref{chap:sop}).
 
@@
 A standardized test that is an admissions requirement for many graduate schools in the US.
-However, it is not required for most CS PhD programs (\autoref{sec:standard-tests}).
+However, it is not required for most CS PhD programs (\autoref{chap:standard-tests}).
 
@@
 
-\paragraph{School \autoref{sec:your-school}} Adcom already knows about your school, which is a plus (see \autoref{sec:your-school}).
+\paragraph{School \autoref{chap:your-school}} Adcom already knows about your school, which is a plus (see \autoref{chap:your-school}).
 You are also more familiar with the US education system and culture (so hopefully you are aware of the cultural issues listed in \autoref{sec:cultural}).
 
-\paragraph{Standard Tests \autoref{sec:standard-tests}} You do not need to take TOEFL or IELTS because you already did your undergrad (or MS) at a US university.
+\paragraph{Standard Tests \autoref{chap:standard-tests}} You do not need to take TOEFL or IELTS because you already did your undergrad (or MS) at a US university.
 You might also be more comfortable communicating in English, e.g., contacting professors (\autoref{sec:contact}).
@@
 
-\paragraph{Research Experience \autoref{sec:research-opportunities}} You might have many opportunities to do research as an undergraduate, e.g., through REUs and internships at your undergrad university.
+\paragraph{Research Experience \autoref{chap:research-opportunities}} You might have many opportunities to do research as an undergraduate, e.g., through REUs and internships at your undergrad university.
 Highlight such experience in your application.
@@
 \subsection{Funding}\label{sec:ms-funding}
-Unlike PhD programs, which often have funding (\autoref{sec:funding}), MS students are typically not funded.
-The reason is because MS students do not focus on research and thus are not funded through RA and PhD students have priority for TA positions.
+Unlike PhD programs, which often have funding (\autoref{sec:funding}), MS students are typically \emph{not} funded.
+This is because MS students do not focus on research and thus are not funded through RA and PhD students have priority for TA positions.
+
 \begin{itemize}
@@
 are not willing to take MS students as RAs because they are not around long enough to be productive.
-It can take a while for a student to get up to speed and start being productive, and by that time, they are about to graduate.
+It can take a while for a student to get used to research and start being productive, and by that time (1--2 years), they are about to graduate.
 Moreover, the goal of most MS students is to get a job, not to do research, so they are not as motivated to do research.
 
-\item[\textbf{TA}] MS students are typically not given TA positions because PhDs are given priority.
+\item[\textbf{TA}] MS students are typically not given TA positions because PhDs are given priority as they will stay much longer (\autoref{sec:time}).
 Many depts.
@@
 However, there are always exceptions, e.g., if you have a strong background and can demonstrate that you can be productive in research, then a prof. might be willing to take you as their RA.
-Moreover, some schools have TAs for MS students, and you can apply for these positions. For example, Mason CS has quite a few TAs for MS students; my courses in the past two years have MS TAs.
+Moreover, some schools have TAs for MS students, and you can apply for these positions. For example, Mason CS has quite a few TAs for MS students, and my courses in the past two years have MS TAs.
 
@@
 
-\item \textbf{Living Cost:} MS students are typically not funded, so living cost can be a big factor.
+\item \textbf{Living Cost:} MS students are typically not funded (\autoref{sec:ms-funding}), so living cost can be a big factor.
 You should consider the tuition and living expenses of the area.
@@
 
-\chapter{Research Opportunities}\label{sec:research-opportunities}
+\chapter{Research Opportunities}\label{chap:research-opportunities}
 \chapterinfo{How to get research experience as an undergrad.}
@@
 
-As discussed in \autoref{sec:research-experience}, having a successful research experience can greatly strengthen a PhD application.
-Research experience gives you opportunities to try out research, determine what research area you're interested in, publish papers (\autoref{sec:research-experience}), connect with researchers, and get strong LoRs (\autoref{sec:lor}).
+As discussed in \autoref{chap:research-experience}, having a successful research experience can greatly strengthen a PhD application.
+Research experience gives you opportunities to try out research, determine what research area you're interested in, publish papers (\autoref{chap:research-experience}), connect with researchers, and get strong LoRs (\autoref{chap:lor}).
 This section provides some guidance on how to gain research experience as an undergraduate student or student at a smaller college where research opportunities might be limited.
@@
 By contributing code, fixing bugs, implementing new features, or providing documentation, students can gain practical research experience and interact with experienced developers and researchers in the field.
-Not only you gain research experience, but you might be able to get a LoR from the project maintainer, and you should write about this experience in your SOP (\autoref{sec:sop}.)
+Not only you gain research experience, but you might be able to get a LoR from the project maintainer, and you should write about this experience in your SOP (\autoref{chap:sop}.)
 
@@
 By creating profiles, joining relevant research groups, and participating in discussions, students can connect with established researchers, contribute to ongoing projects, and potentially collaborate on publications or research proposals as they provide access
+
+% \chapter{Fellowship Applications}\label{sec:fellowship}
+% \chapterinfo{How to apply for fellowships such as NSF GRFP, DOD DSEG to fund your PhD.}
+
+%To give students as well as their LoR writer on insider view.
+Most applications are from students at "top" places where professors are very knowledgable with NSF reviews etc and know how to write a good LoR.
+This is not the case for students at smaller places.
+Level the field , to help ppl who are not familiar with NSF ..
+Of course, while the actual research experience and proposals are dependent ,  having these knowledge help level the field.
+Also, the LoR writer can also get some insights on what to write in the LoR.
+
+% Getting a fellowship such as NSF GRFP can significantly improve your chances of getting into a good PhD program.
+In general, fellowships are competitive, prestigious, and provide you with more freedom in your research.
+This section provides some guidance on how to apply for fellowships and increase your chances of success.
+
+% \begin{commentbox}[Vu:]
+%     I did not apply for NSF GRFP, DOD DSEG or other ``early'' PhD fellowships because I did not know about them.
+I only learned about them after I start my PhD and already have an MS (thus is eligible for some of them).
+I wish I had known about them and applied.
+% \end{commentbox}
+
+% \begin{commentbox}[Vu:]
+
+% \end{commentbox}
+% \section{NSF GRFP}
+% We review applications in December--January and participate in a virtual panel to discuss our reviews.
+
+% \paragraph{Evaluating the application's research potential, \emph{not} the project potential} This is very different than a normal proposal like an NSF one that asks for funding for a specific project.
+In such case the project is evaluated.
+However, in the GRFP, the research potential of the applicant is evaluated, not the project.
+This means that you should focus on your research experience, your potential to do research, and how you can contribute to the field, etc.
+You should also focus on your broader impacts, e.g.,  how you can help improve STEM education through outreach and mentoring, how you can contribute to diversity and inclusion in STEM fields, how you can help society, etc.
 
@@
 The document was initially intended for international students but has been generalized to all students.
-I have put the source code of this document on \href{https://github.com/nguyenthanhvuh/phd-cs-us}{GitHub} so that others can contribute to it.
-
+The document is open-source and available on \href{https://github.com/nguyenthanhvuh/phd-cs-us}{GitHub} so that others can contribute to it.
 \paragraph{About Me} I am an associate professor in the CS dept at George Mason University (Mason).
@@
 Both UNL and Mason are large R1 research universities in the US and have many international students in their CS graduate programs.
-I have been in the PhD admission process at Mason and UNL for many years.
+My personal and lab website is at \href{https://dynaroars.github.io}{dynaroars.github.io}.
+
+I have been involved the PhD admission process at Mason and UNL for many years.
 This allows me to have a good understanding of the process, the challenges students face, and what faculty are looking for.
 Currently I serve as the program director of the MS program in Software Engineering at Mason (thus also have experience with MS admission---which is quite different than PhD as discussed in \autoref{chap:ms}).
-My personal and lab website is at \href{https://dynaroars.github.io}{dynaroars.github.io}.
+I also have been in the panels of PhD fellowships including NSF GRFP and DoD NDSEG, which allows me to provide some insights on external fellowhip applications (\autoref{sec:fellowships}).
 
 Though I was not an international student, many of my students and collaborators are/were.
-I also mentor multiple students from Vietnam and have close colleagues and friends who were once international students.
+I also mentor students from Vietnam and have close colleagues and friends who were once international students.
 I hope to capture the diverse challenges and experiences they've faced in this document so that it can be a valuable resource for prospective international students.

### 2024-12-11 demystify.tex
updates

@@
 
+\part{After You Apply}\label{part:after-apply}
 \chapter{Interview and the Waiting Game}\label{chap:interview}
@@
 
-\part{Funding}\label{part:funding}
+\part{Funding, Schools, and Profs.}\label{part:funding-schools-profs}
+
 \chapter{Funding}\label{sec:funding}
@@
 
-\chapterinfo{Not every university has PhD progarms in CS. Not every professor, even those in CS, can advise or graduate CS PhD students.}
+\chapterinfo{Not every university has a CS PhD program. Not every professor, even those in CS, can advise or graduate CS PhD students.}
 
@@
 
-\section{Are profs. so busy that they completely ignore emails?} \label{sec:busy}
+\subsection{Are profs. so busy that they completely ignore emails?}\label{sec:busy}
 

### 2024-12-12 demystify.tex
updates

@@
 
-\subsection{Waiving Your Right} When you ask someone to write a letter for you, you will be asked if you want to waive your right to see the letter.  You should \emph{always} waive your right.
+\subsection{Waiving Your Right}\label{sec:waive-right}
+
+When you ask someone to write a letter for you, you will be asked if you want to waive your right to see the letter.  You should \emph{always} waive your right.
 
@@
 
-\section{What can you do?}\label{sec:lor-what-to-do}
-
-While you cannot control what your letter writers write, there are multiple things you can do to help them write a strong letter for you.
+%\section{What can you do?}\label{sec:lor-what-to-do}
+
+%While you cannot control what your letter writers write, there are multiple things you can do to help them write a strong letter for you.
 
@@
 
-\subsection{How do you write LoR? Is it lots of work?}\label{sec:lor-writing}
+\subsection{How is a LoR written? Is it lots of work?}\label{sec:lor-writing}
 
 If a student asks me to write letter for them, I will agree as I believe it is part of my job.
-However, I will ask them to waive their right to see the letter (\autoref{sec:lor-what-to-do}), and will not write if they do not do so.
-
-I will also let the student know if I can write a strong letter for them.
-If I can't, I will tell them and suggest they find someone else.
+However, I will ask them to waive their right to see the letter (\autoref{sec:waive-right}), and will not write if they do not do so.
+
+I will also let the student know if I cannot write a strong letter for them (e.g., I don't know them that well), and suggest they find someone else.
 If they insist, then I will write for them.
@@
 
-For students who I know well, i.e., strong letters, then writing will take a lot longer as it will be personalized.
-While I have a general template of such strong letters, it still can take me an hour or more to write such a letter.
-I often ask the students to provide information (\autoref{sec:lor-what-to-do}) and what they think I should highlight in the letter.
+For a student that I know well, I will write them a strong letter, which takes a lot longer as it will be personalized.
+While I have a general template for LoR, it still can take me an hour or more to write it.
+I often ask the students to provide information (\autoref{sec:help-your-LOR-writers}) and what they think I should highlight in the letter.
 They can also provide me their SOP (if they already have written one) so I can complement what they say with my own perspective.
+I often do not share what I write with the students, just to keep it more genuine and honest.
+
+I usually submit a letter for a student in batch (e.g., I submit to all universities that student is applying to at the same time).
+Typically, each submission takes from 5--10 mins, depending on the application system.
+I will also let the student know when I have submitted the letter, and ask if I miss any.
 
@@
 So if you think your school is well-known, mention it (and back up with some evidence like rankings or awards) in your SOP.
-You can ask your LOR writers to talk about your school as well (\autoref{sec:lor-what-to-do}).
+You can ask your LOR writers to talk about your school as well (\autoref{sec:help-your-LOR-writers}).
 Of course, if you're interested in working with Vietnamese, consider  \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{CS programs in the US that have Vietnamese professors}.

### 2024-12-12 demystify.tex
updates

@@
 
-\section{English Tests (IELTS, TOEFL)} Unless your degrees are from specific countries such as \href{https://www.gmu.edu/international/english-language-requirements}{these}, you will need to take standardized English tests.
+\section{English Tests (IELTS, TOEFL)}\label{sec:english-tests}
+
+Unless your degrees are from the US or certain countries such as \href{https://www.gmu.edu/international/english-language-requirements}{these}, you will need to take standardized English tests.
 On one hand you will need to show some level of English proficiency, but on the other hand, you do not need to have very good scores in these tests (many adcom members themselves were once international students and struggled with English).
@@
 \end{commentbox}
+% Note that English exams are not required for all students, e.g., domestic students and those from certain countries are waived from taking these exams.
+For example, Mason waives English exams for students from these \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-Mason#standard-tests-waiver-eligible-countries}{countries}.
 
@@
 are willing to argue for your case, they might be reluctant to go against the requirement of the university.
-
-Note that English exams are not required for all students, e.g., domestic students and those from certain countries are waived from taking these exams.
-For example, Mason waives English exams for students from these \href{https://github.com/dynaroars/dynaroars.github.io/wiki/About-Mason#standard-tests-waiver-eligible-countries}{countries}.
 
@@
   \item Objective: an MS is typically to prepare you for \emph{industry}, while a PhD is to prepare you for research and academia.
-Some MS has thesis option but the research is much lighter compared to a PhD.
+Some MS has thesis option but in general research is not a focus in MS programs.
 
@@
 
-You typically need to submit the similar materials as in a PhD application, e.g., transcripts, LoRs, SOP, and test scores.
+You will submit similar materials as in a PhD application, e.g., transcripts, LoRs, SOP, and test scores.
 However, research experience is not required, and LoRs can be from anyone who can speak about your academic or working abilities.
@@
 If you have a low GPA, a high GRE score can help offset that.
-As with PhD, English proficiency tests (TOEFL/IELTS) are required for international students.
+As with PhD, English proficiency tests (TOEFL/IELTS) are required for international students (but just passing the mimimum requirement is often enough \autoref{sec:english-tests}).
+Note that just as with PhD, Englih tests are are waived if you did your undergrad at a US university.
 

### 2024-12-13 demystify.tex
updates

@@
 
-\chapter{Should You Apply to a CS PhD Program in the US?}\label{sec:should}
+\chapter{Should You Apply?}\label{sec:should}
 
@@
 International students, especially those from less well-known universities and smaller countries, often wonder if they should apply to a CS PhD program in the US.
-Specifically, they are worried about (i) the \textbf{difficulty of getting admitted} and (ii) the \textbf{cost of graduate study} in the US.
-Part of the reasons is the lack of information and guidance on the admission process in the US (let's face it, social places like Facebook and Reddit are full of confusion and contradictory information).
+Common concerns include (i) the \textbf{difficulty of getting admitted}, (ii) the \textbf{cost of graduate study} in the US, and for some, the fear of \textbf{discrimination} based on gender identity, sexual orientation, or cultural background.
+
+Much of these uncertainties stem from the lack of information and guidance on the admission process in the US (social places like Facebook and Reddit are full of confusion and contradictory information).
 This document aims to address these concerns.
 
-First, applying to a good US university \emph{should not} be any harder than at schools in other countries.
+\paragraph{Not any harder than other countries} First, applying to a good US university \emph{should not} be any harder than at schools in other countries.
 It might even be more flexible since CS PhD in the US \emph{do not} require having an MS or a research topic, proposal, or adviser in advance (\autoref{sec:non-us-differences} compares CS PhD study in the US to other countries).
@@
 
-Second, PhD students in CS \emph{do not} need to worry about funding, especially at good R1
+\paragraph{Funding is not an issue} Second, PhD students in CS \emph{do not} need to worry about funding, especially at good R1
 universities in the US. 
@@
 receive \emph{full funding} (\autoref{sec:funding}) to support your study. This includes tuition,
-health insurance, and stipend (i.e., you get paid for your study). 
-For many students, stipends are not only sufficient for them to live but also for their families (e.g., spouse and children) to live comfortably.
+health insurance, and stipend (i.e., in STEM field you get paid for your study!). 
+For many students, the stipends provided are not only sufficient to cover your own living expenses but also enough to support your families (e.g., spouse and children), allowing them to live comfortably.
 
@@
 Full funding for CS PhD students is the norm in the US, and I'd go as far as to say that if you are not admitted with full funding, you might want to not accept the offer.
+
+\paragraph{Discrimination is not tolerated} Third, higher education in the US take a very strong stance against discrimination, especially at the university level.
+Many resources are available to encourage and support students from underrepresented and minority (URM) groups to pursue research, e.g., specific scholarships and fellowships.
+Faculty in CS are also \emph{very} familiar with and are highly encouraged to support URM students in their group.
+There are also incentives, such as dedicated funding, to motivate faculty to recruit and mentor URM students (\autoref{sec:urm}).
 
@@
 
-In the US, you often start your PhD without an adviser or project and find them later.
-You will have two initial years to explore and find an adviser and research topic.
-After that you explore projects with that adviser before settling on one (can take a couple of years).
-
-\emph{Course work}: as mentioned above in the US you will spend the first couple of years taking classes and exploring potential adviser and research topics. 
+In the US, you often start your PhD without an adviser or project and find them later. Usually you have two initial years to take classes, explore and find an adviser and research topic.
+
+\emph{Course work}: In the US ou will spend the first couple of years taking classes and exploring potential adviser and research topics. 
 After that, you have to pass a series of exams during your PhD, e.g., qualifying exam, comprehensive exam, thesis proposal defense\footnote{The word ``ABD'' (all but dissertation) is used in the US to refer to a PhD candidate who has finished all course work and exams and only needs to write and defend their dissertation.}.
 
-In other countries, you start your research right away, e.g., you immediately work on the research project you proposed with the adviser you chose.
+In other countries, you often start your research right away, e.g., you immediately work on the research project you proposed with the adviser you chose.
 Moreover, you might not have exams like those in the US or only have to do a few of them.
@@
 After you submit your PhD application, it will first be checked for general requirements, e.g., did you submit your transcripts and standard scores?
-did your reference writers submit their letters?
 Usually, this screening process is done through a central university system, i.e., not by CS faculty.
@@
 For example, if the student is interested in a research area that no faculty is working on, or an area where no faculty is taking new students (e.g., AI/ML where faculty likely already has many students) then the student will not be admitted.
-This is increasingly common as the number of applicants increases significantly in recent years in CS.
-
-However, if the student has contacted a faculty member and that faculty member is interested in the student and has made this known to the admission committee, then the student is likely to be admitted, \emph{even if the faculty does not review the application or has funding} (thus the importance of contacting faculty).
-Similarly, if the student mentioned a faculty member in their SOP, adcom might ask that faculty member to look at their application and if they are interested in the student.
-Even if the student has a weak profile (but still passes the minimum requirement from the university), they might be admitted if a faculty member is willing to take them.
+This is increasingly common as the number of applicants grows much faster than the number of available faculty.
+Note that not every CS faculty can formally advise and graduate CS PhD students (\autoref{sec:faculty-types}).
+
+However, if the student has contacted a faculty member and that faculty is interested in the student and has made this known to the admission committee, then the student is likely to be admitted, \emph{even if the faculty does not review the application or has funding} (thus the benefit of contacting faculty as discussed in\autoref{sec:contact}).
+Similarly, if the student mentioned a faculty in their SOP, adcom might ask that faculty to look at their application and if they are interested in the student.
+Even if the student has a weak profile (but still passes the minimum requirement from the university), they might be admitted if a faculty is willing to take them.
 Adcom members, especially in the US, are very reluctant to go against the faculty's decision (e.g., if a faculty wants to admit a student, we are not going to reject them).
@@
 We also talk to each other for interesting or strong applications, e.g., how to recruit this student or who should be the adviser.
-If the student mentioned a faculty member in their SOP, we might ask that faculty member if they are interested in the student.
+If the student mentioned a faculty member in their SOP, we might ask that faculty if they are interested in the student.
 
@@
 
-The goal of adcom is to evaluate your research experience, potential, and interest to see if you \textbf{fit into our PhD program}!
+The goal of adcom is to evaluate your research experience, potential, and interest to see if you \textbf{fit into its PhD program}!
 The emphasis here is \emph{fitting}, which varies from school to school, faculty to faculty, and even from year to year.
-The committee will look at various factors, but the most important ones are letters of recommendation (LORs),  statements of purpose (SOP), and research experience and potential, e.g., publications.
+The committee will look at various factors, but the most important ones are letters of recommendation (LORs),  statements of purpose (SOP), and research background and experience, e.g., publications.
 
@@
 
-\paragraph{Remind Your Writers} After entering your writers' information in the application system, you should tell your writers about that and let them know they will soon receive an email from the university to submit their letters.
+\subsection{Remind Your Writers}\label{sec:remind-writers}
+
+After entering your writers' information in the application system, you should tell your writers about that and let them know they will soon receive an email from the university to submit their letters.
 You should also remind them when you submit your application and ask them to submit their letters on time if they haven't done so.
@@
 However, we do not use it this way in academia, e.g., ``who is your PI?'' might only make sense to fellow Reddit applicants, but might not be understood by others, ``who is your adviser?'' is what you want to use.
-    \item[STEM] Science, Technology, Engineering, and Mathematics. A term used to group together these academic disciplines. CS is considered a STEM field.
+
+  \item[STEM] Science, Technology, Engineering, and Mathematics. Fields in or related to these areas are considered STEM fields. CS is a STEM field.
+
   \item[URM] Underrepresented Minority. Refers to groups that are underrepresented in higher education, e.g., Black, Hispanic, and Native American students.

### 2024-12-14 demystify.tex
updates

@@
 \paragraph{Discrimination is not tolerated} Third, higher education in the US take a very strong stance against discrimination, especially at the university level.
-Many resources are available to encourage and support students from underrepresented and minority (URM) groups to pursue research, e.g., specific scholarships and fellowships.
-Faculty in CS are also \emph{very} familiar with and are highly encouraged to support URM students in their group.
-There are also incentives, such as dedicated funding, to motivate faculty to recruit and mentor URM students (\autoref{sec:urm}).
+Many resources are available to encourage and support students from minority groups to pursue research, e.g., specific scholarships and fellowships.
+Faculty in CS are also \emph{very} familiar with and are highly encouraged to support such students in their group.
+There are also incentives, such as dedicated funding, to motivate faculty to recruit and mentor minority students (\autoref{sec:urm}).
 

### 2024-12-14 demystify.tex
updates

@@
 
-\chapter{Letters of Recommendation (LOR)}\label{chap:lor}
+\chapter{Letters of Recommendation (LoR)}\label{chap:lor}
 \chapterinfo{\emph{LORs are very important}, but only if they are personalized and talk about your research ability.}
@@
 
-Most PhD programs will require at least \textbf{two LORs}. 
-The goal of the LoR is to provide the adcom with an assessment of your research ability and potential. 
-The letter should 
-focus on your research background, achievements, and potentials, all of which paint a compelling picture of you as a researcher.
+Most PhD programs will require at least \emph{two} letter of recommendations (LoRs). 
+The goal of the LoRs is to provide the adcom with an assessment of your research ability from someone who knows you.
+The LoRs should focus on your research background, achievements, and potentials, all of which paint a compelling picture of you as a researcher to adcom.
 
@@
 LoR writers should also \emph{have the credibility} to convince adcom members on your research ability and potential.
-Such a person is typically someone who has a PhD (so they know what a PhD is about) and is active in research (so they can evaluate your research ability).
+Such a person is typically someone who has a PhD, so they know what a PhD is about and so can evaluate your research ability.
 LoR writers are often your research advisers, professors, or supervisors who have worked with you on research projects.
 
-\subsection{LoR from Famous People}  Having a strong letter from an internationally recognized researcher will \emph{greatly strengthen} your application.
-A strong LOR from such a researcher can outweigh the lack of publications (or low GPA or unknown school).
-The reason is simple: adcom members trust people that they know and respect.
-If a well-known researcher enthusiastically recommends you, then the adcom members will likely trust that recommendation.
+\subsection{LoR from Famous People}  Having a strong letter from an internationally recognized researcher can \emph{greatly strengthen} your application.
+Such a letter can outweigh the lack of publications (or low GPA or unknown school).
+A well-known researcher has a reputation to maintain (so they won't recommend someone bad), and adcom member trust people that they know, or heard about and respect (compared to someone they don't know).
 
@@
 
-It is also worth noting that a LoR from a ``full'' senior professor who has not been active in research might not be as strong as a LoR from a junior faculty who has been actively mentoring you in research.
+%It is also worth noting that a LoR from a ``full'' senior professor who has not been active in research might not be as strong as a LoR from a junior faculty who has been actively mentoring you in research.
 This also applies to administrative, e.g., chairs, directors, and deans (\autoref{sec:admin-letters}.) The fame of the writer is less important than how well they know you and can write about your research ability.
-
-Many students get letters from supervisors from companies where they did internships or are working.
-It is OK as long as it is a research-based personalized letter.
-Again, the emphasis here is \emph{research}, i.e., the letters should describe your research experiences and potential.
-Letters focusing on your coursework or non-research projects at a company won't carry much weight.
-
-\begin{commentbox}[Vu:]
-  If your grading system is not US standard, or you are from a good school unknown outside your country, you can ask your reference writers to explain that in their letters.
-For example, ``Bach Khoa'' are the top universities in Vietnam for STEM studies but few people outside Vietnam know about them.
-So if you are from there, you should ask your reference writers to mention that.
-\end{commentbox}
 
@@
 
-  \subsection{LoRs from Admin People}\label{sec:admin-letters}
-
-  Many international students try to get LoRs from high-ranking administrators in their universities such as department chair/head, dean, or director.
+  \subsection{LoRs from Dept Chair, Dean, or Boss and Supervisor at Work}\label{sec:admin-letters}
+
+  Many students try to get LoRs from high-ranking administrators in their universities such as department chair/head, dean, or director.
 The students never worked with these people (they might take a class or so with these profs), but mistakenly believe that these LoRs are valuable due to the writer's high position in the university.
@@
   \end{commentbox}
+
+  Many students get letters from supervisors from companies where they did internships or are working.
+It is OK as long as it is a research-based personalized letter.
+Again, the emphasis here is \emph{research}, i.e., the letters should describe your research experiences and potential.
+Letters focusing on your coursework or non-research projects at a company won't carry much weight.
 
@@
 \subsection{Self-written Letters are Bad}\label{sec:self-letters}
-Many students write letters themselves, typically due to the requests of their profs.
-or letter writers.
-Such letters have \emph{little value} and are considered weak by reviewers---why can you not even find someone who cares or knows enough about you to write a candid personal reference letter?
+
+Many profs.
+and letter writers ask students to write their own letters, which is a common practice in some countries.
+Unfortunately, such letters have \emph{little value} and are considered weak by reviewers---why can you not even find someone who cares or knows enough about you to write a candid personal reference letter?
 Instead of the ref.
@@
 
-Adcom members are just too experienced (having reviews hundreds of those letters) to spot and suspect such letters.
-There is a very difference between a letter written by an experienced professor and a student.
+Self-written letters are easy to spot as an experienced professor would write very differently compared to an undergraduate student.
 For example, they can provide convincing and concrete examples based on their experience and compare you to their own students, and of course the writing style of a professor is very different from an undergraduate student.
@@
 As mentioned in~\autoref{sec:lor-writers}, LoR writers should be someone who knows you well and has the credibility to evaluate your research ability.
-You can explicitly ask if they are willing to write a strong letter for you.
+In the US, students usually explicitly ask if the writer would be willing to provide a strong letter, and the writer are also very direct in their response.
 If they are not willing (you should be thankful that they are honest with you), then you should ask someone else.
@@
  As mentioned above in \autoref{sec:self-letters} and \autoref{sec:generic-letters}, do not write your own letter and generic letters do not give much value.
-Thus, to help your writer to write a strong, customized LoR, you should provide them details or unique things about yourself.
+Thus, to help your writer to write a strong, customized LoR, you can provide them details or unique things about yourself.
 For example, let them know about your GPA, research and work experience, papers (if any), or anything you want them to mention.
@@
 Sometimes your writer will explicitly ask you for such information, but if not, you should provide it anyway (especially if you have not interacted with them much or have not done much research with them).
+
+\begin{commentbox}[Vu:]
+    If your grading system is not US standard, or you are from a good school unknown outside your country, you can ask your reference writers to explain that in their letters.
+For example, ``Bach Khoa'' are the top universities in Vietnam for STEM studies but few people outside Vietnam know about them.
+So if you are from there, you should ask your reference writers to mention that.
+\end{commentbox}
 

### 2024-12-14 demystify.tex
updates

@@
 
-% Below are some tips to approach your LoR writers:
-% \begin{itemize}
-% \item \textbf{Ask in advance}: You should ask for LoRs \emph{at least a month} before the deadline.
-People are busy with their own work and likely have many other students asking for LoRs.
-Also, a strong LoR takes time to write, so you should give them enough time to write a strong letter for you.
-
-% \item \textbf{Help your writers}: You should provide your writers with details about the programs you are applying to, the deadlines, and any other information they need to know.
-You should also provide them with details about your research experience and potential so that they can write a strong letter for you.
-% \item \textbf{Follow up}: After you ask for LoRs, you should follow up with your writers to make sure they have submitted the letters on time.
-You should also thank them for writing the letters for you.
-% \item \textbf{Waive your right}: When you ask someone to write a letter for you, you will be asked if you want to waive your right to see the letter.
-You should always waive your right.
+Below are some tips to approach your LoR writers:
+\begin{itemize}
+\item \textbf{Ask in advance} You should ask for LoRs \emph{at least a month} before the deadline.
+People are busy and writing a strong LoR takes time (\autoref{sec:lor-writing}), so you should give them enough time to write a strong letter for you.
+
+\item \textbf{Waive your right (\autoref{sec:waive-right})} You should always waive your right to see the letter.
 This shows that you trust your writers and that you are not trying to twist their words.
-% \item \textbf{Remind your writers}: After you enter your writers' information in the application system, you should tell your writers about that and let them know they will soon receive an email from the university to submit their letters.
-You should also remind them when you submit your application and ask them to submit their letters on time if they haven't done so.
-% \item \textbf{Help your writers}: To help your writers write a strong letter for you, you should provide them with details about your research experience and potential.
-You should also provide them with a draft of your SOP so that they can see what you are saying about yourself and complement that with their own perspective.
-% \item \textbf{Thank your writers}: After your writers have submitted their letters, you should thank them for writing the letters for you.
-You should also let them know the outcome of your applications.
-% \item \textbf{Ask for feedback}: After you have been admitted or rejected, you should ask your writers for feedback on your application.
-This will help you improve your application for future applications.
-% \item \textbf{Stay in touch}: After you have been admitted or rejected, you should stay in touch with your writers.
-This will help you build a relationship with them and will help you in the future.
-% \end{itemize}
+
+\item \textbf{Help your writers (\autoref{sec:help-your-LOR-writers})}: You should tell your writers the programs you are applying to, their deadlines, etc.
+You can also share your SOP with them and other details about your research experience and potential.
+
+\item \textbf{Ask for feedback} If the writer is very close to you, you can ask them for feedback on your application, e.g., your SOP.
+
+\item \textbf{Follow up and Stay in touch} Follow up with your writers to make sure they have submitted the letters on time.
+After your writers have submitted all of their letters, \emph{thank them}!
+Finally, let them know the outcome of your applications and stay in touch with them.
+This will help you build a relationship with them and you might need them to write for you again in the future.
+\end{itemize}
 
@@
 While I try to say something positive, e.g., the student is hardworking, or that I heard they are working on a research project, etc, the letter will still be short and weak (\autoref{sec:generic-letters}).
-Usually it takes me about 5--10 minutes to write such a letter.
-
-For a student that I know well, I will write them a strong letter, which takes a lot longer as it will be personalized.
-While I have a general template for LoR, it still can take me an hour or more to write it.
+Usually it takes me about \emph{5--10} minutes to write such a letter.
+
+Strong letters take a lot longer as it will be personalized.
+While I have a general template for LoR, it still can take me \emph{an hour or more} to write it.
 I often ask the students to provide information (\autoref{sec:help-your-LOR-writers}) and what they think I should highlight in the letter.
@@
 I often do not share what I write with the students, just to keep it more genuine and honest.
+
+My letter always has the university letterhead, my signature, and is signed and dated.
+While using a letterhead or signing the letter is not required (I never paid much attention to these when reading LoRs), it makes the letter more official and professional.
+This website has the template of the letter with letterhead and signature: \url{https://www.overleaf.com/read/xzyrxkdjfxsp#2a1f9e}.
 

### 2024-12-16 demystify.tex
updates

@@
 \item \textbf{Ask in advance} You should ask for LoRs \emph{at least a month} before the deadline.
-People are busy and writing a strong LoR takes time (\autoref{sec:lor-writing}), so you should give them enough time to write a strong letter for you.
+People are busy and writing a strong LoR takes time (\autoref{sec:lor-writing}), so give them enough time to write a strong letter for you.
 
@@
 
-\item \textbf{Ask for feedback} If the writer is very close to you, you can ask them for feedback on your application, e.g., your SOP.
+\item \textbf{Ask for feedback} If the writer is very close to you, you can ask them for feedback on your SOP (\autoref{chap:sop}) and other application materials.
+If the writer is a professor, they might have served in adcom committees, seen many SOPs, and can provide valuable feedback.
 
 \item \textbf{Follow up and Stay in touch} Follow up with your writers to make sure they have submitted the letters on time.
-After your writers have submitted all of their letters, \emph{thank them}!
+Note that their letters might have a different due date than your application (\autoref{sec:remind-writers}).
+
+After your writers have submitted all of their letters, \emph{thank them} (\autoref{sec:thank-writers}).
 Finally, let them know the outcome of your applications and stay in touch with them.
@@
 
-When you ask someone to write a letter for you, you will be asked if you want to waive your right to see the letter.  You should \emph{always} waive your right.
+When you ask someone to write a letter for you, you should \emph{always} waive your right.
 
@@
 
-\subsection{Help your LOR Writer}\label{sec:help-your-LOR-writers}
+\subsection{Helping your LOR Writer}\label{sec:help-your-LOR-writers}
 
@@
 
-\subsection{Remind Your Writers}\label{sec:remind-writers}
+\subsection{Reminding Your Writers}\label{sec:remind-writers}
 
 After entering your writers' information in the application system, you should tell your writers about that and let them know they will soon receive an email from the university to submit their letters.
-You should also remind them when you submit your application and ask them to submit their letters on time if they haven't done so.
+You should also tell them when you submit your application and remind them to submit their letters on time if they haven't done so.
 
@@
 
-\paragraph{Thank Your Writers} After your application is submitted and your writers have submitted their letters (i.e., the wait begins), you should send a quick thank you note, which serves both as an acknowledgement that you know they have submitted the letters and as an appreciate for their help.
+\subsection{Thank Your Writers}\label{sec:thank-writers}
+After your application is submitted and your writers have submitted their letters (i.e., the wait begins), you should send a quick thank you note, which serves both as an acknowledgement that you know they have submitted the letters and as an appreciate for their help.
+
 You should also let them know the outcome of your application, regardless of whether you are admitted or not.

### 2024-12-16 demystify.tex
updates

@@
 \usepackage[T1]{fontenc}
-\usepackage[margin=1.5in]{geometry}
 \usepackage{soul}
@@
 \setlength\intextsep{0pt} % remove extra space above and below in-line float
-\usepackage{hyperref}
-\hypersetup{
-  colorlinks,
-  citecolor=black,
-  filecolor=black,
-  linkcolor=blue,
-  urlcolor=blue,
-}
+
 \usepackage{booktabs}
@@
 \usepackage{xcolor}
-
-\usepackage{anyfontsize}
-\usepackage{sectsty}
+%\usepackage{anyfontsize}
+%\usepackage{sectsty}
 
@@
 
-\def\chapterinfo#1{%
-  \addcontentsline{toc}{chapterinfo}{%
-    \noexpand\numberline{}\color{black}{#1}}%
+%\usepackage{tocloft} % For custom TOC levels
+
+\newcommand{\chapterinfo}[1]{%
+  \phantomsection
+  \addcontentsline{toc}{chapterinfo}{#1}%
 }
-\def\sectioninfo#1{%
-  \addcontentsline{toc}{sectioninfo}{%
-    \noexpand\numberline{}\color{black}{#1}}%
+
+\newcommand{\sectioninfo}[1]{%
+  \phantomsection
+  \addcontentsline{toc}{sectioninfo}{#1}%
 }
-\def\subsectioninfo#1{%
-    \addcontentsline{toc}{subsectioninfo}{%
-    \noexpand\numberline{}\color{black}{#1}}%
+
+\newcommand{\subsectioninfo}[1]{%
+  \phantomsection
+  \addcontentsline{toc}{subsectioninfo}{#1}%
 }
@@
 \newcommand{\red}[1]{{\color{red}{#1}}}
+
+\usepackage{hyperref}
+\hypersetup{
+    hidelinks,
+    linktoc=all,
+  colorlinks,
+  citecolor=black,
+  filecolor=black,
+  linkcolor=blue,
+  urlcolor=blue,
+  pdfpagemode=UseOutlines,
+}
+
+\usepackage[margin=1.5in]{geometry}
 
@@
 \section{PhD in the US vs. Other Countries}\label{sec:non-us-differences}
+
 \sectioninfo{Among several differences, CS PhD in the US does not require an MS degree but has a longer PhD study time.}
@@
 
-\subsection{Why CS depts do not waive the application fee?}\label{sec:fee-waive}
+\section{Why CS depts do not waive the application fee?}\label{sec:fee-waive}
 
@@
 \chapter{Research Experience}\label{chap:research-experience}
-
-\sectioninfo{\emph{Publications can greatly help}. Papers in good venues are concrete evidence that you have successfully engaged in research. But they are not required!}
+\chapterinfo{\emph{Publications can greatly help}. Papers in good venues are concrete evidence that you have successfully engaged in research. But they are not required!}
 
@@
 
+\part{Miscs and FAQs}\label{part:faqs}
 \chapter{Miscs and FAQs}\label{sec:faqs}
@@
 
+\part{Appendices}\label{part:appendices}
 \appendix

### 2024-12-16 demystify.tex
updates

@@
   boxrule=1pt, frame hidden,
-  %borderline = {1pt}{0pt}{blue!75!black},
   colback=gray!20!white,
@@
 
-%\usepackage{tocloft} % For custom TOC levels
-
 \newcommand{\chapterinfo}[1]{%
   \phantomsection
-  \addcontentsline{toc}{chapterinfo}{#1}%
+  \addcontentsline{toc}{chapterinfo}{\color{black}{#1}}%
 }
-
 \newcommand{\sectioninfo}[1]{%
   \phantomsection
-  \addcontentsline{toc}{sectioninfo}{#1}%
+  \addcontentsline{toc}{sectioninfo}{\color{black}{#1}}%
 }
-
 \newcommand{\subsectioninfo}[1]{%
   \phantomsection
-  \addcontentsline{toc}{subsectioninfo}{#1}%
+  \addcontentsline{toc}{subsectioninfo}{\color{black}{#1}}%
 }
@@
 \subsection{How much do you cost?}\label{sec:ra-cost}
-\sectioninfo{Your entire PhD program costs about \$400K in total, but you \emph{do not} pay for it.}
+\subsectioninfo{Your entire PhD program costs about \$400K in total, but you \emph{do not} pay for it.}
 

### 2024-12-16 demystify.tex
updates

@@
 
-\part{Your Application Materials}\label{part:application}
+\part{Application Materials}\label{part:application}
 
@@
 
-\section{Publications} The most concrete evidence of research ability is having \emph{papers in reputable international journals or conferences}.
+\section{Publications}\label{sec:publications}
+
+The most concrete evidence of research ability is having \emph{papers in reputable international journals or conferences}.
 Having published papers, especially at top venues, is a sign that you have been successfully involved in research.
@@
 
-In your SOP, focus on research potential and convince us through your experience, e.g., published papers.
+In your SOP, focus on research potential (\autoref{chap:research-experience}) and convince us through your experience, e.g., published papers (\autoref{sec:publications}).
 Back up your claims with \textbf{concrete evidence}.
 For example, if you say you have experience with teaching, then show what you did, e.g., undergrad TA or mentoring someone.
-If you say you work on a project, then show some results, e.g., paper submitted (or even rejected), achieved certain performance improvement over the state of the art.
+If you say you work on a research project, then show some results, e.g., paper submitted (or even rejected), achieved certain performance improvement over the state of the art.
 
@@
 
-Finally, your SOP reviewed by your LoR writers and professors, especially those who have served in adcom, or even postdocs or PhD students as they have been through this process.
+Finally, have your SOP reviewed by your LoR writers (\autoref{sec:help-your-LOR-writers}) and professors, especially those who have served in adcom, or even postdocs or PhD students as they have been through this process.
 
@@
 One common pitfall is to write about how you are from a poor family, a small town, or a developing country.
-These are not related to DEI and raise concerns about you not understanding or even bother to do some basic research about the topic.
-I have seen people consider themselves as minority because they are Asian or international students studying in the US (have you seen how many Asian or international students are in CS programs in the US?).\\
+These are not related to DEI and show that you do not understand the topic.
+I have seen people consider themselves as minority because they are Asian or international students studying in the US (have you seen how many Asian or international students are in CS programs?).\\
 
@@
 Just show your awareness and understanding of the topic and what you did or how you plan to contribute to it (e.g., mentoring students, helping others).
-One of the best DEI statements I have read starts the acknowledgement that the applicant is not a minority, but a white male in a STEM field (i.e., that make them the majority), and that they do not have much experience in DEI, but show that they are aware of the challenges faced by minority groups and how they can contribute to it.
+One of the best DEI statements I have read starts the acknowledgement that the applicant is not a minority, but a white male in a STEM field (i.e., they are the majority), and that they do not have much experience in DEI, but show that they are aware of the challenges faced by minority groups and how they can contribute to it.
 \end{commentbox}

### 2024-12-17 demystify.tex
updates

@@
   \phantomsection
-  \addcontentsline{toc}{chapterinfo}{\color{black}{#1}}%
+  \addcontentsline{toc}{chapterinfo}{\color{black}{\emph{#1}}}%
 }
@@
   \phantomsection
-  \addcontentsline{toc}{sectioninfo}{\color{black}{#1}}%
+  \addcontentsline{toc}{sectioninfo}{\color{black}{\emph{#1}}}%
 }
@@
   \phantomsection
-  \addcontentsline{toc}{subsectioninfo}{\color{black}{#1}}%
+  \addcontentsline{toc}{subsectioninfo}{\color{black}{\texttt #1}}%
 }
@@
 
-\chapterinfo{\emph{Yes, definitely}.  CS PhD study in the US is fully funded and admission into good universities in the US is not any harder than in other countries.}
+\chapterinfo{Yes, definitely.  CS PhD study in the US is fully funded and admission into good universities in the US is not any harder than in other countries.}
 
@@
 International students, especially those from less well-known universities and smaller countries, often wonder if they should apply to a CS PhD program in the US.
-Common concerns include (i) the \textbf{difficulty of getting admitted}, (ii) the \textbf{cost of graduate study} in the US, and for some, the fear of \textbf{discrimination} based on gender identity, sexual orientation, or cultural background.
-
+Common concerns include (i) the \textbf{difficulty of getting admitted}, (ii) the \textbf{cost of graduate study} in the US, and for some, the fear of \textbf{discrimination} e.g., based on gender identity or cultural background.
 Much of these uncertainties stem from the lack of information and guidance on the admission process in the US (social places like Facebook and Reddit are full of confusion and contradictory information).
@@
 
-\paragraph{Discrimination is not tolerated} Third, higher education in the US take a very strong stance against discrimination, especially at the university level.
-Many resources are available to encourage and support students from minority groups to pursue research, e.g., specific scholarships and fellowships.
+\paragraph{Discrimination is not tolerated} Third, not only is discrimination illegal, diversity and inclusion are highly valued in US universities.
+Many universities take great pride in their diverse student body and faculty.
+
+In fact, many resources are available to encourage and support students from minority and diverse groups to pursue higher education and research, e.g., specific scholarships and fellowships.
 Faculty in CS are also \emph{very} familiar with and are highly encouraged to support such students in their group.
-There are also incentives, such as dedicated funding, to motivate faculty to recruit and mentor minority students (\autoref{sec:urm}).
+There are also incentives, such as dedicated funding and awards, to motivate faculty to recruit and mentor minority students (\autoref{sec:urm}).
 
@@
 
-No, not in the US (other countries often do though~\autoref{sec:non-us-differences}).
-In the US, it is common to apply for a PhD program directly after your undergrad.
+No, while other countries often require an MS for PhD student in CS (~\autoref{sec:non-us-differences}), it is common in he US to apply for a PhD program directly after a 4-year undergrad program (e.g., after getting a B.S degree).
 Most CS PhD programs are designed so that students can get MS degree ``along the way'' to PhD, e.g., after finishing the 2-year course work.
@@
 \begin{table}
-\caption{Comparison of US and Other PhD Programs}\label{tab:us-vs-other}
+\caption{Comparison of the CS PhD program in the US and other countries}\label{tab:us-vs-other}
 \center
@@
 
-\sectioninfo{Applications are evaluated by the \emph{PhD Admission} (adcom) committee and each application is typically reviewed by three faculty members.}
+\chapterinfo{Applications are evaluated by the \emph{PhD Admission} (adcom) committee and each application is typically reviewed by three faculty members.}
 
@@
 \section{Admission Committee (adcom)}\label{sec:adcom}
+\sectioninfo{Adcom members are faculty who evaluate your application.}
 
@@
 \chapter{Letters of Recommendation (LoR)}\label{chap:lor}
-\chapterinfo{\emph{LORs are very important}, but only if they are personalized and talk about your research ability.}
+\chapterinfo{LORs are very important, but only if they are personalized and talk about your research ability.}
 
@@
 \chapter{Research Experience}\label{chap:research-experience}
-\chapterinfo{\emph{Publications can greatly help}. Papers in good venues are concrete evidence that you have successfully engaged in research. But they are not required!}
+\chapterinfo{Publications can greatly help. Papers in good venues are concrete evidence that you have successfully engaged in research. But they are not required!}
 
@@
 \chapter{Statement of Purpose (SOP)}\label{chap:sop}
-\chapterinfo{\emph{SOP is important}. Write it in such a way that makes you \emph{stand out}.}
+\chapterinfo{SOP is important. Write it in such a way that makes you \emph{stand out}.}
 
@@
 \chapter{Your School and Grades}\label{chap:your-school}
-\chapterinfo{\emph{High grades probably won't help much} (unless you're from a very top and well-known school), but bad ones likely will raise concerns.}
+\chapterinfo{High grades probably won't help much, but bad ones likely will raise concerns.}
 
@@
 \chapter{Standard Tests (GRE and IELTS)}\label{chap:standard-tests}
-\chapterinfo{\emph{Standard tests are not important}.
-GRE typically \emph{is not} required.
-For standard English tests (not required for domestic students), just do enough to pass the minimum requirements.}
+\chapterinfo{Standard tests are not important. GRE typically \emph{is not} required. For standard English tests (not required for domestic students), just do enough to pass the minimum requirements.}
 
@@
 \chapter{Interview and the Waiting Game}\label{chap:interview}
-\chapterinfo{Getting an interview is typically a \emph{good sign}; but no interview does not mean rejection}
+\chapterinfo{Getting an interview is typically a good sign; but no interview does not mean rejection}
 
@@
 \section{Notification Timeline: Why rejection letters are sent so late?}\label{sec:late-rejection}
-\sectioninfo{Grad programs often wait for the accepted students to make their decisions, typically by April 15, before sending out rejection letters.}
+\sectioninfo{Grad programs often wait for accepted students to make their decisions, typically by April 15, before sending out rejection letters.}
 

### 2024-12-17 demystify.tex
updates

@@
   \phantomsection
-  \addcontentsline{toc}{chapterinfo}{\color{black}{\emph{#1}}}%
+  \addcontentsline{toc}{chapterinfo}{\textcolor{black}{\emph{#1}}}%
 }
@@
   \phantomsection
-  \addcontentsline{toc}{sectioninfo}{\color{black}{\emph{#1}}}%
+  \addcontentsline{toc}{sectioninfo}{\textcolor{black}{\emph{#1}}}%
 }
@@
   \phantomsection
-  \addcontentsline{toc}{subsectioninfo}{\color{black}{\texttt #1}}%
+  \addcontentsline{toc}{subsectioninfo}{\textcolor{black}{\texttt #1}}%
 }
@@
 \chapter{Letters of Recommendation (LoR)}\label{chap:lor}
-\chapterinfo{LORs are very important, but only if they are personalized and talk about your research ability.}
+\chapterinfo{LORs are very important, but only if they are \textbf{personalized} and \textbf{research focused}.}
 
@@
 \chapter{Research Experience}\label{chap:research-experience}
-\chapterinfo{Publications can greatly help. Papers in good venues are concrete evidence that you have successfully engaged in research. But they are not required!}
+\chapterinfo{Publications are not \textbf{required} but can \textbf{greatly help}. Papers in good venues are concrete evidence that you have successfully engaged in research.}
 
@@
 For example, participating in Math Olympiads if you want to do theory or winning ACM programming contests if you want to ``build'' systems, e.g., software analysis.  
+So do talk about them in your SOP (\autoref{chap:sop}) and have your LoR writers mention them in their letters (\autoref{chap:lor}).
 
 \chapter{Statement of Purpose (SOP)}\label{chap:sop}
-\chapterinfo{SOP is important. Write it in such a way that makes you \emph{stand out}.}
+\chapterinfo{SOP is important. Write it in such a way that makes you \textbf{stand out}.}
 
@@
 \chapter{Standard Tests (GRE and IELTS)}\label{chap:standard-tests}
-\chapterinfo{Standard tests are not important. GRE typically \emph{is not} required. For standard English tests (not required for domestic students), just do enough to pass the minimum requirements.}
+\chapterinfo{Standard tests are \textbf{not} important.
+GRE typically \emph{is not} required.
+For standard English tests (not required for domestic students), just do enough to pass the minimum requirements.}
 

### 2024-12-18 demystify.tex
updates

@@
 \usepackage{soul}
-% \usepackage[small,compact]{titlesec} %very powerful
 \usepackage[most]{tcolorbox}
@@
 \newpage
+\renewcommand{\contentsname}{Contents and Summary}
 \tableofcontents
 
-% \chapter{Summary}\label{sec:summary}
-
-% This section summarizes the main points of this guideline. It also gives you an overview to decide which specific topics you want to explore more thoroughly.
-
-  %\item Should you apply to a PhD CS program in the US?
-
-          %\item \emph{Contacting a prof. is recommended}, but do it \emph{properly} to get a reply (\autoref{sec:contact}).
-
-          %\item  (\autoref{sec:interview}).
-
-    % \item Miscs and FAQS
-    %     \begin{itemize}
-
-    %       \item (\autoref{sec:read-all}).
-
-    %       \item
-    %       \item
-    %       \item (\autoref{sec:msrequirement}), and CS PhD takes abou (\autoref{sec:time}). Pu(\autoref{sec:non-us-differences}).
-    %       \item
-    %       \item
-    %       \item
-    %       %\item Despite some miserable stories on social media, many PhD students have good mentors, supportive lab mates, healthy working environment ... and are happy (\autoref{sec:happy}).
-    %     \end{itemize}
-
-    % \item Appendices
-    % \begin{itemize}
-    %   \item Specific advice for domestic students (\autoref{sec:domestic-students}).
-    %   \item How to find research opportunities (\autoref{sec:research-opportunities}).
-    %   \item Cultural differences between US and other countries (\autoref{sec:cultural}).
-    % \end{itemize}
-
 \mainmatter
 
-\part{Introduction To CS PhD Admission in the US}\label{part:intro}
+\part{Introduction To CS PhD Admission}\label{part:intro}
 
@@
 
-\chapterinfo{Yes, definitely.  CS PhD study in the US is fully funded and admission into good universities in the US is not any harder than in other countries.}
+\chapterinfo{Yes, definitely.  CS PhD study in the US is often fully funded and admission into good universities in the US is not any harder than in other countries.}
 
@@
 
-\chapterinfo{Applications are evaluated by the \emph{PhD Admission} (adcom) committee and each application is typically reviewed by three faculty members.}
+\chapterinfo{Applications are evaluated by the \emph{PhD Admission} (\textbf{adcom}) committee and each application is typically reviewed by three faculty members.}
 
@@
 \section{Admission Committee (adcom)}\label{sec:adcom}
-\sectioninfo{Adcom members are faculty who evaluate your application.}
+\sectioninfo{Adcom members are faculty who evaluate your application. They consider various factors, e.g., research experience, LoRs, SOP}
 
 Your applications are reviewed by a PhD admission committee or \emph{adcom} that consists of faculty members in CS\footnote{In some cases the committee can involve affiliated faculty from different disciplines.}.
-These adcom members have a wide range of expertise and background to ensure diverse perspectives in the evaluation process.
+Adcom members have a wide range of expertise and background to ensure diverse perspectives in the evaluation process.
 The size and the review load of the adcom depend on the department size.
@@
 
-Each application is assigned to about three adcom members, who will evaluate your profile and reach a consensus.
-While the assigned reviewers are the main ones deciding your application, other faculty in the department can also have access to your application and provide inputs and opinions on your profile.
-
 The PhD adcom typically involves assistant professors in the department (see \autoref{sec:faculty-types} for various types of faculty).
 This provides junior faculty the opportunities to recruit students.
-The chair of the committee will likely be a senior faculty, but they will not review individual applications and instead assign them to committee members.
+The adcom \emph{chair} will likely be a senior faculty, but they will not review individual applications and instead assign them to committee members.
 The chair will look at various factors such as research interests or mentioning faculty names to assign the applications to appropriate faculty, e.g., I am often assigned to review applicants interested in software engineering.
+
+Each application is assigned to about three adcom members, who will evaluate your profile and reach a consensus.
+They will consider various factors, e.g., research experience, GPA, LoRs, SOP, and test scores.
+More details about these factors are discussed in \autoref{part:appendices}.
 
@@
 our decision (\autoref{sec:ievaluate}) with a summary of your application, where we list
-strengths, e.g., a well-known school, and weaknesses, e.g., weak
-LORs.
+strengths, e.g., a well-known school, and weaknesses, e.g., weak LoRs.
 \end{commentbox}
@@
 % tvn: the main thing is research experiences
+
+\section{How applications are assigned to adcom members?}\label{sec:applications-assigned}
+\sectioninfo{Adcom members only review applications assigned to them (typically matching their expertise) and rarely get involved in other applications}
+
+While adcom members (in fact, any faculty) can view any application, we only review those that are assigned to us, which are already too many.
+Adcom chair will assign applications to reviewers based on their expertise (e.g., if a student says they want to do SE or interested in working with me), and reviewers will only review those applications.
+Occasionally we might look at other applications, e.g., if we know the student or have some special circumstances (e.g., if that student contacted me, I know that student, they are from Vietnam, etc).
+However, even if we look at them, we usually do not get involved in their evaluation.
+
+Note that while the assigned reviewers are the main ones deciding your application, other faculty in the department can also have access to your application and provide inputs and opinions on your profile.
+Thus, it helps to contact faculty (\autoref{sec:contact}) and mention faculty you're interested in in your SOP (\autoref{chap:sop}).
 
@@
 
-\section{How applications are assigned to adcom members?}\label{sec:applications-assigned}
-\sectioninfo{Adcom members only review applications assigned to them (typically matching their expertise) and rarely get involved in other applications}
-
-While adcom members (in fact, any faculty) can view any application, we only review those that are assigned to us, which are already too many.
-Adcom chair will assign applications to reviewers based on their expertise (e.g., if a student says they want to do SE or interested in working with me), and reviewers will only review those applications.
-Occasionally we might look at other applications, e.g., if we know the student or have some special circumstances (e.g., if that student contacted me, I know that student, they are from Vietnam, etc).
-However, even if we look at them, we usually do not get involved in their evaluation.
-
 \section{Why CS depts do not waive the application fee?}\label{sec:fee-waive}
@@
 
-\section{LoR writers}\label{sec:lor-writers}
+\section{LoR writers}\label{sec:lor-writers} \sectioninfo{LoR writers should be someone who (i) can talk in depth about your research experience and potential and (ii) have the credibility to evaluate your research ability.}
 
@@
 \section{Asking for LORs}\label{sec:asking-lor}
+\sectioninfo{Ask for LoRs at least a month before the deadline.  Waive your right to see the letter. Help your writers by providing details about your research experience and potential.}
 
@@
 
-Typically, an interview takes about 15--30 minutes, and one important aspect of evaluation is your ability to effectively communicate, including speaking and understanding English. 
-You might be asked to talk about your research experience and interests and to read a paper and discuss it.
-In some rare cases you might also be asked to solve a problem (one of my colleagues at Mason often does coding interview).
-The interview is also a good opportunity for you to ask questions, e.g., about the prof, their group, the CS PhD program, and the university.
-
-You should treat the interview as an informal chat.
-Have ``elevator pitches'' about your research experience and interests.
-You might also want to have a 5-minute presentation about your research.
-If a prof.
-asks you to read a paper, do it and be prepared to discuss it.
-You can also ask if you need to prepare for a coding problem.
-Finally, prepare some thoughtful questions to ask, e.g., about the program, the university, or the professor's research (see \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Answers-to-Ph.D-Advisor-Guide}{some questions} you can ask a potential adviser).
+\paragraph{When do interviews happen?}
+
+The timeline for interviews varies.
+Faculty set up interviews based on their (busy and erratic) schedule.
+Do not be surprised if you get an interview invitation at the last minute.
+Some schools do not do interviews at all (\autoref{sec:no-interview}).
 
@@
 
+\paragraph{Preparing for interviews} Typically, an interview takes about 15--30 minutes, and one important aspect of evaluation is your ability to effectively communicate, including speaking and understanding English.
+You might be asked to talk about your research experience and interests and to read a paper and discuss it.
+In some rare cases you might also be asked to solve a problem (one of my colleagues at Mason often does coding interview).
+
+You should treat the interview as an informal chat.
+Prepare an \emph{``elevator pitche''} about your research experience and interests.
+You might also want to have a 5-minute presentation about your research.
+If a prof.
+asks you to read a paper, do it and be prepared to discuss it.
+You should also ask if you need to prepare for coding.
+Finally, the interview gives you an opportunity to ask questions, e.g., about the program and the professor's research.
+You should definitely ask as it shows that you're interested (see \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Answers-to-Ph.D-Advisor-Guide}{some questions} you can ask).
+
 \paragraph{Follow-Up Emails} If you had an interview and have not heard back, you can email to ask about the status of your application.
 See \autoref{sec:accept-postpone-decline} for how to check status and follow-up emails.
-
-\paragraph{When do interviews happen?}
-
-The timeline for interviews varies.
-Faculty set up interviews based on their (busy and erratic) schedule.
-Do not be surprised if you get an interview invitation at the last minute.
-Some schools do not do interviews at all (\autoref{sec:no-interview}).
 
@@
 
-\section{If you do not get admitted}\label{sec:not-accepted}
+\chapter{Dealing with Rejection}\label{chap:not-accepted}
+
+\section{Try Again!}\label{sec:try-again}
 
@@
 
-\subsection{Why did you get rejected?}\label{sec:why-rejected}
+\section{Why did you get rejected?}\label{sec:why-rejected}
 \subsectioninfo{You aim too high, are overqualified, or even because you applied to AI/ML, a super competitive field in recent years with many applicants.}
@@
 Many students simply just go for very top schools (e.g., top 10) and are surprised when they are rejected to all of them, in multiple years, and completely get shut out.
-This is very obvious but many students still do this.
-An analogy is a person who has never hike wanting to climb Everest, which btw if you could, you might have a better admission chance (\autoref{sec:improve-your-chance}).
+This is very obvious but many students still do this (\autoref{sec:chance-me}).
+An analogy is a person who has never hike but wants to climb Mt.
+Everest (which btw if you could, you might have a better admission chance as mentioned in \autoref{sec:improve-your-chance}).
 
@@
 
-\subsection{Increasing your admission chance}\label{sec:improve-your-chance}
+\section{Increasing your admission chance}\label{sec:improve-your-chance}
 \subsectioninfo{You can improve your profile by being unique and standing out.}
@@
 
-\part{Funding, Schools, and Profs.}\label{part:funding-schools-profs}
+\part{Types of Funding and Choosing Schools and Profs.}\label{part:funding-schools-profs}
 

### 2024-12-18 demystify.tex
updates

@@
 \chapterinfo{Getting an interview is typically a good sign; but no interview does not mean rejection}
+
+\epigraph{Bart: ``Are we there yet? Are we there yet? Are we there yet?''\\
+Homer: ``No!''\\
+Bart: ``Okay, how about now?''}{\textsc{The Simpsons}}
 
@@
 
+In most schools I've been at, the application deadline is in December, and adcom meets begins when school starts in mid or end of January.
+Adcom chair sends out review assignments to adcom members, about 30ish per faculty (\autoref{sec:adcom}).
+We usually have about 2 weeks to review all applications.
+As mentioned in~\autoref{sec:adcom-discuss}, adcom members review applications individually and independently.
+We only discuss when there are disagreements (adcom chair will determine which applications to discuss).
+
+I typically reserve a whole day to review all applications (this is to be fair to all applicants).
 On average, I spend about 10--15 minutes reviewing each application (less for clear rejections and more for potential acceptances).
@@
 This is \emph{not necessary} and makes it look like you're trying to show off your connections.
-These nuances represent some cultural differences that you may encounter and will gradually adapt to.
+These nuances represent some cultural and academic differences that you may encounter and will gradually adapt to.
 More on cultural differences in \autoref{sec:cultural}.
@@
 \paragraph{School \autoref{chap:your-school}} Adcom already knows about your school, which is a plus (see \autoref{chap:your-school}).
-You are also more familiar with the US education system and culture (so hopefully you are aware of the cultural issues listed in \autoref{sec:cultural}).
+You are also more familiar with the US education system and academic culture (\autoref{sec:cultural}).
 
@@
 
-\chapter{Cultural Differences}\label{sec:cultural}
+\chapter{Academic and Cultural Differences}\label{sec:cultural}
 
@@
 
-This section lists some general cultural issues that students, especially international ones, might want to pay attention to.
+This section lists some general academic and cultural issues that students, especially international ones, might want to pay attention to.
 

### 2024-12-21 demystify.tex
updates

@@
 \usepackage{epigraph}
+\newcommand{\myepigraphsimpsons}[1]{
+\epigraph{\vspace{-0.2in} \emph{#1}}{\textsc{The Simpsons}}
+}
 \usepackage{cite}
@@
 
-\epigraph{\vspace{-0.2in} Don't make fun of graduate students. They just made a terrible life choice.}{\textsc{The Simpsons}}
+\myepigraphsimpsons{``Don't make fun of graduate students. They just made a terrible life choice.''}
 
@@
 
-\epigraph{How is education supposed to make me feel smarter?
-Besides, every time I learn something new, it pushes some of the old stuff out of my brain.
-Remember when I took that home wine making course, and I forgot how to drive?}{\textsc{The Simpsons}}
-
 After you submit your PhD application, it will first be checked for general requirements, e.g., did you submit your transcripts and standard scores?
@@
 
-\epigraph{Son, if you really want something in this life, you have to work for it. Now quiet! They’re about to announce the lottery numbers.}{\textsc{The Simpsons}}
+\myepigraphsimpsons{``Son, if you really want something in this life, you have to work for it. Now quiet! They’re about to announce the lottery numbers.''}
 
@@
 \chapterinfo{LORs are very important, but only if they are \textbf{personalized} and \textbf{research focused}.}
+
+\myepigraphsimpsons{``To whom it may concern… D’oh!''}
 
@@
 
-\epigraph{Bart: ``Are we there yet? Are we there yet? Are we there yet?''\\
+\myepigraphsimpsons{Bart: ``Are we there yet? Are we there yet? Are we there yet?''\\
 Homer: ``No!''\\
-Bart: ``Okay, how about now?''}{\textsc{The Simpsons}}
+Bart: ``Okay, how about now?''}
 
@@
 
-\epigraph{``Oh...
+\myepigraphsimpsons{``Oh...
 and how is education supposed to make me feel smarter?
 Besides, every time I learn something new, it pushes some old stuff out of my brain.
-Remember when I took that home wine-making course and I forgot how to drive?''}{\textsc{The Simpsons}}
+Remember when I took that home wine-making course and I forgot how to drive?''}
 
@@
 
-\epigraph{Bart, with \$10,000, we’d be millionaires! We could buy all kinds of useful things like…love!’}{\textsc{The Simpsons}}
+\myepigraphsimpsons{``Bart, with \$10,000, we’d be millionaires! We could buy all kinds of useful things like ... love!''}
 
@@
 
-\epigraph{\vspace{-0.2in} The answer to life's problems aren't at the bottom of a bottle. Heh heh! They're on TV!}{\textsc{The Simpsons}}
+\myepigraphsimpsons{``The answer to life's problems aren't at the bottom of a bottle. Heh heh! They're on TV!''}
 
@@
 
-\subsection{Schools that offer PhD in CS}
+\subsection{Schools offering PhD in CS}\label{sec:schools-offering-phd}
 
@@
 
-\epigraph{``I want to share something with you – the three little sentences that will get you through life; number 1: Cover for me, number 2: Oh, good idea, Boss, and number 3: It was like that when I got here.''}{\textsc{The Simpsons}}
+\myepigraphsimpsons{``I want to share something with you – the three little sentences that will get you through life; number 1: Cover for me, number 2: Oh, good idea, Boss, and number 3: It was like that when I got here.''}
 
@@
 
-In most schools I've been at, the application deadline is in December, and adcom meets begins when school starts in mid or end of January.
+In the schools I've been at, the application deadline is in December, and adcom meets when school starts in mid or end of January.
 Adcom chair sends out review assignments to adcom members, about 30ish per faculty (\autoref{sec:adcom}).
@@
 
-I typically reserve a whole day to review all applications (this is to be fair to all applicants).
+I typically reserve a whole day to review all applications.
 On average, I spend about 10--15 minutes reviewing each application (less for clear rejections and more for potential acceptances).
@@
 
-I read carefully \emph{strong LoRs} and skim weaker ones, noting either strong letters from well-known professors/researchers or generic content.
+I read carefully \emph{strong LoRs} and skim weaker ones, noting either strong letters from well-known professors/researchers and talk about research experiences or weak letters with generic content (e.g., ``student was in my class'').
 I skim over \emph{CV} and look for publications, research experiences, and notable achievements.
@@
 
-I skim weak \emph{SOPs} but read strong ones thoroughly.
+I skim weak \emph{SOPs} but read strong ones carefully.
 I note whether the SOP is exciting, research-oriented, standing out, e.g., applicant is minority (\autoref{sec:urm}), and tailored to our program (e.g., if they are familiar with the work of some faculty or have talked to them).
@@
 Finally, I enter my decision, which is usually either a rejection or an offer of admission with full funding (e.g., from TA).
-I also recommend very strong candidates for the University Presidential fellowship.
+I also recommend very strong candidates for the University Presidential fellowship., which is a fellowship from GMU that provides funding similar to an RA (\autoref{sec:funding}).
 Note that while the system has other decision options, eg., admit without funding, provisional admission (e.g., if they need to take some courses), I rarely use them, simply because we want to fully fund our PhD students.
@@
 
-\epigraph{\vspace{-0.2in} I didn’t do it. Nobody saw me do it. You can’t prove anything!}{\textsc{The Simpsons}}
+\myepigraphsimpsons{``I didn’t do it. Nobody saw me do it. You can’t prove anything!''}
 
@@
 
-\epigraph{\vspace{-0.2in} I’m not a bad guy! I work hard, and I love my kids. So why should I spend half my Sunday hearing about how I’m going to Hell?}{\textsc{The Simpsons}}
+\myepigraphsimpsons{``I’m not a bad guy! I work hard, and I love my kids. So why should I spend half my Sunday hearing about how I’m going to Hell?''}
 
@@
 
-\epigraph{\vspace{-0.2in} Kids, you tried your best and you failed miserably. The lesson is: never try.}{\textsc{The Simpsons}}
+\myepigraphsimpsons{``Kids, you tried your best and you failed miserably. The lesson is: never try.''}
 
@@
 
-\epigraph{\vspace{-0.2in} Ay caramba!}{\textsc{The Simpsons}}
+\myepigraphsimpsons{``Ay caramba!''}
 
@@
 
-\epigraph{\vspace{-0.2in} The whole damn system is wrong!}{\textsc{The Simpsons}}
+\myepigraphsimpsons{``The whole damn system is wrong!''}
 

### 2024-12-21 demystify.tex
updates

@@
 Thus, other writings, even those under submissions or rejected, would still be good and better than nothing.
-Be sure to upload your papers with your application and talk about them in your statement (see \autoref{chap:sop}).
+Be sure to upload your papers with your application (\autoref{sec:writing-sample}) and talk about them in your SOP (\autoref{chap:sop}).
 
@@
 \begin{commentbox}[Craig:]
-  Mason and many other universities allow you to upload your published papers and other writing samples.
+  Mason and many other universities allow you to upload your published papers and other writing samples (\autoref{sec:writing-sample}).
 In many cases, even if the papers were not published at top places, we can still determine their quality by simply skimming over the paper.
@@
 \section{Writing Sample}\label{sec:writing-sample}
-TBD TODO
+It is a good idea to upload a writing sample, e.g., a paper you wrote, a report, or a thesis.
+Even if the paper is published, not published, or already on Arxiv, you should upload it.
+Because samples are part of the application package, adcom members can quickly skim through it to see your research and writing abilities.
 
@@
 Attaching CV is OK.
-Sample papers and links to your Arxiv papers or GitHub projects are also OK if they are relevant.
+Sample papers (\autoref{sec:writing-sample}) and links to your Arxiv papers or GitHub projects are also OK if they are relevant.
 

### 2024-12-21 demystify.tex
updates

@@
 
-\part{Introduction To CS PhD Admission}\label{part:intro}
+\part{Introduction}\label{part:intro}
 
@@
 
-\myepigraphsimpsons{``Son, if you really want something in this life, you have to work for it. Now quiet! They’re about to announce the lottery numbers.''}
+\myepigraphsimpsons{``Son, if you really want something in this life, you have to work for it. Now quiet! They're about to announce the lottery numbers.''}
 
@@
 
-\myepigraphsimpsons{``To whom it may concern… D’oh!''}
+\myepigraphsimpsons{``To whom it may concern… D'oh!''}
 
@@
 
+\myepigraphsimpsons{``I've got to study harder and publish faster!''}
+
 Here we look at publications and other research experiences that can strengthen your application.
@@
 \chapterinfo{SOP is important. Write it in such a way that makes you \textbf{stand out}.}
+
+\myepigraphsimpsons{``All my life I’ve had one dream: to achieve my many goals.''}
 
@@
 
+\myepigraphsimpsons{``Woohoo! I'm a college man! I won't need my high school diploma anymore''}
+
 \section{School} Graduating from top universities \emph{that adcom members recognize helps}.
@@
 For standard English tests (not required for domestic students), just do enough to pass the minimum requirements.}
+
+\myepigraphsimpsons{``I've got to study harder. Everyone knows standardized tests are biased against the poor, the lazy, and the stupid.''}
 
@@
 \chapter{Miscellaneous}\label{chap:misc}
+
+\myepigraphsimpsons{``I'm not a college graduate. I'm not even a high school graduate. But I'm a pretty good judge of character.''}
+
 \section{CV/Resume}
@@
 It is a good idea to upload a writing sample, e.g., a paper you wrote, a report, or a thesis.
-Even if the paper is published, not published, or already on Arxiv, you should upload it.
+Whether the paper is published, not published, or already on Arxiv, you should upload it.
 Because samples are part of the application package, adcom members can quickly skim through it to see your research and writing abilities.
@@
 \chapter{Dealing with Rejection}\label{chap:not-accepted}
+\chapterinfo{Rejection is common and part of academia.  Common reasons for rejection include aiming too high, not a good fit, or bad luck. Learn from it and try again.}
+
+\myepigraphsimpsons{``You tried your best and you failed miserably. The lesson is: never try.''}
+
+Rejection is a common part of the PhD application process.
+In fact, rejection is part of academia (e.g., you will get rejected for papers, grants, jobs, etc).
+Don't take rejection personally and to learn from it.
 
@@
 
-\part{Types of Funding and Choosing Schools and Profs.}\label{part:funding-schools-profs}
+\part{Funding, Schools, and Profs.}\label{part:funding-schools-profs}
 
@@
 
-\myepigraphsimpsons{``Bart, with \$10,000, we’d be millionaires! We could buy all kinds of useful things like ... love!''}
+\myepigraphsimpsons{``Bart, with \$10,000, we'd be millionaires! We could buy all kinds of useful things like ... love!''}
 
@@
 
-\chapter{Choosing Schools and Professors}\label{sec:schoolsandprofs}
-
+\chapter{Choosing Schools}\label{sec:schoolsandprofs}
 \chapterinfo{Not every university has a CS PhD program. Not every professor, even those in CS, can advise or graduate CS PhD students.}
 
-\myepigraphsimpsons{``The answer to life's problems aren't at the bottom of a bottle. Heh heh! They're on TV!''}
+\myepigraphsimpsons{``It's not the school you go to, it's what you do while you're there!''}
+%\myepigraphsimpsons{``The answer to life's problems aren't at the bottom of a bottle. Heh heh! They're on TV!''}
 
@@
   I always encourage the students I admit to talk with my students and the students of other faculty in other schools who admitted them.
-You will unlikely hear straight-out complaints from current students in a prof’s group.
-But sometimes what is important are things that they (current students) don’t tell you.
+You will unlikely hear straight-out complaints from current students in a prof's group.
+But sometimes what is important are things that they (current students) don't tell you.
 Pay attention to their ``level of excitement'' being in the group.
@@
 
-\section{Choosing an Adviser}
-\sectioninfo{The best adviser is the one that you can work well with. But you do not know that until you start working with them.}
-
-There is no one-size-fits-all answer to this question. The best adviser is the one that you can work well with.  But how do you find such a person?
+\chapter{Choosing an Adviser}\label{chap:choosing-adviser}
+\chapterinfo{The best adviser is the one that you can work well with. But you do not know that until you start working with them.}
+
+There is no one-size-fits-all answer to finding an adviser. The best adviser is the one that you can work well with.  But how do you find such a person?
 
@@
 
-\subsection{Finding an Adviser}
+\section{Finding an Adviser}
 
@@
 \end{itemize}
-\subsection{Who can serve as a PhD adviser?}\label{sec:faculty-types}
+
+\section{Who can serve as a PhD adviser?}\label{sec:faculty-types}
 
@@
 
-\subsection{Tenured or tenure-track faculty?}\label{sec:tenure-vs-tenure-track}
+\section{Tenured or tenure-track faculty?}\label{sec:tenure-vs-tenure-track}
 
@@
 
-\myepigraphsimpsons{``I didn’t do it. Nobody saw me do it. You can’t prove anything!''}
+\myepigraphsimpsons{``I didn't do it. Nobody saw me do it. You can't prove anything!''}
 
@@
 
-\myepigraphsimpsons{``I’m not a bad guy! I work hard, and I love my kids. So why should I spend half my Sunday hearing about how I’m going to Hell?''}
+\myepigraphsimpsons{``I'm not a bad guy! I work hard, and I love my kids. So why should I spend half my Sunday hearing about how I'm going to Hell?''}
 

### 2024-12-21 demystify.tex
updates

--- 2024-12-21 demystify.tex
@@
 \usepackage[T1]{fontenc}
+\usepackage[margin=1.5in]{geometry}
 \usepackage{soul}
+% \usepackage[small,compact]{titlesec} %very powerful
 \usepackage[most]{tcolorbox}
@@
 \usepackage{epigraph}
-\newcommand{\myepigraphsimpsons}[1]{
-\epigraph{\vspace{-0.2in} \emph{#1}}{\textsc{The Simpsons}}
-}
 \usepackage{cite}
@@
 \setlength\intextsep{0pt} % remove extra space above and below in-line float
-
+\usepackage{hyperref}
+\hypersetup{
+  colorlinks,
+  citecolor=black,
+  filecolor=black,
+  linkcolor=blue,
+  urlcolor=blue,
+}
 \usepackage{booktabs}
@@
 \usepackage{xcolor}
-%\usepackage{anyfontsize}
-%\usepackage{sectsty}
+
+\usepackage{anyfontsize}
+\usepackage{sectsty}
 
@@
   boxrule=1pt, frame hidden,
+  %borderline = {1pt}{0pt}{blue!75!black},
   colback=gray!20!white,
@@
 
-\newcommand{\chapterinfo}[1]{%
-  \phantomsection
-  \addcontentsline{toc}{chapterinfo}{\textcolor{black}{\emph{#1}}}%
+\def\chapterinfo#1{%
+  \addcontentsline{toc}{chapterinfo}{%
+    \noexpand\numberline{}\color{black}{#1}}%
 }
-\newcommand{\sectioninfo}[1]{%
-  \phantomsection
-  \addcontentsline{toc}{sectioninfo}{\textcolor{black}{\emph{#1}}}%
+\def\sectioninfo#1{%
+  \addcontentsline{toc}{sectioninfo}{%
+    \noexpand\numberline{}\color{black}{#1}}%
 }
-\newcommand{\subsectioninfo}[1]{%
-  \phantomsection
-  \addcontentsline{toc}{subsectioninfo}{\textcolor{black}{\texttt #1}}%
+\def\subsectioninfo#1{%
+    \addcontentsline{toc}{subsectioninfo}{%
+    \noexpand\numberline{}\color{black}{#1}}%
 }
@@
 \newcommand{\red}[1]{{\color{red}{#1}}}
-
-\usepackage{hyperref}
-\hypersetup{
-    hidelinks,
-    linktoc=all,
-  colorlinks,
-  citecolor=black,
-  filecolor=black,
-  linkcolor=blue,
-  urlcolor=blue,
-  pdfpagemode=UseOutlines,
-}
-
-\usepackage[margin=1.5in]{geometry}
 
@@
 \newpage
-\renewcommand{\contentsname}{Contents and Summary}
 \tableofcontents
 
+% \chapter{Summary}\label{sec:summary}
+
+% This section summarizes the main points of this guideline. It also gives you an overview to decide which specific topics you want to explore more thoroughly.
+
+  %\item Should you apply to a PhD CS program in the US?
+
+          %\item \emph{Contacting a prof. is recommended}, but do it \emph{properly} to get a reply (\autoref{sec:contact}).
+
+          %\item  (\autoref{sec:interview}).
+
+    % \item Miscs and FAQS
+    %     \begin{itemize}
+
+    %       \item (\autoref{sec:read-all}).
+
+    %       \item
+    %       \item
+    %       \item (\autoref{sec:msrequirement}), and CS PhD takes abou (\autoref{sec:time}). Pu(\autoref{sec:non-us-differences}).
+    %       \item
+    %       \item
+    %       \item
+    %       %\item Despite some miserable stories on social media, many PhD students have good mentors, supportive lab mates, healthy working environment ... and are happy (\autoref{sec:happy}).
+    %     \end{itemize}
+
+    % \item Appendices
+    % \begin{itemize}
+    %   \item Specific advice for domestic students (\autoref{sec:domestic-students}).
+    %   \item How to find research opportunities (\autoref{sec:research-opportunities}).
+    %   \item Cultural differences between US and other countries (\autoref{sec:cultural}).
+    % \end{itemize}
+
 \mainmatter
 
-\part{Introduction To CS PhD Admission}\label{part:intro}
+\part{Introduction To CS PhD Admission in the US}\label{part:intro}
 
@@
 
-\chapterinfo{Yes, definitely.  CS PhD study in the US is often fully funded and admission into good universities in the US is not any harder than in other countries.}
-
-\myepigraphsimpsons{``Don't make fun of graduate students. They just made a terrible life choice.''}
+\chapterinfo{\emph{Yes, definitely}.  CS PhD study in the US is fully funded and admission into good universities in the US is not any harder than in other countries.}
+
+\epigraph{\vspace{-0.2in} Don't make fun of graduate students. They just made a terrible life choice.}{\textsc{The Simpsons}}
 
 International students, especially those from less well-known universities and smaller countries, often wonder if they should apply to a CS PhD program in the US.
-Common concerns include (i) the \textbf{difficulty of getting admitted}, (ii) the \textbf{cost of graduate study} in the US, and for some, the fear of \textbf{discrimination} e.g., based on gender identity or cultural background.
+Common concerns include (i) the \textbf{difficulty of getting admitted}, (ii) the \textbf{cost of graduate study} in the US, and for some, the fear of \textbf{discrimination} based on gender identity, sexual orientation, or cultural background.
+
 Much of these uncertainties stem from the lack of information and guidance on the admission process in the US (social places like Facebook and Reddit are full of confusion and contradictory information).
@@
 
-\paragraph{Discrimination is not tolerated} Third, not only is discrimination illegal, diversity and inclusion are highly valued in US universities.
-Many universities take great pride in their diverse student body and faculty.
-
-In fact, many resources are available to encourage and support students from minority and diverse groups to pursue higher education and research, e.g., specific scholarships and fellowships.
+\paragraph{Discrimination is not tolerated} Third, higher education in the US take a very strong stance against discrimination, especially at the university level.
+Many resources are available to encourage and support students from minority groups to pursue research, e.g., specific scholarships and fellowships.
 Faculty in CS are also \emph{very} familiar with and are highly encouraged to support such students in their group.
-There are also incentives, such as dedicated funding and awards, to motivate faculty to recruit and mentor minority students (\autoref{sec:urm}).
+There are also incentives, such as dedicated funding, to motivate faculty to recruit and mentor minority students (\autoref{sec:urm}).
 
@@
 
-The guideline can be useful for all students, but it is primarily intended for \textbf{international students} from \textbf{smaller countries and less well-known universities} (or domestic students from US universities with no PhD programs or limited research opportunities).
+While this guideline can be useful for all students, it would most benefit \textbf{international students} from \textbf{smaller countries and less well-known universities} (or domestic students from US universities with no PhD programs or limited research opportunities).
 
 Students from top schools with strong research programs and experience might already know most of the information given here.
+They might also receive guidance from their professors or grad students who have gone through the process.
+Students from smaller schools or countries, however, might not have access to such resources and might not know where to start.
 My goal is thus to level the playing field by providing info that is not readily available to less privileged students. 
-I hope to encourage more students with such backgrounds to apply to CS PhD programs in the US.
-
-Another goal of this document is to explain the \emph{``why''} behind the admission process, e.g., why LoRs are important, why you should not draft your own LoR, why you should contact professors, etc.
+I hope to encourage more students with such backgrounds to apply and succeed in CS PhD programs in the US.
+
+This document also aim to explain the \emph{``why''} behind the admission process, e.g., why LoRs are important, why you should not draft your own LoR, why you should contact professors, etc.
+There are numerous \href{https://github.com/nguyenthanhvuh/phd-cs-us}{resources online} from students  that tell you \emph{``what''} to do, but few explain the \emph{``why''}.
 Understanding the reason and mindset of the admission committee and profs.
@@
 
-While this document may not fully meet the needs of students aiming for top-10 schools (\autoref{sec:selecting-ranking-schools} and \autoref{sec:ranking}), which typically have more stringent requirements (e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU), it should be valuable for those targeting other reputable R1 institutions.
-It is certainly useful for students aiming for Mason!
+While this document may not fully meet the needs of students aiming for top-10 schools (\autoref{sec:selecting-ranking-schools} and \autoref{sec:ranking}), which typically have more stringent requirements (e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU), it should be valuable for those targeting other reputable R1 institutions, which are frankly the majority of good schools in the US.
 
@@
 
-No, while other countries often require an MS for PhD student in CS (~\autoref{sec:non-us-differences}), it is common in he US to apply for a PhD program directly after a 4-year undergrad program (e.g., after getting a B.S degree).
+No, not in the US (other countries often do though~\autoref{sec:non-us-differences}).
+In the US, it is common to apply for a PhD program directly after your undergrad.
 Most CS PhD programs are designed so that students can get MS degree ``along the way'' to PhD, e.g., after finishing the 2-year course work.
@@
 \section{PhD in the US vs. Other Countries}\label{sec:non-us-differences}
-
 \sectioninfo{Among several differences, CS PhD in the US does not require an MS degree but has a longer PhD study time.}
@@
 \begin{table}
-\caption{Comparison of the CS PhD program in the US and other countries}\label{tab:us-vs-other}
+\caption{Comparison of US and Other PhD Programs}\label{tab:us-vs-other}
 \center
@@
 
-\chapterinfo{Applications are evaluated by the \emph{PhD Admission} (\textbf{adcom}) committee and each application is typically reviewed by three faculty members.}
+\sectioninfo{Applications are evaluated by the \emph{PhD Admission} (adcom) committee and each application is typically reviewed by three faculty members.}
+
+\epigraph{How is education supposed to make me feel smarter?
+Besides, every time I learn something new, it pushes some of the old stuff out of my brain.
+Remember when I took that home wine making course, and I forgot how to drive?}{\textsc{The Simpsons}}
 
@@
 \section{Admission Committee (adcom)}\label{sec:adcom}
-\sectioninfo{Adcom members are faculty who evaluate your application. They consider various factors, e.g., research experience, LoRs, SOP}
 
 Your applications are reviewed by a PhD admission committee or \emph{adcom} that consists of faculty members in CS\footnote{In some cases the committee can involve affiliated faculty from different disciplines.}.
-Adcom members have a wide range of expertise and background to ensure diverse perspectives in the evaluation process.
+These adcom members have a wide range of expertise and background to ensure diverse perspectives in the evaluation process.
 The size and the review load of the adcom depend on the department size.
@@
 
+Each application is assigned to about three adcom members, who will evaluate your profile and reach a consensus.
+While the assigned reviewers are the main ones deciding your application, other faculty in the department can also have access to your application and provide inputs and opinions on your profile.
+
 The PhD adcom typically involves assistant professors in the department (see \autoref{sec:faculty-types} for various types of faculty).
 This provides junior faculty the opportunities to recruit students.
-The adcom \emph{chair} will likely be a senior faculty, but they will not review individual applications and instead assign them to committee members.
+The chair of the committee will likely be a senior faculty, but they will not review individual applications and instead assign them to committee members.
 The chair will look at various factors such as research interests or mentioning faculty names to assign the applications to appropriate faculty, e.g., I am often assigned to review applicants interested in software engineering.
-
-Each application is assigned to about three adcom members, who will evaluate your profile and reach a consensus.
-They will consider various factors, e.g., research experience, GPA, LoRs, SOP, and test scores.
-More details about these factors are discussed in \autoref{part:appendices}.
 
@@
 our decision (\autoref{sec:ievaluate}) with a summary of your application, where we list
-strengths, e.g., a well-known school, and weaknesses, e.g., weak LoRs.
+strengths, e.g., a well-known school, and weaknesses, e.g., weak
+LORs.
 \end{commentbox}
@@
 % tvn: the main thing is research experiences
+
+\section{How are decisions made?}\label{sec:how-decisions}
+\sectioninfo{Even if all adcom reviewers recommend acceptance, the application can still be rejected. Vice versa, if all reviewers think the application is weak, the student might still be admitted.}
+
+After reviewers have evaluated an application, adcom chair will review all evaluations, look at entered notes, and ask reviewers to discuss and resolve discrepancies to reach a consensus (e.g., a reviewer wants to accept but the other wants to reject).
+Typically, the decision is made entirely by the reviewers; there is no involvement from the adcom chair, department chair, or others.
+In most cases adcom members, even those reviewing the same application, make decisions independently and do not talk to each other (just a common practice to avoid biasing).
+In some rare cases we might (\autoref{sec:adcom-discuss}).
+
+Even if \emph{all} reviewers recommend acceptance, the application is not automatically accepted, especially if no faculty is willing to advise the student.
+For example, if the student is interested in a research area that no faculty is working on, or an area where no faculty is taking new students (e.g., AI/ML where faculty likely already has many students) then the student will not be admitted.
+This is increasingly common as the number of applicants grows much faster than the number of available faculty.
+Note that not every CS faculty can formally advise and graduate CS PhD students (\autoref{sec:faculty-types}).
+
+However, if the student has contacted a faculty member and that faculty is interested in the student and has made this known to the admission committee, then the student is likely to be admitted, \emph{even if the faculty does not review the application or has funding} (thus the benefit of contacting faculty as discussed in\autoref{sec:contact}).
+Similarly, if the student mentioned a faculty in their SOP, adcom might ask that faculty to look at their application and if they are interested in the student.
+Even if the student has a weak profile (but still passes the minimum requirement from the university), they might be admitted if a faculty is willing to take them.
+Adcom members, especially in the US, are very reluctant to go against the faculty's decision (e.g., if a faculty wants to admit a student, we are not going to reject them).
+
+\section{Do adcom members talk to each other?}\label{sec:adcom-discuss}
+\sectioninfo{Sometimes adcom members discuss applicants, but in most cases they make independent decisions.}
+
+However, in some rare cases we do talk to each other.
+For example, when there are discrepancies in evaluations, in which case the adcom chair will ask reviewers to discuss the application to reach a consensus.
+We also talk to each other for interesting or strong applications, e.g., how to recruit this student or who should be the adviser.
+If the student mentioned a faculty member in their SOP, we might ask that faculty if they are interested in the student.
 
@@
 
-Note that while the assigned reviewers are the main ones deciding your application, other faculty in the department can also have access to your application and provide inputs and opinions on your profile.
-Thus, it helps to contact faculty (\autoref{sec:contact}) and mention faculty you're interested in in your SOP (\autoref{chap:sop}).
-
-\section{How are decisions made?}\label{sec:how-decisions}
-\sectioninfo{Even if all adcom reviewers recommend acceptance, the application can still be rejected. Vice versa, if all reviewers think the application is weak, the student might still be admitted.}
-
-After reviewers have evaluated an application, adcom chair will review all evaluations, look at entered notes, and ask reviewers to discuss and resolve discrepancies to reach a consensus (e.g., a reviewer wants to accept but the other wants to reject).
-Typically, the decision is made entirely by the reviewers; there is no involvement from the adcom chair, department chair, or others.
-In most cases adcom members, even those reviewing the same application, make decisions independently and do not talk to each other (just a common practice to avoid biasing).
-In some rare cases we might (\autoref{sec:adcom-discuss}).
-
-Even if \emph{all} reviewers recommend acceptance, the application is not automatically accepted, especially if no faculty is willing to advise the student.
-For example, if the student is interested in a research area that no faculty is working on, or an area where no faculty is taking new students (e.g., AI/ML where faculty likely already has many students) then the student will not be admitted.
-This is increasingly common as the number of applicants grows much faster than the number of available faculty.
-Note that not every CS faculty can formally advise and graduate CS PhD students (\autoref{sec:faculty-types}).
-
-However, if the student has contacted a faculty member and that faculty is interested in the student and has made this known to the admission committee, then the student is likely to be admitted, \emph{even if the faculty does not review the application or has funding} (thus the benefit of contacting faculty as discussed in\autoref{sec:contact}).
-Similarly, if the student mentioned a faculty in their SOP, adcom might ask that faculty to look at their application and if they are interested in the student.
-Even if the student has a weak profile (but still passes the minimum requirement from the university), they might be admitted if a faculty is willing to take them.
-Adcom members, especially in the US, are very reluctant to go against the faculty's decision (e.g., if a faculty wants to admit a student, we are not going to reject them).
-
-\section{Do adcom members talk to each other?}\label{sec:adcom-discuss}
-\sectioninfo{Sometimes adcom members discuss applicants, but in most cases they make independent decisions.}
-
-However, in some rare cases we do talk to each other.
-For example, when there are discrepancies in evaluations, in which case the adcom chair will ask reviewers to discuss the application to reach a consensus.
-We also talk to each other for interesting or strong applications, e.g., how to recruit this student or who should be the adviser.
-If the student mentioned a faculty member in their SOP, we might ask that faculty if they are interested in the student.
-
-\section{Why CS depts do not waive the application fee?}\label{sec:fee-waive}
+\subsection{Why CS depts do not waive the application fee?}\label{sec:fee-waive}
 
@@
 
-\part{Application Materials}\label{part:application}
+\part{Your Application Materials}\label{part:application}
 
@@
 
-\myepigraphsimpsons{``Son, if you really want something in this life, you have to work for it. Now quiet! They’re about to announce the lottery numbers.''}
+\epigraph{Son, if you really want something in this life, you have to work for it. Now quiet! They’re about to announce the lottery numbers.}{\textsc{The Simpsons}}
 
@@
 
-\chapter{Letters of Recommendation (LoR)}\label{chap:lor}
-\chapterinfo{LORs are very important, but only if they are \textbf{personalized} and \textbf{research focused}.}
-
-\myepigraphsimpsons{``To whom it may concern… D’oh!''}
+\chapter{Letters of Recommendation (LOR)}\label{chap:lor}
+\chapterinfo{\emph{LORs are very important}, but only if they are personalized and talk about your research ability.}
 
@@
 
-Most PhD programs will require at least \emph{two} letter of recommendations (LoRs). 
-The goal of the LoRs is to provide the adcom with an assessment of your research ability from someone who knows you.
-The LoRs should focus on your research background, achievements, and potentials, all of which paint a compelling picture of you as a researcher to adcom.
-
-\section{LoR writers}\label{sec:lor-writers} \sectioninfo{LoR writers should be someone who (i) can talk in depth about your research experience and potential and (ii) have the credibility to evaluate your research ability.}
+Most PhD programs will require at least \textbf{two LORs}. 
+The goal of the LoR is to provide the adcom with an assessment of your research ability and potential. 
+The letter should 
+focus on your research background, achievements, and potentials, all of which paint a compelling picture of you as a researcher.
+
+\section{LoR writers}\label{sec:lor-writers}
 
@@
 LoR writers should also \emph{have the credibility} to convince adcom members on your research ability and potential.
-Such a person is typically someone who has a PhD, so they know what a PhD is about and so can evaluate your research ability.
+Such a person is typically someone who has a PhD (so they know what a PhD is about) and is active in research (so they can evaluate your research ability).
 LoR writers are often your research advisers, professors, or supervisors who have worked with you on research projects.
 
-\subsection{LoR from Famous People}  Having a strong letter from an internationally recognized researcher can \emph{greatly strengthen} your application.
-Such a letter can outweigh the lack of publications (or low GPA or unknown school).
-A well-known researcher has a reputation to maintain (so they won't recommend someone bad), and adcom member trust people that they know, or heard about and respect (compared to someone they don't know).
+\subsection{LoR from Famous People}  Having a strong letter from an internationally recognized researcher will \emph{greatly strengthen} your application.
+A strong LOR from such a researcher can outweigh the lack of publications (or low GPA or unknown school).
+The reason is simple: adcom members trust people that they know and respect.
+If a well-known researcher enthusiastically recommends you, then the adcom members will likely trust that recommendation.
 
@@
 
-%It is also worth noting that a LoR from a ``full'' senior professor who has not been active in research might not be as strong as a LoR from a junior faculty who has been actively mentoring you in research.
+It is also worth noting that a LoR from a ``full'' senior professor who has not been active in research might not be as strong as a LoR from a junior faculty who has been actively mentoring you in research.
 This also applies to administrative, e.g., chairs, directors, and deans (\autoref{sec:admin-letters}.) The fame of the writer is less important than how well they know you and can write about your research ability.
+
+Many students get letters from supervisors from companies where they did internships or are working.
+It is OK as long as it is a research-based personalized letter.
+Again, the emphasis here is \emph{research}, i.e., the letters should describe your research experiences and potential.
+Letters focusing on your coursework or non-research projects at a company won't carry much weight.
+
+\begin{commentbox}[Vu:]
+  If your grading system is not US standard, or you are from a good school unknown outside your country, you can ask your reference writers to explain that in their letters.
+For example, ``Bach Khoa'' are the top universities in Vietnam for STEM studies but few people outside Vietnam know about them.
+So if you are from there, you should ask your reference writers to mention that.
+\end{commentbox}
 
@@
 
-  \subsection{LoRs from Dept Chair, Dean, or Boss and Supervisor at Work}\label{sec:admin-letters}
-
-  Many students try to get LoRs from high-ranking administrators in their universities such as department chair/head, dean, or director.
+  \subsection{LoRs from Admin People}\label{sec:admin-letters}
+
+  Many international students try to get LoRs from high-ranking administrators in their universities such as department chair/head, dean, or director.
 The students never worked with these people (they might take a class or so with these profs), but mistakenly believe that these LoRs are valuable due to the writer's high position in the university.
@@
   \end{commentbox}
-
-  Many students get letters from supervisors from companies where they did internships or are working.
-It is OK as long as it is a research-based personalized letter.
-Again, the emphasis here is \emph{research}, i.e., the letters should describe your research experiences and potential.
-Letters focusing on your coursework or non-research projects at a company won't carry much weight.
 
@@
 \subsection{Self-written Letters are Bad}\label{sec:self-letters}
-
-Many profs.
-and letter writers ask students to write their own letters, which is a common practice in some countries.
-Unfortunately, such letters have \emph{little value} and are considered weak by reviewers---why can you not even find someone who cares or knows enough about you to write a candid personal reference letter?
+Many students write letters themselves, typically due to the requests of their profs.
+or letter writers.
+Such letters have \emph{little value} and are considered weak by reviewers---why can you not even find someone who cares or knows enough about you to write a candid personal reference letter?
 Instead of the ref.
@@
 
-Self-written letters are easy to spot as an experienced professor would write very differently compared to an undergraduate student.
+Adcom members are just too experienced (having reviews hundreds of those letters) to spot and suspect such letters.
+There is a very difference between a letter written by an experienced professor and a student.
 For example, they can provide convincing and concrete examples based on their experience and compare you to their own students, and of course the writing style of a professor is very different from an undergraduate student.
@@
 \section{Asking for LORs}\label{sec:asking-lor}
-\sectioninfo{Ask for LoRs at least a month before the deadline.  Waive your right to see the letter. Help your writers by providing details about your research experience and potential.}
 
 As mentioned in~\autoref{sec:lor-writers}, LoR writers should be someone who knows you well and has the credibility to evaluate your research ability.
-In the US, students usually explicitly ask if the writer would be willing to provide a strong letter, and the writer are also very direct in their response.
+You can explicitly ask if they are willing to write a strong letter for you.
 If they are not willing (you should be thankful that they are honest with you), then you should ask someone else.
 
-Below are some tips to approach your LoR writers:
-\begin{itemize}
-\item \textbf{Ask in advance} You should ask for LoRs \emph{at least a month} before the deadline.
-People are busy and writing a strong LoR takes time (\autoref{sec:lor-writing}), so give them enough time to write a strong letter for you.
-
-\item \textbf{Waive your right (\autoref{sec:waive-right})} You should always waive your right to see the letter.
+% Below are some tips to approach your LoR writers:
+% \begin{itemize}
+% \item \textbf{Ask in advance}: You should ask for LoRs \emph{at least a month} before the deadline.
+People are busy with their own work and likely have many other students asking for LoRs.
+Also, a strong LoR takes time to write, so you should give them enough time to write a strong letter for you.
+
+% \item \textbf{Help your writers}: You should provide your writers with details about the programs you are applying to, the deadlines, and any other information they need to know.
+You should also provide them with details about your research experience and potential so that they can write a strong letter for you.
+% \item \textbf{Follow up}: After you ask for LoRs, you should follow up with your writers to make sure they have submitted the letters on time.
+You should also thank them for writing the letters for you.
+% \item \textbf{Waive your right}: When you ask someone to write a letter for you, you will be asked if you want to waive your right to see the letter.
+You should always waive your right.
 This shows that you trust your writers and that you are not trying to twist their words.
-
-\item \textbf{Help your writers (\autoref{sec:help-your-LOR-writers})}: You should tell your writers the programs you are applying to, their deadlines, etc.
-You can also share your SOP with them and other details about your research experience and potential.
-
-\item \textbf{Ask for feedback} If the writer is very close to you, you can ask them for feedback on your SOP (\autoref{chap:sop}) and other application materials.
-If the writer is a professor, they might have served in adcom committees, seen many SOPs, and can provide valuable feedback.
-
-\item \textbf{Follow up and Stay in touch} Follow up with your writers to make sure they have submitted the letters on time.
-Note that their letters might have a different due date than your application (\autoref{sec:remind-writers}).
-
-After your writers have submitted all of their letters, \emph{thank them} (\autoref{sec:thank-writers}).
-Finally, let them know the outcome of your applications and stay in touch with them.
-This will help you build a relationship with them and you might need them to write for you again in the future.
-\end{itemize}
+% \item \textbf{Remind your writers}: After you enter your writers' information in the application system, you should tell your writers about that and let them know they will soon receive an email from the university to submit their letters.
+You should also remind them when you submit your application and ask them to submit their letters on time if they haven't done so.
+% \item \textbf{Help your writers}: To help your writers write a strong letter for you, you should provide them with details about your research experience and potential.
+You should also provide them with a draft of your SOP so that they can see what you are saying about yourself and complement that with their own perspective.
+% \item \textbf{Thank your writers}: After your writers have submitted their letters, you should thank them for writing the letters for you.
+You should also let them know the outcome of your applications.
+% \item \textbf{Ask for feedback}: After you have been admitted or rejected, you should ask your writers for feedback on your application.
+This will help you improve your application for future applications.
+% \item \textbf{Stay in touch}: After you have been admitted or rejected, you should stay in touch with your writers.
+This will help you build a relationship with them and will help you in the future.
+% \end{itemize}
 
@@
 
-When you ask someone to write a letter for you, you should \emph{always} waive your right.
+When you ask someone to write a letter for you, you will be asked if you want to waive your right to see the letter.  You should \emph{always} waive your right.
 
@@
 
-\subsection{Helping your LOR Writer}\label{sec:help-your-LOR-writers}
+\subsection{Help your LOR Writer}\label{sec:help-your-LOR-writers}
 
  As mentioned above in \autoref{sec:self-letters} and \autoref{sec:generic-letters}, do not write your own letter and generic letters do not give much value.
-Thus, to help your writer to write a strong, customized LoR, you can provide them details or unique things about yourself.
+Thus, to help your writer to write a strong, customized LoR, you should provide them details or unique things about yourself.
 For example, let them know about your GPA, research and work experience, papers (if any), or anything you want them to mention.
@@
 
-\begin{commentbox}[Vu:]
-    If your grading system is not US standard, or you are from a good school unknown outside your country, you can ask your reference writers to explain that in their letters.
-For example, ``Bach Khoa'' are the top universities in Vietnam for STEM studies but few people outside Vietnam know about them.
-So if you are from there, you should ask your reference writers to mention that.
-\end{commentbox}
-
-\subsection{Reminding Your Writers}\label{sec:remind-writers}
+\subsection{Remind Your Writers}\label{sec:remind-writers}
 
 After entering your writers' information in the application system, you should tell your writers about that and let them know they will soon receive an email from the university to submit their letters.
-You should also tell them when you submit your application and remind them to submit their letters on time if they haven't done so.
+You should also remind them when you submit your application and ask them to submit their letters on time if they haven't done so.
 
@@
 
-\subsection{Thank Your Writers}\label{sec:thank-writers}
-After your application is submitted and your writers have submitted their letters (i.e., the wait begins), you should send a quick thank you note, which serves both as an acknowledgement that you know they have submitted the letters and as an appreciate for their help.
-
+\paragraph{Thank Your Writers} After your application is submitted and your writers have submitted their letters (i.e., the wait begins), you should send a quick thank you note, which serves both as an acknowledgement that you know they have submitted the letters and as an appreciate for their help.
 You should also let them know the outcome of your application, regardless of whether you are admitted or not.
@@
 While I try to say something positive, e.g., the student is hardworking, or that I heard they are working on a research project, etc, the letter will still be short and weak (\autoref{sec:generic-letters}).
-Usually it takes me about \emph{5--10} minutes to write such a letter.
-
-Strong letters take a lot longer as it will be personalized.
-While I have a general template for LoR, it still can take me \emph{an hour or more} to write it.
+Usually it takes me about 5--10 minutes to write such a letter.
+
+For a student that I know well, I will write them a strong letter, which takes a lot longer as it will be personalized.
+While I have a general template for LoR, it still can take me an hour or more to write it.
 I often ask the students to provide information (\autoref{sec:help-your-LOR-writers}) and what they think I should highlight in the letter.
@@
 
-My letter always has the university letterhead, my signature, and is signed and dated.
-While using a letterhead or signing the letter is not required (I never paid much attention to these when reading LoRs), it makes the letter more official and professional.
-This website has the template of the letter with letterhead and signature: \url{https://www.overleaf.com/read/xzyrxkdjfxsp#2a1f9e}.
-
 I usually submit a letter for a student in batch (e.g., I submit to all universities that student is applying to at the same time).
@@
 \chapter{Research Experience}\label{chap:research-experience}
-\chapterinfo{Publications are not \textbf{required} but can \textbf{greatly help}. Papers in good venues are concrete evidence that you have successfully engaged in research.}
+
+\sectioninfo{\emph{Publications can greatly help}. Papers in good venues are concrete evidence that you have successfully engaged in research. But they are not required!}
 
@@
 
-\section{Publications}\label{sec:publications}
-
-The most concrete evidence of research ability is having \emph{papers in reputable international journals or conferences}.
+\section{Publications} The most concrete evidence of research ability is having \emph{papers in reputable international journals or conferences}.
 Having published papers, especially at top venues, is a sign that you have been successfully involved in research.
@@
 For example, participating in Math Olympiads if you want to do theory or winning ACM programming contests if you want to ``build'' systems, e.g., software analysis.
-So do talk about them in your SOP (\autoref{chap:sop}) and have your LoR writers mention them in their letters (\autoref{chap:lor}).
 
 \chapter{Statement of Purpose (SOP)}\label{chap:sop}
-\chapterinfo{SOP is important. Write it in such a way that makes you \textbf{stand out}.}
+\chapterinfo{\emph{SOP is important}. Write it in such a way that makes you \emph{stand out}.}
 
@@
 
-In your SOP, focus on research potential (\autoref{chap:research-experience}) and convince us through your experience, e.g., published papers (\autoref{sec:publications}).
+In your SOP, focus on research potential and convince us through your experience, e.g., published papers.
 Back up your claims with \textbf{concrete evidence}.
 For example, if you say you have experience with teaching, then show what you did, e.g., undergrad TA or mentoring someone.
-If you say you work on a research project, then show some results, e.g., paper submitted (or even rejected), achieved certain performance improvement over the state of the art.
+If you say you work on a project, then show some results, e.g., paper submitted (or even rejected), achieved certain performance improvement over the state of the art.
 
@@
 
-Finally, have your SOP reviewed by your LoR writers (\autoref{sec:help-your-LOR-writers}) and professors, especially those who have served in adcom, or even postdocs or PhD students as they have been through this process.
+Finally, your SOP reviewed by your LoR writers and professors, especially those who have served in adcom, or even postdocs or PhD students as they have been through this process.
 
@@
 One common pitfall is to write about how you are from a poor family, a small town, or a developing country.
-These are not related to DEI and show that you do not understand the topic.
-I have seen people consider themselves as minority because they are Asian or international students studying in the US (have you seen how many Asian or international students are in CS programs?).\\
+These are not related to DEI and raise concerns about you not understanding or even bother to do some basic research about the topic.
+I have seen people consider themselves as minority because they are Asian or international students studying in the US (have you seen how many Asian or international students are in CS programs in the US?).\\
 
@@
 Just show your awareness and understanding of the topic and what you did or how you plan to contribute to it (e.g., mentoring students, helping others).
-One of the best DEI statements I have read starts the acknowledgement that the applicant is not a minority, but a white male in a STEM field (i.e., they are the majority), and that they do not have much experience in DEI, but show that they are aware of the challenges faced by minority groups and how they can contribute to it.
+One of the best DEI statements I have read starts the acknowledgement that the applicant is not a minority, but a white male in a STEM field (i.e., that make them the majority), and that they do not have much experience in DEI, but show that they are aware of the challenges faced by minority groups and how they can contribute to it.
 \end{commentbox}
@@
 \chapter{Your School and Grades}\label{chap:your-school}
-\chapterinfo{High grades probably won't help much, but bad ones likely will raise concerns.}
+\chapterinfo{\emph{High grades probably won't help much} (unless you're from a very top and well-known school), but bad ones likely will raise concerns.}
 
@@
 \chapter{Standard Tests (GRE and IELTS)}\label{chap:standard-tests}
-\chapterinfo{Standard tests are \textbf{not} important.
+\chapterinfo{\emph{Standard tests are not important}.
 GRE typically \emph{is not} required.
@@
 \chapter{Interview and the Waiting Game}\label{chap:interview}
-\chapterinfo{Getting an interview is typically a good sign; but no interview does not mean rejection}
-
-\myepigraphsimpsons{Bart: ``Are we there yet? Are we there yet? Are we there yet?''\\
-Homer: ``No!''\\
-Bart: ``Okay, how about now?''}
+\chapterinfo{Getting an interview is typically a \emph{good sign}; but no interview does not mean rejection}
 
@@
 
-\paragraph{When do interviews happen?}
-
-The timeline for interviews varies.
-Faculty set up interviews based on their (busy and erratic) schedule.
-Do not be surprised if you get an interview invitation at the last minute.
-Some schools do not do interviews at all (\autoref{sec:no-interview}).
+Typically, an interview takes about 15--30 minutes, and one important aspect of evaluation is your ability to effectively communicate, including speaking and understanding English. 
+You might be asked to talk about your research experience and interests and to read a paper and discuss it.
+In some rare cases you might also be asked to solve a problem (one of my colleagues at Mason often does coding interview).
+The interview is also a good opportunity for you to ask questions, e.g., about the prof, their group, the CS PhD program, and the university.
+
+You should treat the interview as an informal chat.
+Have ``elevator pitches'' about your research experience and interests.
+You might also want to have a 5-minute presentation about your research.
+If a prof.
+asks you to read a paper, do it and be prepared to discuss it.
+You can also ask if you need to prepare for a coding problem.
+Finally, prepare some thoughtful questions to ask, e.g., about the program, the university, or the professor's research (see \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Answers-to-Ph.D-Advisor-Guide}{some questions} you can ask a potential adviser).
 
@@
 
-\paragraph{Preparing for interviews} Typically, an interview takes about 15--30 minutes, and one important aspect of evaluation is your ability to effectively communicate, including speaking and understanding English.
-You might be asked to talk about your research experience and interests and to read a paper and discuss it.
-In some rare cases you might also be asked to solve a problem (one of my colleagues at Mason often does coding interview).
-
-You should treat the interview as an informal chat.
-Prepare an \emph{``elevator pitche''} about your research experience and interests.
-You might also want to have a 5-minute presentation about your research.
-If a prof.
-asks you to read a paper, do it and be prepared to discuss it.
-You should also ask if you need to prepare for coding.
-Finally, the interview gives you an opportunity to ask questions, e.g., about the program and the professor's research.
-You should definitely ask as it shows that you're interested (see \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Answers-to-Ph.D-Advisor-Guide}{some questions} you can ask).
-
 \paragraph{Follow-Up Emails} If you had an interview and have not heard back, you can email to ask about the status of your application.
 See \autoref{sec:accept-postpone-decline} for how to check status and follow-up emails.
+
+\paragraph{When do interviews happen?}
+
+The timeline for interviews varies.
+Faculty set up interviews based on their (busy and erratic) schedule.
+Do not be surprised if you get an interview invitation at the last minute.
+Some schools do not do interviews at all (\autoref{sec:no-interview}).
 
@@
 \section{Notification Timeline: Why rejection letters are sent so late?}\label{sec:late-rejection}
-\sectioninfo{Grad programs often wait for accepted students to make their decisions, typically by April 15, before sending out rejection letters.}
+\sectioninfo{Grad programs often wait for the accepted students to make their decisions, typically by April 15, before sending out rejection letters.}
 
@@
 
-\myepigraphsimpsons{``Oh...
+\epigraph{``Oh...
 and how is education supposed to make me feel smarter?
 Besides, every time I learn something new, it pushes some old stuff out of my brain.
-Remember when I took that home wine-making course and I forgot how to drive?''}
+Remember when I took that home wine-making course and I forgot how to drive?''}{\textsc{The Simpsons}}
 
@@
 
-\chapter{Dealing with Rejection}\label{chap:not-accepted}
-
-\section{Try Again!}\label{sec:try-again}
+\section{If you do not get admitted}\label{sec:not-accepted}
 
@@
 
-\section{Why did you get rejected?}\label{sec:why-rejected}
+\subsection{Why did you get rejected?}\label{sec:why-rejected}
 \subsectioninfo{You aim too high, are overqualified, or even because you applied to AI/ML, a super competitive field in recent years with many applicants.}
@@
 Many students simply just go for very top schools (e.g., top 10) and are surprised when they are rejected to all of them, in multiple years, and completely get shut out.
-This is very obvious but many students still do this (\autoref{sec:chance-me}).
-An analogy is a person who has never hike but wants to climb Mt.
-Everest (which btw if you could, you might have a better admission chance as mentioned in \autoref{sec:improve-your-chance}).
+This is very obvious but many students still do this.
+An analogy is a person who has never hike wanting to climb Everest, which btw if you could, you might have a better admission chance (\autoref{sec:improve-your-chance}).
 
@@
 
-\section{Increasing your admission chance}\label{sec:improve-your-chance}
+\subsection{Increasing your admission chance}\label{sec:improve-your-chance}
 \subsectioninfo{You can improve your profile by being unique and standing out.}
@@
 
-\myepigraphsimpsons{``Bart, with \$10,000, we’d be millionaires! We could buy all kinds of useful things like ... love!''}
+\epigraph{Bart, with \$10,000, we’d be millionaires! We could buy all kinds of useful things like…love!’}{\textsc{The Simpsons}}
 
@@
 \subsection{How much do you cost?}\label{sec:ra-cost}
-\subsectioninfo{Your entire PhD program costs about \$400K in total, but you \emph{do not} pay for it.}
+\sectioninfo{Your entire PhD program costs about \$400K in total, but you \emph{do not} pay for it.}
 
@@
 
-\myepigraphsimpsons{``The answer to life's problems aren't at the bottom of a bottle. Heh heh! They're on TV!''}
+\epigraph{\vspace{-0.2in} The answer to life's problems aren't at the bottom of a bottle. Heh heh! They're on TV!}{\textsc{The Simpsons}}
 
@@
 
-\subsection{Schools offering PhD in CS}\label{sec:schools-offering-phd}
+\subsection{Schools that offer PhD in CS}
 
@@
 
-\part{Miscs and FAQs}\label{part:faqs}
 \chapter{Miscs and FAQs}\label{sec:faqs}
 
-\myepigraphsimpsons{``I want to share something with you – the three little sentences that will get you through life; number 1: Cover for me, number 2: Oh, good idea, Boss, and number 3: It was like that when I got here.''}
+\epigraph{``I want to share something with you – the three little sentences that will get you through life; number 1: Cover for me, number 2: Oh, good idea, Boss, and number 3: It was like that when I got here.''}{\textsc{The Simpsons}}
 
@@
 
-In the schools I've been at, the application deadline is in December, and adcom meets when school starts in mid or end of January.
-Adcom chair sends out review assignments to adcom members, about 30ish per faculty (\autoref{sec:adcom}).
-We usually have about 2 weeks to review all applications.
-As mentioned in~\autoref{sec:adcom-discuss}, adcom members review applications individually and independently.
-We only discuss when there are disagreements (adcom chair will determine which applications to discuss).
-
-I typically reserve a whole day to review all applications.
 On average, I spend about 10--15 minutes reviewing each application (less for clear rejections and more for potential acceptances).
@@
 
-I read carefully \emph{strong LoRs} and skim weaker ones, noting either strong letters from well-known professors/researchers and talk about research experiences or weak letters with generic content (e.g., ``student was in my class'').
+I read carefully \emph{strong LoRs} and skim weaker ones, noting either strong letters from well-known professors/researchers or generic content.
 I skim over \emph{CV} and look for publications, research experiences, and notable achievements.
@@
 
-I skim weak \emph{SOPs} but read strong ones carefully.
+I skim weak \emph{SOPs} but read strong ones thoroughly.
 I note whether the SOP is exciting, research-oriented, standing out, e.g., applicant is minority (\autoref{sec:urm}), and tailored to our program (e.g., if they are familiar with the work of some faculty or have talked to them).
@@
 Finally, I enter my decision, which is usually either a rejection or an offer of admission with full funding (e.g., from TA).
-I also recommend very strong candidates for the University Presidential fellowship., which is a fellowship from GMU that provides funding similar to an RA (\autoref{sec:funding}).
+I also recommend very strong candidates for the University Presidential fellowship.
 Note that while the system has other decision options, eg., admit without funding, provisional admission (e.g., if they need to take some courses), I rarely use them, simply because we want to fully fund our PhD students.
@@
 This is \emph{not necessary} and makes it look like you're trying to show off your connections.
-These nuances represent some cultural and academic differences that you may encounter and will gradually adapt to.
+These nuances represent some cultural differences that you may encounter and will gradually adapt to.
 More on cultural differences in \autoref{sec:cultural}.
@@
 
-\part{Appendices}\label{part:appendices}
 \appendix
@@
 
-\myepigraphsimpsons{``I didn’t do it. Nobody saw me do it. You can’t prove anything!''}
+\epigraph{\vspace{-0.2in} I didn’t do it. Nobody saw me do it. You can’t prove anything!}{\textsc{The Simpsons}}
 
@@
 
-\myepigraphsimpsons{``I’m not a bad guy! I work hard, and I love my kids. So why should I spend half my Sunday hearing about how I’m going to Hell?''}
+\epigraph{\vspace{-0.2in} I’m not a bad guy! I work hard, and I love my kids. So why should I spend half my Sunday hearing about how I’m going to Hell?}{\textsc{The Simpsons}}
 
@@
 \paragraph{School \autoref{chap:your-school}} Adcom already knows about your school, which is a plus (see \autoref{chap:your-school}).
-You are also more familiar with the US education system and academic culture (\autoref{sec:cultural}).
+You are also more familiar with the US education system and culture (so hopefully you are aware of the cultural issues listed in \autoref{sec:cultural}).
 
@@
 
-\myepigraphsimpsons{``Kids, you tried your best and you failed miserably. The lesson is: never try.''}
+\epigraph{\vspace{-0.2in} Kids, you tried your best and you failed miserably. The lesson is: never try.}{\textsc{The Simpsons}}
 
@@
 
-\chapter{Academic and Cultural Differences}\label{sec:cultural}
-
-\myepigraphsimpsons{``Ay caramba!''}
-
-This section lists some general academic and cultural issues that students, especially international ones, might want to pay attention to.
+\chapter{Cultural Differences}\label{sec:cultural}
+
+\epigraph{\vspace{-0.2in} Ay caramba!}{\textsc{The Simpsons}}
+
+This section lists some general cultural issues that students, especially international ones, might want to pay attention to.
 
@@
 
-\myepigraphsimpsons{``The whole damn system is wrong!''}
+\epigraph{\vspace{-0.2in} The whole damn system is wrong!}{\textsc{The Simpsons}}
 

### 2024-12-21 demystify.tex

--- 2024-12-21 demystify.tex
@@
 \newcommand*\l@subsectioninfo{\@nodottedtocline{2}{3.8em}{3.2em}}
-\newcommand*\l@subsubsectioninfo{\@nodottedtocline{3}{7.0em}{4.1em}}
-\newcommand*\l@paragraphinfo{\@nodottedtocline{4}{10em}{5em}}
+%\newcommand*\l@subsubsectioninfo{\@nodottedtocline{3}{7.0em}{4.1em}}
+%\newcommand*\l@paragraphinfo{\@nodottedtocline{4}{10em}{5em}}
 %\newcommand*\l@subparagraphinfo{\@nodottedtocline{5}{12em}{6em}}line{5}{12em}{6em}}

### 2024-12-22 demystify.tex
updates

@@
 
-However, if the student has contacted a faculty member and that faculty is interested in the student and has made this known to the admission committee, then the student is likely to be admitted, \emph{even if the faculty does not review the application or has funding} (thus the benefit of contacting faculty as discussed in\autoref{sec:contact}).
+However, if the student has contacted a faculty member and that faculty is interested in the student and has made this known to the admission committee, then the student is likely to be admitted, \emph{even if the faculty does not review the application or has funding} (thus the benefit of contacting faculty as discussed in~\autoref{sec:contact}).
+
 Similarly, if the student mentioned a faculty in their SOP, adcom might ask that faculty to look at their application and if they are interested in the student.
@@
 
-However, in some rare cases we do talk to each other.
-For example, when there are discrepancies in evaluations, in which case the adcom chair will ask reviewers to discuss the application to reach a consensus.
-We also talk to each other for interesting or strong applications, e.g., how to recruit this student or who should be the adviser.
+Not really, we typically review applications independently and do not talk to each other.
+This is to avoid biasing, e.g., if one reviewer says they want to accept, the other might feel pressured to accept as well.
+
+However, when there are discrepancies in evaluations, the adcom chair will ask reviewers to discuss the application to reach a consensus.
+We might also talk to each other for interesting or strong applications, e.g., how to recruit this student or who should be the adviser.
 If the student mentioned a faculty member in their SOP, we might ask that faculty if they are interested in the student.
+
+Note that other disciplines might have different practices, e.g., they might select a top list of applicants and then discuss them in a meeting to determine who to interview.
+In CS, both the reviews and interviews are often done independently (\autoref{sec:interviews}).
+
+\begin{commentbox}[A prof. in Physics at an R1 university:]
+  We have a pretty well fleshed out grading rubric for applications that has categories like grades, research, writing ability, etc.
+I would say our rubric is weighted about 1/2 on academics (research, LORs, grades) and 1/2 on the idea of ``grit'' or ``resilience'' (engagement, leadership, working through obstacles).\\
+
+  The rubric helps a lot to standardize how committee members grade, and speeds things up a bit because you know what to look for.
+We spent what seemed like forever on the details of the grading system (e.g., what does a a score of `3' vs a `2' in writing look like?) but now it's very helpful.\\
+
+  We also do roughly three rounds of selection: a first ``triage'' round to determine the top ~100 applications, a second round to determine about 25 people to interview, and then a third round to decide the actual offers.
+That also helps to speed things up a bit, since in the first round with all the applications you can move fairly quickly since you just need to sort into ``good'' and ``bad''.
+By the time we're getting into the details and reading everything more closely in rounds 2 and 3 most of the applications have been removed from consideration.
+So for this method I do maybe 5 minutes per app in round 1, but closer to 20 minutes per app in round 2, and usually round 3 is long discussions about specific people.
+\end{commentbox}
 
@@
 
-Sometimes students would go to great lengths just to get letters from well-known professors in their school (e.g., department head or dean).
+Sometimes students would go to great lengths just to get letters from ``top'' professors in their school (e.g., department head or dean).
 But as mentioned, if these professors do not know you, their letters would likely be generic and carry little value (sometimes \red{red flags}).
@@
 In many cases your LOR writers \emph{are not given any deadline}.
-The following lists several LOR invitation emails I received from various universities in the past few years:
-
+\autoref{fig:fig:lor-invitations} lists several LOR invitation emails I received from various universities in the past few years.
+
+\begin{figure}
 \begin{emphasisbox}
@@
 \end{emphasisbox}
+\caption{Examples of LOR invitation emails}\label{fig:lor-invitations}
+\end{figure}
 
@@
 \end{commentbox}
+
 \paragraph{What if I don't have any pubs} It is understandable that many students do not have the opportunity to publish papers.
 Thus, other writings, even those under submissions or rejected, would still be good and better than nothing.
-Be sure to upload your papers with your application (\autoref{sec:writing-sample}) and talk about them in your SOP (\autoref{chap:sop}).
+Be sure to upload your these with your application (\autoref{sec:writing-sample}) and talk about them in your SOP (\autoref{chap:sop}).
 
@@
 \paragraph{I am not the first author} Being the first author is good because it indicates you own the work.
-However, it's perfectly OK to be second or third or last.
-It is difficult to publish a good paper and so being a co-author is still a good sign about your research experience.
-In any case (and especially in the case when you're not the first author), you should explain the work and your contribution.
-Better yet, have your LoR writers talk about your contribution in their letters.
+However, it's \emph{perfectly OK to be second or third or even last}.
+Adcom members know it is difficult to publish a good paper, and so being a co-author is still a good sign about your research experience.
+In any case, especially in the case when you're not the first author, you should explain the work and your contribution.
+Better yet, have your LoR writers (\autoref{sec:help-your-LOR-writers}) talk about your work and contribution in their letters.
 
@@
 
-Note that adcom reviewers might not be familiar with all research labs, especially those outside the US For example, while VinAI is  well-known in Vietnam and potentially in Asia, it might not be very well-known in the US.
+Note that adcom reviewers might not be familiar with all research labs, especially those outside the US.
+For example, while VinAI is  well-known in Vietnam and potentially in Asia, it might not be very well-known in the US.
 So you or your LoR writers should explicitly say something about them in your statement.
@@
 For example, participating in Math Olympiads if you want to do theory or winning ACM programming contests if you want to ``build'' systems, e.g., software analysis.  
-So do talk about them in your SOP (\autoref{chap:sop}) and have your LoR writers mention them in their letters (\autoref{chap:lor}).
+So do talk about them in your SOP (\autoref{chap:sop}) and have your LoR writers mention them in their letters (\autoref{sec:help-your-LOR-writers}).
 
@@
 
-\section{Interviews}
+\section{Interviews}\label{sec:interviews}
 
@@
 
-The timeline for interviews varies.
-Faculty set up interviews based on their (busy and erratic) schedule.
+The timeline for interviews varies.  Faculty set up interviews based on their (busy and erratic) schedule. Some try to get interviews done before the holidays, while others do them after the holidays.
 Do not be surprised if you get an interview invitation at the last minute.
-Some schools do not do interviews at all (\autoref{sec:no-interview}).
+Some profs.
+are informal and may just email you to chat (e.g., \emph{``could you chat in an hr?''}), while others might schedule a formal interview (e.g., \emph{``can you chat at 2 pm on Friday or 10 am on Monday?''}).
+
+Some schools do not do interviews at all (\autoref{sec:no-interview}). They review applications and make decisions based on them.  If you do not get an interview, it does not mean you're out.
 
@@
 
+\myepigraphsimpsons{``It’s not easy being a mentor. But it sure is funny to watch other people screw up.''}
+
 There is no one-size-fits-all answer to finding an adviser. The best adviser is the one that you can work well with.  But how do you find such a person?
@@
 
-\section{Contacting a prof. How to get a desired reply?}\label{sec:contact}
+\section{Contacting a Prof. How to get a desired reply?}\label{sec:contact}
 
@@
 
-    \item \textbf{Don't CALL}.
-Not related to email but sometimes students get desperate and call the prof.
-This is a big no-no, especially for CS ppl who often prefer email over phone calls.
-Seriously though, don't call any prof.
-unless they ask you to.
+    \item \textbf{Do Not Call}. Not related to email but sometimes students get desperate and call the prof.  This is a big no-no, especially for CS ppl who often prefer email over phone calls. 
 \end{itemize}
@@
 
-\item \textbf{I am not taking students but I co-advise/can serve on your committee} While this might sound good, it's generic because it says once you're admitted and have an adviser, then contact me again.
+\item \textbf{I am not taking students but I co-advise/can serve on your committee.} While this might sound good, it's still generic because it says \emph{once you're admitted and have an adviser}, then contact me again.
 
@@
 
-In the schools I've been at, the application deadline is in December, and adcom meets when school starts in mid or end of January.
+In the schools I've been at, the application deadline is in Dec, and adcom meets when school starts in mid or end of Jan.
 Adcom chair sends out review assignments to adcom members, about 30ish per faculty (\autoref{sec:adcom}).
@@
 
-I typically reserve a whole day to review all applications.
+I typically reserve a whole day (or two days) to review all applications.
 On average, I spend about 10--15 minutes reviewing each application (less for clear rejections and more for potential acceptances).
@@
 I also recommend very strong candidates for the University Presidential fellowship., which is a fellowship from GMU that provides funding similar to an RA (\autoref{sec:funding}).
-Note that while the system has other decision options, eg., admit without funding, provisional admission (e.g., if they need to take some courses), I rarely use them, simply because we want to fully fund our PhD students.
+Note that while the system has other decision options, eg., admit without funding, provisional admission (e.g., if they need to take some courses), I do not use them, simply because we either reject or admit with funding.
+
+Note that I \emph{do not} need to interview a student to make a decision.
+I can tell from the application if they are strong or not.
+However, if I want to recruit a student, I will ask them to chat with me.
+This is quite different than in other discipline where reviews, interviews, and selection are done in multiple rounds (see~\autoref{sec:adcom-discuss} for an example in Physics).
+
+Of course my recommendation is just one of the three or four faculty who review the application.
+The adcom chair will compile all recommendations and make a decision based on them (\autoref{sec:adcom-discuss}).
+If there are disagreements, adcom chair will ask the reviewers to discuss the application. 
+Unfortunately, even if all revierwers recommend a student, they might not be admitted (\autoref{sec:why-rejected}) if there are too many students or they are not a good fit for the program (e.g., no one is willing to advise them).
 
@@
 
-\begin{commentbox}[DK:]
+\begin{commentbox}[Kapur:]
   I was amused to read this as if I recall correctly, you never called me by my first name when you were at UNM. You always called me Prof. And, many times, I would jokingly call you back as Prof. Vu.
@@
 
-Do not include ranking (e.g., Assistant, Associate, Scientist, ...) when referring to someone.
+Do not include ranking (\autoref{sec:faculty-types}), e.g., Assistant, Associate, Scientist, ..., when referring to someone.
 I see many international students include a lengthy title of people they know, e.g., \emph{I am advised by Asst.
@@
 X, and I also collaborate with Distinguished Scientist Y}.
+
 This is \emph{not necessary} and makes it look like you're trying to show off your connections.
 These nuances represent some cultural and academic differences that you may encounter and will gradually adapt to.
-More on cultural differences in \autoref{sec:cultural}.
+More on these differences in \autoref{sec:cultural}.
 

### 2024-12-22 demystify.tex
updates

@@
 
-\myepigraphsimpsons{``Ay caramba!''}
+\myepigraphsimpsons{``As intelligent as you are, you have to learn to appreciate other points of view.''}
 
@@
 
-Here are the typical steps: (i) a faculty suspecting a cheating case will report it to the Office of Academic Integrity (OAI) at the university---the report often has supporting evidence and suggested penalty (e.g., a failing grade);  (ii)  OAI will completely take over and investigate the case independently (the prof.
-is not involved); and (iii) OAI will make the final decision.
-It is important to note that after receiving the report from your prof., OAI \emph{completely} takes over and makes its own decision.
-This means begging your professor will not help because they simply are no longer involved in the case and cannot do much.
+Here are the typical steps: (i) a faculty suspecting a cheating case will report it to the Office of Academic Integrity (OAI) at the university---the report often has supporting evidence, e.g., a copy of the exam or assignment; (ii) OAI will investigate the case, e.g., by interviewing the student and faculty, and checking the evidence; (iii) OAI will make a decision, e.g., whether the student is guilty or not; and (iv) if guilty, OAI will decide on the punishment, e.g., a failing grade for the assignment or exam, a failing grade for the course, or even expulsion from the university, especially for repeat offenders.
+
+So after receiving the report from your prof., OAI \emph{completely} takes over and makes its decision.
+This means begging your professor will not help because they simply are no longer involved in the case and cannot do anything.
 

### 2024-12-24 demystify.tex
Updates from Overleaf

@@
 
-\chapterinfo{Yes, definitely.  CS PhD study in the US is often fully funded and admission into good universities in the US is not any harder than in other countries.}
+\chapterinfo{Yes, definitely.  CS PhD study in the US is often \textbf{fully funded} and admission into good universities in the US is \textbf{not} any harder than in other countries.}
 
@@
 
-\paragraph{Not any harder than other countries} First, applying to a good US university \emph{should not} be any harder than at schools in other countries.
+\paragraph*{Not any harder than other countries} First, applying to a good US university \emph{should not} be any harder than at schools in other countries.
 It might even be more flexible since CS PhD in the US \emph{do not} require having an MS or a research topic, proposal, or adviser in advance (\autoref{sec:non-us-differences} compares CS PhD study in the US to other countries).
@@
 
-\paragraph{Core CS topics} Common CS knowledge that you should know are:
+\paragraph*{Core CS topics} Common CS knowledge that you should know are:
+
 \begin{itemize}

### 2024-12-24 demystify.tex
Updates from Overleaf

@@
 
-\paragraph*{Not any harder than other countries} First, applying to a good US university \emph{should not} be any harder than at schools in other countries.
+\paragraph{Not any harder than other countries} First, applying to a good US university \emph{should not} be any harder than at schools in other countries.
 It might even be more flexible since CS PhD in the US \emph{do not} require having an MS or a research topic, proposal, or adviser in advance (\autoref{sec:non-us-differences} compares CS PhD study in the US to other countries).

### 2024-12-26 demystify.tex
Updates from Overleaf
updates

@@
 In many cases your LOR writers \emph{are not given any deadline}.
-\autoref{fig:fig:lor-invitations} lists several LOR invitation emails I received from various universities in the past few years.
+\autoref{fig:lor-invitations} lists several LOR invitation emails I received from various universities in the past few years.
 
@@
     \textbf{Sample 1:}
-
     .. is applying for admission into the Computer Science (Ph.D.) program in the Graduate School at the University of Massachusetts Amherst and has listed you as a reference.
@@
     \textbf{Sample 2:}
-
     Dear ..
@@
     \textbf{Sample 3:}
-
     .. is applying to the University of Nebraska-Lincoln and has named you as a recommender. \textbf{This applicant has waived the right to view your recommendation.}
@@
     \textbf{Sample 2:}
-
     Dear Recommender,
@@
 
-Many international students mistakenly believe that their school is well-known, but in fact, it is not, e.g., many Vietnamese students believe that ``Bach Khoa'' is well-known, but it is not (and their confusing acronyms HUST and HCMUT only make it worse).
+The reason is similar to LoRs (\autoref{sec:ievaluate}).
+If we know your school has a good reputation, then we can trust the education and grades from your school.
+If we do not know your school, then we are uncertain about its quality and the grades you received.
+
+Many international students mistakenly believe that their school is well-known, but in fact, it is not, e.g., many Vietnamese students believe that ``Bach Khoa'', which is one of the best universityies in the country, is well-known, but it is not (and their confusing acronyms HUST and HCMUT only make it worse).
 So if you think your school is well-known, mention it (and back up with some evidence like rankings or awards) in your SOP.
@@
 
-\paragraph{Locally} Start looking for research opportunities at your institution.
+\section{Locally}
+
+Start looking for research opportunities at your institution.
 If you did well or liked a class, you can check with the professor of that class for research opportunities.
@@
 
-There might be also special programs for undergraduate research from the university, e.g., Mason has the OSCAR program, UNL has UCARE, and NSF has REUs (for US citizens and permanent residents).
+Many universities have programs to encourage undergraduate research. For example, Mason has the OSCAR program, UNL has UCARE, and NSF has REUs (for US citizens and permanent residents).
 
@@
 
-\paragraph{Online Research Communities and Open Source Contributions} Places such as GitHub offer many research opportunities, even though they might not appear so (e.g., they do not explicitly mention research).
+\section{Open Source Contributions}
+
+Places such as GitHub offer many research opportunities, even though they might not appear so (e.g., they do not explicitly mention research).
 In many cases, professors or research labs put their projects on GitHub.
@@
 
-\paragraph{Virtual Research Opportunities} This is less common but several places offer virtual internships and research programs aimed at providing hands-on research experience.
+\section{Virtual Research Opportunities} 
+This is less common but several places offer virtual internships and research programs aimed at providing hands-on research experience.
 These programs often involve working remotely under the guidance of experienced mentors and collaborating with a team of fellow researchers.
@@
 
-\begin{table}[h]
+\begin{table}
   \centering

### 2024-12-26 demystify.tex
updates

@@
 So I want to share some details about the admission process and advice for those who are interested in applying for a \textbf{PhD in Computer Science in the US}.
-Originally, this document was intended for \emph{international students}, but has been expanded to include information for \emph{US domestic students} (\autoref{sec:domestic-students}).
+Originally, this book was intended for \emph{international students}, but has been expanded to include information for \emph{US domestic students} (\autoref{sec:domestic-students}).
 Moreover, while this primarily aims at students interested in CS, it might be relevant to students from various STEM (Science, Technologies, Engineering, and Mathematics) disciplines.
-Furthermore, although many examples are specifics for schools that I and other contributors of this document know about, the information should be generalizable to other good R1\footnote{An \href{https://en.wikipedia.org/wiki/List_of_research_universities_in_the_United_States}{R1 institution} in the US is a research-intensive university with a high level of research activity across various disciplines.
+Furthermore, although many examples are specifics for schools that I and other contributors of this book know about, the information should be generalizable to other good R1\footnote{An \href{https://en.wikipedia.org/wiki/List_of_research_universities_in_the_United_States}{R1 institution} in the US is a research-intensive university with a high level of research activity across various disciplines.
 Currently, 146 (out of 4000) US universities are classified as R1.} institutions in the US.
@@
 \begin{mybox}
-This document is open source and updated regularly. Its latest version is available at
+This book is open source and updated regularly. Its latest version is available at
 
@@
 Much of these uncertainties stem from the lack of information and guidance on the admission process in the US (social places like Facebook and Reddit are full of confusion and contradictory information).
-This document aims to address these concerns.
+This book aims to address these concerns.
 
@@
 
-\section{Who is this document for?}
-\sectioninfo{This document is intended for international students from smaller countries and domestic students from less well-known universities.}
-
-While this guideline can be useful for all students, it would most benefit \textbf{international students} from \textbf{smaller countries and less well-known universities} (or domestic students from US universities with no PhD programs or limited research opportunities).
+\section{Who is this handbook for?}
+\sectioninfo{This book is intended for international students from smaller countries and domestic students from less well-known universities.}
+
+While this book can be useful for all students, it would most benefit \textbf{international students} from \textbf{smaller countries and less well-known universities} (or domestic students from US universities with no PhD programs or limited research opportunities).
 
@@
 
-This document also aim to explain the \emph{``why''} behind the admission process, e.g., why LoRs are important, why you should not draft your own LoR, why you should contact professors, etc.
+This book also aim to explain the \emph{``why''} behind the admission process, e.g., why LoRs are important, why you should not draft your own LoR, why you should contact professors, etc.
 There are numerous \href{https://github.com/nguyenthanhvuh/phd-cs-us}{resources online} from students  that tell you \emph{``what''} to do, but few explain the \emph{``why''}.
@@
 
-While this document may not fully meet the needs of students aiming for top-10 schools (\autoref{sec:selecting-ranking-schools} and \autoref{sec:ranking}), which typically have more stringent requirements (e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU), it should be valuable for those targeting other reputable R1 institutions, which are frankly the majority of good schools in the US.
+While this handbook may not fully meet the needs of students aiming for top-10 schools (\autoref{sec:selecting-ranking-schools} and \autoref{sec:ranking}), which typically have more stringent requirements (e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU), it should be valuable for those targeting other reputable R1 institutions, which are frankly the majority of good schools in the US.
 
@@
 
-While being ambitious and aiming high is good, you should understand how PhD admission works (e.g., by reading \emph{this} document and realizing things such as having a good GPA or GRE doesn't mean much to top programs), that in the US there are many good schools, and just be realistic.
+While being ambitious and aiming high is good, you should understand how PhD admission works (e.g., by reading \emph{this} handbook and realizing things such as having a good GPA or GRE doesn't mean much to top programs), that in the US there are many good schools, and just be realistic.
 
@@
 
-Most of what is written in this document applies to both domestic\footnote{For simplicity domestic means you did your undergrad (or MS) at a US university.
+Most of what is written in this handbook applies to both domestic\footnote{For simplicity domestic means you did your undergrad (or MS) at a US university.
 Different universities might have different definitions, e.g., permanent residents or US citizens.} and international students.
@@
 \chapter{History and Acknowledgement}\label{sec:ack}
-\paragraph{History} This document was conceived during a lunch with a faculty at Mason.
+\paragraph{History} This handbook was conceived during a lunch with a faculty at Mason.
 We talked about why Mason was not able to attract good Vietnamese and other international students, despite having a much stronger CS program than many schools that these students want to go to (part of the reason is described in \autoref{sec:selecting-ranking-schools}).
@@
 However, most participants are students in non-CS fields or not in the US, and their answers are not always accurate and can lead to more confusion.
-So I thought it would be useful to have a document that is specific to CS PhD programs in the US from an insider perspective.
-
-I started writing this document in May 2023 and have been updating it since then (mostly around deadline time when I tend to procrastinate, i.e., \emph{productive procrastination}!).
-The document was initially intended for international students but has been generalized to all students.
-The document is open-source and available on \href{https://github.com/nguyenthanhvuh/phd-cs-us}{GitHub} so that others can contribute to it.
+So I thought it would be useful to have a handbook that is specific to CS PhD programs in the US from an insider perspective.
+
+I started writing this book in May 2023 and have been updating it since then (mostly around deadline time when I tend to procrastinate, i.e., \emph{productive procrastination}!).
+The book was initially intended for international students but has been generalized to all students.
+The handbook is open-source and available on \href{https://github.com/nguyenthanhvuh/phd-cs-us}{GitHub} so that others can contribute to it.
 \paragraph{About Me} I am an associate professor in the CS dept at George Mason University (Mason).
@@
 I also mentor students from Vietnam and have close colleagues and friends who were once international students.
-I hope to capture the diverse challenges and experiences they've faced in this document so that it can be a valuable resource for prospective international students.
+I hope to capture the diverse challenges and experiences they've faced in this book so that it can be a valuable resource for prospective international students.
 Finally, my upbringing in the US provides a perspective aligned with American culture, allowing me to shed light on various issues, particularly those related to cultural differences (\autoref{sec:cultural}).
@@
 
-Also thanks to NSF for encouraging faculty to be creative in research and education, which allows me to write this document.
+Also thanks to NSF for encouraging faculty to be creative in research and education, which allows me to write this handbook.
 

### 2024-12-26 demystify.tex
updates

@@
 \appendix
-\chapter{Glossary}\label{sec:glossary}
+\chapter{Glossary}\label{chap:glossary}
 
@@
   \item[Field or Area] A specific research area, e.g., Machine Learning, Computer Vision, Software Engineering, etc.
+
+  \item[International vs.
+Domestic students] International students are loosely defined as those who would need a visa to study in the US (\autoref{chap:visa}).
+Domestic students (\autoref{sec:domestic-students}) are those who do not need a visa, typically US citizens or permanent residents.
 
@@
 
-\chapter{Visa for International Students}\label{sec:visa}
+\chapter{Visa for International Students}\label{chap:visa}
 

### 2024-12-27 demystify.tex
Updates from Overleaf

@@
 
-  \item[Domestic Student] A student who is typically a US citizen or permanent resident (``greencard holder'').
-        But sometimes it can also refer to students who did their undergrad (or MS) at a US university.
-
   \item[Field or Area] A specific research area, e.g., Machine Learning, Computer Vision, Software Engineering, etc.
@@
 
-For international students pursuing a PhD in the US, \textbf{F-1} is the main visa that allows a student to study full-time at an accredited institution.  Here are some key points about the F-1 visa:
+As defined in~\autoref{chap:glossary}, international students are those who would need a visa to study in the US.
+In most cases, \textbf{F-1} is the main visa needed to study full-time at an accredited institution.
+Here are some key points about F-1:
 
@@
 
-Most of what is written in this handbook applies to both domestic\footnote{For simplicity domestic means you did your undergrad (or MS) at a US university.
-Different universities might have different definitions, e.g., permanent residents or US citizens.} and international students.
+Most of what is written in this handbook applies to both domestic\footnote{As mentioned in \autoref{chap:glossary}, domestic means you do not require a visa to study in the US.} and international students.
 However, there are some differences and benefits that domestic students should be aware of and can leverage to improve their chances of admission.

### 2024-12-29 demystify.tex
Updates from Overleaf

@@
 
-\section{CV/Resume}
+\section{CV}
 \sectioninfo{Highlight and summarize major achievements, e.g., Publications, Programming Awards}
 
-The CV/Resume provides a summary of the applicant's achievements. Most schools require you to upload your CV with your application.
-Prepare your CV in such a way that it allows the reviewers to quickly scan to identify major achievements, e.g., Publications, Programming Competition Awards, and Teaching Experience.
+The CV provides a summary of the applicant's achievements. Most schools require you to upload your CV with your application.
+
+Prepare your CV in such a way that allows reviewers to quickly scan to identify major achievements, e.g., Publications, Programming Competition Awards, and Teaching Experience.
+Moreover, you don't need to adhere to the 1-page rule of resume in job searching, CV has no such requirement and is often longer.
 
@@
 Moreover, we care and pay attention more about your LOR, SOP, etc.
-So do not spend too much time on your CV, just make sure it is easy to skim through and well-organized around research activities and achievements.
+So do not spend too much time on your CV, just make sure it is easy to skim through (\autoref{sec:ievaluate}) and well-organized around research activities and achievements.
 \end{commentbox}
@@
 
-After you apply, you might get interviews.
+After you apply, you \emph{might} get interviews.
 The most common case is that a prof.
-is interested in working with you and wants to chat with you, e.g., to offer RA (\autoref{sec:ra}).
+is interested in working with you and wants to chat, e.g., to offer RA (\autoref{sec:ra}).
 In some cases, the interview is done by several professors, e.g., to see if a student fits in their group or to recruit a very strong student to their program.
@@
 
-The timeline for interviews varies.  Faculty set up interviews based on their (busy and erratic) schedule. Some try to get interviews done before the holidays, while others do them after the holidays.
+The timeline for interviews varies.  Faculty set up interviews based on their busy and erratic schedule. Some try to get interviews done before the holidays, while others do them after the holidays.
 Do not be surprised if you get an interview invitation at the last minute.
@@
 
-Some schools do not do interviews at all (\autoref{sec:no-interview}). They review applications and make decisions based on them.  If you do not get an interview, it does not mean you're out.
+Some programs \textbf{do not do interviews} at all (\autoref{sec:no-interview}).
+They review applications and make decisions based on them.
+If you do not get an interview, it does not mean you're out.
 
@@
 Finally, the interview gives you an opportunity to ask questions, e.g., about the program and the professor's research.
-You should definitely ask as it shows that you're interested (see \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Answers-to-Ph.D-Advisor-Guide}{some questions} you can ask).
+You should definitely ask as it shows that you're interested.
+See \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Answers-to-Ph.D-Advisor-Guide}{some questions} you can ask.
 
@@
 
-Some schools send out admission letters in batches, some do rolling admission, and some do not send anything out (e.g., you're on their waitlist).
-You should hear back from most schools by mid-March (though rejection can come out very late).
+Some schools send out admission letters in batches, some do rolling admission, and some do not send anything out e.g., you're on their waitlist.
+You should hear back from most schools by mid-March, though rejection can come out much later.
 
@@
 \paragraph{Response Deadlines} Accepted students are usually given a deadline to make decisions on their offers, often around \textbf{April 15th} in the US.
-This is a standard deadline for PhD programs due to the Council of Graduate Schools (CGS) Resolution (\href{https://cgsnet.org/wp-content/uploads/2024/01/CGS_April15_Resolution_Jan312024.pdf}{see more}.
+This is a standard deadline for PhD programs due to the Council of Graduate Schools Resolution (\href{https://cgsnet.org/wp-content/uploads/2024/01/CGS_April15_Resolution_Jan312024.pdf}{CGS}).
 After this date, CS programs can gauge how many slots remain unfilled.
@@
 You can try to contact the school to ask about your status, but they might not reply, they might say they are still reviewing applications, or give you inaccurate information (e.g., you will hear in two weeks).
-In short, you just have to be patient and wait (and also beware that some schools do not send out rejection letters at all).
+In short, you just have to be patient and wait, and also beware that some schools do not send out rejection letters at all.
 

### 2024-12-29 demystify.tex
updates

--- 2024-12-26 demystify.tex
@@
 
-% \chapter{Fellowship Applications}\label{sec:fellowship}
-% \chapterinfo{How to apply for fellowships such as NSF GRFP, DOD DSEG to fund your PhD.}
+\chapter{Fellowship Applications}\label{sec:fellowship}
+\chapterinfo{How to apply for fellowships such as NSF GRFP, DOD DSEG to fund your PhD.}
 
@@
 
-% \paragraph{Evaluating the application's research potential, \emph{not} the project potential} This is very different than a normal proposal like an NSF one that asks for funding for a specific project.
-In such case the project is evaluated.
-However, in the GRFP, the research potential of the applicant is evaluated, not the project.
-This means that you should focus on your research experience, your potential to do research, and how you can contribute to the field, etc.
-You should also focus on your broader impacts, e.g.,  how you can help improve STEM education through outreach and mentoring, how you can contribute to diversity and inclusion in STEM fields, how you can help society, etc.
+\section{How Applications are Evaluated}
+
+\begin{itemize}
+\item Each applicant is assigned to ~3 reviewers who have expertise in the applicant's research topics (so only your field but also your specific area).
+This means reviewers can be quite knowledgeable in your field and your work, your LoR writers and their work, and thus can evaluate your application quite well.
+
+\item While we review applicants from various levels (e.g., undergrad, 1st year, 2nd year in PhD program), we evaluate and compare applicants \emph{within} their cohort level.
+This means that we do not compare an undergrad, who is still applying to grad school, to a 2nd year PhD student who already has an adviser and published multiple top papers.
+This ensures that each applicant is evaluated fairly based on their level.
+
+\end{itemize}
+
+\section{Intellectual Merit and Broader Impacts}
+Your will be evaluated based on two criteria: Intellectual Merit (IM) and Broader Impacts (BI).
+
+\paragraph{Intellectual merit} is based on your research potential, e.g., your research experience, your potential to do research, and how you can contribute to the field.
+
+Note that unlike a normal project or research proposal, in which the evaluation is based on the project, in the GRFP, the evaluation is \emph{not} about the research plan.
+We are repeatedly reminded to not focus too much on the research plan as many applicants are still applying for PhD programs and might not even get into the field or work with a prof.
+that fit their research plan.
+Instead, the evaluation is based on your \emph{potential} of being a good researcher using various criteria, e.g., your research experience, what others think of you (LoRs), research plan, etc.
+This also means that even if you have very little research experience, you can still get the fellowship if you can convince the reviewers that you have the potential to be a good researcher.
+Vice versa, if you have a lot of research experience and in very top places but cannot convince the reviewers that you have the potential to be a good researcher, you will not get the fellowship.
+
+\paragraph{Broader impacts} based on how you and your research can benefit society.
+Many students just mention how their research is cross-discipline and thus has broader impacts (e.g., improve software quality and thus help society).
+This is too common and not enough.
+You also want to talk about how you, as a grad student with your research, can help society, e.g., through outreach, mentoring, diversity, etc.
+As mentioned below, you will want to be specific and concrete, e.g., ``I will work with highschool students and underrepresented students'' is not as convincing as ``I have had experience in mentoring highschool students X and will continue to do so through Y,Z ...''.
+
+\section{Common Pitfalls}
+
+\paragraph{Too technical} Remember that you're writing to a general audience in CS, not experts in your field.
+While NSF tries to match reviewers with your field, CS fields are simply too broad and you might get a reviewer in the field but not in your specific area.
+So avoid technical jargon and explain things in simple terms. If the reviewer does not understand your proposal, they will likely give you a low score.
+
+You should ask your professor or someone who's familiar with writing NSF proposals to review your reseach plan.  They can tell you if it is too technical and help you simplify it.
+
+\paragraph{Preliminary Work} Your research plan should have some preliminary work to convince reviewers that you have thought about the problem and have some initial promising results.
+It does not have to a lot, e.g., you don't need published results or an implementation, a small experiment done by hand on some small examples would suffice.
+Moreover, it should be concrete and convincing, e.g., \emph{``When being applied to the program in Fig 1.
+of the paper published PLDI'19 that approach X failed, my idea worked and was able to ...''}.
+This will set you apart from others who just have a plan but no results.
+
+% Prelim work does not, and in fact should not, be extensive.
+You do not need a paper or a complete implementation.
+You don't need to show that your idea is better than others or even that it will work.
+You just need to show that you have tried it and it has some promise.
+This can be as simple as a small experiment done by hand on some small examples,  e.g., if you come up with a new abstraction in PL for proving correctness, then you can just manually apply it to a few small programs, preferrably ones that existing works have trouble with.
+This is enough to show that you have thought about the problem and have some initial results.
+
+% To make prelim concrete (and thus more convincing), you might want to use specific names and numbers, e.g.,  \emph{``I tried my prototype on the MNIST dataset, achieving an accuracy of 90\%''}, or \emph{``I used Lean to formalize and prove a specific case of the Fermat's Last Theorem''}.
+In addition, negative results are results and can be very convincing.
+So don't be afraid to say \emph{``I tried my idea on X but it failed because of Y, and I plan to address this by Z''}.
+
+\paragraph{Research} is the emphasis.
+Many made a mistake of focusing on teaching and mentoring and not on research.
+This is especially true for students who have a lot of teaching and mentoring experience, e.g., from a teaching-focused university or a teaching assistant.
+While teaching and mentoring are important, the GRFP is a \emph{research} fellowship which aims to train researchers.
+So you will need to demonstrate your passion for research and your potential to be a good researcher.
 

### 2024-12-29 demystify.tex
updates

@@
 
-\section{GREs Are Optional and Do Not Matter} While a few schools still require taking the GRE (e.g., UCF), most good CS PhD programs in the US \textbf{no longer} require it.
-Thus, if you haven't taken the GRE or have bad scores, then don't waste time (re)taking it.
-Being optional really means optional, and not taking it will not hurt your application.
-
-However, if you took it and have really good scores then it might be worth it to include and talk about them in your application, but don't expect them to make much difference (many adcom reviewers never took it).
+\section{GREs Are Optional and Do Not Matter}\label{sec:gre}
+
+While a few schools still require taking the GRE (e.g., UCF), most good CS PhD programs in the US \textbf{no longer} require it.
+The reason is that GRE scores do not correlate well with research ability, which is the most important factor for PhD admission.
+It is also worth noting that many faculty members themselves did not take the GRE or have bad scores.
+
+Thus, if you have bad GRE scores or haven't taken the GRE, then don't waste time (re)taking it. Being optional really means optional, and not taking it will not hurt your application.
+However, if you took it and have really good scores then it might be worth it to include (and perhaps talk about) them in your application, but don't expect them to make much difference.
 But if your scores are bad, then you should not include them in your application, which can be a \red{red flag}.
@@
 You should just do well enough to pass the minimum requirement set by the university. %, which is required for TA (\autoref{sec:funding}).
-Just as with grades (\autoref{chap:your-school}) and GRE, having high scores in these tests might not help, but having too low scores can be a \red{red flag} and sometimes results in an automatic rejection (\autoref{sec:why-rejected}), e.g., below the minimum requirement.
+Just as with grades (\autoref{chap:your-school}) and GRE (\autoref{sec:gre}), having high scores in these tests might not help, but having too low scores can be a \red{red flag} and sometimes results in an automatic rejection (\autoref{sec:why-rejected}), e.g., below the minimum requirement.
 
@@
 
-While being ambitious and aiming high is good, you should understand how PhD admission works (e.g., by reading \emph{this} handbook and realizing things such as having a good GPA or GRE doesn't mean much to top programs), that in the US there are many good schools, and just be realistic.
+While being ambitious and aiming high is good, you should understand how PhD admission works (e.g., by reading \emph{this} handbook and realizing things such as having a good GPA (\autoref{sec:gpa}) or GRE (\autoref{sec:gre}) doesn't mean much to top programs), that in the US there are many good schools, and just be realistic.
 
@@
 
-\paragraph{School \autoref{chap:your-school}} Adcom already knows about your school, which is a plus (see \autoref{chap:your-school}).
+\paragraph{School \autoref{chap:your-school}} Adcom already knows about your school, which is a plus.
 You are also more familiar with the US education system and academic culture (\autoref{sec:cultural}).
@@
 
-\paragraph{Test Scores} GREs are often not required for MS applications.
-However, some schools might still require them, and you should check with the school you're applying to.
+\paragraph{Test Scores} Unlike PhDs, which often do not require GREs, MS programs often require GREs.
+However, some schools do make them optional, and you should check with the school you're applying to.
 If you have a low GPA, a high GRE score can help offset that.
-As with PhD, English proficiency tests (TOEFL/IELTS) are required for international students (but just passing the mimimum requirement is often enough \autoref{sec:english-tests}).
-Note that just as with PhD, Englih tests are are waived if you did your undergrad at a US university.
+
+As with PhD, English proficiency tests (TOEFL/IELTS) are required for international students (but just passing the minimum requirement is often enough \autoref{sec:english-tests}).
+Note that just as with PhD, English tests are are waived if you did your undergrad at a US university.
 

### 2025-01-07 demystify.tex
updates

@@
 Common concerns include (i) the \textbf{difficulty of getting admitted}, (ii) the \textbf{cost of graduate study} in the US, and for some, the fear of \textbf{discrimination} e.g., based on gender identity or cultural background.
-Much of these uncertainties stem from the lack of information and guidance on the admission process in the US (social places like Facebook and Reddit are full of confusion and contradictory information).
+
+Much of these uncertainties stem from the lack of information and guidance on the admission process in the US.
+Social platforms like Facebook and Reddit are full of confusion and contradictory information.
 This book aims to address these concerns.
@@
 
-This book also aim to explain the \emph{``why''} behind the admission process, e.g., why LoRs are important, why you should not draft your own LoR, why you should contact professors, etc.
-There are numerous \href{https://github.com/nguyenthanhvuh/phd-cs-us}{resources online} from students  that tell you \emph{``what''} to do, but few explain the \emph{``why''}.
+This book also aims to explain the reasons behind the admission process.
+There are numerous \href{https://github.com/nguyenthanhvuh/phd-cs-us}{resources online} from students  that tell you \emph{``what''} to do, but few explain the \emph{``why''}, e.g., why LoRs matter so much, why you should not draft your own LoR, why you should contact professors, etc.
 Understanding the reason and mindset of the admission committee and profs.
@@
 
-Some universities do waive, e.g., Rice does for all PhD applicants and many universities for domestic students (\autoref{sec:domestic-students}), however most do not for international students.
+Some universities do waive, e.g., Rice, TTIC do not have fee for PhD applications and many universities for domestic students (\autoref{sec:domestic-students}), however most do not for international students.
 Some require proof of financial difficulties, e.g., a statement from a financial adviser or a bank statement.
@@
 As mentioned in~\autoref{sec:lor-writers}, LoR writers should be someone who knows you well and has the credibility to evaluate your research ability.
-In the US, students usually explicitly ask if the writer would be willing to provide a strong letter, and the writer are also very direct in their response.
+In the US, it's common for students to explicitly ask if the writer would be willing to provide a strong letter, and the writer are also very direct in their response.
 If they are not willing (you should be thankful that they are honest with you), then you should ask someone else.
@@
 \item \textbf{Ask in advance} You should ask for LoRs \emph{at least a month} before the deadline.
-People are busy and writing a strong LoR takes time (\autoref{sec:lor-writing}), so give them enough time to write a strong letter for you.
+People have commitments and writing a strong LoR takes time (\autoref{sec:lor-writing}), so give them enough time to write a strong letter for you.
 
@@
 
-\item \textbf{Help your writers (\autoref{sec:help-your-LOR-writers})}: You should tell your writers the programs you are applying to, their deadlines, etc.
+\item \textbf{Help your writers (\autoref{sec:help-your-LOR-writers})}You should tell your writers the programs you are applying to, their deadlines, etc.
 You can also share your SOP with them and other details about your research experience and potential.
 
-\item \textbf{Ask for feedback} If the writer is very close to you, you can ask them for feedback on your SOP (\autoref{chap:sop}) and other application materials.
+\item \textbf{Ask for feedback} If the writer is very close to you and willing to, you can ask them for feedback on your SOP (\autoref{chap:sop}) and other application materials.
 If the writer is a professor, they might have served in adcom committees, seen many SOPs, and can provide valuable feedback.
@@
 After your writers have submitted all of their letters, \emph{thank them} (\autoref{sec:thank-writers}).
-Finally, let them know the outcome of your applications and stay in touch with them.
+Let them know the outcome of your applications and stay in touch with them.
 This will help you build a relationship with them and you might need them to write for you again in the future.
@@
 Here we look at publications and other research experiences that can strengthen your application.
-\autoref{chap:research-opportunities} provides more information on how to find research opportunities (e.g., during your undergrad study).
+\autoref{chap:research-opportunities} provides more information on how to find research opportunities, e.g., during your undergrad study.
 
@@
 
-Publications are never required for application, however given the competitiveness of CS admission, they can significantly strengthen your application and are becoming the norm.
+Publications are \emph{never required} for application, however given the competitiveness of CS admission, they can significantly strengthen your application and \emph{are becoming the norm}.
 Applicants admitted to top schools, especially in popular fields such as ML and NLP, often have multiple first-authored papers at top places.
+Fig.~\ref{fig:wu} shows examples of applicants to Stanford CS PhD.
+
+\begin{figure}
+\begin{center}
+    \includegraphics[scale=0.3]{files/wu.jpeg}
+\end{center}
+\caption{Applicants to top CS programs have multiple first-authored papers at top places.}\label{fig:wu}
+\end{figure}
+
+\paragraph{What if I don't have any pubs} It is understandable that many students do not have the opportunity to publish papers.
+Thus, other writings, even those under submissions or even rejected, would still be good and better than nothing.
+Be sure to upload your these with your application (\autoref{sec:writing-sample}) and talk about them in your SOP (\autoref{chap:sop}).
+Adcom members can quickly skim over the paper to determine its quality (\autoref{sec:ievaluate}).
+
+Note that local conferences and non-English journals or conferences do
+not carry as much weight since their quality is often unknown to US faculty.
+However, if you have published in such places, you should still upload them, mention them in your statement, and explain why they are good.
+
+% In CS, conferences are important and top conferences are prestigious.
+You can find top CS conferences at places such as \href{https://csrankings.org}{CSRankings}, which ranks CS programs based on how their faculty publish at top \emph{conferences} (see more in \autoref{sec:ranking}).
+
+\paragraph{I am not the first author} Being the first author is good because it indicates you own the work.
+However, it's \emph{perfectly OK to be second or third or even last}.
+Adcom members know it is difficult to publish a good paper, and so being a co-author is still a good sign about your research experience.
+In any case, especially in the case when you're not the first author, you should explain the work and your contribution.
+Better yet, have your LoR writers (\autoref{sec:help-your-LOR-writers}) talk about your work and contribution in their letters.
+
+\begin{commentbox}[Craig:]
+  Mason and many other universities allow you to upload your published papers and other writing samples (\autoref{sec:writing-sample}).
+In many cases, even if the papers were not published at top places, we can still determine their quality by simply skimming over the paper.
+\end{commentbox}
 
@@
 
-\paragraph{What if I don't have any pubs} It is understandable that many students do not have the opportunity to publish papers.
-Thus, other writings, even those under submissions or rejected, would still be good and better than nothing.
-Be sure to upload your these with your application (\autoref{sec:writing-sample}) and talk about them in your SOP (\autoref{chap:sop}).
-
-Note that local conferences and non-English journals or conferences do
-not carry as much weight since their quality is often unknown to US faculty.
-However, if you have published in such places, you should still upload them, mention them in your statement, and explain why they are good.
-
-% In CS, conferences are important and top conferences are prestigious.
-You can find top CS conferences at places such as \href{https://csrankings.org}{CSRankings}, which ranks CS programs based on how their faculty publish at top \emph{conferences} (see more in \autoref{sec:ranking}).
-
-\paragraph{I am not the first author} Being the first author is good because it indicates you own the work.
-However, it's \emph{perfectly OK to be second or third or even last}.
-Adcom members know it is difficult to publish a good paper, and so being a co-author is still a good sign about your research experience.
-In any case, especially in the case when you're not the first author, you should explain the work and your contribution.
-Better yet, have your LoR writers (\autoref{sec:help-your-LOR-writers}) talk about your work and contribution in their letters.
-
-\begin{commentbox}[Craig:]
-  Mason and many other universities allow you to upload your published papers and other writing samples (\autoref{sec:writing-sample}).
-In many cases, even if the papers were not published at top places, we can still determine their quality by simply skimming over the paper.
-\end{commentbox}
-
 \section{Work Experience} Work experience at \emph{well-known research laboratories}, such as Microsoft Research, can strengthen your
 application.
-Note that the emphasis here is \emph{research} places, not software development or non-research work.
-For example, working at a FAANG (Facebook, Amazon, Apple, Netflix, Google) company as a software engineer is not research experience, though it can be helpful for MS applications discussed in \autoref{chap:ms}.
+The emphasis here is \emph{research} places, not software development or non-research work.
+For example, working at a FAANG (Facebook, Amazon, Apple, Netflix, Google) company as a software engineer does not count as research experience, although it can be helpful for MS applications discussed in \autoref{chap:ms}.
 Similarly, a LoR from your supervisor for non-research experience might not count much.
-So do not spend much time talking about non-research job in your SOP.
+So do not spend much time talking about development job in your SOP.
 
@@
 For example, while VinAI is well-known in Vietnam and potentially in Asia, it might not be very well-known in the US.
-So you or your LoR writers should explicitly say something about them in your statement.
+So you or your LoR writers should explicitly say something about them in your statement or letter.
 In general, if you did some good research work, then you should mention that in your SOP and ask your supervisor to write about it in their LOR (\autoref{chap:lor}).
 
-% \begin{commentbox}[Hung:]
-%   The reputation of VinAI has been increasing steadily over the past few years; many of my colleagues heard about VinAI.
-% \end{commentbox}
-
-\section{Competitions} Winning \emph{internationally recognized competitions} can demonstrate your research potential.
+\section{Competitions}\label{sec:competitions}
+
+Winning \emph{internationally recognized competitions} can demonstrate your research potential.
 For example, participating in Math Olympiads if you want to do theory or winning ACM programming contests if you want to ``build'' systems, e.g., software analysis.  
@@
 
-\paragraph{Not a Good fit and Bad luck}  You could have an excellent profile (e.g., strong research and LoRs), but if you are interested in a research area that the program does not have, you will not be admitted.
+\paragraph{Not a Good Fit and Bad luck}  You could have an excellent profile (e.g., strong research and LoRs), but if you are interested in a research area that the program does not have, you will not be admitted.
 Similarly, if no faculty is willing to advise you (e.g., they are on sabbatical or personal leave, do not have sufficient funding, or already have too many students), you will not be admitted.
@@
     \item \textbf{Send emails to the wrong prof.}.
-Most students often do not pay attention or know about this, but a very common reason why you don't get a reply is that you send email to profs.
-who do not / cannot advise CS PhD students (e.g., adjunct, emeritus, teaching, non-CS, etc.).
+Many students do not pay attention or know about this, but a very common reason why you don't get a reply is that you send email to profs.
+who do not / cannot advise CS PhD students (e.g., adjunct, emeritus, teaching, non-CS).
 See \autoref{sec:faculty-types} for details.
@@
 might be able to co-advise, they typically do not have the bandwidth, funding, and the desire to mentor students for research (they are already overloaded with teaching and other service duties).
+
     So do some homework before sending emails, e.g., most CS profs.
@@
 Most teaching and adjunct profs.
-do not have such a research page.
+do not have a research group or students listed on their website.
 
@@
 You should already know this!
-A copy and paste kind of email or those that can be sent to multiple ppl with very little modifications show the lack of interest and will be treated as spam.
+A copy and paste kind of email or those that can be sent to multiple people with very little modifications show the lack of interest and will be treated as spam.
 Most likely we will not reply to these emails.
@@
     \item \textbf{Flowery greetings and language}.
-Don't use ``Dear esteemed professor'', don't ``hey there''.
+Don't use ``Dear esteemed professor''.
 Do not call the prof.
@@
 Do not use colors, fancy fonts or formats.
-While not really a kiss of death, it is very annoying, especially for ppl in CS (and probably many other fields) who often prefer plain messages.
+While not really a kiss of death, it is very annoying, especially for people in CS (and probably many other fields) who often prefer plain messages.
 
@@
 
-%\section{"Chance me" questions}\label{sec:chance-me}
-
-%Reddit and other grad admission forums are full of ``Chance me'' and ``roast my profile'' questions.
-These are questions where students post their profiles and ask others to evaluate their chances of getting admitted to their choice of schools.
-
-%These questions are not useful and should be avoided.
-First, no one knows!
-Even adcom members can't tell if you would get admitted to their programs or not.
-As mentioned throughout this guide, the admission process is complex and involves many factors that are not part of your profile, e.g., whether the faculty have funding or are interested in your background.
-Second, answers to this questions are often not helpful to you.
-Faculty and people knowledgeable about the process will not answer these questions, and those who do are often not knowledgeable and might give you false hope or despair.
-OK, someone say your GRE is low (which does not even matter for most schools (\autoref{chap:standard-tests}) , are you going to retake it?
-Or someone says your GPA is low, are you going to apply to a different school?
-No, you should apply to the schools you want to go to and see what happens.
-If you don't get in, then you can ask for feedback and improve your profile for the next round.
-
-\section{``Chance me'' on getting admitted to Top 10}\label{sec:chance-me}
-
-Many students, especially international students, aim for very top CS PhD programs such as Stanford and MIT, and Ivy Leagues.
+\section{``Chance me'' on getting admitted to T10 programs}\label{sec:chance-me}
+
+Many students, especially international students, aim for very top schools such as Stanford and MIT, and Ivy Leagues.
 Every year Reddit and other forums have numerous students asking for evaluation of their chances of getting into these schools (the so called \emph{``chance me''} or \emph{``roast my CV/profile''} posts) and then later posts on being \emph{``rejected everywhere''}.
 
-\paragraph{You are unlikely to get in these schools}  While being ambitious is always good, it is important to be realistic, and the harsh reality is that it is very unlikely that you will get in MIT, or Harvard.
-Why?
-people who do get in these schools often were explicitly encouraged by their mentors and LoR writers, who themselves might be alumni of these schools or are academic celebrities who've sent many of their students there.
+\paragraph{You are unlikely to get in these schools}  While being ambitious is always good, it is important to be realistic, and the harsh reality is that it is very unlikely that you will get in MIT or Harvard unless you are very exceptional (in which case you would not be asking for chances on Reddit).
+
+People who do get in these schools often were explicitly encouraged by their mentors and LoR writers, who themselves might be alumni of these schools or are academic celebrities who've sent many of their students there.
 If you are not in this category (doubt you are, because you're asking random people on the internet instead of your trusted LoR writers), then you're likely not going to get in, and MIT and Stanford would glady take your application money.
@@
 Unfortunately, even if all revierwers recommend a student, they might not be admitted (\autoref{sec:why-rejected}) if there are too many students or they are not a good fit for the program (e.g., no one is willing to advise them).
+
+\section{Checking for AI in SOPs}\label{sec:ai-sop}
+
+As AI/LLMs become more popular, many students wonder if they could use it to help with their statements and university/adcom reviewers would check and penalize AI-assisted writing.
+
+While I cannot speak for others, I do not check your statements for AI contents.
+First, I do not have the time to do that.
+It is much easier for me to just read the statement and see if it makes sense and stands out (\autoref{sec:ievaluate}).
+Hint: AI-generated content reads very strangely and faculty is just too experienced in reading essays and SOPs from students
+to not notice it.
+Second, AI-checking technology is very unreliable and inconsistent (e.g., a checker might claim that 80\% of an essay is AI-generated while another says it is 0\%).
+Third, while it is not OK to use AI to write your statement (as it is both easy to detect and just feels wrong), I think it is fine to use it to help you polish your writing, e.g., I find the ``proofread'' feature in Apple's \texttt{Writing Tools} useful to fix writing mistakes.
 
@@
 
-\section{How Applications are Evaluated}
+\section{NSF GRFP}\label{sec:nsf-grfp}
+
+\subsection{How Applications are Evaluated}
 
 \begin{itemize}
-\item Each applicant is assigned to ~3 reviewers who have expertise in the applicant's research topics (so only your field but also your specific area).
+\item Each applicant is assigned to ~3 reviewers who have expertise in the applicant's research topics (so not only your field but also your specific area).
 This means reviewers can be quite knowledgeable in your field and your work, your LoR writers and their work, and thus can evaluate your application quite well.
@@
 
-\section{Intellectual Merit and Broader Impacts}
+\subsection{Intellectual Merit and Broader Impacts}
 Your will be evaluated based on two criteria: Intellectual Merit (IM) and Broader Impacts (BI).
@@
 
-\section{Common Pitfalls}
+\subsection{Common Pitfalls}
 

### 2025-01-11 demystify.tex
updates

@@
 
-It is unlike that you can negotiate TA stipend (\autoref{sec:ta}), which is set by the university.
-Moreover, while RA is set by the adviser (\autoref{sec:ra}), it is often based on the department's standard, and it also would be unfair to other students if you get more (especially since you're new and have not proven yourself yet).
-However, your stipend is often increased every year by the university (automatically by a small amount).
+Negotiating a TA stipend is unlikely, as it is determined by the university (\autoref{sec:ta}).
+For RA (\autoref{sec:ra}), advisers have more leeway as it is funded by their grants.
+However, they will likely do not negotiate RA stipend as they have to be fair to other students and also have to follow departmental standards.
+Typically, the most wiggle room is funding in the summer (\autoref{sec:summer-funding}) where the prof.
+might be able to pay you more from their grants.
+Note that the university typically automatically increases these stipends each year by a small amount.
 
@@
 In short, standard things set by the university or department are unlikely negotiable.
-However, you can ask for things such as computer equipment (\autoref{sec:buying-equipment}), books, (and maybe even furniture), and a conference travel budget.
+However, you can ask for things such as books and computer equipment (\autoref{sec:buying-equipment}).
 
@@
 
-\subsection{How Applications are Evaluated}
+\subsection{How Applications are Evaluate}
 
 \begin{itemize}
-\item Each applicant is assigned to ~3 reviewers who have expertise in the applicant's research topics (so not only your field but also your specific area).
-This means reviewers can be quite knowledgeable in your field and your work, your LoR writers and their work, and thus can evaluate your application quite well.
-
-\item While we review applicants from various levels (e.g., undergrad, 1st year, 2nd year in PhD program), we evaluate and compare applicants \emph{within} their cohort level.
-This means that we do not compare an undergrad, who is still applying to grad school, to a 2nd year PhD student who already has an adviser and published multiple top papers.
+\item Each applicant is assigned to \emph{about 3 reviewers} who all seem to have expertise in the applicant's research topics.
+This means reviewers can be quite knowledgeable in your field and your work, your LoR writers and their work, and thus can evaluate your application quite in depth.
+This is different from PhD applications where adcom reviewers (\autoref{sec:adcom}) might not be in your field.
+
+Note that if sometimes an original reviwer is no longer available then someone who is not in the field might review your application.
+Thus in general you want to make your application accessible to a general audience in CS.
+
+\item Reviewers attempt to compare applicants \emph{within their cohort level}.
+This means that we do not compare an undergrad, who is still applying to grad school, to a 2nd year PhD student who already has an adviser and published multiple papers.
 This ensures that each applicant is evaluated fairly based on their level.
+
+\item Unlike CS PhD applications, which focus on research potentials in which concrete evidence like papers, research experience, and LoRs are important, NSF GRFP also focuses on \emph{potential} but also on broader impacts, e.g., how you can help society, how you can mentor others, etc.
 

### 2025-01-14 demystify.tex
Updates from Overleaf

@@
 \begin{mybox}
-This book is open source and updated regularly. Its latest version is available at
+This book is open source and available at
 
@@
 International students, especially those from less well-known universities and smaller countries, often wonder if they should apply to a CS PhD program in the US.
-Common concerns include (i) the \textbf{difficulty of getting admitted}, (ii) the \textbf{cost of graduate study} in the US, and for some, the fear of \textbf{discrimination} e.g., based on gender identity or cultural background.
+Common concerns include (i) the \textbf{difficulty of getting admitted}, (ii) the \textbf{cost of graduate study} in the US, and for some, (iii) the fear of \textbf{discrimination} e.g., based on gender identity or cultural background.
 
@@
 
-\paragraph{Not any harder than other countries} First, applying to a good US university \emph{should not} be any harder than at schools in other countries.
+\paragraph{Not any harder than other countries} Applying to a good US university \emph{should not} be any harder than at schools in other countries.
 It might even be more flexible since CS PhD in the US \emph{do not} require having an MS or a research topic, proposal, or adviser in advance (\autoref{sec:non-us-differences} compares CS PhD study in the US to other countries).
@@
 
-\paragraph{Funding is not an issue} Second, PhD students in CS \emph{do not} need to worry about funding, especially at good R1
+\paragraph{Funding is not an issue} PhD students in CS \emph{do not} need to worry about funding, especially at good R1
 universities in the US. 
@@
 
-\paragraph{Discrimination is not tolerated} Third, not only is discrimination illegal, diversity and inclusion are highly valued in US universities.
+\paragraph{Discrimination is not tolerated} Not only is discrimination illegal, diversity and inclusion are highly valued in US universities.
 Many universities take great pride in their diverse student body and faculty.
@@
 This book also aims to explain the reasons behind the admission process.
-There are numerous \href{https://github.com/nguyenthanhvuh/phd-cs-us}{resources online} from students  that tell you \emph{``what''} to do, but few explain the \emph{``why''}, e.g., why LoRs matter so much, why you should not draft your own LoR, why you should contact professors, etc.
+There are numerous \href{https://github.com/nguyenthanhvuh/phd-cs-us}{resources online} that tell you \emph{``what''} to do, but few explain the \emph{``why''}, e.g., why LoRs matter so much, why you should not draft your own LoR, why you should contact professors, etc.
 Understanding the reason and mindset of the admission committee and profs.
@@
 
-While this handbook may not fully meet the needs of students aiming for top-10 schools (\autoref{sec:selecting-ranking-schools} and \autoref{sec:ranking}), which typically have more stringent requirements (e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU), it should be valuable for those targeting other reputable R1 institutions, which are frankly the majority of good schools in the US.
+While this handbook may not fully meet the needs of students aiming for top-10 schools (\autoref{sec:selecting-ranking-schools} and \autoref{sec:ranking}), which typically have more stringent requirements (e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU), it should be valuable for those targeting other reputable R1 institutions.
 

### 2025-01-15 demystify.tex
Updates from Overleaf

@@
 
-Moreover, depending on the university, you may receive additional benefits such as summer pay, laptops, and traveling to conferences and workshops. 
+Moreover, you often receive additional benefits such as summer pay (\autoref{sec:summer-funding}), laptops (\autoref{sec:buying-equipment}), and traveling to conferences and workshops. 
 Full funding for CS PhD students is the norm in the US, and I'd go as far as to say that if you are not admitted with full funding, you might want to not accept the offer.
@@
 They might also receive guidance from their professors or grad students who have gone through the process.
-Students from smaller schools or countries, however, might not have access to such resources and might not know where to start.
+Students from smaller schools or countries, however, might not have access to such resources and know where to start.
 My goal is thus to level the playing field by providing info that is not readily available to less privileged students. 
@@
 
-While this handbook may not fully meet the needs of students aiming for top-10 schools (\autoref{sec:selecting-ranking-schools} and \autoref{sec:ranking}), which typically have more stringent requirements (e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU), it should be valuable for those targeting other reputable R1 institutions.
+%While this handbook may not fully meet the needs of students aiming for top-10 schools (\autoref{sec:selecting-ranking-schools} and \autoref{sec:ranking}), which typically have more stringent requirements (e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU), it should be valuable for those targeting other reputable R1 institutions.
 
@@
 
-\begin{center}
+\begin{figure}
+\centering
   \includegraphics[scale=0.5]{files/c4a.png}
-\end{center}
+  \caption{The ``ambition'' level of a PhD student over their years of study (they miss the 6--7th year when the ambition is \emph{``Just let me graduate''}).}
+\end{figure}
 
@@
 
-The PhDComics figure above shows the ``ambition'' level of a PhD student over their years of study (they miss the 6--7th year when the ambition is \emph{``Just let me graduate''}).
-
 \begin{commentbox}[Vu:]
@@
 
-A main question adcom has for a non-CS or non-STEM student is if you have the sufficient technical background obtained through core CS courses.
+A main question adcom (\autoref{sec:adcom}) has for a non-CS or non-STEM student is if you have the sufficient technical background obtained through core CS courses.
 So you need to show that you have such knowledge through your coursework, projects, or research.
@@
 
-No, while other countries often require an MS for PhD student in CS (~\autoref{sec:non-us-differences}), it is common in he US to apply for a PhD program directly after a 4-year undergrad program (e.g., after getting a B.S degree).
+No, while other countries often require an MS for PhD student in CS (\autoref{sec:non-us-differences}), it is common in the US to apply for a PhD program directly after a 4-year undergrad program (e.g., after getting a B.S degree).
 Most CS PhD programs are designed so that students can get MS degree ``along the way'' to PhD, e.g., after finishing the 2-year course work.
@@
 \emph{Course work}: In the US ou will spend the first couple of years taking classes and exploring potential adviser and research topics. 
-After that, you have to pass a series of exams during your PhD, e.g., qualifying exam, comprehensive exam, thesis proposal defense\footnote{The word ``ABD'' (all but dissertation) is used in the US to refer to a PhD candidate who has finished all course work and exams and only needs to write and defend their dissertation.}.
+After that, you have to pass a series of exams during your PhD, e.g., qualifying exam, comprehensive exam, thesis proposal defense\footnote{ABD (all but dissertation) refers to a PhD candidate who has finished all course work and exams and only needs to write and defend their dissertation.}.
 
@@
 
-\section{Admission Committee (adcom)}\label{sec:adcom}
+\section{Admission Committee (Adcom)}\label{sec:adcom}
 \sectioninfo{Adcom members are faculty who evaluate your application. They consider various factors, e.g., research experience, LoRs, SOP}
@@
 Each application is assigned to about three adcom members, who will evaluate your profile and reach a consensus.
-They will consider various factors, e.g., research experience, GPA, LoRs, SOP, and test scores.
+They will consider various factors, e.g., research experience, GPA, LoRs, SOP, test scores, and interviews.
 More details about these factors are discussed in \autoref{part:appendices}.
@@
 our decision (\autoref{sec:ievaluate}) with a summary of your application, where we list
-strengths, e.g., a well-known school, and weaknesses, e.g., weak LoRs.
+strengths, e.g., came from a well-known school, and weaknesses, e.g., weak and generic LoRs.
 \end{commentbox}
@@
 
-Not really, we typically review applications independently and do not talk to each other.
+We typically review applications independently and do not talk to each other.
 This is to avoid biasing, e.g., if one reviewer says they want to accept, the other might feel pressured to accept as well.
@@
 
-\begin{commentbox}[A prof. in Physics at an R1 university:]
+\begin{commentbox}[From a prof. in Physics at an R1 university:]
   We have a pretty well fleshed out grading rubric for applications that has categories like grades, research, writing ability, etc.
@@
 
-\section{Why CS depts do not waive the application fee?}\label{sec:fee-waive}
-
-Some universities do waive, e.g., Rice, TTIC do not have fee for PhD applications and many universities for domestic students (\autoref{sec:domestic-students}), however most do not for international students.
-Some require proof of financial difficulties, e.g., a statement from a financial adviser or a bank statement.
-
-In general, application fee is typically a requirement of the university. Individual departments and programs do not have the flexibility to waive the application fee, even if they want to.
+\section{Waiving the application fee}\label{sec:fee-waive}
+
+Some universities do waive, e.g., Rice, TTIC do not have fee for PhD applications and many universities for domestic students (\autoref{sec:domestic-students}).
+Some programs also waive if the applicants attend some of their opening sessions.
+Some programs waive if the applicant provide proof of financial difficulties, e.g., a statement from a financial adviser or a bank statement.
+
+However most do not waive the application fee, which is typically a requirement of the university.
+Individual departments and programs do not have the flexibility to waive the application fee, even if they want to.
 
@@
 Most CS programs already receive many applications and would be overwhelmed if the application process were free.
-Even with application fee the competition is already very tough, imagine if the application is free and the number of applications doubles or triples.\\
+Even with application fee the competition is already very tough, imagine if the application is free and the number of applications triples or quadruples.\\
 
@@
 Most PhD programs will require at least \emph{two} letter of recommendations (LoRs). 
-The goal of the LoRs is to provide the adcom with an assessment of your research ability from someone who knows you.
+The goal of the LoRs is to provide the adcom with an assessment of your research ability from someone who knows about research and \emph{you} well.
 The LoRs should focus on your research background, achievements, and potentials, all of which paint a compelling picture of you as a researcher to adcom.
@@
 
-\subsection{LoR from Famous People}  Having a strong letter from an internationally recognized researcher can \emph{greatly strengthen} your application.
+\subsection{LoR from Well-Known People}  Having a strong letter from an internationally recognized researcher can \emph{greatly strengthen} your application.
 Such a letter can outweigh the lack of publications (or low GPA or unknown school).

### 2025-01-19 demystify.tex
updates

@@
 If they insist, then I will write for them.
-While I try to say something positive, e.g., the student is hardworking, or that I heard they are working on a research project, etc, the letter will still be short and weak (\autoref{sec:generic-letters}).
-Usually it takes me about \emph{5--10} minutes to write such a letter.
+While I try to say something positive, e.g., the student is hardworking and receive good grades, the letter will still be short and weak (\autoref{sec:generic-letters}).
+Usually it takes me about \emph{5--10} minutes to write such a ``weak'' letter.
 
@@
 
-\section{GREs Are Optional and Do Not Matter}\label{sec:gre}
+\section{GREs Are Optional and Do Not Matter for PhD Admissions}\label{sec:gre}
 
@@
 But if your scores are bad, then you should not include them in your application, which can be a \red{red flag}.
+
+Note that while GRE is not important for CS PhD admission, it might be required or important for MS admission (\autoref{chap:ms}).
+This is because MS programs are more course-based and thus care more about grades and standardized tests.
 
@@
 
-\subsection{How Applications are Evaluate}
+\subsection{How Applications are Evaluated}
 
@@
 So you will need to demonstrate your passion for research and your potential to be a good researcher.
+
+\section{DoD NDSEG}
+
+The DoD National Defense Science and Engineering Graduate (NDSEG) is another prestigious fellowship supporting \emph{U.S.
+citizens} pursuing advanced degrees in STEM fields critical to national defense (such as CS).
+It provides up to three years of funding, including a  stipend, tuition coverage, and health insurance.
+
+\subsection{Evaluation}
+Unlike the NSF GRFP, for NDSEG reviewers \emph{do not} see your LoRs, i.e., we do not know who wrote them or what they wrote.
+Moreover, we only see \emph{redacted versions} of your submitted materials (you are asked to provide a full and redacted version of your research plan, personal essays, and CV).
+Thus we do not know what schools your went to, where you did your internships, your publications or research experience.
+We can see your current transcript, but not your past ones.
+
+Thus, reviewers mainly get to know you and evaluate your application by reading your research plan and personal essays.
+The reason is for fairness and avoiding bias based on the school you went to, places you published at, your LoR writers, etc.
+It could be that external reviewers like myself help with the first cut based on the redacted materials, and then the final committee makes the final decision based on the full materials.
+
+\begin{commentbox}[Vu:]
+I am used to reading full applications and thus find it a bit difficult to evaluate NDSEG applications because I do not have the full picture.
+For example, a student mentioned that they have published in the topic they proposed but redacted all information about the publication, so I have no idea where they published, what the paper is about, etc.
+It is also difficult to evaluate the academic background of the student, e.g., some student just started grad school and so their transcript is essentially blank, and I do not know where they went to undergrad, etc.
+It is also annoying to see CVs with so many blanked out information.
+They might as well should just submit a blank CV one.
+\\
+
+In the beginning I find it surprising that I don't get to see the LoRs, but I understand the reasoning behind it and gradually think it is a good idea.
+As mentioned above, LoR writers, especially those from top schools, often overrate their students and thus can bias the evaluation.
+So I don't find I am missing much by not seeing the LoRs.
+\end{commentbox}
+
+\subsection{Common Pitfalls and Tips}
+Based on my experience reviewing NDSEG applications, here are some common pitfalls to avoid:
+\begin{itemize}
+
+    \item Customize your research to DoD, e.g., through specific project from DARPA or ONR. 
+    Make this prominent, e.g., in the introduction or in its own section.  This is different from NSF GRFP where you want broader impacts to society.
+
+    \item Don't make up things or exaggerate your research.
+Reviewers are experts in the field and can easily spot technical inaccuracies or ideas that have already been studied.
+Understandbly many applicants are new, but you need to do proper research on what you propose.
+In particular, you should aim to have (i) clear problem statement, (ii) why it is important (esp.
+to DoD), (iii) what has been done and their challenges/limitations, and (iv) what you plan to do to address the challenges.
+Moreover, research is often quite technical and ambitious, so you want to show some preliminary work, e.g., a small experiment by hand on some small examples, to convince reviewers that you have thought about the problem and have some initial results.\\
+
+    Unlike GRFP where we have more things to evaluate, in NDSEG we mainly evaluate you based on your personally essays and especially research plan.
+So you want to make sure your research plan is well thought out and convincing.
+
+    \item Personal stories and essays should be unique and interesting.
+Many personal stories are quite generic and do not tell much about the student.
+For example, being an URM, involving in a robotics club or mentoring underrepresented students are quite common and do not tell much.\\
+
+    What you want is to pick a few thing and provide concrete examples and details.
+For example, provide a concrete experience in being minority and how it has affected you, talk about what you have accomplished in your robotics club, e.g., you built a robot that won a competition, or you mentored a specific URM student who then went on to do great things.
+Just a small, interesting detail can make your essay stand out.
+
+    \item Not using all allowed pages.
+Many research plans use only 1/2 or 2/3 of the number of allowed pages, which themselves are already quite short.
+Such documents are often not detailed enough and do not provide enough information for reviewers to understand and evaluate the research plan.
+So use all allowed pages to provide enough details about your research plan.
+This is especially important as we do not know much about you due to redacted materials.
+
+\item 
+personal essay / stories are written in text box and so do not llook very professional and doesn’t seem to be heavily weighted compared to research plan (e.g., very few questions asked about the personal stories compared to questions asked on research plan).
+
+- use latex,  do not write in word and especially use left justification  (see how CS and engineer people publish papers)  
+- If you have an advisor or Lor writer, ask them to look at your writing/essays .
+This is even suggested directly by NSF and DoD.
+You are new to the field and what your ideas might not be accurat or have already been explored — let them help you here.
+While you can BS to PhD adcom because there you don’t really write a research plan,  but for these fellowships you do and reviewers are all experts in the field and would look at these carefully.
+
+\end{itemize}
 

### 2025-02-06 demystify.tex
updates

@@
 In fact, many resources are available to encourage and support students from minority and diverse groups to pursue higher education and research, e.g., specific scholarships and fellowships.
-Faculty in CS are also \emph{very} familiar with and are highly encouraged to support such students in their group.
+Faculty in CS are also familiar with and are encouraged to support such students in their group.
 There are also incentives, such as dedicated funding and awards, to motivate faculty to recruit and mentor minority students (\autoref{sec:urm}).
@@
 
-Most US universities have CS studies, but many of them do not have a PhD program in CS.
+Most US universities have CS programs, but many of them do not have a CS \emph{PhD} program.
 These universities might offer just Bachelor's degrees (e.g., BS) and no graduate studies (i.e., no MS or PhD degrees), or they just offer MS programs (but no PhD).
@@
 
-Many students just put schools into two bins: (i) top schools that they dream about, and (ii) everything else.
-They commonly use rankings from US News, which is not transparent and questionable\footnote{\url{https://cra.org/cra-statement-us-news-world-report-rankings-computer-science-universities/}}.
+Many students put universities into two bins: (i) top schools that they dream about, and (ii) everything else.
+They often use rankings from US News, which is not transparent and questionable\footnote{\url{https://cra.org/cra-statement-us-news-world-report-rankings-computer-science-universities/}}.
 Sometimes they evaluate based on the reputation of the school's undergrad program or the reputation of the school's non-CS programs such as medical, math, or physics.
@@
 
-Many students, especially international students, aim for very top schools such as Stanford and MIT, and Ivy Leagues.
-Every year Reddit and other forums have numerous students asking for evaluation of their chances of getting into these schools (the so called \emph{``chance me''} or \emph{``roast my CV/profile''} posts) and then later posts on being \emph{``rejected everywhere''}.
+Many international students aim for very top schools such as Stanford and MIT, and Ivy Leagues.
+Every year Reddit and other forums have numerous students asking for evaluation of their chances of getting into these schools (the so called \emph{``chance me''} or \emph{``roast my CV/profile''} posts) and then later posts on being \emph{``ghosted and rejected everywhere''}.
 
@@
 However, if someone has this issue, then you might not want to work with them anyway (and such bias is not acceptable in academics and likely would not be tolerated by the university).
-In my experience, it is the opposite: \emph{many profs.
+In my experience, it is often the opposite: \emph{many profs.
 actively seek to work with students from diverse backgrounds and view diversity as an asset to their research group}.
@@
 The DoD National Defense Science and Engineering Graduate (NDSEG) is another prestigious fellowship supporting \emph{U.S.
-citizens} pursuing advanced degrees in STEM fields critical to national defense (such as CS).
+citizens} pursuing advanced degrees in STEM fields \emph{critical to national defense} (such as CS).
 It provides up to three years of funding, including a  stipend, tuition coverage, and health insurance.
 
+You can read more details about NDSEG and its guideline for applicants from the \href{https://www.ndsegfellowships.org/}{official website}.
+Here, I mainly focus on things that stand out to me as reviewers and also compare them with NSF GRFP as many students apply to both.
 \subsection{Evaluation}
-Unlike the NSF GRFP, for NDSEG reviewers \emph{do not} see your LoRs, i.e., we do not know who wrote them or what they wrote.
-Moreover, we only see \emph{redacted versions} of your submitted materials (you are asked to provide a full and redacted version of your research plan, personal essays, and CV).
+
+\paragraph{Redacted info.} Unlike the NSF GRFP, for NDSEG reviewers \emph{do not} see your LoRs, i.e., we do not know who wrote them or what they wrote.
+Moreover, we only see \emph{redacted versions} of your submitted materials (applicants are asked to provide full and redacted versions of your research plan, CV, and other materials).
 Thus we do not know what schools your went to, where you did your internships, your publications or research experience.
-We can see your current transcript, but not your past ones.
 
 Thus, reviewers mainly get to know you and evaluate your application by reading your research plan and personal essays.
-The reason is for fairness and avoiding bias based on the school you went to, places you published at, your LoR writers, etc.
-It could be that external reviewers like myself help with the first cut based on the redacted materials, and then the final committee makes the final decision based on the full materials.
+This is for \emph{fairness and avoiding bias} based on the school you went to, places you published at, your LoR writers, etc.
+Nonethless, I felt I am missing too much information.
+However, it could be that external reviewers like myself help with the first cut based on the redacted materials, and then there are internal reviewers who makes the final decision based on the full materials.
+
+\paragraph{Non-expert} NDSEG reviewers hold advanced degrees in STEM fields.
+However, NDSEG reviewers are more general and might not be in your specific field.
+For example, in GRFP I mostly review applications in PL and FM, while in NDSEG it seems that I review applications in all areas of CS and even non-CS.
+Thus, you should make your research plan \emph{more accessible to a general audience}.
 
 \begin{commentbox}[Vu:]
-I am used to reading full applications and thus find it a bit difficult to evaluate NDSEG applications because I do not have the full picture.
+I find a bit difficult to evaluate NDSEG applications because I do not have the full picture.
 For example, a student mentioned that they have published in the topic they proposed but redacted all information about the publication, so I have no idea where they published, what the paper is about, etc.
@@
 \subsection{Common Pitfalls and Tips}
-Based on my experience reviewing NDSEG applications, here are some common pitfalls to avoid:
+Based on my experience in reviewing, here are some common pitfalls to avoid:
 \begin{itemize}
@@
 
-    \item Don't make up things or exaggerate your research.
-Reviewers are experts in the field and can easily spot technical inaccuracies or ideas that have already been studied.
-Understandbly many applicants are new, but you need to do proper research on what you propose.
-In particular, you should aim to have (i) clear problem statement, (ii) why it is important (esp.
+    \item Unclear research plan. 
+     You should aim to have (i) clear problem statement, (ii) why it is important (esp.
 to DoD), (iii) what has been done and their challenges/limitations, and (iv) what you plan to do to address the challenges.
@@
     Unlike GRFP where we have more things to evaluate, in NDSEG we mainly evaluate you based on your personally essays and especially research plan.
-So you want to make sure your research plan is well thought out and convincing.
+So you should do proper research on what you propose and make it clear and convincing.
 
     \item Personal stories and essays should be unique and interesting.
-Many personal stories are quite generic and do not tell much about the student.
+Many essays and stories I read are quite generic and do not tell much about the student.
 For example, being an URM, involving in a robotics club or mentoring underrepresented students are quite common and do not tell much.\\
 
-    What you want is to pick a few thing and provide concrete examples and details.
+    You should pick a few thing and provide concrete examples and details.
 For example, provide a concrete experience in being minority and how it has affected you, talk about what you have accomplished in your robotics club, e.g., you built a robot that won a competition, or you mentored a specific URM student who then went on to do great things.
@@
 
-\item 
-personal essay / stories are written in text box and so do not llook very professional and doesn’t seem to be heavily weighted compared to research plan (e.g., very few questions asked about the personal stories compared to questions asked on research plan).
-
-- use latex,  do not write in word and especially use left justification  (see how CS and engineer people publish papers)  
-- If you have an advisor or Lor writer, ask them to look at your writing/essays .
-This is even suggested directly by NSF and DoD.
-You are new to the field and what your ideas might not be accurat or have already been explored — let them help you here.
-While you can BS to PhD adcom because there you don’t really write a research plan,  but for these fellowships you do and reviewers are all experts in the field and would look at these carefully.
+%personal essay / stories are written in text box and so do not llook very professional and doesn’t seem to be heavily weighted compared to research plan (e.g., very few questions asked about the personal stories compared to questions asked on research plan).
+
+%- use latex,  do not write in word and especially use left justification  (see how CS and engineer people publish papers)  
+\item If you have an advisor or Lor writer, ask them to look at your writing/essays.
+This is because you are new to the field and research writing, and your ideas might not be accurate or have already been explored.
 

### 2025-02-10 demystify.tex
updates

@@
 Second, AI-checking technology is very unreliable and inconsistent (e.g., a checker might claim that 80\% of an essay is AI-generated while another says it is 0\%).
-Third, while it is not OK to use AI to write your statement (as it is both easy to detect and just feels wrong), I think it is fine to use it to help you polish your writing, e.g., I find the ``proofread'' feature in Apple's \texttt{Writing Tools} useful to fix writing mistakes.
+Third, while it is not OK to use AI to write your statement (as it is both easy to detect and just feels wrong), I think it is fine to use it to help you polish your writing, e.g., the ``proofread'' feature in Apple's \texttt{Writing Tools} is quite useful for fixing writing issues.
 

### 2025-02-10 demystify.tex

@@
     Objective & Industry & Research \\
-    Admission Req & Research experience & No research experience \\
+    Admission Req & No research experience & Research experience \\
     Coursework Req & Yes & Yes (research is much more important)  \\

### 2025-02-10 demystify.tex
updates

--- 2025-02-10 demystify.tex
@@
 If there are disagreements, adcom chair will ask the reviewers to discuss the application. 
-Unfortunately, even if all revierwers recommend a student, they might not be admitted (\autoref{sec:why-rejected}) if there are too many students or they are not a good fit for the program (e.g., no one is willing to advise them).
+Unfortunately, even if all reviewers recommend a student, they might not be admitted (\autoref{sec:why-rejected}) if there are too many students or they are not a good fit for the program (e.g., no one is willing to advise them).
 
@@
 
-As AI/LLMs become more popular, many students wonder if they could use it to help with their statements and university/adcom reviewers would check and penalize AI-assisted writing.
+As AI/LLM's become more popular, many students wonder if they could use it to help with their statements and university/adcom reviewers would check and penalize AI-assisted writing.
 

### 2025-02-10 demystify.tex
updates

@@
 \chapter*{Preface}
-Having been involved in PhD admission committees for many years, I've realized that \textbf{international} students, especially those in smaller countries or less well-known universities, often lack a clear understanding of
+Having been involved in PhD admission committees for many years, I've observed that \textbf{international students}, especially those in smaller countries or less well-known universities, often lack a clear understanding of
 the Computer Science PhD admission process at US universities. This confusion not only
@@
 So I want to share some details about the admission process and advice for those who are interested in applying for a \textbf{PhD in Computer Science in the US}.
-Originally, this book was intended for \emph{international students}, but has been expanded to include information for \emph{US domestic students} (\autoref{sec:domestic-students}).
+Originally, this book was intended for international students, but has been expanded to include information for \textbf{US domestic students} (\autoref{sec:domestic-students}).
 Moreover, while this primarily aims at students interested in CS, it might be relevant to students from various STEM (Science, Technologies, Engineering, and Mathematics) disciplines.
@@
 
-While this book can be useful for all students, it would most benefit \textbf{international students} from \textbf{smaller countries and less well-known universities} (or domestic students from US universities with no PhD programs or limited research opportunities).
+While this book can be useful for all students, it would most benefit \textbf{international students} from \textbf{smaller countries and less well-known universities}.
+It is also useful for \textbf{domestic students} from US universities with \textbf{no PhD programs or limited research opportunities}.
 
 Students from top schools with strong research programs and experience might already know most of the information given here.
-They might also receive guidance from their professors or grad students who have gone through the process.
+They likely have received guidance from their professors or grad students who have gone through the process.
 Students from smaller schools or countries, however, might not have access to such resources and know where to start.
@@
 
-This book also aims to explain the reasons behind the admission process.
+This book also aims to explain the \textbf{reasons behind the admission process}.
 There are numerous \href{https://github.com/nguyenthanhvuh/phd-cs-us}{resources online} that tell you \emph{``what''} to do, but few explain the \emph{``why''}, e.g., why LoRs matter so much, why you should not draft your own LoR, why you should contact professors, etc.
@@
     Admission Req & No research experience & Research experience \\
-    Coursework Req & Yes & Yes (research is much more important)  \\
+    Coursework Req & Yes & Yes (but research is much more important)  \\
     Duration & 2 years & 5--7 years \\
@@
 
-  \item Course requirements: MS has a specific number of courses (typically can be done in 2 years).
+  \item Course requirements: MS has a specific number of courses that typically can be done in 2 years.
 You graduate with an MS when you're done with the courses.
-PhD also has coursework requirements (typically taken in your first 2 years as mentioned in \autoref{sec:time}) but focuses on research after that.
+PhD also has coursework requirements, which are typically taken in your first 2 years  (\autoref{sec:time}).
+However, after (and also during) coursework, you focus mainly on research.
 You graduate with a PhD when you have done enough research and written a dissertation, which usually takes much longer time than coursework.
@@
 
-However, there are always exceptions, e.g., if you have a strong background and can demonstrate that you can be productive in research, then a prof. might be willing to take you as their RA.
+\subsection{Exceptions} 
+While MS students are typically not funded and do not focus on research, there are always exceptions.
+Some MS programs are fully funded, e.g., those that are research-oriented and have a strong research component.
 Moreover, some schools have TAs for MS students, and you can apply for these positions. For example, Mason CS has quite a few TAs for MS students, and my courses in the past two years have MS TAs.
+Finally, if you have a strong background and can demonstrate that you can be productive in research, then a prof.
+might be willing to take you as their RA.
+Some of these students convert to PhD students after having this research experience.
 
@@
 
-Because of the differences between MS and PhD programs, you should consider different factors when selecting and ranking MS programs.
+Because of the differences between MS and PhD programs (\autoref{sec:phd-vs-ms}), you should consider different factors when selecting and ranking MS programs.
 For example, you might want to consider factors such as location, industry connections, and job placement instead of research areas and advisers (\autoref{sec:schoolsandprofs}).
@@
 
-\item \textbf{Ranking:} Unlike PhD programs where the adviser and research are perhaps the most important factor, in MS programs the ranking of the school is more important.
+\item \textbf{Ranking:} Unlike PhD programs where the adviser and research are likely the most important factor, in MS programs the ranking of the school is typically more important.
 Thus, you should consider the ranking of the school in CS, especially in the specific area you're interested in.
@@
 (\href{https://csrankings.org}{CSRankings}, Jan.
-2024).
+2025).
 \red{$^*$} indicates that the university has \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{Vietnamese prof.} that can advise CS PhD students.}\label{tab:ranking}
@@
     \toprule
-    1 & Carnegie Mellon & 26 & Duke University \\
-    2 & Univ. of Illinois at Urbana-Champaign\red{$^*$}  & 27 & Univ. of California - Santa Barbara \\
-    3 & Univ. of California-San Diego & 28 & Rutgers University\red{$^*$} \\
-    4 & Georgia Institute of Technology & 29 & Univ. of California - Riverside\\
-    5 & MIT                            & 30 & Pennsylvania State University  \\
-    6 & Univ. of California - Berkeley& 31 & Northwestern University\\
-    7 & University of Michigan - Ann Arbor\red{$^*$}   & 32& George Mason University\red{$^*$}\\
-    8 & University of Washington      &33 &  Texas A\&M University\red{$^*$} \\
-    9 &  Stanford University  &34&  University of Utah \\
-    10 & Cornell University  & 35 &  North Carolina State University\\\
-    11 & University of Maryland - College Park &  36& Ohio State \\
-    12 & Northeastern University\red{$^*$} &37& University of Virginia  \\
-    13 & Purdue University &38& Yale University \\
-    14 & University of Wisconsin - Madison\red{$^*$} &39 & Univ. of California - Santa Cruz \\
-    15 & University of Texas at Austin &40& Brown University \\
-    16 & University of Pennsylvania &41 & Harvard University \\
-    17 & Columbia University\red{$^*$} &42 & Boston University  \\
-    18 & Princeton University\red{$^*$}  & 43& Rice University\\
-    19 & New York University  & 44&  University at Buffalo\red{$^*$}\\
-    20 &  University of Massachusetts-Amherst\red{$^*$} &45& University of Colorado-Boulder \\
-    21 & Univ. of California - Los Angeles &46& University of Illinois at Chicago  \\
-    22 & University of Southern California &47& University of Minnesota \\
-    23 & University of Chicago &48& University of North Carolina\red{$^*$} \\
-    24 & Stony Brook University\red{$^*$} &49& Arizona State University\red{$^*$} \\
-    25 &  Univ. of California - Irvine&50& Univ. Of California - Davis \\
+    1 & Carnegie Mellon & 26 & Northwestern University\\
+    2 & Univ. of Illinois at Urbana-Champaign\red{$^*$}  & 27 & Pennsylvania State University  \\
+    3 & Univ. of California-San Diego & 27 & Univ. of California - Riverside \\
+    4 & Georgia Institute of Technology & 29 & Univ. of California - Santa Barbara \\
+    5 & MIT                            & 30 &   Duke University \\
+    6 & University of Michigan - Ann Arbor\red{$^*$}  & 30 & Rutgers University\red{$^*$} \\
+    7 & University of Washington    & 32& University of Utah\\
+    8 & Univ. of California - Berkeley    &33 & George Mason University\red{$^*$}\\
+    9 & Cornell University   &34&   Texas A\&M University\red{$^*$}  \\
+    10 & University of Maryland - College Park & 34 &  Univ. of California - Santa Cruz \\
+    11 & Stanford University  &  36& North Carolina State University \\
+    12 & Northeastern University\red{$^*$} &37& Ohio State \\
+    13 & Purdue University &37& University of Virginia  \\
+    14 & New York University  &37 & Yale University \\
+    14 & University of Texas at Austin &40& Boston University  \\
+    16 & Princeton University\red{$^*$}&41 & Univ. Of California - Davis \\
+    16 & University of Pennsylvania  &42 & Brown University \\
+    18 & Columbia University\red{$^*$}  & 42& Harvard \\
+    19 & Univ. of California - Los Angeles  & 42& University of Illinois at Chicago  \\
+    20 &  University of Wisconsin - Madison\red{$^*$}  &45& Arizona State University\red{$^*$} \\
+    21 &  University of Southern California &45& University at Buffalo\red{$^*$} \\
+    22 & University of Massachusetts-Amherst\red{$^*$} &45& University of North Carolina\red{$^*$} \\
+    23 & University of Chicago &48& Oregon State University\red{$^*$}\\
+    24 & Stony Brook University\red{$^*$} &48& Rice University\\
+    25 &  Univ. of California - Irvine&48& University of Colorado-Boulder \\
+    & &48& University of Minnesota\\ 
     \bottomrule
@@
 \chapter{History and Acknowledgement}\label{sec:ack}
-\paragraph{History} This handbook was conceived during a lunch with a faculty at Mason.
+\paragraph{History} This book was conceived during a lunch with a faculty at Mason.
 We talked about why Mason was not able to attract good Vietnamese and other international students, despite having a much stronger CS program than many schools that these students want to go to (part of the reason is described in \autoref{sec:selecting-ranking-schools}).
@@
 There I saw many questions from students about PhD programs.
-However, most participants are students in non-CS fields or not in the US, and their answers are not always accurate and can lead to more confusion.
+However, most participants are students in non-CS fields or not in the US, and their answers are typically not accurate and often lead to more confusion.
 So I thought it would be useful to have a handbook that is specific to CS PhD programs in the US from an insider perspective.
@@
 I started writing this book in May 2023 and have been updating it since then (mostly around deadline time when I tend to procrastinate, i.e., \emph{productive procrastination}!).
-The book was initially intended for international students but has been generalized to all students.
-The handbook is open-source and available on \href{https://github.com/nguyenthanhvuh/phd-cs-us}{GitHub} so that others can contribute to it.
+The book was initially intended for international students but has been generalized to all students (domestic students also get confused by the process!).
+The book is open-source and available on \href{https://github.com/nguyenthanhvuh/phd-cs-us}{GitHub} so that others can contribute to it.
 \paragraph{About Me} I am an associate professor in the CS dept at George Mason University (Mason).
@@
 This allows me to have a good understanding of the process, the challenges students face, and what faculty are looking for.
-Currently I serve as the program director of the MS program in Software Engineering at Mason (thus also have experience with MS admission---which is quite different than PhD as discussed in \autoref{chap:ms}).
-I also have been in the panels of PhD fellowships including NSF GRFP and DoD NDSEG, which allows me to provide some insights on external fellowhip applications (\autoref{sec:fellowships}).
+Currently I am the program director of the MS program in Software Engineering at Mason (thus also have experience with MS admission---which is quite different than PhD as discussed in \autoref{chap:ms}).
+I also have served in the panels of PhD fellowships including NSF GRFP and DoD NDSEG, which allows me to provide some insights on external fellowhip applications (\autoref{sec:fellowships}).
 

### 2025-02-10 demystify.tex
updates

@@
 
-\autoref{tab:us-vs-other} summarizes the main differences between CS PhD in the US and other countries. % This is based on my experience and what I heard from others who did PhD in other countries.
+\autoref{tab:us-vs-other} summarizes the main differences between CS PhD in the US and other countries.
+Note that these differences can vary by institution and country.
+Some countries might have a PhD program that is similar to the US.
+The following are some common differences:
 
@@
 
-In other countries, you often start your research right away, e.g., you immediately work on the research project you proposed with the adviser you chose.
+In other countries, you often start your research right away and work on the research project you proposed with the adviser you chose.
 Moreover, you might not have exams like those in the US or only have to do a few of them.
@@
 \emph{Funding}:  In many countries, funding comes from the university or the gov't.
-This funding often has a fixed duration (e.g., 3 or 4 years).
+This funding often has a fixed duration, e.g., 3 or 4 years.
 In the US (\autoref{sec:funding}), funding (e.g., RA) comes directly from your adviser (no fixed duration).

### 2025-02-15 demystify.tex
updates

@@
 
-\subsection{LoR from Well-Known People}  Having a strong letter from an internationally recognized researcher can \emph{greatly strengthen} your application.
+\subsection{LoR from Well-Known People}\label{sec:famous-lor} 
+ Having a strong letter from an internationally recognized researcher can \emph{greatly strengthen} your application.
 Such a letter can outweigh the lack of publications (or low GPA or unknown school).
@@
 Again, the emphasis here is \emph{research}, i.e., the letters should describe your research experiences and potential.
-Letters focusing on your coursework or non-research projects at a company won't carry much weight.
+Letters focusing on non-research projects at a company won't carry much weight.
+
+Finally, despite best intentions, the writers might not have the experience to write a strong LoR or lack the ability to evaluate your research ability.
+This is unfortunate but common, and if you are in this situation, you should find someone else to write for you (see~\autoref{sec:lor-writers} and~\autoref{sec:famous-lor}).
+
+\begin{commentbox}[Hung:]
+    A sad reality is that most professors in Vietnam \textbf{DO NOT} know how to write a good letter, or are lazy in writing letters hence delegate the writing to the students.
+Unfortunately, there is no easy solution to this problem.
+\end{commentbox}
 
@@
 For example, if a student I don't know well asks me to write a letter for them, I will explicitly tell them I don't know them that well to write much about them, and such a short, generic, and weak letter will not help their case.
-
-\begin{commentbox}[Hung:]
-  A sad reality is that most professors in Vietnam \textbf{DO NOT} know how to write a good letter, or are lazy in writing letters hence delegate the writing to the students.
-Unfortunately, there is no easy solution to this problem.
-\end{commentbox}
 
@@
 
-The most concrete evidence of research ability is having \emph{papers in reputable international journals or conferences}.
-Having published papers, especially at top venues, is a sign that you have been successfully involved in research.
-
-Publications are \emph{never required} for application, however given the competitiveness of CS admission, they can significantly strengthen your application and \emph{are becoming the norm}.
-Applicants admitted to top schools, especially in popular fields such as ML and NLP, often have multiple first-authored papers at top places.
-Fig.~\ref{fig:wu} shows examples of applicants to Stanford CS PhD.
-
 \begin{figure}
@@
 
+The most concrete evidence of research ability is having \emph{papers in reputable international journals or conferences}.
+Having published papers, especially at top venues, is a sign that you have been successfully involved in research.
+
+Publications are \emph{never required} for application, however given the competitiveness of CS admission, they can significantly strengthen your application and \emph{are becoming the norm}.
+Applicants admitted to top schools, especially in popular fields such as ML and NLP, often have multiple first-authored papers at top places.
+Fig.~\ref{fig:wu} shows examples of applicants to Stanford CS PhD.
+
 \paragraph{What if I don't have any pubs} It is understandable that many students do not have the opportunity to publish papers.
@@
 The emphasis here is \emph{research} places, not software development or non-research work.
-For example, working at a FAANG (Facebook, Amazon, Apple, Netflix, Google) company as a software engineer does not count as research experience, although it can be helpful for MS applications discussed in \autoref{chap:ms}.
-Similarly, a LoR from your supervisor for non-research experience might not count much.
+For example, working at a FAANG (Facebook, Amazon, Apple, Netflix, Google) company as a software engineer does not count as research experience, although it can be helpful for MS applications ()\autoref{chap:ms}).
+Similarly, a LoR from your supervisor for non-research experience might not count much (\autoref{sec:admin-letters}).
 So do not spend much time talking about development job in your SOP.
@@
 
-If you do not get admitted to any schools or don't want to go to the ones that admit you, try again next time.
-Graduate admission can involve randomness and noise.
+If you do not get admitted to any schools or don't want to go to the ones that admit you, try again next year!
+Grad.
+admission can involve randomness and noise.
 In the meantime, you can work on improving your profile, e.g., get more research experiences, publish more papers, improve your connections for stronger LoRs, etc.
-See rejection reasons (\autoref{sec:why-rejected}) and additional advice to improve your chance next time (\autoref{sec:improve-your-chance}).
+See rejection reasons (\autoref{sec:why-rejected}) and additional advice to improve your chance (\autoref{sec:improve-your-chance}).
 
@@
 Many students lament that they get no interviews or are rejected and that the admission process seems random.  However, while it is true that the process is not perfect, it is not random.
-Other than not having a strong profile (e.g., research potential, GPA, LoRs, SOP, poor interviews), here are some other reasons for rejections.
+It might be that your profile is not as strong as you \emph{think} it is.  Even if it is strong, you still can get rejected, e.g., not a good fit, overqualified, having redflags, etc.
+
+\subsection{Your profile is not as strong as you think}  
+I have seen multiple cases where students think they have a strong profile, but they do not. Here are some common examples.
+
+Your \emph{LoRs} (\autoref{sec:lor-writers}) might not be as strong as you think. Your ref writers might not be very well-known, or active in research, or know how to write a good letter.
+
+Your \emph{school} might not be as good as you think.
+It might be T1 in your country (or in your own definition), but US faculty does not know about it, e.g., most US faculty would not know about schools in Vietnam.
+
+Your \emph{research} might not be as strong as you think.
+Doing research does not mean much (or anything) if there is no results, e.g,.
+publications.
+Publications might not have much value if they were published localled or through unknown places.
+Even papers at top venues might not be as good as you think, e.g., many AI/ML/NLP faculty are not too interested in ``dataset'' papers at NeurIPS.
+
+\subsection{Other Common Reasons for Rejection}
+Even with strong profile (e.g., research potential, GPA, LoRs, SOP, interviews), you can still get rejected for various other reasons as follows.
 
@@
 Do you participate in outreach activities that help increase diversity and inclusion in CS?
-Diversity is highly valued in CS programs in the US (\autoref{sec:urm}), and experiences in these areas can make you stand out and get noticed from reviewers.
+Diversity is valued in CS programs in the US (\autoref{sec:urm}), and experiences in these areas can make you stand out and get noticed from reviewers.
 

### 2025-02-15 demystify.tex
updates

@@
 Moreover, a top professor at your university might not be well-known to US faculty (see more details in \autoref{sec:admin-letters} and \autoref{chap:your-school}).
-So save the trouble and get letters from \emph{any} professors/supervisors who know you well and can write a good letter about \emph{your} research ability.
+So save the trouble and get letters from \emph{any} professors/supervisors who know you well and can write a good letter about \emph{your} research ability (\autoref{sec:famous-lor}).
 It's better to have a good personalized letter about your own research ability from someone who is less well-known than a generic/weak letter from a well-known person.
@@
 
-Your \emph{school} might not be as good as you think.
+Your \emph{school} (\autoref{chap:your-school}) might not be as good as you think.
 It might be T1 in your country (or in your own definition), but US faculty does not know about it, e.g., most US faculty would not know about schools in Vietnam.
 
-Your \emph{research} might not be as strong as you think.
+Your \emph{research} (\autoref{chap:research-experience}) might not be as strong as you think.
 Doing research does not mean much (or anything) if there is no results, e.g,.
-publications.
+publications (\autoref{sec:publications}).
 Publications might not have much value if they were published localled or through unknown places.

### 2025-02-16 demystify.tex
updates

@@
 \section{CV}
-\sectioninfo{Highlight and summarize major achievements, e.g., Publications, Programming Awards}
+\sectioninfo{Highlight and summarize major achievements such as major publications and competitive programming awards.}
 
@@
 \chapter{Interview and the Waiting Game}\label{chap:interview}
-\chapterinfo{Getting an interview is typically a good sign; but no interview does not mean rejection}
+\chapterinfo{Getting an interview is typically a good sign; but no interview does not mean rejection.}
 
@@
 \chapter{Dealing with Rejection}\label{chap:not-accepted}
-\chapterinfo{Rejection is common and part of academia.  Common reasons for rejection include aiming too high, not a good fit, or bad luck. Learn from it and try again.}
+\chapterinfo{Rejection is part of academia.  Common reasons for rejection include aiming too high, not a good fit, and bad luck.  Do not take rejection personally. Learn from it and try again.}
 

### 2025-02-16 demystify.tex
updates

@@
 
+The National Science Foundation (NSF) Graduate Research Fellowship Program (GRFP) is a prestigious fellowship for US citizens, nationals, and permanent residents who are in their first or second year of graduate study.
+The fellowship provides three years of financial support, whih includes a stipend and tuition.
+The GRFP is highly competitive, with many applicants from top schools, with strong LoR and research experience.
+
+The GRFP website has detailed information on the application process and eligibility requirements.
+This document focuses on how your application is evaluated and how you can improve your chances of success.
+
 \subsection{How Applications are Evaluated}
@@
 \begin{itemize}
-\item Each applicant is assigned to \emph{about 3 reviewers} who all seem to have expertise in the applicant's research topics.
+\item Each applicant is assigned to \emph{about 3 reviewers} who have expertise in the applicant's research topics.
 This means reviewers can be quite knowledgeable in your field and your work, your LoR writers and their work, and thus can evaluate your application quite in depth.
@@
 Note that if sometimes an original reviwer is no longer available then someone who is not in the field might review your application.
-Thus in general you want to make your application accessible to a general audience in CS.
-
-\item Reviewers attempt to compare applicants \emph{within their cohort level}.
+Thus you want to make your statements accessible to a general CS audience.
+
+\item Reviewers typically try to compare applicants \emph{within their cohort level}.
 This means that we do not compare an undergrad, who is still applying to grad school, to a 2nd year PhD student who already has an adviser and published multiple papers.
@@
 
-\item Unlike CS PhD applications, which focus on research potentials in which concrete evidence like papers, research experience, and LoRs are important, NSF GRFP also focuses on \emph{potential} but also on broader impacts, e.g., how you can help society, how you can mentor others, etc.
+\item Unlike CS PhD applications that focus on research potentials in which papers, research experience, and LoRs are important, NSF GRFP looks at both \emph{research potential} and \emph{broader impacts}, e.g., how you can help society, how you can mentor others, etc.
+It is not expected that you have published papers or have a lot of research experience (recall that MS students can also get the fellowship).
+Instead, you need to convince the reviewers that you have the potential to be a good researcher (technical merits) and that you can help society (broader impacts).
+More in \autoref{sec:grfp-im-bi}.
 
@@
 
-\subsection{Intellectual Merit and Broader Impacts}
-Your will be evaluated based on two criteria: Intellectual Merit (IM) and Broader Impacts (BI).
-
-\paragraph{Intellectual merit} is based on your research potential, e.g., your research experience, your potential to do research, and how you can contribute to the field.
-
-Note that unlike a normal project or research proposal, in which the evaluation is based on the project, in the GRFP, the evaluation is \emph{not} about the research plan.
+\subsection{Intellectual Merit and Broader Impacts}\label{sec:grfp-im-bi}
+
+Your application will be evaluated based on two criteria: Intellectual Merit (IM) and Broader Impacts (BI).
+
+\paragraph{Intellectual merit (IM)} IM evaluates your research potential, e.g., your research experience, your potential to do research, and how you can contribute to the field.
+Unlike a normal project or research proposal, in which the evaluation is entirely based on the project, in the GRFP, the evaluation is \emph{not} entirely about the research plan.
 We are repeatedly reminded to not focus too much on the research plan as many applicants are still applying for PhD programs and might not even get into the field or work with a prof.
 that fit their research plan.
+
 Instead, the evaluation is based on your \emph{potential} of being a good researcher using various criteria, e.g., your research experience, what others think of you (LoRs), research plan, etc.
@@
 
-\paragraph{Broader impacts} based on how you and your research can benefit society.
+\paragraph{Broader impacts (BM)} BM is a common and important component in NSF application.
+It is about how you and your research can \emph{benefit society}.
 Many students just mention how their research is cross-discipline and thus has broader impacts (e.g., improve software quality and thus help society).
 This is too common and not enough.
-You also want to talk about how you, as a grad student with your research, can help society, e.g., through outreach, mentoring, diversity, etc.
-As mentioned below, you will want to be specific and concrete, e.g., ``I will work with highschool students and underrepresented students'' is not as convincing as ``I have had experience in mentoring highschool students X and will continue to do so through Y,Z ...''.
+You also want to talk about how you, as a grad student with your research, can help society, e.g., through outreach, mentoring, etc.
+As mentioned below, you will want to be specific and concrete, e.g., \emph{``I will work with highschool students ..''} is not as convincing as \emph{``I have had experience in mentoring highschool students X and will continue to do so through Y,Z ...''}.
+
+Writing BM is often difficult, not only for students but also new faculty.  You should ask someone who has experience writing NSF proposals to review your BM.
 
@@
 So you will need to demonstrate your passion for research and your potential to be a good researcher.
+
+\paragraph{Overexaggeration Lors} Many students have LoRs that overrate them, e.g., \emph{``the best student I have ever had''}, \emph{``better than most PhD students''}, etc.
+This is not only not convincing but can also hurt your application.
+Reviewers are often professors themselves and know that such statements are often not true.
+Moreover, such statements can bias the evaluation, e.g., if a reviewer sees that a student is from a top school and has a LoR from a top professor, they might give the student a higher score than they deserve.
+So you should ask your LoR writers to be honest and specific, e.g., \emph{``the student did well in my class and I think they have the potential to be a good researcher''}.
 

### 2025-02-16 demystify.tex
updates

@@
 
-% Prelim work does not, and in fact should not, be extensive.
-You do not need a paper or a complete implementation.
-You don't need to show that your idea is better than others or even that it will work.
-You just need to show that you have tried it and it has some promise.
-This can be as simple as a small experiment done by hand on some small examples,  e.g., if you come up with a new abstraction in PL for proving correctness, then you can just manually apply it to a few small programs, preferrably ones that existing works have trouble with.
-This is enough to show that you have thought about the problem and have some initial results.
-
-% To make prelim concrete (and thus more convincing), you might want to use specific names and numbers, e.g.,  \emph{``I tried my prototype on the MNIST dataset, achieving an accuracy of 90\%''}, or \emph{``I used Lean to formalize and prove a specific case of the Fermat's Last Theorem''}.
-In addition, negative results are results and can be very convincing.
-So don't be afraid to say \emph{``I tried my idea on X but it failed because of Y, and I plan to address this by Z''}.
-
-\paragraph{Research} is the emphasis.
-Many made a mistake of focusing on teaching and mentoring and not on research.
-This is especially true for students who have a lot of teaching and mentoring experience, e.g., from a teaching-focused university or a teaching assistant.
-While teaching and mentoring are important, the GRFP is a \emph{research} fellowship which aims to train researchers.
-So you will need to demonstrate your passion for research and your potential to be a good researcher.
-
-\paragraph{Overexaggeration Lors} Many students have LoRs that overrate them, e.g., \emph{``the best student I have ever had''}, \emph{``better than most PhD students''}, etc.
-This is not only not convincing but can also hurt your application.
-Reviewers are often professors themselves and know that such statements are often not true.
-Moreover, such statements can bias the evaluation, e.g., if a reviewer sees that a student is from a top school and has a LoR from a top professor, they might give the student a higher score than they deserve.
-So you should ask your LoR writers to be honest and specific, e.g., \emph{``the student did well in my class and I think they have the potential to be a good researcher''}.
+\paragraph{Related Work} Many research plans motivate the problem well but do not discuss limitations of existing work, making it questionable if the problem is important or if you are aware of the challenges.
+Thus, you should do a thorough literature review and discuss what people have done and their limitations.
+Your research plan should then talk about how you plan to address these limitations, i.e., fill the gap in the literature.
+
+\paragraph{Do NOT BS} Many research plans include technical details or proposed work that are vauge, unrealistic, or even impossible.
+This is similar to a student taking an exam and write a lot of nonsense to fill the page and hope for partial credits.
+As mentioned, reviewers are often chosen based on their expertise in your field and can detect B.S.
+Again, do your homework and ask someone who knows the field, e.g., your advisor or LoR writer, to review your research plan.
+
+\paragraph{Overexaggeration Lors} Many GRFP applicants are from top schools with well-known professors that are very good at writing.
+The issue is that they often use flowery language and overrate their students.
+Reviewers are often professors themselves and can detect B.S.
+So you should ask your LoR writers to be honest, tone down the language (similar to overclaiming research results when writing papers), and be specific with concrete examples to demonstrate why you're ``the best''.
 

### 2025-02-16 demystify.tex
updates

@@
 
-\subsection{Common Pitfalls}
+\subsection{Common Pitfalls}\label{sec:grfp-pitfalls}
 
@@
 \paragraph{Overexaggeration Lors} Many GRFP applicants are from top schools with well-known professors that are very good at writing.
-The issue is that they often use flowery language and overrate their students.
-Reviewers are often professors themselves and can detect B.S.
-So you should ask your LoR writers to be honest, tone down the language (similar to overclaiming research results when writing papers), and be specific with concrete examples to demonstrate why you're ``the best''.
+The issue is that they often use flowery language and completely overrate their students.
+Reviewers are often professors themselves and can sense this overexaggeration.
+Moreover, they often know the LoR writers and can compare the LoR with the student's application and see if they match.
+So you should ask your LoR writers to be honest, tone down the language (similar to overclaiming research conributions when writing papers), and be specific with concrete examples to demonstrate why you're \emph{``the best''}.
 
@@
 Moreover, we only see \emph{redacted versions} of your submitted materials (applicants are asked to provide full and redacted versions of your research plan, CV, and other materials).
-Thus we do not know what schools your went to, where you did your internships, your publications or research experience.
-
-Thus, reviewers mainly get to know you and evaluate your application by reading your research plan and personal essays.
+Thus, we do not know what schools your went to, where you did your internships, your publications or research experience.
+
+Reviewers mainly get to know you and evaluate your application by reading your research plan and personal essays.
 This is for \emph{fairness and avoiding bias} based on the school you went to, places you published at, your LoR writers, etc.
@@
 For example, a student mentioned that they have published in the topic they proposed but redacted all information about the publication, so I have no idea where they published, what the paper is about, etc.
-It is also difficult to evaluate the academic background of the student, e.g., some student just started grad school and so their transcript is essentially blank, and I do not know where they went to undergrad, etc.
+It is also difficult to evaluate the academic background of the student, e.g., some student just started grad school and so their transcript is essentially blank, so I have no idea on their academic background.
 It is also annoying to see CVs with so many blanked out information.
@@
 In the beginning I find it surprising that I don't get to see the LoRs, but I understand the reasoning behind it and gradually think it is a good idea.
-As mentioned above, LoR writers, especially those from top schools, often overrate their students and thus can bias the evaluation.
+As mentioned above, LoR writers for NSF GRFP (\autoref{sec:grfp-pitfalls}), especially those from top schools, often overrate their students and thus can bias the evaluation.
 So I don't find I am missing much by not seeing the LoRs.
@@
 
-\subsection{Common Pitfalls and Tips}
-Based on my experience in reviewing, here are some common pitfalls to avoid:
+\subsection{Common Pitfalls and Tips}\label{sec:ndseg-pitfalls}
+
+These are some common pitfalls I see in NDSEG applications and some tips to avoid them.  While some of these are similar to NSF GRFP (\autoref{sec:grfp-pitfalls}), some are specific to NDSEG.
+
 \begin{itemize}
@@
 %- use latex,  do not write in word and especially use left justification  (see how CS and engineer people publish papers)
-\item If you have an advisor or Lor writer, ask them to look at your writing/essays.
-This is because you are new to the field and research writing, and your ideas might not be accurate or have already been explored.
 
 \end{itemize}
+
+Finally, just as with the NSF GRFP and even PhD SOPs, ask your LoR writers and adviser to look at your writing/essays.
+You are new to the field and research writing, and your ideas might not be accurate or have already been explored.
+It's better that they tell you now than for you to find out later from reviewers.
 

### 2025-02-16 demystify.tex
updates

@@
 
+%Securing a prestigious fellowship such as the National Science Foundation Graduate Research Fellowship Program (NSF GRFP) or the Department of Defense National Defense Science and Engineering Graduate Fellowship (DoD NDSEG) can significantly enhance a student’s prospects for admission to a strong PhD program.
+These fellowships not only provide financial support but also offer academic freedom and prestige that can set an applicant apart from their peers.
+However, successfully obtaining one of these fellowships is no easy feat—they are highly competitive, and applications are scrutinized rigorously by expert reviewers.
+
+%One of the biggest advantages that applicants from elite institutions often have is access to faculty members who are well-versed in the review process.
+These faculty members understand the nuances of writing strong letters of recommendation (LoRs) and guiding students through the proposal-writing process.
+Unfortunately, students from smaller institutions or those with mentors unfamiliar with NSF or DoD fellowship reviews may find themselves at a disadvantage—not because of their potential or research ability, but simply due to a lack of insider knowledge about what makes a strong application.
+
+%This chapter aims to level the playing field by providing an insider’s perspective on how these applications are reviewed.
+Drawing from my experience as a reviewer for both NSF GRFP and DoD NDSEG, I will demystify the evaluation process, highlight key elements that make applications stand out, and offer strategic advice for applicants and their letter writers.
+While research experience and proposal quality are, of course, crucial to success, understanding the review criteria and expectations can be just as important.
+
+%Moreover, this chapter will provide guidance for LoR writers, helping them craft letters that effectively support their students’ applications.
+Many well-intentioned letters fail to address the aspects that reviewers prioritize, which can unintentionally weaken an otherwise strong application.
+By offering insights into what reviewers look for, I hope to empower both applicants and their mentors to present their strongest possible case for these fellowships.
+
+%Whether you are an aspiring PhD student preparing your application or a faculty member writing a recommendation, this chapter will equip you with the knowledge needed to navigate the NSF GRFP and DoD NDSEG review process with confidence.
+
 %To give students as well as their LoR writer on insider view.
@@
 
-The GRFP website has detailed information on the application process and eligibility requirements.
-This document focuses on how your application is evaluated and how you can improve your chances of success.
+The \href{https://www.nsfgrfp.org/}{GRFP website} has detailed information on the application process and eligibility requirements.
+This document focuses on how your application is evaluated and tips on improving your application.
 
@@
 
-\subsection{Common Pitfalls}\label{sec:grfp-pitfalls}
-
-\paragraph{Too technical} Remember that you're writing to a general audience in CS, not experts in your field.
-While NSF tries to match reviewers with your field, CS fields are simply too broad and you might get a reviewer in the field but not in your specific area.
-So avoid technical jargon and explain things in simple terms. If the reviewer does not understand your proposal, they will likely give you a low score.
+\subsection{Common Pitfalls and Tips}\label{sec:grfp-pitfalls}
+
+These are common pitfalls that many applicants make in their GRFP applications and tips on how to avoid them.
+
+\begin{itemize}
+
+\item \textbf{Too Technical.} Remember that you're writing to a general audience in CS, not experts in your field.
+While NSF tries to match reviewers with your field, CS fields are simply too broad and you might get a reviewer in the field but not in your specific area or familiar with the problem you're working on.
+So avoid technical jargon and explain things in simple terms. If the reviewer cannot understand your proposal, they will likely give you a low score.
 
@@
 
-\paragraph{Preliminary Work} Your research plan should have some preliminary work to convince reviewers that you have thought about the problem and have some initial promising results.
+\item \textbf{Preliminary Work.} Your research plan should have some preliminary work to convince reviewers that you have thought about the problem and have some initial promising results.
 It does not have to a lot, e.g., you don't need published results or an implementation, a small experiment done by hand on some small examples would suffice.
@@
 
-\paragraph{Related Work} Many research plans motivate the problem well but do not discuss limitations of existing work, making it questionable if the problem is important or if you are aware of the challenges.
+\item \textbf{Related Work and Challenges.} Many research plans motivate the problem well but do not discuss limitations of existing work, making it questionable if the problem is important or if you are aware of the challenges.
 Thus, you should do a thorough literature review and discuss what people have done and their limitations.
@@
 
-\paragraph{Do NOT BS} Many research plans include technical details or proposed work that are vauge, unrealistic, or even impossible.
-This is similar to a student taking an exam and write a lot of nonsense to fill the page and hope for partial credits.
+\item \textbf{Do not BS.} Many research plans include technical details or proposed work that are vauge, unrealistic, or even impossible.
+This is similar to a student taking an exam and write a lot of nonsense and hope for partial credits.
 As mentioned, reviewers are often chosen based on their expertise in your field and can detect B.S.
@@
 
-\paragraph{Overexaggeration Lors} Many GRFP applicants are from top schools with well-known professors that are very good at writing.
-The issue is that they often use flowery language and completely overrate their students.
-Reviewers are often professors themselves and can sense this overexaggeration.
-Moreover, they often know the LoR writers and can compare the LoR with the student's application and see if they match.
-So you should ask your LoR writers to be honest, tone down the language (similar to overclaiming research conributions when writing papers), and be specific with concrete examples to demonstrate why you're \emph{``the best''}.
+\item \textbf{Overexaggeration Lors.} Many GRFP applicants are from top schools with well-known professors that are very good at writing.
+However, they often use flowery language and completely overrate their students.
+Reviewers are  professors themselves and can sense this overexaggeration.
+So you should ask your LoR writers to tone down the language (similar to overclaiming research conributions when writing papers) and and be specific with concrete examples to demonstrate why you're \emph{``the best''}.
+
+\end{itemize}
 
@@
 citizens} pursuing advanced degrees in STEM fields \emph{critical to national defense} (such as CS).
-It provides up to three years of funding, including a  stipend, tuition coverage, and health insurance.
+It provides up to three years of funding, including a  stipend and tuition coverage.
 
@@
 
-\paragraph{Redacted info.} Unlike the NSF GRFP, for NDSEG reviewers \emph{do not} see your LoRs, i.e., we do not know who wrote them or what they wrote.
-Moreover, we only see \emph{redacted versions} of your submitted materials (applicants are asked to provide full and redacted versions of your research plan, CV, and other materials).
+\paragraph{Redacted info.} Unlike the NSF GRFP, reviewers only see \emph{redacted versions} of your submitted materials (applicants are asked to provide full and redacted versions of your research plan, CV, and other materials).
 Thus, we do not know what schools your went to, where you did your internships, your publications or research experience.
+We also \emph{do not} see your LoRs, i.e., we do not know who wrote them or what they wrote.
 
@@
 For example, a student mentioned that they have published in the topic they proposed but redacted all information about the publication, so I have no idea where they published, what the paper is about, etc.
-It is also difficult to evaluate the academic background of the student, e.g., some student just started grad school and so their transcript is essentially blank, so I have no idea on their academic background.
+It is also difficult to evaluate the academic background of the student, e.g., some student just started grad school and so their transcript is essentially blank.
 It is also annoying to see CVs with so many blanked out information.
@@
 
-These are some common pitfalls I see in NDSEG applications and some tips to avoid them.  While some of these are similar to NSF GRFP (\autoref{sec:grfp-pitfalls}), some are specific to NDSEG.
+These are some common pitfalls I see in NDSEG applications and some tips to avoid them.  Some of these are similar to NSF GRFP (\autoref{sec:grfp-pitfalls}) while some are specific to NDSEG.
 
@@
 
-    \item Customize your research to DoD, e.g., through specific project from DARPA or ONR. 
+    \item \textbf{Customize your research to DoD}, e.g., through specific project from DARPA or ONR. 
     Make this prominent, e.g., in the introduction or in its own section.  This is different from NSF GRFP where you want broader impacts to society.
 
-    \item Unclear research plan. 
+    \item \textbf{Unclear research plan.}
      You should aim to have (i) clear problem statement, (ii) why it is important (esp.
 to DoD), (iii) what has been done and their challenges/limitations, and (iv) what you plan to do to address the challenges.
-Moreover, research is often quite technical and ambitious, so you want to show some preliminary work, e.g., a small experiment by hand on some small examples, to convince reviewers that you have thought about the problem and have some initial results.\\
+Moreover, research is often quite technical and ambitious, so you want to show some preliminary work, e.g., a small experiment by hand on some small examples, to convince reviewers that you have thought about the problem and have some initial results.
 
@@
 
-    \item Personal stories and essays should be unique and interesting.
+    \item \textbf{Personal stories and essays should be unique and interesting}.
 Many essays and stories I read are quite generic and do not tell much about the student.
-For example, being an URM, involving in a robotics club or mentoring underrepresented students are quite common and do not tell much.\\
+For example, being an URM, involving in a robotics club or mentoring underrepresented students are quite common and do not tell much.
 
@@
 
-    \item Not using all allowed pages.
+    \item \textbf{Not using all allowed pages}.
 Many research plans use only 1/2 or 2/3 of the number of allowed pages, which themselves are already quite short.

### 2025-02-16 demystify.tex
updates

@@
 So I want to share some details about the admission process and advice for those who are interested in applying for a \textbf{PhD in Computer Science in the US}.
-Originally, this book was intended for international students, but has been expanded to include information for \textbf{US domestic students} (\autoref{sec:domestic-students}).
+Originally, this book was written for international students, but has been expanded to include information for \textbf{US domestic students} (\autoref{chap:domestic-students}).
 Moreover, while this primarily aims at students interested in CS, it might be relevant to students from various STEM (Science, Technologies, Engineering, and Mathematics) disciplines.
@@
 
-\chapter{Should You Apply?}\label{sec:should}
+\chapter{Should You Apply?}\label{chap:should}
 
@@
 If you are admitted, you will almost certainly
-receive \emph{full funding} (\autoref{sec:funding}) to support your study. This includes tuition,
+receive \emph{full funding} (\autoref{chap:funding}) to support your study. This includes tuition,
 health insurance, and stipend (i.e., in STEM field you get paid for your study!). 
@@
 
-%While this handbook may not fully meet the needs of students aiming for top-10 schools (\autoref{sec:selecting-ranking-schools} and \autoref{sec:ranking}), which typically have more stringent requirements (e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU), it should be valuable for those targeting other reputable R1 institutions.
+%While this handbook may not fully meet the needs of students aiming for top-10 schools (\autoref{sec:selecting-ranking-schools} and \autoref{chap:rankings}), which typically have more stringent requirements (e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU), it should be valuable for those targeting other reputable R1 institutions.
 
 %\begin{commentbox}[Vu:]
-%    One of the reasons I created this document is that my colleagues at Mason are interested in recruiting Vietnamese students and are surprised when seeing very few applications in Vietnam (see \autoref{sec:ack}).
+%    One of the reasons I created this document is that my colleagues at Mason are interested in recruiting Vietnamese students and are surprised when seeing very few applications in Vietnam (see \autoref{chap:ack}).
 Each year our CS PhD program receives 500+ PhD applications, most of which are international but only 5--6 are from Vietnam.
@@
 This funding often has a fixed duration, e.g., 3 or 4 years.
-In the US (\autoref{sec:funding}), funding (e.g., RA) comes directly from your adviser (no fixed duration).
+In the US (\autoref{chap:funding}), funding (e.g., RA) comes directly from your adviser (no fixed duration).
 There are also fewer TA opportunities in European universities compared to the US.
@@
 
-\chapter{How is Your Application Evaluated?}\label{sec:evalapps}
+\chapter{How is Your Application Evaluated?}\label{chap:evalapps}
 
@@
 \begin{commentbox}{Vu:}
-At Mason, we usually decide that a full-time PhD candidate is either (i) admitted with funding (\autoref{sec:funding}) or (ii) rejected. In other words, in most cases, we either
+At Mason, we usually decide that a full-time PhD candidate is either (i) admitted with funding (\autoref{chap:funding}) or (ii) rejected. In other words, in most cases, we either
 admit you with full funding or reject your application. In some rare cases, we admit
@@
 
-Some universities do waive, e.g., Rice, TTIC do not have fee for PhD applications and many universities for domestic students (\autoref{sec:domestic-students}).
+Some universities do waive, e.g., Rice, TTIC do not have fee for PhD applications and many universities for domestic students (\autoref{chap:domestic-students}).
 Some programs also waive if the applicants attend some of their opening sessions.
@@
 % In CS, conferences are important and top conferences are prestigious.
-You can find top CS conferences at places such as \href{https://csrankings.org}{CSRankings}, which ranks CS programs based on how their faculty publish at top \emph{conferences} (see more in \autoref{sec:ranking}).
+You can find top CS conferences at places such as \href{https://csrankings.org}{CSRankings}, which ranks CS programs based on how their faculty publish at top \emph{conferences} (see more in \autoref{chap:rankings}).
 
@@
 statement of purpose (SOP) or personal statement\footnote{Few schools separate these documents and ask you to write both: SOP, which focuses on research experiences, and personal statement, which is everything more personal, e.g., why PhD, challenges, diversity, etc}!
-A well-written SOP also shows that you can communicate, which is very important in research, and that you can effectively teach and communicate with students, which is important for TA funding (see \autoref{sec:funding}).
+A well-written SOP also shows that you can communicate, which is very important in research, and that you can effectively teach and communicate with students, which is important for TA funding (see \autoref{chap:funding}).
 Many SOP samples for CS are \href{https://cs-sop.org/}{available here}.
@@
 This can raise concerns about your ability to handle the PhD.
-Don't talk about religious or political beliefs (\autoref{sec:cultural}).
+Don't talk about religious or political beliefs (\autoref{chap:cultural}).
 \item \textbf{Criticizing} your current or previous institution or professors. Just like in a job interview, don't badmouth your current or previous employer. 
@@
 On one hand you will need to show some level of English proficiency, but on the other hand, you do not need to have very good scores in these tests (many adcom members themselves were once international students and struggled with English).
-You should just do well enough to pass the minimum requirement set by the university. %, which is required for TA (\autoref{sec:funding}).
+You should just do well enough to pass the minimum requirement set by the university. %, which is required for TA (\autoref{chap:funding}).
 Just as with grades (\autoref{chap:your-school}) and GRE (\autoref{sec:gre}), having high scores in these tests might not help, but having too low scores can be a \red{red flag} and sometimes results in an automatic rejection (\autoref{sec:why-rejected}), e.g., below the minimum requirement.
@@
 
-\chapter{Getting Admitted}\label{sec:accepted}
+\chapter{Getting Admitted}\label{chap:accepted}
 \chapterinfo{Congrats! Now it is your turn to evaluate the school!
@@
 Now you're at a different game because the schools that have admitted you will try to get you to accept them!
-Important factors to consider include the reputation of schools and professors (\autoref{sec:schoolsandprofs}), and funding availability (\autoref{sec:funding}).
+Important factors to consider include the reputation of schools and professors (\autoref{chap:schoolsandprofs}), and funding availability (\autoref{chap:funding}).
 You will likely have to make your decision (\autoref{sec:accept-postpone-decline}) by around \emph{April 15}, which is the deadline for most schools
@@
 
-\chapter{Funding}\label{sec:funding}
+\chapter{Funding}\label{chap:funding}
 \chapterinfo{TAs, RAs, and fellowships are main funding sources for PhDs.
@@
   Having health insurance is required at many US universities.
-Do not assume that you're young and healthy and ignore insurance (\autoref{sec:cultural-misc}).
+Do not assume that you're young and healthy and ignore insurance (\autoref{chap:cultural-misc}).
 At Mason, and most CS PhD programs, your GTA or GRA comes with full insurance.
@@
 
-For external fellowships including those from the US government (e.g., NSF, DOD) and tech companies (e.g., Google, Microsoft, Facebook), you will need to apply.
+For external fellowships including those from the US government (e.g., NSF GRFP, DOD NDSEG) and tech companies (e.g., Google, Microsoft, Facebook), you will need to apply.
 % However, tech companies including Google, Microsoft, Facebook have fellowships that international students can apply for.
-Prestigious external fellowships typically require a clear and good research plan (the GRFP also requires outreach activities and plans).
+Prestigious external fellowships typically require a clear and good research plan (the GRFP also requires broader impacts discussion).
 So it might be a good idea to wait until your second year to have research experience and even publication before applying.
@@
 
-In addition to general fellowships that are available to all students, there are also specific fellowships, e.g., for US citizens and permanent residents (\autoref{sec:domestic-students}), for underrepresented minority (URM) students (\autoref{sec:urm}), and students from specific countries.
+In addition to general fellowships that are available to all students, there are also specific fellowships, e.g., for US citizens and permanent residents (\autoref{chap:domestic-students}), for underrepresented minority (URM) students (\autoref{sec:urm}), and students from specific countries.
 There are many fellowships available, so you should look for them and apply.
@@
 In fact, these are so prestigious that even if you didn't get it but make it to the final round or even \emph{``honorable mentioning''}, your school will still mention you on their website, and you should put it on your CV.
+\autoref{chap:fellowships} discusses the evaluation processes of the NSF GRFP and DoD NDSEG.
 
@@
 
-\chapter{Choosing Schools}\label{sec:schoolsandprofs}
+\chapter{Choosing Schools}\label{chap:schoolsandprofs}
 \chapterinfo{Not every university has a CS PhD program. Not every professor, even those in CS, can advise or graduate CS PhD students.}
@@
 While this can be confusing due to the large number of universities in the US, a little research, e.g., searching for PhD in CS from the school website, will help you find out.
-Schools listed in \autoref{sec:ranking} have PhD programs in CS, so you can start there.
+Schools listed in \autoref{chap:rankings} have PhD programs in CS, so you can start there.
 
@@
 This is also a good way to learn about individual faculty, e.g., who works on what, and well-known CS conferences\footnote{In CS (and probably only in CS), conferences, not journals, are often the main venue to publish research findings (see \href{https://homes.cs.washington.edu/~mernst/advice/conferences-vs-journals.html}{why here}).}.
-%\autoref{sec:ranking} gives the top 50 CS programs in the US according to CSRankings.
+%\autoref{chap:rankings} gives the top 50 CS programs in the US according to CSRankings.
 
@@
 
-If you get admission to several places, you should consider attending Open Houses (\autoref{sec:accepted}) and contact profs.
+If you get admission to several places, you should consider attending Open Houses (\autoref{chap:accepted}) and contact profs.
 that you're interested in at those places and talk to them.
@@
 They typically teach 3--4 classes per semester (which is quite a lot) and \emph{do not} have research responsibilities, e.g., they do not have to worry much about publishing papers or obtaining grants.
-As such they typically do not have funding (e.g., RA~\autoref{sec:funding}) to support PhD students.
+As such they typically do not have funding (e.g., RA~\autoref{chap:funding}) to support PhD students.
 Depending on the university, they might be able to \emph{co-advise} and mentor PhD students with a tenure-line faculty, but they typically do not formally advise and graduate PhD students.
@@
 to \emph{ask about your admission chance}, please \textbf{don't}.
-We don't know and can't answer because as explained in \autoref{sec:evalapps}, we don't make individual decisions and might not even be assigned to evaluate your application.
+We don't know and can't answer because as explained in \autoref{chap:evalapps}, we don't make individual decisions and might not even be assigned to evaluate your application.
 It is the same as sending a paper draft to a journal editor and asking them if your paper has a chance.
@@
 \part{Miscs and FAQs}\label{part:faqs}
-\chapter{Miscs and FAQs}\label{sec:faqs}
+\chapter{Miscs and FAQs}\label{chap:faqs}
 
@@
 Finally, I enter my decision, which is usually either a rejection or an offer of admission with full funding (e.g., from TA).
-I also recommend very strong candidates for the University Presidential fellowship., which is a fellowship from GMU that provides funding similar to an RA (\autoref{sec:funding}).
+I also recommend very strong candidates for the University Presidential fellowship., which is a fellowship from GMU that provides funding similar to an RA (\autoref{chap:funding}).
 Note that while the system has other decision options, eg., admit without funding, provisional admission (e.g., if they need to take some courses), I do not use them, simply because we either reject or admit with funding.
@@
 These nuances represent some cultural and academic differences that you may encounter and will gradually adapt to.
-More on these differences in \autoref{sec:cultural}.
+More on these differences in \autoref{chap:cultural}.
 
@@
 
-  \item[Adcom~\autoref{sec:evalapps}] Admission Committee. The group of faculty members who review applications and make admission decisions.
-
-  \item [Adcom members~\autoref{sec:evalapps}] Faculty members who are part of the admission committee. These are the people who review applications and make admission decisions.
-
-  \item[Adcom chair~\autoref{sec:evalapps}] The faculty member who leads the admission committee.
+  \item[Adcom~\autoref{chap:evalapps}] Admission Committee. The group of faculty members who review applications and make admission decisions.
+
+  \item [Adcom members~\autoref{chap:evalapps}] Faculty members who are part of the admission committee. These are the people who review applications and make admission decisions.
+
+  \item[Adcom chair~\autoref{chap:evalapps}] The faculty member who leads the admission committee.
 This person is often not involved in individual admission decisions but oversees the process (e.g., resolve disputes, ensure fairness).
@@
 Domestic students] International students are loosely defined as those who would need a visa to study in the US (\autoref{chap:visa}).
-Domestic students (\autoref{sec:domestic-students}) are those who do not need a visa, typically US citizens or permanent residents.
+Domestic students (\autoref{chap:domestic-students}) are those who do not need a visa, typically US citizens or permanent residents.
 
@@
 
-  \item[Open House \autoref{sec:accepted}] An event where admitted students visit the university to meet faculty, students, and staff, and learn more about the program.
+  \item[Open House \autoref{chap:accepted}] An event where admitted students visit the university to meet faculty, students, and staff, and learn more about the program.
 This is a great opportunity to get a feel for the university and the department and decide if it is a good fit for you.
@@
 
-\chapter{Domestic Students}\label{sec:domestic-students}
+\chapter{Domestic Students}\label{chap:domestic-students}
 \chapterinfo{Specific benefits and opportunities for domestic students applying to CS PhD programs.}
@@
 \paragraph{School \autoref{chap:your-school}} Adcom already knows about your school, which is a plus.
-You are also more familiar with the US education system and academic culture (\autoref{sec:cultural}).
+You are also more familiar with the US education system and academic culture (\autoref{chap:cultural}).
 
@@
 
-\paragraph{Funding \autoref{sec:funding}} You have more opportunities for funding, e.g., through government scholarships for US citizens and permanent residents.
+\paragraph{Funding \autoref{chap:funding}} You have more opportunities for funding, e.g., through government scholarships for US citizens and permanent residents.
 You can also apply to specific programs \emph{before} you start your PhD, e.g., NSF Graduate Research Fellowship Program (GRFP) and Hertz Foundation Fellowship.
-These fellowships are very competitive and can significantly improve your admission chances.
+These fellowships are very competitive (\autoref{chap:fellowships}) and can significantly improve your admission chances.
 
@@
 
-\paragraph{Open House \autoref{sec:accepted}} It is easier for you to attend open house events in person.  This can help you make a better decision on which school to attend.
+\paragraph{Open House \autoref{chap:accepted}} It is easier for you to attend open house events in person.  This can help you make a better decision on which school to attend.
 
@@
 
-  \item Funding: MS is typically \emph{not funded}, while PhD is (\autoref{sec:funding}). See \autoref{sec:ms-funding} for more details on MS funding.
+  \item Funding: MS is typically \emph{not funded}, while PhD is (\autoref{chap:funding}). See \autoref{sec:ms-funding} for more details on MS funding.
 
@@
 \subsection{Funding}\label{sec:ms-funding}
-Unlike PhD programs, which often have funding (\autoref{sec:funding}), MS students are typically \emph{not} funded.
+Unlike PhD programs, which often have funding (\autoref{chap:funding}), MS students are typically \emph{not} funded.
 This is because MS students do not focus on research and thus are not funded through RA and PhD students have priority for TA positions.
@@
 Because of the differences between MS and PhD programs (\autoref{sec:phd-vs-ms}), you should consider different factors when selecting and ranking MS programs.
-For example, you might want to consider factors such as location, industry connections, and job placement instead of research areas and advisers (\autoref{sec:schoolsandprofs}).
+For example, you might want to consider factors such as location, industry connections, and job placement instead of research areas and advisers (\autoref{chap:schoolsandprofs}).
 
@@
 
-\chapter{Fellowship Applications}\label{sec:fellowship}
-\chapterinfo{How to apply for fellowships such as NSF GRFP, DOD DSEG to fund your PhD.}
+\chapter{Fellowship Applications}\label{chap:fellowships}
+\chapterinfo{How fellowships such as NSF GRFP and DOD DSEG are evaluated.}
+
+Having an external and major fellowship can significantly improve your chances of getting into a good PhD program (\autoref{sec:fellowships}).
+The National Science Foundation Graduate Research Fellowship Program (NSF GRFP) and the Department of Defense National Defense Science and Engineering Graduate Fellowship (DoD NDSEG) are two of the most prestigious fellowships for CS PhD students who are US citizens or permanent residents.
+It is rumored that several universities would revert their admission decision on a student if they get one of these fellowships.
+
+The \href{https://www.nsfgrfp.org/}{GRFP} and \href{https://www.ndsegfellowships.org/}{NDSEG} website have detailed information on the application process and eligibility requirements.
+This section focuses on how your application is evaluated and tips on improving your application.
+Note that while these are specific GRFP and NDSEG, the general advice can be applied to other major fellowships (those for non-US citizens or permanent residents).
 
@@
 
-The National Science Foundation (NSF) Graduate Research Fellowship Program (GRFP) is a prestigious fellowship for US citizens, nationals, and permanent residents who are in their first or second year of graduate study.
-The fellowship provides three years of financial support, whih includes a stipend and tuition.
-The GRFP is highly competitive, with many applicants from top schools, with strong LoR and research experience.
-
-The \href{https://www.nsfgrfp.org/}{GRFP website} has detailed information on the application process and eligibility requirements.
-This document focuses on how your application is evaluated and tips on improving your application.
+The NSF GRFP is a prestigious fellowship for US citizens, nationals, and permanent residents who are in the early stages of their graduate study in STEM fields, including both MS and PhD students.
+The fellowship provides three years of financial support, which includes a stipend and tuition.
+The GRFP is highly competitive, with many applicants from top schools, with strong LoRs and research experience.
 
@@
 
-The DoD National Defense Science and Engineering Graduate (NDSEG) is another prestigious fellowship supporting \emph{U.S.
-citizens} pursuing advanced degrees in STEM fields \emph{critical to national defense} (such as CS).
+The NDSEG is another prestigious fellowship supporting \emph{U.S. citizens} pursuing advanced degrees in STEM fields \emph{critical to national defense} (such as CS).
 It provides up to three years of funding, including a  stipend and tuition coverage.
 
-You can read more details about NDSEG and its guideline for applicants from the \href{https://www.ndsegfellowships.org/}{official website}.
-Here, I mainly focus on things that stand out to me as reviewers and also compare them with NSF GRFP as many students apply to both.
 \subsection{Evaluation}
@@
 
-\chapter{Academic and Cultural Differences}\label{sec:cultural}
+\chapter{Academic and Cultural Differences}\label{chap:cultural}
 
@@
 
-  \section{Miscs}\label{sec:cultural-misc}
+  \section{Miscs}\label{chap:cultural-misc}
 
@@
   \paragraph{Healthcare System}  You (and your spouse) will need health insurance!
-Otherwise you will be charged a lot for healthcare services.
-However, as mentioned in \autoref{sec:funding}, your  TA/RA  will cover health insurance (and your spouse/children also get insurance or discount for it).
-
-  However, be aware that heathcare services can still be expensive even with insurance.
-So  know what your insurance covers and be prepared for unexpected costs.
+Otherwise you will be charged a lot for healthcare services when you need them.
+However, as mentioned in \autoref{chap:funding}, your  TA/RA (and fellowships)  will cover health insurance.
+Your spouse/children also get health insurance or significant discounts under your plan.
+
+  Note that even with insurance, heathcare services might still be expensive.
+So you should know what your insurance covers and be prepared for unexpected costs.
 Moreover, healthcare system has many confusing jargons such as HMO, PPO, deductibles, co-pays, and coinsurance (try take a look at the Explanation of Benefits or EoB statement you received from your insurance company).
 It's arguably the most complicated system in the US and even Americans often do not understand it (and politicians often exploit this to their advantage).
-So do not hesitate to ask your HR or the insurance company for help.
+Do not hesitate to ask your HR or the insurance company for help.
 
@@
 So adding an extra 15--20\% to your bills is common, especially in restaurants.
+
   You should also tip for other services, e.g., Uber, taxi, haircuts, and hotel services.
@@
 
-\chapter{CSRankings: Rankings of CS PhD programs}\label{sec:ranking}
+\chapter{CSRankings: Rankings of CS PhD programs}\label{chap:rankings}
 \chapterinfo{CSRankings.org is a ranking system based on faculty publications at top CS conferences.}
@@
 %\autoref{tab:ranking} lists .
-\chapter{History and Acknowledgement}\label{sec:ack}
+\chapter{History and Acknowledgement}\label{chap:ack}
 \paragraph{History} This book was conceived during a lunch with a faculty at Mason.
@@
 There I saw many questions from students about PhD programs.
-However, most participants are students in non-CS fields or not in the US, and their answers are typically not accurate and often lead to more confusion.
+However, most participants are students, many of whom in non-CS fields or not in the US, and their answers are typically not accurate and often lead to more confusion.
 So I thought it would be useful to have a handbook that is specific to CS PhD programs in the US from an insider perspective.
 
-I started writing this book in May 2023 and have been updating it since then (mostly around deadline time when I tend to procrastinate, i.e., \emph{productive procrastination}!).
-The book was initially intended for international students but has been generalized to all students (domestic students also get confused by the process!).
+I started writing this book in May 2023 and have been updating it since then (mostly around deadline time when I procrastinate, i.e., \emph{productive procrastination}!).
+The book was initially intended for international students but has been updated to include domestic students (\autoref{chap:domestic-students} and \autoref{chap:fellowships}).
 The book is open-source and available on \href{https://github.com/nguyenthanhvuh/phd-cs-us}{GitHub} so that others can contribute to it.
@@
 Currently I am the program director of the MS program in Software Engineering at Mason (thus also have experience with MS admission---which is quite different than PhD as discussed in \autoref{chap:ms}).
-I also have served in the panels of PhD fellowships including NSF GRFP and DoD NDSEG, which allows me to provide some insights on external fellowhip applications (\autoref{sec:fellowships}).
+I also have served in the panels of PhD fellowships including NSF GRFP and DoD NDSEG, which allows me to provide some insights on external fellowhip applications (\autoref{chap:fellowships}).
 
@@
 I hope to capture the diverse challenges and experiences they've faced in this book so that it can be a valuable resource for prospective international students.
-Finally, my upbringing in the US provides a perspective aligned with American culture, allowing me to shed light on various issues, particularly those related to cultural differences (\autoref{sec:cultural}).
+Finally, my upbringing in the US provides a perspective aligned with American culture, allowing me to shed light on various issues, particularly those related to cultural differences (\autoref{chap:cultural}).
 
@@
 
-Also thanks to NSF for encouraging faculty to be creative in research and education, which allows me to write this handbook.
+Also thanks to NSF for encouraging faculty to be creative in research and education, which allows me to work on this book.
 

### 2025-02-16 demystify.tex
updates

@@
 The National Science Foundation Graduate Research Fellowship Program (NSF GRFP) and the Department of Defense National Defense Science and Engineering Graduate Fellowship (DoD NDSEG) are two of the most prestigious fellowships for CS PhD students who are US citizens or permanent residents.
-It is rumored that several universities would revert their admission decision on a student if they get one of these fellowships.
+These fellowship provide multiple years of full financial support, which includes a stipend and tuition. 
+It is rumored that several programs would revert their admission decision on a student if they get one of these fellowships.
 
@@
 The NSF GRFP is a prestigious fellowship for US citizens, nationals, and permanent residents who are in the early stages of their graduate study in STEM fields, including both MS and PhD students.
-The fellowship provides three years of financial support, which includes a stipend and tuition.
 The GRFP is highly competitive, with many applicants from top schools, with strong LoRs and research experience.
@@
 
-\item \textbf{Overexaggeration Lors.} Many GRFP applicants are from top schools with well-known professors that are very good at writing.
+\item \textbf{Overexaggeration Lors.} Many GRFP applicants have LoRs from professors that are very good at writing.
 However, they often use flowery language and completely overrate their students.
-Reviewers are  professors themselves and can sense this overexaggeration.
-So you should ask your LoR writers to tone down the language (similar to overclaiming research conributions when writing papers) and and be specific with concrete examples to demonstrate why you're \emph{``the best''}.
+Reviewers can sense this overexaggeration and might not trust the LoRs.
+This is similar to overclaiming research conributions when writing papers.
+So you should mention this to your LoR writers so that they can tone down the language and and be specific with concrete examples to demonstrate why you're \emph{``the best''}.
 
@@
 
-The NDSEG is another prestigious fellowship supporting \emph{U.S. citizens} pursuing advanced degrees in STEM fields \emph{critical to national defense} (such as CS).
-It provides up to three years of funding, including a  stipend and tuition coverage.
+The NDSEG is another prestigious fellowship supporting \emph{U.S.
+citizens} pursuing advanced degrees in STEM fields \emph{critical to national defense} (such as CS).
+NDSEG has several differences from NSF GRFP, e.g., it redacts more information from the reviewers and focuses more on defense-related research.
 

### 2025-02-16 demystify.tex
updates

@@
 \paragraph{Intellectual merit (IM)} IM evaluates your research potential, e.g., your research experience, your potential to do research, and how you can contribute to the field.
-Unlike a normal project or research proposal, in which the evaluation is entirely based on the project, in the GRFP, the evaluation is \emph{not} entirely about the research plan.
-We are repeatedly reminded to not focus too much on the research plan as many applicants are still applying for PhD programs and might not even get into the field or work with a prof.
+Unlike a normal project or research proposal, in which the evaluation is entirely based on the merits of the proposed work, in the GRFP, the evaluation is \emph{not} entirely about the research plan.
+Reviewers are repeatedly reminded to not focus too much on the research plan as many applicants are still applying for PhD programs and might not even get into the field or work with a prof.
 that fit their research plan.
@@
 citizens} pursuing advanced degrees in STEM fields \emph{critical to national defense} (such as CS).
-NDSEG has several differences from NSF GRFP, e.g., it redacts more information from the reviewers and focuses more on defense-related research.
+NDSEG has several differences from NSF GRFP, e.g., it redacts more information from the reviewers and focuses more on application to DoD.
 

### 2025-02-17 demystify.tex
updates

--- 2025-02-16 demystify.tex
@@
 So I want to share some details about the admission process and advice for those who are interested in applying for a \textbf{PhD in Computer Science in the US}.
-Originally, this book was written for international students, but has been expanded to include information for \textbf{US domestic students} (\autoref{chap:domestic-students}).
+Originally, this book was intended for international students, but has been expanded to include information for \textbf{US domestic students} (\autoref{sec:domestic-students}).
 Moreover, while this primarily aims at students interested in CS, it might be relevant to students from various STEM (Science, Technologies, Engineering, and Mathematics) disciplines.
@@
 
-\chapter{Should You Apply?}\label{chap:should}
+\chapter{Should You Apply?}\label{sec:should}
 
@@
 If you are admitted, you will almost certainly
-receive \emph{full funding} (\autoref{chap:funding}) to support your study. This includes tuition,
+receive \emph{full funding} (\autoref{sec:funding}) to support your study. This includes tuition,
 health insurance, and stipend (i.e., in STEM field you get paid for your study!). 
@@
 
-%While this handbook may not fully meet the needs of students aiming for top-10 schools (\autoref{sec:selecting-ranking-schools} and \autoref{chap:rankings}), which typically have more stringent requirements (e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU), it should be valuable for those targeting other reputable R1 institutions.
+%While this handbook may not fully meet the needs of students aiming for top-10 schools (\autoref{sec:selecting-ranking-schools} and \autoref{sec:ranking}), which typically have more stringent requirements (e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU), it should be valuable for those targeting other reputable R1 institutions.
 
 %\begin{commentbox}[Vu:]
-%    One of the reasons I created this document is that my colleagues at Mason are interested in recruiting Vietnamese students and are surprised when seeing very few applications in Vietnam (see \autoref{chap:ack}).
+%    One of the reasons I created this document is that my colleagues at Mason are interested in recruiting Vietnamese students and are surprised when seeing very few applications in Vietnam (see \autoref{sec:ack}).
 Each year our CS PhD program receives 500+ PhD applications, most of which are international but only 5--6 are from Vietnam.
@@
 This funding often has a fixed duration, e.g., 3 or 4 years.
-In the US (\autoref{chap:funding}), funding (e.g., RA) comes directly from your adviser (no fixed duration).
+In the US (\autoref{sec:funding}), funding (e.g., RA) comes directly from your adviser (no fixed duration).
 There are also fewer TA opportunities in European universities compared to the US.
@@
 
-\chapter{How is Your Application Evaluated?}\label{chap:evalapps}
+\chapter{How is Your Application Evaluated?}\label{sec:evalapps}
 
@@
 \begin{commentbox}{Vu:}
-At Mason, we usually decide that a full-time PhD candidate is either (i) admitted with funding (\autoref{chap:funding}) or (ii) rejected. In other words, in most cases, we either
+At Mason, we usually decide that a full-time PhD candidate is either (i) admitted with funding (\autoref{sec:funding}) or (ii) rejected. In other words, in most cases, we either
 admit you with full funding or reject your application. In some rare cases, we admit
@@
 
-Some universities do waive, e.g., Rice, TTIC do not have fee for PhD applications and many universities for domestic students (\autoref{chap:domestic-students}).
+Some universities do waive, e.g., Rice, TTIC do not have fee for PhD applications and many universities for domestic students (\autoref{sec:domestic-students}).
 Some programs also waive if the applicants attend some of their opening sessions.
@@
 % In CS, conferences are important and top conferences are prestigious.
-You can find top CS conferences at places such as \href{https://csrankings.org}{CSRankings}, which ranks CS programs based on how their faculty publish at top \emph{conferences} (see more in \autoref{chap:rankings}).
+You can find top CS conferences at places such as \href{https://csrankings.org}{CSRankings}, which ranks CS programs based on how their faculty publish at top \emph{conferences} (see more in \autoref{sec:ranking}).
 
@@
 statement of purpose (SOP) or personal statement\footnote{Few schools separate these documents and ask you to write both: SOP, which focuses on research experiences, and personal statement, which is everything more personal, e.g., why PhD, challenges, diversity, etc}!
-A well-written SOP also shows that you can communicate, which is very important in research, and that you can effectively teach and communicate with students, which is important for TA funding (see \autoref{chap:funding}).
+A well-written SOP also shows that you can communicate, which is very important in research, and that you can effectively teach and communicate with students, which is important for TA funding (see \autoref{sec:funding}).
 Many SOP samples for CS are \href{https://cs-sop.org/}{available here}.
@@
 This can raise concerns about your ability to handle the PhD.
-Don't talk about religious or political beliefs (\autoref{chap:cultural}).
+Don't talk about religious or political beliefs (\autoref{sec:cultural}).
 \item \textbf{Criticizing} your current or previous institution or professors. Just like in a job interview, don't badmouth your current or previous employer. 
@@
 On one hand you will need to show some level of English proficiency, but on the other hand, you do not need to have very good scores in these tests (many adcom members themselves were once international students and struggled with English).
-You should just do well enough to pass the minimum requirement set by the university. %, which is required for TA (\autoref{chap:funding}).
+You should just do well enough to pass the minimum requirement set by the university. %, which is required for TA (\autoref{sec:funding}).
 Just as with grades (\autoref{chap:your-school}) and GRE (\autoref{sec:gre}), having high scores in these tests might not help, but having too low scores can be a \red{red flag} and sometimes results in an automatic rejection (\autoref{sec:why-rejected}), e.g., below the minimum requirement.
@@
 \section{CV}
-\sectioninfo{Highlight and summarize major achievements such as major publications and competitive programming awards.}
+\sectioninfo{Highlight and summarize major achievements, e.g., Publications, Programming Awards}
 
@@
 \chapter{Interview and the Waiting Game}\label{chap:interview}
-\chapterinfo{Getting an interview is typically a good sign; but no interview does not mean rejection.}
+\chapterinfo{Getting an interview is typically a good sign; but no interview does not mean rejection}
 
@@
 
-\chapter{Getting Admitted}\label{chap:accepted}
+\chapter{Getting Admitted}\label{sec:accepted}
 \chapterinfo{Congrats! Now it is your turn to evaluate the school!
@@
 Now you're at a different game because the schools that have admitted you will try to get you to accept them!
-Important factors to consider include the reputation of schools and professors (\autoref{chap:schoolsandprofs}), and funding availability (\autoref{chap:funding}).
+Important factors to consider include the reputation of schools and professors (\autoref{sec:schoolsandprofs}), and funding availability (\autoref{sec:funding}).
 You will likely have to make your decision (\autoref{sec:accept-postpone-decline}) by around \emph{April 15}, which is the deadline for most schools
@@
 \chapter{Dealing with Rejection}\label{chap:not-accepted}
-\chapterinfo{Rejection is part of academia.  Common reasons for rejection include aiming too high, not a good fit, and bad luck.  Do not take rejection personally. Learn from it and try again.}
+\chapterinfo{Rejection is common and part of academia.  Common reasons for rejection include aiming too high, not a good fit, or bad luck. Learn from it and try again.}
 
@@
 
-\chapter{Funding}\label{chap:funding}
+\chapter{Funding}\label{sec:funding}
 \chapterinfo{TAs, RAs, and fellowships are main funding sources for PhDs.
@@
   Having health insurance is required at many US universities.
-Do not assume that you're young and healthy and ignore insurance (\autoref{chap:cultural-misc}).
+Do not assume that you're young and healthy and ignore insurance (\autoref{sec:cultural-misc}).
 At Mason, and most CS PhD programs, your GTA or GRA comes with full insurance.
@@
 
-For external fellowships including those from the US government (e.g., NSF GRFP, DOD NDSEG) and tech companies (e.g., Google, Microsoft, Facebook), you will need to apply.
+For external fellowships including those from the US government (e.g., NSF, DOD) and tech companies (e.g., Google, Microsoft, Facebook), you will need to apply.
 % However, tech companies including Google, Microsoft, Facebook have fellowships that international students can apply for.
-Prestigious external fellowships typically require a clear and good research plan (the GRFP also requires broader impacts discussion).
+Prestigious external fellowships typically require a clear and good research plan (the GRFP also requires outreach activities and plans).
 So it might be a good idea to wait until your second year to have research experience and even publication before applying.
@@
 
-In addition to general fellowships that are available to all students, there are also specific fellowships, e.g., for US citizens and permanent residents (\autoref{chap:domestic-students}), for underrepresented minority (URM) students (\autoref{sec:urm}), and students from specific countries.
+In addition to general fellowships that are available to all students, there are also specific fellowships, e.g., for US citizens and permanent residents (\autoref{sec:domestic-students}), for underrepresented minority (URM) students (\autoref{sec:urm}), and students from specific countries.
 There are many fellowships available, so you should look for them and apply.
@@
 In fact, these are so prestigious that even if you didn't get it but make it to the final round or even \emph{``honorable mentioning''}, your school will still mention you on their website, and you should put it on your CV.
-\autoref{chap:fellowships} discusses the evaluation processes of the NSF GRFP and DoD NDSEG.
 
@@
 
-\chapter{Choosing Schools}\label{chap:schoolsandprofs}
+\chapter{Choosing Schools}\label{sec:schoolsandprofs}
 \chapterinfo{Not every university has a CS PhD program. Not every professor, even those in CS, can advise or graduate CS PhD students.}
@@
 While this can be confusing due to the large number of universities in the US, a little research, e.g., searching for PhD in CS from the school website, will help you find out.
-Schools listed in \autoref{chap:rankings} have PhD programs in CS, so you can start there.
+Schools listed in \autoref{sec:ranking} have PhD programs in CS, so you can start there.
 
@@
 This is also a good way to learn about individual faculty, e.g., who works on what, and well-known CS conferences\footnote{In CS (and probably only in CS), conferences, not journals, are often the main venue to publish research findings (see \href{https://homes.cs.washington.edu/~mernst/advice/conferences-vs-journals.html}{why here}).}.
-%\autoref{chap:rankings} gives the top 50 CS programs in the US according to CSRankings.
+%\autoref{sec:ranking} gives the top 50 CS programs in the US according to CSRankings.
 
@@
 
-If you get admission to several places, you should consider attending Open Houses (\autoref{chap:accepted}) and contact profs.
+If you get admission to several places, you should consider attending Open Houses (\autoref{sec:accepted}) and contact profs.
 that you're interested in at those places and talk to them.
@@
 They typically teach 3--4 classes per semester (which is quite a lot) and \emph{do not} have research responsibilities, e.g., they do not have to worry much about publishing papers or obtaining grants.
-As such they typically do not have funding (e.g., RA~\autoref{chap:funding}) to support PhD students.
+As such they typically do not have funding (e.g., RA~\autoref{sec:funding}) to support PhD students.
 Depending on the university, they might be able to \emph{co-advise} and mentor PhD students with a tenure-line faculty, but they typically do not formally advise and graduate PhD students.
@@
 to \emph{ask about your admission chance}, please \textbf{don't}.
-We don't know and can't answer because as explained in \autoref{chap:evalapps}, we don't make individual decisions and might not even be assigned to evaluate your application.
+We don't know and can't answer because as explained in \autoref{sec:evalapps}, we don't make individual decisions and might not even be assigned to evaluate your application.
 It is the same as sending a paper draft to a journal editor and asking them if your paper has a chance.
@@
 \part{Miscs and FAQs}\label{part:faqs}
-\chapter{Miscs and FAQs}\label{chap:faqs}
+\chapter{Miscs and FAQs}\label{sec:faqs}
 
@@
 Finally, I enter my decision, which is usually either a rejection or an offer of admission with full funding (e.g., from TA).
-I also recommend very strong candidates for the University Presidential fellowship., which is a fellowship from GMU that provides funding similar to an RA (\autoref{chap:funding}).
+I also recommend very strong candidates for the University Presidential fellowship., which is a fellowship from GMU that provides funding similar to an RA (\autoref{sec:funding}).
 Note that while the system has other decision options, eg., admit without funding, provisional admission (e.g., if they need to take some courses), I do not use them, simply because we either reject or admit with funding.
@@
 These nuances represent some cultural and academic differences that you may encounter and will gradually adapt to.
-More on these differences in \autoref{chap:cultural}.
+More on these differences in \autoref{sec:cultural}.
 
@@
 
-  \item[Adcom~\autoref{chap:evalapps}] Admission Committee. The group of faculty members who review applications and make admission decisions.
-
-  \item [Adcom members~\autoref{chap:evalapps}] Faculty members who are part of the admission committee. These are the people who review applications and make admission decisions.
-
-  \item[Adcom chair~\autoref{chap:evalapps}] The faculty member who leads the admission committee.
+  \item[Adcom~\autoref{sec:evalapps}] Admission Committee. The group of faculty members who review applications and make admission decisions.
+
+  \item [Adcom members~\autoref{sec:evalapps}] Faculty members who are part of the admission committee. These are the people who review applications and make admission decisions.
+
+  \item[Adcom chair~\autoref{sec:evalapps}] The faculty member who leads the admission committee.
 This person is often not involved in individual admission decisions but oversees the process (e.g., resolve disputes, ensure fairness).
@@
 Domestic students] International students are loosely defined as those who would need a visa to study in the US (\autoref{chap:visa}).
-Domestic students (\autoref{chap:domestic-students}) are those who do not need a visa, typically US citizens or permanent residents.
+Domestic students (\autoref{sec:domestic-students}) are those who do not need a visa, typically US citizens or permanent residents.
 
@@
 
-  \item[Open House \autoref{chap:accepted}] An event where admitted students visit the university to meet faculty, students, and staff, and learn more about the program.
+  \item[Open House \autoref{sec:accepted}] An event where admitted students visit the university to meet faculty, students, and staff, and learn more about the program.
 This is a great opportunity to get a feel for the university and the department and decide if it is a good fit for you.
@@
 
-\chapter{Domestic Students}\label{chap:domestic-students}
+\chapter{Domestic Students}\label{sec:domestic-students}
 \chapterinfo{Specific benefits and opportunities for domestic students applying to CS PhD programs.}
@@
 \paragraph{School \autoref{chap:your-school}} Adcom already knows about your school, which is a plus.
-You are also more familiar with the US education system and academic culture (\autoref{chap:cultural}).
+You are also more familiar with the US education system and academic culture (\autoref{sec:cultural}).
 
@@
 
-\paragraph{Funding \autoref{chap:funding}} You have more opportunities for funding, e.g., through government scholarships for US citizens and permanent residents.
+\paragraph{Funding \autoref{sec:funding}} You have more opportunities for funding, e.g., through government scholarships for US citizens and permanent residents.
 You can also apply to specific programs \emph{before} you start your PhD, e.g., NSF Graduate Research Fellowship Program (GRFP) and Hertz Foundation Fellowship.
-These fellowships are very competitive (\autoref{chap:fellowships}) and can significantly improve your admission chances.
+These fellowships are very competitive and can significantly improve your admission chances.
 
@@
 
-\paragraph{Open House \autoref{chap:accepted}} It is easier for you to attend open house events in person.  This can help you make a better decision on which school to attend.
+\paragraph{Open House \autoref{sec:accepted}} It is easier for you to attend open house events in person.  This can help you make a better decision on which school to attend.
 
@@
 
-  \item Funding: MS is typically \emph{not funded}, while PhD is (\autoref{chap:funding}). See \autoref{sec:ms-funding} for more details on MS funding.
+  \item Funding: MS is typically \emph{not funded}, while PhD is (\autoref{sec:funding}). See \autoref{sec:ms-funding} for more details on MS funding.
 
@@
 \subsection{Funding}\label{sec:ms-funding}
-Unlike PhD programs, which often have funding (\autoref{chap:funding}), MS students are typically \emph{not} funded.
+Unlike PhD programs, which often have funding (\autoref{sec:funding}), MS students are typically \emph{not} funded.
 This is because MS students do not focus on research and thus are not funded through RA and PhD students have priority for TA positions.
@@
 Because of the differences between MS and PhD programs (\autoref{sec:phd-vs-ms}), you should consider different factors when selecting and ranking MS programs.
-For example, you might want to consider factors such as location, industry connections, and job placement instead of research areas and advisers (\autoref{chap:schoolsandprofs}).
+For example, you might want to consider factors such as location, industry connections, and job placement instead of research areas and advisers (\autoref{sec:schoolsandprofs}).
 
@@
 
-\chapter{Fellowship Applications}\label{chap:fellowships}
-\chapterinfo{How fellowships such as NSF GRFP and DOD DSEG are evaluated.}
-
-Having an external and major fellowship can significantly improve your chances of getting into a good PhD program (\autoref{sec:fellowships}).
-The National Science Foundation Graduate Research Fellowship Program (NSF GRFP) and the Department of Defense National Defense Science and Engineering Graduate Fellowship (DoD NDSEG) are two of the most prestigious fellowships for CS PhD students who are US citizens or permanent residents.
-It is rumored that several universities would revert their admission decision on a student if they get one of these fellowships.
-
-The \href{https://www.nsfgrfp.org/}{GRFP} and \href{https://www.ndsegfellowships.org/}{NDSEG} website have detailed information on the application process and eligibility requirements.
-This section focuses on how your application is evaluated and tips on improving your application.
-Note that while these are specific GRFP and NDSEG, the general advice can be applied to other major fellowships (those for non-US citizens or permanent residents).
-
-%Securing a prestigious fellowship such as the National Science Foundation Graduate Research Fellowship Program (NSF GRFP) or the Department of Defense National Defense Science and Engineering Graduate Fellowship (DoD NDSEG) can significantly enhance a student’s prospects for admission to a strong PhD program.
-These fellowships not only provide financial support but also offer academic freedom and prestige that can set an applicant apart from their peers.
-However, successfully obtaining one of these fellowships is no easy feat—they are highly competitive, and applications are scrutinized rigorously by expert reviewers.
-
-%One of the biggest advantages that applicants from elite institutions often have is access to faculty members who are well-versed in the review process.
-These faculty members understand the nuances of writing strong letters of recommendation (LoRs) and guiding students through the proposal-writing process.
-Unfortunately, students from smaller institutions or those with mentors unfamiliar with NSF or DoD fellowship reviews may find themselves at a disadvantage—not because of their potential or research ability, but simply due to a lack of insider knowledge about what makes a strong application.
-
-%This chapter aims to level the playing field by providing an insider’s perspective on how these applications are reviewed.
-Drawing from my experience as a reviewer for both NSF GRFP and DoD NDSEG, I will demystify the evaluation process, highlight key elements that make applications stand out, and offer strategic advice for applicants and their letter writers.
-While research experience and proposal quality are, of course, crucial to success, understanding the review criteria and expectations can be just as important.
-
-%Moreover, this chapter will provide guidance for LoR writers, helping them craft letters that effectively support their students’ applications.
-Many well-intentioned letters fail to address the aspects that reviewers prioritize, which can unintentionally weaken an otherwise strong application.
-By offering insights into what reviewers look for, I hope to empower both applicants and their mentors to present their strongest possible case for these fellowships.
-
-%Whether you are an aspiring PhD student preparing your application or a faculty member writing a recommendation, this chapter will equip you with the knowledge needed to navigate the NSF GRFP and DoD NDSEG review process with confidence.
+\chapter{Fellowship Applications}\label{sec:fellowship}
+\chapterinfo{How to apply for fellowships such as NSF GRFP, DOD DSEG to fund your PhD.}
 
@@
 
-The NSF GRFP is a prestigious fellowship for US citizens, nationals, and permanent residents who are in the early stages of their graduate study in STEM fields, including both MS and PhD students.
-The fellowship provides three years of financial support, which includes a stipend and tuition.
-The GRFP is highly competitive, with many applicants from top schools, with strong LoRs and research experience.
-
 \subsection{How Applications are Evaluated}
@@
 \begin{itemize}
-\item Each applicant is assigned to \emph{about 3 reviewers} who have expertise in the applicant's research topics.
+\item Each applicant is assigned to \emph{about 3 reviewers} who all seem to have expertise in the applicant's research topics.
 This means reviewers can be quite knowledgeable in your field and your work, your LoR writers and their work, and thus can evaluate your application quite in depth.
@@
 Note that if sometimes an original reviwer is no longer available then someone who is not in the field might review your application.
-Thus you want to make your statements accessible to a general CS audience.
-
-\item Reviewers typically try to compare applicants \emph{within their cohort level}.
+Thus in general you want to make your application accessible to a general audience in CS.
+
+\item Reviewers attempt to compare applicants \emph{within their cohort level}.
 This means that we do not compare an undergrad, who is still applying to grad school, to a 2nd year PhD student who already has an adviser and published multiple papers.
@@
 
-\item Unlike CS PhD applications that focus on research potentials in which papers, research experience, and LoRs are important, NSF GRFP looks at both \emph{research potential} and \emph{broader impacts}, e.g., how you can help society, how you can mentor others, etc.
-It is not expected that you have published papers or have a lot of research experience (recall that MS students can also get the fellowship).
-Instead, you need to convince the reviewers that you have the potential to be a good researcher (technical merits) and that you can help society (broader impacts).
-More in \autoref{sec:grfp-im-bi}.
+\item Unlike CS PhD applications, which focus on research potentials in which concrete evidence like papers, research experience, and LoRs are important, NSF GRFP also focuses on \emph{potential} but also on broader impacts, e.g., how you can help society, how you can mentor others, etc.
 
@@
 
-\subsection{Intellectual Merit and Broader Impacts}\label{sec:grfp-im-bi}
-
-Your application will be evaluated based on two criteria: Intellectual Merit (IM) and Broader Impacts (BI).
-
-\paragraph{Intellectual merit (IM)} IM evaluates your research potential, e.g., your research experience, your potential to do research, and how you can contribute to the field.
-Unlike a normal project or research proposal, in which the evaluation is entirely based on the project, in the GRFP, the evaluation is \emph{not} entirely about the research plan.
+\subsection{Intellectual Merit and Broader Impacts}
+Your will be evaluated based on two criteria: Intellectual Merit (IM) and Broader Impacts (BI).
+
+\paragraph{Intellectual merit} is based on your research potential, e.g., your research experience, your potential to do research, and how you can contribute to the field.
+
+Note that unlike a normal project or research proposal, in which the evaluation is based on the project, in the GRFP, the evaluation is \emph{not} about the research plan.
 We are repeatedly reminded to not focus too much on the research plan as many applicants are still applying for PhD programs and might not even get into the field or work with a prof.
 that fit their research plan.
-
 Instead, the evaluation is based on your \emph{potential} of being a good researcher using various criteria, e.g., your research experience, what others think of you (LoRs), research plan, etc.
@@
 
-\paragraph{Broader impacts (BM)} BM is a common and important component in NSF application.
-It is about how you and your research can \emph{benefit society}.
+\paragraph{Broader impacts} based on how you and your research can benefit society.
 Many students just mention how their research is cross-discipline and thus has broader impacts (e.g., improve software quality and thus help society).
 This is too common and not enough.
-You also want to talk about how you, as a grad student with your research, can help society, e.g., through outreach, mentoring, etc.
-As mentioned below, you will want to be specific and concrete, e.g., \emph{``I will work with highschool students ..''} is not as convincing as \emph{``I have had experience in mentoring highschool students X and will continue to do so through Y,Z ...''}.
-
-Writing BM is often difficult, not only for students but also new faculty.  You should ask someone who has experience writing NSF proposals to review your BM.
-
-\subsection{Common Pitfalls and Tips}\label{sec:grfp-pitfalls}
-
-These are common pitfalls that many applicants make in their GRFP applications and tips on how to avoid them.
-
-\begin{itemize}
-
-\item \textbf{Too Technical.} Remember that you're writing to a general audience in CS, not experts in your field.
-While NSF tries to match reviewers with your field, CS fields are simply too broad and you might get a reviewer in the field but not in your specific area or familiar with the problem you're working on.
-So avoid technical jargon and explain things in simple terms. If the reviewer cannot understand your proposal, they will likely give you a low score.
+You also want to talk about how you, as a grad student with your research, can help society, e.g., through outreach, mentoring, diversity, etc.
+As mentioned below, you will want to be specific and concrete, e.g., ``I will work with highschool students and underrepresented students'' is not as convincing as ``I have had experience in mentoring highschool students X and will continue to do so through Y,Z ...''.
+
+\subsection{Common Pitfalls}
+
+\paragraph{Too technical} Remember that you're writing to a general audience in CS, not experts in your field.
+While NSF tries to match reviewers with your field, CS fields are simply too broad and you might get a reviewer in the field but not in your specific area.
+So avoid technical jargon and explain things in simple terms. If the reviewer does not understand your proposal, they will likely give you a low score.
 
@@
 
-\item \textbf{Preliminary Work.} Your research plan should have some preliminary work to convince reviewers that you have thought about the problem and have some initial promising results.
+\paragraph{Preliminary Work} Your research plan should have some preliminary work to convince reviewers that you have thought about the problem and have some initial promising results.
 It does not have to a lot, e.g., you don't need published results or an implementation, a small experiment done by hand on some small examples would suffice.
@@
 
-\item \textbf{Related Work and Challenges.} Many research plans motivate the problem well but do not discuss limitations of existing work, making it questionable if the problem is important or if you are aware of the challenges.
-Thus, you should do a thorough literature review and discuss what people have done and their limitations.
-Your research plan should then talk about how you plan to address these limitations, i.e., fill the gap in the literature.
-
-\item \textbf{Do not BS.} Many research plans include technical details or proposed work that are vauge, unrealistic, or even impossible.
-This is similar to a student taking an exam and write a lot of nonsense and hope for partial credits.
-As mentioned, reviewers are often chosen based on their expertise in your field and can detect B.S.
-Again, do your homework and ask someone who knows the field, e.g., your advisor or LoR writer, to review your research plan.
-
-\item \textbf{Overexaggeration Lors.} Many GRFP applicants are from top schools with well-known professors that are very good at writing.
-However, they often use flowery language and completely overrate their students.
-Reviewers are  professors themselves and can sense this overexaggeration.
-So you should ask your LoR writers to tone down the language (similar to overclaiming research conributions when writing papers) and and be specific with concrete examples to demonstrate why you're \emph{``the best''}.
-
-\end{itemize}
+% Prelim work does not, and in fact should not, be extensive.
+You do not need a paper or a complete implementation.
+You don't need to show that your idea is better than others or even that it will work.
+You just need to show that you have tried it and it has some promise.
+This can be as simple as a small experiment done by hand on some small examples,  e.g., if you come up with a new abstraction in PL for proving correctness, then you can just manually apply it to a few small programs, preferrably ones that existing works have trouble with.
+This is enough to show that you have thought about the problem and have some initial results.
+
+% To make prelim concrete (and thus more convincing), you might want to use specific names and numbers, e.g.,  \emph{``I tried my prototype on the MNIST dataset, achieving an accuracy of 90\%''}, or \emph{``I used Lean to formalize and prove a specific case of the Fermat's Last Theorem''}.
+In addition, negative results are results and can be very convincing.
+So don't be afraid to say \emph{``I tried my idea on X but it failed because of Y, and I plan to address this by Z''}.
+
+\paragraph{Research} is the emphasis.
+Many made a mistake of focusing on teaching and mentoring and not on research.
+This is especially true for students who have a lot of teaching and mentoring experience, e.g., from a teaching-focused university or a teaching assistant.
+While teaching and mentoring are important, the GRFP is a \emph{research} fellowship which aims to train researchers.
+So you will need to demonstrate your passion for research and your potential to be a good researcher.
 
@@
 
-The NDSEG is another prestigious fellowship supporting \emph{U.S. citizens} pursuing advanced degrees in STEM fields \emph{critical to national defense} (such as CS).
-It provides up to three years of funding, including a  stipend and tuition coverage.
-
+The DoD National Defense Science and Engineering Graduate (NDSEG) is another prestigious fellowship supporting \emph{U.S.
+citizens} pursuing advanced degrees in STEM fields \emph{critical to national defense} (such as CS).
+It provides up to three years of funding, including a  stipend, tuition coverage, and health insurance.
+
+You can read more details about NDSEG and its guideline for applicants from the \href{https://www.ndsegfellowships.org/}{official website}.
+Here, I mainly focus on things that stand out to me as reviewers and also compare them with NSF GRFP as many students apply to both.
 \subsection{Evaluation}
 
-\paragraph{Redacted info.} Unlike the NSF GRFP, reviewers only see \emph{redacted versions} of your submitted materials (applicants are asked to provide full and redacted versions of your research plan, CV, and other materials).
-Thus, we do not know what schools your went to, where you did your internships, your publications or research experience.
-We also \emph{do not} see your LoRs, i.e., we do not know who wrote them or what they wrote.
-
-Reviewers mainly get to know you and evaluate your application by reading your research plan and personal essays.
+\paragraph{Redacted info.} Unlike the NSF GRFP, for NDSEG reviewers \emph{do not} see your LoRs, i.e., we do not know who wrote them or what they wrote.
+Moreover, we only see \emph{redacted versions} of your submitted materials (applicants are asked to provide full and redacted versions of your research plan, CV, and other materials).
+Thus we do not know what schools your went to, where you did your internships, your publications or research experience.
+
+Thus, reviewers mainly get to know you and evaluate your application by reading your research plan and personal essays.
 This is for \emph{fairness and avoiding bias} based on the school you went to, places you published at, your LoR writers, etc.
@@
 For example, a student mentioned that they have published in the topic they proposed but redacted all information about the publication, so I have no idea where they published, what the paper is about, etc.
-It is also difficult to evaluate the academic background of the student, e.g., some student just started grad school and so their transcript is essentially blank.
+It is also difficult to evaluate the academic background of the student, e.g., some student just started grad school and so their transcript is essentially blank, and I do not know where they went to undergrad, etc.
 It is also annoying to see CVs with so many blanked out information.
@@
 In the beginning I find it surprising that I don't get to see the LoRs, but I understand the reasoning behind it and gradually think it is a good idea.
-As mentioned above, LoR writers for NSF GRFP (\autoref{sec:grfp-pitfalls}), especially those from top schools, often overrate their students and thus can bias the evaluation.
+As mentioned above, LoR writers, especially those from top schools, often overrate their students and thus can bias the evaluation.
 So I don't find I am missing much by not seeing the LoRs.
@@
 
-\subsection{Common Pitfalls and Tips}\label{sec:ndseg-pitfalls}
-
-These are some common pitfalls I see in NDSEG applications and some tips to avoid them.  Some of these are similar to NSF GRFP (\autoref{sec:grfp-pitfalls}) while some are specific to NDSEG.
-
+\subsection{Common Pitfalls and Tips}
+Based on my experience in reviewing, here are some common pitfalls to avoid:
 \begin{itemize}
 
-    \item \textbf{Customize your research to DoD}, e.g., through specific project from DARPA or ONR. 
+    \item Customize your research to DoD, e.g., through specific project from DARPA or ONR. 
     Make this prominent, e.g., in the introduction or in its own section.  This is different from NSF GRFP where you want broader impacts to society.
 
-    \item \textbf{Unclear research plan.}
+    \item Unclear research plan. 
      You should aim to have (i) clear problem statement, (ii) why it is important (esp.
 to DoD), (iii) what has been done and their challenges/limitations, and (iv) what you plan to do to address the challenges.
-Moreover, research is often quite technical and ambitious, so you want to show some preliminary work, e.g., a small experiment by hand on some small examples, to convince reviewers that you have thought about the problem and have some initial results.
+Moreover, research is often quite technical and ambitious, so you want to show some preliminary work, e.g., a small experiment by hand on some small examples, to convince reviewers that you have thought about the problem and have some initial results.\\
 
@@
 
-    \item \textbf{Personal stories and essays should be unique and interesting}.
+    \item Personal stories and essays should be unique and interesting.
 Many essays and stories I read are quite generic and do not tell much about the student.
-For example, being an URM, involving in a robotics club or mentoring underrepresented students are quite common and do not tell much.
+For example, being an URM, involving in a robotics club or mentoring underrepresented students are quite common and do not tell much.\\
 
@@
 
-    \item \textbf{Not using all allowed pages}.
+    \item Not using all allowed pages.
 Many research plans use only 1/2 or 2/3 of the number of allowed pages, which themselves are already quite short.
@@
 %- use latex,  do not write in word and especially use left justification  (see how CS and engineer people publish papers)  
+\item If you have an advisor or Lor writer, ask them to look at your writing/essays.
+This is because you are new to the field and research writing, and your ideas might not be accurate or have already been explored.
 
@@
 
-Finally, just as with the NSF GRFP and even PhD SOPs, ask your LoR writers and adviser to look at your writing/essays.
-You are new to the field and research writing, and your ideas might not be accurate or have already been explored.
-It's better that they tell you now than for you to find out later from reviewers.
-
-\chapter{Academic and Cultural Differences}\label{chap:cultural}
+\chapter{Academic and Cultural Differences}\label{sec:cultural}
 
@@
 
-  \section{Miscs}\label{chap:cultural-misc}
+  \section{Miscs}\label{sec:cultural-misc}
 
@@
   \paragraph{Healthcare System}  You (and your spouse) will need health insurance!
-Otherwise you will be charged a lot for healthcare services when you need them.
-However, as mentioned in \autoref{chap:funding}, your  TA/RA (and fellowships)  will cover health insurance.
-Your spouse/children also get health insurance or significant discounts under your plan.
-
-  Note that even with insurance, heathcare services might still be expensive.
-So you should know what your insurance covers and be prepared for unexpected costs.
+Otherwise you will be charged a lot for healthcare services.
+However, as mentioned in \autoref{sec:funding}, your  TA/RA  will cover health insurance (and your spouse/children also get insurance or discount for it).
+
+  However, be aware that heathcare services can still be expensive even with insurance.
+So  know what your insurance covers and be prepared for unexpected costs.
 Moreover, healthcare system has many confusing jargons such as HMO, PPO, deductibles, co-pays, and coinsurance (try take a look at the Explanation of Benefits or EoB statement you received from your insurance company).
 It's arguably the most complicated system in the US and even Americans often do not understand it (and politicians often exploit this to their advantage).
-Do not hesitate to ask your HR or the insurance company for help.
+So do not hesitate to ask your HR or the insurance company for help.
 
@@
 So adding an extra 15--20\% to your bills is common, especially in restaurants.
-
   You should also tip for other services, e.g., Uber, taxi, haircuts, and hotel services.
@@
 
-\chapter{CSRankings: Rankings of CS PhD programs}\label{chap:rankings}
+\chapter{CSRankings: Rankings of CS PhD programs}\label{sec:ranking}
 \chapterinfo{CSRankings.org is a ranking system based on faculty publications at top CS conferences.}
@@
 %\autoref{tab:ranking} lists .
-\chapter{History and Acknowledgement}\label{chap:ack}
+\chapter{History and Acknowledgement}\label{sec:ack}
 \paragraph{History} This book was conceived during a lunch with a faculty at Mason.
@@
 There I saw many questions from students about PhD programs.
-However, most participants are students, many of whom in non-CS fields or not in the US, and their answers are typically not accurate and often lead to more confusion.
+However, most participants are students in non-CS fields or not in the US, and their answers are typically not accurate and often lead to more confusion.
 So I thought it would be useful to have a handbook that is specific to CS PhD programs in the US from an insider perspective.
 
-I started writing this book in May 2023 and have been updating it since then (mostly around deadline time when I procrastinate, i.e., \emph{productive procrastination}!).
-The book was initially intended for international students but has been updated to include domestic students (\autoref{chap:domestic-students} and \autoref{chap:fellowships}).
+I started writing this book in May 2023 and have been updating it since then (mostly around deadline time when I tend to procrastinate, i.e., \emph{productive procrastination}!).
+The book was initially intended for international students but has been generalized to all students (domestic students also get confused by the process!).
 The book is open-source and available on \href{https://github.com/nguyenthanhvuh/phd-cs-us}{GitHub} so that others can contribute to it.
@@
 Currently I am the program director of the MS program in Software Engineering at Mason (thus also have experience with MS admission---which is quite different than PhD as discussed in \autoref{chap:ms}).
-I also have served in the panels of PhD fellowships including NSF GRFP and DoD NDSEG, which allows me to provide some insights on external fellowhip applications (\autoref{chap:fellowships}).
+I also have served in the panels of PhD fellowships including NSF GRFP and DoD NDSEG, which allows me to provide some insights on external fellowhip applications (\autoref{sec:fellowships}).
 
@@
 I hope to capture the diverse challenges and experiences they've faced in this book so that it can be a valuable resource for prospective international students.
-Finally, my upbringing in the US provides a perspective aligned with American culture, allowing me to shed light on various issues, particularly those related to cultural differences (\autoref{chap:cultural}).
+Finally, my upbringing in the US provides a perspective aligned with American culture, allowing me to shed light on various issues, particularly those related to cultural differences (\autoref{sec:cultural}).
 
@@
 
-Also thanks to NSF for encouraging faculty to be creative in research and education, which allows me to work on this book.
+Also thanks to NSF for encouraging faculty to be creative in research and education, which allows me to write this handbook.
 

### 2025-02-19 demystify.tex
updates

--- 2025-02-16 demystify.tex
@@
 International students, especially those from less well-known universities and smaller countries, often wonder if they should apply to a CS PhD program in the US.
-Common concerns include (i) the \textbf{difficulty of getting admitted}, (ii) the \textbf{cost of graduate study} in the US, and for some, (iii) the fear of \textbf{discrimination} e.g., based on gender identity or cultural background.
+Common concerns include (i) the \textbf{difficulty of getting admitted} and (ii) the \textbf{cost of graduate study} in the US.
+%, and for some, (iii) the fear of \textbf{discrimination} e.g., based on gender identity or cultural background.
 
@@
 
-\paragraph{Discrimination is not tolerated} Not only is discrimination illegal, diversity and inclusion are highly valued in US universities.
+% \paragraph{Discrimination is not tolerated} Not only is discrimination illegal, diversity and inclusion are highly valued in US universities.
 Many universities take great pride in their diverse student body and faculty.
 
-In fact, many resources are available to encourage and support students from minority and diverse groups to pursue higher education and research, e.g., specific scholarships and fellowships.
+% In fact, many resources are available to encourage and support students from minority and diverse groups to pursue higher education and research, e.g., specific scholarships and fellowships.
 Faculty in CS are also familiar with and are encouraged to support such students in their group.
@@
 
-You should talk about things that adcom members might not know about and can help make you \textbf{stand out} in the application pool of thousands of applicants, e.g., you are member of a minority or LGBTQ+ group (\autoref{sec:urm}), your personal Github project with hundreds of stars or your regular contributions to well-known open-source projects (see \autoref{sec:improve-your-chance} for increasing your admission chance).
+%you are member of a minority or LGBTQ+ group (\autoref{sec:urm}),
+
+You should talk about things that adcom members might not know about and can help make you \textbf{stand out} in the application pool of thousands of applicants, e.g., your personal Github project with hundreds of stars or your regular contributions to well-known open-source projects (see \autoref{sec:improve-your-chance} for increasing your admission chance).
 
@@
 
-Many universities now require a diversity, equity, and inclusion (DEI) statement as part of the application (also see \autoref{sec:urm}).
+Many universities now require a diversity, equity, and inclusion (DEI) statement as part of the application.
 Students, especially international ones, often struggle with what to write, especially when they are not from a minority group or do not have much experience in DEI.
@@
 Do you participate in outreach activities that help increase diversity and inclusion in CS?
-Diversity is valued in CS programs in the US (\autoref{sec:urm}), and experiences in these areas can make you stand out and get noticed from reviewers.
+Diversity is valued in CS programs in the US, and experiences in these areas can make you stand out and get noticed from reviewers.
 
@@
 
-For external fellowships including those from the US government (e.g., NSF GRFP, DOD NDSEG) and tech companies (e.g., Google, Microsoft, Facebook), you will need to apply.
+For external fellowships including those from the US government, e.g., see~\autoref{chap:fellowships} for NSF GRFP and  DOD NDSEG, and tech companies (e.g., Google, Microsoft, Facebook), you will need to apply.
 % However, tech companies including Google, Microsoft, Facebook have fellowships that international students can apply for.
@@
 
-In addition to general fellowships that are available to all students, there are also specific fellowships, e.g., for US citizens and permanent residents (\autoref{chap:domestic-students}), for underrepresented minority (URM) students (\autoref{sec:urm}), and students from specific countries.
+%In addition to general fellowships that are available to all students, there are also specific fellowships, e.g., for US citizens and permanent residents (\autoref{chap:domestic-students}), for underrepresented minority (URM) students (\autoref{sec:urm}), and students from specific countries.
 There are many fellowships available, so you should look for them and apply.
@@
 I skim weak \emph{SOPs} but read strong ones carefully.
-I note whether the SOP is exciting, research-oriented, standing out, e.g., applicant is minority (\autoref{sec:urm}), and tailored to our program (e.g., if they are familiar with the work of some faculty or have talked to them).
+I note whether the SOP is exciting, research-oriented, standing out, and tailored to our program (e.g., if they are familiar with the work of some faculty or have talked to them).
 These notes are entered into the evaluation system.
@@
 
-\section{Member of an underrepresented or LGBTQ+ groups}\label{sec:urm}
-\sectioninfo{Being a minority can make you stand out and help your application.}
-
-A common question is whether you should mention, e.g., in your SOP, that you are a member of an underrepresented and minority (URM) group\footnote{URM in CS in the US include African Americans, Hispanic Americans, American Indians/Native Alaskans, Native Hawaiians/Pacific Islanders (excluding Asian Americans), and multiracial students identifying with at least one of the previously listed URM categories.} or LGBTQ+ group.
-The main concern is that you can be discriminated against or that you are asking for special treatment.
-
-%This is a strong concern for international (and even domestic) students, who may have encountered discrimination in their countries or personal experiences. 
-%This question arises often for international (and even domestic) students.
+% \section{Member of an underrepresented or LGBTQ+ groups}\label{sec:urm}
+% \sectioninfo{Being a minority can make you stand out and help your application.}
+
+% A common question is whether you should mention, e.g., in your SOP, that you are a member of an underrepresented and minority (URM) group\footnote{URM in CS in the US include African Americans, Hispanic Americans, American Indians/Native Alaskans, Native Hawaiians/Pacific Islanders (excluding Asian Americans), and multiracial students identifying with at least one of the previously listed URM categories.} or LGBTQ+ group.
+% The main concern is that you can be discriminated against or that you are asking for special treatment.
+
+% %This is a strong concern for international (and even domestic) students, who may have encountered discrimination in their countries or personal experiences. 
+% %This question arises often for international (and even domestic) students.
 The main worry is that they might get discriminated against, as often happen in their countries or personal experiences.
@@
 
-In my personal opinion, highlighting URM or LGBTQ+ identity can \emph{be beneficial}, especially if you can articulate how your diverse experiences contribute to diversity and inclusion in academics.
+% In my personal opinion, highlighting URM or LGBTQ+ identity can \emph{be beneficial}, especially if you can articulate how your diverse experiences contribute to diversity and inclusion in academics.
 Many universities have a strong commitment to diversity and inclusion and actively
-\emph{recruit students from underepresentative backgrounds} (e.g., some \href{https://cse-climate.engin.umich.edu/reports/climate-dei-reports/cse-climate-and-dei-report-2022-2023/#grad-ethnicity}{stats from UMich CSE} and GMU often touts itself as one of the most diverse universities in the US).
+% \emph{recruit students from underepresentative backgrounds} (e.g., some \href{https://cse-climate.engin.umich.edu/reports/climate-dei-reports/cse-climate-and-dei-report-2022-2023/#grad-ethnicity}{stats from UMich CSE} and GMU often touts itself as one of the most diverse universities in the US).
 Moreover, many scholarships and fellowships are created specifically for URM and LGBTQ+ students, which you should explore if qualify.
 
-Understandably, you might feel uncomfortable disclosing this, fearing you would lose a chance of working with a faculty due to their bias.
+% Understandably, you might feel uncomfortable disclosing this, fearing you would lose a chance of working with a faculty due to their bias.
 However, if someone has this issue, then you might not want to work with them anyway (and such bias is not acceptable in academics and likely would not be tolerated by the university).
@@
 
-Moreover, promoting diversity aligns with the goal of broadening participation in computing (BPC) that is strongly supported within academia (e.g., see the \href{https://plans.bpcnet.org/GeorgeMasonUniversity_ComputerScience_DepartmentalBPCPlan.pdf}{BPC plan at Mason CS}) and government funding agencies such as NSF (e.g., see \href{https://new.nsf.gov/cise/broadening-participation}{here} and \href{https://www.nsf.gov/pubs/2022/nsf22125/nsf22125.jsp}{here}).
-Providing this information can also make it easier for faculty to apply for certain scholarships or fellowships to support you, e.g., from a private donor that has offered a fellowship only to students from certain backgrounds.
-
-Ultimately, this is a personal choice and you do what you feel comfortable.
-If you do not feel comfortable disclosing this, you should not feel pressured to do so.
-However, if you feel comfortable but are unsure if it is accepted in higher education in the US, then I hope this helps you understand that it is not only accepted but also valued and appreciated.
+% Moreover, promoting diversity aligns with the goal of broadening participation in computing (BPC) that is strongly supported within academia (e.g., see the \href{https://plans.bpcnet.org/GeorgeMasonUniversity_ComputerScience_DepartmentalBPCPlan.pdf}{BPC plan at Mason CS}) and government funding agencies such as NSF (e.g., see \href{https://new.nsf.gov/cise/broadening-participation}{here} and \href{https://www.nsf.gov/pubs/2022/nsf22125/nsf22125.jsp}{here}).
+% Providing this information can also make it easier for faculty to apply for certain scholarships or fellowships to support you, e.g., from a private donor that has offered a fellowship only to students from certain backgrounds.
+
+% Ultimately, this is a personal choice and you do what you feel comfortable.
+% If you do not feel comfortable disclosing this, you should not feel pressured to do so.
+% However, if you feel comfortable but are unsure if it is accepted in higher education in the US, then I hope this helps you understand that it is not only accepted but also valued and appreciated.
 
@@
   \item [Diversity Statement~\autoref{sec:diversity-statement}] A document sometimes required in applications where applicants describe their contributions to diversity and inclusion, or how their background adds to the diversity of the academic community.
-This document is often required for faculty positions, but more universities are requiring it for graduate admissions.
+This document is often required for faculty positions, but some universities are requiring it for graduate admissions.
   Note that you do not need to be from an underrepresented group to write a diversity statement.
@@
 However, there are some differences and benefits that domestic students should be aware of and can leverage to improve their chances of admission.
-\paragraph{Standing out \autoref{sec:improve-your-chance}} There are \emph{few} domestic applications compared to international ones, i.e., domestic students are the minority (\autoref{sec:urm}) in the CS PhD application pool.
+\paragraph{Standing out \autoref{sec:improve-your-chance}} There are \emph{few} domestic applications compared to international ones, i.e., domestic students are the minority in the CS PhD application pool.
 Many US universities want to increase the number of domestic students in their programs (and as mentioned later, there are specific fellowships and funding for domestic students).

### 2025-02-19 demystify.tex
updates

@@
 Just show your awareness and understanding of the topic and what you did or how you plan to contribute to it (e.g., mentoring students, helping others).
-One of the best DEI statements I have read starts the acknowledgement that the applicant is not a minority, but a white male in a STEM field (i.e., they are the majority), and that they do not have much experience in DEI, but show that they are aware of the challenges faced by minority groups and how they can contribute to it.
+One of the best statements I have read starts the acknowledgement that the applicant is not a minority, but a white male in a STEM field (i.e., they are the majority), and that they do not have much experience in DEI, but show that they are aware of the challenges faced by minority groups and how they can contribute to it.
 \end{commentbox}

### 2025-02-23 demystify.tex
updates

@@
 Don't use too much coloring or fancy fonts (like those in Words).
-CS academics like using LaTeX (common way to write our papers and other documents), so write your SOP using LaTeX (with Times or default font, 11pt, and 1-inch margin).
+Don't use left alignment (seems to be default in Words) as it is hard to read.
+CS academics like using \textbf{LaTeX} (common way to write our papers and other documents), so write your SOP using LaTeX (with Times or default font, 11pt, and 1-inch margin).
 \end{itemize}
@@
 
-Many universities now require a diversity, equity, and inclusion (DEI) statement as part of the application.
+Some universities require a diversity, equity, and inclusion (DEI) statement as part of the application.
 Students, especially international ones, often struggle with what to write, especially when they are not from a minority group or do not have much experience in DEI.
@@
 
-Having an external and major fellowship can significantly improve your chances of getting into a good PhD program (\autoref{sec:fellowships}).
-The National Science Foundation Graduate Research Fellowship Program (NSF GRFP) and the Department of Defense National Defense Science and Engineering Graduate Fellowship (DoD NDSEG) are two of the most prestigious fellowships for CS PhD students who are US citizens or permanent residents.
-These fellowship provide multiple years of full financial support, which includes a stipend and tuition. 
-It is rumored that several programs would revert their admission decision on a student if they get one of these fellowships.
-
-The \href{https://www.nsfgrfp.org/}{GRFP} and \href{https://www.ndsegfellowships.org/}{NDSEG} website have detailed information on the application process and eligibility requirements.
-This section focuses on how your application is evaluated and tips on improving your application.
+Having an external and major fellowship can significantly relieve the financial burden of your PhD and improve your chances of getting into a good PhD program (\autoref{sec:fellowships}). 
+Among the most prestigious fellowships for CS PhD students who are US citizens or permanent residents are the National Science Foundation Graduate Research Fellowship Program (NSF GRFP) and the Department of Defense National Defense Science and Engineering Graduate Fellowship (DoD NDSEG).
+These fellowships provide multiple years of full financial support, which includes a stipend and tuition.
+
+The \href{https://www.nsfgrfp.org/}{GRFP} and \href{https://www.ndsegfellowships.org/}{NDSEG} websites have detailed information on the application process and eligibility requirements.
+This section focuses on how your application is evaluated and tips on improving it.
 Note that while these are specific GRFP and NDSEG, the general advice can be applied to other major fellowships (those for non-US citizens or permanent residents).
@@
 
-\item \textbf{Too Technical.} Remember that you're writing to a general audience in CS, not experts in your field.
+\item \textbf{Too Technical or Narrow.} Remember that you're writing to a general audience in CS, not experts in your field.
 While NSF tries to match reviewers with your field, CS fields are simply too broad and you might get a reviewer in the field but not in your specific area or familiar with the problem you're working on.
-So avoid technical jargon and explain things in simple terms. If the reviewer cannot understand your proposal, they will likely give you a low score.
-
-You should ask your professor or someone who's familiar with writing NSF proposals to review your reseach plan.  They can tell you if it is too technical and help you simplify it.
+So avoid technical jargon and explain things in simple terms. If the reviewer cannot understand your research plan, they will likely give you a low score.
+
+You should ask your professor or someone who's familiar with writing NSF proposals to review what you wrote.  They can tell you if it is too technical or narrow and help you improve it.
 
@@
 
-Finally, just as with the NSF GRFP and even PhD SOPs, ask your LoR writers and adviser to look at your writing/essays.
+\section{Common Tips for Both Fellowships}
+
+\begin{itemize}
+  \item \textbf{Use LaTeX to write your research plan and essays}.
+This is the standard in CS and engineering and will make your application look more professional.
+Moreover, do not use the default Word settings, e.g., left justification, as it is not standard in CS and engineering and difficult to read.
+
+  \item \textbf{Ask your LoR writers and adviser to look at your writing/essays}.
 You are new to the field and research writing, and your ideas might not be accurate or have already been explored.
 It's better that they tell you now than for you to find out later from reviewers.
+\end{itemize}
 

### 2025-02-23 phd-cs-us/demystify.tex
updates

--- 2025-02-23 demystify.tex
@@
     George Mason University, Dept. of Computer Science\\[0.1in]
-    \today{} (latest version available on  \href{https://github.com/nguyenthanhvuh/phd-cs-us}{Github})
+    \today{} (latest version available on  \href{https://github.com/dynaroars/books/phd-cs-us}{Github})
     };
@@
 \begin{center}
-  \href{https://nguyenthanhvuh.github.io/phd-cs-us/demystify.pdf}{nguyenthanhvuh.github.io/phd-cs-us/demystify.pdf},
+  \href{https://dynaroars.github.io/books/phd-cs-us/demystify.pdf}{dynaroars.github.io/books/phd-cs-us/demystify.pdf},
 \end{center}
 
-\noindent and its \LaTeX{} source is also on \href{https://github.com/nguyenthanhvuh/phd-cs-us}{GitHub}.
-If you have questions or comments, feel free to create new \href{https://github.com/nguyenthanhvuh/phd-cs-us/issues}{GitHub issues} or \href{https://github.com/nguyenthanhvuh/phd-cs-us/discussions}{discussions}.
+\noindent and its \LaTeX{} source is also on \href{https://github.com/dynaroars/books/phd-cs-us}{GitHub}.
+If you have questions or comments, feel free to create new \href{https://github.com/dynaroars/books/issues}{GitHub issues} or \href{https://github.com/dynaroars/books/discussions}{discussions}.
 

### 2025-02-23 demystify.tex
updates

@@
     George Mason University, Dept. of Computer Science\\[0.1in]
-    \today{} (latest version available on  \href{https://github.com/nguyenthanhvuh/phd-cs-us}{Github})
+    \today{} (latest version available on  \href{https://github.com/dynaroars/phd-cs-us}{Github})
     };
@@
 \begin{center}
-  \href{https://nguyenthanhvuh.github.io/phd-cs-us/demystify.pdf}{nguyenthanhvuh.github.io/phd-cs-us/demystify.pdf},
+  \href{https://dynaroars.github.io/phd-cs-us/demystify.pdf}{dynaroars.github.io/phd-cs-us/demystify.pdf},
 \end{center}
 
-\noindent and its \LaTeX{} source is also on \href{https://github.com/nguyenthanhvuh/phd-cs-us}{GitHub}.
-If you have questions or comments, feel free to create new \href{https://github.com/nguyenthanhvuh/phd-cs-us/issues}{GitHub issues} or \href{https://github.com/nguyenthanhvuh/phd-cs-us/discussions}{discussions}.
+\noindent and its \LaTeX{} source is also on \href{https://github.com/dynaroars/phd-cs-us}{GitHub}.
+If you have questions or comments, feel free to create new \href{https://github.com/dynaroars/phd-cs-us/issues}{GitHub issues} or \href{https://github.com/dynaroars/phd-cs-us/discussions}{discussions}.
 
@@
 This book also aims to explain the \textbf{reasons behind the admission process}.
-There are numerous \href{https://github.com/nguyenthanhvuh/phd-cs-us}{resources online} that tell you \emph{``what''} to do, but few explain the \emph{``why''}, e.g., why LoRs matter so much, why you should not draft your own LoR, why you should contact professors, etc.
+There are numerous \href{https://github.com/dynaroars/phd-cs-us}{resources online} that tell you \emph{``what''} to do, but few explain the \emph{``why''}, e.g., why LoRs matter so much, why you should not draft your own LoR, why you should contact professors, etc.
 Understanding the reason and mindset of the admission committee and profs.
@@
 The book was initially intended for international students but has been updated to include domestic students (\autoref{chap:domestic-students} and \autoref{chap:fellowships}).
-The book is open-source and available on \href{https://github.com/nguyenthanhvuh/phd-cs-us}{GitHub} so that others can contribute to it.
+The book is open-source and available on \href{https://github.com/dynaroars/phd-cs-us}{GitHub} so that others can contribute to it.
 \paragraph{About Me} I am an associate professor in the CS dept at George Mason University (Mason).
